### 2023-06-02 popl24/hyperprop.tex
adding manuscript source directory

\section{Hyperproperties}

Finally, consider our secure communications example, that we call program
$\prog_3$. We will assume that $H = \{ 1,2 \}$ and $C = \{ 0 \}$, so
the adversary has access to $\vx{0}{0}$ in particular:
\begin{verbatimtab}
v[2,0] := f[1,0];
v[0,0] := f[1,0] xor s[1,0];
v[2,1] := v[0,0];
v[2,2] := v[2,1] xor v[2,0]
\end{verbatimtab}
First, let's focus on information available to the adversary
about the secret $\sx{1}{0}$ (aside from its publicly-known marginal
distribution). We will consider the marginal distribution
of $\sx{1}{0}$ and $\vx{0}{0}$:
\begin{mathpar}
X \defeq \{ \vx{0}{0}, \sx{1}{0} \}

\pdf{X} \defeq \margd{\progd(\prog_3)}{X}
\end{mathpar}
Since the adversary can directly observe the value in $\vx{0}{0}$ in
any given run, we therefore consider the distribution conditioned on
that value. As we can see, this shows that the protocol reveals
no information about the secret:
\begin{mathpar}
  \condd{(\pdf{X}}{\{ \vx{0}{0} \mto 0\})}(\{ \sx{1}{0} \mapsto 0\}) = .5

  \condd{(\pdf{X}}{\{ \vx{0}{0} \mto 0\})}(\{ \sx{1}{0} \mapsto 1\}) = .5

  \condd{(\pdf{X}}{\{ \vx{0}{0} \mto 1\})}(\{ \sx{1}{0} \mapsto 0\}) = .5

  \condd{(\pdf{X}}{\{ \vx{0}{0} \mto 1\})}(\{ \sx{1}{0} \mapsto 1\}) = .5
\end{mathpar}

Now, let's focus on the information directly available to honest parties,
and specifically the information available through $\vx{2}{2}$ about
$\sx{1}{0}$. We first define the relevant marginal distribution:
\begin{mathpar}
X \defeq \{ \vx{2}{2}, \sx{1}{0} \}

\pdf{X} \defeq \margd{\progd(\prog_3)}{X}
\end{mathpar}
Then we consider the marginal distribution of the secret conditioned
on the view $\vx{2}{2}$, which shows the direct correlation and therefore
correctness. 
\begin{mathpar}
  \condd{(\pdf{X}}{\{ \vx{0}{0} \mto 0\})}(\{ \sx{1}{0} \mapsto 0\}) = 1

  \condd{(\pdf{X}}{\{ \vx{0}{0} \mto 0\})}(\{ \sx{1}{0} \mapsto 1\}) = 0

  \condd{(\pdf{X}}{\{ \vx{0}{0} \mto 1\})}(\{ \sx{1}{0} \mapsto 0\}) = 0

  \condd{(\pdf{X}}{\{ \vx{0}{0} \mto 1\})}(\{ \sx{1}{0} \mapsto 1\}) = 1
\end{mathpar}

\subsection{Probabilistic Noninterference} The discussion above suggests a
security semantics- that is, a hyperproperty- analagous to
noninterference in a probabilistic setting. First, let's
define low equivalence of memories in the current setting:
\begin{definition}[Memory Low Equivalence]
  Given $\store^1$ and $\store^2$ with $\dom(\store^1) = \dom(\store^2) = X$,
  we write $\store^1 =_C \store^2$ iff $\store^1_{X_C} = \store^2_{X_C}$.
\end{definition}
Now, we can define probabilistic noninterference as a property of
programs ensuring that low-equivalent initial memories produce
the same low-observable trace (i.e., corrupt views) \emph{with the
same probability}. 
\begin{definition}[Probabilistic Noninterference]
  Given a program $\prog$ with $\iov(\prog) = (S,V)$, we say that
  $\prog$ satisfies \emph{probabilistic noninterference} iff for all
  $\store_1, \store_2 \in \mems(S)$:
  $$\store_1 =_C \store_2 \implies
    \margd{(\condd{\progd(\prog)}{\store_1})}{V_C} =
    \margd{(\condd{\progd(\prog)}{\store_2})}{V_C}$$
\end{definition}

An important consequence of probabilistic noninterference refines an
idea we introduced in Lecture Notes 12, which is that noninterering
programs do not improve adversarial knowledge about secrets.
More precisely, in $\fedcat$ the initial probability of a correct
guess by the adversary is the marginal distribution of honest secrets
in the program distribution conditioned by corrupt inputs.
The final (and thus best, due to monotonicity) adversarial
knowledge is also the marginal distribution of honest secrets in the
program distribution, but conditioned by corrupt inputs \emph{and}
corrupt views. Noninterference in the probabilistic setting implies
that these are the same.
\begin{lemma}
  Given a program $\prog$ with $\iov(\prog) = (S,V)$, 
  $\prog$ satisfies probabilistic noninterference iff for all
  $\store \in \mems(S \cup V)$:
  $$\margd{(\condd{\progd(\prog)}{\store_{S_C}})}{S_H} =
    \margd{(\condd{\progd(\prog)}{\store_{(S\cup V)_C}})}{S_H} $$
\end{lemma}

\example{Example: Additive Secret Sharing} A nice ``Hello World''
example of an MPC protocol is additive secret sharing, where $k$
parties split up their secrets into \emph{shares}, with the property
that all $k$ shares are required to reconstruct the secret and any
fewer reveals nothing. In $\fedcat$, we can use $\ttt{xor}$, again, to
generate shares. Here, party 1 generates 3 shares of the secret
$\sx{1}{1}$- one is given to party 2 through view $\vx{2}{s1}$, one is
given to party 3 through view $\vx{3}{s1}$, and party 1 keeps
$\fx{1}{s1}$ for itself:
\begin{verbatimtab}
v[2,s1] := flip[1,1] xor flip[1,s1] xor s[1,1];
v[3,s1] := flip[1,1];
\end{verbatimtab}
Note that $\vx{2}{s1}$ and $\vx{3}{s1}$ are both in uniform
random distributions.

Another property of $\ttt{xor}$ is that it is additively homomorphic,
while functioning as addition in the boolean algebra (i.e., the binary
field). That is, given:
\begin{mathpar}
  x_1\ \ttt{xor}\ y_1 = z_1

  x_2\ \ttt{xor}\ y_2 = z_2
\end{mathpar}
then we have:
$$
(x_1\ \ttt{xor}\ x_2)\ \ttt{xor}\ (y_1\ \ttt{xor}\ y_2) = z_1\ \ttt{xor}\ z_2
$$
Thus, we can obtain secret 3-party summation (xor) by continuing
the protocol begun above- parties 2 and 3 can share their secrets
in the same fashion:
\begin{verbatimtab}
v[1,s2] := flip[2,1] xor flip[2,s2] xor s[2,1];
v[3,s2] := flip[2,1];

v[1,s3] := flip[3,1] xor flip[3,s3] xor s[3,1];
v[2,s3] := flip[3,1];
\end{verbatimtab}
Then, assuming that party 0 is the ``public party'', each party sums
its own shares, and then the sum of sum of shares is revealed
in the output view $\vx{0}{output}$. By additive homomorphism of
$\ttt{xor}$, this output is the sum of the 3 secrets, but
the protocol reveals nothing to the adversary, even if one
(or even two, in fact) of $\{1,2,3\}$ are corrupt. 
\begin{verbatimtab}
v[0,ss1] := v[1,s2] xor v[1,s3] xor flip[1,s1];
v[0,ss2] := v[2,s1] xor v[2,s3] xor flip[2,s2];
v[0,ss3] := v[3,s1] xor v[3,s2] xor flip[3,s3];

v[0,output] := v[0,ss1] xor v[0,ss2] xor v[0,ss3]
\end{verbatimtab}

Because this protocol correctly implements 3-party addition in the
binary field, it can reveal information to the adversary. For example,
suppose that $1 \in C$, and $\sx{1}{1}$ is $1$, and after running the
protocol suppose that $\vx{0}{output}$ is 0. Then the adversary knows
that 2 and 3's secrets are either 1 and 0, or 0 and 1, with 50/50
probability, and definitely neither 0 and 0, nor 1 and 1. However, the
protocol does not reveal anything more through adversarial views than
what the output reveals.

\medskip

\subsection{Passive Security and the Real/Ideal Model} The traditional
security model of MPC is formulated by positing a \emph{simulator} that
exists in the ideal world, and is trying to reconstruct observable
information in a real world run of a protocol $\prog$ that
correctly implements an ideal functionality $\idealf$- that is,
execution of a configuration $(\store,\prog)$. We will assume
that any protocol $\prog$ includes a public \emph{output view}
$x$ where the result is published. The protocol is
\emph{correct} iff the resulting memory contains $x \mapsto
\idealf(\store_S)$, where $S$ are the secret inputs (honest
and corrupt) in $\prog$.

The simulator is given just the corrupt inputs to the run- that is,
$\store_{S_C}$- and the output of the protocol- that is,
$\idealf(\store)$- and tries to reconstruct the \emph{distribution of
adversarial views} observed in the real world. The intuition here is
that the simulator is given the information that is necessarily
available to the adversary by observing a run of the protocl, and if
the simulator can reconstruct whatever information appears in the real
world adversarial views, by reconstructing their distributions, then
those views leak no additional information, i.e., \emph{have no
additional dependencies on honest input secrets than what is revealed
by the output and conditionings on corrupt inputs}.

The simulator is represented by a probabilistic algorithm $\SIM_C$,
aka a \emph{simulation}, that is parameterized by a corrupt inputs and
the output of an ideal functionality, and that returns a set of
adversarial views (as a memory) with some probability. Given
corrupt inputs $\store$ and ideal functionality output $v$,  
we write:
$$
\prob(\SIM_C(\store,v) = \store')
$$
to denote the probability that $\SIM_C(\store,v)$
returns corrupt views $\store'$ as a result. We can then define the
probability distribution of corrupt views reconstructed
by the simulator as follows:
\begin{definition}
  Given $C$, $\store$, and $v$, we write $\progd(\SIM_C(\store,v))$ to
  denote the distribution of corrupt views reconstructed by the
  simulation, which is a distribution $\pdf{V}$ where for
  all $\store' \in \mems(V)$:
  $$
  \pdf{V}(\store')\ \defeq\ \prob(\SIM_C(\store,v) = \store') 
  $$
\end{definition}

Then we can define passive security in the real/ideal
model as follows. One detail here is the assumption of an
\emph{honest majority} which is a standard bound on the
capabilities of the adversary in this threat model.
\begin{definition}[Passive Security]
  Assume given a program $\prog$ that correctly implements an ideal
  functionality $\idealf$, with $\iov(\prog) = (S,V)$.  Then $\prog$
  is \emph{passive secure in the simulator model} iff for all
  partitions of the federation into honest and corrupt sets $H$ and $C$
  with $|C| < |H|$ and for all $\store \in \mems(S)$ there exists a
  simulation $\SIM_C$ such that:
  $$
  \progd(\SIM_C(\store_{S_C},\idealf(\store))) = \margd{(\condd{\progd(\prog)}{\store})}{V_C}
  $$
\end{definition}

%\newcommand{\outv}{\mathit{out}}

\subsection{A Sound Hyperproperty for Passive Security} The real/ideal
model can be hard to understand and passive security of protocols can
be difficult to prove, much less enforce automatically. Formulating
a hyperproperty that implies passive security can provide some
insights into the meaning of passive security, provide a connection
to the foundational theories we've discussed, and can provide
methodological support for proofs and even enforcement.

The intuition that passive security means that the protocol releases
no more information than what is released by the output suggests a
refinement of noninterference. Specifically, it suggests that in
passive secure protocols, the adversarial knowledge (in the Lecture
Notes 12 sense) obtained from the output does not change during
protocol execution, which otherwise should be noninterfering.  That
is, the set of low equivalent initial memories that can produce the
same output should generate the same distributions of adversarial
views. 
\begin{definition}[Noninterference Modulo Output]
  Given $H$, $C$, and a program $\prog$ with $\iov(\prog) = (S,V)$ and output
  view $x \in V$, we say that
  $\prog$ satisfies \emph{noninterference modulo output} iff for all
  $\store_1,\store_2 \in \mems(S)$:
  \begin{eqnarray*}
    & \store_1 =_C \store_2 \wedge
     (\margd{(\condd{\progd(\prog)}{\store_1})}{\{ x \}} =
     \margd{(\condd{\progd(\prog)}{\store_2})}{\{ x \}}) \\
    & \implies \\
    & \margd{(\condd{\progd(\prog)}{\store_1})}{V_C} =
    \margd{(\condd{\progd(\prog)}{\store_2})}{V_C}
  \end{eqnarray*}
\end{definition}
We can show, for example, that the 3-party xor protocol defined above
satisfies noninterference modulo output.

What is the formal connection with passive security? Note in
particular that the hyperproperty implies that the simulator can use
the protocol itself to reconstruct the distribution of adversarial
views by making a random guess about honest secrets given public
knowledge about the ideal functionality $\idealf$ a protocol
implements. However, there is a subtlety here in that the adversary
must be able to tractably reconstruct this knowledge. This is not
generally true, in particular if $\idealf$ is a 1-way
function. Define:
\begin{definition}[Ideal Knowledge]
  Given a functionality $\idealf$ and output value $v$, the associated
  \emph{ideal knowledge}, denoted $\ik(\idealf,v)$ is:
  $$
  \{ \store | \idealf(\store) = v \}
  $$
  We say that $\idealf$ is \emph{invertible} iff $\ik(\idealf, v)$ for all
  $v$ can be computed tractably.
\end{definition}
Now we can show our main result, that noninterference modulo output,
together with the assumption that $\idealf$ is invertible, implies
passive security of a protocol\footnote{We emphasize that the converse
is not asserted.}. 
\begin{theorem}
  Assume given invertible $\idealf$ and a protocol $\prog$ that
  correctly implements $\idealf$.  If program $\prog$ satisfies
  noninterference modulo output for all $H$ and $C$ with $|H| < |C|$,
  then $\prog$ is passive secure.
\end{theorem}

\begin{proof}
  Let $H$ and $C$ be an arbitrary partition of the federation and
  suppose $\prog$ satifies noninterference modulo output. Let
  $\iov(\prog) = (V,S)$ and $\flips(\prog) = F$ with output view $x
  \in V$ and let $\store$ be an arbitrary member of $\mems(S \cup
  F)$. Then the distribution of adversarial views in the real world
  is, by definition:
  $$\margd{(\condd{\progd(\prog)}{\store_S})}{V_C}$$

  In the ideal world, the simulator is given both $\store_{S_C}$ and
  $\idealf(\store_S)$.  The simulation $\SIM_C(\store_{S_C},
  \idealf(\store_S))$ can be defined as follows. First, some $\store'
  \in \ik(\idealf,\idealf(\store))$ is randomly chosen such that
  $\store' =_C \store_S$, as is a random tape $\store'' \in
  \mems(F)$\footnote{The real/ideal model allows consultation of a
  ``Random Oracle''.}. Then, the run of $(\store' \cup \store'',
  \prog)$ is evaluated in simulation, yielding $(\store^{\SIM},\varnothing)$.
  The simulation returns $\store^{\SIM}_{V_C}$ as a result.

  Now, since the random tape is selected in simulation from the same distribution
  that we assume for the real world, after selection of $\store'$ the
  probability that any particular $\store^{\SIM}_{V_C}$ is returned is by definition:
  $$
   (\margd{(\condd{\progd(\prog)}{\store'})}{V_C})(\store^{\SIM}_{V_C})
  $$
  Furthermore, since we assume that $\prog$ correctly implements $\idealf$, this
  means:
  $$
  \store_S =_C \store' \wedge
     (\margd{(\condd{\progd(\prog)}{\store_S})}{\{ x \}} =
      \margd{(\condd{\progd(\prog)}{\store'})}{\{ x \}})
  $$
  so by the definition of noninterference modulo output the choice
  any choice of $\store'$ yields the same distribution, i.e.:
  $$\margd{(\condd{\progd(\prog)}{\store_S})}{V_C}$$
  Thus, by definition:
  $$
   \progd(\SIM_C(\store_{S_C},\idealf(\store))) = \margd{(\condd{\progd(\prog)}{\store_S})}{V_C}
  $$
  \qed
\end{proof}

\begin{lemma}
  Given $H$, $C$, and a program $\prog$ with $\iov(\prog) = (S,V)$ and
  output view $x \in V$
  $\prog$ satisfies noninterference modulo output iff for all
  $\store \in \mems(S \cup V)$:
  $$\margd{(\condd{\progd(\prog)}{\store_{(S \cup \{x\})_C}})}{S_H} =
    \margd{(\condd{\progd(\prog)}{\store_{(S\cup V)_C}})}{S_H} $$
\end{lemma}

### 2023-06-02 popl24/main.tex
adding manuscript source directory

\documentclass[acmsmall,screen,review]{acmart}

\usepackage{amsmath}
\usepackage{amstext}
\usepackage{fp-frame}
\usepackage[latin1]{inputenc}
\usepackage{mathpartir}
\usepackage{fancyvrb}
\usepackage{moreverb}
\usepackage{stmaryrd}
\usepackage{enumerate}

\input{macros}

\begin{document}

\title{Automating MPC Security}

\author{Author names withheld for double-blind reviewing.}

\begin{abstract}
  We show how to formulate a sound hyperproperty of MPC passive security, and how this
  supports automated verification.
\end{abstract}

\maketitle

\input{minicat}

\input{pmf}

\input{hyperprop}

\bibliographystyle{ACM-Reference-Format}
\bibliography{sample-base}

\end{document}
\endinput

### 2023-06-02 popl24/minicat.tex
adding manuscript source directory

\section{The $\minicat$ Language}

The language $\minicat$ comprises a standard syntax of
boolean algebra and arithmetic expressions over integers. In addition, we allow reference
to \emph{variables} in expressions. \emph{Programs} are possibly empty
lists of assignments of expressions to variables.
$$
\begin{array}{rcl@{\qquad\qquad}r}
x,y,z &\in& \mathcal{V} & \textit{variables}\\
n &\in& \mathbb{Z} & \textit{integers}\\
b &::=& \true \mid \false & \textit{booleans}\\
v &::=& n \mid b & \textit{values} \\[2mm]
\mathit{unop} &::=& \mathrm{not} \mid - & \textit{unary operations} \\
\mathit{binop} &::=& \mathrm{and} \mid \mathrm{or} \mid + \mid * \mid \mod \mid\ =\ \mid & \textit{binary operations} \\[2mm]
e &::=& x \mid v \mid\, \mathit{unop}\, e \mid e\, \mathit{binop}\, e \mid  \eite{e}{e}{e}  & \textit{expressions}\\[2mm]
\prog &::=& \varnothing\quad \mid\quad \eassign{x}{e};\ \prog & \textit{programs}
\end{array}
$$
So for example, the following is a program- note that it assume the existence of an
\emph{input} variable $x$, which is undefined by the program itself:
$$
\eassign{y}{x + 1}; \eassign{z}{2 * y}
$$

For variable-free expressions in $\minicat$,
we can extend our interpretation of
boolean algebra $\cod{\cdot}$ to integer arithmetic in the obvious
manner, so for example $\cod{2 * 4} = 8$. However, to resolve
expressions that include variables, we need to define the semantics in
terms of references to \emph{memories}. We take the view that memories
map variables to values.
\begin{definition}[Memories]
A \emph{memory}, aka \emph{store} $\store$ is a finite (partial) mapping from variables to values.
The \emph{domain} of a memory is written $\dom(\store)$ and is the finite
set of variables on which the memory is defined. We write $\store\{ x \mapsto v\}$
to denote the memory $\store'$ such that $\store'(x) = v$ and otherwise
$\store'(y) = \store(y)$ for all $y \in (\dom(\store) - \{ x \}) $.
\end{definition}
Now we can modify our interpretation of expressions to take into account a given memory.
We write $\cod{\store, e}$ to mean the interpretation of $e$ where each variable $x$ in $e$
is replaced by $\store(x)$. In full detail, we define this as follows:
\begin{eqnarray*}
\cod{\store, x} &=& \store(x)\\
\cod{\store, \mathit{unop}\ e} &=& \cod{\mathit{unop}\ \cod{\store,e}}\\
\cod{\store, e_1\ \mathit{binop}\ e_2} &=& \cod{\cod{\store,e_1}\ \mathit{binop}\ \cod{\store,e_2}}\\
\cod{\store, \eite{e_1}{e_2}{e_3}} &=&
             \begin{cases}
                \cod{\store,e_2} & \text{if\ } \cod{\store,e_1}\\
                \cod{\store,e_3} & \text{otherwise}
             \end{cases}
\end{eqnarray*}

A critical point is that the semantics of $\minicat$ is defined not
just with respect to programs, but with respect to a pair of a memory
and a program $(\store, \prog)$, which we call
a \emph{configuration}. Together with the interpretation of
expressions that reference memories, we can obtain
an \emph{operational} semantics for configurations. This semantics
specifies how memories evolve during computation as a result of
assignments to variables in programs. The fundamental relation is
the \emph{reduction} aka \emph{evaluation} relation $\redx$,
which is a relation on configurations, that we define as follows:
$$
(\store, \eassign{x}{e};\prog) \redx (\extend{\store}{x}{\cod{\store,e}}, \prog)
$$

\subsection{Synchronous Distributed Protocols} With some minor syntactic
modifications to $\minicat$, we can model and reason about security in
protocols for distributed communications. We will call this modified
system $\fedcat$, alluding to the existence of a group of participants
we call a \emph{federation}. We restrict ourselves to boolean values
in $\fedcat$, which we will denote $1$ and $0$ for convenience.  We
assume the full set of the usual boolean operations including xor.

We identify participants by integers $n$, and assume that the
federation can be partitioned into two sets $H$ and $C$ where $H$ are
all \emph{honest} and $C$ are \emph{corrupt}. Then, we modify
variables $x$ to be of the form $\sx{n}{s}$ denoting a \emph{secret}
belonging to party $n$ and distinguished by a string $s$, or of
the form $\vx{n}{s}$ denoting a \emph{view} belonging to party $n$ and
distinguished by $s$. A view is a memory location that records a value
received by a party. So for example, in the following program, party 1
sends the negation of their secret $\ttt{foo}$ to party 2, with the
value recorded in party 2's $\ttt{bar}$ view:
$$
\eassign{\vx{2}{bar}}{\Dnot (\sx{1}{foo})}
$$ In $\fedcat$ we require that all assignments are to views, and in
any assignment $\eassign{\vx{n}{s}}{e}$ we require all variables that
occur in $e$ to belong to the same party. We will also assume that
secrets are defined in initial memories, that views are not, and that
views are never redefined (overwritten). This models communication
between parties through views, where each party can compute locally on
its own values, and ensures that the final memory after computation
contains a record of all secrets and views. We will use
\emph{protocol} as a synonym for program in $\fedcat$ given our intent
to model synchronous distributed computations.

 Although $\fedcat$ is a simple
language we can explore some surprisingly complex distributed
protocols with the introduction of randomized elements. In $\fedcat$
we accomplish this by introducing a new form of variables called
\emph{flips}, written $\fx{n}{s}$, which are owned by a party $n$ and
distinguished by string $s$. We will assume that flips can be either
$1$ or $0$ with equal probability.  So for example, in this
program, party 1 sends their flip $\ttt{foo}$ to party 2,
with the value recorded in party 2's $\ttt{bar}$ view:
$$
\eassign{\vx{2}{bar}}{\fx{1}{foo}}
$$
We don't change anything about the semantics of $\fedcat$
to model this randomization, rather we will assume that all
flip values are randomly assigned in initial memories. This is
the so-called ``random tape'' approach.

### 2023-06-02 popl24/pmf.tex
adding manuscript source directory

\section{Probability Mass Functions and Program Distributions}

\begin{definition}
  A \emph{memory probability distribution} $\pdf{X}$ is a function mapping
  memories in $\mems(X)$ to values in $[0..1]$, such that:
  $$
  \sum_{\store \in \mems(X)} \pdf{X}(\store) \  = \ 1
  $$
  We call $X$ the \emph{domain} of $\pdf{X}$.
\end{definition}

\begin{definition}
  Given $X$ the \emph{marginal distribution} of $Y \subseteq X$
  in a distribution $\pdf{X}$, denoted $\margd{\pdf{X}}{Y}$,
  is a distribution with domain $Y$ where for all
  $\store \in \mems(Y)$:
  $$
  (\margd{\pdf{X}}{Y})(\store) =
  \sum_{\store' \in \mems(X-Y)} \pdf{X}(\store \cup \store')
  $$
\end{definition}

\begin{definition}
  Given $\pdf{X}$, its \emph{conditional distribution given
  $\store$} with $\dom(\store) = Y$ and $Y \subseteq X$, denoted
  $\condd{\pdf{X}}{\store}$, is a distribution with domain $X - Y$ where for all
  $\store' \in \mems(X - Y)$:
  $$
  (\condd{\pdf{X}}{\store})(\store') =
      \pdf{X}(\store \cup \store') / ((\margd{\pdf{X}}{Y})(\store))
  $$
\end{definition}

Now we can define the probability distribution of a program $\prog$,
that we denote $\progd(\prog)$. Recall that $\fedcat$ is
deterministic, so a run is determined by the input secrets together
with random flip assignments. And since we constrain programs to not
overwrite memories, we can be sure that final memories resulting from
computation contain both a complete record of all initial secrets and
flips as well as views of communicated information. Thus a final
memory represents a run of the program, and since we consider and set
of \emph{input} secrets and flips to be equally likely, we consider
each memory in the set of final memories of a program to be equally
likely, which establishes the basic probability distribution for the
program. And since we don't care about flip values which are only
obversable insofar as they affect views, the probability distribution
we are really interested in is the marginal distribution of secrets
and views in this basic program distribution.
\begin{definition}
  Given progam $\prog$ with $\iov(\prog) = (S,V)$ and $\flips(\prog) = F$, define $M$ as:
  $$
  M \defeq \{ \store \mid (\dom(\store) = S \cup V \cup F) \wedge (\store_{S \cup F},\prog) \redxs (\store,\varnothing) \}
  $$
  Define also $\pdf{S \cup V \cup F}$ as the program's \emph{basic distribution} such that for all
  $\store \in \mems(S \cup V \cup F)$:
  $$
  \pdf{S \cup V \cup F}(\store) =
  \begin{cases}
    1 / |M| & \text{if}\ \store \in M\\
    0 & \text{otherwise}
  \end{cases}
  $$
  Then the \emph{program distribution of $\prog$}, denoted $\progd(\prog)$, is the
  marginal distribution of $S \cup V$ in $\prog$'s basic distribution:
  $$
  \progd(\prog) =  \margd{\pdf{S \cup V \cup F}}{S\cup V}
  $$
\end{definition}

### 2023-06-02 popl24/metalang.tex
checkpoint

\section{A Protocol Metalanguage}

$$
\begin{array}{rcl@{\hspace{8mm}}r}
b &\in& \{ \etrue, \efalse \} \\
w &\in& \mathrm{String} \\ 
\cid &\in& \mathrm{Clients} \qquad  (\mathrm{Clients} \subset \mathbb{N} \text{\ in\ this\ presentation})\\[2mm]
\bop &\in& \{ \eand, \eor, \exor \} \\[2mm]
\be &::=& b \mid \flip{\cid}{w} \mid \secret{\cid}{w} \mid \view{\cid}{w} \mid \oracle{w} \mid & \textit{boolean expressions}\\
& &  \enot\ \be \mid \be\ \bop\ \be \mid \OT{\be}{\be}{\be} \mid \select{\be}{\be}{\be} \\[2mm]
\instr &::=& \eassign{\view{\cid}{w}}{\be} \mid \eassign{\view{\cid}{w}}{\OT{\be}{\be}{\be}} & \textit{view assignments} \\[2mm]
\prog &::=& \instr \mid \instr; \prog & \textit{protocols}
\end{array}
$$

\begin{eqnarray*}
\lcod{\store, \etrue}{\cid} &=& \etrue\\
\lcod{\store, \efalse}{\cid} &=& \efalse\\
\lcod{\store, \flip{\cid}{w}}{\cid} &=& \store(\flip{\cid}{w})\\
\lcod{\store, \secret{\cid}{w}}{\cid} &=& \store(\secret{\cid}{w})\\
\lcod{\store, \view{\cid}{w}}{\cid} &=& \store(\view{\cid}{w})\\
\lcod{\store, \oracle{w}}{\cid} &=& \store(\oracle{w})\\
\lcod{\store, \enot\ \be}{\cid} &=& \cod{\enot\ \lcod{\store,\be}{\cid}}\\
\lcod{\store, \be_1\ \mathit{binop}\ \be_2}{\cid} &=&
    \cod{\lcod{\store,\be_1}{\cid}\ \mathit{binop}\ \lcod{\store,\be_2}{\cid}}\\
\lcod{\store, \select{\be_1}{\be_2}{\be_3}}{\cid} &=&
             \begin{cases}
                \lcod{\store,\be_2}{\cid} & \text{if\ } \lcod{\store,\be_1}{\cid}\\
                \lcod{\store,\be_3}{\cid} & \text{otherwise}
             \end{cases}
\end{eqnarray*}

\begin{mathpar}
  (\store, \eassign{\view{\cid}{w}}{\be};\prog) \redx (\extend{\store}{\view{\cid}{w}}{\lcod{\store,\be}{\cid}}, \prog)

  \inferrule
  {b = \text{if\ } \lcod{\store,\be_1}{\cid}  \text{\ then\ } \lcod{\store,\be_2}{\cid'} \text{\ else\ } \lcod{\store,\be_3}{\cid'}}
      {(\store, \eassign{\view{\cid}{w}}{\OT{\be_1}{\be_2}{\be_3}};\prog) \redx (\extend{\store}{\view{\cid}{w}}{b}, \prog)}
\end{mathpar}

$$
\begin{array}{rcl@{\hspace{8mm}}r}
b &\in& \{ \etrue, \efalse \} \\
w &\in& \mathrm{String} \\ 
\flab &\in& \mathrm{Field}\\
s,x &\in& \mathrm{EVar}\\
f &\in& \mathrm{FName}\\
\cid &\in& \mathrm{Clients} \qquad  \mathrm{Clients} = \mathbb{N} \text{\ is\ OK}\\
\alpha,\beta &\in& \mathrm{DVar} \qquad \ttt{'a}, \ttt{'b} \text{\ etc.\ is\ OK}\\[5mm]
% \str &::=& w \ \mid \ s \ \mid \ \str||\str & \textit{string terms}\\[2mm]
e &::=& b \mid \flip{e}{e} \mid \secret{e}{e} \mid \view{e}{e} \mid \oracle{e} \mid \enot\ e \mid e\ \eand\ e \mid e\ \exor\ e \mid & \textit{expressions}\\[1mm]
& & \OT{e}{e}{e} \mid \select{e}{e}{e} \mid 
\send{\view{e}{e}}{e} \mid e;e \mid \elet{x}{e}{e} \mid\\[1mm]
& & x \mid f(e,\ldots,e) \mid \{ \flab = e; \ldots; \flab = e \}
\mid e.\flab \mid e\ttt{||}e \mid (e) \\[2mm]
v &::=& w \mid b \mid \cid \mid \{ \flab = v;\ldots;\flab = v \} 
 \mid (\,) & \textit{values}
\end{array}
$$

The syntax of functions is as follows. We require type annotations 
formal parameters. These types support the categories of values. 
The full type syntax is richer than the syntax of type annotations, 
as we show below, but a restricted language here simplifies type 
checking.
$$
\begin{array}{rcl@{\hspace{15mm}}r}
{fn} &::=& f(x : \srct,\ldots,x : \srct) \{ e \} & \textit{functions}\\[3mm]
\srct &::=& \cidty{x} \mid \stringty{x} \mid \jpdty{\alpha}{\alpha} \mid
    \{ \flab : \srct;\ldots;\flab : \srct \} 
& \textit{annotated types} \end{array}
$$

\begin{definition}
A \emph{codebase} $\codebase$ is a list of function 
declarations. We write $ \codebase(f) = x_1,\ldots,x_n,\ e$
iff $f(x : \srct,\ldots,x : \srct) \{ e \} \in \codebase$.
A \emph{program}, aka \emph{protocol} $\prot$ is a pair of a 
codebase and expression $\codebase, e$. We may omit
$\codebase$ if it is clear from context.  
\end{definition}

### 2023-06-03 popl24/metalang.tex
checkpoint

@@
 \be &::=& b \mid \flip{\cid}{w} \mid \secret{\cid}{w} \mid \view{\cid}{w} \mid \oracle{w} \mid & \textit{boolean expressions}\\
-& &  \enot\ \be \mid \be\ \bop\ \be \mid \OT{\be}{\be}{\be} \mid \select{\be}{\be}{\be} \\[2mm]
+& &  \enot\ \be \mid \be\ \bop\ \be \mid \select{\be}{\be}{\be} \\[2mm]
 \instr &::=& \eassign{\view{\cid}{w}}{\be} \mid \eassign{\view{\cid}{w}}{\OT{\be}{\be}{\be}} & \textit{view assignments} \\[2mm]
@@
 \begin{array}{rcl@{\hspace{8mm}}r}
-b &\in& \{ \etrue, \efalse \} \\
-w &\in& \mathrm{String} \\ 
 \flab &\in& \mathrm{Field}\\
-s,x &\in& \mathrm{EVar}\\
-f &\in& \mathrm{FName}\\
-\cid &\in& \mathrm{Clients} \qquad  \mathrm{Clients} = \mathbb{N} \text{\ is\ OK}\\
-\alpha,\beta &\in& \mathrm{DVar} \qquad \ttt{'a}, \ttt{'b} \text{\ etc.\ is\ OK}\\[5mm]
-% \str &::=& w \ \mid \ s \ \mid \ \str||\str & \textit{string terms}\\[2mm]
-e &::=& b \mid \flip{e}{e} \mid \secret{e}{e} \mid \view{e}{e} \mid \oracle{e} \mid \enot\ e \mid e\ \eand\ e \mid e\ \exor\ e \mid & \textit{expressions}\\[1mm]
-& & \OT{e}{e}{e} \mid \select{e}{e}{e} \mid 
-\send{\view{e}{e}}{e} \mid e;e \mid \elet{x}{e}{e} \mid\\[1mm]
-& & x \mid f(e,\ldots,e) \mid \{ \flab = e; \ldots; \flab = e \}
+x &\in& \mathrm{EVar}\\
+f &\in& \mathrm{FName}\\[2mm]
+e &::=& b \mid \flip{e}{e} \mid \secret{e}{e} \mid \view{e}{e} \mid \oracle{e} \mid \enot\ e \mid e\ \eand\ e \mid e\ \exor\ e \mid & \textit{expressions}\\[0mm]
+& & \select{e}{e}{e} \mid 
+\send{\view{e}{e}}{e} \mid \send{\view{e}{e}}{\OT{e}{e}{e}} \mid e;e \mid \\[0mm]
+& & x \mid \elet{x}{e}{e} \mid f(e,\ldots,e) \mid \{ \flab = e; \ldots; \flab = e \}
 \mid e.\flab \mid e\ttt{||}e \mid (e) \\[2mm]
 v &::=& w \mid b \mid \cid \mid \{ \flab = v;\ldots;\flab = v \} 
- \mid (\,) & \textit{values}
+\mid (\,) & \textit{values}\\[2mm]
+{fn} &::=& f(x,\ldots,x) \{ e \} & \textit{functions}
 \end{array}
-$$
-
-The syntax of functions is as follows. We require type annotations 
-formal parameters. These types support the categories of values. 
-The full type syntax is richer than the syntax of type annotations, 
-as we show below, but a restricted language here simplifies type 
-checking.
-$$
-\begin{array}{rcl@{\hspace{15mm}}r}
-{fn} &::=& f(x : \srct,\ldots,x : \srct) \{ e \} & \textit{functions}\\[3mm]
-\srct &::=& \cidty{x} \mid \stringty{x} \mid \jpdty{\alpha}{\alpha} \mid
-    \{ \flab : \srct;\ldots;\flab : \srct \} 
-& \textit{annotated types} \end{array}
 $$
@@
 declarations. We write $ \codebase(f) = x_1,\ldots,x_n,\ e$
-iff $f(x : \srct,\ldots,x : \srct) \{ e \} \in \codebase$.
-A \emph{program}, aka \emph{protocol} $\prot$ is a pair of a 
+iff $f(x_1,\ldots,x_n) \{ e \} \in \codebase$.
+A \emph{metaprogram}, aka \emph{mataprotocol} is a pair of a 
 codebase and expression $\codebase, e$. We may omit
@@
 \end{definition}
+
+$$
+\begin{array}{rcl@{\hspace{3mm}}r}
+E &::=& [\,] \mid \enot\ E \mid E\ \bop\ e \mid v\ \bop\ E \mid & \textit{evaluation contexts}\\[1mm]
+& & \flip{E}{e} \mid \secret{E}{e} \mid \view{E}{e} \mid \oracle{E} \mid  \\[1mm]
+& & \flip{\cid}{E} \mid \secret{\cid}{E} \mid \view{\cid}{E} \mid \send{E}{e} \mid \send{\view{\cid}{w}}{E}\\[1mm]
+& & \OT{E}{e}{e} \mid \OT{v}{E}{e} \mid \OT{v}{v}{E} \mid\\[1mm]
+& & \select{E}{e}{e} \mid \select{v}{E}{e} \mid \select{v}{v}{E} \mid\\[1mm]
+& & \elet{x}{E}{e} \mid f(v,\ldots,v,E,e,\ldots,e) \mid \\[1mm]
+& & \{ \flab = v;\ldots;\flab = v;\flab = E;\flab = e;\ldots;\flab = e \} \mid E.\flab \mid E\ttt{||}e \mid v \ttt{||}E
+\end{array}
+$$
+
+$$
+\begin{array}{rcl@{\hspace{10mm}}r}
+\config{\prog}{\elet{x}{v}{e}} &\redx& \config{\prog}{e[v/x]}\\
+\config{\prog}{f(v_1,...,v_n)} &\redx&
+\config{\prog}{e[v_1/x_1,\ldots,v_n/x_n]} & 
+ \codebase(f) = x_1,\ldots,x_n,\ e\\
+\config{\prog}{\{\ldots; \flab = v; \ldots\}.\flab} &\redx&
+ \config{\prog}{v}\\
+ \config{\prog}{w_1\ttt{||}w_2} &\redx& \config{\prog}{w_1w_2}\\
+ \config{\prog}{v;e} &\redx& \config{\prog}{e}\\
+\config{\prog}{\instr} &\redx& \config{\prog;\instr}{()}\\
+\config{\prog}{E[e]} &\redx& \config{\prog'}{E[e']} & \text{if}\ \config{\prog}{e} \redx \config{\prog'}{e'} 
+\end{array}
+$$

### 2023-06-03 popl24/bruteforce.tex
checkpoint

\section{Brute Force Verification}

\begin{eqnarray*}
\solve{\stores}{\etrue} &=& \stores\\
\solve{\stores}{\efalse} &=& \varnothing\\
\solve{\stores}{\flip{\cid}{w}} &=& \{ \store \in \stores \mid \store(\flip{\cid}{w}) \} \\
\solve{\stores}{\secret{\cid}{w}} &=& \{ \store \in \stores \mid \store(\secret{\cid}{w}) \} \\
\solve{\stores}{\view{\cid}{w}} &=& \{ \store \in \stores \mid \store(\view{\cid}{w}) \} \\
\solve{\stores}{\oracle{w}} &=& \{ \store \in \stores \mid \store(\oracle{w}) \} \\
\solve{\stores}{(\enot\ \be)} &=& \stores - (\solve{\stores}{\be})\\
\solve{\stores}{(\be_1\ \eand\ \be_2)} &=& (\solve{\stores}{\be_1}) \cap (\solve{\stores}{\be_2}) \\
\solve{\stores}{(\be_1\ \eor\ \be_2)} &=& (\solve{\stores}{\be_1}) \cup (\solve{\stores}{\be_2}) \\
\solve{\stores}{(\be_1\ \exor\ \be_2)} &=&
 ((\solve{\stores}{\be_1}) \cap (\stores - \solve{\stores}{\be_2})) \cup
 ((\stores - \solve{\stores}{\be_1}) \cap (\solve{\stores}{\be_2})) \\
\solve{\stores}{\select{\be_1}{\be_2}{\be_3}} &=&
 ((\solve{\stores}{\be_1}) \cap (\solve{\stores}{\be_2})) \cup
 ((\stores - \solve{\stores}{\be_1}) \cap (\solve{\stores}{\be_3})) \\
\solve{\stores}{\OT{\be_1}{\be_2}{\be_3}} &=&
 ((\solve{\stores}{\be_1}) \cap (\solve{\stores}{\be_2})) \cup
 ((\stores - \solve{\stores}{\be_1}) \cap (\solve{\stores}{\be_3}))
\end{eqnarray*}

$$
g\ \ \stores\ (\eassign{\view{\cid}{w}}{\be}) =\begin{array}[t]{l}
\elet{\stores'}{\solve{(\mems(\vars\ \be))}{\be}}{}\\
  \{\extend{\store}{\view{\cid}{w}}{\etrue} \mid \store \in \stores' \} \cup
  \{\extend{\store}{\view{\cid}{w}}{\efalse} \mid \store \in \stores - \stores' \}\end{array}
$$

$$
\mathit{foldl}\ g\ \mems(S \cup F)\ \prog
$$

$$
f(\eassign{\view{\cid}{w}}{\be}) = (\view{\cid}{w}, (\solve{(\mems(\vars\ \be))}{\be}))
$$

$$
\mathit{map}\ f\ \prog
$$

### 2023-06-03 popl24/minicat.tex
checkpoint

@@
-\section{The $\minicat$ Language}
+\section{The $\minifed$ Language}
 
-The language $\minicat$ comprises a standard syntax of
-boolean algebra and arithmetic expressions over integers. In addition, we allow reference
-to \emph{variables} in expressions. \emph{Programs} are possibly empty
-lists of assignments of expressions to variables.
 $$
-\begin{array}{rcl@{\qquad\qquad}r}
-x,y,z &\in& \mathcal{V} & \textit{variables}\\
-n &\in& \mathbb{Z} & \textit{integers}\\
-b &::=& \true \mid \false & \textit{booleans}\\
-v &::=& n \mid b & \textit{values} \\[2mm]
-\mathit{unop} &::=& \mathrm{not} \mid - & \textit{unary operations} \\
-\mathit{binop} &::=& \mathrm{and} \mid \mathrm{or} \mid + \mid * \mid \mod \mid\ =\ \mid & \textit{binary operations} \\[2mm]
-e &::=& x \mid v \mid\, \mathit{unop}\, e \mid e\, \mathit{binop}\, e \mid  \eite{e}{e}{e}  & \textit{expressions}\\[2mm]
-\prog &::=& \varnothing\quad \mid\quad \eassign{x}{e};\ \prog & \textit{programs}
+\begin{array}{rcl@{\hspace{8mm}}r}
+b &\in& \{ \etrue, \efalse \} \\
+w &\in& \mathrm{String} \\ 
+\cid &\in& \mathrm{Clients} \qquad  (\mathrm{Clients} \subset \mathbb{N} \text{\ in\ this\ presentation})\\[2mm]
+\bop &\in& \{ \eand, \eor, \exor \} \\[2mm]
+\be &::=& b \mid \flip{\cid}{w} \mid \secret{\cid}{w} \mid \view{\cid}{w} \mid \oracle{w} \mid & \textit{boolean expressions}\\
+& &  \enot\ \be \mid \be\ \bop\ \be \mid \select{\be}{\be}{\be} \mid \OT{\be}{\be}{\be} \\[2mm]
+\instr &::=& \eassign{\view{\cid}{w}}{\be} & \textit{view assignments} \\[2mm]
+\prog &::=& \varnothing \mid \instr; \prog & \textit{protocols}
 \end{array}
 $$
-So for example, the following is a program- note that it assume the existence of an
-\emph{input} variable $x$, which is undefined by the program itself:
-$$
-\eassign{y}{x + 1}; \eassign{z}{2 * y}
-$$
 
-For variable-free expressions in $\minicat$,
-we can extend our interpretation of
-boolean algebra $\cod{\cdot}$ to integer arithmetic in the obvious
-manner, so for example $\cod{2 * 4} = 8$. However, to resolve
-expressions that include variables, we need to define the semantics in
-terms of references to \emph{memories}. We take the view that memories
-map variables to values.
 \begin{definition}[Memories]
@@
 set of variables on which the memory is defined. We write $\store\{ x \mapsto v\}$
-to denote the memory $\store'$ such that $\store'(x) = v$ and otherwise
-$\store'(y) = \store(y)$ for all $y \in (\dom(\store) - \{ x \}) $.
+for  $x\not\in\dom(\store)$ to denote the memory $\store'$ such that $\store'(x) = v$ and otherwise
+$\store'(y) = \store(y)$ for all $y \in \dom(\store)$.
 \end{definition}
-Now we can modify our interpretation of expressions to take into account a given memory.
-We write $\cod{\store, e}$ to mean the interpretation of $e$ where each variable $x$ in $e$
-is replaced by $\store(x)$. In full detail, we define this as follows:
+
 \begin{eqnarray*}
-\cod{\store, x} &=& \store(x)\\
-\cod{\store, \mathit{unop}\ e} &=& \cod{\mathit{unop}\ \cod{\store,e}}\\
-\cod{\store, e_1\ \mathit{binop}\ e_2} &=& \cod{\cod{\store,e_1}\ \mathit{binop}\ \cod{\store,e_2}}\\
-\cod{\store, \eite{e_1}{e_2}{e_3}} &=&
+\lcod{\store, \etrue}{\cid} &=& \etrue\\
+\lcod{\store, \efalse}{\cid} &=& \efalse\\
+\lcod{\store, \flip{\cid}{w}}{\cid} &=& \store(\flip{\cid}{w})\\
+\lcod{\store, \secret{\cid}{w}}{\cid} &=& \store(\secret{\cid}{w})\\
+\lcod{\store, \view{\cid}{w}}{\cid} &=& \store(\view{\cid}{w})\\
+\lcod{\store, \oracle{w}}{\cid} &=& \store(\oracle{w})\\
+\lcod{\store, \enot\ \be}{\cid} &=& \cod{\enot\ \lcod{\store,\be}{\cid}}\\
+\lcod{\store, \be_1\ \mathit{binop}\ \be_2}{\cid} &=&
+    \cod{\lcod{\store,\be_1}{\cid}\ \mathit{binop}\ \lcod{\store,\be_2}{\cid}}\\
+\lcod{\store, \select{\be_1}{\be_2}{\be_3}}{\cid} &=&
              \begin{cases}
-                \cod{\store,e_2} & \text{if\ } \cod{\store,e_1}\\
-                \cod{\store,e_3} & \text{otherwise}
-             \end{cases}
+                \lcod{\store,\be_2}{\cid} & \text{if\ } \lcod{\store,\be_1}{\cid}\\
+                \lcod{\store,\be_3}{\cid} & \text{if\ } \neg\lcod{\store,\be_1}{\cid}
+             \end{cases}%\\
+%\lcod{\store, \OT{\be_1}{\be_2}{\be_3}}{\cid} &=&
+%             \begin{cases}
+%                \lcod{\store,\be_2}{\cid'} & \text{if\ } \lcod{\store,\be_1}{\cid}\\
+%                \lcod{\store,\be_3}{\cid'} & \text{if\ } \neg\lcod{\store,\be_1}{\cid}
+%             \end{cases}
 \end{eqnarray*}
 
-A critical point is that the semantics of $\minicat$ is defined not
-just with respect to programs, but with respect to a pair of a memory
-and a program $(\store, \prog)$, which we call
-a \emph{configuration}. Together with the interpretation of
-expressions that reference memories, we can obtain
-an \emph{operational} semantics for configurations. This semantics
-specifies how memories evolve during computation as a result of
-assignments to variables in programs. The fundamental relation is
-the \emph{reduction} aka \emph{evaluation} relation $\redx$,
-which is a relation on configurations, that we define as follows:
-$$
-(\store, \eassign{x}{e};\prog) \redx (\extend{\store}{x}{\cod{\store,e}}, \prog)
-$$
+\begin{mathpar}
+  (\store, \eassign{\view{\cid}{w}}{\be};\prog) \redx (\extend{\store}{\view{\cid}{w}}{\lcod{\store,\be}{\cid'}}, \prog)
 
-\subsection{Synchronous Distributed Protocols} With some minor syntactic
-modifications to $\minicat$, we can model and reason about security in
-protocols for distributed communications. We will call this modified
-system $\fedcat$, alluding to the existence of a group of participants
-we call a \emph{federation}. We restrict ourselves to boolean values
-in $\fedcat$, which we will denote $1$ and $0$ for convenience.  We
-assume the full set of the usual boolean operations including xor.
-
-We identify participants by integers $n$, and assume that the
-federation can be partitioned into two sets $H$ and $C$ where $H$ are
-all \emph{honest} and $C$ are \emph{corrupt}. Then, we modify
-variables $x$ to be of the form $\sx{n}{s}$ denoting a \emph{secret}
-belonging to party $n$ and distinguished by a string $s$, or of
-the form $\vx{n}{s}$ denoting a \emph{view} belonging to party $n$ and
-distinguished by $s$. A view is a memory location that records a value
-received by a party. So for example, in the following program, party 1
-sends the negation of their secret $\ttt{foo}$ to party 2, with the
-value recorded in party 2's $\ttt{bar}$ view:
-$$
-\eassign{\vx{2}{bar}}{\Dnot (\sx{1}{foo})}
-$$ In $\fedcat$ we require that all assignments are to views, and in
-any assignment $\eassign{\vx{n}{s}}{e}$ we require all variables that
-occur in $e$ to belong to the same party. We will also assume that
-secrets are defined in initial memories, that views are not, and that
-views are never redefined (overwritten). This models communication
-between parties through views, where each party can compute locally on
-its own values, and ensures that the final memory after computation
-contains a record of all secrets and views. We will use
-\emph{protocol} as a synonym for program in $\fedcat$ given our intent
-to model synchronous distributed computations.
-
- Although $\fedcat$ is a simple
-language we can explore some surprisingly complex distributed
-protocols with the introduction of randomized elements. In $\fedcat$
-we accomplish this by introducing a new form of variables called
-\emph{flips}, written $\fx{n}{s}$, which are owned by a party $n$ and
-distinguished by string $s$. We will assume that flips can be either
-$1$ or $0$ with equal probability.  So for example, in this
-program, party 1 sends their flip $\ttt{foo}$ to party 2,
-with the value recorded in party 2's $\ttt{bar}$ view:
-$$
-\eassign{\vx{2}{bar}}{\fx{1}{foo}}
-$$
-We don't change anything about the semantics of $\fedcat$
-to model this randomization, rather we will assume that all
-flip values are randomly assigned in initial memories. This is
-the so-called ``random tape'' approach.
+  \inferrule
+  {b = \text{if\ } \lcod{\store,\be_1}{\cid}  \text{\ then\ } \lcod{\store,\be_2}{\cid'} \text{\ else\ } \lcod{\store,\be_3}{\cid'}}
+      {(\store, \eassign{\view{\cid}{w}}{\OT{\be_1}{\be_2}{\be_3}};\prog) \redx (\extend{\store}{\view{\cid}{w}}{b}, \prog)}
+\end{mathpar}

### 2023-06-03 popl24/pmf.tex
checkpoint

@@
 \begin{definition}
-  Given progam $\prog$ with $\iov(\prog) = (S,V)$ and $\flips(\prog) = F$, define $M$ as:
+  Given progam $\prog$ with $\iov(\prog) = (S,V)$ and $\flips(\prog) = F$, define $\stores$ as:
   $$
-  M \defeq \{ \store \mid (\dom(\store) = S \cup V \cup F) \wedge (\store_{S \cup F},\prog) \redxs (\store,\varnothing) \}
+  \stores \defeq \{ \store \mid (\dom(\store) = S \cup V \cup F) \wedge (\store_{S \cup F},\prog) \redxs (\store,\varnothing) \}
   $$
@@
   \begin{cases}
-    1 / |M| & \text{if}\ \store \in M\\
+    1 / |\stores| & \text{if}\ \store \in \stores\\
     0 & \text{otherwise}

### 2023-06-03 popl24/minicat.tex
checkpoint

@@
 \end{mathpar}
+
+\paragraph{Variable and Memory Conventions} Given a program $\prog$, we
+write $\iov(\prog)$ to denote the pair $(S,V)$
+where $S$ is the set of secret variables in $\prog$ and
+$V$ is the set of views in $\prog$, and we write $\flips(\prog)$ to
+denote the set $F$ of flip variables in $\prog$. For any set of variables $X$
+and parties $P$, we write $X_P$ to denote the subset
+of $X$ owned by any party in $P$, and we write $X_H$ and
+$X_C$ to denote the subsets belonging to honest and
+corrupt parties, respectively.
+
+Given a set of variables $X$, we write $\store_X$ to mean the memory
+$\store$ restricted to the domain $X$, and
+and we define $\mems(X)$ as:
+$$
+\mems(X) \defeq \{ \store \mid \dom(\store) = X \}
+$$
+We will also write $\store_1 \cup \store_2$ to denote
+the memory resulting by combining disjoint
+memories $\store_1$ and $\store_2$, i.e.,
+under the assumption that $\dom(\store_1) \cap \dom(\store_2) = \varnothing$.

### 2023-06-04 popl24/minicat.tex
checkpoint

@@
 
+The $\minifed$ language provides a simple model of synchronous
+protocols between a federation of \emph{clients} exchanging values in
+the binary field. We will identify clients by natural numbers, and
+federations- finite sets of clients- are always given statically.
+As we will see, our threat model assumes a partition of the federation
+into \emph{honest} and \emph{corrupt} subsets.
+
+We model probabilistic programming via the \emph{random tape}
+approach. That is, we will assume that programs can make reference to
+values chosen from a uniform random distribution- coin ``flips''- via
+values arbitrarily assigned in the initial program memory- the random
+tape. Programs execute deterministically given the random tape.
+
+\subsection{Syntax} The syntax of $\minifed$ includes a standard
+boolean algebra with $\eand$, $\eor$, $\exor$, and $\enot$ as
+primitives. In addition programs have different kinds of variables,
+including \emph{secrets} $\secret{\cid}{w}$, \emph{flips}
+$\flip{\cid}{w}$, and \emph{views} $\view{\cid}{w}$.  Each of these
+variables are indexed by an ``owner'' client $\cid$ and distinguishing
+string $w$. So for example, $\secret{\ttt{1}}{\ttt{foo}}$ is client
+$\ttt{1}$'s secret called $\ttt{foo}$. All clients can make reference
+to the \emph{oracle} through variables $\oracle{w}$- this is a form
+of shared randomness that is standard in the MPC setting \cite{XXX}.
+Each client can only compute on their own variables in
+\emph{expressions} $\be$, and share values with other clients by
+assignment to their views as in $\eassign{\view{\cid}{w}}{\be}$.  A
+\emph{protocol} $\prog$ is a possibly empty sequence of view
+assignments.
+
+Expressions also include $\select{\be_1}{\be_2}{\be_3}$ which is
+like a conditional expression in the binary field with:
+$$
+\select{\be_1}{\be_2}{\be_3} \equiv (\be_1\ \eand\ \be_2)\ \eor\ (\enot\ \be_1\ \eand\ \be_3)
+$$
+We also include \emph{oblivious transfer} $\OT{\be_1}{\be_2}{\be_3}$ as a primitive,
+with semantics similar to $\ttt{select}$ but with important nuances
+related to communication between clients as we discuss more below.
+As we will demonstrate, it is not necessary to include $\ttt{OT}$
+as a primitive since we can implement it in a provably secure
+fashion. However its inclusion simplifies our presentation and is a useful
+convenience.
+
+\subsection{Semantics}
+
+\begin{fpfig}[t]{$\minifed$ source code syntax.}{fig-minifed}
 $$
@@
 w &\in& \mathrm{String} \\ 
-\cid &\in& \mathrm{Clients} \qquad  (\mathrm{Clients} \subset \mathbb{N} \text{\ in\ this\ presentation})\\[2mm]
+\cid &\in& \mathrm{Clients} \subset  \mathbb{N} \\[2mm]%\qquad  (\mathrm{Clients} \subset \mathbb{N} \text{\ in\ this\ presentation})\\[2mm]
 \bop &\in& \{ \eand, \eor, \exor \} \\[2mm]
@@
 $$ 
+\end{fpfig}
 

### 2023-06-05 popl24/minicat.tex
checkpoint

@@
 values chosen from a uniform random distribution- coin ``flips''- via
-values arbitrarily assigned in the initial program memory- the random
-tape. Programs execute deterministically given the random tape.
+values arbitrarily assigned in the initial program memory.  Programs
+execute deterministically given the random tape. The random tape
+formulation supports our automated analysis as we discuss in
+section
 

### 2023-06-06 popl24/minicat.tex
checkpoint

@@
 As we will see, our threat model assumes a partition of the federation
-into \emph{honest} and \emph{corrupt} subsets.
+into \emph{honest} $H$ and \emph{corrupt} $C$ subsets.
 
@@
 
-Expressions also include $\select{\be_1}{\be_2}{\be_3}$ which is
-like a conditional expression in the binary field with:
+We let $X$ range over sets of variables and $S,V,F$ to range over sets
+of secrets, views, and flips respectively. Given a program $\prog$, we
+write $\iov(\prog)$ to denote the pair $(S,V)$ where $S$ is the set of
+secret variables in $\prog$ and $V$ is the set of views in $\prog$,
+and we write $\flips(\prog)$ to denote the set $F$ of flip variables
+in $\prog$. For any set of variables $X$ and parties $P$, we write
+$X_P$ to denote the subset of $X$ owned by any party in $P$, and we
+write $X_H$ and $X_C$ to denote the subsets belonging to honest and
+corrupt parties, respectively.
+
+Expressions forms include the convenience
+$\select{\be_1}{\be_2}{\be_3}$ which is like a conditional expression
+in the binary field with:
 $$
@@
 
-\begin{definition}[Memories]
-A \emph{memory}, aka \emph{store} $\store$ is a finite (partial) mapping from variables to values.
-The \emph{domain} of a memory is written $\dom(\store)$ and is the finite
-set of variables on which the memory is defined. We write $\store\{ x \mapsto v\}$
-for  $x\not\in\dom(\store)$ to denote the memory $\store'$ such that $\store'(x) = v$ and otherwise
-$\store'(y) = \store(y)$ for all $y \in \dom(\store)$.
-\end{definition}
+\emph{Memories} are fundamental to the semantics of $\fedcat$ and
+provide the random tape and secret inputs to protocols, and record
+view assignments. Memories $\store$ are finite (partial) mapping from
+variables to values. The \emph{domain} of a memory is written
+$\dom(\store)$ and is the finite set of variables on which the memory
+is defined. We write $\store\{ x \mapsto v\}$ for
+$x\not\in\dom(\store)$ to denote the memory $\store'$ such that
+$\store'(x) = v$ and otherwise $\store'(y) = \store(y)$ for all $y \in
+\dom(\store)$.
+
+Given a set of variables $X$, we write $\store_X$ to denoted the
+memory $\store$ restricted to the domain $X$, and and we define
+$\mems(X)$ as the set of all memories with domain $X$:
+$$
+\mems(X) \defeq \{ \store \mid \dom(\store) = X \}
+$$
+Thus, given a protocol $\prog$, the set of all random tapes for
+$\prog$ is $\mems(\flips(\prog))$. We let $\stores$ range
+over sets of memories.
+
+Given a variable-free boolean expression $\be$, we write
+$\cod{\be}$ to denote the standard interpretation of $\be$.
+With the introduction of variables to expressions, we
+have two concerns. First, we need to interpret expressions
+with respect to a specific memory, and second, we need
+to ensure that variables are used ``legally''. That is,
+since expressions define local computation, all variables
+used in an expression must belong to the same client.
 
@@
 \end{mathpar}
-
-\paragraph{Variable and Memory Conventions} Given a program $\prog$, we
-write $\iov(\prog)$ to denote the pair $(S,V)$
-where $S$ is the set of secret variables in $\prog$ and
-$V$ is the set of views in $\prog$, and we write $\flips(\prog)$ to
-denote the set $F$ of flip variables in $\prog$. For any set of variables $X$
-and parties $P$, we write $X_P$ to denote the subset
-of $X$ owned by any party in $P$, and we write $X_H$ and
-$X_C$ to denote the subsets belonging to honest and
-corrupt parties, respectively.
-
-Given a set of variables $X$, we write $\store_X$ to mean the memory
-$\store$ restricted to the domain $X$, and
-and we define $\mems(X)$ as:
-$$
-\mems(X) \defeq \{ \store \mid \dom(\store) = X \}
-$$
-We will also write $\store_1 \cup \store_2$ to denote
-the memory resulting by combining disjoint
-memories $\store_1$ and $\store_2$, i.e.,
-under the assumption that $\dom(\store_1) \cap \dom(\store_2) = \varnothing$.

### 2023-06-06 popl24/metalang.tex
checkpoint

@@
 \section{A Protocol Metalanguage}
-
-$$
-\begin{array}{rcl@{\hspace{8mm}}r}
-b &\in& \{ \etrue, \efalse \} \\
-w &\in& \mathrm{String} \\ 
-\cid &\in& \mathrm{Clients} \qquad  (\mathrm{Clients} \subset \mathbb{N} \text{\ in\ this\ presentation})\\[2mm]
-\bop &\in& \{ \eand, \eor, \exor \} \\[2mm]
-\be &::=& b \mid \flip{\cid}{w} \mid \secret{\cid}{w} \mid \view{\cid}{w} \mid \oracle{w} \mid & \textit{boolean expressions}\\
-& &  \enot\ \be \mid \be\ \bop\ \be \mid \select{\be}{\be}{\be} \\[2mm]
-\instr &::=& \eassign{\view{\cid}{w}}{\be} \mid \eassign{\view{\cid}{w}}{\OT{\be}{\be}{\be}} & \textit{view assignments} \\[2mm]
-\prog &::=& \instr \mid \instr; \prog & \textit{protocols}
-\end{array}
-$$
-
-\begin{eqnarray*}
-\lcod{\store, \etrue}{\cid} &=& \etrue\\
-\lcod{\store, \efalse}{\cid} &=& \efalse\\
-\lcod{\store, \flip{\cid}{w}}{\cid} &=& \store(\flip{\cid}{w})\\
-\lcod{\store, \secret{\cid}{w}}{\cid} &=& \store(\secret{\cid}{w})\\
-\lcod{\store, \view{\cid}{w}}{\cid} &=& \store(\view{\cid}{w})\\
-\lcod{\store, \oracle{w}}{\cid} &=& \store(\oracle{w})\\
-\lcod{\store, \enot\ \be}{\cid} &=& \cod{\enot\ \lcod{\store,\be}{\cid}}\\
-\lcod{\store, \be_1\ \mathit{binop}\ \be_2}{\cid} &=&
-    \cod{\lcod{\store,\be_1}{\cid}\ \mathit{binop}\ \lcod{\store,\be_2}{\cid}}\\
-\lcod{\store, \select{\be_1}{\be_2}{\be_3}}{\cid} &=&
-             \begin{cases}
-                \lcod{\store,\be_2}{\cid} & \text{if\ } \lcod{\store,\be_1}{\cid}\\
-                \lcod{\store,\be_3}{\cid} & \text{otherwise}
-             \end{cases}
-\end{eqnarray*}
-
-\begin{mathpar}
-  (\store, \eassign{\view{\cid}{w}}{\be};\prog) \redx (\extend{\store}{\view{\cid}{w}}{\lcod{\store,\be}{\cid}}, \prog)
-
-  \inferrule
-  {b = \text{if\ } \lcod{\store,\be_1}{\cid}  \text{\ then\ } \lcod{\store,\be_2}{\cid'} \text{\ else\ } \lcod{\store,\be_3}{\cid'}}
-      {(\store, \eassign{\view{\cid}{w}}{\OT{\be_1}{\be_2}{\be_3}};\prog) \redx (\extend{\store}{\view{\cid}{w}}{b}, \prog)}
-\end{mathpar}
 
@@
 \mid e.\flab \mid e\ttt{||}e \mid (e) \\[2mm]
-v &::=& w \mid b \mid \cid \mid \{ \flab = v;\ldots;\flab = v \} 
+v &::=& w \mid \cid \mid \be \mid \{ \flab = v;\ldots;\flab = v \} 
 \mid (\,) & \textit{values}\\[2mm]
@@
 $$
+
+\begin{verbatimtab}
+andgg(g, wla, wlb, wlc, wva, wvb) {
+andgate(g, wla, wlb, wlc);
+evalgate(g, wva, wvb)
+}
+\end{verbatimtab}
+
+\begin{lemma}[Input Encoding]
+Assume given the following evaluation relation:
+$$
+\config{\prog_1}{E[\ttt{encode}(w,v_1,v_2)]} \redxs \config{\prog_1;\prog_2}{E[v]}
+$$
+If there exists gate identifiers $g_1, g_2$ such that:
+\begin{enumerate}
+  \item $g_1 \ne g_2 \wedge g_1,g_2 \not\in \gates(\prog_1)$
+  \item $\owl(\prog_1,v_1,g_1) \wedge \owl(\prog_1,v_2,g_2)$
+  \item $\gNIMO(\prog_1,v.\ttt{wv1},g_1) \wedge \gNIMO(\prog_1,v.\ttt{wv2},g_2)$
+\end{enumerate}
+then:
+\begin{enumerate}
+\item $g_1,g_2 \in  \gates(\prog_2)$
+\item $\gNIMO(\prog_1;\prog_2,v.\ttt{wv1},g_1) \wedge \gNIMO(\prog_1;\prog_2,v.\ttt{wv2},g_2)$
+\item $$(\margd{\progd(\prog_1;\prog_2)}{\iov(\prog_1)}) * (\margd{\progd(\prog_1;\prog_2)}{\iov(\prog_2)})$$
+\end{enumerate}
+\end{lemma}
+
+\begin{lemma}[Composition ($\eand$)]
+Assume given the following evaluation relation:
+$$
+\config{\prog_1}{E[\ttt{andgg}(g,v_1,\ldots,v_5)]} \redxs \config{\prog_1;\prog_2}{E[v_6]}
+$$
+If there exist gate identifiers $g_1, g_2$ such that:
+\begin{enumerate}
+  \item $g_1 \ne g_2 \ne g \wedge g_1,g_2 \in \gates(\prog_1) \wedge g \not\in \gates(\prog_1)$
+  \item $\owl(\prog_1,v_1,g_1) \wedge \owl(\prog_1,v_2,g_2) \wedge \owl(\prog_1,v_3,g)$
+  \item $\gNIMO(\prog_1,v_4,g_1) \wedge \gNIMO(\prog_1,v_5,g_2)$
+\end{enumerate}
+then:
+\begin{enumerate}
+  \item $g \in \gates(\prog_2)$
+  \item $\gNIMO(\prog_1;\prog_2,v_6,g)$
+  \item $(\margd{\progd(\prog_1;\prog_2)}{\iov(\prog_1)}) * (\margd{\progd(\prog_1;\prog_2)}{\iov(\prog_2)})$
+\end{enumerate}
+\end{lemma}
+
+\begin{lemma}[Decoding]
+Assume given the following evaluation relation:
+$$
+\config{\prog_1}{E[\ttt{decoder}(v_1,v_2)]} \redxs \config{\prog_1;\prog_2}{E[\be]}
+$$
+If there exists gate identifier $g \in \gates(\prog_1)$ such that:
+\begin{enumerate}
+  \item $\owl(\prog_1,v_1,g)$
+  \item $\gNIMO(\prog_1,v_2,g)$
+\end{enumerate}
+Then $\NIMO(\prog_1;\prog_2;\eassign{x}{\be})$.
+\end{lemma}

### 2023-06-06 popl24/metalang.tex
checkpoint

@@
 
+\subsection{Compositional Properties: Garbled $\NIMO$}
+
 \begin{verbatimtab}
@@
 \end{verbatimtab}
+
+\begin{definition}[Garbled $\NIMO$]
+  We write $\gNIMO(\prog,\{ \ttt{k} = \be_{\ttt{k}}; \ttt{p} = \be_{\ttt{p}}\}, g)$ iff for all $H$, $C$,
+  both $w \in \{\ttt{p},\ttt{k}\}$, and all 
+  memories $\store \in \mems(\iov(\prog) \cup \{ \itv \})$ where
+  $
+  \store(\itv) = \store(\flip{2}{g.w})$ or $\store(\itv) = \neg\store(\flip{2}{g.w})
+  $
+  it is the case that:
+  $$
+  \margd{(\condd{\progd(\prog,\be_w)}{\store_{S_C \cup \{\itv \}}})}{S_H} =
+  \margd{(\condd{\progd(\prog,\be_w)}{\store_{S_C \cup V_C \cup \{\itv \}}})}{S_H}
+  $$
+\end{definition}
+
+\begin{definition}
+  We write $\owl(\prog, \{ \ttt{k} = \be_{\ttt{k}}; \ttt{p} = \be_{\ttt{p}}\}, g)$
+  iff for all $b$ and both $w \in \{\ttt{p},\ttt{k}\}$:
+  $$
+  (\margd{(\condd{\progd(\prog,\be_{w})}{\{ \flip{2}{g.w} \mapsto b \}})}{\itv})(\{ \itv \mapsto b\}) = 1
+  $$
+\end{definition}
 
@@
   \item $\owl(\prog_1,v_1,g_1) \wedge \owl(\prog_1,v_2,g_2)$
-  \item $\gNIMO(\prog_1,v.\ttt{wv1},g_1) \wedge \gNIMO(\prog_1,v.\ttt{wv2},g_2)$
 \end{enumerate}
@@
 \item $\gNIMO(\prog_1;\prog_2,v.\ttt{wv1},g_1) \wedge \gNIMO(\prog_1;\prog_2,v.\ttt{wv2},g_2)$
-\item $$(\margd{\progd(\prog_1;\prog_2)}{\iov(\prog_1)}) * (\margd{\progd(\prog_1;\prog_2)}{\iov(\prog_2)})$$
+\item $(\margd{\progd(\prog_1;\prog_2)}{\iov(\prog_1)}) * (\margd{\progd(\prog_1;\prog_2)}{\iov(\prog_2)})$
 \end{enumerate}
@@
   \item $g_1 \ne g_2 \ne g \wedge g_1,g_2 \in \gates(\prog_1) \wedge g \not\in \gates(\prog_1)$
-  \item $\owl(\prog_1,v_1,g_1) \wedge \owl(\prog_1,v_2,g_2) \wedge \owl(\prog_1,v_3,g)$
+  \item $\owl(\prog_1,v_1,g_1) \wedge \owl(\prog_1,v_2,g_2)$
   \item $\gNIMO(\prog_1,v_4,g_1) \wedge \gNIMO(\prog_1,v_5,g_2)$
@@
 \begin{enumerate}
-  \item $g \in \gates(\prog_2)$
+  \item $g \in \gates(\prog_2) \wedge \owl(\prog_1;\prog_2,v_3,g)$
   \item $\gNIMO(\prog_1;\prog_2,v_6,g)$

### 2023-06-06 popl24/minicat.tex
checkpoint

@@
 
-\subsection{Syntax} The syntax of $\minifed$ includes a standard
+\subsection{Syntax} The syntax of $\minifed$, defined in
+Figure \ref{fig-minifed-syntax}, includes a standard
 boolean algebra with $\eand$, $\eor$, $\exor$, and $\enot$ as
@@
 of secrets, views, and flips respectively. Given a program $\prog$, we
-write $\iov(\prog)$ to denote the pair $(S,V)$ where $S$ is the set of
+write $\iov(\prog)$ to denote $S \cup V$ where $S$ is the set of
 secret variables in $\prog$ and $V$ is the set of views in $\prog$,
@@
 Expressions forms include the convenience
-$\select{\be_1}{\be_2}{\be_3}$ which is like a conditional expression
-in the binary field with:
+$\select{\be_1}{\be_2}{\be_3}$ which is essentiall a conditional
+expression with:
 $$
@@
 
-\begin{fpfig}[t]{$\minifed$ source code syntax.}{fig-minifed}
+\begin{fpfig}[t]{$\minifed$ source code syntax.}{fig-minifed-syntax}
 $$
@@
 
-Given a variable-free boolean expression $\be$, we write
-$\cod{\be}$ to denote the standard interpretation of $\be$.
-With the introduction of variables to expressions, we
-have two concerns. First, we need to interpret expressions
-with respect to a specific memory, and second, we need
-to ensure that variables are used ``legally''. That is,
-since expressions define local computation, all variables
-used in an expression must belong to the same client.
+Given a variable-free boolean expression $\be$, we write $\cod{\be}$
+to denote the standard interpretation of $\be$.  With the introduction
+of variables to expressions, we have two concerns. First, we need to
+interpret variables with respect to a specific memory, and second,
+we need to ensure that variables are used ``legally''. That is, since
+expressions define local computation, all variables used in an
+expression must belong to the same client.  Thus, we denote
+interpretation of expressions $\be$ possibly containing variables as
+$\lcod{\store,\be}{\cid}$, where $\store$ associates variables with
+values and all variables must be owned by client $\cid$. This is
+defined in Figure \ref{fig-minifed-interp}.
 
+\begin{fpfig}[t]{$\minifed$ expression interpretation.}{fig-minifed-interp}
 \begin{eqnarray*}
@@
 \end{eqnarray*}
+\end{fpfig}
 
+Evaluation of configurations is then defined via a small-step reduction relation $\redx$.
+This is defined in the obvious manner for view assignments other than through
+$\ttt{OT}$- however note that reduction requires that views are never reassigned. 
 \begin{mathpar}
   (\store, \eassign{\view{\cid}{w}}{\be};\prog) \redx (\extend{\store}{\view{\cid}{w}}{\lcod{\store,\be}{\cid'}}, \prog)
-
+\end{mathpar}
+For view assignments through $\ttt{OT}$, we define a different reduction rule that
+captures the appropriate semantics- that is, the selection bit is \emph{not} communicated
+to the sender (through a view assignment), and only the selected bit is sent to the receiver:
+\begin{mathpar}
   \inferrule
@@
 \end{mathpar}
+We define $\redxs$ as the reflexive, transitive closure of $\redx$.
+
+Given $\prog$ with $\iov(\prog) = S \cup V$ and $\flips(\prog) = F$,
+any execution of $\prog$ is assumed to be an evaluation of a
+configuration $\config{\store}{\prog}$ where $\store \in \mems(S \cup
+F)$- that is, with secrets and the random tape as inputs- and where
+$\config{\store}{\prog} \redxs \config{\store'}{\varnothing}$ it is
+the case that $\store' \in \mems(S \cup F \cup V)$.
+
+\subsection{Examples}
+
+\begin{verbatimtab}
+v[2,0] := f[1,0];
+v[0,0] := f[1,0] xor s[1,0];
+v[2,1] := v[0,0];
+v[2,2] := v[2,1] xor v[2,0]
+\end{verbatimtab}
+
+\begin{verbatimtab}
+v[2, rd] := select[flip[2, d], H[r1], H[r0]];
+v[1, e] := s[2, c] xor flip[2, d];
+v[2, f0] := s[1, m0] xor select[v[1, e], H[r1], H[r0]];
+v[2, f1] := s[1, m1] xor select[not v[1, e], H[r1], H[r0]];
+v[2, mc] := (select[s[2, c], v[2, f1], v[2, f0]] xor v[2, rd])
+\end{verbatimtab}

### 2023-06-07 popl24/hyperprop.tex
checkpoint

@@
-\section{Hyperproperties}
+\section{Hyperproperties of MPC Passive Security}
 
@@
 \end{lemma}
-
-\example{Example: Additive Secret Sharing} A nice ``Hello World''
-example of an MPC protocol is additive secret sharing, where $k$
-parties split up their secrets into \emph{shares}, with the property
-that all $k$ shares are required to reconstruct the secret and any
-fewer reveals nothing. In $\fedcat$, we can use $\ttt{xor}$, again, to
-generate shares. Here, party 1 generates 3 shares of the secret
-$\sx{1}{1}$- one is given to party 2 through view $\vx{2}{s1}$, one is
-given to party 3 through view $\vx{3}{s1}$, and party 1 keeps
-$\fx{1}{s1}$ for itself:
-\begin{verbatimtab}
-v[2,s1] := flip[1,1] xor flip[1,s1] xor s[1,1];
-v[3,s1] := flip[1,1];
-\end{verbatimtab}
-Note that $\vx{2}{s1}$ and $\vx{3}{s1}$ are both in uniform
-random distributions.
-
-Another property of $\ttt{xor}$ is that it is additively homomorphic,
-while functioning as addition in the boolean algebra (i.e., the binary
-field). That is, given:
-\begin{mathpar}
-  x_1\ \ttt{xor}\ y_1 = z_1
-
-  x_2\ \ttt{xor}\ y_2 = z_2
-\end{mathpar}
-then we have:
-$$
-(x_1\ \ttt{xor}\ x_2)\ \ttt{xor}\ (y_1\ \ttt{xor}\ y_2) = z_1\ \ttt{xor}\ z_2
-$$
-Thus, we can obtain secret 3-party summation (xor) by continuing
-the protocol begun above- parties 2 and 3 can share their secrets
-in the same fashion:
-\begin{verbatimtab}
-v[1,s2] := flip[2,1] xor flip[2,s2] xor s[2,1];
-v[3,s2] := flip[2,1];
-
-v[1,s3] := flip[3,1] xor flip[3,s3] xor s[3,1];
-v[2,s3] := flip[3,1];
-\end{verbatimtab}
-Then, assuming that party 0 is the ``public party'', each party sums
-its own shares, and then the sum of sum of shares is revealed
-in the output view $\vx{0}{output}$. By additive homomorphism of
-$\ttt{xor}$, this output is the sum of the 3 secrets, but
-the protocol reveals nothing to the adversary, even if one
-(or even two, in fact) of $\{1,2,3\}$ are corrupt. 
-\begin{verbatimtab}
-v[0,ss1] := v[1,s2] xor v[1,s3] xor flip[1,s1];
-v[0,ss2] := v[2,s1] xor v[2,s3] xor flip[2,s2];
-v[0,ss3] := v[3,s1] xor v[3,s2] xor flip[3,s3];
-
-v[0,output] := v[0,ss1] xor v[0,ss2] xor v[0,ss3]
-\end{verbatimtab}
-
-Because this protocol correctly implements 3-party addition in the
-binary field, it can reveal information to the adversary. For example,
-suppose that $1 \in C$, and $\sx{1}{1}$ is $1$, and after running the
-protocol suppose that $\vx{0}{output}$ is 0. Then the adversary knows
-that 2 and 3's secrets are either 1 and 0, or 0 and 1, with 50/50
-probability, and definitely neither 0 and 0, nor 1 and 1. However, the
-protocol does not reveal anything more through adversarial views than
-what the output reveals.
-
-\medskip
 

### 2023-06-07 popl24/minicat.tex
checkpoint

@@
 Expressions forms include the convenience
-$\select{\be_1}{\be_2}{\be_3}$ which is essentiall a conditional
+$\select{\be_1}{\be_2}{\be_3}$ which is essentially a conditional
 expression with:
@@
 
+\begin{example}[One-Time Pad]
+  \label{example-otp}
+We can model secure communication between two parties
+using symmetric key encryption over an insecure channel, for
+example. In the following program, we assume $H = \{ 1,2 \}$ and
+$C = \{ 0 \}$. Party 1 first shares the randomly generated
+key $\fx{1}{0}$ with party 2. Then, party 1 encrypts
+its secret $\sx{1}{0}$ and sends it to party 0,
+which relays the ciphertext in $\vx{0}{0}$ to
+party 2. Finally, party 2 decrypts\footnote{Recall we are using
+a random tape approach, so the value of $\fx{1}{0}$ remains
+consistent during any run the program, though that value
+is randomly determined by the initial memory in a given
+run} the ciphertext and stores the plaintext in $\vx{2}{2}$. 
 \begin{verbatimtab}
@@
 \end{verbatimtab}
+This protocol is \emph{correct} because $\vx{2}{2}$ is exactly
+correlated with the secret $\sx{1}{0}$, and it is \emph{secure}
+because the corrupt view $\vx{0}{0}$ in isolation is probabilistically
+independent of the secret $\sx{1}{0}$, while serving as a
+communication channel for its encryption. 
+\end{example}
+
+\begin{example}[Lambda-obliv]
+  \label{example-lambda-obliv}
+
+  \begin{verbatimtab}
+v[0,0] := select(s[1,0], flip[1,0], flip[1,1] );
+v[0,1] := flip[1,0]
+\end{verbatimtab}
+
+\begin{verbatimtab}
+v[0,0] := select(s[1,0],select(flip[1,0], flip[1,0], flip[1,1]),flip[1,2])
+\end{verbatimtab}
+\end{example}
+
+\begin{example}[Oblivious Transfer]
+  \label{example-lambda-OT}
 
@@
 \end{verbatimtab}
+\end{example}
+
+\begin{example}[Additive Secret Sharing]
+    \label{example-lambda-he}
+Another 
+example of an MPC protocol is additive secret sharing, where $k$
+parties split up their secrets into \emph{shares}, with the property
+that all $k$ shares are required to reconstruct the secret and any
+fewer reveals nothing. In $\fedcat$, we can use $\ttt{xor}$, again, to
+generate shares. Here, party 1 generates 3 shares of the secret
+$\sx{1}{1}$- one is given to party 2 through view $\vx{2}{s1}$, one is
+given to party 3 through view $\vx{3}{s1}$, and party 1 keeps
+$\fx{1}{s1}$ for itself:
+\begin{verbatimtab}
+v[2,s1] := flip[1,1] xor flip[1,s1] xor s[1,1];
+v[3,s1] := flip[1,1];
+\end{verbatimtab}
+Note that $\vx{2}{s1}$ and $\vx{3}{s1}$ are both in uniform
+random distributions.
+
+Another property of $\ttt{xor}$ is that it is additively homomorphic,
+while functioning as addition in the boolean algebra (i.e., the binary
+field). That is, given:
+\begin{mathpar}
+  x_1\ \ttt{xor}\ y_1 = z_1
+
+  x_2\ \ttt{xor}\ y_2 = z_2
+\end{mathpar}
+then we have:
+$$
+(x_1\ \ttt{xor}\ x_2)\ \ttt{xor}\ (y_1\ \ttt{xor}\ y_2) = z_1\ \ttt{xor}\ z_2
+$$
+Thus, we can obtain secret 3-party summation (xor) by continuing
+the protocol begun above- parties 2 and 3 can share their secrets
+in the same fashion:
+\begin{verbatimtab}
+v[1,s2] := flip[2,1] xor flip[2,s2] xor s[2,1];
+v[3,s2] := flip[2,1];
+
+v[1,s3] := flip[3,1] xor flip[3,s3] xor s[3,1];
+v[2,s3] := flip[3,1];
+\end{verbatimtab}
+Then, assuming that party 0 is the ``public party'', each party sums
+its own shares, and then the sum of sum of shares is revealed
+in the output view $\vx{0}{output}$. By additive homomorphism of
+$\ttt{xor}$, this output is the sum of the 3 secrets, but
+the protocol reveals nothing to the adversary, even if one
+(or even two, in fact) of $\{1,2,3\}$ are corrupt. 
+\begin{verbatimtab}
+v[0,ss1] := v[1,s2] xor v[1,s3] xor flip[1,s1];
+v[0,ss2] := v[2,s1] xor v[2,s3] xor flip[2,s2];
+v[0,ss3] := v[3,s1] xor v[3,s2] xor flip[3,s3];
+
+v[0,output] := v[0,ss1] xor v[0,ss2] xor v[0,ss3]
+\end{verbatimtab}
+
+Because this protocol correctly implements 3-party addition in the
+binary field, it can reveal information to the adversary. For example,
+suppose that $1 \in C$, and $\sx{1}{1}$ is $1$, and after running the
+protocol suppose that $\vx{0}{output}$ is 0. Then the adversary knows
+that 2 and 3's secrets are either 1 and 0, or 0 and 1, with 50/50
+probability, and definitely neither 0 and 0, nor 1 and 1. However, the
+protocol does not reveal anything more through adversarial views than
+what the output reveals.
+\end{example}

### 2023-06-07 popl24/minicat.tex
checkpoint

@@
 view assignments. Memories $\store$ are finite (partial) mapping from
-variables to values. The \emph{domain} of a memory is written
-$\dom(\store)$ and is the finite set of variables on which the memory
-is defined. We write $\store\{ x \mapsto v\}$ for
+variables to binary values $\beta \in \{0,1\}$. The \emph{domain} of a
+memory is written $\dom(\store)$ and is the finite set of variables on
+which the memory is defined. We write $\store\{ x \mapsto \beta\}$ for
 $x\not\in\dom(\store)$ to denote the memory $\store'$ such that
-$\store'(x) = v$ and otherwise $\store'(y) = \store(y)$ for all $y \in
+$\store'(x) = \beta$ and otherwise $\store'(y) = \store(y)$ for all $y \in
 \dom(\store)$.
@@
 Given a variable-free boolean expression $\be$, we write $\cod{\be}$
-to denote the standard interpretation of $\be$.  With the introduction
-of variables to expressions, we have two concerns. First, we need to
-interpret variables with respect to a specific memory, and second,
-we need to ensure that variables are used ``legally''. That is, since
-expressions define local computation, all variables used in an
-expression must belong to the same client.  Thus, we denote
-interpretation of expressions $\be$ possibly containing variables as
-$\lcod{\store,\be}{\cid}$, where $\store$ associates variables with
-values and all variables must be owned by client $\cid$. This is
-defined in Figure \ref{fig-minifed-interp}.
+to denote the standard interpretation of $\be$ in the binary field.
+With the introduction of variables to expressions, we have two
+concerns. First, we need to interpret variables with respect to a
+specific memory, and second, we need to ensure that variables are used
+``legally''. That is, since expressions define local computation, all
+variables used in an expression must belong to the same client.  Thus,
+we denote interpretation of expressions $\be$ possibly containing
+variables as $\lcod{\store,\be}{\cid}$, where $\store$ associates
+variables with values and all variables must be owned by client
+$\cid$. This is defined in Figure \ref{fig-minifed-interp}.
 
@@
 \begin{eqnarray*}
-\lcod{\store, \etrue}{\cid} &=& \etrue\\
-\lcod{\store, \efalse}{\cid} &=& \efalse\\
+\lcod{\store, \etrue}{\cid} &=& 1\\
+\lcod{\store, \efalse}{\cid} &=& 0\\
 \lcod{\store, \flip{\cid}{w}}{\cid} &=& \store(\flip{\cid}{w})\\
@@
   \inferrule
-  {b = \text{if\ } \lcod{\store,\be_1}{\cid}  \text{\ then\ } \lcod{\store,\be_2}{\cid'} \text{\ else\ } \lcod{\store,\be_3}{\cid'}}
-      {(\store, \eassign{\view{\cid}{w}}{\OT{\be_1}{\be_2}{\be_3}};\prog) \redx (\extend{\store}{\view{\cid}{w}}{b}, \prog)}
+  {\beta = \text{if\ } \lcod{\store,\be_1}{\cid}  \text{\ then\ } \lcod{\store,\be_2}{\cid'} \text{\ else\ } \lcod{\store,\be_3}{\cid'}}
+      {(\store, \eassign{\view{\cid}{w}}{\OT{\be_1}{\be_2}{\be_3}};\prog) \redx (\extend{\store}{\view{\cid}{w}}{\beta}, \prog)}
 \end{mathpar}

### 2023-06-07 popl24/pmf.tex
checkpoint

@@
 
+In order to precisely characterize distributions of secrets and views
+in programs we define probability mass functions in a standard manner,
+following the formulation in \ref{XXX}. In this formulation, memories
+denote conjunctions of variable assignments. Intuitively, given a
+program $\prog$ with $\iov(\prog) = S \cup V$ and $\store \in
+\mems(\iov(\prog))$, when considering the execution of $\prog$
+initialized with secrets $\store_S$, we are fundamentally concerned
+with the joint probability of the assignments in $\store_V$
+conditioned by the assignments $\store_S$. This is because
+probabilistic dependence between honest secrets and corrupt views
+\emph{is} information leakage.
+
+\subsection{Basic Definitions} We begin by defining memory probability distributions as follows.
 \begin{definition}
-  A \emph{memory probability distribution} $\pdf{X}$ is a function mapping
-  memories in $\mems(X)$ to values in $[0..1]$, such that:
+  A \emph{memory probability distribution} $\pdf{X}$ is a function
+  mapping memories in $\mems(X)$ to values in $[0..1]$, such that:
   $$
@@
 \end{definition}
-
+To recover succinct and familiar notation, we may omit the domain of a
+distribution when it is clear from an application context-
+i.e., we allow the following sugaring:
+$$
+\pdf{}(\store) \defeq \pdf{\dom(\store)}(\store)
+$$
+Now, we can define a notion of marginal and conditional
+distributions as follows, which are standard for discrete
+probability mass functions. 
 \begin{definition}
@@
 \end{definition}
+To recover familiar notation we allow the following syntactic
+sugarings:
+\begin{eqnarray*}
+\pdf{X}(\store|\store') &\defeq& (\condd{\pdf{X}}{\store})(\store')\\
+\pdf{X}(\store) &\defeq& (\margd{\pdf{X}}{\dom(\store)})(\store) \qquad \dom(\store) \subset X
+\end{eqnarray*}
 
+\subsection{Program Distributions}
 Now we can define the probability distribution of a program $\prog$,
-that we denote $\progd(\prog)$. Recall that $\fedcat$ is
-deterministic, so a run is determined by the input secrets together
-with random flip assignments. And since we constrain programs to not
-overwrite memories, we can be sure that final memories resulting from
-computation contain both a complete record of all initial secrets and
-flips as well as views of communicated information. Thus a final
-memory represents a run of the program, and since we consider and set
-of \emph{input} secrets and flips to be equally likely, we consider
-each memory in the set of final memories of a program to be equally
-likely, which establishes the basic probability distribution for the
-program. And since we don't care about flip values which are only
-obversable insofar as they affect views, the probability distribution
-we are really interested in is the marginal distribution of secrets
-and views in this basic program distribution.
+that we denote $\progd(\prog)$. Since $\fedcat$ is deterministic the
+results (view assignments) of any run are determined by the input
+secrets together with the random tape. And since we constrain programs
+to not overwrite views we are assured that \emph{final} memories
+contain both a complete record of all initial secrets as well as views
+resulting from communicated information.
+
+We consider random tapes to be selected from a uniform distribution.
+In our threat model we will also assume that any possible initial
+secrets are equally likely when formulating the program distribution.
+In this setting, given a program $\prog$ with $\iov(\prog) = S \cup
+V$ we will consider all $\store
+\in \mems(\iov(\prog) \cup \flips(\prog))$ such that
+$
+\config{\store_{S \cup \flips(\prog)}}{\prog} \redxs \config{\store_}{\varnothing}
+$
+to be equally probable, establishing the basic distribution of the
+program. From this we can immediately derive the marginal distribution
+of $S \cup V$ to reason about dependencies between secrets and views. 
 \begin{definition}
+  \label{def-progd}
   Given progam $\prog$ with $\iov(\prog) = (S,V)$ and $\flips(\prog) = F$, define $\stores$ as:

### 2023-06-07 popl24/minicat.tex
checkpoint

@@
 \subsection{Syntax} The syntax of $\minifed$, defined in
-Figure \ref{fig-minifed-syntax}, includes a standard
-boolean algebra with $\eand$, $\eor$, $\exor$, and $\enot$ as
-primitives. In addition programs have different kinds of variables,
-including \emph{secrets} $\secret{\cid}{w}$, \emph{flips}
-$\flip{\cid}{w}$, and \emph{views} $\view{\cid}{w}$.  Each of these
-variables are indexed by an ``owner'' client $\cid$ and distinguishing
-string $w$. So for example, $\secret{\ttt{1}}{\ttt{foo}}$ is client
-$\ttt{1}$'s secret called $\ttt{foo}$. All clients can make reference
-to the \emph{oracle} through variables $\oracle{w}$- this is a form
-of shared randomness that is standard in the MPC setting \cite{XXX}.
-Each client can only compute on their own variables in
-\emph{expressions} $\be$, and share values with other clients by
-assignment to their views as in $\eassign{\view{\cid}{w}}{\be}$.  A
-\emph{protocol} $\prog$ is a possibly empty sequence of view
-assignments.
+Figure \ref{fig-minifed-syntax}, includes a standard boolean algebra
+with $\eand$, $\eor$, $\exor$, and $\enot$ as primitives. In addition
+programs have different kinds of variables, including \emph{secrets}
+$\secret{\cid}{w}$, \emph{flips} $\flip{\cid}{w}$, and \emph{views}
+$\view{\cid}{w}$.  Each of these variables are indexed by an ``owner''
+client $\cid$ and distinguishing string $w$. So for example,
+$\secret{\ttt{1}}{\ttt{foo}}$ is client $\ttt{1}$'s secret called
+$\ttt{foo}$. All clients can make reference to the \emph{oracle}
+through variables $\oracle{w}$- this is a form of shared randomness
+that is standard in the MPC setting \cite{XXX}.  Each client can only
+compute on their own variables in \emph{expressions} $\be$, and share
+values with other clients by assignment to their views as in
+$\eassign{\view{\cid}{w}}{\be}$.  A \emph{protocol} $\prog$ is a
+possibly empty sequence of view assignments. We will generally omit
+$\varnothing$ from example code, i.e., writing $c_1;\ldots;\c_n$
+instead of $c_1;\ldots;\c_n;\varnothing$, and abusing notation we will
+write $\prog_1;\prog_2$ to denote the concatenation of $\prog_1$
+and $\prog_2$.
 

### 2023-06-07 popl24/minicat.tex
checkpoint

@@
 possibly empty sequence of view assignments. We will generally omit
-$\varnothing$ from example code, i.e., writing $c_1;\ldots;\c_n$
-instead of $c_1;\ldots;\c_n;\varnothing$, and abusing notation we will
+$\varnothing$ from example code, writing $\instr_1;\ldots;\instr_n$
+instead of $\instr_1;\ldots;\instr_n;\varnothing$, and abusing notation we will
 write $\prog_1;\prog_2$ to denote the concatenation of $\prog_1$

### 2023-06-07 popl24/pmf.tex
checkpoint

@@
 In this setting, given a program $\prog$ with $\iov(\prog) = S \cup
-V$ we will consider all $\store
-\in \mems(\iov(\prog) \cup \flips(\prog))$ such that
+V$ and $\flips(\prog) = F$ we will consider all $\store
+\in \mems(S \cup V \cup F)$ such that
 $
-\config{\store_{S \cup \flips(\prog)}}{\prog} \redxs \config{\store_}{\varnothing}
+\config{\store_{S \cup F}}{\prog} \redxs \config{\store_}{\varnothing}
 $
@@
   $$
+  In some cases we will also be concerned with the (joint)
+  probabilities of expression interpretation given a preceding program
+  execution, and we write $\progd(\prog, \be)$ to denote the program
+  distribution $\progd(\prog;\eassign{\itv}{\be})$ where $\itv$ is a
+  special variable that is never used in programs.
 \end{definition}

### 2023-06-08 popl24/minicat.tex
checkpoint

@@
 
+Here we introduce some examples to illustrate $\minifed$ and how it
+can be used to model secure communication protocols for values in the
+binary field. Most of these are translations (into $\minifed$) of
+object examples in previous related work \cite{XXX}. We will return to
+these as running examples throughout the paper to illustrate various
+concepts and advantages of our methods.
 \begin{example}[One-Time Pad]
   \label{example-otp}
-We can model secure communication between two parties
-using symmetric key encryption over an insecure channel, for
-example. In the following program, we assume $H = \{ 1,2 \}$ and
-$C = \{ 0 \}$. Party 1 first shares the randomly generated
-key $\fx{1}{0}$ with party 2. Then, party 1 encrypts
-its secret $\sx{1}{0}$ and sends it to party 0,
-which relays the ciphertext in $\vx{0}{0}$ to
-party 2. Finally, party 2 decrypts\footnote{Recall we are using
-a random tape approach, so the value of $\fx{1}{0}$ remains
-consistent during any run the program, though that value
-is randomly determined by the initial memory in a given
-run} the ciphertext and stores the plaintext in $\vx{2}{2}$. 
+As observed in \cite{XXX} and elsewhere, coin flips with $\exor$ can
+be used to model one-time pad encryption in the binary field. On the
+basis of this we can model secure communication between over an
+insecure channel using symmetric key encryption. In the following
+program, we assume $H = \{ 1,2 \}$ and $C = \{ 0 \}$. Client 1 first
+shares the randomly generated ``key'' $\fx{1}{0}$ with client 2
+through view $\view{2}{0}$. Then, client 1 encrypts its secret
+$\sx{1}{0}$ and sends it to party 0, which relays the ciphertext in
+$\vx{0}{0}$ to client 2. Finally, client 2 decrypts the the ciphertext
+and stores the plaintext in $\vx{2}{2}$.
 \begin{verbatimtab}
@@
 \end{verbatimtab}
-This protocol is \emph{correct} because $\vx{2}{2}$ is exactly
-correlated with the secret $\sx{1}{0}$, and it is \emph{secure}
-because the corrupt view $\vx{0}{0}$ in isolation is probabilistically
-independent of the secret $\sx{1}{0}$, while serving as a
-communication channel for its encryption. 
+Intuitively, this protocol is \emph{correct} because $\vx{2}{2}$ is
+exactly correlated with the secret $\sx{1}{0}$, and it is
+\emph{secure} because the corrupt view $\vx{0}{0}$ in isolation is
+probabilistically independent of the secret $\sx{1}{0}$.
 \end{example}

### 2023-06-08 popl24/minicat.tex
checkpoint

@@
 
-\begin{example}[Lambda-obliv]
+\begin{example}[$\lambda_{\mathrm{obliv}}$]
   \label{example-lambda-obliv}
-
+  In \cite{XXX}, Figure 3, two examples are introduced that illustrate
+  subtleties of probabilistic programming in
+  $\lambda_{\mathrm{obliv}}$, a language for oblivious memory
+  management. We consider their translations here where we
+  assume $H = \{ 1 \}$ and $C = \{ 0 \}$. First, the following
+  examples shows that reuse of coin flips can result in release of
+  information, even though flips are chosen from a uniform
+  distribution. This is because the adversary can deduce that
+  $\secret{1}{s}$ is more likely to be 1 in case their views are the
+  same value:
   \begin{verbatimtab}
-v[0,0] := select(s[1,0], flip[1,0], flip[1,1] );
-v[0,1] := flip[1,0]
-\end{verbatimtab}
-
+    v[0,0] := select(s[1,s], flip[1,sx], flip[1,sy]);
+    v[0,1] := flip[1,sx] \end{verbatimtab}
+  In the next example, since the nested select is more likely to evaluate
+  to $1$ than $0$, the adversary can deduce that $\secret{1}{s}$ is
+  more likely to be 1 in case $\view{0}{0}$ is 1:
   \begin{verbatimtab}
-v[0,0] := select(s[1,0],select(flip[1,0], flip[1,0], flip[1,1]),flip[1,2])
-\end{verbatimtab}
+  v[0,0] := select(s[1,s],select(flip[1,sx], flip[1,sx], flip[1,sy]),flip[1,z]) \end{verbatimtab}
+  We note that these examples are considered ``bad'' in \cite{XXX} since there
+  no information about honest secrets should be revealed to the adversary in
+  their threat model. For this reason they disallow reuse of flips through a
+  linear type discipline, which is a culprit in both of these examples. However,
+  as we have shown in Example \ref{example-otp} the reuse of flips is useful in
+  the modeling of symmetric key encryption, and generally we will not enforce
+  the same linear type discipline on the use of flips.
 \end{example}
@@
   \label{example-lambda-OT}
-
+  In \cite{XXX} a protocol for $\ttt{OT}$ is defined that we recreate as
+  follows, where client 2 is the receiver and client 1 is the sender. Whereas
+  in \cite{XXX} a trusted third party provides the sender and receiver with
+  shared randomness, in our version both clients refer to the oracle $\ttt{H}$
+  to obtain shared random values. 
 \begin{verbatimtab}
-v[2, rd] := select[flip[2, d], H[r1], H[r0]];
+  v[2, rd] := select(flip[2, d], H[r1], H[r0]);
   v[1, e] := s[2, c] xor flip[2, d];
-v[2, f0] := s[1, m0] xor select[v[1, e], H[r1], H[r0]];
-v[2, f1] := s[1, m1] xor select[not v[1, e], H[r1], H[r0]];
-v[2, mc] := (select[s[2, c], v[2, f1], v[2, f0]] xor v[2, rd])
+  v[2, f0] := s[1, m0] xor select(v[1, e], H[r1], H[r0]);
+  v[2, f1] := s[1, m1] xor select(not v[1, e], H[r1], H[r0]);
+  v[2, mc] := select(s[2, c], v[2, f1], v[2, f0]) xor v[2, rd]
+\end{verbatimtab}
+Below we will discuss how to automatically verify security properties
+of this protocol that correspond to the properties proven manually in
+\cite{XXX}, and additionally how to automatically verify its passive
+security in the simulator model. With respect to our language primitives,
+we note that this protocol is probabilistically equivalent to the
+following:
+\begin{verbatimtab}  
+  view[2, mc] := OT(s[2, c], s[1, m1], s[1, m0])
 \end{verbatimtab}
@@
     \label{example-lambda-he}
-Another 
-example of an MPC protocol is additive secret sharing, where $k$
-parties split up their secrets into \emph{shares}, with the property
-that all $k$ shares are required to reconstruct the secret and any
-fewer reveals nothing. In $\fedcat$, we can use $\ttt{xor}$, again, to
-generate shares. Here, party 1 generates 3 shares of the secret
-$\sx{1}{1}$- one is given to party 2 through view $\vx{2}{s1}$, one is
-given to party 3 through view $\vx{3}{s1}$, and party 1 keeps
-$\fx{1}{s1}$ for itself:
+Additive secret sharing is an MPC protocol where $k$ parties split up
+their secrets into \emph{shares}, with the property that all $k$
+shares are required to reconstruct the secret and any fewer reveals
+nothing. Additionally, this sharing enjoys additive homomorphism. In
+the binary field it is well-known that we can extend our use
+of $\ttt{xor}$ and flips to generate an arbitrary number of shares
+with additive homomorphism.
+
+For example, to generate shares for three-party additive sharing,
+clients 1,2, and 3 can break up and distribute their shares to
+each other as follows:
 \begin{verbatimtab}
@@
   v[3,s1] := flip[1,1];
-\end{verbatimtab}
-Note that $\vx{2}{s1}$ and $\vx{3}{s1}$ are both in uniform
-random distributions.
 
-Another property of $\ttt{xor}$ is that it is additively homomorphic,
-while functioning as addition in the boolean algebra (i.e., the binary
-field). That is, given:
-\begin{mathpar}
-  x_1\ \ttt{xor}\ y_1 = z_1
-
-  x_2\ \ttt{xor}\ y_2 = z_2
-\end{mathpar}
-then we have:
-$$
-(x_1\ \ttt{xor}\ x_2)\ \ttt{xor}\ (y_1\ \ttt{xor}\ y_2) = z_1\ \ttt{xor}\ z_2
-$$
-Thus, we can obtain secret 3-party summation (xor) by continuing
-the protocol begun above- parties 2 and 3 can share their secrets
-in the same fashion:
-\begin{verbatimtab}
   v[1,s2] := flip[2,1] xor flip[2,s2] xor s[2,1];
@@
 \end{verbatimtab}
-Then, assuming that party 0 is the ``public party'', each party sums
+Then, assuming that party 0 is a ``public'' client, each party sums
 its own shares, and then the sum of sum of shares is revealed
 in the output view $\vx{0}{output}$. By additive homomorphism of
-$\ttt{xor}$, this output is the sum of the 3 secrets, but
-the protocol reveals nothing to the adversary, even if one
-(or even two, in fact) of $\{1,2,3\}$ are corrupt. 
+$\ttt{xor}$, this output is the sum of the 3 secrets.
 \begin{verbatimtab}
@@
 \end{verbatimtab}
-
-Because this protocol correctly implements 3-party addition in the
-binary field, it can reveal information to the adversary. For example,
-suppose that $1 \in C$, and $\sx{1}{1}$ is $1$, and after running the
-protocol suppose that $\vx{0}{output}$ is 0. Then the adversary knows
-that 2 and 3's secrets are either 1 and 0, or 0 and 1, with 50/50
-probability, and definitely neither 0 and 0, nor 1 and 1. However, the
-protocol does not reveal anything more through adversarial views than
-what the output reveals.
+We note that, as for MPC protocols generally, the notion of
+``secrecy'' here is subtle and not the same as for encryption as in
+Example \ref{example-otp}, particularly when the threat model allows
+protocol participants to be corrupt. For example, suppose that $1 \in C$, and
+$\sx{1}{1}$ is $1$, and after running the protocol suppose that
+$\vx{0}{output}$ is 0. Then the adversary knows that 2 and 3's secrets
+are either 1 and 0, or 0 and 1, with 50/50 probability, and definitely
+neither 0 and 0, nor 1 and 1. 
 \end{example}

### 2023-06-08 popl24/minicat.tex
checkpoint

@@
 these as running examples throughout the paper to illustrate various
-concepts and advantages of our methods.
+concepts and advantages of our methods. In Section \ref{section-metalang}
+we will formulate a larger example implementation of Yao's
+Garbled Circuits (Example \ref{example-ygc}). 
 \begin{example}[One-Time Pad]
@@
   assume $H = \{ 1 \}$ and $C = \{ 0 \}$. First, the following
-  examples shows that reuse of coin flips can result in release of
+  example shows that the reuse of coin flips can result in release of
   information, even though flips are chosen from a uniform
@@
   $\secret{1}{s}$ is more likely to be 1 in case their views are the
-  same value:
+  same value. We call this subexample $(a)$:
   \begin{verbatimtab}
@@
   to $1$ than $0$, the adversary can deduce that $\secret{1}{s}$ is
-  more likely to be 1 in case $\view{0}{0}$ is 1:
+  more likely to be 1 in case $\view{0}{0}$ is 1. We call this subexample $(b)$:
   \begin{verbatimtab}

### 2023-06-08 popl24/minicat.tex
checkpoint

@@
 \subsection{Examples}
+\label{section-minicat-examples}
 
@@
 \begin{example}[Additive Secret Sharing]
-    \label{example-lambda-he}
+    \label{example-he}
 Additive secret sharing is an MPC protocol where $k$ parties split up

### 2023-06-08 popl24/pmf.tex
checkpoint

@@
   $$
+  We note immediately that $|\stores| = 2^{S \cup F}$ given the (deterministic) definition of $\redx$. 
   Define also $\pdf{S \cup V \cup F}$ as the program's \emph{basic distribution} such that for all
@@
 \end{definition}
+
+\subsection{Examples}
+
+In the following, we will refer to program examples defined in
+Section \ref{section-minicat-examples} as $\prog_x$, where $x$
+is the Example number.
+
+\begin{fpfig}[t]{Truth table visualization of basic distributions for Examples \ref{example-otp} (L) and \ref{example-lambda-obliv}(a) (R).}{fig-basic-distributions}
+{\footnotesize
+  $$
+  \begin{array}{cc}
+    \begin{array}{cccccc}
+      \verb+s[1,0]+ & \verb+f[1,0]+ & \verb+v[2,0]+  & \verb+v[2,1]+ & \verb+v[2,2]+ & \verb+v[0,0]+\\
+      \hline
+      0 & 0 & 0 & 0 & 0 & 0 \\ 
+      0 & 1 & 1 & 1 & 0 & 1 \\ 
+      1 & 0 & 0 & 1 & 1 & 1 \\ 
+      1 & 1 & 1 & 0 & 1 & 0
+    \end{array}
+    & 
+    \begin{array}{ccccc}
+      \verb+s[1,s]+ & \verb+f[1,sx]+ & \verb+f[1,sy]+ & \verb+v[0,0]+ & \verb+v[0,1]+ \\
+      \hline
+      0 & 0 & 0 & 0 & 0 \\ 
+      0 & 0 & 1 & 1 & 0 \\ 
+      0 & 1 & 0 & 0 & 1 \\ 
+      0 & 1 & 1 & 1 & 1 \\
+      1 & 0 & 0 & 0 & 0 \\ 
+      1 & 0 & 1 & 0 & 0 \\ 
+      1 & 1 & 0 & 1 & 1 \\ 
+      1 & 1 & 1 & 1 & 1  
+    \end{array}
+  \end{array}
+  $$
+}
+\end{fpfig}
+
+\paragraph{Example \ref{example-otp} revisited} For small programs it is
+helpful to visualize their basic distributions graphically as their
+truth tables, where each row corresponds to an equally likely final
+memory for the given program. In Figure \ref{fig-basic-distributions}
+we show this for Example \ref{example-otp}. In this truth table we can
+see that $\sx{1}{0}$ and $\vx{2}{2}$ are perfectly correlated. But
+from the adversarial view, $\sx{1}{0}$ is 1 or 0 with equal
+probability given any value of $\vx{0}{0}$. More precisely, for any
+$\beta_1,\beta_2$ we have:
+\begin{mathpar}
+(\progd(\prog_{\ref{example-otp}}))(\{\sx{1}{0} \mapsto \beta_1 \}|\{ \vx{0}{0} \mapsto \beta_2 \}) = .5
+\end{mathpar}
+Note that this corresponds to \emph{perfect secrecy} as described
+in \cite{XXX}.
+
+\paragraph{Example \ref{example-lambda-obliv}(a) revisited} As demonstrated
+by Example \ref{example-lambda-obliv} taken from \cite{XXX}, it is possible
+to leak information probabilistically through obvservable dependencies
+in views. In the basic distribution for Example \ref{example-lambda-obliv}(a)
+illustrated in Figure \ref{fig-basic-distributions}, we can see that
+the adversarial views $\vx{0}{0}$ and $\vx{0}{1}$ are both 1 in
+three different runs of the protocol, and in 2 out of 3 the
+honest secret $\sx{1}{0}$ is 1. More precisely:
+$$
+(\progd(\prog_{\ref{example-lambda-obliv}(a)}))(\{\sx{1}{0} \mapsto 1 \}|\{ \vx{0}{0} \mapsto 1, \vx{0}{1} \mapsto 1, \}) = 2/3
+$$
+
+\paragraph{Example \ref{example-he} revisited} This standard example
+of an MPC secure protocol illustrates how a functionality- in this
+case addition (in the binary field)- can be implemented securely. As
+mentioned previously, it is \emph{not} the case that \emph{no}
+information about input secrets can be revealed to the adversary in
+the MPC threat model, since the result of the functionality can itself
+reveal information. For example, if the result of addition is $0$,
+the adversary knows that all clients must have the same value.
+More precisely:
+\begin{mathpar}
+  (\sum_{\beta}(\progd(\prog_{\ref{example-he}}))(\{\sx{1}{1} \mapsto \beta, \sx{2}{1} \mapsto \beta, \sx{3}{1} \mapsto \beta \}|\{ \vx{0}{output} \mapsto 0 \})) = 1
+\end{mathpar}

### 2023-06-08 popl24/pmf.tex
checkpoint

@@
 \end{mathpar}
+If we assume that a client is corrupted we can formally consider the
+impacts on information released to the adversary by conditioning the
+program distribution on corrupt secrets.  For example, if we assume
+client 2 with input secret 1 is corrupted and the output is 0, then
+adversary knows with certainty that clients 1 and 3 also have input
+secrets 1. We can express this as:

### 2023-06-09 popl24/pmf.tex
checkpoint

--- 2023-06-08 popl24/pmf.tex
@@
 reveal information. For example, if the result of addition is $0$,
-the adversary knows that all clients must have the same value.
-More precisely:
+the adversary knows that two of the clients' secrets must be 0
+and the third's must be 1, though not whose is whose- so we can
+observe:
 \begin{mathpar}
-  (\sum_{\beta}(\progd(\prog_{\ref{example-he}}))(\{\sx{1}{1} \mapsto \beta, \sx{2}{1} \mapsto \beta, \sx{3}{1} \mapsto \beta \}|\{ \vx{0}{output} \mapsto 0 \})) = 1
+  (\progd(\prog_{\ref{example-he}}))(\{\sx{1}{1} \mapsto 0, \sx{2}{1} \mapsto 1, \sx{3}{1} \mapsto 0 \}|\{ \vx{0}{output} \mapsto 0 \}) = .33
 \end{mathpar}
+Further, since the standard MPC threat model allows partipating
+clients to be corrupted we can formally consider the impacts on
+information released to the adversary by conditioning the program
+distribution on corrupt secrets.  So if we assume client 2
+with input secret 1 is corrupted, and we assume the output is 0, then
+adversary knows with certainty that clients 1 and 3 both have input
+secrets 0. We can express this as:
+$$
+\progd(\prog_{\ref{example-he}}))(\{\sx{1}{1} \mapsto 0, \sx{3}{1} \mapsto 0 \}|\{ \sx{2}{1} \mapsto 1, \vx{0}{output} \mapsto 0 \})
+= 1
+$$ A critical point to note is that this information revealed to the
+adversary does not violate MPC security- under the assumption of
+public release of the functionality result, and under the assumption
+of the possible corruption of clients, this is the cost of doing
+business. Rather, the concern is that the protocol does not release
+any additional information through information communicated to views
+of corrupted parties. Additive secret sharing as realized in
+$\prog_{\ref{example-he}}$ is MPC secure in this sense- so for
+example, for any $\beta_1$ and $\beta_2$ we have:
+$$
+\begin{array}{c}
+\progd(\prog_{\ref{example-he}}))(\{\sx{1}{1} \mapsto 0 \}|\{ \sx{2}{1} \mapsto 0, \vx{0}{output} \mapsto 0 \})
+= \\
+\progd(\prog_{\ref{example-he}}))(\{\sx{1}{1} \mapsto 0 \}|\{ \vx{2}{s1} \mapsto \beta_1, \vx{2}{s3} \mapsto \beta_2, \sx{2}{1} \mapsto 0, \vx{0}{output} \mapsto 0 \})
+= \\
+.5
+\end{array}
+$$

### 2023-06-09 popl24/pmf.tex
checkpoint

@@
 $$
+An important point illustrated by this example is that the
+probability of values in views, and not just the values themselves,
+can release information to the adversary.
 

### 2023-06-09 popl24/composition.tex
checkpoint

--- 2023-06-06 popl24/metalang.tex
@@
-\section{A Protocol Metalanguage}
-
-$$
-\begin{array}{rcl@{\hspace{8mm}}r}
-\flab &\in& \mathrm{Field}\\
-x &\in& \mathrm{EVar}\\
-f &\in& \mathrm{FName}\\[2mm]
-e &::=& b \mid \flip{e}{e} \mid \secret{e}{e} \mid \view{e}{e} \mid \oracle{e} \mid \enot\ e \mid e\ \eand\ e \mid e\ \exor\ e \mid & \textit{expressions}\\[0mm]
-& & \select{e}{e}{e} \mid 
-\send{\view{e}{e}}{e} \mid \send{\view{e}{e}}{\OT{e}{e}{e}} \mid e;e \mid \\[0mm]
-& & x \mid \elet{x}{e}{e} \mid f(e,\ldots,e) \mid \{ \flab = e; \ldots; \flab = e \}
-\mid e.\flab \mid e\ttt{||}e \mid (e) \\[2mm]
-v &::=& w \mid \cid \mid \be \mid \{ \flab = v;\ldots;\flab = v \} 
-\mid (\,) & \textit{values}\\[2mm]
-{fn} &::=& f(x,\ldots,x) \{ e \} & \textit{functions}
-\end{array}
-$$
-
-\begin{definition}
-A \emph{codebase} $\codebase$ is a list of function 
-declarations. We write $ \codebase(f) = x_1,\ldots,x_n,\ e$
-iff $f(x_1,\ldots,x_n) \{ e \} \in \codebase$.
-A \emph{metaprogram}, aka \emph{mataprotocol} is a pair of a 
-codebase and expression $\codebase, e$. We may omit
-$\codebase$ if it is clear from context.  
-\end{definition}
-
-$$
-\begin{array}{rcl@{\hspace{3mm}}r}
-E &::=& [\,] \mid \enot\ E \mid E\ \bop\ e \mid v\ \bop\ E \mid & \textit{evaluation contexts}\\[1mm]
-& & \flip{E}{e} \mid \secret{E}{e} \mid \view{E}{e} \mid \oracle{E} \mid  \\[1mm]
-& & \flip{\cid}{E} \mid \secret{\cid}{E} \mid \view{\cid}{E} \mid \send{E}{e} \mid \send{\view{\cid}{w}}{E}\\[1mm]
-& & \OT{E}{e}{e} \mid \OT{v}{E}{e} \mid \OT{v}{v}{E} \mid\\[1mm]
-& & \select{E}{e}{e} \mid \select{v}{E}{e} \mid \select{v}{v}{E} \mid\\[1mm]
-& & \elet{x}{E}{e} \mid f(v,\ldots,v,E,e,\ldots,e) \mid \\[1mm]
-& & \{ \flab = v;\ldots;\flab = v;\flab = E;\flab = e;\ldots;\flab = e \} \mid E.\flab \mid E\ttt{||}e \mid v \ttt{||}E
-\end{array}
-$$
-
-$$
-\begin{array}{rcl@{\hspace{10mm}}r}
-\config{\prog}{\elet{x}{v}{e}} &\redx& \config{\prog}{e[v/x]}\\
-\config{\prog}{f(v_1,...,v_n)} &\redx&
-\config{\prog}{e[v_1/x_1,\ldots,v_n/x_n]} & 
- \codebase(f) = x_1,\ldots,x_n,\ e\\
-\config{\prog}{\{\ldots; \flab = v; \ldots\}.\flab} &\redx&
- \config{\prog}{v}\\
- \config{\prog}{w_1\ttt{||}w_2} &\redx& \config{\prog}{w_1w_2}\\
- \config{\prog}{v;e} &\redx& \config{\prog}{e}\\
-\config{\prog}{\instr} &\redx& \config{\prog;\instr}{()}\\
-\config{\prog}{E[e]} &\redx& \config{\prog'}{E[e']} & \text{if}\ \config{\prog}{e} \redx \config{\prog'}{e'} 
-\end{array}
-$$
-
-\subsection{Compositional Properties: Garbled $\NIMO$}
+\section{Compositional Properties: Garbled $\NIMO$}
+\label{section-composition}
 

### 2023-06-09 popl24/hyperprop.tex
checkpoint

@@
 
-Finally, consider our secure communications example, that we call program
-$\prog_3$. We will assume that $H = \{ 1,2 \}$ and $C = \{ 0 \}$, so
-the adversary has access to $\vx{0}{0}$ in particular:
-\begin{verbatimtab}
-v[2,0] := f[1,0];
-v[0,0] := f[1,0] xor s[1,0];
-v[2,1] := v[0,0];
-v[2,2] := v[2,1] xor v[2,0]
-\end{verbatimtab}
-First, let's focus on information available to the adversary
-about the secret $\sx{1}{0}$ (aside from its publicly-known marginal
-distribution). We will consider the marginal distribution
-of $\sx{1}{0}$ and $\vx{0}{0}$:
-\begin{mathpar}
-X \defeq \{ \vx{0}{0}, \sx{1}{0} \}
+As we've discussed in our section on Related Work
+(\ref{section-related}), previous authors have noted connections
+between the traditional formulation of the MPC security model- aka the
+simulator security model- and hyperproperty-style formulations of
+program trace behavior \cite{XXX} including probabilistic
+noninterference. This formulation of hyperproperties that are at least
+sound with respect to simulator security bridges a conceptual gap
+between foundational approaches to security, and also motivates and
+supports automated analysis \cite{XXX}.
 
-\pdf{X} \defeq \margd{\progd(\prog_3)}{X}
-\end{mathpar}
-Since the adversary can directly observe the value in $\vx{0}{0}$ in
-any given run, we therefore consider the distribution conditioned on
-that value. As we can see, this shows that the protocol reveals
-no information about the secret:
-\begin{mathpar}
-  \condd{(\pdf{X}}{\{ \vx{0}{0} \mto 0\})}(\{ \sx{1}{0} \mapsto 0\}) = .5
-
-  \condd{(\pdf{X}}{\{ \vx{0}{0} \mto 0\})}(\{ \sx{1}{0} \mapsto 1\}) = .5
-
-  \condd{(\pdf{X}}{\{ \vx{0}{0} \mto 1\})}(\{ \sx{1}{0} \mapsto 0\}) = .5
-
-  \condd{(\pdf{X}}{\{ \vx{0}{0} \mto 1\})}(\{ \sx{1}{0} \mapsto 1\}) = .5
-\end{mathpar}
-
-Now, let's focus on the information directly available to honest parties,
-and specifically the information available through $\vx{2}{2}$ about
-$\sx{1}{0}$. We first define the relevant marginal distribution:
-\begin{mathpar}
-X \defeq \{ \vx{2}{2}, \sx{1}{0} \}
-
-\pdf{X} \defeq \margd{\progd(\prog_3)}{X}
-\end{mathpar}
-Then we consider the marginal distribution of the secret conditioned
-on the view $\vx{2}{2}$, which shows the direct correlation and therefore
-correctness. 
-\begin{mathpar}
-  \condd{(\pdf{X}}{\{ \vx{0}{0} \mto 0\})}(\{ \sx{1}{0} \mapsto 0\}) = 1
-
-  \condd{(\pdf{X}}{\{ \vx{0}{0} \mto 0\})}(\{ \sx{1}{0} \mapsto 1\}) = 0
-
-  \condd{(\pdf{X}}{\{ \vx{0}{0} \mto 1\})}(\{ \sx{1}{0} \mapsto 0\}) = 0
-
-  \condd{(\pdf{X}}{\{ \vx{0}{0} \mto 1\})}(\{ \sx{1}{0} \mapsto 1\}) = 1
-\end{mathpar}
-
-\subsection{Probabilistic Noninterference} The discussion above suggests a
-security semantics- that is, a hyperproperty- analagous to
-noninterference in a probabilistic setting. First, let's
-define low equivalence of memories in the current setting:
-\begin{definition}[Memory Low Equivalence]
-  Given $\store^1$ and $\store^2$ with $\dom(\store^1) = \dom(\store^2) = X$,
-  we write $\store^1 =_C \store^2$ iff $\store^1_{X_C} = \store^2_{X_C}$.
-\end{definition}
-Now, we can define probabilistic noninterference as a property of
-programs ensuring that low-equivalent initial memories produce
-the same low-observable trace (i.e., corrupt views) \emph{with the
-same probability}. 
-\begin{definition}[Probabilistic Noninterference]
-  Given a program $\prog$ with $\iov(\prog) = (S,V)$, we say that
-  $\prog$ satisfies \emph{probabilistic noninterference} iff for all
-  $\store_1, \store_2 \in \mems(S)$:
-  $$\store_1 =_C \store_2 \implies
-    \margd{(\condd{\progd(\prog)}{\store_1})}{V_C} =
-    \margd{(\condd{\progd(\prog)}{\store_2})}{V_C}$$
-\end{definition}
-
-An important consequence of probabilistic noninterference refines an
-idea we introduced in Lecture Notes 12, which is that noninterering
-programs do not improve adversarial knowledge about secrets.
-More precisely, in $\fedcat$ the initial probability of a correct
-guess by the adversary is the marginal distribution of honest secrets
-in the program distribution conditioned by corrupt inputs.
-The final (and thus best, due to monotonicity) adversarial
-knowledge is also the marginal distribution of honest secrets in the
-program distribution, but conditioned by corrupt inputs \emph{and}
-corrupt views. Noninterference in the probabilistic setting implies
-that these are the same.
-\begin{lemma}
-  Given a program $\prog$ with $\iov(\prog) = (S,V)$, 
-  $\prog$ satisfies probabilistic noninterference iff for all
-  $\store \in \mems(S \cup V)$:
-  $$\margd{(\condd{\progd(\prog)}{\store_{S_C}})}{S_H} =
-    \margd{(\condd{\progd(\prog)}{\store_{(S\cup V)_C}})}{S_H} $$
-\end{lemma}
+In this section we formulate a new hyperproperty that we show is sound
+for passive security in the real/ideal model, called
+\emph{noninterference modulo output ($\NIMO$)}. We do not claim
+completeness with respect to passive security, but soundness
+guarantees that enforcement of $\NIMO$ in our analyses implies passive
+security. In Section \section{section-bruteforce} we show how $\NIMO$ can
+be automatically enforced in a brute-force manner, and in
+Section \section{section-bruteforce}
 
 \subsection{Passive Security and the Real/Ideal Model} The traditional
-security model of MPC is formulated by positing a \emph{simulator} that
-exists in the ideal world, and is trying to reconstruct observable
-information in a real world run of a protocol $\prog$ that
+security model of MPC is formulated by positing a \emph{simulator}
+that exists in the ideal world, and is trying to reconstruct
+observable information in a real world run of a protocol $\prog$ that
 correctly implements an ideal functionality $\idealf$- that is,
-execution of a configuration $(\store,\prog)$. We will assume
-that any protocol $\prog$ includes a public \emph{output view}
-$x$ where the result is published. The protocol is
-\emph{correct} iff the resulting memory contains $x \mapsto
-\idealf(\store_S)$, where $S$ are the secret inputs (honest
-and corrupt) in $\prog$.
+execution of a configuration $(\store,\prog)$. We will assume that any
+protocol $\prog$ includes a public \emph{output view} $x$ where the
+result is published. The protocol is \emph{correct} iff the resulting
+memory contains $x \mapsto \idealf(\store_S)$, where $S$ are the
+secret inputs (honest and corrupt) in $\prog$.
 
@@
 
-%\newcommand{\outv}{\mathit{out}}
+\subsection{Probabilistic Noninterference} The discussion above suggests a
+security semantics- that is, a hyperproperty- analagous to
+noninterference in a probabilistic setting. First, let's
+define low equivalence of memories in the current setting:
+\begin{definition}[Memory Low Equivalence]
+  Given $\store^1$ and $\store^2$ with $\dom(\store^1) = \dom(\store^2) = X$,
+  we write $\store^1 =_C \store^2$ iff $\store^1_{X_C} = \store^2_{X_C}$.
+\end{definition}
+Now, we can define probabilistic noninterference as a property of
+programs ensuring that low-equivalent initial memories produce
+the same low-observable trace (i.e., corrupt views) \emph{with the
+same probability}. 
+\begin{definition}[Probabilistic Noninterference]
+  Given a program $\prog$ with $\iov(\prog) = (S,V)$, we say that
+  $\prog$ satisfies \emph{probabilistic noninterference} iff for all
+  $\store_1, \store_2 \in \mems(S)$:
+  $$\store_1 =_C \store_2 \implies
+    \margd{(\condd{\progd(\prog)}{\store_1})}{V_C} =
+    \margd{(\condd{\progd(\prog)}{\store_2})}{V_C}$$
+\end{definition}
+
+An important consequence of probabilistic noninterference refines an
+idea we introduced in Lecture Notes 12, which is that noninterering
+programs do not improve adversarial knowledge about secrets.
+More precisely, in $\fedcat$ the initial probability of a correct
+guess by the adversary is the marginal distribution of honest secrets
+in the program distribution conditioned by corrupt inputs.
+The final (and thus best, due to monotonicity) adversarial
+knowledge is also the marginal distribution of honest secrets in the
+program distribution, but conditioned by corrupt inputs \emph{and}
+corrupt views. Noninterference in the probabilistic setting implies
+that these are the same.
+\begin{lemma}
+  Given a program $\prog$ with $\iov(\prog) = (S,V)$, 
+  $\prog$ satisfies probabilistic noninterference iff for all
+  $\store \in \mems(S \cup V)$:
+  $$\margd{(\condd{\progd(\prog)}{\store_{S_C}})}{S_H} =
+    \margd{(\condd{\progd(\prog)}{\store_{(S\cup V)_C}})}{S_H} $$
+\end{lemma}
 

### 2023-06-09 popl24/metalang.tex
checkpoint

--- 2023-06-06 popl24/metalang.tex
@@
 \section{A Protocol Metalanguage}
+\label{section-metalang}
 
@@
 $$
-
-\begin{verbatimtab}
-andgg(g, wla, wlb, wlc, wva, wvb) {
-andgate(g, wla, wlb, wlc);
-evalgate(g, wva, wvb)
-}
-\end{verbatimtab}
-
-\begin{lemma}[Input Encoding]
-Assume given the following evaluation relation:
-$$
-\config{\prog_1}{E[\ttt{encode}(w,v_1,v_2)]} \redxs \config{\prog_1;\prog_2}{E[v]}
-$$
-If there exists gate identifiers $g_1, g_2$ such that:
-\begin{enumerate}
-  \item $g_1 \ne g_2 \wedge g_1,g_2 \not\in \gates(\prog_1)$
-  \item $\owl(\prog_1,v_1,g_1) \wedge \owl(\prog_1,v_2,g_2)$
-  \item $\gNIMO(\prog_1,v.\ttt{wv1},g_1) \wedge \gNIMO(\prog_1,v.\ttt{wv2},g_2)$
-\end{enumerate}
-then:
-\begin{enumerate}
-\item $g_1,g_2 \in  \gates(\prog_2)$
-\item $\gNIMO(\prog_1;\prog_2,v.\ttt{wv1},g_1) \wedge \gNIMO(\prog_1;\prog_2,v.\ttt{wv2},g_2)$
-\item $$(\margd{\progd(\prog_1;\prog_2)}{\iov(\prog_1)}) * (\margd{\progd(\prog_1;\prog_2)}{\iov(\prog_2)})$$
-\end{enumerate}
-\end{lemma}
-
-\begin{lemma}[Composition ($\eand$)]
-Assume given the following evaluation relation:
-$$
-\config{\prog_1}{E[\ttt{andgg}(g,v_1,\ldots,v_5)]} \redxs \config{\prog_1;\prog_2}{E[v_6]}
-$$
-If there exist gate identifiers $g_1, g_2$ such that:
-\begin{enumerate}
-  \item $g_1 \ne g_2 \ne g \wedge g_1,g_2 \in \gates(\prog_1) \wedge g \not\in \gates(\prog_1)$
-  \item $\owl(\prog_1,v_1,g_1) \wedge \owl(\prog_1,v_2,g_2) \wedge \owl(\prog_1,v_3,g)$
-  \item $\gNIMO(\prog_1,v_4,g_1) \wedge \gNIMO(\prog_1,v_5,g_2)$
-\end{enumerate}
-then:
-\begin{enumerate}
-  \item $g \in \gates(\prog_2)$
-  \item $\gNIMO(\prog_1;\prog_2,v_6,g)$
-  \item $(\margd{\progd(\prog_1;\prog_2)}{\iov(\prog_1)}) * (\margd{\progd(\prog_1;\prog_2)}{\iov(\prog_2)})$
-\end{enumerate}
-\end{lemma}
-
-\begin{lemma}[Decoding]
-Assume given the following evaluation relation:
-$$
-\config{\prog_1}{E[\ttt{decoder}(v_1,v_2)]} \redxs \config{\prog_1;\prog_2}{E[\be]}
-$$
-If there exists gate identifier $g \in \gates(\prog_1)$ such that:
-\begin{enumerate}
-  \item $\owl(\prog_1,v_1,g)$
-  \item $\gNIMO(\prog_1,v_2,g)$
-\end{enumerate}
-Then $\NIMO(\prog_1;\prog_2;\eassign{x}{\be})$.
-\end{lemma}

### 2023-06-09 popl24/pmf.tex
checkpoint

--- 2023-06-09 popl24/pmf.tex
@@
 $$
-(\progd(\prog_{\ref{example-lambda-obliv}(a)}))(\{\sx{1}{0} \mapsto 1 \}|\{ \vx{0}{0} \mapsto 1, \vx{0}{1} \mapsto 1, \}) = 2/3
+(\progd(\prog_{\ref{example-lambda-obliv}(a)}))(\{\sx{1}{0} \mapsto 1 \}|\{ \vx{0}{0} \mapsto 1, \vx{0}{1} \mapsto 1 \}) = 2/3
 $$
+An important point illustrated by this example is that the
+probability of values in views, and not just the values themselves,
+can release information to the adversary.
 

### 2023-06-09 popl24/hyperprop.tex
checkpoint

@@
 guarantees that enforcement of $\NIMO$ in our analyses implies passive
-security. In Section \section{section-bruteforce} we show how $\NIMO$ can
+security. In Section \ref{section-bruteforce} we show how $\NIMO$ can
 be automatically enforced in a brute-force manner, and in
-Section \section{section-bruteforce}
+Section \ref{section-composition}
 

### 2023-06-09 popl24/hyperprop.tex
checkpoint

@@
 security. In Section \ref{section-bruteforce} we show how $\NIMO$ can
-be automatically enforced in a brute-force manner, and in
-Section \ref{section-composition}
+be automatically enforced in a brute-force manner, and in Section
+\ref{section-composition} we show how $\NIMO$ can be refined to obtain
+compositional properties that can be verified in a scalable manner for
+protocols that incorporate modular sub-protocols, with a
+specific use-case of Yao's Garbled Circuits (YGC) as an example.
+The latter methodology demonstrates a particular advantage of our
+hyperproperty formulation.
 

### 2023-06-09 popl24/hyperprop.tex
checkpoint

@@
 \ref{section-composition} we show how $\NIMO$ can be refined to obtain
-compositional properties that can be verified in a scalable manner for
-protocols that incorporate modular sub-protocols, with a
-specific use-case of Yao's Garbled Circuits (YGC) as an example.
-The latter methodology demonstrates a particular advantage of our
-hyperproperty formulation.
+compositional properties supporting efficient, scalable enforcement of
+complex protocols that incorporate modular subprotocols, with a
+specific use-case of Yao's Garbled Circuits (YGC) as an example where
+garbled gates are the compositional units.  The latter methodology
+demonstrates a particular advantage of our hyperproperty formulation.
 
-\subsection{Passive Security and the Real/Ideal Model} The traditional
-security model of MPC is formulated by positing a \emph{simulator}
-that exists in the ideal world, and is trying to reconstruct
-observable information in a real world run of a protocol $\prog$ that
-correctly implements an ideal functionality $\idealf$- that is,
-execution of a configuration $(\store,\prog)$. We will assume that any
-protocol $\prog$ includes a public \emph{output view} $x$ where the
-result is published. The protocol is \emph{correct} iff the resulting
-memory contains $x \mapsto \idealf(\store_S)$, where $S$ are the
+\subsection{Passive Security and the Real/Ideal Model} Here we
+briefly summarize the formal definition of MPC security. For a
+more detailed discussion the reader is referred to \cite{XXX}.
+In this paper we will consider the \emph{passive} aka
+semi-honest aka honest-but-curious variant. In this threat
+model, protocol participants can be corrupted, but all clients
+follow the rules of the protocol- as opposed to the
+\emph{active} aka maliciouos variant where participants
+are allowed to break these rules. In the passive model, the
+vulnerability is through information leaked to the adversary
+through the views of corrupted participants. While the passive
+model also assumes that any clients can potentially be
+corrupted, it assumes that no more than half will be-
+i.e., it assumes an \emph{honest majority}.
+
+The real/ideal model posits a \emph{simulator} that exists in an ideal
+world, whose goal is to reconstruct observable information in a real
+world run of a protocol $\prog$ that correctly implements an ideal
+functionality $\idealf$- that is, execution of a configuration
+$\config{\store}{\prog}$. We will assume that any protocol $\prog$
+includes a public \emph{output view} $\outv$ where the result is
+published, where $0$ is the ``public'' client and thus $\outv$
+is of the form $\vx{0}{w}$ for some $w$, i.e., is owned by 0.
+Cl
+Note that the protocol is \emph{correct} iff the resulting
+memory contains $\outv \mapsto \idealf(\store_S)$, where $S$ are the
 secret inputs (honest and corrupt) in $\prog$.
@@
 world adversarial views, by reconstructing their distributions, then
-those views leak no additional information, i.e., \emph{have no
+those views leak no additional information, i.e., they have no
 additional dependencies on honest input secrets than what is revealed
-by the output and conditionings on corrupt inputs}.
+by the output and conditionings on corrupt inputs.
 
@@
 Then we can define passive security in the real/ideal
-model as follows. One detail here is the assumption of an
-\emph{honest majority} which is a standard bound on the
-capabilities of the adversary in this threat model.
+model as follows. 
 \begin{definition}[Passive Security]
@@
 
-\subsection{Probabilistic Noninterference} The discussion above suggests a
-security semantics- that is, a hyperproperty- analagous to
-noninterference in a probabilistic setting. First, let's
-define low equivalence of memories in the current setting:
+\subsection{Probabilistic Noninterference} Previous work has noted
+relations between probabilistic noninterference and security
+properties such as memory trace obliviousness \cite{XXX} and passive
+security \cite{XXX}. And above (Section \ref{section-pmf-examples}) we
+have demonstrated that our basic encryption scheme enjoys a noninterference
+property. To formulate probabilistic noninterference in
+our setting we begin by defining low equivalence of memories.
 \begin{definition}[Memory Low Equivalence]
@@
 
-An important consequence of probabilistic noninterference refines an
-idea we introduced in Lecture Notes 12, which is that noninterering
-programs do not improve adversarial knowledge about secrets.
-More precisely, in $\fedcat$ the initial probability of a correct
-guess by the adversary is the marginal distribution of honest secrets
-in the program distribution conditioned by corrupt inputs.
-The final (and thus best, due to monotonicity) adversarial
-knowledge is also the marginal distribution of honest secrets in the
-program distribution, but conditioned by corrupt inputs \emph{and}
-corrupt views. Noninterference in the probabilistic setting implies
-that these are the same.
+Despite the intuitive resonance between noninterference
+and passive security, since MPC
+protocols implement an ideal functionality, they cannot generally satisfy
+noninterference. We can easily demonstrate this by using
+Example \ref{example-he} as a failure witness, where we'll
+assume $C = \{0,2\}$ and $H = \{1,3\}$:
+$$
+(\progd(\prog_{\ref{example-he}}))({\outv \mapsto 1}|\{\{\sx{1}{1} \mapsto 0, \sx{2}{1} \mapsto 1, \sx{3}{1} \mapsto 0 \} \})
+= 1
+$$
+whereas:
+$$
+(\progd(\prog_{\ref{example-he}}))({\outv \mapsto 1}|\{\{\sx{1}{1} \mapsto 1, \sx{2}{1} \mapsto 1, \sx{3}{1} \mapsto 0 \} \})
+= 0
+$$
+
+Probabilistic noninterference can also be restated to formally capture a
+basic intuition that noninterfering programs do not provide
+the adversary with any clues about honest secrets through corrupt
+views, as in the following Lemma. This is similar to asserting that
+adversarial knowledge is not increased by the execution of a program as in formulations
+of so-called gradual release \cite{XXX}. 
 \begin{lemma}

### 2023-06-09 popl24/hyperprop.tex
checkpoint

@@
 
-\subsection{A Sound Hyperproperty for Passive Security} The real/ideal
-model can be hard to understand and passive security of protocols can
-be difficult to prove, much less enforce automatically. Formulating
-a hyperproperty that implies passive security can provide some
-insights into the meaning of passive security, provide a connection
-to the foundational theories we've discussed, and can provide
-methodological support for proofs and even enforcement.
-
+\subsection{A Sound Hyperproperty for Passive Security}
 The intuition that passive security means that the protocol releases
 no more information than what is released by the output suggests a
-refinement of noninterference. Specifically, it suggests that in
-passive secure protocols, the adversarial knowledge (in the Lecture
-Notes 12 sense) obtained from the output does not change during
-protocol execution, which otherwise should be noninterfering.  That
-is, the set of low equivalent initial memories that can produce the
-same output should generate the same distributions of adversarial
-views. 
+refinement of noninterference. That is, the set of low equivalent
+initial memories that can produce the same output should generate the
+same distributions of adversarial views.
 \begin{definition}[Noninterference Modulo Output]

### 2023-06-09 popl24/hyperprop.tex
checkpoint

@@
 published, where $0$ is the ``public'' client and thus $\outv$
-is of the form $\vx{0}{w}$ for some $w$, i.e., is owned by 0.
-Cl
+is of the form $\view{0}{w}$ for some $w$, i.e., is owned by 0.
 Note that the protocol is \emph{correct} iff the resulting
@@
 $\store_{S_C}$- and the output of the protocol- that is,
-$\idealf(\store)$- and tries to reconstruct the \emph{distribution of
-adversarial views} observed in the real world. The intuition here is
+$\idealf(\store)$- and aims to reconstruct the distribution of
+adversarial views observed in the real world. The intuition here is
 that the simulator is given the information that is necessarily
@@
 
-\subsection{Probabilistic Noninterference} Previous work has noted
+\subsection{Probabilistic Noninterference}
+\label{section-ni}
+Previous work has noted
 relations between probabilistic noninterference and security
@@
 \subsection{A Sound Hyperproperty for Passive Security}
+\label{section-nimo}
 The intuition that passive security means that the protocol releases
-no more information than what is released by the output suggests a
+no more information than what is released by conditioning on the
+public output and corrupt inputs suggests a
 refinement of noninterference. That is, the set of low equivalent
-initial memories that can produce the same output should generate the
+initial memories \emph{that can produce the same output} should generate the
 same distributions of adversarial views.
 \begin{definition}[Noninterference Modulo Output]
-  Given $H$, $C$, and a program $\prog$ with $\iov(\prog) = (S,V)$ and output
-  view $x \in V$, we say that
-  $\prog$ satisfies \emph{noninterference modulo output} iff for all
-  $\store_1,\store_2 \in \mems(S)$:
+  We say that a program  $\prog$ satisfies \emph{noninterference modulo output},
+  written $\NIMO(\prog)$, iff for all $H$, $C$, and 
+  $\store_1,\store_2 \in \mems(S)$ we have:
   \begin{eqnarray*}
     & \store_1 =_C \store_2 \wedge
-     (\margd{(\condd{\progd(\prog)}{\store_1})}{\{ x \}} =
-     \margd{(\condd{\progd(\prog)}{\store_2})}{\{ x \}}) \\
+     (\margd{(\condd{\progd(\prog)}{\store_1})}{\{ \outv \}} =
+     \margd{(\condd{\progd(\prog)}{\store_2})}{\{ \outv \}}) \\
     & \implies \\
@@
   \end{eqnarray*}
+where $\iov(\prog) = (S,V)$.
 \end{definition}
-We can show, for example, that the 3-party xor protocol defined above
-satisfies noninterference modulo output.
 
-What is the formal connection with passive security? Note in
-particular that the hyperproperty implies that the simulator can use
+From an information flow perspective, we can say that a $\NIMO$
+protocol does not release information beyond what is declassified
+via the public output. From a real/ideal perspective, we can
+observe that $\NIMO$ implies that the simulator can use
 the protocol itself to reconstruct the distribution of adversarial
-views by making a random guess about honest secrets given public
-knowledge about the ideal functionality $\idealf$ a protocol
-implements. However, there is a subtlety here in that the adversary
+views. That is, given the output of the ideal functionality
+$\idealf(\store_S)$, the simulator can randomly choose any
+$\store'$ such that $\store' =_C \store$ and $\idealf(\store'_S) =
+\idealf(\store_S)$ and run the protocol with initial memory $\store'$-
+$\NIMO$ guarantees ideal reconstruction of the real-world corrupt views. 
+ However, there is a subtlety here in that the adversary
 must be able to tractably reconstruct this knowledge. This is not
@@
 together with the assumption that $\idealf$ is invertible, implies
-passive security of a protocol\footnote{We emphasize that the converse
-is not asserted.}. 
+passive security of a protocol.
 \begin{theorem}
@@
   $$
-  \qed
 \end{proof}

### 2023-06-09 popl24/hyperprop.tex
checkpoint

@@
   We say that a program  $\prog$ satisfies \emph{noninterference modulo output},
-  written $\NIMO(\prog)$, iff for all $H$, $C$, and 
+  written $\NIMO(\prog)$, iff for all $H$ and $C$ with $|C|\le|H|$ and 
   $\store_1,\store_2 \in \mems(S)$ we have:
@@
   $$
-  \{ \store | \idealf(\store) = v \}
+  \{ \store\ |\ \idealf(\store) = v \}
   $$
@@
   Assume given invertible $\idealf$ and a protocol $\prog$ that
-  correctly implements $\idealf$.  If program $\prog$ satisfies
-  noninterference modulo output for all $H$ and $C$ with $|H| < |C|$,
+  correctly implements $\idealf$.  If $\NIMO(\prog)$
   then $\prog$ is passive secure.
@@
 
+As for noninterference, we can show that $\NIMO$ is equivalent
+to saying that a given program does not change the probability
+of honest secrets conditioned on corrupt secrets and views
+(including the output view), as compared to conditioning on
+just the corrupt secrets and input. In addition to providing
+insights about $\NIMO$, this form is helpful for
+verification techniques. 
 \begin{lemma}

### 2023-06-09 popl24/hyperprop.tex
checkpoint

@@
 \begin{lemma}
-  Given $H$, $C$, and a program $\prog$ with $\iov(\prog) = (S,V)$ and
-  output view $x \in V$
-  $\prog$ satisfies noninterference modulo output iff for all
+  $\NIMO(\prog)$ iff for all $H$ and $C$ with $|C| \le |H|$ and
   $\store \in \mems(S \cup V)$:
-  $$\margd{(\condd{\progd(\prog)}{\store_{(S \cup \{x\})_C}})}{S_H} =
+  $$\margd{(\condd{\progd(\prog)}{\store_{(S \cup \{\outv\})_C}})}{S_H} =
   \margd{(\condd{\progd(\prog)}{\store_{(S\cup V)_C}})}{S_H} $$
+  where  $\iov(\prog) = (S,V)$.
 \end{lemma}

### 2023-06-09 popl24/bruteforce.tex
checkpoint

@@
 \section{Brute Force Verification}
+\label{section-bruteforce}
 
@@
 g\ \ \stores\ (\eassign{\view{\cid}{w}}{\be}) =\begin{array}[t]{l}
-\elet{\stores'}{\solve{(\mems(\vars\ \be))}{\be}}{}\\
+\elet{\stores'}{\solve{\stores}{\be}}{}\\
   \{\extend{\store}{\view{\cid}{w}}{\etrue} \mid \store \in \stores' \} \cup
@@
 
+\begin{lemma}
+  For all $\stores$ and $\be$ with $\vars(\be) \subseteq \dom(\stores)$,
+  $(\solve{\stores}{\be}) = \{ \store \in \store \ \mid\ \lcod{\store,\be}{\cid} = 1 \}$
+  for some $\cid$.
+\end{lemma}
+
+\begin{lemma}
+  For all $\prog$ where $\iov(\prog) = S \cup V$ and $\flips(\prog) = F$:
+  $$
+  (\mathit{foldl}\ g\ \mems(S \cup F)\ \prog) =
+  \{ \store \in \mems(S\cup F \cup V) \mid \config{\store_{S \cup F}}{\prog} \redxs \config{\store}{\varnothing} \}
+  $$
+\end{lemma}
+
 $$
@@
 $$
+
+\begin{mathpar}
+  \logit{x \mapsto 1} = x
+
+  \logit{x \mapsto 0} = \neg x
+
+  \logit{\{ x_1 \mapsto \beta_1, \ldots, x_n \mapsto \beta_n\}} =
+  \logit{x_1 \mapsto \beta_1} \wedge \cdots \wedge \logit{x_n \mapsto \beta_n}
+
+  \mathit{clauses}(x,\{ \store_1,...,\store_n \}) = x \gets \logit{\store_1} \vee \cdots \vee x \gets \logit{\store_n}
+\end{mathpar}
+
+$$
+\mathit{map}\ \mathit{clauses}\ (\mathit{map}\ f\ \prog)
+$$

### 2023-06-09 popl24/bruteforce.tex
checkpoint

@@
   For all $\stores$ and $\be$ with $\vars(\be) \subseteq \dom(\stores)$,
-  $(\solve{\stores}{\be}) = \{ \store \in \store \ \mid\ \lcod{\store,\be}{\cid} = 1 \}$
+  $(\solve{\stores}{\be}) = \{ \store \in \stores \ \mid\ \lcod{\store,\be}{\cid} = 1 \}$
   for some $\cid$.

### 2023-06-10 popl24/bruteforce.tex
checkpoint

@@
 
+We have shown how our formulation of program distributions, a precise
+definition of intensional protocol behavior, can be used to enforce
+the extensional property of MPC security. This intensional description
+also has the benefit of being flexible- we can use the same mechanisms
+to desribe $\NIMO$, noninterference, perfect secrecy, violations of
+obliviousness, etc. These mechanisms can also be exploited to support
+automated program analysis.
+
+We initially explore a brute force approach to automatically verifying
+program behaviour, the first step of which is to compute the basic
+distribution of a program $\prog$. The program distribution
+$\progd(\prog)$ can obviously be automatically derived from that, as
+can be any of its marginal or conditional distributions. In this
+Section we consider two automated mechanisms for deriving
+$\progd(\prog)$ in this way. The first is through a straightforward
+computation of all final program memories.  The second approach
+rewrites $\minifed$ programs to stratified Datalog programs, that are
+amenable HPC optimizations such as parallelization and GPU matrix
+computations as shown in recent work \cite{XXX}.
+
+The basic distribution of a program $\prog$ is based on the
+final memories resulting from any run of $\prog$, denoted
+$\runs(\prog)$.
+\begin{definition}
+  Given $\prog$ with $\iov(\prog) = S \cup V$ and $\flips(\prog) = F$:
+  $$
+  \runs(\prog) \defeq \{ \store \in \mems(S\cup F \cup V) \mid \config{\store_{S \cup F}}{\prog} \redxs \config{\store}{\varnothing} \}
+  $$
+\end{definition}
+As we've observed, $|\runs(\prog)| = 2^{|S \cup F|}$, so the size of
+$\runs(\prog)$ is exponential in the amount of randomness used in the
+protocol which generally can be used as a proxy for the protocol
+size. In Section \ref{section-composition} we will show how the
+methods described here can be used to individually verify
+compositional properties of small program components, that guarantee
+security properties in larger protocols that combine them. Thus, while
+scalability to program size is a clear challenge for our brute force
+techniques alone, our ultimate goal is to use them for verification of
+compositional properties in larger programs.
+
+\begin{fpfig}[t]{Filtering memories that satisfy a boolean expression.}{fig-solve}
 \begin{eqnarray*}
@@
 \solve{\stores}{(\be_1\ \exor\ \be_2)} &=&
- ((\solve{\stores}{\be_1}) \cap (\stores - \solve{\stores}{\be_2})) \cup
- ((\stores - \solve{\stores}{\be_1}) \cap (\solve{\stores}{\be_2})) \\
+ ((\solve{\stores}{\be_1}) \cap (\stores - \solve{\stores}{\be_2})) \cup\\
+ && ((\stores - \solve{\stores}{\be_1}) \cap (\solve{\stores}{\be_2})) \\
 \solve{\stores}{\select{\be_1}{\be_2}{\be_3}} &=&
- ((\solve{\stores}{\be_1}) \cap (\solve{\stores}{\be_2})) \cup
- ((\stores - \solve{\stores}{\be_1}) \cap (\solve{\stores}{\be_3})) \\
+ ((\solve{\stores}{\be_1}) \cap (\solve{\stores}{\be_2})) \cup \\
+ && ((\stores - \solve{\stores}{\be_1}) \cap (\solve{\stores}{\be_3})) \\
 \solve{\stores}{\OT{\be_1}{\be_2}{\be_3}} &=&
- ((\solve{\stores}{\be_1}) \cap (\solve{\stores}{\be_2})) \cup
- ((\stores - \solve{\stores}{\be_1}) \cap (\solve{\stores}{\be_3}))
+ ((\solve{\stores}{\be_1}) \cap (\solve{\stores}{\be_2})) \cup\\
+ && ((\stores - \solve{\stores}{\be_1}) \cap (\solve{\stores}{\be_3}))
 \end{eqnarray*}
+\end{fpfig}
+
+\subsection{Enumerating $\runs(\prog)$}
 
@@
 
+\begin{lemma}
+  Given $\prog$ where $\iov(\prog) = S \cup V$ and $\flips(\prog) = F$. Define:
+  $$
+  \pdf{S \cup F \cup V}(\store) = \begin{cases}1/2^{|S \cup F|} & \text{if\ } \store \in
+    (\mathit{foldl}\ g\ \mems(S \cup F)\ \prog) \\ 0 & \text{otherwise} \end{cases} 
+  $$
+  Then:
+  $$
+  \margd{\pdf{S \cup F \cup V}}{S \cup V} = \progd(\prog)
+  $$
+\end{lemma}
+
 $$

### 2023-06-10 popl24/bruteforce.tex
checkpoint

@@
 
-\subsection{Enumerating $\runs(\prog)$}
+\subsection{Method 1: Enumerating $\runs(\prog)$}
 
-$$
-g\ \ \stores\ (\eassign{\view{\cid}{w}}{\be}) =\begin{array}[t]{l}
-\elet{\stores'}{\solve{\stores}{\be}}{}\\
-  \{\extend{\store}{\view{\cid}{w}}{\etrue} \mid \store \in \stores' \} \cup
-  \{\extend{\store}{\view{\cid}{w}}{\efalse} \mid \store \in \stores - \stores' \}\end{array}
-$$
+Here we describe a straightforward technique for obtaining program
+distributions $\progd(\prog)$ through direct computation of
+$\runs(\prog)$. An implementation of these techniques along with
+Examples from Sections \ref{section-minicat-examples} and
+\ref{section-metalang-ygc} is available online \cite{XXX}.
 
-$$
-\mathit{foldl}\ g\ \mems(S \cup F)\ \prog
-$$
-
+In Figure \ref{fig-solve} we define the algorithm $\solve{\stores}{\be}$
+which filters $\stores$ based on satisfaction of $\be$. This algorithm
+is fundamental to our brute force methods. It's correctness is characterized
+as follows. 
 \begin{lemma}
@@
 \end{lemma}
+\begin{proof}
+  Since we assume safety of programs we can assume that all variables in $\be$ have the
+  same owner $\cid$. The result otherwise follows in a straightforward manner by induction
+  on $\be$. 
+\end{proof}
 
 \begin{lemma}
-  For all $\prog$ where $\iov(\prog) = S \cup V$ and $\flips(\prog) = F$:
-  $$
-  (\mathit{foldl}\ g\ \mems(S \cup F)\ \prog) =
-  \{ \store \in \mems(S\cup F \cup V) \mid \config{\store_{S \cup F}}{\prog} \redxs \config{\store}{\varnothing} \}
-  $$
+  Given $\prog$ where $\iov(\prog) = S \cup V$ and $\flips(\prog) = F$. Define:
+  \begin{eqnarray*}
+    {tt}\ \ \stores\ (\eassign{\view{\cid}{w}}{\be}) &\defeq& \begin{array}{l}
+      \mathrm{let}\ \stores' = \solve{\stores}{\be} \ \mathrm{in}\\
+      \ \ \{\extend{\store}{\view{\cid}{w}}{\etrue} \mid \store \in \stores' \}\ \cup\\
+      \ \ \{\extend{\store}{\view{\cid}{w}}{\efalse} \mid \store \in \stores - \stores' \}\end{array}\\[2mm]
+    \cruns(\prog) &\defeq& \mathit{foldl}\ {tt}\ \mems(S \cup F)\ \prog
+  \end{eqnarray*}
+  Then $\cruns(\prog) = \runs(\prog)$.
 \end{lemma}
@@
   \pdf{S \cup F \cup V}(\store) = \begin{cases}1/2^{|S \cup F|} & \text{if\ } \store \in
-    (\mathit{foldl}\ g\ \mems(S \cup F)\ \prog) \\ 0 & \text{otherwise} \end{cases} 
+    \cruns(\prog) \\ 0 & \text{otherwise} \end{cases} 
   $$
@@
 \end{lemma}
+
+\subsection{Method 2: Rewriting to Datalog}
 

### 2023-06-10 popl24/minicat.tex
checkpoint

@@
 $\prog$ is $\mems(\flips(\prog))$. We let $\stores$ range
-over sets of memories.
+over sets of memories with the same domain, and abusing notation
+we write $\dom(\stores)$ to denote the common domain.
 
@@
 $\config{\store}{\prog} \redxs \config{\store'}{\varnothing}$ it is
-the case that $\store' \in \mems(S \cup F \cup V)$.
+the case that $\store' \in \mems(S \cup F \cup V)$. We define a
+program as \emph{safe} iff for all $\store \in \mems(S \cup F)$ there
+exists $\store'$ where $\config{\store}{\prog} \redxs
+\config{\store'}{\varnothing}$. Unsafe programs are those where views
+are used before they're defined, or where expressions mix owners of
+variables. As a sanity condition we will only consider safe programs
+in our presentation, though in Section \ref{section-composition} we
+define a type system that guarantees program safety.
 
@@
 \begin{verbatimtab}  
-  view[2, mc] := OT(s[2, c], s[1, m1], s[1, m0])
+  v[2, mc] := OT(s[2, c], s[1, m1], s[1, m0])
 \end{verbatimtab}

### 2023-06-10 popl24/bruteforce.tex
checkpoint

@@
 \begin{lemma}
+  \label{lemma-solves}
   For all $\stores$ and $\be$ with $\vars(\be) \subseteq \dom(\stores)$,
@@
 
+In order to automatically obtain $\runs(\prog)$, we can do a left
+folding of $\solve$ across $\prog$, inductively extending memories
+with valid view assignments in the order of their definition.
+We can then use this to compute $\progd(\prog)$. We denote
+these computations as $\cruns$ and $\progd$ respectively.
 \begin{lemma}
+  \label{lemma-cruns}
   Given $\prog$ where $\iov(\prog) = S \cup V$ and $\flips(\prog) = F$. Define:
@@
       \mathrm{let}\ \stores' = \solve{\stores}{\be} \ \mathrm{in}\\
-      \ \ \{\extend{\store}{\view{\cid}{w}}{\etrue} \mid \store \in \stores' \}\ \cup\\
-      \ \ \{\extend{\store}{\view{\cid}{w}}{\efalse} \mid \store \in \stores - \stores' \}\end{array}\\[2mm]
+      \ \ \{\extend{\store}{\view{\cid}{w}}{1} \mid \store \in \stores' \}\ \cup\\
+      \ \ \{\extend{\store}{\view{\cid}{w}}{0} \mid \store \in \stores - \stores' \}\end{array}\\[2mm]
     \cruns(\prog) &\defeq& \mathit{foldl}\ {tt}\ \mems(S \cup F)\ \prog
@@
 \end{lemma}
-
+\begin{proof}
+  By Lemma \ref{lemma-solves} and induction on the length of $\prog$. 
+\end{proof}
 \begin{lemma}
-  Given $\prog$ where $\iov(\prog) = S \cup V$ and $\flips(\prog) = F$. Define:
+  \label{lemma-cprogd}
+  Given $\prog$ where $\iov(\prog) = S \cup V$ and $\flips(\prog) = F$. Define
+  $\cprogd(\prog)$ as the distribution $\margd{\pdf{S \cup F \cup V}}{S\cup V}$,
+  where:
   $$
@@
   $$
-  \margd{\pdf{S \cup F \cup V}}{S \cup V} = \progd(\prog)
+  \cprogd(\prog) = \progd(\prog)
   $$
 \end{lemma}
+\begin{proof}
+  Immediately by Lemma \ref{lemma-cruns} and Definition \ref{def-progd}.
+\end{proof}
+
+We have used the code for $\cruns$ and $\cprogd$ extensively to design and
+verify the examples in this paper. We will discuss other examples in
+Section \ref{section-composition}, but here we note that we've verified
+passive security of Examples \ref{example-he} and \ref{example-OT}-
+for the former this is well-known so the result is a simple proof-of-concept,
+but the latter is a new (though unsurprising) result, having not been
+directly proven in \cite{XXX}. The method for
+verifying $\NIMO$ described here generalizes to any $\prog$. 
+\begin{lemma}[Automated Verification of Passive Security]
+  $\prog_{\ref{example-he}}$ and $\prog_{\ref{example-OT}};\eassign{\outv}{\vx{2}{mc}}$ are
+  passive secure.
+\end{lemma}
+\begin{proof}
+  For both $\prog \in \{
+  \prog_{\ref{example-he}},\prog_{\ref{example-OT}};\eassign{\outv}{\vx{2}{mc}}\}$
+  we compute $\cruns(\prog)$ and $\cprogd(\prog)$ and then verify
+  $\NIMO(\prog)$ using the following algorithm. Given $\iov(\prog) = S
+  \cup V$ and $\flips(\prog) = F$, we enumerate all memories $\store
+  \in \cruns(\prog)$, then enumerate all partitions of clients into
+  sets $H$,$C$ such that $|C| \le |H|$ and $0 \in C$, and then check:
+  $$
+  \cprogd(\prog)(\store_{S_H}|\store_{S_C \cup \{\outv\}}) =
+  \cprogd(\prog)(\store_{S_H}|\store_{S_C \cup V_C})
+  $$
+  for each $H,C,\store$. By Lemma \ref{lemma-cprogd} and Theorem \ref{theorem-nimo}
+  the result follows.
+\end{proof}
 

### 2023-06-10 popl24/hyperprop.tex
checkpoint

@@
 \begin{theorem}
+  \label{theorem-nimo}
   Assume given invertible $\idealf$ and a protocol $\prog$ that
@@
 
-  In the ideal world, the simulator is given both $\store_{S_C}$ and
+  The simulator is given both $\store_{S_C}$ and
   $\idealf(\store_S)$.  The simulation $\SIM_C(\store_{S_C},

### 2023-06-10 popl24/minicat.tex
checkpoint

@@
 \begin{example}[Oblivious Transfer]
-  \label{example-lambda-OT}
+  \label{example-OT}
   In \cite{XXX} a protocol for $\ttt{OT}$ is defined that we recreate as

### 2023-06-10 popl24/hyperprop.tex
checkpoint

@@
 \begin{lemma}
+  \label{lemma-nimo}
   $\NIMO(\prog)$ iff for all $H$ and $C$ with $|C| \le |H|$ and

### 2023-06-10 popl24/bruteforce.tex
checkpoint

@@
 distributions $\progd(\prog)$ through direct computation of
-$\runs(\prog)$. An implementation of these techniques along with
-Examples from Sections \ref{section-minicat-examples} and
-\ref{section-metalang-ygc} is available online \cite{XXX}.
+$\runs(\prog)$. Although we could obtain $\runs(\prog)$ simply by
+executing $\prog$ given all random tapes and input secrets, the method
+we describe here is empirically more efficient (though still
+exponential in the length of the random tape), and also supports the
+translation to Datalog described in Section
+\ref{section-bruteforce-datalog}. An implementation of these
+techniques along with Examples from Sections
+\ref{section-minicat-examples} and \ref{section-metalang-ygc} is
+available online \cite{XXX}.
 
@@
   $$
-  for each $H,C,\store$. By Lemma \ref{lemma-cprogd} and Theorem \ref{theorem-nimo}
-  the result follows.
+  for each $H,C,\store$. The result follows by Lemma \ref{lemma-cprogd}, Lemma \ref{lemma-nimo}, and
+  Theorem \ref{theorem-nimo}.
 \end{proof}
@@
 \subsection{Method 2: Rewriting to Datalog}
+\label{section-bruteforce-datalog}
 
 $$
-f(\eassign{\view{\cid}{w}}{\be}) = (\view{\cid}{w}, (\solve{(\mems(\vars\ \be))}{\be}))
-$$
-
-$$
-\mathit{map}\ f\ \prog
+{vtt}(\eassign{\view{\cid}{w}}{\be}) = (\view{\cid}{w}, (\solve{(\mems(\vars\ \be))}{\be}))
 $$
@@
 $$
-\mathit{map}\ \mathit{clauses}\ (\mathit{map}\ f\ \prog)
+\datalog(\prog) \defeq  \mathit{map}\ \mathit{clauses}\ (\mathit{map}\ {vtt}\ \prog)
 $$
+
+For all $\prog$, $\datalog(\prog)$ is a \emph{normal}, \emph{stratified}
+program \cite{aspis2018linear}, and $\store$ is the unique least Herbrand
+model of $\mathit{facts}(\store_{S \cup F}),\datalog(\prog)$
+iff $\store \in \runs(\prog)$.

### 2023-06-11 popl24/ygc.tex
checkpoint

\begin{fpfig}[t]{Yao's Garbled Circuits, auxiliary functions.}{fig-ygc-aux}
{\footnotesize
\begin{verbatimtab}
  keygen(gid, b1, b2) { select4(b1,b2,H[gid || "1"],H[gid || "2"],H[gid || "3"],H[gid || "4"]) }

  keysgen(gid, b1, b2)
  {
    let k11 = keygen(gid,b1,b2) in
    let k10 = keygen(gid,b1,not b2) in
    let k01 = keygen(gid,not b1,b2) in
    let k00 = keygen(gid,not b1,not b2) in
    {k11 = k11; k10 = k10; k01 = k01; k00 = k00}
  }

  andtable(keys, bt, ap, bp)
  {
    let r11 = (keys.k11 xor bt) in 
    let r10 = (keys.k10 xor (not bt)) in
    let r01 = (keys.k01 xor (not bt)) in
    let r00 = (keys.k00 xor (not bt)) in
    permute4(ap,bp,r11,r10,r01,r00)
  }

  sharetable(gid, vid, table)
  {   
    v[1, "gate:" || gid || vid || "1"] := table.v1;
    v[1, "gate:" || gid || vid || "2"] := table.v2;
    v[1, "gate:" || gid || vid || "3"] := table.v3;
    v[1, "gate:" || gid || vid || "4"] := table.v4
  }
\end{verbatimtab}
}
\end{fpfig}

\begin{fpfig}[t]{Yao's Garbled Circuits, garbled gates and evaluation code.}{fig-ygc-gates}
{\footnotesize
\begin{verbatimtab}
  garbledecode(wl)
  {
    let r1 = wl.k xor true in
    let r0 = (not wl.k) xor false in
    v[1,"OUTtt1"] := select[wl.p,r1,r0];
    v[1,"OUTtt2"] := select[not wl.p,r1,r0]
  }

  evaldecode(wl, p) { wl.k xor select[wl.p,v[1,"OUTtt1"],v[1,"OUTtt2"]] }

  evalgate(gid, wla, wlb)
  {
    let k = keygen(gid,wla.k,wlb.k) in
    let ct = select4(wla.p,wlb.p,
               v[1,gid || "tt1"],v[1,gid || "tt2"],v[1,gid || "tt3"],v[1,gid || "tt4"]) in
    let cp = select4(wla.p,wlb.p,
               v[1,gid || "pt1"],v[1,gid || "pt2"],v[1,gid || "pt3"],v[1,gid || "pt4"]) in
    { k = k xor ct; p = k xor cp }
  }

  andgate(gid, wla, wlb, wlc) 
  {
    let keys = keysgen(gid,wla.k,wlb.k) in
    sharetable(gid,"tt",andtable(keys,wlc.k,wla.p,wlb.p));
    sharetable(gid,"pt",andtable(keys,wlc.p,wla.p,wlb.p))
  }

  encode(gid, wla,wlb)
  {
    let wla = { k = flip[2,"fwl1"]; p = flip[2,"pwl1"] } in
    let wlb = { k = flip[2,"fwl2"]; p = flip[2,"pwl2"] } in
    { wv1 = { k = OT[s[1,"0"],wla.k,(not wla.k)]; p = OT[s[1,"0"],wla.p,(not wla.p)]}; 
      wv2 = { k = select[s[2,"0"],wlb.k,(not wlb.k)]; p = select[s[2,"0"],wlb.p,(not wlb.p)] } }
  }
\end{verbatimtab}
}
\end{fpfig}

%  andtable(keys, bt, ap, bp)
%  {
%    let r11 = (keys.k11 xor bt) in 
%    let r10 = (keys.k10 xor (not bt)) in
%    let r01 = (keys.k01 xor (not bt)) in
%    let r00 = (keys.k00 xor (not bt)) in
%    permute4(ap,bp,r11,r10,r01,r00)
%  }

### 2023-06-11 popl24/bruteforce.tex
checkpoint

@@
 
+Here we show how to rewrite any $\prog$ to an equivalent Datalog
+program, which supports application of recent work in linear algebraic
+interpretation of Datalog and optimizations of model computation on
+high performance computers \cite{XXX}. The method here also enumerates
+$\runs(\prog)$ memory-by-memory, rather than in a ``batched'' manner
+as in our first method, allowing parallelization of model
+computation. The rewriting we describe here is to Datalog with
+negation, with a negation-as-failure model. Here is the syntax,
+where \emph{atoms} are $\minicat$ variables, \emph{literals}
+are atoms or negated atoms, and clause bodies are lists of
+literals that we interpret as conjoined. A Datalog program
+is a list of clauses:
+$$
+\begin{array}{rclr}
+  \alpha &::=& \view{\cid}{w} \mid  \secret{\cid}{w} \mid \flip{\cid}{w} \mid \oracle{w} & \qquad \textit{(atoms)}\\
+  \mathit{body} &::=&  \alpha,\mathit{body} \mid \neg\alpha,\mathit{body} \mid \varnothing \\
+  \mathit{clause} &::=& \alpha \gets \mathit{body}
+\end{array}
+$$
+We can use techniques in \cite{XXX} to also eliminate negation from
+resulting programs, this presentation is just to summarize our
+fundamental approach.
+
 $$
@@
 \begin{mathpar}
-  \logit{x \mapsto 1} = x
+  \logit{\alpha \mapsto 1} = \alpha
 
-  \logit{x \mapsto 0} = \neg x
+  \logit{\alpha \mapsto 0} = \neg \alpha
 
-  \logit{\{ x_1 \mapsto \beta_1, \ldots, x_n \mapsto \beta_n\}} =
-  \logit{x_1 \mapsto \beta_1} \wedge \cdots \wedge \logit{x_n \mapsto \beta_n}
+  \logit{\{ \alpha_1 \mapsto \beta_1, \ldots, \alpha_n \mapsto \beta_n\}} =
+  \logit{\alpha_1 \mapsto \beta_1} \wedge \cdots \wedge \logit{\alpha_n \mapsto \beta_n}
 
-  \mathit{clauses}(x,\{ \store_1,...,\store_n \}) = x \gets \logit{\store_1} \vee \cdots \vee x \gets \logit{\store_n}
+  \mathit{clauses}(\alpha,\{ \store_1,...,\store_n \}) = \alpha \gets \logit{\store_1} \vee \cdots \vee \alpha \gets \logit{\store_n}
 \end{mathpar}

### 2023-06-11 popl24/bruteforce.tex
checkpoint

@@
 
+The first step in converting a protocol $\prot$ to a Datalog
+program is to apply $\solve$ ``locally'' to each view definition
+in $\prog$, obtaining constraints on memories that satisfy each
+view in isolation.  
+\begin{lemma} Let $\vars(\be)$ be the variables in $\be$, and define:
 $$
-{vtt}(\eassign{\view{\cid}{w}}{\be}) = (\view{\cid}{w}, (\solve{(\mems(\vars\ \be))}{\be}))
+{vtt}(\eassign{\view{\cid}{w}}{\be}) \defeq (\view{\cid}{w}, (\solve{(\mems(\vars\ \be))}{\be}))
 $$
+Then ${vtt}(\eassign{\view{\cid}{w}}{\be}) = (\view{\cid}{w},\stores)$ where $\stores
+  = \{ \store \in \mems(\vars\ \be) \mid \lcod{\store,\be}{\cid} = 1 \}$ for some $\cid$.
+\end{lemma}
+Given this definition, the mapping of ${vtt}$ across a program
+$\prog$- i.e., $(\mathit{map}\ {vtt}\ \prog)$- essentially defines the
+logic program for ``view atoms'' modulo syntactic conversion. We can
+accomplish the latter as follows, where $\datalog(\prog)$ defines the
+full conversion.
 

### 2023-06-11 popl24/bruteforce.tex
checkpoint

@@
 full conversion.
-
+\begin{definition}
 \begin{mathpar}
@@
 $$
+\end{definition}
 
+\begin{lemma}
 For all $\prog$, $\datalog(\prog)$ is a \emph{normal}, \emph{stratified}
@@
 iff $\store \in \runs(\prog)$.
+\end{lemma}

### 2023-06-11 popl24/bruteforce.tex
checkpoint

--- 2023-06-11 popl24/bruteforce.tex
@@
 where \emph{atoms} are $\minicat$ variables, \emph{literals}
-are atoms or negated atoms, and clause bodies are lists of
-literals that we interpret as conjoined. A Datalog program
-is a list of clauses:
+are atoms or negated atoms, and clause bodies are conjunctions of literals.
+A \emph{Datalog program} is a list of clauses.
 $$
@@
   \alpha &::=& \view{\cid}{w} \mid  \secret{\cid}{w} \mid \flip{\cid}{w} \mid \oracle{w} & \qquad \textit{(atoms)}\\
-  \mathit{body} &::=&  \alpha,\mathit{body} \mid \neg\alpha,\mathit{body} \mid \varnothing \\
+  \mathit{body} &::=&  \alpha \mid \neg \alpha \mid alpha \wedge \mathit{body} \mid \neg\alpha \wedge \mathit{body} \mid \varnothing \\
   \mathit{clause} &::=& \alpha \gets \mathit{body}
@@
 full conversion.
-
+\begin{definition} We define the conversion from memories to
+  literals and clause bodies as follows:
 \begin{mathpar}
@@
   \logit{\alpha_1 \mapsto \beta_1} \wedge \cdots \wedge \logit{\alpha_n \mapsto \beta_n}
-
+\end{mathpar}
+Given pairs $(\alpha,\stores)$ in the range of ${vtt}$, we define the conversion
+to clauses as follows:
+\begin{mathpar}
   \mathit{clauses}(\alpha,\{ \store_1,...,\store_n \}) = \alpha \gets \logit{\store_1} \vee \cdots \vee \alpha \gets \logit{\store_n}
 \end{mathpar}
-
+The $\minicat$-to-Datalog conversion is then defined as:
 $$
@@
 $$
+\end{definition}
 
-For all $\prog$, $\datalog(\prog)$ is a \emph{normal}, \emph{stratified}
-program \cite{aspis2018linear}, and $\store$ is the unique least Herbrand
-model of $\mathit{facts}(\store_{S \cup F}),\datalog(\prog)$
+In addition to converting view definitions to logic clauses, we also need to convert
+secrets and random tapes. Since we assume given values for these in an arbitrary run of
+the program, we can capture these as a ``fact base'' in our program, where
+a fact is a clause of the form $\alpha \gets \varnothing$ and means that $\alpha$
+is true in any model of the program. 
+\begin{definition}
+  Given $\store$, let $\{\alpha_1,\ldots,\alpha_n \} =
+  \{ \alpha \in \dom(\store) \mid \store(\alpha) = 1 \}$.
+  Then define:
+  $$
+  \mathit{facts}(\store) = \alpha_1 \gets \varnothing, \ldots, \alpha_n \gets \varnothing
+  $$
+\end{definition}
+
+For any safe $\prog$ with $\iov(\prog) = S \cup V$ and $\flips(\prog) = F$, and
+$\store \in \mems(S \cup F)$, it is the case
+that $(\mathit{facts}(\store),\datalog(\prog))$ is a \emph{normal}, \emph{stratified}
+(in fact, non-recursive) Datalog program, and so has a unique Least Herbrand Model
+and is thus amenable to HPC optimization techniques \cite{aspis2018linear}. 
+To compute $\runs(\prog)$, and thus $\progd(\prog)$, we compute
+the Least Herbrand Model $\store$ of $(\mathit{facts}(\store'),\datalog(\prog))$
+for all $\store' \in \mems(S \cup F)$, observing that model computation for
+each $\store'$ can be done in parallel. The following establishes
+correctness of this approach.
+\begin{lemma}
+  For all $\prog$ with $\iov(\prog) = S \cup V$ and $\flips(\prog) = F$,
+  $\datalog(\prog)$ is a \emph{normal}, \emph{stratified}
+  program \cite{aspis2018linear}, and $\store$ is the unique Least Herbrand
+  Model of $(\mathit{facts}(\store_{S \cup F}),\datalog(\prog))$
   iff $\store \in \runs(\prog)$.
+\end{lemma}
+A full empirical exploration of the application of HPC optimizations
+as in \cite{XXX} is beyond the scope of this paper but is a compelling
+topic for future work. The reader is referred to \cite{XXX} for empirical
+results of HPC optimizations in other logic programming contexts.

### 2023-06-11 popl24/bruteforce.tex
checkpoint

@@
 
-\subsection{Method 1: Enumerating $\runs(\prog)$}
+\subsection{Method 1: Inductive Computation of $\runs(\prog)$}
 
@@
 We can then use this to compute $\progd(\prog)$. We denote
-these computations as $\cruns$ and $\progd$ respectively.
+these computations as $\cruns$ and $\cprogd$ respectively.
 \begin{lemma}
@@
 
-\subsection{Method 2: Rewriting to Datalog}
+\subsection{Method 2: Conversion to Datalog}
 \label{section-bruteforce-datalog}
@@
 computation. The rewriting we describe here is to Datalog with
-negation, with a negation-as-failure model. Here is the syntax,
-where \emph{atoms} are $\minicat$ variables, \emph{literals}
+negation, with a negation-as-failure model, though we can use also
+use techniques in \cite{XXX} to also eliminate negation from
+resulting programs. \emph{Atoms} are $\minicat$ variables, \emph{literals}
 are atoms or negated atoms, and clause bodies are conjunctions of literals.
@@
   \alpha &::=& \view{\cid}{w} \mid  \secret{\cid}{w} \mid \flip{\cid}{w} \mid \oracle{w} & \qquad \textit{(atoms)}\\
-  \mathit{body} &::=&  \alpha \mid \neg \alpha \mid alpha \wedge \mathit{body} \mid \neg\alpha \wedge \mathit{body} \mid \varnothing \\
+  \mathit{body} &::=&  \alpha \mid \neg \alpha \mid \alpha \wedge \mathit{body} \mid \neg\alpha \wedge \mathit{body} \mid \varnothing \\
   \mathit{clause} &::=& \alpha \gets \mathit{body}
@@
 $$
-We can use techniques in \cite{XXX} to also eliminate negation from
-resulting programs, this presentation is just to summarize our
-fundamental approach.
 

### 2023-06-11 popl24/hyperprop.tex
checkpoint

@@
 that the simulator is given the information that is necessarily
-available to the adversary by observing a run of the protocl, and if
+available to the adversary by observing a run of the protocol, and if
 the simulator can reconstruct whatever information appears in the real

### 2023-06-11 popl24/bruteforce.tex
checkpoint

@@
 use techniques in \cite{XXX} to also eliminate negation from
-resulting programs. \emph{Atoms} are $\minicat$ variables, \emph{literals}
+resulting programs. \emph{Atoms} are $\minifed$ variables, \emph{literals}
 are atoms or negated atoms, and clause bodies are conjunctions of literals.
@@
 \end{mathpar}
-The $\minicat$-to-Datalog conversion is then defined as:
+The $\minifed$-to-Datalog conversion is then defined as:
 $$

### 2023-06-11 popl24/metalang.tex
checkpoint

@@
 
+\begin{fpfig}[t]{Syntax of $\metaprot$.}{fig-metaprot-syntax}
 $$
@@
 & & x \mid \elet{x}{e}{e} \mid f(e,\ldots,e) \mid \{ \flab = e; \ldots; \flab = e \}
-\mid e.\flab \mid e\ttt{||}e \mid (e) \\[2mm]
+\mid e.\flab \mid e\concat e \mid (e) \\[2mm]
 v &::=& w \mid \cid \mid \be \mid \{ \flab = v;\ldots;\flab = v \} 
@@
 $$
+\end{fpfig}
 
@@
 
+\begin{fpfig}[t]{Evaluation contexts and operational semantics of $\metaprot$.}{fig-metaprot-semantics}
 $$
 \begin{array}{rcl@{\hspace{3mm}}r}
-E &::=& [\,] \mid \enot\ E \mid E\ \bop\ e \mid v\ \bop\ E \mid & \textit{evaluation contexts}\\[1mm]
-& & \flip{E}{e} \mid \secret{E}{e} \mid \view{E}{e} \mid \oracle{E} \mid  \\[1mm]
-& & \flip{\cid}{E} \mid \secret{\cid}{E} \mid \view{\cid}{E} \mid \send{E}{e} \mid \send{\view{\cid}{w}}{E}\\[1mm]
-& & \OT{E}{e}{e} \mid \OT{v}{E}{e} \mid \OT{v}{v}{E} \mid\\[1mm]
-& & \select{E}{e}{e} \mid \select{v}{E}{e} \mid \select{v}{v}{E} \mid\\[1mm]
-& & \elet{x}{E}{e} \mid f(v,\ldots,v,E,e,\ldots,e) \mid \\[1mm]
-& & \{ \flab = v;\ldots;\flab = v;\flab = E;\flab = e;\ldots;\flab = e \} \mid E.\flab \mid E\ttt{||}e \mid v \ttt{||}E
+E &::=& [\,] \mid \enot\ E \mid E\ \bop\ e \mid v\ \bop\ E \mid  \flip{E}{e} \mid \secret{E}{e} \mid \view{E}{e} \mid \oracle{E} \mid  \\[1mm]
+& & \flip{\cid}{E} \mid \secret{\cid}{E} \mid \view{\cid}{E} \mid \send{E}{e} \mid \send{\view{\cid}{w}}{E} \mid \OT{E}{e}{e} \\[1mm]
+& & \mid \OT{v}{E}{e} \mid \OT{v}{v}{E} \mid \select{E}{e}{e} \mid \select{v}{E}{e} \mid \\[1mm]
+& & \select{v}{v}{E} \mid \elet{x}{E}{e} \mid f(v,\ldots,v,E,e,\ldots,e) \mid \\[1mm]
+& & \{ \flab = v;\ldots;\flab = v;\flab = E;\flab = e;\ldots;\flab = e \} \mid E.\flab \mid E\concat e \mid v \concat E
 \end{array}
 $$
-
+\medskip
 $$
@@
  \config{\prog}{v}\\
- \config{\prog}{w_1\ttt{||}w_2} &\redx& \config{\prog}{w_1w_2}\\
+ \config{\prog}{w_1\concat w_2} &\redx& \config{\prog}{w_1w_2}\\
  \config{\prog}{v;e} &\redx& \config{\prog}{e}\\
@@
 $$
+\end{fpfig}
+
+\subsection{An Implementation of Yao's Garbled Circuits}
+\label{section-metalang-ygc}
+
+\input{ygc}

### 2023-06-12 popl24/metalang.tex
checkpoint

@@
 
+Large practical MPC computations are based on much larger protocols
+than the examples we've considered so far. These larger protocols are
+typically based on compositional units. An example of this is Yao's
+Garbled Circuits (YGC), which are composed of so-called garbled gates.
+Languages for defined garbled circuits, beginning with Fairplay \cite{XXX},
+treat gates as compositional units that are wired together by the programmer
+to generate a complete circuit. The $\fedprot$ language is low-level
+and does not include abstractions for defining composable elements.
+
+In this Section we introduce the $\metaprot$ language which includes
+structured data and function definitions, which are sufficiently
+expressive to define composable protocol elements such as garbled
+gates. The $\metaprot$ language is a \emph{metalanguage}- it produces
+$\fedprot$ protocols as a result of computation. That is, $\metaprot$ is
+a high-level language that generates low-level protocol code.
+
+\subsection{Syntax}
+
 \begin{definition}
@@
 \end{definition}
+
+\begin{example} Consider the following function definition,
+  that splits a client's secret into 3 shares as in
+  Example \ref{example-he}. 
+  \begin{verbatimtab}
+    share3(client, secretid)
+    {
+      let s1 = flip[client, share1] in
+      let s2 = flip[client, share2] in
+      let s3 = (s1 xor s2) xor s[client, s: || secretid] in
+      {s1 = s1;s2 = s2;s3 = s3}
+    } \end{verbatimtab}
+  Here is $\metaprot$ program that uses this this function definition:
+  \begin{verbatimtab}
+    let shares = share3(1, mysecret) in
+    v[2,s1] := shares.s2;
+    v[3,s1] := shares.s3 \end{verbatimtab}
+  which generates the following $\minifed$ program:
+  \begin{verbatimtab}
+    v[2,s1] := flip[1, share2];
+    v[3,s1] := flip[1, share1] xor flip[1, share2] xor s[client, s:mysecret] \end{verbatimtab}
+\end{example}
+
+\subsection{Semantics}
 

### 2023-06-12 popl24/metalang.tex
checkpoint

@@
 
+The syntax of $\metaprot$ is defined in Figure
+\ref{fig-metaprot-syntax}.  It includes a syntax of function
+definitions and records, and values include client ids, identifier
+strings, and boolean expressions.  Expression forms allow dynamic
+construction of boolean expression forms and view assignments. When
+$\metaprot$ programs construct a $\fedprot$ assignment, a side effect
+occurs whereby the assignment is added to the end of the $\fedprot$
+program accumulated during evaluation.
+
+Formally, we consider a complete metaprogram to include both a
+codebase and a ``main'' program that uses the codebase. 
 \begin{definition}
@@
 
-\begin{example} Consider the following function definition,
-  that splits a client's secret into 3 shares as in
-  Example \ref{example-he}. 
+When we consider the example of YGC in detail below, our focus will be
+on developing a codebase that can be used to define arbitrary
+circuits, i.e., complete and concrete protocols. Since strings and
+identifiers can be constructed manually, and expressions can occur
+inside assignments and boolean expression forms, function definitions
+can generalize over $\fedprot$-level patterns to obtain composable
+program units. As a simple example, consider 3 party secret
+sharing as illustrated in Example \ref{example-he}. We can
+define a function $\ttt{share3}$ that abstracts the process
+of splitting a given client's secret into 3 separate shares.
+\begin{example} The function $\ttt{share3}$ 
+  splits a client's secret into 3 shares returned as a record
+  with fields $\ttt{s1-3}$:
   \begin{verbatimtab}

### 2023-06-12 popl24/metalang.tex
checkpoint

@@
 of splitting a given client's secret into 3 separate shares.
-\begin{example} The function $\ttt{share3}$ 
+\begin{example} \label{example-share3} The function $\ttt{share3}$ 
   splits a client's secret into 3 shares returned as a record
@@
     v[3,s1] := shares.s3 \end{verbatimtab}
-  which generates the following $\minifed$ program:
+  which generates the following $\minifed$ program, as we formalize in Example \ref{example-share3-eval}
+  below:
   \begin{verbatimtab}
@@
 
+\begin{example}
+  \label{example-share3-eval}
+  Let $\codebase,e_{\ref{example-share3}}$ be the $\metaprot$ program, and let 
+  $\prog_{\ref{example-share3}}$ be the  $\minifed$ program defined
+  in Example \ref{example-share3}. We refer to the latter as ``accumulated''
+  by evaluation in the sense that $\config{\varnothing}{e_{\ref{example-share3}}}
+  \redxs \config{\prog_{\ref{example-share3}}}{\varnothing}$.
+\end{example}
+
 \subsection{An Implementation of Yao's Garbled Circuits}

### 2023-06-12 popl24/metalang.tex
checkpoint

@@
 expressive to define composable protocol elements such as garbled
-gates. The $\metaprot$ language is a \emph{metalanguage}- it produces
-$\fedprot$ protocols as a result of computation. That is, $\metaprot$ is
-a high-level language that generates low-level protocol code.
+gates. The $\metaprot$ language is a \emph{metalanguage}, in the sense
+that it produces $\fedprot$ protocols as a result of computation. That
+is, $\metaprot$ is a high-level language that generates low-level
+protocol code.
 
@@
 
+We define a small-step evaluation aka reduction relation $\redx$ in
+Figure \ref{fig-metaprot-semantics}.  We write $\redxs$ to denote the
+reflexive, transitive closure of $\redx$. Reduction is defined on
+\emph{configurations} which are pairs of the form $\config{\prog}{e}$,
+where $\prog$ is the $\minifed$ program accumulated during evaluation.
+In this definition we write $e[v/x]$ to denote the substitution of $v$
+for free occurences of $x$ in $e$. The rules are mostly standard,
+except why a concrete $\minifed$ assignment is encountered it is added
+to the end of $\prog$.
+
+The rules rely on a definition of \emph{evaluation contexts} $E$
+allowing computation within a larger program context, where $E[e]$
+denotes an expression with $e$ in the hole $[]$ of $E$. Evaluation
+contexts include boolean expression forms, allowing generalization
+and instantiation of compositional program elements.
 \begin{example}

### 2023-06-12 popl24/metalang.tex
checkpoint

@@
 \input{ygc}
+
+In Figures \ref{fig-ygc-aux} and \ref{fig-ygc-gates} we define a
+codebase for garbled circuits. This definition follows the
+\emph{point-and-permute} method described in \cite{XXX} and elsewhere,
+to which the reader is referred for more in-depth discussion.
+In this implementation client 2 is the \emph{garbler} and
+client 1 is the \emph{evaluator}. The garbler builds the garbled
+tables and shares them with the evaluator, who then evaluates
+the gate in an oblivious fashion until the final public output is
+generated through decryption.

### 2023-06-13 popl24/metalang.tex
checkpoint

@@
     v[2,s1] := flip[1, share2];
-    v[3,s1] := flip[1, share1] xor flip[1, share2] xor s[client, s:mysecret] \end{verbatimtab}
+    v[3,s1] := flip[1, share1] xor flip[1, share2] xor s[1, s:mysecret] \end{verbatimtab}
 \end{example}
@@
   \label{example-share3-eval}
-  Let $\codebase,e_{\ref{example-share3}}$ be the $\metaprot$ program, and let 
+  Let $\codebase,e_{\ref{example-share3}}$ be the $\metaprot$ program and let 
   $\prog_{\ref{example-share3}}$ be the  $\minifed$ program defined
   in Example \ref{example-share3}. We refer to the latter as ``accumulated''
-  by evaluation in the sense that $\config{\varnothing}{e_{\ref{example-share3}}}
+  by evaluation of the former in the sense that $\config{\varnothing}{e_{\ref{example-share3}}}
   \redxs \config{\prog_{\ref{example-share3}}}{\varnothing}$.
@@
 generated through decryption.
+
+\begin{verbatimtab}
+  let wla = owl(s1) in
+  let wlb = owl(s2) in
+  let wlc = owl(0) in
+  andgate(0,wla,wlb,wlc);
+  garbledecode(wlc);
+  let secrets = encode(secrets,wla,wlb) in
+  v[0,output] := decode(evalgate(0, secrets.wv1, secrets.wv2))
+\end{verbatimtab}

### 2023-06-13 popl24/ygc.tex
checkpoint

@@
 \begin{verbatimtab}
-  keygen(gid, b1, b2) { select4(b1,b2,H[gid || "1"],H[gid || "2"],H[gid || "3"],H[gid || "4"]) }
+  keygen(gid, b1, b2) { select4(b1,b2,H[gid || 1],H[gid || 2],H[gid || 3],H[gid || 4]) }
 
@@
 
-  sharetable(gid, vid, table)
+  sharetable(gid, tid, table)
   {   
-    v[1, "gate:" || gid || vid || "1"] := table.v1;
-    v[1, "gate:" || gid || vid || "2"] := table.v2;
-    v[1, "gate:" || gid || vid || "3"] := table.v3;
-    v[1, "gate:" || gid || vid || "4"] := table.v4
+    v[1, gate: || gid || tid || 1] := table.v1;
+    v[1, gate: || gid || tid || 2] := table.v2;
+    v[1, gate: || gid || tid || 3] := table.v3;
+    v[1, gate: || gid || tid || 4] := table.v4
   }
+
+  owl(gid) {  { k = flip[2,gate: || gid || .k]; p = flip[2,gate: || gid || .p] }  }
 \end{verbatimtab}
@@
     let r0 = (not wl.k) xor false in
-    v[1,"OUTtt1"] := select[wl.p,r1,r0];
-    v[1,"OUTtt2"] := select[not wl.p,r1,r0]
+    v[1,OUTtt1] := select[wl.p,r1,r0];
+    v[1,OUTtt2] := select[not wl.p,r1,r0]
   }
 
-  evaldecode(wl, p) { wl.k xor select[wl.p,v[1,"OUTtt1"],v[1,"OUTtt2"]] }
+  evaldecode(wl, p) { wl.k xor select[wl.p,v[1,OUTtt1],v[1,OUTtt2]] }
 
@@
     let ct = select4(wla.p,wlb.p,
-               v[1,gid || "tt1"],v[1,gid || "tt2"],v[1,gid || "tt3"],v[1,gid || "tt4"]) in
+               v[1,gid || tt1],v[1,gid || tt2],v[1,gid || tt3],v[1,gid || tt4]) in
     let cp = select4(wla.p,wlb.p,
-               v[1,gid || "pt1"],v[1,gid || "pt2"],v[1,gid || "pt3"],v[1,gid || "pt4"]) in
+               v[1,gid || pt1],v[1,gid || pt2],v[1,gid || pt3],v[1,gid || pt4]) in
     { k = k xor ct; p = k xor cp }
@@
     let keys = keysgen(gid,wla.k,wlb.k) in
-    sharetable(gid,"tt",andtable(keys,wlc.k,wla.p,wlb.p));
-    sharetable(gid,"pt",andtable(keys,wlc.p,wla.p,wlb.p))
+    sharetable(gid,tt,andtable(keys,wlc.k,wla.p,wlb.p));
+    sharetable(gid,pt,andtable(keys,wlc.p,wla.p,wlb.p))
   }
 
-  encode(gid, wla,wlb)
+  encode(sid, owl1, owl2)
   {
-    let wla = { k = flip[2,"fwl1"]; p = flip[2,"pwl1"] } in
-    let wlb = { k = flip[2,"fwl2"]; p = flip[2,"pwl2"] } in
-    { wv1 = { k = OT[s[1,"0"],wla.k,(not wla.k)]; p = OT[s[1,"0"],wla.p,(not wla.p)]}; 
-      wv2 = { k = select[s[2,"0"],wlb.k,(not wlb.k)]; p = select[s[2,"0"],wlb.p,(not wlb.p)] } }
+    v[1,gate: || sid || 1.k] := OT(s[1,0],owl1.k,(not owl1.k));
+    v[1,gate: || sid || 1.p] := OT(s[1,0],owl1.p,(not owl1.p));
+    v[1,gate: || sid || 2.k] := select(s[2,0],owl2.k,(not owl2.k));
+    v[1,gate: || sid || 2.p] := select(s[2,0],owl2.p,(not owl2.p));
+    { wv1 = { k = v[1,gate: || sid || 1.k]; p = v[1,gate: || sid || 1.p] };
+      wv2 = { k = v[1,gate: || sid || 2.k]; p = v[1,gate: || sid || 2.p] } }
   }
@@
 %  }
+
+%  encode(gid, wla,wlb)
+%  {
+%    let wla = { k = flip[2,fwl1]; p = flip[2,pwl1] } in
+%    let wlb = { k = flip[2,fwl2]; p = flip[2,pwl2] } in
+%    
+%    { wv1 = { k = OT[s[1,0],wla.k,(not wla.k)]; p = OT[s[1,0],wla.p,(not wla.p)]}; 
+%      wv2 = { k = select[s[2,0],wlb.k,(not wlb.k)]; p = select[s[2,0],wlb.p,(not wlb.p)] } }
+%  }

### 2023-06-13 popl24/bruteforce.tex
checkpoint

@@
 \begin{lemma}[Automated Verification of Passive Security]
+  \label{lemma-bruteforce-nimo}
   $\prog_{\ref{example-he}}$ and $\prog_{\ref{example-OT}};\eassign{\outv}{\vx{2}{mc}}$ are

### 2023-06-13 popl24/metalang.tex
checkpoint

@@
 
-\subsection{An Implementation of Yao's Garbled Circuits}
+\subsection{A Library for Yao's Garbled Circuits}
 \label{section-metalang-ygc}
@@
 
+\emph{Wire labels} are fundamental to YGC, and essentially represent
+gate output values in an encrypted form. In our definition, wire
+labels are representated by records $\ttt{\{ k = }\beta_1\ttt{; p =
+}\beta_2\ttt{ \}}$, where $\ttt{k}$ is the \emph{key bit} and
+$\ttt{p}$ is the \emph{pointer bit}, and $\beta_1$ and $\beta_2$ are
+flips. Flips in each output wire label are owned by the garbler and
+are unique per gate by definition of their identifying string, and the
+representation of $0$ is the negation of $1$. For example, here is the
+representation of 1 and 0 respectively in the output wire label for a
+hypothetical gate 6:
+\begin{mathpar}
+  \ttt{\{ k = flip[2,gate:6.k]; p =  flip[2,gate:6.p]] \}}
+
+  \ttt{\{ k = not flip[2,gate:6.k]; p =  not flip[2,gate:6.p]] \}}
+\end{mathpar}
+The pointer bits in wire labels are used to select permuted rows in
+table garblings. The key bits are used to identify a unique key for
+table row in each garbled gate. Intuitively, if $\beta_1$ and
+$\beta_2$ are either key or pointer bits encoding 1 on two input wire
+labels to a binary gate, rows and keys in the gate are enumerated in
+the order:
+$$
+\neg\beta_1\neg\beta_2,\ \neg\beta_1\beta_2,\ \beta_1\neg\beta_2,\ \beta_1\beta_2
+$$
+
+In our implementation, gates are wired together using gate
+identifiers, which are strings $w$. Top-level functionality in Figures
+\ref{fig-ygc-aux} and \ref{fig-ygc-gates} includes the following:
+\begin{itemize}
+\item \ttt{andgate}: This defines a subprotocol for the garbler
+  to define a garbled gate $\ttt{gid}$ with input wires from gates
+  $\ttt{ga}$ and $\ttt{gb}$. The garbler generates keys and garbles
+  the rows in YGC fashion, them with client $1$ in
+  views in a standard form. For example, the view for
+  a hypothetical gate 6, row 2 garbled truth table is $\ttt{v[1,gate:6tt2]}$.
+  We note that garbled gates of other binary operators can be obtained with
+  replacement of $\ttt{andtable}$ with appropriate garbled  table definitions. 
+\item \ttt{evalgate}: This defines a subprotocol for the evaluator to
+  evaluate gate $\ttt{gid}$ given input wire values $\ttt{wva}$ and
+  $\ttt{wvb}$.
+\item \ttt{garbledecode} and \ttt{evaldecode}: The former function
+  defines the garbler's protocol for encrypting the circuit
+  output from final gate $\ttt{gid}$, and the latter defines
+  the evaluator's output decryption protocol.
+\item \ttt{encode}: This defines the initial phase of the protocol,
+  where the evaluator receives the wire value from their own
+  secret $\sx{1}{sa}$ via $\ttt{OT}$, and the garbler communicates
+  the wire value for their own secret $\sx{2}{sb}$ directly.
+\end{itemize}
+Here is an example program that uses this library to define
+a ciruit with a single and gate and input secrets $\ttt{s1}$ and
+$\ttt{s2}$ from client's 1 and 2 respectively. 
 \begin{verbatimtab}
-  let wla = owl(s1) in
-  let wlb = owl(s2) in
-  let wlc = owl(0) in
-  andgate(0,wla,wlb,wlc);
-  garbledecode(wlc);
-  let secrets = encode(secrets,wla,wlb) in
+  andgate(0,s1,s2);
+  garbledecode(0);
+  let secrets = encode(s1,s2) in
   v[0,output] := decode(evalgate(0, secrets.wv1, secrets.wv2))
 \end{verbatimtab}
+We have verified passive security of the $\fedprot$ protocol
+generated by this and other small circuits using the
+technique described in Lemma \ref{lemma-bruteforce-nimo}.
+But obviously large circuits with thousands of gates would be
+intractable to verify with this method. In the next Section
+we discuss compositional methods to address this issue.

### 2023-06-13 popl24/ygc.tex
checkpoint

@@
 \begin{verbatimtab}
-  garbledecode(wl)
+  garbledecode(gid)
   {
+    let wl = owl(gid) in
     let r1 = wl.k xor true in
     let r0 = (not wl.k) xor false in
-    v[1,OUTtt1] := select[wl.p,r1,r0];
-    v[1,OUTtt2] := select[not wl.p,r1,r0]
+    v[1,OUTtt1] := select(wl.p,r1,r0);
+    v[1,OUTtt2] := select(not wl.p,r1,r0)
   }
 
-  evaldecode(wl, p) { wl.k xor select[wl.p,v[1,OUTtt1],v[1,OUTtt2]] }
+  evaldecode(wv, p) { wv.k xor select[wv.p,v[1,OUTtt1],v[1,OUTtt2]] }
 
-  evalgate(gid, wla, wlb)
+  evalgate(gid, wva, wvb)
   {
-    let k = keygen(gid,wla.k,wlb.k) in
-    let ct = select4(wla.p,wlb.p,
+    let k = keygen(gid,wva.k,wvb.k) in
+    let ct = select4(wva.p,wvb.p,
                v[1,gid || tt1],v[1,gid || tt2],v[1,gid || tt3],v[1,gid || tt4]) in
-    let cp = select4(wla.p,wlb.p,
+    let cp = select4(wva.p,wvb.p,
                v[1,gid || pt1],v[1,gid || pt2],v[1,gid || pt3],v[1,gid || pt4]) in
@@
 
-  andgate(gid, wla, wlb, wlc) 
+  andgate(gid, ga, gb) 
   {
+    let wla = owl(ga) in
+    let wlb = owl(gb) in
+    let wlc = owl(gid) in
     let keys = keysgen(gid,wla.k,wlb.k) in
@@
 
-  encode(sid, owl1, owl2)
+  encode(sa, sb)
   {
-    v[1,gate: || sid || 1.k] := OT(s[1,0],owl1.k,(not owl1.k));
-    v[1,gate: || sid || 1.p] := OT(s[1,0],owl1.p,(not owl1.p));
-    v[1,gate: || sid || 2.k] := select(s[2,0],owl2.k,(not owl2.k));
-    v[1,gate: || sid || 2.p] := select(s[2,0],owl2.p,(not owl2.p));
-    { wv1 = { k = v[1,gate: || sid || 1.k]; p = v[1,gate: || sid || 1.p] };
-      wv2 = { k = v[1,gate: || sid || 2.k]; p = v[1,gate: || sid || 2.p] } }
+    let owl1 = owl(sa) in
+    let owl2 = owl(sb) in
+    v[1,gate: || sa || 1.k] := OT(s[1,sa],owl1.k,(not owl1.k));
+    v[1,gate: || sa || 1.p] := OT(s[1,sa],owl1.p,(not owl1.p));
+    v[1,gate: || sb || 2.k] := select(s[2,sb],owl2.k,(not owl2.k));
+    v[1,gate: || sb || 2.p] := select(s[2,sb],owl2.p,(not owl2.p));
+    { wv1 = { k = v[1,gate: || sa || 1.k]; p = v[1,gate: || sa || 1.p] };
+      wv2 = { k = v[1,gate: || sb || 2.k]; p = v[1,gate: || sb || 2.p] } }
   }

### 2023-06-13 popl24/composition.tex
checkpoint

@@
 \label{section-composition}
+
+\subsection{A Library for Yao's Garbled Circuits}
+\label{section-metalang-ygc}
+
+\input{ygc}
+
+In Figures \ref{fig-ygc-aux} and \ref{fig-ygc-gates} we define a
+codebase for garbled circuits. This definition follows the
+\emph{point-and-permute} method described in \cite{XXX} and elsewhere,
+to which the reader is referred for more in-depth discussion.
+In this implementation client 2 is the \emph{garbler} and
+client 1 is the \emph{evaluator}. The garbler builds the garbled
+tables and shares them with the evaluator, who then evaluates
+the gate in an oblivious fashion until the final public output is
+generated through decryption.
+
+\emph{Wire labels} are fundamental to YGC, and essentially represent
+gate output values in an encrypted form. In our definition, wire
+labels are representated by records $\ttt{\{ k = }\beta_1\ttt{; p =
+}\beta_2\ttt{ \}}$, where $\ttt{k}$ is the \emph{key bit} and
+$\ttt{p}$ is the \emph{pointer bit}, and $\beta_1$ and $\beta_2$ are
+flips. Flips in each output wire label are owned by the garbler and
+are unique per gate by definition of their identifying string, and the
+representation of $0$ is the negation of $1$. For example, here is the
+representation of 1 and 0 respectively in the output wire label for a
+hypothetical gate 6:
+\begin{mathpar}
+  \ttt{\{ k = flip[2,gate:6.k]; p =  flip[2,gate:6.p]] \}}
+
+  \ttt{\{ k = not flip[2,gate:6.k]; p =  not flip[2,gate:6.p]] \}}
+\end{mathpar}
+The pointer bits in wire labels are used to select permuted rows in
+table garblings. The key bits are used to identify a unique key for
+table row in each garbled gate. Intuitively, if $\beta_1$ and
+$\beta_2$ are either key or pointer bits encoding 1 on two input wire
+labels to a binary gate, rows and keys in the gate are enumerated in
+the order:
+$$
+\neg\beta_1\neg\beta_2,\ \neg\beta_1\beta_2,\ \beta_1\neg\beta_2,\ \beta_1\beta_2
+$$
+
+In our implementation, gates are wired together using gate
+identifiers, which are strings $w$. Top-level functionality in Figures
+\ref{fig-ygc-aux} and \ref{fig-ygc-gates} includes the following:
+\begin{itemize}
+\item \ttt{andgate}: This defines a subprotocol for the garbler
+  to define a garbled gate $\ttt{gid}$ with input wires from gates
+  $\ttt{ga}$ and $\ttt{gb}$. The garbler generates keys and garbles
+  the rows in YGC fashion, them with client $1$ in
+  views in a standard form. For example, the view for
+  a hypothetical gate 6, row 2 garbled truth table is $\ttt{v[1,gate:6tt2]}$.
+  We note that garbled gates of other binary operators can be obtained with
+  replacement of $\ttt{andtable}$ with appropriate garbled  table definitions. 
+\item \ttt{evalgate}: This defines a subprotocol for the evaluator to
+  evaluate gate $\ttt{gid}$ given input wire values $\ttt{wva}$ and
+  $\ttt{wvb}$.
+\item \ttt{garbledecode} and \ttt{evaldecode}: The former function
+  defines the garbler's protocol for encrypting the circuit
+  output from final gate $\ttt{gid}$, and the latter defines
+  the evaluator's output decryption protocol.
+\item \ttt{encode}: This defines the initial phase of the protocol,
+  where the evaluator receives the wire value from their own
+  secret $\sx{1}{sa}$ via $\ttt{OT}$, and the garbler communicates
+  the wire value for their own secret $\sx{2}{sb}$ directly.
+\end{itemize}
+Here is an example program that uses this library to define
+a ciruit with a single and gate and input secrets $\ttt{s1}$ and
+$\ttt{s2}$ from client's 1 and 2 respectively. 
+\begin{verbatimtab}
+  andgate(0,s1,s2);
+  garbledecode(0);
+  let secrets = encode(s1,s2) in
+  v[0,output] := decode(evalgate(0, secrets.wv1, secrets.wv2))
+\end{verbatimtab}
+We have verified passive security of the $\fedprot$ protocol
+generated by this and other small circuits using the
+technique described in Lemma \ref{lemma-bruteforce-nimo}.
+But obviously large circuits with thousands of gates would be
+intractable to verify with this method. In the next Section
+we discuss compositional methods to address this issue.
+
+\subsection{Garbled $\NIMO$ as Pre-~and Post-Conditions}
 

### 2023-06-13 popl24/metalang.tex
checkpoint

--- 2023-06-13 popl24/metalang.tex
@@
 v &::=& w \mid \cid \mid \be \mid \{ \flab = v;\ldots;\flab = v \} 
-\mid (\,) & \textit{values}\\[2mm]
+\mid \ttt{()} & \textit{values}\\[2mm]
 {fn} &::=& f(x,\ldots,x) \{ e \} & \textit{functions}
@@
   by evaluation of the former in the sense that $\config{\varnothing}{e_{\ref{example-share3}}}
-  \redxs \config{\prog_{\ref{example-share3}}}{\varnothing}$.
+  \redxs \config{\prog_{\ref{example-share3}}}{\ttt{()}}$.
 \end{example}
 
-\subsection{An Implementation of Yao's Garbled Circuits}
-\label{section-metalang-ygc}
+\subsection{Type Theory and Static Type Safety}
 
-\input{ygc}
+It is desirable to statically enforce safety of both $\metaprot$
+programs and the safety of the $\fedprot$ programs they
+generate. Although safety of the latter could be enforced
+post-generation by a direct analysis, for large programs this can be
+much more expensive and it is also better to not wast time on
+resource-intensive compilation of programs with known errors
+\cite{XXX}. Some consequences of safety errors, for example accidental
+reuse of one-time pads, can also undermine security.
 
-In Figures \ref{fig-ygc-aux} and \ref{fig-ygc-gates} we define a
-codebase for garbled circuits. This definition follows the
-\emph{point-and-permute} method described in \cite{XXX} and elsewhere,
-to which the reader is referred for more in-depth discussion.
-In this implementation client 2 is the \emph{garbler} and
-client 1 is the \emph{evaluator}. The garbler builds the garbled
-tables and shares them with the evaluator, who then evaluates
-the gate in an oblivious fashion until the final public output is
-generated through decryption.
+The type syntax of $\metaprot$ is defined in Figure
+\ref{fig-metaprot-tsyntax}. It includes a weak form of dependency:
+string types $\stringty{e}$ and client types $\cidty{e}$
+are parameterized by expressions $e$ that precisely reflect the
+type of the value. Boolean expression forms have the type
+$\bet{e}$ indexed by expressions $e$ indicating the client
+id type of the expression.
 
-\begin{verbatimtab}
-  let wla = owl(s1) in
-  let wlb = owl(s2) in
-  let wlc = owl(0) in
-  andgate(0,wla,wlb,wlc);
-  garbledecode(wlc);
-  let secrets = encode(secrets,wla,wlb) in
-  v[0,output] := decode(evalgate(0, secrets.wv1, secrets.wv2))
-\end{verbatimtab}
+\begin{example}
+  Given $\ttt{share3}$ as defined in Example \ref{example-share3} and
+  annotation:
+  $$\tas(\ttt{share3}) =  \ttt{cid(client)}\ *\ \ttt{string(sid)}$$
+  the type of $\ttt{share3}$ is $\tas(\ttt{share3}) \rightarrow \tau,\varnothing$
+  where $\tau$ is:
+  $$
+  \ttt{\{ s1 : bool[client]; s2 : bool[client]; s3 : bool[client] \}}
+  $$
+\end{example}
+
+\begin{fpfig}[t]{Type Syntax of $\metaprot$.}{fig-metaprot-tsyntax}
+$$
+\begin{array}{rcl@{\hspace{2mm}}r}
+\srct &::=& \cidty{e} \mid \stringty{e} \mid \bet{e} \mid  & \gdesc{types}\\ 
+ &&  \{ \flab : \srct;\ldots;\flab : \srct \} \mid \tau * \cdots * \tau \rightarrow \tau,\viewst \\[1mm]
+\viewst  &::=& \view{e}{e};\viewst \mid \varnothing   & \gdesc{view effects}\\[1mm]
+\Gamma &::=& \Gamma; x : \tau \mid \varnothing & \gdesc{type environments}    
+\end{array}
+$$
+\end{fpfig}
+
+\begin{fpfig}[t]{$\metaprot$ type judgement rules.}{fig-metaprot-tjudge}
+\begin{mathpar}
+\inferrule[\TirName{VarT}]
+{}
+{\tjudge{\viewst}{\Gamma}{x}{\Gamma(x)}{\viewst}}
+
+\inferrule[\TirName{CidT}]
+{}
+{\tjudge{\viewst}{\Gamma}{\cid}{\cidty{\cid}}{\viewst}}
+
+\inferrule[\TirName{StringT}]
+{}
+{\tjudge{\viewst}{\Gamma}{w}{\stringty{w}}{\viewst}}
+
+\inferrule[\TirName{ConcatT}]
+{\tjudge{\viewst}{\Gamma}{e_1}{\stringty{e_1'}}{\viewst_1}\\
+\tjudge{\viewst_1}{\Gamma}{e_2}{\stringty{e_2'}}{\viewst_2}
+}
+{\tjudge{\viewst}{\Gamma}{e_1||e_2}{\stringty{e_1' ||e_2'}}{\viewst_2}}
+
+\inferrule[\TirName{BoolT}]
+{}
+{\tjudge{\viewst}{\Gamma}{\etrue}{\bet{\cid}}{\viewst}}
+
+\inferrule[\TirName{OracleT}]
+{\tjudge{\viewst}{\Gamma}{e}{\stringty{e'}}{\viewst'}}
+{\tjudge{\viewst}{\Gamma}{\oracle{e}}{\bet{\cid}}{\viewst'}}
+
+\inferrule[\TirName{SecretT}]
+{\tjudge{\viewst}{\Gamma}{e_1}{\cidty{e_1'}}{\viewst_1}\\
+\tjudge{\viewst_1}{\Gamma}{e_2}{\stringty{e_2'}}{\viewst_2}}
+{\tjudge{\viewst}{\Gamma}{\secret{e_1}{e_2}}{\bet{e_1'}}}
+
+\inferrule[\TirName{AndT}]
+{
+\tjudge{\viewst}{\Gamma}{e_1}{\bet{e}}{\viewst_1}\\
+\tjudge{\viewst_1}{\Gamma}{e_2}{\bet{e}}{\viewst_2}
+}
+{\tjudge{\viewst}{\Gamma}{e_1\ \eand\ e_2}{\bet{e}}{\viewst_2}}
+
+\inferrule[\TirName{AssignT}]
+{
+\tjudge{\viewst}{\Gamma}{e_1}{\cidty{e_1'}}{\viewst_1}\\
+\tjudge{\viewst_1}{\Gamma}{e_2}{\stringty{e_2'}}{\viewst_2}\\
+\tjudge{\viewst_2}{\Gamma}{e_3}{\bet{e_3'}}{\viewst_3}
+}
+{
+\tjudge{\viewst}{\Gamma}{\eassign{\view{e_1}{e_2}}{e_3}}{\unity}{(\viewst_3 ; \view{e_1'}{e_2'} )}
+}
+
+\inferrule[AppT]
+{\Gamma(f) =  
+ \tau_1 * \cdots * \tau_n \rightarrow \tau, \viewst_f \\ 
+ \tjudge{\viewst}{\Gamma}{e_1}{\sigma(\tau_1)}{\viewst_1}
+ \ \cdots\  
+ \tjudge{\viewst_{n-1}}{\Gamma}{e_n}{\sigma(\tau_n)}{\viewst_n}}
+{\tjudge{\viewst}{\Gamma}{f(e_1,\ldots,e_n)}{\sigma(\tau)}{(\viewst_n ; \sigma(\viewst_f))}}
+
+\inferrule[FnT]
+{
+  \codebase(f) = x_1,\ldots,x_n,\ e \\ \tas(f) = \tau_1 * \cdots * \tau_n \\
+  \tjudge{\varnothing}{\Gamma; x_1 : \tau_1; \ldots; x_n : \tau_n}{e}{\tau}{\viewst_f}
+}
+{ \Gamma \vdash f : \tau_1 * \cdots * \tau_n \rightarrow \tau, \viewst_f }
+
+\inferrule[ProgT]
+{
+\forall f \in \dom(\codebase)\ .\ \Gamma \vdash f : \Gamma(f) \\ \Gamma \vdash e : \tau, \view{\cid_1}{w_1};\ldots;\view{\cid_1}{w_1}
+}
+{
+\Gamma \vdash \codebase,e : \tau,\{\view{\cid_1}{w_1}\} \sqcup \cdots \sqcup \{ \view{\cid_n}{w_n}\}
+}
+\end{mathpar}
+\end{fpfig}
+
+\begin{theorem}
+  \label{theorem-metalang-safety}
+  Given $\codebase$, $e$, and $\Gamma$ with $\Gamma \vdash \codebase,e : \unity : V$,
+  then $\config{\varnothing}{e} \redxs \config{\prog}{\ttt{()}}$ where
+  $\prog$ is safe with $\iov(\prog) = S \cup V$ for some $S$.
+\end{theorem}

### 2023-06-13 popl24/minicat.tex
checkpoint

@@
 We let $X$ range over sets of variables and $S,V,F$ to range over sets
-of secrets, views, and flips respectively. Given a program $\prog$, we
-write $\iov(\prog)$ to denote $S \cup V$ where $S$ is the set of
-secret variables in $\prog$ and $V$ is the set of views in $\prog$,
-and we write $\flips(\prog)$ to denote the set $F$ of flip variables
-in $\prog$. For any set of variables $X$ and parties $P$, we write
-$X_P$ to denote the subset of $X$ owned by any party in $P$, and we
-write $X_H$ and $X_C$ to denote the subsets belonging to honest and
-corrupt parties, respectively.
+of secrets, views, and flips (including invocations of the oracle)
+respectively. Given a program $\prog$, we write $\iov(\prog)$ to
+denote $S \cup V$ where $S$ is the set of secret variables in $\prog$
+and $V$ is the set of views in $\prog$, and we write $\flips(\prog)$
+to denote the set $F$ of flip variables in $\prog$. For any set of
+variables $X$ and parties $P$, we write $X_P$ to denote the subset of
+$X$ owned by any party in $P$, and we write $X_H$ and $X_C$ to denote
+the subsets belonging to honest and corrupt parties, respectively.
 
@@
 are used before they're defined, or where expressions mix owners of
-variables. As a sanity condition we will only consider safe programs
-in our presentation, though in Section \ref{section-composition} we
-define a type system that guarantees program safety.
+variables. As a sanity condition we will only consider safe 
+programs in our presentation of $\fedprot$, though in Section
+\ref{section-metalang} we define a type system that guarantees
+program safety by construction, as demonstrated in Theorem
+\ref{theorem-metalang-safety}.
 

### 2023-06-13 popl24/metalang.tex
checkpoint

@@
 
+Type judgements for expressions are of the form
+$\tjudge{\viewst_1}{\gamma}{e}{\viewst_2}$ where the \emph{view
+effect} $\viewst_1$ denotes the views that have been defined so far,
+and $\viewst_2$ records new views defined as the effect of the
+expression on the residual $\minifed$ program.  Type judgements are
+syntax-directed- selected rules are shown in Figure
+\ref{fig-metaprot-tjudge}. The $\TirName{AssignT}$ rule
+captures the effect of new view definitions. The
+$\TirName{And}$ rule illustrates how program safety is enforced,
+by ensuring that subexpressions of boolean expressions have the
+same owner.
+
+The $\TirName{FnT}$ and $\TirName{Appt}$ rules apply to function
+definition and application respectively, and rely on the
+definition of function input type annotations $\tas$ and
+type term substitutions $\sigma$. 
+\begin{definition}
+  A \emph{function input type annotation} $\tas$ is a mapping from
+  function names $f$ to type products $\tau_1 * \cdots * \tau_n$.
+  A \emph{type term substitution} $\sigma$ is a mapping from
+  $\minifed$ variables $x$ to values, where $\sigma(\tau)$ denotes
+  the replacement of occurences of $x$ in $\tau$ with $\sigma(x)$. 
+\end{definition}
+We assume that input type annotations $\tas$ are provided by the
+programmer for all function definitions. This guarantees that
+$\metaprot$ type checking is straightforward and efficient.
+Function types are of the form:
+$$
+\tau_1 * \cdots * \tau_n \rightarrow \tau,\viewst
+$$
+where $\viewst$ denotes the effect of the function on the residual
+program.  The function type can be understood as a dependent $\Pi$
+type, with every term variable bound. When applied, these variables
+are instantiated with a type term substitution $\sigma$. In our
+implementation, we adapt \emph{synthesis} as defined in MetaML \cite{XXX}
+to obtain this $\sigma$- essentially this is a match on the syntactic
+structure of types and expressions. 
 \begin{example}
@@
 
-\begin{fpfig}[t]{$\metaprot$ type judgement rules.}{fig-metaprot-tjudge}
+\begin{fpfig}[t]{Selected $\metaprot$ type judgement rules.}{fig-metaprot-tjudge}
 \begin{mathpar}
@@
 
-\begin{theorem}
+Our $\metaprot$ type safety result is as follows. In addition to safe execution of the metaprogram,
+it also guarantees safety of the residual $\fedprot$ program. 
+\begin{theorem}[$\metaprot$ Type Safety]
   \label{theorem-metalang-safety}

### 2023-06-13 popl24/ygc.tex
checkpoint

@@
-\begin{fpfig}[t]{Yao's Garbled Circuits, auxiliary functions.}{fig-ygc-aux}
+\begin{fpfig}[t]{Yao's Garbled Circuits, selected auxiliary functions.}{fig-ygc-aux}
 {\footnotesize
@@
 
-  evaldecode(wv, p) { wv.k xor select[wv.p,v[1,OUTtt1],v[1,OUTtt2]] }
+  evaldecode(wv, p) { wv.k xor select(wv.p,v[1,OUTtt1],v[1,OUTtt2]) }
 

### 2023-06-14 popl24/metalang.tex
checkpoint

@@
 $\bet{e}$ indexed by expressions $e$ indicating the client
-id type of the expression.
+id type of the expression. The dependency is weak in the
+sense that expressions in types are a strict subset of
+expression forms- for any $\stringty{e}$ the expression $e$
+is either a variable, a string, or a concatenation form,
+and for any $\cidty{e}$ or $\bet{e}$ the expression $e$
+is either a variable or a client id $\cid$.
 

### 2023-06-14 popl24/metalang.tex
checkpoint

@@
 
-Our $\metaprot$ type safety result is as follows. In addition to safe execution of the metaprogram,
-it also guarantees safety of the residual $\fedprot$ program. 
+Top-level type judgements are of the form $\Gamma \vdash \codebase, e
+: \tau, V$, where all the functions in $\codebase$ are well-typed in
+$\Gamma$, the top level view effect $V$ is a set of concrete
+$\fedprot$ views which are constructed by disjoint union of the views
+in the effect of $e$- the disjointness requirement guarantees that
+views are uniquely defined. Our $\metaprot$ type safety result is
+formulated as follows. In addition to safe execution of the
+metaprogram, it also guarantees safety of the residual $\fedprot$
+program.
 \begin{theorem}[$\metaprot$ Type Safety]

### 2023-06-14 popl24/ygc.tex
checkpoint

@@
 
-  evaldecode(wv, p) { wv.k xor select(wv.p,v[1,OUTtt1],v[1,OUTtt2]) }
+  evaldecode(wv) { wv.k xor select(wv.p,v[1,OUTtt1],v[1,OUTtt2]) }
 
@@
 
+\begin{fpfig}[t]{Yao's Garbled Circuits, gate library type annotations.}{fig-ygc-types}
+{\footnotesize
+\begin{verbatimtab}
+  garbledecode : string(gid)
+
+  evaldecode   : { k = bool[1]; p = bool[1] }
+
+  evalgate     : string(gid) *  { k = bool[1]; p = bool[1] } * { k = bool[1]; p = bool[1] }
+
+  andgate      : string(gid) * string(ga) * string(gb)
+
+  encode       : string(sa) * string(sb)
+\end{verbatimtab}
+}
+\end{fpfig}
 %  andtable(keys, bt, ap, bp)

### 2023-06-15 popl24/composition.tex
checkpoint

@@
 \label{section-composition}
+
+Using YGC as a case study, in this Section we will discuss
+compositional properties that can be automatically verified on small
+program components using techniques developed in Section
+\ref{section-bruteforce}. These properties compose to yield $\NIMO$
+and scale to arbitrary programs. We will express these properties
+manually as pre-~and post-conditions on components, so our
+general approach is in a proof assistance style. The properties
+we use are refinements of $\NIMO$, and the overall technique is
+enabled by the intensional character of our program distribution
+formulation.
 
@@
 the gate in an oblivious fashion until the final public output is
-generated through decryption.
+generated through decryption. This definition is well-typed,
+with input type annotations for top-level functions listed in
+Figure \ref{fig-ygc-types}. Well-typed programs using these
+libraries are therefore guaranteed to yield safe $\minifed$
+programs.
 
@@
 \end{itemize}
-Here is an example program that uses this library to define
+\begin{example}
+  \label{example-andcircuit}
+The following program uses our YGC library to define
 a ciruit with a single and gate and input secrets $\ttt{s1}$ and
@@
 \end{verbatimtab}
+\end{example}
 We have verified passive security of the $\fedprot$ protocol
@@
 
-\subsection{Garbled $\NIMO$ as Pre-~and Post-Conditions}
+\subsection{YGC Gate Pre-~and Post-Conditions}
+
+A basic intuition about how garbled gates work is that they are
+``nearly'' passive secure, modulo the feature that outputs of
+individual logic gates in circuits are encryped. That is, each gate
+enjoys $\NIMO$, except the output is not in the clear, but correlated
+with the flips that the garbler sampled for the gate's output wire
+label, both key and pointer bits. We call this \emph{Garbled $\NIMO$},
+or $\gNIMO$. To clarify this definition, recall that Definition \ref{def-progd}
+posits a variable $\itv$ to refer to the expression $\be$ when
+considering the distribution $\progd(\prog,\be)$.
+\begin{definition}[Garbled $\NIMO$]
+  We write $\gNIMO(\prog,\{ \ttt{k} = \be_{\ttt{k}}; \ttt{p} =
+  \be_{\ttt{p}}\}, g)$ with $\iov(\prog) = S \cup V$ iff for all $H$,
+  $C$ with $|C| \le |H|$, for all $\store \in
+  \mems(S_C \cup V_C)$, and for all $\stores$ with domain $X = S_C \cup V_C
+    \cup \{ \itv, \flip{2}{\ttt{gate:}g.w} \}$ for both $w \in \{\ttt{p},\ttt{k}\}$, such that:
+  $$
+  \stores = \{ \store' \mid \store \subset \store' \wedge 
+  \store(\itv) = \store(\flip{2}{\ttt{gate:}g.w}) \} 
+  $$
+  or
+  $$
+  \stores = \{ \store' \mid \store \subset \store' \wedge 
+  \store(\itv) = \neg\store(\flip{2}{\ttt{gate:}g.w}) \} 
+  $$
+  %where for any $\beta$:
+  %$$\store^2 = \store\{ \flip{2}{\ttt{gate:}g.w} \mapsto \beta \}\{ \itv \mapsto \beta \}$$
+  %or
+  %$$
+  %\store^2 = \store\{ \flip{2}{\ttt{gate:}g.w} \mapsto \beta \}\{ \itv \mapsto \neg\beta \}
+  %$$
+  we have:
+  $$
+  \margd{(\condd{\progd(\prog,\be_w)}{\stores_{X-{V_C}}})}{S_H} =
+  \margd{(\condd{\progd(\prog,\be_w)}{\stores})}{S_H}
+  $$
+\end{definition}
+
+Our aim now is to express that $\gNIMO$ is an invariant preserved throughout
+the circuit by demonstrating pre-~and post-conditions for component execution.
+Our formulation will assume a protocol definition where the garbler and
+evaluator coordinate garbling, sharing, and evaluation of gates iteratively,
+although the generated $\fedprot$ program could be rewritten so the garbler
+shares the entire circuit before execution by the evaluator. As a sanity
+condition we will assume all $\metaprot$ programs we consider in this
+section are well-typed.
+
+First, we show that encoding of input secrets for use by the evaluator
+establishes the $\gNIMO$ invariant.
 
@@
 
-\begin{definition}[Garbled $\NIMO$]
-  We write $\gNIMO(\prog,\{ \ttt{k} = \be_{\ttt{k}}; \ttt{p} = \be_{\ttt{p}}\}, g)$ iff for all $H$, $C$,
-  both $w \in \{\ttt{p},\ttt{k}\}$, and all 
-  memories $\store \in \mems(\iov(\prog) \cup \{ \itv \})$ where
-  $
-  \store(\itv) = \store(\flip{2}{g.w})$ or $\store(\itv) = \neg\store(\flip{2}{g.w})
-  $
-  it is the case that:
-  $$
-  \margd{(\condd{\progd(\prog,\be_w)}{\store_{S_C \cup \{\itv \}}})}{S_H} =
-  \margd{(\condd{\progd(\prog,\be_w)}{\store_{S_C \cup V_C \cup \{\itv \}}})}{S_H}
-  $$
-\end{definition}
-
 \begin{definition}
   We write $\owl(\prog, \{ \ttt{k} = \be_{\ttt{k}}; \ttt{p} = \be_{\ttt{p}}\}, g)$
-  iff for all $b$ and both $w \in \{\ttt{p},\ttt{k}\}$:
+w  iff for all $b$ and both $w \in \{\ttt{p},\ttt{k}\}$:
   $$

### 2023-06-15 popl24/ygc.tex
checkpoint

@@
     let k = keygen(gid,wva.k,wvb.k) in
+    let g = gate: || gid in
     let ct = select4(wva.p,wvb.p,
-               v[1,gid || tt1],v[1,gid || tt2],v[1,gid || tt3],v[1,gid || tt4]) in
+               v[1,g || tt1],v[1,g || tt2],v[1,g || tt3],v[1,g || tt4]) in
     let cp = select4(wva.p,wvb.p,
-               v[1,gid || pt1],v[1,gid || pt2],v[1,gid || pt3],v[1,gid || pt4]) in
+               v[1,g || pt1],v[1,g || pt2],v[1,g || pt3],v[1,g || pt4]) in
     { k = k xor ct; p = k xor cp }
@@
 
-\begin{fpfig}[t]{Yao's Garbled Circuits, gate library type annotations.}{fig-ygc-types}
+\begin{fpfig}[t]{Yao's Garbled Circuits, selected gate library type annotations.}{fig-ygc-types}
 {\footnotesize

### 2023-06-15 popl24/minicat.tex
checkpoint

@@
 $x\not\in\dom(\store)$ to denote the memory $\store'$ such that
-$\store'(x) = \beta$ and otherwise $\store'(y) = \store(y)$ for all $y \in
-\dom(\store)$.
+$\store'(x) = \beta$ and otherwise $\store'(y) = \store(y)$ for all $y
+\in \dom(\store)$. We write $\store_1 \subseteq \store_2$ iff
+$\dom(\store_1) \subseteq \dom(\store_2)$ and $\store_1(x) =
+\store_2(x)$ for all $x \in \dom(\store_1)$. We write $\store_1 \cap
+\store_2$ to denote the combination of $\store_1$ and $\store_2$
+assuming $\store_1(x) = \store_2(x)$ for all $x \in \dom(\store_1)
+\cap \dom(\store_2)$, otherwise $\store_1 \cap \store_2$ is undefined.
 
@@
 over sets of memories with the same domain, and abusing notation
-we write $\dom(\stores)$ to denote the common domain.
+we write $\dom(\stores)$ to denote the common domain,
+and $\stores_X = \{ \store_X | \store \in \stores \}$.
 

### 2023-06-15 popl24/pmf.tex
checkpoint

@@
   (\margd{\pdf{X}}{Y})(\store) =
-  \sum_{\store' \in \mems(X-Y)} \pdf{X}(\store \cup \store')
+  \sum_{\store' \in \mems(X-Y)} \pdf{X}(\store \cap \store')
   $$
@@
   Given $\pdf{X}$, its \emph{conditional distribution given
-  $\store$} with $\dom(\store) = Y$ and $Y \subseteq X$, denoted
-  $\condd{\pdf{X}}{\store}$, is a distribution with domain $X - Y$ where for all
-  $\store' \in \mems(X - Y)$:
+  $\stores$} with $\dom(\stores) = Y$ and $Y \subseteq X$, denoted
+  $\condd{\pdf{X}}{\stores}$, is a distribution with domain $X$ where for all
+  $\store \in \mems(X)$:
   $$
-  (\condd{\pdf{X}}{\store})(\store') =
-      \pdf{X}(\store \cup \store') / ((\margd{\pdf{X}}{Y})(\store))
+  (\condd{\pdf{X}}{\stores})(\store) =
+  (\sum_{\store' \in \stores} \pdf{X}(\store \cap \store')) /
+  (\sum_{\store' \in \stores} \pdf{X}(\store'))
   $$
+  where $\pdf{X}(\store \cap \store')) = 0$ if $\store \cap \store'$ is undefined.
+  we write $\condd{\pdf{X}}{\store}$ as an abbreviation for $\condd{\pdf{X}}{\{ \store \}}$.
 \end{definition}
@@
 \subsection{Examples}
+\label{section-pmf-examples}
 

### 2023-06-15 popl24/minicat.tex
checkpoint

@@
 we write $\dom(\stores)$ to denote the common domain,
-and $\stores_X = \{ \store_X | \store \in \stores \}$.
+and $\stores_X \defeq \{ \store_X | \store \in \stores \}$.
 

### 2023-06-15 popl24/pmf.tex
checkpoint

@@
   Given $\pdf{X}$, its \emph{conditional distribution given
-  $\stores$} with $\dom(\stores) = Y$ and $Y \subseteq X$, denoted
+  $\stores$} with $\dom(\stores) = Y$ with $Y \subseteq X$, denoted
   $\condd{\pdf{X}}{\stores}$, is a distribution with domain $X$ where for all
@@
   (\sum_{\store' \in \stores} \pdf{X}(\store \cap \store')) /
-  (\sum_{\store' \in \stores} \pdf{X}(\store'))
+  (\sum_{\store' \in \stores} (\margd{\pdf{X}}{Y})(\store'))
   $$

### 2023-06-15 popl24/composition.tex
checkpoint

@@
   \be_{\ttt{p}}\}, g)$ with $\iov(\prog) = S \cup V$ iff for all $H$,
-  $C$ with $|C| \le |H|$, for all $\store \in
-  \mems(S_C \cup V_C)$, and for all $\stores$ with domain $X = S_C \cup V_C
-    \cup \{ \itv, \flip{2}{\ttt{gate:}g.w} \}$ for both $w \in \{\ttt{p},\ttt{k}\}$, such that:
+  $C$ with $|C| \le |H|$, for all $\store \in \mems(S_C \cup V_C)$,
+  for both $w \in \{\ttt{p},\ttt{k}\}$, and for all $\stores$ with
+  domain $X = S_C \cup V_C \cup \{ \itv, \flip{2}{\ttt{gate:}g.w} \}$
+  such that:
   $$
   \stores = \{ \store' \mid \store \subset \store' \wedge 
-  \store(\itv) = \store(\flip{2}{\ttt{gate:}g.w}) \} 
+  \store'(\itv) = \store'(\flip{2}{\ttt{gate:}g.w}) \} 
   $$
@@
   \stores = \{ \store' \mid \store \subset \store' \wedge 
-  \store(\itv) = \neg\store(\flip{2}{\ttt{gate:}g.w}) \} 
+  \store'(\itv) = \neg\store'(\flip{2}{\ttt{gate:}g.w}) \} 
   $$
@@
 \end{definition}
+
+\begin{lemma}
+  \label{lemma-scope}
+  If $(\margd{\progd(\prog_1;\prog_2)}{\iov(\prog_1)}) * (\margd{\progd(\prog_1;\prog_2)}{\iov(\prog_2)})$
+  and $\flips(\prog_1) \cap \flips(\prog_2) = \varnothing$ then
+  $(\margd{\progd(\prog_1';\prog_2)}{\iov(\prog_1)}) * (\margd{\progd(\prog_1';\prog_2)}{\iov(\prog_2)})$
+  for any $\prog_1'$ where $\flips(\prog_1') \cap \flips(\prog_2) = \varnothing$.
+\end{lemma}
 

### 2023-06-15 popl24/composition.tex
checkpoint

@@
   \label{lemma-scope}
+  If $(\margd{\progd(\prog_1;\prog_2)}{\iov(\prog_1)}) *
+  (\margd{\progd(\prog_1;\prog_2)}{\iov(\prog_2)})$ and
+  $\flips(\prog_1) \cap \flips(\prog_2) = \varnothing$ then
+  $(\margd{\progd(\prog_1';\prog_2)}{\iov(\prog_1)}) *
+  (\margd{\progd(\prog_1';\prog_2)}{\iov(\prog_2)})$ for any
+  $\prog_1'$ where $\prog_1';\prog_2$ is safe and $\flips(\prog_1')
+  \cap \flips(\prog_2) = \varnothing$.
+\end{lemma}
+
+\begin{definition}
+  We write $g \in \prog$ iff $\view{1}{\ttt{gate:}g\ttt{tt}i} \in \iov(\prog)$
+  and $\view{1}{\ttt{gate:}g\ttt{pt}i} \in \iov(\prog)$ for all $0 < i \le 4$,
+  or  $\view{1}{\ttt{gate:}gi\ttt{.k}} \in \iov(\prog)$
+  and $\view{1}{\ttt{gate:}gi\ttt{.p}} \in \iov(\prog)$ for all $0 < i \le 2$.
+\end{definition}
+
+\begin{lemma}
+  \label{lemma-gnimo}
   If $(\margd{\progd(\prog_1;\prog_2)}{\iov(\prog_1)}) * (\margd{\progd(\prog_1;\prog_2)}{\iov(\prog_2)})$
-  and $\flips(\prog_1) \cap \flips(\prog_2) = \varnothing$ then
-  $(\margd{\progd(\prog_1';\prog_2)}{\iov(\prog_1)}) * (\margd{\progd(\prog_1';\prog_2)}{\iov(\prog_2)})$
-  for any $\prog_1'$ where $\flips(\prog_1') \cap \flips(\prog_2) = \varnothing$.
+  and $\flips(\prog_1) \cap \flips(\prog_2) = \varnothing$ and
+  $\gNIMO(\prog_1;\prog_2,\{ \ttt{k} = \be_{\ttt{k}}; \ttt{p} = \be_{\ttt{p}}\}, g)$ for
+  $g \in \prog_1$,
+  then $\gNIMO(\prog_1';\prog_2,\{ \ttt{k} = \be_{\ttt{k}}; \ttt{p} = \be_{\ttt{p}}\}, g)$
+  for any $\prog_1'$ where $\prog_1';\prog_2$ is safe and
+  $\flips(\prog_1') \cap \flips(\prog_2) = \varnothing$.
+\end{lemma}
+
+\begin{lemma}[Input Encoding]
+  Given well-typed $\prog_1;E[\ttt{encode}(s_1,s_2)]$ where:
+  $$
+  \config{\prog_1}{E[\ttt{encode}(s_1,s_2)]} \redxs
+  \config{\prog_1;\prog_2}{E[\ttt{\{ wv1 =} v_1\ttt{; wv2 = } v_2 \ttt{\}}]}
+  $$
+  Then:
+  $$\gNIMO(\prog_1;\prog_2,v_1,s_1) \wedge \gNIMO(\prog_1;\prog_2,v_2,s_2)$$
 \end{lemma}
@@
   We write $\owl(\prog, \{ \ttt{k} = \be_{\ttt{k}}; \ttt{p} = \be_{\ttt{p}}\}, g)$
-w  iff for all $b$ and both $w \in \{\ttt{p},\ttt{k}\}$:
+ iff for all $b$ and both $w \in \{\ttt{p},\ttt{k}\}$:
   $$

### 2023-06-17 popl24/composition.tex
checkpoint

@@
 technique described in Lemma \ref{lemma-bruteforce-nimo}.
-But obviously large circuits with thousands of gates would be
+But large circuits with thousands of gates would be
 intractable to verify with this method. In the next Section
@@
 
-\subsection{YGC Gate Pre-~and Post-Conditions}
+\subsection{Verification of Pre-~and Post-Conditions}
 
-A basic intuition about how garbled gates work is that they are
-``nearly'' passive secure, modulo the feature that outputs of
-individual logic gates in circuits are encryped. That is, each gate
-enjoys $\NIMO$, except the output is not in the clear, but correlated
-with the flips that the garbler sampled for the gate's output wire
-label, both key and pointer bits. We call this \emph{Garbled $\NIMO$},
-or $\gNIMO$. To clarify this definition, recall that Definition \ref{def-progd}
-posits a variable $\itv$ to refer to the expression $\be$ when
-considering the distribution $\progd(\prog,\be)$.
+A basic intuition about YGC is that garbled gates considered
+individually are ``nearly'' passive secure, modulo the feature that
+outputs of individual logic gates in circuits are encryped. That is,
+each gate enjoys an encrypted variant of$\NIMO$ such that the relevant
+output is not in the clear but correlated flips sampled for the gate's
+output wire label by the garbler, both key and pointer bits. We call
+this \emph{Garbled $\NIMO$}, or $\gNIMO$. In the definition of
+$\gNIMO$ we consider whether a given value $\itv$ correlates
+positively or negatively to the flip chosen for a wire lable, which
+can be either $1$ or $0$ uniformly- recall that
+Definition \ref{def-progd} posits a variable $\itv$ to refer to the
+expression $\be$ when considering the distribution
+$\progd(\prog,\be)$.
 \begin{definition}[Garbled $\NIMO$]
@@
 
+\subsubsection{Verification Witnesses}
+
+Our aim is to demonstrated that $\gNIMO$ is preserved by gates within circuits,
+where preservation is expressed via pre-~and post-conditions. We can use
+techniques in Section \ref{section-bruteforce} to automatically verify postconditions
+for components, using witness programs that satisfy preconditions prior to
+component execution. For example, we can automatically verify that $\ttt{encode}$
+generates a program and wire values the enjoy $\gNIMO$. Then, to verify
+that we can verify
+that the $\minifed$ program and wire value generated by the following program satisfies
+$\gNIMO$:
+\begin{verbatimtab}
+  let secrets = encode(s1,s2) in
+  andgate(0,s1,s2); evalgate(secrets.wv1,secrets.wv2)
+\end{verbatimtab}
+However, we need to establish that $\ttt{encode(s1,s2)}$ is an
+adequate witness to correctness, so our automated verification generalizes to
+arbitrary circuit contexts.
+
+An important part of this argument is probabilistic independence,
+and in related reasoning we can leverage prior work on probabilistic
+separation logic \cite{barthe2019probabilistic}. We will use an
+analog of their separation formula, which we define in our setting as follows:
+\begin{definition}
+  We write $(\margd{\pdf{X}}{Y}) * (\margd{\pdf{X}}{X-Y})$ iff
+  $\pdf{X}(\store) = \pdf{X}(\store_{Y}) * \pdf{X}(\store_{X-Y})$
+  for all $\store \in \mems(X)$.
+\end{definition}
+
+We aim to show that gates are probabilistically independent of
+each other (i.e., are separated), and this independence supports
+correctness-preserving composition. Noting that gates have
+naming conventions, and that gates are distinguished by gate
+identifiers $g$, we can define what we mean by a gate
+$g$ being in a given program $\prog$, written $g \in \prog$. 
+\begin{definition}
+  We write $g \in \prog$ iff $\view{1}{\ttt{gate:}g\ttt{tt}i} \in \iov(\prog)$
+  and $\view{1}{\ttt{gate:}g\ttt{pt}i} \in \iov(\prog)$ for all $0 < i \le 4$,
+  or  $\view{1}{\ttt{gate:}gi\ttt{.k}} \in \iov(\prog)$
+  and $\view{1}{\ttt{gate:}gi\ttt{.p}} \in \iov(\prog)$ for all $0 < i \le 2$.
+\end{definition}
+
+A critical property is that separation is preserved by safe
+(in the sense of Theorem \ref{theorem-metalang-safety}) substitution
+of programs, assuming disjointness of flips used in programs separated
+programs. 
 \begin{lemma}
@@
 \end{lemma}
+\begin{proof}
+  \cnote{The Lemma statement may be slightly off, there may be more restrictions on $\prog_1$.
+  But generally this result follows by results from \cite{barthe2019probabilistic.}}
+\end{proof}
 
-\begin{definition}
-  We write $g \in \prog$ iff $\view{1}{\ttt{gate:}g\ttt{tt}i} \in \iov(\prog)$
-  and $\view{1}{\ttt{gate:}g\ttt{pt}i} \in \iov(\prog)$ for all $0 < i \le 4$,
-  or  $\view{1}{\ttt{gate:}gi\ttt{.k}} \in \iov(\prog)$
-  and $\view{1}{\ttt{gate:}gi\ttt{.p}} \in \iov(\prog)$ for all $0 < i \le 2$.
-\end{definition}
-
+Another important property is that $\gNIMO$ is preserved by substitution
+for a witness program. This follows by Lemma \ref{lemma-scope} and definition
+of $\gNIMO$. 
 \begin{lemma}
@@
 
+\subsubsection{Pre-~ and Post-Conditions}
+\label{section-pre-post}
+Now we can establish relevant pre-~and post-conditions for the
+top-level YGC functions defined in Figure \ref{fig-ygc-gates}. We note
+that in addition to the automated verification mechanisms described in
+Section \ref{section-bruteforce}, type safety in $\metaprot$ plays an
+important role, in part by ensuring that gates are uniquely defined
+(as views).
+
+Our strategy is to consider input encoding, internal gates, and
+output decoding separately. To begin, we show that input encoding
+establishes generates a program and wire values that initialize
+the $\gNIMO$ invariant in the circuit. There is no precondition
+witness needed for this result, which holds in any type-safe
+program context since the variables used by $\ttt{encode}$
+are all fresh. 
 \begin{lemma}[Input Encoding]
+  \label{lemma-encode}
   Given well-typed $\prog_1;E[\ttt{encode}(s_1,s_2)]$ where:
@@
   \config{\prog_1}{E[\ttt{encode}(s_1,s_2)]} \redxs
-  \config{\prog_1;\prog_2}{E[\ttt{\{ wv1 =} v_1\ttt{; wv2 = } v_2 \ttt{\}}]}
+  \config{\prog_1;\prog_2}{E[\ttt{\{wv1=}v_1\ttt{;wv2=}v_2\ttt{\}}]}
   $$
@@
 
-Our aim now is to express that $\gNIMO$ is an invariant preserved throughout
-the circuit by demonstrating pre-~and post-conditions for component execution.
-Our formulation will assume a protocol definition where the garbler and
-evaluator coordinate garbling, sharing, and evaluation of gates iteratively,
-although the generated $\fedprot$ program could be rewritten so the garbler
-shares the entire circuit before execution by the evaluator. As a sanity
-condition we will assume all $\metaprot$ programs we consider in this
-section are well-typed.
+\begin{proof}
+  We have automatically verified
+  $\gNIMO(\prog_2,v_1,\ttt{s1})$ and $\gNIMO(\prog_2,v_1,\ttt{s1})$
+  where:
+  $$
+  \config{\varnothing}{\ttt{encode(s1,s2)}} \redxs
+  \config{\prog_2}{\ttt{\{wv1=}v_1\ttt{;wv2=}v_2\ttt{\}}}
+  $$
+  using the method described in Section \ref{section-bruteforce-tables}.
+  Given arbitrary $\prog_1$ such that $\prog_1;\ttt{encode}(s_1,s_2)$
+  is well-typed, hence $\prog_1;\prog_2$ is safe, 
+  since $\ttt{encode}$ uses unique secrets and flips per gates and
+  gates (views) are not redefined in safe programs,
+  $(\iov(\prog_1) \cup \flips(\prog_1)) \cap (\iov(\prog_2) \cup
+  \flips(\prog_2)) = \varnothing$. This implies 
+  $(\margd{\progd(\prog_1;\prog_2)}{\iov(\prog_1)}) *
+  (\margd{\progd(\prog_1;\prog_2)}{\iov(\prog_2)})$, so the result follows
+  by Lemma \ref{lemma-gnimo}.
+\end{proof}
 
-First, we show that encoding of input secrets for use by the evaluator
-establishes the $\gNIMO$ invariant.
+One nuance in our results for internal gates and for decoding is that the relevant
+compositional units are the complete related subprotocols, including both
+the garbler's construction of the gates and the evaluator's evaluation of
+them. We therefore make the following definitions. However, we note that
+our formal results are applicable to the YGC style where the garbler shares
+the entire circuit prior to evaluation, rather than gate-by-gate during
+evaluation. 
+\begin{definition}
+  \label{ygc-modules}
+  We define a composable internal \ttt{and} gate sharing and evaluation function
+  called $\ttt{andgg}$, and an output decoder gate share and evaluation function called
+  $\ttt{decode}$, as follows. 
+\begin{verbatimtab}
+  andgg(g, ga, gb, wva, wvb) { andgate(g, ga, gb); evalgate(g, wva, wvb) }
 
-\begin{verbatimtab}
-andgg(g, wla, wlb, wlc, wva, wvb) {
-andgate(g, wla, wlb, wlc);
-evalgate(g, wva, wvb)
-}
+  decode(g, v) { garbledecode(g); evaldecode(v) }
 \end{verbatimtab}
-
-\begin{definition}
-  We write $\owl(\prog, \{ \ttt{k} = \be_{\ttt{k}}; \ttt{p} = \be_{\ttt{p}}\}, g)$
- iff for all $b$ and both $w \in \{\ttt{p},\ttt{k}\}$:
-  $$
-  (\margd{(\condd{\progd(\prog,\be_{w})}{\{ \flip{2}{g.w} \mapsto b \}})}{\itv})(\{ \itv \mapsto b\}) = 1
-  $$
 \end{definition}
 
-\begin{lemma}[Input Encoding]
-Assume given the following evaluation relation:
+Now we can demonstrate that our $\ttt{and}$ gate definition preserves
+$\gNIMO$, and additionally that internale gates are probabilistically independent.
+This latter fact is crucial to compositionality, and allows us for automatically
+verify the circuit in isolation, with input encoding serving as an adequate input
+witness for generalization. 
+\begin{lemma}[Composition ($\eand$)]
+  \label{lemma-compose}
+Given well-typed $\prog_1;E[\ttt{andgg}(g,g_1,g_2,v_1,v_2)]$ where:
 $$
-\config{\prog_1}{E[\ttt{encode}(w,v_1,v_2)]} \redxs \config{\prog_1;\prog_2}{E[v]}
+\config{\prog_1}{E[\ttt{andgg}(g,g_1,g_2,v_1,v_2)]} \redxs \config{\prog_1;\prog_2}{E[v]}
 $$
-If there exists gate identifiers $g_1, g_2$ such that:
+If the following preconditions hold:
 \begin{enumerate}
-  \item $g_1 \ne g_2 \wedge g_1,g_2 \not\in \gates(\prog_1)$
-  \item $\owl(\prog_1,v_1,g_1) \wedge \owl(\prog_1,v_2,g_2)$
+\item $g_1 \ne g_2$
+\item $\gNIMO(\prog_1,v_1,g_1) \wedge \gNIMO(\prog_1,v_2,g_2)$ 
 \end{enumerate}
-then:
+then we have as postconditions:
 \begin{enumerate}
-\item $g_1,g_2 \in  \gates(\prog_2)$
-\item $\gNIMO(\prog_1;\prog_2,v.\ttt{wv1},g_1) \wedge \gNIMO(\prog_1;\prog_2,v.\ttt{wv2},g_2)$
+  \item $\gNIMO(\prog_1;\prog_2,v,g)$
   \item $(\margd{\progd(\prog_1;\prog_2)}{\iov(\prog_1)}) * (\margd{\progd(\prog_1;\prog_2)}{\iov(\prog_2)})$
@@
 \end{lemma}
-
-\begin{lemma}[Composition ($\eand$)]
-Assume given the following evaluation relation:
+\begin{proof}
+  We have automatically verified the postconditions (1) and (2) for the particular
+  program $\prog_1;\prog_2$ resulting from execution of:
+  $$e\quad \defeq\quad \ttt{let gs = encode(s1,s2) in andgg(0,s1,s2,gs.wv1,gs.wv2)}$$
+  where:
   $$
-\config{\prog_1}{E[\ttt{andgg}(g,v_1,\ldots,v_5)]} \redxs \config{\prog_1;\prog_2}{E[v_6]}
+  \config{\varnothing}{e} \redxs
+  \config{\prog_1}{\ttt{andgg}(\ttt{0},\ttt{s1},\ttt{s2},v_1,v_2)} \redxs \config{\prog_1;\prog_2}{v}
   $$
-If there exist gate identifiers $g_1, g_2$ such that:
-\begin{enumerate}
-  \item $g_1 \ne g_2 \ne g \wedge g_1,g_2 \in \gates(\prog_1) \wedge g \not\in \gates(\prog_1)$
-  \item $\owl(\prog_1,v_1,g_1) \wedge \owl(\prog_1,v_2,g_2)$
-  \item $\gNIMO(\prog_1,v_4,g_1) \wedge \gNIMO(\prog_1,v_5,g_2)$
-\end{enumerate}
-then:
-\begin{enumerate}
-  \item $g \in \gates(\prog_2) \wedge \owl(\prog_1;\prog_2,v_3,g)$
-  \item $\gNIMO(\prog_1;\prog_2,v_6,g)$
-  \item $(\margd{\progd(\prog_1;\prog_2)}{\iov(\prog_1)}) * (\margd{\progd(\prog_1;\prog_2)}{\iov(\prog_2)})$
-\end{enumerate}
+  By Lemma \ref{lemma-encode} we have that
+  $\gNIMO(\prog_1,v_1,\ttt{s1}) \wedge \gNIMO(\prog_1,v_2,\ttt{s2})$. 
+  And since $\ttt{andgate}$ generates unique flips per gate, and well-typedness ensures
+  uniqueness of gates, we can generalize the result to replace $\prog_1$ with arbitrary $\prog_1'$
+  such that $\prog_1';\prog_2$ is safe,
+  applying Lemma \ref{lemma-scope} to obtain postcondition (2) and Lemma \ref{lemma-gnimo}
+  to obtain postcondition (1). 
+\end{proof}
+Finally, we can demonstrate that decoding correctly casts a garbled output into the
+clear, and thus establishes $\NIMO$ for the garbled circuit that precedes it.
+\begin{lemma}[Decoding]
+  \label{lemma-decode}
+Given well-typed $\prog_1;E[\ttt{decode}(g,v)]$ where
+$$
+\config{\prog_1}{E[\ttt{decode}(g,v)]} \redxs \config{\prog_1;\prog_2}{E[\be]}
+$$
+If $\gNIMO(\prog_1,v,g)$, then $\NIMO(\prog_1;\prog_2;\eassign{\outv}{\be})$.
 \end{lemma}
-
-\begin{lemma}[Decoding]
-Assume given the following evaluation relation:
-$$
-\config{\prog_1}{E[\ttt{decoder}(v_1,v_2)]} \redxs \config{\prog_1;\prog_2}{E[\be]}
-$$
-If there exists gate identifier $g \in \gates(\prog_1)$ such that:
-\begin{enumerate}
-  \item $\owl(\prog_1,v_1,g)$
-  \item $\gNIMO(\prog_1,v_2,g)$
-\end{enumerate}
-Then $\NIMO(\prog_1;\prog_2;\eassign{x}{\be})$.
-\end{lemma}
+\begin{proof}
+  We have automatically verified $
+  \progd(\prog,\be)(\{ \itv \mapsto 1 \}) = 1$ where:
+  $$\config{\varnothing}{\ttt{decode(g,\{ k = flip[2, gate:g.k]; p = flip[2,gate:g.p] \})}}
+    \redxs \config{\prog}{\be}$$
+    and also $
+  \progd(\prog,\be)(\{ \itv \mapsto 0 \}) = 1$ where:
+  $$\config{\varnothing}{\ttt{decode(g,\{ k = not flip[2, gate:g.k]; p = not flip[2,gate:g.p] \})}}
+    \redxs \config{\prog}{\be}$$
+  The result follows by definition of $\gNIMO$ and $\NIMO$.
+\end{proof}

### 2023-06-18 popl24/composition.tex
checkpoint

@@
 \end{proof}
+
+\begin{fpfig}[t]{YGC copy gate definitions.}{fig-ygc-copy}
+  {\footnotesize
+    \begin{verbatimtab}
+      sharetab2(gid, tid, k, p, b)
+      {
+        let r1 = k xor b in
+        let r0 = (not k) xor (not b) in
+        v[1,gate: || gid || tid || 1] := select(p,r1,r0);
+        v[1,gate: || gid || tid || 2] := select(not p,r1,r0);
+      }
+
+      copygate(ca,cb,g)
+      {
+        let wl = owl(g) in
+        let owl1 = owl(ca) in
+        let owl2 = owl(cb) in
+        sharetab2(ca,tt,wl.k,wl.p,owl1.k); sharetab2(ca,pt,wl.k,wl.p,owl1.p);
+        sharetab2(cb,tt,wl.k,wl.p,owl2.k); sharetab2(cb,pt,wl.k,wl.p,owl2.p)
+      }
+
+      evalcopy(ca,cb,wv)
+      {
+        let wv1k = wv.k xor select(wv.p,v[1,gate: || ca || tt1], v[1,gate: || ca || tt2]) in
+        let wv1p = wv.k xor select(wv.p,v[1,gate: || ca || pt1], v[1,gate: || ca || pt1]) in
+        let wv2k = wv.k xor select(wv.p,v[1,gate: || cb || tt1], v[1,gate: || cb || tt2]) in
+        let wv2p = wv.k xor select(wv.p,v[1,gate: || cb || pt1], v[1,gate: || cb || pt1]) in
+        { wv1 = { k = wv1k; p = wv1p }; wv2 = { k = wv2k; p = wv2p } }  
+      }  \end{verbatimtab}
+  }
+\end{fpfig}
+
+\subsection{Building and Extending Circuits: Copy Gates}
+
+Our results in Section \ref{section-pre-post} and Theorem
+\ref{theorem-nimo} imply that any well-typed program using
+$\ttt{encode}$, $\ttt{andgg}$, and $\ttt{decode}$ will generate a
+circuit that is passive secure. To add an $\ttt{or}$ gate to our
+library, we can use $\ttt{andgg}$ as a template, in fact simple
+modification of $\ttt{andtable}$ is all that would be needed. Thus,
+this library embodies a Fairplay-like language where well-typedness
+of metaprograms guarantees safety and security of generated protocols.
+
+The techniques in Section \ref{section-pre-post} also establish an
+analysis that can be used when extending the language. Various
+previous work has proposed different implementation of gates, such as
+optimized $\ttt{xor}$ and $\ttt{and}$ gates \cite{XXX} and our
+verification methods can automatically check their security. For
+example, the naive implementation of copy gates turns out to be
+unsound in YGC \cite{XXX}.
+
+\begin{verbatimtab}
+  let s = encode(s1,s2) in
+  let s1c = copy(c1s1,c2s1,s1,s.wv1) in
+  let s2c = copy(c1s2,c2s2,s2,s.wv2) in
+  let wv1 = andgg(1,c1s1,c1s2,s1c.wv1,s2c.wv1) in
+  let wv2 = andgg(2,c2s1,c2s2,s1c.wv2,s2c.wv2) in
+  decode(3,andgg(3,1,2,wv1,wv2))
+\end{verbatimtab}

### 2023-06-18 popl24/composition.tex
checkpoint

--- 2023-06-17 popl24/composition.tex
@@
 generates a program and wire values the enjoy $\gNIMO$. Then, to verify
-that we can verify
-that the $\minifed$ program and wire value generated by the following program satisfies
-$\gNIMO$:
+correctness of $\ttt{and}$ gate construction and evaluation- i.e., that it
+preserves $\gNIMO$- we can verify
+that the $\minifed$ program and wire value generated by the following program
+$\metaprot$ program satisfies $\gNIMO$:
 \begin{verbatimtab}
@@
 However, we need to establish that $\ttt{encode(s1,s2)}$ is an
-adequate witness to correctness, so our automated verification generalizes to
+adequate witness, in that it ensures that automated verification generalizes to
 arbitrary circuit contexts.
@@
 
-We aim to show that gates are probabilistically independent of
-each other (i.e., are separated), and this independence supports
-correctness-preserving composition. Noting that gates have
-naming conventions, and that gates are distinguished by gate
-identifiers $g$, we can define what we mean by a gate
-$g$ being in a given program $\prog$, written $g \in \prog$. 
 \begin{definition}
-  We write $g \in \prog$ iff $\view{1}{\ttt{gate:}g\ttt{tt}i} \in \iov(\prog)$
-  and $\view{1}{\ttt{gate:}g\ttt{pt}i} \in \iov(\prog)$ for all $0 < i \le 4$,
-  or  $\view{1}{\ttt{gate:}gi\ttt{.k}} \in \iov(\prog)$
-  and $\view{1}{\ttt{gate:}gi\ttt{.p}} \in \iov(\prog)$ for all $0 < i \le 2$.
+  We say that $\prog$ \emph{covers} $X$ iff
+  $\progd(\prog)(\store) > 0$ for all $\store\in\mems(X)$.
 \end{definition}
 
-A critical property is that separation is preserved by safe
-(in the sense of Theorem \ref{theorem-metalang-safety}) substitution
-of programs, assuming disjointness of flips used in programs separated
-programs. 
+We aim to show that gates are probabilistically independent of each
+other (i.e., are separated), and this independence supports
+correctness-preserving composition. A critical property is that
+separation is preserved by safe (in the sense of Theorem
+\ref{theorem-metalang-safety}) substitution of programs, assuming
+disjointness of flips used in programs separated programs.
 \begin{lemma}
   \label{lemma-scope}
+  Let $X = \vars(\prog_1) \cap \vars(\prog_2)$.
   If $(\margd{\progd(\prog_1;\prog_2)}{\iov(\prog_1)}) *
-  (\margd{\progd(\prog_1;\prog_2)}{\iov(\prog_2)})$ and
-  $\flips(\prog_1) \cap \flips(\prog_2) = \varnothing$ then
+  (\margd{\progd(\prog_1;\prog_2)}{\iov(\prog_2)})$ and $\prog_1$
+  covers $X$, then 
   $(\margd{\progd(\prog_1';\prog_2)}{\iov(\prog_1)}) *
   (\margd{\progd(\prog_1';\prog_2)}{\iov(\prog_2)})$ for any
-  $\prog_1'$ where $\prog_1';\prog_2$ is safe and $\flips(\prog_1')
-  \cap \flips(\prog_2) = \varnothing$.
+  $\prog_1'$ such that $\vars(\prog_1') \cap \vars(\prog_2) = X$ and
+  $\prog_1';\prog_2$ is safe.
 \end{lemma}
 \begin{proof}
-  \cnote{The Lemma statement may be slightly off, there may be more restrictions on $\prog_1$.
-  But generally this result follows by results from \cite{barthe2019probabilistic.}}
+  Suppose on the contrary that there exists $\prog_1'$ that is not
+  separable from $\prog_2$. By results in
+  \cite{barthe2019probabilistic}, this implies that there
+  exists $\store^n$ and $\store^n$ for $0 \le n \le 1$ such that
+  $\store^1_X = \store^{1'}_X$ where:
+  $$
+  \config{\store^0}{\prog_1;\prog_2} \redxs \config{\store^1}{\prog_2} \redxs
+  \config{\store^2}{\varnothing}
+  $$
+  and:
+  $$
+  \config{\store^{0'}}{\prog_1;\prog_2} \redxs \config{\store^{1'}}{\prog_2} \redxs
+  \config{\store^{2'}}{\varnothing}
+  $$
+  where $\store^2_{\vars(\prog_2)} \ne \store^{2'}_{\vars(\prog_2)}$, which is a
+  contradiction.
+%   This implies that there is some
+%  $\store \in \mems(\vars(\prog_1';\prog_2))$ such that
+%  $\store_{\vars{\prog_{1}'}} \in \runs{\prog_1'}$ and
+%  \mems(\vars(\prog_1))$ with   
 \end{proof}
 
-Another important property is that $\gNIMO$ is preserved by substitution
-for a witness program. This follows by Lemma \ref{lemma-scope} and definition
-of $\gNIMO$. 
+We must also choose witnesses that allow us to generalize verification
+of $\gNIMO$ preservation through gates. We need to consider separately
+the definition of gates and their use. Our type discipline ensures that
+gates can only be defined once, but it is also known that gate output
+wire labels should only be used once \cite{XXX}, in which case we
+say they are \emph{wired}. We return to this issue in more detail in
+Section \ref{section-composition-copy}.
+\begin{definition}
+  We write $g \in \prog$ iff all views for $g$'s table are fully
+  defined in $\prog$. We write $\wired(\prog)$ to denote the
+  set of gates $g$ whose views occur on the rhs of an assignment
+  in $\prog$.
+  %$\view{1}{\ttt{gate:}g\ttt{tt}i} \in \iov(\prog)$
+  %and $\view{1}{\ttt{gate:}g\ttt{pt}i} \in \iov(\prog)$ for all $0 < i \le 4$,
+  %or  $\view{1}{\ttt{gate:}gi\ttt{.k}} \in \iov(\prog)$
+  %and $\view{1}{\ttt{gate:}gi\ttt{.p}} \in \iov(\prog)$ for all $0 < i \le 2$.
+\end{definition}
+Given the above definition, the needed result follows by Lemma \ref{lemma-scope} and definition
+of $\gNIMO$:
 \begin{lemma}
   \label{lemma-gnimo}
+  Let $X = \vars(\prog_1) \cap \vars(\prog_2)$.
   If $(\margd{\progd(\prog_1;\prog_2)}{\iov(\prog_1)}) * (\margd{\progd(\prog_1;\prog_2)}{\iov(\prog_2)})$
-  and $\flips(\prog_1) \cap \flips(\prog_2) = \varnothing$ and
+  and
   $\gNIMO(\prog_1;\prog_2,\{ \ttt{k} = \be_{\ttt{k}}; \ttt{p} = \be_{\ttt{p}}\}, g)$ for
-  $g \in \prog_1$,
+  $g \in \prog_2$,
   then $\gNIMO(\prog_1';\prog_2,\{ \ttt{k} = \be_{\ttt{k}}; \ttt{p} = \be_{\ttt{p}}\}, g)$
   for any $\prog_1'$ where $\prog_1';\prog_2$ is safe and
-  $\flips(\prog_1') \cap \flips(\prog_2) = \varnothing$.
+  $\vars(\prog_1') \cap \vars(\prog_2) = X$.
 \end{lemma}
@@
 \begin{enumerate}
-\item $g_1 \ne g_2$
+\item $(\{ g_1 \} \sqcup \{ g_2 \}) \cap \wired(\prog_1) = \varnothing$
 \item $\gNIMO(\prog_1,v_1,g_1) \wedge \gNIMO(\prog_1,v_2,g_2)$ 
@@
 \end{proof}
+
+\begin{fpfig}[t]{YGC copy gate definitions.}{fig-ygc-copy}
+  {\footnotesize
+    \begin{verbatimtab}
+      sharetab2(gid, tid, k, p, b)
+      {
+        let r1 = k xor b in
+        let r0 = (not k) xor (not b) in
+        v[1,gate: || gid || tid || 1] := select(p,r1,r0);
+        v[1,gate: || gid || tid || 2] := select(not p,r1,r0);
+      }
+
+      copygate(ca,cb,g)
+      {
+        let wl = owl(g) in
+        let owl1 = owl(ca) in
+        let owl2 = owl(cb) in
+        sharetab2(ca,tt,wl.k,wl.p,owl1.k); sharetab2(ca,pt,wl.k,wl.p,owl1.p);
+        sharetab2(cb,tt,wl.k,wl.p,owl2.k); sharetab2(cb,pt,wl.k,wl.p,owl2.p)
+      }
+
+      evalcopy(ca,cb,wv)
+      {
+        let wv1k = wv.k xor select(wv.p,v[1,gate: || ca || tt1], v[1,gate: || ca || tt2]) in
+        let wv1p = wv.k xor select(wv.p,v[1,gate: || ca || pt1], v[1,gate: || ca || pt1]) in
+        let wv2k = wv.k xor select(wv.p,v[1,gate: || cb || tt1], v[1,gate: || cb || tt2]) in
+        let wv2p = wv.k xor select(wv.p,v[1,gate: || cb || pt1], v[1,gate: || cb || pt1]) in
+        { wv1 = { k = wv1k; p = wv1p }; wv2 = { k = wv2k; p = wv2p } }  
+      }
+
+      copy(ca,cb,g,wv) { copygate(ca,cb,g); evalcopy(ca,cb,wv) } 
+    \end{verbatimtab}
+  }
+\end{fpfig}
+
+\subsection{Building and Extending Automatically Secure Circuits}
+\label{section-composition-copy}
+
+Our results in Section \ref{section-pre-post} and Theorem
+\ref{theorem-nimo} imply that any well-typed program using
+$\ttt{encode}$, $\ttt{andgg}$, and $\ttt{decode}$ will generate a
+circuit that is passive secure, provided that gate outputs are
+not used more than once. To add an $\ttt{or}$ gate to our
+library, we can use $\ttt{andgg}$ as a template, in fact simple
+modification of $\ttt{andtable}$ is all that would be needed. Thus,
+this library embodies a Fairplay-like language where well-typedness
+of metaprograms guarantees safety and security of generated protocols.
+
+The requirements that output wire labels are not used more than once
+reflects a known result, that naive copy gate output is unsound in
+YGC- i.e., passive security fails. In fact, we are unable to
+automatically verify $\gNIMO$ for $\ttt{andgg}$ without the assumption
+that input wire lables are in independent uniform distributions. We
+can capture the necessary fix as in Figure \ref{fig-ygc-copy}. The
+garbler creates a copy gate with two new output wire labels in
+independent uniform distributions (in \ttt{copygate}), and permutes
+them using the input wire label pointer after encrypting with the
+input wire label keys. The evaluator then recovers the encrypted
+output wire values during evaluation (in \ttt{evalcopy}). Given an
+$\ttt{orgg}$ gate defined as outlined abotve, we can then use
+secure $\ttt{copy}$ as follows: 
+\begin{verbatimtab}
+  let s = encode(s1,s2) in
+  let s1c = copy(c1s1,c2s1,s1,s.wv1) in
+  let s2c = copy(c1s2,c2s2,s2,s.wv2) in
+  let wv1 = andgg(1,c1s1,c1s2,s1c.wv1,s2c.wv1) in
+  let wv2 = orgg(2,c2s1,c2s2,s1c.wv2,s2c.wv2) in
+  decode(3,andgg(3,1,2,wv1,wv2))
+\end{verbatimtab}
+These definitions of $\ttt{copy}$ and $\ttt{orgg}$ have been
+automatically verified as in Section \ref{section-pre-post} for
+$\ttt{andgg}$. In general, any new gate form- such as optimized gates
+\cite{XXX}- can be verified automatically in a similar way.
+
+While the copy gate supports secure programming patterns, security
+still relies on programmer discipline, i.e., not wiring a gate more
+than once. However, we can keep track of wired gates in type effects,
+similar to the manner in which we keep track of defined views- type
+dependence in our system is able to accurately precisely track
+gate identifiers. This is also reminiscent of the use of linearity
+in other related systems for enforcing obliviousness \cite{darais2019language}.

### 2023-06-18 popl24/minicat.tex
checkpoint

@@
 denote $S \cup V$ where $S$ is the set of secret variables in $\prog$
-and $V$ is the set of views in $\prog$, and we write $\flips(\prog)$
-to denote the set $F$ of flip variables in $\prog$. For any set of
+and $V$ is the set of views in $\prog$, we write $\flips(\prog)$
+to denote the set $F$ of flip variables in $\prog$, and we
+write $\vars(\prog)$ to denote $\iov(\prog) \cup \flips(\prog)$. For any set of
 variables $X$ and parties $P$, we write $X_P$ to denote the subset of

### 2023-06-19 popl24/bruteforce.tex
checkpoint

@@
 amenable HPC optimizations such as parallelization and GPU matrix
-computations as shown in recent work \cite{XXX}.
+computations as shown in recent work
+\cite{sakama2017linear,aspis2018linear,nguyen2022enhancing,nguyen2021efficient}.
 
@@
 \subsection{Method 1: Inductive Computation of $\runs(\prog)$}
+\label{section-bruteforce-tables}
 
@@
 \ref{section-minicat-examples} and \ref{section-metalang-ygc} is
-available online \cite{XXX}.
+available online\footnote{URL redacted for double-blind review.}.
 
@@
 but the latter is a new (though unsurprising) result, having not been
-directly proven in \cite{XXX}. The method for
+directly proven in \cite{barthe2019probabilistic}. The method for
 verifying $\NIMO$ described here generalizes to any $\prog$. 
@@
 interpretation of Datalog and optimizations of model computation on
-high performance computers \cite{XXX}. The method here also enumerates
+high performance computers \cite{sakama2017linear}. The method here also enumerates
 $\runs(\prog)$ memory-by-memory, rather than in a ``batched'' manner
@@
 negation, with a negation-as-failure model, though we can use also
-use techniques in \cite{XXX} to also eliminate negation from
+use techniques in \cite{sakama2017linear} to eliminate negation from
 resulting programs. \emph{Atoms} are $\minifed$ variables, \emph{literals}
@@
 A full empirical exploration of the application of HPC optimizations
-as in \cite{XXX} is beyond the scope of this paper but is a compelling
-topic for future work. The reader is referred to \cite{XXX} for empirical
-results of HPC optimizations in other logic programming contexts.
+is beyond the scope of this paper but is a compelling topic for future
+work. The reader is referred to \cite{nguyen2022enhancing} for
+empirical results of HPC optimizations in other logic programming
+contexts.

### 2023-06-19 popl24/composition.tex
checkpoint

@@
 codebase for garbled circuits. This definition follows the
-\emph{point-and-permute} method described in \cite{XXX} and elsewhere,
+\emph{point-and-permute} method described in \cite{evans2018pragmatic} and elsewhere,
 to which the reader is referred for more in-depth discussion.
@@
 
+\cnote{Lemmas \ref{lemma-scope} and \ref{lemma-gnimo} in this Section may need some
+refinement but I think the general argument is reasonably developed.}
+
 Our aim is to demonstrated that $\gNIMO$ is preserved by gates within circuits,
@@
 gates can only be defined once, but it is also known that gate output
-wire labels should only be used once \cite{XXX}, in which case we
+wire labels should only be used once \cite{nieminen2023breaking}, in which case we
 say they are \emph{wired}. We return to this issue in more detail in

### 2023-06-19 popl24/intro.tex
checkpoint

\section{Introduction}

In this paper we make the following contributions.
\begin{itemize}
\item A new probabilistic programming language $\minifed$ for defining
  synchronous distributed protocols over the binary field. We consider example
  protocols including OT, communications encryption with one-time
  pads, and additive secret sharing (Section \ref{section-minicat}).
  \cnote{I'm hoping $\minifed$ hasn't been used already.}
\item A formulation of $\minifed$ program distributions, including
  distributions of secrets and views, supporting expression of security
  properties (Section \ref{section-pmf}). This intensional characterization
  is useful for expressing internal characteristics of protocols, and
  is needed for MPC due to the fact that obliviousness alone is not sufficient
  to capture characteristics of the real/ideal model, in particular the
  allowance of some information leakage through public outputs. 
\item A novel hyperproperty of program execution traces, called
  \emph{nonintererence modulo output ($\NIMO$)}, based on our
  formulation of program distributions, that implies passive security
  (Section \ref{section-hyperprop}.
\item A brute force mechanism for verifying passive security in
  $\minifed$ protocols, that is also amenable to HPC optimizations via
  translation of $\minifed$ protocols into Datalog (Sectior \ref{section-bruteforce}). 
\item A new metaprogramming language $\metaprot$ that dynamically
  generates $\minifed$ protocols. It includes control and data structures
  and is able to express logical protocol components, and enjoys a
  type safety result that guarantees that generated protocols
  are semantically well-defined (Section \ref{section-metalang}).
\item A formulation of Yao's Garbled Circuits, and a formulation of a
  compositional property of an extensible library of garbled gate
  components that is verifiable automatically. This property
  guarantees that any well-formed circuits using the library are
  passive secure (Section \ref{section-composition}).
\end{itemize}

### 2023-06-19 popl24/metalang.tex
checkpoint

@@
 Garbled Circuits (YGC), which are composed of so-called garbled gates.
-Languages for defined garbled circuits, beginning with Fairplay \cite{XXX},
+Languages for defined garbled circuits, beginning with Fairplay \cite{269581},
 treat gates as compositional units that are wired together by the programmer
@@
 resource-intensive compilation of programs with known errors
-\cite{XXX}. Some consequences of safety errors, for example accidental
+\cite{kreuter2012billion}. Some consequences of safety errors, for example accidental
 reuse of one-time pads, can also undermine security.
@@
 are instantiated with a type term substitution $\sigma$. In our
-implementation, we adapt \emph{synthesis} as defined in MetaML \cite{XXX}
+implementation, we adapt \emph{synthesis} as defined in Dependent ML \cite{10.1145/292540.292560}
 to obtain this $\sigma$- essentially this is a match on the syntactic
@@
 metaprogram, it also guarantees safety of the residual $\fedprot$
-program.
+program. Of course, it is important to note that ``safety'' here
+means the usual type safety property that programs do not get stuck,
+it does not imply any security hyperproperties. 
 \begin{theorem}[$\metaprot$ Type Safety]

### 2023-06-19 popl24/minicat.tex
checkpoint

@@
 through variables $\oracle{w}$- this is a form of shared randomness
-that is standard in the MPC setting \cite{XXX}.  Each client can only
+that is standard in the MPC setting \cite{evans2018pragmatic}.  Each client can only
 compute on their own variables in \emph{expressions} $\be$, and share
@@
 and $\prog_2$.
+
+\begin{fpfig}[t]{$\minifed$ source code syntax.}{fig-minifed-syntax}
+$$
+\begin{array}{rcl@{\hspace{8mm}}r}
+b &\in& \{ \etrue, \efalse \} \\
+w &\in& \mathrm{String} \\ 
+\cid &\in& \mathrm{Clients} \subset  \mathbb{N} \\[2mm]%\qquad  (\mathrm{Clients} \subset \mathbb{N} \text{\ in\ this\ presentation})\\[2mm]
+\bop &\in& \{ \eand, \eor, \exor \} \\[2mm]
+\be &::=& b \mid \flip{\cid}{w} \mid \secret{\cid}{w} \mid \view{\cid}{w} \mid \oracle{w} \mid & \textit{boolean expressions}\\
+& &  \enot\ \be \mid \be\ \bop\ \be \mid \select{\be}{\be}{\be} \mid \OT{\be}{\be}{\be} \\[2mm]
+\instr &::=& \eassign{\view{\cid}{w}}{\be} & \textit{view assignments} \\[2mm]
+\prog &::=& \varnothing \mid \instr; \prog & \textit{protocols}
+\end{array}
+$$ 
+\end{fpfig}
 
@@
 \subsection{Semantics}
-
-\begin{fpfig}[t]{$\minifed$ source code syntax.}{fig-minifed-syntax}
-$$
-\begin{array}{rcl@{\hspace{8mm}}r}
-b &\in& \{ \etrue, \efalse \} \\
-w &\in& \mathrm{String} \\ 
-\cid &\in& \mathrm{Clients} \subset  \mathbb{N} \\[2mm]%\qquad  (\mathrm{Clients} \subset \mathbb{N} \text{\ in\ this\ presentation})\\[2mm]
-\bop &\in& \{ \eand, \eor, \exor \} \\[2mm]
-\be &::=& b \mid \flip{\cid}{w} \mid \secret{\cid}{w} \mid \view{\cid}{w} \mid \oracle{w} \mid & \textit{boolean expressions}\\
-& &  \enot\ \be \mid \be\ \bop\ \be \mid \select{\be}{\be}{\be} \mid \OT{\be}{\be}{\be} \\[2mm]
-\instr &::=& \eassign{\view{\cid}{w}}{\be} & \textit{view assignments} \\[2mm]
-\prog &::=& \varnothing \mid \instr; \prog & \textit{protocols}
-\end{array}
-$$ 
-\end{fpfig}
 
@@
 binary field. Most of these are translations (into $\minifed$) of
-object examples in previous related work \cite{XXX}. We will return to
+object examples in previous related work
+\cite{barthe2019probabilistic,darais2019language}.  We will return to
 these as running examples throughout the paper to illustrate various
-concepts and advantages of our methods. In Section \ref{section-metalang}
-we will formulate a larger example implementation of Yao's
-Garbled Circuits (Example \ref{example-ygc}). 
+concepts and advantages of our methods. In Section
+\ref{section-metalang} we will formulate a larger example
+implementation of Yao's Garbled Circuits (Example \ref{example-ygc}).
 \begin{example}[One-Time Pad]
   \label{example-otp}
-As observed in \cite{XXX} and elsewhere, coin flips with $\exor$ can
+As observed in \cite{barthe2019probabilistic} and elsewhere, coin flips with $\exor$ can
 be used to model one-time pad encryption in the binary field. On the
@@
   \label{example-lambda-obliv}
-  In \cite{XXX}, Figure 3, two examples are introduced that illustrate
+  In \cite{darais2019language}, Figure 3, two examples are introduced that illustrate
   subtleties of probabilistic programming in
@@
   v[0,0] := select(s[1,s],select(flip[1,sx], flip[1,sx], flip[1,sy]),flip[1,z]) \end{verbatimtab}
-  We note that these examples are considered ``bad'' in \cite{XXX} since there
+  We note that these examples are considered ``bad'' in \cite{darais2019language} since there
   no information about honest secrets should be revealed to the adversary in
@@
   \label{example-OT}
-  In \cite{XXX} a protocol for $\ttt{OT}$ is defined that we recreate as
+  In \cite{barthe2019probabilistic} a protocol for $\ttt{OT}$ is defined that we recreate as
   follows, where client 2 is the receiver and client 1 is the sender. Whereas
-  in \cite{XXX} a trusted third party provides the sender and receiver with
+  in \cite{barthe2019probabilistic} a trusted third party provides the sender and receiver with
   shared randomness, in our version both clients refer to the oracle $\ttt{H}$
@@
 of this protocol that correspond to the properties proven manually in
-\cite{XXX}, and additionally how to automatically verify its passive
+\cite{barthe2019probabilistic}, and additionally how to automatically verify its passive
 security in the simulator model. With respect to our language primitives,

### 2023-06-19 popl24/pmf.tex
checkpoint

@@
 \section{Probability Mass Functions and Program Distributions}
+\label{section-pmf}
 
@@
 in programs we define probability mass functions in a standard manner,
-following the formulation in \ref{XXX}. In this formulation, memories
+following the formulation in \ref{barthe2019probabilistic}. In this formulation, memories
 denote conjunctions of variable assignments. Intuitively, given a
@@
 Note that this corresponds to \emph{perfect secrecy} as described
-in \cite{XXX}.
+in \cite{barthe2019probabilistic}.
 
 \paragraph{Example \ref{example-lambda-obliv}(a) revisited} As demonstrated
-by Example \ref{example-lambda-obliv} taken from \cite{XXX}, it is possible
+by Example \ref{example-lambda-obliv} taken from \cite{darais2019language}, it is possible
 to leak information probabilistically through obvservable dependencies

### 2023-06-19 popl24/intro.tex
checkpoint

@@
   formulation of program distributions, that implies passive security
-  (Section \ref{section-hyperprop}.
+  (Section \ref{section-hyperprop}).
 \item A brute force mechanism for verifying passive security in

### 2023-06-19 popl24/intro.tex
checkpoint

@@
   $\minifed$ protocols, that is also amenable to HPC optimizations via
-  translation of $\minifed$ protocols into Datalog (Sectior \ref{section-bruteforce}). 
+  translation of $\minifed$ protocols into Datalog (Section \ref{section-bruteforce}). 
 \item A new metaprogramming language $\metaprot$ that dynamically

### 2023-06-20 popl24/bruteforce.tex
typos

@@
 computation. The rewriting we describe here is to Datalog with
-negation, with a negation-as-failure model, though we can use also
+negation, with a negation-as-failure model, though we can also
 use techniques in \cite{sakama2017linear} to eliminate negation from

### 2023-06-20 popl24/composition.tex
checkpoint
typos

@@
 gate output values in an encrypted form. In our definition, wire
-labels are representated by records $\ttt{\{ k = }\beta_1\ttt{; p =
+labels are represented by records $\ttt{\{ k = }\beta_1\ttt{; p =
 }\beta_2\ttt{ \}}$, where $\ttt{k}$ is the \emph{key bit} and
@@
 The following program uses our YGC library to define
-a ciruit with a single and gate and input secrets $\ttt{s1}$ and
+a circuit with a single and gate and input secrets $\ttt{s1}$ and
 $\ttt{s2}$ from client's 1 and 2 respectively. 
@@
 individually are ``nearly'' passive secure, modulo the feature that
-outputs of individual logic gates in circuits are encryped. That is,
-each gate enjoys an encrypted variant of$\NIMO$ such that the relevant
+outputs of individual logic gates in circuits are encrypted. That is,
+each gate enjoys an encrypted variant of $\NIMO$ such that the relevant
 output is not in the clear but correlated flips sampled for the gate's
@@
 $\gNIMO$ we consider whether a given value $\itv$ correlates
-positively or negatively to the flip chosen for a wire lable, which
+positively or negatively to the flip chosen for a wire label, which
 can be either $1$ or $0$ uniformly- recall that
@@
   $$
-  \config{\store^{0'}}{\prog_1;\prog_2} \redxs \config{\store^{1'}}{\prog_2} \redxs
+  \config{\store^{0}}{\prog_1;\prog_2} \redxs \config{\store^{1'}}{\prog_2} \redxs
   \config{\store^{2'}}{\varnothing}

### 2023-06-20 popl24/hyperprop.tex
checkpoint
typos

@@
 \section{Hyperproperties of MPC Passive Security}
+\label{section-hyperprop}
 
@@
 
-\subsection{Passive Security and the Real/Ideal Model} Here we
+\subsection{Passive Security and the Real/Ideal Model} \dnote{would shortening the background info on how simulator model works be a bad idea?} Here we
 briefly summarize the formal definition of MPC security. For a
@@
 follow the rules of the protocol- as opposed to the
-\emph{active} aka maliciouos variant where participants
+\emph{active} aka malicious variant where participants
 are allowed to break these rules. In the passive model, the

### 2023-06-20 popl24/intro.tex
typos

@@
 \item A novel hyperproperty of program execution traces, called
-  \emph{nonintererence modulo output ($\NIMO$)}, based on our
+  \emph{noninterference modulo output ($\NIMO$)}, based on our
   formulation of program distributions, that implies passive security

### 2023-06-20 popl24/metalang.tex
checkpoint
typos

@@
 construction of boolean expression forms and view assignments. When
-$\metaprot$ programs construct a $\fedprot$ assignment, a side effect
+$\metaprot$ programs construct an $\fedprot$ assignment, a side effect
 occurs whereby the assignment is added to the end of the $\fedprot$
@@
 iff $f(x_1,\ldots,x_n) \{ e \} \in \codebase$.
-A \emph{metaprogram}, aka \emph{mataprotocol} is a pair of a 
+A \emph{metaprogram}, aka \emph{mataprotocol} \dnote{meta?} is a pair of a 
 codebase and expression $\codebase, e$. We may omit
@@
 for free occurences of $x$ in $e$. The rules are mostly standard,
-except why a concrete $\minifed$ assignment is encountered it is added
+except why \dnote{should this be a when?} a concrete $\minifed$ assignment is encountered it is added
 to the end of $\prog$.
@@
 post-generation by a direct analysis, for large programs this can be
-much more expensive and it is also better to not wast time on
+much more expensive and it is also better to not waste time on
 resource-intensive compilation of programs with known errors
@@
 \tjudge{\viewst_1}{\Gamma}{e_2}{\stringty{e_2'}}{\viewst_2}}
-{\tjudge{\viewst}{\Gamma}{\secret{e_1}{e_2}}{\bet{e_1'}}}
+{\tjudge{\viewst}{\Gamma}{\secret{e_1}{e_2}}{\bet{e_1'}}{\views_2}}
 

### 2023-06-20 popl24/minicat.tex
checkpoint
typos

@@
 \section{The $\minifed$ Language}
+\label{section-minicat}
 
@@
 formulation supports our automated analysis as we discuss in
-section
+section \dnote{5?}
 
@@
 Figure \ref{fig-minifed-syntax}, includes a standard boolean algebra
-with $\eand$, $\eor$, $\exor$, and $\enot$ as primitives. In addition
-programs have different kinds of variables, including \emph{secrets}
+with $\eand$, $\eor$, $\exor$, and $\enot$ as primitives. In addition,
+programs have different kinds of variables including \emph{secrets}
 $\secret{\cid}{w}$, \emph{flips} $\flip{\cid}{w}$, and \emph{views}
@@
 
-Expressions forms include the convenience
-$\select{\be_1}{\be_2}{\be_3}$ which is essentially a conditional
+Expression forms include the convenience
+$\select{\be_1}{\be_2}{\be_3}$, which is essentially a conditional
 expression with:
@@
 
-Given a set of variables $X$, we write $\store_X$ to denoted the
-memory $\store$ restricted to the domain $X$, and and we define
+Given a set of variables $X$, we write $\store_X$ to denote the
+memory $\store$ restricted to the domain $X$, and we define
 $\mems(X)$ as the set of all memories with domain $X$:
@@
 \ref{section-metalang} we will formulate a larger example
-implementation of Yao's Garbled Circuits (Example \ref{example-ygc}).
+implementation of Yao's Garbled Circuits.
 \begin{example}[One-Time Pad]

### 2023-06-20 popl24/pmf.tex
typos

@@
 In order to precisely characterize distributions of secrets and views
-in programs we define probability mass functions in a standard manner,
+in programs, we define probability mass functions in a standard manner,
 following the formulation in \ref{barthe2019probabilistic}. In this formulation, memories
@@
 secrets together with the random tape. And since we constrain programs
-to not overwrite views we are assured that \emph{final} memories
+to not overwrite views, we are assured that \emph{final} memories
 contain both a complete record of all initial secrets as well as views

### 2023-06-20 popl24/pmf.tex
typos pmf.tex

@@
 We consider random tapes to be selected from a uniform distribution.
-In our threat model we will also assume that any possible initial
+In our threat model, we will also assume that any possible initial
 secrets are equally likely when formulating the program distribution.
@@
 to be equally probable, establishing the basic distribution of the
-program. From this we can immediately derive the marginal distribution
+program. From this, we can immediately derive the marginal distribution
 of $S \cup V$ to reason about dependencies between secrets and views. 
@@
   \label{def-progd}
-  Given progam $\prog$ with $\iov(\prog) = (S,V)$ and $\flips(\prog) = F$, define $\stores$ as:
+  Given program $\prog$ with $\iov(\prog) = (S,V)$ and $\flips(\prog) = F$, define $\stores$ as:
   $$
@@
   $$
-  In some cases we will also be concerned with the (joint)
+  In some cases, we will also be concerned with the (joint)
   probabilities of expression interpretation given a preceding program
@@
 by Example \ref{example-lambda-obliv} taken from \cite{darais2019language}, it is possible
-to leak information probabilistically through obvservable dependencies
+to leak information probabilistically through observable dependencies
 in views. In the basic distribution for Example \ref{example-lambda-obliv}(a)
@@
 \end{mathpar}
-Further, since the standard MPC threat model allows partipating
+Further, since the standard MPC threat model allows participating
 clients to be corrupted we can formally consider the impacts on
@@
 with input secret 1 is corrupted, and we assume the output is 0, then
-adversary knows with certainty that clients 1 and 3 both have input
+the adversary knows with certainty that clients 1 and 3 both have input
 secrets 0. We can express this as:

### 2023-06-20 popl24/metalang.tex
typo

@@
     } \end{verbatimtab}
-  Here is $\metaprot$ program that uses this this function definition:
+  Here is $\metaprot$ program that uses this function definition:
   \begin{verbatimtab}

### 2023-06-20 popl24/hyperprop.tex
typos hyperprop.tex

@@
 
-In this section we formulate a new hyperproperty that we show is sound
+In this section, we formulate a new hyperproperty that we show is sound
 for passive security in the real/ideal model, called
@@
 more detailed discussion the reader is referred to \cite{XXX}.
-In this paper we will consider the \emph{passive} aka
+In this paper, we will consider the \emph{passive} aka
 semi-honest aka honest-but-curious variant. In this threat
@@
 must be able to tractably reconstruct this knowledge. This is not
-generally true, in particular if $\idealf$ is a 1-way
+generally true, in particular, if $\idealf$ is a 1-way
 function. Define:
@@
   Let $H$ and $C$ be an arbitrary partition of the federation and
-  suppose $\prog$ satifies noninterference modulo output. Let
+  suppose $\prog$ satisfies noninterference modulo output. Let
   $\iov(\prog) = (V,S)$ and $\flips(\prog) = F$ with output view $x

### 2023-06-20 popl24/bruteforce.tex
typos bruteforce.tex

@@
 also has the benefit of being flexible- we can use the same mechanisms
-to desribe $\NIMO$, noninterference, perfect secrecy, violations of
+to describe $\NIMO$, noninterference, perfect secrecy, violations of
 obliviousness, etc. These mechanisms can also be exploited to support
@@
 can be any of its marginal or conditional distributions. In this
-Section we consider two automated mechanisms for deriving
+Section, we consider two automated mechanisms for deriving
 $\progd(\prog)$ in this way. The first is through a straightforward
@@
 passive security of Examples \ref{example-he} and \ref{example-OT}-
-for the former this is well-known so the result is a simple proof-of-concept,
+for the former, this is well-known so the result is a simple proof-of-concept,
 but the latter is a new (though unsurprising) result, having not been

### 2023-06-25 popl24/intro.tex
Update intro.tex

@@
 
-In this paper we make the following contributions.
+In this paper, we make the following contributions.
 \begin{itemize}
@@
   synchronous distributed protocols over the binary field. We consider example
-  protocols including OT, communications encryption with one-time
+  protocols including oblivious transfer, communications encryption with one-time
   pads, and additive secret sharing (Section \ref{section-minicat}).
@@
   is needed for MPC due to the fact that obliviousness alone is not sufficient
-  to capture characteristics of the real/ideal model, in particular the
+  to capture characteristics of the real/ideal model, in particular, the
   allowance of some information leakage through public outputs. 

### 2023-06-28 popl24/composition.tex
checkpoint

@@
-\section{Compositional Properties: Garbled $\NIMO$}
+\section{Component Certification and Composition}
 \label{section-composition}
@@
 \ref{section-bruteforce}. These properties compose to yield $\NIMO$
-and scale to arbitrary programs. We will express these properties
-manually as pre-~and post-conditions on components, so our
-general approach is in a proof assistance style. The properties
-we use are refinements of $\NIMO$, and the overall technique is
-enabled by the intensional character of our program distribution
-formulation.
+and scale to arbitrary programs. In our metatheory we will express
+these properties as pre-~and post-conditions on components, so our
+general approach is in a proof assistant style.
 
@@
 
-\subsection{Verification of Pre-~and Post-Conditions}
+\subsection{Component Certification for $\NIMO$}
+\label{section-composition-certification}
 
-A basic intuition about YGC is that garbled gates considered
-individually are ``nearly'' passive secure, modulo the feature that
-outputs of individual logic gates in circuits are encrypted. That is,
-each gate enjoys an encrypted variant of $\NIMO$ such that the relevant
-output is not in the clear but correlated flips sampled for the gate's
-output wire label by the garbler, both key and pointer bits. We call
-this \emph{Garbled $\NIMO$}, or $\gNIMO$. In the definition of
-$\gNIMO$ we consider whether a given value $\itv$ correlates
-positively or negatively to the flip chosen for a wire label, which
-can be either $1$ or $0$ uniformly- recall that
-Definition \ref{def-progd} posits a variable $\itv$ to refer to the
-expression $\be$ when considering the distribution
-$\progd(\prog,\be)$.
-\begin{definition}[Garbled $\NIMO$]
-  We write $\gNIMO(\prog,\{ \ttt{k} = \be_{\ttt{k}}; \ttt{p} =
-  \be_{\ttt{p}}\}, g)$ with $\iov(\prog) = S \cup V$ iff for all $H$,
-  $C$ with $|C| \le |H|$, for all $\store \in \mems(S_C \cup V_C)$,
-  for both $w \in \{\ttt{p},\ttt{k}\}$, and for all $\stores$ with
-  domain $X = S_C \cup V_C \cup \{ \itv, \flip{2}{\ttt{gate:}g.w} \}$
-  such that:
-  $$
-  \stores = \{ \store' \mid \store \subset \store' \wedge 
-  \store'(\itv) = \store'(\flip{2}{\ttt{gate:}g.w}) \} 
-  $$
-  or
-  $$
-  \stores = \{ \store' \mid \store \subset \store' \wedge 
-  \store'(\itv) = \neg\store'(\flip{2}{\ttt{gate:}g.w}) \} 
-  $$
-  %where for any $\beta$:
-  %$$\store^2 = \store\{ \flip{2}{\ttt{gate:}g.w} \mapsto \beta \}\{ \itv \mapsto \beta \}$$
-  %or
-  %$$
-  %\store^2 = \store\{ \flip{2}{\ttt{gate:}g.w} \mapsto \beta \}\{ \itv \mapsto \neg\beta \}
-  %$$
-  we have:
-  $$
-  \margd{(\condd{\progd(\prog,\be_w)}{\stores_{X-{V_C}}})}{S_H} =
-  \margd{(\condd{\progd(\prog,\be_w)}{\stores})}{S_H}
-  $$
-\end{definition}
+We can use the automated verification methods developed in Section
+\ref{section-brute-force} to \emph{certify} YGC
+components. Certification obtains properties that we can then use
+higher-level reasoning-- inspired in part by techniques in
+\cite{barthe2019probabilistic}-- to prove that circuits built with
+certified components satisfy $\NIMO$ (Theorem \ref{theorem-ygc-NIMO})
+and thus are passive secure.
 
-\subsubsection{Verification Witnesses}
-
-\cnote{Lemmas \ref{lemma-scope} and \ref{lemma-gnimo} in this Section may need some
-refinement but I think the general argument is reasonably developed.}
-
-Our aim is to demonstrated that $\gNIMO$ is preserved by gates within circuits,
-where preservation is expressed via pre-~and post-conditions. We can use
-techniques in Section \ref{section-bruteforce} to automatically verify postconditions
-for components, using witness programs that satisfy preconditions prior to
-component execution. For example, we can automatically verify that $\ttt{encode}$
-generates a program and wire values the enjoy $\gNIMO$. Then, to verify
-correctness of $\ttt{and}$ gate construction and evaluation- i.e., that it
-preserves $\gNIMO$- we can verify
-that the $\minifed$ program and wire value generated by the following program
-$\metaprot$ program satisfies $\gNIMO$:
-\begin{verbatimtab}
-  let secrets = encode(s1,s2) in
-  andgate(0,s1,s2); evalgate(secrets.wv1,secrets.wv2)
-\end{verbatimtab}
-However, we need to establish that $\ttt{encode(s1,s2)}$ is an
-adequate witness, in that it ensures that automated verification generalizes to
-arbitrary circuit contexts.
-
-An important part of this argument is probabilistic independence,
-and in related reasoning we can leverage prior work on probabilistic
-separation logic \cite{barthe2019probabilistic}. We will use an
-analog of their separation formula, which we define in our setting as follows:
-\begin{definition}
-  We write $(\margd{\pdf{X}}{Y}) * (\margd{\pdf{X}}{X-Y})$ iff
-  $\pdf{X}(\store) = \pdf{X}(\store_{Y}) * \pdf{X}(\store_{X-Y})$
-  for all $\store \in \mems(X)$.
-\end{definition}
-
-\begin{definition}
-  We say that $\prog$ \emph{covers} $X$ iff
-  $\progd(\prog)(\store) > 0$ for all $\store\in\mems(X)$.
-\end{definition}
-
-We aim to show that gates are probabilistically independent of each
-other (i.e., are separated), and this independence supports
-correctness-preserving composition. A critical property is that
-separation is preserved by safe (in the sense of Theorem
-\ref{theorem-metalang-safety}) substitution of programs, assuming
-disjointness of flips used in programs separated programs.
-\begin{lemma}
-  \label{lemma-scope}
-  Let $X = \vars(\prog_1) \cap \vars(\prog_2)$.
-  If $(\margd{\progd(\prog_1;\prog_2)}{\iov(\prog_1)}) *
-  (\margd{\progd(\prog_1;\prog_2)}{\iov(\prog_2)})$ and $\prog_1$
-  covers $X$, then 
-  $(\margd{\progd(\prog_1';\prog_2)}{\iov(\prog_1)}) *
-  (\margd{\progd(\prog_1';\prog_2)}{\iov(\prog_2)})$ for any
-  $\prog_1'$ such that $\vars(\prog_1') \cap \vars(\prog_2) = X$ and
-  $\prog_1';\prog_2$ is safe.
-\end{lemma}
-\begin{proof}
-  Suppose on the contrary that there exists $\prog_1'$ that is not
-  separable from $\prog_2$. By results in
-  \cite{barthe2019probabilistic}, this implies that there
-  exists $\store^n$ and $\store^n$ for $0 \le n \le 1$ such that
-  $\store^1_X = \store^{1'}_X$ where:
-  $$
-  \config{\store^0}{\prog_1;\prog_2} \redxs \config{\store^1}{\prog_2} \redxs
-  \config{\store^2}{\varnothing}
-  $$
-  and:
-  $$
-  \config{\store^{0}}{\prog_1;\prog_2} \redxs \config{\store^{1'}}{\prog_2} \redxs
-  \config{\store^{2'}}{\varnothing}
-  $$
-  where $\store^2_{\vars(\prog_2)} \ne \store^{2'}_{\vars(\prog_2)}$, which is a
-  contradiction.
-%   This implies that there is some
-%  $\store \in \mems(\vars(\prog_1';\prog_2))$ such that
-%  $\store_{\vars{\prog_{1}'}} \in \runs{\prog_1'}$ and
-%  \mems(\vars(\prog_1))$ with   
-\end{proof}
-
-We must also choose witnesses that allow us to generalize verification
-of $\gNIMO$ preservation through gates. We need to consider separately
-the definition of gates and their use. Our type discipline ensures that
-gates can only be defined once, but it is also known that gate output
-wire labels should only be used once \cite{nieminen2023breaking}, in which case we
-say they are \emph{wired}. We return to this issue in more detail in
-Section \ref{section-composition-copy}.
-\begin{definition}
-  We write $g \in \prog$ iff all views for $g$'s table are fully
-  defined in $\prog$. We write $\wired(\prog)$ to denote the
-  set of gates $g$ whose views occur on the rhs of an assignment
-  in $\prog$.
-  %$\view{1}{\ttt{gate:}g\ttt{tt}i} \in \iov(\prog)$
-  %and $\view{1}{\ttt{gate:}g\ttt{pt}i} \in \iov(\prog)$ for all $0 < i \le 4$,
-  %or  $\view{1}{\ttt{gate:}gi\ttt{.k}} \in \iov(\prog)$
-  %and $\view{1}{\ttt{gate:}gi\ttt{.p}} \in \iov(\prog)$ for all $0 < i \le 2$.
-\end{definition}
-Given the above definition, the needed result follows by Lemma \ref{lemma-scope} and definition
-of $\gNIMO$:
-\begin{lemma}
-  \label{lemma-gnimo}
-  Let $X = \vars(\prog_1) \cap \vars(\prog_2)$.
-  If $(\margd{\progd(\prog_1;\prog_2)}{\iov(\prog_1)}) * (\margd{\progd(\prog_1;\prog_2)}{\iov(\prog_2)})$
-  and
-  $\gNIMO(\prog_1;\prog_2,\{ \ttt{k} = \be_{\ttt{k}}; \ttt{p} = \be_{\ttt{p}}\}, g)$ for
-  $g \in \prog_2$,
-  then $\gNIMO(\prog_1';\prog_2,\{ \ttt{k} = \be_{\ttt{k}}; \ttt{p} = \be_{\ttt{p}}\}, g)$
-  for any $\prog_1'$ where $\prog_1';\prog_2$ is safe and
-  $\vars(\prog_1') \cap \vars(\prog_2) = X$.
-\end{lemma}
-
-\subsubsection{Pre-~ and Post-Conditions}
-\label{section-pre-post}
-Now we can establish relevant pre-~and post-conditions for the
-top-level YGC functions defined in Figure \ref{fig-ygc-gates}. We note
-that in addition to the automated verification mechanisms described in
-Section \ref{section-bruteforce}, type safety in $\metaprot$ plays an
-important role, in part by ensuring that gates are uniquely defined
-(as views).
-
-Our strategy is to consider input encoding, internal gates, and
-output decoding separately. To begin, we show that input encoding
-establishes generates a program and wire values that initialize
-the $\gNIMO$ invariant in the circuit. There is no precondition
-witness needed for this result, which holds in any type-safe
-program context since the variables used by $\ttt{encode}$
-are all fresh. 
-\begin{lemma}[Input Encoding]
-  \label{lemma-encode}
-  Given well-typed $\prog_1;E[\ttt{encode}(s_1,s_2)]$ where:
-  $$
-  \config{\prog_1}{E[\ttt{encode}(s_1,s_2)]} \redxs
-  \config{\prog_1;\prog_2}{E[\ttt{\{wv1=}v_1\ttt{;wv2=}v_2\ttt{\}}]}
-  $$
-  Then:
-  $$\gNIMO(\prog_1;\prog_2,v_1,s_1) \wedge \gNIMO(\prog_1;\prog_2,v_2,s_2)$$
-\end{lemma}
-
-\begin{proof}
-  We have automatically verified
-  $\gNIMO(\prog_2,v_1,\ttt{s1})$ and $\gNIMO(\prog_2,v_1,\ttt{s1})$
-  where:
-  $$
-  \config{\varnothing}{\ttt{encode(s1,s2)}} \redxs
-  \config{\prog_2}{\ttt{\{wv1=}v_1\ttt{;wv2=}v_2\ttt{\}}}
-  $$
-  using the method described in Section \ref{section-bruteforce-tables}.
-  Given arbitrary $\prog_1$ such that $\prog_1;\ttt{encode}(s_1,s_2)$
-  is well-typed, hence $\prog_1;\prog_2$ is safe, 
-  since $\ttt{encode}$ uses unique secrets and flips per gates and
-  gates (views) are not redefined in safe programs,
-  $(\iov(\prog_1) \cup \flips(\prog_1)) \cap (\iov(\prog_2) \cup
-  \flips(\prog_2)) = \varnothing$. This implies 
-  $(\margd{\progd(\prog_1;\prog_2)}{\iov(\prog_1)}) *
-  (\margd{\progd(\prog_1;\prog_2)}{\iov(\prog_2)})$, so the result follows
-  by Lemma \ref{lemma-gnimo}.
-\end{proof}
-
-One nuance in our results for internal gates and for decoding is that the relevant
-compositional units are the complete related subprotocols, including both
-the garbler's construction of the gates and the evaluator's evaluation of
-them. We therefore make the following definitions. However, we note that
-our formal results are applicable to the YGC style where the garbler shares
-the entire circuit prior to evaluation, rather than gate-by-gate during
-evaluation. 
+Our method relies on a definition of compositional units comprising
+complete related subprotocols-- i.e., both the garbler's
+construction of the gates and the evaluator's evaluation of them. We
+therefore make the following definitions. However, we note that our
+formal results are applicable to the YGC style where the garbler
+shares the entire circuit prior to evaluation.
 \begin{definition}
@@
   andgg(g, ga, gb, wva, wvb) { andgate(g, ga, gb); evalgate(g, wva, wvb) }
-
   decode(g, v) { garbledecode(g); evaldecode(v) }
@@
 
-Now we can demonstrate that our $\ttt{and}$ gate definition preserves
-$\gNIMO$, and additionally that internale gates are probabilistically independent.
-This latter fact is crucial to compositionality, and allows us for automatically
-verify the circuit in isolation, with input encoding serving as an adequate input
-witness for generalization. 
-\begin{lemma}[Composition ($\eand$)]
+Our certification techniques are defined generally wrt implementations
+of input encoding, decoding, and internal gates. In the following we
+will refer to arbitrary decode, encode, and gate functions with the
+restriction that any function in each of these categories has the same
+valid type signature as $\ttt{decode}$, $\ttt{encode}$, and
+$\ttt{andgg}$ respectively.
+
+Certifications will demonstrate probabilistic independence of input
+wire lables and garbled tables in a program's basic distribution, so
+we need to consider dependencies of flips and gate views. Thus we make
+the following definitions:
+\begin{definition}
+  We write $\progtt(\prog)$ to denote the \emph{basic distribution} of
+  $\prog$ as defined in Definition \ref{def-progd}. For any program $\prog =
+  (\eassign{v_1}{\be_1};\ldots;\eassign{v_n}{\be_n})$, define
+  $\vdefs(\prog) \defeq \{ v_1,\ldots,v_n \}$.
+\end{definition}
+We also need to formalize notions of probabilistic independence, aka
+\emph{separation}, and correlation. We borrow the symbols $*$ and $\sim$
+from \cite{barthe2019probabilistic} with the same denotations.
+\begin{definition}
+  We write $\vc{\pdf{X}}{x}{y}$ iff $\pdf{X}(\{ x \mapsto 0\}\ |\ \{ y \mapsto 0 \}) =
+  \pdf{X}(\{ x \mapsto 1\}\ |\ \{ y \mapsto 1 \}) = 1$.
+  We write $\sep{\pdf{X}}{Y}{Z}$ iff for all
+    $\store \in \mems(X \cup Y)$:
+  $$(\margd{\pdf{X}}{Y \cup Z})(\store) =
+  (\margd{\pdf{X}}{Y})(\store_Y) * (\margd{\pdf{X}}{Z})(\store_Z)$$ 
+\end{definition}
+It is also necessary to define notions of correlation with wire
+labels, since wire values flow into and out of gates in either
+positive of negative correlation with them.
+\begin{definition}
+  We write $\gc{\prog}{\{ \ttt{k = }\be_{k}; \ttt{p = }\be_{p} \}}{g}$ iff one of the
+  following hold:
+  \begin{enumerate}[\hspace{5mm}i.]
+  \item $\vc{\progtt(\prog,\be_k)}{\itv}{\ttt{flip[2,gate:}g\ttt{.k]}}\quad \text{and} \quad
+    \vc{\progtt(\prog,\be_p)}{\itv}{\ttt{flip[2,gate:}g\ttt{.p]}}$
+  \item $\vc{\progtt(\prog,\enot\ \be_k)}{\itv}{\ttt{flip[2,gate:}g\ttt{.k]}}\quad \text{and}\quad
+    \vc{\progtt(\prog,\enot\ \be_p)}{\itv}{\ttt{flip[2,gate:}g\ttt{.p]}}$
+  \end{enumerate}
+\end{definition}
+Thus, if a value is in correlation with a wire label, it is either
+correlated with the wire label or with the inverse of both the key and
+pointer bits. For subsequent discussion we make the following addition
+to the codebase:
+$$
+\ttt{invert(\{ k = bk; p = bp \}) \{ \{ k = not bk; p = not bp \} \}}
+$$
+Gate certification establishes separation of input labels from the
+gate table and output value, and correlation of the output with
+the output wire label under any input wire value combination.
+\begin{definition}[Gate Certification]
+  \label{definition-gate-certification}
+Let $\mathit{gate}$ be a gate function. For $1 \le i \le 4$  let:
+  $$
+  \config{\varnothing}{\ttt{v[0,out] := } e_i}\redxs\config{\prog_i}{\ttt{()}}
+  $$
+  where
+  \begin{enumerate}[\hspace{5mm}$e_1 \defeq$]
+    \item $\mathit{gate}\ttt{(c,a,b,owl(a),owl(b));}$
+    \item $\mathit{gate}\ttt{(c,a,b,owl(a),invert(owl(b)));}$
+    \item $\mathit{gate}\ttt{(c,a,b,invert(owl(a)),owl(b));}$
+    \item $\mathit{gate}\ttt{(c,a,b,invert(owl(a)),invert(owl(b)))}$
+  \end{enumerate}
+  Define:
+  $${\small X \defeq \ttt{\{ flip[2,gate:a.k], flip[2,gate:a.p], flip[2,gate:b.k], flip[2,gate:b.p] \}}}$$
+  Then $\mathit{gate}$ is \emph{certified} iff for all $1 \le i \le 4$:
+  $$
+  \sep{\progtt(\prog_i)}{X}{\vdefs(\prog_i)} 
+  \quad \text{and} \quad \gc{\progd(\prog_i)}{\vx{0}{out}}{\ttt{owl(c)}}
+  $$
+\end{definition}
+Encode certification establishes correlation of the encoded input values with
+the secret wire labels, and separation of client 2's secret from client 1's views. 
+\begin{definition}[Encode Certification]
+  Let $\mathit{encode}$ be an encoding function, and let:
+  $$
+  \config{\varnothing}{\ttt{encode(s1,s2)}} \redxs
+  \config{\prog}{\ttt{\{wv1=}v_1\ttt{;wv2=}v_2\ttt{\}}}
+  $$
+  Then $\mathit{encode}$ is \emph{certified} iff each of the following conditions hold:
+  \begin{enumerate}[\hspace{5mm}i.]
+  \item $\gc{\progd(\prog,v_1)}{\itv}{\ttt{owl(s1)}} \quad \text{and} \quad
+    \gc{\progd(\prog,v_2)}{\itv}{\ttt{owl(s2)}}$
+  \item $\sep{\progd(\prog,v_2)}{\{\ttt{s[2,s2]\}}}{(\vdefs(\prog) \cup \itv)}$
+  \end{enumerate}
+\end{definition}
+Decode certification establishes separation of the input wire
+label from the decoding tables under any input value condition. 
+\begin{definition}[Decode Certification]
+  Let $\mathit{decode}$ be a decoding function, and let:
+  $$
+  \config{\varnothing}{\mathit{decode}(\ttt{c}, \ttt{owl(c)})}
+  \redxs \config{\prog_1}{\be_1}
+  \quad \text{and} \quad
+  \config{\varnothing}{\mathit{decode}(\ttt{c}, \ttt{invert(owl(c))})}
+  \redxs \config{\prog_0}{\be_0}
+  $$
+  Then $\mathit{decode}$ is certified iff:
+  $$
+  \sep{\progtt(\prog_i)}{\ttt{\{flip[2,gate:c.k],flip[2,gate:c.p]\}}}{\vdefs(\prog_i)}
+  $$
+  %Then $\mathit{decode}$ is certified iff both of the following conditions hold:
+  %\begin{enumerate}[\hspace{5mm}i.]
+  %\item $\progd(\prog_1,\be_1)(\{ \itv \mapsto 1 \}) = 1$ and
+  %  $\progd(\prog_0,\be_0)(\{ \itv \mapsto 0 \}) = 1$
+  %\item $\sep{\progtt(\prog_i)}{\ttt{\{flip[2,gate:c.k],flip[2,gate:c.p]\}}}{\vdefs(\prog_i)}$
+  %\end{enumerate}
+\end{definition}
+We have certified each of the YGC components detailed above. 
+\begin{lemma}
+  \label{lemma-certification}
+  Each of $\ttt{decode}$, $\ttt{encode}$, and $\ttt{andgg}$ are certified.
+\end{lemma}
+
+\subsection{Correctness of Certified Components}
+\label{section-composition-metatheory}
+
+To demonstrate correctness of certification, we need to show that
+isolated component certificates preserve relevant properties
+when integrated into larger programs. This resonates with
+\emph{framing} as developed in \cite{barthe2019probabilistic}.
+In the following we will use $f$ to range over flips.
+First, we show that if a flip $f$ is independent from the views
+in a protocol (as is the case with input labels used to
+encode a garbled table), then separation from views that
+use $f$ (e.g., in an output label) is preserved. 
+\begin{lemma}[Wire Framing]
+  \label{lemma-wire-framing}
+  If $\sep{\progtt(\prog)}{\{f\}}{\vdefs(\prog)}$ for flip $f$ and
+  $\vars(\prog) \cap \vars(\prog') = \{ f \}$ then
+  $\sep{\progtt(\prog';\prog)}{\vdefs(\prog')}{\vdefs(\prog)}$.
+\end{lemma}
+Next, we show that output correlations (e.g., between output wire values and
+labels) are preserved by substitution of correlated values for flips (e.g.,
+between input wire values and labels).
+\begin{lemma}[Substitution]
+  \label{lemma-substitution}
+  If $\vc{\progtt(\prog,\be)}{\itv}{f}$ and $\vc{\progtt(\prog',\be')}{\itv}{f'}$
+  and $\sep{\progtt(\prog';\prog)}{\vdefs(\prog')}{\vdefs(\prog)}$
+  then $\vc{\progtt(\prog';\prog,\be[\be'/f'])}{\itv}{f}$.
+\end{lemma}
+Our wire framing and substitution Lemmas then allow us to generalize
+certification of gates and encoding to arbitrary circuit contexts,
+assuming that outputs are not wired to multiple inputs \cite{tate2003garbled,nieminen2023breaking}.
+\begin{lemma}[Gate Composition]
   \label{lemma-compose}
-Given well-typed $\prog_1;E[\ttt{andgg}(g,g_1,g_2,v_1,v_2)]$ where:
+Given well-typed $\prog_1;E[\mathit{gate}(g,g_1,g_2,v_1,v_2)]$ and certified $\mathit{gate}$ where:
 $$
-\config{\prog_1}{E[\ttt{andgg}(g,g_1,g_2,v_1,v_2)]} \redxs \config{\prog_1;\prog_2}{E[v]}
+\config{\prog_1}{E[\mathit{gate}(g,g_1,g_2,v_1,v_2)]} \redxs \config{\prog_1;\prog_2}{E[v]}
 $$
@@
 \begin{enumerate}
-\item $(\{ g_1 \} \sqcup \{ g_2 \}) \cap \wired(\prog_1) = \varnothing$
-\item $\gNIMO(\prog_1,v_1,g_1) \wedge \gNIMO(\prog_1,v_2,g_2)$ 
+\item $\{ g_1 \} \cap \{ g_2 \} \cap \wired(\prog_1) = \varnothing$ and $g \not\in \prog_1$
+\item $\gc{\progd(\prog_1)}{v_1}{\owl(g_1)}$ and $\gc{\progd(\prog_1)}{v_2}{\owl(g_2)}$
 \end{enumerate}
@@
 \begin{enumerate}
-  \item $\gNIMO(\prog_1;\prog_2,v,g)$
-  \item $(\margd{\progd(\prog_1;\prog_2)}{\iov(\prog_1)}) * (\margd{\progd(\prog_1;\prog_2)}{\iov(\prog_2)})$
+  \item $\gc{\progd(\prog_1;\prog_2)}{v}{\owl(g_2)}$
+  \item $\sep{\progd(\prog_1;\prog_2,v)}{\vdefs(\prog_1)}{\vdefs(\prog_2) \cup \{\itv\}}$
 \end{enumerate}
@@
 \begin{proof}
-  We have automatically verified the postconditions (1) and (2) for the particular
-  program $\prog_1;\prog_2$ resulting from execution of:
-  $$e\quad \defeq\quad \ttt{let gs = encode(s1,s2) in andgg(0,s1,s2,gs.wv1,gs.wv2)}$$
-  where:
-  $$
-  \config{\varnothing}{e} \redxs
-  \config{\prog_1}{\ttt{andgg}(\ttt{0},\ttt{s1},\ttt{s2},v_1,v_2)} \redxs \config{\prog_1;\prog_2}{v}
-  $$
-  By Lemma \ref{lemma-encode} we have that
-  $\gNIMO(\prog_1,v_1,\ttt{s1}) \wedge \gNIMO(\prog_1,v_2,\ttt{s2})$. 
-  And since $\ttt{andgate}$ generates unique flips per gate, and well-typedness ensures
-  uniqueness of gates, we can generalize the result to replace $\prog_1$ with arbitrary $\prog_1'$
-  such that $\prog_1';\prog_2$ is safe,
-  applying Lemma \ref{lemma-scope} to obtain postcondition (2) and Lemma \ref{lemma-gnimo}
-  to obtain postcondition (1). 
+  Given that $g_1$ and $g_2$ are distinct and not wired in $\prog_1$
+  we are assured that $\ttt{owl}(g_1)$ and $\ttt{owl}(g_1)$ are in
+  independent uniform distributions, and given that $g \not\in
+  \prog_1$ we are assured that $\ttt{owl}(g)$ contains entirely fresh
+  flips. Thus, precondition 2, Lemma \ref{lemma-wire-framing}, and
+  Definition \ref{definition-gate-certification} imply postcondition
+  2.  Furthermore, since gate certification guarantees that
+  correlation with $\ttt{owl}(g)$ is guaranteed under all possible gate
+  input wire values, precondition 2 and Lemma \ref{lemma-substitution}
+  guarantee postcondition 2.
 \end{proof}
-Finally, we can demonstrate that decoding correctly casts a garbled output into the
-clear, and thus establishes $\NIMO$ for the garbled circuit that precedes it.
-\begin{lemma}[Decoding]
+
+\begin{lemma}[Decode Composition]
   \label{lemma-decode}
-Given well-typed $\prog_1;E[\ttt{decode}(g,v)]$ where
+Given well-typed $\prog_1;E[\mathit{decode}(g,v)]$ and certified $\mathit{decode}$ where 
 $$
@@
 $$
-If $\gNIMO(\prog_1,v,g)$, then $\NIMO(\prog_1;\prog_2;\eassign{\outv}{\be})$.
+If the following preconditions hold:
+\begin{enumerate}
+\item $g \not\in \cap \wired(\prog_1)$
+\item $\gc{\progd(\prog_1)}{v}{\owl(g)}$
+\end{enumerate}
+then we have as postcondition:
+\begin{enumerate}
+  \item $\sep{\progd(\prog_1;\prog_2)}{\vdefs(\prog_1)}{\vdefs(\prog_2)}$
+\end{enumerate}
 \end{lemma}
+Given the above, we can establish that any circuit built with certified
+components is passive secure. 
+\begin{theorem}
+  \label{theorem-ygc-NIMO}
+  If $e$ is a well-typed circuit definition using certified components
+  where no gate output is wired more than once, and $\config{\varnothing}{e}
+  \redxs \config{\prog}{\varnothing}$, then $\NIMO(\prog)$. 
+\end{theorem}
+
 \begin{proof}
-  We have automatically verified $
-  \progd(\prog,\be)(\{ \itv \mapsto 1 \}) = 1$ where:
-  $$\config{\varnothing}{\ttt{decode(g,\{ k = flip[2, gate:g.k]; p = flip[2,gate:g.p] \})}}
-    \redxs \config{\prog}{\be}$$
-    and also $
-  \progd(\prog,\be)(\{ \itv \mapsto 0 \}) = 1$ where:
-  $$\config{\varnothing}{\ttt{decode(g,\{ k = not flip[2, gate:g.k]; p = not flip[2,gate:g.p] \})}}
-    \redxs \config{\prog}{\be}$$
-  The result follows by definition of $\gNIMO$ and $\NIMO$.
+  Let $\iov(\prog) = S \cup (V \cup \outv)$. Noting that any
+  encoding necessarily uses entirely fresh variables by Definition and
+  type safety, by applying Lemmas \ref{lemma-compose} and
+  \ref{lemma-decode} and the frame rule of
+  \cite{barthe2019probabilistic} we obtain:
+  $$
+  \sep{\progd(\prog)}{(S_{\{2\}}\cup\{\outv\})}{V}
+  $$
+  Noting that the garbler (client 2) has no views, this means that
+  the garbler's secrets are independent of the evaluator's views. The
+  result follows by Definition \ref{definition-NIMO}.
 \end{proof}

### 2023-06-28 popl24/hyperprop.tex
checkpoint

@@
 \begin{definition}[Noninterference Modulo Output]
+  \label{definition-NIMO}
   We say that a program  $\prog$ satisfies \emph{noninterference modulo output},

### 2023-06-28 popl24/pmf.tex
checkpoint

@@
 in programs, we define probability mass functions in a standard manner,
-following the formulation in \ref{barthe2019probabilistic}. In this formulation, memories
+following the formulation in \cite{barthe2019probabilistic}. In this formulation, memories
 denote conjunctions of variable assignments. Intuitively, given a

### 2023-07-01 popl24/composition.tex
checkpoint

@@
   \label{theorem-ygc-NIMO}
-  If $e$ is a well-typed circuit definition using certified components
+  If $e$ is a well-typed YGC circuit definition using certified components
   where no gate output is wired more than once, and $\config{\varnothing}{e}
@@
 in other related systems for enforcing obliviousness \cite{darais2019language}.
+
+\subsection{Compositional Properties of GMW}
+
+\input{gmw}
+
+\begin{definition}[GMW Gate Certification]
+  \label{definition-gmwgate-certification}
+  Let $\mathit{gate}$ be a gate function and let $e$ be the following
+  $\metaprot$ program:
+  $$
+  \begin{array}{l}
+    \ttt{let in1 = \{c1 = flip[1,s11];c2 = flip[2,s12]\} in}\\
+    \ttt{let in2 = \{c1 = flip[1,s21];c2 = flip[2,s22]\} in}\\
+    \mathit{gate}\ttt{(g,in1,in2)}
+  \end{array}
+  $$
+  and let $
+  \config{\varnothing}{e}\redxs\config{\prog}{\ttt{\{c1 = v[1,gout]; c2 = v[2,gout]\}}}
+  $.
+  Then $\mathit{gate}$ is \emph{certified} iff both of the following conditions hold:
+  \begin{enumerate}
+  \item $\sep{\progtt(\prog)}{\ttt{\{flip[1,s11],flip[1,s21]\}}}{\ttt{\{v[2,gout]\}}}$
+  \item $\sep{\progtt(\prog)}{\ttt{\{flip[2,s12],flip[2,s22]\}}}{\ttt{\{v[1,gout]\}}}$
+  \end{enumerate}
+\end{definition}
+
+\begin{lemma}[GMW Gate Composition]
+  \label{lemma-compose-gmw}
+  Given well-typed $\prog_1;E[\mathit{gate}(\ttt{g},v_1,v_2)]$ and certified $\mathit{gate}$ where:
+  $$
+  \begin{array}{lcr}
+   v_1 &\defeq& \ttt{\{c1 = v[1,g1out]; c2 = v[2,g1out]\}}\\
+   v_2 &\defeq& \ttt{\{c1 = v[1,g2out]; c2 = v[2,g2out]\}}
+  \end{array}
+  $$
+  and:
+  $$
+  \config{\prog_1}{E[\mathit{gate}(\ttt{g},v_1,v_2)]} \redxs \config{\prog_1;\prog_2}{E[\ttt{\{c1 = v[1,gout]; c2 = v[2,gout]\}}]}
+  $$
+  Then the following postconditions hold:
+  \begin{enumerate}
+  \item $\sep{\progtt(\prog)}{\ttt{\{v[1,g1out],v[1,g2out]\}}}{\ttt{\{v[2,gout]\}}}$
+  \item $\sep{\progtt(\prog)}{\ttt{\{v[2,g1out],v[2,g2out]\}}}{\ttt{\{v[1,gout]\}}}$
+  \end{enumerate}
+\end{lemma}
+
+\begin{definition}[GMW Encode Certification]
+  Let $\mathit{encode}$ be an encoding function, and let:
+  $$
+  \config{\varnothing}{\ttt{encode(s1,s2)}} \redxs
+  \config{\prog}{\ttt{\{shares1 = }v_1\ttt{; shares2 = }v_2\ttt{\}}}
+  $$
+  where:
+  $$
+  \begin{array}{rcl}
+    v_1 &=& \ttt{\{ c1 = v[1,s1out]; c2 = v[2,s1out] \}}\\
+    v_2 &=& \ttt{\{ c1 = v[1,s2out]; c2 = v[2,s2out] \}}
+  \end{array}
+  $$
+  Then $\mathit{encode}$ is \emph{certified} iff each of the following conditions hold:
+  \begin{enumerate}[\hspace{5mm}i.]
+  \item $\sep{\progtt(\prog)}{\ttt{\{ s[1,s1] \}}}{\ttt{\{ v[2,s1out], v[2,s2out] \}}}$
+  \item $\sep{\progtt(\prog)}{\ttt{\{ s[2,s2] \}}}{\ttt{\{ v[1,s1out], v[1,s2out] \}}}$
+  \end{enumerate}
+\end{definition}
+
+\begin{theorem}
+  \label{theorem-gmw-NIMO}
+  If $e$ is a well-typed GMW circuit definition using certified
+  components and $\config{\varnothing}{e} \redxs
+  \config{\prog}{\varnothing}$, then $\NIMO(\prog)$.
+\end{theorem}
+\begin{proof}
+  Let $\iov(\prog) = S \cup (V \cup \outv)$. Noting that any encoding
+  necessarily uses entirely fresh variables by Definition and type
+  safety, by applying Lemmas \ref{lemma-compose-gmw} and
+  \ref{lemma-wire-framing} and the frame rule of
+  \cite{barthe2019probabilistic} we obtain:
+  $$
+  \sep{\progd(\prog)}{(S_{\{2\}}\cup\{\outv\})}{V_{\{1\}}}
+  \quad \text{and} \quad \sep{\progd(\prog)}{(S_{\{1\}}\cup\{\outv\})}{V_{\{2\}}}
+  $$
+  That is, the marginal distribution of client 1's secrets and the output are independent
+  of client 2's views, and vice-versa. The result follows. 
+\end{proof}

### 2023-07-01 popl24/main.tex
checkpoint

@@
 
-\title{Automating MPC Security}
+\title{Automating Verification of MPC Security: $\metaprot$ to $\minifed$}
 
@@
 \begin{abstract}
-  We show how to formulate a sound hyperproperty of MPC passive security, and how this
-  supports automated verification.
+  We formulate a sound hyperproperty of MPC passive security and show how this
+  supports automated and semi-automated verification.
 \end{abstract}
@@
 \maketitle
+
+\input{intro}
 
@@
 
+\input{bruteforce}
+
+\input{metalang}
+
+\input{composition}
+
 \bibliographystyle{ACM-Reference-Format}
-\bibliography{sample-base}
+\bibliography{logic-bibliography,secure-computation-bibliography}
 

### 2023-07-01 popl24/pmf.tex
checkpoint

@@
   A \emph{memory probability distribution} $\pdf{X}$ is a function
-  mapping memories in $\mems(X)$ to values in $[0..1]$, such that:
+  mapping memories in $\mems(X)$ to values in $[0,1]$, such that:
   $$

### 2023-07-01 popl24/composition.tex
checkpoint

@@
 
-Our results in Section \ref{section-pre-post} and Theorem
-\ref{theorem-nimo} imply that any well-typed program using
+Theorem \ref{theorem-nimo} means that any well-typed program using
 $\ttt{encode}$, $\ttt{andgg}$, and $\ttt{decode}$ will generate a
-circuit that is passive secure, provided that gate outputs are
-not used more than once. To add an $\ttt{or}$ gate to our
-library, we can use $\ttt{andgg}$ as a template, in fact simple
-modification of $\ttt{andtable}$ is all that would be needed. Thus,
-this library embodies a Fairplay-like language where well-typedness
-of metaprograms guarantees safety and security of generated protocols.
+circuit that is passive secure, provided that gate outputs are not
+used more than once. To add an $\ttt{or}$ gate to our library, we can
+use $\ttt{andgg}$ as a template, in fact simple modification of
+$\ttt{andtable}$ is all that would be needed. Thus, this library
+embodies a Fairplay-like language where well-typedness of metaprograms
+guarantees safety and security of generated protocols.
 

### 2023-07-01 popl24/gmw.tex
checkpoint

\begin{fpfig}[t]{2-Party GMW functions.}{fig-gmw}
{\footnotesize
  \begin{verbatimtab}
    encodegmw(in1,in2)
    {
      v[1, in1 || out] := s[1,in1] xor flip[1,in1];
      v[2, in1 || out] := flip[1,in1];
      v[1, in2 || out] := s[2,in2] xor flip[2,in2];
      v[2, in2 || out] := flip[2,in2]
      { shares1 = { c1 = v[1, in1 || out]; c2 = v[2, in1 || out] };
        shares2 = { c1 = v[1, in2 || out]; c2 = v[2, in2 || out]} } 
    }

    andtablegmw(b1, b2)
    {
      let r11 = (b1 xor true) and (b2 xor true) in
      let r10 = (b1 xor true) and (b2 xor false) in
      let r01 = (b1 xor false) and (b2 xor true) in
      let r00 = (bl xor false) and (b2 xor false) in
      { v1 = r11; v2 = r10; v3 = r01; v4 = r00 }
    }

    andgmw(g, shares1, shares2)
    {
      let r = flip[1,g || .r] in
      let table = andtablegmw(shares1.c1, shares2.c1) in
      let r11 =  r xor table.v1 in
      let r10 =  r xor table.v2 in
      let r01 =  r xor table.v3 in
      let r00 =  r xor table.v4 in
      v[2,g || out] := OT4(shares1.c2, shares2.c2, r11, r10, r01, r00);
      v[1,g || out] := r;
      { c1 = v[1,g || out]; c2 = v[2,g || out]}
    }

    decodegmw(shares)
    {
      v[0,output] := shares.c1 xor shares.c2
    }
  \end{verbatimtab}
}
\end{fpfig}

### 2023-07-03 popl24/hyperprop.tex
checkpoint

@@
 
-\subsection{Passive Security and the Real/Ideal Model} \dnote{would shortening the background info on how simulator model works be a bad idea?} Here we
-briefly summarize the formal definition of MPC security. For a
+\subsection{Passive Security and the Real/Ideal Model}
+\label{section-hyperprop-passive}
+Here we briefly summarize the formal definition of MPC security. For a
 more detailed discussion the reader is referred to \cite{XXX}.
@@
 \subsection{Probabilistic Noninterference}
-\label{section-ni}
+\label{section-hyperprop-ni}
 Previous work has noted

### 2023-07-03 popl24/intro.tex
checkpoint

@@
 \section{Introduction}
+
+Secure Multi-Party Computation (MPC) protocols support data privacy in
+important distributed applications such as privacy-preserving machine
+learning and Zero-Knowledge proofs in blockchains. MPC methods have
+been developed by the cryptography community for many years, while
+receiving the attention of the programming languages community only
+relatively recently. In contrast, information flow security has
+received significant attention in PL theory and practice especially
+since the turn of the century \cite{1159651}, including a menagerie of
+variants, enforcement mechanisms, and programming frameworks. Much of
+this has been enabled by the unified metatheory of
+\emph{hyperproperties} \cite{10.5555/1891823.1891830} that establishes
+a common conceptual framework for reasoning about and implementing
+systems with information flow security.
+
+Our goal is to explore and establish connections between the security
+model of MPC-- the \emph{real/ideal} aka \emph{simulator security}
+model -- and trace-based hyperproperties, and to leverage these
+connections to obtain automated enforcement mechanisms for MPC
+protocol development. Currently, proof methods for MPC protocol
+development are well-studied \cite{Lindell2017} but manual and 
+thus tedious and error-prone, and are also non-standardized and
+unfamiliar to most PL theorists. Therefore our exploration will
+make both theoretical and practical contributions by bridging a gap between
+information flow and simulator security methodologies.
+
+\subsection{The Analysis Challenge of MPC} MPC protocols involve communication
+between a group of distributed participants called a \emph{federation}
+that collaboratively compute and publish the result of some known
+\emph{ideal functionality} $\idealf$, while keeping each party's input
+``secret'', without the use of a trusted third party. This last part
+is critical. For example, if we take $\idealf$ to be the majority vote
+function, a protocol for computing $\idealf$ is MPC-secure if, given
+any set of input votes, it correctly computes and publishes the voting
+result without directly revealing the votes of individual participants
+to the public or to other participants. However, note that by
+publishing the output, some information about individual votes may be
+implicitly declassified.  For example, in the case of a majority vote
+in a federation of size 3, if the motion carries and party 1 has voted
+no, then party 1 knows exactly the votes of parties 2 and 3. This
+cannot be avoided due to the nature of $\idealf$.
+
+Security in the MPC setting thus means that protocols cannot reveal
+any more information than the output of the ideal functionality, which
+will be published. The security model also assumes that some subset of
+participants can be corrupted and collude adversarially to infer
+secret information. The accepted method of demonstrating protocol
+security in this setting is to define a \emph{simulation} algorithm
+that runs in the ``ideal'' world which, given just the inputs of
+corrupted parties and the output of the ideal functionality, is able
+to reconstruct information that corrupted parties receive in their
+so-called \emph{views} of the protocol running in the real world.
+This implies that adversarial views provide no information beyond what
+is provided by the ideal output alone. Simulation is defined
+probabilistically since MPC protocols typically rely on cryptographic
+and probabilistic methods.
+
+In Section \ref{section-hyperprop-passive} we formalize real/ideal
+security, and we define and discuss simple examples of MPC protocols
+in Section \ref{section-minicat-examples} and more complex ones in
+Section \ref{section-composition}. But an immediate and main point in
+relation to security hyperproperties is that, due to the potential for
+information release in MPC, simulator security is \emph{not} a strict
+probabilistic noninterference or trace obliviousness property, as we
+show in Section \ref{section-hyperprop-ni}- rather, the public output
+allows and sets an upper bound on declassification.
+
+\subsection{Related Work}
+
+We distinguish between \emph{extensional} vs.~\emph{intensional}
+properties and analysis of MPC protocols. By extensional, we mean
+analysis concerned with the use and interaction of complete protocols
+that are proved to be secure. By intensional, we mean analysis of the
+inner workings of protocols themselves.
+
+\subsection{Contributions}
 

### 2023-07-03 popl24/intro.tex
checkpoint

@@
 Secure Multi-Party Computation (MPC) protocols support data privacy in
-important distributed applications such as privacy-preserving machine
-learning and Zero-Knowledge proofs in blockchains. MPC methods have
-been developed by the cryptography community for many years, while
-receiving the attention of the programming languages community only
-relatively recently. In contrast, information flow security has
+important modern, distributed applications such as privacy-preserving
+machine learning and Zero-Knowledge proofs in blockchains \cite{XXX}
+\cnote{Need some banger cites here, maybe more MPC hype.}. MPC methods
+have been developed by the cryptography community for many years,
+while receiving the attention of the programming languages community
+only relatively recently. In contrast, information flow security has
 received significant attention in PL theory and practice especially
@@
 any set of input votes, it correctly computes and publishes the voting
-result without directly revealing the votes of individual participants
-to the public or to other participants. However, note that by
-publishing the output, some information about individual votes may be
-implicitly declassified.  For example, in the case of a majority vote
-in a federation of size 3, if the motion carries and party 1 has voted
-no, then party 1 knows exactly the votes of parties 2 and 3. This
-cannot be avoided due to the nature of $\idealf$.
+result but reveals no other information to the public or to other
+participants. However, by publishing the result, some information
+about individual votes may be implicitly declassified.  For example,
+in the case of a majority vote in a federation of size 3, if the
+motion carries and party 1 has voted no, then party 1 knows exactly
+the votes of parties 2 and 3. This cannot be avoided due to the nature
+of $\idealf$.
 
 Security in the MPC setting thus means that protocols cannot reveal
-any more information than the output of the ideal functionality, which
-will be published. The security model also assumes that some subset of
-participants can be corrupted and collude adversarially to infer
-secret information. The accepted method of demonstrating protocol
-security in this setting is to define a \emph{simulation} algorithm
-that runs in the ``ideal'' world which, given just the inputs of
-corrupted parties and the output of the ideal functionality, is able
-to reconstruct information that corrupted parties receive in their
-so-called \emph{views} of the protocol running in the real world.
-This implies that adversarial views provide no information beyond what
-is provided by the ideal output alone. Simulation is defined
-probabilistically since MPC protocols typically rely on cryptographic
-and probabilistic methods.
+any secret information other than what is impicitly declassified by
+the publicized output of the ideal functionality. The security model
+also assumes that some subset of participants can be corrupted and
+collude adversarially to possibly infer more secret information. The
+accepted method of demonstrating protocol security in this setting is
+to define a \emph{simulation} algorithm that runs in the ``ideal''
+world which, given just the inputs of corrupted parties and the output
+of the ideal functionality, is able to reconstruct information that
+corrupted parties receive in their so-called \emph{views} of the
+protocol running in the real world.  This implies that adversarial
+views provide no information beyond what is provided by the ideal
+output alone. Simulation is defined probabilistically since MPC
+protocols typically rely on cryptographic and probabilistic methods.
 
@@
 properties and analysis of MPC protocols. By extensional, we mean
-analysis concerned with the use and interaction of complete protocols
-that are proved to be secure. By intensional, we mean analysis of the
-inner workings of protocols themselves.
+simulator security itself as well as analysis concerned with the use
+and interaction of complete secure protocols. By intensional, we mean
+analysis and inner workings of protocols themselves. While the
+probabilistic formulations we develop could conceivably used
+extensionally, our main focus is on intensional properties of
+protocols, both through whole-program analysis and compositional
+properties of program components that support security.
+
+Previous work on analysis for the SecreC language
+\cite{almeida2018enforcing,10.1145/2637113.2637119} is concerned with
+extensional properties of MPC, in particular the specification and
+enforcement of declassification bounds in programs that use MPC in
+subprograms. This work is explicitly reminiscent of information flow
+approaches such as delimited information release
+\cite{10.1007/978-3-540-37621-7_9} and downgrading policies and
+relaxed noninterference \cite{10.1145/1040305.1040319}. However, their
+program logic assumes correctness of the underlying MPC protocols.
+The Wys$^\star$ language \cite{wysstar}, based on Wysteria
+\cite{rastogi2014wysteria}, has similar goals and includes a
+trace-based semantics for reasoning about the extensional interactions
+of MPC protocols. Their compiler also enforces some intensional
+properties, in particular that underlying multi-threaded protocols
+enforce the single-threaded source language semantics.
 

### 2023-07-03 popl24/intro.tex
checkpoint

@@
 probabilistic formulations we develop could conceivably used
-extensionally, our main focus is on intensional properties of
-protocols, both through whole-program analysis and compositional
-properties of program components that support security.
+extensionally, our main focus is on automated and semi-automated
+reasoning about intensional properties of protocols, both through
+whole-program analysis and compositional properties of program
+components that support security.
 
@@
 trace-based semantics for reasoning about the extensional interactions
-of MPC protocols. Their compiler also enforces some intensional
-properties, in particular that underlying multi-threaded protocols
-enforce the single-threaded source language semantics.
+of MPC protocols. Their compiler also guarantees that underlying
+multi-threaded protocols enforce the single-threaded source language
+semantics which is an intensional property. All of these systems
+leverage type systems to obtain automated security analyses.
 
-\subsection{Contributions}
+In the realm of intensional analysis, previous work has shown how to
+reduce proofs of simulator security to program equivalence assertions,
+which supports applications known techniques in automated verification
+\cite{8429300}. However, they focus on noninterference properties,
+which are arguably sound but notably incomplete wrt MPC functionality
+as we discuss above and in Section
+\ref{section-hyperprop-ni}. Probabilistic Separation Logic
+\cite{barthe2019probabilistic} develops a logical framework for
+reasoning about probabilistic independence in programs.  They consider
+several MPC protocols and show how to systematically deduce, e.g.,
+probabilistic independence of input secrets and adversarial views that
+are critical intensional properties of simulator security.  However,
+their focus on independence does not address the subtleties of
+dependence that arise during MPC protocol execution, and the authors
+do not formulate or prove simulator security properties in
+\cite{barthe2019probabilistic}. The $\lambda_{\mathrm{obliv}}$
+language \cite{darais2019language} is perhaps most closely related to
+our work. Here, a type system is used to automatically enforce
+so-called probabilistic trace obliviousness and they show how to use
+their system to implement several variants of oblivious RAM. However,
+obliviousness is related to noninterference and does not address
+allowable information leakage as in MPC, and their use of type
+linearity disallows reuse of sampled probabilistic values which is
+common in the MPC idiom, as our examples will illustrate. Similarly,
+previous work on oblivious data structures \cite{10.1145/3498713},
+while intended for sound integration into MPC protocols, focus on
+expression and enforcement of obliviousness and do not account
+for allowable information leakage.
+
+\subsection{Overview and Contributions}
 

### 2023-07-03 popl24/intro.tex
checkpoint

@@
 
+In \ref{section-minicat} we develop a new probabilistic programming
+language $\minifed$ for defining synchronous distributed protocols
+over the binary field. The syntax and semantics provides a succinct
+account of \emph{views} and (synchronous) messaging between protocol
+\emph{clients}. In this Section we consider small example protocols
+including oblivious transfer, communications encryption with one-time
+pads, and additive secret sharing. In Section \ref{section-pmf} we
+formulate \emph{program distributions}, a formalism for expressing
+dependencies between secrets and views, useful for expressing
+intensional characteristics of protocols and precisely quantifying
+allowable information leakage in terms of protocol outputs.
+
 In this paper, we make the following contributions.

### 2023-07-03 popl24/intro.tex
checkpoint

@@
 intensional characteristics of protocols and precisely quantifying
-allowable information leakage in terms of protocol outputs.
-
-In this paper, we make the following contributions.
-\begin{itemize}
-\item A new probabilistic programming language $\minifed$ for defining
-  synchronous distributed protocols over the binary field. We consider example
-  protocols including oblivious transfer, communications encryption with one-time
-  pads, and additive secret sharing (Section \ref{section-minicat}).
-  \cnote{I'm hoping $\minifed$ hasn't been used already.}
-\item A formulation of $\minifed$ program distributions, including
-  distributions of secrets and views, supporting expression of security
-  properties (Section \ref{section-pmf}). This intensional characterization
-  is useful for expressing internal characteristics of protocols, and
-  is needed for MPC due to the fact that obliviousness alone is not sufficient
-  to capture characteristics of the real/ideal model, in particular, the
-  allowance of some information leakage through public outputs. 
-\item A novel hyperproperty of program execution traces, called
-  \emph{noninterference modulo output ($\NIMO$)}, based on our
-  formulation of program distributions, that implies passive security
-  (Section \ref{section-hyperprop}).
-\item A brute force mechanism for verifying passive security in
-  $\minifed$ protocols, that is also amenable to HPC optimizations via
-  translation of $\minifed$ protocols into Datalog (Section \ref{section-bruteforce}). 
-\item A new metaprogramming language $\metaprot$ that dynamically
-  generates $\minifed$ protocols. It includes control and data structures
-  and is able to express logical protocol components, and enjoys a
-  type safety result that guarantees that generated protocols
-  are semantically well-defined (Section \ref{section-metalang}).
-\item A formulation of Yao's Garbled Circuits, and a formulation of a
-  compositional property of an extensible library of garbled gate
-  components that is verifiable automatically. This property
-  guarantees that any well-formed circuits using the library are
-  passive secure (Section \ref{section-composition}).
-\end{itemize}
+allowable information leakage in terms of protocol outputs. On this
+basis we then formulate a novel hyperproperty of program execution
+traces, called \emph{noninterference modulo output ($\NIMO$)} that we
+demonstrate implies passive simulator security. In Section
+\ref{section-bruteforce} we define a brute force mechanism for
+verifying passive security in $\minifed$ protocols, that is also
+amenable to HPC optimizations via translation of $\minifed$ protocols
+into Datalog \cnote{I think this Section can be on the chopping block
+  for space savings, since it's based on truth table
+  computation/Datalog semantics.}. In Section \ref{section-metalang}
+we define a new metaprogramming language $\metaprot$ that dynamically
+generates $\minifed$ protocols. It includes control and data
+structures and is able to express logical protocol components, and
+enjoys a type safety result that guarantees that generated protocols
+are semantically well-defined. Our key analysis results are presented
+in Section \ref{section-composition}, where we show how intensional,
+automatically certifiable properties of program components can be
+composed to obtain whole-program MPC security. We consider the GMW and
+YGC protocols as examplars, both of which use \emph{circuit}
+abstractions comprising \emph{gates} in arbitrarily large
+configurations, and show how gates can be automatically certified to
+ensure security in circuits that use them in a manner that scales to
+any circuit size. This approach is extensible, in that new gate
+implementations can be automatically certified to guarantee safe
+inclusion in circuit libraries.

### 2023-07-03 popl24/intro.tex
checkpoint

@@
   for space savings, since it's based on truth table
-  computation/Datalog semantics.}. In Section \ref{section-metalang}
-we define a new metaprogramming language $\metaprot$ that dynamically
-generates $\minifed$ protocols. It includes control and data
-structures and is able to express logical protocol components, and
-enjoys a type safety result that guarantees that generated protocols
-are semantically well-defined. Our key analysis results are presented
-in Section \ref{section-composition}, where we show how intensional,
+  computation/Datalog semantics.}.
+
+In Section \ref{section-metalang} we define a new metaprogramming
+language $\metaprot$ that dynamically generates $\minifed$
+protocols. It includes control and data structures and is able to
+express logical protocol components, and enjoys a type safety result
+that guarantees that generated protocols are semantically
+well-defined. Our key analysis results are presented in Section
+\ref{section-composition}, where we show how intensional,
 automatically certifiable properties of program components can be

### 2023-07-03 popl24/intro.tex
checkpoint

@@
 
-In \ref{section-minicat} we develop a new probabilistic programming
+In Section \ref{section-minicat} we develop a new probabilistic programming
 language $\minifed$ for defining synchronous distributed protocols

### 2023-07-03 popl24/intro.tex
checkpoint

@@
 inclusion in circuit libraries.
+
+In summary, our analysis takes a ``just do it'' approach- rather than
+using, e.g., traditional security type or linear type disciplines to
+prevent ``bad things'' from happening, we base our analysis on precise
+quantification of intensional protocol behavior and especially
+dependencies between secret inputs, public outputs, and adversarial
+views of protocols. We ameliorate scalability issues associated with
+fine-grained analysis by manually establishing compositional
+properties inspired by probabilistic separation logic
+\cite{barthe2019probabilistic}, obtaining a middle-ground, scalable
+``proof assistant'' style approach enabled by MPC protocol paradigms.

### 2023-07-04 popl24/scratch.tex
checkpoint

\subsection{Compositional Properties of GMW}

\input{gmw}
\cnote{Need to define type signatures.}

Our certification techniques are defined generally wrt implementations
of input encoding, decoding, and internal gates. In the following we
will refer to arbitrary decode, encode, and gate functions with the
restriction that any function in each of these categories has the same
valid type signature as $\ttt{decodegmw}$, $\ttt{encodegmw}$, and
$\ttt{andgmw}$ respectively.
\begin{definition}
  We write $\progtt(\prog)$ to denote the \emph{basic distribution} of
  $\prog$ as defined in Definition \ref{def-progd}. For any program $\prog =
  (\eassign{v_1}{\be_1};\ldots;\eassign{v_n}{\be_n})$, define
  $\vdefs(\prog) \defeq \{ v_1,\ldots,v_n \}$.
\end{definition}
We also need to formalize notions of probabilistic independence, aka
\emph{separation}, and correlation. We borrow the symbols $*$ and $\sim$
from \cite{barthe2019probabilistic} with the same denotations.
\begin{definition}
  We write $\vc{\pmf}{x}{y}$ iff $\pmf(\{ x \mapsto 0\}\ |\ \{ y \mapsto 0 \}) =
  \pmf(\{ x \mapsto 1\}\ |\ \{ y \mapsto 1 \}) = 1$.
  We write $\sep{\pmf}{Y}{Z}$ iff for all
    $\store \in \mems(X \cup Y)$:
  $$(\margd{\pmf}{Y \cup Z})(\store) =
  \pmf(\store_Y) * \pmf(\store_Z)$$ 
\end{definition}

\begin{lemma}
  \label{lemma-separation}
  The following properties hold:
  \begin{enumerate}
  \item $\sep{\pmf}{Y}{Z}$ iff $\sep{\pmf}{Z}{Y}$
  \item $\vc{\pmf}{x}{y}$ if  $\vc{\pmf}{y}{x}$
  \item $\vc{\pmf}{x}{y}$ and $\vc{\pmf}{y}{z}$ imply $\vc{\pmf}{x}{z}$
  \item $\sep{\pmf}{X}{Y}$ and $\sep{\pmf}{X}{Z}$ imply $\sep{\pmf}{X}{(Y \cup Z)}$
  \item If $\prog_1;\prog_2$ is safe and $\vars(\prog_1) \cap \vars(\prog_2) = \varnothing$
    then $\sep{\progtt(\prog_1;\prog_2)}{\vars(\prog_1)}{\vars(\prog_2)}$.
  \end{enumerate}
\end{lemma}

\begin{definition}[GMW Gate Certification]
  \label{definition-gmwgate-certification}
  Let $\mathit{gate}$ be a gate function and let $e$ be the following
  $\metaprot$ program:
  $$
  \begin{array}{l}
    \ttt{let in1 = \{c1 = flip[1,s11];c2 = flip[2,s12]\} in}\\
    \ttt{let in2 = \{c1 = flip[1,s21];c2 = flip[2,s22]\} in}\\
    \mathit{gate}\ttt{(g,in1,in2)}
  \end{array}
  $$
  and let $
  \config{\varnothing}{e}\redxs\config{\prog}{\ttt{\{c1 = v[1,gout]; c2 = v[2,gout]\}}}
  $.
  Then $\mathit{gate}$ is \emph{certified} iff the following conditions hold:
  \begin{enumerate}[\hspace{5mm}i.]
  \item $\sep{\progtt(\prog)}{\ttt{\{flip[1,s11],flip[1,s21]\}}}{\ttt{\{v[2,gout]\}}}$
  \item $\sep{\progtt(\prog)}{\ttt{\{flip[2,s12],flip[2,s22]\}}}{\ttt{\{v[1,gout]\}}}$
  \item All variables in:$$\vars(\prog) - \{ \ttt{flip[1,s11],flip[1,s21],flip[2,s12],flip[2,s22]}\}$$
    are distinguished by $\ttt{g}$ (i.e., contain $\ttt{g}$ as an identifier substring). 
  \end{enumerate}
\end{definition}

\begin{lemma}
  \label{lemma-sepjoin}
  $\sep{\progtt(\prog)}{X}{Y}$ iff for all 
  $\store^1, \store^2 \in \runs(\prog)$ there exists
  $\store \in \runs(\prog)$ with
  $\store^1_{X} \cap \store^2_{Y} \subseteq \store$.
\end{lemma}

\begin{lemma}
  \label{lemma-presub}
  Given $\be_1$, $\be_2$, and $i$ where $\vars(\be_1) \cap
  \vars(\be_2) = \varnothing$ and $\lcod{\store_1,\be_1}{i} = \beta$
  with $\store_1(x) = \lcod{\store_2,\be_2}{i}$ and
  $\dom(\store_1) = \vars(\be_1)$ and $\dom(\store_2) = \vars(\be_2)$.
  Then $\lcod{\store_1\cap\store_2,\be_1[\be_2/x]}{i} = \beta$.
\end{lemma}
\begin{proof}
By straightforward structural induction on $\be$.
\end{proof}

\begin{lemma}[Substitution]
  \label{lemma-substitution}
  If $\sep{\progtt(\prog_2)}{\{ f \}}{\vdefs(\prog_2)}$ and
  $\prog_1;\prog_2[\be/f]$ is safe with $\vars(\prog_1,\be) \cap
  \vars(\prog_2) = \varnothing$ then
  $\sep{\progtt(\prog_1;\prog_2[\be/f])}{\vdefs(\prog_1)}{\vdefs(\prog_2[\be/f])}$.
\end{lemma}
\begin{proof}
  Suppose on the contrary it was not the case that
  $\sep{\progtt(\prog_1;\prog_2[\be/f])}{\vdefs(\prog_1)}{\vdefs(\prog_2[\be/f])}$.
  Then by Lemma \ref{lemma-sepjoin} there exists $\store_1,\store_2
  \in \runs(\prog_1;\prog_2[\be/f])$ with no $\store \in
  \runs(\prog_1;\prog_2[\be/f])$ such that $\store^1_{\vdefs(\prog_1)}
  \cap \store^2_{\vdefs(\prog_2)} \subseteq \store$.  But also by
  assumption and Lemma \ref{lemma-sepjoin} for any $\beta$ there
  exists $\store' \in \runs(\prog_2)$ with $\{ f \mapsto \beta \} \cap
  \store^2_{\vdefs(\prog_2)} \subseteq \store'$. So in particular, we
  have $\{ f \mapsto \lcod{\store^1_{\vdefs(\prog_1)},\be}{i} \} \cap
  \store^2_{\vdefs(\prog_2)} \subseteq \store'$ for any $i$. This, the
  assumption $\vars(\prog_1,\be) \cap \vars(\prog_2) = \varnothing$,
  and application of Lemma \ref{lemma-presub} leads to the consequence
  that there exists $\store \in \runs(\prog_1;\prog_2[\be/f])$ with
  $\store \supseteq \store^1_{\vdefs(\prog_1)} \cap
  \store^2_{\vdefs(\prog_2)}$ given the assumption $\vars(\prog_1,\be)
  \cap \vars(\prog_2) = \varnothing$, which is a contradiction.
\end{proof}

\begin{lemma}[GMW Gate Composition]
  \label{lemma-composition-gmw}
  Given well-typed $\prog_1;E[\mathit{gate}(\ttt{g},v_1,v_2)]$ and certified $\mathit{gate}$ where:
  $$
  \begin{array}{lcr}
   v_1 &=& \ttt{\{c1 = v[1,g1out]; c2 = v[2,g1out]\}}\\
   v_2 &=& \ttt{\{c1 = v[1,g2out]; c2 = v[2,g2out]\}}
  \end{array}
  $$
  for some $\ttt{g1}$ and $\ttt{g2}$ and:
  $$
  \config{\prog_1}{E[\mathit{gate}(\ttt{g},v_1,v_2)]} \redxs \config{\prog_1;\prog_2}{E[\ttt{\{c1 = v[1,gout]; c2 = v[2,gout]\}}]}
  $$
  Then the following postconditions hold:
  \begin{enumerate}
  \item $\sep{\progtt(\prog_1;\prog_2)}{\ttt{\{v[1,g1out],v[1,g2out]\}}}{\ttt{\{v[2,gout]\}}}$
  \item $\sep{\progtt(\prog_1;\prog_2)}{\ttt{\{v[2,g1out],v[2,g2out]\}}}{\ttt{\{v[1,gout]\}}}$
  \end{enumerate}
\end{lemma}

\begin{proof}
  Let $\prog$ be as defined in Definition \ref{definition-gmwgate-certification}.
  We observe:
  $$
  {\footnotesize
    \begin{array}{c}
      \prog_2 = \\
      \prog[\ttt{v[1,g1out]}/\ttt{flip[1,s11]}][\ttt{v[2,g1out]}/\ttt{flip[2,s12]}][\ttt{v[1,g2out]}/\ttt{flip[1,s21]}][\ttt{v[2,g2out]}/\ttt{flip[2,s22]}]
    \end{array}
  }
  $$
  The assumption of well-typedness of $\prog_1;E[\mathit{gate}(\ttt{g},v_1,v_2)]$ assures
  that $\ttt{g}$ has not been previously used as a gate identifier, since otherwise the
  views $\ttt{v[1,gout]}$ and $\ttt{v[1,gout]}$ would have been previously defined. This
  ensures $\vars(\prog_1,x) \cap
  \vars(\prog_2) = \varnothing$ for:
  $$x \in \{ \ttt{v[1,g1out]}, \ttt{v[2,g1out]}, \ttt{v[1,g2out]}, \ttt{v[2,g2out]}\}$$
  given condition (iii) in Definition \ref{definition-gmwgate-certification}.
  The result follows by Lemma \ref{lemma-substitution}. 
\end{proof}

\begin{lemma}[GMW Gate Invariance]
  \label{lemma-preservation-gmw}
  Given well-typed $\prog_1;E[\mathit{gate}(\ttt{g},v_1,v_2)]$ and certified $\mathit{gate}$ where
  for some $\ttt{g1}$ and $\ttt{g2}$, :
  $$
  \begin{array}{lcr}
   v_1 &=& \ttt{\{c1 = v[1,g1out]; c2 = v[2,g1out]\}}\\
   v_2 &=& \ttt{\{c1 = v[1,g2out]; c2 = v[2,g2out]\}}
  \end{array}
  $$
  and where $\iov(\prog_1) = S \cup V$ we have:
  $$
  \sep{\progd(\prog_1)}{S_{\{2\}}}{V_{\{1\}}}
  \quad \text{and} \quad \sep{\progd(\prog_1)}{S_{\{1\}}}{V_{\{2\}}}
  $$
  and:
  $$
  \config{\prog_1}{E[\mathit{gate}(\ttt{g},v_1,v_2)]} \redxs \config{\prog_1;\prog_2}{E[\ttt{\{c1 = v[1,gout]; c2 = v[2,gout]\}}]}
  $$
  then we have as a postcondition:
  %\begin{enumerate}
  $$\sep{\progd(\prog_1;\prog_2)}{S_{\{2\}}}{(V_{\{1\}} \cup \ttt{\{v[1,gout]\}})}
  \quad \text{and}\quad \sep{\progd(\prog_1;\prog_2)}{S_{\{1\}}}{(V_{\{2\}} \cup \ttt{\{v[2,gout]\}})}$$
  %\end{enumerate}
\end{lemma}

\begin{proof}
By Lemma \ref{lemma-composition-gmw} we have:
$$\sep{\progtt(\prog_1;\prog_2)}{\ttt{\{v[1,g1out],v[1,g2out]\}}}{\ttt{\{v[2,gout]\}}}$$
and
$$\sep{\progtt(\prog_1;\prog_2)}{\ttt{\{v[2,g1out],v[2,g2out]\}}}{\ttt{\{v[1,gout]\}}}$$
\end{proof}

\begin{definition}[GMW Encode Certification]
  Let $\mathit{encode}$ be an encoding function, and let:
  $$
  \config{\varnothing}{\ttt{encode(s1,s2)}} \redxs
  \config{\prog}{\ttt{\{shares1 = }v_1\ttt{; shares2 = }v_2\ttt{\}}}
  $$
  where:
  $$
  \begin{array}{rcl}
    v_1 &=& \ttt{\{ c1 = v[1,s1out]; c2 = v[2,s1out] \}}\\
    v_2 &=& \ttt{\{ c1 = v[1,s2out]; c2 = v[2,s2out] \}}
  \end{array}
  $$
  Then $\mathit{encode}$ is \emph{certified} iff each of the following conditions hold:
  \begin{enumerate}[\hspace{5mm}i.]
  \item $\sep{\progtt(\prog)}{\ttt{\{ s[1,s1] \}}}{\ttt{\{ v[2,s1out], v[2,s2out] \}}}$
  \item $\sep{\progtt(\prog)}{\ttt{\{ s[2,s2] \}}}{\ttt{\{ v[1,s1out], v[1,s2out] \}}}$
  \item All $x \in \vars(\prog)$ are distinguished by $\ttt{in1}$ or $\ttt{in2}$. 
  \end{enumerate}
\end{definition}

\begin{theorem}
  \label{theorem-gmw-NIMO}
  If $e$ is a well-typed GMW circuit definition using certified
  components and $\config{\varnothing}{e} \redxs
  \config{\prog}{\varnothing}$, then $\NIMO(\prog)$.
\end{theorem}
\begin{proof}
  Let $\iov(\prog) = S \cup (V \cup \outv)$. Noting that any encoding
  necessarily uses entirely fresh variables by Definition and type
  safety, by applying Lemmas \ref{lemma-compose-gmw} and
  \ref{lemma-wire-framing} and the frame rule of
  \cite{barthe2019probabilistic} we obtain:
  $$
  \sep{\progd(\prog)}{(S_{\{2\}}\cup\{\outv\})}{V_{\{1\}}}
  \quad \text{and} \quad \sep{\progd(\prog)}{(S_{\{1\}}\cup\{\outv\})}{V_{\{2\}}}
  $$
  That is, the marginal distribution of client 1's secrets and the output are independent
  of client 2's views, and vice-versa. The result follows. 
\end{proof}

### 2023-07-04 popl24/scratch.tex
checkpoint

@@
   \item $\vc{\pmf}{x}{y}$ and $\vc{\pmf}{y}{z}$ imply $\vc{\pmf}{x}{z}$
-  \item $\sep{\pmf}{X}{Y}$ and $\sep{\pmf}{X}{Z}$ imply $\sep{\pmf}{X}{(Y \cup Z)}$
-  \item If $\prog_1;\prog_2$ is safe and $\vars(\prog_1) \cap \vars(\prog_2) = \varnothing$
-    then $\sep{\progtt(\prog_1;\prog_2)}{\vars(\prog_1)}{\vars(\prog_2)}$.
+  \item ($\sep{\pmf}{X}{Y}$ and $\sep{\pmf}{X}{Z}$) iff $\sep{\pmf}{X}{(Y \cup Z)}$
+  %\item If $\prog_1;\prog_2$ is safe and $\vars(\prog_1) \cap \vars(\prog_2) = \varnothing$
+  %  then $\sep{\progtt(\prog_1;\prog_2)}{\vars(\prog_1)}{\vars(\prog_2)}$.
   \end{enumerate}
 \end{lemma}
+
+\begin{lemma}
+  \label{lemma-sepjoin}
+  $\sep{\progtt(\prog)}{X}{Y}$ iff for all 
+  $\store^1, \store^2 \in \runs(\prog)$ there exists
+  $\store \in \runs(\prog)$ with
+  $\store^1_{X} \cap \store^2_{Y} \subseteq \store$.
+\end{lemma}
+
+%\begin{lemma}[Wire Framing]
+%  \label{lemma-wire-framing}
+%  If $\sep{\progtt(\prog_2)}{X}{Z}$ and
+%  $\vars(\prog_1) \cap \vars(\prog_2) \subseteq X$ then
+%  $\sep{\progtt(\prog_1;\prog_2)}{Y}{Z}$ for
+%  $Y \subseteq \vars(\prog_1)$.
+%\end{lemma}
 
@@
 \begin{lemma}
-  \label{lemma-sepjoin}
-  $\sep{\progtt(\prog)}{X}{Y}$ iff for all 
-  $\store^1, \store^2 \in \runs(\prog)$ there exists
-  $\store \in \runs(\prog)$ with
-  $\store^1_{X} \cap \store^2_{Y} \subseteq \store$.
+  \label{lemma-gmw-frame}
+  If $\sep{\progtt(\prog_2)}{Y}{Z}$ and $\sep{\progtt(\prog_1)}{X}{Y}$ for
+  $X \not\in \vars(\prog_2)$ then $\sep{\progtt(\prog_1;\prog_2)}{X}{Z}$. 
 \end{lemma}
@@
   \label{lemma-substitution}
-  If $\sep{\progtt(\prog_2)}{\{ f \}}{\vdefs(\prog_2)}$ and
+  If $\sep{\progtt(\prog_2)}{\{ f \}}{X}$ and
   $\prog_1;\prog_2[\be/f]$ is safe with $\vars(\prog_1,\be) \cap
-  \vars(\prog_2) = \varnothing$ then
-  $\sep{\progtt(\prog_1;\prog_2[\be/f])}{\vdefs(\prog_1)}{\vdefs(\prog_2[\be/f])}$.
+  X = \varnothing$ then
+  $\sep{\progtt(\prog_1;\prog_2[\be/f])}{\vars(\be)}{X}$.
 \end{lemma}
@@
   Suppose on the contrary it was not the case that
-  $\sep{\progtt(\prog_1;\prog_2[\be/f])}{\vdefs(\prog_1)}{\vdefs(\prog_2[\be/f])}$.
+  $\sep{\progtt(\prog_1;\prog_2[\be/f])}{\vars(\be)}{X}$.
   Then by Lemma \ref{lemma-sepjoin} there exists $\store_1,\store_2
   \in \runs(\prog_1;\prog_2[\be/f])$ with no $\store \in
-  \runs(\prog_1;\prog_2[\be/f])$ such that $\store^1_{\vdefs(\prog_1)}
-  \cap \store^2_{\vdefs(\prog_2)} \subseteq \store$.  But also by
+  \runs(\prog_1;\prog_2[\be/f])$ such that $\store^1_{\vars(\be)}
+  \cap \store^2_X \subseteq \store$.  But also by
   assumption and Lemma \ref{lemma-sepjoin} for any $\beta$ there
   exists $\store' \in \runs(\prog_2)$ with $\{ f \mapsto \beta \} \cap
-  \store^2_{\vdefs(\prog_2)} \subseteq \store'$. So in particular, we
-  have $\{ f \mapsto \lcod{\store^1_{\vdefs(\prog_1)},\be}{i} \} \cap
-  \store^2_{\vdefs(\prog_2)} \subseteq \store'$ for any $i$. This, the
+  \store^2_X \subseteq \store'$. So in particular, we
+  have $\{ f \mapsto \lcod{\store^1_{\vars(\be)},\be}{i} \} \cap
+  \store^2_X \subseteq \store'$ for any $i$. This, the
   assumption $\vars(\prog_1,\be) \cap \vars(\prog_2) = \varnothing$,
@@
   that there exists $\store \in \runs(\prog_1;\prog_2[\be/f])$ with
-  $\store \supseteq \store^1_{\vdefs(\prog_1)} \cap
-  \store^2_{\vdefs(\prog_2)}$ given the assumption $\vars(\prog_1,\be)
+  $\store \supseteq \store^1_{\vars(\be)} \cap
+  \store^2_X$ given the assumption $\vars(\prog_1,\be)
   \cap \vars(\prog_2) = \varnothing$, which is a contradiction.
@@
 
-\begin{lemma}[GMW Gate Composition]
-  \label{lemma-composition-gmw}
-  Given well-typed $\prog_1;E[\mathit{gate}(\ttt{g},v_1,v_2)]$ and certified $\mathit{gate}$ where:
+\begin{lemma}[GMW Gate Invariant]
+  \label{lemma-preservation-gmw}
+  Given well-typed $\prog_1;E[\mathit{gate}(\ttt{g},v_1,v_2)]$ and certified $\mathit{gate}$ where
+  for some $\ttt{g1}$ and $\ttt{g2}$, :
   $$
@@
   $$
-  for some $\ttt{g1}$ and $\ttt{g2}$ and:
+  and let:
   $$
@@
   $$
-  Then the following postconditions hold:
-  \begin{enumerate}
-  \item $\sep{\progtt(\prog_1;\prog_2)}{\ttt{\{v[1,g1out],v[1,g2out]\}}}{\ttt{\{v[2,gout]\}}}$
-  \item $\sep{\progtt(\prog_1;\prog_2)}{\ttt{\{v[2,g1out],v[2,g2out]\}}}{\ttt{\{v[1,gout]\}}}$
-  \end{enumerate}
+  If we have the following precondition, where $\iov(\prog_1) = S \cup V$:
+  $$
+  \sep{\progd(\prog_1)}{(S_{\{2\}} \cup V_{\{2\}})}{(S_{\{1\}} \cup V_{\{1\}})}
+  $$
+  then we have as a postcondition:
+  $$
+  \sep{\progd(\prog_1)}{(S_{\{2\}} \cup V_{\{2\}} \cup \{ \ttt{v[2,gout]} \})}{(S_{\{1\}} \cup V_{\{1\}} \cup \{ \ttt{v[1,gout]} \})}
+  $$
 \end{lemma}
@@
   }
-  $$
-  The assumption of well-typedness of $\prog_1;E[\mathit{gate}(\ttt{g},v_1,v_2)]$ assures
-  that $\ttt{g}$ has not been previously used as a gate identifier, since otherwise the
-  views $\ttt{v[1,gout]}$ and $\ttt{v[1,gout]}$ would have been previously defined. This
-  ensures $\vars(\prog_1,x) \cap
-  \vars(\prog_2) = \varnothing$ for:
+  $$ Now, the assumption of well-typedness of
+  $\prog_1;E[\mathit{gate}(\ttt{g},v_1,v_2)]$ assures that $\ttt{g}$
+  has not been previously used as a gate identifier in $\prog_1$,
+  since otherwise the views $\ttt{v[1,gout]}$ and $\ttt{v[2,gout]}$
+  would have been previously defined. This ensures $\vars(\prog_1,x)
+  \cap \vars(\prog_2) = \varnothing$ for:
   $$x \in \{ \ttt{v[1,g1out]}, \ttt{v[2,g1out]}, \ttt{v[1,g2out]}, \ttt{v[2,g2out]}\}$$
   given condition (iii) in Definition \ref{definition-gmwgate-certification}.
-  The result follows by Lemma \ref{lemma-substitution}. 
-\end{proof}
-
-\begin{lemma}[GMW Gate Invariance]
-  \label{lemma-preservation-gmw}
-  Given well-typed $\prog_1;E[\mathit{gate}(\ttt{g},v_1,v_2)]$ and certified $\mathit{gate}$ where
-  for some $\ttt{g1}$ and $\ttt{g2}$, :
-  $$
-  \begin{array}{lcr}
-   v_1 &=& \ttt{\{c1 = v[1,g1out]; c2 = v[2,g1out]\}}\\
-   v_2 &=& \ttt{\{c1 = v[1,g2out]; c2 = v[2,g2out]\}}
-  \end{array}
-  $$
-  and where $\iov(\prog_1) = S \cup V$ we have:
-  $$
-  \sep{\progd(\prog_1)}{S_{\{2\}}}{V_{\{1\}}}
-  \quad \text{and} \quad \sep{\progd(\prog_1)}{S_{\{1\}}}{V_{\{2\}}}
-  $$
-  and:
-  $$
-  \config{\prog_1}{E[\mathit{gate}(\ttt{g},v_1,v_2)]} \redxs \config{\prog_1;\prog_2}{E[\ttt{\{c1 = v[1,gout]; c2 = v[2,gout]\}}]}
-  $$
-  then we have as a postcondition:
-  %\begin{enumerate}
-  $$\sep{\progd(\prog_1;\prog_2)}{S_{\{2\}}}{(V_{\{1\}} \cup \ttt{\{v[1,gout]\}})}
-  \quad \text{and}\quad \sep{\progd(\prog_1;\prog_2)}{S_{\{1\}}}{(V_{\{2\}} \cup \ttt{\{v[2,gout]\}})}$$
-  %\end{enumerate}
-\end{lemma}
-
-\begin{proof}
-By Lemma \ref{lemma-composition-gmw} we have:
+  Thus by Lemma \ref{lemma-substitution} we have:
   $$\sep{\progtt(\prog_1;\prog_2)}{\ttt{\{v[1,g1out],v[1,g2out]\}}}{\ttt{\{v[2,gout]\}}}$$
@@
   $$\sep{\progtt(\prog_1;\prog_2)}{\ttt{\{v[2,g1out],v[2,g2out]\}}}{\ttt{\{v[1,gout]\}}}$$
+  Since $\vars(\prog_1) \cap \vars(\prog_2)$ are only the wire inputs to $\ttt{g}$,
+  therefore by assumption and Lemma \ref{lemma-gmw-frame}:
+  $$\sep{\progtt(\prog_1;\prog_2)}{(S_{\{1\}} \cup V_{\{1\}})}{\ttt{\{v[2,gout]\}}}$$
+  and also:
+  $$\sep{\progtt(\prog_1;\prog_2)}{\ttt{\{v[1,g1out],v[1,g2out]\}}}{(S_{\{2\}}\cup V_{\{2\}})}$$
+  etc., so the result follows. 
 \end{proof}

### 2023-07-04 popl24/scratch.tex
checkpoint

@@
   Since $\vars(\prog_1) \cap \vars(\prog_2)$ are only the wire inputs to $\ttt{g}$,
-  therefore by assumption and Lemma \ref{lemma-gmw-frame}:
-  $$\sep{\progtt(\prog_1;\prog_2)}{(S_{\{1\}} \cup V_{\{1\}})}{\ttt{\{v[2,gout]\}}}$$
-  and also:
-  $$\sep{\progtt(\prog_1;\prog_2)}{\ttt{\{v[1,g1out],v[1,g2out]\}}}{(S_{\{2\}}\cup V_{\{2\}})}$$
-  etc., so the result follows. 
+  therefore by assumption and Lemma \ref{lemma-gmw-frame} the following assertions are
+  valid:
+  \begin{mathpar}
+    \sep{\progtt(\prog_1;\prog_2)}{(S_{\{1\}} \cup V_{\{1\}})}{\ttt{\{v[2,gout]\}}}
+
+    \sep{\progtt(\prog_1;\prog_2)}{\ttt{\{v[1,g1out],v[1,g2out]\}}}{(S_{\{2\}}\cup V_{\{2\}})}
+
+    \sep{\progtt(\prog_1;\prog_2)}{(S_{\{2\}} \cup V_{\{2\}})}{\ttt{\{v[1,gout]\}}}
+
+    \sep{\progtt(\prog_1;\prog_2)}{\ttt{\{v[2,g1out],v[2,g2out]\}}}{(S_{\{1\}}\cup V_{\{1\}})}
+  \end{mathpar}
+  The result follows by Lemma \ref{lemma-separation}.
 \end{proof}

### 2023-07-05 popl24/scratch.tex
checkpoint

@@
   \label{lemma-gmw-frame}
-  If $\sep{\progtt(\prog_2)}{Y}{Z}$ and $\sep{\progtt(\prog_1)}{X}{Y}$ for
-  $X \not\in \vars(\prog_2)$ then $\sep{\progtt(\prog_1;\prog_2)}{X}{Z}$. 
+  If $\sep{\progtt(\prog_1;\prog_2)}{Y}{V}$ for $V \in \vdefs(\prog_2)$ and $\sep{\progtt(\prog_1)}{X}{Y}$ for
+  $X \not\in \vars(\prog_2)$ then $\sep{\progtt(\prog_1;\prog_2)}{X}{V}$. 
 \end{lemma}
+\begin{proof}
+  By assumption and Lemma \ref{lemma-sepjoin}, for all
+  $\store^1,\store^2 \in \runs(\prog_1;\prog_2)$ we have $\store^1_Y
+  \cap \store^2_V \subseteq \store$ for some $\store_a \in
+  \runs(\prog_1;\prog_2)$ and $\store^1_X \cap \store^2_Y \subseteq
+  \store$ for some $\store_b \in \runs(\prog_1;\prog_2)$.  Thus, since
+  we assume $X \not\in \vars(\prog_2)$ we have $\store_a \cap \store_b
+  \in \runs(\prog_1;\prog_2)$ and $\store^1_X \cap \store^2_V =
+  \store_a \cap \store_b$, obtaining the result by
+  \ref{lemma-sepjoin}.
+\end{proof}
 
@@
   \label{lemma-preservation-gmw}
-  Given well-typed $\prog_1;E[\mathit{gate}(\ttt{g},v_1,v_2)]$ and certified $\mathit{gate}$ where
+  Given well-typed $\prog_1;E[\mathit{gate}(\ttt{g},v_1,v_2)]$ for certified $\mathit{gate}$ where
   for some $\ttt{g1}$ and $\ttt{g2}$, :
@@
   $$
-  \sep{\progd(\prog_1)}{(S_{\{2\}} \cup V_{\{2\}} \cup \{ \ttt{v[2,gout]} \})}{(S_{\{1\}} \cup V_{\{1\}} \cup \{ \ttt{v[1,gout]} \})}
+  \sep{\progd(\prog_1;\prog_2)}{(S_{\{2\}} \cup V_{\{2\}} \cup \{ \ttt{v[2,gout]} \})}{(S_{\{1\}} \cup V_{\{1\}} \cup \{ \ttt{v[1,gout]} \})}
   $$
@@
 \begin{definition}[GMW Encode Certification]
+  \label{definition-gmwencode-certification}
   Let $\mathit{encode}$ be an encoding function, and let:
@@
   \begin{enumerate}[\hspace{5mm}i.]
-  \item $\sep{\progtt(\prog)}{\ttt{\{ s[1,s1] \}}}{\ttt{\{ v[2,s1out], v[2,s2out] \}}}$
-  \item $\sep{\progtt(\prog)}{\ttt{\{ s[2,s2] \}}}{\ttt{\{ v[1,s1out], v[1,s2out] \}}}$
+  \item $\sep{\progtt(\prog)}{\ttt{\{ s[1,s1],  v[1,s1out], v[1,s2out] \}}}{\ttt{\{ s[2,s2], v[2,s1out], v[2,s2out] \}}}$
   \item All $x \in \vars(\prog)$ are distinguished by $\ttt{in1}$ or $\ttt{in2}$. 
@@
 \end{definition}
+
+\begin{lemma}[GMW Encode Invariant] 
+  Given well-typed $\prog_1;E[\mathit{encode}\ttt{(s1,s2)}]$ for certified $\mathit{encode}$, let:
+  $$
+  \config{\prog_1}{E[\mathit{encode}(\ttt{s1},\ttt{s2})]} \redxs
+  \config{\prog_1;\prog_2}{E[\ttt{\{shares1 = }v_1\ttt{; shares2 = }v_2\ttt{\}}]}
+  $$
+  If we have the following precondition, where $\iov(\prog_1) = S \cup V$:
+  $$
+  \sep{\progd(\prog_1)}{(S_{\{2\}} \cup V_{\{2\}})}{(S_{\{1\}} \cup V_{\{1\}})}
+  $$
+  then we have as a postcondition, where $\iov(\prog_1;\prog_2) = S' \cup V'$:
+  $$
+  \sep{\progd(\prog_1;\prog_2)}{(S'_{\{2\}} \cup V'_{\{2\}})}{(S'_{\{1\}} \cup V'_{\{1\}})}
+  $$
+\end{lemma}
+
+\begin{proof}
+  By Definition \ref{definition-gmwencode-certification} we have:
+  $$\sep{\progtt(\prog)}{\ttt{\{ s[1,s1],  v[1,s1out], v[1,s2out] \}}}{\ttt{\{ s[2,s2], v[2,s1out], v[2,s2out] \}}}$$
+  Since we assume that secrets are in uniform and independent marginal distributions a priori,
+  and $\vars(\prog_1) \cap \vars(\prog_2)$ by condition (ii) of
+  Definition \ref{definition-gmwencode-certification}, thus by assumption and Lemma \ref{lemma-separation}:
+  $$\sep{\progtt(\prog)}{(S_1 \cup V_1 \cup \ttt{\{ s[1,s1],  v[1,s1out], v[1,s2out] \}})}{(S_2 \cup V_2 \cup \ttt{\{ s[2,s2], v[2,s1out], v[2,s2out] \}}}$$
+  which was to be proven.
+\end{proof}
 

### 2023-07-05 popl24/scratch.tex
checkpoint

--- 2023-07-04 popl24/scratch.tex
@@
   \item $\vc{\pmf}{x}{y}$ and $\vc{\pmf}{y}{z}$ imply $\vc{\pmf}{x}{z}$
-  \item ($\sep{\pmf}{X}{Y}$ and $\sep{\pmf}{X}{Z}$) iff $\sep{\pmf}{X}{(Y \cup Z)}$
+  \item $\sep{\pmf}{X}{(Y \cup Z)}$ implies $(\sep{\pmf}{X}{Y}$ and $\sep{\pmf}{X}{Z})$
+  \item $(\sep{\pmf}{X}{Y}$ and $\sep{\pmf}{X}{Z}$ and $\sep{\pmf}{Y}{Z})$ implies $\sep{\pmf}{X}{(Y \cup Z)}$ 
   %\item If $\prog_1;\prog_2$ is safe and $\vars(\prog_1) \cap \vars(\prog_2) = \varnothing$
@@
   \begin{enumerate}[\hspace{5mm}i.]
-  \item $\sep{\progtt(\prog)}{\ttt{\{flip[1,s11],flip[1,s21]\}}}{\ttt{\{v[2,gout]\}}}$
-  \item $\sep{\progtt(\prog)}{\ttt{\{flip[2,s12],flip[2,s22]\}}}{\ttt{\{v[1,gout]\}}}$
+  \item $\sep{\progtt(\prog)}{\ttt{\{flip[1,s11],flip[1,s21]\}}}{\ttt{\{v[1,gout],v[2,gout]\}}}$
+  \item $\sep{\progtt(\prog)}{\ttt{\{flip[2,s12],flip[2,s22]\}}}{\ttt{\{v[1,gout],v[2,gout]\}}}$
+  % \item $\sep{\progtt(\prog)}{\ttt{\{v[1,gout]\}}}{\ttt{\{v[2,gout]\}}}$
   \item All variables in:$$\vars(\prog) - \{ \ttt{flip[1,s11],flip[1,s21],flip[2,s12],flip[2,s22]}\}$$
@@
 \begin{lemma}[GMW Gate Invariant]
-  \label{lemma-preservation-gmw}
-  Given well-typed $\prog_1;E[\mathit{gate}(\ttt{g},v_1,v_2)]$ and certified $\mathit{gate}$ where
-  for some $\ttt{g1}$ and $\ttt{g2}$, :
+  \label{lemma-gmw-preservation}
+  Given well-typed $\prog_1;E[\mathit{gate}(\ttt{g},v_1,v_2)]$ for certified $\mathit{gate}$ where:
   $$
-  \begin{array}{lcr}
-   v_1 &=& \ttt{\{c1 = v[1,g1out]; c2 = v[2,g1out]\}}\\
-   v_2 &=& \ttt{\{c1 = v[1,g2out]; c2 = v[2,g2out]\}}
-  \end{array}
+  \config{\prog_1}{E[\mathit{gate}(\ttt{g},v_1,v_2)]} \redxs \config{\prog_1;\prog_2}{E[v]}
   $$
-  and let:
-  $$
-  \config{\prog_1}{E[\mathit{gate}(\ttt{g},v_1,v_2)]} \redxs \config{\prog_1;\prog_2}{E[\ttt{\{c1 = v[1,gout]; c2 = v[2,gout]\}}]}
-  $$
-  If we have the following precondition, where $\iov(\prog_1) = S \cup V$:
-  $$
-  \sep{\progd(\prog_1)}{(S_{\{2\}} \cup V_{\{2\}})}{(S_{\{1\}} \cup V_{\{1\}})}
-  $$
-  then we have as a postcondition:
-  $$
-  \sep{\progd(\prog_1)}{(S_{\{2\}} \cup V_{\{2\}} \cup \{ \ttt{v[2,gout]} \})}{(S_{\{1\}} \cup V_{\{1\}} \cup \{ \ttt{v[1,gout]} \})}
-  $$
+  If we have the following preconditions, where $\iov(\prog_1) = S \cup V$:
+  \begin{enumerate}[\hspace{5mm}i.]
+  \item $\sep{\progd(\prog_1)}{S_{\{2\}}}{V_{\{1\}}}$
+  \item $\sep{\progd(\prog_1)}{S_{\{1\}}}{V_{\{2\}}}$
+  \end{enumerate}
+  then we have as a postconditions, where $\iov(\prog_1;\prog_2) = S \cup V'$:
+  \begin{enumerate}[\hspace{5mm}i.]
+  \item $\sep{\progd(\prog_1;\prog_2)}{S_{\{2\}}}{V'_{\{1\}}}$
+  \item $\sep{\progd(\prog_1;\prog_2)}{S_{\{1\}}}{V'_{\{2\}}}$
+  \end{enumerate}
 \end{lemma}
@@
 \begin{proof}
+  Given well-typedness of $\prog_1;E[\mathit{gate}(\ttt{g},v_1,v_2)]$, we have for some $\ttt{g1}$ and $\ttt{g2}$:
+  $$
+  \begin{array}{lcl}
+   v_1 &=& \ttt{\{c1 = v[1,g1out]; c2 = v[2,g1out]\}}\\
+   v_2 &=& \ttt{\{c1 = v[1,g2out]; c2 = v[2,g2out]\}}\\
+    v &=& \ttt{\{c1 = v[1,gout]; c2 = v[2,gout]\}}
+  \end{array}
+  $$
   Let $\prog$ be as defined in Definition \ref{definition-gmwgate-certification}.
@@
   }
-  $$ Now, the assumption of well-typedness of
-  $\prog_1;E[\mathit{gate}(\ttt{g},v_1,v_2)]$ assures that $\ttt{g}$
+  $$
+  Now, the assumption of well-typedness also assures that $\ttt{g}$
   has not been previously used as a gate identifier in $\prog_1$,
@@
   would have been previously defined. This ensures $\vars(\prog_1,x)
-  \cap \vars(\prog_2) = \varnothing$ for:
+  \cap \vars(\prog) = \varnothing$ for:
   $$x \in \{ \ttt{v[1,g1out]}, \ttt{v[2,g1out]}, \ttt{v[1,g2out]}, \ttt{v[2,g2out]}\}$$
-  given condition (iii) in Definition \ref{definition-gmwgate-certification}.
+  given condition (iii) of Definition \ref{definition-gmwgate-certification}.
   Thus by Lemma \ref{lemma-substitution} we have:
-  $$\sep{\progtt(\prog_1;\prog_2)}{\ttt{\{v[1,g1out],v[1,g2out]\}}}{\ttt{\{v[2,gout]\}}}$$
+  $$\sep{\progtt(\prog_1;\prog_2)}{\ttt{\{v[1,g1out],v[1,g2out]\}}}{\ttt{\{v[1,gout],v[2,gout]\}}}$$
   and
-  $$\sep{\progtt(\prog_1;\prog_2)}{\ttt{\{v[2,g1out],v[2,g2out]\}}}{\ttt{\{v[1,gout]\}}}$$
-  Since $\vars(\prog_1) \cap \vars(\prog_2)$ are only the wire inputs to $\ttt{g}$,
-  therefore by assumption and Lemma \ref{lemma-gmw-frame}:
-  $$\sep{\progtt(\prog_1;\prog_2)}{(S_{\{1\}} \cup V_{\{1\}})}{\ttt{\{v[2,gout]\}}}$$
-  and also:
-  $$\sep{\progtt(\prog_1;\prog_2)}{\ttt{\{v[1,g1out],v[1,g2out]\}}}{(S_{\{2\}}\cup V_{\{2\}})}$$
-  etc., so the result follows. 
+  $$\sep{\progtt(\prog_1;\prog_2)}{\ttt{\{v[2,g1out],v[2,g2out]\}}}{\ttt{\{v[1,gout],v[2,gout]\}}}$$
+  And by Lemma \ref{lemma-gmw-frame} and condition (iii) of Definition
+  \ref{definition-gmwgate-certification} we then have:
+  $$\sep{\progtt(\prog_1;\prog_2)}{S_{\{1\}}}{\ttt{\{v[2,gout]\}}} \quad \text{and} \quad
+  \sep{\progtt(\prog_1;\prog_2)}{S_{\{2\}}}{\ttt{\{v[1,gout]\}}}$$
+  so by precondition assumptions and Lemma \ref{lemma-separation} we have:
+  $$\sep{\progtt(\prog_1;\prog_2)}{S_{\{1\}}}{V'_{\{2\}}} \quad \text{and} \quad
+  \sep{\progtt(\prog_1;\prog_2)}{S_{\{2\}}}{V'_{\{1\}}}$$
 \end{proof}
@@
   \begin{enumerate}[\hspace{5mm}i.]
-  \item $\sep{\progtt(\prog)}{\ttt{\{ s[1,s1] \}}}{\ttt{\{ v[2,s1out], v[2,s2out] \}}}$
-  \item $\sep{\progtt(\prog)}{\ttt{\{ s[2,s2] \}}}{\ttt{\{ v[1,s1out], v[1,s2out] \}}}$
+    %\item $\sep{\progtt(\prog)}{\ttt{\{ s[1,s1],  v[1,s1out], v[1,s2out] \}}}{\ttt{\{ s[2,s2], v[2,s1out], v[2,s2out] \}}}$
+  \item $\sep{\progtt(\prog)}{\ttt{\{s[1,s1]\}}}{\ttt{\{v[2,s1out],v[2,s2out]\}}}$
+  \item $\sep{\progtt(\prog)}{\ttt{\{s[2,s2]\}}}{\ttt{\{v[1,s1out],v[1,s2out]\}}}$
   \item All $x \in \vars(\prog)$ are distinguished by $\ttt{in1}$ or $\ttt{in2}$. 
   \end{enumerate}
+
 \end{definition}
+
+\begin{lemma}[GMW Encode Invariant]
+  \label{lemma-gmw-encode}
+  Given well-typed $\prog_1;E[\mathit{encode}\ttt{(s1,s2)}]$ for certified $\mathit{encode}$, let:
+  $$
+  \config{\prog_1}{E[\mathit{encode}(\ttt{s1},\ttt{s2})]} \redxs
+  \config{\prog_1;\prog_2}{E[v]}
+  $$
+  If we have the following preconditions, where $\iov(\prog_1) = S' \cup V$:
+  \begin{enumerate}[\hspace{5mm}i.]
+  \item $\sep{\progd(\prog_1)}{S_{\{2\}}}{V_{\{1\}}}$
+  \item $\sep{\progd(\prog_1)}{S_{\{1\}}}{V_{\{2\}}}$
+  \end{enumerate}
+  then we have as a postconditions, where $\iov(\prog_1;\prog_2) = S' \cup V'$: 
+  \begin{enumerate}[\hspace{5mm}i.]
+  \item $\sep{\progd(\prog_1;\prog_2)}{S'_{\{2\}}}{V'_{\{1\}}}$
+  \item $\sep{\progd(\prog_1;\prog_2)}{S'_{\{1\}}}{V'_{\{2\}}}$
+  \end{enumerate}
+\end{lemma}
+
+\begin{proof}
+  By assumptions of well-typedness we have:
+  $$
+  \begin{array}{rcl}
+    v &=&  \ttt{\{shares1 = }v_1\ttt{; shares2 = }v_2\ttt{\}}\\
+    v_1 &=& \ttt{\{ c1 = v[1,s1out]; c2 = v[2,s1out] \}}\\
+    v_2 &=& \ttt{\{ c1 = v[1,s2out]; c2 = v[2,s2out] \}}
+  \end{array}
+  $$
+  And by Definition \ref{definition-gmwencode-certification} we have:
+  \begin{mathpar}
+    \sep{\progtt(\prog)}{\ttt{\{s[1,s1]\}}}{\ttt{\{v[2,s1out],v[2,s2out]\}}}
+
+    \sep{\progtt(\prog)}{\ttt{\{s[2,s2]\}}}{\ttt{\{v[1,s1out],v[1,s2out]\}}}
+  \end{mathpar}
+  Since we assume that secrets are in uniform and independent marginal
+  distributions a priori, and $\vars(\prog_1) \cap \vars(\prog_2) =
+  \varnothing$ by condition (iii) of Definition 
+  \ref{definition-gmwencode-certification} and assumptions of well-typedness, 
+  the result follows by preconditions and Lemma \ref{lemma-separation}.
+\end{proof}
 
@@
   \label{theorem-gmw-NIMO}
-  If $e$ is a well-typed GMW circuit definition using certified
-  components and $\config{\varnothing}{e} \redxs
+  If $\eassign{\outv}{\ttt{decode}(e)}$ is a well-typed GMW circuit
+  definition using certified components and
+  $\config{\varnothing}{\eassign{\outv}{\ttt{decode}(e)}} \redxs
   \config{\prog}{\varnothing}$, then $\NIMO(\prog)$.
@@
 \begin{proof}
-  Let $\iov(\prog) = S \cup (V \cup \outv)$. Noting that any encoding
-  necessarily uses entirely fresh variables by Definition and type
-  safety, by applying Lemmas \ref{lemma-compose-gmw} and
-  \ref{lemma-wire-framing} and the frame rule of
-  \cite{barthe2019probabilistic} we obtain:
+  Let $\iov(\prog) = S \cup (V \cup \outv)$. By definition of $\ttt{decode}$
+  and assumptions of well-typedness we have for some $\ttt{g}$:
   $$
-  \sep{\progd(\prog)}{(S_{\{2\}}\cup\{\outv\})}{V_{\{1\}}}
-  \quad \text{and} \quad \sep{\progd(\prog)}{(S_{\{1\}}\cup\{\outv\})}{V_{\{2\}}}
+  \config{\varnothing}{\eassign{\outv}{\ttt{decode}(e)}} \redxs
+  \config{\prog;\eassign{\outv}{\ttt{v[1,gout] xor v[2,gout]}}}{\varnothing}
   $$
-  That is, the marginal distribution of client 1's secrets and the output are independent
-  of client 2's views, and vice-versa. The result follows. 
+  where by Lemmas \ref{lemma-gmw-encode} and \ref{lemma-gmw-preservation} we
+  have $\sep{\progd(\prog)}{S_{\{1\}}}{V_{\{2\}}}$ and
+  $\sep{\progd(\prog)}{S_{\{2\}}}{V_{\{1\}}}$. 
 \end{proof}

### 2023-07-06 popl24/composition.tex
checkpoint

--- 2023-06-28 popl24/composition.tex
@@
 
-Using YGC as a case study, in this Section we will discuss
-compositional properties that can be automatically verified on small
-program components using techniques developed in Section
-\ref{section-bruteforce}. These properties compose to yield $\NIMO$
-and scale to arbitrary programs. In our metatheory we will express
-these properties as pre-~and post-conditions on components, so our
-general approach is in a proof assistant style.
+\input{gmw}
 
-\subsection{A Library for Yao's Garbled Circuits}
-\label{section-metalang-ygc}
+The Yao's Garbled Circuits (YGC) and GMW protocols are examples of a
+common idiom in MPC protocol design- arbitrarily large \emph{circuits}
+are built up from smaller boolean or arithmetic \emph{gates} that
+operate on encodings of input secrets. These circuits can become quite
+large in practice \cite{kreuter2012billion}, but are built from
+well-defined components. Brute force verification techniques as
+described in Section \ref{section-bruteforce} are not scalable to
+large circuits, but in this Section we will demonstrate that correctly
+designed circuit components enjoy compositional properties
+that can be automatically certification on small program components
+using brute force methods. We prove that these properties preserve
+invariants throughout circuit definition that yield $\NIMO$ and thus
+scale to arbitrary programs. Furthermore, the certifications can be
+used on arbitrary gate definitions, allowing new gate implementations
+to be added soundly to libraries.
 
-\input{ygc}
+In this Section we develop and use implementations of both 2-party GMW
+and YGC to illustrate our approach and demonstrate its generality.
+Main results are in Theorems \ref{theorem-gmw-NIMO} and
+\ref{theorem-ygc-NIMO}.
 
-In Figures \ref{fig-ygc-aux} and \ref{fig-ygc-gates} we define a
-codebase for garbled circuits. This definition follows the
-\emph{point-and-permute} method described in \cite{evans2018pragmatic} and elsewhere,
-to which the reader is referred for more in-depth discussion.
-In this implementation client 2 is the \emph{garbler} and
-client 1 is the \emph{evaluator}. The garbler builds the garbled
-tables and shares them with the evaluator, who then evaluates
-the gate in an oblivious fashion until the final public output is
-generated through decryption. This definition is well-typed,
-with input type annotations for top-level functions listed in
-Figure \ref{fig-ygc-types}. Well-typed programs using these
-libraries are therefore guaranteed to yield safe $\minifed$
-programs.
+\subsection{Compositional Properties of GMW}
 
-\emph{Wire labels} are fundamental to YGC, and essentially represent
-gate output values in an encrypted form. In our definition, wire
-labels are represented by records $\ttt{\{ k = }\beta_1\ttt{; p =
-}\beta_2\ttt{ \}}$, where $\ttt{k}$ is the \emph{key bit} and
-$\ttt{p}$ is the \emph{pointer bit}, and $\beta_1$ and $\beta_2$ are
-flips. Flips in each output wire label are owned by the garbler and
-are unique per gate by definition of their identifying string, and the
-representation of $0$ is the negation of $1$. For example, here is the
-representation of 1 and 0 respectively in the output wire label for a
-hypothetical gate 6:
-\begin{mathpar}
-  \ttt{\{ k = flip[2,gate:6.k]; p =  flip[2,gate:6.p]] \}}
+The GMW protocol uses secret sharing to represent data flowing through
+circuits. In the 2-party case, clients 1 and 2 each share their input
+secrets, and use those shares to represent inputs to gates. Outputs
+are also represented as shares. We refer to the pair of shares
+representing any particular value as a \emph{wire value}, and
+we represent them via records of the form
+$
+\ttt{\{c1 = } v_1\ttt{;c2 = } v_2\ttt{\}} 
+$
+where $v_1$ and $v_2$ are client 1 and 2's shares respectively.
 
-  \ttt{\{ k = not flip[2,gate:6.k]; p =  not flip[2,gate:6.p]] \}}
-\end{mathpar}
-The pointer bits in wire labels are used to select permuted rows in
-table garblings. The key bits are used to identify a unique key for
-table row in each garbled gate. Intuitively, if $\beta_1$ and
-$\beta_2$ are either key or pointer bits encoding 1 on two input wire
-labels to a binary gate, rows and keys in the gate are enumerated in
-the order:
-$$
-\neg\beta_1\neg\beta_2,\ \neg\beta_1\beta_2,\ \beta_1\neg\beta_2,\ \beta_1\beta_2
-$$
+For full details of the GMW protocol the reader is referred to
+\cite{evans2018pragmatic}. Our implementation libary is shown in
+Figure \ref{fig-gmw}, with type signatures for the library functions
+shown in Figure \ref{fig-gmw-types}. We show the And gate since it is
+an interesting component. The Figure includes the
+following top-level functions:
+\begin{itemize}
+\item \ttt{encodegmw}: This function encodes client 1's and client 2's
+  secret bits called $\ttt{s[1,s1]}$ and $\ttt{s[1,s2]}$ into two
+  distinct wire values (pairs of shares).
+\item \ttt{andgmw}: This function defines the gate $\ttt{g}$, the
+  identifier $\ttt{g}$ being used to distinguish randomness used
+  within.  In our version client 1 builds the output table (using
+  \ttt{andtablegmw})and transfers the correct output share to client 2
+  using 1-out-of-4 OT as per standard GMW protocol.
+\item \ttt{decode}: This function decodes and publishes a wire value
+  by $\ttt{xor}$ing the shares. Note that this requires both client 1
+  and 2 to publicize their shares.
+\end{itemize}
 
-In our implementation, gates are wired together using gate
-identifiers, which are strings $w$. Top-level functionality in Figures
-\ref{fig-ygc-aux} and \ref{fig-ygc-gates} includes the following:
-\begin{itemize}
-\item \ttt{andgate}: This defines a subprotocol for the garbler
-  to define a garbled gate $\ttt{gid}$ with input wires from gates
-  $\ttt{ga}$ and $\ttt{gb}$. The garbler generates keys and garbles
-  the rows in YGC fashion, them with client $1$ in
-  views in a standard form. For example, the view for
-  a hypothetical gate 6, row 2 garbled truth table is $\ttt{v[1,gate:6tt2]}$.
-  We note that garbled gates of other binary operators can be obtained with
-  replacement of $\ttt{andtable}$ with appropriate garbled table definitions. 
-\item \ttt{evalgate}: This defines a subprotocol for the evaluator to
-  evaluate gate $\ttt{gid}$ given input wire values $\ttt{wva}$ and
-  $\ttt{wvb}$.
-\item \ttt{garbledecode} and \ttt{evaldecode}: The former function
-  defines the garbler's protocol for encrypting the circuit
-  output from final gate $\ttt{gid}$, and the latter defines
-  the evaluator's output decryption protocol.
-\item \ttt{encode}: This defines the initial phase of the protocol,
-  where the evaluator receives the wire value from their own
-  secret $\sx{1}{sa}$ via $\ttt{OT}$, and the garbler communicates
-  the wire value for their own secret $\sx{2}{sb}$ directly.
-\end{itemize}
-\begin{example}
-  \label{example-andcircuit}
-The following program uses our YGC library to define
-a circuit with a single and gate and input secrets $\ttt{s1}$ and
-$\ttt{s2}$ from client's 1 and 2 respectively. 
-\begin{verbatimtab}
-  andgate(0,s1,s2);
-  garbledecode(0);
-  let secrets = encode(s1,s2) in
-  v[0,output] := decode(evalgate(0, secrets.wv1, secrets.wv2))
-\end{verbatimtab}
-\end{example}
-We have verified passive security of the $\fedprot$ protocol
-generated by this and other small circuits using the
-technique described in Lemma \ref{lemma-bruteforce-nimo}.
-But large circuits with thousands of gates would be
-intractable to verify with this method. In the next Section
-we discuss compositional methods to address this issue.
-
-\subsection{Component Certification for $\NIMO$}
-\label{section-composition-certification}
-
-We can use the automated verification methods developed in Section
-\ref{section-brute-force} to \emph{certify} YGC
-components. Certification obtains properties that we can then use
-higher-level reasoning-- inspired in part by techniques in
-\cite{barthe2019probabilistic}-- to prove that circuits built with
-certified components satisfy $\NIMO$ (Theorem \ref{theorem-ygc-NIMO})
-and thus are passive secure.
-
-Our method relies on a definition of compositional units comprising
-complete related subprotocols-- i.e., both the garbler's
-construction of the gates and the evaluator's evaluation of them. We
-therefore make the following definitions. However, we note that our
-formal results are applicable to the YGC style where the garbler
-shares the entire circuit prior to evaluation.
-\begin{definition}
-  \label{ygc-modules}
-  We define a composable internal \ttt{and} gate sharing and evaluation function
-  called $\ttt{andgg}$, and an output decoder gate share and evaluation function called
-  $\ttt{decode}$, as follows. 
-\begin{verbatimtab}
-  andgg(g, ga, gb, wva, wvb) { andgate(g, ga, gb); evalgate(g, wva, wvb) }
-  decode(g, v) { garbledecode(g); evaldecode(v) }
-\end{verbatimtab}
-\end{definition}
+\subsubsection{Gate Certification}
 
@@
 restriction that any function in each of these categories has the same
-valid type signature as $\ttt{decode}$, $\ttt{encode}$, and
-$\ttt{andgg}$ respectively.
+valid type signature as $\ttt{decodegmw}$, $\ttt{encodegmw}$, and
+$\ttt{andgmw}$ respectively. When gates are used in ciruits they are
+parameterized by wire values-- in certification we will use carefully
+chosen witness parameters.
 
-Certifications will demonstrate probabilistic independence of input
-wire lables and garbled tables in a program's basic distribution, so
-we need to consider dependencies of flips and gate views. Thus we make
-the following definitions:
+When we consider internal components and hence intensional properties
+of circuits, we will consider behaviors with respect to particular bits
+of randomness. Hence we will consider basic program distributions. 
 \begin{definition}
@@
 from \cite{barthe2019probabilistic} with the same denotations.
+\cnote{I switched here to using $\pmf$ to denote an arbitrary probability mass
+function, will update earlier sections with this notation.}
 \begin{definition}
-  We write $\vc{\pdf{X}}{x}{y}$ iff $\pdf{X}(\{ x \mapsto 0\}\ |\ \{ y \mapsto 0 \}) =
-  \pdf{X}(\{ x \mapsto 1\}\ |\ \{ y \mapsto 1 \}) = 1$.
-  We write $\sep{\pdf{X}}{Y}{Z}$ iff for all
+  We write $\vc{\pmf}{x}{y}$ iff $\pmf(\{ x \mapsto 0\}\ |\ \{ y \mapsto 0 \}) =
+  \pmf(\{ x \mapsto 1\}\ |\ \{ y \mapsto 1 \}) = 1$.
+  We write $\sep{\pmf}{Y}{Z}$ iff for all
     $\store \in \mems(X \cup Y)$:
-  $$(\margd{\pdf{X}}{Y \cup Z})(\store) =
-  (\margd{\pdf{X}}{Y})(\store_Y) * (\margd{\pdf{X}}{Z})(\store_Z)$$ 
+  $$(\margd{\pmf}{Y \cup Z})(\store) =
+  \pmf(\store_Y) * \pmf(\store_Z)$$ 
 \end{definition}
-It is also necessary to define notions of correlation with wire
-labels, since wire values flow into and out of gates in either
-positive of negative correlation with them.
-\begin{definition}
-  We write $\gc{\prog}{\{ \ttt{k = }\be_{k}; \ttt{p = }\be_{p} \}}{g}$ iff one of the
-  following hold:
+
+\begin{fpfig}[t]{GMW library type annotations.}{fig-gmw-types}
+{\footnotesize
+  \begin{verbatimtab}
+   encodegmw   : string(gid) * string(gid)
+
+   andtablegmw : { k = bool[i]; p = bool[i] }
+
+   andgmw      : string(gid) *  { c1 = bool[1]; c2 = bool[2] } * { c1 = bool[1]; c2 = bool[2] }
+
+   decodegmw   : { c1 = bool[1]; c2 = bool[2] }  \end{verbatimtab}
+}
+\end{fpfig}
+
+Now we can define GMW gate certification as follows. Intuitively, the
+input wire shares of each client should be independent of the gate output
+shares. This ensures independence of clients' shares (views).
+\begin{definition}[GMW Gate Certification]
+  \label{definition-gmwgate-certification}
+  Let $\mathit{gate}$ be a gate function and let $e$ be the following
+  $\metaprot$ program:
+  $$
+  \begin{array}{l}
+    \ttt{let in1 = \{c1 = flip[1,s11];c2 = flip[2,s12]\} in}\\
+    \ttt{let in2 = \{c1 = flip[1,s21];c2 = flip[2,s22]\} in}\\
+    \mathit{gate}\ttt{(g,in1,in2)}
+  \end{array}
+  $$
+  and let $
+  \config{\varnothing}{e}\redxs\config{\prog}{\ttt{\{c1 = v[1,gout]; c2 = v[2,gout]\}}}
+  $.
+  Then $\mathit{gate}$ is \emph{certified} iff the following conditions hold:
   \begin{enumerate}[\hspace{5mm}i.]
-  \item $\vc{\progtt(\prog,\be_k)}{\itv}{\ttt{flip[2,gate:}g\ttt{.k]}}\quad \text{and} \quad
-    \vc{\progtt(\prog,\be_p)}{\itv}{\ttt{flip[2,gate:}g\ttt{.p]}}$
-  \item $\vc{\progtt(\prog,\enot\ \be_k)}{\itv}{\ttt{flip[2,gate:}g\ttt{.k]}}\quad \text{and}\quad
-    \vc{\progtt(\prog,\enot\ \be_p)}{\itv}{\ttt{flip[2,gate:}g\ttt{.p]}}$
+  \item $\sep{\progtt(\prog)}{\ttt{\{flip[1,s11],flip[1,s21]\}}}{\ttt{\{v[1,gout],v[2,gout]\}}}$
+  \item $\sep{\progtt(\prog)}{\ttt{\{flip[2,s12],flip[2,s22]\}}}{\ttt{\{v[1,gout],v[2,gout]\}}}$
+  % \item $\sep{\progtt(\prog)}{\ttt{\{v[1,gout]\}}}{\ttt{\{v[2,gout]\}}}$
+  \item All variables in:$$\vars(\prog) - \{ \ttt{flip[1,s11],flip[1,s21],flip[2,s12],flip[2,s22]}\}$$
+    are distinguished by $\ttt{g}$ (i.e., contain $\ttt{g}$ as an identifier substring). 
   \end{enumerate}
 \end{definition}
-Thus, if a value is in correlation with a wire label, it is either
-correlated with the wire label or with the inverse of both the key and
-pointer bits. For subsequent discussion we make the following addition
-to the codebase:
-$$
-\ttt{invert(\{ k = bk; p = bp \}) \{ \{ k = not bk; p = not bp \} \}}
-$$
-Gate certification establishes separation of input labels from the
-gate table and output value, and correlation of the output with
-the output wire label under any input wire value combination.
-\begin{definition}[Gate Certification]
-  \label{definition-gate-certification}
-Let $\mathit{gate}$ be a gate function. For $1 \le i \le 4$  let:
-  $$
-  \config{\varnothing}{\ttt{v[0,out] := } e_i}\redxs\config{\prog_i}{\ttt{()}}
-  $$
-  where
-  \begin{enumerate}[\hspace{5mm}$e_1 \defeq$]
-    \item $\mathit{gate}\ttt{(c,a,b,owl(a),owl(b));}$
-    \item $\mathit{gate}\ttt{(c,a,b,owl(a),invert(owl(b)));}$
-    \item $\mathit{gate}\ttt{(c,a,b,invert(owl(a)),owl(b));}$
-    \item $\mathit{gate}\ttt{(c,a,b,invert(owl(a)),invert(owl(b)))}$
-  \end{enumerate}
-  Define:
-  $${\small X \defeq \ttt{\{ flip[2,gate:a.k], flip[2,gate:a.p], flip[2,gate:b.k], flip[2,gate:b.p] \}}}$$
-  Then $\mathit{gate}$ is \emph{certified} iff for all $1 \le i \le 4$:
-  $$
-  \sep{\progtt(\prog_i)}{X}{\vdefs(\prog_i)} 
-  \quad \text{and} \quad \gc{\progd(\prog_i)}{\vx{0}{out}}{\ttt{owl(c)}}
-  $$
-\end{definition}
-Encode certification establishes correlation of the encoded input values with
-the secret wire labels, and separation of client 2's secret from client 1's views. 
-\begin{definition}[Encode Certification]
+
+Encoding certification establishes the desired property in the protocol-- that
+input secrets of each client are independent of the other client's views. 
+\begin{definition}[GMW Encode Certification]
+  \label{definition-gmwencode-certification}
   Let $\mathit{encode}$ be an encoding function, and let:
@@
   \config{\varnothing}{\ttt{encode(s1,s2)}} \redxs
-  \config{\prog}{\ttt{\{wv1=}v_1\ttt{;wv2=}v_2\ttt{\}}}
+  \config{\prog}{\ttt{\{shares1 = }v_1\ttt{; shares2 = }v_2\ttt{\}}}
+  $$
+  where:
+  $$
+  \begin{array}{rcl}
+    v_1 &=& \ttt{\{ c1 = v[1,s1out]; c2 = v[2,s1out] \}}\\
+    v_2 &=& \ttt{\{ c1 = v[1,s2out]; c2 = v[2,s2out] \}}
+  \end{array}
   $$
@@
   \begin{enumerate}[\hspace{5mm}i.]
-  \item $\gc{\progd(\prog,v_1)}{\itv}{\ttt{owl(s1)}} \quad \text{and} \quad
-    \gc{\progd(\prog,v_2)}{\itv}{\ttt{owl(s2)}}$
-  \item $\sep{\progd(\prog,v_2)}{\{\ttt{s[2,s2]\}}}{(\vdefs(\prog) \cup \itv)}$
+    %\item $\sep{\progtt(\prog)}{\ttt{\{ s[1,s1],  v[1,s1out], v[1,s2out] \}}}{\ttt{\{ s[2,s2], v[2,s1out], v[2,s2out] \}}}$
+  \item $\sep{\progtt(\prog)}{\ttt{\{s[1,s1]\}}}{\ttt{\{v[2,s1out],v[2,s2out]\}}}$
+  \item $\sep{\progtt(\prog)}{\ttt{\{s[2,s2]\}}}{\ttt{\{v[1,s1out],v[1,s2out]\}}}$
+  \item All $x \in \vars(\prog)$ are distinguished by $\ttt{in1}$ or $\ttt{in2}$. 
   \end{enumerate}
   \end{definition}
-Decode certification establishes separation of the input wire
-label from the decoding tables under any input value condition. 
-\begin{definition}[Decode Certification]
-  Let $\mathit{decode}$ be a decoding function, and let:
-  $$
-  \config{\varnothing}{\mathit{decode}(\ttt{c}, \ttt{owl(c)})}
-  \redxs \config{\prog_1}{\be_1}
-  \quad \text{and} \quad
-  \config{\varnothing}{\mathit{decode}(\ttt{c}, \ttt{invert(owl(c))})}
-  \redxs \config{\prog_0}{\be_0}
-  $$
-  Then $\mathit{decode}$ is certified iff:
-  $$
-  \sep{\progtt(\prog_i)}{\ttt{\{flip[2,gate:c.k],flip[2,gate:c.p]\}}}{\vdefs(\prog_i)}
-  $$
-  %Then $\mathit{decode}$ is certified iff both of the following conditions hold:
-  %\begin{enumerate}[\hspace{5mm}i.]
-  %\item $\progd(\prog_1,\be_1)(\{ \itv \mapsto 1 \}) = 1$ and
-  %  $\progd(\prog_0,\be_0)(\{ \itv \mapsto 0 \}) = 1$
-  %\item $\sep{\progtt(\prog_i)}{\ttt{\{flip[2,gate:c.k],flip[2,gate:c.p]\}}}{\vdefs(\prog_i)}$
-  %\end{enumerate}
-\end{definition}
-We have certified each of the YGC components detailed above. 
+
+We have certified both $\ttt{encodegmw}$ and $\ttt{andgmw}$. Note that
+an Or or Xor gate, for example, could be certified with this same
+method to extend our library.
+
+\subsubsection{Gate Invariants} We can show that composition maintains an
+important invariant in a circuit $\prog$- namely that client 1's
+secrets remain independent of client 2's views, and vice-versa, prior
+to decoding.  That is, where $\iov(\prog) = S \cup V$, we show that
+$\sep{\progd(\prog)}{S_{\{1\}}}{V_{\{2\}}}$ and
+$\sep{\progd(\prog)}{S_{\{2\}}}{V_{\{1\}}}$ are preserved prior to
+decoding.
+
+We observe the following properties of separation which are borrowed from prior work
+\cite{barthe2019probabilistic} and which will be used frequently in proofs. 
 \begin{lemma}
-  \label{lemma-certification}
-  Each of $\ttt{decode}$, $\ttt{encode}$, and $\ttt{andgg}$ are certified.
+  \label{lemma-separation}
+  The following properties hold:
+  \begin{enumerate}
+  \item $\sep{\pmf}{Y}{Z}$ iff $\sep{\pmf}{Z}{Y}$
+  \item $\vc{\pmf}{x}{y}$ if  $\vc{\pmf}{y}{x}$
+  \item $\vc{\pmf}{x}{y}$ and $\vc{\pmf}{y}{z}$ imply $\vc{\pmf}{x}{z}$
+  \item $\sep{\pmf}{X}{(Y \cup Z)}$ implies $(\sep{\pmf}{X}{Y}$ and $\sep{\pmf}{X}{Z})$
+  \item $(\sep{\pmf}{X}{Y}$ and $\sep{\pmf}{X}{Z}$ and $\sep{\pmf}{Y}{Z})$ implies $\sep{\pmf}{X}{(Y \cup Z)}$ 
+  %\item If $\prog_1;\prog_2$ is safe and $\vars(\prog_1) \cap \vars(\prog_2) = \varnothing$
+  %  then $\sep{\progtt(\prog_1;\prog_2)}{\vars(\prog_1)}{\vars(\prog_2)}$.
+  \end{enumerate}
+\end{lemma}
+The following property also follows by results in \cite{barthe2019probabilistic} and will
+be useful to make constructions that demonstrate variable separation (especially in Lemma
+\ref{lemma-gmw-frame}).
+\begin{lemma}
+  \label{lemma-sepjoin}
+  $\sep{\progtt(\prog)}{X}{Y}$ iff for all 
+  $\store^1, \store^2 \in \runs(\prog)$ there exists
+  $\store \in \runs(\prog)$ with
+  $\store^1_{X} \cap \store^2_{Y} \subseteq \store$.
 \end{lemma}
 
-\subsection{Correctness of Certified Components}
-\label{section-composition-metatheory}
+The following Lemma allows us to separate variables in programs that may not occur
+in the same subprogram, and hence consider gate certification properties in various
+contexts. It is reminiscent of the \emph{frame} rule of \cite{barthe2019probabilistic}.
+\begin{lemma}
+  \label{lemma-gmw-frame}
+  If $\sep{\progtt(\prog_1;\prog_2)}{Y}{V}$ for $V \in \vdefs(\prog_2)$ and $\sep{\progtt(\prog_1)}{X}{Y}$ for
+  $X \not\in \vars(\prog_2)$ then $\sep{\progtt(\prog_1;\prog_2)}{X}{V}$. 
+\end{lemma}
+\begin{proof}
+  By assumption and Lemma \ref{lemma-sepjoin}, for all
+  $\store^1,\store^2 \in \runs(\prog_1;\prog_2)$ we have $\store^1_Y
+  \cap \store^2_V \subseteq \store$ for some $\store_a \in
+  \runs(\prog_1;\prog_2)$ and $\store^1_X \cap \store^2_Y \subseteq
+  \store$ for some $\store_b \in \runs(\prog_1;\prog_2)$.  Thus, since
+  we assume $X \not\in \vars(\prog_2)$ we have $\store_a \cap \store_b
+  \in \runs(\prog_1;\prog_2)$ and $\store^1_X \cap \store^2_V =
+  \store_a \cap \store_b$, obtaining the result by
+  \ref{lemma-sepjoin}.
+\end{proof}
 
-To demonstrate correctness of certification, we need to show that
-isolated component certificates preserve relevant properties
-when integrated into larger programs. This resonates with
-\emph{framing} as developed in \cite{barthe2019probabilistic}.
-In the following we will use $f$ to range over flips.
-First, we show that if a flip $f$ is independent from the views
-in a protocol (as is the case with input labels used to
-encode a garbled table), then separation from views that
-use $f$ (e.g., in an output label) is preserved. 
-\begin{lemma}[Wire Framing]
-  \label{lemma-wire-framing}
-  If $\sep{\progtt(\prog)}{\{f\}}{\vdefs(\prog)}$ for flip $f$ and
-  $\vars(\prog) \cap \vars(\prog') = \{ f \}$ then
-  $\sep{\progtt(\prog';\prog)}{\vdefs(\prog')}{\vdefs(\prog)}$.
+The next two Lemmas allow us to use the properties of gates established by
+certification and replace the ``dummy'' parameters used there with actual
+wire values in full circuit contexts. Intuitively we show that substitution
+for dummy flips preserves separation.
+\begin{lemma}
+  \label{lemma-presub}
+  Given $\be_1$, $\be_2$, and $i$ where $\vars(\be_1) \cap
+  \vars(\be_2) = \varnothing$ and $\lcod{\store_1,\be_1}{i} = \beta$
+  with $\store_1(x) = \lcod{\store_2,\be_2}{i}$ and
+  $\dom(\store_1) = \vars(\be_1)$ and $\dom(\store_2) = \vars(\be_2)$.
+  Then $\lcod{\store_1\cap\store_2,\be_1[\be_2/x]}{i} = \beta$.
 \end{lemma}
-Next, we show that output correlations (e.g., between output wire values and
-labels) are preserved by substitution of correlated values for flips (e.g.,
-between input wire values and labels).
+\begin{proof}
+By straightforward structural induction on $\be$.
+\end{proof}
+
 \begin{lemma}[Substitution]
   \label{lemma-substitution}
-  If $\vc{\progtt(\prog,\be)}{\itv}{f}$ and $\vc{\progtt(\prog',\be')}{\itv}{f'}$
-  and $\sep{\progtt(\prog';\prog)}{\vdefs(\prog')}{\vdefs(\prog)}$
-  then $\vc{\progtt(\prog';\prog,\be[\be'/f'])}{\itv}{f}$.
+  If $\sep{\progtt(\prog_2)}{\{ f \}}{X}$ and
+  $\prog_1;\prog_2[\be/f]$ is safe with $\vars(\prog_1,\be) \cap
+  X = \varnothing$ then
+  $\sep{\progtt(\prog_1;\prog_2[\be/f])}{\vars(\be)}{X}$.
 \end{lemma}
-Our wire framing and substitution Lemmas then allow us to generalize
-certification of gates and encoding to arbitrary circuit contexts,
-assuming that outputs are not wired to multiple inputs \cite{tate2003garbled,nieminen2023breaking}.
-\begin{lemma}[Gate Composition]
-  \label{lemma-compose}
-Given well-typed $\prog_1;E[\mathit{gate}(g,g_1,g_2,v_1,v_2)]$ and certified $\mathit{gate}$ where:
+\begin{proof}
+  Suppose on the contrary it was not the case that
+  $\sep{\progtt(\prog_1;\prog_2[\be/f])}{\vars(\be)}{X}$.
+  Then by Lemma \ref{lemma-sepjoin} there exists $\store_1,\store_2
+  \in \runs(\prog_1;\prog_2[\be/f])$ with no $\store \in
+  \runs(\prog_1;\prog_2[\be/f])$ such that $\store^1_{\vars(\be)}
+  \cap \store^2_X \subseteq \store$.  But also by
+  assumption and Lemma \ref{lemma-sepjoin} for any $\beta$ there
+  exists $\store' \in \runs(\prog_2)$ with $\{ f \mapsto \beta \} \cap
+  \store^2_X \subseteq \store'$. So in particular, we
+  have $\{ f \mapsto \lcod{\store^1_{\vars(\be)},\be}{i} \} \cap
+  \store^2_X \subseteq \store'$ for any $i$. This, the
+  assumption $\vars(\prog_1,\be) \cap \vars(\prog_2) = \varnothing$,
+  and application of Lemma \ref{lemma-presub} leads to the consequence
+  that there exists $\store \in \runs(\prog_1;\prog_2[\be/f])$ with
+  $\store \supseteq \store^1_{\vars(\be)} \cap
+  \store^2_X$ given the assumption $\vars(\prog_1,\be)
+  \cap \vars(\prog_2) = \varnothing$, which is a contradiction.
+\end{proof}
+
+Now we can prove that certified gates preserve the desired invariant
+during execution in arbitrary circuit contexts.
+We formulate this in terms of pre-~and post-conditions of gate
+evaluation.
+\begin{lemma}[GMW Gate Invariant]
+  \label{lemma-gmw-preservation}
+  Given well-typed $\prog_1;E[\mathit{gate}(\ttt{g},v_1,v_2)]$ for certified $\mathit{gate}$ where:
   $$
-\config{\prog_1}{E[\mathit{gate}(g,g_1,g_2,v_1,v_2)]} \redxs \config{\prog_1;\prog_2}{E[v]}
+  \config{\prog_1}{E[\mathit{gate}(\ttt{g},v_1,v_2)]} \redxs \config{\prog_1;\prog_2}{E[v]}
   $$
-If the following preconditions hold:
-\begin{enumerate}
-\item $\{ g_1 \} \cap \{ g_2 \} \cap \wired(\prog_1) = \varnothing$ and $g \not\in \prog_1$
-\item $\gc{\progd(\prog_1)}{v_1}{\owl(g_1)}$ and $\gc{\progd(\prog_1)}{v_2}{\owl(g_2)}$
+  If we have the following preconditions, where $\iov(\prog_1) = S \cup V$:
+  \begin{enumerate}[\hspace{5mm}i.]
+  \item $\sep{\progd(\prog_1)}{S_{\{2\}}}{V_{\{1\}}}$
+  \item $\sep{\progd(\prog_1)}{S_{\{1\}}}{V_{\{2\}}}$
   \end{enumerate}
-then we have as postconditions:
-\begin{enumerate}
-  \item $\gc{\progd(\prog_1;\prog_2)}{v}{\owl(g_2)}$
-  \item $\sep{\progd(\prog_1;\prog_2,v)}{\vdefs(\prog_1)}{\vdefs(\prog_2) \cup \{\itv\}}$
+  then we have as a postconditions, where $\iov(\prog_1;\prog_2) = S \cup V'$:
+  \begin{enumerate}[\hspace{5mm}i.]
+  \item $\sep{\progd(\prog_1;\prog_2)}{S_{\{2\}}}{V'_{\{1\}}}$
+  \item $\sep{\progd(\prog_1;\prog_2)}{S_{\{1\}}}{V'_{\{2\}}}$
   \end{enumerate}
 \end{lemma}
+
 \begin{proof}
-  Given that $g_1$ and $g_2$ are distinct and not wired in $\prog_1$
-  we are assured that $\ttt{owl}(g_1)$ and $\ttt{owl}(g_1)$ are in
-  independent uniform distributions, and given that $g \not\in
-  \prog_1$ we are assured that $\ttt{owl}(g)$ contains entirely fresh
-  flips. Thus, precondition 2, Lemma \ref{lemma-wire-framing}, and
-  Definition \ref{definition-gate-certification} imply postcondition
-  2.  Furthermore, since gate certification guarantees that
-  correlation with $\ttt{owl}(g)$ is guaranteed under all possible gate
-  input wire values, precondition 2 and Lemma \ref{lemma-substitution}
-  guarantee postcondition 2.
+  Given well-typedness of $\prog_1;E[\mathit{gate}(\ttt{g},v_1,v_2)]$, we have for some $\ttt{g1}$ and $\ttt{g2}$:
+  $$
+  \begin{array}{lcl}
+   v_1 &=& \ttt{\{c1 = v[1,g1out]; c2 = v[2,g1out]\}}\\
+   v_2 &=& \ttt{\{c1 = v[1,g2out]; c2 = v[2,g2out]\}}\\
+    v &=& \ttt{\{c1 = v[1,gout]; c2 = v[2,gout]\}}
+  \end{array}
+  $$
+  Let $\prog$ be as defined in Definition \ref{definition-gmwgate-certification}.
+  We observe:
+  $$
+  {\footnotesize
+    \begin{array}{c}
+      \prog_2 = \\
+      \prog[\ttt{v[1,g1out]}/\ttt{flip[1,s11]}][\ttt{v[2,g1out]}/\ttt{flip[2,s12]}][\ttt{v[1,g2out]}/\ttt{flip[1,s21]}][\ttt{v[2,g2out]}/\ttt{flip[2,s22]}]
+    \end{array}
+  }
+  $$
+  Now, the assumption of well-typedness also assures that $\ttt{g}$
+  has not been previously used as a gate identifier in $\prog_1$,
+  since otherwise the views $\ttt{v[1,gout]}$ and $\ttt{v[2,gout]}$
+  would have been previously defined. This ensures $\vars(\prog_1,x)
+  \cap \vars(\prog) = \varnothing$ for:
+  $$x \in \{ \ttt{v[1,g1out]}, \ttt{v[2,g1out]}, \ttt{v[1,g2out]}, \ttt{v[2,g2out]}\}$$
+  given condition (iii) of Definition \ref{definition-gmwgate-certification}.
+  Thus by Lemma \ref{lemma-substitution} we have:
+  $$\sep{\progtt(\prog_1;\prog_2)}{\ttt{\{v[1,g1out],v[1,g2out]\}}}{\ttt{\{v[1,gout],v[2,gout]\}}}$$
+  and
+  $$\sep{\progtt(\prog_1;\prog_2)}{\ttt{\{v[2,g1out],v[2,g2out]\}}}{\ttt{\{v[1,gout],v[2,gout]\}}}$$
+  And by Lemma \ref{lemma-gmw-frame} and condition (iii) of Definition
+  \ref{definition-gmwgate-certification} we then have:
+  $$\sep{\progtt(\prog_1;\prog_2)}{S_{\{1\}}}{\ttt{\{v[2,gout]\}}} \quad \text{and} \quad
+  \sep{\progtt(\prog_1;\prog_2)}{S_{\{2\}}}{\ttt{\{v[1,gout]\}}}$$
+  so by precondition assumptions and Lemma \ref{lemma-separation} we have:
+  $$\sep{\progtt(\prog_1;\prog_2)}{S_{\{1\}}}{V'_{\{2\}}} \quad \text{and} \quad
+  \sep{\progtt(\prog_1;\prog_2)}{S_{\{2\}}}{V'_{\{1\}}}$$
 \end{proof}
 
-\begin{lemma}[Decode Composition]
-  \label{lemma-decode}
-Given well-typed $\prog_1;E[\mathit{decode}(g,v)]$ and certified $\mathit{decode}$ where 
+Here we show that encoding also preserves the desired invariant.
+The result here is more straightforward than it is for gates, because
+the variables used during encoding are guaranteed to be distinct from
+the rest of the program.
+\begin{lemma}[GMW Encode Invariant]
+  \label{lemma-gmw-encode}
+  Given well-typed $\prog_1;E[\mathit{encode}\ttt{(s1,s2)}]$ for certified $\mathit{encode}$, let:
   $$
-\config{\prog_1}{E[\ttt{decode}(g,v)]} \redxs \config{\prog_1;\prog_2}{E[\be]}
+  \config{\prog_1}{E[\mathit{encode}(\ttt{s1},\ttt{s2})]} \redxs
+  \config{\prog_1;\prog_2}{E[v]}
   $$
-If the following preconditions hold:
-\begin{enumerate}
-\item $g \not\in \cap \wired(\prog_1)$
-\item $\gc{\progd(\prog_1)}{v}{\owl(g)}$
+  If we have the following preconditions, where $\iov(\prog_1) = S' \cup V$:
+  \begin{enumerate}[\hspace{5mm}i.]
+  \item $\sep{\progd(\prog_1)}{S_{\{2\}}}{V_{\{1\}}}$
+  \item $\sep{\progd(\prog_1)}{S_{\{1\}}}{V_{\{2\}}}$
   \end{enumerate}
-then we have as postcondition:
-\begin{enumerate}
-  \item $\sep{\progd(\prog_1;\prog_2)}{\vdefs(\prog_1)}{\vdefs(\prog_2)}$
+  then we have as a postconditions, where $\iov(\prog_1;\prog_2) = S' \cup V'$: 
+  \begin{enumerate}[\hspace{5mm}i.]
+  \item $\sep{\progd(\prog_1;\prog_2)}{S'_{\{2\}}}{V'_{\{1\}}}$
+  \item $\sep{\progd(\prog_1;\prog_2)}{S'_{\{1\}}}{V'_{\{2\}}}$
   \end{enumerate}
 \end{lemma}
-Given the above, we can establish that any circuit built with certified
-components is passive secure. 
-\begin{theorem}
-  \label{theorem-ygc-NIMO}
-  If $e$ is a well-typed circuit definition using certified components
-  where no gate output is wired more than once, and $\config{\varnothing}{e}
-  \redxs \config{\prog}{\varnothing}$, then $\NIMO(\prog)$. 
-\end{theorem}
 
 \begin{proof}
-  Let $\iov(\prog) = S \cup (V \cup \outv)$. Noting that any
-  encoding necessarily uses entirely fresh variables by Definition and
-  type safety, by applying Lemmas \ref{lemma-compose} and
-  \ref{lemma-decode} and the frame rule of
-  \cite{barthe2019probabilistic} we obtain:
+  By assumptions of well-typedness we have:
   $$
-  \sep{\progd(\prog)}{(S_{\{2\}}\cup\{\outv\})}{V}
+  \begin{array}{rcl}
+    v &=&  \ttt{\{shares1 = }v_1\ttt{; shares2 = }v_2\ttt{\}}\\
+    v_1 &=& \ttt{\{ c1 = v[1,s1out]; c2 = v[2,s1out] \}}\\
+    v_2 &=& \ttt{\{ c1 = v[1,s2out]; c2 = v[2,s2out] \}}
+  \end{array}
   $$
-  Noting that the garbler (client 2) has no views, this means that
-  the garbler's secrets are independent of the evaluator's views. The
-  result follows by Definition \ref{definition-NIMO}.
+  And by Definition \ref{definition-gmwencode-certification} we have:
+  \begin{mathpar}
+    \sep{\progtt(\prog)}{\ttt{\{s[1,s1]\}}}{\ttt{\{v[2,s1out],v[2,s2out]\}}}
+
+    \sep{\progtt(\prog)}{\ttt{\{s[2,s2]\}}}{\ttt{\{v[1,s1out],v[1,s2out]\}}}
+  \end{mathpar}
+  Since we assume that secrets are in uniform and independent marginal
+  distributions a priori, and $\vars(\prog_1) \cap \vars(\prog_2) =
+  \varnothing$ by condition (iii) of Definition 
+  \ref{definition-gmwencode-certification} and assumptions of well-typedness, 
+  the result follows by preconditions and Lemma \ref{lemma-separation}.
 \end{proof}
 
-\begin{fpfig}[t]{YGC copy gate definitions.}{fig-ygc-copy}
-  {\footnotesize
-    \begin{verbatimtab}
-      sharetab2(gid, tid, k, p, b)
-      {
-        let r1 = k xor b in
-        let r0 = (not k) xor (not b) in
-        v[1,gate: || gid || tid || 1] := select(p,r1,r0);
-        v[1,gate: || gid || tid || 2] := select(not p,r1,r0);
-      }
+On the basis of the preceding we can prove our main result as
+follows. Note that we assume a normal form of programs where the last
+instruction is a decoding and public reveal of the output.
+\begin{theorem}
+  \label{theorem-gmw-NIMO}
+  If $\eassign{\outv}{\ttt{decode}(e)}$ is a well-typed GMW circuit
+  definition using certified components and
+  $\config{\varnothing}{\eassign{\outv}{\ttt{decode}(e)}} \redxs
+  \config{\prog}{\varnothing}$, then $\NIMO(\prog)$.
+\end{theorem}
+\begin{proof}
+  Let $\iov(\prog) = S \cup (V \cup \outv)$. By definition of $\ttt{decode}$
+  and assumptions of well-typedness we have for some $\ttt{g}$:
+  $$
+  \config{\varnothing}{\eassign{\outv}{\ttt{decode}(e)}} \redxs
+  \config{\prog;\eassign{\outv}{\ttt{v[1,gout] xor v[2,gout]}}}{\varnothing}
+  $$
+  where by Lemmas \ref{lemma-gmw-encode} and \ref{lemma-gmw-preservation} we
+  have $\sep{\progd(\prog)}{S_{\{1\}}}{V_{\{2\}}}$ and
+  $\sep{\progd(\prog)}{S_{\{2\}}}{V_{\{1\}}}$. Since by properties of \ttt{xor} we have:
+  $$
+  \sep{\progd(\prog;\eassign{\outv}{\ttt{v[1,gout] xor v[2,gout]}})}{\{\outv\}}{\{\ttt{v[1,gout]}\}}
+  $$
+  and:
+  $$
+  \sep{\progd(\prog;\eassign{\outv}{\ttt{v[1,gout] xor v[2,gout]}})}{\{\outv\}}{\{\ttt{v2,gout]\}}
+  $$
+  therefore by Lemma \ref{lemma-gmw-frame}}:
 
-      copygate(ca,cb,g)
-      {
-        let wl = owl(g) in
-        let owl1 = owl(ca) in
-        let owl2 = owl(cb) in
-        sharetab2(ca,tt,wl.k,wl.p,owl1.k); sharetab2(ca,pt,wl.k,wl.p,owl1.p);
-        sharetab2(cb,tt,wl.k,wl.p,owl2.k); sharetab2(cb,pt,wl.k,wl.p,owl2.p)
-      }
-
-      evalcopy(ca,cb,wv)
-      {
-        let wv1k = wv.k xor select(wv.p,v[1,gate: || ca || tt1], v[1,gate: || ca || tt2]) in
-        let wv1p = wv.k xor select(wv.p,v[1,gate: || ca || pt1], v[1,gate: || ca || pt1]) in
-        let wv2k = wv.k xor select(wv.p,v[1,gate: || cb || tt1], v[1,gate: || cb || tt2]) in
-        let wv2p = wv.k xor select(wv.p,v[1,gate: || cb || pt1], v[1,gate: || cb || pt1]) in
-        { wv1 = { k = wv1k; p = wv1p }; wv2 = { k = wv2k; p = wv2p } }  
-      }
-
-      copy(ca,cb,g,wv) { copygate(ca,cb,g); evalcopy(ca,cb,wv) } 
-    \end{verbatimtab}
-  }
-\end{fpfig}
-
-\subsection{Building and Extending Automatically Secure Circuits}
-\label{section-composition-copy}
-
-Our results in Section \ref{section-pre-post} and Theorem
-\ref{theorem-nimo} imply that any well-typed program using
-$\ttt{encode}$, $\ttt{andgg}$, and $\ttt{decode}$ will generate a
-circuit that is passive secure, provided that gate outputs are
-not used more than once. To add an $\ttt{or}$ gate to our
-library, we can use $\ttt{andgg}$ as a template, in fact simple
-modification of $\ttt{andtable}$ is all that would be needed. Thus,
-this library embodies a Fairplay-like language where well-typedness
-of metaprograms guarantees safety and security of generated protocols.
-
-The requirements that output wire labels are not used more than once
-reflects a known result, that naive copy gate output is unsound in
-YGC- i.e., passive security fails. In fact, we are unable to
-automatically verify $\gNIMO$ for $\ttt{andgg}$ without the assumption
-that input wire lables are in independent uniform distributions. We
-can capture the necessary fix as in Figure \ref{fig-ygc-copy}. The
-garbler creates a copy gate with two new output wire labels in
-independent uniform distributions (in \ttt{copygate}), and permutes
-them using the input wire label pointer after encrypting with the
-input wire label keys. The evaluator then recovers the encrypted
-output wire values during evaluation (in \ttt{evalcopy}). Given an
-$\ttt{orgg}$ gate defined as outlined abotve, we can then use
-secure $\ttt{copy}$ as follows: 
-\begin{verbatimtab}
-  let s = encode(s1,s2) in
-  let s1c = copy(c1s1,c2s1,s1,s.wv1) in
-  let s2c = copy(c1s2,c2s2,s2,s.wv2) in
-  let wv1 = andgg(1,c1s1,c1s2,s1c.wv1,s2c.wv1) in
-  let wv2 = orgg(2,c2s1,c2s2,s1c.wv2,s2c.wv2) in
-  decode(3,andgg(3,1,2,wv1,wv2))
-\end{verbatimtab}
-These definitions of $\ttt{copy}$ and $\ttt{orgg}$ have been
-automatically verified as in Section \ref{section-pre-post} for
-$\ttt{andgg}$. In general, any new gate form- such as optimized gates
-\cite{XXX}- can be verified automatically in a similar way.
-
-While the copy gate supports secure programming patterns, security
-still relies on programmer discipline, i.e., not wiring a gate more
-than once. However, we can keep track of wired gates in type effects,
-similar to the manner in which we keep track of defined views- type
-dependence in our system is able to accurately precisely track
-gate identifiers. This is also reminiscent of the use of linearity
-in other related systems for enforcing obliviousness \cite{darais2019language}.
+  $$
+  \sep{\progd(\prog;\eassign{\outv}{\ttt{v[1,gout] xor v[2,gout]}})}{(S_{\{1\}} \cup \{ \outv \})}{V_{\{2\}}}
+  $$
+  and:
+  $$
+  \sep{\progd(\prog;\eassign{\outv}{\ttt{v[1,gout] xor v[2,gout]}})}{(S_{\{2\}} \cup \{ \outv \})}{V_{\{1\}}}
+  $$
+  The result follows.
+\end{proof}

### 2023-07-06 popl24/hyperprop.tex
checkpoint

@@
 simulator security model- and hyperproperty-style formulations of
-program trace behavior \cite{XXX} including probabilistic
+program trace behavior \cite{8429300,wysstar} including probabilistic
 noninterference. This formulation of hyperproperties that are at least
@@
 between foundational approaches to security, and also motivates and
-supports automated analysis \cite{XXX}.
+supports automated analysis.
 
@@
 Here we briefly summarize the formal definition of MPC security. For a
-more detailed discussion the reader is referred to \cite{XXX}.
+more detailed discussion the reader is referred to \cite{evans2018pragmatic}.
 In this paper, we will consider the \emph{passive} aka
@@
 relations between probabilistic noninterference and security
-properties such as memory trace obliviousness \cite{XXX} and passive
-security \cite{XXX}. And above (Section \ref{section-pmf-examples}) we
+properties such as memory trace obliviousness \cite{darais2019language} and passive
+security \cite{barthe2019probabilistic}. And above (Section \ref{section-pmf-examples}) we
 have demonstrated that our basic encryption scheme enjoys a noninterference
@@
 adversarial knowledge is not increased by the execution of a program as in formulations
-of so-called gradual release \cite{XXX}. 
+of so-called gradual release \cite{4223226}. 
 \begin{lemma}

### 2023-07-06 popl24/scratch.tex
checkpoint

--- 2023-07-04 popl24/scratch.tex
@@
-\subsection{Compositional Properties of GMW}
+\section{Component Certification and Composition}
+\label{section-composition}
 
-\input{gmw}
-\cnote{Need to define type signatures.}
+Using YGC as a case study, in this Section we will discuss
+compositional properties that can be automatically verified on small
+program components using techniques developed in Section
+\ref{section-bruteforce}. These properties compose to yield $\NIMO$
+and scale to arbitrary programs. In our metatheory we will express
+these properties as pre-~and post-conditions on components, so our
+general approach is in a proof assistant style.
+
+\subsection{A Library for Yao's Garbled Circuits}
+\label{section-composition-ygc}
+
+\input{ygc}
+
+In Figures \ref{fig-ygc-aux} and \ref{fig-ygc-gates} we define a
+codebase for garbled circuits. This definition follows the
+\emph{point-and-permute} method described in \cite{evans2018pragmatic} and elsewhere,
+to which the reader is referred for more in-depth discussion.
+In this implementation client 2 is the \emph{garbler} and
+client 1 is the \emph{evaluator}. The garbler builds the garbled
+tables and shares them with the evaluator, who then evaluates
+the gate in an oblivious fashion until the final public output is
+generated through decryption. This definition is well-typed,
+with input type annotations for top-level functions listed in
+Figure \ref{fig-ygc-types}. Well-typed programs using these
+libraries are therefore guaranteed to yield safe $\minifed$
+programs.
+
+\emph{Wire labels} are fundamental to YGC, and essentially represent
+gate output values in an encrypted form. In our definition, wire
+labels are represented by records $\ttt{\{ k = }\beta_1\ttt{; p =
+}\beta_2\ttt{ \}}$, where $\ttt{k}$ is the \emph{key bit} and
+$\ttt{p}$ is the \emph{pointer bit}, and $\beta_1$ and $\beta_2$ are
+flips. Flips in each output wire label are owned by the garbler and
+are unique per gate by definition of their identifying string, and the
+representation of $0$ is the negation of $1$. For example, here is the
+representation of 1 and 0 respectively in the output wire label for a
+hypothetical gate 6:
+\begin{mathpar}
+  \ttt{\{ k = flip[2,gate:6.k]; p =  flip[2,gate:6.p]] \}}
+
+  \ttt{\{ k = not flip[2,gate:6.k]; p =  not flip[2,gate:6.p]] \}}
+\end{mathpar}
+The pointer bits in wire labels are used to select permuted rows in
+table garblings. The key bits are used to identify a unique key for
+table row in each garbled gate. Intuitively, if $\beta_1$ and
+$\beta_2$ are either key or pointer bits encoding 1 on two input wire
+labels to a binary gate, rows and keys in the gate are enumerated in
+the order:
+$$
+\neg\beta_1\neg\beta_2,\ \neg\beta_1\beta_2,\ \beta_1\neg\beta_2,\ \beta_1\beta_2
+$$
+
+In our implementation, gates are wired together using gate
+identifiers, which are strings $w$. Top-level functionality in Figures
+\ref{fig-ygc-aux} and \ref{fig-ygc-gates} includes the following:
+\begin{itemize}
+\item \ttt{andgate}: This defines a subprotocol for the garbler
+  to define a garbled gate $\ttt{gid}$ with input wires from gates
+  $\ttt{ga}$ and $\ttt{gb}$. The garbler generates keys and garbles
+  the rows in YGC fashion, them with client $1$ in
+  views in a standard form. For example, the view for
+  a hypothetical gate 6, row 2 garbled truth table is $\ttt{v[1,gate:6tt2]}$.
+  We note that garbled gates of other binary operators can be obtained with
+  replacement of $\ttt{andtable}$ with appropriate garbled table definitions. 
+\item \ttt{evalgate}: This defines a subprotocol for the evaluator to
+  evaluate gate $\ttt{gid}$ given input wire values $\ttt{wva}$ and
+  $\ttt{wvb}$.
+\item \ttt{garbledecode} and \ttt{evaldecode}: The former function
+  defines the garbler's protocol for encrypting the circuit
+  output from final gate $\ttt{gid}$, and the latter defines
+  the evaluator's output decryption protocol.
+\item \ttt{encode}: This defines the initial phase of the protocol,
+  where the evaluator receives the wire value from their own
+  secret $\sx{1}{sa}$ via $\ttt{OT}$, and the garbler communicates
+  the wire value for their own secret $\sx{2}{sb}$ directly.
+\end{itemize}
+\begin{example}
+  \label{example-andcircuit}
+The following program uses our YGC library to define
+a circuit with a single and gate and input secrets $\ttt{s1}$ and
+$\ttt{s2}$ from client's 1 and 2 respectively. 
+\begin{verbatimtab}
+  andgate(0,s1,s2);
+  garbledecode(0);
+  let secrets = encode(s1,s2) in
+  v[0,output] := decode(evalgate(0, secrets.wv1, secrets.wv2))
+\end{verbatimtab}
+\end{example}
+We have verified passive security of the $\fedprot$ protocol
+generated by this and other small circuits using the
+technique described in Lemma \ref{lemma-bruteforce-nimo}.
+But large circuits with thousands of gates would be
+intractable to verify with this method. In the next Section
+we discuss compositional methods to address this issue.
+
+\subsection{Component Certification for $\NIMO$}
+\label{section-composition-certification}
+
+We can use the automated verification methods developed in Section
+\ref{section-brute-force} to \emph{certify} YGC
+components. Certification obtains properties that we can then use
+higher-level reasoning-- inspired in part by techniques in
+\cite{barthe2019probabilistic}-- to prove that circuits built with
+certified components satisfy $\NIMO$ (Theorem \ref{theorem-ygc-NIMO})
+and thus are passive secure.
+
+Our method relies on a definition of compositional units comprising
+complete related subprotocols-- i.e., both the garbler's
+construction of the gates and the evaluator's evaluation of them. We
+therefore make the following definitions. However, we note that our
+formal results are applicable to the YGC style where the garbler
+shares the entire circuit prior to evaluation.
+\begin{definition}
+  \label{ygc-modules}
+  We define a composable internal \ttt{and} gate sharing and evaluation function
+  called $\ttt{andgg}$, and an output decoder gate share and evaluation function called
+  $\ttt{decode}$, as follows. 
+\begin{verbatimtab}
+  andgg(g, ga, gb, wva, wvb) { andgate(g, ga, gb); evalgate(g, wva, wvb) }
+  decode(g, v) { garbledecode(g); evaldecode(v) }
+\end{verbatimtab}
+\end{definition}
 
@@
 restriction that any function in each of these categories has the same
-valid type signature as $\ttt{decodegmw}$, $\ttt{encodegmw}$, and
-$\ttt{andgmw}$ respectively.
+valid type signature as $\ttt{decode}$, $\ttt{encode}$, and
+$\ttt{andgg}$ respectively.
+
+Certifications will demonstrate probabilistic independence of input
+wire lables and garbled tables in a program's basic distribution, so
+we need to consider dependencies of flips and gate views. Thus we make
+the following definitions:
 \begin{definition}
@@
 \begin{definition}
-  We write $\vc{\pmf}{x}{y}$ iff $\pmf(\{ x \mapsto 0\}\ |\ \{ y \mapsto 0 \}) =
-  \pmf(\{ x \mapsto 1\}\ |\ \{ y \mapsto 1 \}) = 1$.
-  We write $\sep{\pmf}{Y}{Z}$ iff for all
+  We write $\vc{\pdf{X}}{x}{y}$ iff $\pdf{X}(\{ x \mapsto 0\}\ |\ \{ y \mapsto 0 \}) =
+  \pdf{X}(\{ x \mapsto 1\}\ |\ \{ y \mapsto 1 \}) = 1$.
+  We write $\sep{\pdf{X}}{Y}{Z}$ iff for all
     $\store \in \mems(X \cup Y)$:
-  $$(\margd{\pmf}{Y \cup Z})(\store) =
-  \pmf(\store_Y) * \pmf(\store_Z)$$ 
+  $$(\margd{\pdf{X}}{Y \cup Z})(\store) =
+  (\margd{\pdf{X}}{Y})(\store_Y) * (\margd{\pdf{X}}{Z})(\store_Z)$$ 
 \end{definition}
-
-\begin{lemma}
-  \label{lemma-separation}
-  The following properties hold:
-  \begin{enumerate}
-  \item $\sep{\pmf}{Y}{Z}$ iff $\sep{\pmf}{Z}{Y}$
-  \item $\vc{\pmf}{x}{y}$ if  $\vc{\pmf}{y}{x}$
-  \item $\vc{\pmf}{x}{y}$ and $\vc{\pmf}{y}{z}$ imply $\vc{\pmf}{x}{z}$
-  \item ($\sep{\pmf}{X}{Y}$ and $\sep{\pmf}{X}{Z}$) iff $\sep{\pmf}{X}{(Y \cup Z)}$
-  %\item If $\prog_1;\prog_2$ is safe and $\vars(\prog_1) \cap \vars(\prog_2) = \varnothing$
-  %  then $\sep{\progtt(\prog_1;\prog_2)}{\vars(\prog_1)}{\vars(\prog_2)}$.
-  \end{enumerate}
-\end{lemma}
-
-\begin{lemma}
-  \label{lemma-sepjoin}
-  $\sep{\progtt(\prog)}{X}{Y}$ iff for all 
-  $\store^1, \store^2 \in \runs(\prog)$ there exists
-  $\store \in \runs(\prog)$ with
-  $\store^1_{X} \cap \store^2_{Y} \subseteq \store$.
-\end{lemma}
-
-%\begin{lemma}[Wire Framing]
-%  \label{lemma-wire-framing}
-%  If $\sep{\progtt(\prog_2)}{X}{Z}$ and
-%  $\vars(\prog_1) \cap \vars(\prog_2) \subseteq X$ then
-%  $\sep{\progtt(\prog_1;\prog_2)}{Y}{Z}$ for
-%  $Y \subseteq \vars(\prog_1)$.
-%\end{lemma}
-
-\begin{definition}[GMW Gate Certification]
-  \label{definition-gmwgate-certification}
-  Let $\mathit{gate}$ be a gate function and let $e$ be the following
-  $\metaprot$ program:
-  $$
-  \begin{array}{l}
-    \ttt{let in1 = \{c1 = flip[1,s11];c2 = flip[2,s12]\} in}\\
-    \ttt{let in2 = \{c1 = flip[1,s21];c2 = flip[2,s22]\} in}\\
-    \mathit{gate}\ttt{(g,in1,in2)}
-  \end{array}
-  $$
-  and let $
-  \config{\varnothing}{e}\redxs\config{\prog}{\ttt{\{c1 = v[1,gout]; c2 = v[2,gout]\}}}
-  $.
-  Then $\mathit{gate}$ is \emph{certified} iff the following conditions hold:
+It is also necessary to define notions of correlation with wire
+labels, since wire values flow into and out of gates in either
+positive of negative correlation with them.
+\begin{definition}
+  We write $\gc{\prog}{\{ \ttt{k = }\be_{k}; \ttt{p = }\be_{p} \}}{g}$ iff one of the
+  following hold:
   \begin{enumerate}[\hspace{5mm}i.]
-  \item $\sep{\progtt(\prog)}{\ttt{\{flip[1,s11],flip[1,s21]\}}}{\ttt{\{v[2,gout]\}}}$
-  \item $\sep{\progtt(\prog)}{\ttt{\{flip[2,s12],flip[2,s22]\}}}{\ttt{\{v[1,gout]\}}}$
-  \item All variables in:$$\vars(\prog) - \{ \ttt{flip[1,s11],flip[1,s21],flip[2,s12],flip[2,s22]}\}$$
-    are distinguished by $\ttt{g}$ (i.e., contain $\ttt{g}$ as an identifier substring). 
+  \item $\vc{\progtt(\prog,\be_k)}{\itv}{\ttt{flip[2,gate:}g\ttt{.k]}}\quad \text{and} \quad
+    \vc{\progtt(\prog,\be_p)}{\itv}{\ttt{flip[2,gate:}g\ttt{.p]}}$
+  \item $\vc{\progtt(\prog,\enot\ \be_k)}{\itv}{\ttt{flip[2,gate:}g\ttt{.k]}}\quad \text{and}\quad
+    \vc{\progtt(\prog,\enot\ \be_p)}{\itv}{\ttt{flip[2,gate:}g\ttt{.p]}}$
   \end{enumerate}
 \end{definition}
-
-\begin{lemma}
-  \label{lemma-gmw-frame}
-  If $\sep{\progtt(\prog_2)}{Y}{Z}$ and $\sep{\progtt(\prog_1)}{X}{Y}$ for
-  $X \not\in \vars(\prog_2)$ then $\sep{\progtt(\prog_1;\prog_2)}{X}{Z}$. 
-\end{lemma}
-
-\begin{lemma}
-  \label{lemma-presub}
-  Given $\be_1$, $\be_2$, and $i$ where $\vars(\be_1) \cap
-  \vars(\be_2) = \varnothing$ and $\lcod{\store_1,\be_1}{i} = \beta$
-  with $\store_1(x) = \lcod{\store_2,\be_2}{i}$ and
-  $\dom(\store_1) = \vars(\be_1)$ and $\dom(\store_2) = \vars(\be_2)$.
-  Then $\lcod{\store_1\cap\store_2,\be_1[\be_2/x]}{i} = \beta$.
-\end{lemma}
-\begin{proof}
-By straightforward structural induction on $\be$.
-\end{proof}
-
-\begin{lemma}[Substitution]
-  \label{lemma-substitution}
-  If $\sep{\progtt(\prog_2)}{\{ f \}}{X}$ and
-  $\prog_1;\prog_2[\be/f]$ is safe with $\vars(\prog_1,\be) \cap
-  X = \varnothing$ then
-  $\sep{\progtt(\prog_1;\prog_2[\be/f])}{\vars(\be)}{X}$.
-\end{lemma}
-\begin{proof}
-  Suppose on the contrary it was not the case that
-  $\sep{\progtt(\prog_1;\prog_2[\be/f])}{\vars(\be)}{X}$.
-  Then by Lemma \ref{lemma-sepjoin} there exists $\store_1,\store_2
-  \in \runs(\prog_1;\prog_2[\be/f])$ with no $\store \in
-  \runs(\prog_1;\prog_2[\be/f])$ such that $\store^1_{\vars(\be)}
-  \cap \store^2_X \subseteq \store$.  But also by
-  assumption and Lemma \ref{lemma-sepjoin} for any $\beta$ there
-  exists $\store' \in \runs(\prog_2)$ with $\{ f \mapsto \beta \} \cap
-  \store^2_X \subseteq \store'$. So in particular, we
-  have $\{ f \mapsto \lcod{\store^1_{\vars(\be)},\be}{i} \} \cap
-  \store^2_X \subseteq \store'$ for any $i$. This, the
-  assumption $\vars(\prog_1,\be) \cap \vars(\prog_2) = \varnothing$,
-  and application of Lemma \ref{lemma-presub} leads to the consequence
-  that there exists $\store \in \runs(\prog_1;\prog_2[\be/f])$ with
-  $\store \supseteq \store^1_{\vars(\be)} \cap
-  \store^2_X$ given the assumption $\vars(\prog_1,\be)
-  \cap \vars(\prog_2) = \varnothing$, which is a contradiction.
-\end{proof}
-
-\begin{lemma}[GMW Gate Invariant]
-  \label{lemma-preservation-gmw}
-  Given well-typed $\prog_1;E[\mathit{gate}(\ttt{g},v_1,v_2)]$ and certified $\mathit{gate}$ where
-  for some $\ttt{g1}$ and $\ttt{g2}$, :
+Thus, if a value is in correlation with a wire label, it is either
+correlated with the wire label or with the inverse of both the key and
+pointer bits. For subsequent discussion we make the following addition
+to the codebase:
 $$
-  \begin{array}{lcr}
-   v_1 &=& \ttt{\{c1 = v[1,g1out]; c2 = v[2,g1out]\}}\\
-   v_2 &=& \ttt{\{c1 = v[1,g2out]; c2 = v[2,g2out]\}}
-  \end{array}
+\ttt{invert(\{ k = bk; p = bp \}) \{ \{ k = not bk; p = not bp \} \}}
 $$
-  and let:
+Gate certification establishes separation of input labels from the
+gate table and output value, and correlation of the output with
+the output wire label under any input wire value combination.
+\begin{definition}[Gate Certification]
+  \label{definition-gate-certification}
+Let $\mathit{gate}$ be a gate function. For $1 \le i \le 4$  let:
   $$
-  \config{\prog_1}{E[\mathit{gate}(\ttt{g},v_1,v_2)]} \redxs \config{\prog_1;\prog_2}{E[\ttt{\{c1 = v[1,gout]; c2 = v[2,gout]\}}]}
+  \config{\varnothing}{\ttt{v[0,out] := } e_i}\redxs\config{\prog_i}{\ttt{()}}
   $$
-  If we have the following precondition, where $\iov(\prog_1) = S \cup V$:
+  where
+  \begin{enumerate}[\hspace{5mm}$e_1 \defeq$]
+    \item $\mathit{gate}\ttt{(c,a,b,owl(a),owl(b));}$
+    \item $\mathit{gate}\ttt{(c,a,b,owl(a),invert(owl(b)));}$
+    \item $\mathit{gate}\ttt{(c,a,b,invert(owl(a)),owl(b));}$
+    \item $\mathit{gate}\ttt{(c,a,b,invert(owl(a)),invert(owl(b)))}$
+  \end{enumerate}
+  Define:
+  $${\small X \defeq \ttt{\{ flip[2,gate:a.k], flip[2,gate:a.p], flip[2,gate:b.k], flip[2,gate:b.p] \}}}$$
+  Then $\mathit{gate}$ is \emph{certified} iff for all $1 \le i \le 4$:
   $$
-  \sep{\progd(\prog_1)}{(S_{\{2\}} \cup V_{\{2\}})}{(S_{\{1\}} \cup V_{\{1\}})}
+  \sep{\progtt(\prog_i)}{X}{\vdefs(\prog_i)} 
+  \quad \text{and} \quad \gc{\progd(\prog_i)}{\vx{0}{out}}{\ttt{owl(c)}}
   $$
-  then we have as a postcondition:
-  $$
-  \sep{\progd(\prog_1)}{(S_{\{2\}} \cup V_{\{2\}} \cup \{ \ttt{v[2,gout]} \})}{(S_{\{1\}} \cup V_{\{1\}} \cup \{ \ttt{v[1,gout]} \})}
-  $$
-\end{lemma}
-
-\begin{proof}
-  Let $\prog$ be as defined in Definition \ref{definition-gmwgate-certification}.
-  We observe:
-  $$
-  {\footnotesize
-    \begin{array}{c}
-      \prog_2 = \\
-      \prog[\ttt{v[1,g1out]}/\ttt{flip[1,s11]}][\ttt{v[2,g1out]}/\ttt{flip[2,s12]}][\ttt{v[1,g2out]}/\ttt{flip[1,s21]}][\ttt{v[2,g2out]}/\ttt{flip[2,s22]}]
-    \end{array}
-  }
-  $$ Now, the assumption of well-typedness of
-  $\prog_1;E[\mathit{gate}(\ttt{g},v_1,v_2)]$ assures that $\ttt{g}$
-  has not been previously used as a gate identifier in $\prog_1$,
-  since otherwise the views $\ttt{v[1,gout]}$ and $\ttt{v[2,gout]}$
-  would have been previously defined. This ensures $\vars(\prog_1,x)
-  \cap \vars(\prog_2) = \varnothing$ for:
-  $$x \in \{ \ttt{v[1,g1out]}, \ttt{v[2,g1out]}, \ttt{v[1,g2out]}, \ttt{v[2,g2out]}\}$$
-  given condition (iii) in Definition \ref{definition-gmwgate-certification}.
-  Thus by Lemma \ref{lemma-substitution} we have:
-  $$\sep{\progtt(\prog_1;\prog_2)}{\ttt{\{v[1,g1out],v[1,g2out]\}}}{\ttt{\{v[2,gout]\}}}$$
-  and
-  $$\sep{\progtt(\prog_1;\prog_2)}{\ttt{\{v[2,g1out],v[2,g2out]\}}}{\ttt{\{v[1,gout]\}}}$$
-  Since $\vars(\prog_1) \cap \vars(\prog_2)$ are only the wire inputs to $\ttt{g}$,
-  therefore by assumption and Lemma \ref{lemma-gmw-frame} the following assertions are
-  valid:
-  \begin{mathpar}
-    \sep{\progtt(\prog_1;\prog_2)}{(S_{\{1\}} \cup V_{\{1\}})}{\ttt{\{v[2,gout]\}}}
-
-    \sep{\progtt(\prog_1;\prog_2)}{\ttt{\{v[1,g1out],v[1,g2out]\}}}{(S_{\{2\}}\cup V_{\{2\}})}
-
-    \sep{\progtt(\prog_1;\prog_2)}{(S_{\{2\}} \cup V_{\{2\}})}{\ttt{\{v[1,gout]\}}}
-
-    \sep{\progtt(\prog_1;\prog_2)}{\ttt{\{v[2,g1out],v[2,g2out]\}}}{(S_{\{1\}}\cup V_{\{1\}})}
-  \end{mathpar}
-  The result follows by Lemma \ref{lemma-separation}.
-\end{proof}
-
-\begin{definition}[GMW Encode Certification]
+\end{definition}
+Encode certification establishes correlation of the encoded input values with
+the secret wire labels, and separation of client 2's secret from client 1's views. 
+\begin{definition}[Encode Certification]
   Let $\mathit{encode}$ be an encoding function, and let:
@@
   \config{\varnothing}{\ttt{encode(s1,s2)}} \redxs
-  \config{\prog}{\ttt{\{shares1 = }v_1\ttt{; shares2 = }v_2\ttt{\}}}
-  $$
-  where:
-  $$
-  \begin{array}{rcl}
-    v_1 &=& \ttt{\{ c1 = v[1,s1out]; c2 = v[2,s1out] \}}\\
-    v_2 &=& \ttt{\{ c1 = v[1,s2out]; c2 = v[2,s2out] \}}
-  \end{array}
+  \config{\prog}{\ttt{\{wv1=}v_1\ttt{;wv2=}v_2\ttt{\}}}
   $$
@@
   \begin{enumerate}[\hspace{5mm}i.]
-  \item $\sep{\progtt(\prog)}{\ttt{\{ s[1,s1] \}}}{\ttt{\{ v[2,s1out], v[2,s2out] \}}}$
-  \item $\sep{\progtt(\prog)}{\ttt{\{ s[2,s2] \}}}{\ttt{\{ v[1,s1out], v[1,s2out] \}}}$
-  \item All $x \in \vars(\prog)$ are distinguished by $\ttt{in1}$ or $\ttt{in2}$. 
+  \item $\gc{\progd(\prog,v_1)}{\itv}{\ttt{owl(s1)}} \quad \text{and} \quad
+    \gc{\progd(\prog,v_2)}{\itv}{\ttt{owl(s2)}}$
+  \item $\sep{\progd(\prog,v_2)}{\{\ttt{s[2,s2]\}}}{(\vdefs(\prog) \cup \itv)}$
   \end{enumerate}
 \end{definition}
+Decode certification establishes separation of the input wire
+label from the decoding tables under any input value condition. 
+\begin{definition}[Decode Certification]
+  Let $\mathit{decode}$ be a decoding function, and let:
+  $$
+  \config{\varnothing}{\mathit{decode}(\ttt{c}, \ttt{owl(c)})}
+  \redxs \config{\prog_1}{\be_1}
+  \quad \text{and} \quad
+  \config{\varnothing}{\mathit{decode}(\ttt{c}, \ttt{invert(owl(c))})}
+  \redxs \config{\prog_0}{\be_0}
+  $$
+  Then $\mathit{decode}$ is certified iff:
+  $$
+  \sep{\progtt(\prog_i)}{\ttt{\{flip[2,gate:c.k],flip[2,gate:c.p]\}}}{\vdefs(\prog_i)}
+  $$
+  %Then $\mathit{decode}$ is certified iff both of the following conditions hold:
+  %\begin{enumerate}[\hspace{5mm}i.]
+  %\item $\progd(\prog_1,\be_1)(\{ \itv \mapsto 1 \}) = 1$ and
+  %  $\progd(\prog_0,\be_0)(\{ \itv \mapsto 0 \}) = 1$
+  %\item $\sep{\progtt(\prog_i)}{\ttt{\{flip[2,gate:c.k],flip[2,gate:c.p]\}}}{\vdefs(\prog_i)}$
+  %\end{enumerate}
+\end{definition}
+We have certified each of the YGC components detailed above. 
+\begin{lemma}
+  \label{lemma-certification}
+  Each of $\ttt{decode}$, $\ttt{encode}$, and $\ttt{andgg}$ are certified.
+\end{lemma}
 
+\subsection{Correctness of Certified Components}
+\label{section-composition-metatheory}
+
+To demonstrate correctness of certification, we need to show that
+isolated component certificates preserve relevant properties
+when integrated into larger programs. This resonates with
+\emph{framing} as developed in \cite{barthe2019probabilistic}.
+In the following we will use $f$ to range over flips.
+First, we show that if a flip $f$ is independent from the views
+in a protocol (as is the case with input labels used to
+encode a garbled table), then separation from views that
+use $f$ (e.g., in an output label) is preserved. 
+\begin{lemma}[Wire Framing]
+  \label{lemma-wire-framing}
+  If $\sep{\progtt(\prog)}{\{f\}}{\vdefs(\prog)}$ for flip $f$ and
+  $\vars(\prog) \cap \vars(\prog') = \{ f \}$ then
+  $\sep{\progtt(\prog';\prog)}{\vdefs(\prog')}{\vdefs(\prog)}$.
+\end{lemma}
+Next, we show that output correlations (e.g., between output wire values and
+labels) are preserved by substitution of correlated values for flips (e.g.,
+between input wire values and labels).
+\begin{lemma}[Substitution]
+  \label{lemma-substitution}
+  If $\vc{\progtt(\prog,\be)}{\itv}{f}$ and $\vc{\progtt(\prog',\be')}{\itv}{f'}$
+  and $\sep{\progtt(\prog';\prog)}{\vdefs(\prog')}{\vdefs(\prog)}$
+  then $\vc{\progtt(\prog';\prog,\be[\be'/f'])}{\itv}{f}$.
+\end{lemma}
+Our wire framing and substitution Lemmas then allow us to generalize
+certification of gates and encoding to arbitrary circuit contexts,
+assuming that outputs are not wired to multiple inputs \cite{tate2003garbled,nieminen2023breaking}.
+\begin{lemma}[Gate Composition]
+  \label{lemma-compose}
+Given well-typed $\prog_1;E[\mathit{gate}(g,g_1,g_2,v_1,v_2)]$ and certified $\mathit{gate}$ where:
+$$
+\config{\prog_1}{E[\mathit{gate}(g,g_1,g_2,v_1,v_2)]} \redxs \config{\prog_1;\prog_2}{E[v]}
+$$
+If the following preconditions hold:
+\begin{enumerate}
+\item $\{ g_1 \} \cap \{ g_2 \} \cap \wired(\prog_1) = \varnothing$ and $g \not\in \prog_1$
+\item $\gc{\progd(\prog_1)}{v_1}{\owl(g_1)}$ and $\gc{\progd(\prog_1)}{v_2}{\owl(g_2)}$
+\end{enumerate}
+then we have as postconditions:
+\begin{enumerate}
+  \item $\gc{\progd(\prog_1;\prog_2)}{v}{\owl(g_2)}$
+  \item $\sep{\progd(\prog_1;\prog_2,v)}{\vdefs(\prog_1)}{\vdefs(\prog_2) \cup \{\itv\}}$
+\end{enumerate}
+\end{lemma}
+\begin{proof}
+  Given that $g_1$ and $g_2$ are distinct and not wired in $\prog_1$
+  we are assured that $\ttt{owl}(g_1)$ and $\ttt{owl}(g_1)$ are in
+  independent uniform distributions, and given that $g \not\in
+  \prog_1$ we are assured that $\ttt{owl}(g)$ contains entirely fresh
+  flips. Thus, precondition 2, Lemma \ref{lemma-wire-framing}, and
+  Definition \ref{definition-gate-certification} imply postcondition
+  2.  Furthermore, since gate certification guarantees that
+  correlation with $\ttt{owl}(g)$ is guaranteed under all possible gate
+  input wire values, precondition 2 and Lemma \ref{lemma-substitution}
+  guarantee postcondition 2.
+\end{proof}
+
+\begin{lemma}[Decode Composition]
+  \label{lemma-decode}
+Given well-typed $\prog_1;E[\mathit{decode}(g,v)]$ and certified $\mathit{decode}$ where 
+$$
+\config{\prog_1}{E[\ttt{decode}(g,v)]} \redxs \config{\prog_1;\prog_2}{E[\be]}
+$$
+If the following preconditions hold:
+\begin{enumerate}
+\item $g \not\in \cap \wired(\prog_1)$
+\item $\gc{\progd(\prog_1)}{v}{\owl(g)}$
+\end{enumerate}
+then we have as postcondition:
+\begin{enumerate}
+  \item $\sep{\progd(\prog_1;\prog_2)}{\vdefs(\prog_1)}{\vdefs(\prog_2)}$
+\end{enumerate}
+\end{lemma}
+Given the above, we can establish that any circuit built with certified
+components is passive secure. 
 \begin{theorem}
-  \label{theorem-gmw-NIMO}
-  If $e$ is a well-typed GMW circuit definition using certified
-  components and $\config{\varnothing}{e} \redxs
-  \config{\prog}{\varnothing}$, then $\NIMO(\prog)$.
+  \label{theorem-ygc-NIMO}
+  If $e$ is a well-typed YGC circuit definition using certified components
+  where no gate output is wired more than once, and $\config{\varnothing}{e}
+  \redxs \config{\prog}{\varnothing}$, then $\NIMO(\prog)$. 
 \end{theorem}
+
 \begin{proof}
-  Let $\iov(\prog) = S \cup (V \cup \outv)$. Noting that any encoding
-  necessarily uses entirely fresh variables by Definition and type
-  safety, by applying Lemmas \ref{lemma-compose-gmw} and
-  \ref{lemma-wire-framing} and the frame rule of
+  Let $\iov(\prog) = S \cup (V \cup \outv)$. Noting that any
+  encoding necessarily uses entirely fresh variables by Definition and
+  type safety, by applying Lemmas \ref{lemma-compose} and
+  \ref{lemma-decode} and the frame rule of
   \cite{barthe2019probabilistic} we obtain:
   $$
-  \sep{\progd(\prog)}{(S_{\{2\}}\cup\{\outv\})}{V_{\{1\}}}
-  \quad \text{and} \quad \sep{\progd(\prog)}{(S_{\{1\}}\cup\{\outv\})}{V_{\{2\}}}
+  \sep{\progd(\prog)}{(S_{\{2\}}\cup\{\outv\})}{V}
   $$
-  That is, the marginal distribution of client 1's secrets and the output are independent
-  of client 2's views, and vice-versa. The result follows. 
+  Noting that the garbler (client 2) has no views, this means that
+  the garbler's secrets are independent of the evaluator's views. The
+  result follows by Definition \ref{definition-NIMO}.
 \end{proof}
+
+\begin{fpfig}[t]{YGC copy gate definitions.}{fig-ygc-copy}
+  {\footnotesize
+    \begin{verbatimtab}
+      sharetab2(gid, tid, k, p, b)
+      {
+        let r1 = k xor b in
+        let r0 = (not k) xor (not b) in
+        v[1,gate: || gid || tid || 1] := select(p,r1,r0);
+        v[1,gate: || gid || tid || 2] := select(not p,r1,r0);
+      }
+
+      copygate(ca,cb,g)
+      {
+        let wl = owl(g) in
+        let owl1 = owl(ca) in
+        let owl2 = owl(cb) in
+        sharetab2(ca,tt,wl.k,wl.p,owl1.k); sharetab2(ca,pt,wl.k,wl.p,owl1.p);
+        sharetab2(cb,tt,wl.k,wl.p,owl2.k); sharetab2(cb,pt,wl.k,wl.p,owl2.p)
+      }
+
+      evalcopy(ca,cb,wv)
+      {
+        let wv1k = wv.k xor select(wv.p,v[1,gate: || ca || tt1], v[1,gate: || ca || tt2]) in
+        let wv1p = wv.k xor select(wv.p,v[1,gate: || ca || pt1], v[1,gate: || ca || pt1]) in
+        let wv2k = wv.k xor select(wv.p,v[1,gate: || cb || tt1], v[1,gate: || cb || tt2]) in
+        let wv2p = wv.k xor select(wv.p,v[1,gate: || cb || pt1], v[1,gate: || cb || pt1]) in
+        { wv1 = { k = wv1k; p = wv1p }; wv2 = { k = wv2k; p = wv2p } }  
+      }
+
+      copy(ca,cb,g,wv) { copygate(ca,cb,g); evalcopy(ca,cb,wv) } 
+    \end{verbatimtab}
+  }
+\end{fpfig}
+
+\subsection{Building and Extending Automatically Secure Circuits}
+\label{section-composition-copy}
+
+Theorem \ref{theorem-nimo} means that any well-typed program using
+$\ttt{encode}$, $\ttt{andgg}$, and $\ttt{decode}$ will generate a
+circuit that is passive secure, provided that gate outputs are not
+used more than once. To add an $\ttt{or}$ gate to our library, we can
+use $\ttt{andgg}$ as a template, in fact simple modification of
+$\ttt{andtable}$ is all that would be needed. Thus, this library
+embodies a Fairplay-like language where well-typedness of metaprograms
+guarantees safety and security of generated protocols.
+
+The requirements that output wire labels are not used more than once
+reflects a known result, that naive copy gate output is unsound in
+YGC- i.e., passive security fails. In fact, we are unable to
+automatically verify $\gNIMO$ for $\ttt{andgg}$ without the assumption
+that input wire lables are in independent uniform distributions. We
+can capture the necessary fix as in Figure \ref{fig-ygc-copy}. The
+garbler creates a copy gate with two new output wire labels in
+independent uniform distributions (in \ttt{copygate}), and permutes
+them using the input wire label pointer after encrypting with the
+input wire label keys. The evaluator then recovers the encrypted
+output wire values during evaluation (in \ttt{evalcopy}). Given an
+$\ttt{orgg}$ gate defined as outlined abotve, we can then use
+secure $\ttt{copy}$ as follows: 
+\begin{verbatimtab}
+  let s = encode(s1,s2) in
+  let s1c = copy(c1s1,c2s1,s1,s.wv1) in
+  let s2c = copy(c1s2,c2s2,s2,s.wv2) in
+  let wv1 = andgg(1,c1s1,c1s2,s1c.wv1,s2c.wv1) in
+  let wv2 = orgg(2,c2s1,c2s2,s1c.wv2,s2c.wv2) in
+  decode(3,andgg(3,1,2,wv1,wv2))
+\end{verbatimtab}
+These definitions of $\ttt{copy}$ and $\ttt{orgg}$ have been
+automatically verified as in Section \ref{section-pre-post} for
+$\ttt{andgg}$. In general, any new gate form- such as optimized gates
+\cite{XXX}- can be verified automatically in a similar way.
+
+While the copy gate supports secure programming patterns, security
+still relies on programmer discipline, i.e., not wiring a gate more
+than once. However, we can keep track of wired gates in type effects,
+similar to the manner in which we keep track of defined views- type
+dependence in our system is able to accurately precisely track
+gate identifiers. This is also reminiscent of the use of linearity
+in other related systems for enforcing obliviousness \cite{darais2019language}.

### 2023-07-06 popl24/minicat.tex
checkpoint

@@
 \cap \dom(\store_2)$, otherwise $\store_1 \cap \store_2$ is undefined.
+We write $\store_1 \subseteq \store_2$ iff $\store_1 \cap \store_2
+= \store_1$.
 
@@
   \label{example-OT}
-  In \cite{barthe2019probabilistic} a protocol for $\ttt{OT}$ is defined that we recreate as
-  follows, where client 2 is the receiver and client 1 is the sender. Whereas
-  in \cite{barthe2019probabilistic} a trusted third party provides the sender and receiver with
-  shared randomness, in our version both clients refer to the oracle $\ttt{H}$
-  to obtain shared random values. 
+  In \cite{barthe2019probabilistic} a protocol for $\ttt{OT}$ is
+  defined that we recreate as follows, where client 2 is the receiver
+  and client 1 is the sender, and a trusted third party provides the
+  sender and receiver with shared randomness. The secret \ttt{s[2,c]}
+  is the receiver's choice bit, and \ttt{s[1,m0]} and \ttt{s[1,m1]}
+  are the sender's two message options that the receiver chooses from. 
 \begin{verbatimtab}
-  v[2, rd] := select(flip[2, d], H[r1], H[r0]);
-  v[1, e] := s[2, c] xor flip[2, d];
-  v[2, f0] := s[1, m0] xor select(v[1, e], H[r1], H[r0]);
-  v[2, f1] := s[1, m1] xor select(not v[1, e], H[r1], H[r0]);
-  v[2, mc] := select(s[2, c], v[2, f1], v[2, f0]) xor v[2, rd]
+  v[1,r0] := s[3,r0];
+  v[1,r1] := s[3,r1];
+  v[2,d] := flip[3,d];
+  v[2,rd] := select(flip[3,d], s[3,r1], s[3,r0]);
+  v[1,e] := s[2,c] xor v[2,d];
+  v[2,f0] := s[1,m0] xor select(v[1, e],v[1,r1],v[1,r0]);
+  v[2,f1] := s[1,m1] xor select(not v[1, e],v[1,r1],v[1,r0]);
+  v[2,mc] := select(s[2,c],v[2,f1],v[2,f0]] xor v[2,rd)
 \end{verbatimtab}
@@
 of this protocol that correspond to the properties proven manually in
-\cite{barthe2019probabilistic}, and additionally how to automatically verify its passive
-security in the simulator model. With respect to our language primitives,
-we note that this protocol is probabilistically equivalent to the
-following:
+\cite{barthe2019probabilistic}, and additionally how to automatically
+verify its passive security in the simulator model. With respect to
+our language primitives, we note that this protocol is equivalent to
+the following:
 \begin{verbatimtab}  
-  v[2, mc] := OT(s[2, c], s[1, m1], s[1, m0])
+  v[2,mc] := OT(s[2,c], s[1,m1], s[1,m0])
 \end{verbatimtab}

### 2023-07-06 popl24/hyperprop.tex
checkpoint

@@
 
-In this section, we formulate a new hyperproperty that we show is sound
-for passive security in the real/ideal model, called
+In this section, we formulate a new hyperproperty that we show is
+sound for passive security in the real/ideal model, called
 \emph{noninterference modulo output ($\NIMO$)}. We do not claim
@@
 compositional properties supporting efficient, scalable enforcement of
-complex protocols that incorporate modular subprotocols, with a
-specific use-case of Yao's Garbled Circuits (YGC) as an example where
-garbled gates are the compositional units.  The latter methodology
-demonstrates a particular advantage of our hyperproperty formulation.
+complex protocols that incorporate modular subprotocols, using
+both YGC and GMW circuits as examples where gates are
+subprotocols. The latter methodology demonstrates a particular
+advantage of our intensional formulation.
 
@@
   denote the distribution of corrupt views reconstructed by the
-  simulation, which is a distribution $\pdf{V}$ where for
+  simulation, where for
   all $\store' \in \mems(V)$:
   $$
-  \pdf{V}(\store')\ \defeq\ \prob(\SIM_C(\store,v) = \store') 
+  \progd(\SIM_C(\store,v))(\store')\ \defeq\ \prob(\SIM_C(\store,v) = \store') 
   $$
@@
   $$
-  \progd(\SIM_C(\store_{S_C},\idealf(\store))) = \margd{(\condd{\progd(\prog)}{\store})}{V_C}
+  \progd(\SIM_C(\store_{S_C},\idealf(\store))) = \condd{\progd(\prog)}{V_C}{\store}
   $$
@@
   $$\store_1 =_C \store_2 \implies
-    \margd{(\condd{\progd(\prog)}{\store_1})}{V_C} =
-    \margd{(\condd{\progd(\prog)}{\store_2})}{V_C}$$
+    \condd{\progd(\prog)}{V_C}{\store_1} = \condd{\progd(\prog)}{V_C}{\store_2}$$
 \end{definition}
@@
   $\store \in \mems(S \cup V)$:
-  $$\margd{(\condd{\progd(\prog)}{\store_{S_C}})}{S_H} =
-    \margd{(\condd{\progd(\prog)}{\store_{(S\cup V)_C}})}{S_H} $$
+  $$\condd{\progd(\prog)}{S_H}{\store_{S_C}} =
+    \condd{\progd(\prog)}{S_H}{\store_{(S\cup V)_C}} $$
 \end{lemma}
@@
   $\store_1,\store_2 \in \mems(S)$ we have:
-  \begin{eqnarray*}
-    & \store_1 =_C \store_2 \wedge
-     (\margd{(\condd{\progd(\prog)}{\store_1})}{\{ \outv \}} =
-     \margd{(\condd{\progd(\prog)}{\store_2})}{\{ \outv \}}) \\
-    & \implies \\
-    & \margd{(\condd{\progd(\prog)}{\store_1})}{V_C} =
-    \margd{(\condd{\progd(\prog)}{\store_2})}{V_C}
-  \end{eqnarray*}
+  $$
+  (\store_1 =_C \store_2 \ \wedge \ 
+  (\condd{\progd(\prog)}{\{ \outv \}}{\store_1} = \condd{\progd(\prog)}{\{ \outv \}}{\store_2}))
+  \implies 
+  (\condd{\progd(\prog)}{V_C}{\store_1} = \condd{\progd(\prog)}{V_C}{\store_2})
+  $$
   where $\iov(\prog) = (S,V)$.
@@
   $$
-  We say that $\idealf$ is \emph{invertible} iff $\ik(\idealf, v)$ for all
+  We say that $\idealf$ is \emph{pre-imageable} iff $\ik(\idealf, v)$ for all
   $v$ can be computed tractably.
@@
 Now we can show our main result, that noninterference modulo output,
-together with the assumption that $\idealf$ is invertible, implies
+together with the assumption that $\idealf$ is pre-imageable, implies
 passive security of a protocol.
@@
   \label{theorem-nimo}
-  Assume given invertible $\idealf$ and a protocol $\prog$ that
+  Assume given pre-imageable $\idealf$ and a protocol $\prog$ that
   correctly implements $\idealf$.  If $\NIMO(\prog)$
@@
   $\iov(\prog) = (V,S)$ and $\flips(\prog) = F$ with output view $x
-  \in V$ and let $\store$ be an arbitrary member of $\mems(S \cup
+  \in V$ and let $\store$ be an arbitrary element of $\mems(S \cup
   F)$. Then the distribution of adversarial views in the real world
   is, by definition:
-  $$\margd{(\condd{\progd(\prog)}{\store_S})}{V_C}$$
+  $$\condd{\progd(\prog)}{V_C}{\store_S}$$
 
@@
   $$
-   (\margd{(\condd{\progd(\prog)}{\store'})}{V_C})(\store^{\SIM}_{V_C})
+  \progd(\prog)(\store^{\SIM}_{V_C} | \store')
   $$
@@
   \store_S =_C \store' \wedge
-     (\margd{(\condd{\progd(\prog)}{\store_S})}{\{ x \}} =
-      \margd{(\condd{\progd(\prog)}{\store'})}{\{ x \}})
+     \condd{\progd(\prog)}{\{ \outv \}}{\store_S} =
+     \condd{\progd(\prog)}{\{ \outv \}}{\store'}
   $$
-  so by the definition of noninterference modulo output the choice
-  any choice of $\store'$ yields the same distribution, i.e.:
-  $$\margd{(\condd{\progd(\prog)}{\store_S})}{V_C}$$
-  Thus, by definition:
-  $$
-   \progd(\SIM_C(\store_{S_C},\idealf(\store))) = \margd{(\condd{\progd(\prog)}{\store_S})}{V_C}
-  $$
+  so by the definition of noninterference modulo output
+  any choice of $\store'$ yields the same distribution $\condd{\progd(\prog)}{V_C}{\store_S}$.
+  Therefore by definition
+  $
+   \progd(\SIM_C(\store_{S_C},\idealf(\store))) = \condd{\progd(\prog)}{V_C}{\store_S}
+  $.
 \end{proof}
@@
 just the corrupt secrets and input. In addition to providing
-insights about $\NIMO$, this form is helpful for
+insights about $\NIMO$, this form is convenient for
 verification techniques. 
@@
   $\store \in \mems(S \cup V)$:
-  $$\margd{(\condd{\progd(\prog)}{\store_{(S \cup \{\outv\})_C}})}{S_H} =
-  \margd{(\condd{\progd(\prog)}{\store_{(S\cup V)_C}})}{S_H} $$
+  $$\condd{\progd(\prog)}{S_H}{\store_{(S \cup \{\outv\})_C}} =
+  \condd{\progd(\prog)}{S_H}{\store_{(S\cup V)_C}} $$
   where  $\iov(\prog) = (S,V)$.

### 2023-07-06 popl24/pmf.tex
checkpoint

@@
 
-In order to precisely characterize distributions of secrets and views
-in programs, we define probability mass functions in a standard manner,
-following the formulation in \cite{barthe2019probabilistic}. In this formulation, memories
-denote conjunctions of variable assignments. Intuitively, given a
-program $\prog$ with $\iov(\prog) = S \cup V$ and $\store \in
-\mems(\iov(\prog))$, when considering the execution of $\prog$
-initialized with secrets $\store_S$, we are fundamentally concerned
-with the joint probability of the assignments in $\store_V$
-conditioned by the assignments $\store_S$. This is because
-probabilistic dependence between honest secrets and corrupt views
-\emph{is} information leakage.
+Given a program $\prog$ with $\iov(\prog) = S \cup V$ and $\store \in
+\mems(\iov(\prog))$, we are fundamentally concerned with the joint
+probability of the assignments in $\store_V$ conditioned by the
+assignments $\store_S$ in any given run, since probabilistic
+dependence between honest secrets and corrupt views is our metric of
+information leakage.
 
-\subsection{Basic Definitions} We begin by defining memory probability distributions as follows.
+\subsection{Basic Definitions}
+
+We begin by defining joint probability mass functions in the standard
+manner, though following \cite{barthe2019probabilistic} we use
+memories to denote mappings of variables to values (i.e., outcomes),
+so for example given a pmf $\pmf$ we will write $\pmf(\{ \sx{1}{a}
+\mapsto 0, \vx{2}{b} \mapsto 1 \})$ to denote the probability that
+$\sx{1}{a} = 0$ and $\vx{2}{b} = 1$ in the final memory of a
+program containing these variables.
 \begin{definition}
-  A \emph{memory probability distribution} $\pdf{X}$ is a function
-  mapping memories in $\mems(X)$ to values in $[0,1]$, such that:
+  A \emph{probability mass function} $\pmf$ is a function
+  mapping memories in $\mems(X)$ for some $X$ to values in $[0,1]$, such that:
   $$
-  \sum_{\store \in \mems(X)} \pdf{X}(\store) \  = \ 1
+  \sum_{\store \in \mems(X)} \pmf(\store) \  = \ 1
   $$
-  We call $X$ the \emph{domain} of $\pdf{X}$.
 \end{definition}
-To recover succinct and familiar notation, we may omit the domain of a
-distribution when it is clear from an application context-
-i.e., we allow the following sugaring:
-$$
-\pdf{}(\store) \defeq \pdf{\dom(\store)}(\store)
-$$
+%To recover succinct and familiar notation, we may omit the domain of a
+%distribution when it is clear from an application context-
+%i.e., we allow the following sugaring:
+%$$
+%\pdf{}(\store) \defeq \pdf{\dom(\store)}(\store)
+%$$
 Now, we can define a notion of marginal and conditional
@@
 \begin{definition}
-  Given $X$ the \emph{marginal distribution} of $Y \subseteq X$
-  in a distribution $\pdf{X}$, denoted $\margd{\pdf{X}}{Y}$,
-  is a distribution with domain $Y$ where for all
-  $\store \in \mems(Y)$:
+  Given $\pmf$ the \emph{marginal distribution} of variables $X$
+  in $\pmf$, denoted $\margd{\pmf}{X}$, is defined as follows:
   $$
-  (\margd{\pdf{X}}{Y})(\store) =
-  \sum_{\store' \in \mems(X-Y)} \pdf{X}(\store \cap \store')
+  \forall \store \in \mems(X) \quad . \quad \margd{\pmf}{X}(\store) =
+  \sum_{\store' \in \mems(X-\dom(\dom(\pmf)))} \pmf(\store \cap \store')
   $$
@@
 \begin{definition}
-  Given $\pdf{X}$, its \emph{conditional distribution given
+  Given $\margd{\pmf}{X}$, its \emph{conditional distribution given
   $\stores$} with $\dom(\stores) = Y$ with $Y \subseteq X$, denoted
-  $\condd{\pdf{X}}{\stores}$, is a distribution with domain $X$ where for all
+  $\condd{\pmf}{X}{\stores}$, is a distribution with domain $X$ where for all
   $\store \in \mems(X)$:
   $$
-  (\condd{\pdf{X}}{\stores})(\store) =
-  (\sum_{\store' \in \stores} \pdf{X}(\store \cap \store')) /
-  (\sum_{\store' \in \stores} (\margd{\pdf{X}}{Y})(\store'))
+  \condd{\pmf}{X}{\stores}(\store) =
+  (\sum_{\store' \in \stores} \margd{\pmf}{X}(\store \cap \store')) /
+  (\sum_{\store' \in \stores} \margd{\pmf}{Y}(\store'))
   $$
-  where $\pdf{X}(\store \cap \store')) = 0$ if $\store \cap \store'$ is undefined.
-  we write $\condd{\pdf{X}}{\store}$ as an abbreviation for $\condd{\pdf{X}}{\{ \store \}}$.
+  where $\margd{\pmf}{X}(\store \cap \store')) = 0$ if $\store \cap \store'$ is undefined.
 \end{definition}
@@
 \begin{eqnarray*}
-\pdf{X}(\store|\store') &\defeq& (\condd{\pdf{X}}{\store})(\store')\\
-\pdf{X}(\store) &\defeq& (\margd{\pdf{X}}{\dom(\store)})(\store) \qquad \dom(\store) \subset X
+  \condd{\pmf}{X}{\store}  &\defeq& \condd{\pmf}{X}{\{ \store\}}\\
+  \pmf(\store)  &\defeq& \margd{\pmf}{X}(\store)  \qquad \dom(\store) = X\\
+  \pmf(\store|\stores)  &\defeq& \condd{\pmf}{X}{\stores}(\store) \qquad \dom(\store) = X
 \end{eqnarray*}
@@
   We note immediately that $|\stores| = 2^{S \cup F}$ given the (deterministic) definition of $\redx$. 
-  Define also $\pdf{S \cup V \cup F}$ as the program's \emph{basic distribution} such that for all
+  Define also $\pmf$ as the program's \emph{basic distribution} such that for all
   $\store \in \mems(S \cup V \cup F)$:
   $$
-  \pdf{S \cup V \cup F}(\store) =
+  \pmf(\store) =
   \begin{cases}
@@
   $$
-  \progd(\prog) =  \margd{\pdf{S \cup V \cup F}}{S\cup V}
+  \progd(\prog) =  \margd{\pmf}{S\cup V}
   $$

### 2023-07-06 popl24/composition.tex
checkpoint

@@
 from \cite{barthe2019probabilistic} with the same denotations.
-\cnote{I switched here to using $\pmf$ to denote an arbitrary probability mass
-function, will update earlier sections with this notation.}
 \begin{definition}
@@
     $\store \in \mems(X \cup Y)$:
-  $$(\margd{\pmf}{Y \cup Z})(\store) =
+  $$\margd{\pmf}{Y \cup Z}(\store) =
   \pmf(\store_Y) * \pmf(\store_Z)$$ 
@@
   \label{lemma-gmw-frame}
-  If $\sep{\progtt(\prog_1;\prog_2)}{Y}{V}$ for $V \in \vdefs(\prog_2)$ and $\sep{\progtt(\prog_1)}{X}{Y}$ for
+  If $\sep{\progtt(\prog_1;\prog_2)}{Y}{V}$ for $Y = \vars(\prog_1) \cap \vars(\prog_2)$ and
+  $V \in \vdefs(\prog_2)$ and $\sep{\progtt(\prog_1)}{X}{Y}$ for
   $X \not\in \vars(\prog_2)$ then $\sep{\progtt(\prog_1;\prog_2)}{X}{V}$. 

### 2023-07-06 popl24/metalang.tex
checkpoint

@@
 iff $f(x_1,\ldots,x_n) \{ e \} \in \codebase$.
-A \emph{metaprogram}, aka \emph{mataprotocol} \dnote{meta?} is a pair of a 
+A \emph{metaprogram}, aka \emph{metaprotocol}  is a pair of a 
 codebase and expression $\codebase, e$. We may omit
@@
 for free occurences of $x$ in $e$. The rules are mostly standard,
-except why \dnote{should this be a when?} a concrete $\minifed$ assignment is encountered it is added
+except when a concrete $\minifed$ assignment is encountered it is added
 to the end of $\prog$.

### 2023-07-06 popl24/composition.tex
checkpoint

@@
 
-Now we can define GMW gate certification as follows. Intuitively, the
-input wire shares of each client should be independent of the gate output
-shares. This ensures independence of clients' shares (views).
+Now we can define GMW gate certification as follows. Intuitively, we certify
+that GMW gates satisfy a probabilistic noninterference property wrt each output
+share considered individually. This ensures that input dependencies remain
+encrypted from each client's perspective during circuit evaluation.
 \begin{definition}[GMW Gate Certification]
@@
   \begin{enumerate}[\hspace{5mm}i.]
-  \item $\sep{\progtt(\prog)}{\ttt{\{flip[1,s11],flip[1,s21]\}}}{\ttt{\{v[1,gout],v[2,gout]\}}}$
-  \item $\sep{\progtt(\prog)}{\ttt{\{flip[2,s12],flip[2,s22]\}}}{\ttt{\{v[1,gout],v[2,gout]\}}}$
+  \item $\sep{\progtt(\prog)}{\ttt{\{flip[1,s11],flip[1,s21],flip[2,s12],flip[2,s22]\}}}{\ttt{\v[1,gout]\}}}$
+  \item $\sep{\progtt(\prog)}{\ttt{\{flip[1,s11],flip[1,s21],flip[2,s12],flip[2,s22]\}}}{\ttt{\v[2,gout]\}}}$
   % \item $\sep{\progtt(\prog)}{\ttt{\{v[1,gout]\}}}{\ttt{\{v[2,gout]\}}}$
@@
   \item $\sep{\pmf}{X}{(Y \cup Z)}$ implies $(\sep{\pmf}{X}{Y}$ and $\sep{\pmf}{X}{Z})$
-  \item $(\sep{\pmf}{X}{Y}$ and $\sep{\pmf}{X}{Z}$ and $\sep{\pmf}{Y}{Z})$ implies $\sep{\pmf}{X}{(Y \cup Z)}$ 
+  \item $(\sep{\pmf}{X}{Y}$ and $\sep{\pmf}{(X \cup Y)}{Z})$ implies $\sep{\pmf}{X}{(Y \cup Z)}$ 
   %\item If $\prog_1;\prog_2$ is safe and $\vars(\prog_1) \cap \vars(\prog_2) = \varnothing$
@@
 The following property also follows by results in \cite{barthe2019probabilistic} and will
-be useful to make constructions that demonstrate variable separation (especially in Lemma
-\ref{lemma-gmw-frame}).
+be useful to make constructions that demonstrate variable separation.
 \begin{lemma}
@@
 
-The following Lemma allows us to separate variables in programs that may not occur
-in the same subprogram, and hence consider gate certification properties in various
-contexts. It is reminiscent of the \emph{frame} rule of \cite{barthe2019probabilistic}.
-\begin{lemma}
-  \label{lemma-gmw-frame}
-  If $\sep{\progtt(\prog_1;\prog_2)}{Y}{V}$ for $Y = \vars(\prog_1) \cap \vars(\prog_2)$ and
-  $V \in \vdefs(\prog_2)$ and $\sep{\progtt(\prog_1)}{X}{Y}$ for
-  $X \not\in \vars(\prog_2)$ then $\sep{\progtt(\prog_1;\prog_2)}{X}{V}$. 
+%The following Lemma allows us to separate variables in programs that may not occur
+%in the same subprogram, and hence consider gate certification properties in various
+%contexts. It is reminiscent of the \emph{frame} rule of \cite{barthe2019probabilistic}.
+%\begin{lemma}
+%  \label{lemma-gmw-frame}
+%  If $\sep{\progtt(\prog_1;\prog_2)}{Y}{V}$ for $Y = \vars(\prog_1) \cap \vars(\prog_2)$ and
+%  $V \in \vdefs(\prog_2)$ and $\sep{\progtt(\prog_1)}{X}{Y}$ for
+%  $X \not\in \vars(\prog_2)$ then $\sep{\progtt(\prog_1;\prog_2)}{X}{V}$. 
+%\end{lemma}
+%\begin{proof}
+%  By assumption and Lemma \ref{lemma-sepjoin}, for all
+%  $\store^1,\store^2 \in \runs(\prog_1;\prog_2)$ we have $\store^1_Y
+%  \cap \store^2_V \subseteq \store_a$ for some $\store_a \in
+%  \runs(\prog_1;\prog_2)$ and $\store^1_X \cap \store^2_Y \subseteq
+%  \store_b$ for some $\store_b \in \runs(\prog_1;\prog_2)$.  Thus, since
+%  we assume $X \not\in \vars(\prog_2)$ we have $\store_a \cap \store_b
+%  \in \runs(\prog_1;\prog_2)$ and $\store^1_X \cap \store^2_V =
+%  \store_a \cap \store_b$, obtaining the result by
+%  \ref{lemma-sepjoin}.
+%\end{proof}
+
+The next ``noninterference'' Lemma is important to prove the desired invariance
+preservation property of GMW gates, given certification.
+\begin{lemma}[Noninterference]
+  \label{lemma-noninterference}
+  Given $\prog_1;\prog_2$ and $X = \iov(\prog_2) - \vdefs(\prog_2)$ and
+  $Y \subseteq \vdefs(\prog_2)$. If $\sep{\progd(\prog_1;\prog_2)}{X}{Y}$
+  then $\sep{\progd(\prog_1;\prog_2)}{\iov(\prog_1)}{Y}$.
 \end{lemma}
-\begin{proof}
-  By assumption and Lemma \ref{lemma-sepjoin}, for all
-  $\store^1,\store^2 \in \runs(\prog_1;\prog_2)$ we have $\store^1_Y
-  \cap \store^2_V \subseteq \store$ for some $\store_a \in
-  \runs(\prog_1;\prog_2)$ and $\store^1_X \cap \store^2_Y \subseteq
-  \store$ for some $\store_b \in \runs(\prog_1;\prog_2)$.  Thus, since
-  we assume $X \not\in \vars(\prog_2)$ we have $\store_a \cap \store_b
-  \in \runs(\prog_1;\prog_2)$ and $\store^1_X \cap \store^2_V =
-  \store_a \cap \store_b$, obtaining the result by
-  \ref{lemma-sepjoin}.
-\end{proof}
 
@@
   Thus by Lemma \ref{lemma-substitution} we have:
-  $$\sep{\progtt(\prog_1;\prog_2)}{\ttt{\{v[1,g1out],v[1,g2out]\}}}{\ttt{\{v[1,gout],v[2,gout]\}}}$$
+  $$\sep{\progtt(\prog_1;\prog_2)}{\ttt{\{v[1,g1out],v[1,g2out],v[2,g1out],v[2,g2out]\}}}{\ttt{\{v[1,gout]\}}}$$
   and
-  $$\sep{\progtt(\prog_1;\prog_2)}{\ttt{\{v[2,g1out],v[2,g2out]\}}}{\ttt{\{v[1,gout],v[2,gout]\}}}$$
-  And by Lemma \ref{lemma-gmw-frame} and condition (iii) of Definition
+  $$\sep{\progtt(\prog_1;\prog_2)}{\ttt{\{v[1,g1out],v[1,g2out],v[2,g1out],v[2,g2out]\}}}{\ttt{\{v[2,gout]\}}}$$
+  Now, by Lemmas \ref{lemma-noninterference} and \ref{lemma-separation} and condition (iii) of Definition
   \ref{definition-gmwgate-certification} we then have:
-  $$\sep{\progtt(\prog_1;\prog_2)}{S_{\{1\}}}{\ttt{\{v[2,gout]\}}} \quad \text{and} \quad
-  \sep{\progtt(\prog_1;\prog_2)}{S_{\{2\}}}{\ttt{\{v[1,gout]\}}}$$
+  $$\sep{\progtt(\prog_1;\prog_2)}{(S_{\{1\}} \cup V_{\{2\}})}{\ttt{\{v[2,gout]\}}} \quad \text{and} \quad
+  \sep{\progtt(\prog_1;\prog_2)}{(S_{\{2\}} \cup V_{\{1\}})}{\ttt{\{v[1,gout]\}}}$$
   so by precondition assumptions and Lemma \ref{lemma-separation} we have:
@@
   therefore by Lemma \ref{lemma-gmw-frame}}:    
-
   $$

### 2023-07-06 popl24/scratch.tex
checkpoint

@@
-\section{Component Certification and Composition}
-\label{section-composition}
-
-Using YGC as a case study, in this Section we will discuss
-compositional properties that can be automatically verified on small
-program components using techniques developed in Section
-\ref{section-bruteforce}. These properties compose to yield $\NIMO$
-and scale to arbitrary programs. In our metatheory we will express
-these properties as pre-~and post-conditions on components, so our
-general approach is in a proof assistant style.
-
-\subsection{A Library for Yao's Garbled Circuits}
+\subsection{Compositional Properties of YGC}
 \label{section-composition-ygc}
@@
 
-\subsection{Component Certification for $\NIMO$}
+\subsubsection{Component Certification for $\NIMO$}
 \label{section-composition-certification}
@@
 
-\subsection{Correctness of Certified Components}
+\subsubsection{Correctness of Certified Components}
 \label{section-composition-metatheory}

### 2023-07-06 popl24/intro.tex
checkpoint

@@
 Security in the MPC setting thus means that protocols cannot reveal
-any secret information other than what is impicitly declassified by
+any secret information other than what is implicitly declassified by
 the publicized output of the ideal functionality. The security model
@@
 analysis and inner workings of protocols themselves. While the
-probabilistic formulations we develop could conceivably used
+probabilistic formulations we develop could conceivably be used
 extensionally, our main focus is on automated and semi-automated
@@
 approaches such as delimited information release
-\cite{10.1007/978-3-540-37621-7_9} and downgrading policies and
+\cite{10.1007/978-3-540-37621-7_9}, downgrading policies, and
 relaxed noninterference \cite{10.1145/1040305.1040319}. However, their
@@
 multi-threaded protocols enforce the single-threaded source language
-semantics which is an intensional property. All of these systems
+semantics which is an intensional property. \mnote{We need a clearer statement here of the flaws with Wys-star specifically} All of these systems
 leverage type systems to obtain automated security analyses.
@@
 reduce proofs of simulator security to program equivalence assertions,
-which supports applications known techniques in automated verification
+which supports the application of known techniques in automated verification
 \cite{8429300}. However, they focus on noninterference properties,
@@
 allowable information leakage as in MPC, and their use of type
-linearity disallows reuse of sampled probabilistic values which is
+linearity disallows the reuse of sampled probabilistic values which is
 common in the MPC idiom, as our examples will illustrate. Similarly,
@@
 account of \emph{views} and (synchronous) messaging between protocol
-\emph{clients}. In this Section we consider small example protocols
+\emph{clients}. In this Section, we consider small example protocols
 including oblivious transfer, communications encryption with one-time
@@
 \cite{barthe2019probabilistic}, obtaining a middle-ground, scalable
-``proof assistant'' style approach enabled by MPC protocol paradigms.
+``proof assistant style'' approach enabled by MPC circuit paradigms.

### 2023-07-06 popl24/composition.tex
checkpoint

@@
   have $\sep{\progd(\prog)}{S_{\{1\}}}{V_{\{2\}}}$ and
-  $\sep{\progd(\prog)}{S_{\{2\}}}{V_{\{1\}}}$. Since by properties of \ttt{xor} we have:
-  $$
-  \sep{\progd(\prog;\eassign{\outv}{\ttt{v[1,gout] xor v[2,gout]}})}{\{\outv\}}{\{\ttt{v[1,gout]}\}}
-  $$
+  $\sep{\progd(\prog)}{S_{\{2\}}}{V_{\{1\}}}$. This implies:
+  $$\sep{\progd(\prog;\eassign{\outv}{\ttt{v[1,gout] xor v[2,gout]}})}{S_{\{1\}}}{V_{\{2\}}}$$ and:
+  $$\sep{\progd(\prog;\eassign{\outv}{\ttt{v[1,gout] xor v[2,gout]}})}{S_{\{2\}}}{V_{\{1\}}}$$
+  Thus:
+  $\condd{\progd(\prog;\eassign{\outv}{\ttt{v[1,gout] xor v[2,gout]}})}{S_{\{1\}}}{V_{\{2\}} \cup \{ \outv \}}
+  = \condd{\progd(\prog;\eassign{\outv}{\ttt{v[1,gout] xor v[2,gout]}})}{S_{\{1\}}}{\{ \outv \}}$
   and:
-  $$
-  \sep{\progd(\prog;\eassign{\outv}{\ttt{v[1,gout] xor v[2,gout]}})}{\{\outv\}}{\{\ttt{v2,gout]\}}
-  $$
-  therefore by Lemma \ref{lemma-gmw-frame}}:    
-  $$
-  \sep{\progd(\prog;\eassign{\outv}{\ttt{v[1,gout] xor v[2,gout]}})}{(S_{\{1\}} \cup \{ \outv \})}{V_{\{2\}}}
-  $$
-  and:
-  $$
-  \sep{\progd(\prog;\eassign{\outv}{\ttt{v[1,gout] xor v[2,gout]}})}{(S_{\{2\}} \cup \{ \outv \})}{V_{\{1\}}}
-  $$
-  The result follows.
+  $\condd{\progd(\prog;\eassign{\outv}{\ttt{v[1,gout] xor v[2,gout]}})}{S_{\{2\}}}{V_{\{1\}} \cup \{ \outv \}}
+  = \condd{\progd(\prog;\eassign{\outv}{\ttt{v[1,gout] xor v[2,gout]}})}{S_{\{2\}}}{\{ \outv \}}$
 \end{proof}

### 2023-07-06 popl24/composition.tex
checkpoint

--- 2023-07-06 popl24/composition.tex
@@
   have $\sep{\progd(\prog)}{S_{\{1\}}}{V_{\{2\}}}$ and
-  $\sep{\progd(\prog)}{S_{\{2\}}}{V_{\{1\}}}$. Since by properties of \ttt{xor} we have:
-  $$
-  \sep{\progd(\prog;\eassign{\outv}{\ttt{v[1,gout] xor v[2,gout]}})}{\{\outv\}}{\{\ttt{v[1,gout]}\}}
-  $$
-  and:
-  $$
-  \sep{\progd(\prog;\eassign{\outv}{\ttt{v[1,gout] xor v[2,gout]}})}{\{\outv\}}{\{\ttt{v2,gout]\}}
-  $$
-  therefore by Lemma \ref{lemma-gmw-frame}}:    
-  $$
-  \sep{\progd(\prog;\eassign{\outv}{\ttt{v[1,gout] xor v[2,gout]}})}{(S_{\{1\}} \cup \{ \outv \})}{V_{\{2\}}}
-  $$
-  and:
-  $$
-  \sep{\progd(\prog;\eassign{\outv}{\ttt{v[1,gout] xor v[2,gout]}})}{(S_{\{2\}} \cup \{ \outv \})}{V_{\{1\}}}
-  $$
-  The result follows.
+  $\sep{\progd(\prog)}{S_{\{2\}}}{V_{\{1\}}}$.
+  Let $\prog' \defeq \prog;\eassign{\outv}{\ttt{v[1,gout] xor v[2,gout]}}$. The
+  preceding then implies
+  $\sep{\progd(\prog')}{S_{\{1\}}}{V_{\{2\}}}$ and
+  $\sep{\progd(\prog')}{S_{\{2\}}}{V_{\{1\}}}$,
+  so that for all $\beta$:
+  $$\sep{\condd{\progd(\prog')}{S_{\{1\}} \cup V_{\{2\}}}{\{ \outv \mapsto \beta \}}}{S_{\{1\}}}{V_{\{2\}}}
+  \quad \text{and} \quad
+    \sep{\condd{\progd(\prog')}{S_{\{1\}} \cup V_{\{2\}}}{\{ \outv \mapsto \beta \}}}{S_{\{1\}}}{V_{\{2\}}}$$
+  thus for all $\store \in \mems(V_2 \cup \{ \outv \})$:
+  $$\condd{\progd(\prog')}{S_{\{1\}}}{\store}
+  = \condd{\progd(\prog')}{S_{\{1\}}}{\store_{\{\outv\}}}$$
+  and for all $\store \in \mems(V_1 \cup \{ \outv \})$:
+  $$\condd{\progd(\prog')}{S_{\{2\}}}{\store}
+  = \condd{\progd(\prog')}{S_{\{2\}}}{\store_{\{\outv\}}}$$
+  establishing the result by Lemma \ref{lemma-nimo} and Definition \ref{definition-NIMO}. 
 \end{proof}

### 2023-07-06 popl24/composition.tex
checkpoint

@@
   \quad \text{and} \quad
-    \sep{\condd{\progd(\prog')}{S_{\{1\}} \cup V_{\{2\}}}{\{ \outv \mapsto \beta \}}}{S_{\{1\}}}{V_{\{2\}}}$$
+    \sep{\condd{\progd(\prog')}{S_{\{2\}} \cup V_{\{1\}}}{\{ \outv \mapsto \beta \}}}{S_{\{2\}}}{V_{\{1\}}}$$
   thus for all $\store \in \mems(V_2 \cup \{ \outv \})$:

### 2023-07-07 popl24/scratch.tex
checkpoint

@@
 In Figures \ref{fig-ygc-aux} and \ref{fig-ygc-gates} we define a
-codebase for garbled circuits. This definition follows the
-\emph{point-and-permute} method described in \cite{evans2018pragmatic} and elsewhere,
-to which the reader is referred for more in-depth discussion.
+codebase for Yao's garbled circuits (YGC). This definition follows the
+\emph{point-and-permute} method described in \cite{evans2018pragmatic}
+and elsewhere, to which the reader is referred for more in-depth discussion.
 In this implementation client 2 is the \emph{garbler} and
@@
 
-\subsubsection{Component Certification for $\NIMO$}
-\label{section-composition-certification}
+\subsubsection{Gate Certification}
+\label{section-ygc-certification}
 
-We can use the automated verification methods developed in Section
-\ref{section-brute-force} to \emph{certify} YGC
-components. Certification obtains properties that we can then use
-higher-level reasoning-- inspired in part by techniques in
-\cite{barthe2019probabilistic}-- to prove that circuits built with
-certified components satisfy $\NIMO$ (Theorem \ref{theorem-ygc-NIMO})
-and thus are passive secure.
-
-Our method relies on a definition of compositional units comprising
+Our certification method relies on a definition of compositional units comprising
 complete related subprotocols-- i.e., both the garbler's
@@
 
-Certifications will demonstrate probabilistic independence of input
-wire lables and garbled tables in a program's basic distribution, so
-we need to consider dependencies of flips and gate views. Thus we make
-the following definitions:
-\begin{definition}
-  We write $\progtt(\prog)$ to denote the \emph{basic distribution} of
-  $\prog$ as defined in Definition \ref{def-progd}. For any program $\prog =
-  (\eassign{v_1}{\be_1};\ldots;\eassign{v_n}{\be_n})$, define
-  $\vdefs(\prog) \defeq \{ v_1,\ldots,v_n \}$.
-\end{definition}
-We also need to formalize notions of probabilistic independence, aka
-\emph{separation}, and correlation. We borrow the symbols $*$ and $\sim$
-from \cite{barthe2019probabilistic} with the same denotations.
-\begin{definition}
-  We write $\vc{\pdf{X}}{x}{y}$ iff $\pdf{X}(\{ x \mapsto 0\}\ |\ \{ y \mapsto 0 \}) =
-  \pdf{X}(\{ x \mapsto 1\}\ |\ \{ y \mapsto 1 \}) = 1$.
-  We write $\sep{\pdf{X}}{Y}{Z}$ iff for all
-    $\store \in \mems(X \cup Y)$:
-  $$(\margd{\pdf{X}}{Y \cup Z})(\store) =
-  (\margd{\pdf{X}}{Y})(\store_Y) * (\margd{\pdf{X}}{Z})(\store_Z)$$ 
-\end{definition}
-It is also necessary to define notions of correlation with wire
+It is necessary to define notions of correlation with wire
 labels, since wire values flow into and out of gates in either

### 2023-07-07 popl24/composition.tex
checkpoint

@@
 \end{itemize}
+\begin{example}
+  \label{example-gmw-andcircuit}
+The following program uses our GMW library to define
+a circuit with a single And gate and input secrets $\ttt{s1}$ and
+$\ttt{s2}$ from client's 1 and 2 respectively. 
+\begin{verbatimtab}
+  let ss = encodegmw(s1,s2) in
+  v[0,output] := decode(andgmw(0,ss.shares1,ss.shares2))
+\end{verbatimtab}
+\end{example}
 
@@
 method to extend our library.
+\begin{lemma}
+  \label{lemma-gmw-certification}
+  Each of $\ttt{encodegmw}$ and $\ttt{andgmw}$ are certified.
+\end{lemma}
 
-\subsubsection{Gate Invariants} We can show that composition maintains an
-important invariant in a circuit $\prog$- namely that client 1's
-secrets remain independent of client 2's views, and vice-versa, prior
-to decoding.  That is, where $\iov(\prog) = S \cup V$, we show that
+\subsubsection{Gate Invariants and Circuit Security}
+We can show that composition maintains an important invariant in a
+circuit $\prog$- namely that client 1's secrets remain independent of
+client 2's views, and vice-versa, prior to decoding.  That is, where
+$\iov(\prog) = S \cup V$, we show that
 $\sep{\progd(\prog)}{S_{\{1\}}}{V_{\{2\}}}$ and
@@
 for dummy flips preserves separation. 
+In these results and subsequently we will use $f$ to range over flips.
 \begin{lemma}
@@
 
-\begin{lemma}[Substitution]
+\begin{lemma}[Substitution$*$]
   \label{lemma-substitution}
@@
 \end{proof}
+
+\subsection{Compositional Properties of YGC}
+\label{section-composition-ygc}
+
+\input{ygc}
+
+In Figures \ref{fig-ygc-aux} and \ref{fig-ygc-gates} we define a
+codebase for Yao's garbled circuits (YGC). This definition follows the
+\emph{point-and-permute} method described in \cite{evans2018pragmatic}
+and elsewhere, to which the reader is referred for more in-depth discussion.
+In this implementation client 2 is the \emph{garbler} and
+client 1 is the \emph{evaluator}. The garbler builds the garbled
+tables and shares them with the evaluator, who then evaluates
+the gate in an oblivious fashion until the final public output is
+generated through decryption. This definition is well-typed,
+with input type annotations for top-level functions listed in
+Figure \ref{fig-ygc-types}. Well-typed programs using these
+libraries are therefore guaranteed to yield safe $\minifed$
+programs.
+
+\emph{Wire labels} are fundamental to YGC, and essentially represent
+gate output values in an encrypted form. In our definition, wire
+labels are represented by records $\ttt{\{ k = }\beta_1\ttt{; p =
+}\beta_2\ttt{ \}}$, where $\ttt{k}$ is the \emph{key bit} and
+$\ttt{p}$ is the \emph{pointer bit}, and $\beta_1$ and $\beta_2$ are
+flips. Flips in each output wire label are owned by the garbler and
+are unique per gate by definition of their identifying string, and the
+representation of $0$ is the negation of $1$. For example, here is the
+representation of 1 and 0 respectively in the output wire label for a
+hypothetical gate 6:
+\begin{mathpar}
+  \ttt{\{ k = flip[2,gate:6.k]; p =  flip[2,gate:6.p]] \}}
+
+  \ttt{\{ k = not flip[2,gate:6.k]; p =  not flip[2,gate:6.p]] \}}
+\end{mathpar}
+The pointer bits in wire labels are used to select permuted rows in
+table garblings. The key bits are used to identify a unique key for
+table row in each garbled gate. Intuitively, if $\beta_1$ and
+$\beta_2$ are either key or pointer bits encoding 1 on two input wire
+labels to a binary gate, rows and keys in the gate are enumerated in
+the order:
+$$
+\neg\beta_1\neg\beta_2,\ \neg\beta_1\beta_2,\ \beta_1\neg\beta_2,\ \beta_1\beta_2
+$$
+
+In our implementation, gates are wired together using gate
+identifiers, which are strings $w$. Top-level functionality in Figures
+\ref{fig-ygc-aux} and \ref{fig-ygc-gates} includes the following:
+\begin{itemize}
+\item \ttt{andgate}: This defines a subprotocol for the garbler
+  to define a garbled gate $\ttt{gid}$ with input wires from gates
+  $\ttt{ga}$ and $\ttt{gb}$. The garbler generates keys and garbles
+  the rows in YGC fashion, them with client $1$ in
+  views in a standard form. For example, the view for
+  a hypothetical gate 6, row 2 garbled truth table is $\ttt{v[1,gate:6tt2]}$.
+  We note that garbled gates of other binary operators can be obtained with
+  replacement of $\ttt{andtable}$ with appropriate garbled table definitions. 
+\item \ttt{evalgate}: This defines a subprotocol for the evaluator to
+  evaluate gate $\ttt{gid}$ given input wire values $\ttt{wva}$ and
+  $\ttt{wvb}$.
+\item \ttt{garbledecode} and \ttt{evaldecode}: The former function
+  defines the garbler's protocol for encrypting the circuit
+  output from final gate $\ttt{gid}$, and the latter defines
+  the evaluator's output decryption protocol.
+\item \ttt{encode}: This defines the initial phase of the protocol,
+  where the evaluator receives the wire value from their own
+  secret $\sx{1}{sa}$ via $\ttt{OT}$, and the garbler communicates
+  the wire value for their own secret $\sx{2}{sb}$ directly.
+\end{itemize}
+\begin{example}
+  \label{example-andcircuit}
+The following program uses our YGC library to define
+a circuit with a single and gate and input secrets $\ttt{s1}$ and
+$\ttt{s2}$ from client's 1 and 2 respectively. 
+\begin{verbatimtab}
+  andgate(0,s1,s2);
+  garbledecode(0);
+  let secrets = encode(s1,s2) in
+  v[0,output] := decode(evalgate(0, secrets.wv1, secrets.wv2))
+\end{verbatimtab}
+\end{example}
+We have verified passive security of the $\fedprot$ protocol
+generated by this and other small circuits using the
+technique described in Lemma \ref{lemma-bruteforce-nimo}.
+But large circuits with thousands of gates would be
+intractable to verify with this method. In the next Section
+we discuss compositional methods to address this issue.
+
+\subsubsection{Gate Certification}
+\label{section-ygc-certification}
+
+Our certification method relies on a definition of compositional units comprising
+complete related subprotocols-- i.e., both the garbler's
+construction of the gates and the evaluator's evaluation of them. We
+therefore make the following definitions. However, we note that our
+formal results are applicable to the YGC style where the garbler
+shares the entire circuit prior to evaluation.
+\begin{definition}
+  \label{ygc-modules}
+  We define a composable internal \ttt{and} gate sharing and evaluation function
+  called $\ttt{andgg}$, and an output decoder gate share and evaluation function called
+  $\ttt{decode}$, as follows. 
+\begin{verbatimtab}
+  andgg(g, ga, gb, wva, wvb) { andgate(g, ga, gb); evalgate(g, wva, wvb) }
+  decode(g, v) { garbledecode(g); evaldecode(v) }
+\end{verbatimtab}
+\end{definition}
+
+Our certification techniques are defined generally wrt implementations
+of input encoding, decoding, and internal gates. In the following we
+will refer to arbitrary decode, encode, and gate functions with the
+restriction that any function in each of these categories has the same
+valid type signature as $\ttt{decode}$, $\ttt{encode}$, and
+$\ttt{andgg}$ respectively.
+
+It is necessary to define notions of correlation with wire
+labels, since wire values flow into and out of gates in either
+positive of negative correlation with them.
+\begin{definition}
+  \label{definition-gc}
+  We write $\gc{\prog}{\{ \ttt{k = }\be_{k}; \ttt{p = }\be_{p} \}}{g}$ iff one of the
+  following hold:
+  \begin{enumerate}[\hspace{5mm}i.]
+  \item $\vc{\progtt(\prog,\be_k)}{\itv}{\ttt{flip[2,gate:}g\ttt{.k]}}\quad \text{and} \quad
+    \vc{\progtt(\prog,\be_p)}{\itv}{\ttt{flip[2,gate:}g\ttt{.p]}}$
+  \item $\vc{\progtt(\prog,\enot\ \be_k)}{\itv}{\ttt{flip[2,gate:}g\ttt{.k]}}\quad \text{and}\quad
+    \vc{\progtt(\prog,\enot\ \be_p)}{\itv}{\ttt{flip[2,gate:}g\ttt{.p]}}$
+  \end{enumerate}
+\end{definition}
+Thus, if a value is in correlation with a wire label, it is either
+correlated with the wire label or with the inverse of both the key and
+pointer bits. For subsequent discussion we make the following addition
+to the codebase:
+$$
+\ttt{invert(\{ k = bk; p = bp \}) \{ \{ k = not bk; p = not bp \} \}}
+$$
+Gate certification establishes separation of input labels from the
+gate table and output value, and correlation of the output with
+the output wire label under any input wire value combination.
+\begin{definition}[YGC Gate Certification]
+  \label{definition-ygcgate-certification}
+  Let $\mathit{gate}$ be a gate function.
+  For $1 \le i \le 4$  let
+  $
+  \config{\varnothing}{e_i}\redxs\config{\prog}{v_i}
+  $
+  where:
+  \begin{enumerate}[\hspace{5mm}$e_1 \defeq$]
+    \item $\mathit{gate}\ttt{(c,a,b,owl(a),owl(b));}$
+    \item $\mathit{gate}\ttt{(c,a,b,owl(a),invert(owl(b)));}$
+    \item $\mathit{gate}\ttt{(c,a,b,invert(owl(a)),owl(b));}$
+    \item $\mathit{gate}\ttt{(c,a,b,invert(owl(a)),invert(owl(b)))}$
+  \end{enumerate}
+  Define:
+  $${\small X \defeq \ttt{\{ flip[2,gate:a.k], flip[2,gate:a.p], flip[2,gate:b.k], flip[2,gate:b.p] \}}}$$
+  Then $\mathit{gate}$ is \emph{certified} iff the following conditions hold:
+  \begin{enumerate}[\hspace{5mm}i.]
+  \item $\sep{\progtt(\prog)}{X}{\vdefs(\prog)}$
+  \item For all $1 \le i \le 4$, $\gc{\prog}{v_i}{\ttt{c}}$
+  \end{enumerate}
+\end{definition}
+Encode certification establishes correlation of the encoded input values with
+the secret wire labels, and separation of client 2's secret from client 1's views. 
+\begin{definition}[YGC Encode Certification]
+  \label{definition-ygcencode-certification}
+  Let $\mathit{encode}$ be an encoding function, and let:
+  $$
+  \config{\varnothing}{\ttt{encode(s1,s2)}} \redxs
+  \config{\prog}{\ttt{\{wv1=}v_1\ttt{;wv2=}v_2\ttt{\}}}
+  $$
+  Then $\mathit{encode}$ is \emph{certified} iff each of the following conditions hold:
+  \begin{enumerate}[\hspace{5mm}i.]
+  \item $\gc{\prog}{v_1}{\ttt{s1}} \quad \text{and} \quad \gc{\prog}{v_2}{\ttt{s2}}$
+  \item $\sep{\progd(\prog)}{\{\ttt{s[1,s1]\}}}{\vdefs(\prog)_{\{2\}}}$
+  \item $\sep{\progd(\prog)}{\{\ttt{s[2,s2]\}}}{\vdefs(\prog)_{\{1\}}}$    
+  \item All $x \in \vars(\prog)$ are distinguished by $\ttt{in1}$ or $\ttt{in2}$. 
+  \end{enumerate}
+\end{definition}
+Decode certification establishes separation of the input wire
+label from the decoding tables under any input value condition. 
+\begin{definition}[YGC Decode Certification]
+  \label{definition-ygdecode-certification}
+  Let $\mathit{decode}$ be a decoding function, and let:
+  $$
+  \config{\varnothing}{\mathit{decode}(\ttt{c}, \ttt{owl(c)})}
+  \redxs \config{\prog}{\be}
+  %\quad \text{and} \quad
+  %\config{\varnothing}{\mathit{decode}(\ttt{c}, \ttt{invert(owl(c))})}
+  %\redxs \config{\prog}{\be_0}
+  $$
+  Then $\mathit{decode}$ is certified iff:
+  $$
+  \sep{\progtt(\prog)}{\ttt{\{flip[2,gate:c.k],flip[2,gate:c.p]\}}}{\vdefs(\prog)}
+  $$
+  %Then $\mathit{decode}$ is certified iff both of the following conditions hold:
+  %\begin{enumerate}[\hspace{5mm}i.]
+  %\item $\progd(\prog_1,\be_1)(\{ \itv \mapsto 1 \}) = 1$ and
+  %  $\progd(\prog_0,\be_0)(\{ \itv \mapsto 0 \}) = 1$
+  %\item $\sep{\progtt(\prog_i)}{\ttt{\{flip[2,gate:c.k],flip[2,gate:c.p]\}}}{\vdefs(\prog_i)}$
+  %\end{enumerate}
+\end{definition}
+We have certified each of the YGC components detailed above. 
+\begin{lemma}
+  \label{lemma-ygc-certification}
+  Each of $\ttt{decode}$, $\ttt{encode}$, and $\ttt{andgg}$ are certified.
+\end{lemma}
+
+\subsubsection{Gate Invariants and Circuit Security}
+\label{section-composition-metatheory}
+
+To demonstrate correctness of certification, we need to show that
+isolated component certificates preserve relevant properties
+when integrated into larger programs. 
+%First, we show that if a flip $f$ is independent from the views
+%in a protocol (as is the case with input labels used to
+%encode a garbled table), then separation from views that
+%use $f$ (e.g., in an output label) is preserved. 
+%\begin{lemma}[Wire Framing]
+%  \label{lemma-wire-framing}
+%  If $\sep{\progtt(\prog)}{\{f\}}{\vdefs(\prog)}$ for flip $f$ and
+%  $\vars(\prog) \cap \vars(\prog') = \{ f \}$ then
+%  $\sep{\progtt(\prog';\prog)}{\vdefs(\prog')}{\vdefs(\prog)}$.
+%\end{lemma}
+%Next,
+First we establish another substitution result, showing that
+correlations (e.g., between output wire values and labels) are
+preserved by substitution of correlated values for flips (e.g.,
+between input wire values and labels).
+\begin{lemma}[Substitution$\sim$]
+  \label{lemma-substitution-sim}
+  If $\vc{\progtt(\prog_2,\be)}{\itv}{f}$ and $\vc{\progtt(\prog_1,\be')}{\itv}{f'}$
+  and $\vars(\be') \cap (\vars(\be) - \{ f \}) = \varnothing$
+  then $\vc{\progtt(\prog_1;\prog_2,\be[\be'/f'])}{\itv}{f}$.
+\end{lemma}
+Now we can show how to splice gates into arbitrary circuits in an
+invariant-preserving manner, assuming that outputs are not wired to
+multiple inputs \cite{tate2003garbled,nieminen2023breaking}.
+\begin{lemma}[YGC Gate Invariant]
+  \label{lemma-ygc-preservation}
+Given well-typed $\prog_1;E[\mathit{gate}(g,g_1,g_2,v_1,v_2)]$ and certified $\mathit{gate}$ where:
+$$
+\config{\prog_1}{E[\mathit{gate}(g,g_1,g_2,v_1,v_2)]} \redxs \config{\prog_1;\prog_2}{E[v]}
+$$
+If the following preconditions hold where $\iov(\prog_1) = S \cup V$:
+\begin{enumerate}
+\item $\{ g_1 \} \cap \{ g_2 \} \cap \wired(\prog_1) = \varnothing$ and $g \not\in \prog_1$
+\item $\gc{\prog_1}{v_1}{g_1}$ and $\gc{\prog_1}{v_2}{g_2}$
+\item $\sep{\progd(\prog_1)}{S_{\{1\}}}{V_{\{2\}}}$
+\item $\sep{\progd(\prog_1)}{S_{\{2\}}}{V_{\{1\}}}$
+\end{enumerate}
+then we have as postconditions where $\iov(\prog_1) = S \cup V'$:
+\begin{enumerate}
+\item $\gc{\prog_1;\prog_2}{v}{g_2}$
+\item $\sep{\progd(\prog_1,\prog_2)}{S_{\{1\}}}{V'_{\{2\}}}$
+\item $\sep{\progd(\prog_1,\prog_2)}{S_{\{2\}}}{V'_{\{1\}}}$
+\end{enumerate}
+\end{lemma}
+\begin{proof}
+  Given preconditions we have $v_1 =  \{ \ttt{k = }\be^1_{k}; \ttt{p = }\be^1_{p} \}$ and
+  $v_2 = \{ \ttt{k = }\be^2_{k}; \ttt{p = }\be^2_{p} \}$ for some
+  $\be^1_{k}$,$\be^2_{k}$,$\be^1_{p}$,$\be^2_{p}$ with correlations as per Definition
+  \ref{definition-gc}.
+  %\begin{eqnarray*}
+  %  v_1 &=& \{ \ttt{k = }\be^1_{k}; \ttt{p = }\be^1_{p} \} \\
+  %  v_2 &=& \{ \ttt{k = }\be^2_{k}; \ttt{p = }\be^2_{p} \}
+  %\end{eqnarray*}
+  Let $\prog$ be as defined in Definition \ref{definition-ygcgate-certification}.
+  We observe:
+  \begin{eqnarray*}
+    &\prog_2 = \\
+    &{\small \prog[\be^1_{k}/\ttt{flip[2,gate:a.k]}][\be^1_{p}/\ttt{flip[2,gate:a.p]}][\be^2_{k}/\ttt{flip[2,gate:b.k]}][\be^1_{p}/\ttt{flip[2,gate:b.p]}]}
+  \end{eqnarray*}
+  Given that $g_1$ and $g_2$ are distinct and not wired in $\prog_1$
+  we are assured that $\ttt{owl}(g_1)$ and $\ttt{owl}(g_1)$ are in
+  independent uniform distributions, and given that $g \not\in
+  \prog_1$ we are assured that $\ttt{owl}(g)$ contains entirely fresh
+  flips. Thus by condition (i) of Definition \ref{definition-ygcgate-certification} and
+  Lemma \ref{lemma-substitution} we have:
+  $$
+  \sep{\prog_1;\prog_2}{\vars(\be^1_{k},\be^2_{k},\be^1_{p},\be^2_{p})}{\vdefs(\prog_2)}
+  $$
+  Thus by Lemmas \ref{lemma-noninterference} and \ref{lemma-separation} we establish
+  postconditions (ii) and (iii).
+
+  Also since $v_1$ and $v_2$ are correlated either positively or negatively with
+  $\ttt{owl}(g_1)$ and $\ttt{owl}(g_2)$ respectively by precondition (ii),
+  by Definition \ref{definition-gc}, precondition (i), and Lemma \ref{lemma-substitution-sim}
+  we establish postcondition (i), since Definition \ref{definition-ygcgate-certification}
+  requires gate output correlation with $\ttt{owl}(g)$ given any input valence conditions.
+\end{proof}
+
+\begin{lemma}[YGC Encode Invariant]
+  \label{lemma-ygc-encode}
+Given well-typed $\prog_1;E[\mathit{encode}(s_1,s_2)]$ and certified $\mathit{encode}$ where 
+$$
+\config{\prog_1}{E[\ttt{encode}(s_1,s_2)]} \redxs \config{\prog_1;\prog_2}{E[\ttt{\{wv1=}v_1\ttt{;wv2=}v_2\ttt{\}}]}
+$$
+If the following preconditions hold where $\iov(\prog_1) = S \cup V$:
+\begin{enumerate}
+\item $\sep{\progd(\prog_1)}{S_{\{1\}}}{V_{\{2\}}}$
+\item $\sep{\progd(\prog_1)}{S_{\{2\}}}{V_{\{1\}}}$
+\end{enumerate}
+then we have as postconditions where $\iov(\prog_1) = S' \cup V'$:
+\begin{enumerate}
+\item $\gc{\prog_1;\prog_2}{v_1}{s_1}$ and $\gc{\prog_1}{v_2}{s_2}$
+\item $\sep{\progd(\prog_1;\prog_2)}{S'_{\{1\}}}{V'_{\{2\}}}$
+\item $\sep{\progd(\prog_1;\prog_2)}{S'_{\{2\}}}{V'_{\{1\}}}$
+\end{enumerate}
+\end{lemma}
+\begin{proof}
+  By Definition \ref{definition-ygcencode-certification} we have:
+  \begin{mathpar}
+    \sep{\progtt(\prog_2)}{\ttt{\{s[1,s1]\}}}{\vdefs(\prog_2)_{\{2\}}}
+
+    \sep{\progtt(\prog_2)}{\ttt{\{s[2,s2]\}}}{\vdefs(\prog_2)_{\{1\}}}
+  \end{mathpar}
+  Since we assume that secrets are in uniform and independent marginal
+  distributions a priori, and $\vars(\prog_1) \cap \vars(\prog_2) =
+  \varnothing$ by condition (iv) of Definition
+  \ref{definition-ygcencode-certification} and assumptions of
+  well-typedness, conditions (ii-iii) follow by Lemmas
+  \ref{lemma-noninterference} and \ref{lemma-separation}. Also by
+  Definition \ref{definition-ygcencode-certification} we have
+  $\gc{\prog_2}{v_1}{s_1}$ and $\gc{\prog_2}{v_1}{s_1}$, so also
+  $\gc{\prog_1;\prog_2}{v_1}{s_1}$ and
+  $\gc{\prog_1;\prog_2}{v_1}{s_1}$ by condition (iv) of Definition
+  \ref{definition-ygcencode-certification} and Lemmas
+  \ref{lemma-noninterference} and \ref{lemma-separation}.
+\end{proof}
+
+Given the above, we can now establish that any circuit built with
+certified components is passive secure.  Note that we assume a normal
+form of programs where the last instruction is a decoding and public
+reveal of the output, and we require that no gate is wired more than once.
+\cnote{There is another subtlety here, that gates are wired correctly
+  in the sense that if the garbler generates a table with input wire label
+  $g_1$ then only $g_1$ is wired to that input. I'll modify the
+  typing to enforce this.}
+\begin{theorem}
+  \label{theorem-ygc-NIMO}
+  If $\eassign{\outv}{\ttt{decode}(g,e)}$ is a well-typed YGC circuit
+  definition using certified components where no gate output is wired
+  more than once, and
+  $\config{\varnothing}{\eassign{\outv}{\ttt{decode}(g,e)}} \redxs
+  \config{\prog}{\varnothing}$, then $\NIMO(\prog)$.
+\end{theorem}
+
+\begin{proof}
+  By assumptions of well-typedness we
+  have for some $\be_k$, $\be_p$, and $\be$:
+  $$
+  \config{\varnothing}{\eassign{\outv}{\ttt{decode}(g,e)}} \redxs
+  \config{\prog_1}{\eassign{\outv}{\ttt{decode}(g,\{ \ttt{k = }\be_k;  \ttt{p = }\be_p\})}}
+  \redxs \config{\prog_1;\prog_2}{\eassign{\outv}{\be}}
+  $$
+  Let $\iov(\prog_1) = S \cup V$.
+  By Lemmas \ref{lemma-ygc-encode} and \ref{lemma-ygc-preservation} we
+  have $\sep{\progd(\prog_1)}{S_{\{1\}}}{V_{\{2\}}}$ and
+  $\sep{\progd(\prog_1)}{S_{\{2\}}}{V_{\{1\}}}$.
+  Let $\prog$ be as defined in Definition \ref{definition-ygdecode-certification}.
+  We observe:
+  \begin{eqnarray*}
+    & \prog_2 = \\ 
+    & \prog[\be_k/\ttt{\{flip[2,gate:c.k]}][\be_p/\ttt{flip[2,gate:c.p]}]
+  \end{eqnarray*}
+  so also by  Definition \ref{definition-ygdecode-certification} and
+  Lemma \ref{lemma-substitution} we have:
+  $$
+  \sep{\progtt(\prog_1;\prog_2)}{\vars(\be_k,\be_p)}{\vdefs(\prog_2)}
+  $$
+  so, letting $\iov(\prog_1;\prog_2) = S \cup V'$ by
+  Lemmas \ref{lemma-noninterference} \ref{lemma-separation} we have
+  $\sep{\progd(\prog_1;\prog_2)}{S_{\{1\}}}{V'_{\{2\}}}$ and
+  $\sep{\progd(\prog_1;\prog_2)}{S_{\{2\}}}{V'_{\{1\}}}$.
+  Thus, letting $\prog' \defeq (\prog_1;\prog_2;\eassign{\outv}{\be})$, for all $\beta$:
+  $$\sep{\condd{\progd(\prog')}{S_{\{1\}} \cup V_{\{2\}}}{\{ \outv \mapsto \beta \}}}{S_{\{1\}}}{V_{\{2\}}}
+  \quad \text{and} \quad
+    \sep{\condd{\progd(\prog')}{S_{\{2\}} \cup V_{\{1\}}}{\{ \outv \mapsto \beta \}}}{S_{\{2\}}}{V_{\{1\}}}$$
+  thus for all $\store \in \mems(V_2 \cup \{ \outv \})$:
+  $$\condd{\progd(\prog')}{S_{\{1\}}}{\store}
+  = \condd{\progd(\prog')}{S_{\{1\}}}{\store_{\{\outv\}}}$$
+  and for all $\store \in \mems(V_1 \cup \{ \outv \})$:
+  $$\condd{\progd(\prog')}{S_{\{2\}}}{\store}
+  = \condd{\progd(\prog')}{S_{\{2\}}}{\store_{\{\outv\}}}$$
+  establishing the result by Lemma \ref{lemma-nimo} and Definition \ref{definition-NIMO}.
+\end{proof}
+
+\begin{fpfig}[t]{YGC copy gate definitions.}{fig-ygc-copy}
+  {\footnotesize
+    \begin{verbatimtab}
+      sharetab2(gid, tid, k, p, b)
+      {
+        let r1 = k xor b in
+        let r0 = (not k) xor (not b) in
+        v[1,gate: || gid || tid || 1] := select(p,r1,r0);
+        v[1,gate: || gid || tid || 2] := select(not p,r1,r0);
+      }
+
+      copygate(ca,cb,g)
+      {
+        let wl = owl(g) in
+        let owl1 = owl(ca) in
+        let owl2 = owl(cb) in
+        sharetab2(ca,tt,wl.k,wl.p,owl1.k); sharetab2(ca,pt,wl.k,wl.p,owl1.p);
+        sharetab2(cb,tt,wl.k,wl.p,owl2.k); sharetab2(cb,pt,wl.k,wl.p,owl2.p)
+      }
+
+      evalcopy(ca,cb,wv)
+      {
+        let wv1k = wv.k xor select(wv.p,v[1,gate: || ca || tt1], v[1,gate: || ca || tt2]) in
+        let wv1p = wv.k xor select(wv.p,v[1,gate: || ca || pt1], v[1,gate: || ca || pt1]) in
+        let wv2k = wv.k xor select(wv.p,v[1,gate: || cb || tt1], v[1,gate: || cb || tt2]) in
+        let wv2p = wv.k xor select(wv.p,v[1,gate: || cb || pt1], v[1,gate: || cb || pt1]) in
+        { wv1 = { k = wv1k; p = wv1p }; wv2 = { k = wv2k; p = wv2p } }  
+      }
+
+      copy(ca,cb,g,wv) { copygate(ca,cb,g); evalcopy(ca,cb,wv) } 
+    \end{verbatimtab}
+  }
+\end{fpfig}
+
+\subsection{Building and Extending Automatically Secure Circuits}
+\label{section-composition-copy}
+
+Theorems \ref{theorem-gmw-NIMO} \ref{theorem-ygc-NIMO} means that any
+well-typed program using certified components will generate a circuit
+that is passive secure. To add an $\ttt{or}$ gate to our YGC library,
+for example, we can use $\ttt{andgg}$ as a template, in fact simple
+modification of $\ttt{andtable}$ is all that would be needed. Thus,
+this library embodies a Fairplay-like language where well-typedness of
+metaprograms guarantees safety and security of generated protocols.
+
+The requirements that output wire labels are not used more than once
+in YGC reflects a known result, that naive copy gate output is unsound
+in YGC- i.e., passive security fails. In fact, we are unable to
+automatically verify $\gNIMO$ for $\ttt{andgg}$ without the assumption
+that input wire lables are in independent uniform distributions. We
+can capture the necessary fix as in Figure \ref{fig-ygc-copy}. The
+garbler creates a copy gate with two new output wire labels in
+independent uniform distributions (in \ttt{copygate}), and permutes
+them using the input wire label pointer after encrypting with the
+input wire label keys. The evaluator then recovers the encrypted
+output wire values during evaluation (in \ttt{evalcopy}). Given an
+$\ttt{orgg}$ gate defined as outlined abotve, we can then use secure
+$\ttt{copy}$ as follows:
+\begin{verbatimtab}
+  let s = encode(s1,s2) in
+  let s1c = copy(c1s1,c2s1,s1,s.wv1) in
+  let s2c = copy(c1s2,c2s2,s2,s.wv2) in
+  let wv1 = andgg(1,c1s1,c1s2,s1c.wv1,s2c.wv1) in
+  let wv2 = orgg(2,c2s1,c2s2,s1c.wv2,s2c.wv2) in
+  decode(3,andgg(3,1,2,wv1,wv2))
+\end{verbatimtab}
+These definitions of $\ttt{copy}$ and $\ttt{orgg}$ have been
+automatically verified as in Section \ref{section-pre-post} for
+$\ttt{andgg}$. In general, any new gate form- such as optimized gates
+\cite{XXX}- can be verified automatically in a similar way.
+
+While the copy gate supports secure programming patterns, security
+still relies on programmer discipline, i.e., not wiring a gate more
+than once. However, we can keep track of wired gates in type effects,
+similar to the manner in which we keep track of defined views- type
+dependence in our system is able to accurately precisely track
+gate identifiers. This is also reminiscent of the use of linearity
+in other related systems for enforcing obliviousness \cite{darais2019language}.

### 2023-07-07 popl24/scratch.tex
checkpoint

--- 2023-07-06 popl24/scratch.tex
@@
 \begin{definition}
+  \label{definition-gc}
   We write $\gc{\prog}{\{ \ttt{k = }\be_{k}; \ttt{p = }\be_{p} \}}{g}$ iff one of the
@@
 the output wire label under any input wire value combination.
-\begin{definition}[Gate Certification]
-  \label{definition-gate-certification}
-Let $\mathit{gate}$ be a gate function. For $1 \le i \le 4$  let:
-  $$
-  \config{\varnothing}{\ttt{v[0,out] := } e_i}\redxs\config{\prog_i}{\ttt{()}}
-  $$
-  where
+\begin{definition}[YGC Gate Certification]
+  \label{definition-ygcgate-certification}
+  Let $\mathit{gate}$ be a gate function.
+  For $1 \le i \le 4$  let
+  $
+  \config{\varnothing}{e_i}\redxs\config{\prog}{v_i}
+  $
+  where:
   \begin{enumerate}[\hspace{5mm}$e_1 \defeq$]
@@
   $${\small X \defeq \ttt{\{ flip[2,gate:a.k], flip[2,gate:a.p], flip[2,gate:b.k], flip[2,gate:b.p] \}}}$$
-  Then $\mathit{gate}$ is \emph{certified} iff for all $1 \le i \le 4$:
-  $$
-  \sep{\progtt(\prog_i)}{X}{\vdefs(\prog_i)} 
-  \quad \text{and} \quad \gc{\progd(\prog_i)}{\vx{0}{out}}{\ttt{owl(c)}}
-  $$
+  Then $\mathit{gate}$ is \emph{certified} iff the following conditions hold:
+  \begin{enumerate}[\hspace{5mm}i.]
+  \item $\sep{\progtt(\prog)}{X}{\vdefs(\prog)}$
+  \item For all $1 \le i \le 4$, $\gc{\prog}{v_i}{\ttt{c}}$
+  \end{enumerate}
 \end{definition}
@@
 the secret wire labels, and separation of client 2's secret from client 1's views. 
-\begin{definition}[Encode Certification]
+\begin{definition}[YGC Encode Certification]
+  \label{definition-ygcencode-certification}
   Let $\mathit{encode}$ be an encoding function, and let:
@@
   \begin{enumerate}[\hspace{5mm}i.]
-  \item $\gc{\progd(\prog,v_1)}{\itv}{\ttt{owl(s1)}} \quad \text{and} \quad
-    \gc{\progd(\prog,v_2)}{\itv}{\ttt{owl(s2)}}$
-  \item $\sep{\progd(\prog,v_2)}{\{\ttt{s[2,s2]\}}}{(\vdefs(\prog) \cup \itv)}$
+  \item $\gc{\prog}{v_1}{\ttt{s1}} \quad \text{and} \quad \gc{\prog}{v_2}{\ttt{s2}}$
+  \item $\sep{\progd(\prog)}{\{\ttt{s[1,s1]\}}}{\vdefs(\prog)_{\{2\}}}$
+  \item $\sep{\progd(\prog)}{\{\ttt{s[2,s2]\}}}{\vdefs(\prog)_{\{1\}}}$    
+  \item All $x \in \vars(\prog)$ are distinguished by $\ttt{in1}$ or $\ttt{in2}$. 
   \end{enumerate}
@@
 label from the decoding tables under any input value condition. 
-\begin{definition}[Decode Certification]
+\begin{definition}[YGC Decode Certification]
+  \label{definition-ygdecode-certification}
   Let $\mathit{decode}$ be a decoding function, and let:
@@
   \config{\varnothing}{\mathit{decode}(\ttt{c}, \ttt{owl(c)})}
-  \redxs \config{\prog_1}{\be_1}
-  \quad \text{and} \quad
-  \config{\varnothing}{\mathit{decode}(\ttt{c}, \ttt{invert(owl(c))})}
-  \redxs \config{\prog_0}{\be_0}
+  \redxs \config{\prog}{\be}
+  %\quad \text{and} \quad
+  %\config{\varnothing}{\mathit{decode}(\ttt{c}, \ttt{invert(owl(c))})}
+  %\redxs \config{\prog}{\be_0}
   $$
@@
   $$
-  \sep{\progtt(\prog_i)}{\ttt{\{flip[2,gate:c.k],flip[2,gate:c.p]\}}}{\vdefs(\prog_i)}
+  \sep{\progtt(\prog)}{\ttt{\{flip[2,gate:c.k],flip[2,gate:c.p]\}}}{\vdefs(\prog)}
   $$
@@
 \begin{lemma}
-  \label{lemma-certification}
+  \label{lemma-ygc-certification}
   Each of $\ttt{decode}$, $\ttt{encode}$, and $\ttt{andgg}$ are certified.
@@
 
-\subsubsection{Correctness of Certified Components}
+\subsubsection{Gate Invariants and Circuit Security}
 \label{section-composition-metatheory}
@@
 isolated component certificates preserve relevant properties
-when integrated into larger programs. This resonates with
-\emph{framing} as developed in \cite{barthe2019probabilistic}.
-In the following we will use $f$ to range over flips.
-First, we show that if a flip $f$ is independent from the views
-in a protocol (as is the case with input labels used to
-encode a garbled table), then separation from views that
-use $f$ (e.g., in an output label) is preserved. 
-\begin{lemma}[Wire Framing]
-  \label{lemma-wire-framing}
-  If $\sep{\progtt(\prog)}{\{f\}}{\vdefs(\prog)}$ for flip $f$ and
-  $\vars(\prog) \cap \vars(\prog') = \{ f \}$ then
-  $\sep{\progtt(\prog';\prog)}{\vdefs(\prog')}{\vdefs(\prog)}$.
+when integrated into larger programs. 
+%First, we show that if a flip $f$ is independent from the views
+%in a protocol (as is the case with input labels used to
+%encode a garbled table), then separation from views that
+%use $f$ (e.g., in an output label) is preserved. 
+%\begin{lemma}[Wire Framing]
+%  \label{lemma-wire-framing}
+%  If $\sep{\progtt(\prog)}{\{f\}}{\vdefs(\prog)}$ for flip $f$ and
+%  $\vars(\prog) \cap \vars(\prog') = \{ f \}$ then
+%  $\sep{\progtt(\prog';\prog)}{\vdefs(\prog')}{\vdefs(\prog)}$.
+%\end{lemma}
+%Next,
+First we establish another substitution result, showing that
+correlations (e.g., between output wire values and labels) are
+preserved by substitution of correlated values for flips (e.g.,
+between input wire values and labels).
+\begin{lemma}[Substitution$\sim$]
+  \label{lemma-substitution-sim}
+  If $\vc{\progtt(\prog_2,\be)}{\itv}{f}$ and $\vc{\progtt(\prog_1,\be')}{\itv}{f'}$
+  and $\vars(\be') \cap (\vars(\be) - \{ f \}) = \varnothing$
+  then $\vc{\progtt(\prog_1;\prog_2,\be[\be'/f'])}{\itv}{f}$.
 \end{lemma}
-Next, we show that output correlations (e.g., between output wire values and
-labels) are preserved by substitution of correlated values for flips (e.g.,
-between input wire values and labels).
-\begin{lemma}[Substitution]
-  \label{lemma-substitution}
-  If $\vc{\progtt(\prog,\be)}{\itv}{f}$ and $\vc{\progtt(\prog',\be')}{\itv}{f'}$
-  and $\sep{\progtt(\prog';\prog)}{\vdefs(\prog')}{\vdefs(\prog)}$
-  then $\vc{\progtt(\prog';\prog,\be[\be'/f'])}{\itv}{f}$.
-\end{lemma}
-Our wire framing and substitution Lemmas then allow us to generalize
-certification of gates and encoding to arbitrary circuit contexts,
-assuming that outputs are not wired to multiple inputs \cite{tate2003garbled,nieminen2023breaking}.
-\begin{lemma}[Gate Composition]
-  \label{lemma-compose}
+Now we can show how to splice gates into arbitrary circuits in an
+invariant-preserving manner, assuming that outputs are not wired to
+multiple inputs \cite{tate2003garbled,nieminen2023breaking}.
+\begin{lemma}[YGC Gate Invariant]
+  \label{lemma-ygc-preservation}
 Given well-typed $\prog_1;E[\mathit{gate}(g,g_1,g_2,v_1,v_2)]$ and certified $\mathit{gate}$ where:
@@
 $$
-If the following preconditions hold:
+If the following preconditions hold where $\iov(\prog_1) = S \cup V$:
 \begin{enumerate}
 \item $\{ g_1 \} \cap \{ g_2 \} \cap \wired(\prog_1) = \varnothing$ and $g \not\in \prog_1$
-\item $\gc{\progd(\prog_1)}{v_1}{\owl(g_1)}$ and $\gc{\progd(\prog_1)}{v_2}{\owl(g_2)}$
+\item $\gc{\prog_1}{v_1}{g_1}$ and $\gc{\prog_1}{v_2}{g_2}$
+\item $\sep{\progd(\prog_1)}{S_{\{1\}}}{V_{\{2\}}}$
+\item $\sep{\progd(\prog_1)}{S_{\{2\}}}{V_{\{1\}}}$
 \end{enumerate}
-then we have as postconditions:
+then we have as postconditions where $\iov(\prog_1) = S \cup V'$:
 \begin{enumerate}
-  \item $\gc{\progd(\prog_1;\prog_2)}{v}{\owl(g_2)}$
-  \item $\sep{\progd(\prog_1;\prog_2,v)}{\vdefs(\prog_1)}{\vdefs(\prog_2) \cup \{\itv\}}$
+\item $\gc{\prog_1;\prog_2}{v}{g_2}$
+\item $\sep{\progd(\prog_1,\prog_2)}{S_{\{1\}}}{V'_{\{2\}}}$
+\item $\sep{\progd(\prog_1,\prog_2)}{S_{\{2\}}}{V'_{\{1\}}}$
 \end{enumerate}
@@
 \begin{proof}
+  Given preconditions we have $v_1 =  \{ \ttt{k = }\be^1_{k}; \ttt{p = }\be^1_{p} \}$ and
+  $v_2 = \{ \ttt{k = }\be^2_{k}; \ttt{p = }\be^2_{p} \}$ for some
+  $\be^1_{k}$,$\be^2_{k}$,$\be^1_{p}$,$\be^2_{p}$ with correlations as per Definition
+  \ref{definition-gc}.
+  %\begin{eqnarray*}
+  %  v_1 &=& \{ \ttt{k = }\be^1_{k}; \ttt{p = }\be^1_{p} \} \\
+  %  v_2 &=& \{ \ttt{k = }\be^2_{k}; \ttt{p = }\be^2_{p} \}
+  %\end{eqnarray*}
+  Let $\prog$ be as defined in Definition \ref{definition-ygcgate-certification}.
+  We observe:
+  \begin{eqnarray*}
+    &\prog_2 = \\
+    &{\small \prog[\be^1_{k}/\ttt{flip[2,gate:a.k]}][\be^1_{p}/\ttt{flip[2,gate:a.p]}][\be^2_{k}/\ttt{flip[2,gate:b.k]}][\be^1_{p}/\ttt{flip[2,gate:b.p]}]}
+  \end{eqnarray*}
   Given that $g_1$ and $g_2$ are distinct and not wired in $\prog_1$
@@
   \prog_1$ we are assured that $\ttt{owl}(g)$ contains entirely fresh
-  flips. Thus, precondition 2, Lemma \ref{lemma-wire-framing}, and
-  Definition \ref{definition-gate-certification} imply postcondition
-  2.  Furthermore, since gate certification guarantees that
-  correlation with $\ttt{owl}(g)$ is guaranteed under all possible gate
-  input wire values, precondition 2 and Lemma \ref{lemma-substitution}
-  guarantee postcondition 2.
+  flips. Thus by condition (i) of Definition \ref{definition-ygcgate-certification} and
+  Lemma \ref{lemma-substitution} we have:
+  $$
+  \sep{\prog_1;\prog_2}{\vars(\be^1_{k},\be^2_{k},\be^1_{p},\be^2_{p})}{\vdefs(\prog_2)}
+  $$
+  Thus by Lemmas \ref{lemma-noninterference} and \ref{lemma-separation} we establish
+  postconditions (ii) and (iii).
+
+  Also since $v_1$ and $v_2$ are correlated either positively or negatively with
+  $\ttt{owl}(g_1)$ and $\ttt{owl}(g_2)$ respectively by precondition (ii),
+  by Definition \ref{definition-gc}, precondition (i), and Lemma \ref{lemma-substitution-sim}
+  we establish postcondition (i), since Definition \ref{definition-ygcgate-certification}
+  requires gate output correlation with $\ttt{owl}(g)$ given any input valence conditions.
 \end{proof}
 
-\begin{lemma}[Decode Composition]
-  \label{lemma-decode}
-Given well-typed $\prog_1;E[\mathit{decode}(g,v)]$ and certified $\mathit{decode}$ where 
+\begin{lemma}[YGC Encode Invariant]
+  \label{lemma-ygc-encode}
+Given well-typed $\prog_1;E[\mathit{encode}(s_1,s_2)]$ and certified $\mathit{encode}$ where 
 $$
-\config{\prog_1}{E[\ttt{decode}(g,v)]} \redxs \config{\prog_1;\prog_2}{E[\be]}
+\config{\prog_1}{E[\ttt{encode}(s_1,s_2)]} \redxs \config{\prog_1;\prog_2}{E[\ttt{\{wv1=}v_1\ttt{;wv2=}v_2\ttt{\}}]}
 $$
-If the following preconditions hold:
+If the following preconditions hold where $\iov(\prog_1) = S \cup V$:
 \begin{enumerate}
-\item $g \not\in \cap \wired(\prog_1)$
-\item $\gc{\progd(\prog_1)}{v}{\owl(g)}$
+\item $\sep{\progd(\prog_1)}{S_{\{1\}}}{V_{\{2\}}}$
+\item $\sep{\progd(\prog_1)}{S_{\{2\}}}{V_{\{1\}}}$
 \end{enumerate}
-then we have as postcondition:
+then we have as postconditions where $\iov(\prog_1) = S' \cup V'$:
 \begin{enumerate}
-  \item $\sep{\progd(\prog_1;\prog_2)}{\vdefs(\prog_1)}{\vdefs(\prog_2)}$
+\item $\gc{\prog_1;\prog_2}{v_1}{s_1}$ and $\gc{\prog_1}{v_2}{s_2}$
+\item $\sep{\progd(\prog_1;\prog_2)}{S'_{\{1\}}}{V'_{\{2\}}}$
+\item $\sep{\progd(\prog_1;\prog_2)}{S'_{\{2\}}}{V'_{\{1\}}}$
 \end{enumerate}
 \end{lemma}
-Given the above, we can establish that any circuit built with certified
-components is passive secure. 
+\begin{proof}
+  By Definition \ref{definition-ygcencode-certification} we have:
+  \begin{mathpar}
+    \sep{\progtt(\prog_2)}{\ttt{\{s[1,s1]\}}}{\vdefs(\prog_2)_{\{2\}}}
+
+    \sep{\progtt(\prog_2)}{\ttt{\{s[2,s2]\}}}{\vdefs(\prog_2)_{\{1\}}}
+  \end{mathpar}
+  Since we assume that secrets are in uniform and independent marginal
+  distributions a priori, and $\vars(\prog_1) \cap \vars(\prog_2) =
+  \varnothing$ by condition (iv) of Definition
+  \ref{definition-ygcencode-certification} and assumptions of
+  well-typedness, conditions (ii-iii) follow by Lemmas
+  \ref{lemma-noninterference} and \ref{lemma-separation}. Also by
+  Definition \ref{definition-ygcencode-certification} we have
+  $\gc{\prog_2}{v_1}{s_1}$ and $\gc{\prog_2}{v_1}{s_1}$, so also
+  $\gc{\prog_1;\prog_2}{v_1}{s_1}$ and
+  $\gc{\prog_1;\prog_2}{v_1}{s_1}$ by condition (iv) of Definition
+  \ref{definition-ygcencode-certification} and Lemmas
+  \ref{lemma-noninterference} and \ref{lemma-separation}.
+\end{proof}
+
+Given the above, we can now establish that any circuit built with
+certified components is passive secure.  Note that we assume a normal
+form of programs where the last instruction is a decoding and public
+reveal of the output, and we require that no gate is wired more than once.
+\cnote{There is another subtlety here, that gates are wired correctly
+  in the sense that if the garbler generates a table with input wire label
+  $g_1$ then only $g_1$ is wired to that input. I'll modify the
+  typing to enforce this.}
 \begin{theorem}
   \label{theorem-ygc-NIMO}
-  If $e$ is a well-typed YGC circuit definition using certified components
-  where no gate output is wired more than once, and $\config{\varnothing}{e}
-  \redxs \config{\prog}{\varnothing}$, then $\NIMO(\prog)$. 
+  If $\eassign{\outv}{\ttt{decode}(g,e)}$ is a well-typed YGC circuit
+  definition using certified components where no gate output is wired
+  more than once, and
+  $\config{\varnothing}{\eassign{\outv}{\ttt{decode}(g,e)}} \redxs
+  \config{\prog}{\varnothing}$, then $\NIMO(\prog)$.
 \end{theorem}
@@
 \begin{proof}
-  Let $\iov(\prog) = S \cup (V \cup \outv)$. Noting that any
-  encoding necessarily uses entirely fresh variables by Definition and
-  type safety, by applying Lemmas \ref{lemma-compose} and
-  \ref{lemma-decode} and the frame rule of
-  \cite{barthe2019probabilistic} we obtain:
+  By assumptions of well-typedness we
+  have for some $\be_k$, $\be_p$, and $\be$:
   $$
-  \sep{\progd(\prog)}{(S_{\{2\}}\cup\{\outv\})}{V}
+  \config{\varnothing}{\eassign{\outv}{\ttt{decode}(g,e)}} \redxs
+  \config{\prog_1}{\eassign{\outv}{\ttt{decode}(g,\{ \ttt{k = }\be_k;  \ttt{p = }\be_p\})}}
+  \redxs \config{\prog_1;\prog_2}{\eassign{\outv}{\be}}
   $$
-  Noting that the garbler (client 2) has no views, this means that
-  the garbler's secrets are independent of the evaluator's views. The
-  result follows by Definition \ref{definition-NIMO}.
+  Let $\iov(\prog_1) = S \cup V$.
+  By Lemmas \ref{lemma-ygc-encode} and \ref{lemma-ygc-preservation} we
+  have $\sep{\progd(\prog_1)}{S_{\{1\}}}{V_{\{2\}}}$ and
+  $\sep{\progd(\prog_1)}{S_{\{2\}}}{V_{\{1\}}}$.
+  Let $\prog$ be as defined in Definition \ref{definition-ygdecode-certification}.
+  We observe:
+  \begin{eqnarray*}
+    & \prog_2 = \\ 
+    & \prog[\be_k/\ttt{\{flip[2,gate:c.k]}][\be_p/\ttt{flip[2,gate:c.p]}]
+  \end{eqnarray*}
+  so also by  Definition \ref{definition-ygdecode-certification} and
+  Lemma \ref{lemma-substitution} we have:
+  $$
+  \sep{\progtt(\prog_1;\prog_2)}{\vars(\be_k,\be_p)}{\vdefs(\prog_2)}
+  $$
+  so, letting $\iov(\prog_1;\prog_2) = S \cup V'$ by
+  Lemmas \ref{lemma-noninterference} \ref{lemma-separation} we have
+  $\sep{\progd(\prog_1;\prog_2)}{S_{\{1\}}}{V'_{\{2\}}}$ and
+  $\sep{\progd(\prog_1;\prog_2)}{S_{\{2\}}}{V'_{\{1\}}}$.
+  Thus, letting $\prog' \defeq (\prog_1;\prog_2;\eassign{\outv}{\be})$, for all $\beta$:
+  $$\sep{\condd{\progd(\prog')}{S_{\{1\}} \cup V_{\{2\}}}{\{ \outv \mapsto \beta \}}}{S_{\{1\}}}{V_{\{2\}}}
+  \quad \text{and} \quad
+    \sep{\condd{\progd(\prog')}{S_{\{2\}} \cup V_{\{1\}}}{\{ \outv \mapsto \beta \}}}{S_{\{2\}}}{V_{\{1\}}}$$
+  thus for all $\store \in \mems(V_2 \cup \{ \outv \})$:
+  $$\condd{\progd(\prog')}{S_{\{1\}}}{\store}
+  = \condd{\progd(\prog')}{S_{\{1\}}}{\store_{\{\outv\}}}$$
+  and for all $\store \in \mems(V_1 \cup \{ \outv \})$:
+  $$\condd{\progd(\prog')}{S_{\{2\}}}{\store}
+  = \condd{\progd(\prog')}{S_{\{2\}}}{\store_{\{\outv\}}}$$
+  establishing the result by Lemma \ref{lemma-nimo} and Definition \ref{definition-NIMO}.
 \end{proof}
@@
 
-Theorem \ref{theorem-nimo} means that any well-typed program using
-$\ttt{encode}$, $\ttt{andgg}$, and $\ttt{decode}$ will generate a
-circuit that is passive secure, provided that gate outputs are not
-used more than once. To add an $\ttt{or}$ gate to our library, we can
-use $\ttt{andgg}$ as a template, in fact simple modification of
-$\ttt{andtable}$ is all that would be needed. Thus, this library
-embodies a Fairplay-like language where well-typedness of metaprograms
-guarantees safety and security of generated protocols.
+Theorems \ref{theorem-gmw-NIMO} \ref{theorem-ygc-NIMO} means that any
+well-typed program using certified components will generate a circuit
+that is passive secure. To add an $\ttt{or}$ gate to our YGC library,
+for example, we can use $\ttt{andgg}$ as a template, in fact simple
+modification of $\ttt{andtable}$ is all that would be needed. Thus,
+this library embodies a Fairplay-like language where well-typedness of
+metaprograms guarantees safety and security of generated protocols.
 
 The requirements that output wire labels are not used more than once
-reflects a known result, that naive copy gate output is unsound in
-YGC- i.e., passive security fails. In fact, we are unable to
+in YGC reflects a known result, that naive copy gate output is unsound
+in YGC- i.e., passive security fails. In fact, we are unable to
 automatically verify $\gNIMO$ for $\ttt{andgg}$ without the assumption
@@
 output wire values during evaluation (in \ttt{evalcopy}). Given an
-$\ttt{orgg}$ gate defined as outlined abotve, we can then use
-secure $\ttt{copy}$ as follows: 
+$\ttt{orgg}$ gate defined as outlined abotve, we can then use secure
+$\ttt{copy}$ as follows:
 \begin{verbatimtab}

### 2023-07-07 popl24/intro.tex
suggestion in the intro

@@
 
-\subsection{The Analysis Challenge of MPC} MPC protocols involve communication
+\subsection{The Analysis Challenge of MPC}
+\jnote{On its surface, security for MPC protocols seems similar to probabilistic noninterference.
+However, MPC security differs from noninterference in a subtle but fundamental way, and approaches for verifying noninterference cannot naturally extend to MPC security.}
+MPC protocols involve communication
 between a group of distributed participants called a \emph{federation}

### 2023-07-08 popl24/minicat.tex
Update minicat.tex

@@
 $\secret{\cid}{w}$, \emph{flips} $\flip{\cid}{w}$, and \emph{views}
-$\view{\cid}{w}$.  Each of these variables are indexed by an ``owner''
+$\view{\cid}{w}$.  Each of these variables is indexed by an ``owner''
 client $\cid$ and distinguishing string $w$. So for example,
@@
 as a primitive since we can implement it in a provably secure
-fashion. However its inclusion simplifies our presentation and is a useful
+fashion. However, its inclusion simplifies our presentation and is a useful
 convenience.
@@
 variables as $\lcod{\store,\be}{\cid}$, where $\store$ associates
-variables with values and all variables must be owned by client
+variables with values, and all variables must be owned by client
 $\cid$. This is defined in Figure \ref{fig-minifed-interp}.
@@
 are used before they're defined, or where expressions mix owners of
-variables. As a sanity condition we will only consider safe 
+variables. As a sanity condition, we will only consider safe 
 programs in our presentation of $\fedprot$, though in Section
@@
 $\sx{1}{0}$ and sends it to party 0, which relays the ciphertext in
-$\vx{0}{0}$ to client 2. Finally, client 2 decrypts the the ciphertext
+$\vx{0}{0}$ to client 2. Finally, client 2 decrypts the ciphertext
 and stores the plaintext in $\vx{2}{2}$.
 \begin{verbatimtab}
-v[2,0] := f[1,0];
-v[0,0] := f[1,0] xor s[1,0];
-v[2,1] := v[0,0];
-v[2,2] := v[2,1] xor v[2,0]
+v[2,"0"] := f[1,"0"];
+v[0,"0"] := f[1,"0"] xor s[1,"0"];
+v[2,"1"] := v[0,"0"];
+v[2,"2"] := v[2,"1"] xor v[2,"0"]
 \end{verbatimtab}
@@
   assume $H = \{ 1 \}$ and $C = \{ 0 \}$. First, the following
-  example shows that the reuse of coin flips can result in release of
+  example shows that the reuse of coin flips can result in the release of
   information, even though flips are chosen from a uniform
@@
   \begin{verbatimtab}
-    v[0,0] := select(s[1,s], flip[1,sx], flip[1,sy]);
-    v[0,1] := flip[1,sx] \end{verbatimtab}
+    v[0,"0"] := select(s[1,"s"], flip[1,"sx"], flip[1,"sy"]);
+    v[0,"1"] := flip[1,"sx"] \end{verbatimtab}
   In the next example, since the nested select is more likely to evaluate
@@
   \begin{verbatimtab}
-  v[0,0] := select(s[1,s],select(flip[1,sx], flip[1,sx], flip[1,sy]),flip[1,z]) \end{verbatimtab}
+  v[0,"0"] := select(s[1,"s"],select(flip[1,"sx"], flip[1,"sx"], flip[1,"sy"]),flip[1,"z"]) \end{verbatimtab}
   We note that these examples are considered ``bad'' in \cite{darais2019language} since there
   no information about honest secrets should be revealed to the adversary in
-  their threat model. For this reason they disallow reuse of flips through a
+  their threat model. For this reason, they disallow the reuse of flips through a
   linear type discipline, which is a culprit in both of these examples. However,
@@
 \begin{verbatimtab}
-  v[1,r0] := s[3,r0];
-  v[1,r1] := s[3,r1];
-  v[2,d] := flip[3,d];
-  v[2,rd] := select(flip[3,d], s[3,r1], s[3,r0]);
-  v[1,e] := s[2,c] xor v[2,d];
-  v[2,f0] := s[1,m0] xor select(v[1, e],v[1,r1],v[1,r0]);
-  v[2,f1] := s[1,m1] xor select(not v[1, e],v[1,r1],v[1,r0]);
-  v[2,mc] := select(s[2,c],v[2,f1],v[2,f0]] xor v[2,rd)
+  v[1,"r0"] := s[3,"r0"];
+  v[1,"r1"] := s[3,"r1"];
+  v[2,"d"] := flip[3,"d"];
+  v[2,"rd"] := select(flip[3,"d"], s[3,"r1"], s[3,"r0"]);
+  v[1,"e"] := s[2,"c"] xor v[2,"d"];
+  v[2,"f0"] := s[1,"m0"] xor select(v[1,"e"],v[1,"r1"],v[1,"r0"]);
+  v[2,"f1"] := s[1,"m1"] xor select(not v[1,"e"],v[1,"r1"],v[1,"r0"]);
+  v[2,"mc"] := select(s[2,"c"],v[2,"f1"],v[2,"f0"]] xor v[2,"rd")
 \end{verbatimtab}
@@
 \begin{verbatimtab}  
-  v[2,mc] := OT(s[2,c], s[1,m1], s[1,m0])
+  v[2,"mc"] := OT(s[2,"c"], s[1,"m1"], s[1,"m0"])
 \end{verbatimtab}
@@
 \begin{verbatimtab}
-  v[2,s1] := flip[1,1] xor flip[1,s1] xor s[1,1];
-  v[3,s1] := flip[1,1];
+  v[2,"s1"] := flip[1,"1"] xor flip[1,"s1"] xor s[1,"1"];
+  v[3,"s1"] := flip[1,"1"];
 
-  v[1,s2] := flip[2,1] xor flip[2,s2] xor s[2,1];
-  v[3,s2] := flip[2,1];
+  v[1,"s2"] := flip[2,"1"] xor flip[2,"s2"] xor s[2,"1"];
+  v[3,"s2"] := flip[2,"1"];
 
-  v[1,s3] := flip[3,1] xor flip[3,s3] xor s[3,1];
-  v[2,s3] := flip[3,1];
+  v[1,"s3"] := flip[3,"1"] xor flip[3,"s3"] xor s[3,"1"];
+  v[2,"s3"] := flip[3,"1"];
 \end{verbatimtab}
 Then, assuming that party 0 is a ``public'' client, each party sums
-its own shares, and then the sum of sum of shares is revealed
+its own shares, and then the sum of the sum of shares is revealed
 in the output view $\vx{0}{output}$. By additive homomorphism of
@@
 \begin{verbatimtab}
-  v[0,ss1] := v[1,s2] xor v[1,s3] xor flip[1,s1];
-  v[0,ss2] := v[2,s1] xor v[2,s3] xor flip[2,s2];
-  v[0,ss3] := v[3,s1] xor v[3,s2] xor flip[3,s3];
+  v[0,"ss1"] := v[1,"s2"] xor v[1,"s3"] xor flip[1,"s1"];
+  v[0,"ss2"] := v[2,"s1"] xor v[2,"s3"] xor flip[2,"s2"];
+  v[0,"ss3"] := v[3,"s1"] xor v[3,"s2"] xor flip[3,"s3"];
 
-  v[0,output] := v[0,ss1] xor v[0,ss2] xor v[0,ss3]
+  v[0,"output"] := v[0,"ss1"] xor v[0,"ss2"] xor v[0,"ss3"]
 \end{verbatimtab}

### 2023-07-08 popl24/metalang.tex
Update metalang.tex

@@
 
-In this Section we introduce the $\metaprot$ language which includes
+In this Section, we introduce the $\metaprot$ language which includes
 structured data and function definitions, which are sufficiently
@@
     } \end{verbatimtab}
-  Here is $\metaprot$ program that uses this function definition:
+  Here is a $\metaprot$ program that uses this function definition:
   \begin{verbatimtab}
@@
 resource-intensive compilation of programs with known errors
-\cite{kreuter2012billion}. Some consequences of safety errors, for example accidental
-reuse of one-time pads, can also undermine security.
+\cite{kreuter2012billion}. Some consequences of safety errors, for example, accidental
+reuse of one-time pads can also undermine security.
 
@@
 $\Gamma$, the top level view effect $V$ is a set of concrete
-$\fedprot$ views which are constructed by disjoint union of the views
+$\fedprot$ views which are constructed by a disjoint union of the views
 in the effect of $e$- the disjointness requirement guarantees that

### 2023-07-08 popl24/hyperprop.tex
Update hyperprop.tex

@@
 security. In Section \ref{section-bruteforce} we show how $\NIMO$ can
-be automatically enforced in a brute-force manner, and in Section
+be automatically enforced in a brute-force manner and in Section
 \ref{section-composition} we show how $\NIMO$ can be refined to obtain
@@
 Here we briefly summarize the formal definition of MPC security. For a
-more detailed discussion the reader is referred to \cite{evans2018pragmatic}.
+more detailed discussion, the reader is referred to \cite{evans2018pragmatic}.
 In this paper, we will consider the \emph{passive} aka
@@
 The simulator is represented by a probabilistic algorithm $\SIM_C$,
-aka a \emph{simulation}, that is parameterized by a corrupt inputs and
+aka a \emph{simulation}, that is parameterized by corrupt inputs and
 the output of an ideal functionality, and that returns a set of
@@
   $$
-  so by the definition of noninterference modulo output
+  so by the definition of noninterference modulo output,
   any choice of $\store'$ yields the same distribution $\condd{\progd(\prog)}{V_C}{\store_S}$.

### 2023-07-08 popl24/composition.tex
Update composition.tex

@@
 described in Section \ref{section-bruteforce} are not scalable to
-large circuits, but in this Section we will demonstrate that correctly
+large circuits, but in this Section, we will demonstrate that correctly
 designed circuit components enjoy compositional properties
@@
 
-In this Section we develop and use implementations of both 2-party GMW
+In this Section, we develop and use implementations of both 2-party GMW
 and YGC to illustrate our approach and demonstrate its generality.
@@
 
-For full details of the GMW protocol the reader is referred to
-\cite{evans2018pragmatic}. Our implementation libary is shown in
+For full details of the GMW protocol, the reader is referred to
+\cite{evans2018pragmatic}. Our implementation library is shown in
 Figure \ref{fig-gmw}, with type signatures for the library functions
@@
   identifier $\ttt{g}$ being used to distinguish randomness used
-  within.  In our version client 1 builds the output table (using
+  within.  In our version, client 1 builds the output table (using
   \ttt{andtablegmw})and transfers the correct output share to client 2
@@
 a circuit with a single And gate and input secrets $\ttt{s1}$ and
-$\ttt{s2}$ from client's 1 and 2 respectively. 
+$\ttt{s2}$ from clients 1 and 2 respectively. 
 \begin{verbatimtab}
@@
 valid type signature as $\ttt{decodegmw}$, $\ttt{encodegmw}$, and
-$\ttt{andgmw}$ respectively. When gates are used in ciruits they are
+$\ttt{andgmw}$ respectively. When gates are used in circuits they are
 parameterized by wire values-- in certification we will use carefully
@@
 for dummy flips preserves separation. 
-In these results and subsequently we will use $f$ to range over flips.
+In these results and subsequently, we will use $f$ to range over flips.
 \begin{lemma}
@@
 Here we show that encoding also preserves the desired invariant.
-The result here is more straightforward than it is for gates, because
+The result here is more straightforward than it is for gates because
 the variables used during encoding are guaranteed to be distinct from
@@
 
-On the basis of the preceding we can prove our main result as
+On the basis of the preceding, we can prove our main result as
 follows. Note that we assume a normal form of programs where the last
@@
 and elsewhere, to which the reader is referred for more in-depth discussion.
-In this implementation client 2 is the \emph{garbler} and
+In this implementation, client 2 is the \emph{garbler} and
 client 1 is the \emph{evaluator}. The garbler builds the garbled
@@
 
-\emph{Wire labels} are fundamental to YGC, and essentially represent
+\emph{Wire labels} are fundamental to YGC and essentially represent
 gate output values in an encrypted form. In our definition, wire
@@
 a circuit with a single and gate and input secrets $\ttt{s1}$ and
-$\ttt{s2}$ from client's 1 and 2 respectively. 
+$\ttt{s2}$ from clients 1 and 2 respectively. 
 \begin{verbatimtab}
@@
 labels, since wire values flow into and out of gates in either
-positive of negative correlation with them.
+positive or negative correlation with them.
 \begin{definition}
@@
 correlated with the wire label or with the inverse of both the key and
-pointer bits. For subsequent discussion we make the following addition
+pointer bits. For subsequent discussion, we make the following addition
 to the codebase:
@@
 %Next,
-First we establish another substitution result, showing that
+First, we establish another substitution result, showing that
 correlations (e.g., between output wire values and labels) are
@@
 
-The requirements that output wire labels are not used more than once
+The requirement that output wire labels are not used more than once
 in YGC reflects a known result, that naive copy gate output is unsound
@@
 automatically verify $\gNIMO$ for $\ttt{andgg}$ without the assumption
-that input wire lables are in independent uniform distributions. We
+that input wire labels are in independent uniform distributions. We
 can capture the necessary fix as in Figure \ref{fig-ygc-copy}. The
@@
 output wire values during evaluation (in \ttt{evalcopy}). Given an
-$\ttt{orgg}$ gate defined as outlined abotve, we can then use secure
+$\ttt{orgg}$ gate defined as outlined above, we can then use secure
 $\ttt{copy}$ as follows:

### 2023-07-08 popl24/minicat.tex
Update minicat.tex

@@
 client $\cid$ and distinguishing string $w$. So for example,
-$\secret{\ttt{1}}{\ttt{foo}}$ is client $\ttt{1}$'s secret called
-$\ttt{foo}$. All clients can make reference to the \emph{oracle}
+$\secret{\ttt{1}}{\ttt{"foo"}}$ is client $\ttt{1}$'s secret 
+$\ttt{"foo"}$. All clients can make reference to the \emph{oracle}
 through variables $\oracle{w}$- this is a form of shared randomness
@@
 shares the randomly generated ``key'' $\fx{1}{0}$ with client 2
-through view $\view{2}{0}$. Then, client 1 encrypts its secret
+through view $\vx{2}{0}$. Then, client 1 encrypts its secret
 $\sx{1}{0}$ and sends it to party 0, which relays the ciphertext in
@@
   distribution. This is because the adversary can deduce that
-  $\secret{1}{s}$ is more likely to be 1 in case their views are the
+  $\sx{1}{s}$ is more likely to be 1 in case their views are the
   same value. We call this subexample $(a)$:
@@
   In the next example, since the nested select is more likely to evaluate
-  to $1$ than $0$, the adversary can deduce that $\secret{1}{s}$ is
-  more likely to be 1 in case $\view{0}{0}$ is 1. We call this subexample $(b)$:
+  to $1$ than $0$, the adversary can deduce that $\sx{1}{s}$ is
+  more likely to be 1 in case $\vx{0}{0}$ is 1. We call this subexample $(b)$:
   \begin{verbatimtab}
-  v[0,"0"] := select(s[1,"s"],select(flip[1,"sx"], flip[1,"sx"], flip[1,"sy"]),flip[1,"z"]) \end{verbatimtab}
+  v[0,"0"] := select(s[1,"s"], 
+  select(flip[1,"sx"], flip[1,"sx"], flip[1,"sy"]),flip[1,"z"]) \end{verbatimtab}
   We note that these examples are considered ``bad'' in \cite{darais2019language} since there
@@
   and client 1 is the sender, and a trusted third party provides the
-  sender and receiver with shared randomness. The secret \ttt{s[2,c]}
-  is the receiver's choice bit, and \ttt{s[1,m0]} and \ttt{s[1,m1]}
+  sender and receiver with shared randomness. The secret $\sx{2}{c}$
+  is the receiver's choice bit, and $\sx{1}{m0}$ and $\sx{1}{m1}$
   are the sender's two message options that the receiver chooses from. 

### 2023-07-08 popl24/intro.tex
Update intro.tex

@@
 important modern, distributed applications such as privacy-preserving
-machine learning and Zero-Knowledge proofs in blockchains \cite{XXX}
+machine learning \cite{li2021privacy, knott2021crypten, koch2020privacy, liu2020privacy} and Zero-Knowledge proofs in blockchains \cite{ishai2009zero, lu2019honeybadgermpc, gao2022symmeproof, tomaz2020preserving}
 \cnote{Need some banger cites here, maybe more MPC hype.}. MPC methods
@@
 approaches such as delimited information release
-\cite{10.1007/978-3-540-37621-7_9}, downgrading policies, and
+\cite{10.1007/978-3-540-37621-7_9}, downgrading policies \cite{li2005downgrading}, and
 relaxed noninterference \cite{10.1145/1040305.1040319}. However, their

### 2023-07-09 popl24/automation.tex
checkpoint

\section{Automating Security Verification}

We will use \emph{certification} to refer to algorithmic methods of
verifying properties of $\minifed$ protocols, vs.~manual methods. For
the sake of clarity we make the following definition:
\begin{definition}
  A predicate $\phi$ on $\minifed$ protocols is \emph{certifiable} iff
  there exists an algorithm $A$ such that $A(\prog)$ iff $\phi(\prog)$,
  and a safe protocol $\prog$ is \emph{certified for $\phi$} iff
  $A(\prog)$ holds.
\end{definition}

We base certification of protocols $\prog$ on a calculation of its
program distribution. Recall from Definition \ref{definition-progd} that
program distributions are calculated from runs of a protocol given
any random tape and input secrets. That is:
\begin{definition}
  Given $\prog$ with $\iov(\prog) = S \cup V$ and $\flips(\prog) = F$, define:
  $$
  \runs(\prog) \defeq \{ \store \in \mems(S\cup F \cup V) \mid \config{\store_{S \cup F}}{\prog} \redxs \config{\store}{\varnothing} \}
  $$
\end{definition}

Given any $\prog$, the set $\runs(\prog)$ can be computed via brute
force methods. One is to fold a standard interpretation of logical
formulas across $\prog$, obtaining a truth table representation of
$\runs(\prog)$. We can also interpret $\prog$ as a stratified Datalog
program, with $\runs(\prog)$ being the set of least Herbrand models
obtained by representing different random tapes and input secrets as
different fact bases. This representation is amenable HPC
optimizations such as parallelization and GPU matrix computations as
shown in recent work
\cite{sakama2017linear,aspis2018linear,nguyen2022enhancing,nguyen2021efficient}.
We provide more details of brute force computation of $\runs(\prog)$ in
Appendix \ref{section-bruteforce} and an implementation of
the tabular method is available in our online repository\footnote{Link
redacted for double-blind reviewing.}.

Given a calculation of $\runs(\prog)$, we can easily calculate its
basic distribution as in Definition \ref{definition-progd)}, as well
as any marginalization (such as $\progd(\prog)$) or conditioning of it.

\subsection{Certification of Extensional Properties}

Given computation of $\prog(\prog)$ for programs $\prog$, we can
immediately observe that the extensional hyperproperties discussed in
Section \ref{section-hyperprop} are certifiable.
\begin{lemma}
  Both $\pni$ (Definition \ref{definition-PNI}) and $\NIMO$ (Definition \ref{definition-NIMO}) are certifiable. 
\end{lemma}

\begin{proof}
  Let $\progd(\prog)$ be computed as described above, and let $\iov(\prog) = S
  \cup V$. To verify $\pni(\prog)$ for
  some given $H$ and $C$, we enumerate $\runs(\prog)$ and check:
  $$
  \progd(\prog)(\store_{S_H}|\store_{S_C}) =
  \progd(\prog)(\store_{S_H}|\store_{S_C \cup V_C})
  $$
  for all $\store \in \mems(S \cup V)$, and return success $\pni(\prog)$ iff every such
  check succeeds. The result follows by Lemmas \ref{lemma-cprogd} and \ref{lemma-pni}.
  To verify $\nimo(\prog)$, we first enumerate all partitions of clients
  into set $H$ and $C$ where $0 \in C$, then enumerate $\runs(\prog)$, and
  then check:
  $$
  \progd(\prog)(\store_{S_H}|\store_{S_C \cup \{\outv\}}) =
  \progd(\prog)(\store_{S_H}|\store_{S_C \cup V_C})
  $$
  The result follows by Lemma \ref{lemma-cprogd} and Lemma \ref{lemma-nimo}.
\end{proof}

We can thus automatically verify security properties of examples considered
in previous sections. 
\begin{lemma}
  $\prog_{\ref{example-he}}$ is passive secure,
  $\prog_{\ref{example-OT}}$ is passive secure assuming client 3 is
  honest, and $\prog_{\ref{example-otp}}$ satisfies probabilistic
  noninterference and hence perfect secrecy as defined in
  \ref{barthe2019probabilistic}.
\end{lemma}
\begin{proof}
  We have certified $\nimo(prog_{\ref{example-he}})$, $\nimo(prog_{\ref{example-OT}})$ assuming
  $3 \in H$, and $\pni(\prog_{\ref{example-otp}})$. The result follows by Lemmas
  \ref{lemma-pni} and \ref{lemma-ni}.
\end{proof}

\subsection{Certification of Composable Intensional Properties}

While the brute force method is feasible for smaller programs, it is
not feasible for larger programs such as large YGC or GMW circuits
since the size of $\runs(\prog)$ is exponential in the size of the
random tape plus the number of input secrets.

However, circuit abstractions are based on component gates which
define subprotocols that we can consider in isolation. Out approach to
verifying circuits strategy is thus to show that any circuit built
from independently certified gates satisfies $\NIMO$. Intuitively, we
can consider any gate as a mini-program and certify intensional
properties of its isolated program distribution. These intensional
properties are guaranteed secure in circuit compositions, as we
demonstrate using methods based on probabilistic separation logic
\ref{barthe2019probabilistic}. While this last step is manual, our
gate certifications are defined generally and can be used to certify
new gate implementations, and verification applies to a circuit of any
size. In Section \ref{section-composition} we will demonstrate our
approach using YGC and GMW as examples.

When we consider intensional properties
of gates in isolation, we will explicitly consider flip conditions in
distributions. Thus we will be directly concerned with basic distributions,
which we denote $\progtt(\prog)$, and we will often focus on views
defined in subprotocols. 
\begin{definition}
  We write $\progtt(\prog)$ to denote the \emph{basic distribution} of
  $\prog$ as defined in Definition \ref{def-progd}. For any program $\prog =
  (\eassign{v_1}{\be_1};\ldots;\eassign{v_n}{\be_n})$, define
  $\vdefs(\prog) \defeq \{ v_1,\ldots,v_n \}$.
\end{definition}
Notions of probabilistic independence, aka \emph{separation}, as well
as correlation, are important to certify. We borrow the symbols $*$
and $\sim$ from \cite{barthe2019probabilistic} with the same
denotations.
\begin{definition}
  We write $\vc{\pmf}{x}{y}$ iff $\pmf(\{ x \mapsto 0\}\ |\ \{ y \mapsto 0 \}) =
  \pmf(\{ x \mapsto 1\}\ |\ \{ y \mapsto 1 \}) = 1$.
  We write $\sep{\pmf}{X}{Y}$ iff for all
    $\store \in \mems(X \cup Y)$ we have
  $\margd{\pmf}{X \cup Y}(\store) =
  \pmf(\store_X) * \pmf(\store_Y)$
\end{definition}
We immediately observe that separation and correlation are certifiable,
since they are based on marginalization and conditioning of program
distributions. 
\begin{lemma}
  Both of the following are true:
  \begin{enumerate}
  \item $\sep{\progtt(\prog)}{X}{Y}$ is certifiable for any $X$ and $Y$.
  \item $\vc{\progtt(\prog)}{x}{y}$ is certifiable for any $x$ and $y$.
  \end{enumerate}
\end{lemma}

\subsection{Compositional Metatheory}

Since we certify properties of gates in isolation with witness inputs,
our metatheory is concerned with preservation of important invariants
when certified components are ``spliced'' into larger program
contexts.  One important reasoning principle is related to
noninterference properties- specifically, if we show that gate views
enjoy a noninterference property wrt their free (input) variables,
probabilistic independence propagates to other variables in the
preceding program.
\begin{lemma}[Noninterference]
  \label{lemma-noninterference}
  Given $\prog_1;\prog_2$ and $X = \iov(\prog_2) - \vdefs(\prog_2)$ and
  $Y \subseteq \vdefs(\prog_2)$. If $\sep{\progd(\prog_1;\prog_2)}{X}{Y}$
  then $\sep{\progd(\prog_1;\prog_2)}{\iov(\prog_1)}{Y}$.
\end{lemma}

Also, when we certify gates in isolation with witness inputs, we can
``splice'' these properties into larger program contexts using the
following substitution Lemmas, showing that the substitution of
wire inputs for witness inputs preserves both separation and correlation.
\begin{lemma}[Substitution$*$]
  \label{lemma-substitution}
  If $\sep{\progtt(\prog_2)}{\{ f \}}{X}$ and
  $\prog_1;\prog_2[\be/f]$ is safe with $\vars(\prog_1,\be) \cap
  X = \varnothing$ then
  $\sep{\progtt(\prog_1;\prog_2[\be/f])}{\vars(\be)}{X}$.
\end{lemma}

\begin{lemma}[Substitution$\sim$]
  \label{lemma-substitution-sim}
  If $\vc{\progtt(\prog_2,\be)}{\itv}{f}$ and $\vc{\progtt(\prog_1,\be')}{\itv}{f'}$
  and $\vars(\be') \cap (\vars(\be) - \{ f \}) = \varnothing$
  then $\vc{\progtt(\prog_1;\prog_2,\be[\be'/f'])}{\itv}{f}$.
\end{lemma}

We note that while the Frame rule of \cite{barthe2019probabilistic}
supports local reasoning in larger program contexts, it does not
support this splicing method, and our Lemmas \ref{lemma-substitution}
and \ref{lemma-substitution-sim} are both novel.

### 2023-07-09 popl24/hyperprop.tex
checkpoint

@@
 \begin{definition}[Probabilistic Noninterference]
+  \label{definition-PNI}
   Given a program $\prog$ with $\iov(\prog) = (S,V)$, we say that
-  $\prog$ satisfies \emph{probabilistic noninterference} iff for all
+  $\prog$ satisfies \emph{probabilistic noninterference}, and we write
+  $\pni(\prog)$, iff for all
   $\store_1, \store_2 \in \mems(S)$:
@@
 \begin{lemma}
+  \label{lemma-pni)
   Given a program $\prog$ with $\iov(\prog) = (S,V)$, 

### 2023-07-09 popl24/automation.tex
checkpoint

@@
 We can thus automatically verify security properties of examples considered
-in previous sections. 
+in previous sections, observing that passive security of $\prog_{\ref{example-OT}}$
+has not been previously demonstrated.
 \begin{lemma}
-  $\prog_{\ref{example-he}}$ is passive secure,
+  $\prog_{\ref{example-he}}$, $\prog_{\ref{example-gmw-andcircuit}}$,
+  $\prog_{\ref{example-ygc-andcircuit}}$ are all passive secure,
   $\prog_{\ref{example-OT}}$ is passive secure assuming client 3 is
@@
 \begin{proof}
-  We have certified $\nimo(prog_{\ref{example-he}})$, $\nimo(prog_{\ref{example-OT}})$ assuming
+  We have certified $\nimo(prog_{\ref{example-he}})$, $\nimo(\prog_{\ref{example-gmw-andcircuit}})$,
+  $\nimo(\prog_{\ref{example-ygc-andcircuit}})$, and $\nimo(prog_{\ref{example-OT}})$ assuming
   $3 \in H$, and $\pni(\prog_{\ref{example-otp}})$. The result follows by Lemmas
-  \ref{lemma-pni} and \ref{lemma-ni}.
+  \ref{lemma-pni} and \ref{lemma-nimo}.
 \end{proof}
@@
 
-While the brute force method is feasible for smaller programs, it is
-not feasible for larger programs such as large YGC or GMW circuits
-since the size of $\runs(\prog)$ is exponential in the size of the
-random tape plus the number of input secrets.
+While the brute force method is feasible for smaller programs,
+including small circuits such as Examples
+$\prog_{\ref{example-gmw-andcircuit}}$ and
+$\prog_{\ref{example-ygc-andcircuit}}$, it is not for larger programs
+such as YGC or GMW circuits since the size of $\runs(\prog)$ is
+exponential in the size of the random tape plus the number of input
+secrets, and circuits can be quite large in practice
+\cite{kreuter2012billion}.  However, the YGC and GMW protocols are
+examples of a common idiom in MPC protocol design- arbitrarily large
+circuits are built up from smaller boolean or arithmetic gates that
+operate on encodings of input secrets.
 
-However, circuit abstractions are based on component gates which
-define subprotocols that we can consider in isolation. Out approach to
-verifying circuits strategy is thus to show that any circuit built
-from independently certified gates satisfies $\NIMO$. Intuitively, we
-can consider any gate as a mini-program and certify intensional
-properties of its isolated program distribution. These intensional
-properties are guaranteed secure in circuit compositions, as we
-demonstrate using methods based on probabilistic separation logic
+Our approach is thus to design gate certifications that demonstrably
+guarantee security for circuits built from them. Intuitively, we can
+consider any gate as a mini-program and certify intensional properties
+of its isolated program distribution. These intensional properties are
+guaranteed secure in circuit compositions, as we demonstrate using
+methods based on probabilistic separation logic
 \ref{barthe2019probabilistic}. While this last step is manual, our
@@
 preceding program.
-\begin{lemma}[Noninterference]
+\begin{restatable}[Noninterference]{lemma}{noninterference}
   \label{lemma-noninterference}
@@
   then $\sep{\progd(\prog_1;\prog_2)}{\iov(\prog_1)}{Y}$.
-\end{lemma}
+\end{restatable}
 
@@
 wire inputs for witness inputs preserves both separation and correlation.
-\begin{lemma}[Substitution$*$]
+\begin{restatable}[Substitution$*$]{lemma}{substar}
   \label{lemma-substitution}
@@
   $\sep{\progtt(\prog_1;\prog_2[\be/f])}{\vars(\be)}{X}$.
-\end{lemma}
+\end{restatable}
 
@@
 support this splicing method, and our Lemmas \ref{lemma-substitution}
-and \ref{lemma-substitution-sim} are both novel.
+and \ref{lemma-substitution-sim} are both novel. Supplementary proofs
+are demonstrated in Appendix \ref{section-proofs}.

### 2023-07-09 popl24/bruteforce.tex
checkpoint

@@
-\section{Brute Force Verification}
+\section{Brute Force Computation of Program Distributions}
 \label{section-bruteforce}
 
-We have shown how our formulation of program distributions, a precise
-definition of intensional protocol behavior, can be used to enforce
-the extensional property of MPC security. This intensional description
-also has the benefit of being flexible- we can use the same mechanisms
-to describe $\NIMO$, noninterference, perfect secrecy, violations of
-obliviousness, etc. These mechanisms can also be exploited to support
-automated program analysis.
-
-We initially explore a brute force approach to automatically verifying
+We first explore a brute force approach to automatically verifying
 program behaviour, the first step of which is to compute the basic
@@
 
-The basic distribution of a program $\prog$ is based on the
-final memories resulting from any run of $\prog$, denoted
-$\runs(\prog)$.
-\begin{definition}
-  Given $\prog$ with $\iov(\prog) = S \cup V$ and $\flips(\prog) = F$:
-  $$
-  \runs(\prog) \defeq \{ \store \in \mems(S\cup F \cup V) \mid \config{\store_{S \cup F}}{\prog} \redxs \config{\store}{\varnothing} \}
-  $$
-\end{definition}
+%The basic distribution of a program $\prog$ is based on the
+%final memories resulting from any run of $\prog$, denoted
+%$\runs(\prog)$.
+%\begin{definition}
+%  Given $\prog$ with $\iov(\prog) = S \cup V$ and $\flips(\prog) = F$:
+%  $$
+%  \runs(\prog) \defeq \{ \store \in \mems(S\cup F \cup V) \mid \config{\store_{S \cup F}}{\prog} \redxs \config{\store}{\varnothing} \}
+%  $$
+%\end{definition}
+
 As we've observed, $|\runs(\prog)| = 2^{|S \cup F|}$, so the size of
@@
   Given $\prog$ where $\iov(\prog) = S \cup V$ and $\flips(\prog) = F$. Define
-  $\cprogd(\prog)$ as the distribution $\margd{\pdf{S \cup F \cup V}}{S\cup V}$,
-  where:
+  $\cprogd(\prog)$ as the distribution $\pmf$
+  where for all $\store \in \mems(S\cup F \cup V)$:
   $$
-  \pdf{S \cup F \cup V}(\store) = \begin{cases}1/2^{|S \cup F|} & \text{if\ } \store \in
+  \pmf(\store) = \begin{cases}1/2^{|S \cup F|} & \text{if\ } \store \in
     \cruns(\prog) \\ 0 & \text{otherwise} \end{cases} 
@@
 
+%%% I moved the following to automation.tex - chris
+\begin{comment}
 We have used the code for $\cruns$ and $\cprogd$ extensively to design and
@@
 \end{proof}
+
+\end{comment}
 

### 2023-07-09 popl24/hyperprop.tex
checkpoint

@@
 \begin{lemma}
-  \label{lemma-pni)
+  \label{lemma-pni}
   Given a program $\prog$ with $\iov(\prog) = (S,V)$, 

### 2023-07-09 popl24/intro.tex
checkpoint

@@
 important modern, distributed applications such as privacy-preserving
-machine learning \cite{li2021privacy, knott2021crypten, koch2020privacy, liu2020privacy} and Zero-Knowledge proofs in blockchains \cite{ishai2009zero, lu2019honeybadgermpc, gao2022symmeproof, tomaz2020preserving}
-\cnote{Need some banger cites here, maybe more MPC hype.}. MPC methods
+machine learning \cite{li2021privacy, knott2021crypten, koch2020privacy, liu2020privacy} and Zero-Knowledge proofs in blockchains \cite{ishai2009zero, lu2019honeybadgermpc, gao2022symmeproof, tomaz2020preserving}.
+MPC methods
 have been developed by the cryptography community for many years,
@@
 amenable to HPC optimizations via translation of $\minifed$ protocols
-into Datalog \cnote{I think this Section can be on the chopping block
-  for space savings, since it's based on truth table
-  computation/Datalog semantics.}.
+into Datalog.
 

### 2023-07-09 popl24/metalang.tex
checkpoint

@@
-\section{A Protocol Metalanguage}
+\section{The $\metaprot$ Metalanguage}
 \label{section-metalang}
@@
 \end{theorem}
+
+\subsection{Example: GMW}
+\label{section-metalang-gmw}
+
+\input{gmw}
+
+\begin{fpfig}[t]{GMW library type annotations.}{fig-gmw-types}
+{\footnotesize
+  \begin{verbatimtab}
+   encodegmw   : string(gid) * string(gid)
+
+   andtablegmw : { k = bool[i]; p = bool[i] }
+
+   andgmw      : string(gid) *  { c1 = bool[1]; c2 = bool[2] } * { c1 = bool[1]; c2 = bool[2] }
+
+   decodegmw   : { c1 = bool[1]; c2 = bool[2] }  \end{verbatimtab}
+}
+\end{fpfig}
+
+The GMW protocol uses secret sharing to represent data flowing through
+circuits. In the 2-party case, clients 1 and 2 each share their input
+secrets, and use those shares to represent inputs to gates. Outputs
+are also represented as shares. We refer to the pair of shares
+representing any particular value as a \emph{wire value}, and
+we represent them via records of the form
+$
+\ttt{\{c1 = } v_1\ttt{;c2 = } v_2\ttt{\}} 
+$
+where $v_1$ and $v_2$ are client 1 and 2's shares respectively.
+
+For full details of the GMW protocol the reader is referred to
+\cite{evans2018pragmatic}. Our implementation libary is shown in
+Figure \ref{fig-gmw}, with type signatures for the library functions
+shown in Figure \ref{fig-gmw-types}. We show the And gate since it is
+an interesting component. The Figure includes the
+following top-level functions:
+\begin{itemize}
+\item \ttt{encodegmw}: This function encodes client 1's and client 2's
+  secret bits called $\ttt{s[1,s1]}$ and $\ttt{s[1,s2]}$ into two
+  distinct wire values (pairs of shares).
+\item \ttt{andgmw}: This function defines the gate $\ttt{g}$, the
+  identifier $\ttt{g}$ being used to distinguish randomness used
+  within.  In our version client 1 builds the output table (using
+  \ttt{andtablegmw})and transfers the correct output share to client 2
+  using 1-out-of-4 OT as per standard GMW protocol.
+\item \ttt{decode}: This function decodes and publishes a wire value
+  by $\ttt{xor}$ing the shares. Note that this requires both client 1
+  and 2 to publicize their shares.
+\end{itemize}
+\begin{example}
+  \label{example-gmw-andcircuit}
+The following program uses our GMW library to define
+a circuit with a single And gate and input secrets $\ttt{s1}$ and
+$\ttt{s2}$ from client's 1 and 2 respectively. 
+\begin{verbatimtab}
+  let ss = encodegmw(s1,s2) in
+  v[0,output] := decode(andgmw(0,ss.shares1,ss.shares2))
+\end{verbatimtab}
+\end{example}
+
+\subsection{Example: YGC}
+\label{section-metalang-ygc}
+
+\input{ygc}
+
+In Figures \ref{fig-ygc-aux} and \ref{fig-ygc-gates} we define a
+codebase for Yao's garbled circuits (YGC). This definition follows the
+\emph{point-and-permute} method described in \cite{evans2018pragmatic}
+and elsewhere, to which the reader is referred for more in-depth discussion.
+In this implementation client 2 is the \emph{garbler} and
+client 1 is the \emph{evaluator}. The garbler builds the garbled
+tables and shares them with the evaluator, who then evaluates
+the gate in an oblivious fashion until the final public output is
+generated through decryption. This definition is well-typed,
+with input type annotations for top-level functions listed in
+Figure \ref{fig-ygc-types}. Well-typed programs using these
+libraries are therefore guaranteed to yield safe $\minifed$
+programs.
+
+\emph{Wire labels} are fundamental to YGC, and essentially represent
+gate output values in an encrypted form. In our definition, wire
+labels are represented by records $\ttt{\{ k = }\beta_1\ttt{; p =
+}\beta_2\ttt{ \}}$, where $\ttt{k}$ is the \emph{key bit} and
+$\ttt{p}$ is the \emph{pointer bit}, and $\beta_1$ and $\beta_2$ are
+flips. Flips in each output wire label are owned by the garbler and
+are unique per gate by definition of their identifying string, and the
+representation of $0$ is the negation of $1$. For example, here is the
+representation of 1 and 0 respectively in the output wire label for a
+hypothetical gate 6:
+\begin{mathpar}
+  \ttt{\{ k = flip[2,gate:6.k]; p =  flip[2,gate:6.p]] \}}
+
+  \ttt{\{ k = not flip[2,gate:6.k]; p =  not flip[2,gate:6.p]] \}}
+\end{mathpar}
+The pointer bits in wire labels are used to select permuted rows in
+table garblings. The key bits are used to identify a unique key for
+table row in each garbled gate. Intuitively, if $\beta_1$ and
+$\beta_2$ are either key or pointer bits encoding 1 on two input wire
+labels to a binary gate, rows and keys in the gate are enumerated in
+the order:
+$$
+\neg\beta_1\neg\beta_2,\ \neg\beta_1\beta_2,\ \beta_1\neg\beta_2,\ \beta_1\beta_2
+$$
+
+In our implementation, gates are wired together using gate
+identifiers, which are strings $w$. Top-level functionality in Figures
+\ref{fig-ygc-aux} and \ref{fig-ygc-gates} includes the following:
+\begin{itemize}
+\item \ttt{andgate}: This defines a subprotocol for the garbler
+  to define a garbled gate $\ttt{gid}$ with input wires from gates
+  $\ttt{ga}$ and $\ttt{gb}$. The garbler generates keys and garbles
+  the rows in YGC fashion, them with client $1$ in
+  views in a standard form. For example, the view for
+  a hypothetical gate 6, row 2 garbled truth table is $\ttt{v[1,gate:6tt2]}$.
+  We note that garbled gates of other binary operators can be obtained with
+  replacement of $\ttt{andtable}$ with appropriate garbled table definitions. 
+\item \ttt{evalgate}: This defines a subprotocol for the evaluator to
+  evaluate gate $\ttt{gid}$ given input wire values $\ttt{wva}$ and
+  $\ttt{wvb}$.
+\item \ttt{garbledecode} and \ttt{evaldecode}: The former function
+  defines the garbler's protocol for encrypting the circuit
+  output from final gate $\ttt{gid}$, and the latter defines
+  the evaluator's output decryption protocol.
+\item \ttt{encode}: This defines the initial phase of the protocol,
+  where the evaluator receives the wire value from their own
+  secret $\sx{1}{sa}$ via $\ttt{OT}$, and the garbler communicates
+  the wire value for their own secret $\sx{2}{sb}$ directly.
+\end{itemize}
+\begin{example}
+  \label{example-ygc-andcircuit}
+The following program uses our YGC library to define
+a circuit with a single and gate and input secrets $\ttt{s1}$ and
+$\ttt{s2}$ from client's 1 and 2 respectively. 
+\begin{verbatimtab}
+  andgate(0,s1,s2);
+  garbledecode(0);
+  let secrets = encode(s1,s2) in
+  v[0,output] := decode(evalgate(0, secrets.wv1, secrets.wv2))
+\end{verbatimtab}
+\end{example}

### 2023-07-09 popl24/proofs.tex
checkpoint

--- 2023-07-06 popl24/composition.tex
@@
-\section{Component Certification and Composition}
-\label{section-composition}
-
-\input{gmw}
-
-The Yao's Garbled Circuits (YGC) and GMW protocols are examples of a
-common idiom in MPC protocol design- arbitrarily large \emph{circuits}
-are built up from smaller boolean or arithmetic \emph{gates} that
-operate on encodings of input secrets. These circuits can become quite
-large in practice \cite{kreuter2012billion}, but are built from
-well-defined components. Brute force verification techniques as
-described in Section \ref{section-bruteforce} are not scalable to
-large circuits, but in this Section we will demonstrate that correctly
-designed circuit components enjoy compositional properties
-that can be automatically certification on small program components
-using brute force methods. We prove that these properties preserve
-invariants throughout circuit definition that yield $\NIMO$ and thus
-scale to arbitrary programs. Furthermore, the certifications can be
-used on arbitrary gate definitions, allowing new gate implementations
-to be added soundly to libraries.
-
-In this Section we develop and use implementations of both 2-party GMW
-and YGC to illustrate our approach and demonstrate its generality.
-Main results are in Theorems \ref{theorem-gmw-NIMO} and
-\ref{theorem-ygc-NIMO}.
-
-\subsection{Compositional Properties of GMW}
-
-The GMW protocol uses secret sharing to represent data flowing through
-circuits. In the 2-party case, clients 1 and 2 each share their input
-secrets, and use those shares to represent inputs to gates. Outputs
-are also represented as shares. We refer to the pair of shares
-representing any particular value as a \emph{wire value}, and
-we represent them via records of the form
-$
-\ttt{\{c1 = } v_1\ttt{;c2 = } v_2\ttt{\}} 
-$
-where $v_1$ and $v_2$ are client 1 and 2's shares respectively.
-
-For full details of the GMW protocol the reader is referred to
-\cite{evans2018pragmatic}. Our implementation libary is shown in
-Figure \ref{fig-gmw}, with type signatures for the library functions
-shown in Figure \ref{fig-gmw-types}. We show the And gate since it is
-an interesting component. The Figure includes the
-following top-level functions:
-\begin{itemize}
-\item \ttt{encodegmw}: This function encodes client 1's and client 2's
-  secret bits called $\ttt{s[1,s1]}$ and $\ttt{s[1,s2]}$ into two
-  distinct wire values (pairs of shares).
-\item \ttt{andgmw}: This function defines the gate $\ttt{g}$, the
-  identifier $\ttt{g}$ being used to distinguish randomness used
-  within.  In our version client 1 builds the output table (using
-  \ttt{andtablegmw})and transfers the correct output share to client 2
-  using 1-out-of-4 OT as per standard GMW protocol.
-\item \ttt{decode}: This function decodes and publishes a wire value
-  by $\ttt{xor}$ing the shares. Note that this requires both client 1
-  and 2 to publicize their shares.
-\end{itemize}
-
-\subsubsection{Gate Certification}
-
-Our certification techniques are defined generally wrt implementations
-of input encoding, decoding, and internal gates. In the following we
-will refer to arbitrary decode, encode, and gate functions with the
-restriction that any function in each of these categories has the same
-valid type signature as $\ttt{decodegmw}$, $\ttt{encodegmw}$, and
-$\ttt{andgmw}$ respectively. When gates are used in ciruits they are
-parameterized by wire values-- in certification we will use carefully
-chosen witness parameters.
-
-When we consider internal components and hence intensional properties
-of circuits, we will consider behaviors with respect to particular bits
-of randomness. Hence we will consider basic program distributions. 
-\begin{definition}
-  We write $\progtt(\prog)$ to denote the \emph{basic distribution} of
-  $\prog$ as defined in Definition \ref{def-progd}. For any program $\prog =
-  (\eassign{v_1}{\be_1};\ldots;\eassign{v_n}{\be_n})$, define
-  $\vdefs(\prog) \defeq \{ v_1,\ldots,v_n \}$.
-\end{definition}
-We also need to formalize notions of probabilistic independence, aka
-\emph{separation}, and correlation. We borrow the symbols $*$ and $\sim$
-from \cite{barthe2019probabilistic} with the same denotations.
-\begin{definition}
-  We write $\vc{\pmf}{x}{y}$ iff $\pmf(\{ x \mapsto 0\}\ |\ \{ y \mapsto 0 \}) =
-  \pmf(\{ x \mapsto 1\}\ |\ \{ y \mapsto 1 \}) = 1$.
-  We write $\sep{\pmf}{Y}{Z}$ iff for all
-    $\store \in \mems(X \cup Y)$:
-  $$\margd{\pmf}{Y \cup Z}(\store) =
-  \pmf(\store_Y) * \pmf(\store_Z)$$ 
-\end{definition}
-
-\begin{fpfig}[t]{GMW library type annotations.}{fig-gmw-types}
-{\footnotesize
-  \begin{verbatimtab}
-   encodegmw   : string(gid) * string(gid)
-
-   andtablegmw : { k = bool[i]; p = bool[i] }
-
-   andgmw      : string(gid) *  { c1 = bool[1]; c2 = bool[2] } * { c1 = bool[1]; c2 = bool[2] }
-
-   decodegmw   : { c1 = bool[1]; c2 = bool[2] }  \end{verbatimtab}
-}
-\end{fpfig}
-
-Now we can define GMW gate certification as follows. Intuitively, we certify
-that GMW gates satisfy a probabilistic noninterference property wrt each output
-share considered individually. This ensures that input dependencies remain
-encrypted from each client's perspective during circuit evaluation.
-\begin{definition}[GMW Gate Certification]
-  \label{definition-gmwgate-certification}
-  Let $\mathit{gate}$ be a gate function and let $e$ be the following
-  $\metaprot$ program:
-  $$
-  \begin{array}{l}
-    \ttt{let in1 = \{c1 = flip[1,s11];c2 = flip[2,s12]\} in}\\
-    \ttt{let in2 = \{c1 = flip[1,s21];c2 = flip[2,s22]\} in}\\
-    \mathit{gate}\ttt{(g,in1,in2)}
-  \end{array}
-  $$
-  and let $
-  \config{\varnothing}{e}\redxs\config{\prog}{\ttt{\{c1 = v[1,gout]; c2 = v[2,gout]\}}}
-  $.
-  Then $\mathit{gate}$ is \emph{certified} iff the following conditions hold:
-  \begin{enumerate}[\hspace{5mm}i.]
-  \item $\sep{\progtt(\prog)}{\ttt{\{flip[1,s11],flip[1,s21],flip[2,s12],flip[2,s22]\}}}{\ttt{\v[1,gout]\}}}$
-  \item $\sep{\progtt(\prog)}{\ttt{\{flip[1,s11],flip[1,s21],flip[2,s12],flip[2,s22]\}}}{\ttt{\v[2,gout]\}}}$
-  % \item $\sep{\progtt(\prog)}{\ttt{\{v[1,gout]\}}}{\ttt{\{v[2,gout]\}}}$
-  \item All variables in:$$\vars(\prog) - \{ \ttt{flip[1,s11],flip[1,s21],flip[2,s12],flip[2,s22]}\}$$
-    are distinguished by $\ttt{g}$ (i.e., contain $\ttt{g}$ as an identifier substring). 
-  \end{enumerate}
-\end{definition}
-
-Encoding certification establishes the desired property in the protocol-- that
-input secrets of each client are independent of the other client's views. 
-\begin{definition}[GMW Encode Certification]
-  \label{definition-gmwencode-certification}
-  Let $\mathit{encode}$ be an encoding function, and let:
-  $$
-  \config{\varnothing}{\ttt{encode(s1,s2)}} \redxs
-  \config{\prog}{\ttt{\{shares1 = }v_1\ttt{; shares2 = }v_2\ttt{\}}}
-  $$
-  where:
-  $$
-  \begin{array}{rcl}
-    v_1 &=& \ttt{\{ c1 = v[1,s1out]; c2 = v[2,s1out] \}}\\
-    v_2 &=& \ttt{\{ c1 = v[1,s2out]; c2 = v[2,s2out] \}}
-  \end{array}
-  $$
-  Then $\mathit{encode}$ is \emph{certified} iff each of the following conditions hold:
-  \begin{enumerate}[\hspace{5mm}i.]
-    %\item $\sep{\progtt(\prog)}{\ttt{\{ s[1,s1],  v[1,s1out], v[1,s2out] \}}}{\ttt{\{ s[2,s2], v[2,s1out], v[2,s2out] \}}}$
-  \item $\sep{\progtt(\prog)}{\ttt{\{s[1,s1]\}}}{\ttt{\{v[2,s1out],v[2,s2out]\}}}$
-  \item $\sep{\progtt(\prog)}{\ttt{\{s[2,s2]\}}}{\ttt{\{v[1,s1out],v[1,s2out]\}}}$
-  \item All $x \in \vars(\prog)$ are distinguished by $\ttt{in1}$ or $\ttt{in2}$. 
-  \end{enumerate}
-  \end{definition}
-
-We have certified both $\ttt{encodegmw}$ and $\ttt{andgmw}$. Note that
-an Or or Xor gate, for example, could be certified with this same
-method to extend our library.
-
-\subsubsection{Gate Invariants} We can show that composition maintains an
-important invariant in a circuit $\prog$- namely that client 1's
-secrets remain independent of client 2's views, and vice-versa, prior
-to decoding.  That is, where $\iov(\prog) = S \cup V$, we show that
-$\sep{\progd(\prog)}{S_{\{1\}}}{V_{\{2\}}}$ and
-$\sep{\progd(\prog)}{S_{\{2\}}}{V_{\{1\}}}$ are preserved prior to
-decoding.
+\section{Supplementary Proofs}
+\label{section-proofs}
 
@@
 
-%The following Lemma allows us to separate variables in programs that may not occur
-%in the same subprogram, and hence consider gate certification properties in various
-%contexts. It is reminiscent of the \emph{frame} rule of \cite{barthe2019probabilistic}.
-%\begin{lemma}
-%  \label{lemma-gmw-frame}
-%  If $\sep{\progtt(\prog_1;\prog_2)}{Y}{V}$ for $Y = \vars(\prog_1) \cap \vars(\prog_2)$ and
-%  $V \in \vdefs(\prog_2)$ and $\sep{\progtt(\prog_1)}{X}{Y}$ for
-%  $X \not\in \vars(\prog_2)$ then $\sep{\progtt(\prog_1;\prog_2)}{X}{V}$. 
-%\end{lemma}
-%\begin{proof}
-%  By assumption and Lemma \ref{lemma-sepjoin}, for all
-%  $\store^1,\store^2 \in \runs(\prog_1;\prog_2)$ we have $\store^1_Y
-%  \cap \store^2_V \subseteq \store_a$ for some $\store_a \in
-%  \runs(\prog_1;\prog_2)$ and $\store^1_X \cap \store^2_Y \subseteq
-%  \store_b$ for some $\store_b \in \runs(\prog_1;\prog_2)$.  Thus, since
-%  we assume $X \not\in \vars(\prog_2)$ we have $\store_a \cap \store_b
-%  \in \runs(\prog_1;\prog_2)$ and $\store^1_X \cap \store^2_V =
-%  \store_a \cap \store_b$, obtaining the result by
-%  \ref{lemma-sepjoin}.
-%\end{proof}
-
-The next ``noninterference'' Lemma is important to prove the desired invariance
-preservation property of GMW gates, given certification.
-\begin{lemma}[Noninterference]
-  \label{lemma-noninterference}
-  Given $\prog_1;\prog_2$ and $X = \iov(\prog_2) - \vdefs(\prog_2)$ and
-  $Y \subseteq \vdefs(\prog_2)$. If $\sep{\progd(\prog_1;\prog_2)}{X}{Y}$
-  then $\sep{\progd(\prog_1;\prog_2)}{\iov(\prog_1)}{Y}$.
-\end{lemma}
-
-The next two Lemmas allow us to use the properties of gates established by
-certification and replace the ``dummy'' parameters used there with actual
-wire values in full circuit contexts. Intuitively we show that substitution
-for dummy flips preserves separation.
 \begin{lemma}
@@
 
-\begin{lemma}[Substitution]
-  \label{lemma-substitution}
-  If $\sep{\progtt(\prog_2)}{\{ f \}}{X}$ and
-  $\prog_1;\prog_2[\be/f]$ is safe with $\vars(\prog_1,\be) \cap
-  X = \varnothing$ then
-  $\sep{\progtt(\prog_1;\prog_2[\be/f])}{\vars(\be)}{X}$.
-\end{lemma}
+\substar*
+
 \begin{proof}
@@
 \end{proof}
-
-Now we can prove that certified gates preserve the desired invariant
-during execution in arbitrary circuit contexts.
-We formulate this in terms of pre-~and post-conditions of gate
-evaluation.
-\begin{lemma}[GMW Gate Invariant]
-  \label{lemma-gmw-preservation}
-  Given well-typed $\prog_1;E[\mathit{gate}(\ttt{g},v_1,v_2)]$ for certified $\mathit{gate}$ where:
-  $$
-  \config{\prog_1}{E[\mathit{gate}(\ttt{g},v_1,v_2)]} \redxs \config{\prog_1;\prog_2}{E[v]}
-  $$
-  If we have the following preconditions, where $\iov(\prog_1) = S \cup V$:
-  \begin{enumerate}[\hspace{5mm}i.]
-  \item $\sep{\progd(\prog_1)}{S_{\{2\}}}{V_{\{1\}}}$
-  \item $\sep{\progd(\prog_1)}{S_{\{1\}}}{V_{\{2\}}}$
-  \end{enumerate}
-  then we have as a postconditions, where $\iov(\prog_1;\prog_2) = S \cup V'$:
-  \begin{enumerate}[\hspace{5mm}i.]
-  \item $\sep{\progd(\prog_1;\prog_2)}{S_{\{2\}}}{V'_{\{1\}}}$
-  \item $\sep{\progd(\prog_1;\prog_2)}{S_{\{1\}}}{V'_{\{2\}}}$
-  \end{enumerate}
-\end{lemma}
-
-\begin{proof}
-  Given well-typedness of $\prog_1;E[\mathit{gate}(\ttt{g},v_1,v_2)]$, we have for some $\ttt{g1}$ and $\ttt{g2}$:
-  $$
-  \begin{array}{lcl}
-   v_1 &=& \ttt{\{c1 = v[1,g1out]; c2 = v[2,g1out]\}}\\
-   v_2 &=& \ttt{\{c1 = v[1,g2out]; c2 = v[2,g2out]\}}\\
-    v &=& \ttt{\{c1 = v[1,gout]; c2 = v[2,gout]\}}
-  \end{array}
-  $$
-  Let $\prog$ be as defined in Definition \ref{definition-gmwgate-certification}.
-  We observe:
-  $$
-  {\footnotesize
-    \begin{array}{c}
-      \prog_2 = \\
-      \prog[\ttt{v[1,g1out]}/\ttt{flip[1,s11]}][\ttt{v[2,g1out]}/\ttt{flip[2,s12]}][\ttt{v[1,g2out]}/\ttt{flip[1,s21]}][\ttt{v[2,g2out]}/\ttt{flip[2,s22]}]
-    \end{array}
-  }
-  $$
-  Now, the assumption of well-typedness also assures that $\ttt{g}$
-  has not been previously used as a gate identifier in $\prog_1$,
-  since otherwise the views $\ttt{v[1,gout]}$ and $\ttt{v[2,gout]}$
-  would have been previously defined. This ensures $\vars(\prog_1,x)
-  \cap \vars(\prog) = \varnothing$ for:
-  $$x \in \{ \ttt{v[1,g1out]}, \ttt{v[2,g1out]}, \ttt{v[1,g2out]}, \ttt{v[2,g2out]}\}$$
-  given condition (iii) of Definition \ref{definition-gmwgate-certification}.
-  Thus by Lemma \ref{lemma-substitution} we have:
-  $$\sep{\progtt(\prog_1;\prog_2)}{\ttt{\{v[1,g1out],v[1,g2out],v[2,g1out],v[2,g2out]\}}}{\ttt{\{v[1,gout]\}}}$$
-  and
-  $$\sep{\progtt(\prog_1;\prog_2)}{\ttt{\{v[1,g1out],v[1,g2out],v[2,g1out],v[2,g2out]\}}}{\ttt{\{v[2,gout]\}}}$$
-  Now, by Lemmas \ref{lemma-noninterference} and \ref{lemma-separation} and condition (iii) of Definition
-  \ref{definition-gmwgate-certification} we then have:
-  $$\sep{\progtt(\prog_1;\prog_2)}{(S_{\{1\}} \cup V_{\{2\}})}{\ttt{\{v[2,gout]\}}} \quad \text{and} \quad
-  \sep{\progtt(\prog_1;\prog_2)}{(S_{\{2\}} \cup V_{\{1\}})}{\ttt{\{v[1,gout]\}}}$$
-  so by precondition assumptions and Lemma \ref{lemma-separation} we have:
-  $$\sep{\progtt(\prog_1;\prog_2)}{S_{\{1\}}}{V'_{\{2\}}} \quad \text{and} \quad
-  \sep{\progtt(\prog_1;\prog_2)}{S_{\{2\}}}{V'_{\{1\}}}$$
-\end{proof}
-
-Here we show that encoding also preserves the desired invariant.
-The result here is more straightforward than it is for gates, because
-the variables used during encoding are guaranteed to be distinct from
-the rest of the program.
-\begin{lemma}[GMW Encode Invariant]
-  \label{lemma-gmw-encode}
-  Given well-typed $\prog_1;E[\mathit{encode}\ttt{(s1,s2)}]$ for certified $\mathit{encode}$, let:
-  $$
-  \config{\prog_1}{E[\mathit{encode}(\ttt{s1},\ttt{s2})]} \redxs
-  \config{\prog_1;\prog_2}{E[v]}
-  $$
-  If we have the following preconditions, where $\iov(\prog_1) = S' \cup V$:
-  \begin{enumerate}[\hspace{5mm}i.]
-  \item $\sep{\progd(\prog_1)}{S_{\{2\}}}{V_{\{1\}}}$
-  \item $\sep{\progd(\prog_1)}{S_{\{1\}}}{V_{\{2\}}}$
-  \end{enumerate}
-  then we have as a postconditions, where $\iov(\prog_1;\prog_2) = S' \cup V'$: 
-  \begin{enumerate}[\hspace{5mm}i.]
-  \item $\sep{\progd(\prog_1;\prog_2)}{S'_{\{2\}}}{V'_{\{1\}}}$
-  \item $\sep{\progd(\prog_1;\prog_2)}{S'_{\{1\}}}{V'_{\{2\}}}$
-  \end{enumerate}
-\end{lemma}
-
-\begin{proof}
-  By assumptions of well-typedness we have:
-  $$
-  \begin{array}{rcl}
-    v &=&  \ttt{\{shares1 = }v_1\ttt{; shares2 = }v_2\ttt{\}}\\
-    v_1 &=& \ttt{\{ c1 = v[1,s1out]; c2 = v[2,s1out] \}}\\
-    v_2 &=& \ttt{\{ c1 = v[1,s2out]; c2 = v[2,s2out] \}}
-  \end{array}
-  $$
-  And by Definition \ref{definition-gmwencode-certification} we have:
-  \begin{mathpar}
-    \sep{\progtt(\prog)}{\ttt{\{s[1,s1]\}}}{\ttt{\{v[2,s1out],v[2,s2out]\}}}
-
-    \sep{\progtt(\prog)}{\ttt{\{s[2,s2]\}}}{\ttt{\{v[1,s1out],v[1,s2out]\}}}
-  \end{mathpar}
-  Since we assume that secrets are in uniform and independent marginal
-  distributions a priori, and $\vars(\prog_1) \cap \vars(\prog_2) =
-  \varnothing$ by condition (iii) of Definition 
-  \ref{definition-gmwencode-certification} and assumptions of well-typedness, 
-  the result follows by preconditions and Lemma \ref{lemma-separation}.
-\end{proof}
-
-On the basis of the preceding we can prove our main result as
-follows. Note that we assume a normal form of programs where the last
-instruction is a decoding and public reveal of the output.
-\begin{theorem}
-  \label{theorem-gmw-NIMO}
-  If $\eassign{\outv}{\ttt{decode}(e)}$ is a well-typed GMW circuit
-  definition using certified components and
-  $\config{\varnothing}{\eassign{\outv}{\ttt{decode}(e)}} \redxs
-  \config{\prog}{\varnothing}$, then $\NIMO(\prog)$.
-\end{theorem}
-\begin{proof}
-  Let $\iov(\prog) = S \cup (V \cup \outv)$. By definition of $\ttt{decode}$
-  and assumptions of well-typedness we have for some $\ttt{g}$:
-  $$
-  \config{\varnothing}{\eassign{\outv}{\ttt{decode}(e)}} \redxs
-  \config{\prog;\eassign{\outv}{\ttt{v[1,gout] xor v[2,gout]}}}{\varnothing}
-  $$
-  where by Lemmas \ref{lemma-gmw-encode} and \ref{lemma-gmw-preservation} we
-  have $\sep{\progd(\prog)}{S_{\{1\}}}{V_{\{2\}}}$ and
-  $\sep{\progd(\prog)}{S_{\{2\}}}{V_{\{1\}}}$.
-  Let $\prog' \defeq \prog;\eassign{\outv}{\ttt{v[1,gout] xor v[2,gout]}}$. The
-  preceding then implies
-  $\sep{\progd(\prog')}{S_{\{1\}}}{V_{\{2\}}}$ and
-  $\sep{\progd(\prog')}{S_{\{2\}}}{V_{\{1\}}}$,
-  so that for all $\beta$:
-  $$\sep{\condd{\progd(\prog')}{S_{\{1\}} \cup V_{\{2\}}}{\{ \outv \mapsto \beta \}}}{S_{\{1\}}}{V_{\{2\}}}
-  \quad \text{and} \quad
-    \sep{\condd{\progd(\prog')}{S_{\{2\}} \cup V_{\{1\}}}{\{ \outv \mapsto \beta \}}}{S_{\{2\}}}{V_{\{1\}}}$$
-  thus for all $\store \in \mems(V_2 \cup \{ \outv \})$:
-  $$\condd{\progd(\prog')}{S_{\{1\}}}{\store}
-  = \condd{\progd(\prog')}{S_{\{1\}}}{\store_{\{\outv\}}}$$
-  and for all $\store \in \mems(V_1 \cup \{ \outv \})$:
-  $$\condd{\progd(\prog')}{S_{\{2\}}}{\store}
-  = \condd{\progd(\prog')}{S_{\{2\}}}{\store_{\{\outv\}}}$$
-  establishing the result by Lemma \ref{lemma-nimo} and Definition \ref{definition-NIMO}. 
-\end{proof}

### 2023-07-09 popl24/scratch.tex
checkpoint

@@
-\subsection{Compositional Properties of YGC}
-\label{section-composition-ygc}
+\section{Automating Verification of Circuit Security}
 
-\input{ygc}
+Main results are in Theorems \ref{theorem-gmw-NIMO} and
+\ref{theorem-ygc-NIMO}.
 
-In Figures \ref{fig-ygc-aux} and \ref{fig-ygc-gates} we define a
-codebase for Yao's garbled circuits (YGC). This definition follows the
-\emph{point-and-permute} method described in \cite{evans2018pragmatic}
-and elsewhere, to which the reader is referred for more in-depth discussion.
-In this implementation client 2 is the \emph{garbler} and
-client 1 is the \emph{evaluator}. The garbler builds the garbled
-tables and shares them with the evaluator, who then evaluates
-the gate in an oblivious fashion until the final public output is
-generated through decryption. This definition is well-typed,
-with input type annotations for top-level functions listed in
-Figure \ref{fig-ygc-types}. Well-typed programs using these
-libraries are therefore guaranteed to yield safe $\minifed$
-programs.
+\subsection{GMW Component Certification}
 
-\emph{Wire labels} are fundamental to YGC, and essentially represent
-gate output values in an encrypted form. In our definition, wire
-labels are represented by records $\ttt{\{ k = }\beta_1\ttt{; p =
-}\beta_2\ttt{ \}}$, where $\ttt{k}$ is the \emph{key bit} and
-$\ttt{p}$ is the \emph{pointer bit}, and $\beta_1$ and $\beta_2$ are
-flips. Flips in each output wire label are owned by the garbler and
-are unique per gate by definition of their identifying string, and the
-representation of $0$ is the negation of $1$. For example, here is the
-representation of 1 and 0 respectively in the output wire label for a
-hypothetical gate 6:
+Now we can define GMW gate certification as follows. Intuitively, we certify
+that GMW gates satisfy a probabilistic noninterference property wrt each output
+share considered individually. This ensures that input dependencies remain
+encrypted from each client's perspective during circuit evaluation.
+\begin{definition}[GMW Gate Certification]
+  \label{definition-gmwgate-certification}
+  Let $\mathit{gate}$ be a gate function and let $e$ be the following
+  $\metaprot$ program:
+  $$
+  \begin{array}{l}
+    \ttt{let in1 = \{c1 = flip[1,s11];c2 = flip[2,s12]\} in}\\
+    \ttt{let in2 = \{c1 = flip[1,s21];c2 = flip[2,s22]\} in}\\
+    \mathit{gate}\ttt{(g,in1,in2)}
+  \end{array}
+  $$
+  and let $
+  \config{\varnothing}{e}\redxs\config{\prog}{\ttt{\{c1 = v[1,gout]; c2 = v[2,gout]\}}}
+  $.
+  Then $\mathit{gate}$ is \emph{certified} iff the following conditions hold:
+  \begin{enumerate}[\hspace{5mm}i.]
+  \item $\sep{\progtt(\prog)}{\ttt{\{flip[1,s11],flip[1,s21],flip[2,s12],flip[2,s22]\}}}{\ttt{\v[1,gout]\}}}$
+  \item $\sep{\progtt(\prog)}{\ttt{\{flip[1,s11],flip[1,s21],flip[2,s12],flip[2,s22]\}}}{\ttt{\v[2,gout]\}}}$
+  % \item $\sep{\progtt(\prog)}{\ttt{\{v[1,gout]\}}}{\ttt{\{v[2,gout]\}}}$
+  \item All variables in:$$\vars(\prog) - \{ \ttt{flip[1,s11],flip[1,s21],flip[2,s12],flip[2,s22]}\}$$
+    are distinguished by $\ttt{g}$ (i.e., contain $\ttt{g}$ as an identifier substring). 
+  \end{enumerate}
+\end{definition}
+
+Encoding certification establishes the desired property in the protocol-- that
+input secrets of each client are independent of the other client's views. 
+\begin{definition}[GMW Encode Certification]
+  \label{definition-gmwencode-certification}
+  Let $\mathit{encode}$ be an encoding function, and let:
+  $$
+  \config{\varnothing}{\ttt{encode(s1,s2)}} \redxs
+  \config{\prog}{\ttt{\{shares1 = }v_1\ttt{; shares2 = }v_2\ttt{\}}}
+  $$
+  where:
+  $$
+  \begin{array}{rcl}
+    v_1 &=& \ttt{\{ c1 = v[1,s1out]; c2 = v[2,s1out] \}}\\
+    v_2 &=& \ttt{\{ c1 = v[1,s2out]; c2 = v[2,s2out] \}}
+  \end{array}
+  $$
+  Then $\mathit{encode}$ is \emph{certified} iff each of the following conditions hold:
+  \begin{enumerate}[\hspace{5mm}i.]
+    %\item $\sep{\progtt(\prog)}{\ttt{\{ s[1,s1],  v[1,s1out], v[1,s2out] \}}}{\ttt{\{ s[2,s2], v[2,s1out], v[2,s2out] \}}}$
+  \item $\sep{\progtt(\prog)}{\ttt{\{s[1,s1]\}}}{\ttt{\{v[2,s1out],v[2,s2out]\}}}$
+  \item $\sep{\progtt(\prog)}{\ttt{\{s[2,s2]\}}}{\ttt{\{v[1,s1out],v[1,s2out]\}}}$
+  \item All $x \in \vars(\prog)$ are distinguished by $\ttt{in1}$ or $\ttt{in2}$. 
+  \end{enumerate}
+  \end{definition}
+
+We have certified both $\ttt{encodegmw}$ and $\ttt{andgmw}$. Note that
+an Or or Xor gate, for example, could be certified with this same
+method to extend our library.
+\begin{lemma}
+  \label{lemma-gmw-certification}
+  Each of $\ttt{encodegmw}$ and $\ttt{andgmw}$ are certified.
+\end{lemma}
+
+\subsection{GMW Gate Invariants and Circuit Security}
+We can show that composition maintains an important invariant in a
+circuit $\prog$- namely that client 1's secrets remain independent of
+client 2's views, and vice-versa, prior to decoding.  That is, where
+$\iov(\prog) = S \cup V$, we show that
+$\sep{\progd(\prog)}{S_{\{1\}}}{V_{\{2\}}}$ and
+$\sep{\progd(\prog)}{S_{\{2\}}}{V_{\{1\}}}$ are preserved prior to
+decoding.
+
+%The following Lemma allows us to separate variables in programs that may not occur
+%in the same subprogram, and hence consider gate certification properties in various
+%contexts. It is reminiscent of the \emph{frame} rule of \cite{barthe2019probabilistic}.
+%\begin{lemma}
+%  \label{lemma-gmw-frame}
+%  If $\sep{\progtt(\prog_1;\prog_2)}{Y}{V}$ for $Y = \vars(\prog_1) \cap \vars(\prog_2)$ and
+%  $V \in \vdefs(\prog_2)$ and $\sep{\progtt(\prog_1)}{X}{Y}$ for
+%  $X \not\in \vars(\prog_2)$ then $\sep{\progtt(\prog_1;\prog_2)}{X}{V}$. 
+%\end{lemma}
+%\begin{proof}
+%  By assumption and Lemma \ref{lemma-sepjoin}, for all
+%  $\store^1,\store^2 \in \runs(\prog_1;\prog_2)$ we have $\store^1_Y
+%  \cap \store^2_V \subseteq \store_a$ for some $\store_a \in
+%  \runs(\prog_1;\prog_2)$ and $\store^1_X \cap \store^2_Y \subseteq
+%  \store_b$ for some $\store_b \in \runs(\prog_1;\prog_2)$.  Thus, since
+%  we assume $X \not\in \vars(\prog_2)$ we have $\store_a \cap \store_b
+%  \in \runs(\prog_1;\prog_2)$ and $\store^1_X \cap \store^2_V =
+%  \store_a \cap \store_b$, obtaining the result by
+%  \ref{lemma-sepjoin}.
+%\end{proof}
+
+Now we can prove that certified gates preserve the desired invariant
+during execution in arbitrary circuit contexts.
+We formulate this in terms of pre-~and post-conditions of gate
+evaluation.
+\begin{lemma}[GMW Gate Invariant]
+  \label{lemma-gmw-preservation}
+  Given well-typed $\prog_1;E[\mathit{gate}(\ttt{g},v_1,v_2)]$ for certified $\mathit{gate}$ where:
+  $$
+  \config{\prog_1}{E[\mathit{gate}(\ttt{g},v_1,v_2)]} \redxs \config{\prog_1;\prog_2}{E[v]}
+  $$
+  If we have the following preconditions, where $\iov(\prog_1) = S \cup V$:
+  \begin{enumerate}[\hspace{5mm}i.]
+  \item $\sep{\progd(\prog_1)}{S_{\{2\}}}{V_{\{1\}}}$
+  \item $\sep{\progd(\prog_1)}{S_{\{1\}}}{V_{\{2\}}}$
+  \end{enumerate}
+  then we have as a postconditions, where $\iov(\prog_1;\prog_2) = S \cup V'$:
+  \begin{enumerate}[\hspace{5mm}i.]
+  \item $\sep{\progd(\prog_1;\prog_2)}{S_{\{2\}}}{V'_{\{1\}}}$
+  \item $\sep{\progd(\prog_1;\prog_2)}{S_{\{1\}}}{V'_{\{2\}}}$
+  \end{enumerate}
+\end{lemma}
+
+\begin{proof}
+  Given well-typedness of $\prog_1;E[\mathit{gate}(\ttt{g},v_1,v_2)]$, we have for some $\ttt{g1}$ and $\ttt{g2}$:
+  $$
+  \begin{array}{lcl}
+   v_1 &=& \ttt{\{c1 = v[1,g1out]; c2 = v[2,g1out]\}}\\
+   v_2 &=& \ttt{\{c1 = v[1,g2out]; c2 = v[2,g2out]\}}\\
+    v &=& \ttt{\{c1 = v[1,gout]; c2 = v[2,gout]\}}
+  \end{array}
+  $$
+  Let $\prog$ be as defined in Definition \ref{definition-gmwgate-certification}.
+  We observe:
+  $$
+  {\footnotesize
+    \begin{array}{c}
+      \prog_2 = \\
+      \prog[\ttt{v[1,g1out]}/\ttt{flip[1,s11]}][\ttt{v[2,g1out]}/\ttt{flip[2,s12]}][\ttt{v[1,g2out]}/\ttt{flip[1,s21]}][\ttt{v[2,g2out]}/\ttt{flip[2,s22]}]
+    \end{array}
+  }
+  $$
+  Now, the assumption of well-typedness also assures that $\ttt{g}$
+  has not been previously used as a gate identifier in $\prog_1$,
+  since otherwise the views $\ttt{v[1,gout]}$ and $\ttt{v[2,gout]}$
+  would have been previously defined. This ensures $\vars(\prog_1,x)
+  \cap \vars(\prog) = \varnothing$ for:
+  $$x \in \{ \ttt{v[1,g1out]}, \ttt{v[2,g1out]}, \ttt{v[1,g2out]}, \ttt{v[2,g2out]}\}$$
+  given condition (iii) of Definition \ref{definition-gmwgate-certification}.
+  Thus by Lemma \ref{lemma-substitution} we have:
+  $$\sep{\progtt(\prog_1;\prog_2)}{\ttt{\{v[1,g1out],v[1,g2out],v[2,g1out],v[2,g2out]\}}}{\ttt{\{v[1,gout]\}}}$$
+  and
+  $$\sep{\progtt(\prog_1;\prog_2)}{\ttt{\{v[1,g1out],v[1,g2out],v[2,g1out],v[2,g2out]\}}}{\ttt{\{v[2,gout]\}}}$$
+  Now, by Lemmas \ref{lemma-noninterference} and \ref{lemma-separation} and condition (iii) of Definition
+  \ref{definition-gmwgate-certification} we then have:
+  $$\sep{\progtt(\prog_1;\prog_2)}{(S_{\{1\}} \cup V_{\{2\}})}{\ttt{\{v[2,gout]\}}} \quad \text{and} \quad
+  \sep{\progtt(\prog_1;\prog_2)}{(S_{\{2\}} \cup V_{\{1\}})}{\ttt{\{v[1,gout]\}}}$$
+  so by precondition assumptions and Lemma \ref{lemma-separation} we have:
+  $$\sep{\progtt(\prog_1;\prog_2)}{S_{\{1\}}}{V'_{\{2\}}} \quad \text{and} \quad
+  \sep{\progtt(\prog_1;\prog_2)}{S_{\{2\}}}{V'_{\{1\}}}$$
+\end{proof}
+
+Here we show that encoding also preserves the desired invariant.
+The result here is more straightforward than it is for gates, because
+the variables used during encoding are guaranteed to be distinct from
+the rest of the program.
+\begin{lemma}[GMW Encode Invariant]
+  \label{lemma-gmw-encode}
+  Given well-typed $\prog_1;E[\mathit{encode}\ttt{(s1,s2)}]$ for certified $\mathit{encode}$, let:
+  $$
+  \config{\prog_1}{E[\mathit{encode}(\ttt{s1},\ttt{s2})]} \redxs
+  \config{\prog_1;\prog_2}{E[v]}
+  $$
+  If we have the following preconditions, where $\iov(\prog_1) = S' \cup V$:
+  \begin{enumerate}[\hspace{5mm}i.]
+  \item $\sep{\progd(\prog_1)}{S_{\{2\}}}{V_{\{1\}}}$
+  \item $\sep{\progd(\prog_1)}{S_{\{1\}}}{V_{\{2\}}}$
+  \end{enumerate}
+  then we have as a postconditions, where $\iov(\prog_1;\prog_2) = S' \cup V'$: 
+  \begin{enumerate}[\hspace{5mm}i.]
+  \item $\sep{\progd(\prog_1;\prog_2)}{S'_{\{2\}}}{V'_{\{1\}}}$
+  \item $\sep{\progd(\prog_1;\prog_2)}{S'_{\{1\}}}{V'_{\{2\}}}$
+  \end{enumerate}
+\end{lemma}
+
+\begin{proof}
+  By assumptions of well-typedness we have:
+  $$
+  \begin{array}{rcl}
+    v &=&  \ttt{\{shares1 = }v_1\ttt{; shares2 = }v_2\ttt{\}}\\
+    v_1 &=& \ttt{\{ c1 = v[1,s1out]; c2 = v[2,s1out] \}}\\
+    v_2 &=& \ttt{\{ c1 = v[1,s2out]; c2 = v[2,s2out] \}}
+  \end{array}
+  $$
+  And by Definition \ref{definition-gmwencode-certification} we have:
   \begin{mathpar}
-  \ttt{\{ k = flip[2,gate:6.k]; p =  flip[2,gate:6.p]] \}}
+    \sep{\progtt(\prog)}{\ttt{\{s[1,s1]\}}}{\ttt{\{v[2,s1out],v[2,s2out]\}}}
 
-  \ttt{\{ k = not flip[2,gate:6.k]; p =  not flip[2,gate:6.p]] \}}
+    \sep{\progtt(\prog)}{\ttt{\{s[2,s2]\}}}{\ttt{\{v[1,s1out],v[1,s2out]\}}}
   \end{mathpar}
-The pointer bits in wire labels are used to select permuted rows in
-table garblings. The key bits are used to identify a unique key for
-table row in each garbled gate. Intuitively, if $\beta_1$ and
-$\beta_2$ are either key or pointer bits encoding 1 on two input wire
-labels to a binary gate, rows and keys in the gate are enumerated in
-the order:
+  Since we assume that secrets are in uniform and independent marginal
+  distributions a priori, and $\vars(\prog_1) \cap \vars(\prog_2) =
+  \varnothing$ by condition (iii) of Definition 
+  \ref{definition-gmwencode-certification} and assumptions of well-typedness, 
+  the result follows by preconditions and Lemma \ref{lemma-separation}.
+\end{proof}
+
+On the basis of the preceding we can prove our main result as
+follows. Note that we assume a normal form of programs where the last
+instruction is a decoding and public reveal of the output.
+\begin{theorem}
+  \label{theorem-gmw-NIMO}
+  If $\eassign{\outv}{\ttt{decode}(e)}$ is a well-typed GMW circuit
+  definition using certified components and
+  $\config{\varnothing}{\eassign{\outv}{\ttt{decode}(e)}} \redxs
+  \config{\prog}{\varnothing}$, then $\NIMO(\prog)$.
+\end{theorem}
+\begin{proof}
+  Let $\iov(\prog) = S \cup (V \cup \outv)$. By definition of $\ttt{decode}$
+  and assumptions of well-typedness we have for some $\ttt{g}$:
   $$
-\neg\beta_1\neg\beta_2,\ \neg\beta_1\beta_2,\ \beta_1\neg\beta_2,\ \beta_1\beta_2
+  \config{\varnothing}{\eassign{\outv}{\ttt{decode}(e)}} \redxs
+  \config{\prog;\eassign{\outv}{\ttt{v[1,gout] xor v[2,gout]}}}{\varnothing}
   $$
+  where by Lemmas \ref{lemma-gmw-encode} and \ref{lemma-gmw-preservation} we
+  have $\sep{\progd(\prog)}{S_{\{1\}}}{V_{\{2\}}}$ and
+  $\sep{\progd(\prog)}{S_{\{2\}}}{V_{\{1\}}}$.
+  Let $\prog' \defeq \prog;\eassign{\outv}{\ttt{v[1,gout] xor v[2,gout]}}$. The
+  preceding then implies
+  $\sep{\progd(\prog')}{S_{\{1\}}}{V_{\{2\}}}$ and
+  $\sep{\progd(\prog')}{S_{\{2\}}}{V_{\{1\}}}$,
+  so that for all $\beta$:
+  $$\sep{\condd{\progd(\prog')}{S_{\{1\}} \cup V_{\{2\}}}{\{ \outv \mapsto \beta \}}}{S_{\{1\}}}{V_{\{2\}}}
+  \quad \text{and} \quad
+    \sep{\condd{\progd(\prog')}{S_{\{2\}} \cup V_{\{1\}}}{\{ \outv \mapsto \beta \}}}{S_{\{2\}}}{V_{\{1\}}}$$
+  thus for all $\store \in \mems(V_2 \cup \{ \outv \})$:
+  $$\condd{\progd(\prog')}{S_{\{1\}}}{\store}
+  = \condd{\progd(\prog')}{S_{\{1\}}}{\store_{\{\outv\}}}$$
+  and for all $\store \in \mems(V_1 \cup \{ \outv \})$:
+  $$\condd{\progd(\prog')}{S_{\{2\}}}{\store}
+  = \condd{\progd(\prog')}{S_{\{2\}}}{\store_{\{\outv\}}}$$
+  establishing the result by Lemma \ref{lemma-nimo} and Definition \ref{definition-NIMO}. 
+\end{proof}
 
-In our implementation, gates are wired together using gate
-identifiers, which are strings $w$. Top-level functionality in Figures
-\ref{fig-ygc-aux} and \ref{fig-ygc-gates} includes the following:
-\begin{itemize}
-\item \ttt{andgate}: This defines a subprotocol for the garbler
-  to define a garbled gate $\ttt{gid}$ with input wires from gates
-  $\ttt{ga}$ and $\ttt{gb}$. The garbler generates keys and garbles
-  the rows in YGC fashion, them with client $1$ in
-  views in a standard form. For example, the view for
-  a hypothetical gate 6, row 2 garbled truth table is $\ttt{v[1,gate:6tt2]}$.
-  We note that garbled gates of other binary operators can be obtained with
-  replacement of $\ttt{andtable}$ with appropriate garbled table definitions. 
-\item \ttt{evalgate}: This defines a subprotocol for the evaluator to
-  evaluate gate $\ttt{gid}$ given input wire values $\ttt{wva}$ and
-  $\ttt{wvb}$.
-\item \ttt{garbledecode} and \ttt{evaldecode}: The former function
-  defines the garbler's protocol for encrypting the circuit
-  output from final gate $\ttt{gid}$, and the latter defines
-  the evaluator's output decryption protocol.
-\item \ttt{encode}: This defines the initial phase of the protocol,
-  where the evaluator receives the wire value from their own
-  secret $\sx{1}{sa}$ via $\ttt{OT}$, and the garbler communicates
-  the wire value for their own secret $\sx{2}{sb}$ directly.
-\end{itemize}
-\begin{example}
-  \label{example-andcircuit}
-The following program uses our YGC library to define
-a circuit with a single and gate and input secrets $\ttt{s1}$ and
-$\ttt{s2}$ from client's 1 and 2 respectively. 
-\begin{verbatimtab}
-  andgate(0,s1,s2);
-  garbledecode(0);
-  let secrets = encode(s1,s2) in
-  v[0,output] := decode(evalgate(0, secrets.wv1, secrets.wv2))
-\end{verbatimtab}
-\end{example}
-We have verified passive security of the $\fedprot$ protocol
-generated by this and other small circuits using the
-technique described in Lemma \ref{lemma-bruteforce-nimo}.
-But large circuits with thousands of gates would be
-intractable to verify with this method. In the next Section
-we discuss compositional methods to address this issue.
-
-\subsubsection{Gate Certification}
+\subsection{YGC Component Certification}
 \label{section-ygc-certification}
@@
 
-\subsubsection{Gate Invariants and Circuit Security}
+\subsection{YGC Gate Invariants and Circuit Security}
 \label{section-composition-metatheory}
@@
 %Next,
-First we establish another substitution result, showing that
-correlations (e.g., between output wire values and labels) are
-preserved by substitution of correlated values for flips (e.g.,
-between input wire values and labels).
-\begin{lemma}[Substitution$\sim$]
-  \label{lemma-substitution-sim}
-  If $\vc{\progtt(\prog_2,\be)}{\itv}{f}$ and $\vc{\progtt(\prog_1,\be')}{\itv}{f'}$
-  and $\vars(\be') \cap (\vars(\be) - \{ f \}) = \varnothing$
-  then $\vc{\progtt(\prog_1;\prog_2,\be[\be'/f'])}{\itv}{f}$.
-\end{lemma}
+
 Now we can show how to splice gates into arbitrary circuits in an

### 2023-07-09 popl24/metalang.tex
checkpoint

@@
 \begin{fpfig}[t]{Syntax of $\metaprot$.}{fig-metaprot-syntax}
+  {\small
     $$
     \begin{array}{rcl@{\hspace{8mm}}r}
-\flab &\in& \mathrm{Field}\\
-x &\in& \mathrm{EVar}\\
-f &\in& \mathrm{FName}\\[2mm]
+      \multicolumn{3}{l}{\flab \in \mathrm{Field},\   x \in \mathrm{EVar}, \  f \in \mathrm{FName}}\\[2mm]
+      %x &\in& \mathrm{EVar}\\
+      %f &\in& \mathrm{FName}\\[2mm]
       e &::=& b \mid \flip{e}{e} \mid \secret{e}{e} \mid \view{e}{e} \mid \oracle{e} \mid \enot\ e \mid e\ \eand\ e \mid e\ \exor\ e \mid & \textit{expressions}\\[0mm]
@@
     $$
+  }
 \end{fpfig}
@@
   with fields $\ttt{s1-3}$:
+  {\small
   \begin{verbatimtab}
-    share3(client, secretid)
-    {
+      share3(client, secretid) {
         let s1 = flip[client, share1] in
@@
       } \end{verbatimtab}
+  }
   Here is a $\metaprot$ program that uses this function definition:
+  {\small
     \begin{verbatimtab}
@@
       v[3,s1] := shares.s3 \end{verbatimtab}
+  }
   which generates the following $\minifed$ program, as we formalize in Example \ref{example-share3-eval}
   below:
+  {\small
   \begin{verbatimtab}
@@
       v[3,s1] := flip[1, share1] xor flip[1, share2] xor s[1, s:mysecret] \end{verbatimtab}
+  }
 \end{example}
@@
 \begin{fpfig}[t]{Evaluation contexts and operational semantics of $\metaprot$.}{fig-metaprot-semantics}
+  {\small
     $$
@@
 \begin{fpfig}[t]{Type Syntax of $\metaprot$.}{fig-metaprot-tsyntax}
+  {\small
     $$
     \begin{array}{rcl@{\hspace{2mm}}r}
-\srct &::=& \cidty{e} \mid \stringty{e} \mid \bet{e} \mid  & \gdesc{types}\\ 
- &&  \{ \flab : \srct;\ldots;\flab : \srct \} \mid \tau * \cdots * \tau \rightarrow \tau,\viewst \\[1mm]
+      \srct &::=& \cidty{e} \mid \stringty{e} \mid \bet{e} \mid  \{ \flab : \srct;\ldots;\flab : \srct \} \mid \tau * \cdots * \tau \rightarrow \tau,\viewst & \gdesc{types}\\ [1mm]
+      %&&  \{ \flab : \srct;\ldots;\flab : \srct \} \mid \tau * \cdots * \tau \rightarrow \tau,\viewst \\[1mm]
       \viewst  &::=& \view{e}{e};\viewst \mid \varnothing   & \gdesc{view effects}\\[1mm]
@@
 \begin{fpfig}[t]{Selected $\metaprot$ type judgement rules.}{fig-metaprot-tjudge}
+{\small
 \begin{mathpar}
@@
 \end{mathpar}
+}
 \end{fpfig}
@@
 $\ttt{s2}$ from client's 1 and 2 respectively. 
+{\small
   \begin{verbatimtab}
-  let ss = encodegmw(s1,s2) in
-  v[0,output] := decode(andgmw(0,ss.shares1,ss.shares2))
+  let ss = encodegmw(s1,s2) in v[0,output] := decode(andgmw(0,ss.shares1,ss.shares2))
   \end{verbatimtab}
+}
 \end{example}

### 2023-07-09 popl24/minicat.tex
checkpoint

@@
 values arbitrarily assigned in the initial program memory.  Programs
-execute deterministically given the random tape. The random tape
-formulation supports our automated analysis as we discuss in
-section \dnote{5?}
+execute deterministically given the random tape.
 
@@
 \begin{fpfig}[t]{$\minifed$ source code syntax.}{fig-minifed-syntax}
+  {\small
     $$
     \begin{array}{rcl@{\hspace{8mm}}r}
-b &\in& \{ \etrue, \efalse \} \\
-w &\in& \mathrm{String} \\ 
-\cid &\in& \mathrm{Clients} \subset  \mathbb{N} \\[2mm]%\qquad  (\mathrm{Clients} \subset \mathbb{N} \text{\ in\ this\ presentation})\\[2mm]
-\bop &\in& \{ \eand, \eor, \exor \} \\[2mm]
+      \multicolumn{4}{c}{b \in \{ \etrue, \efalse \},\  w \in \mathrm{String},\ \cid \in \mathrm{Clients} \subset  \mathbb{N},\ \bop \in \{ \eand, \eor, \exor \}} \\[2mm]
       \be &::=& b \mid \flip{\cid}{w} \mid \secret{\cid}{w} \mid \view{\cid}{w} \mid \oracle{w} \mid & \textit{boolean expressions}\\
@@
 \begin{fpfig}[t]{$\minifed$ expression interpretation.}{fig-minifed-interp}
-\begin{eqnarray*}
+  {\small
+  $$
+  \begin{array}{cc}
+    \begin{array}{rcl}
       \lcod{\store, \etrue}{\cid} &=& 1\\
@@
       \lcod{\store, \flip{\cid}{w}}{\cid} &=& \store(\flip{\cid}{w})\\
-\lcod{\store, \secret{\cid}{w}}{\cid} &=& \store(\secret{\cid}{w})\\
+      \lcod{\store, \secret{\cid}{w}}{\cid} &=& \store(\secret{\cid}{w})
+    \end{array} & 
+    \begin{array}{rcl}
       \lcod{\store, \view{\cid}{w}}{\cid} &=& \store(\view{\cid}{w})\\
@@
       \lcod{\store, \be_1\ \mathit{binop}\ \be_2}{\cid} &=&
-    \cod{\lcod{\store,\be_1}{\cid}\ \mathit{binop}\ \lcod{\store,\be_2}{\cid}}\\
-\lcod{\store, \select{\be_1}{\be_2}{\be_3}}{\cid} &=&
+      \cod{\lcod{\store,\be_1}{\cid}\ \mathit{binop}\ \lcod{\store,\be_2}{\cid}}
+    \end{array}
+  \end{array}
+  $$
+  $$
+  \lcod{\store, \select{\be_1}{\be_2}{\be_3}}{\cid} =
   \begin{cases}
@@
       %             \end{cases}
-\end{eqnarray*}
+  $$
+  }
 \end{fpfig}
@@
 and stores the plaintext in $\vx{2}{2}$.
+{\small
   \begin{verbatimtab}
@@
     v[2,"1"] := v[0,"0"];
-v[2,"2"] := v[2,"1"] xor v[2,"0"]
-\end{verbatimtab}
+    v[2,"2"] := v[2,"1"] xor v[2,"0"]  \end{verbatimtab}
+}
 Intuitively, this protocol is \emph{correct} because $\vx{2}{2}$ is
@@
   same value. We call this subexample $(a)$:
+  {\small
     \begin{verbatimtab}
@@
     v[0,"1"] := flip[1,"sx"] \end{verbatimtab}
+  }
   In the next example, since the nested select is more likely to evaluate
@@
   more likely to be 1 in case $\vx{0}{0}$ is 1. We call this subexample $(b)$:
+  {\small
     \begin{verbatimtab}
@@
   select(flip[1,"sx"], flip[1,"sx"], flip[1,"sy"]),flip[1,"z"]) \end{verbatimtab}
+  }
   We note that these examples are considered ``bad'' in \cite{darais2019language} since there
@@
   are the sender's two message options that the receiver chooses from. 
+  {\small
     \begin{verbatimtab}
@@
       v[2,"f1"] := s[1,"m1"] xor select(not v[1,"e"],v[1,"r1"],v[1,"r0"]);
-  v[2,"mc"] := select(s[2,"c"],v[2,"f1"],v[2,"f0"]] xor v[2,"rd")
-\end{verbatimtab}
+      v[2,"mc"] := select(s[2,"c"],v[2,"f1"],v[2,"f0"]] xor v[2,"rd") \end{verbatimtab}
+  }
 Below we will discuss how to automatically verify security properties
@@
 the following:
+{\small
 \begin{verbatimtab}  
-  v[2,"mc"] := OT(s[2,"c"], s[1,"m1"], s[1,"m0"])
-\end{verbatimtab}
+      v[2,"mc"] := OT(s[2,"c"], s[1,"m1"], s[1,"m0"]) \end{verbatimtab}
+}
 \end{example}
@@
 each other as follows:
+{\small
 \begin{verbatimtab}
@@
      v[3,"s1"] := flip[1,"1"];
-
      v[1,"s2"] := flip[2,"1"] xor flip[2,"s2"] xor s[2,"1"];
      v[3,"s2"] := flip[2,"1"];
-
      v[1,"s3"] := flip[3,"1"] xor flip[3,"s3"] xor s[3,"1"];
-  v[2,"s3"] := flip[3,"1"];
-\end{verbatimtab}
+     v[2,"s3"] := flip[3,"1"] \end{verbatimtab}
+}
 Then, assuming that party 0 is a ``public'' client, each party sums
@@
 $\ttt{xor}$, this output is the sum of the 3 secrets.
+{\small
 \begin{verbatimtab}
@@
      v[0,"ss3"] := v[3,"s1"] xor v[3,"s2"] xor flip[3,"s3"];
-
-  v[0,"output"] := v[0,"ss1"] xor v[0,"ss2"] xor v[0,"ss3"]
-\end{verbatimtab}
+     v[0,"output"] := v[0,"ss1"] xor v[0,"ss2"] xor v[0,"ss3"] \end{verbatimtab}
+}
 We note that, as for MPC protocols generally, the notion of

### 2023-07-10 popl24/metalang.tex
checkpoint

@@
 than the examples we've considered so far. These larger protocols are
-typically based on compositional units. An example of this is Yao's
-Garbled Circuits (YGC), which are composed of so-called garbled gates.
-Languages for defined garbled circuits, beginning with Fairplay \cite{269581},
-treat gates as compositional units that are wired together by the programmer
-to generate a complete circuit. The $\fedprot$ language is low-level
-and does not include abstractions for defining composable elements.
+typically based on compositional units. Examples include GMW circuits
+and Yao's Garbled Circuits (YGC), which are composed of so-called
+garbled gates.  Languages for defined garbled circuits, beginning with
+Fairplay \cite{269581}, treat gates as compositional units that are
+wired together by the programmer to generate a complete circuit. The
+$\fedprot$ language is low-level and does not include abstractions for
+defining composable elements.
 
@@
 
-When we consider the example of YGC in detail below, our focus will be
+When we consider the examples of GMW and YGC in detail below, our focus will be
 on developing a codebase that can be used to define arbitrary
@@
 
-\input{ygc}
+{\footnotesize
+\begin{verbatimtab}
+  owl(gid) { { k = flip[2,"gate:" || gid || ".k"]; p = flip[2,"gate:" || gid || ".p"] } }
+
+  encode(sa, sb) {
+    let owl1 = owl(sa) in
+    let owl2 = owl(sb) in
+    v[1,gate: || sa || 1.k] := OT(s[1,sa],owl1.k,(not owl1.k));
+    v[1,gate: || sa || 1.p] := OT(s[1,sa],owl1.p,(not owl1.p));
+    v[1,gate: || sb || 2.k] := select(s[2,sb],owl2.k,(not owl2.k));
+    v[1,gate: || sb || 2.p] := select(s[2,sb],owl2.p,(not owl2.p));
+    { wv1 = { k = v[1,"gate:" || sa || "1.k"]; p = v[1,"gate:" || sa || "1.p"] };
+      wv2 = { k = v[1,"gate:" || sb || "2.k"]; p = v[1,"gate:" || sb || "2.p"] } }
+  }
+
+  andgg(g, ga, gb, wva, wvb) { andgate(g, ga, gb); evalgate(g, wva, wvb) }
+
+  decode(g, wv) { garbledecode(wv.g); evaldecode(v) }
+\end{verbatimtab}
+}
+
+{\footnotesize
+\begin{verbatimtab}
+  let secrets = encode("s1","s2") in v[0,output] := decode(andgg("0", secrets.wv1, secrets.wv2)) \end{verbatimtab}
+}
+
+{\footnotesize
+\begin{verbatimtab}
+  owl    :   string(gid)
+
+  encode :   string(sa) * string(sb)
+
+  andgg  :  string(g) * { k : bool[1]; p : bool[1] } *  { k : bool[1]; p : bool[1] }
+
+  decode :  string(g) * { k : bool[1]; p : bool[1] }
+\end{verbatimtab}
+}
 
@@
 Figure \ref{fig-ygc-types}. Well-typed programs using these
-libraries are therefore guaranteed to yield safe $\minifed$
+libraries are therguaranteed to yield safe $\minifed$
 programs.

### 2023-07-10 popl24/scratch.tex
checkpoint

--- 2023-07-07 popl24/scratch.tex
@@
-\subsection{Compositional Properties of YGC}
-\label{section-composition-ygc}
-
-\input{ygc}
-
-In Figures \ref{fig-ygc-aux} and \ref{fig-ygc-gates} we define a
-codebase for Yao's garbled circuits (YGC). This definition follows the
-\emph{point-and-permute} method described in \cite{evans2018pragmatic}
-and elsewhere, to which the reader is referred for more in-depth discussion.
-In this implementation client 2 is the \emph{garbler} and
-client 1 is the \emph{evaluator}. The garbler builds the garbled
-tables and shares them with the evaluator, who then evaluates
-the gate in an oblivious fashion until the final public output is
-generated through decryption. This definition is well-typed,
-with input type annotations for top-level functions listed in
-Figure \ref{fig-ygc-types}. Well-typed programs using these
-libraries are therefore guaranteed to yield safe $\minifed$
-programs.
-
-\emph{Wire labels} are fundamental to YGC, and essentially represent
-gate output values in an encrypted form. In our definition, wire
-labels are represented by records $\ttt{\{ k = }\beta_1\ttt{; p =
-}\beta_2\ttt{ \}}$, where $\ttt{k}$ is the \emph{key bit} and
-$\ttt{p}$ is the \emph{pointer bit}, and $\beta_1$ and $\beta_2$ are
-flips. Flips in each output wire label are owned by the garbler and
-are unique per gate by definition of their identifying string, and the
-representation of $0$ is the negation of $1$. For example, here is the
-representation of 1 and 0 respectively in the output wire label for a
-hypothetical gate 6:
-\begin{mathpar}
-  \ttt{\{ k = flip[2,gate:6.k]; p =  flip[2,gate:6.p]] \}}
-
-  \ttt{\{ k = not flip[2,gate:6.k]; p =  not flip[2,gate:6.p]] \}}
-\end{mathpar}
-The pointer bits in wire labels are used to select permuted rows in
-table garblings. The key bits are used to identify a unique key for
-table row in each garbled gate. Intuitively, if $\beta_1$ and
-$\beta_2$ are either key or pointer bits encoding 1 on two input wire
-labels to a binary gate, rows and keys in the gate are enumerated in
-the order:
-$$
-\neg\beta_1\neg\beta_2,\ \neg\beta_1\beta_2,\ \beta_1\neg\beta_2,\ \beta_1\beta_2
-$$
-
-In our implementation, gates are wired together using gate
-identifiers, which are strings $w$. Top-level functionality in Figures
-\ref{fig-ygc-aux} and \ref{fig-ygc-gates} includes the following:
-\begin{itemize}
-\item \ttt{andgate}: This defines a subprotocol for the garbler
-  to define a garbled gate $\ttt{gid}$ with input wires from gates
-  $\ttt{ga}$ and $\ttt{gb}$. The garbler generates keys and garbles
-  the rows in YGC fashion, them with client $1$ in
-  views in a standard form. For example, the view for
-  a hypothetical gate 6, row 2 garbled truth table is $\ttt{v[1,gate:6tt2]}$.
-  We note that garbled gates of other binary operators can be obtained with
-  replacement of $\ttt{andtable}$ with appropriate garbled table definitions. 
-\item \ttt{evalgate}: This defines a subprotocol for the evaluator to
-  evaluate gate $\ttt{gid}$ given input wire values $\ttt{wva}$ and
-  $\ttt{wvb}$.
-\item \ttt{garbledecode} and \ttt{evaldecode}: The former function
-  defines the garbler's protocol for encrypting the circuit
-  output from final gate $\ttt{gid}$, and the latter defines
-  the evaluator's output decryption protocol.
-\item \ttt{encode}: This defines the initial phase of the protocol,
-  where the evaluator receives the wire value from their own
-  secret $\sx{1}{sa}$ via $\ttt{OT}$, and the garbler communicates
-  the wire value for their own secret $\sx{2}{sb}$ directly.
-\end{itemize}
-\begin{example}
-  \label{example-andcircuit}
-The following program uses our YGC library to define
-a circuit with a single and gate and input secrets $\ttt{s1}$ and
-$\ttt{s2}$ from client's 1 and 2 respectively. 
-\begin{verbatimtab}
-  andgate(0,s1,s2);
-  garbledecode(0);
-  let secrets = encode(s1,s2) in
-  v[0,output] := decode(evalgate(0, secrets.wv1, secrets.wv2))
-\end{verbatimtab}
-\end{example}
-We have verified passive security of the $\fedprot$ protocol
-generated by this and other small circuits using the
-technique described in Lemma \ref{lemma-bruteforce-nimo}.
-But large circuits with thousands of gates would be
-intractable to verify with this method. In the next Section
-we discuss compositional methods to address this issue.
-
-\subsubsection{Gate Certification}
-\label{section-ygc-certification}
-
-Our certification method relies on a definition of compositional units comprising
-complete related subprotocols-- i.e., both the garbler's
-construction of the gates and the evaluator's evaluation of them. We
-therefore make the following definitions. However, we note that our
-formal results are applicable to the YGC style where the garbler
-shares the entire circuit prior to evaluation.
-\begin{definition}
-  \label{ygc-modules}
-  We define a composable internal \ttt{and} gate sharing and evaluation function
-  called $\ttt{andgg}$, and an output decoder gate share and evaluation function called
-  $\ttt{decode}$, as follows. 
-\begin{verbatimtab}
-  andgg(g, ga, gb, wva, wvb) { andgate(g, ga, gb); evalgate(g, wva, wvb) }
-  decode(g, v) { garbledecode(g); evaldecode(v) }
-\end{verbatimtab}
-\end{definition}
-
-Our certification techniques are defined generally wrt implementations
-of input encoding, decoding, and internal gates. In the following we
-will refer to arbitrary decode, encode, and gate functions with the
-restriction that any function in each of these categories has the same
-valid type signature as $\ttt{decode}$, $\ttt{encode}$, and
-$\ttt{andgg}$ respectively.
-
-It is necessary to define notions of correlation with wire
-labels, since wire values flow into and out of gates in either
-positive of negative correlation with them.
-\begin{definition}
-  We write $\gc{\prog}{\{ \ttt{k = }\be_{k}; \ttt{p = }\be_{p} \}}{g}$ iff one of the
-  following hold:
-  \begin{enumerate}[\hspace{5mm}i.]
-  \item $\vc{\progtt(\prog,\be_k)}{\itv}{\ttt{flip[2,gate:}g\ttt{.k]}}\quad \text{and} \quad
-    \vc{\progtt(\prog,\be_p)}{\itv}{\ttt{flip[2,gate:}g\ttt{.p]}}$
-  \item $\vc{\progtt(\prog,\enot\ \be_k)}{\itv}{\ttt{flip[2,gate:}g\ttt{.k]}}\quad \text{and}\quad
-    \vc{\progtt(\prog,\enot\ \be_p)}{\itv}{\ttt{flip[2,gate:}g\ttt{.p]}}$
-  \end{enumerate}
-\end{definition}
-Thus, if a value is in correlation with a wire label, it is either
-correlated with the wire label or with the inverse of both the key and
-pointer bits. For subsequent discussion we make the following addition
-to the codebase:
-$$
-\ttt{invert(\{ k = bk; p = bp \}) \{ \{ k = not bk; p = not bp \} \}}
-$$
-Gate certification establishes separation of input labels from the
-gate table and output value, and correlation of the output with
-the output wire label under any input wire value combination.
-\begin{definition}[Gate Certification]
-  \label{definition-gate-certification}
-Let $\mathit{gate}$ be a gate function. For $1 \le i \le 4$  let:
-  $$
-  \config{\varnothing}{\ttt{v[0,out] := } e_i}\redxs\config{\prog_i}{\ttt{()}}
-  $$
-  where
-  \begin{enumerate}[\hspace{5mm}$e_1 \defeq$]
-    \item $\mathit{gate}\ttt{(c,a,b,owl(a),owl(b));}$
-    \item $\mathit{gate}\ttt{(c,a,b,owl(a),invert(owl(b)));}$
-    \item $\mathit{gate}\ttt{(c,a,b,invert(owl(a)),owl(b));}$
-    \item $\mathit{gate}\ttt{(c,a,b,invert(owl(a)),invert(owl(b)))}$
-  \end{enumerate}
-  Define:
-  $${\small X \defeq \ttt{\{ flip[2,gate:a.k], flip[2,gate:a.p], flip[2,gate:b.k], flip[2,gate:b.p] \}}}$$
-  Then $\mathit{gate}$ is \emph{certified} iff for all $1 \le i \le 4$:
-  $$
-  \sep{\progtt(\prog_i)}{X}{\vdefs(\prog_i)} 
-  \quad \text{and} \quad \gc{\progd(\prog_i)}{\vx{0}{out}}{\ttt{owl(c)}}
-  $$
-\end{definition}
-Encode certification establishes correlation of the encoded input values with
-the secret wire labels, and separation of client 2's secret from client 1's views. 
-\begin{definition}[Encode Certification]
-  Let $\mathit{encode}$ be an encoding function, and let:
-  $$
-  \config{\varnothing}{\ttt{encode(s1,s2)}} \redxs
-  \config{\prog}{\ttt{\{wv1=}v_1\ttt{;wv2=}v_2\ttt{\}}}
-  $$
-  Then $\mathit{encode}$ is \emph{certified} iff each of the following conditions hold:
-  \begin{enumerate}[\hspace{5mm}i.]
-  \item $\gc{\progd(\prog,v_1)}{\itv}{\ttt{owl(s1)}} \quad \text{and} \quad
-    \gc{\progd(\prog,v_2)}{\itv}{\ttt{owl(s2)}}$
-  \item $\sep{\progd(\prog,v_2)}{\{\ttt{s[2,s2]\}}}{(\vdefs(\prog) \cup \itv)}$
-  \end{enumerate}
-\end{definition}
-Decode certification establishes separation of the input wire
-label from the decoding tables under any input value condition. 
-\begin{definition}[Decode Certification]
-  Let $\mathit{decode}$ be a decoding function, and let:
-  $$
-  \config{\varnothing}{\mathit{decode}(\ttt{c}, \ttt{owl(c)})}
-  \redxs \config{\prog_1}{\be_1}
-  \quad \text{and} \quad
-  \config{\varnothing}{\mathit{decode}(\ttt{c}, \ttt{invert(owl(c))})}
-  \redxs \config{\prog_0}{\be_0}
-  $$
-  Then $\mathit{decode}$ is certified iff:
-  $$
-  \sep{\progtt(\prog_i)}{\ttt{\{flip[2,gate:c.k],flip[2,gate:c.p]\}}}{\vdefs(\prog_i)}
-  $$
-  %Then $\mathit{decode}$ is certified iff both of the following conditions hold:
-  %\begin{enumerate}[\hspace{5mm}i.]
-  %\item $\progd(\prog_1,\be_1)(\{ \itv \mapsto 1 \}) = 1$ and
-  %  $\progd(\prog_0,\be_0)(\{ \itv \mapsto 0 \}) = 1$
-  %\item $\sep{\progtt(\prog_i)}{\ttt{\{flip[2,gate:c.k],flip[2,gate:c.p]\}}}{\vdefs(\prog_i)}$
-  %\end{enumerate}
-\end{definition}
-We have certified each of the YGC components detailed above. 
-\begin{lemma}
-  \label{lemma-certification}
-  Each of $\ttt{decode}$, $\ttt{encode}$, and $\ttt{andgg}$ are certified.
-\end{lemma}
-
-\subsubsection{Correctness of Certified Components}
-\label{section-composition-metatheory}
-
-To demonstrate correctness of certification, we need to show that
-isolated component certificates preserve relevant properties
-when integrated into larger programs. This resonates with
-\emph{framing} as developed in \cite{barthe2019probabilistic}.
-In the following we will use $f$ to range over flips.
-First, we show that if a flip $f$ is independent from the views
-in a protocol (as is the case with input labels used to
-encode a garbled table), then separation from views that
-use $f$ (e.g., in an output label) is preserved. 
-\begin{lemma}[Wire Framing]
-  \label{lemma-wire-framing}
-  If $\sep{\progtt(\prog)}{\{f\}}{\vdefs(\prog)}$ for flip $f$ and
-  $\vars(\prog) \cap \vars(\prog') = \{ f \}$ then
-  $\sep{\progtt(\prog';\prog)}{\vdefs(\prog')}{\vdefs(\prog)}$.
-\end{lemma}
-Next, we show that output correlations (e.g., between output wire values and
-labels) are preserved by substitution of correlated values for flips (e.g.,
-between input wire values and labels).
-\begin{lemma}[Substitution]
-  \label{lemma-substitution}
-  If $\vc{\progtt(\prog,\be)}{\itv}{f}$ and $\vc{\progtt(\prog',\be')}{\itv}{f'}$
-  and $\sep{\progtt(\prog';\prog)}{\vdefs(\prog')}{\vdefs(\prog)}$
-  then $\vc{\progtt(\prog';\prog,\be[\be'/f'])}{\itv}{f}$.
-\end{lemma}
-Our wire framing and substitution Lemmas then allow us to generalize
-certification of gates and encoding to arbitrary circuit contexts,
-assuming that outputs are not wired to multiple inputs \cite{tate2003garbled,nieminen2023breaking}.
-\begin{lemma}[Gate Composition]
-  \label{lemma-compose}
-Given well-typed $\prog_1;E[\mathit{gate}(g,g_1,g_2,v_1,v_2)]$ and certified $\mathit{gate}$ where:
-$$
-\config{\prog_1}{E[\mathit{gate}(g,g_1,g_2,v_1,v_2)]} \redxs \config{\prog_1;\prog_2}{E[v]}
-$$
-If the following preconditions hold:
-\begin{enumerate}
-\item $\{ g_1 \} \cap \{ g_2 \} \cap \wired(\prog_1) = \varnothing$ and $g \not\in \prog_1$
-\item $\gc{\progd(\prog_1)}{v_1}{\owl(g_1)}$ and $\gc{\progd(\prog_1)}{v_2}{\owl(g_2)}$
-\end{enumerate}
-then we have as postconditions:
-\begin{enumerate}
-  \item $\gc{\progd(\prog_1;\prog_2)}{v}{\owl(g_2)}$
-  \item $\sep{\progd(\prog_1;\prog_2,v)}{\vdefs(\prog_1)}{\vdefs(\prog_2) \cup \{\itv\}}$
-\end{enumerate}
-\end{lemma}
-\begin{proof}
-  Given that $g_1$ and $g_2$ are distinct and not wired in $\prog_1$
-  we are assured that $\ttt{owl}(g_1)$ and $\ttt{owl}(g_1)$ are in
-  independent uniform distributions, and given that $g \not\in
-  \prog_1$ we are assured that $\ttt{owl}(g)$ contains entirely fresh
-  flips. Thus, precondition 2, Lemma \ref{lemma-wire-framing}, and
-  Definition \ref{definition-gate-certification} imply postcondition
-  2.  Furthermore, since gate certification guarantees that
-  correlation with $\ttt{owl}(g)$ is guaranteed under all possible gate
-  input wire values, precondition 2 and Lemma \ref{lemma-substitution}
-  guarantee postcondition 2.
-\end{proof}
-
-\begin{lemma}[Decode Composition]
-  \label{lemma-decode}
-Given well-typed $\prog_1;E[\mathit{decode}(g,v)]$ and certified $\mathit{decode}$ where 
-$$
-\config{\prog_1}{E[\ttt{decode}(g,v)]} \redxs \config{\prog_1;\prog_2}{E[\be]}
-$$
-If the following preconditions hold:
-\begin{enumerate}
-\item $g \not\in \cap \wired(\prog_1)$
-\item $\gc{\progd(\prog_1)}{v}{\owl(g)}$
-\end{enumerate}
-then we have as postcondition:
-\begin{enumerate}
-  \item $\sep{\progd(\prog_1;\prog_2)}{\vdefs(\prog_1)}{\vdefs(\prog_2)}$
-\end{enumerate}
-\end{lemma}
-Given the above, we can establish that any circuit built with certified
-components is passive secure. 
-\begin{theorem}
-  \label{theorem-ygc-NIMO}
-  If $e$ is a well-typed YGC circuit definition using certified components
-  where no gate output is wired more than once, and $\config{\varnothing}{e}
-  \redxs \config{\prog}{\varnothing}$, then $\NIMO(\prog)$. 
-\end{theorem}
-
-\begin{proof}
-  Let $\iov(\prog) = S \cup (V \cup \outv)$. Noting that any
-  encoding necessarily uses entirely fresh variables by Definition and
-  type safety, by applying Lemmas \ref{lemma-compose} and
-  \ref{lemma-decode} and the frame rule of
-  \cite{barthe2019probabilistic} we obtain:
-  $$
-  \sep{\progd(\prog)}{(S_{\{2\}}\cup\{\outv\})}{V}
-  $$
-  Noting that the garbler (client 2) has no views, this means that
-  the garbler's secrets are independent of the evaluator's views. The
-  result follows by Definition \ref{definition-NIMO}.
-\end{proof}
-
-\begin{fpfig}[t]{YGC copy gate definitions.}{fig-ygc-copy}
 {\footnotesize
 \begin{verbatimtab}
-      sharetab2(gid, tid, k, p, b)
-      {
-        let r1 = k xor b in
-        let r0 = (not k) xor (not b) in
-        v[1,gate: || gid || tid || 1] := select(p,r1,r0);
-        v[1,gate: || gid || tid || 2] := select(not p,r1,r0);
+  owl(gid) { { k = flip[2,"gate:" || gid || ".k"]; p = flip[2,"gate:" || gid || ".p"] } }
+
+  encode(sa, sb) {
+    let owl1 = owl(sa) in
+    let owl2 = owl(sb) in
+    v[1,gate: || sa || 1.k] := OT(s[1,sa],owl1.k,(not owl1.k));
+    v[1,gate: || sa || 1.p] := OT(s[1,sa],owl1.p,(not owl1.p));
+    v[1,gate: || sb || 2.k] := select(s[2,sb],owl2.k,(not owl2.k));
+    v[1,gate: || sb || 2.p] := select(s[2,sb],owl2.p,(not owl2.p));
+    { wv1 = { k = v[1,"gate:" || sa || "1.k"]; p = v[1,"gate:" || sa || "1.p"] };
+      wv2 = { k = v[1,"gate:" || sb || "2.k"]; p = v[1,"gate:" || sb || "2.p"] } }
   }
 
-      copygate(ca,cb,g)
-      {
-        let wl = owl(g) in
-        let owl1 = owl(ca) in
-        let owl2 = owl(cb) in
-        sharetab2(ca,tt,wl.k,wl.p,owl1.k); sharetab2(ca,pt,wl.k,wl.p,owl1.p);
-        sharetab2(cb,tt,wl.k,wl.p,owl2.k); sharetab2(cb,pt,wl.k,wl.p,owl2.p)
+  andgg(g, ga, gb, wva, wvb) { andgate(g, ga, gb); evalgate(g, wva, wvb) }
+
+  decode(g, wv) { garbledecode(wv.g); evaldecode(v) }
+\end{verbatimtab}
 }
 
-      evalcopy(ca,cb,wv)
-      {
-        let wv1k = wv.k xor select(wv.p,v[1,gate: || ca || tt1], v[1,gate: || ca || tt2]) in
-        let wv1p = wv.k xor select(wv.p,v[1,gate: || ca || pt1], v[1,gate: || ca || pt1]) in
-        let wv2k = wv.k xor select(wv.p,v[1,gate: || cb || tt1], v[1,gate: || cb || tt2]) in
-        let wv2p = wv.k xor select(wv.p,v[1,gate: || cb || pt1], v[1,gate: || cb || pt1]) in
-        { wv1 = { k = wv1k; p = wv1p }; wv2 = { k = wv2k; p = wv2p } }  
+{\footnotesize
+\begin{verbatimtab}
+  let secrets = encode("s1","s2") in v[0,output] := decode(andgg("0", secrets.wv1, secrets.wv2)) \end{verbatimtab}
 }
 
-      copy(ca,cb,g,wv) { copygate(ca,cb,g); evalcopy(ca,cb,wv) } 
+{\footnotesize
+\begin{verbatimtab}
+  owl    :   string(gid)
+
+  encode :   string(sa) * string(sb)
+
+  andgg  :  string(g) * { k : bool[1]; p : bool[1] } *  { k : bool[1]; p : bool[1] }
+
+  decode :  string(g) * { k : bool[1]; p : bool[1] }
 \end{verbatimtab}
 }
-\end{fpfig}
-
-\subsection{Building and Extending Automatically Secure Circuits}
-\label{section-composition-copy}
-
-Theorem \ref{theorem-nimo} means that any well-typed program using
-$\ttt{encode}$, $\ttt{andgg}$, and $\ttt{decode}$ will generate a
-circuit that is passive secure, provided that gate outputs are not
-used more than once. To add an $\ttt{or}$ gate to our library, we can
-use $\ttt{andgg}$ as a template, in fact simple modification of
-$\ttt{andtable}$ is all that would be needed. Thus, this library
-embodies a Fairplay-like language where well-typedness of metaprograms
-guarantees safety and security of generated protocols.
-
-The requirements that output wire labels are not used more than once
-reflects a known result, that naive copy gate output is unsound in
-YGC- i.e., passive security fails. In fact, we are unable to
-automatically verify $\gNIMO$ for $\ttt{andgg}$ without the assumption
-that input wire lables are in independent uniform distributions. We
-can capture the necessary fix as in Figure \ref{fig-ygc-copy}. The
-garbler creates a copy gate with two new output wire labels in
-independent uniform distributions (in \ttt{copygate}), and permutes
-them using the input wire label pointer after encrypting with the
-input wire label keys. The evaluator then recovers the encrypted
-output wire values during evaluation (in \ttt{evalcopy}). Given an
-$\ttt{orgg}$ gate defined as outlined abotve, we can then use
-secure $\ttt{copy}$ as follows: 
-\begin{verbatimtab}
-  let s = encode(s1,s2) in
-  let s1c = copy(c1s1,c2s1,s1,s.wv1) in
-  let s2c = copy(c1s2,c2s2,s2,s.wv2) in
-  let wv1 = andgg(1,c1s1,c1s2,s1c.wv1,s2c.wv1) in
-  let wv2 = orgg(2,c2s1,c2s2,s1c.wv2,s2c.wv2) in
-  decode(3,andgg(3,1,2,wv1,wv2))
-\end{verbatimtab}
-These definitions of $\ttt{copy}$ and $\ttt{orgg}$ have been
-automatically verified as in Section \ref{section-pre-post} for
-$\ttt{andgg}$. In general, any new gate form- such as optimized gates
-\cite{XXX}- can be verified automatically in a similar way.
-
-While the copy gate supports secure programming patterns, security
-still relies on programmer discipline, i.e., not wiring a gate more
-than once. However, we can keep track of wired gates in type effects,
-similar to the manner in which we keep track of defined views- type
-dependence in our system is able to accurately precisely track
-gate identifiers. This is also reminiscent of the use of linearity
-in other related systems for enforcing obliviousness \cite{darais2019language}.

### 2023-07-10 popl24/ygc.tex
checkpoint

@@
-\begin{fpfig}[t]{Yao's Garbled Circuits, selected auxiliary functions.}{fig-ygc-aux}
+\begin{fpfig}{Yao's Garbled Circuits, selected auxiliary functions.}{fig-ygc-aux}
 {\footnotesize
@@
 
-\begin{fpfig}[t]{Yao's Garbled Circuits, garbled gates and evaluation code.}{fig-ygc-gates}
+\begin{fpfig}{Yao's Garbled Circuits, garbled gates and evaluation code.}{fig-ygc-gates}
 {\footnotesize
@@
 
-\begin{fpfig}[t]{Yao's Garbled Circuits, selected gate library type annotations.}{fig-ygc-types}
+\begin{fpfig}{Yao's Garbled Circuits, selected gate library type annotations.}{fig-ygc-types}
 {\footnotesize
@@
 %  }
+
+\section{YGC Code in Detail}
+
+In Figures \ref{fig-ygc-aux} and \ref{fig-ygc-gates} we define a
+codebase for Yao's garbled circuits (YGC). This definition follows the
+\emph{point-and-permute} method described in \cite{evans2018pragmatic}
+and elsewhere, to which the reader is referred for more in-depth discussion.
+In this implementation client 2 is the \emph{garbler} and
+client 1 is the \emph{evaluator}. The garbler builds the garbled
+tables and shares them with the evaluator, who then evaluates
+the gate in an oblivious fashion until the final public output is
+generated through decryption. This definition is well-typed,
+with input type annotations for top-level functions listed in
+Figure \ref{fig-ygc-types}. Well-typed programs using these
+libraries are therguaranteed to yield safe $\minifed$
+programs.

### 2023-07-10 popl24/metalang.tex
checkpoint

@@
 
-When we consider the examples of GMW and YGC in detail below, our focus will be
-on developing a codebase that can be used to define arbitrary
-circuits, i.e., complete and concrete protocols. Since strings and
-identifiers can be constructed manually, and expressions can occur
-inside assignments and boolean expression forms, function definitions
-can generalize over $\fedprot$-level patterns to obtain composable
-program units. As a simple example, consider 3 party secret
-sharing as illustrated in Example \ref{example-he}. We can
-define a function $\ttt{share3}$ that abstracts the process
-of splitting a given client's secret into 3 separate shares.
+When we consider the examples of GMW and YGC in detail below, our
+focus will be on developing a codebase that can be used to define
+arbitrary circuits, i.e., complete and concrete protocols. Since
+strings and identifiers can be constructed manually, and expressions
+can occur inside assignments and boolean expression forms, function
+definitions can generalize over $\fedprot$-level patterns to obtain
+composable program units. \cnote{Cur the following example to save
+  space?.}  As a simple example, consider 3 party secret sharing as
+illustrated in Example \ref{example-he}. We can define a function
+$\ttt{share3}$ that abstracts the process of splitting a given
+client's secret into 3 separate shares.
 \begin{example} \label{example-share3} The function $\ttt{share3}$ 
@@
 
-{\footnotesize
-\begin{verbatimtab}
-  owl    :   string(gid)
-
-  encode :   string(sa) * string(sb)
-
-  andgg  :  string(g) * { k : bool[1]; p : bool[1] } *  { k : bool[1]; p : bool[1] }
-
-  decode :  string(g) * { k : bool[1]; p : bool[1] }
-\end{verbatimtab}
-}
-
 In Figures \ref{fig-ygc-aux} and \ref{fig-ygc-gates} we define a
@@
 $\ttt{s2}$ from client's 1 and 2 respectively. 
+{\footnotesize
 \begin{verbatimtab}
-  andgate(0,s1,s2);
-  garbledecode(0);
-  let secrets = encode(s1,s2) in
-  v[0,output] := decode(evalgate(0, secrets.wv1, secrets.wv2))
-\end{verbatimtab}
+  let secrets = encode("s1","s2") in v[0,output] := decode(andgg("0", secrets.wv1, secrets.wv2)) \end{verbatimtab}
+}
+%\begin{verbatimtab}
+%  andgate(0,s1,s2);
+%  garbledecode(0);
+%  let secrets = encode(s1,s2) in
+%  v[0,output] := decode(evalgate(0, secrets.wv1, secrets.wv2))
+%\end{verbatimtab}
 \end{example}

### 2023-07-10 popl24/composition.tex
checkpoint

--- 2023-07-07 popl24/composition.tex
@@
-\section{Component Certification and Composition}
-\label{section-composition}
+\section{Automating Verification of Circuit Security}
 
-\input{gmw}
-
-The Yao's Garbled Circuits (YGC) and GMW protocols are examples of a
-common idiom in MPC protocol design- arbitrarily large \emph{circuits}
-are built up from smaller boolean or arithmetic \emph{gates} that
-operate on encodings of input secrets. These circuits can become quite
-large in practice \cite{kreuter2012billion}, but are built from
-well-defined components. Brute force verification techniques as
-described in Section \ref{section-bruteforce} are not scalable to
-large circuits, but in this Section we will demonstrate that correctly
-designed circuit components enjoy compositional properties
-that can be automatically certification on small program components
-using brute force methods. We prove that these properties preserve
-invariants throughout circuit definition that yield $\NIMO$ and thus
-scale to arbitrary programs. Furthermore, the certifications can be
-used on arbitrary gate definitions, allowing new gate implementations
-to be added soundly to libraries.
-
-In this Section we develop and use implementations of both 2-party GMW
-and YGC to illustrate our approach and demonstrate its generality.
 Main results are in Theorems \ref{theorem-gmw-NIMO} and
@@
 
-\subsection{Compositional Properties of GMW}
-
-The GMW protocol uses secret sharing to represent data flowing through
-circuits. In the 2-party case, clients 1 and 2 each share their input
-secrets, and use those shares to represent inputs to gates. Outputs
-are also represented as shares. We refer to the pair of shares
-representing any particular value as a \emph{wire value}, and
-we represent them via records of the form
-$
-\ttt{\{c1 = } v_1\ttt{;c2 = } v_2\ttt{\}} 
-$
-where $v_1$ and $v_2$ are client 1 and 2's shares respectively.
-
-For full details of the GMW protocol the reader is referred to
-\cite{evans2018pragmatic}. Our implementation libary is shown in
-Figure \ref{fig-gmw}, with type signatures for the library functions
-shown in Figure \ref{fig-gmw-types}. We show the And gate since it is
-an interesting component. The Figure includes the
-following top-level functions:
-\begin{itemize}
-\item \ttt{encodegmw}: This function encodes client 1's and client 2's
-  secret bits called $\ttt{s[1,s1]}$ and $\ttt{s[1,s2]}$ into two
-  distinct wire values (pairs of shares).
-\item \ttt{andgmw}: This function defines the gate $\ttt{g}$, the
-  identifier $\ttt{g}$ being used to distinguish randomness used
-  within.  In our version client 1 builds the output table (using
-  \ttt{andtablegmw})and transfers the correct output share to client 2
-  using 1-out-of-4 OT as per standard GMW protocol.
-\item \ttt{decode}: This function decodes and publishes a wire value
-  by $\ttt{xor}$ing the shares. Note that this requires both client 1
-  and 2 to publicize their shares.
-\end{itemize}
-\begin{example}
-  \label{example-gmw-andcircuit}
-The following program uses our GMW library to define
-a circuit with a single And gate and input secrets $\ttt{s1}$ and
-$\ttt{s2}$ from client's 1 and 2 respectively. 
-\begin{verbatimtab}
-  let ss = encodegmw(s1,s2) in
-  v[0,output] := decode(andgmw(0,ss.shares1,ss.shares2))
-\end{verbatimtab}
-\end{example}
-
-\subsubsection{Gate Certification}
-
-Our certification techniques are defined generally wrt implementations
-of input encoding, decoding, and internal gates. In the following we
-will refer to arbitrary decode, encode, and gate functions with the
-restriction that any function in each of these categories has the same
-valid type signature as $\ttt{decodegmw}$, $\ttt{encodegmw}$, and
-$\ttt{andgmw}$ respectively. When gates are used in ciruits they are
-parameterized by wire values-- in certification we will use carefully
-chosen witness parameters.
-
-When we consider internal components and hence intensional properties
-of circuits, we will consider behaviors with respect to particular bits
-of randomness. Hence we will consider basic program distributions. 
-\begin{definition}
-  We write $\progtt(\prog)$ to denote the \emph{basic distribution} of
-  $\prog$ as defined in Definition \ref{def-progd}. For any program $\prog =
-  (\eassign{v_1}{\be_1};\ldots;\eassign{v_n}{\be_n})$, define
-  $\vdefs(\prog) \defeq \{ v_1,\ldots,v_n \}$.
-\end{definition}
-We also need to formalize notions of probabilistic independence, aka
-\emph{separation}, and correlation. We borrow the symbols $*$ and $\sim$
-from \cite{barthe2019probabilistic} with the same denotations.
-\begin{definition}
-  We write $\vc{\pmf}{x}{y}$ iff $\pmf(\{ x \mapsto 0\}\ |\ \{ y \mapsto 0 \}) =
-  \pmf(\{ x \mapsto 1\}\ |\ \{ y \mapsto 1 \}) = 1$.
-  We write $\sep{\pmf}{Y}{Z}$ iff for all
-    $\store \in \mems(X \cup Y)$:
-  $$\margd{\pmf}{Y \cup Z}(\store) =
-  \pmf(\store_Y) * \pmf(\store_Z)$$ 
-\end{definition}
-
-\begin{fpfig}[t]{GMW library type annotations.}{fig-gmw-types}
-{\footnotesize
-  \begin{verbatimtab}
-   encodegmw   : string(gid) * string(gid)
-
-   andtablegmw : { k = bool[i]; p = bool[i] }
-
-   andgmw      : string(gid) *  { c1 = bool[1]; c2 = bool[2] } * { c1 = bool[1]; c2 = bool[2] }
-
-   decodegmw   : { c1 = bool[1]; c2 = bool[2] }  \end{verbatimtab}
-}
-\end{fpfig}
+\subsection{GMW Component Certification}
 
@@
 
-\subsubsection{Gate Invariants and Circuit Security}
+\subsection{GMW Gate Invariants and Circuit Security}
 We can show that composition maintains an important invariant in a
@@
 decoding.
-
-We observe the following properties of separation which are borrowed from prior work
-\cite{barthe2019probabilistic} and which will be used frequently in proofs. 
-\begin{lemma}
-  \label{lemma-separation}
-  The following properties hold:
-  \begin{enumerate}
-  \item $\sep{\pmf}{Y}{Z}$ iff $\sep{\pmf}{Z}{Y}$
-  \item $\vc{\pmf}{x}{y}$ if  $\vc{\pmf}{y}{x}$
-  \item $\vc{\pmf}{x}{y}$ and $\vc{\pmf}{y}{z}$ imply $\vc{\pmf}{x}{z}$
-  \item $\sep{\pmf}{X}{(Y \cup Z)}$ implies $(\sep{\pmf}{X}{Y}$ and $\sep{\pmf}{X}{Z})$
-  \item $(\sep{\pmf}{X}{Y}$ and $\sep{\pmf}{(X \cup Y)}{Z})$ implies $\sep{\pmf}{X}{(Y \cup Z)}$ 
-  %\item If $\prog_1;\prog_2$ is safe and $\vars(\prog_1) \cap \vars(\prog_2) = \varnothing$
-  %  then $\sep{\progtt(\prog_1;\prog_2)}{\vars(\prog_1)}{\vars(\prog_2)}$.
-  \end{enumerate}
-\end{lemma}
-The following property also follows by results in \cite{barthe2019probabilistic} and will
-be useful to make constructions that demonstrate variable separation.
-\begin{lemma}
-  \label{lemma-sepjoin}
-  $\sep{\progtt(\prog)}{X}{Y}$ iff for all 
-  $\store^1, \store^2 \in \runs(\prog)$ there exists
-  $\store \in \runs(\prog)$ with
-  $\store^1_{X} \cap \store^2_{Y} \subseteq \store$.
-\end{lemma}
-
-%The following Lemma allows us to separate variables in programs that may not occur
-%in the same subprogram, and hence consider gate certification properties in various
-%contexts. It is reminiscent of the \emph{frame} rule of \cite{barthe2019probabilistic}.
-%\begin{lemma}
-%  \label{lemma-gmw-frame}
-%  If $\sep{\progtt(\prog_1;\prog_2)}{Y}{V}$ for $Y = \vars(\prog_1) \cap \vars(\prog_2)$ and
-%  $V \in \vdefs(\prog_2)$ and $\sep{\progtt(\prog_1)}{X}{Y}$ for
-%  $X \not\in \vars(\prog_2)$ then $\sep{\progtt(\prog_1;\prog_2)}{X}{V}$. 
-%\end{lemma}
-%\begin{proof}
-%  By assumption and Lemma \ref{lemma-sepjoin}, for all
-%  $\store^1,\store^2 \in \runs(\prog_1;\prog_2)$ we have $\store^1_Y
-%  \cap \store^2_V \subseteq \store_a$ for some $\store_a \in
-%  \runs(\prog_1;\prog_2)$ and $\store^1_X \cap \store^2_Y \subseteq
-%  \store_b$ for some $\store_b \in \runs(\prog_1;\prog_2)$.  Thus, since
-%  we assume $X \not\in \vars(\prog_2)$ we have $\store_a \cap \store_b
-%  \in \runs(\prog_1;\prog_2)$ and $\store^1_X \cap \store^2_V =
-%  \store_a \cap \store_b$, obtaining the result by
-%  \ref{lemma-sepjoin}.
-%\end{proof}
-
-The next ``noninterference'' Lemma is important to prove the desired invariance
-preservation property of GMW gates, given certification.
-\begin{lemma}[Noninterference]
-  \label{lemma-noninterference}
-  Given $\prog_1;\prog_2$ and $X = \iov(\prog_2) - \vdefs(\prog_2)$ and
-  $Y \subseteq \vdefs(\prog_2)$. If $\sep{\progd(\prog_1;\prog_2)}{X}{Y}$
-  then $\sep{\progd(\prog_1;\prog_2)}{\iov(\prog_1)}{Y}$.
-\end{lemma}
-
-The next two Lemmas allow us to use the properties of gates established by
-certification and replace the ``dummy'' parameters used there with actual
-wire values in full circuit contexts. Intuitively we show that substitution
-for dummy flips preserves separation. 
-In these results and subsequently we will use $f$ to range over flips.
-\begin{lemma}
-  \label{lemma-presub}
-  Given $\be_1$, $\be_2$, and $i$ where $\vars(\be_1) \cap
-  \vars(\be_2) = \varnothing$ and $\lcod{\store_1,\be_1}{i} = \beta$
-  with $\store_1(x) = \lcod{\store_2,\be_2}{i}$ and
-  $\dom(\store_1) = \vars(\be_1)$ and $\dom(\store_2) = \vars(\be_2)$.
-  Then $\lcod{\store_1\cap\store_2,\be_1[\be_2/x]}{i} = \beta$.
-\end{lemma}
-\begin{proof}
-By straightforward structural induction on $\be$.
-\end{proof}
-
-\begin{lemma}[Substitution$*$]
-  \label{lemma-substitution}
-  If $\sep{\progtt(\prog_2)}{\{ f \}}{X}$ and
-  $\prog_1;\prog_2[\be/f]$ is safe with $\vars(\prog_1,\be) \cap
-  X = \varnothing$ then
-  $\sep{\progtt(\prog_1;\prog_2[\be/f])}{\vars(\be)}{X}$.
-\end{lemma}
-\begin{proof}
-  Suppose on the contrary it was not the case that
-  $\sep{\progtt(\prog_1;\prog_2[\be/f])}{\vars(\be)}{X}$.
-  Then by Lemma \ref{lemma-sepjoin} there exists $\store_1,\store_2
-  \in \runs(\prog_1;\prog_2[\be/f])$ with no $\store \in
-  \runs(\prog_1;\prog_2[\be/f])$ such that $\store^1_{\vars(\be)}
-  \cap \store^2_X \subseteq \store$.  But also by
-  assumption and Lemma \ref{lemma-sepjoin} for any $\beta$ there
-  exists $\store' \in \runs(\prog_2)$ with $\{ f \mapsto \beta \} \cap
-  \store^2_X \subseteq \store'$. So in particular, we
-  have $\{ f \mapsto \lcod{\store^1_{\vars(\be)},\be}{i} \} \cap
-  \store^2_X \subseteq \store'$ for any $i$. This, the
-  assumption $\vars(\prog_1,\be) \cap \vars(\prog_2) = \varnothing$,
-  and application of Lemma \ref{lemma-presub} leads to the consequence
-  that there exists $\store \in \runs(\prog_1;\prog_2[\be/f])$ with
-  $\store \supseteq \store^1_{\vars(\be)} \cap
-  \store^2_X$ given the assumption $\vars(\prog_1,\be)
-  \cap \vars(\prog_2) = \varnothing$, which is a contradiction.
-\end{proof}
 
@@
 the rest of the program.
-\begin{lemma}[GMW Encode Invariant]
+\begin{restatable}[GMW Encode Invariant]{lemma}{gmwencode}
   \label{lemma-gmw-encode}
@@
   \end{enumerate}
-\end{lemma}
-
-\begin{proof}
-  By assumptions of well-typedness we have:
-  $$
-  \begin{array}{rcl}
-    v &=&  \ttt{\{shares1 = }v_1\ttt{; shares2 = }v_2\ttt{\}}\\
-    v_1 &=& \ttt{\{ c1 = v[1,s1out]; c2 = v[2,s1out] \}}\\
-    v_2 &=& \ttt{\{ c1 = v[1,s2out]; c2 = v[2,s2out] \}}
-  \end{array}
-  $$
-  And by Definition \ref{definition-gmwencode-certification} we have:
-  \begin{mathpar}
-    \sep{\progtt(\prog)}{\ttt{\{s[1,s1]\}}}{\ttt{\{v[2,s1out],v[2,s2out]\}}}
-
-    \sep{\progtt(\prog)}{\ttt{\{s[2,s2]\}}}{\ttt{\{v[1,s1out],v[1,s2out]\}}}
-  \end{mathpar}
-  Since we assume that secrets are in uniform and independent marginal
-  distributions a priori, and $\vars(\prog_1) \cap \vars(\prog_2) =
-  \varnothing$ by condition (iii) of Definition 
-  \ref{definition-gmwencode-certification} and assumptions of well-typedness, 
-  the result follows by preconditions and Lemma \ref{lemma-separation}.
-\end{proof}
+\end{restatable}
 
@@
 
-\subsection{Compositional Properties of YGC}
-\label{section-composition-ygc}
-
-\input{ygc}
-
-In Figures \ref{fig-ygc-aux} and \ref{fig-ygc-gates} we define a
-codebase for Yao's garbled circuits (YGC). This definition follows the
-\emph{point-and-permute} method described in \cite{evans2018pragmatic}
-and elsewhere, to which the reader is referred for more in-depth discussion.
-In this implementation client 2 is the \emph{garbler} and
-client 1 is the \emph{evaluator}. The garbler builds the garbled
-tables and shares them with the evaluator, who then evaluates
-the gate in an oblivious fashion until the final public output is
-generated through decryption. This definition is well-typed,
-with input type annotations for top-level functions listed in
-Figure \ref{fig-ygc-types}. Well-typed programs using these
-libraries are therefore guaranteed to yield safe $\minifed$
-programs.
-
-\emph{Wire labels} are fundamental to YGC, and essentially represent
-gate output values in an encrypted form. In our definition, wire
-labels are represented by records $\ttt{\{ k = }\beta_1\ttt{; p =
-}\beta_2\ttt{ \}}$, where $\ttt{k}$ is the \emph{key bit} and
-$\ttt{p}$ is the \emph{pointer bit}, and $\beta_1$ and $\beta_2$ are
-flips. Flips in each output wire label are owned by the garbler and
-are unique per gate by definition of their identifying string, and the
-representation of $0$ is the negation of $1$. For example, here is the
-representation of 1 and 0 respectively in the output wire label for a
-hypothetical gate 6:
-\begin{mathpar}
-  \ttt{\{ k = flip[2,gate:6.k]; p =  flip[2,gate:6.p]] \}}
-
-  \ttt{\{ k = not flip[2,gate:6.k]; p =  not flip[2,gate:6.p]] \}}
-\end{mathpar}
-The pointer bits in wire labels are used to select permuted rows in
-table garblings. The key bits are used to identify a unique key for
-table row in each garbled gate. Intuitively, if $\beta_1$ and
-$\beta_2$ are either key or pointer bits encoding 1 on two input wire
-labels to a binary gate, rows and keys in the gate are enumerated in
-the order:
-$$
-\neg\beta_1\neg\beta_2,\ \neg\beta_1\beta_2,\ \beta_1\neg\beta_2,\ \beta_1\beta_2
-$$
-
-In our implementation, gates are wired together using gate
-identifiers, which are strings $w$. Top-level functionality in Figures
-\ref{fig-ygc-aux} and \ref{fig-ygc-gates} includes the following:
-\begin{itemize}
-\item \ttt{andgate}: This defines a subprotocol for the garbler
-  to define a garbled gate $\ttt{gid}$ with input wires from gates
-  $\ttt{ga}$ and $\ttt{gb}$. The garbler generates keys and garbles
-  the rows in YGC fashion, them with client $1$ in
-  views in a standard form. For example, the view for
-  a hypothetical gate 6, row 2 garbled truth table is $\ttt{v[1,gate:6tt2]}$.
-  We note that garbled gates of other binary operators can be obtained with
-  replacement of $\ttt{andtable}$ with appropriate garbled table definitions. 
-\item \ttt{evalgate}: This defines a subprotocol for the evaluator to
-  evaluate gate $\ttt{gid}$ given input wire values $\ttt{wva}$ and
-  $\ttt{wvb}$.
-\item \ttt{garbledecode} and \ttt{evaldecode}: The former function
-  defines the garbler's protocol for encrypting the circuit
-  output from final gate $\ttt{gid}$, and the latter defines
-  the evaluator's output decryption protocol.
-\item \ttt{encode}: This defines the initial phase of the protocol,
-  where the evaluator receives the wire value from their own
-  secret $\sx{1}{sa}$ via $\ttt{OT}$, and the garbler communicates
-  the wire value for their own secret $\sx{2}{sb}$ directly.
-\end{itemize}
-\begin{example}
-  \label{example-andcircuit}
-The following program uses our YGC library to define
-a circuit with a single and gate and input secrets $\ttt{s1}$ and
-$\ttt{s2}$ from client's 1 and 2 respectively. 
-\begin{verbatimtab}
-  andgate(0,s1,s2);
-  garbledecode(0);
-  let secrets = encode(s1,s2) in
-  v[0,output] := decode(evalgate(0, secrets.wv1, secrets.wv2))
-\end{verbatimtab}
-\end{example}
-We have verified passive security of the $\fedprot$ protocol
-generated by this and other small circuits using the
-technique described in Lemma \ref{lemma-bruteforce-nimo}.
-But large circuits with thousands of gates would be
-intractable to verify with this method. In the next Section
-we discuss compositional methods to address this issue.
-
-\subsubsection{Gate Certification}
+\subsection{YGC Component Certification}
 \label{section-ygc-certification}
@@
 multiple inputs \cite{tate2003garbled,nieminen2023breaking}.
-\begin{lemma}[YGC Gate Invariant]
+\begin{restatable}[YGC Gate Invariant]{lemma}{ygcgate}
   \label{lemma-ygc-preservation}
@@
 \end{enumerate}
-\end{lemma}
-\begin{proof}
-  Given preconditions we have $v_1 =  \{ \ttt{k = }\be^1_{k}; \ttt{p = }\be^1_{p} \}$ and
-  $v_2 = \{ \ttt{k = }\be^2_{k}; \ttt{p = }\be^2_{p} \}$ for some
-  $\be^1_{k}$,$\be^2_{k}$,$\be^1_{p}$,$\be^2_{p}$ with correlations as per Definition
-  \ref{definition-gc}.
-  %\begin{eqnarray*}
-  %  v_1 &=& \{ \ttt{k = }\be^1_{k}; \ttt{p = }\be^1_{p} \} \\
-  %  v_2 &=& \{ \ttt{k = }\be^2_{k}; \ttt{p = }\be^2_{p} \}
-  %\end{eqnarray*}
-  Let $\prog$ be as defined in Definition \ref{definition-ygcgate-certification}.
-  We observe:
-  \begin{eqnarray*}
-    &\prog_2 = \\
-    &{\small \prog[\be^1_{k}/\ttt{flip[2,gate:a.k]}][\be^1_{p}/\ttt{flip[2,gate:a.p]}][\be^2_{k}/\ttt{flip[2,gate:b.k]}][\be^1_{p}/\ttt{flip[2,gate:b.p]}]}
-  \end{eqnarray*}
-  Given that $g_1$ and $g_2$ are distinct and not wired in $\prog_1$
-  we are assured that $\ttt{owl}(g_1)$ and $\ttt{owl}(g_1)$ are in
-  independent uniform distributions, and given that $g \not\in
-  \prog_1$ we are assured that $\ttt{owl}(g)$ contains entirely fresh
-  flips. Thus by condition (i) of Definition \ref{definition-ygcgate-certification} and
-  Lemma \ref{lemma-substitution} we have:
-  $$
-  \sep{\prog_1;\prog_2}{\vars(\be^1_{k},\be^2_{k},\be^1_{p},\be^2_{p})}{\vdefs(\prog_2)}
-  $$
-  Thus by Lemmas \ref{lemma-noninterference} and \ref{lemma-separation} we establish
-  postconditions (ii) and (iii).
+\end{restatable}
 
-  Also since $v_1$ and $v_2$ are correlated either positively or negatively with
-  $\ttt{owl}(g_1)$ and $\ttt{owl}(g_2)$ respectively by precondition (ii),
-  by Definition \ref{definition-gc}, precondition (i), and Lemma \ref{lemma-substitution-sim}
-  we establish postcondition (i), since Definition \ref{definition-ygcgate-certification}
-  requires gate output correlation with $\ttt{owl}(g)$ given any input valence conditions.
-\end{proof}
-
-\begin{lemma}[YGC Encode Invariant]
+\begin{restatable}[YGC Encode Invariant]{lemma}{ygcencode}
   \label{lemma-ygc-encode}
@@
 \end{enumerate}
-\end{lemma}
-\begin{proof}
-  By Definition \ref{definition-ygcencode-certification} we have:
-  \begin{mathpar}
-    \sep{\progtt(\prog_2)}{\ttt{\{s[1,s1]\}}}{\vdefs(\prog_2)_{\{2\}}}
-
-    \sep{\progtt(\prog_2)}{\ttt{\{s[2,s2]\}}}{\vdefs(\prog_2)_{\{1\}}}
-  \end{mathpar}
-  Since we assume that secrets are in uniform and independent marginal
-  distributions a priori, and $\vars(\prog_1) \cap \vars(\prog_2) =
-  \varnothing$ by condition (iv) of Definition
-  \ref{definition-ygcencode-certification} and assumptions of
-  well-typedness, conditions (ii-iii) follow by Lemmas
-  \ref{lemma-noninterference} and \ref{lemma-separation}. Also by
-  Definition \ref{definition-ygcencode-certification} we have
-  $\gc{\prog_2}{v_1}{s_1}$ and $\gc{\prog_2}{v_1}{s_1}$, so also
-  $\gc{\prog_1;\prog_2}{v_1}{s_1}$ and
-  $\gc{\prog_1;\prog_2}{v_1}{s_1}$ by condition (iv) of Definition
-  \ref{definition-ygcencode-certification} and Lemmas
-  \ref{lemma-noninterference} and \ref{lemma-separation}.
-\end{proof}
+\end{restatable}
 
@@
   in the sense that if the garbler generates a table with input wire label
-  $g_1$ then only $g_1$ is wired to that input. I'll modify the
-  typing to enforce this.}
-\begin{theorem}
+  $g_1$ then only $g_1$ is wired to that input. }
+\begin{restatable}{theorem}{ygcnimo}
   \label{theorem-ygc-NIMO}
@@
   \config{\prog}{\varnothing}$, then $\NIMO(\prog)$.
-\end{theorem}
-
-\begin{proof}
-  By assumptions of well-typedness we
-  have for some $\be_k$, $\be_p$, and $\be$:
-  $$
-  \config{\varnothing}{\eassign{\outv}{\ttt{decode}(g,e)}} \redxs
-  \config{\prog_1}{\eassign{\outv}{\ttt{decode}(g,\{ \ttt{k = }\be_k;  \ttt{p = }\be_p\})}}
-  \redxs \config{\prog_1;\prog_2}{\eassign{\outv}{\be}}
-  $$
-  Let $\iov(\prog_1) = S \cup V$.
-  By Lemmas \ref{lemma-ygc-encode} and \ref{lemma-ygc-preservation} we
-  have $\sep{\progd(\prog_1)}{S_{\{1\}}}{V_{\{2\}}}$ and
-  $\sep{\progd(\prog_1)}{S_{\{2\}}}{V_{\{1\}}}$.
-  Let $\prog$ be as defined in Definition \ref{definition-ygdecode-certification}.
-  We observe:
-  \begin{eqnarray*}
-    & \prog_2 = \\ 
-    & \prog[\be_k/\ttt{\{flip[2,gate:c.k]}][\be_p/\ttt{flip[2,gate:c.p]}]
-  \end{eqnarray*}
-  so also by  Definition \ref{definition-ygdecode-certification} and
-  Lemma \ref{lemma-substitution} we have:
-  $$
-  \sep{\progtt(\prog_1;\prog_2)}{\vars(\be_k,\be_p)}{\vdefs(\prog_2)}
-  $$
-  so, letting $\iov(\prog_1;\prog_2) = S \cup V'$ by
-  Lemmas \ref{lemma-noninterference} \ref{lemma-separation} we have
-  $\sep{\progd(\prog_1;\prog_2)}{S_{\{1\}}}{V'_{\{2\}}}$ and
-  $\sep{\progd(\prog_1;\prog_2)}{S_{\{2\}}}{V'_{\{1\}}}$.
-  Thus, letting $\prog' \defeq (\prog_1;\prog_2;\eassign{\outv}{\be})$, for all $\beta$:
-  $$\sep{\condd{\progd(\prog')}{S_{\{1\}} \cup V_{\{2\}}}{\{ \outv \mapsto \beta \}}}{S_{\{1\}}}{V_{\{2\}}}
-  \quad \text{and} \quad
-    \sep{\condd{\progd(\prog')}{S_{\{2\}} \cup V_{\{1\}}}{\{ \outv \mapsto \beta \}}}{S_{\{2\}}}{V_{\{1\}}}$$
-  thus for all $\store \in \mems(V_2 \cup \{ \outv \})$:
-  $$\condd{\progd(\prog')}{S_{\{1\}}}{\store}
-  = \condd{\progd(\prog')}{S_{\{1\}}}{\store_{\{\outv\}}}$$
-  and for all $\store \in \mems(V_1 \cup \{ \outv \})$:
-  $$\condd{\progd(\prog')}{S_{\{2\}}}{\store}
-  = \condd{\progd(\prog')}{S_{\{2\}}}{\store_{\{\outv\}}}$$
-  establishing the result by Lemma \ref{lemma-nimo} and Definition \ref{definition-NIMO}.
-\end{proof}
+\end{restatable}
 

### 2023-07-10 popl24/metalang.tex
checkpoint

@@
 
+\begin{fpfig}{Yao's Garbled Circuit component functions.}{fig-ygc-lib}
 {\footnotesize
@@
 
-  encode(sa, sb) {
+  encodeygc(sa, sb) {
     let owl1 = owl(sa) in
@@
 
-  andgg(g, ga, gb, wva, wvb) { andgate(g, ga, gb); evalgate(g, wva, wvb) }
+  andygc(g, ga, gb, wva, wvb) { andgate(g, ga, gb); evalgate(g, wva, wvb) }
 
-  decode(g, wv) { garbledecode(wv.g); evaldecode(v) }
+  decodeygc(g, wv) { garbledecode(g); evaldecode(wv) }
 \end{verbatimtab}
 }
+\end{fpfig}
 
-{\footnotesize
-\begin{verbatimtab}
-  let secrets = encode("s1","s2") in v[0,output] := decode(andgg("0", secrets.wv1, secrets.wv2)) \end{verbatimtab}
-}
-
-In Figures \ref{fig-ygc-aux} and \ref{fig-ygc-gates} we define a
-codebase for Yao's garbled circuits (YGC). This definition follows the
-\emph{point-and-permute} method described in \cite{evans2018pragmatic}
-and elsewhere, to which the reader is referred for more in-depth discussion.
-In this implementation client 2 is the \emph{garbler} and
-client 1 is the \emph{evaluator}. The garbler builds the garbled
-tables and shares them with the evaluator, who then evaluates
-the gate in an oblivious fashion until the final public output is
-generated through decryption. This definition is well-typed,
-with input type annotations for top-level functions listed in
-Figure \ref{fig-ygc-types}. Well-typed programs using these
-libraries are therguaranteed to yield safe $\minifed$
-programs.
+Our YGC implementation follows the \emph{point-and-permute} method
+described in \cite{evans2018pragmatic} and elsewhere, to which the
+reader is referred for more in-depth discussion.  In this
+implementation client 2 is the \emph{garbler} and client 1 is the
+\emph{evaluator}. The garbler builds the garbled tables and shares
+them with the evaluator, who then evaluates the gate in an oblivious
+fashion until the final public output is generated through
+decryption.
 
@@
 \end{mathpar}
-The pointer bits in wire labels are used to select permuted rows in
-table garblings. The key bits are used to identify a unique key for
-table row in each garbled gate. Intuitively, if $\beta_1$ and
-$\beta_2$ are either key or pointer bits encoding 1 on two input wire
-labels to a binary gate, rows and keys in the gate are enumerated in
-the order:
-$$
-\neg\beta_1\neg\beta_2,\ \neg\beta_1\beta_2,\ \beta_1\neg\beta_2,\ \beta_1\beta_2
-$$
-
 In our implementation, gates are wired together using gate
-identifiers, which are strings $w$. Top-level functionality in Figures
-\ref{fig-ygc-aux} and \ref{fig-ygc-gates} includes the following:
+identifiers, which are strings $w$. Top-level functionality
+shown in Figure \ref{fig-ygc-lib} includes:
 \begin{itemize}
+\item \ttt{owl}: This function returns the output wire label for any
+  gate identified by \ttt{gid}.
 \item \ttt{andgate}: This defines a subprotocol for the garbler
@@
   $\ttt{wvb}$.
+\item \ttt{andygc}: This composes the garbler and evaluator subprotocols
+  for construction and evaluation of an And gate. 
 \item \ttt{garbledecode} and \ttt{evaldecode}: The former function
@@
   output from final gate $\ttt{gid}$, and the latter defines
-  the evaluator's output decryption protocol.
-\item \ttt{encode}: This defines the initial phase of the protocol,
+  the evaluator's output decryption protocol. These are composed with \ttt{decodeygc}.
+\item \ttt{encodegmw}: This defines the initial phase of the protocol,
   where the evaluator receives the wire value from their own
@@
 \end{itemize}
+We provide more code details in Appendix \ref{section-ygc}, Figures
+\ref{fig-ygc-aux} and \ref{fig-ygc-gates}, and complete code in our
+online repository.
 \begin{example}
@@
 $\ttt{s2}$ from client's 1 and 2 respectively. 
-{\footnotesize
+{\small
 \begin{verbatimtab}
-  let secrets = encode("s1","s2") in v[0,output] := decode(andgg("0", secrets.wv1, secrets.wv2)) \end{verbatimtab}
+  let secrets = encode("s1","s2") in
+  v[0,output] := decode(andgg("0", secrets.wv1, secrets.wv2)) \end{verbatimtab}
 }

### 2023-07-10 popl24/proofs.tex
checkpoint

@@
 \end{proof}
+
+\gmwencode*
+
+\begin{proof}
+  By assumptions of well-typedness we have:
+  $$
+  \begin{array}{rcl}
+    v &=&  \ttt{\{shares1 = }v_1\ttt{; shares2 = }v_2\ttt{\}}\\
+    v_1 &=& \ttt{\{ c1 = v[1,s1out]; c2 = v[2,s1out] \}}\\
+    v_2 &=& \ttt{\{ c1 = v[1,s2out]; c2 = v[2,s2out] \}}
+  \end{array}
+  $$
+  And by Definition \ref{definition-gmwencode-certification} we have:
+  \begin{mathpar}
+    \sep{\progtt(\prog)}{\ttt{\{s[1,s1]\}}}{\ttt{\{v[2,s1out],v[2,s2out]\}}}
+
+    \sep{\progtt(\prog)}{\ttt{\{s[2,s2]\}}}{\ttt{\{v[1,s1out],v[1,s2out]\}}}
+  \end{mathpar}
+  Since we assume that secrets are in uniform and independent marginal
+  distributions a priori, and $\vars(\prog_1) \cap \vars(\prog_2) =
+  \varnothing$ by condition (iii) of Definition 
+  \ref{definition-gmwencode-certification} and assumptions of well-typedness, 
+  the result follows by preconditions and Lemma \ref{lemma-separation}.
+\end{proof}
+
+\ygcgate*
+
+\begin{proof}
+  Given preconditions we have $v_1 =  \{ \ttt{k = }\be^1_{k}; \ttt{p = }\be^1_{p} \}$ and
+  $v_2 = \{ \ttt{k = }\be^2_{k}; \ttt{p = }\be^2_{p} \}$ for some
+  $\be^1_{k}$,$\be^2_{k}$,$\be^1_{p}$,$\be^2_{p}$ with correlations as per Definition
+  \ref{definition-gc}.
+  %\begin{eqnarray*}
+  %  v_1 &=& \{ \ttt{k = }\be^1_{k}; \ttt{p = }\be^1_{p} \} \\
+  %  v_2 &=& \{ \ttt{k = }\be^2_{k}; \ttt{p = }\be^2_{p} \}
+  %\end{eqnarray*}
+  Let $\prog$ be as defined in Definition \ref{definition-ygcgate-certification}.
+  We observe:
+  \begin{eqnarray*}
+    &\prog_2 = \\
+    &{\small \prog[\be^1_{k}/\ttt{flip[2,gate:a.k]}][\be^1_{p}/\ttt{flip[2,gate:a.p]}][\be^2_{k}/\ttt{flip[2,gate:b.k]}][\be^1_{p}/\ttt{flip[2,gate:b.p]}]}
+  \end{eqnarray*}
+  Given that $g_1$ and $g_2$ are distinct and not wired in $\prog_1$
+  we are assured that $\ttt{owl}(g_1)$ and $\ttt{owl}(g_1)$ are in
+  independent uniform distributions, and given that $g \not\in
+  \prog_1$ we are assured that $\ttt{owl}(g)$ contains entirely fresh
+  flips. Thus by condition (i) of Definition \ref{definition-ygcgate-certification} and
+  Lemma \ref{lemma-substitution} we have:
+  $$
+  \sep{\prog_1;\prog_2}{\vars(\be^1_{k},\be^2_{k},\be^1_{p},\be^2_{p})}{\vdefs(\prog_2)}
+  $$
+  Thus by Lemmas \ref{lemma-noninterference} and \ref{lemma-separation} we establish
+  postconditions (ii) and (iii).
+
+  Also since $v_1$ and $v_2$ are correlated either positively or negatively with
+  $\ttt{owl}(g_1)$ and $\ttt{owl}(g_2)$ respectively by precondition (ii),
+  by Definition \ref{definition-gc}, precondition (i), and Lemma \ref{lemma-substitution-sim}
+  we establish postcondition (i), since Definition \ref{definition-ygcgate-certification}
+  requires gate output correlation with $\ttt{owl}(g)$ given any input valence conditions.
+\end{proof}
+
+\ygcencode*
+
+\begin{proof}
+  By Definition \ref{definition-ygcencode-certification} we have:
+  \begin{mathpar}
+    \sep{\progtt(\prog_2)}{\ttt{\{s[1,s1]\}}}{\vdefs(\prog_2)_{\{2\}}}
+
+    \sep{\progtt(\prog_2)}{\ttt{\{s[2,s2]\}}}{\vdefs(\prog_2)_{\{1\}}}
+  \end{mathpar}
+  Since we assume that secrets are in uniform and independent marginal
+  distributions a priori, and $\vars(\prog_1) \cap \vars(\prog_2) =
+  \varnothing$ by condition (iv) of Definition
+  \ref{definition-ygcencode-certification} and assumptions of
+  well-typedness, conditions (ii-iii) follow by Lemmas
+  \ref{lemma-noninterference} and \ref{lemma-separation}. Also by
+  Definition \ref{definition-ygcencode-certification} we have
+  $\gc{\prog_2}{v_1}{s_1}$ and $\gc{\prog_2}{v_1}{s_1}$, so also
+  $\gc{\prog_1;\prog_2}{v_1}{s_1}$ and
+  $\gc{\prog_1;\prog_2}{v_1}{s_1}$ by condition (iv) of Definition
+  \ref{definition-ygcencode-certification} and Lemmas
+  \ref{lemma-noninterference} and \ref{lemma-separation}.
+\end{proof}
+
+\ygcnimo*
+
+\begin{proof}
+  By assumptions of well-typedness we
+  have for some $\be_k$, $\be_p$, and $\be$:
+  $$
+  \config{\varnothing}{\eassign{\outv}{\ttt{decode}(g,e)}} \redxs
+  \config{\prog_1}{\eassign{\outv}{\ttt{decode}(g,\{ \ttt{k = }\be_k;  \ttt{p = }\be_p\})}}
+  \redxs \config{\prog_1;\prog_2}{\eassign{\outv}{\be}}
+  $$
+  Let $\iov(\prog_1) = S \cup V$.
+  By Lemmas \ref{lemma-ygc-encode} and \ref{lemma-ygc-preservation} we
+  have $\sep{\progd(\prog_1)}{S_{\{1\}}}{V_{\{2\}}}$ and
+  $\sep{\progd(\prog_1)}{S_{\{2\}}}{V_{\{1\}}}$.
+  Let $\prog$ be as defined in Definition \ref{definition-ygdecode-certification}.
+  We observe:
+  \begin{eqnarray*}
+    & \prog_2 = \\ 
+    & \prog[\be_k/\ttt{\{flip[2,gate:c.k]}][\be_p/\ttt{flip[2,gate:c.p]}]
+  \end{eqnarray*}
+  so also by  Definition \ref{definition-ygdecode-certification} and
+  Lemma \ref{lemma-substitution} we have:
+  $$
+  \sep{\progtt(\prog_1;\prog_2)}{\vars(\be_k,\be_p)}{\vdefs(\prog_2)}
+  $$
+  so, letting $\iov(\prog_1;\prog_2) = S \cup V'$ by
+  Lemmas \ref{lemma-noninterference} \ref{lemma-separation} we have
+  $\sep{\progd(\prog_1;\prog_2)}{S_{\{1\}}}{V'_{\{2\}}}$ and
+  $\sep{\progd(\prog_1;\prog_2)}{S_{\{2\}}}{V'_{\{1\}}}$.
+  Thus, letting $\prog' \defeq (\prog_1;\prog_2;\eassign{\outv}{\be})$, for all $\beta$:
+  $$\sep{\condd{\progd(\prog')}{S_{\{1\}} \cup V_{\{2\}}}{\{ \outv \mapsto \beta \}}}{S_{\{1\}}}{V_{\{2\}}}
+  \quad \text{and} \quad
+    \sep{\condd{\progd(\prog')}{S_{\{2\}} \cup V_{\{1\}}}{\{ \outv \mapsto \beta \}}}{S_{\{2\}}}{V_{\{1\}}}$$
+  thus for all $\store \in \mems(V_2 \cup \{ \outv \})$:
+  $$\condd{\progd(\prog')}{S_{\{1\}}}{\store}
+  = \condd{\progd(\prog')}{S_{\{1\}}}{\store_{\{\outv\}}}$$
+  and for all $\store \in \mems(V_1 \cup \{ \outv \})$:
+  $$\condd{\progd(\prog')}{S_{\{2\}}}{\store}
+  = \condd{\progd(\prog')}{S_{\{2\}}}{\store_{\{\outv\}}}$$
+  establishing the result by Lemma \ref{lemma-nimo} and Definition \ref{definition-NIMO}.
+\end{proof}

### 2023-07-10 popl24/ygc.tex
checkpoint

@@
 \section{YGC Code in Detail}
+\label{section-ygc}
 
@@
 with input type annotations for top-level functions listed in
-Figure \ref{fig-ygc-types}. Well-typed programs using these
-libraries are therguaranteed to yield safe $\minifed$
-programs.
+Figure \ref{fig-ygc-types}.
+
+The pointer bits in wire labels are used to select permuted rows in
+table garblings. The key bits are used to identify a unique key for
+table row in each garbled gate. Intuitively, if $\beta_1$ and
+$\beta_2$ are either key or pointer bits encoding 1 on two input wire
+labels to a binary gate, rows and keys in the gate are enumerated in
+the order:
+$$
+\neg\beta_1\neg\beta_2,\ \neg\beta_1\beta_2,\ \beta_1\neg\beta_2,\ \beta_1\beta_2
+$$

### 2023-07-10 popl24/composition.tex
checkpoint

@@
 \label{section-ygc-certification}
-
-Our certification method relies on a definition of compositional units comprising
-complete related subprotocols-- i.e., both the garbler's
-construction of the gates and the evaluator's evaluation of them. We
-therefore make the following definitions. However, we note that our
-formal results are applicable to the YGC style where the garbler
-shares the entire circuit prior to evaluation.
-\begin{definition}
-  \label{ygc-modules}
-  We define a composable internal \ttt{and} gate sharing and evaluation function
-  called $\ttt{andgg}$, and an output decoder gate share and evaluation function called
-  $\ttt{decode}$, as follows. 
-\begin{verbatimtab}
-  andgg(g, ga, gb, wva, wvb) { andgate(g, ga, gb); evalgate(g, wva, wvb) }
-  decode(g, v) { garbledecode(g); evaldecode(v) }
-\end{verbatimtab}
-\end{definition}
 

### 2023-07-10 popl24/ygc.tex
checkpoint

@@
 \emph{point-and-permute} method described in \cite{evans2018pragmatic}
-and elsewhere, to which the reader is referred for more in-depth discussion.
-In this implementation client 2 is the \emph{garbler} and
+and elsewhere, to which the reader is referred for more in-depth
+discussion.  In this implementation client 2 is the \emph{garbler} and
 client 1 is the \emph{evaluator}. The garbler builds the garbled
-tables and shares them with the evaluator, who then evaluates
-the gate in an oblivious fashion until the final public output is
-generated through decryption. This definition is well-typed,
-with input type annotations for top-level functions listed in
-Figure \ref{fig-ygc-types}.
+tables and shares them with the evaluator, who then evaluates the gate
+in an oblivious fashion until the final public output is generated
+through decryption. Our certification method leverages this definition
+of compositional units comprising complete related subprotocols--
+i.e., both the garbler's construction of the gates and the evaluator's
+evaluation of them. However, we note that our formal results are applicable
+to the YGC style where the garbler shares the entire circuit prior to
+evaluation. Our implementation is well-typed, with input type annotations
+for top-level functions listed in Figure \ref{fig-ygc-types}.
 

### 2023-07-10 popl24/bruteforce.tex
checkpoint

@@
 
+\begin{fpfig}[t]{Truth table visualizations of $\runs(\prog_{\ref{example-otp}})$ (L) and
+    $\runs(\prog_{\ref{example-lambda-obliv}(a)})$ (R).}{fig-basic-distributions}
+{\footnotesize
+  $$
+  \begin{array}{cc}
+    \begin{array}{cccccc}
+      \verb+s[1,0]+ & \verb+f[1,0]+ & \verb+v[2,0]+  & \verb+v[2,1]+ & \verb+v[2,2]+ & \verb+v[0,0]+\\
+      \hline
+      0 & 0 & 0 & 0 & 0 & 0 \\ 
+      0 & 1 & 1 & 1 & 0 & 1 \\ 
+      1 & 0 & 0 & 1 & 1 & 1 \\ 
+      1 & 1 & 1 & 0 & 1 & 0
+    \end{array}
+    & 
+    \begin{array}{ccccc}
+      \verb+s[1,s]+ & \verb+f[1,sx]+ & \verb+f[1,sy]+ & \verb+v[0,0]+ & \verb+v[0,1]+ \\
+      \hline
+      0 & 0 & 0 & 0 & 0 \\ 
+      0 & 0 & 1 & 1 & 0 \\ 
+      0 & 1 & 0 & 0 & 1 \\ 
+      0 & 1 & 1 & 1 & 1 \\
+      1 & 0 & 0 & 0 & 0 \\ 
+      1 & 0 & 1 & 0 & 0 \\ 
+      1 & 1 & 0 & 1 & 1 \\ 
+      1 & 1 & 1 & 1 & 1  
+    \end{array}
+  \end{array}
+  $$
+}
+\end{fpfig}
+
 %The basic distribution of a program $\prog$ is based on the
@@
 
-As we've observed, $|\runs(\prog)| = 2^{|S \cup F|}$, so the size of
-$\runs(\prog)$ is exponential in the amount of randomness used in the
-protocol which generally can be used as a proxy for the protocol
-size. In Section \ref{section-composition} we will show how the
-methods described here can be used to individually verify
-compositional properties of small program components, that guarantee
-security properties in larger protocols that combine them. Thus, while
-scalability to program size is a clear challenge for our brute force
-techniques alone, our ultimate goal is to use them for verification of
-compositional properties in larger programs.
+As we've observed, $\runs(\prog)$ can be represented in a truth table
+form, as illustrated in Figure \ref{fig-basic-distributions}.
 

### 2023-07-10 popl24/pmf.tex
checkpoint

@@
   \label{def-progd}
+  \label{definition-progd}
   Given program $\prog$ with $\iov(\prog) = (S,V)$ and $\flips(\prog) = F$, define $\stores$ as:
@@
 
-\begin{fpfig}[t]{Truth table visualization of basic distributions for Examples \ref{example-otp} (L) and \ref{example-lambda-obliv}(a) (R).}{fig-basic-distributions}
-{\footnotesize
-  $$
-  \begin{array}{cc}
-    \begin{array}{cccccc}
-      \verb+s[1,0]+ & \verb+f[1,0]+ & \verb+v[2,0]+  & \verb+v[2,1]+ & \verb+v[2,2]+ & \verb+v[0,0]+\\
-      \hline
-      0 & 0 & 0 & 0 & 0 & 0 \\ 
-      0 & 1 & 1 & 1 & 0 & 1 \\ 
-      1 & 0 & 0 & 1 & 1 & 1 \\ 
-      1 & 1 & 1 & 0 & 1 & 0
-    \end{array}
-    & 
-    \begin{array}{ccccc}
-      \verb+s[1,s]+ & \verb+f[1,sx]+ & \verb+f[1,sy]+ & \verb+v[0,0]+ & \verb+v[0,1]+ \\
-      \hline
-      0 & 0 & 0 & 0 & 0 \\ 
-      0 & 0 & 1 & 1 & 0 \\ 
-      0 & 1 & 0 & 0 & 1 \\ 
-      0 & 1 & 1 & 1 & 1 \\
-      1 & 0 & 0 & 0 & 0 \\ 
-      1 & 0 & 1 & 0 & 0 \\ 
-      1 & 1 & 0 & 1 & 1 \\ 
-      1 & 1 & 1 & 1 & 1  
-    \end{array}
-  \end{array}
-  $$
-}
-\end{fpfig}
-
-\paragraph{Example \ref{example-otp} revisited} For small programs it is
-helpful to visualize their basic distributions graphically as their
-truth tables, where each row corresponds to an equally likely final
-memory for the given program. In Figure \ref{fig-basic-distributions}
-we show this for Example \ref{example-otp}. In this truth table we can
-see that $\sx{1}{0}$ and $\vx{2}{2}$ are perfectly correlated. But
+\paragraph{Example \ref{example-otp} revisited}
+%For small programs it is
+%helpful to visualize their basic distributions graphically as their
+%truth tables, where each row corresponds to an equally likely final
+%memory for the given program. In Figure \ref{fig-basic-distributions}
+%we show this for Example \ref{example-otp}. In this truth table we can
+%see that $\sx{1}{0}$ and $\vx{2}{2}$ are perfectly correlated.
+In this example, 
 from the adversarial view, $\sx{1}{0}$ is 1 or 0 with equal
@@
 to leak information probabilistically through observable dependencies
-in views. In the basic distribution for Example \ref{example-lambda-obliv}(a)
-illustrated in Figure \ref{fig-basic-distributions}, we can see that
+in views.
+%In the basic distribution for Example \ref{example-lambda-obliv}(a)
+%illustrated in Figure \ref{fig-basic-distributions},
+We can see that
 the adversarial views $\vx{0}{0}$ and $\vx{0}{1}$ are both 1 in

### 2023-07-10 popl24/metalang.tex
checkpoint

@@
 hypothetical gate 6:
+{\small
   \begin{mathpar}
@@
   \end{mathpar}
+}
 In our implementation, gates are wired together using gate

### 2023-07-10 popl24/hyperprop.tex
checkpoint

@@
 corrupt inputs $\store$ and ideal functionality output $v$,  
-we write:
-$$
+we write
+$
 \prob(\SIM_C(\store,v) = \store')
-$$
+$
 to denote the probability that $\SIM_C(\store,v)$
@@
   $\prog$ satisfies probabilistic noninterference iff for all
-  $\store \in \mems(S \cup V)$:
-  $$\condd{\progd(\prog)}{S_H}{\store_{S_C}} =
-    \condd{\progd(\prog)}{S_H}{\store_{(S\cup V)_C}} $$
+  $\store \in \mems(S \cup V)$,
+  $\condd{\progd(\prog)}{S_H}{\store_{S_C}} =
+    \condd{\progd(\prog)}{S_H}{\store_{(S\cup V)_C}}$.
 \end{lemma}
@@
   Given a functionality $\idealf$ and output value $v$, the associated
-  \emph{ideal knowledge}, denoted $\ik(\idealf,v)$ is:
-  $$
+  \emph{ideal knowledge}, denoted $\ik(\idealf,v)$ is
+  $
   \{ \store\ |\ \idealf(\store) = v \}
-  $$
+  $.
   We say that $\idealf$ is \emph{pre-imageable} iff $\ik(\idealf, v)$ for all
@@
   F)$. Then the distribution of adversarial views in the real world
-  is, by definition:
-  $$\condd{\progd(\prog)}{V_C}{\store_S}$$
+  is, by definition,
+  $\condd{\progd(\prog)}{V_C}{\store_S}$.
 
@@
   that we assume for the real world, after selection of $\store'$ the
-  probability that any particular $\store^{\SIM}_{V_C}$ is returned is by definition:
-  $$
+  probability that any particular $\store^{\SIM}_{V_C}$ is returned is by definition
+  $
   \progd(\prog)(\store^{\SIM}_{V_C} | \store')
-  $$
+  $.
   Furthermore, since we assume that $\prog$ correctly implements $\idealf$, this
-  means:
-  $$
+  means
+  $
   \store_S =_C \store' \wedge
@@
      \condd{\progd(\prog)}{\{ \outv \}}{\store'}
-  $$
+  $,
   so by the definition of noninterference modulo output,

### 2023-07-10 popl24/metalang.tex
checkpoint

@@
 $\metaprot$ type checking is straightforward and efficient.
-Function types are of the form:
-$$
-\tau_1 * \cdots * \tau_n \rightarrow \tau,\viewst
-$$
+Function types are of the form
+$
+(\tau_1 * \cdots * \tau_n \rightarrow \tau,\viewst)
+$,
 where $\viewst$ denotes the effect of the function on the residual

### 2023-07-10 popl24/minicat.tex
checkpoint

@@
 $\select{\be_1}{\be_2}{\be_3}$, which is essentially a conditional
-expression with:
-$$
+expression with
+$
 \select{\be_1}{\be_2}{\be_3} \equiv (\be_1\ \eand\ \be_2)\ \eor\ (\enot\ \be_1\ \eand\ \be_3)
-$$
+$.
 We also include \emph{oblivious transfer} $\OT{\be_1}{\be_2}{\be_3}$ as a primitive,
@@
 
-\begin{fpfig}[t]{$\minifed$ expression interpretation.}{fig-minifed-interp}
+\begin{fpfig}[t]{$\minifed$ expression interpretation (T) and reduction (B) rules.}{fig-minifed-interp}
   {\small
@@
   $$
-  }
-\end{fpfig}
+  \vspace{1mm}
 
-Evaluation of configurations is then defined via a small-step reduction relation $\redx$.
-This is defined in the obvious manner for view assignments other than through
-$\ttt{OT}$- however note that reduction requires that views are never reassigned. 
+  \rule{130mm}{0.5pt}
+
   \begin{mathpar}
     (\store, \eassign{\view{\cid}{w}}{\be};\prog) \redx (\extend{\store}{\view{\cid}{w}}{\lcod{\store,\be}{\cid'}}, \prog)
-\end{mathpar}
-For view assignments through $\ttt{OT}$, we define a different reduction rule that
-captures the appropriate semantics- that is, the selection bit is \emph{not} communicated
-to the sender (through a view assignment), and only the selected bit is sent to the receiver:
-\begin{mathpar}
+
       \inferrule
@@
 \end{mathpar}
+  }
+\end{fpfig}
+
+Evaluation of configurations is then defined via a small-step reduction relation $\redx$,
+also shown in Figure \ref{fig-minifed-interp}.
+This is defined in the obvious manner for view assignments other than through
+$\ttt{OT}$- however note that reduction requires that views are never reassigned. 
+For view assignments through $\ttt{OT}$, we define a different reduction rule that
+captures the appropriate semantics- that is, the selection bit is \emph{not} communicated
+to the sender (through a view assignment), and only the selected bit is sent to the receiver:
 We define $\redxs$ as the reflexive, transitive closure of $\redx$.

### 2023-07-10 popl24/pmf.tex
checkpoint

@@
 \end{definition}
-To recover familiar notation we allow the following syntactic
-sugarings:
-\begin{eqnarray*}
-  \condd{\pmf}{X}{\store}  &\defeq& \condd{\pmf}{X}{\{ \store\}}\\
-  \pmf(\store)  &\defeq& \margd{\pmf}{X}(\store)  \qquad \dom(\store) = X\\
-  \pmf(\store|\stores)  &\defeq& \condd{\pmf}{X}{\stores}(\store) \qquad \dom(\store) = X
-\end{eqnarray*}
+To recover familiar notation we allow the syntactic
+sugarings $\condd{\pmf}{X}{\store}  \defeq \condd{\pmf}{X}{\{ \store\}}$, and
+$\pmf(\store)  \defeq \margd{\pmf}{X}(\store)$ and $\pmf(\store|\stores) \defeq
+\condd{\pmf}{X}{\stores}(\store)$ where $\dom(\store) = X$.
+%\begin{eqnarray*}
+%  \condd{\pmf}{X}{\store}  &\defeq& \condd{\pmf}{X}{\{ \store\}}\\
+%  \pmf(\store)  &\defeq& \margd{\pmf}{X}(\store)  \qquad \dom(\store) = X\\
+%  \pmf(\store|\stores)  &\defeq& \condd{\pmf}{X}{\stores}(\store) \qquad \dom(\store) = X
+%\end{eqnarray*}
 
@@
   $$
-  \pmf(\store) =
-  \begin{cases}
-    1 / |\stores| & \text{if}\ \store \in \stores\\
-    0 & \text{otherwise}
-  \end{cases}
+  \pmf(\store) =  1 / |\stores| \ \text{if}\ \store \in \stores, \text{otherwise}\ 0
+  %\begin{cases}
+  %  1 / |\stores| & \text{if}\ \store \in \stores\\
+  %  0 & \text{otherwise}
+  %\end{cases}
   $$
   Then the \emph{program distribution of $\prog$}, denoted $\progd(\prog)$, is the
-  marginal distribution of $S \cup V$ in $\prog$'s basic distribution:
-  $$
-  \progd(\prog) =  \margd{\pmf}{S\cup V}
-  $$
+  marginal distribution of $S \cup V$ in $\prog$'s basic distribution, i.e., 
+  $
+  \progd(\prog) \defeq  \margd{\pmf}{S\cup V}
+  $.
   In some cases, we will also be concerned with the (joint)

### 2023-07-10 popl24/gmw.tex
checkpoint

@@
-\begin{fpfig}[t]{2-Party GMW functions.}{fig-gmw}
+\begin{fpfig}[t]{2-Party GMW Circuit Library with And gate.}{fig-gmw}
 {\footnotesize
   \begin{verbatimtab}
-    encodegmw(in1,in2)
-    {
-      v[1, in1 || out] := s[1,in1] xor flip[1,in1];
-      v[2, in1 || out] := flip[1,in1];
-      v[1, in2 || out] := s[2,in2] xor flip[2,in2];
-      v[2, in2 || out] := flip[2,in2]
-      { shares1 = { c1 = v[1, in1 || out]; c2 = v[2, in1 || out] };
-        shares2 = { c1 = v[1, in2 || out]; c2 = v[2, in2 || out]} } 
+    encodegmw(in1,in2) {
+      v[1, in1 || "out"] := s[1,in1] xor flip[1,in1];
+      v[2, in1 || "out"] := flip[1,in1];
+      v[1, in2 || "out"] := s[2,in2] xor flip[2,in2];
+      v[2, in2 || "out"] := flip[2,in2]
+      { shares1 = { c1 = v[1, in1 || "out"]; c2 = v[2, in1 || "out"] };
+        shares2 = { c1 = v[1, in2 || "out"]; c2 = v[2, in2 || "out"]} } 
     }
 
-    andtablegmw(b1, b2)
-    {
+    andtablegmw(b1, b2) {
       let r11 = (b1 xor true) and (b2 xor true) in
@@
 
-    andgmw(g, shares1, shares2)
-    {
-      let r = flip[1,g || .r] in
+    andgmw(g, shares1, shares2) {
+      let r = flip[1,g || ".r"] in
       let table = andtablegmw(shares1.c1, shares2.c1) in
@@
       let r00 =  r xor table.v4 in
-      v[2,g || out] := OT4(shares1.c2, shares2.c2, r11, r10, r01, r00);
-      v[1,g || out] := r;
-      { c1 = v[1,g || out]; c2 = v[2,g || out]}
+      v[2,g || "out"] := OT4(shares1.c2, shares2.c2, r11, r10, r01, r00);
+      v[1,g || "out"] := r;
+      { c1 = v[1,g || "out"]; c2 = v[2,g || "out"]}
     }
 
-    decodegmw(shares)
-    {
-      v[0,output] := shares.c1 xor shares.c2
-    }
+    decodegmw(shares) { v[0,"output"] := shares.c1 xor shares.c2 }
   \end{verbatimtab}

### 2023-07-10 popl24/metalang.tex
checkpoint

@@
       share3(client, secretid) {
-        let s1 = flip[client, share1] in
-        let s2 = flip[client, share2] in
-        let s3 = (s1 xor s2) xor s[client, s: || secretid] in
+        let s1 = flip[client, "share1"] in
+        let s2 = flip[client, "share2"] in
+        let s3 = (s1 xor s2) xor s[client, "s:" || secretid] in
         {s1 = s1;s2 = s2;s3 = s3}
@@
     \begin{verbatimtab}
-      let shares = share3(1, mysecret) in
+      let shares = share3(1, "mysecret") in
       v[2,s1] := shares.s2;
@@
   \begin{verbatimtab}
-      v[2,s1] := flip[1, share2];
-      v[3,s1] := flip[1, share1] xor flip[1, share2] xor s[1, s:mysecret] \end{verbatimtab}
+      v[2,s1] := flip[1, "share2"];
+      v[3,s1] := flip[1, "share1"] xor flip[1, "share2"] xor s[1, "s:mysecret"] \end{verbatimtab}
   }
@@
 \item \ttt{encodegmw}: This function encodes client 1's and client 2's
-  secret bits called $\ttt{s[1,s1]}$ and $\ttt{s[1,s2]}$ into two
+  secret bits called $\ttt{s[1,"s1"]}$ and $\ttt{s[1,"s2"]}$ into two
   distinct wire values (pairs of shares).
-\item \ttt{andgmw}: This function defines the gate $\ttt{g}$, the
-  identifier $\ttt{g}$ being used to distinguish randomness used
+\item \ttt{andgmw}: This function defines the gate $\ttt{"g"}$, the
+  identifier $\ttt{"g"}$ being used to distinguish randomness used
   within.  In our version client 1 builds the output table (using
@@
   \begin{verbatimtab}
-  let ss = encodegmw(s1,s2) in v[0,output] := decode(andgmw(0,ss.shares1,ss.shares2))
-  \end{verbatimtab}
+  let ss = encodegmw("s1","s2") in v[0,"output"] := decode(andgmw(0,ss.shares1,ss.shares2)) \end{verbatimtab}
 }
@@
   \begin{mathpar}
-    \ttt{\{ k = flip[2,gate:6.k]; p =  flip[2,gate:6.p]] \}}
+    \ttt{\{ k = flip[2,"gate:6.k"]; p =  flip[2,"gate:6.p"]] \}}
 
-    \ttt{\{ k = not flip[2,gate:6.k]; p =  not flip[2,gate:6.p]] \}}
+    \ttt{\{ k = not flip[2,"gate:6.k"]; p =  not flip[2,"gate:6.p"]] \}}
   \end{mathpar}

### 2023-07-10 popl24/main.tex
checkpoint

@@
 \usepackage{enumerate}
+\usepackage{thmtools,thm-restate}
+\usepackage{comment}
 
@@
 
-\author{Author names withheld for double-blind reviewing.}
+\author{Author names withheld for double-blind reviewing}
 
 \begin{abstract}
-  We formulate a sound hyperproperty of MPC passive security and show how this
-  supports automated and semi-automated verification.
+Secure Multi-Party Computation (MPC) protocols support data privacy in
+important modern distributed applications. Security for MPC protocols
+is superficially similar to probabilistic noninterference, but differs
+in a subtle but fundamental way, and approaches for verifying
+noninterference cannot naturally extend to MPC security.  Currently,
+proof methods for MPC protocols are well-studied but manual
+and thus tedious and error-prone, and are also non-standardized and
+unfamiliar to most PL theorists.  Our goal is to strengthen
+connections between the security model of MPC and trace-based
+hyperproperties, and to leverage them to obtain automated
+enforcement mechanisms for MPC protocol development.  We define a
+novel hyperproperty that is provably sound for MPC security, as a
+foundation for fully and partially automated verification of passive
+MPC security in protocols including arbitrarily large YGC and GMW
+circuits.
 \end{abstract}
@@
 
+\input{metalang}
+
 \input{pmf}
@@
 
-\input{bruteforce}
-
-\input{metalang}
+\input{automation}
 
 \input{composition}
+
+%% This is scratch pad to develop ideas before integration.
+%\input{scratch}
 
@@
 
+\appendix
+
+\input{ygc}
+
+\input{bruteforce}
+
+\input{proofs}
+
 \end{document}

### 2023-07-10 popl24/intro.tex
checkpoint

@@
 important modern, distributed applications such as privacy-preserving
-machine learning \cite{li2021privacy, knott2021crypten, koch2020privacy, liu2020privacy} and Zero-Knowledge proofs in blockchains \cite{ishai2009zero, lu2019honeybadgermpc, gao2022symmeproof, tomaz2020preserving}.
-MPC methods
-have been developed by the cryptography community for many years,
-while receiving the attention of the programming languages community
-only relatively recently. In contrast, information flow security has
-received significant attention in PL theory and practice especially
-since the turn of the century \cite{1159651}, including a menagerie of
-variants, enforcement mechanisms, and programming frameworks. Much of
-this has been enabled by the unified metatheory of
-\emph{hyperproperties} \cite{10.5555/1891823.1891830} that establishes
-a common conceptual framework for reasoning about and implementing
-systems with information flow security.
+machine learning \cite{li2021privacy, knott2021crypten,
+  koch2020privacy, liu2020privacy} and Zero-Knowledge proofs in
+blockchains \cite{ishai2009zero, lu2019honeybadgermpc,
+  gao2022symmeproof, tomaz2020preserving}. MPC methods have been
+developed by the cryptography community for many years, while
+receiving the attention of the programming languages community more
+recently. In contrast, information flow security has received
+significant attention in PL theory and practice especially since the
+turn of the century \cite{1159651}, including a menagerie of variants,
+enforcement mechanisms, and programming frameworks. Much of this has
+been enabled by the unified metatheory of \emph{hyperproperties}
+\cite{10.5555/1891823.1891830}, including noninterference, that
+establishes a common conceptual framework for reasoning about and
+implementing systems with information flow security.
 
-Our goal is to explore and establish connections between the security
+On its surface, security for MPC protocols seems similar to
+probabilistic noninterference. However, MPC security differs from
+noninterference in a subtle but fundamental way, and approaches for
+verifying noninterference cannot naturally extend to MPC security. Our
+goal is to explore and establish connections between the security
 model of MPC-- the \emph{real/ideal} aka \emph{simulator security}
@@
 protocol development. Currently, proof methods for MPC protocol
-development are well-studied \cite{Lindell2017} but manual and 
-thus tedious and error-prone, and are also non-standardized and
-unfamiliar to most PL theorists. Therefore our exploration will
-make both theoretical and practical contributions by bridging a gap between
+development are well-studied \cite{Lindell2017} but manual and thus
+tedious and error-prone, and are also non-standardized and unfamiliar
+to most PL theorists. Therefore our exploration will make both
+theoretical and practical contributions by bridging a gap between
 information flow and simulator security methodologies.
@@
 \subsection{The Analysis Challenge of MPC}
-\jnote{On its surface, security for MPC protocols seems similar to probabilistic noninterference.
-However, MPC security differs from noninterference in a subtle but fundamental way, and approaches for verifying noninterference cannot naturally extend to MPC security.}
 MPC protocols involve communication
@@
 in Section \ref{section-minicat-examples} and more complex ones in
-Section \ref{section-composition}. But an immediate and main point in
+Section \ref{section-metalang}. But an immediate and main point in
 relation to security hyperproperties is that, due to the potential for
@@
 
-In Section \ref{section-minicat} we develop a new probabilistic programming
-language $\minifed$ for defining synchronous distributed protocols
-over the binary field. The syntax and semantics provides a succinct
-account of \emph{views} and (synchronous) messaging between protocol
-\emph{clients}. In this Section, we consider small example protocols
-including oblivious transfer, communications encryption with one-time
-pads, and additive secret sharing. In Section \ref{section-pmf} we
+\paragraph{Language design.} In Section \ref{section-minicat} we
+develop a new probabilistic programming language $\minifed$ for
+defining synchronous distributed protocols over the binary field. The
+syntax and semantics provides a succinct account of \emph{views} and
+(synchronous) messaging between protocol \emph{clients}. In Section
+\ref{section-metalang} we define a new metaprogramming language
+$\metaprot$ that dynamically generates $\minifed$ protocols. It
+includes control and data structures and is able to express logical
+protocol components, and enjoys a type safety result that guarantees
+that generated protocols are semantically well-defined.
+
+\paragraph{Hyperproperty formulation.} In Section \ref{section-pmf} we
 formulate \emph{program distributions}, a formalism for expressing
@@
 traces, called \emph{noninterference modulo output ($\NIMO$)} that we
-demonstrate implies passive simulator security. In Section
-\ref{section-bruteforce} we define a brute force mechanism for
-verifying passive security in $\minifed$ protocols, that is also
-amenable to HPC optimizations via translation of $\minifed$ protocols
-into Datalog.
+demonstrate implies passive simulator security.
 
-In Section \ref{section-metalang} we define a new metaprogramming
-language $\metaprot$ that dynamically generates $\minifed$
-protocols. It includes control and data structures and is able to
-express logical protocol components, and enjoys a type safety result
-that guarantees that generated protocols are semantically
-well-defined. Our key analysis results are presented in Section
-\ref{section-composition}, where we show how intensional,
-automatically certifiable properties of program components can be
-composed to obtain whole-program MPC security. We consider the GMW and
-YGC protocols as examplars, both of which use \emph{circuit}
-abstractions comprising \emph{gates} in arbitrarily large
-configurations, and show how gates can be automatically certified to
-ensure security in circuits that use them in a manner that scales to
-any circuit size. This approach is extensible, in that new gate
-implementations can be automatically certified to guarantee safe
-inclusion in circuit libraries.
-
-In summary, our analysis takes a ``just do it'' approach- rather than
-using, e.g., traditional security type or linear type disciplines to
-prevent ``bad things'' from happening, we base our analysis on precise
-quantification of intensional protocol behavior and especially
-dependencies between secret inputs, public outputs, and adversarial
-views of protocols. We ameliorate scalability issues associated with
-fine-grained analysis by manually establishing compositional
-properties inspired by probabilistic separation logic
-\cite{barthe2019probabilistic}, obtaining a middle-ground, scalable
-``proof assistant style'' approach enabled by MPC circuit paradigms.
+\paragraph{Fully and partially automated verification.} In Section
+\ref{section-automation} we develop an automated technique, aka
+certification, that verifies extensional properties of protocols such
+as $\NIMO$, and also intensional properties of protocol
+components. The latter include properties of protocol components that
+provably guarantee security through composition in large
+circuits. These compositional proofs are based on separation logic but
+include novel results to support our semi-automated method for large
+circuits where we (1) define certification for components, (2) certify
+components, (3) prove $\NIMO$ based on the certifiedness of
+components. We develop novel methods partially based on separation
+logic \cite{barthe2019probabilistic} to enable (3).  In Section
+\ref{section-composition} we illustrate our method to obtain
+semi-automated verification methods for the GMW and Yao's Garbled
+Circuit protocols, where certification can be re-used to validate new
+gate implementations.

### 2023-07-11 popl24/automation.tex
checkpoint

@@
 \section{Automating Security Verification}
+\label{section-automation}
+
+\subsection{Brute Force Computation of Program Distributions}
+\label{section-bruteforce}
+
+We first explore a brute force approach to automatically verifying
+program behaviour, the first step of which is to compute the basic
+distribution of a program $\prog$. The program distribution
+$\progd(\prog)$ can obviously be automatically derived from that, as
+can be any of its marginal or conditional distributions. In this
+Section, we consider two automated mechanisms for deriving
+$\progd(\prog)$ in this way. The first is through a straightforward
+computation of all final program memories.  The second approach
+rewrites $\minifed$ programs to stratified Datalog programs, that are
+amenable HPC optimizations such as parallelization and GPU matrix
+computations as shown in recent work
+\cite{sakama2017linear,aspis2018linear,nguyen2022enhancing,nguyen2021efficient}.
+
+\begin{fpfig}[t]{Truth table visualizations of $\runs(\prog_{\ref{example-otp}})$ (L) and
+    $\runs(\prog_{\ref{example-lambda-obliv}(a)})$ (R).}{fig-basic-distributions}
+{\footnotesize
+  $$
+  \begin{array}{cc}
+    \begin{array}{cccccc}
+      \verb+s[1,0]+ & \verb+f[1,0]+ & \verb+v[2,0]+  & \verb+v[2,1]+ & \verb+v[2,2]+ & \verb+v[0,0]+\\
+      \hline
+      0 & 0 & 0 & 0 & 0 & 0 \\ 
+      0 & 1 & 1 & 1 & 0 & 1 \\ 
+      1 & 0 & 0 & 1 & 1 & 1 \\ 
+      1 & 1 & 1 & 0 & 1 & 0
+    \end{array}
+    & 
+    \begin{array}{ccccc}
+      \verb+s[1,s]+ & \verb+f[1,sx]+ & \verb+f[1,sy]+ & \verb+v[0,0]+ & \verb+v[0,1]+ \\
+      \hline
+      0 & 0 & 0 & 0 & 0 \\ 
+      0 & 0 & 1 & 1 & 0 \\ 
+      0 & 1 & 0 & 0 & 1 \\ 
+      0 & 1 & 1 & 1 & 1 \\
+      1 & 0 & 0 & 0 & 0 \\ 
+      1 & 0 & 1 & 0 & 0 \\ 
+      1 & 1 & 0 & 1 & 1 \\ 
+      1 & 1 & 1 & 1 & 1  
+    \end{array}
+  \end{array}
+  $$
+}
+\end{fpfig}
+
+%The basic distribution of a program $\prog$ is based on the
+%final memories resulting from any run of $\prog$, denoted
+%$\runs(\prog)$.
+%\begin{definition}
+%  Given $\prog$ with $\iov(\prog) = S \cup V$ and $\flips(\prog) = F$:
+%  $$
+%  \runs(\prog) \defeq \{ \store \in \mems(S\cup F \cup V) \mid \config{\store_{S \cup F}}{\prog} \redxs \config{\store}{\varnothing} \}
+%  $$
+%\end{definition}
+
+As we've observed, $\runs(\prog)$ can be represented in a truth table
+form, as illustrated in Figure \ref{fig-basic-distributions}.
+
+\begin{fpfig}[t]{Filtering memories that satisfy a boolean expression.}{fig-solve}
+{\small
+\begin{eqnarray*}
+\solve{\stores}{\etrue} &=& \stores\\
+\solve{\stores}{\efalse} &=& \varnothing\\
+\solve{\stores}{\flip{\cid}{w}} &=& \{ \store \in \stores \mid \store(\flip{\cid}{w}) \} \\
+\solve{\stores}{\secret{\cid}{w}} &=& \{ \store \in \stores \mid \store(\secret{\cid}{w}) \} \\
+\solve{\stores}{\view{\cid}{w}} &=& \{ \store \in \stores \mid \store(\view{\cid}{w}) \} \\
+\solve{\stores}{\oracle{w}} &=& \{ \store \in \stores \mid \store(\oracle{w}) \} \\
+\solve{\stores}{(\enot\ \be)} &=& \stores - (\solve{\stores}{\be})\\
+\solve{\stores}{(\be_1\ \eand\ \be_2)} &=& (\solve{\stores}{\be_1}) \cap (\solve{\stores}{\be_2}) \\
+\solve{\stores}{(\be_1\ \eor\ \be_2)} &=& (\solve{\stores}{\be_1}) \cup (\solve{\stores}{\be_2}) \\
+\solve{\stores}{(\be_1\ \exor\ \be_2)} &=&
+ ((\solve{\stores}{\be_1}) \cap (\stores - \solve{\stores}{\be_2})) \cup\\
+ && ((\stores - \solve{\stores}{\be_1}) \cap (\solve{\stores}{\be_2})) \\
+\solve{\stores}{\select{\be_1}{\be_2}{\be_3}} &=&
+ ((\solve{\stores}{\be_1}) \cap (\solve{\stores}{\be_2})) \cup \\
+ && ((\stores - \solve{\stores}{\be_1}) \cap (\solve{\stores}{\be_3})) \\
+\solve{\stores}{\OT{\be_1}{\be_2}{\be_3}} &=&
+ ((\solve{\stores}{\be_1}) \cap (\solve{\stores}{\be_2})) \cup\\
+ && ((\stores - \solve{\stores}{\be_1}) \cap (\solve{\stores}{\be_3}))
+\end{eqnarray*}
+}
+\end{fpfig}
+
+Here we describe a straightforward technique for obtaining program
+distributions $\progd(\prog)$ through direct computation of
+$\runs(\prog)$. Although we could obtain $\runs(\prog)$ simply by
+executing $\prog$ given all random tapes and input secrets, the method
+we describe here is empirically more efficient (though still
+exponential in the length of the random tape), and also supports the
+translation to Datalog described in Section
+\ref{section-bruteforce-datalog}. An implementation of these
+techniques along with Examples from Sections
+\ref{section-minicat-examples} and \ref{section-metalang-ygc} is
+available online\footnote{URL redacted for double-blind review.}.
+
+In Figure \ref{fig-solve} we define the algorithm $\solve{\stores}{\be}$
+which filters $\stores$ based on satisfaction of $\be$. This algorithm
+is fundamental to our brute force methods. It's correctness is characterized
+as follows. 
+\begin{lemma}
+  \label{lemma-solves}
+  For all $\stores$ and $\be$ with $\vars(\be) \subseteq \dom(\stores)$,
+  $(\solve{\stores}{\be}) = \{ \store \in \stores \ \mid\ \lcod{\store,\be}{\cid} = 1 \}$
+  for some $\cid$.
+\end{lemma}
+\begin{proof}
+  Since we assume safety of programs we can assume that all variables in $\be$ have the
+  same owner $\cid$. The result otherwise follows in a straightforward manner by induction
+  on $\be$. 
+\end{proof}
+
+In order to automatically obtain $\runs(\prog)$, we can do a left
+folding of $\solve$ across $\prog$, inductively extending memories
+with valid view assignments in the order of their definition.
+We can then use this to compute $\progd(\prog)$. We denote
+these computations as $\cruns$ and $\cprogd$ respectively.
+\begin{lemma}
+  \label{lemma-cruns}
+  Given $\prog$ where $\iov(\prog) = S \cup V$ and $\flips(\prog) = F$. Define:
+  \begin{eqnarray*}
+    {tt}\ \ \stores\ (\eassign{\view{\cid}{w}}{\be}) &\defeq& \begin{array}{l}
+      \mathrm{let}\ \stores' = \solve{\stores}{\be} \ \mathrm{in}\\
+      \ \ \{\extend{\store}{\view{\cid}{w}}{1} \mid \store \in \stores' \}\ \cup\\
+      \ \ \{\extend{\store}{\view{\cid}{w}}{0} \mid \store \in \stores - \stores' \}\end{array}\\[2mm]
+    \cruns(\prog) &\defeq& \mathit{foldl}\ {tt}\ \mems(S \cup F)\ \prog
+  \end{eqnarray*}
+  Then $\cruns(\prog) = \runs(\prog)$.
+\end{lemma}
+\begin{proof}
+  By Lemma \ref{lemma-solves} and induction on the length of $\prog$. 
+\end{proof}
+\begin{lemma}
+  \label{lemma-cprogd}
+  Given $\prog$ where $\iov(\prog) = S \cup V$ and $\flips(\prog) = F$. Define
+  $\cprogd(\prog)$ as the distribution $\pmf$
+  where for all $\store \in \mems(S\cup F \cup V)$:
+  $$
+  \pmf(\store) = \begin{cases}1/2^{|S \cup F|} & \text{if\ } \store \in
+    \cruns(\prog) \\ 0 & \text{otherwise} \end{cases} 
+  $$
+  Then:
+  $$
+  \cprogd(\prog) = \progd(\prog)
+  $$
+\end{lemma}
+\begin{proof}
+  Immediately by Lemma \ref{lemma-cruns} and Definition \ref{def-progd}.
+\end{proof}
+
+\subsection{Protocol Certification}
 

### 2023-07-11 popl24/examples.tex
checkpoint

--- 2023-07-10 popl24/pmf.tex
@@
-\section{Probability Mass Functions and Program Distributions}
-\label{section-pmf}
+\section{Supplemental Examples}
 
-Given a program $\prog$ with $\iov(\prog) = S \cup V$ and $\store \in
-\mems(\iov(\prog))$, we are fundamentally concerned with the joint
-probability of the assignments in $\store_V$ conditioned by the
-assignments $\store_S$ in any given run, since probabilistic
-dependence between honest secrets and corrupt views is our metric of
-information leakage.
+\subsection{$\metaprot$ and $\minicat$ Protocol Examples}
 
-\subsection{Basic Definitions}
+\begin{example}[Additive Secret Sharing]
+    \label{example-he}
+Additive secret sharing is an MPC protocol where $k$ parties split up
+their secrets into \emph{shares}, with the property that all $k$
+shares are required to reconstruct the secret and any fewer reveals
+nothing. Additionally, this sharing enjoys additive homomorphism. In
+the binary field it is well-known that we can extend our use
+of $\ttt{xor}$ and flips to generate an arbitrary number of shares
+with additive homomorphism.
 
-We begin by defining joint probability mass functions in the standard
-manner, though following \cite{barthe2019probabilistic} we use
-memories to denote mappings of variables to values (i.e., outcomes),
-so for example given a pmf $\pmf$ we will write $\pmf(\{ \sx{1}{a}
-\mapsto 0, \vx{2}{b} \mapsto 1 \})$ to denote the probability that
-$\sx{1}{a} = 0$ and $\vx{2}{b} = 1$ in the final memory of a
-program containing these variables.
-\begin{definition}
-  A \emph{probability mass function} $\pmf$ is a function
-  mapping memories in $\mems(X)$ for some $X$ to values in $[0,1]$, such that:
+For example, to generate shares for three-party additive sharing,
+clients 1,2, and 3 can break up and distribute their shares to
+each other as follows:
+{\small
+\begin{verbatimtab}
+     v[2,"s1"] := flip[1,"1"] xor flip[1,"s1"] xor s[1,"1"];
+     v[3,"s1"] := flip[1,"1"];
+     v[1,"s2"] := flip[2,"1"] xor flip[2,"s2"] xor s[2,"1"];
+     v[3,"s2"] := flip[2,"1"];
+     v[1,"s3"] := flip[3,"1"] xor flip[3,"s3"] xor s[3,"1"];
+     v[2,"s3"] := flip[3,"1"] \end{verbatimtab}
+}
+Then, assuming that party 0 is a ``public'' client, each party sums
+its own shares, and then the sum of the sum of shares is revealed
+in the output view $\vx{0}{output}$. By additive homomorphism of
+$\ttt{xor}$, this output is the sum of the 3 secrets.
+{\small
+\begin{verbatimtab}
+     v[0,"ss1"] := v[1,"s2"] xor v[1,"s3"] xor flip[1,"s1"];
+     v[0,"ss2"] := v[2,"s1"] xor v[2,"s3"] xor flip[2,"s2"];
+     v[0,"ss3"] := v[3,"s1"] xor v[3,"s2"] xor flip[3,"s3"];
+     v[0,"output"] := v[0,"ss1"] xor v[0,"ss2"] xor v[0,"ss3"] \end{verbatimtab}
+}
+We note that, as for MPC protocols generally, the notion of
+``secrecy'' here is subtle and not the same as for encryption as in
+Example \ref{example-otp}, particularly when the threat model allows
+protocol participants to be corrupt. For example, suppose that $1 \in C$, and
+$\sx{1}{1}$ is $1$, and after running the protocol suppose that
+$\vx{0}{output}$ is 0. Then the adversary knows that 2 and 3's secrets
+are either 1 and 0, or 0 and 1, with 50/50 probability, and definitely
+neither 0 and 0, nor 1 and 1. 
+\end{example}
+
+Buliding on Example \ref{example-he}, we can define a function
+$\ttt{share3}$ that abstracts the process of splitting a given
+client's secret into 3 separate shares.
+\begin{example} \label{example-share3} The function $\ttt{share3}$ 
+  splits a client's secret into 3 shares returned as a record
+  with fields $\ttt{s1-3}$:
+  {\small
+  \begin{verbatimtab}
+      share3(client, secretid) {
+        let s1 = flip[client, "share1"] in
+        let s2 = flip[client, "share2"] in
+        let s3 = (s1 xor s2) xor s[client, "s:" || secretid] in
+        {s1 = s1;s2 = s2;s3 = s3}
+      } \end{verbatimtab}
+  }
+  Here is a $\metaprot$ program that uses this function definition:
+  {\small
+    \begin{verbatimtab}
+      let shares = share3(1, "mysecret") in
+      v[2,s1] := shares.s2;
+      v[3,s1] := shares.s3 \end{verbatimtab}
+  }
+  which generates the following $\minifed$ program, as we formalize in Example \ref{example-share3-eval}
+  below:
+  {\small
+  \begin{verbatimtab}
+      v[2,s1] := flip[1, "share2"];
+      v[3,s1] := flip[1, "share1"] xor flip[1, "share2"] xor s[1, "s:mysecret"] \end{verbatimtab}
+  }
+\end{example}
+
+\begin{example}
+  \label{example-share3-eval}
+  Let $\codebase,e_{\ref{example-share3}}$ be the $\metaprot$ program and let 
+  $\prog_{\ref{example-share3}}$ be the  $\minifed$ program defined
+  in Example \ref{example-share3}. We refer to the latter as ``accumulated''
+  by evaluation of the former in the sense that $\config{\varnothing}{e_{\ref{example-share3}}}
+  \redxs \config{\prog_{\ref{example-share3}}}{\ttt{()}}$.
+\end{example}
+
+\begin{example}
+  Given $\ttt{share3}$ as defined in Example \ref{example-share3} and
+  annotation:
+  $$\tas(\ttt{share3}) =  \ttt{cid(client)}\ *\ \ttt{string(sid)}$$
+  the type of $\ttt{share3}$ is $\tas(\ttt{share3}) \rightarrow \tau,\varnothing$
+  where $\tau$ is:
   $$
-  \sum_{\store \in \mems(X)} \pmf(\store) \  = \ 1
+  \ttt{\{ s1 : bool[client]; s2 : bool[client]; s3 : bool[client] \}}
   $$
-\end{definition}
-%To recover succinct and familiar notation, we may omit the domain of a
-%distribution when it is clear from an application context-
-%i.e., we allow the following sugaring:
-%$$
-%\pdf{}(\store) \defeq \pdf{\dom(\store)}(\store)
-%$$
-Now, we can define a notion of marginal and conditional
-distributions as follows, which are standard for discrete
-probability mass functions. 
-\begin{definition}
-  Given $\pmf$ the \emph{marginal distribution} of variables $X$
-  in $\pmf$, denoted $\margd{\pmf}{X}$, is defined as follows:
-  $$
-  \forall \store \in \mems(X) \quad . \quad \margd{\pmf}{X}(\store) =
-  \sum_{\store' \in \mems(X-\dom(\dom(\pmf)))} \pmf(\store \cap \store')
-  $$
-\end{definition}
+\end{example}
 
-\begin{definition}
-  Given $\margd{\pmf}{X}$, its \emph{conditional distribution given
-  $\stores$} with $\dom(\stores) = Y$ with $Y \subseteq X$, denoted
-  $\condd{\pmf}{X}{\stores}$, is a distribution with domain $X$ where for all
-  $\store \in \mems(X)$:
-  $$
-  \condd{\pmf}{X}{\stores}(\store) =
-  (\sum_{\store' \in \stores} \margd{\pmf}{X}(\store \cap \store')) /
-  (\sum_{\store' \in \stores} \margd{\pmf}{Y}(\store'))
-  $$
-  where $\margd{\pmf}{X}(\store \cap \store')) = 0$ if $\store \cap \store'$ is undefined.
-\end{definition}
-To recover familiar notation we allow the following syntactic
-sugarings:
-\begin{eqnarray*}
-  \condd{\pmf}{X}{\store}  &\defeq& \condd{\pmf}{X}{\{ \store\}}\\
-  \pmf(\store)  &\defeq& \margd{\pmf}{X}(\store)  \qquad \dom(\store) = X\\
-  \pmf(\store|\stores)  &\defeq& \condd{\pmf}{X}{\stores}(\store) \qquad \dom(\store) = X
-\end{eqnarray*}
-
-\subsection{Program Distributions}
-Now we can define the probability distribution of a program $\prog$,
-that we denote $\progd(\prog)$. Since $\fedcat$ is deterministic the
-results (view assignments) of any run are determined by the input
-secrets together with the random tape. And since we constrain programs
-to not overwrite views, we are assured that \emph{final} memories
-contain both a complete record of all initial secrets as well as views
-resulting from communicated information.
-
-We consider random tapes to be selected from a uniform distribution.
-In our threat model, we will also assume that any possible initial
-secrets are equally likely when formulating the program distribution.
-In this setting, given a program $\prog$ with $\iov(\prog) = S \cup
-V$ and $\flips(\prog) = F$ we will consider all $\store
-\in \mems(S \cup V \cup F)$ such that
-$
-\config{\store_{S \cup F}}{\prog} \redxs \config{\store_}{\varnothing}
-$
-to be equally probable, establishing the basic distribution of the
-program. From this, we can immediately derive the marginal distribution
-of $S \cup V$ to reason about dependencies between secrets and views. 
-\begin{definition}
-  \label{def-progd}
-  \label{definition-progd}
-  Given program $\prog$ with $\iov(\prog) = (S,V)$ and $\flips(\prog) = F$, define $\stores$ as:
-  $$
-  \stores \defeq \{ \store \mid (\dom(\store) = S \cup V \cup F) \wedge (\store_{S \cup F},\prog) \redxs (\store,\varnothing) \}
-  $$
-  We note immediately that $|\stores| = 2^{S \cup F}$ given the (deterministic) definition of $\redx$. 
-  Define also $\pmf$ as the program's \emph{basic distribution} such that for all
-  $\store \in \mems(S \cup V \cup F)$:
-  $$
-  \pmf(\store) =
-  \begin{cases}
-    1 / |\stores| & \text{if}\ \store \in \stores\\
-    0 & \text{otherwise}
-  \end{cases}
-  $$
-  Then the \emph{program distribution of $\prog$}, denoted $\progd(\prog)$, is the
-  marginal distribution of $S \cup V$ in $\prog$'s basic distribution:
-  $$
-  \progd(\prog) =  \margd{\pmf}{S\cup V}
-  $$
-  In some cases, we will also be concerned with the (joint)
-  probabilities of expression interpretation given a preceding program
-  execution, and we write $\progd(\prog, \be)$ to denote the program
-  distribution $\progd(\prog;\eassign{\itv}{\be})$ where $\itv$ is a
-  special variable that is never used in programs.
-\end{definition}
-
-\subsection{Examples}
+\subsection{Program Distribution Examples}
 \label{section-pmf-examples}

### 2023-07-11 popl24/metalang.tex
checkpoint

@@
 \label{section-metalang}
-
-\begin{fpfig}[t]{Syntax of $\metaprot$.}{fig-metaprot-syntax}
-  {\small
-    $$
-    \begin{array}{rcl@{\hspace{8mm}}r}
-      \multicolumn{3}{l}{\flab \in \mathrm{Field},\   x \in \mathrm{EVar}, \  f \in \mathrm{FName}}\\[2mm]
-      %x &\in& \mathrm{EVar}\\
-      %f &\in& \mathrm{FName}\\[2mm]
-      e &::=& b \mid \flip{e}{e} \mid \secret{e}{e} \mid \view{e}{e} \mid \oracle{e} \mid \enot\ e \mid e\ \eand\ e \mid e\ \exor\ e \mid & \textit{expressions}\\[0mm]
-      & & \select{e}{e}{e} \mid 
-      \send{\view{e}{e}}{e} \mid \send{\view{e}{e}}{\OT{e}{e}{e}} \mid e;e \mid \\[0mm]
-      & & x \mid \elet{x}{e}{e} \mid f(e,\ldots,e) \mid \{ \flab = e; \ldots; \flab = e \}
-      \mid e.\flab \mid e\concat e \mid (e) \\[2mm]
-      v &::=& w \mid \cid \mid \be \mid \{ \flab = v;\ldots;\flab = v \} 
-      \mid \ttt{()} & \textit{values}\\[2mm]
-           {fn} &::=& f(x,\ldots,x) \{ e \} & \textit{functions}
-    \end{array}
-    $$
-  }
-\end{fpfig}
 
@@
 The syntax of $\metaprot$ is defined in Figure
-\ref{fig-metaprot-syntax}.  It includes a syntax of function
+\ref{fig-metaprot}.  It includes a syntax of function
 definitions and records, and values include client ids, identifier
@@
 definitions can generalize over $\fedprot$-level patterns to obtain
-composable program units. \cnote{Cur the following example to save
-  space?.}  As a simple example, consider 3 party secret sharing as
-illustrated in Example \ref{example-he}. We can define a function
-$\ttt{share3}$ that abstracts the process of splitting a given
-client's secret into 3 separate shares.
-\begin{example} \label{example-share3} The function $\ttt{share3}$ 
-  splits a client's secret into 3 shares returned as a record
-  with fields $\ttt{s1-3}$:
-  {\small
-  \begin{verbatimtab}
-      share3(client, secretid) {
-        let s1 = flip[client, "share1"] in
-        let s2 = flip[client, "share2"] in
-        let s3 = (s1 xor s2) xor s[client, "s:" || secretid] in
-        {s1 = s1;s2 = s2;s3 = s3}
-      } \end{verbatimtab}
-  }
-  Here is a $\metaprot$ program that uses this function definition:
-  {\small
-    \begin{verbatimtab}
-      let shares = share3(1, "mysecret") in
-      v[2,s1] := shares.s2;
-      v[3,s1] := shares.s3 \end{verbatimtab}
-  }
-  which generates the following $\minifed$ program, as we formalize in Example \ref{example-share3-eval}
-  below:
-  {\small
-  \begin{verbatimtab}
-      v[2,s1] := flip[1, "share2"];
-      v[3,s1] := flip[1, "share1"] xor flip[1, "share2"] xor s[1, "s:mysecret"] \end{verbatimtab}
-  }
-\end{example}
+composable program units.
 
@@
 
-\begin{fpfig}[t]{Evaluation contexts and operational semantics of $\metaprot$.}{fig-metaprot-semantics}
+\begin{fpfig}[t]{$\metaprot$ syntax (T), evaluation contexts (M), and operational semantics (B).}{fig-metaprot}
+    {\small
+    $$
+    \begin{array}{rcl@{\hspace{8mm}}r}
+      \multicolumn{3}{l}{\flab \in \mathrm{Field},\   x \in \mathrm{EVar}, \  f \in \mathrm{FName}}\\[2mm]
+      %x &\in& \mathrm{EVar}\\
+      %f &\in& \mathrm{FName}\\[2mm]
+      e &::=& b \mid \flip{e}{e} \mid \secret{e}{e} \mid \view{e}{e} \mid \oracle{e} \mid \enot\ e \mid e\ \eand\ e \mid e\ \exor\ e \mid & \textit{expressions}\\[0mm]
+      & & \select{e}{e}{e} \mid 
+      \send{\view{e}{e}}{e} \mid \send{\view{e}{e}}{\OT{e}{e}{e}} \mid e;e \mid \\[0mm]
+      & & x \mid \elet{x}{e}{e} \mid f(e,\ldots,e) \mid \{ \flab = e; \ldots; \flab = e \}
+      \mid e.\flab \mid e\concat e \mid (e) \\[2mm]
+      v &::=& w \mid \cid \mid \be \mid \{ \flab = v;\ldots;\flab = v \} 
+      \mid \ttt{()} & \textit{values}\\[2mm]
+           {fn} &::=& f(x,\ldots,x) \{ e \} & \textit{functions}
+    \end{array}
+    $$
+    }
+
+    \rule{130mm}{0.5pt}
+
     {\small
@@
     $$
-    \medskip
+    \vspace{.5mm}
+
+    \rule{130mm}{0.5pt}
+
     $$
@@
 We define a small-step evaluation aka reduction relation $\redx$ in
-Figure \ref{fig-metaprot-semantics}.  We write $\redxs$ to denote the
+Figure \ref{fig-metaprot}.  We write $\redxs$ to denote the
 reflexive, transitive closure of $\redx$. Reduction is defined on
@@
 and instantiation of compositional program elements.
-\begin{example}
-  \label{example-share3-eval}
-  Let $\codebase,e_{\ref{example-share3}}$ be the $\metaprot$ program and let 
-  $\prog_{\ref{example-share3}}$ be the  $\minifed$ program defined
-  in Example \ref{example-share3}. We refer to the latter as ``accumulated''
-  by evaluation of the former in the sense that $\config{\varnothing}{e_{\ref{example-share3}}}
-  \redxs \config{\prog_{\ref{example-share3}}}{\ttt{()}}$.
-\end{example}
 
@@
 and for any $\cidty{e}$ or $\bet{e}$ the expression $e$
-is either a variable or a client id $\cid$.
+is either a variable or a client id $\cid$. We define
+the equivalence $w_1 \concat w_2 \equiv w_1w_2$ in typings.
 
@@
 structure of types and expressions.
-\begin{example}
-  Given $\ttt{share3}$ as defined in Example \ref{example-share3} and
-  annotation:
-  $$\tas(\ttt{share3}) =  \ttt{cid(client)}\ *\ \ttt{string(sid)}$$
-  the type of $\ttt{share3}$ is $\tas(\ttt{share3}) \rightarrow \tau,\varnothing$
-  where $\tau$ is:
-  $$
-  \ttt{\{ s1 : bool[client]; s2 : bool[client]; s3 : bool[client] \}}
-  $$
-\end{example}
 

### 2023-07-11 popl24/minicat.tex
checkpoint

@@
 \subsection{Syntax} The syntax of $\minifed$, defined in
-Figure \ref{fig-minifed-syntax}, includes a standard boolean algebra
+Figure \ref{fig-minifed}, includes a standard boolean algebra
 with $\eand$, $\eor$, $\exor$, and $\enot$ as primitives. In addition,
@@
 and $\prog_2$.
-
-\begin{fpfig}[t]{$\minifed$ source code syntax.}{fig-minifed-syntax}
-  {\small
-    $$
-    \begin{array}{rcl@{\hspace{8mm}}r}
-      \multicolumn{4}{c}{b \in \{ \etrue, \efalse \},\  w \in \mathrm{String},\ \cid \in \mathrm{Clients} \subset  \mathbb{N},\ \bop \in \{ \eand, \eor, \exor \}} \\[2mm]
-      \be &::=& b \mid \flip{\cid}{w} \mid \secret{\cid}{w} \mid \view{\cid}{w} \mid \oracle{w} \mid & \textit{boolean expressions}\\
-      & &  \enot\ \be \mid \be\ \bop\ \be \mid \select{\be}{\be}{\be} \mid \OT{\be}{\be}{\be} \\[2mm]
-      \instr &::=& \eassign{\view{\cid}{w}}{\be} & \textit{view assignments} \\[2mm]
-      \prog &::=& \varnothing \mid \instr; \prog & \textit{protocols}
-    \end{array}
-    $$
-  }
-\end{fpfig}
 
@@
 variables with values, and all variables must be owned by client
-$\cid$. This is defined in Figure \ref{fig-minifed-interp}.
+$\cid$. This is defined in Figure \ref{fig-minifed}.
 
-\begin{fpfig}[t]{$\minifed$ expression interpretation (T) and reduction (B) rules.}{fig-minifed-interp}
+\begin{fpfig}[t]{$\minifed$ syntax (T), expression interpretation (M),  and reduction (B) rules.}{fig-minifed}
+  {\small
+    $$
+    \begin{array}{rcl@{\hspace{8mm}}r}
+      \multicolumn{4}{c}{b \in \{ \etrue, \efalse \},\  w \in \mathrm{String},\ \cid \in \mathrm{Clients} \subset  \mathbb{N},\ \bop \in \{ \eand, \eor, \exor \}} \\[2mm]
+      \be &::=& b \mid \flip{\cid}{w} \mid \secret{\cid}{w} \mid \view{\cid}{w} \mid \oracle{w} \mid & \textit{boolean expressions}\\
+      & &  \enot\ \be \mid \be\ \bop\ \be \mid \select{\be}{\be}{\be} \mid \OT{\be}{\be}{\be} \\[2mm]
+      \instr &::=& \eassign{\view{\cid}{w}}{\be} & \textit{view assignments} \\[2mm]
+      \prog &::=& \varnothing \mid \instr; \prog & \textit{protocols}
+    \end{array}
+    $$
+  }
+
+  \rule{130mm}{0.5pt}
+
   {\small
@@
 Evaluation of configurations is then defined via a small-step reduction relation $\redx$,
-also shown in Figure \ref{fig-minifed-interp}.
+also shown in Figure \ref{fig-minifed}.
 This is defined in the obvious manner for view assignments other than through
@@
 \end{example}
-
-\begin{example}[Additive Secret Sharing]
-    \label{example-he}
-Additive secret sharing is an MPC protocol where $k$ parties split up
-their secrets into \emph{shares}, with the property that all $k$
-shares are required to reconstruct the secret and any fewer reveals
-nothing. Additionally, this sharing enjoys additive homomorphism. In
-the binary field it is well-known that we can extend our use
-of $\ttt{xor}$ and flips to generate an arbitrary number of shares
-with additive homomorphism.
-
-For example, to generate shares for three-party additive sharing,
-clients 1,2, and 3 can break up and distribute their shares to
-each other as follows:
-{\small
-\begin{verbatimtab}
-     v[2,"s1"] := flip[1,"1"] xor flip[1,"s1"] xor s[1,"1"];
-     v[3,"s1"] := flip[1,"1"];
-     v[1,"s2"] := flip[2,"1"] xor flip[2,"s2"] xor s[2,"1"];
-     v[3,"s2"] := flip[2,"1"];
-     v[1,"s3"] := flip[3,"1"] xor flip[3,"s3"] xor s[3,"1"];
-     v[2,"s3"] := flip[3,"1"] \end{verbatimtab}
-}
-Then, assuming that party 0 is a ``public'' client, each party sums
-its own shares, and then the sum of the sum of shares is revealed
-in the output view $\vx{0}{output}$. By additive homomorphism of
-$\ttt{xor}$, this output is the sum of the 3 secrets.
-{\small
-\begin{verbatimtab}
-     v[0,"ss1"] := v[1,"s2"] xor v[1,"s3"] xor flip[1,"s1"];
-     v[0,"ss2"] := v[2,"s1"] xor v[2,"s3"] xor flip[2,"s2"];
-     v[0,"ss3"] := v[3,"s1"] xor v[3,"s2"] xor flip[3,"s3"];
-     v[0,"output"] := v[0,"ss1"] xor v[0,"ss2"] xor v[0,"ss3"] \end{verbatimtab}
-}
-We note that, as for MPC protocols generally, the notion of
-``secrecy'' here is subtle and not the same as for encryption as in
-Example \ref{example-otp}, particularly when the threat model allows
-protocol participants to be corrupt. For example, suppose that $1 \in C$, and
-$\sx{1}{1}$ is $1$, and after running the protocol suppose that
-$\vx{0}{output}$ is 0. Then the adversary knows that 2 and 3's secrets
-are either 1 and 0, or 0 and 1, with 50/50 probability, and definitely
-neither 0 and 0, nor 1 and 1. 
-\end{example}

### 2023-07-11 popl24/pmf.tex
checkpoint

@@
 \end{definition}
-
-\subsection{Examples}
-\label{section-pmf-examples}
-
-In the following, we will refer to program examples defined in
-Section \ref{section-minicat-examples} as $\prog_x$, where $x$
-is the Example number.
-
-\paragraph{Example \ref{example-otp} revisited}
-%For small programs it is
-%helpful to visualize their basic distributions graphically as their
-%truth tables, where each row corresponds to an equally likely final
-%memory for the given program. In Figure \ref{fig-basic-distributions}
-%we show this for Example \ref{example-otp}. In this truth table we can
-%see that $\sx{1}{0}$ and $\vx{2}{2}$ are perfectly correlated.
-In this example, 
-from the adversarial view, $\sx{1}{0}$ is 1 or 0 with equal
-probability given any value of $\vx{0}{0}$. More precisely, for any
-$\beta_1,\beta_2$ we have:
-\begin{mathpar}
-(\progd(\prog_{\ref{example-otp}}))(\{\sx{1}{0} \mapsto \beta_1 \}|\{ \vx{0}{0} \mapsto \beta_2 \}) = .5
-\end{mathpar}
-Note that this corresponds to \emph{perfect secrecy} as described
-in \cite{barthe2019probabilistic}.
-
-\paragraph{Example \ref{example-lambda-obliv}(a) revisited} As demonstrated
-by Example \ref{example-lambda-obliv} taken from \cite{darais2019language}, it is possible
-to leak information probabilistically through observable dependencies
-in views.
-%In the basic distribution for Example \ref{example-lambda-obliv}(a)
-%illustrated in Figure \ref{fig-basic-distributions},
-We can see that
-the adversarial views $\vx{0}{0}$ and $\vx{0}{1}$ are both 1 in
-three different runs of the protocol, and in 2 out of 3 the
-honest secret $\sx{1}{0}$ is 1. More precisely:
-$$
-(\progd(\prog_{\ref{example-lambda-obliv}(a)}))(\{\sx{1}{0} \mapsto 1 \}|\{ \vx{0}{0} \mapsto 1, \vx{0}{1} \mapsto 1 \}) = 2/3
-$$
-An important point illustrated by this example is that the
-probability of values in views, and not just the values themselves,
-can release information to the adversary.
-
-\paragraph{Example \ref{example-he} revisited} This standard example
-of an MPC secure protocol illustrates how a functionality- in this
-case addition (in the binary field)- can be implemented securely. As
-mentioned previously, it is \emph{not} the case that \emph{no}
-information about input secrets can be revealed to the adversary in
-the MPC threat model, since the result of the functionality can itself
-reveal information. For example, if the result of addition is $0$,
-the adversary knows that two of the clients' secrets must be 0
-and the third's must be 1, though not whose is whose- so we can
-observe:
-\begin{mathpar}
-  (\progd(\prog_{\ref{example-he}}))(\{\sx{1}{1} \mapsto 0, \sx{2}{1} \mapsto 1, \sx{3}{1} \mapsto 0 \}|\{ \vx{0}{output} \mapsto 0 \}) = .33
-\end{mathpar}
-Further, since the standard MPC threat model allows participating
-clients to be corrupted we can formally consider the impacts on
-information released to the adversary by conditioning the program
-distribution on corrupt secrets.  So if we assume client 2
-with input secret 1 is corrupted, and we assume the output is 0, then
-the adversary knows with certainty that clients 1 and 3 both have input
-secrets 0. We can express this as:
-$$
-\progd(\prog_{\ref{example-he}}))(\{\sx{1}{1} \mapsto 0, \sx{3}{1} \mapsto 0 \}|\{ \sx{2}{1} \mapsto 1, \vx{0}{output} \mapsto 0 \})
-= 1
-$$ A critical point to note is that this information revealed to the
-adversary does not violate MPC security- under the assumption of
-public release of the functionality result, and under the assumption
-of the possible corruption of clients, this is the cost of doing
-business. Rather, the concern is that the protocol does not release
-any additional information through information communicated to views
-of corrupted parties. Additive secret sharing as realized in
-$\prog_{\ref{example-he}}$ is MPC secure in this sense- so for
-example, for any $\beta_1$ and $\beta_2$ we have:
-$$
-\begin{array}{c}
-\progd(\prog_{\ref{example-he}}))(\{\sx{1}{1} \mapsto 0 \}|\{ \sx{2}{1} \mapsto 0, \vx{0}{output} \mapsto 0 \})
-= \\
-\progd(\prog_{\ref{example-he}}))(\{\sx{1}{1} \mapsto 0 \}|\{ \vx{2}{s1} \mapsto \beta_1, \vx{2}{s3} \mapsto \beta_2, \sx{2}{1} \mapsto 0, \vx{0}{output} \mapsto 0 \})
-= \\
-.5
-\end{array}
-$$

### 2023-07-11 popl24/automation.tex
Update automation.tex

@@
 Since we certify properties of gates in isolation with witness inputs,
-our metatheory is concerned with preservation of important invariants
+our metatheory is concerned with the preservation of important invariants
 when certified components are ``spliced'' into larger program

### 2023-07-11 popl24/ygc.tex
add quotes ygc.tex

@@
 \begin{verbatimtab}
-  keygen(gid, b1, b2) { select4(b1,b2,H[gid || 1],H[gid || 2],H[gid || 3],H[gid || 4]) }
+  keygen(gid, b1, b2) { select4(b1,b2,H[gid || "1"],H[gid || "2"],H[gid || "3"],H[gid || "4"]) }
 
@@
   {   
-    v[1, gate: || gid || tid || 1] := table.v1;
-    v[1, gate: || gid || tid || 2] := table.v2;
-    v[1, gate: || gid || tid || 3] := table.v3;
-    v[1, gate: || gid || tid || 4] := table.v4
+    v[1, gate: || gid || tid || "1"] := table.v1;
+    v[1, gate: || gid || tid || "2"] := table.v2;
+    v[1, gate: || gid || tid || "3"] := table.v3;
+    v[1, gate: || gid || tid || "4"] := table.v4
   }
@@
     let r0 = (not wl.k) xor false in
-    v[1,OUTtt1] := select(wl.p,r1,r0);
-    v[1,OUTtt2] := select(not wl.p,r1,r0)
+    v[1,"OUTtt1"] := select(wl.p,r1,r0);
+    v[1,"OUTtt2"] := select(not wl.p,r1,r0)
   }
 
-  evaldecode(wv) { wv.k xor select(wv.p,v[1,OUTtt1],v[1,OUTtt2]) }
+  evaldecode(wv) { wv.k xor select(wv.p,v[1,"OUTtt1"],v[1,"OUTtt2"]) }
 
@@
     let ct = select4(wva.p,wvb.p,
-               v[1,g || tt1],v[1,g || tt2],v[1,g || tt3],v[1,g || tt4]) in
+               v[1,g || "tt1"],v[1,g || "tt2"],v[1,g || "tt3"],v[1,g || "tt4"]) in
     let cp = select4(wva.p,wvb.p,
-               v[1,g || pt1],v[1,g || pt2],v[1,g || pt3],v[1,g || pt4]) in
+               v[1,g || "pt1"],v[1,g || "pt2"],v[1,g || "pt3"],v[1,g || "pt4"]) in
     { k = k xor ct; p = k xor cp }

### 2023-07-11 popl24/intro.tex
add related work for PPLs

@@
 
+Our work also shares many ideas with probabilistic programming
+languages designed to perform (exact or approximate)
+inference~\cite{holtzen2020scaling, carpenter2017stan, wood2014new,
+  bingham2019pyro, albarghouthi2017fairsquare, de2007problog,
+  pfeffer2009figaro, saad2021sppl}. Our setting, however, requires
+verifying properties beyond inference, including conditional
+independence. Recent work by Li et al.~\cite{li2023lilac} proposes a
+manual approach for proving such properties, but does not provide
+automation.
+
 \subsection{Overview and Contributions}

### 2023-07-11 popl24/automation.tex
checkpoint

--- 2023-07-11 popl24/automation.tex
@@
 
-\subsection{Brute Force Computation of Program Distributions}
-\label{section-bruteforce}
+%We first explore a brute force approach to automatically verifying
+%program behaviour, the first step of which is to compute the basic
+%distribution of a program $\prog$. The program distribution
+%$\progd(\prog)$ can obviously be automatically derived from that, as
+%can be any of its marginal or conditional distributions. In this
+%Section, we consider two automated mechanisms for deriving
+%$\progd(\prog)$ in this way. The first is through a straightforward
+%computation of all final program memories.  The second approach
+%rewrites $\minifed$ programs to stratified Datalog programs, that are
+%amenable HPC optimizations such as parallelization and GPU matrix
+%computations as shown in recent work
+%\cite{sakama2017linear,aspis2018linear,nguyen2022enhancing,nguyen2021efficient}.
 
-We first explore a brute force approach to automatically verifying
-program behaviour, the first step of which is to compute the basic
-distribution of a program $\prog$. The program distribution
-$\progd(\prog)$ can obviously be automatically derived from that, as
-can be any of its marginal or conditional distributions. In this
-Section, we consider two automated mechanisms for deriving
-$\progd(\prog)$ in this way. The first is through a straightforward
-computation of all final program memories.  The second approach
-rewrites $\minifed$ programs to stratified Datalog programs, that are
-amenable HPC optimizations such as parallelization and GPU matrix
-computations as shown in recent work
-\cite{sakama2017linear,aspis2018linear,nguyen2022enhancing,nguyen2021efficient}.
+%Here we describe a straightforward technique for obtaining program
+%distributions $\progd(\prog)$ through direct computation of
+%$\runs(\prog)$. Although we could obtain $\runs(\prog)$ simply by
+%executing $\prog$ given all random tapes and input secrets, the method
+%we describe here is empirically more efficient (though still
+%exponential in the length of the random tape), and also supports the
+%translation to Datalog described in Section
+%\ref{section-bruteforce-datalog}. An implementation of these
+%techniques along with Examples from Sections
+%\ref{section-minicat-examples} and \ref{section-metalang-ygc} is
+%available online\footnote{URL redacted for double-blind review.}.
+
+Our first step in automating security analysis is to show how to
+compute the basic program distribution. Recall that basic
+distributions are based on the memories resulting from execution of an
+$\minicat$ program, which contain a record of all client views and the
+public output.  Given $\prog$ we denote this as $\runs(\prog)$, and we
+denote the basic distribution of a program as $\progtt(\prog)$.
+\begin{definition}
+  Given $\prog$ with $\iov(\prog) = S \cup V$ and $\flips(\prog) = F$:
+  $$
+  \runs(\prog) \defeq \{ \store \in \mems(S\cup F \cup V) \mid \config{\store_{S \cup F}}{\prog} \redxs \config{\store}{\varnothing} \}
+  $$
+  We write $\progtt(\prog)$ to denote the \emph{basic distribution} of
+  $\prog$, where given $\iov(\prog) = S \cup V$ and $\flips(\prog) = F$
+  we have for all $\store \in \mems(S \cup V \cup F)$:
+  $$
+  \progtt(\prog)(\store) =  1 / 2^{|S\cup F|} \ \text{if}\ \store \in \runs(\prog), \text{otherwise}\ 0
+  $$
+\end{definition}
+
+Any marginalization or conditioning of $\progtt(\prog)$ can be easily
+obtained algorithmically, in particular $\progd(\prog) =
+\margd{\progtt(\prog)}{\iov(\prog)}$.  As already discussed in Section
+\ref{section-hyperprop} and as we will discuss more below, extensional
+and intensional security properties are predicated on marginalizations
+and conditionings of basic distributions. So calculating
+$\runs(\prog)$ establishes the foundation for our analysis.
 
@@
 
-%The basic distribution of a program $\prog$ is based on the
-%final memories resulting from any run of $\prog$, denoted
-%$\runs(\prog)$.
-%\begin{definition}
-%  Given $\prog$ with $\iov(\prog) = S \cup V$ and $\flips(\prog) = F$:
-%  $$
-%  \runs(\prog) \defeq \{ \store \in \mems(S\cup F \cup V) \mid \config{\store_{S \cup F}}{\prog} \redxs \config{\store}{\varnothing} \}
-%  $$
-%\end{definition}
+Intuitively, we can represent $\runs(\prog)$ in tabular form, as
+illustrated in Figure \ref{fig-basic-distributions} for Examples
+\ref{example-otp} and \ref{example-lambda-obliv}(a). So a basic
+strategy is to directly compute this truth table for any given
+program. We first describe an approach where we compute
+$\runs(\instr_1,\ldots,\instr_n)$ iteratively for the subprograms
+$\instr_1,\ldots,\instr_i$ for $1 \le i \le n$, by individually
+computing the truth table of expressions $\be_i$ for each $\instr_i =
+(\eassign{v_i}{\be_i})$, and joining this with
+$\runs(\instr_1,\ldots,\instr_{i-1})$.
 
-As we've observed, $\runs(\prog)$ can be represented in a truth table
-form, as illustrated in Figure \ref{fig-basic-distributions}.
+In Figure \ref{fig-solve} we define the algorithm
+$\solve{\stores}{\be}$ which filters a given $\stores$ to obtain the
+subset whose elements satisfy $\be$ (make it true). It's correctness
+is characterized as follows.
+\begin{lemma}
+  \label{lemma-solves}
+  For all $\stores$ and $\be$ with $\vars(\be) \subseteq \dom(\stores)$,
+  $(\solve{\stores}{\be}) = \{ \store \in \stores \ \mid\ \lcod{\store,\be}{\cid} = 1 \}$
+  for some $\cid$.
+\end{lemma}
+\begin{proof}
+  Since we assume safety of programs we can assume that all variables in $\be$ have the
+  same owner $\cid$. The result otherwise follows in a straightforward manner by induction
+  on $\be$. 
+\end{proof}
 
@@
 
-Here we describe a straightforward technique for obtaining program
-distributions $\progd(\prog)$ through direct computation of
-$\runs(\prog)$. Although we could obtain $\runs(\prog)$ simply by
-executing $\prog$ given all random tapes and input secrets, the method
-we describe here is empirically more efficient (though still
-exponential in the length of the random tape), and also supports the
-translation to Datalog described in Section
-\ref{section-bruteforce-datalog}. An implementation of these
-techniques along with Examples from Sections
-\ref{section-minicat-examples} and \ref{section-metalang-ygc} is
-available online\footnote{URL redacted for double-blind review.}.
-
-In Figure \ref{fig-solve} we define the algorithm $\solve{\stores}{\be}$
-which filters $\stores$ based on satisfaction of $\be$. This algorithm
-is fundamental to our brute force methods. It's correctness is characterized
-as follows. 
-\begin{lemma}
-  \label{lemma-solves}
-  For all $\stores$ and $\be$ with $\vars(\be) \subseteq \dom(\stores)$,
-  $(\solve{\stores}{\be}) = \{ \store \in \stores \ \mid\ \lcod{\store,\be}{\cid} = 1 \}$
-  for some $\cid$.
-\end{lemma}
-\begin{proof}
-  Since we assume safety of programs we can assume that all variables in $\be$ have the
-  same owner $\cid$. The result otherwise follows in a straightforward manner by induction
-  on $\be$. 
-\end{proof}
-
-In order to automatically obtain $\runs(\prog)$, we can do a left
-folding of $\solve$ across $\prog$, inductively extending memories
-with valid view assignments in the order of their definition.
-We can then use this to compute $\progd(\prog)$. We denote
-these computations as $\cruns$ and $\cprogd$ respectively.
+To obtain $\runs(\prog)$, we can perform a left folding of $\solve$
+across $\prog$ and each view definition, inductively extending
+memories with valid view assignments in the order of their definition.
+We denote this computation as $\cruns(\prog)$.
 \begin{lemma}
@@
 \end{proof}
+The above yields an algorithm for computing basic distributions and
+thus program distributions. 
 \begin{lemma}
@@
   Given $\prog$ where $\iov(\prog) = S \cup V$ and $\flips(\prog) = F$. Define
-  $\cprogd(\prog)$ as the distribution $\pmf$
+  $\cprogtt(\prog)$ as the distribution $\pmf$
   where for all $\store \in \mems(S\cup F \cup V)$:
   $$
-  \pmf(\store) = \begin{cases}1/2^{|S \cup F|} & \text{if\ } \store \in
-    \cruns(\prog) \\ 0 & \text{otherwise} \end{cases} 
+  \pmf(\store) = 1 / 2^{|S\cup F|} \ \text{if}\ \store \in \cruns(\prog), \text{otherwise}\ 0
+  %\begin{cases}1/2^{|S \cup F|} & \text{if\ } \store \in
+  %  \cruns(\prog) \\ 0 & \text{otherwise} \end{cases} 
   $$
-  Then:
-  $$
-  \cprogd(\prog) = \progd(\prog)
-  $$
+  Then $\cprogtt(\prog) = \progtt(\prog)$ and $\margd{\cprogtt(\prog)}{S\cup V} = \progd(\prog)$.
 \end{lemma}
@@
 \end{proof}
+A full implementation of these methods is available in our online
+repository\footnote{Link redacted for double-blind reviewing.}. To
+avoid notational confusion we will continue to refer to $\progtt$ and
+$\cruns$ in our subsequent discussion of automated verification but
+with the understanding that both can be computed using the above
+algorithms.
 
-\subsection{Protocol Certification}
+Since $|\runs(\prog)|$ grows exponentially in the size of
+$\flips(\prog)$, scalability to larger programs is a concern.  However
+computation of $\runs$ is amenable to optimizations, especially
+parallelization- rather than computing the truth table
+``horizontally'', it can be computed vertically. We can rewrite
+$\minifed$ programs $\prog$ to stratified Datalog programs, where
+$\runs(\prog)$ is equivalent to set of least Herbrand models obtained
+by representing different random tapes and input secrets as different
+fact bases. Details of this rewrite are provided in Appendix
+\ref{section-bruteforce}. The model for each fact base can be computed
+independently and in parallel, with the potential to apply other
+matrix optimizations developed in recent work on HPC for Datalog
+\cite{sakama2017linear,aspis2018linear,nguyen2022enhancing,nguyen2021efficient}.
+
+\subsection{Certification of Extensional Properties}
 
@@
 \end{definition}
-
-We base certification of protocols $\prog$ on a calculation of its
-program distribution. Recall from Definition \ref{definition-progd} that
-program distributions are calculated from runs of a protocol given
-any random tape and input secrets. That is:
-\begin{definition}
-  Given $\prog$ with $\iov(\prog) = S \cup V$ and $\flips(\prog) = F$, define:
-  $$
-  \runs(\prog) \defeq \{ \store \in \mems(S\cup F \cup V) \mid \config{\store_{S \cup F}}{\prog} \redxs \config{\store}{\varnothing} \}
-  $$
-\end{definition}
-
-Given any $\prog$, the set $\runs(\prog)$ can be computed via brute
-force methods. One is to fold a standard interpretation of logical
-formulas across $\prog$, obtaining a truth table representation of
-$\runs(\prog)$. We can also interpret $\prog$ as a stratified Datalog
-program, with $\runs(\prog)$ being the set of least Herbrand models
-obtained by representing different random tapes and input secrets as
-different fact bases. This representation is amenable HPC
-optimizations such as parallelization and GPU matrix computations as
-shown in recent work
-\cite{sakama2017linear,aspis2018linear,nguyen2022enhancing,nguyen2021efficient}.
-We provide more details of brute force computation of $\runs(\prog)$ in
-Appendix \ref{section-bruteforce} and an implementation of
-the tabular method is available in our online repository\footnote{Link
-redacted for double-blind reviewing.}.
-
-Given a calculation of $\runs(\prog)$, we can easily calculate its
-basic distribution as in Definition \ref{definition-progd)}, as well
-as any marginalization (such as $\progd(\prog)$) or conditioning of it.
-
-\subsection{Certification of Extensional Properties}
 
@@
   honest, and $\prog_{\ref{example-otp}}$ satisfies probabilistic
-  noninterference and hence perfect secrecy as defined in
-  \ref{barthe2019probabilistic}.
+  noninterference and hence perfect secrecy (as defined in
+  \cite{barthe2019probabilistic}).
 \end{lemma}
@@
 
-Our approach is thus to design gate certifications that demonstrably
+Our solution is to design gate certifications that demonstrably
 guarantee security for circuits built from them. Intuitively, we can
@@
 defined in subprotocols. 
-\begin{definition}
-  We write $\progtt(\prog)$ to denote the \emph{basic distribution} of
-  $\prog$ as defined in Definition \ref{def-progd}. For any program $\prog =
-  (\eassign{v_1}{\be_1};\ldots;\eassign{v_n}{\be_n})$, define
-  $\vdefs(\prog) \defeq \{ v_1,\ldots,v_n \}$.
-\end{definition}
 Notions of probabilistic independence, aka \emph{separation}, as well
@@
 when certified components are ``spliced'' into larger program
-contexts.  One important reasoning principle is related to
-noninterference properties- specifically, if we show that gate views
-enjoy a noninterference property wrt their free (input) variables,
+contexts. This technique often requires us to focus on views defined
+in subprograms, so we define the following.
+\begin{definition}
+For any program $\prog =
+  (\eassign{v_1}{\be_1};\ldots;\eassign{v_n}{\be_n})$, define
+$\vdefs(\prog) \defeq \{ v_1,\ldots,v_n \}$.
+\end{definition}
+
+An important reasoning principle is related to component
+noninterference- specifically, if we show that gate views enjoy a
+noninterference property wrt their free (input) variables,
 probabilistic independence propagates to other variables in the

### 2023-07-11 popl24/composition.tex
checkpoint

@@
 \section{Automating Verification of Circuit Security}
+\label{section-composition}
 
@@
 
+\begin{comment}
+
 \begin{fpfig}[t]{YGC copy gate definitions.}{fig-ygc-copy}
@@
 in other related systems for enforcing obliviousness \cite{darais2019language}.
+
+\end{comment}

### 2023-07-11 popl24/metalang.tex
checkpoint

@@
 The type syntax of $\metaprot$ is defined in Figure
-\ref{fig-metaprot-tsyntax}. It includes a weak form of dependency:
+\ref{fig-metaprot-types}. It includes a weak form of dependency:
 string types $\stringty{e}$ and client types $\cidty{e}$
@@
 syntax-directed- selected rules are shown in Figure
-\ref{fig-metaprot-tjudge}. The $\TirName{AssignT}$ rule
-captures the effect of new view definitions. The
-$\TirName{And}$ rule illustrates how program safety is enforced,
-by ensuring that subexpressions of boolean expressions have the
-same owner.
+\ref{fig-metaprot-types}
+%, with a more complete set provided in
+%Appendix Figure \ref{fig-metaprot-tjudge}
+and a full implementation is available in our online repository \cite{jpdf-github}. The
+$\TirName{AssignT}$ rule captures the effect of new view
+definitions. The $\TirName{And}$ rule illustrates how program safety
+is enforced, by ensuring that subexpressions of boolean expressions
+have the same owner.
 
-The $\TirName{FnT}$ and $\TirName{Appt}$ rules apply to function
-definition and application respectively, and rely on the
+The $\TirName{Appt}$ rule applies to function
+definition and application respectively. Function typings rely on the
 definition of function input type annotations $\tas$ and
@@
 
-\begin{fpfig}[t]{Type Syntax of $\metaprot$.}{fig-metaprot-tsyntax}
+\begin{fpfig}[t]{Type Syntax of $\metaprot$ (T) and selected type judgement rules.}{fig-metaprot-types}
 {\small
@@
 }
-\end{fpfig}
 
-\begin{fpfig}[t]{Selected $\metaprot$ type judgement rules.}{fig-metaprot-tjudge}
+\rule{130mm}{0.5pt}
+
 {\small
 \begin{mathpar}
-\inferrule[\TirName{VarT}]
-{}
-{\tjudge{\viewst}{\Gamma}{x}{\Gamma(x)}{\viewst}}
-
-\inferrule[\TirName{CidT}]
-{}
-{\tjudge{\viewst}{\Gamma}{\cid}{\cidty{\cid}}{\viewst}}
-
-\inferrule[\TirName{StringT}]
-{}
-{\tjudge{\viewst}{\Gamma}{w}{\stringty{w}}{\viewst}}
-
-\inferrule[\TirName{ConcatT}]
-{\tjudge{\viewst}{\Gamma}{e_1}{\stringty{e_1'}}{\viewst_1}\\
-\tjudge{\viewst_1}{\Gamma}{e_2}{\stringty{e_2'}}{\viewst_2}
-}
-{\tjudge{\viewst}{\Gamma}{e_1||e_2}{\stringty{e_1' ||e_2'}}{\viewst_2}}
-
-\inferrule[\TirName{BoolT}]
-{}
-{\tjudge{\viewst}{\Gamma}{\etrue}{\bet{\cid}}{\viewst}}
-
-\inferrule[\TirName{OracleT}]
-{\tjudge{\viewst}{\Gamma}{e}{\stringty{e'}}{\viewst'}}
-{\tjudge{\viewst}{\Gamma}{\oracle{e}}{\bet{\cid}}{\viewst'}}
-
 \inferrule[\TirName{SecretT}]
@@
 {\tjudge{\viewst}{\Gamma}{f(e_1,\ldots,e_n)}{\sigma(\tau)}{(\viewst_n ; \sigma(\viewst_f))}}
-
-\inferrule[FnT]
-{
-  \codebase(f) = x_1,\ldots,x_n,\ e \\ \tas(f) = \tau_1 * \cdots * \tau_n \\
-  \tjudge{\varnothing}{\Gamma; x_1 : \tau_1; \ldots; x_n : \tau_n}{e}{\tau}{\viewst_f}
-}
-{ \Gamma \vdash f : \tau_1 * \cdots * \tau_n \rightarrow \tau, \viewst_f }
-
-\inferrule[ProgT]
-{
-\forall f \in \dom(\codebase)\ .\ \Gamma \vdash f : \Gamma(f) \\ \Gamma \vdash e : \tau, \view{\cid_1}{w_1};\ldots;\view{\cid_1}{w_1}
-}
-{
-\Gamma \vdash \codebase,e : \tau,\{\view{\cid_1}{w_1}\} \sqcup \cdots \sqcup \{ \view{\cid_n}{w_n}\}
-}
 \end{mathpar}
@@
 
-\begin{fpfig}{Yao's Garbled Circuit component functions.}{fig-ygc-lib}
+\begin{fpfig}[t]{Yao's Garbled Circuit component functions.}{fig-ygc-lib}
 {\footnotesize
@@
 \ref{fig-ygc-aux} and \ref{fig-ygc-gates}, and complete code in our
-online repository.
+online repository \cite{jpdf-github}.
 \begin{example}

### 2023-07-11 popl24/intro.tex
checkpoint

@@
 verifying noninterference cannot naturally extend to MPC security. Our
-goal is to explore and establish connections between the security
-model of MPC-- the \emph{real/ideal} aka \emph{simulator security}
-model -- and trace-based hyperproperties, and to leverage these
-connections to obtain automated enforcement mechanisms for MPC
-protocol development. Currently, proof methods for MPC protocol
-development are well-studied \cite{Lindell2017} but manual and thus
-tedious and error-prone, and are also non-standardized and unfamiliar
-to most PL theorists. Therefore our exploration will make both
-theoretical and practical contributions by bridging a gap between
-information flow and simulator security methodologies.
+goal is to explore connections between the security model of MPC-- the
+\emph{real/ideal} aka \emph{simulator security} model -- and
+trace-based hyperproperties, and to leverage these connections to
+obtain automated enforcement mechanisms for MPC protocol
+development. Currently, proof methods for MPC protocol development are
+well-studied \cite{Lindell2017} with some previous work leveraging
+connections with hyperproperties \cite{8429300}. But these methods
+require fully manual proof construction and are thus tedious and
+error-prone. Therefore our exploration makes both theoretical and
+practical contributions.
 
@@
 
-\subsection{Related Work}
+\subsection{Overview and Contributions}
 
@@
 and interaction of complete secure protocols. By intensional, we mean
-analysis and inner workings of protocols themselves. While the
-probabilistic formulations we develop could conceivably be used
-extensionally, our main focus is on automated and semi-automated
-reasoning about intensional properties of protocols, both through
-whole-program analysis and compositional properties of program
-components that support security.
+analysis and inner workings of protocols themselves
+
+\paragraph{Language design.} In Section \ref{section-minicat} we
+develop a new probabilistic programming language $\minifed$ for
+defining synchronous distributed protocols over the binary field. The
+syntax and semantics provides a succinct account of \emph{views} and
+(synchronous) messaging between protocol \emph{clients}. In Section
+\ref{section-metalang} we define a new metaprogramming language
+$\metaprot$ that dynamically generates $\minifed$ protocols. It
+includes control and data structures and is able to express logical
+protocol components, and enjoys a type safety result that guarantees
+that generated protocols are semantically well-defined.
+
+\paragraph{Hyperproperty formulation.} In Section \ref{section-pmf} we
+formulate \emph{program distributions}, a formalism for expressing
+dependencies between secrets and views, useful for expressing
+intensional characteristics of protocols and precisely quantifying
+allowable information leakage in terms of protocol outputs. On this
+basis we can then formulate extensional hyperproperties, including
+\emph{noninterference modulo output ($\NIMO$)} that implies passive
+simulator security. But further, we can formulate intensional
+protocol properties such as probabilistic conditioning of program
+distributions.
+
+\paragraph{Fully and partially automated verification.} In Section
+\ref{section-automation} we develop an automated technique, aka
+certification, that verifies extensional properties of protocols such
+as $\NIMO$, and also intensional properties of protocol
+components. The latter include properties of protocol components that
+provably guarantee security through composition in large
+circuits. These compositional proofs are based on separation logic but
+include novel results to support our semi-automated method for large
+circuits where we (1) define certification for components, (2) certify
+components, (3) prove $\NIMO$ based on the certifiedness of
+components. We develop novel methods partially based on separation
+logic \cite{barthe2019probabilistic} to enable (3).  In Section
+\ref{section-composition} we illustrate our method to obtain
+semi-automated verification methods for the GMW and Yao's Garbled
+Circuit protocols, where certification can be re-used to validate new
+gate implementations.
+
+\subsection{Related Work}
+\label{section-related-work}
+
+Our main focus is on automated and semi-automated reasoning about
+intensional properties of protocols, both through whole-program
+analysis and compositional properties of program components that
+support security.
 
@@
 reduce proofs of simulator security to program equivalence assertions,
-which supports the application of known techniques in automated verification
-\cite{8429300}. However, they focus on noninterference properties,
-which are arguably sound but notably incomplete wrt MPC functionality
-as we discuss above and in Section
-\ref{section-hyperprop-ni}. Probabilistic Separation Logic
+which supports computer-aided verification of MPC protocols in
+EasyCrypt \cite{8429300}. This work also leveraged hyperproperty
+characterizations of simulator security, both passive and
+active. However, their EasyCrpyt security proof constructions are
+entirely manual. Probabilistic Separation Logic
 \cite{barthe2019probabilistic} develops a logical framework for
@@
 while intended for sound integration into MPC protocols, focus on
-expression and enforcement of obliviousness and do not account
-for allowable information leakage.
+expression and enforcement of obliviousness and do not account for
+allowable information leakage.
 
@@
 automation.
-
-\subsection{Overview and Contributions}
-
-\paragraph{Language design.} In Section \ref{section-minicat} we
-develop a new probabilistic programming language $\minifed$ for
-defining synchronous distributed protocols over the binary field. The
-syntax and semantics provides a succinct account of \emph{views} and
-(synchronous) messaging between protocol \emph{clients}. In Section
-\ref{section-metalang} we define a new metaprogramming language
-$\metaprot$ that dynamically generates $\minifed$ protocols. It
-includes control and data structures and is able to express logical
-protocol components, and enjoys a type safety result that guarantees
-that generated protocols are semantically well-defined.
-
-\paragraph{Hyperproperty formulation.} In Section \ref{section-pmf} we
-formulate \emph{program distributions}, a formalism for expressing
-dependencies between secrets and views, useful for expressing
-intensional characteristics of protocols and precisely quantifying
-allowable information leakage in terms of protocol outputs. On this
-basis we then formulate a novel hyperproperty of program execution
-traces, called \emph{noninterference modulo output ($\NIMO$)} that we
-demonstrate implies passive simulator security.
-
-\paragraph{Fully and partially automated verification.} In Section
-\ref{section-automation} we develop an automated technique, aka
-certification, that verifies extensional properties of protocols such
-as $\NIMO$, and also intensional properties of protocol
-components. The latter include properties of protocol components that
-provably guarantee security through composition in large
-circuits. These compositional proofs are based on separation logic but
-include novel results to support our semi-automated method for large
-circuits where we (1) define certification for components, (2) certify
-components, (3) prove $\NIMO$ based on the certifiedness of
-components. We develop novel methods partially based on separation
-logic \cite{barthe2019probabilistic} to enable (3).  In Section
-\ref{section-composition} we illustrate our method to obtain
-semi-automated verification methods for the GMW and Yao's Garbled
-Circuit protocols, where certification can be re-used to validate new
-gate implementations.

### 2023-07-11 popl24/main.tex
checkpoint

@@
 noninterference cannot naturally extend to MPC security.  Currently,
-proof methods for MPC protocols are well-studied but manual
-and thus tedious and error-prone, and are also non-standardized and
-unfamiliar to most PL theorists.  Our goal is to strengthen
-connections between the security model of MPC and trace-based
-hyperproperties, and to leverage them to obtain automated
-enforcement mechanisms for MPC protocol development.  We define a
-novel hyperproperty that is provably sound for MPC security, as a
-foundation for fully and partially automated verification of passive
-MPC security in protocols including arbitrarily large YGC and GMW
-circuits.
+proof methods for MPC protocols are well-studied but manual and thus
+tedious and error-prone, and are also non-standardized and unfamiliar
+to most PL theorists.  Our goal is to leverage connections between the
+security model of MPC and trace-based hyperproperties to obtain
+automated proof methods for MPC protocol development.  We develop a
+language model with a tightly coupled notion of probabilistic program
+distributions, as a foundation for fully and partially automated
+verification of passive MPC security in protocols including
+arbitrarily large YGC and GMW circuits.
 \end{abstract}
@@
 
+\input{examples}
+
+%\input{types}
+
 \input{ygc}

### 2023-07-11 popl24/hyperprop.tex
checkpoint

@@
 program trace behavior \cite{8429300,wysstar} including probabilistic
-noninterference. This formulation of hyperproperties that are at least
-sound with respect to simulator security bridges a conceptual gap
-between foundational approaches to security, and also motivates and
-supports automated analysis.
-
-In this section, we formulate a new hyperproperty that we show is
-sound for passive security in the real/ideal model, called
-\emph{noninterference modulo output ($\NIMO$)}. We do not claim
-completeness with respect to passive security, but soundness
-guarantees that enforcement of $\NIMO$ in our analyses implies passive
-security. In Section \ref{section-bruteforce} we show how $\NIMO$ can
-be automatically enforced in a brute-force manner and in Section
-\ref{section-composition} we show how $\NIMO$ can be refined to obtain
-compositional properties supporting efficient, scalable enforcement of
-complex protocols that incorporate modular subprotocols, using
-both YGC and GMW circuits as examples where gates are
-subprotocols. The latter methodology demonstrates a particular
-advantage of our intensional formulation.
+noninterference. In this Section we show for to formulate extensional
+properties relevant to MPC passive security in our language model by
+direct appeal to program distributions. In particular,
+\emph{noninterference modulo output ($\NIMO$)} is shown to be sound
+for MPC passive security. In Section \ref{section-automation} we will
+show how $\NIMO$ can be automatically enforced directly via
+whole-program analysis, or by automatically verifying composable,
+intensional properties of circuit components.
 
@@
 noninterference. We can easily demonstrate this by using
-Example \ref{example-he} as a failure witness, where we'll
-assume $C = \{0,2\}$ and $H = \{1,3\}$:
+Example \ref{example-gmw-andcircuit} as a failure witness, where we'll
+assume $C = \{0,2\}$ and $H = \{1\}$:
 $$
-(\progd(\prog_{\ref{example-he}}))({\outv \mapsto 1}|\{\{\sx{1}{1} \mapsto 0, \sx{2}{1} \mapsto 1, \sx{3}{1} \mapsto 0 \} \})
-= 1
+(\progd(\prog_{\ref{example-gmw-andcircuit}}))({\outv \mapsto 1}|\{\{\sx{1}{s1} \mapsto 0, \sx{2}{s2} \mapsto 1 \} \})
+= 0
 $$
@@
 $$
-(\progd(\prog_{\ref{example-he}}))({\outv \mapsto 1}|\{\{\sx{1}{1} \mapsto 1, \sx{2}{1} \mapsto 1, \sx{3}{1} \mapsto 0 \} \})
-= 0
+(\progd(\prog_{\ref{example-gmw-andcircuit}}))({\outv \mapsto 1}|\{\{\sx{1}{s1} \mapsto 1, \sx{2}{s2} \mapsto 1 \} \})
+= 1
 $$
@@
 together with the assumption that $\idealf$ is pre-imageable, implies
-passive security of a protocol.
-\begin{theorem}
+passive security of a protocol. The proof is provided in Appendix
+\ref{section-proofs}, and follows by essentially the same argument as
+for the analogous property of ``perfect passive NI security'' in
+\cite{8429300}.  
+\begin{restatable}{theorem}{nimosecure}
   \label{theorem-nimo}
@@
   then $\prog$ is passive secure.
-\end{theorem}
-
-\begin{proof}
-  Let $H$ and $C$ be an arbitrary partition of the federation and
-  suppose $\prog$ satisfies noninterference modulo output. Let
-  $\iov(\prog) = (V,S)$ and $\flips(\prog) = F$ with output view $x
-  \in V$ and let $\store$ be an arbitrary element of $\mems(S \cup
-  F)$. Then the distribution of adversarial views in the real world
-  is, by definition,
-  $\condd{\progd(\prog)}{V_C}{\store_S}$.
-
-  The simulator is given both $\store_{S_C}$ and
-  $\idealf(\store_S)$.  The simulation $\SIM_C(\store_{S_C},
-  \idealf(\store_S))$ can be defined as follows. First, some $\store'
-  \in \ik(\idealf,\idealf(\store))$ is randomly chosen such that
-  $\store' =_C \store_S$, as is a random tape $\store'' \in
-  \mems(F)$\footnote{The real/ideal model allows consultation of a
-  ``Random Oracle''.}. Then, the run of $(\store' \cup \store'',
-  \prog)$ is evaluated in simulation, yielding $(\store^{\SIM},\varnothing)$.
-  The simulation returns $\store^{\SIM}_{V_C}$ as a result.
-
-  Now, since the random tape is selected in simulation from the same distribution
-  that we assume for the real world, after selection of $\store'$ the
-  probability that any particular $\store^{\SIM}_{V_C}$ is returned is by definition
-  $
-  \progd(\prog)(\store^{\SIM}_{V_C} | \store')
-  $.
-  Furthermore, since we assume that $\prog$ correctly implements $\idealf$, this
-  means
-  $
-  \store_S =_C \store' \wedge
-     \condd{\progd(\prog)}{\{ \outv \}}{\store_S} =
-     \condd{\progd(\prog)}{\{ \outv \}}{\store'}
-  $,
-  so by the definition of noninterference modulo output,
-  any choice of $\store'$ yields the same distribution $\condd{\progd(\prog)}{V_C}{\store_S}$.
-  Therefore by definition
-  $
-   \progd(\SIM_C(\store_{S_C},\idealf(\store))) = \condd{\progd(\prog)}{V_C}{\store_S}
-  $.
-\end{proof}
-
+\end{restatable}
 As for noninterference, we can show that $\NIMO$ is equivalent

### 2023-07-11 popl24/proofs.tex
checkpoint

@@
 \label{section-proofs}
+
+\nimosecure*
+
+\begin{proof}
+  Let $H$ and $C$ be an arbitrary partition of the federation and
+  suppose $\prog$ satisfies noninterference modulo output. Let
+  $\iov(\prog) = (V,S)$ and $\flips(\prog) = F$ with output view $\outv
+  \in V$ and let $\store$ be an arbitrary element of $\mems(S \cup
+  F)$. Then the distribution of adversarial views in the real world
+  is, by definition,
+  $\condd{\progd(\prog)}{V_C}{\store_S}$.
+
+  The simulator is given both $\store_{S_C}$ and
+  $\idealf(\store_S)$.  The simulation $\SIM_C(\store_{S_C},
+  \idealf(\store_S))$ can be defined as follows. First, some $\store'
+  \in \ik(\idealf,\idealf(\store))$ is randomly chosen such that
+  $\store' =_C \store_S$, as is a random tape $\store'' \in
+  \mems(F)$\footnote{The real/ideal model allows consultation of a
+  ``Random Oracle''.}. Then, the run of $(\store' \cup \store'',
+  \prog)$ is evaluated in simulation, yielding $(\store^{\SIM},\varnothing)$.
+  The simulation returns $\store^{\SIM}_{V_C}$ as a result.
+
+  Now, since the random tape is selected in simulation from the same distribution
+  that we assume for the real world, after selection of $\store'$ the
+  probability that any particular $\store^{\SIM}_{V_C}$ is returned is by definition
+  $
+  \progd(\prog)(\store^{\SIM}_{V_C} | \store')
+  $.
+  Furthermore, since we assume that $\prog$ correctly implements $\idealf$, this
+  means
+  $
+  \store_S =_C \store' \wedge
+     \condd{\progd(\prog)}{\{ \outv \}}{\store_S} =
+     \condd{\progd(\prog)}{\{ \outv \}}{\store'}
+  $,
+  so by the definition of noninterference modulo output,
+  any choice of $\store'$ yields the same distribution $\condd{\progd(\prog)}{V_C}{\store_S}$.
+  Therefore by definition
+  $
+   \progd(\SIM_C(\store_{S_C},\idealf(\store))) = \condd{\progd(\prog)}{V_C}{\store_S}
+  $.
+\end{proof}
 

### 2023-07-11 popl24/hyperprop.tex
checkpoint

@@
 
-As we've discussed in our section on Related Work
-(\ref{section-related}), previous authors have noted connections
-between the traditional formulation of the MPC security model- aka the
-simulator security model- and hyperproperty-style formulations of
-program trace behavior \cite{8429300,wysstar} including probabilistic
-noninterference. In this Section we show for to formulate extensional
-properties relevant to MPC passive security in our language model by
-direct appeal to program distributions. In particular,
-\emph{noninterference modulo output ($\NIMO$)} is shown to be sound
-for MPC passive security. In Section \ref{section-automation} we will
-show how $\NIMO$ can be automatically enforced directly via
-whole-program analysis, or by automatically verifying composable,
-intensional properties of circuit components.
+As discussed in Section \ref{section-related-work} previous
+authors have noted connections between the traditional formulation of
+the MPC security model- aka the simulator security model- and
+hyperproperty-style formulations of program trace behavior
+\cite{8429300,wysstar} including probabilistic noninterference. In
+this Section we show for to formulate extensional properties relevant
+to MPC passive security in our language model by direct appeal to
+program distributions. In particular, \emph{noninterference modulo
+output ($\NIMO$)} is shown to be sound for MPC passive security. In
+Section \ref{section-automation} we will show how $\NIMO$ can be
+automatically enforced directly via whole-program analysis, or by
+automatically verifying composable, intensional properties of circuit
+components.
 
@@
 functionality $\idealf$- that is, execution of a configuration
-$\config{\store}{\prog}$. We will assume that any protocol $\prog$
-includes a public \emph{output view} $\outv$ where the result is
-published, where $0$ is the ``public'' client and thus $\outv$
-is of the form $\view{0}{w}$ for some $w$, i.e., is owned by 0.
-Note that the protocol is \emph{correct} iff the resulting
-memory contains $\outv \mapsto \idealf(\store_S)$, where $S$ are the
-secret inputs (honest and corrupt) in $\prog$.
+$\config{\store}{\prog}$. We will assume that any $\idealf$ is
+deterministic- while MPC generally allows for probabilistic ideal
+functionalities, this restriction captures a large and important class
+of protocols. We will assume that any protocol $\prog$ includes a
+public \emph{output view} $\outv$ where the result is published, where
+$0$ is the ``public'' client and thus $\outv$ is of the form
+$\view{0}{w}$ for some $w$, i.e., is owned by 0.  Note that the
+protocol is \emph{correct} iff the resulting memory contains $\outv
+\mapsto \idealf(\store_S)$, where $S$ are the secret inputs (honest
+and corrupt) in $\prog$. Note that this implies that a correct
+protocol publicizes the same output on the same secret inputs given
+any choice of random tape.
 
@@
 
-\subsection{Probabilistic Noninterference}
+\subsection{Probabilistic Noninterference ($\pni$)} 
 \label{section-hyperprop-ni}
@@
 
-Despite the intuitive resonance between noninterference
-and passive security, since MPC
-protocols implement an ideal functionality, they cannot generally satisfy
-noninterference. We can easily demonstrate this by using
-Example \ref{example-gmw-andcircuit} as a failure witness, where we'll
-assume $C = \{0,2\}$ and $H = \{1\}$:
+Despite the intuitive resonance between noninterference and passive
+security, since MPC protocols implement an ideal functionality, they
+cannot generally satisfy noninterference. Noting that the public
+output is always a corrupt view, we can easily demonstrate this by
+using Example \ref{example-gmw-andcircuit} as a failure witness, where
+we'll assume $C = \{0,2\}$ and $H = \{1\}$:
 $$
@@
 
-\subsection{A Sound Hyperproperty for Passive Security}
+\subsection{Noninteference Modulo Output ($\NIMO$)}
 \label{section-nimo}

### 2023-07-11 popl24/examples.tex
checkpoint

@@
 
-\paragraph{Example \ref{example-lambda-obliv}(a) revisited} As demonstrated
-by Example \ref{example-lambda-obliv} taken from \cite{darais2019language}, it is possible
-to leak information probabilistically through observable dependencies
-in views.
-%In the basic distribution for Example \ref{example-lambda-obliv}(a)
-%illustrated in Figure \ref{fig-basic-distributions},
-We can see that
-the adversarial views $\vx{0}{0}$ and $\vx{0}{1}$ are both 1 in
-three different runs of the protocol, and in 2 out of 3 the
-honest secret $\sx{1}{0}$ is 1. More precisely:
-$$
-(\progd(\prog_{\ref{example-lambda-obliv}(a)}))(\{\sx{1}{0} \mapsto 1 \}|\{ \vx{0}{0} \mapsto 1, \vx{0}{1} \mapsto 1 \}) = 2/3
-$$
-An important point illustrated by this example is that the
-probability of values in views, and not just the values themselves,
-can release information to the adversary.
-
 \paragraph{Example \ref{example-he} revisited} This standard example

### 2023-07-11 popl24/pmf.tex
checkpoint

@@
 \end{definition}
+
+\paragraph{Example \ref{example-lambda-obliv}(a) revisited} As demonstrated
+by Example \ref{example-lambda-obliv} taken from \cite{darais2019language}, it is possible
+to leak information probabilistically through observable dependencies
+in views. We can see that the adversarial views $\vx{0}{0}$ and $\vx{0}{1}$
+are both 1 in three different runs of the protocol, and in 2 out of 3
+the honest secret $\sx{1}{0}$ is 1 (clearly illustrated later in
+Figure \ref{fig-basic-distributions}). More precisely:
+$$
+(\progd(\prog_{\ref{example-lambda-obliv}(a)}))(\{\sx{1}{0} \mapsto 1 \}|\{ \vx{0}{0} \mapsto 1, \vx{0}{1} \mapsto 1 \}) = 2/3
+$$ An important point illustrated by this example is that the
+probability of values in views, and not just the values themselves,
+can release information to the adversary.  More examples are provided
+in Appendix \ref{section-exmaples}.

### 2023-07-11 popl24/metalang.tex
checkpoint

@@
 is, $\metaprot$ is a high-level language that generates low-level
-protocol code.
+protocol code. We endow $\metaprot$ with a type system that enjoys a
+type safety result guaranteeing safety of generated protocols. We
+consider library implementations for GMW and YGC circuits as
+examples. Additional examples are provided in Appendix
+\ref{section-examples-lang}.
 
@@
 
-\begin{fpfig}[t]{Type Syntax of $\metaprot$ (T) and selected type judgement rules.}{fig-metaprot-types}
+\begin{fpfig}[t]{Type syntax of $\metaprot$ (T) and selected type judgement rules (B).}{fig-metaprot-types}
 {\small
@@
 metaprogram, it also guarantees safety of the residual $\fedprot$
-program. Of course, it is important to note that ``safety'' here
-means the usual type safety property that programs do not get stuck,
-it does not imply any security hyperproperties. 
+program (Definition \ref{definition-safety}).
 \begin{theorem}[$\metaprot$ Type Safety]

### 2023-07-11 popl24/minicat.tex
checkpoint

@@
 
-We let $X$ range over sets of variables and $S,V,F$ to range over sets
-of secrets, views, and flips (including invocations of the oracle)
+We let $x$ range over variables, and will use $s$, $v$, and $f$ to
+range over individual secrets, views, and flips respecitively. We let
+$X$ range over sets of variables and $S,V,F$ to range over sets of
+secrets, views, and flips (including invocations of the oracle)
 respectively. Given a program $\prog$, we write $\iov(\prog)$ to
 denote $S \cup V$ where $S$ is the set of secret variables in $\prog$
-and $V$ is the set of views in $\prog$, we write $\flips(\prog)$
-to denote the set $F$ of flip variables in $\prog$, and we
-write $\vars(\prog)$ to denote $\iov(\prog) \cup \flips(\prog)$. For any set of
-variables $X$ and parties $P$, we write $X_P$ to denote the subset of
-$X$ owned by any party in $P$, and we write $X_H$ and $X_C$ to denote
-the subsets belonging to honest and corrupt parties, respectively.
+and $V$ is the set of views in $\prog$, we write $\flips(\prog)$ to
+denote the set $F$ of flip variables in $\prog$, and we write
+$\vars(\prog)$ to denote $\iov(\prog) \cup \flips(\prog)$. For any set
+of variables $X$ and parties $P$, we write $X_P$ to denote the subset
+of $X$ owned by any party in $P$, and we write $X_H$ and $X_C$ to
+denote the subsets belonging to honest and corrupt parties,
+respectively.
 
@@
 the case that $\store' \in \mems(S \cup F \cup V)$. We define a
-program as \emph{safe} iff for all $\store \in \mems(S \cup F)$ there
+notion of protocol safety as follows, which is a well-formedness
+property.
+\begin{definition}[Protocol Safety]
+  \label{definition-safety}
+A program is \emph{safe} iff for all $\store \in \mems(S \cup F)$ there
 exists $\store'$ where $\config{\store}{\prog} \redxs
-\config{\store'}{\varnothing}$. Unsafe programs are those where views
+\config{\store'}{\varnothing}$.
+\end{definition}
+Observe that unsafe programs are those where views
 are used before they're defined, or where expressions mix owners of

### 2023-07-11 popl24/automation.tex
checkpoint

@@
 \begin{lemma}
-  \label{lemma-cprogd}
+  \label{lemma-cprogtt}
   Given $\prog$ where $\iov(\prog) = S \cup V$ and $\flips(\prog) = F$. Define
@@
 A full implementation of these methods is available in our online
-repository\footnote{Link redacted for double-blind reviewing.}. To
+repository \cite{jpdf-github}. To
 avoid notational confusion we will continue to refer to $\progtt$ and
@@
 \subsection{Certification of Extensional Properties}
+\label{section-automation-extensional}
 
@@
   for all $\store \in \mems(S \cup V)$, and return success $\pni(\prog)$ iff every such
-  check succeeds. The result follows by Lemmas \ref{lemma-cprogd} and \ref{lemma-pni}.
+  check succeeds. The result follows by Lemmas \ref{lemma-cprogtt} and \ref{lemma-pni}.
   To verify $\nimo(\prog)$, we first enumerate all partitions of clients
@@
   $$
-  The result follows by Lemma \ref{lemma-cprogd} and Lemma \ref{lemma-nimo}.
+  The result follows by Lemma \ref{lemma-cprogtt} and Lemma \ref{lemma-nimo}.
 \end{proof}
@@
 \subsection{Certification of Composable Intensional Properties}
+\label{section-automation-intensional}
 
@@
 \begin{lemma}
+  \label{lemma-autosep}
   Both of the following are true:
@@
 \subsection{Compositional Metatheory}
+\label{section-automation-logic}
 

### 2023-07-11 popl24/composition.tex
checkpoint

@@
-\section{Automating Verification of Circuit Security}
+\section{Certification and Invariants for Circuit Security}
 \label{section-composition}
 
-Main results are in Theorems \ref{theorem-gmw-NIMO} and
+Here we show how to apply the certification method described in
+Section \ref{section-automation-intensional} to obtain passive
+security for any 2-party GMW or YGC circuits built using the libraries
+described in Sections \ref{section-metalang-gmw} and
+\ref{section-metalang-ygc}.  Main results are in Theorems
+\ref{theorem-gmw-NIMO} and \ref{theorem-ygc-NIMO}.
+
+Our certification techniques are defined generally wrt implementations
+of input encoding, decoding, and internal gates for both GMW and YGC
+circuits. In the following we will refer to arbitrary decode, encode,
+and gate functions with the restriction that any function in each of
+these categories has the same valid type signatures as their existing
+instantiations in our GMW and YGC libaries. All certifications can be
+automatically verified as a consequence of Lemmas \ref{lemma-cprogtt}
+and \ref{lemma-autosep}, and complete implementations of the
+certifications are available in our repository \cite{jpdf-github}.
+
+We also demonstrate that intenstional properties of certified
+components maintain a crucial separation invariant within circuits
+using methods developed in Section \ref{section-automation-logic}.
+Specifically, we show that within a circuit $\prog$, the views of each
+client remain probabilistically independent of the other client's
+secrets, i.e., $\sep{\progd(\prog)}{S_{\{1\}}}{V_{\{2\}}}$ and
+%\quad \text{and} \quad
+$\sep{\progd(\prog)}{S_{\{2\}}}{V_{\{1\}}}$.
+Through certification of decodings we can finally obtain $\NIMO$ in
+circuits as in Theorems \ref{theorem-gmw-NIMO} and
 \ref{theorem-ygc-NIMO}.
@@
 
-Now we can define GMW gate certification as follows. Intuitively, we certify
-that GMW gates satisfy a probabilistic noninterference property wrt each output
-share considered individually. This ensures that input dependencies remain
-encrypted from each client's perspective during circuit evaluation.
+In keeping with the standard definition of GMW, we can certify that
+GMW gates satisfy a probabilistic noninterference property wrt each
+output share considered individually. This ensures that input
+dependencies remain encrypted from each client's perspective during
+circuit evaluation.
 \begin{definition}[GMW Gate Certification]
@@
   \begin{array}{l}
-    \ttt{let in1 = \{c1 = flip[1,s11];c2 = flip[2,s12]\} in}\\
-    \ttt{let in2 = \{c1 = flip[1,s21];c2 = flip[2,s22]\} in}\\
-    \mathit{gate}\ttt{(g,in1,in2)}
+    \ttt{let in1 = \{c1 = flip[1,"s11"];c2 = flip[2,"s12"]\} in}\\
+    \ttt{let in2 = \{c1 = flip[1,"s21"];c2 = flip[2,"s22"]\} in}\\
+    \mathit{gate}\ttt{("g",in1,in2)}
   \end{array}
@@
   and let $
-  \config{\varnothing}{e}\redxs\config{\prog}{\ttt{\{c1 = v[1,gout]; c2 = v[2,gout]\}}}
+  \config{\varnothing}{e}\redxs\config{\prog}{\ttt{\{c1 = v[1,"gout"]; c2 = v[2,"gout"]\}}}
   $.
@@
   \begin{enumerate}[\hspace{5mm}i.]
-  \item $\sep{\progtt(\prog)}{\ttt{\{flip[1,s11],flip[1,s21],flip[2,s12],flip[2,s22]\}}}{\ttt{\v[1,gout]\}}}$
-  \item $\sep{\progtt(\prog)}{\ttt{\{flip[1,s11],flip[1,s21],flip[2,s12],flip[2,s22]\}}}{\ttt{\v[2,gout]\}}}$
+  \item {\small$\sep{\progtt(\prog)}{\ttt{\{flip[1,"s11"],flip[1,"s21"],flip[2,"s12"],flip[2,"s22"]\}}}{\ttt{v[1,"gout"]\}}}$}
+  \item {\small$\sep{\progtt(\prog)}{\ttt{\{flip[1,"s11"],flip[1,"s21"],flip[2,"s12"],flip[2,"s22"]\}}}{\ttt{v[2,"gout"]\}}}$}
   % \item $\sep{\progtt(\prog)}{\ttt{\{v[1,gout]\}}}{\ttt{\{v[2,gout]\}}}$
-  \item All variables in:$$\vars(\prog) - \{ \ttt{flip[1,s11],flip[1,s21],flip[2,s12],flip[2,s22]}\}$$
-    are distinguished by $\ttt{g}$ (i.e., contain $\ttt{g}$ as an identifier substring). 
+  \item All variables in:$$\vars(\prog) - \{ \ttt{flip[1,"s11"],flip[1,"s21"],flip[2,"s12"],flip[2,"s22"]}\}$$
+    are distinguished by $\ttt{"g"}$ (i.e., contain $\ttt{"g"}$ as an identifier substring). 
   \end{enumerate}
@@
   $$
-  \config{\varnothing}{\ttt{encode(s1,s2)}} \redxs
+  \config{\varnothing}{\mathit{encode}\ttt{("s1","s2")}} \redxs
   \config{\prog}{\ttt{\{shares1 = }v_1\ttt{; shares2 = }v_2\ttt{\}}}
@@
   \begin{array}{rcl}
-    v_1 &=& \ttt{\{ c1 = v[1,s1out]; c2 = v[2,s1out] \}}\\
-    v_2 &=& \ttt{\{ c1 = v[1,s2out]; c2 = v[2,s2out] \}}
+    v_1 &=& \ttt{\{ c1 = v[1,"s1out"]; c2 = v[2,"s1out"] \}}\\
+    v_2 &=& \ttt{\{ c1 = v[1,"s2out"]; c2 = v[2,"s2out"] \}}
   \end{array}
@@
     %\item $\sep{\progtt(\prog)}{\ttt{\{ s[1,s1],  v[1,s1out], v[1,s2out] \}}}{\ttt{\{ s[2,s2], v[2,s1out], v[2,s2out] \}}}$
-  \item $\sep{\progtt(\prog)}{\ttt{\{s[1,s1]\}}}{\ttt{\{v[2,s1out],v[2,s2out]\}}}$
-  \item $\sep{\progtt(\prog)}{\ttt{\{s[2,s2]\}}}{\ttt{\{v[1,s1out],v[1,s2out]\}}}$
-  \item All $x \in \vars(\prog)$ are distinguished by $\ttt{in1}$ or $\ttt{in2}$. 
+  \item $\sep{\progtt(\prog)}{\ttt{\{s[1,"s1"]\}}}{\ttt{\{v[2,"s1out"],v[2,"s2out"]\}}}$
+  \item $\sep{\progtt(\prog)}{\ttt{\{s[2,"s2"]\}}}{\ttt{\{v[1,"s1out"],v[1,"s2out"]\}}}$
+  \item All $x \in \vars(\prog)$ are distinguished by $\ttt{"s1"}$ or $\ttt{"s2"}$. 
   \end{enumerate}
@@
   \label{lemma-gmw-preservation}
-  Given well-typed $\prog_1;E[\mathit{gate}(\ttt{g},v_1,v_2)]$ for certified $\mathit{gate}$ where:
+  Given well-typed $\prog_1;E[\mathit{gate}(\ttt{"g"},v_1,v_2)]$ for certified $\mathit{gate}$ where:
   $$
-  \config{\prog_1}{E[\mathit{gate}(\ttt{g},v_1,v_2)]} \redxs \config{\prog_1;\prog_2}{E[v]}
+  \config{\prog_1}{E[\mathit{gate}(\ttt{"g"},v_1,v_2)]} \redxs \config{\prog_1;\prog_2}{E[v]}
   $$
@@
 \begin{proof}
-  Given well-typedness of $\prog_1;E[\mathit{gate}(\ttt{g},v_1,v_2)]$, we have for some $\ttt{g1}$ and $\ttt{g2}$:
+  Given well-typedness of $\prog_1;E[\mathit{gate}(\ttt{g},v_1,v_2)]$, we have for some $\ttt{"g1"}$ and $\ttt{"g2"}$:
   $$
   \begin{array}{lcl}
-   v_1 &=& \ttt{\{c1 = v[1,g1out]; c2 = v[2,g1out]\}}\\
-   v_2 &=& \ttt{\{c1 = v[1,g2out]; c2 = v[2,g2out]\}}\\
-    v &=& \ttt{\{c1 = v[1,gout]; c2 = v[2,gout]\}}
+   v_1 &=& \ttt{\{c1 = v[1,"g1out"]; c2 = v[2,"g1out"]\}}\\
+   v_2 &=& \ttt{\{c1 = v[1,"g2out"]; c2 = v[2,"g2out"]\}}\\
+    v &=& \ttt{\{c1 = v[1,"gout"]; c2 = v[2,"gout"]\}}
   \end{array}
@@
       \prog_2 = \\
-      \prog[\ttt{v[1,g1out]}/\ttt{flip[1,s11]}][\ttt{v[2,g1out]}/\ttt{flip[2,s12]}][\ttt{v[1,g2out]}/\ttt{flip[1,s21]}][\ttt{v[2,g2out]}/\ttt{flip[2,s22]}]
+      \prog[\ttt{v[1,"g1out"]}/\ttt{flip[1,"s11"]}][\ttt{v[2,"g1out"]}/\ttt{flip[2,"s12"]}][\ttt{v[1,"g2out"]}/\ttt{flip[1,"s21"]}][\ttt{v[2,"g2out"]}/\ttt{flip[2,"s22"]}]
     \end{array}
@@
   $$
-  Now, the assumption of well-typedness also assures that $\ttt{g}$
+  Now, the assumption of well-typedness also assures that $\ttt{"g"}$
   has not been previously used as a gate identifier in $\prog_1$,
-  since otherwise the views $\ttt{v[1,gout]}$ and $\ttt{v[2,gout]}$
+  since otherwise the views $\ttt{v[1,"gout"]}$ and $\ttt{v[2,"gout"]}$
   would have been previously defined. This ensures $\vars(\prog_1,x)
   \cap \vars(\prog) = \varnothing$ for:
-  $$x \in \{ \ttt{v[1,g1out]}, \ttt{v[2,g1out]}, \ttt{v[1,g2out]}, \ttt{v[2,g2out]}\}$$
+  $$x \in \{ \ttt{v[1,"g1out"]}, \ttt{v[2,"g1out"]}, \ttt{v[1,"g2out"]}, \ttt{v[2,"g2out"]}\}$$
   given condition (iii) of Definition \ref{definition-gmwgate-certification}.
   Thus by Lemma \ref{lemma-substitution} we have:
-  $$\sep{\progtt(\prog_1;\prog_2)}{\ttt{\{v[1,g1out],v[1,g2out],v[2,g1out],v[2,g2out]\}}}{\ttt{\{v[1,gout]\}}}$$
+  $$\sep{\progtt(\prog_1;\prog_2)}{\ttt{\{v[1,"g1out"],v[1,"g2out"],v[2,"g1out"],v[2,"g2out"]\}}}{\ttt{\{v[1,"gout"]\}}}$$
   and
-  $$\sep{\progtt(\prog_1;\prog_2)}{\ttt{\{v[1,g1out],v[1,g2out],v[2,g1out],v[2,g2out]\}}}{\ttt{\{v[2,gout]\}}}$$
+  $$\sep{\progtt(\prog_1;\prog_2)}{\ttt{\{v[1,"g1out"],v[1,"g2out"],v[2,"g1out"],v[2,"g2out"]\}}}{\ttt{\{v[2,"gout"]\}}}$$
   Now, by Lemmas \ref{lemma-noninterference} and \ref{lemma-separation} and condition (iii) of Definition
   \ref{definition-gmwgate-certification} we then have:
-  $$\sep{\progtt(\prog_1;\prog_2)}{(S_{\{1\}} \cup V_{\{2\}})}{\ttt{\{v[2,gout]\}}} \quad \text{and} \quad
-  \sep{\progtt(\prog_1;\prog_2)}{(S_{\{2\}} \cup V_{\{1\}})}{\ttt{\{v[1,gout]\}}}$$
+  $$\sep{\progtt(\prog_1;\prog_2)}{(S_{\{1\}} \cup V_{\{2\}})}{\ttt{\{v[2,"gout"]\}}} \quad \text{and} \quad
+  \sep{\progtt(\prog_1;\prog_2)}{(S_{\{2\}} \cup V_{\{1\}})}{\ttt{\{v[1,"gout"]\}}}$$
   so by precondition assumptions and Lemma \ref{lemma-separation} we have:
@@
   \config{\varnothing}{\eassign{\outv}{\ttt{decode}(e)}} \redxs
-  \config{\prog;\eassign{\outv}{\ttt{v[1,gout] xor v[2,gout]}}}{\varnothing}
+  \config{\prog;\eassign{\outv}{\ttt{v[1,"gout"] xor v[2,"gout"]}}}{\varnothing}
   $$
@@
   $\sep{\progd(\prog)}{S_{\{2\}}}{V_{\{1\}}}$.
-  Let $\prog' \defeq \prog;\eassign{\outv}{\ttt{v[1,gout] xor v[2,gout]}}$. The
+  Let $\prog' \defeq \prog;\eassign{\outv}{\ttt{v[1,"gout"] xor v[2,"gout"]}}$. The
   preceding then implies
@@
   \item $\gc{\prog}{v_1}{\ttt{s1}} \quad \text{and} \quad \gc{\prog}{v_2}{\ttt{s2}}$
-  \item $\sep{\progd(\prog)}{\{\ttt{s[1,s1]\}}}{\vdefs(\prog)_{\{2\}}}$
-  \item $\sep{\progd(\prog)}{\{\ttt{s[2,s2]\}}}{\vdefs(\prog)_{\{1\}}}$    
+  \item $\sep{\progd(\prog)}{\{\ttt{s[1,"s1"]\}}}{\vdefs(\prog)_{\{2\}}}$
+  \item $\sep{\progd(\prog)}{\{\ttt{s[2,"s2"]\}}}{\vdefs(\prog)_{\{1\}}}$    
   \item All $x \in \vars(\prog)$ are distinguished by $\ttt{in1}$ or $\ttt{in2}$. 

### 2023-07-11 popl24/composition.tex
checkpoint

@@
 \subsection{GMW Gate Invariants and Circuit Security}
-We can show that composition maintains an important invariant in a
-circuit $\prog$- namely that client 1's secrets remain independent of
-client 2's views, and vice-versa, prior to decoding.  That is, where
+
+Now we demonstrate that certified gates preserve the desired invariant
+during execution in arbitrary circuit contexts.That is, where
 $\iov(\prog) = S \cup V$, we show that
@@
 $\sep{\progd(\prog)}{S_{\{2\}}}{V_{\{1\}}}$ are preserved prior to
-decoding.
-
-Now we can prove that certified gates preserve the desired invariant
-during execution in arbitrary circuit contexts.
-We formulate this in terms of pre-~and post-conditions of gate
-evaluation.
+decoding.  We formulate this in terms of pre-~and post-conditions of
+gate evaluation. Proofs omitted here are provided in Appendix
+\ref{section-proofs}.
 \begin{lemma}[GMW Gate Invariant]

### 2023-07-11 popl24/automation.tex
checkpoint

@@
 parallelization- rather than computing the truth table
-``horizontally'', it can be computed vertically. We can rewrite
+``horizontally'', it can be computed ``vertically''. We can rewrite
 $\minifed$ programs $\prog$ to stratified Datalog programs, where

### 2023-07-11 popl24/composition.tex
checkpoint

@@
 \begin{proof}
-  Given well-typedness of $\prog_1;E[\mathit{gate}(\ttt{g},v_1,v_2)]$, we have for some $\ttt{"g1"}$ and $\ttt{"g2"}$:
+  Given well-typedness of $\prog_1;E[\mathit{gate}(\ttt{"g"},v_1,v_2)]$, we have for some $\ttt{"g1"}$ and $\ttt{"g2"}$:
   $$
@@
   $$
-  {\footnotesize
-    \begin{array}{c}
-      \prog_2 = \\
-      \prog[\ttt{v[1,"g1out"]}/\ttt{flip[1,"s11"]}][\ttt{v[2,"g1out"]}/\ttt{flip[2,"s12"]}][\ttt{v[1,"g2out"]}/\ttt{flip[1,"s21"]}][\ttt{v[2,"g2out"]}/\ttt{flip[2,"s22"]}]
+    \begin{array}{lcl}
+      \prog_2 &=&
+           \prog[\ttt{v[1,"g1out"]}/\ttt{flip[1,"s11"]}][\ttt{v[2,"g1out"]}/\ttt{flip[2,"s12"]}]\\
+      & & \phantom{\prog}[\ttt{v[1,"g2out"]}/\ttt{flip[1,"s21"]}][\ttt{v[2,"g2out"]}/\ttt{flip[2,"s22"]}]
     \end{array}
-  }
   $$
@@
   \label{lemma-gmw-encode}
-  Given well-typed $\prog_1;E[\mathit{encode}\ttt{(s1,s2)}]$ for certified $\mathit{encode}$, let:
+  Given well-typed $\prog_1;E[\mathit{encode}\ttt{("s1","s2")}]$ for certified $\mathit{encode}$, let:
   $$
-  \config{\prog_1}{E[\mathit{encode}(\ttt{s1},\ttt{s2})]} \redxs
+  \config{\prog_1}{E[\mathit{encode}(\ttt{"s1"},\ttt{"s2"})]} \redxs
   \config{\prog_1;\prog_2}{E[v]}

### 2023-07-11 popl24/automation.tex
checkpoint

@@
 
-To obtain $\runs(\prog)$, we can perform a left folding of $\solve$
+To obtain $\runs(\prog)$, we can perform a standard left folding of $\solve$
 across $\prog$ and each view definition, inductively extending

### 2023-07-11 popl24/main.tex
Add keywords and classifications

@@
 
+%%
+%% The code below is generated by the tool at http://dl.acm.org/ccs.cfm.
+%% Please copy and paste the code instead of the example below.
+%%
+\begin{CCSXML}
+<ccs2012>
+<concept>
+<concept_id>10002978.10002986.10002990</concept_id>
+<concept_desc>Security and privacy~Logic and verification</concept_desc>
+<concept_significance>500</concept_significance>
+</concept>
+<concept>
+<concept_id>10003752.10003753.10003757</concept_id>
+<concept_desc>Theory of computation~Probabilistic computation</concept_desc>
+<concept_significance>300</concept_significance>
+</concept>
+</ccs2012>
+\end{CCSXML}
+
+\ccsdesc[500]{Security and privacy~Logic and verification}
+\ccsdesc[500]{Theory of computation~Probabilistic computation}
+
+%%
+%% Keywords. The author(s) should pick words that accurately describe
+%% the work being presented. Separate the keywords with commas.
+\keywords{Multiparty computation, program verification, probabilistic programming}
+
 \maketitle

### 2023-07-11 popl24/automation.tex
checkpoint

--- 2023-07-11 popl24/automation.tex
@@
 
-To obtain $\runs(\prog)$, we can perform a left folding of $\solve$
-across $\prog$ and each view definition, inductively extending
+To obtain $\runs(\prog)$, we can perform a standard left folding of $\solve{\!}{\!}$
+across $\prog$ as a list of view definitions, inductively extending
 memories with valid view assignments in the order of their definition.

### 2023-07-11 popl24/composition.tex
checkpoint

@@
 \subsection{GMW Component Certification}
+\label{section-gmw-certification}
 
@@
 \subsection{GMW Gate Invariants and Circuit Security}
+\label{section-gmw-composition}
 
@@
 
-Our certification techniques are defined generally wrt implementations
-of input encoding, decoding, and internal gates. In the following we
-will refer to arbitrary decode, encode, and gate functions with the
-restriction that any function in each of these categories has the same
-valid type signature as $\ttt{decode}$, $\ttt{encode}$, and
-$\ttt{andgg}$ respectively.
-
-It is necessary to define notions of correlation with wire
+In the case of YGC, wires are defined via wire labels as opposed
+to shares as in GMW. It is necessary to define notions of correlation with wire
 labels, since wire values flow into and out of gates in either
-positive of negative correlation with them.
+positive of negative correlation with them, and this correlation is
+integral to both pre-~and post-conditions of certification. We
+write $\gc{\prog}{v}{g}$ to denote this, where $g$
+is the gate with which $v$ has \emph{\bf g}arbled \emph{\bf o}utput
+\emph{\bf c}orrelation.
 \begin{definition} 
   \label{definition-gc}
-  We write $\gc{\prog}{\{ \ttt{k = }\be_{k}; \ttt{p = }\be_{p} \}}{g}$ iff one of the
+  We write $\gc{\prog}{\{ \ttt{k = }\be_{k}; \ttt{p = }\be_{p} \}}{\ttt{"g"}}$ iff one of the
   following hold:
-  \begin{enumerate}[\hspace{5mm}i.]
-  \item $\vc{\progtt(\prog,\be_k)}{\itv}{\ttt{flip[2,gate:}g\ttt{.k]}}\quad \text{and} \quad
-    \vc{\progtt(\prog,\be_p)}{\itv}{\ttt{flip[2,gate:}g\ttt{.p]}}$
-  \item $\vc{\progtt(\prog,\enot\ \be_k)}{\itv}{\ttt{flip[2,gate:}g\ttt{.k]}}\quad \text{and}\quad
-    \vc{\progtt(\prog,\enot\ \be_p)}{\itv}{\ttt{flip[2,gate:}g\ttt{.p]}}$
+  \begin{enumerate}[\hspace{2mm}i.]
+  \item $\vc{\progtt(\prog,\be_k)}{\itv}{\ttt{"flip[2,gate:g.k]"}}\ \, \text{and} \ \,
+    \vc{\progtt(\prog,\be_p)}{\itv}{\ttt{"flip[2,gate:g.p]"}}$
+  \item $\vc{\progtt(\prog,\enot\ \be_k)}{\itv}{\ttt{"flip[2,gate:g.k]"}}\ \, \text{and} \ \,
+    \vc{\progtt(\prog,\enot\ \be_p)}{\itv}{\ttt{"flip[2,gate:g.p]"}}$
   \end{enumerate}
@@
 the output wire label under any input wire value combination.
+Recall from Figure \ref{fig-ygc-lib} that $\ttt{owl("g")}$ returns
+the output wire label of gate $\ttt{g}$ in a positive valence.
 \begin{definition}[YGC Gate Certification]
@@
   \begin{enumerate}[\hspace{5mm}$e_1 \defeq$]
-    \item $\mathit{gate}\ttt{(c,a,b,owl(a),owl(b));}$
-    \item $\mathit{gate}\ttt{(c,a,b,owl(a),invert(owl(b)));}$
-    \item $\mathit{gate}\ttt{(c,a,b,invert(owl(a)),owl(b));}$
-    \item $\mathit{gate}\ttt{(c,a,b,invert(owl(a)),invert(owl(b)))}$
+    \item $\mathit{gate}\ttt{("c","a","b",owl("a"),owl("b"));}$
+    \item $\mathit{gate}\ttt{("c","a","b",owl("a"),invert(owl("b")));}$
+    \item $\mathit{gate}\ttt{("c","a","b",invert(owl("a")),owl("b"));}$
+    \item $\mathit{gate}\ttt{("c","a","b",invert(owl("a")),invert(owl("b")))}$
   \end{enumerate}
   Define:
-  $${\small X \defeq \ttt{\{ flip[2,gate:a.k], flip[2,gate:a.p], flip[2,gate:b.k], flip[2,gate:b.p] \}}}$$
+  $${\small X \defeq \ttt{\{ flip[2,"gate:a.k"], flip[2,"gate:a.p"], flip[2,"gate:b.k"], flip[2,"gate:b.p"] \}}}$$
   Then $\mathit{gate}$ is \emph{certified} iff the following conditions hold:
@@
   \item $\sep{\progtt(\prog)}{X}{\vdefs(\prog)}$
-  \item For all $1 \le i \le 4$, $\gc{\prog}{v_i}{\ttt{c}}$
+  \item For all $1 \le i \le 4$, $\gc{\prog}{v_i}{\ttt{"c"}}$
   \end{enumerate}
@@
   $$
-  \config{\varnothing}{\ttt{encode(s1,s2)}} \redxs
+  \config{\varnothing}{\mathit{encode}\ttt{("s1","s2")}} \redxs
   \config{\prog}{\ttt{\{wv1=}v_1\ttt{;wv2=}v_2\ttt{\}}}
@@
   \begin{enumerate}[\hspace{5mm}i.]
-  \item $\gc{\prog}{v_1}{\ttt{s1}} \quad \text{and} \quad \gc{\prog}{v_2}{\ttt{s2}}$
+  \item $\gc{\prog}{v_1}{\ttt{"s1"}} \quad \text{and} \quad \gc{\prog}{v_2}{\ttt{"s2"}}$
   \item $\sep{\progd(\prog)}{\{\ttt{s[1,"s1"]\}}}{\vdefs(\prog)_{\{2\}}}$
   \item $\sep{\progd(\prog)}{\{\ttt{s[2,"s2"]\}}}{\vdefs(\prog)_{\{1\}}}$    
-  \item All $x \in \vars(\prog)$ are distinguished by $\ttt{in1}$ or $\ttt{in2}$. 
+  \item All $x \in \vars(\prog)$ are distinguished by $\ttt{"s1"}$ or $\ttt{"s2"}$. 
   \end{enumerate}
@@
   $$
-  \config{\varnothing}{\mathit{decode}(\ttt{c}, \ttt{owl(c)})}
+  \config{\varnothing}{\mathit{decode}(\ttt{"c"}, \ttt{owl("c")})}
   \redxs \config{\prog}{\be}
@@
   $$
-  \sep{\progtt(\prog)}{\ttt{\{flip[2,gate:c.k],flip[2,gate:c.p]\}}}{\vdefs(\prog)}
+  \sep{\progtt(\prog)}{\ttt{\{flip[2,"gate:c.k"],flip[2,"gate:c.p"]\}}}{\vdefs(\prog)}
   $$
@@
 \subsection{YGC Gate Invariants and Circuit Security}
-\label{section-composition-metatheory}
+\label{section-ygc-composition}
 
-To demonstrate correctness of certification, we need to show that
-isolated component certificates preserve relevant properties
-when integrated into larger programs. 
-%First, we show that if a flip $f$ is independent from the views
-%in a protocol (as is the case with input labels used to
-%encode a garbled table), then separation from views that
-%use $f$ (e.g., in an output label) is preserved. 
-%\begin{lemma}[Wire Framing]
-%  \label{lemma-wire-framing}
-%  If $\sep{\progtt(\prog)}{\{f\}}{\vdefs(\prog)}$ for flip $f$ and
-%  $\vars(\prog) \cap \vars(\prog') = \{ f \}$ then
-%  $\sep{\progtt(\prog';\prog)}{\vdefs(\prog')}{\vdefs(\prog)}$.
-%\end{lemma}
-%Next,
+Now we demonstrate that certified gates preserve the desired invariant
+during execution in arbitrary circuit contexts, assuming that outputs are not wired to
+multiple inputs \cite{tate2003garbled,nieminen2023breaking}. That is, where
+$\iov(\prog) = S \cup V$, we show that
+$\sep{\progd(\prog)}{S_{\{1\}}}{V_{\{2\}}}$ and
+$\sep{\progd(\prog)}{S_{\{2\}}}{V_{\{1\}}}$ are preserved prior to
+decoding.  We formulate this in terms of pre-~and post-conditions of
+gate evaluation. Proofs omitted here are provided in Appendix
+\ref{section-proofs}.
 
-Now we can show how to splice gates into arbitrary circuits in an
-invariant-preserving manner, assuming that outputs are not wired to
-multiple inputs \cite{tate2003garbled,nieminen2023breaking}.
 \begin{restatable}[YGC Gate Invariant]{lemma}{ygcgate}
@@
 reveal of the output, and we require that no gate is wired more than once.
-\cnote{There is another subtlety here, that gates are wired correctly
-  in the sense that if the garbler generates a table with input wire label
-  $g_1$ then only $g_1$ is wired to that input. }
 \begin{restatable}{theorem}{ygcnimo}

### 2023-07-11 popl24/main.tex
Update classification

@@
        </concept>
+   <concept>
+       <concept_id>10003752.10003790.10003806</concept_id>
+       <concept_desc>Theory of computation~Programming logic</concept_desc>
+       <concept_significance>500</concept_significance>
+       </concept>
  </ccs2012>
@@
 \ccsdesc[500]{Theory of computation~Probabilistic computation}
+\ccsdesc[500]{Theory of computation~Programming logic}
 

### 2023-07-11 popl24/proofs.tex
Update proofs.tex

@@
     v &=&  \ttt{\{shares1 = }v_1\ttt{; shares2 = }v_2\ttt{\}}\\
-    v_1 &=& \ttt{\{ c1 = v[1,s1out]; c2 = v[2,s1out] \}}\\
-    v_2 &=& \ttt{\{ c1 = v[1,s2out]; c2 = v[2,s2out] \}}
+    v_1 &=& \ttt{\{ c1 = v[1,"s1out"]; c2 = v[2,"s1out"] \}}\\
+    v_2 &=& \ttt{\{ c1 = v[1,"s2out"]; c2 = v[2,"s2out"] \}}
   \end{array}
@@
   \begin{mathpar}
-    \sep{\progtt(\prog)}{\ttt{\{s[1,s1]\}}}{\ttt{\{v[2,s1out],v[2,s2out]\}}}
+    \sep{\progtt(\prog)}{\ttt{\{s[1,"s1"]\}}}{\ttt{\{v[2,"s1out"],v[2,"s2out"]\}}}
 
-    \sep{\progtt(\prog)}{\ttt{\{s[2,s2]\}}}{\ttt{\{v[1,s1out],v[1,s2out]\}}}
+    \sep{\progtt(\prog)}{\ttt{\{s[2,"s2"]\}}}{\ttt{\{v[1,"s1out"],v[1,"s2out"]\}}}
   \end{mathpar}
@@
   \begin{mathpar}
-    \sep{\progtt(\prog_2)}{\ttt{\{s[1,s1]\}}}{\vdefs(\prog_2)_{\{2\}}}
+    \sep{\progtt(\prog_2)}{\ttt{\{s[1,"s1"]\}}}{\vdefs(\prog_2)_{\{2\}}}
 
-    \sep{\progtt(\prog_2)}{\ttt{\{s[2,s2]\}}}{\vdefs(\prog_2)_{\{1\}}}
+    \sep{\progtt(\prog_2)}{\ttt{\{s[2,"s2"]\}}}{\vdefs(\prog_2)_{\{1\}}}
   \end{mathpar}

### 2023-07-12 popl24/hyperprop.tex
checkpoint

@@
 
-As discussed in Section \ref{section-related-work} previous
-authors have noted connections between the traditional formulation of
-the MPC security model- aka the simulator security model- and
+As discussed in Section \ref{section-related-work} previous authors
+have noted connections between the traditional formulation of the MPC
+security model- aka the simulator security model- and
 hyperproperty-style formulations of program trace behavior
 \cite{8429300,wysstar} including probabilistic noninterference. In
-this Section we show for to formulate extensional properties relevant
-to MPC passive security in our language model by direct appeal to
-program distributions. In particular, \emph{noninterference modulo
-output ($\NIMO$)} is shown to be sound for MPC passive security. In
-Section \ref{section-automation} we will show how $\NIMO$ can be
-automatically enforced directly via whole-program analysis, or by
-automatically verifying composable, intensional properties of circuit
-components.
+this Section we show how to formulate security properties in our
+language model by direct appeal to program distributions. In
+particular, \emph{noninterference modulo output ($\NIMO$)} is shown to
+be sound for MPC passive security. In Section \ref{section-automation}
+we will show how $\NIMO$ can be automatically enforced directly via
+whole-program analysis, or by automatically verifying composable
+properties of circuit components.
 

### 2023-07-12 popl24/automation.tex
checkpoint

@@
 \margd{\progtt(\prog)}{\iov(\prog)}$.  As already discussed in Section
-\ref{section-hyperprop} and as we will discuss more below, extensional
-and intensional security properties are predicated on marginalizations
+\ref{section-hyperprop} and as we will discuss more below,
+crucial properties are predicated on marginalizations
 and conditionings of basic distributions. So calculating
@@
 Given computation of $\prog(\prog)$ for programs $\prog$, we can
-immediately observe that the extensional hyperproperties discussed in
+immediately observe that the security hyperproperties discussed in
 Section \ref{section-hyperprop} are certifiable.
@@
 
-\subsection{Certification of Composable Intensional Properties}
+\subsection{Certification of Composable Properties}
 \label{section-automation-intensional}
@@
 guarantee security for circuits built from them. Intuitively, we can
-consider any gate as a mini-program and certify intensional properties
-of its isolated program distribution. These intensional properties are
+consider any gate as a mini-program and certify properties
+of its isolated program distribution. These component properties are
 guaranteed secure in circuit compositions, as we demonstrate using
@@
 
-When we consider intensional properties
+When we consider properties
 of gates in isolation, we will explicitly consider flip conditions in

### 2023-07-12 popl24/composition.tex
checkpoint

@@
 and \ref{lemma-autosep}, and complete implementations of the
-certifications are available in our repository \cite{jpdf-github}.
+certifications are available in our repository \cite{jpdf-github},
+that we've used to certify our GMW and YGC libraries (Lemmas
+\ref{lemma-gmw-certification} and \ref{lemma-ygc-certification}.)
 
-We also demonstrate that intenstional properties of certified
+We also demonstrate that properties of certified
 components maintain a crucial separation invariant within circuits

### 2023-07-12 popl24/metalang.tex
Update metalang.tex

@@
     let owl2 = owl(sb) in
-    v[1,gate: || sa || 1.k] := OT(s[1,sa],owl1.k,(not owl1.k));
-    v[1,gate: || sa || 1.p] := OT(s[1,sa],owl1.p,(not owl1.p));
-    v[1,gate: || sb || 2.k] := select(s[2,sb],owl2.k,(not owl2.k));
-    v[1,gate: || sb || 2.p] := select(s[2,sb],owl2.p,(not owl2.p));
+    v[1,"gate:" || sa || "1.k"] := OT(s[1,sa],owl1.k,(not owl1.k));
+    v[1,"gate:" || sa || "1.p"] := OT(s[1,sa],owl1.p,(not owl1.p));
+    v[1,"gate:" || sb || "2.k"] := select(s[2,sb],owl2.k,(not owl2.k));
+    v[1,"gate:" || sb || "2.p"] := select(s[2,sb],owl2.p,(not owl2.p));
     { wv1 = { k = v[1,"gate:" || sa || "1.k"]; p = v[1,"gate:" || sa || "1.p"] };
@@
   views in a standard form. For example, the view for
-  a hypothetical gate 6, row 2 garbled truth table is $\ttt{v[1,gate:6tt2]}$.
+  a hypothetical gate 6, row 2 garbled truth table is $\ttt{v[1,"gate:6tt2"]}$.
   We note that garbled gates of other binary operators can be obtained with

### 2023-07-12 popl24/metalang.tex
checkpoint

--- 2023-07-11 popl24/metalang.tex
@@
 
-\subsection{Example: GMW}
+\subsection{Example: 2-Party GMW}
 \label{section-metalang-gmw}
@@
 \input{gmw}
-
-\begin{fpfig}[t]{GMW library type annotations.}{fig-gmw-types}
-{\footnotesize
-  \begin{verbatimtab}
-   encodegmw   : string(gid) * string(gid)
-
-   andtablegmw : { k = bool[i]; p = bool[i] }
-
-   andgmw      : string(gid) *  { c1 = bool[1]; c2 = bool[2] } * { c1 = bool[1]; c2 = bool[2] }
-
-   decodegmw   : { c1 = bool[1]; c2 = bool[2] }  \end{verbatimtab}
-}
-\end{fpfig}
 
@@
 Figure \ref{fig-gmw}, with type signatures for the library functions
-shown in Figure \ref{fig-gmw-types}. We show the And gate since it is
+also shown. We present the And gate since it is
 an interesting component. The Figure includes the
@@
   \ttt{andtablegmw})and transfers the correct output share to client 2
-  using 1-out-of-4 OT as per standard GMW protocol.
+  using 1-out-of-4 OT ($\ttt{OT4}$), definable using $\ttt{OT}$, as
+  per standard GMW protocol.
 \item \ttt{decode}: This function decodes and publishes a wire value

### 2023-07-12 popl24/conclusion.tex
checkpoint

\section{Conclusion and Future Work}
\label{section-conclusion}

We have developed a new probabilistic programming language called
$\minifed$ for defining synchronous distributed MPC protocols, and a new
metaprogramming language $\metaprot$ that is able to express logical
protocol components such as circuit gates and that dynamically
generates $\minifed$ protocols. We have developed \emph{program
distributions}, a formalism for expressing dependencies between
secrets and views in protocols and precisely quantifying allowable
information leakage in terms of protocol outputs. 
A hyperproperty called $\NIMO$, predicated on program distributions,
was shown to guarantee MPC passive security.

In the $\metaprot$-to-$\minicat$ model we've developed an automated
technique, aka \emph{certification} method, for automatically
verifying MPC security properties such as $\NIMO$ and other supporting
properties of protocol components.  While certification has high
complexity, a compositional approach allows certification of program
components in isolation.  This enables a semi-automated method for
large circuits where we (1) define certification for components, (2)
certify components, (3) prove $\NIMO$ based on the certifiedness of
components. We developed novel methods partially based on separation
to enable (3), and we've illustrated our approach by certifying GMW
and Yao's Garbled Circuit libraries in a manner that guarantees
security in arbitrarily large circuits. Certification tests can also
be re-used to validate circuit extensions.

We believe this line of research is promising for future work,
including extensions to malicious security, and to larger fields and
arithmetic circuits. HPC optimizations of program distribution
calculations, to support analysis of larger circuit components, is
another interesting direction for future work.

### 2023-07-12 popl24/proofs.tex
Update proofs.tex

@@
     &\prog_2 = \\
-    &{\small \prog[\be^1_{k}/\ttt{flip[2,gate:a.k]}][\be^1_{p}/\ttt{flip[2,gate:a.p]}][\be^2_{k}/\ttt{flip[2,gate:b.k]}][\be^1_{p}/\ttt{flip[2,gate:b.p]}]}
+    &{\small \prog[\be^1_{k}/\ttt{flip[2,"gate:a.k"]}][\be^1_{p}/\ttt{flip[2,"gate:a.p"]}][\be^2_{k}/\ttt{flip[2,"gate:b.k"]}][\be^1_{p}/\ttt{flip[2,"gate:b.p"]}]}
   \end{eqnarray*}
@@
     & \prog_2 = \\ 
-    & \prog[\be_k/\ttt{\{flip[2,gate:c.k]}][\be_p/\ttt{flip[2,gate:c.p]}]
+    & \prog[\be_k/\ttt{\{flip[2,"gate:c.k"]}][\be_p/\ttt{flip[2,"gate:c.p"]}]
   \end{eqnarray*}

### 2023-07-12 popl24/composition.tex
Update composition.tex

@@
   \label{lemma-ygc-encode}
-Given well-typed $\prog_1;E[\mathit{encode}(s_1,s_2)]$ and certified $\mathit{encode}$ where 
+Given well-typed $\prog_1;E[\mathit{encode}("s_1","s_2")]$ and certified $\mathit{encode}$ where 
 $$
-\config{\prog_1}{E[\ttt{encode}(s_1,s_2)]} \redxs \config{\prog_1;\prog_2}{E[\ttt{\{wv1=}v_1\ttt{;wv2=}v_2\ttt{\}}]}
+\config{\prog_1}{E[\ttt{encode}("s_1","s_2")]} \redxs \config{\prog_1;\prog_2}{E[\ttt{\{wv1=}v_1\ttt{;wv2=}v_2\ttt{\}}]}
 $$

### 2023-07-12 popl24/automation.tex
checkpoint
Update automation.tex

@@
 compute the basic program distribution. Recall that basic
-distributions are based on the memories resulting from execution of an
+distributions are based on the memories resulting from the execution of an
 $\minicat$ program, which contain a record of all client views and the
@@
 $\minifed$ programs $\prog$ to stratified Datalog programs, where
-$\runs(\prog)$ is equivalent to set of least Herbrand models obtained
+$\runs(\prog)$ is equivalent to the set of least Herbrand models obtained
 by representing different random tapes and input secrets as different
@@
 
-\subsection{Certification of Extensional Properties}
+\subsection{Certification of Whole-Program Security Properties}
 \label{section-automation-extensional}
@@
 verifying properties of $\minifed$ protocols, vs.~manual methods. For
-the sake of clarity we make the following definition:
+the sake of clarity, we make the following definition:
 \begin{definition}

### 2023-07-12 popl24/composition.tex
checkpoint

--- 2023-07-12 popl24/composition.tex
@@
   \begin{enumerate}[\hspace{2mm}i.]
-  \item $\vc{\progtt(\prog,\be_k)}{\itv}{\ttt{"flip[2,gate:g.k]"}}\ \, \text{and} \ \,
-    \vc{\progtt(\prog,\be_p)}{\itv}{\ttt{"flip[2,gate:g.p]"}}$
-  \item $\vc{\progtt(\prog,\enot\ \be_k)}{\itv}{\ttt{"flip[2,gate:g.k]"}}\ \, \text{and} \ \,
-    \vc{\progtt(\prog,\enot\ \be_p)}{\itv}{\ttt{"flip[2,gate:g.p]"}}$
+  \item $\vc{\progtt(\prog,\be_k)}{\itv}{\ttt{flip[2,"gate:g.k"]}}\ \, \text{and} \ \,
+    \vc{\progtt(\prog,\be_p)}{\itv}{\ttt{flip[2,"gate:g.p"]}}$
+  \item $\vc{\progtt(\prog,\enot\ \be_k)}{\itv}{\ttt{flip[2,"gate:g.k"]}}\ \, \text{and} \ \,
+    \vc{\progtt(\prog,\enot\ \be_p)}{\itv}{\ttt{flip[2,"gate:g.p"]}}$
   \end{enumerate}
@@
   \label{lemma-ygc-encode}
-Given well-typed $\prog_1;E[\mathit{encode}(s_1,s_2)]$ and certified $\mathit{encode}$ where 
-$$
-\config{\prog_1}{E[\ttt{encode}(s_1,s_2)]} \redxs \config{\prog_1;\prog_2}{E[\ttt{\{wv1=}v_1\ttt{;wv2=}v_2\ttt{\}}]}
-$$
+Given well-typed $\prog_1;E[\mathit{encode}(s1,s_2)]$ and certified $\mathit{encode}$ where 
+%$$
+%\config{\prog_1}{E[\mathit{encode}(s_1,s_2)]} \redxs \config{\prog_1;\prog_2}{E[\ttt{\{wv1=}v_1\ttt{;wv2=}v_2\ttt{\}}]}
+%$$
 If the following preconditions hold where $\iov(\prog_1) = S \cup V$:

### 2023-07-12 popl24/composition.tex
grammar fix composition.tex

@@
 $\sep{\progd(\prog)}{S_{\{2\}}}{V_{\{1\}}}$.
-Through certification of decodings we can finally obtain $\NIMO$ in
+Through certification of decodings, we can finally obtain $\NIMO$ in
 circuits as in Theorems \ref{theorem-gmw-NIMO} and
@@
 Now we demonstrate that certified gates preserve the desired invariant
-during execution in arbitrary circuit contexts.That is, where
+during execution in arbitrary circuit contexts. That is, where
 $\iov(\prog) = S \cup V$, we show that

### 2023-07-12 popl24/composition.tex
checkpoint

--- 2023-07-12 popl24/composition.tex
@@
   \label{lemma-gmw-encode}
-  Given well-typed $\prog_1;E[\mathit{encode}\ttt{("s1","s2")}]$ for certified $\mathit{encode}$, let:
+  Given well-typed $\prog_1;E[\mathit{encode}(s_1,s_2)]$ for certified $\mathit{encode}$, let:
   $$
-  \config{\prog_1}{E[\mathit{encode}(\ttt{"s1"},\ttt{"s2"})]} \redxs
+  \config{\prog_1}{E[\mathit{encode}(s_1,s_2)]} \redxs
   \config{\prog_1;\prog_2}{E[v]}
@@
   \label{lemma-ygc-encode}
-Given well-typed $\prog_1;E[\mathit{encode}(s1,s_2)]$ and certified $\mathit{encode}$ where 
+Given well-typed $\prog_1;E[\mathit{encode}(s_1,s_2)]$ and certified $\mathit{encode}$ where 
 %$$
@@
 
-Given the above, we can now establish that any circuit built with
+Given the above, we can establish that any circuit built with
 certified components is passive secure.  Note that we assume a normal

### 2023-07-12 popl24/proofs.tex
checkpoint

@@
 \begin{proof}
-  By assumptions of well-typedness we have:
+  By assumptions of well-typedness we have for some $\ttt{"s1"}$ and $\ttt{"s2"}$:
   $$
@@
 \begin{proof}
-  By Definition \ref{definition-ygcencode-certification} we have:
+  By Definition \ref{definition-ygcencode-certification} we have for some
+  $\ttt{"s1"}$ and $\ttt{"s2"}$: 
   \begin{mathpar}

### 2023-11-05 popl24/scratch.tex
checkpoint

@@
-{\footnotesize
-\begin{verbatimtab}
-  owl(gid) { { k = flip[2,"gate:" || gid || ".k"]; p = flip[2,"gate:" || gid || ".p"] } }
+\documentclass[acmsmall,screen,review]{acmart}
 
-  encode(sa, sb) {
-    let owl1 = owl(sa) in
-    let owl2 = owl(sb) in
-    v[1,gate: || sa || 1.k] := OT(s[1,sa],owl1.k,(not owl1.k));
-    v[1,gate: || sa || 1.p] := OT(s[1,sa],owl1.p,(not owl1.p));
-    v[1,gate: || sb || 2.k] := select(s[2,sb],owl2.k,(not owl2.k));
-    v[1,gate: || sb || 2.p] := select(s[2,sb],owl2.p,(not owl2.p));
-    { wv1 = { k = v[1,"gate:" || sa || "1.k"]; p = v[1,"gate:" || sa || "1.p"] };
-      wv2 = { k = v[1,"gate:" || sb || "2.k"]; p = v[1,"gate:" || sb || "2.p"] } }
-  }
+\usepackage{amsmath}
+\usepackage{amstext}
+\usepackage{fp-frame}
+\usepackage[latin1]{inputenc}
+\usepackage{mathpartir}
+\usepackage{fancyvrb}
+\usepackage{moreverb}
+\usepackage{stmaryrd}
+\usepackage{enumerate}
+\usepackage{thmtools,thm-restate}
+\usepackage{comment}
 
-  andgg(g, ga, gb, wva, wvb) { andgate(g, ga, gb); evalgate(g, wva, wvb) }
+\input{macros}
 
-  decode(g, wv) { garbledecode(wv.g); evaldecode(v) }
-\end{verbatimtab}
-}
+\begin{document}
 
-{\footnotesize
-\begin{verbatimtab}
-  let secrets = encode("s1","s2") in v[0,output] := decode(andgg("0", secrets.wv1, secrets.wv2)) \end{verbatimtab}
-}
+\title{Response to Ian's comments on $\metaprot$ to $\minifed$}
 
-{\footnotesize
-\begin{verbatimtab}
-  owl    :   string(gid)
+\maketitle
 
-  encode :   string(sa) * string(sb)
+\section{Model (Over)simplifications}
 
-  andgg  :  string(g) * { k : bool[1]; p : bool[1] } *  { k : bool[1]; p : bool[1] }
+Ian comments that reviewers may push back on (1) our definition of OT
+(actually borrowed from \cite{barthe2019probabilistic}) relies on a
+trusted 3rd party for shared randomness, and (2) reviewers may push
+back on our linear program and reasoning models that don't capture
+concurrency that is a common element of MPC settings.
 
-  decode :  string(g) * { k : bool[1]; p : bool[1] }
-\end{verbatimtab}
-}
+In fact, these were both strong criticisms of the POPL reviewers.
+They are definitely valid criticisms, but I think we'd need to enrich
+our framework to add them. For example, we could add a notion of
+public key encryption as a language primitive and encode semi-honest
+OT using the protocol described in \cite{evans2018pragmatic}, unless
+there's a way to express public key cryptography in the binary field
+(as we encode OTP symmetric key encryption) that I'm not immediately
+seeing. Concurrency would be a much bigger lift- though this could
+perhaps be addressed using techniques in \cite{MITCHELL2006118}.  I
+note that \cite{barthe2019probabilistic} also provides no support for
+concurrent reasoning, and they were the source for the OT protocol we
+used, and honestly I sort of hoping that we would get the same
+indulgences from reviewers on these points, especially since we were
+avoiding any claims about UC. But I guess Barthe et al.~were targeting
+probabilistic reasoning generally, rather than MPC specifically where
+these issues are more central.
+
+\section{Bridging Formal and Computational Views}
+
+Ian cites \cite{10.1007/3-540-44929-9_1,MITCHELL2006118} as examples
+of ``proving cryptographic security using PL techniques for
+traditional cryptography''. I believe the main point here is about
+the relation of our work to previous work on bridging the gap between
+formal and computational views of security. Since we have similar
+aims the relation is clear at a high level.
+
+Technically speaking there are probably some strong contrasts with
+these citations explicitly. For example,
+in \cite{10.1007/3-540-44929-9_1} the formal view is characterized by
+an equivalence relations on expressions that axiomitizes $\{
+0 \}_K \simeq \{ 1 \}_{K'}$ based on primitize assumptions of PK
+encryption. However, our goal is a lower level analysis that could
+conceivably establish these primitive properties of underlying
+protocols.
+
+The methods developed in \cite{10.1007/3-540-44929-9_1} are very
+sophisticated and perhaps more powerful than needed for the context
+we've been working in. For example, they develop a contextual
+congruence relation that looks as if it could support reasoning
+about UC properties of protocols. But we have definitely been
+considering UC to be out of scope (at least initially). On the
+other hand, their machinery for concurrent reasoning could be
+extremely helpful as disussed above.
+
+\section{Dimensions of Declassification}
+
+I agree it would be interesting to do a ``who what were when''
+analysis for MPC. Another very interesting feature of this work
+is the partial equivalence relation (PER) model.  Our notion of
+``ideal knowledge'' is adapted from the notion of knowledge as defined
+in \cite{10.1007/978-3-540-37621-7_9}, and the PER model is clearly
+related to this. More generally, any ideal functionality $F$ induces a
+PER (an extensional PER, more precisely) I'll call $\mathit{IdealDeclass}$
+that characterizes the downgrading/declassification/delimitation policy,
+where $=_C$ is low equivalence (up to equivalence of inputs from corrupt
+parties C):
+$$
+\forall \bar{x}_1,\bar{x}_2\ .\ F(\bar{x}_1) =_C F(\bar{x}_1) \iff \bar{x}_1
+\ \mathit{IdealDeclass}\ \bar{x}_2  
+$$
+
+However, I'll note that our use of joint probability distributions
+provides more subtle machinery that could be more useful, perhaps
+when integrating subroutines (OT) that enjoy computational hardness
+(vs.~purely formal) correctness properties. For example, if an
+MPC output $y$ is known, then any $\bar{x}$ with $F(\bar{x}) = y$
+is equally likely according to the ``Sabelfeld style'' formulation
+of knowledge, whereas we can assign different $\bar{x}$ different
+probabilities in a joint probability distribution.
+
+\section{Comparison with Downgrading Policies}
+
+Ian mentions Viaduct \cite{10.1145/3453483.3454074} which arguably
+represents the current state-of-the-art in information flow frameworks
+with declassition and endorsement. This was missed related work on our
+part, but I would argue that it is really not directly related to our
+work, for two critical reasons that we are having trouble explaining.
+
+\emph{First}, one general comment about this body of work is that
+downgrading policies as such are defined by the declassifications and
+endorsements themselves. For example, gradual release \cite{4223226}
+is the policy that says only explicitly declassified information is
+released, and robust declassification \cite{930133} is a policy that
+guarantees that an active adversary can't learn any more information
+through declassifications than a passive adversary. Some work shows
+how ``intentional'' release policies are derived syntactically from
+program declassifications as reviewed in
+\cite{10.5555/1662658.1662659}, but these policy expressions are sort
+of awkward and course-grained.
+
+In contrast, MPC establishes an ``upper bound'' on knowledge
+(specifically the PER $\mathit{IdealDeclass}$ defined above if we are
+thinking that way), and our goal is to ensure that declassification
+does not overstep that bound. In fact, we have no explicit
+``declassify'' construct (though perhaps view assignments aka message
+passing are arguably declassifications). Rather, we argue that the
+probability mass function (pmf) is an easily understood ``knowledge
+metric'', and we aim to compute the actual mass function associated
+with the protocol, for a direct comparison of the (ideal) upper bound
+and actual (real) bound.
+
+Going back to Viaduct, my understanding is that this system in the robust
+declassification school of thought, where labeling and explicit
+downgrading and endorsement establish the policy. They also include
+``nonmalleability'' as the dual of robust declassification
+\cite{10.1145/3133956.3134054}. A main result is that their compiler
+can exploit MPC and other protocols in the back end to support the
+policy defined by the security annotations in programs. But they never
+aim to compare this ``syntactic'' policy against an extensional policy
+(as we do). Another point is that they don't need to consider multiple
+partitionings of corrupt and honest parties as is needed in MPC.
+
+\emph{Second}, another general difference is that Viaduct for example
+assumes correctness of the MPC protocols they compile to. In this way
+Viaduct is similar to Wysteria, for example
+\cite{rastogi2014wysteria}. Our goal is to establish a general
+frameword for (semi)-automated verification of the MPC protocols
+themselves.  This is why our mass function formulation, and methods
+for iterating through corrupt/honest partitions, for example, are
+important technical details of our system.
+
+Unfortunately, I think we were not able to persuade reviewers on this
+point because the automated component of the analysis is exponential
+and the manual component looks a lot like probabilistic separation
+logic, and is manual. Not sure how to remedy that- though formulating
+this within some mechanized metatheory, as a reviewer suggested, would
+be an obvious way.
+
+\section{Integrity}
+
+This is a bit speculative but since we are working in a space where
+ideal functionalities $F$ are deterministic, we can formulate
+integrity via the PER $\mathit{IdealEndorse}$ that characterizes the
+following endorsement policy, where $=_H$ is high equivalence (up to
+equivalence of inputs from corrupt parties H):
+$$
+\forall \bar{x}_1,\bar{x}_2\ .\ F(\bar{x}_1) =_H F(\bar{x}_1) \iff \bar{x}_1
+\ \mathit{IdealEndorse}\ \bar{x}_2  
+$$
+
+However, to capture either nonmalleability or robust declassification
+we would need to extend our threat model to capture ``attacks'' as
+defined in \cite{930133}. Our model is essentially passive. This is
+non-trivial but I think can clearly be accomplished.
+
+The potential application to verifiable computing is an interesting
+suggestion.  I frankly need to better understand the relevant security
+properties, but in any case, I do think it's possible to formulate notions
+of both robust declassification and nonmalleability in our
+(probabilistic) language setting. And I suspect there are relations to
+security for verifiable computing and nonmalleability, as well as
+malicious security and robust declassification. In particular I
+speculate that NIMO + robust declassification implies malicious
+security.
+
+\section{Generalizations to Realistic Settings}
+
+I need to spend some time understanding the implementation of SPDZ
+with SHE that Ian cites. However as a general comment I think our
+approach is modular- for example if we replaced our OT primitive where
+we can express its information-leakage characteristics as a pmf, we
+can drop that directly into our framework. Anyway we need to reflect
+on this more, especially related to the main critiques of
+oversimplification especially wrt OT.
+
+\bibliographystyle{ACM-Reference-Format}
+%\bibliographystyle{alpha}
+\bibliography{logic-bibliography,secure-computation-bibliography}
+
+\end{document}

### 2023-11-05 revision-1/bruteforce.tex
Add files via upload

--- 2023-07-09 popl24/bruteforce.tex
@@
-\section{Brute Force Computation of Program Distributions}
+\section{Rewriting $\minicat$ to Stratified Datalog}
 \label{section-bruteforce}
-
-We first explore a brute force approach to automatically verifying
-program behaviour, the first step of which is to compute the basic
-distribution of a program $\prog$. The program distribution
-$\progd(\prog)$ can obviously be automatically derived from that, as
-can be any of its marginal or conditional distributions. In this
-Section, we consider two automated mechanisms for deriving
-$\progd(\prog)$ in this way. The first is through a straightforward
-computation of all final program memories.  The second approach
-rewrites $\minifed$ programs to stratified Datalog programs, that are
-amenable HPC optimizations such as parallelization and GPU matrix
-computations as shown in recent work
-\cite{sakama2017linear,aspis2018linear,nguyen2022enhancing,nguyen2021efficient}.
-
-%The basic distribution of a program $\prog$ is based on the
-%final memories resulting from any run of $\prog$, denoted
-%$\runs(\prog)$.
-%\begin{definition}
-%  Given $\prog$ with $\iov(\prog) = S \cup V$ and $\flips(\prog) = F$:
-%  $$
-%  \runs(\prog) \defeq \{ \store \in \mems(S\cup F \cup V) \mid \config{\store_{S \cup F}}{\prog} \redxs \config{\store}{\varnothing} \}
-%  $$
-%\end{definition}
-
-As we've observed, $|\runs(\prog)| = 2^{|S \cup F|}$, so the size of
-$\runs(\prog)$ is exponential in the amount of randomness used in the
-protocol which generally can be used as a proxy for the protocol
-size. In Section \ref{section-composition} we will show how the
-methods described here can be used to individually verify
-compositional properties of small program components, that guarantee
-security properties in larger protocols that combine them. Thus, while
-scalability to program size is a clear challenge for our brute force
-techniques alone, our ultimate goal is to use them for verification of
-compositional properties in larger programs.
-
-\begin{fpfig}[t]{Filtering memories that satisfy a boolean expression.}{fig-solve}
-\begin{eqnarray*}
-\solve{\stores}{\etrue} &=& \stores\\
-\solve{\stores}{\efalse} &=& \varnothing\\
-\solve{\stores}{\flip{\cid}{w}} &=& \{ \store \in \stores \mid \store(\flip{\cid}{w}) \} \\
-\solve{\stores}{\secret{\cid}{w}} &=& \{ \store \in \stores \mid \store(\secret{\cid}{w}) \} \\
-\solve{\stores}{\view{\cid}{w}} &=& \{ \store \in \stores \mid \store(\view{\cid}{w}) \} \\
-\solve{\stores}{\oracle{w}} &=& \{ \store \in \stores \mid \store(\oracle{w}) \} \\
-\solve{\stores}{(\enot\ \be)} &=& \stores - (\solve{\stores}{\be})\\
-\solve{\stores}{(\be_1\ \eand\ \be_2)} &=& (\solve{\stores}{\be_1}) \cap (\solve{\stores}{\be_2}) \\
-\solve{\stores}{(\be_1\ \eor\ \be_2)} &=& (\solve{\stores}{\be_1}) \cup (\solve{\stores}{\be_2}) \\
-\solve{\stores}{(\be_1\ \exor\ \be_2)} &=&
- ((\solve{\stores}{\be_1}) \cap (\stores - \solve{\stores}{\be_2})) \cup\\
- && ((\stores - \solve{\stores}{\be_1}) \cap (\solve{\stores}{\be_2})) \\
-\solve{\stores}{\select{\be_1}{\be_2}{\be_3}} &=&
- ((\solve{\stores}{\be_1}) \cap (\solve{\stores}{\be_2})) \cup \\
- && ((\stores - \solve{\stores}{\be_1}) \cap (\solve{\stores}{\be_3})) \\
-\solve{\stores}{\OT{\be_1}{\be_2}{\be_3}} &=&
- ((\solve{\stores}{\be_1}) \cap (\solve{\stores}{\be_2})) \cup\\
- && ((\stores - \solve{\stores}{\be_1}) \cap (\solve{\stores}{\be_3}))
-\end{eqnarray*}
-\end{fpfig}
-
-\subsection{Method 1: Inductive Computation of $\runs(\prog)$}
-\label{section-bruteforce-tables}
-
-Here we describe a straightforward technique for obtaining program
-distributions $\progd(\prog)$ through direct computation of
-$\runs(\prog)$. Although we could obtain $\runs(\prog)$ simply by
-executing $\prog$ given all random tapes and input secrets, the method
-we describe here is empirically more efficient (though still
-exponential in the length of the random tape), and also supports the
-translation to Datalog described in Section
-\ref{section-bruteforce-datalog}. An implementation of these
-techniques along with Examples from Sections
-\ref{section-minicat-examples} and \ref{section-metalang-ygc} is
-available online\footnote{URL redacted for double-blind review.}.
-
-In Figure \ref{fig-solve} we define the algorithm $\solve{\stores}{\be}$
-which filters $\stores$ based on satisfaction of $\be$. This algorithm
-is fundamental to our brute force methods. It's correctness is characterized
-as follows. 
-\begin{lemma}
-  \label{lemma-solves}
-  For all $\stores$ and $\be$ with $\vars(\be) \subseteq \dom(\stores)$,
-  $(\solve{\stores}{\be}) = \{ \store \in \stores \ \mid\ \lcod{\store,\be}{\cid} = 1 \}$
-  for some $\cid$.
-\end{lemma}
-\begin{proof}
-  Since we assume safety of programs we can assume that all variables in $\be$ have the
-  same owner $\cid$. The result otherwise follows in a straightforward manner by induction
-  on $\be$. 
-\end{proof}
-
-In order to automatically obtain $\runs(\prog)$, we can do a left
-folding of $\solve$ across $\prog$, inductively extending memories
-with valid view assignments in the order of their definition.
-We can then use this to compute $\progd(\prog)$. We denote
-these computations as $\cruns$ and $\cprogd$ respectively.
-\begin{lemma}
-  \label{lemma-cruns}
-  Given $\prog$ where $\iov(\prog) = S \cup V$ and $\flips(\prog) = F$. Define:
-  \begin{eqnarray*}
-    {tt}\ \ \stores\ (\eassign{\view{\cid}{w}}{\be}) &\defeq& \begin{array}{l}
-      \mathrm{let}\ \stores' = \solve{\stores}{\be} \ \mathrm{in}\\
-      \ \ \{\extend{\store}{\view{\cid}{w}}{1} \mid \store \in \stores' \}\ \cup\\
-      \ \ \{\extend{\store}{\view{\cid}{w}}{0} \mid \store \in \stores - \stores' \}\end{array}\\[2mm]
-    \cruns(\prog) &\defeq& \mathit{foldl}\ {tt}\ \mems(S \cup F)\ \prog
-  \end{eqnarray*}
-  Then $\cruns(\prog) = \runs(\prog)$.
-\end{lemma}
-\begin{proof}
-  By Lemma \ref{lemma-solves} and induction on the length of $\prog$. 
-\end{proof}
-\begin{lemma}
-  \label{lemma-cprogd}
-  Given $\prog$ where $\iov(\prog) = S \cup V$ and $\flips(\prog) = F$. Define
-  $\cprogd(\prog)$ as the distribution $\pmf$
-  where for all $\store \in \mems(S\cup F \cup V)$:
-  $$
-  \pmf(\store) = \begin{cases}1/2^{|S \cup F|} & \text{if\ } \store \in
-    \cruns(\prog) \\ 0 & \text{otherwise} \end{cases} 
-  $$
-  Then:
-  $$
-  \cprogd(\prog) = \progd(\prog)
-  $$
-\end{lemma}
-\begin{proof}
-  Immediately by Lemma \ref{lemma-cruns} and Definition \ref{def-progd}.
-\end{proof}
-
-%%% I moved the following to automation.tex - chris
-\begin{comment}
-We have used the code for $\cruns$ and $\cprogd$ extensively to design and
-verify the examples in this paper. We will discuss other examples in
-Section \ref{section-composition}, but here we note that we've verified
-passive security of Examples \ref{example-he} and \ref{example-OT}-
-for the former, this is well-known so the result is a simple proof-of-concept,
-but the latter is a new (though unsurprising) result, having not been
-directly proven in \cite{barthe2019probabilistic}. The method for
-verifying $\NIMO$ described here generalizes to any $\prog$. 
-\begin{lemma}[Automated Verification of Passive Security]
-  \label{lemma-bruteforce-nimo}
-  $\prog_{\ref{example-he}}$ and $\prog_{\ref{example-OT}};\eassign{\outv}{\vx{2}{mc}}$ are
-  passive secure.
-\end{lemma}
-\begin{proof}
-  For both $\prog \in \{
-  \prog_{\ref{example-he}},\prog_{\ref{example-OT}};\eassign{\outv}{\vx{2}{mc}}\}$
-  we compute $\cruns(\prog)$ and $\cprogd(\prog)$ and then verify
-  $\NIMO(\prog)$ using the following algorithm. Given $\iov(\prog) = S
-  \cup V$ and $\flips(\prog) = F$, we enumerate all memories $\store
-  \in \cruns(\prog)$, then enumerate all partitions of clients into
-  sets $H$,$C$ such that $|C| \le |H|$ and $0 \in C$, and then check:
-  $$
-  \cprogd(\prog)(\store_{S_H}|\store_{S_C \cup \{\outv\}}) =
-  \cprogd(\prog)(\store_{S_H}|\store_{S_C \cup V_C})
-  $$
-  for each $H,C,\store$. The result follows by Lemma \ref{lemma-cprogd}, Lemma \ref{lemma-nimo}, and
-  Theorem \ref{theorem-nimo}.
-\end{proof}
-
-\end{comment}
-
-\subsection{Method 2: Conversion to Datalog}
-\label{section-bruteforce-datalog}
 

### 2023-11-05 revision-1/conclusion.tex
Add files via upload

--- 2023-07-12 popl24/conclusion.tex
@@
 including extensions to malicious security, and to larger fields and
-arithmetic circuits. HPC optimizations of program distribution
+arithmetic circuits. HPC optimization of program distribution
 calculations, to support analysis of larger circuit components, is

### 2023-11-05 revision-1/gmw.tex
Add files via upload

--- 2023-07-10 popl24/gmw.tex
@@
-\begin{fpfig}[t]{2-Party GMW Circuit Library with And gate.}{fig-gmw}
+\begin{fpfig}[t]{2-Party GMW circuit library with And gate (T), and type annotations (B).}{fig-gmw}
 {\footnotesize
@@
 
-    decodegmw(shares) { v[0,"output"] := shares.c1 xor shares.c2 }
-  \end{verbatimtab}
+    decodegmw(shares) { v[0,"output"] := shares.c1 xor shares.c2 }   \end{verbatimtab}
+}
+
+\rule{130mm}{0.5pt}
+
+{\footnotesize
+  \begin{verbatimtab}
+   encodegmw   : string(gid) * string(gid)
+
+   andtablegmw : { k = bool[i]; p = bool[i] }
+
+   andgmw      : string(gid) *  { c1 = bool[1]; c2 = bool[2] } * { c1 = bool[1]; c2 = bool[2] }
+
+   decodegmw   : { c1 = bool[1]; c2 = bool[2] }  \end{verbatimtab}
 }

### 2023-11-05 revision-1/intro.tex
Add files via upload

--- 2023-07-11 popl24/intro.tex
@@
 
-We distinguish between \emph{extensional} vs.~\emph{intensional}
-properties and analysis of MPC protocols. By extensional, we mean
-simulator security itself as well as analysis concerned with the use
-and interaction of complete secure protocols. By intensional, we mean
-analysis and inner workings of protocols themselves
-
 \paragraph{Language design.} In Section \ref{section-minicat} we
@@
 dependencies between secrets and views, useful for expressing
-intensional characteristics of protocols and precisely quantifying
-allowable information leakage in terms of protocol outputs. On this
-basis we can then formulate extensional hyperproperties, including
-\emph{noninterference modulo output ($\NIMO$)} that implies passive
-simulator security. But further, we can formulate intensional
-protocol properties such as probabilistic conditioning of program
-distributions.
+characteristics of protocols and precisely quantifying allowable
+information leakage in terms of protocol outputs. On this basis we can
+formulate security hyperproperties, including \emph{noninterference
+modulo output ($\NIMO$)} that implies passive simulator security. But
+program distributions are flexible and can be used to certify other
+important properties of protocol components, such as probabilistic
+independence, correlation, and distributions under conditioning.
 
@@
 \ref{section-automation} we develop an automated technique, aka
-certification, that verifies extensional properties of protocols such
-as $\NIMO$, and also intensional properties of protocol
-components. The latter include properties of protocol components that
-provably guarantee security through composition in large
-circuits. These compositional proofs are based on separation logic but
-include novel results to support our semi-automated method for large
-circuits where we (1) define certification for components, (2) certify
-components, (3) prove $\NIMO$ based on the certifiedness of
-components. We develop novel methods partially based on separation
-logic \cite{barthe2019probabilistic} to enable (3).  In Section
+certification, for automatically verifying security properties such as
+$\NIMO$ and other supporting properties of protocol components.  While
+certification has high complexity, a compositional approach allows
+certification of program components in isolation.  This enables a
+semi-automated method for large circuits where we (1) define
+certification for components, (2) certify components, (3) prove
+$\NIMO$ based on the certifiedness of components. We develop novel
+methods partially based on separation logic
+\cite{barthe2019probabilistic} to enable (3).  In Section
 \ref{section-composition} we illustrate our method to obtain
-semi-automated verification methods for the GMW and Yao's Garbled
+semi-automated verification for the GMW and Yao's Garbled
 Circuit protocols, where certification can be re-used to validate new
@@
 Our main focus is on automated and semi-automated reasoning about
-intensional properties of protocols, both through whole-program
+security properties of specific protocols, both through whole-program
 analysis and compositional properties of program components that
@@
 \cite{almeida2018enforcing,10.1145/2637113.2637119} is concerned with
-extensional properties of MPC, in particular the specification and
-enforcement of declassification bounds in programs that use MPC in
-subprograms. This work is explicitly reminiscent of information flow
-approaches such as delimited information release
-\cite{10.1007/978-3-540-37621-7_9}, downgrading policies \cite{li2005downgrading}, and
-relaxed noninterference \cite{10.1145/1040305.1040319}. However, their
-program logic assumes correctness of the underlying MPC protocols.
-The Wys$^\star$ language \cite{wysstar}, based on Wysteria
-\cite{rastogi2014wysteria}, has similar goals and includes a
-trace-based semantics for reasoning about the extensional interactions
-of MPC protocols. Their compiler also guarantees that underlying
-multi-threaded protocols enforce the single-threaded source language
-semantics which is an intensional property. \mnote{We need a clearer statement here of the flaws with Wys-star specifically} All of these systems
-leverage type systems to obtain automated security analyses.
+properties of complex MPC circuits, in particular a user-friendly
+specification and automated enforcement of declassification bounds in
+programs that use MPC in subprograms. This work is explicitly
+reminiscent of information flow approaches such as delimited
+information release \cite{10.1007/978-3-540-37621-7_9}, downgrading
+policies \cite{li2005downgrading}, and relaxed noninterference
+\cite{10.1145/1040305.1040319}. However, their program logic assumes
+correctness of the underlying MPC protocols.  The Wys$^\star$ language
+\cite{wysstar}, based on Wysteria \cite{rastogi2014wysteria}, has
+similar goals and includes a trace-based semantics for reasoning about
+the interactions of MPC protocols. Their compiler also guarantees that
+underlying multi-threaded protocols enforce the single-threaded source
+language semantics. However their language ultimately compiles to
+protocols assumed to be secure.  All of these systems leverage type
+systems to obtain automated security analyses.
 
-In the realm of intensional analysis, previous work has shown how to
+In the realm of protocol verification, previous work has shown how to
 reduce proofs of simulator security to program equivalence assertions,
@@
 probabilistic independence of input secrets and adversarial views that
-are critical intensional properties of simulator security.  However,
-their focus on independence does not address the subtleties of
+are critical security properties.  However, their methods are manual. 
+Also, their focus on independence does not address the subtleties of
 dependence that arise during MPC protocol execution, and the authors

### 2023-11-05 revision-1/metalang.tex
Add files via upload

--- 2023-07-12 popl24/metalang.tex
@@
   let secrets = encode("s1","s2") in
-  v[0,output] := decode(andgg("0", secrets.wv1, secrets.wv2)) \end{verbatimtab}
+  v[0,"output"] := decode(andgg("0", secrets.wv1, secrets.wv2)) \end{verbatimtab}
 }

### 2023-11-05 revision-1/minicat.tex
Add files via upload

--- 2023-07-11 popl24/minicat.tex
@@
 We let $x$ range over variables, and will use $s$, $v$, and $f$ to
-range over individual secrets, views, and flips respecitively. We let
+range over individual secrets, views, and flips respectively. We let
 $X$ range over sets of variables and $S,V,F$ to range over sets of
@@
 \ref{section-metalang} we will formulate a larger example
-implementation of Yao's Garbled Circuits.
+implementation of Yao's Garbled Circuits. Additional examples
+are provided in Appendix \ref{section-examples-lang} including 3-party
+additive secret sharing.
 \begin{example}[One-Time Pad]

### 2023-11-05 revision-1/proofs.tex
Add files via upload

--- 2023-07-12 popl24/proofs.tex
@@
     &\prog_2 = \\
-    &{\small \prog[\be^1_{k}/\ttt{flip[2,"gate:a.k"]}][\be^1_{p}/\ttt{flip[2,"gate:a.p"]}][\be^2_{k}/\ttt{flip[2,"gate:b.k"]}][\be^1_{p}/\ttt{flip[2,"gate:b.p"]}]}
+    &\begin{array}{l}\prog[\be^1_{k}/\ttt{flip[2,"gate:a.k"]}][\be^1_{p}/\ttt{flip[2,"gate:a.p"]}]\\
+       \phantom{\prog}[\be^2_{k}/\ttt{flip[2,"gate:b.k"]}][\be^1_{p}/\ttt{flip[2,"gate:b.p"]}]\end{array}
   \end{eqnarray*}

### 2023-11-05 revision-1/ygc.tex
Add files via upload

--- 2023-07-11 popl24/ygc.tex
@@
   {   
-    v[1, gate: || gid || tid || "1"] := table.v1;
-    v[1, gate: || gid || tid || "2"] := table.v2;
-    v[1, gate: || gid || tid || "3"] := table.v3;
-    v[1, gate: || gid || tid || "4"] := table.v4
+    v[1, "gate:" || gid || tid || "1"] := table.v1;
+    v[1, "gate:" || gid || tid || "2"] := table.v2;
+    v[1, "gate:" || gid || tid || "3"] := table.v3;
+    v[1, "gate:" || gid || tid || "4"] := table.v4
   }
 
-  owl(gid) {  { k = flip[2,gate: || gid || .k]; p = flip[2,gate: || gid || .p] }  }
+  owl(gid) {  { k = flip[2,"gate:" || gid || ".k"]; p = flip[2,"gate:" || gid || ".p"] }  }
 \end{verbatimtab}
@@
     let k = keygen(gid,wva.k,wvb.k) in
-    let g = gate: || gid in
+    let g = "gate:" || gid in
     let ct = select4(wva.p,wvb.p,
@@
     let owl2 = owl(sb) in
-    v[1,gate: || sa || 1.k] := OT(s[1,sa],owl1.k,(not owl1.k));
-    v[1,gate: || sa || 1.p] := OT(s[1,sa],owl1.p,(not owl1.p));
-    v[1,gate: || sb || 2.k] := select(s[2,sb],owl2.k,(not owl2.k));
-    v[1,gate: || sb || 2.p] := select(s[2,sb],owl2.p,(not owl2.p));
-    { wv1 = { k = v[1,gate: || sa || 1.k]; p = v[1,gate: || sa || 1.p] };
-      wv2 = { k = v[1,gate: || sb || 2.k]; p = v[1,gate: || sb || 2.p] } }
+    v[1,"gate:" || sa || "1.k"] := OT(s[1,sa],owl1.k,(not owl1.k));
+    v[1,"gate:" || sa || "1.p"] := OT(s[1,sa],owl1.p,(not owl1.p));
+    v[1,"gate:" || sb || "2.k"] := select(s[2,sb],owl2.k,(not owl2.k));
+    v[1,"gate:" || sb || "2.p"] := select(s[2,sb],owl2.p,(not owl2.p));
+    { wv1 = { k = v[1,"gate:" || sa || "1.k"]; p = v[1,"gate:" || sa || "1.p"] };
+      wv2 = { k = v[1,"gate:" || sb || "2.k"]; p = v[1,"gate:" || sb || "2.p"] } }
   }

### 2024-04-14 revision-1/lang.tex
checkpoint

\section{The $\minicat$ Protocol Language}

\begin{fpfig}[t]{Top-to-bottom: $\minifed$ expression interpretation, source code restrictions on keys, and reduction rules.}{fig-minifed}
  {\small
    $$
    \begin{array}{rcl@{\hspace{8mm}}r}
      \multicolumn{4}{l}{v \in \mathbb{Z}_p,\ k \in \mathrm{Key},\ w \in \mathrm{String},\ \cid \in \mathrm{Clients} \subset  \mathbb{N} }\\[2mm] %, \bop \in \{ \eand, \eor, \exor \}} \\[2mm]
      \be &::=& v \mid \flip{w} \mid \secret{w} \mid \key{w} \mid \mesg{w} \mid \elab{\be}{\Oracle} \mid \fminus \be \mid \be \fplus \be \mid \be \ftimes \be \mid & \textit{expressions}\\[1mm]
      %& &  \select{\be}{\be}{\be} \mid \ctxt{v}{k} \mid \key{w} \mid \sk{\be}(\be) \mid \pk{\be}{\be}(\be) \mid \pk{\be}{\be} \\[2mm]
      & &  \select{\fp(\be)}{\be}{\be} \ctxt{v,\be}{k}  \mid \sk{\be}(\be) \mid \pk{\be}{\be}(\be) \mid \pk{\be}{\be} \\[2mm]
      \instr &::=& \eassign{\mesg{w}}{\cid}{\be}{\cid} & \textit{message send} \\[2mm]
      \prog &::=& \varnothing \mid \instr; \prog & \textit{protocols}
    \end{array}
    $$

  \rule{130mm}{0.5pt}

  $$
  \begin{array}{rcl}
    \lcod{\store, v}{\cid} &=& v\\
    \lcod{\store, \fminus \be}{\cid} & = &  \fcod{\fminus \lcod{\store, \be}{\cid}}\\ 
    \lcod{\store, \be_1 \fplus \be_2}{\cid} &=& \fcod{\lcod{\store, \be_1}{\cid} \fplus \lcod{\store, \be_2}{\cid}}\\ 
    \lcod{\store, \be_1 \ftimes \be_2}{\cid} &=& \fcod{\lcod{\store, \be_1}{\cid} \ftimes \lcod{\store, \be_2}{\cid}}\\
    \lcod{\store, \select{\fp(v)}{\be_2}{\be_3}}{\cid} &=&
    %\begin{cases}
      \lcod{\store, \be_2}{\cid}\ \text{if}\ \fcod{\fp(v)}\ \text{valid}, 
      \text{otherwise}\ \lcod{\store, \be_3}{\cid}\\
    %\end{cases}
    \lcod{\store, \select{\fp(\be_1)}{\be_2}{\be_3}}{\cid} &=& \lcod{\store, \select{\fp(\lcod{\be_1}{\cid})}{\be_2}{\be_3}}{\cid}\\
    \lcod{\store, \flip{w}}{\cid} &=& \store_\cid(\flip{w})\\
    \lcod{\store, \secret{w}}{\cid} &=& \store_\cid(\secret{w})\\
    \lcod{\store, \mesg{w}}{\cid} &=& \store_\cid(\mesg{w})\\
    \lcod{\store, \key{w}}{\cid}  &=&  \store_\cid(\key{w})\\
    \lcod{\store, \elab{\flip{w}}{\Oracle}}{\cid} &=& \store_\Oracle(\flip{w})\\
    \lcod{\store, \elab{\key{w}}{\Oracle}}{\cid} &=& \store_\Oracle(\key{w})\\
    \lcod{\store, \sk{k}(\ctxt{v,\_}{k})}{\cid} &=& v\\
    \lcod{\store, \sk{\be_1}(\be_2)}{\cid} &=& \lcod{\store, \sk{\lcod{\store,\be_1}{\cid}}(\lcod{\store, \be_2}{\cid})}{\cid}\\
    \lcod{\store, \pk{k}{\_}(v)}{\cid} &=& \ctxt{v,\elab{\flip{\mathit{fresh}}}{\Oracle}}{k}\\
    \lcod{\store, \pk{\be_1}{\be_2}(\be_3)}{\cid} &=&  
       \lcod{\store, \pk{\lcod{\store,\be_1}{\cid}}{\lcod{\store, \be_2}{\cid}}(\lcod{\store, \be_3}{\cid})}{\cid}\\
    \lcod{\store, \pk{k}{v}}{\cid} &=& \pk{k}{v}\\
    \lcod{\store, \pk{\be_1}{\be_2}}{\cid} &=& \lcod{\store, \pk{\lcod{\store,\be_1}{\cid}}{\lcod{\store, \be_2}{\cid}}}{\cid}
  \end{array} 
  $$

  \vspace{4mm}

  \rule{130mm}{0.5pt}
  \begin{mathpar}
    \sk{\key{w}}

    \pk{\key{w}}{\elab{\flip{w}}{\Oracle}}

    \pk{\elab{\key{w}}{\Oracle}}{\elab{\flip{w}}{\Oracle}}
  \end{mathpar}  
  \rule{130mm}{0.5pt}

  \begin{mathpar}
    (\store, \eassign{\mesg{w}}{\cid_1}{\be}{\cid_2};\prog) \redx (\extend{\store}{\mesg{w}_{\cid_1}}{\lcod{\store,\be}{\cid_2}}, \prog)
  \end{mathpar}
  }
\end{fpfig}

### 2024-04-17 revision-1/hyper.tex
checkpoint

\section{Security Hyperproperties}

\begin{definition}[Passive Confidentiality]
  \label{definition-NIMO}
  We say that a program $\prog$ satisfies \emph{passive confidentiality}
  iff for all $H$ and $C$ with $|C|\le|H|$ and 
  $\store_1,\store_2 \in \mems(S)$ we have:
  $$
  (\store_1 =_C \store_2 \ \wedge \ 
  (\condd{\progtt(\prog)}{\{ \outv \}}{\store_1} = \condd{\progtt(\prog)}{\{ \outv \}}{\store_2}))
  \implies 
  (\condd{\progtt(\prog)}{\houtputs}{\store_1} = \condd{\progtt(\prog)}{\houtputs}{\store_2})
  $$
  where $\iov(\prog) = (S,M)$.
\end{definition}

\begin{definition}
  We say that a protocol $\prog$ with $\iov(\prog) = (S,M)$ satisfies \emph{active confidentiality} iff the following conditions hold
  for all adversaries $\adversary$:
  \begin{enumerate}
  \item $\ \,\forall \store \in \mems(S_H) \ .\ \support(\progtt(\prog)(\{ \outv \}|\store)) =
    \support(\progtt(\prog,\adversary)(\{ \outv \}|\store))$
  \item $\begin{array}[t]{l}\forall \store_1, \store_2 \in \mems(S_H), \store \in \mems(\cinputs)\ . \\
    \quad
    \condd{\progtt(\prog,\adversary)}{\{ \outv \}}{\store_1 \cup \store} =
    \condd{\progtt(\prog,\adversary)}{\{ \outv \}}{\store_2 \cup \store} \\
    \qquad \implies\\
    \quad
    \condd{\progtt(\prog,\adversary)}{\houtputs}{\store_1 \cup \store} =
    \condd{\progtt(\prog,\adversary)}{\houtputs}{\store_2 \cup \store}\end{array}$
  \end{enumerate}
\end{definition}

\begin{theorem}
  If $\prog$ satisfies active confidentiality for all $H,C$ then it is malicious secure.
\end{theorem}

\begin{definition}
  A protocol $\prog$ with $\iov(\prog) = (S,M)$ has \emph{integrity} iff 
  $\forall \adversary . \runs_\adversary(\prog) \subseteq \runs(\prog)$.
\end{definition}

\begin{definition}
  A protocol $\prog$ with $\iov(\prog) = (S,M)$ is \emph{malicious correct} iff:
  $$
  \forall \adversary, \store \in \mems(S_H) \ .\ \support(\progtt(\prog)(\{ \outv \}|\store)) =
    \support(\progtt(\prog,\adversary)(\{ \outv \}|\store))
  $$
\end{definition}

\begin{theorem}
  If a protocol has integrity it is malicious correct.
\end{theorem}

\begin{theorem}
  If a protocol is passive secure with integrity, then it satisfies active confidentiality.
\end{theorem}

\begin{theorem}
  If a protocol is passive secure with integrity, then it is malicious secure.
\end{theorem}

### 2024-04-17 revision-1/model.tex
checkpoint

\section{Security Model}

\subsection{Passive Security}

The simulator is represented by a probabilistic algorithm $\SIM_C$,
aka a \emph{simulation}, that is parameterized by corrupt inputs and
the output of an ideal functionality, and that returns a set of
adversarial views (as a memory) with some probability. Given
corrupt inputs $\store$ and ideal functionality output $v$,  
we write
$
\prob(\SIM(\store,v) = \store')
$
to denote the probability that $\SIM(\store,v)$
returns corrupt views $\store'$ as a result. We can then define the
probability distribution of corrupt views reconstructed
by the simulator as follows:
\begin{definition}
  Given $C$, $\store$, and $v$, we write $\dist(\SIM(\store,v))$ to
  denote the distribution of corrupt views reconstructed by the
  simulation, where for
  all $\store' \in \mems(V)$:
  $$
  \dist(\SIM(\store,v))(\store')\ \defeq\ \prob(\SIM(\store,v) = \store') 
  $$
\end{definition}

Then we can define passive security in the real/ideal
model as follows. 
\begin{definition}[Passive Security]
  Assume given a program $\prog$ that correctly implements an ideal
  functionality $\idealf$, with $\iov(\prog) = (S,M,O)$.  Then $\prog$
  is \emph{passive secure in the simulator model} iff for all
  partitions of the federation into honest and corrupt sets $H$ and $C$
  with $|C| < |H|$ and for all $\store \in \mems(S)$ there exists a
  simulation $\SIM$ such that:
  $$
  \dist(\SIM(\store_{S_C},\idealf(\store))) = \condd{\progtt(\prog)}{M_C}{\store}
  $$
\end{definition}

\subsection{Malicious Security}

$$
\begin{array}{rclr}
  (\store, \eassign{\mesg{w}}{\cid_1}{\be}{\cid_2};\prog) &\aredx&
  (\extend{\store}{\mesg{w}_{\cid_1}}{\lcod{\store,\be}{\cid_2}}, \prog) & \cid_2 \in H\\
  (\store, \eassign{\mesg{w}}{\cid_1}{\be}{\cid_2};\prog) &\aredx&
  (\extend{\store}{\mesg{w}_{\cid_1}}{\lcod{\arewrite(\store_C,\be)}{\cid_2}}, \prog) & \cid_2 \in C
\end{array}
$$

\begin{definition}[Corrupt Inputs, Honest Outputs]
  Given a program $\prog$ with $\iov(\prog) = (S,M,O)$ , define $\cinputs$ as the
  messages in $M$ sent from corrupt to honest parties:
  $$
  \cinputs = \{\ \elab{\mesg{w}}{\cid}\ \mid\  \elab{\mesg{w}}{\cid} \in M \wedge \eassign{\mesg{w}}{\cid}{\be}{\cid'} \in \prog
  \wedge \cid \in H \wedge \cid' \in C \ \} 
  $$
  and similarly define $\houtputs$ as the messages in $M$ sent from honest to corrupt parties.
  %Define also $(\afilter\ \prog)$ as $\prog$ with all instructions of the form $\eassign{\mesg{w}}{\cid}{\be}{\cid'}$ removed
  %where $\cid \in H \wedge \cid' \in C$.
\end{definition}

$$
\dist(\SIM(\store_{S_H})) = \condd{\progtt(\prog)}{\houtputs \cup O_H}{\store_{S_H}}
$$

### 2024-04-17 revision-1/hyper.tex
checkpoint

@@
 \section{Security Hyperproperties}
+
+\subsection{Passive Security and Noninterference}
+
+\cnote{Fix the $O$ output variable things, and scoping of $H,C$.}
 
@@
 \end{definition}
+
+\subsection{Gradual Release as a Design Pattern}
+
+\begin{definition}
+  Given $H,C$, a protocol $\prog$ with $\iov(\prog) = (S,M,O)$ satisfies \emph{gradual release} iff
+  $\sep{\progtt(\prog)}{M_C}{S_H}$.
+\end{definition}
+
+\begin{theorem}
+  Gradual release implies passive security.
+\end{theorem}
+
+\subsection{Malicious Security and Robust Declassification}
 
@@
 \begin{definition}
-  A protocol $\prog$ with $\iov(\prog) = (S,M)$ has \emph{integrity} iff 
+  A protocol $\prog$ with has \emph{integrity} iff 
   $\forall \adversary . \runs_\adversary(\prog) \subseteq \runs(\prog)$.
@@
 \begin{definition}
-  A protocol $\prog$ with $\iov(\prog) = (S,M)$ is \emph{malicious correct} iff:
+  A protocol $\prog$ with $\iov(\prog) = (S,M,O)$ is \emph{malicious correct} iff:
   $$
-  \forall \adversary, \store \in \mems(S_H) \ .\ \support(\progtt(\prog)(\{ \outv \}|\store)) =
-    \support(\progtt(\prog,\adversary)(\{ \outv \}|\store))
+  \forall \adversary, \store \in \mems(S_H) \ .\ \support(\progtt(\prog)(O_H|\store)) =
+    \support(\progtt(\prog,\adversary)(O_H|\store))
   $$
@@
 \end{theorem}
+
+\begin{definition}[Robust Declassification]
+  A protocol satisfies \emph{robust declassification} iff it has integrity and
+  satisfies gradual release. 
+\end{definition}
+
+\begin{theorem}
+  Robust declassification implies malicious security.
+\end{theorem}

### 2024-04-18 revision-1/lang.tex
checkpoint

@@
 
-\begin{fpfig}[t]{Top-to-bottom: $\minifed$ expression interpretation, source code restrictions on keys, and reduction rules.}{fig-minifed}
-  {\small
+\begin{fpfig}[t]{Top-to-bottom: Basic $\minifed$ syntax, expression interpretation, and reduction rules.}{fig-minifed}
+  {
     $$
     \begin{array}{rcl@{\hspace{8mm}}r}
-      \multicolumn{4}{l}{v \in \mathbb{Z}_p,\ k \in \mathrm{Key},\ w \in \mathrm{String},\ \cid \in \mathrm{Clients} \subset  \mathbb{N} }\\[2mm] %, \bop \in \{ \eand, \eor, \exor \}} \\[2mm]
-      \be &::=& v \mid \flip{w} \mid \secret{w} \mid \key{w} \mid \mesg{w} \mid \elab{\be}{\Oracle} \mid \fminus \be \mid \be \fplus \be \mid \be \ftimes \be \mid & \textit{expressions}\\[1mm]
+      \multicolumn{4}{l}{v \in \mathbb{Z}_p,\ w \in \mathrm{String},\ \cid \in \mathrm{Clients} \subset  \mathbb{N} }\\[2mm] %, \bop \in \{ \eand, \eor, \exor \}} \\[2mm]
+      \be &::=& v \mid \flip{w} \mid \secret{w} \mid \mesg{w} \mid \rvl{w} \mid \be \fminus \be \mid \be \fplus \be \mid \be \ftimes \be \mid f \mid \be\,\be & \textit{expressions}\\[2mm]
       %& &  \select{\be}{\be}{\be} \mid \ctxt{v}{k} \mid \key{w} \mid \sk{\be}(\be) \mid \pk{\be}{\be}(\be) \mid \pk{\be}{\be} \\[2mm]
-      & &  \select{\fp(\be)}{\be}{\be} \ctxt{v,\be}{k}  \mid \sk{\be}(\be) \mid \pk{\be}{\be}(\be) \mid \pk{\be}{\be} \\[2mm]
-      \instr &::=& \eassign{\mesg{w}}{\cid}{\be}{\cid} & \textit{message send} \\[2mm]
+      %& &  \select{\fp(\be)}{\be}{\be} \ctxt{v,\be}{k}  \mid \sk{\be}(\be) \mid \pk{\be}{\be}(\be) \mid \pk{\be}{\be} \\[2mm]
+      \instr &::=& \eassign{\mesg{w}}{\cid}{\be}{\cid} \mid
+      \reveal{w}{e}{\cid} \mid \pubout{\cid}{\be}{\cid} & \textit{commands} \\[2mm]
       \prog &::=& \varnothing \mid \instr; \prog & \textit{protocols}
@@
   $$
+  \begin{array}{c@{\hspace{5mm}}c}
   \begin{array}{rcl}
     \lcod{\store, v}{\cid} &=& v\\
-    \lcod{\store, \fminus \be}{\cid} & = &  \fcod{\fminus \lcod{\store, \be}{\cid}}\\ 
     \lcod{\store, \be_1 \fplus \be_2}{\cid} &=& \fcod{\lcod{\store, \be_1}{\cid} \fplus \lcod{\store, \be_2}{\cid}}\\ 
-    \lcod{\store, \be_1 \ftimes \be_2}{\cid} &=& \fcod{\lcod{\store, \be_1}{\cid} \ftimes \lcod{\store, \be_2}{\cid}}\\
-    \lcod{\store, \select{\fp(v)}{\be_2}{\be_3}}{\cid} &=&
-    %\begin{cases}
-      \lcod{\store, \be_2}{\cid}\ \text{if}\ \fcod{\fp(v)}\ \text{valid}, 
-      \text{otherwise}\ \lcod{\store, \be_3}{\cid}\\
-    %\end{cases}
-    \lcod{\store, \select{\fp(\be_1)}{\be_2}{\be_3}}{\cid} &=& \lcod{\store, \select{\fp(\lcod{\be_1}{\cid})}{\be_2}{\be_3}}{\cid}\\
-    \lcod{\store, \flip{w}}{\cid} &=& \store_\cid(\flip{w})\\
-    \lcod{\store, \secret{w}}{\cid} &=& \store_\cid(\secret{w})\\
-    \lcod{\store, \mesg{w}}{\cid} &=& \store_\cid(\mesg{w})\\
-    \lcod{\store, \key{w}}{\cid}  &=&  \store_\cid(\key{w})\\
-    \lcod{\store, \elab{\flip{w}}{\Oracle}}{\cid} &=& \store_\Oracle(\flip{w})\\
-    \lcod{\store, \elab{\key{w}}{\Oracle}}{\cid} &=& \store_\Oracle(\key{w})\\
-    \lcod{\store, \sk{k}(\ctxt{v,\_}{k})}{\cid} &=& v\\
-    \lcod{\store, \sk{\be_1}(\be_2)}{\cid} &=& \lcod{\store, \sk{\lcod{\store,\be_1}{\cid}}(\lcod{\store, \be_2}{\cid})}{\cid}\\
-    \lcod{\store, \pk{k}{\_}(v)}{\cid} &=& \ctxt{v,\elab{\flip{\mathit{fresh}}}{\Oracle}}{k}\\
-    \lcod{\store, \pk{\be_1}{\be_2}(\be_3)}{\cid} &=&  
-       \lcod{\store, \pk{\lcod{\store,\be_1}{\cid}}{\lcod{\store, \be_2}{\cid}}(\lcod{\store, \be_3}{\cid})}{\cid}\\
-    \lcod{\store, \pk{k}{v}}{\cid} &=& \pk{k}{v}\\
-    \lcod{\store, \pk{\be_1}{\be_2}}{\cid} &=& \lcod{\store, \pk{\lcod{\store,\be_1}{\cid}}{\lcod{\store, \be_2}{\cid}}}{\cid}
+    \lcod{\store, \be_1 \fminus \be_2}{\cid} &=& \fcod{\lcod{\store, \be_1}{\cid} \fminus \lcod{\store, \be_2}{\cid}}\\ 
+    \lcod{\store, \be_1 \ftimes \be_2}{\cid} &=& \fcod{\lcod{\store, \be_1}{\cid} \ftimes \lcod{\store, \be_2}{\cid}}
+  \end{array} & 
+  \begin{array}{rcl}
+    \lcod{\store, \flip{w}}{\cid} &=& \store(\elab{\flip{w}}{\cid})\\
+    \lcod{\store, \secret{w}}{\cid} &=& \store(\elab{\secret{w}}{\cid})\\
+    \lcod{\store, \mesg{w}}{\cid} &=& \store(\elab{\mesg{w}}{\cid})\\
+    \lcod{\store, \rvl{w}}{\cid} &=& \store(\rvl{w})\\
+    \lcod{\store, f\,e_1\,\cdots\, e_n}{\cid} &=& \delta(f,\lcod{\store, e_1}{\cid},\ldots,\lcod{\store,e_n}{\cid})
+  \end{array}
   \end{array}
@@
   \rule{130mm}{0.5pt}
-  \begin{mathpar}
-    \sk{\key{w}}
-
-    \pk{\key{w}}{\elab{\flip{w}}{\Oracle}}
-
-    \pk{\elab{\key{w}}{\Oracle}}{\elab{\flip{w}}{\Oracle}}
-  \end{mathpar}  
-  \rule{130mm}{0.5pt}
 
@@
     (\store, \eassign{\mesg{w}}{\cid_1}{\be}{\cid_2};\prog) \redx (\extend{\store}{\mesg{w}_{\cid_1}}{\lcod{\store,\be}{\cid_2}}, \prog)
+
+    (\store, \reveal{w}{\be}{\cid};\prog) \redx (\extend{\store}{\rvl{w}}{\lcod{\store,\be}{\cid}}, \prog)
+
+    (\store, \pubout{\cid}{\be}{\cid};\prog) \redx (\extend{\store}{\out{\cid}}{\lcod{\store,\be}{\cid}}, \prog)
   \end{mathpar}
@@
 \end{fpfig}
+
+The $\minifed$ language provides a simple model of synchronous
+protocols between a federation of \emph{clients} exchanging values in
+the binary field. We will identify clients by natural numbers, and
+federations- finite sets of clients- are always given statically.
+As we will see, our threat model assumes a partition of the federation
+into \emph{honest} $H$ and \emph{corrupt} $C$ subsets.
+
+We model probabilistic programming via a \emph{random tape}
+semantics. That is, we will assume that programs can make reference to
+values chosen from a uniform random distributions defined in the
+initial program memory.  Programs aka protocols execute
+deterministically given the random tape.
+
+\subsection{Syntax} The syntax of $\minifed$, defined in
+Figure \ref{fig-minifed}, includes values $v$ and standard
+operations of addition, subtraction, and multiplication in
+a finite field $\mathbb{Z}_p$ with $p$ prime. 
+Protocols are given input secret values $\secret{w}$
+as well as random samples $\flip{w}$ on the input
+tape, both of which are distinguished by
+strings $w$. Protocols are sequences of assignment
+commands of three different forms:
+\begin{itemize}
+\item $\eassign{\mesg{w}}{\cid_2}{\be}{\cid_1}$: This
+  is a \emph{message send} where expression $\be$ is computed
+  by client $\cid_1$ and sent to client $\cid_2$ as message
+  $\mesg{w}$.
+\item $\reveal{w}{\be}{\cid}$: This
+  is a \emph{public reveal} where expression $\be$ is computed
+  by client $\cid$ and publicly shared with (broadcast to) the federation.
+\item $\pubout{\cid}{\be}{\cid}$: This
+  is an \emph{output} where expression $\be$ is computed
+  by client $\cid$ and reported as its output.
+\end{itemize}
+Both messages $\mesg{w}$ and reveals $\rvl{w}$ can be
+referenced in expressions, once they've been assigned.
+
+We let $x$ range over variables, and will use $s$, $v$, and $f$ to
+range over individual secrets, views, and flips respectively. We let
+$X$ range over sets of variables and $S,V,F$ to range over sets of
+secrets, views, and flips (including invocations of the oracle)
+respectively. Given a program $\prog$, we write $\iov(\prog)$ to
+denote $S \cup V$ where $S$ is the set of secret variables in $\prog$
+and $V$ is the set of views in $\prog$, we write $\flips(\prog)$ to
+denote the set $F$ of flip variables in $\prog$, and we write
+$\vars(\prog)$ to denote $\iov(\prog) \cup \flips(\prog)$. For any set
+of variables $X$ and parties $P$, we write $X_P$ to denote the subset
+of $X$ owned by any party in $P$, and we write $X_H$ and $X_C$ to
+denote the subsets belonging to honest and corrupt parties,
+respectively.
+
+\subsection{Semantics}
+
+\emph{Memories} are fundamental to the semantics of $\fedcat$ and
+provide the random tape and secret inputs to protocols, and record
+view assignments. Memories $\store$ are finite (partial) mapping from
+variables to binary values $\beta \in \{0,1\}$. The \emph{domain} of a
+memory is written $\dom(\store)$ and is the finite set of variables on
+which the memory is defined. We write $\store\{ x \mapsto \beta\}$ for
+$x\not\in\dom(\store)$ to denote the memory $\store'$ such that
+$\store'(x) = \beta$ and otherwise $\store'(y) = \store(y)$ for all $y
+\in \dom(\store)$. We write $\store_1 \subseteq \store_2$ iff
+$\dom(\store_1) \subseteq \dom(\store_2)$ and $\store_1(x) =
+\store_2(x)$ for all $x \in \dom(\store_1)$. We write $\store_1 \cap
+\store_2$ to denote the combination of $\store_1$ and $\store_2$
+assuming $\store_1(x) = \store_2(x)$ for all $x \in \dom(\store_1)
+\cap \dom(\store_2)$, otherwise $\store_1 \cap \store_2$ is undefined.
+We write $\store_1 \subseteq \store_2$ iff $\store_1 \cap \store_2
+= \store_1$.
+
+Given a set of variables $X$, we write $\store_X$ to denote the
+memory $\store$ restricted to the domain $X$, and we define
+$\mems(X)$ as the set of all memories with domain $X$:
+$$
+\mems(X) \defeq \{ \store \mid \dom(\store) = X \}
+$$
+Thus, given a protocol $\prog$, the set of all random tapes for
+$\prog$ is $\mems(\flips(\prog))$. We let $\stores$ range
+over sets of memories with the same domain, and abusing notation
+we write $\dom(\stores)$ to denote the common domain,
+and $\stores_X \defeq \{ \store_X | \store \in \stores \}$.
+
+Given a variable-free boolean expression $\be$, we write $\cod{\be}$
+to denote the standard interpretation of $\be$ in the binary field.
+With the introduction of variables to expressions, we have two
+concerns. First, we need to interpret variables with respect to a
+specific memory, and second, we need to ensure that variables are used
+``legally''. That is, since expressions define local computation, all
+variables used in an expression must belong to the same client.  Thus,
+we denote interpretation of expressions $\be$ possibly containing
+variables as $\lcod{\store,\be}{\cid}$, where $\store$ associates
+variables with values, and all variables must be owned by client
+$\cid$. This is defined in Figure \ref{fig-minifed}.
+
+$$
+\delta(\otp,v_1,v_2) \defeq v_1 \fplus v_2
+$$
+
+$$
+\begin{array}{lll}
+  \elab{\mesg{s1}}{2} &:=& \elab{(\otp\ \locflip\ (\flip{x} \fplus \secret{1})}{1} \\ 
+  \elab{\mesg{s1}}{3} &:=& \elab{\flip{x}}{1} \\ 
+  \elab{\mesg{s2}}{1} &:=& \elab{(\otp\ \locflip\ (\flip{x} \fplus \secret{2})}{2} \\ 
+  \elab{\mesg{s2}}{3} &:=& \elab{\flip{x}}{2} \\ 
+  \elab{\mesg{s3}}{1} &:=& \elab{(\otp\ \locflip\ (\flip{x} \fplus \secret{3})}{3} \\ 
+  \elab{\mesg{s3}}{2} &:=& \elab{\flip{x}}{3} \\ 
+  \rvl{1} &:=& \elab{(\locflip \fplus \mesg{s2} \fplus \mesg{s3})}{1} \\ 
+  \rvl{2} &:=& \elab{(\mesg{s1} \fplus \locflip \fplus \mesg{s3})}{2} \\
+  \rvl{3} &:=& \elab{(\mesg{s1} \fplus \mesg{s2} \fplus \locflip)}{3} \\ 
+  \out{1} &:=& \elab{(\rvl{1} \fplus \rvl{2} + \rvl{3})}{1}\\
+  \out{2} &:=& \elab{(\rvl{1} \fplus \rvl{2} + \rvl{3})}{2}\\
+  \out{3} &:=& \elab{(\rvl{1} \fplus \rvl{2} + \rvl{3})}{3}
+  %\elab{\mesg{o1}}{2} &:=& \elab{(\locflip \fplus \mesg{s2} \fplus \mesg{s3})}{1} \\ 
+  %\elab{\mesg{o1}}{3} &:=& \elab{(\locflip \fplus \mesg{s2} \fplus \mesg{s3})}{1} \\ 
+  %\elab{\mesg{o2}}{1} &:=& \elab{(\mesg{s1} \fplus \locflip \fplus \mesg{s3})}{2} \\
+  %\elab{\mesg{o2}}{3} &:=& \elab{(\mesg{s1} \fplus \locflip \fplus \mesg{s3})}{2} \\ 
+  %\elab{\mesg{o3}}{1} &:=& \elab{(\mesg{s1} \fplus \mesg{s2} \fplus \locflip)}{3} \\ 
+  %\elab{\mesg{o3}}{2} &:=& \elab{(\mesg{s1} \fplus \mesg{s2} \fplus \locflip)}{3}\\ 
+  %\pubout{1} &:=& \elab{(\locflip \fplus \mesg{s2} \fplus \mesg{s3} + \mesg{o2} + \mesg{o3})}{1}
+\end{array}
+$$
+
+$$
+(v, [m_1,\ldots,m_n])
+$$
+$$
+k_1,\ldots,k_n
+$$
+$$
+m_\cid = k_\cid + (k_\Delta * v)
+$$
+
+$$
+\delta(\macplus,(v^1, [m_1^1,\ldots,m^1_n]),(v^2, [m_1^2,\ldots,m^2_n]))
+\defeq
+(v^1 \fplus v^2, [m_1^1 \fplus m_1^2 ,\ldots,m^1_n \fplus m_n^2])
+$$
+
+$$
+\elab{\macv{\elab{v}{\cid}}}{1} \macplus \cdots \macplus \elab{\macv{\elab{v}{\cid}}}{n} =
+(v,\ldots)
+$$
+
+$$
+\delta(\macotp,v_1,v_2) \defeq v_1 \macplus v_2
+$$
+
+$$
+\delta(\macauth, (v, [\ldots,m_\cid,\ldots]), k_\cid) \defeq
+     (v, [\ldots,m_\cid,\ldots]) \text{\ if\ } m_i = k_i + (k_\Delta * v)
+$$
+
+$$
+\begin{array}{lcl}
+  \elab{\mesg{a}}{2} &:=&
+  \elab{(\macotp\ \macv{\elab{\secret{x}}{1}}\ \macv{\elab{\flip{a}}{\Oracle}})}{1}\\
+  \elab{\mesg{a}}{1} &:=&
+  \elab{(\macotp\ \macv{\elab{\secret{x}}{1}}\ \macv{\elab{\flip{a}}{\Oracle}})}{2}\\
+  \elab{\mesg{b}}{2} &:=&
+  \elab{(\macotp\ \macv{\elab{\secret{y}}{2}}\ \macv{\elab{\flip{b}}{\Oracle}})}{1}\\
+  \elab{\mesg{b}}{1} &:=&
+  \elab{(\macotp\ \macv{\elab{\secret{y}}{2}}\ \macv{\elab{\flip{b}}{\Oracle}})}{2}\\
+  \elab{\mesg{d}}{1} &:=&
+  \elab{(\macauth(\mesg{a}, \mack{\elab{\secret{x}}{1}}{2} \fminus \mack{\elab{\flip{a}}{\Oracle}}{2}) \macplus (\macv{\elab{\secret{x}}{1}}\macminus\macv{\elab{\flip{a}}{\Oracle}}))}{1}\\
+  \elab{\mesg{e}}{1}&:=&
+  \elab{(\macauth(\mesg{b}, \mack{\elab{\secret{y}}{2}}{2} \fminus \mack{\elab{\flip{b}}{\Oracle}}{2}) \macplus (\macv{\elab{\secret{y}}{2}}\macminus\macv{\elab{\flip{b}}{\Oracle}}))}{1}\\
+  \rvl{1} &:=&
+  \elab{( (\mesg{d} \mactimes \mesg{e}) \macplus
+          (\mesg{d} \mactimes \macv{\elab{\flip{b}}{\Oracle}}) \macplus
+          (\mesg{e} \mactimes \macv{\elab{\flip{a}}{\Oracle}}) \macplus \macv{\elab{\secret{c}}{\Oracle}}
+    )}{1}\\
+  \elab{\mesg{d}}{2} &:=&
+  \elab{(\macauth(\mesg{a}, \mack{\elab{\secret{x}}{1}}{1} \fminus \mack{\elab{\flip{a}}{\Oracle}}{1}) \macplus (\macv{\elab{\secret{x}}{1}}\macminus\macv{\elab{\flip{a}}{\Oracle}}))}{2}\\
+  \elab{\mesg{e}}{2}&:=&
+  \elab{(\macauth(\mesg{b}, \mack{\elab{\secret{y}}{2}}{1} \fminus \mack{\elab{\flip{b}}{\Oracle}}{1}) \macplus (\macv{\elab{\secret{y}}{2}}\macminus\macv{\elab{\flip{b}}{\Oracle}}))}{2}\\
+  \rvl{2} &:=&
+  \elab{( (\mesg{d} \mactimes \mesg{e}) \macplus
+          (\mesg{d} \mactimes \macv{\elab{\flip{b}}{\Oracle}}) \macplus
+          (\mesg{e} \mactimes \macv{\elab{\flip{a}}{\Oracle}}) \macplus \macv{\elab{\secret{c}}{\Oracle}}
+    )}{2}\\
+  \out{1} &:=& \elab{(\rvl{1} \macplus \macauth(\rvl{2},\ldots))}{1} \\
+  \out{2} &:=& \elab{(\macauth(\rvl{1},\ldots) \macplus \rvl{2})}{2}
+\end{array}
+$$

### 2024-04-18 revision-1/lang.tex
checkpoint

@@
       \be &::=& v \mid \flip{w} \mid \secret{w} \mid \mesg{w} \mid \rvl{w} \mid \be \fminus \be \mid \be \fplus \be \mid \be \ftimes \be \mid f \mid \be\,\be & \textit{expressions}\\[2mm]
+      x &::=& \elab{\flip{w}}{\cid} \mid \elab{\secret{w}}{\cid} \mid \elab{\mesg{w}}{\cid} \mid \rvl{w} \mid \out{\cid} & \textit{protocl variables} \\[2mm]
       %& &  \select{\be}{\be}{\be} \mid \ctxt{v}{k} \mid \key{w} \mid \sk{\be}(\be) \mid \pk{\be}{\be}(\be) \mid \pk{\be}{\be} \\[2mm]
@@
   is a \emph{public reveal} where expression $\be$ is computed
-  by client $\cid$ and publicly shared with (broadcast to) the federation.
+  by client $\cid$ and broadcast to the federation.
 \item $\pubout{\cid}{\be}{\cid}$: This
@@
 
-We let $x$ range over variables, and will use $s$, $v$, and $f$ to
-range over individual secrets, views, and flips respectively. We let
-$X$ range over sets of variables and $S,V,F$ to range over sets of
+We let $x$ range over \emph{variables} which are
+identifiers 
+\begin{itemize}
+  \item $S$ ranges over sets of 
+\end{itemize}
+We let $x$ range over variables.%, and will use $s$, $v$, and $f$ to
+%range over individual secrets, views, and flips respectively.
+We let $X$ range over sets of variables,
+and $S,R,M,P,O$$ to range over sets of
 secrets, views, and flips (including invocations of the oracle)

### 2024-04-18 revision-1/lang.tex
checkpoint

@@
 
-We let $x$ range over \emph{variables} which are
-identifiers 
-\begin{itemize}
-  \item $S$ ranges over sets of 
-\end{itemize}
-We let $x$ range over variables.%, and will use $s$, $v$, and $f$ to
-%range over individual secrets, views, and flips respectively.
-We let $X$ range over sets of variables,
-and $S,R,M,P,O$$ to range over sets of
-secrets, views, and flips (including invocations of the oracle)
-respectively. Given a program $\prog$, we write $\iov(\prog)$ to
-denote $S \cup V$ where $S$ is the set of secret variables in $\prog$
-and $V$ is the set of views in $\prog$, we write $\flips(\prog)$ to
-denote the set $F$ of flip variables in $\prog$, and we write
+We let $x$ range over \emph{variables}  which are identifiers
+where client ownership is specified- e.g., $\elab{\mesg{\mathit{foo}}}{\cid}$
+is a message $\mathit{foo}$ that was sent to $\cid$. We let $X$
+range over sets of variables, and more specifically, $S$ ranges over sets of secret variables $\elab{\secret{w}}{\cid}$, $R$ ranges over sets of random variables $\elab{\flip{w}}{\cid}$, $M$ ranges over sets of message variables $\elab{\mesg{w}}{\cid}$, $P$ ranges over sets of public variables $\rvl{w}$, and $O$ ranges over sets of output variabels $\out{\cid}$.
+Given a program $\prog$, we write $\iov(\prog)$ to
+denote the set of $S \cup M \cup P \cup O$ of variables in $\prog$
+with ownership made explicit, and we write $\flips(\prog)$ to
+denote the set $R$ of random samplings in $\prog$ with ownership
+made explicit. We write
 $\vars(\prog)$ to denote $\iov(\prog) \cup \flips(\prog)$. For any set
 of variables $X$ and parties $P$, we write $X_P$ to denote the subset
-of $X$ owned by any party in $P$, and we write $X_H$ and $X_C$ to
+of $X$ owned by any party in $P$, in particular we write $X_H$ and $X_C$ to
 denote the subsets belonging to honest and corrupt parties,
@@
 view assignments. Memories $\store$ are finite (partial) mapping from
-variables to binary values $\beta \in \{0,1\}$. The \emph{domain} of a
+variables to values $v \in \mathbb{Z}_p$. The \emph{domain} of a
 memory is written $\dom(\store)$ and is the finite set of variables on
-which the memory is defined. We write $\store\{ x \mapsto \beta\}$ for
+which the memory is defined. We write $\store\{ x \mapsto v\}$ for
 $x\not\in\dom(\store)$ to denote the memory $\store'$ such that
-$\store'(x) = \beta$ and otherwise $\store'(y) = \store(y)$ for all $y
+$\store'(x) = v$ and otherwise $\store'(y) = \store(y)$ for all $y
 \in \dom(\store)$. We write $\store_1 \subseteq \store_2$ iff
@@
 Thus, given a protocol $\prog$, the set of all random tapes for
-$\prog$ is $\mems(\flips(\prog))$. We let $\stores$ range
-over sets of memories with the same domain, and abusing notation
-we write $\dom(\stores)$ to denote the common domain,
-and $\stores_X \defeq \{ \store_X | \store \in \stores \}$.
+$\prog$ is $\mems(\flips(\prog))$.
+%We let $\stores$ range
+%over sets of memories with the same domain, and abusing notation
+%we write $\dom(\stores)$ to denote the common domain,
+%and $\stores_X \defeq \{ \store_X | \store \in \stores \}$.
 
 Given a variable-free boolean expression $\be$, we write $\cod{\be}$
-to denote the standard interpretation of $\be$ in the binary field.
-With the introduction of variables to expressions, we have two
-concerns. First, we need to interpret variables with respect to a
-specific memory, and second, we need to ensure that variables are used
-``legally''. That is, since expressions define local computation, all
-variables used in an expression must belong to the same client.  Thus,
+to denote the standard interpretation of $\be$ in the arithmetic field
+$\mathbb{Z}{p}$.
+With the introduction of variables to expressions, we need to interpret
+variables with respect to a specific memory, and all
+variables used in an expression must belong to a specified client.  Thus,
 we denote interpretation of expressions $\be$ possibly containing
@@
 variables with values, and all variables must be owned by client
-$\cid$. This is defined in Figure \ref{fig-minifed}.
+$\cid$. This is interpretation is defined in Figure \ref{fig-minifed}.
 

### 2024-04-18 revision-1/lang.tex
checkpoint

@@
 
+\subsubsection{Library Functions}
+
 \subsection{Semantics}
@@
 \emph{Memories} are fundamental to the semantics of $\fedcat$ and
-provide the random tape and secret inputs to protocols, and record
-view assignments. Memories $\store$ are finite (partial) mapping from
-variables to values $v \in \mathbb{Z}_p$. The \emph{domain} of a
+provide random tape and secret inputs to protocols, and also record
+message sends, public broadcast, and client outputs. Memories $\store$ are finite
+(partial) mapping from variables $x$ to values $v \in \mathbb{Z}_p$. The \emph{domain} of a
 memory is written $\dom(\store)$ and is the finite set of variables on

### 2024-04-19 revision-1/lang.tex
checkpoint

@@
 
-\subsubsection{Library Functions}
+\subsubsection{Library Functions} $\minifed$ expression syntax also supports
+calls to library functions $f$. This allows encapsulation and separate
+definition of primitive operations such as one-time-pads and message
+authentication, as we will illustrate with examples. This approach is
+useful since it parameterizes these definitions, and 
+supports verification of behavior specified with types, as we
+discuss in Section \ref{section-types}.
 
@@
 
-Given a variable-free boolean expression $\be$, we write $\cod{\be}$
-to denote the standard interpretation of $\be$ in the arithmetic field
-$\mathbb{Z}{p}$.
-With the introduction of variables to expressions, we need to interpret
-variables with respect to a specific memory, and all
-variables used in an expression must belong to a specified client.  Thus,
-we denote interpretation of expressions $\be$ possibly containing
-variables as $\lcod{\store,\be}{\cid}$, where $\store$ associates
-variables with values, and all variables must be owned by client
-$\cid$. This is interpretation is defined in Figure \ref{fig-minifed}.
+Given a variable-free expression $\be$, we write $\cod{\be}$ to denote
+the standard interpretation of $\be$ in the arithmetic field
+$\mathbb{Z}_{p}$. With the introduction of variables to expressions,
+we need to interpret variables with respect to a specific memory, and
+all variables used in an expression must belong to a specified client.
+Thus, we denote interpretation of expressions $\be$ computed on a
+client $\cid$ as $\lcod{\store,\be}{\cid}$. This interpretation is
+defined in Figure \ref{fig-minifed}. It is also parameterized by
+$\delta$ which defines the semantics of library functions $f$.
 
+The small-step reduction relation $\redx$ is then defined in Figure
+\ref{fig-minifed} to evaluate commands. Reduction is a relation on
+\emph{configurations} $(\store, \prog)$ where all three command forms-
+message send, broadcast, and output- are implemented as updates to the
+memory $\store$. We write $\redxs$ to denote the reflexive, transitive
+closure of\ $\redx$.
+
+\subsection{Example: Passive-Secure Addition}
+
+Shamir addition leverages homomorphic properties of addition in
+arithmetic fields to implement secret addition. If a field value $v_1$
+is in a uniform random distribution with other variables in a program,
+then $v_1 \fplus v_2$ is an encryption of $v_2$ where $v_1$ is an
+information theoretically secure one-time-pad, which is exploited for
+secret sharing. Of course, $\fplus$ is also a meaningful operation
+over any two field values regardless of their distributions.
+
+To capture this distinction we introduce a function $\otp$
+with the following specification:
 $$
@@
 $$
+Although the semantics are the same as addition, the use of $\otp$
+makes a declarative distinction, but more importantly we will be
+able to assign a type to $\otp$ that enforces the one-time discipline
+on its first argument via type linearity as will be discussed in Section
+\ref{section-types}.
 

### 2024-04-19 revision-1/lang.tex
checkpoint

@@
 \subsubsection{Library Functions} $\minifed$ expression syntax also supports
-calls to library functions $f$. This allows encapsulation and separate
+calls to library functions $f$ which can be applied to muliple arguments in a
+curried style. This allows encapsulation and separate
 definition of primitive operations such as one-time-pads and message
@@
 
+To sum their secret values $\secret{\cid}$, each client $\cid$ in
+the federation $\{ 1, 2, 3 \}$  samples a value $\locflip$
+that can be used as a one-time pad in summation with another
+random sample $\flip{x}$ and $\secret{\cid}$. This yields
+two secret shares communicated as messages to the other clients,
+while each client keeps $\locflip$ as its own share.
 $$
@@
   \elab{\mesg{s3}}{1} &:=& \elab{(\otp\ \locflip\ (\flip{x} \fplus \secret{3})}{3} \\ 
-  \elab{\mesg{s3}}{2} &:=& \elab{\flip{x}}{3} \\ 
+  \elab{\mesg{s3}}{2} &:=& \elab{\flip{x}}{3}
+\end{array}
+$$
+Due to field properties of $\fplus$ this scheme guarantees that messages
+are viewed as random noise by any adversary \cite{barthe2019probabilistic}.
+$$
+\begin{array}{lll}
   \rvl{1} &:=& \elab{(\locflip \fplus \mesg{s2} \fplus \mesg{s3})}{1} \\ 

### 2024-04-19 revision-1/lang.tex
checkpoint

@@
 $\store'(x) = v$ and otherwise $\store'(y) = \store(y)$ for all $y
-\in \dom(\store)$. We write $\store_1 \subseteq \store_2$ iff
-$\dom(\store_1) \subseteq \dom(\store_2)$ and $\store_1(x) =
-\store_2(x)$ for all $x \in \dom(\store_1)$. We write $\store_1 \cap
-\store_2$ to denote the combination of $\store_1$ and $\store_2$
-assuming $\store_1(x) = \store_2(x)$ for all $x \in \dom(\store_1)
-\cap \dom(\store_2)$, otherwise $\store_1 \cap \store_2$ is undefined.
-We write $\store_1 \subseteq \store_2$ iff $\store_1 \cap \store_2
-= \store_1$.
+\in \dom(\store)$. We write $\store \subseteq \store'$ iff
+$\dom(\store) \subseteq \dom(\store')$ and $\store(x) =
+\store'(x)$ for all $x \in \dom(\store)$. We write $\store \cap
+\store'$ to denote the combination of $\store$ and $\store'$
+assuming $\store(x) = \store'(x)$ for all $x \in \dom(\store)
+\cap \dom(\store')$, otherwise $\store \cap \store'$ is undefined.
+We write $\store \subseteq \store'$ iff $\store \cap \store'
+= \store$.
 
@@
 Due to field properties of $\fplus$ this scheme guarantees that messages
-are viewed as random noise by any adversary \cite{barthe2019probabilistic}.
+are viewed as random noise by any observer 
+besides $\cid$ \cite{barthe2019probabilistic}. Next, each client
+publicly reveals the sum of all of its shares, including its local
+share. This does reveal information about secrets. Further there
+is no one-time-pad to use in this summation.
 $$
@@
   \rvl{2} &:=& \elab{(\mesg{s1} \fplus \locflip \fplus \mesg{s3})}{2} \\
-  \rvl{3} &:=& \elab{(\mesg{s1} \fplus \mesg{s2} \fplus \locflip)}{3} \\ 
-  \out{1} &:=& \elab{(\rvl{1} \fplus \rvl{2} + \rvl{3})}{1}\\
-  \out{2} &:=& \elab{(\rvl{1} \fplus \rvl{2} + \rvl{3})}{2}\\
-  \out{3} &:=& \elab{(\rvl{1} \fplus \rvl{2} + \rvl{3})}{3}
+  \rvl{3} &:=& \elab{(\mesg{s1} \fplus \mesg{s2} \fplus \locflip)}{3} 
+\end{array}
+$$
+Finally, each client outputs the sum of each sum of shares, yielding
+the sum of secrets. Note that this stage exposes no more information
+than the previous public reveals. 
+$$
 %\elab{\mesg{o1}}{2} &:=& \elab{(\locflip \fplus \mesg{s2} \fplus \mesg{s3})}{1} \\ 
@@
   %\pubout{1} &:=& \elab{(\locflip \fplus \mesg{s2} \fplus \mesg{s3} + \mesg{o2} + \mesg{o3})}{1}
+\begin{array}{lll}
+  \out{1} &:=& \elab{(\rvl{1} \fplus \rvl{2} + \rvl{3})}{1}\\
+  \out{2} &:=& \elab{(\rvl{1} \fplus \rvl{2} + \rvl{3})}{2}\\
+  \out{3} &:=& \elab{(\rvl{1} \fplus \rvl{2} + \rvl{3})}{3}
 \end{array}
 $$
+It is well-known that additive secret sharing is passive
+secure. That is, any adversarial observer can gain no more information
+from the messages exchanged in the protocol than what is exposed by
+the output alone. However, malicious adversaries can corrupt this
+protocol by injecting ``fake'' sums of shares in their public reveals.
+
+\subsection{Example: Malicious Secure Product}
 

### 2024-04-19 revision-1/lang.tex
checkpoint

@@
 
-\subsection{Example: Passive-Secure Addition}
+\subsection{Example: Passive Secure Addition}
 
@@
 Although the semantics are the same as addition, the use of $\otp$
-makes a declarative distinction, but more importantly we will be
+makes a declarative distinction. But more importantly we will be
 able to assign a type to $\otp$ that enforces the one-time discipline
-on its first argument via type linearity as will be discussed in Section
+on its first argument via type linearity, and reflects hiding of
+its second argument via security type labels, as will be discussed in Section
 \ref{section-types}.
@@
 
+\begin{fpfig}[t]{Top-to-bottom: Extended $\minifed$ syntax with BDOZ-style shares,
+    interpretation of operations, and pre-processing properties of memories.}{fig-bdoz}
+  {
 $$
-(v, [m_1,\ldots,m_n])
+\begin{array}{rcl@{\hspace{8mm}}r}
+  \multicolumn{3}{l}{m,k \in \mathbb{Z}_p \qquad \macv ::= (v,[m_1,\ldots,m_n])} &\\[2mm]
+  %\textit{MACed\ values}\\[2mm]
+  \be &::=& \cdots \mid \macv \mid \be \macplus \be \mid \be \mactimes \be \mid \be \macminus \be\\[2mm]
+  x &::=& \cdots \mid \macx{\secret{w}}{\cid} \mid \macx{\flip{w}}{\cid} \mid \mack{x}{\cid}
+\end{array}
+$$
+
+\rule{130mm}{0.5pt}
+
+$$
+\begin{array}{rcl}
+  \fcod{(v^1, [m_1^1,\ldots,m^1_n]) \macplus (v^2, [m_1^2,\ldots,m^2_n])} &=&
+  (v^1 \fplus v^2, [m_1^1 \fplus m_1^2 ,\ldots,m^1_n \fplus m_n^2])\\[2mm]
+  \fcod{(v^1, [m_1^1,\ldots,m^1_n]) \mactimes v} &=&
+   (v^1 \ftimes v, [m_1^1 \ftimes v ,\ldots,m^1_n \ftimes v])
+\end{array}
 $$
 $$
-k_1,\ldots,k_n
+\begin{array}{rcl}
+  \lcod{\store, \be_1 \macplus \be_2}{\cid} &=&
+  \fcod{\lcod{\store,\be_1}{\cid} \macplus \lcod{\store,\be_2}{\cid}}\\[2mm]  
+  \lcod{\store, \be_1 \mactimes \be_2}{\cid} &=&
+  \fcod{\lcod{\store,\be_1}{\cid} \mactimes \lcod{\store,\be_2}{\cid}}
+\end{array}
 $$
+
+\rule{130mm}{0.5pt}
+
+\begin{mathpar}
+  \inferrule
+      {\store(x) = v \\ \store(\macx{x}{\cid_1}) = (v_1,\ldots)\  \cdots\ 
+        \store(\macx{x}{\cid_n}) = (v_n,\ldots)}
+      {v = \fcod{v_1 \fplus \cdots \fplus v_n}}
+
+  \inferrule
+      {\store(\macx{x}{\cid'}) = (v,[m_1,\ldots,m_n]) \\  \cid \in [1..n]}
+      {m_\cid = \fcod{\store(\mack{x}{\cid}) \fplus (\store(\elab{\flip{\ttt{delta}}}{\cid'}) \ftimes v)}}
+
+  \store(\elab{\secret{c}}{\Oracle}) = \lcod{\store,\flip{a} \ftimes \flip{b}}{\Oracle}
+\end{mathpar}
+}
+\end{fpfig}
+
 $$
 m_\cid = k_\cid + (k_\Delta * v)
-$$
-
-$$
-\delta(\macplus,(v^1, [m_1^1,\ldots,m^1_n]),(v^2, [m_1^2,\ldots,m^2_n]))
-\defeq
-(v^1 \fplus v^2, [m_1^1 \fplus m_1^2 ,\ldots,m^1_n \fplus m_n^2])
-$$
-
-$$
-\elab{\macv{\elab{v}{\cid}}}{1} \macplus \cdots \macplus \elab{\macv{\elab{v}{\cid}}}{n} =
-(v,\ldots)
 $$
@@
 \delta(\macauth, (v, [\ldots,m_\cid,\ldots]), k_\cid) \defeq
-     (v, [\ldots,m_\cid,\ldots]) \text{\ if\ } m_i = k_i + (k_\Delta * v)
+     (v, [\ldots,m_\cid,\ldots]) \text{\ if\ } m_\cid = k_\cid + (k_\Delta * v)
 $$
@@
   \elab{\mesg{a}}{2} &:=&
-  \elab{(\macotp\ \macv{\elab{\secret{x}}{1}}\ \macv{\elab{\flip{a}}{\Oracle}})}{1}\\
+  \elab{(\macotp\ \macgv{\elab{\secret{x}}{1}}\ \macgv{\elab{\flip{a}}{\Oracle}})}{1}\\
   \elab{\mesg{a}}{1} &:=&
-  \elab{(\macotp\ \macv{\elab{\secret{x}}{1}}\ \macv{\elab{\flip{a}}{\Oracle}})}{2}\\
+  \elab{(\macotp\ \macgv{\elab{\secret{x}}{1}}\ \macgv{\elab{\flip{a}}{\Oracle}})}{2}\\
   \elab{\mesg{b}}{2} &:=&
-  \elab{(\macotp\ \macv{\elab{\secret{y}}{2}}\ \macv{\elab{\flip{b}}{\Oracle}})}{1}\\
+  \elab{(\macotp\ \macgv{\elab{\secret{y}}{2}}\ \macgv{\elab{\flip{b}}{\Oracle}})}{1}\\
   \elab{\mesg{b}}{1} &:=&
-  \elab{(\macotp\ \macv{\elab{\secret{y}}{2}}\ \macv{\elab{\flip{b}}{\Oracle}})}{2}\\
+  \elab{(\macotp\ \macgv{\elab{\secret{y}}{2}}\ \macgv{\elab{\flip{b}}{\Oracle}})}{2}\\
   \elab{\mesg{d}}{1} &:=&
-  \elab{(\macauth(\mesg{a}, \mack{\elab{\secret{x}}{1}}{2} \fminus \mack{\elab{\flip{a}}{\Oracle}}{2}) \macplus (\macv{\elab{\secret{x}}{1}}\macminus\macv{\elab{\flip{a}}{\Oracle}}))}{1}\\
+  \elab{(\macauth(\mesg{a}, \mack{\macx{\secret{x}}{1}}{2} \fminus \mack{\macx{\flip{a}}{\Oracle}}{2}) \macplus (\macgv{\elab{\secret{x}}{1}}\macminus\macgv{\elab{\flip{a}}{\Oracle}}))}{1}\\
   \elab{\mesg{e}}{1}&:=&
-  \elab{(\macauth(\mesg{b}, \mack{\elab{\secret{y}}{2}}{2} \fminus \mack{\elab{\flip{b}}{\Oracle}}{2}) \macplus (\macv{\elab{\secret{y}}{2}}\macminus\macv{\elab{\flip{b}}{\Oracle}}))}{1}\\
+  \elab{(\macauth(\mesg{b}, \mack{\macx{\secret{y}}{2}}{2} \fminus \mack{\macx{\flip{b}}{\Oracle}}{2}) \macplus (\macgv{\elab{\secret{y}}{2}}\macminus\macgv{\elab{\flip{b}}{\Oracle}}))}{1}\\
   \rvl{1} &:=&
   \elab{( (\mesg{d} \mactimes \mesg{e}) \macplus
-          (\mesg{d} \mactimes \macv{\elab{\flip{b}}{\Oracle}}) \macplus
-          (\mesg{e} \mactimes \macv{\elab{\flip{a}}{\Oracle}}) \macplus \macv{\elab{\secret{c}}{\Oracle}}
+          (\mesg{d} \mactimes \macgv{\elab{\flip{b}}{\Oracle}}) \macplus
+          (\mesg{e} \mactimes \macgv{\elab{\flip{a}}{\Oracle}}) \macplus \macgv{\elab{\secret{c}}{\Oracle}}
     )}{1}\\
   \elab{\mesg{d}}{2} &:=&
-  \elab{(\macauth(\mesg{a}, \mack{\elab{\secret{x}}{1}}{1} \fminus \mack{\elab{\flip{a}}{\Oracle}}{1}) \macplus (\macv{\elab{\secret{x}}{1}}\macminus\macv{\elab{\flip{a}}{\Oracle}}))}{2}\\
+  \elab{(\macauth(\mesg{a}, \mack{\macx{\secret{x}}{1}}{1} \fminus \mack{\macx{\flip{a}}{\Oracle}}{1}) \macplus (\macgv{\elab{\secret{x}}{1}}\macminus\macgv{\elab{\flip{a}}{\Oracle}}))}{2}\\
   \elab{\mesg{e}}{2}&:=&
-  \elab{(\macauth(\mesg{b}, \mack{\elab{\secret{y}}{2}}{1} \fminus \mack{\elab{\flip{b}}{\Oracle}}{1}) \macplus (\macv{\elab{\secret{y}}{2}}\macminus\macv{\elab{\flip{b}}{\Oracle}}))}{2}\\
+  \elab{(\macauth(\mesg{b}, \mack{\macx{\secret{y}}{2}}{1} \fminus \mack{\macx{\flip{b}}{\Oracle}}{1}) \macplus (\macgv{\elab{\secret{y}}{2}}\macminus\macgv{\elab{\flip{b}}{\Oracle}}))}{2}\\
   \rvl{2} &:=&
   \elab{( (\mesg{d} \mactimes \mesg{e}) \macplus
-          (\mesg{d} \mactimes \macv{\elab{\flip{b}}{\Oracle}}) \macplus
-          (\mesg{e} \mactimes \macv{\elab{\flip{a}}{\Oracle}}) \macplus \macv{\elab{\secret{c}}{\Oracle}}
+          (\mesg{d} \mactimes \macgv{\elab{\flip{b}}{\Oracle}}) \macplus
+          (\mesg{e} \mactimes \macgv{\elab{\flip{a}}{\Oracle}}) \macplus \macgv{\elab{\secret{c}}{\Oracle}}
     )}{2}\\

### 2024-04-19 revision-1/lang.tex
checkpoint

@@
 
+Eliminating cheating by adding information theoretically secure MAC
+codes to shares is a fundamental approach accomplished by protocols
+including BDOZ \cite{XXX} and SPDZ \cite{XXX}.  These protocols assume
+pre-processing that distributes shares with MAC codes to all clients.
+This integrates well with pre-processed Beaver Triples to implement
+malicious secure and relatively efficient multiplication \cite{XXX}.
+Here we show how to implement malicious secure two party multiplication
+in the BDOZ style via some extensions of $\minifed$.
+
 $$

### 2024-04-20 revision-1/lang.tex
checkpoint

@@
 $$
-\delta(\otp,v_1,v_2) \defeq v_1 \fplus v_2
+\delta(\otp,v_1,v_2) \defeq \fcod{v_1 \fplus v_2}
 $$
@@
 
+A field value $v$ is secret shared among $n$ clientz in BDOZ as follows.
+Each client $\cid$ in a federation of size $n$ gets a value of the form
+$(v_\cid,[m_1,\ldots,m_n])$, where $v_\cid$ is a secret
+share of $v$, i.e., $v = \fcod{v_1 \fplus \cdots \fplus v_n}$. Further, 
+values $[m_1,\ldots,m_n]$ are MACs of $v_\cid$, such that
+$m_{\cid'} = k_{\cid'} + (k_\Delta * v_\cid)$ for each $\cid'\in [1..n]$,
+where keys $k_{\cid'}$ and $k_\Delta$ are known only to $\cid$
+and $k_\Delta$ is common to all MACs used by $\cid$. By
+extending certain field operations to these forms, denoted $\macv$,
+a form of semi-homorphic encryption is obtained that is adequate to compute
+Beaver Triples. For example, if keys $k_1$ and $k_2$ authenticate
+$\macv_1$ and $\macv_2$, then $k_1 \fplus k_2$ authenticates
+$\macv_1 \macplus \macv_2$,  where $\macplus$ is addition
+extended to MACed values. Relevant semantic rules for this, as
+well as multiplication of MACed and field values, are given in
+Figure \ref{fig-bdoz}- for more details the reader is referred to
+\cite{XXX}.
+
+We define variable forms $\macx{\secret{w}}{\cid}$ and
+$\macx{\flip{w}}{\cid}$ range over local MACed shares of a secret and
+random sample owned by client $\cid$, respectively. We let
+$\mack{\macx{x}{\cid_1}}{\cid_2}$ denote the key that
+authenticates $\cid_n$'s share of $\cid_1$'s value $x$.
+We also assume a key $\flip{\ttt{delta}}$ on each client
+$\cid$, which is that client's $k_\Delta$, 
+In Figure \ref{fig-bdoz} we define the relevant properties
+of pre-processed stores.
+
+Similarly to field values, we can use MACed shares in
+uniform independent distributions as one-time-pads for
+hiding other values, using subtraction over MACed shares.
+So we introduce a function $\macotp$ to do this
+declaratively.
 $$
-m_\cid = k_\cid + (k_\Delta * v)
+\delta(\macotp,v_1,v_2) \defeq \fcod{v_1 \macminus v_2}
 $$
+Also, we can define a $\macauth$ function that authenticates
+a client $\cid$'s share, and an $\macopen$ function
+that extracts a field value from its shares.
+\begin{mathpar}
+\delta(\macauth, (v, [\ldots,m_\cid,\ldots]), k_\cid) \defeq
+     (v, [\ldots,m_\cid,\ldots]) \text{\ if\ } m_\cid = \fcod{k_\cid \fplus (k_\Delta \ftimes v)}
 
-$$
-\delta(\macotp,v_1,v_2) \defeq v_1 \macplus v_2
-$$
+\delta(\macopen, (v_1, [\ldots]), \ldots, (v_n, [\ldots]),) \defeq
+\fcod{v_1 \fplus \cdots \fplus v_n}
+\end{mathpar}
+This authentication eliminates cheating and guarantees protocol
+integrity. As we will show in Section \ref{section-types}, this
+can be reflected in a security type for $\macauth$.
 
-$$
-\delta(\macauth, (v, [\ldots,m_\cid,\ldots]), k_\cid) \defeq
-     (v, [\ldots,m_\cid,\ldots]) \text{\ if\ } m_\cid = k_\cid + (k_\Delta * v)
-$$
-
+Putting these pieces together, we can implement 2-party multiplication with
+Beaver Triples as follows. Recall that Beaver Triples are values
+$(a,b,c)$ such that $c = \fcod{b \ftimes a}$- we implement them here
+as a secret $\secret{c}$ and samples $\flip{a}$ and $\flip{b}$ owned
+by an \emph{oracle} $\Oracle$ that we always assume is honest but is
+otherwise not involved in the protocol. We define pre-processing
+properties of Beaver Triples in Figure \ref{figure-bdoz}. First,
+clients 1 and 2 compute (aka \emph{open}) $\secret{x} \fplus \flip{a}$ and
+$\secret{y} \fplus \flip{b}$ where
+$\secret{x}$ and $\secret{y}$ are 1 and 2's secrets respectively.
+Note that both perform a \emph{secure opening} by authenticating
+each others' shares and record $\secret{x} \fplus \flip{a}$ and
+$\secret{y} \fplus \flip{b}$ as $\mesg{d}$ and $\mesg{e}$ respectively.
 $$
@@
   \elab{\mesg{a}}{2} &:=&
-  \elab{(\macotp\ \macgv{\elab{\secret{x}}{1}}\ \macgv{\elab{\flip{a}}{\Oracle}})}{1}\\
+  \elab{(\macotp\ \macgv{\elab{\flip{a}}{\Oracle}}\ \macgv{\elab{\secret{x}}{1}})}{1}\\
   \elab{\mesg{a}}{1} &:=&
-  \elab{(\macotp\ \macgv{\elab{\secret{x}}{1}}\ \macgv{\elab{\flip{a}}{\Oracle}})}{2}\\
+  \elab{(\macotp\ \macgv{\elab{\flip{a}}{\Oracle}}\ \macgv{\elab{\secret{x}}{1}})}{2}\\
   \elab{\mesg{b}}{2} &:=&
-  \elab{(\macotp\ \macgv{\elab{\secret{y}}{2}}\ \macgv{\elab{\flip{b}}{\Oracle}})}{1}\\
+  \elab{(\macotp\ \macgv{\elab{\flip{b}}{\Oracle}}\ \macgv{\elab{\secret{y}}{2}})}{1}\\
   \elab{\mesg{b}}{1} &:=&
-  \elab{(\macotp\ \macgv{\elab{\secret{y}}{2}}\ \macgv{\elab{\flip{b}}{\Oracle}})}{2}\\
+  \elab{(\macotp\ \macgv{\elab{\flip{b}}{\Oracle}}\ \macgv{\elab{\secret{y}}{2}})}{2}\\
   \elab{\mesg{d}}{1} &:=&
-  \elab{(\macauth(\mesg{a}, \mack{\macx{\secret{x}}{1}}{2} \fminus \mack{\macx{\flip{a}}{\Oracle}}{2}) \macplus (\macgv{\elab{\secret{x}}{1}}\macminus\macgv{\elab{\flip{a}}{\Oracle}}))}{1}\\
+  \elab{(\macopen\ \macauth(\mesg{a}, \mack{\macx{\secret{x}}{1}}{2} \fminus \mack{\macx{\flip{a}}{\Oracle}}{2}) \ (\macgv{\elab{\secret{x}}{1}}\macminus\macgv{\elab{\flip{a}}{\Oracle}}))}{1}\\
   \elab{\mesg{e}}{1}&:=&
-  \elab{(\macauth(\mesg{b}, \mack{\macx{\secret{y}}{2}}{2} \fminus \mack{\macx{\flip{b}}{\Oracle}}{2}) \macplus (\macgv{\elab{\secret{y}}{2}}\macminus\macgv{\elab{\flip{b}}{\Oracle}}))}{1}\\
+  \elab{(\macopen\ \macauth(\mesg{b}, \mack{\macx{\secret{y}}{2}}{2} \fminus \mack{\macx{\flip{b}}{\Oracle}}{2}) \ (\macgv{\elab{\secret{y}}{2}}\macminus\macgv{\elab{\flip{b}}{\Oracle}}))}{1}\\
+  \elab{\mesg{d}}{2} &:=&
+  \elab{(\macopen\ \macauth(\mesg{a}, \mack{\macx{\secret{x}}{1}}{1} \fminus \mack{\macx{\flip{a}}{\Oracle}}{1}) \ (\macgv{\elab{\secret{x}}{1}}\macminus\macgv{\elab{\flip{a}}{\Oracle}}))}{2}\\
+  \elab{\mesg{e}}{2}&:=&
+  \elab{(\macopen\ \macauth(\mesg{b}, \mack{\macx{\secret{y}}{2}}{1} \fminus \mack{\macx{\flip{b}}{\Oracle}}{1}) \ (\macgv{\elab{\secret{y}}{2}}\macminus\macgv{\elab{\flip{b}}{\Oracle}}))}{2}
+\end{array}
+$$
+Now, each client uses their share of $\secret{c}$ to generate their
+share of the final output, which is again reported via a secure opening.
+Homomorphic key calculations are elided for brevity.
+$$
+\begin{array}{lcl}
   \rvl{1} &:=&
-  \elab{( (\mesg{d} \mactimes \mesg{e}) \macplus
+  \elab{( (\mesg{d} \ftimes \mesg{e}) \macplus
           (\mesg{d} \mactimes \macgv{\elab{\flip{b}}{\Oracle}}) \macplus
@@
     )}{1}\\
-  \elab{\mesg{d}}{2} &:=&
-  \elab{(\macauth(\mesg{a}, \mack{\macx{\secret{x}}{1}}{1} \fminus \mack{\macx{\flip{a}}{\Oracle}}{1}) \macplus (\macgv{\elab{\secret{x}}{1}}\macminus\macgv{\elab{\flip{a}}{\Oracle}}))}{2}\\
-  \elab{\mesg{e}}{2}&:=&
-  \elab{(\macauth(\mesg{b}, \mack{\macx{\secret{y}}{2}}{1} \fminus \mack{\macx{\flip{b}}{\Oracle}}{1}) \macplus (\macgv{\elab{\secret{y}}{2}}\macminus\macgv{\elab{\flip{b}}{\Oracle}}))}{2}\\
   \rvl{2} &:=&
-  \elab{( (\mesg{d} \mactimes \mesg{e}) \macplus
+  \elab{( (\mesg{d} \ftimes \mesg{e}) \macplus
           (\mesg{d} \mactimes \macgv{\elab{\flip{b}}{\Oracle}}) \macplus
@@
     )}{2}\\
-  \out{1} &:=& \elab{(\rvl{1} \macplus \macauth(\rvl{2},\ldots))}{1} \\
-  \out{2} &:=& \elab{(\macauth(\rvl{1},\ldots) \macplus \rvl{2})}{2}
+  \out{1} &:=& \elab{(\macopen\ \rvl{1} \ \macauth(\rvl{2},\ldots))}{1} \\
+  \out{2} &:=& \elab{(\macopen\ \macauth(\rvl{1},\ldots) \ \macauth(\rvl{2},\ldots))}{2} 
 \end{array}
 $$
+Some information about input secrets is revealed by the product, but
+secure opening ensures high integrity and hence robustness to
+corruption, aka robust declassification.

### 2024-04-20 revision-1/lang.tex
checkpoint

@@
     \lcod{\store, \rvl{w}}{\cid} &=& \store(\rvl{w})\\
-    \lcod{\store, f\,e_1\,\cdots\, e_n}{\cid} &=& \delta(f,\lcod{\store, e_1}{\cid},\ldots,\lcod{\store,e_n}{\cid})
+    \lcod{\store, f\,\be_1\,\cdots\, \be_n}{\cid} &=& \delta(f,\lcod{\store, \be_1}{\cid},\ldots,\lcod{\store,\be_n}{\cid})
   \end{array}
@@
 \begin{array}{rcl@{\hspace{8mm}}r}
-  \multicolumn{3}{l}{m,k \in \mathbb{Z}_p \qquad \macv ::= (v,[m_1,\ldots,m_n])} &\\[2mm]
+  \multicolumn{3}{l}{m,k \in \mathbb{Z}_p \qquad \macv ::= (v,m)} &\\[2mm]
   %\textit{MACed\ values}\\[2mm]
   \be &::=& \cdots \mid \macv \mid \be \macplus \be \mid \be \mactimes \be \mid \be \macminus \be\\[2mm]
-  x &::=& \cdots \mid \macx{\secret{w}}{\cid} \mid \macx{\flip{w}}{\cid} \mid \mack{x}{\cid}
+  x &::=& \cdots \mid \macx{\secret{w}}{\cid} \mid \macx{\flip{w}}{\cid} \mid \mack{x}{}
 \end{array}
@@
 \rule{130mm}{0.5pt}
-
+\smallskip
 $$
 \begin{array}{rcl}
-  \fcod{(v^1, [m_1^1,\ldots,m^1_n]) \macplus (v^2, [m_1^2,\ldots,m^2_n])} &=&
-  (v^1 \fplus v^2, [m_1^1 \fplus m_1^2 ,\ldots,m^1_n \fplus m_n^2])\\[2mm]
-  \fcod{(v^1, [m_1^1,\ldots,m^1_n]) \mactimes v} &=&
-   (v^1 \ftimes v, [m_1^1 \ftimes v ,\ldots,m^1_n \ftimes v])
+  \fcod{(v_1, m_1) \macplus (v_2, m_2)} &=& (v_1 \fplus v_2, m_1 \fplus m_2)\\[2mm]
+  \fcod{(v, m) \mactimes v'} &=& (v \ftimes v', m \ftimes v')
 \end{array}
@@
 
+\medskip
+
 \rule{130mm}{0.5pt}
@@
   \inferrule
-      {\store(x) = v \\ \store(\macx{x}{\cid_1}) = (v_1,\ldots)\  \cdots\ 
-        \store(\macx{x}{\cid_n}) = (v_n,\ldots)}
-      {v = \fcod{v_1 \fplus \cdots \fplus v_n}}
+      {\store(\macx{x}{\cid_1}) = (v_1,m_1) \\
+        \store(\macx{x}{\cid_2}) = (v_2,m_2)}
+      {\store(x) = \fcod{v_1 \fplus v_2}}
 
   \inferrule
-      {\store(\macx{x}{\cid'}) = (v,[m_1,\ldots,m_n]) \\  \cid \in [1..n]}
-      {m_\cid = \fcod{\store(\mack{x}{\cid}) \fplus (\store(\elab{\flip{\ttt{delta}}}{\cid'}) \ftimes v)}}
+      {\store(\elab{x}{\cid'}) = (v,m) \\ \cid \ne \cid'}
+      {m = \lcod{\store, \mack{x}{} \fplus \flip{\ttt{delta}} \ftimes v}{\cid}}
 
@@
 
-Eliminating cheating by adding information theoretically secure MAC
-codes to shares is a fundamental approach accomplished by protocols
-including BDOZ \cite{XXX} and SPDZ \cite{XXX}.  These protocols assume
-pre-processing that distributes shares with MAC codes to all clients.
-This integrates well with pre-processed Beaver Triples to implement
-malicious secure and relatively efficient multiplication \cite{XXX}.
-Here we show how to implement malicious secure two party multiplication
-in the BDOZ style via some extensions of $\minifed$.
+Eliminating cheating by adding information theoretic secure MAC codes
+to shares is a fundamental approach realized by protocols such as BDOZ
+\cite{XXX} and SPDZ \cite{XXX}.  These protocols assume a
+pre-processing phase that distributes shares with MAC codes to
+clients.  This integrates well with pre-processed Beaver Triples to
+implement malicious secure, and relatively efficient, multiplication
+\cite{XXX}.  Here we show how to implement malicious secure two party
+multiplication in the BDOZ style via some extensions of $\minifed$.
 
-A field value $v$ is secret shared among $n$ clientz in BDOZ as follows.
-Each client $\cid$ in a federation of size $n$ gets a value of the form
-$(v_\cid,[m_1,\ldots,m_n])$, where $v_\cid$ is a secret
-share of $v$, i.e., $v = \fcod{v_1 \fplus \cdots \fplus v_n}$. Further, 
-values $[m_1,\ldots,m_n]$ are MACs of $v_\cid$, such that
-$m_{\cid'} = k_{\cid'} + (k_\Delta * v_\cid)$ for each $\cid'\in [1..n]$,
-where keys $k_{\cid'}$ and $k_\Delta$ are known only to $\cid$
-and $k_\Delta$ is common to all MACs used by $\cid$. By
-extending certain field operations to these forms, denoted $\macv$,
-a form of semi-homorphic encryption is obtained that is adequate to compute
-Beaver Triples. For example, if keys $k_1$ and $k_2$ authenticate
-$\macv_1$ and $\macv_2$, then $k_1 \fplus k_2$ authenticates
-$\macv_1 \macplus \macv_2$,  where $\macplus$ is addition
-extended to MACed values. Relevant semantic rules for this, as
-well as multiplication of MACed and field values, are given in
-Figure \ref{fig-bdoz}- for more details the reader is referred to
-\cite{XXX}.
+A field value $v$ is secret shared among 2 clientz in BDOZ as follows.
+Each client $\cid$ gets a pair of the form $(v_\cid,m_\cid)$, where
+$v_\cid$ is a secret share of $v$, i.e., $v = \fcod{v_1 \fplus v_2}$.
+Further, field values $m_\cid$ are MACs of $v_\cid$ that are authenticated
+by the other client's key. In particular, $m_\cid = k + (k_\Delta *
+v_\cid)$ where \emph{keys} $k$ and $k_\Delta$ are known only to $\cid'
+\ne \cid$ and $k_\Delta$. Further, there is a unique \emph{local key}
+$k$ for each shared value, while the \emph{global key} $k_\Delta$ is
+common to all MACs authenticated by $\cid'$. By extending certain
+field operations to these MACed values, a semi-homorphic encryption
+scheme is obtained that is adequate to compute Beaver Triples. For
+example, if local keys $k_1$ and $k_2$ authenticate $\macv_1$ and
+$\macv_2$, then $k_1 \fplus k_2$ authenticates $\macv_1 \macplus
+\macv_2$, where $\macplus$ is addition extended to MACed
+values. Semantics for summation of MACed shares, and multipication of
+shares and plain field values, are given in Figure \ref{fig-bdoz}.
+For more details the reader is referred to \cite{XXX}. We note
+that while we restrict values $v$, $m$, and $k$ to the same
+field in this presentation for simplicity, in general $m$ and
+$k$ can be in extensions of $\mathbb{Z}_p$.
 
 We define variable forms $\macx{\secret{w}}{\cid}$ and
-$\macx{\flip{w}}{\cid}$ range over local MACed shares of a secret and
-random sample owned by client $\cid$, respectively. We let
-$\mack{\macx{x}{\cid_1}}{\cid_2}$ denote the key that
-authenticates $\cid_n$'s share of $\cid_1$'s value $x$.
-We also assume a key $\flip{\ttt{delta}}$ on each client
-$\cid$, which is that client's $k_\Delta$, 
-In Figure \ref{fig-bdoz} we define the relevant properties
-of pre-processed stores.
+$\macx{\flip{w}}{\cid}$ to range over local MACed shares of a secret
+and random sample owned by client $\cid$, respectively. We let
+$\mack{x}{}$ denote the local key for authenticating shares of $x$.
+We also assume a global key $\flip{\ttt{delta}}$ on each client.
+Pre-processing is modeled by assuming that shares are provided on
+input random tapes (memories) with the required properties, as defined
+on the bottom of Figure \ref{fig-bdoz}.
 
@@
 $$
-Also, we can define a $\macauth$ function that authenticates
-a client $\cid$'s share, and an $\macopen$ function
-that extracts a field value from its shares.
+
+Another key idea in BDOZ is \emph{secure opening}, where distributed
+shares of a value $v$ are communicated and combined to ``open'' the
+same $v$ on each client. This relies on authentication to detect
+cheating in the malicious setting. To implement secure opening we
+define a $\macauth$ function that authenticates shares, and a
+$\macopen$ function that extracts a field value from its shares.
 \begin{mathpar}
-\delta(\macauth, (v, [\ldots,m_\cid,\ldots]), k_\cid) \defeq
-     (v, [\ldots,m_\cid,\ldots]) \text{\ if\ } m_\cid = \fcod{k_\cid \fplus (k_\Delta \ftimes v)}
+\delta(\macauth, (v, m), k) \defeq
+     (v, m) \text{\ if\ } m = k \fplus (\flip{\ttt{delta}} \ftimes v)
 
-\delta(\macopen, (v_1, [\ldots]), \ldots, (v_n, [\ldots]),) \defeq
-\fcod{v_1 \fplus \cdots \fplus v_n}
+\delta(\macopen, (v_1, m_1), (v_2,m_2)) \defeq
+\fcod{v_1 \fplus v_2}
 \end{mathpar}
@@
 otherwise not involved in the protocol. We define pre-processing
-properties of Beaver Triples in Figure \ref{figure-bdoz}. First,
-clients 1 and 2 compute (aka \emph{open}) $\secret{x} \fplus \flip{a}$ and
-$\secret{y} \fplus \flip{b}$ where
-$\secret{x}$ and $\secret{y}$ are 1 and 2's secrets respectively.
-Note that both perform a \emph{secure opening} by authenticating
-each others' shares and record $\secret{x} \fplus \flip{a}$ and
-$\secret{y} \fplus \flip{b}$ as $\mesg{d}$ and $\mesg{e}$ respectively.
+properties of Beaver Triples in Figure \ref{fig-bdoz}. Only one
+Beaver triple can be used per multiplication to ensure secure
+openings, but our scheme can be extended with more Beaver triples
+and reuse of them will be rule out by type linearity as we will
+discuss in Section \ref{section-types}.
+
+First, clients 1 and 2 compute (aka \emph{open}) $\secret{x} \fplus
+\flip{a}$ and $\secret{y} \fplus \flip{b}$ where $\secret{x}$ and
+$\secret{y}$ are 1 and 2's secrets respectively.  Then they both
+perform a secure opening of $\secret{x} \fplus \flip{a}$ and
+$\secret{y} \fplus \flip{b}$, by authenticating each others' shares,
+and record them as $\mesg{d}$ and $\mesg{e}$ respectively. Note
+that the secrets remain hidden since $\flip{a}$ and $\flip{b}$
+are random values.
 $$
@@
   \out{1} &:=& \elab{(\macopen\ \rvl{1} \ \macauth(\rvl{2},\ldots))}{1} \\
-  \out{2} &:=& \elab{(\macopen\ \macauth(\rvl{1},\ldots) \ \macauth(\rvl{2},\ldots))}{2} 
+  \out{2} &:=& \elab{(\macopen\ \macauth(\rvl{1},\ldots) \ \rvl{2})}{2} 
 \end{array}

### 2024-04-20 revision-1/model.tex
checkpoint

@@
 \section{Security Model}
+
+\subsection{Probability Mass Functions and Program Distributions}
+\label{section-pmf}
+
+Given a program $\prog$ with $\iov(\prog) = S \cup V$ and $\store \in
+\mems(\iov(\prog))$, we are fundamentally concerned with the joint
+probability of the assignments in $\store_V$ conditioned by the
+assignments $\store_S$ in any given run, since probabilistic
+dependence between honest secrets and corrupt views is our metric of
+information leakage.
+
+\subsection{Basic Definitions}
+
+We begin by defining joint probability mass functions in the standard
+manner, though following \cite{barthe2019probabilistic} we use
+memories to denote mappings of variables to values (i.e., outcomes),
+so for example given a pmf $\pmf$ we will write $\pmf(\{ \sx{1}{a}
+\mapsto 0, \vx{2}{b} \mapsto 1 \})$ to denote the probability that
+$\sx{1}{a} = 0$ and $\vx{2}{b} = 1$ in the final memory of a
+program containing these variables.
+\begin{definition}
+  A \emph{probability mass function} $\pmf$ is a function
+  mapping memories in $\mems(X)$ for some $X$ to values in $[0,1]$, such that:
+  $$
+  \sum_{\store \in \mems(X)} \pmf(\store) \  = \ 1
+  $$
+\end{definition}
+%To recover succinct and familiar notation, we may omit the domain of a
+%distribution when it is clear from an application context-
+%i.e., we allow the following sugaring:
+%$$
+%\pdf{}(\store) \defeq \pdf{\dom(\store)}(\store)
+%$$
+Now, we can define a notion of marginal and conditional
+distributions as follows, which are standard for discrete
+probability mass functions. 
+\begin{definition}
+  Given $\pmf$ the \emph{marginal distribution} of variables $X$
+  in $\pmf$, denoted $\margd{\pmf}{X}$, is defined as follows:
+  $$
+  \forall \store \in \mems(X) \quad . \quad \margd{\pmf}{X}(\store) =
+  \sum_{\store' \in \mems(X-\dom(\dom(\pmf)))} \pmf(\store \cap \store')
+  $$
+\end{definition}
+
+\begin{definition}
+  Given $\margd{\pmf}{X}$, its \emph{conditional distribution given
+  $\stores$} with $\dom(\stores) = Y$ with $Y \subseteq X$, denoted
+  $\condd{\pmf}{X}{\stores}$, is a distribution with domain $X$ where for all
+  $\store \in \mems(X)$:
+  $$
+  \condd{\pmf}{X}{\stores}(\store) =
+  (\sum_{\store' \in \stores} \margd{\pmf}{X}(\store \cap \store')) /
+  (\sum_{\store' \in \stores} \margd{\pmf}{Y}(\store'))
+  $$
+  where $\margd{\pmf}{X}(\store \cap \store')) = 0$ if $\store \cap \store'$ is undefined.
+\end{definition}
+To recover familiar notation we allow the syntactic
+sugarings $\condd{\pmf}{X}{\store}  \defeq \condd{\pmf}{X}{\{ \store\}}$, and
+$\pmf(\store)  \defeq \margd{\pmf}{X}(\store)$ and $\pmf(\store|\stores) \defeq
+\condd{\pmf}{X}{\stores}(\store)$ where $\dom(\store) = X$.
+%\begin{eqnarray*}
+%  \condd{\pmf}{X}{\store}  &\defeq& \condd{\pmf}{X}{\{ \store\}}\\
+%  \pmf(\store)  &\defeq& \margd{\pmf}{X}(\store)  \qquad \dom(\store) = X\\
+%  \pmf(\store|\stores)  &\defeq& \condd{\pmf}{X}{\stores}(\store) \qquad \dom(\store) = X
+%\end{eqnarray*}
+
+\subsection{Program Distributions}
+Now we can define the probability distribution of a program $\prog$,
+that we denote $\progd(\prog)$. Since $\fedcat$ is deterministic the
+results (view assignments) of any run are determined by the input
+secrets together with the random tape. And since we constrain programs
+to not overwrite views, we are assured that \emph{final} memories
+contain both a complete record of all initial secrets as well as views
+resulting from communicated information.
+
+We consider random tapes to be selected from a uniform distribution.
+In our threat model, we will also assume that any possible initial
+secrets are equally likely when formulating the program distribution.
+In this setting, given a program $\prog$ with $\iov(\prog) = S \cup
+V$ and $\flips(\prog) = F$ we will consider all $\store
+\in \mems(S \cup V \cup F)$ such that
+$
+\config{\store_{S \cup F}}{\prog} \redxs \config{\store_}{\varnothing}
+$
+to be equally probable, establishing the basic distribution of the
+program. From this, we can immediately derive the marginal distribution
+of $S \cup V$ to reason about dependencies between secrets and views. 
+\begin{definition}
+  \label{def-progd}
+  \label{definition-progd}
+  Given program $\prog$ with $\iov(\prog) = (S,V)$ and $\flips(\prog) = F$, define $\stores$ as:
+  $$
+  \stores \defeq \{ \store \mid (\dom(\store) = S \cup V \cup F) \wedge (\store_{S \cup F},\prog) \redxs (\store,\varnothing) \}
+  $$
+  We note immediately that $|\stores| = 2^{S \cup F}$ given the (deterministic) definition of $\redx$. 
+  Define also $\pmf$ as the program's \emph{basic distribution} such that for all
+  $\store \in \mems(S \cup V \cup F)$:
+  $$
+  \pmf(\store) =  1 / |\stores| \ \text{if}\ \store \in \stores, \text{otherwise}\ 0
+  %\begin{cases}
+  %  1 / |\stores| & \text{if}\ \store \in \stores\\
+  %  0 & \text{otherwise}
+  %\end{cases}
+  $$
+  Then the \emph{program distribution of $\prog$}, denoted $\progd(\prog)$, is the
+  marginal distribution of $S \cup V$ in $\prog$'s basic distribution, i.e., 
+  $
+  \progd(\prog) \defeq  \margd{\pmf}{S\cup V}
+  $.
+  In some cases, we will also be concerned with the (joint)
+  probabilities of expression interpretation given a preceding program
+  execution, and we write $\progd(\prog, \be)$ to denote the program
+  distribution $\progd(\prog;\itv := \be)$ where $\itv$ is a
+  special variable that is never used in programs.
+\end{definition}
+
+\begin{definition}
+  We write $\vc{\pmf}{x}{y}$ iff $\pmf(\{ x \mapsto 0\}\ |\ \{ y \mapsto 0 \}) =
+  \pmf(\{ x \mapsto 1\}\ |\ \{ y \mapsto 1 \}) = 1$.
+  We write $\sep{\pmf}{X}{Y}$ iff for all
+    $\store \in \mems(X \cup Y)$ we have
+  $\margd{\pmf}{X \cup Y}(\store) =
+  \pmf(\store_X) * \pmf(\store_Y)$
+\end{definition}
 

### 2024-04-20 revision-1/hyper.tex
checkpoint

@@
 \section{Security Hyperproperties}
+
+\begin{definition}
+  We write $\vc{\pmf}{x}{y}$ iff $\pmf(\{ x \mapsto 0\}\ |\ \{ y \mapsto 0 \}) =
+  \pmf(\{ x \mapsto 1\}\ |\ \{ y \mapsto 1 \}) = 1$.
+  We write $\sep{\pmf}{X}{Y}$ iff for all
+    $\store \in \mems(X \cup Y)$ we have
+  $\margd{\pmf}{X \cup Y}(\store) =
+  \pmf(\store_X) * \pmf(\store_Y)$
+\end{definition}
 
@@
 \end{definition}
-
-\begin{theorem}
-  Gradual release implies passive security.
-\end{theorem}
 
@@
 \begin{theorem}
-  Robust declassification implies malicious security.
+  Passive security with robust declassification implies malicious security.
 \end{theorem}

### 2024-04-20 revision-1/lang.tex
checkpoint

@@
 
-We let $x$ range over \emph{variables}  which are identifiers
-where client ownership is specified- e.g., $\elab{\mesg{\mathit{foo}}}{\cid}$
-is a message $\mathit{foo}$ that was sent to $\cid$. We let $X$
-range over sets of variables, and more specifically, $S$ ranges over sets of secret variables $\elab{\secret{w}}{\cid}$, $R$ ranges over sets of random variables $\elab{\flip{w}}{\cid}$, $M$ ranges over sets of message variables $\elab{\mesg{w}}{\cid}$, $P$ ranges over sets of public variables $\rvl{w}$, and $O$ ranges over sets of output variabels $\out{\cid}$.
-Given a program $\prog$, we write $\iov(\prog)$ to
-denote the set of $S \cup M \cup P \cup O$ of variables in $\prog$
-with ownership made explicit, and we write $\flips(\prog)$ to
-denote the set $R$ of random samplings in $\prog$ with ownership
-made explicit. We write
-$\vars(\prog)$ to denote $\iov(\prog) \cup \flips(\prog)$. For any set
-of variables $X$ and parties $P$, we write $X_P$ to denote the subset
-of $X$ owned by any party in $P$, in particular we write $X_H$ and $X_C$ to
-denote the subsets belonging to honest and corrupt parties,
-respectively.
+We let $x$ range over \emph{variables} which are identifiers where
+client ownership is specified- e.g.,
+$\elab{\mesg{\mathit{foo}}}{\cid}$ is a message $\mathit{foo}$ that
+was sent to $\cid$. We let $X$ range over sets of variables, and more
+specifically, $S$ ranges over sets of secret variables
+$\elab{\secret{w}}{\cid}$, $R$ ranges over sets of random variables
+$\elab{\flip{w}}{\cid}$, $M$ ranges over sets of message variables
+$\elab{\mesg{w}}{\cid}$, $P$ ranges over sets of public variables
+$\rvl{w}$, and $O$ ranges over sets of output variables $\out{\cid}$.
+Given a program $\prog$, we write $\iov(\prog)$ to denote the set $S
+\cup M \cup P \cup O$ of variables in $\prog$ with ownership made
+explicit and $\secrets(\prog)$ to denote $S$, and we write
+$\flips(\prog)$ to denote the set $R$ of random samplings in $\prog$
+with ownership made explicit. We write $\vars(\prog)$ to denote
+$\iov(\prog) \cup \flips(\prog)$. For any set of variables $X$ and
+parties $P$, we write $X_P$ to denote the subset of $X$ owned by any
+party in $P$, in particular we write $X_H$ and $X_C$ to denote the
+subsets belonging to honest and corrupt parties, respectively.
 
@@
   \elab{\mesg{d}}{1} &:=&
-  \elab{(\macopen\ \macauth(\mesg{a}, \mack{\macx{\secret{x}}{1}}{2} \fminus \mack{\macx{\flip{a}}{\Oracle}}{2}) \ (\macgv{\elab{\secret{x}}{1}}\macminus\macgv{\elab{\flip{a}}{\Oracle}}))}{1}\\
+  \elab{(\macopen\ \macauth(\mesg{a}, \mack{\macx{\secret{x}}{1}}{} \fminus \mack{\macx{\flip{a}}{\Oracle}}{}) \ (\macgv{\elab{\secret{x}}{1}}\macminus\macgv{\elab{\flip{a}}{\Oracle}}))}{1}\\
   \elab{\mesg{e}}{1}&:=&
-  \elab{(\macopen\ \macauth(\mesg{b}, \mack{\macx{\secret{y}}{2}}{2} \fminus \mack{\macx{\flip{b}}{\Oracle}}{2}) \ (\macgv{\elab{\secret{y}}{2}}\macminus\macgv{\elab{\flip{b}}{\Oracle}}))}{1}\\
+  \elab{(\macopen\ \macauth(\mesg{b}, \mack{\macx{\secret{y}}{2}}{} \fminus \mack{\macx{\flip{b}}{\Oracle}}{}) \ (\macgv{\elab{\secret{y}}{2}}\macminus\macgv{\elab{\flip{b}}{\Oracle}}))}{1}\\
   \elab{\mesg{d}}{2} &:=&
-  \elab{(\macopen\ \macauth(\mesg{a}, \mack{\macx{\secret{x}}{1}}{1} \fminus \mack{\macx{\flip{a}}{\Oracle}}{1}) \ (\macgv{\elab{\secret{x}}{1}}\macminus\macgv{\elab{\flip{a}}{\Oracle}}))}{2}\\
+  \elab{(\macopen\ \macauth(\mesg{a}, \mack{\macx{\secret{x}}{1}}{} \fminus \mack{\macx{\flip{a}}{\Oracle}}{}) \ (\macgv{\elab{\secret{x}}{1}}\macminus\macgv{\elab{\flip{a}}{\Oracle}}))}{2}\\
   \elab{\mesg{e}}{2}&:=&
-  \elab{(\macopen\ \macauth(\mesg{b}, \mack{\macx{\secret{y}}{2}}{1} \fminus \mack{\macx{\flip{b}}{\Oracle}}{1}) \ (\macgv{\elab{\secret{y}}{2}}\macminus\macgv{\elab{\flip{b}}{\Oracle}}))}{2}
+  \elab{(\macopen\ \macauth(\mesg{b}, \mack{\macx{\secret{y}}{2}}{} \fminus \mack{\macx{\flip{b}}{\Oracle}}{}) \ (\macgv{\elab{\secret{y}}{2}}\macminus\macgv{\elab{\flip{b}}{\Oracle}}))}{2}
 \end{array}

### 2024-04-20 revision-1/model.tex
checkpoint

--- 2024-04-17 revision-1/model.tex
@@
 \section{Security Model}
+\label{section-pmf}
+
+\emph{Simulator security} in the MPC setting means that protocols do
+not reveal any secret information other than what is implicitly
+declassified by the output of the ideal functionality that the
+protocol implements. Characterization of both real world protocol
+execution and simulation is defined probabilistically. Following
+previous work \cite{barthe2019probabilistic} we use probability mass
+functions to express joint dependencies between input and output
+variables, as a metric of information leakage.
+
+\subsection{Probability Mass Functions}
+
+We define joint probability mass functions (pmfs) in the standard
+manner, though following \cite{barthe2019probabilistic} we use
+memories to denote mappings of variables to values (i.e., outcomes),
+so for example given a pmf $\pmf$ we will write $\pmf(\{ \sx{1}{a}
+\mapsto 0, \vx{2}{b} \mapsto 1 \})$ to denote the probability that
+$\sx{1}{a} = 0$ and $\vx{2}{b} = 1$ in the final memory of a
+program containing these variables.
+\begin{definition}
+  A \emph{probability mass function} $\pmf$ is a function
+  mapping memories in $\mems(X)$ for some $X$ to values in $\mathbb{F}_p$, such that:
+  $$
+  \sum_{\store \in \mems(X)} \pmf(\store) \  = \ 1
+  $$
+\end{definition}
+%To recover succinct and familiar notation, we may omit the domain of a
+%distribution when it is clear from an application context-
+%i.e., we allow the following sugaring:
+%$$
+%\pdf{}(\store) \defeq \pdf{\dom(\store)}(\store)
+%$$
+Now, we can define a notion of marginal and conditional
+distributions as follows, which are standard for discrete
+probability mass functions. 
+\begin{definition}
+  Given $\pmf$ the \emph{marginal distribution} of variables $X$
+  in $\pmf$, denoted $\margd{\pmf}{X}$, is defined as follows:
+  $$
+  \forall \store \in \mems(X) \quad . \quad \margd{\pmf}{X}(\store) =
+  \sum_{\store' \in \mems(X-\dom(\dom(\pmf)))} \pmf(\store \cap \store')
+  $$
+\end{definition}
+
+\begin{definition}
+  Given $\margd{\pmf}{X}$, let $\stores$ be a set of memories with the
+  same domain $Y \subseteq X$. Then the \emph{conditional distribution given
+  $\stores$}  denoted
+  $\condd{\pmf}{X}{\stores}$ is a distribution with domain $X$ where for all
+  $\store \in \mems(X)$:
+  $$
+  \condd{\pmf}{X}{\stores}(\store) =
+  (\sum_{\store' \in \stores} \margd{\pmf}{X}(\store \cap \store')) /
+  (\sum_{\store' \in \stores} \margd{\pmf}{Y}(\store'))
+  $$
+  where $\margd{\pmf}{X}(\store \cap \store')) = 0$ if $\store \cap \store'$ is undefined.
+\end{definition}
+To recover familiar notation we allow the syntactic
+sugarings $\condd{\pmf}{X}{\store}  \defeq \condd{\pmf}{X}{\{ \store\}}$, and
+$\pmf(\store)  \defeq \margd{\pmf}{X}(\store)$ and $\pmf(\store|\stores) \defeq
+\condd{\pmf}{X}{\stores}(\store)$ where $\dom(\store) = X$.
+%\begin{eqnarray*}
+%  \condd{\pmf}{X}{\store}  &\defeq& \condd{\pmf}{X}{\{ \store\}}\\
+%  \pmf(\store)  &\defeq& \margd{\pmf}{X}(\store)  \qquad \dom(\store) = X\\
+%  \pmf(\store|\stores)  &\defeq& \condd{\pmf}{X}{\stores}(\store) \qquad \dom(\store) = X
+%\end{eqnarray*}
+
+\subsection{Basic Distribution of a Protocol}
+Now we can define the probability distribution of a program $\prog$,
+that we denote $\progtt(\prog)$. Since $\fedcat$ is deterministic the
+results of any run are determined by the input secrets together with
+the random tape. And since we constrain programs to not overwrite
+views, we are assured that \emph{final} memories contain both a
+complete record of all initial secrets as well as views resulting from
+communicated information.
+
+We consider random tapes to be selected from a uniform distribution
+modulo relevant pre-processing, e.g., Beaver triples and MACed share
+distributions.  We will also assume that any possible initial secrets
+are equally likely when formulating the program distribution.
+We define $\runs(\prog)$ as the set of final memories resulting
+from execution of $\prog$ given any initial memory, and
+treat all elements of $\runs(\prog)$ as equally likely.
+This establishes the basic program distribution that can be
+marginalized and conditioned to quantify input/output
+information dependencies. 
+%In this
+%setting, given a program $\prog$ with $\iov(\prog) = S \cup V$ and
+%$\flips(\prog) = F$ we will consider all $\store \in \mems(S \cup V
+%\cup F)$ such that $ \config{\store_{S \cup F}}{\prog} \redxs
+%\config{\store_}{\varnothing} $ to be equally probable, establishing
+%the basic distribution of the program. %From this, we can immediately
+%derive the marginal distribution of $S \cup V$ to reason about
+%dependencies between secrets and views.
+\begin{definition}
+  \label{def-progd}
+  \label{definition-progd}
+  Given program $\prog$ with $\secrets(\prog) = S$ and $\flips(\prog) = R$, and
+  any assumed pre-processing predicate $\preproc$ on memories, define:
+  $$
+  \runs(\prog) \defeq \{ \store \mid \preproc(\store) \wedge
+  (\dom(\store) = \iov(\prog) \cup R) \wedge (\store_{S \cup R},\prog) \redxs (\store,\varnothing) \}
+  $$
+  Then the \emph{basic distribution of $\prog$}, written $\progtt(\prog)$, is
+  defined such that for all $\store \in \mems(\iov(\prog) \cup R)$:
+  $$
+  \progtt(\store) =  1 / |\runs(\prog)| \ \text{if}\ \store \in \runs(\prog), \text{otherwise}\ 0
+  $$
+  In some cases, we will also be concerned with the (joint)
+  probabilities of expression interpretation given a preceding program
+  execution, and we write $\progtt(\prog, \be)$ to denote the program
+  distribution $\progtt(\prog;\itv := \be)$ where $\itv$ is a
+  special variable that is never used in programs.
+\end{definition}
 
@@
 
-The simulator is represented by a probabilistic algorithm $\SIM_C$,
-aka a \emph{simulation}, that is parameterized by corrupt inputs and
-the output of an ideal functionality, and that returns a set of
-adversarial views (as a memory) with some probability. Given
-corrupt inputs $\store$ and ideal functionality output $v$,  
-we write
-$
-\prob(\SIM(\store,v) = \store')
-$
-to denote the probability that $\SIM(\store,v)$
-returns corrupt views $\store'$ as a result. We can then define the
-probability distribution of corrupt views reconstructed
-by the simulator as follows:
+Information about honest secrets can be revealed to corrupt clients
+through messages sent from honest to corrupt clients, and through
+publicly broadcast information from honest clients. Dually,
+corrupt clients can impact protocol integrity through the messages
+sent from corrupt to honest clients, and through publicly broadcast information
+from corrupt clients. We call the former \emph{corrupt views}, and
+the latter \emph{honest views}. Generally we let $V$ range over sets
+of views.
+\begin{definition}[Corrupt and Honest Views]
+  Given a program $\prog$ with $\iov(\prog) = S \cup M \cup P \cup O$,
+  define $\views(\prog) \defeq M \cup P$, and define $\houtputs$ as the messages in
+  $V = \views(\prog)$ sent from honest
+  to corrupt parties, called \emph{corrupt views}:
+  $$
+  \houtputs = \{\ \rvl{w} \mid\ \reveal{w}{\be}{\cid} \in \prog
+  \wedge \cid \in H \ \} 
+  \cup \{\ \elab{\mesg{w}}{\cid}\ \mid\  \eassign{\mesg{w}}{\cid}{\be}{\cid'} \in \prog
+  \wedge \cid \in C \wedge \cid' \in H \ \} 
+  $$
+  and similarly define $\cinputs$ as the messages in $V$ sent from corrupt to honest
+  parties, called \emph{honest views}:
+  $$
+  \cinputs = \{\ \rvl{w} \mid\ \reveal{w}{\be}{\cid} \in \prog
+  \wedge \cid \in C \ \} 
+  \cup \{\ \elab{\mesg{w}}{\cid}\ \mid\  \eassign{\mesg{w}}{\cid}{\be}{\cid'} \in \prog
+  \wedge \cid \in H \wedge \cid' \in C \ \} 
+  $$
+\end{definition}
+
+In the passive setting the simulator must construct a probabilistic
+algorithm $\SIM$, aka a \emph{simulation}, that is parameterized by
+corrupt inputs and the output of an ideal functionality, and that
+returns a distribution of corrupt views that is indistinguishable
+from the distribution of corrupt views in the real world
+protocol execution.
+
+Given inputs $\store$ and ideal
+functionality output $v$, we write $ \prob(\SIM(\store,v) = \store') $
+to denote the probability that $\SIM(\store,v)$ returns corrupt views
+$\store'$ as a result. We can then define the probability distribution
+of corrupt views reconstructed by the simulator as follows:
 \begin{definition}
@@
 \end{definition}
-
 Then we can define passive security in the real/ideal
@@
   Assume given a program $\prog$ that correctly implements an ideal
-  functionality $\idealf$, with $\iov(\prog) = (S,M,O)$.  Then $\prog$
+  functionality $\idealf$, with $\views(\prog) = V$.  Then $\prog$
   is \emph{passive secure in the simulator model} iff for all
   partitions of the federation into honest and corrupt sets $H$ and $C$
-  with $|C| < |H|$ and for all $\store \in \mems(S)$ there exists a
+  and for all $\store \in \mems(S)$ there exists a
   simulation $\SIM$ such that:
   $$
-  \dist(\SIM(\store_{S_C},\idealf(\store))) = \condd{\progtt(\prog)}{M_C}{\store}
+  \dist(\SIM(\store_{C},\idealf(\store))) = \condd{\progtt(\prog)}{\houtputs}{\store}
   $$
@@
 
-\begin{definition}[Corrupt Inputs, Honest Outputs]
-  Given a program $\prog$ with $\iov(\prog) = (S,M,O)$ , define $\cinputs$ as the
-  messages in $M$ sent from corrupt to honest parties:
-  $$
-  \cinputs = \{\ \elab{\mesg{w}}{\cid}\ \mid\  \elab{\mesg{w}}{\cid} \in M \wedge \eassign{\mesg{w}}{\cid}{\be}{\cid'} \in \prog
-  \wedge \cid \in H \wedge \cid' \in C \ \} 
-  $$
-  and similarly define $\houtputs$ as the messages in $M$ sent from honest to corrupt parties.
-  %Define also $(\afilter\ \prog)$ as $\prog$ with all instructions of the form $\eassign{\mesg{w}}{\cid}{\be}{\cid'}$ removed
-  %where $\cid \in H \wedge \cid' \in C$.
-\end{definition}
-
 $$

### 2024-04-20 revision-1/model.tex
checkpoint

@@
 \end{definition}
+Observe that this definition relies on protocol correctness, which
+we can define as follows.
+\begin{definition}[Passive Correctness]
+  Given $\prog$ with output variables $\out{1},\ldots,\out{n}$ and ideal
+  functionality $\idealf$, we say that $\prog$ is \emph{passive correct}
+  for $\idealf$ iff for all $\store \in \mems(\secrets(\prog))$
+  with $\idealf(\store) = v_1,\ldots,v_n$ we have
+  $\progtt(\prog)(\out{1} \mapsto v_1,\ldots,\out{n} \mapsto v_n \mid \store) = 1$.
+\end{definition}
 

### 2024-04-20 revision-1/model.tex
checkpoint

@@
 
+In the malicious model we assume that corrupt clients are in the thrall of an
+adversary $\adversary$ who does not necessarily follow the rules of the protocol.
+We model this by positing a $\arewrite$ function which is given a corrupt memory
+$\store_C$ and expression $\be$, and returns a rewritten expression that can
+be interpreted to yield a corrupt input. Then we define an evaluation
+relation that incorporates the adversary:
 $$
@@
 $$
+Given this we define the $\adversary$ distribution of $\prog$. 
+\begin{definition}
+  \label{def-progd}
+  \label{definition-progd}
+  Given program $\prog$ with $\secrets(\prog) = S$ and $\flips(\prog) = R$, and
+  any assumed pre-processing predicate $\preproc$ on memories, define:
+  $$
+  \aruns(\prog) \defeq \{ \store \mid \preproc(\store) \wedge
+  (\dom(\store) = \iov(\prog) \cup R) \wedge (\store_{S \cup R},\prog) \aredx^* (\store,\varnothing) \}
+  $$
+  Then the \emph{$\adversary$ distribution of $\prog$}, written $\progtt(\prog,\adversary)$, is
+  defined such that for all $\store \in \mems(\iov(\prog) \cup R)$:
+  $$
+  \progtt(\store) =  1 / |\runs(\prog)| \ \text{if}\ \store \in \aruns(\prog), \text{otherwise}\ 0
+  $$
+\end{definition}
 
+\begin{definition}[Malicious Security]
+  Let $\prog$ be a protocol with $\iov(\prog) = S \cup V \cup O$. We say that
+  $\prog$ is \emph{malicious secure} iff for all
+  partitions of the federation into honest and corrupt sets $H$ and $C$,
+  adversaries $\adversary$, 
+  and all $\store \in \mems(S_H)$ there exists a
+  simulation $\SIM$ which is given $\store$ and has two phases:
+  \begin{enumerate}
+  \item In the first phase (aka input extraction), $\adversary$ gives the
+    simulator some $\store' \in \mems(S_C)$, and the simulator consults an
+    oracle to compute $\idealf(\store \cup \store')$.
+  \item In the second phase, $\SIM$ reconstructs corrupt views $\houtputs$,
+    which it returns along with the honest outputs in $\idealf(\store \cup \store')$, such that:
     $$
-\dist(\SIM(\store_{S_H})) = \condd{\progtt(\prog)}{\houtputs \cup O_H}{\store_{S_H}}
+    \dist(\SIM(\store)) = \condd{\progtt(\prog,\adversary)}{\houtputs \cup O_H}{\store}
     $$
+  \end{enumerate}
+\end{definition}

### 2024-04-21 revision-1/hyper.tex
checkpoint

@@
 
+In this Section we formulate probabilistic versions of well-studied
+hyperproperties of confidentiality and integrity, including noninterference,
+gradual release, declassification, and robust declassification.
+We demonstrate a soundness relation between noninterference and
+passive security, and between robust declassification and malicious
+security. We subsequently leverage this relation to enforce
+malicious security using ``traditional'' security type methods
+in Section \ref{section-types}. Previous work has explored
+a similar approach to security type enforcement
+\cite{6266151,almeida2018enforcing} but mainly
+for aspects of passive security.
+
+\subsection{Passive Security and Noninterference}
+
+Since MPC protocols release some information about secrets through
+outputs of $\idealf$, they do not enjoy strict noninterference.  As
+discussed in Section \ref{section-lang}, public reveals and protocol
+outputs are fundamentally forms of declassification.  But consistent
+with other work \cite{8429300}, we can formulate a version of
+probabilistic noninterence conditioned on output that is sound
+for passive security. It says that if two low-equivalent secret
+inputs generate the same output, then the distributions of corrupt
+views are the same. 
+\begin{definition}[Noninterference modulo output]
+  \label{definition-NIMO}
+  We say that a program $\prog$ satisfies \emph{noninterference modulo output}
+  iff for all $H$ and $C$ and 
+  $\store_1,\store_2 \in \mems(S)$ we have:
+  $$
+  (\store_1 =_C \store_2 \ \wedge \ 
+  (\condd{\progtt(\prog)}{O}{\store_1} = \condd{\progtt(\prog)}{O}{\store_2}))
+  \implies 
+  (\condd{\progtt(\prog)}{\houtputs}{\store_1} = \condd{\progtt(\prog)}{\houtputs}{\store_2})
+  $$
+  where $\iov(\prog) = S \cup V \cup O$.
+\end{definition}
+Intuitively, this conditional noninterference property implies that
+the simulator can just run the protocol in simulation to
+reconstruct real world corrupt view distributions. But it requires
+that the simulator can tractably ``pre-image'' a given output of
+a functionality $\idealf$, to determine the inputs that
+could have produced it. This pre-image is called a
+\emph{kernel} in recent work \cite{XXX}.
 \begin{definition}
-  We write $\vc{\pmf}{x}{y}$ iff $\pmf(\{ x \mapsto 0\}\ |\ \{ y \mapsto 0 \}) =
-  \pmf(\{ x \mapsto 1\}\ |\ \{ y \mapsto 1 \}) = 1$.
+  Given a functionality $\idealf$ and output value $v$, its
+  \emph{kernel}, denoted $\ik(\idealf,v)$ is
+  $
+  \{ \store\ |\ \idealf(\store) = v \}
+  $.
+  We say that $\idealf$ is \emph{pre-imageable} iff $\ik(\idealf, v)$ for all
+  $v$ can be computed tractably.
+\end{definition}
+A soundness result for passive security can then be stated as follows.
+We prove this in a separate manuscript \cite{XXX}, and it is also
+essentially the same as ``perfect passive NI security'' which
+has a similar soundness property \cite{8429300}.  
+\begin{restatable}{theorem}{nimosecure}
+  \label{theorem-nimo}
+  Assume given pre-imageable $\idealf$ and a protocol $\prog$ that
+  correctly implements $\idealf$.  If $\prog$ satisfies noninference modulo output
+  then $\prog$ is passive secure.
+\end{restatable}
+
+\subsection{Gradual Release as a Design Pattern}
+
+Previous work has discussed how MPC security is not noninterference,
+but rather how ideal functional sets an upper bound on
+declassification \cite{6266151,almeida2018enforcing}. Nevertheless,
+probabilistic noninterference is preserved by components of
+cryptographic protocols generally, and can be expressed using
+\emph{probabilistic independence} \cite{darais2019language,barthe2019probabilistic}
+We introduce important notation to express independence:
+\begin{definition}
+%  We write $\vc{\pmf}{x}{y}$ iff $\pmf(\{ x \mapsto 0\}\ |\ \{ y \mapsto 0 \}) =
+%  \pmf(\{ x \mapsto 1\}\ |\ \{ y \mapsto 1 \}) = 1$.
   We write $\sep{\pmf}{X}{Y}$ iff for all
@@
 
-\subsection{Passive Security and Noninterference}
-
-\cnote{Fix the $O$ output variable things, and scoping of $H,C$.}
-
-\begin{definition}[Passive Confidentiality]
-  \label{definition-NIMO}
-  We say that a program $\prog$ satisfies \emph{passive confidentiality}
-  iff for all $H$ and $C$ with $|C|\le|H|$ and 
-  $\store_1,\store_2 \in \mems(S)$ we have:
-  $$
-  (\store_1 =_C \store_2 \ \wedge \ 
-  (\condd{\progtt(\prog)}{\{ \outv \}}{\store_1} = \condd{\progtt(\prog)}{\{ \outv \}}{\store_2}))
-  \implies 
-  (\condd{\progtt(\prog)}{\houtputs}{\store_1} = \condd{\progtt(\prog)}{\houtputs}{\store_2})
-  $$
-  where $\iov(\prog) = (S,M)$.
-\end{definition}
-
-\subsection{Gradual Release as a Design Pattern}
-
+In fact, MPC protocols typically satisfy a \emph{gradual release}
+property\cite{XXX}, where messages exchanged remain probabilistically separable
+from secrets, with only declassification events (reveals and outputs)
+releasing information about honest secrets. 
 \begin{definition}
-  Given $H,C$, a protocol $\prog$ with $\iov(\prog) = (S,M,O)$ satisfies \emph{gradual release} iff
-  $\sep{\progtt(\prog)}{M_C}{S_H}$.
+  Given $H,C$, a protocol $\prog$ with $\iov(\prog) = S \cup M \cup P \cup O$
+  satisfies \emph{gradual release} iff
+  $\sep{\progtt(\prog)}{\houtputs}{S_H}$.
 \end{definition}

### 2024-04-21 revision-1/lang.tex
checkpoint

@@
 \section{The $\minicat$ Protocol Language}
+\label{section-lang}
 
@@
   is a \emph{public reveal} where expression $\be$ is computed
-  by client $\cid$ and broadcast to the federation.
+  by client $\cid$ and broadcast to the federation, typically
+  to communicate intermediate results for use in final output
+  computations.
 \item $\pubout{\cid}{\be}{\cid}$: This
@@
 referenced in expressions, once they've been assigned.
+This distinction between messages and broadcast public
+reveal is consistent with previous formulations, e.g.,
+\cite{6266151}.
 

### 2024-04-21 revision-1/hyper.tex
checkpoint

@@
   satisfies \emph{gradual release} iff
-  $\sep{\progtt(\prog)}{\houtputs}{S_H}$.
+  $\sep{\progtt(\prog)}{M_C}{S_H}$.
 \end{definition}

### 2024-04-21 revision-1/hyper.tex
checkpoint

@@
 
+\begin{definition}[Malicious Correctness]
+  A protocol $\prog$ with $\iov(\prog) = S \cup V \cup O$ is \emph{malicious correct} iff
+  $
+  \forall \adversary, \store \in \mems(S_H) \ .\ \support(\progtt(\prog)(O_H|\store)) =
+    \support(\progtt(\prog,\adversary)(O_H|\store))
+  $.
+\end{definition}
+
 \begin{definition}
-  We say that a protocol $\prog$ with $\iov(\prog) = (S,M)$ satisfies \emph{active confidentiality} iff the following conditions hold
-  for all adversaries $\adversary$:
-  \begin{enumerate}
-  \item $\ \,\forall \store \in \mems(S_H) \ .\ \support(\progtt(\prog)(\{ \outv \}|\store)) =
-    \support(\progtt(\prog,\adversary)(\{ \outv \}|\store))$
-  \item $\begin{array}[t]{l}\forall \store_1, \store_2 \in \mems(S_H), \store \in \mems(\cinputs)\ . \\
+  We say that a protocol $\prog$ with $\iov(\prog) = S \cup V \cup O$ satisfies
+  \emph{active confidentiality} iff
+  it is malicious correct, and for all adversaries $\adversary$:
+  %\begin{enumerate}
+  %\item $\ \,\forall \store \in \mems(S_H) \ .\ \support(\progtt(\prog)(\{ \outv \}|\store)) =
+    %\support(\progtt(\prog,\adversary)(\{ \outv \}|\store))$
+    %\item
+    $$\begin{array}[t]{l}\forall \store_1, \store_2 \in \mems(S_H), \store \in \mems(\cinputs)\ . \\
     \quad
-    \condd{\progtt(\prog,\adversary)}{\{ \outv \}}{\store_1 \cup \store} =
-    \condd{\progtt(\prog,\adversary)}{\{ \outv \}}{\store_2 \cup \store} \\
-    \qquad \implies\\
-    \quad
+    \condd{\progtt(\prog,\adversary)}{O_H}{\store_1 \cup \store} =
+    \condd{\progtt(\prog,\adversary)}{O_H}{\store_2 \cup \store} 
+    %\qquad
+    \implies
+    %\quad
     \condd{\progtt(\prog,\adversary)}{\houtputs}{\store_1 \cup \store} =
-    \condd{\progtt(\prog,\adversary)}{\houtputs}{\store_2 \cup \store}\end{array}$
-  \end{enumerate}
+    \condd{\progtt(\prog,\adversary)}{\houtputs}{\store_2 \cup \store}
+    \end{array}$$
+  %\end{enumerate}
 \end{definition}
@@
   $\forall \adversary . \runs_\adversary(\prog) \subseteq \runs(\prog)$.
-\end{definition}
-
-\begin{definition}
-  A protocol $\prog$ with $\iov(\prog) = (S,M,O)$ is \emph{malicious correct} iff:
-  $$
-  \forall \adversary, \store \in \mems(S_H) \ .\ \support(\progtt(\prog)(O_H|\store)) =
-    \support(\progtt(\prog,\adversary)(O_H|\store))
-  $$
 \end{definition}

### 2024-04-21 revision-1/hyper.tex
checkpoint

@@
 
-\begin{definition}
-  We say that a protocol $\prog$ with $\iov(\prog) = S \cup V \cup O$ satisfies
-  \emph{active confidentiality} iff
-  it is malicious correct, and for all adversaries $\adversary$:
-  %\begin{enumerate}
-  %\item $\ \,\forall \store \in \mems(S_H) \ .\ \support(\progtt(\prog)(\{ \outv \}|\store)) =
-    %\support(\progtt(\prog,\adversary)(\{ \outv \}|\store))$
-    %\item
-    $$\begin{array}[t]{l}\forall \store_1, \store_2 \in \mems(S_H), \store \in \mems(\cinputs)\ . \\
-    \quad
-    \condd{\progtt(\prog,\adversary)}{O_H}{\store_1 \cup \store} =
-    \condd{\progtt(\prog,\adversary)}{O_H}{\store_2 \cup \store} 
-    %\qquad
-    \implies
-    %\quad
-    \condd{\progtt(\prog,\adversary)}{\houtputs}{\store_1 \cup \store} =
-    \condd{\progtt(\prog,\adversary)}{\houtputs}{\store_2 \cup \store}
-    \end{array}$$
-  %\end{enumerate}
+\begin{definition}[Integrity]
+  We say that a protocol $\prog$ with $\iov(\prog) = S \cup V \cup O$ has
+  \emph{integrity} iff for all $H$ and $C$ 
+  it is malicious correct and for all adversaries $\adversary_1,\adversary_2$
+  and memories $\store \in \mems((S \cup O)_H)$:
+    $$
+    \condd{\progtt(\prog,\adversary_1)}{\houtputs}{\store} =
+    \condd{\progtt(\prog,\adversary_2)}{\houtputs}{\store}
+    $$
 \end{definition}
 
+\begin{lemma}
+  \label{lemma-integrity}
+  If a protocol $\prog$ with $\iov(\prog) = S \cup V \cup O$ has
+  integrity, then for all adversaries $\adversary$
+  and memories $\store \in \mems((S \cup O)_H)$:
+  $$
+  \condd{\progtt(\prog,\adversary)}{\houtputs}{\store} =
+  \condd{\progtt(\prog)}{\houtputs}{\store}
+  $$
+\end{lemma}
+
 \begin{theorem}
-  If $\prog$ satisfies active confidentiality for all $H,C$ then it is malicious secure.
+  If a protocol satisfies noninterference modulo output with
+  integrity, then it is malicious secure.
 \end{theorem}
 
+\begin{proof}
+  Let $\prog$ be some protocol with passive security and integrity
+  where $\iov(\prog) = S \cup V \cup O$, and let $\adversary$ be some
+  adversary.  Given some $\store\in S_H$, by definition of malicious
+  correctness the simulator can extract inputs $\store' \in (S_C)$
+  from $\adversary$ where $\idealf(\store \cup \store')$ is equivalent
+  to the real world outputs. These can then be used in a simulated run
+  of $\prog$ along with $\store \in S_H$ to generate a distribution
+  of corrupt views of $\margd{\progtt(\prog)}{\houtputs}$.
+  By assumption of integrity and Lemma \ref{lemma-integrity},
+  this is the same as the real-world distribution of corrupt views
+  in the presence of $\adversary$. 
+\end{proof}
+
 \begin{definition}
-  A protocol $\prog$ with has \emph{integrity} iff 
-  $\forall \adversary . \runs_\adversary(\prog) \subseteq \runs(\prog)$.
+  An adversary $\adversary$ is \emph{passive} if it runs $\prog$ honestly,
+  except for possible substitution of corrupt input secrets.
 \end{definition}
 
-\begin{theorem}
-  If a protocol has integrity it is malicious correct.
-\end{theorem}
-
-\begin{theorem}
-  If a protocol is passive secure with integrity, then it satisfies active confidentiality.
-\end{theorem}
-
-\begin{theorem}
-  If a protocol is passive secure with integrity, then it is malicious secure.
-\end{theorem}
+\begin{lemma}
+  Protocols with only passive adversaries have integrity.
+\end{lemma}
 

### 2024-05-04 revision-1/model.tex
checkpoint
first folder commit

--- 2024-04-20 revision-1/model.tex
@@
 %\end{eqnarray*}
+
+We also define the \emph{support} of a distribution in the usual manner-
+it is the set of values a set of variables can take on with non-zero
+probability.
+\begin{definition}[Support]
+  $\support(\pmf) \defeq \{ (v_1,\ldots,v_n) \mid
+  \pmf(x_1 \mapsto v_1, \ldots, x_n \mapsto v_n) > 0 \} $
+\end{definition}
 
@@
 \begin{definition}
-  Given $C$, $\store$, and $v$, we write $\dist(\SIM(\store,v))$ to
+  Given $\store$, and $v$, we write $\dist(\SIM(\store,v))$ to
   denote the distribution of corrupt views reconstructed by the
@@
 
+In the malicious model we assume that corrupt clients are in the thrall of an
+adversary $\adversary$ who does not necessarily follow the rules of the protocol.
+We model this by positing a $\arewrite$ function which is given a corrupt memory
+$\store_C$ and expression $\be$, and returns a rewritten expression that can
+be interpreted to yield a corrupt input. Then we define an evaluation
+relation that incorporates the adversary:
 $$
 \begin{array}{rclr}
-  (\store, \eassign{\mesg{w}}{\cid_1}{\be}{\cid_2};\prog) &\aredx&
-  (\extend{\store}{\mesg{w}_{\cid_1}}{\lcod{\store,\be}{\cid_2}}, \prog) & \cid_2 \in H\\
-  (\store, \eassign{\mesg{w}}{\cid_1}{\be}{\cid_2};\prog) &\aredx&
-  (\extend{\store}{\mesg{w}_{\cid_1}}{\lcod{\arewrite(\store_C,\be)}{\cid_2}}, \prog) & \cid_2 \in C
+  (\store, \xassign{x}{\be}{\cid};\prog) &\aredx&
+  (\extend{\store}{x}{\lcod{\store,\be}{\cid}}, \prog) & \cid \in H\\
+  (\store, \xassign{x}{\be}{\cid};\prog) &\aredx&
+  (\extend{\store}{x}{\lcod{\arewrite(\store_C,\be)}{\cid}}, \prog) & \cid \in C
 \end{array}
@@
 $$
-\dist(\SIM(\store_{S_H})) = \condd{\progtt(\prog)}{\houtputs \cup O_H}{\store_{S_H}}
+\begin{array}{rcl@{\qquad}r}
+  (\store,\elab{\assert{\phi(\be)}}{\cid};\prog) &\redx& (\store,\prog) & \text{if\ } \phi(\lcod{\store,\be}{\cid}) \text{\ or\ } \cid \in C\\
+  (\store,\elab{\assert{\phi(\be)}}{\cid};\prog) &\redx& (\store,\bot)  & \text{if\ } \neg\phi(\store,\lcod{\store,\be}{\cid})
+\end{array}
 $$
+
+Given this we define the $\adversary$ distribution of $\prog$. 
+\begin{definition}
+  \label{def-progd}
+  \label{definition-progd}
+  Given program $\prog$ with $\iov(\prog) = S \cup V \cup O$ and $\flips(\prog) = R$, and
+  any assumed pre-processing predicate $\preproc$ on memories, define:
+  $$
+  \aruns(\prog) \defeq \{ \store \mid \preproc(\store)  \wedge
+  ((\store_{S\cup R},\prog) \aredx^* (\store,\varnothing) \vee (\store_{S \cup R},\prog) \aredx^* (\store,\bot)) \}
+  $$
+  and also define:
+  $$
+  \botruns(\prog) \defeq \{ \store\{ x_1 \mapsto \bot, \ldots, x_n \mapsto \bot \} \mid
+  \store \in \aruns(\prog) \wedge \{ x_1,\ldots,x_n\} = (V \cup O) - \dom(\store) \}
+  $$
+  Then the \emph{$\adversary$ distribution of $\prog$}, written $\progtt(\prog,\adversary)$, is
+  defined such that for all $\store \in \mems(\iov(\prog) \cup R)$:
+  $$
+  \progtt(\prog,\adversary) =  1 / |\botruns(\prog)| \ \text{if}\ \store \in \botruns(\prog), \text{otherwise}\ 0
+  $$
+\end{definition}
+
+\begin{definition}[Malicious Simulation]
+  Given a protocol $\prog$ with $\iov(\prog) = S \cup V \cup O$, honest and corrupt 
+  clients $H$ and $C$, adversary $\adversary$, and honest inputs
+  $\store \in \mems(S_H)$, the \emph{malicious simulation}  $\SIM(\store)$ has three phases:
+  \begin{enumerate}
+  \item In the first phase $\SIM_1$, $\adversary$ gives the
+    simulator some $\store' \in \mems(S_C)$, and the simulator consults an
+    oracle to compute $\idealf(\store \cup \store') \in \mems(O)$.
+  \item In the second phase $\SIM_2$, the simulator is given the corrupt
+    outputs $\idealf(\store \cup \store')_C$, which are again given to
+    $\adversary$, who decides either to abort or not. If so, then the
+    simulator is given $\sigma_{\mathit{out}} \defeq \{ \out{\cid} \mapsto \bot \mid \cid \in H \}$
+    and arbitrary internal state $\Sigma$.
+    Otherwise the simulator is given $\sigma_{\mathit{out}} \defeq \idealf(\store \cup \store')_H$
+    and $\Sigma$.
+  \item In the third phase $\SIM_3$, given $\store_{\mathit{out}}$ and $\Sigma$, the simulator
+    finally outputs
+    $\store_{\mathit{out}} \cup \store_{\mathit{views}}$ for some
+    calculated $\store_{\mathit{views}} \in \mems(\houtputs)$.
+  \end{enumerate}
+\end{definition}
+
+\begin{definition}[Malicious Security]
+  We write $\dist(\SIM(\store))$ to
+  denote the distribution of honest outputs and corrupt views reconstructed by the
+  malicious simulation, where for
+  all $\store'$:
+  $$
+  \dist(\SIM(\store))(\store')\ \defeq\ \prob(\SIM(\store) = \store') 
+  $$
+  Then a protocol $\prog$ with $\iov(\prog) = S \cup V \cup O$ is \emph{malicious
+  secure} iff for all $H$, $C$, $\adversary$, and $\store \in \mems(S_H)$:
+  $$
+  \dist(\SIM(\store)) = \condd{\progtt(\prog,\adversary)}{\houtputs \cup O_H}{\store}
+  $$  
+\end{definition}

### 2024-05-04 ppdp24/main.tex
first folder commit
checkpoint

--- 2023-07-11 popl24/main.tex
@@
-\documentclass[acmsmall,screen,review]{acmart}
+\documentclass[sigconf]{acmart}
 
@@
 \usepackage{comment}
+\usepackage{booktabs,array}
 
 \input{macros}
+\input{figures}
 
@@
 
-\title{Automating Verification of MPC Security: $\metaprot$ to $\minifed$}
+\title{Language-Based Low-Level MPC Security: $\metaprot$ to $\minifed$}
 
-\author{Author names withheld for double-blind reviewing}
+\author{Christian Skalka}
+\affiliation{
+  \institution{University of Vermont}
+  \city{}
+  \country{}
+%  \city{Burlington}
+%  \country{USA}
+}
+\email{ceskalka@uvm.edu}
+
+\author{Joseph Near}
+\affiliation{
+  \institution{University of Vermont}
+  \city{}
+  \country{}
+%  \city{Burlington}
+%  \country{USA}
+}
+\email{jnear@uvm.edu}
 
@@
 %% the work being presented. Separate the keywords with commas.
-\keywords{Multiparty computation, program verification, probabilistic programming}
+\keywords{Secure multiparty computation, security verification, probabilistic programming, static analysis.}
 
@@
 
-\input{intro}
+\input{lang}
 
-\input{minicat}
+\input{model}
 
-\input{metalang}
-
-\input{pmf}
-
-\input{hyperprop}
-
-\input{automation}
-
-\input{composition}
-
-%% This is scratch pad to develop ideas before integration.
-%\input{scratch}
+\input{hyper}
 
@@
 
-\appendix
-
-\input{examples}
-
-%\input{types}
-
-\input{ygc}
-
-\input{bruteforce}
-
-\input{proofs}
-
 \end{document}

### 2024-05-04 revision-1/model.tex
checkpoint

@@
     finally outputs
-    $\store_{\mathit{out}} \cup \store_{\mathit{views}}$ for some
+    $\store_{\mathit{out}} \cap \store_{\mathit{views}}$ for some
     calculated $\store_{\mathit{views}} \in \mems(\houtputs)$.

### 2024-05-04 ppdp24/hyper.tex
checkpoint

--- 2024-04-21 revision-1/hyper.tex
@@
 
-\subsection{Gradual Release as a Design Pattern}
+\subsection{Delimited Release}
 
@@
 
-In fact, MPC protocols typically satisfy a \emph{gradual release}
+In fact, MPC protocols typically satisfy a \emph{delimited release}
 property\cite{XXX}, where messages exchanged remain probabilistically separable
 from secrets, with only declassification events (reveals and outputs)
-releasing information about honest secrets. 
+releasing information about honest secrets. A key difference is that while
+these declassification events essentially define the policy in delimited
+release, the ideal functionality sets the release policy for MPC
+passive security, so its necessary to show that declassification events respect
+these bounds.
 \begin{definition}
   Given $H,C$, a protocol $\prog$ with $\iov(\prog) = S \cup M \cup P \cup O$
-  satisfies \emph{gradual release} iff
+  satisfies \emph{delimited release} iff
   $\sep{\progtt(\prog)}{M_C}{S_H}$.
@@
   $
-  \forall \adversary, \store \in \mems(S_H) \ .\ \support(\progtt(\prog)(O_H|\store)) =
+  \forall \adversary, \store \in \mems(S_H) \ .\ \support(\progtt(\prog)(O_H|\store)) \supseteq
     \support(\progtt(\prog,\adversary)(O_H|\store))
   $.
+\end{definition}
+
+\begin{definition}[Integrity]
+  We say that a protocol $\prog$ with $\iov(\prog) = S \cup V \cup O$ has
+  \emph{integrity} iff for all $H$ and $C$ 
+  it is malicious correct and for all adversaries $\adversary$,
+  if $\store \in \aruns(\prog)$ 
+  then there exists $\store' \in \mems(S)$ with $\store_{S_H} = \store'_{S_H} $ and:
+    $$
+    \condd{\progtt(\prog,\adversary)}{X}{\store_{S_H \cup \cinputs}} =
+    \condd{\progtt(\prog)}{X}{\store'}
+    $$ 
+  where $X \defeq (\houtputs \cup O_H) \cap \dom(\store)$. 
 \end{definition}

### 2024-05-04 ppdp24/lang.tex
checkpoint

--- 2024-04-21 revision-1/lang.tex
@@
 
-\begin{fpfig}[t]{Top-to-bottom: Basic $\minifed$ syntax, expression interpretation, and reduction rules.}{fig-minifed}
-  {
-    $$
-    \begin{array}{rcl@{\hspace{8mm}}r}
-      \multicolumn{4}{l}{v \in \mathbb{Z}_p,\ w \in \mathrm{String},\ \cid \in \mathrm{Clients} \subset  \mathbb{N} }\\[2mm] %, \bop \in \{ \eand, \eor, \exor \}} \\[2mm]
-      \be &::=& v \mid \flip{w} \mid \secret{w} \mid \mesg{w} \mid \rvl{w} \mid \be \fminus \be \mid \be \fplus \be \mid \be \ftimes \be \mid f \mid \be\,\be & \textit{expressions}\\[2mm]
-      x &::=& \elab{\flip{w}}{\cid} \mid \elab{\secret{w}}{\cid} \mid \elab{\mesg{w}}{\cid} \mid \rvl{w} \mid \out{\cid} & \textit{protocl variables} \\[2mm]
-      %& &  \select{\be}{\be}{\be} \mid \ctxt{v}{k} \mid \key{w} \mid \sk{\be}(\be) \mid \pk{\be}{\be}(\be) \mid \pk{\be}{\be} \\[2mm]
-      %& &  \select{\fp(\be)}{\be}{\be} \ctxt{v,\be}{k}  \mid \sk{\be}(\be) \mid \pk{\be}{\be}(\be) \mid \pk{\be}{\be} \\[2mm]
-      \instr &::=& \eassign{\mesg{w}}{\cid}{\be}{\cid} \mid
-      \reveal{w}{e}{\cid} \mid \pubout{\cid}{\be}{\cid} & \textit{commands} \\[2mm]
-      \prog &::=& \varnothing \mid \instr; \prog & \textit{protocols}
-    \end{array}
-    $$
-
-  \rule{130mm}{0.5pt}
-
-  $$
-  \begin{array}{c@{\hspace{5mm}}c}
-  \begin{array}{rcl}
-    \lcod{\store, v}{\cid} &=& v\\
-    \lcod{\store, \be_1 \fplus \be_2}{\cid} &=& \fcod{\lcod{\store, \be_1}{\cid} \fplus \lcod{\store, \be_2}{\cid}}\\ 
-    \lcod{\store, \be_1 \fminus \be_2}{\cid} &=& \fcod{\lcod{\store, \be_1}{\cid} \fminus \lcod{\store, \be_2}{\cid}}\\ 
-    \lcod{\store, \be_1 \ftimes \be_2}{\cid} &=& \fcod{\lcod{\store, \be_1}{\cid} \ftimes \lcod{\store, \be_2}{\cid}}
-  \end{array} & 
-  \begin{array}{rcl}
-    \lcod{\store, \flip{w}}{\cid} &=& \store(\elab{\flip{w}}{\cid})\\
-    \lcod{\store, \secret{w}}{\cid} &=& \store(\elab{\secret{w}}{\cid})\\
-    \lcod{\store, \mesg{w}}{\cid} &=& \store(\elab{\mesg{w}}{\cid})\\
-    \lcod{\store, \rvl{w}}{\cid} &=& \store(\rvl{w})\\
-    \lcod{\store, f\,\be_1\,\cdots\, \be_n}{\cid} &=& \delta(f,\lcod{\store, \be_1}{\cid},\ldots,\lcod{\store,\be_n}{\cid})
-  \end{array}
-  \end{array}
-  $$
-
-  \vspace{4mm}
-
-  \rule{130mm}{0.5pt}
-
-  \begin{mathpar}
-    (\store, \eassign{\mesg{w}}{\cid_1}{\be}{\cid_2};\prog) \redx (\extend{\store}{\mesg{w}_{\cid_1}}{\lcod{\store,\be}{\cid_2}}, \prog)
-
-    (\store, \reveal{w}{\be}{\cid};\prog) \redx (\extend{\store}{\rvl{w}}{\lcod{\store,\be}{\cid}}, \prog)
-
-    (\store, \pubout{\cid}{\be}{\cid};\prog) \redx (\extend{\store}{\out{\cid}}{\lcod{\store,\be}{\cid}}, \prog)
-  \end{mathpar}
-  }
-\end{fpfig}
+\minifedfig
 
@@
 operations of addition, subtraction, and multiplication in
-a finite field $\mathbb{Z}_p$ with $p$ prime. 
+a finite field $\mathbb{F}_p$ where $p$ is some prime. 
 Protocols are given input secret values $\secret{w}$
@@
 subsets belonging to honest and corrupt parties, respectively.
-
-\subsubsection{Library Functions} $\minifed$ expression syntax also supports
-calls to library functions $f$ which can be applied to muliple arguments in a
-curried style. This allows encapsulation and separate
-definition of primitive operations such as one-time-pads and message
-authentication, as we will illustrate with examples. This approach is
-useful since it parameterizes these definitions, and 
-supports verification of behavior specified with types, as we
-discuss in Section \ref{section-types}.
 
@@
 is in a uniform random distribution with other variables in a program,
-then $v_1 \fplus v_2$ is an encryption of $v_2$ where $v_1$ is an
-information theoretically secure one-time-pad, which is exploited for
-secret sharing. Of course, $\fplus$ is also a meaningful operation
-over any two field values regardless of their distributions.
-
-To capture this distinction we introduce a function $\otp$
-with the following specification:
-$$
-\delta(\otp,v_1,v_2) \defeq \fcod{v_1 \fplus v_2}
-$$
-Although the semantics are the same as addition, the use of $\otp$
-makes a declarative distinction. But more importantly we will be
-able to assign a type to $\otp$ that enforces the one-time discipline
-on its first argument via type linearity, and reflects hiding of
-its second argument via security type labels, as will be discussed in Section
-\ref{section-types}.
+then either  $v_1 \fplus v_2$ or $v_1 \fminus v_2$  is an encryption of $v_2$
+where $v_1$ is an information theoretically secure one-time-pad, which
+is exploited for secret sharing. Of course, $\fplus$ and $\fminus$ are also
+meaningful operations over any two field values regardless of their distributions.
 
@@
 the federation $\{ 1, 2, 3 \}$  samples a value $\locflip$
-that can be used as a one-time pad in summation with another
+that can be used as a one-time pad with another
 random sample $\flip{x}$ and $\secret{\cid}$. This yields
@@
 \begin{array}{lll}
-  \elab{\mesg{s1}}{2} &:=& \elab{(\otp\ \locflip\ (\flip{x} \fplus \secret{1})}{1} \\ 
+  \elab{\mesg{s1}}{2} &:=& \elab{(\locflip \fminus \flip{x} \fminus \secret{1})}{1} \\ 
   \elab{\mesg{s1}}{3} &:=& \elab{\flip{x}}{1} \\ 
-  \elab{\mesg{s2}}{1} &:=& \elab{(\otp\ \locflip\ (\flip{x} \fplus \secret{2})}{2} \\ 
+  \elab{\mesg{s2}}{1} &:=& \elab{(\locflip \fminus \flip{x} \fminus \secret{2})}{2} \\ 
   \elab{\mesg{s2}}{3} &:=& \elab{\flip{x}}{2} \\ 
-  \elab{\mesg{s3}}{1} &:=& \elab{(\otp\ \locflip\ (\flip{x} \fplus \secret{3})}{3} \\ 
+  \elab{\mesg{s3}}{1} &:=& \elab{(\locflip \fminus \flip{x} \fminus \secret{3})}{3} \\ 
   \elab{\mesg{s3}}{2} &:=& \elab{\flip{x}}{3}
@@
 $$
-Due to field properties of $\fplus$ this scheme guarantees that messages
+Due to field properties of $\fminus$ this scheme guarantees that messages
 are viewed as random noise by any observer 
@@
 publicly reveals the sum of all of its shares, including its local
-share. This does reveal information about secrets. Further there
-is no one-time-pad to use in this summation.
+share. This step does reveal information about secrets-- note in
+particular that $\locflip$ is reused and is no longer a one-time-pad:
 $$
@@
 Finally, each client outputs the sum of each sum of shares, yielding
-the sum of secrets. Note that this stage exposes no more information
-than the previous public reveals. 
+the sum of secrets. The protocol is correct because the outputs are all the
+true sum of secrets, and it is secure because no more information about the
+secrets other than that revealed by their sum is exposed.
 $$
@@
 $$
-It is well-known that additive secret sharing is passive
-secure. That is, any adversarial observer can gain no more information
-from the messages exchanged in the protocol than what is exposed by
-the output alone. However, malicious adversaries can corrupt this
-protocol by injecting ``fake'' sums of shares in their public reveals.
-
-\subsection{Example: Malicious Secure Product}
-
-\begin{fpfig}[t]{Top-to-bottom: Extended $\minifed$ syntax with BDOZ-style shares,
-    interpretation of operations, and pre-processing properties of memories.}{fig-bdoz}
-  {
-$$
-\begin{array}{rcl@{\hspace{8mm}}r}
-  \multicolumn{3}{l}{m,k \in \mathbb{Z}_p \qquad \macv ::= (v,m)} &\\[2mm]
-  %\textit{MACed\ values}\\[2mm]
-  \be &::=& \cdots \mid \macv \mid \be \macplus \be \mid \be \mactimes \be \mid \be \macminus \be\\[2mm]
-  x &::=& \cdots \mid \macx{\secret{w}}{\cid} \mid \macx{\flip{w}}{\cid} \mid \mack{x}{}
-\end{array}
-$$
-
-\rule{130mm}{0.5pt}
-\smallskip
-$$
-\begin{array}{rcl}
-  \fcod{(v_1, m_1) \macplus (v_2, m_2)} &=& (v_1 \fplus v_2, m_1 \fplus m_2)\\[2mm]
-  \fcod{(v, m) \mactimes v'} &=& (v \ftimes v', m \ftimes v')
-\end{array}
-$$
-$$
-\begin{array}{rcl}
-  \lcod{\store, \be_1 \macplus \be_2}{\cid} &=&
-  \fcod{\lcod{\store,\be_1}{\cid} \macplus \lcod{\store,\be_2}{\cid}}\\[2mm]  
-  \lcod{\store, \be_1 \mactimes \be_2}{\cid} &=&
-  \fcod{\lcod{\store,\be_1}{\cid} \mactimes \lcod{\store,\be_2}{\cid}}
-\end{array}
-$$
-
-\medskip
-
-\rule{130mm}{0.5pt}
-
-\begin{mathpar}
-  \inferrule
-      {\store(\macx{x}{\cid_1}) = (v_1,m_1) \\
-        \store(\macx{x}{\cid_2}) = (v_2,m_2)}
-      {\store(x) = \fcod{v_1 \fplus v_2}}
-
-  \inferrule
-      {\store(\elab{x}{\cid'}) = (v,m) \\ \cid \ne \cid'}
-      {m = \lcod{\store, \mack{x}{} \fplus \flip{\ttt{delta}} \ftimes v}{\cid}}
-
-  \store(\elab{\secret{c}}{\Oracle}) = \lcod{\store,\flip{a} \ftimes \flip{b}}{\Oracle}
-\end{mathpar}
-}
-\end{fpfig}
-
-Eliminating cheating by adding information theoretic secure MAC codes
-to shares is a fundamental approach realized by protocols such as BDOZ
-\cite{XXX} and SPDZ \cite{XXX}.  These protocols assume a
-pre-processing phase that distributes shares with MAC codes to
-clients.  This integrates well with pre-processed Beaver Triples to
-implement malicious secure, and relatively efficient, multiplication
-\cite{XXX}.  Here we show how to implement malicious secure two party
-multiplication in the BDOZ style via some extensions of $\minifed$.
-
-A field value $v$ is secret shared among 2 clientz in BDOZ as follows.
-Each client $\cid$ gets a pair of the form $(v_\cid,m_\cid)$, where
-$v_\cid$ is a secret share of $v$, i.e., $v = \fcod{v_1 \fplus v_2}$.
-Further, field values $m_\cid$ are MACs of $v_\cid$ that are authenticated
-by the other client's key. In particular, $m_\cid = k + (k_\Delta *
-v_\cid)$ where \emph{keys} $k$ and $k_\Delta$ are known only to $\cid'
-\ne \cid$ and $k_\Delta$. Further, there is a unique \emph{local key}
-$k$ for each shared value, while the \emph{global key} $k_\Delta$ is
-common to all MACs authenticated by $\cid'$. By extending certain
-field operations to these MACed values, a semi-homorphic encryption
-scheme is obtained that is adequate to compute Beaver Triples. For
-example, if local keys $k_1$ and $k_2$ authenticate $\macv_1$ and
-$\macv_2$, then $k_1 \fplus k_2$ authenticates $\macv_1 \macplus
-\macv_2$, where $\macplus$ is addition extended to MACed
-values. Semantics for summation of MACed shares, and multipication of
-shares and plain field values, are given in Figure \ref{fig-bdoz}.
-For more details the reader is referred to \cite{XXX}. We note
-that while we restrict values $v$, $m$, and $k$ to the same
-field in this presentation for simplicity, in general $m$ and
-$k$ can be in extensions of $\mathbb{Z}_p$.
-
-We define variable forms $\macx{\secret{w}}{\cid}$ and
-$\macx{\flip{w}}{\cid}$ to range over local MACed shares of a secret
-and random sample owned by client $\cid$, respectively. We let
-$\mack{x}{}$ denote the local key for authenticating shares of $x$.
-We also assume a global key $\flip{\ttt{delta}}$ on each client.
-Pre-processing is modeled by assuming that shares are provided on
-input random tapes (memories) with the required properties, as defined
-on the bottom of Figure \ref{fig-bdoz}.
-
-Similarly to field values, we can use MACed shares in
-uniform independent distributions as one-time-pads for
-hiding other values, using subtraction over MACed shares.
-So we introduce a function $\macotp$ to do this
-declaratively.
-$$
-\delta(\macotp,v_1,v_2) \defeq \fcod{v_1 \macminus v_2}
-$$
-
-Another key idea in BDOZ is \emph{secure opening}, where distributed
-shares of a value $v$ are communicated and combined to ``open'' the
-same $v$ on each client. This relies on authentication to detect
-cheating in the malicious setting. To implement secure opening we
-define a $\macauth$ function that authenticates shares, and a
-$\macopen$ function that extracts a field value from its shares.
-\begin{mathpar}
-\delta(\macauth, (v, m), k) \defeq
-     (v, m) \text{\ if\ } m = k \fplus (\flip{\ttt{delta}} \ftimes v)
-
-\delta(\macopen, (v_1, m_1), (v_2,m_2)) \defeq
-\fcod{v_1 \fplus v_2}
-\end{mathpar}
-This authentication eliminates cheating and guarantees protocol
-integrity. As we will show in Section \ref{section-types}, this
-can be reflected in a security type for $\macauth$.
-
-Putting these pieces together, we can implement 2-party multiplication with
-Beaver Triples as follows. Recall that Beaver Triples are values
-$(a,b,c)$ such that $c = \fcod{b \ftimes a}$- we implement them here
-as a secret $\secret{c}$ and samples $\flip{a}$ and $\flip{b}$ owned
-by an \emph{oracle} $\Oracle$ that we always assume is honest but is
-otherwise not involved in the protocol. We define pre-processing
-properties of Beaver Triples in Figure \ref{fig-bdoz}. Only one
-Beaver triple can be used per multiplication to ensure secure
-openings, but our scheme can be extended with more Beaver triples
-and reuse of them will be rule out by type linearity as we will
-discuss in Section \ref{section-types}.
-
-First, clients 1 and 2 compute (aka \emph{open}) $\secret{x} \fplus
-\flip{a}$ and $\secret{y} \fplus \flip{b}$ where $\secret{x}$ and
-$\secret{y}$ are 1 and 2's secrets respectively.  Then they both
-perform a secure opening of $\secret{x} \fplus \flip{a}$ and
-$\secret{y} \fplus \flip{b}$, by authenticating each others' shares,
-and record them as $\mesg{d}$ and $\mesg{e}$ respectively. Note
-that the secrets remain hidden since $\flip{a}$ and $\flip{b}$
-are random values.
-$$
-\begin{array}{lcl}
-  \elab{\mesg{a}}{2} &:=&
-  \elab{(\macotp\ \macgv{\elab{\flip{a}}{\Oracle}}\ \macgv{\elab{\secret{x}}{1}})}{1}\\
-  \elab{\mesg{a}}{1} &:=&
-  \elab{(\macotp\ \macgv{\elab{\flip{a}}{\Oracle}}\ \macgv{\elab{\secret{x}}{1}})}{2}\\
-  \elab{\mesg{b}}{2} &:=&
-  \elab{(\macotp\ \macgv{\elab{\flip{b}}{\Oracle}}\ \macgv{\elab{\secret{y}}{2}})}{1}\\
-  \elab{\mesg{b}}{1} &:=&
-  \elab{(\macotp\ \macgv{\elab{\flip{b}}{\Oracle}}\ \macgv{\elab{\secret{y}}{2}})}{2}\\
-  \elab{\mesg{d}}{1} &:=&
-  \elab{(\macopen\ \macauth(\mesg{a}, \mack{\macx{\secret{x}}{1}}{} \fminus \mack{\macx{\flip{a}}{\Oracle}}{}) \ (\macgv{\elab{\secret{x}}{1}}\macminus\macgv{\elab{\flip{a}}{\Oracle}}))}{1}\\
-  \elab{\mesg{e}}{1}&:=&
-  \elab{(\macopen\ \macauth(\mesg{b}, \mack{\macx{\secret{y}}{2}}{} \fminus \mack{\macx{\flip{b}}{\Oracle}}{}) \ (\macgv{\elab{\secret{y}}{2}}\macminus\macgv{\elab{\flip{b}}{\Oracle}}))}{1}\\
-  \elab{\mesg{d}}{2} &:=&
-  \elab{(\macopen\ \macauth(\mesg{a}, \mack{\macx{\secret{x}}{1}}{} \fminus \mack{\macx{\flip{a}}{\Oracle}}{}) \ (\macgv{\elab{\secret{x}}{1}}\macminus\macgv{\elab{\flip{a}}{\Oracle}}))}{2}\\
-  \elab{\mesg{e}}{2}&:=&
-  \elab{(\macopen\ \macauth(\mesg{b}, \mack{\macx{\secret{y}}{2}}{} \fminus \mack{\macx{\flip{b}}{\Oracle}}{}) \ (\macgv{\elab{\secret{y}}{2}}\macminus\macgv{\elab{\flip{b}}{\Oracle}}))}{2}
-\end{array}
-$$
-Now, each client uses their share of $\secret{c}$ to generate their
-share of the final output, which is again reported via a secure opening.
-Homomorphic key calculations are elided for brevity.
-$$
-\begin{array}{lcl}
-  \rvl{1} &:=&
-  \elab{( (\mesg{d} \ftimes \mesg{e}) \macplus
-          (\mesg{d} \mactimes \macgv{\elab{\flip{b}}{\Oracle}}) \macplus
-          (\mesg{e} \mactimes \macgv{\elab{\flip{a}}{\Oracle}}) \macplus \macgv{\elab{\secret{c}}{\Oracle}}
-    )}{1}\\
-  \rvl{2} &:=&
-  \elab{( (\mesg{d} \ftimes \mesg{e}) \macplus
-          (\mesg{d} \mactimes \macgv{\elab{\flip{b}}{\Oracle}}) \macplus
-          (\mesg{e} \mactimes \macgv{\elab{\flip{a}}{\Oracle}}) \macplus \macgv{\elab{\secret{c}}{\Oracle}}
-    )}{2}\\
-  \out{1} &:=& \elab{(\macopen\ \rvl{1} \ \macauth(\rvl{2},\ldots))}{1} \\
-  \out{2} &:=& \elab{(\macopen\ \macauth(\rvl{1},\ldots) \ \rvl{2})}{2} 
-\end{array}
-$$
-Some information about input secrets is revealed by the product, but
-secure opening ensures high integrity and hence robustness to
-corruption, aka robust declassification.

### 2024-05-04 revision-1/hyper.tex
checkpoint

@@
   $
-  \forall \adversary, \store \in \mems(S_H) \ .\ \support(\progtt(\prog)(O_H|\store)) =
+  \forall \adversary, \store \in \mems(S_H) \ .\ \support(\progtt(\prog)(O_H|\store)) \supseteq
     \support(\progtt(\prog,\adversary)(O_H|\store))
@@
   \emph{integrity} iff for all $H$ and $C$ 
-  it is malicious correct and for all adversaries $\adversary_1,\adversary_2$
-  and memories $\store \in \mems((S \cup O)_H)$:
+  it is malicious correct and for all adversaries $\adversary$,
+  if $\store \in \aruns(\prog)$ 
+  then there exists $\store' \in \mems(S)$ with $\store_{S_H} = \store'_{S_H} $ and:
     $$
-    \condd{\progtt(\prog,\adversary_1)}{\houtputs}{\store} =
-    \condd{\progtt(\prog,\adversary_2)}{\houtputs}{\store}
+    \condd{\progtt(\prog,\adversary)}{X}{\store_{S_H \cup \cinputs}} =
+    \condd{\progtt(\prog)}{X}{\store'}
     $$ 
+  where $X \defeq (\houtputs \cup O_H) \cap \dom(\store)$. 
 \end{definition}
 
-\begin{lemma}
-  \label{lemma-integrity}
-  If a protocol $\prog$ with $\iov(\prog) = S \cup V \cup O$ has
-  integrity, then for all adversaries $\adversary$
-  and memories $\store \in \mems((S \cup O)_H)$:
-  $$
-  \condd{\progtt(\prog,\adversary)}{\houtputs}{\store} =
-  \condd{\progtt(\prog)}{\houtputs}{\store}
-  $$
-\end{lemma}
-
 \begin{theorem}
-  If a protocol satisfies noninterference modulo output with
+  If a protocol satisfies noninterference modulo output and has 
   integrity, then it is malicious secure.
@@
 
-\begin{definition}
-  An adversary $\adversary$ is \emph{passive} if it runs $\prog$ honestly,
-  except for possible substitution of corrupt input secrets.
-\end{definition}
-
-\begin{lemma}
-  Protocols with only passive adversaries have integrity.
-\end{lemma}
-
 \begin{definition}[Robust Declassification]

### 2024-05-05 ppdp24/hyper.tex
checkpoint

@@
 \end{definition}
-A soundness result for passive security can then be stated as follows.
-We prove this in a separate manuscript \cite{XXX}, and it is also
-essentially the same as ``perfect passive NI security'' which
-has a similar soundness property \cite{8429300}.  
+A soundness result for passive security can then be given as follows.
+It is essentially the same as ``perfect passive NI security'' explored
+in previous work \cite{8429300}.  
 \begin{restatable}{theorem}{nimosecure}
@@
 \end{restatable}
+
+\begin{proof}
+  Let $H$ and $C$ be an arbitrary partition of the federation and
+  suppose $\prog$ satisfies noninterference modulo output. Let
+  $\iov(\prog) = (V,S)$ and $\flips(\prog) = F$ with output view $\outv
+  \in V$ and let $\store$ be an arbitrary element of $\mems(S \cup
+  F)$. Then the distribution of adversarial views in the real world
+  is, by definition,
+  $\condd{\progd(\prog)}{V_C}{\store_S}$.
+
+  The simulator is given both $\store_{S_C}$ and
+  $\idealf(\store_S)$.  The simulation $\SIM_C(\store_{S_C},
+  \idealf(\store_S))$ can be defined as follows. First, some $\store'
+  \in \ik(\idealf,\idealf(\store))$ is randomly chosen such that
+  $\store' =_C \store_S$, as is a random tape $\store'' \in
+  \mems(F)$\footnote{The real/ideal model allows consultation of a
+  ``Random Oracle''.}. Then, the run of $(\store' \cup \store'',
+  \prog)$ is evaluated in simulation, yielding $(\store^{\SIM},\varnothing)$.
+  The simulation returns $\store^{\SIM}_{V_C}$ as a result.
+
+  Now, since the random tape is selected in simulation from the same distribution
+  that we assume for the real world, after selection of $\store'$ the
+  probability that any particular $\store^{\SIM}_{V_C}$ is returned is by definition
+  $
+  \progd(\prog)(\store^{\SIM}_{V_C} | \store')
+  $.
+  Furthermore, since we assume that $\prog$ correctly implements $\idealf$, this
+  means
+  $
+  \store_S =_C \store' \wedge
+     \condd{\progd(\prog)}{\{ \outv \}}{\store_S} =
+     \condd{\progd(\prog)}{\{ \outv \}}{\store'}
+  $,
+  so by the definition of noninterference modulo output,
+  any choice of $\store'$ yields the same distribution $\condd{\progd(\prog)}{V_C}{\store_S}$.
+  Therefore by definition
+  $
+   \progd(\SIM_C(\store_{S_C},\idealf(\store))) = \condd{\progd(\prog)}{V_C}{\store_S}
+  $.
+\end{proof}
 
@@
 
-\begin{definition}[Integrity]
-  We say that a protocol $\prog$ with $\iov(\prog) = S \cup V \cup O$ has
-  \emph{integrity} iff for all $H$ and $C$ 
-  it is malicious correct and for all adversaries $\adversary_1,\adversary_2$
-  and memories $\store \in \mems((S \cup O)_H)$:
-    $$
-    \condd{\progtt(\prog,\adversary_1)}{\houtputs}{\store} =
-    \condd{\progtt(\prog,\adversary_2)}{\houtputs}{\store}
-    $$
-\end{definition}
-
-\begin{lemma}
-  \label{lemma-integrity}
-  If a protocol $\prog$ with $\iov(\prog) = S \cup V \cup O$ has
-  integrity, then for all adversaries $\adversary$
-  and memories $\store \in \mems((S \cup O)_H)$:
-  $$
-  \condd{\progtt(\prog,\adversary)}{\houtputs}{\store} =
-  \condd{\progtt(\prog)}{\houtputs}{\store}
-  $$
-\end{lemma}
-
 \begin{theorem}
-  If a protocol satisfies noninterference modulo output with
+  If a protocol satisfies noninterference modulo output and has
   integrity, then it is malicious secure.
@@
 \begin{theorem}
-  Passive security with robust declassification implies malicious security.
+  Robust declassification does not imply malicious security, but
+  passive security with robust declassification implies malicious security.
 \end{theorem}

### 2024-05-05 ppdp24/model.tex
checkpoint

--- 2024-05-04 revision-1/model.tex
@@
 
-\emph{Simulator security} in the MPC setting means that protocols do
-not reveal any secret information other than what is implicitly
-declassified by the output of the ideal functionality that the
-protocol implements. Characterization of both real world protocol
-execution and simulation is defined probabilistically. Following
-previous work \cite{barthe2019probabilistic} we use probability mass
-functions to express joint dependencies between input and output
-variables, as a metric of information leakage.
+MPC protocols are intended to implement some \emph{ideal
+functionality} $\idealf$. Given a protocol $\prog$ that implements
+$\idealf$, with $\iov(\prog) = S \cup V \cup O$, we will assume
+that the domain of $\idealf$ is $\mems(S)$ and its range is
+$\mems(O)$. Recall the assumption of a partitioning of a federation
+into honest and corrupt subsets $H$ and $C$.  \emph{Simulator
+security} in the MPC setting means that, given $\store \in \mems(S)$,
+a secure protocol $\prog$ does not reveal any more information about
+honest secrets $\store_H$ to parties in $C$ beyond what is
+implicitly declassified by $\idealf(\sigma)$. Security comes in
+\emph{passive} and \emph{malicious} flavors, wherein the adversary
+either follows the rules or not, respectively. Characterization of
+both real world protocol execution and simulation is defined
+probabilistically. Following previous work
+\cite{barthe2019probabilistic} we use probability mass functions to
+express joint dependencies between input and output variables, as a
+metric of information leakage.
 
@@
 memories to denote mappings of variables to values (i.e., outcomes),
-so for example given a pmf $\pmf$ we will write $\pmf(\{ \sx{1}{a}
-\mapsto 0, \vx{2}{b} \mapsto 1 \})$ to denote the probability that
-$\sx{1}{a} = 0$ and $\vx{2}{b} = 1$ in the final memory of a
+so for example given a pmf $\pmf$ we will write $\pmf(\{ \elab{\secret{x}}{1}
+\mapsto 0, \elab{\mesg{y}}{2} \mapsto 1 \})$ to denote the (joint) probability that
+$\elab{\secret{x}}{1} = 0 \wedge \elab{\mesg{y}}{2} = 1$ in the final memory of a
 program containing these variables.
@@
 We consider random tapes to be selected from a uniform distribution
-modulo relevant pre-processing, e.g., Beaver triples and MACed share
-distributions.  We will also assume that any possible initial secrets
+modulo relevant pre-processing, e.g., Beaver triples for efficient
+multiplication, and/or MACed share distributions as in SPDZ \cite{XXX}.
+We will also assume that any possible initial secrets
 are equally likely when formulating the program distribution.
@@
   \label{definition-progd}
-  Given program $\prog$ with $\secrets(\prog) = S$ and $\flips(\prog) = R$, and
-  any assumed pre-processing predicate $\preproc$ on memories, define:
+  Given $\prog$ with $\secrets(\prog) = S$ and $\flips(\prog) = R$ and
+  pre-processing predicate $\preproc$ on memories, define:
   $$
   \runs(\prog) \defeq \{ \store \mid \preproc(\store) \wedge
-  (\dom(\store) = \iov(\prog) \cup R) \wedge (\store_{S \cup R},\prog) \redxs (\store,\varnothing) \}
+  %(\dom(\store) = \iov(\prog) \cup R) \wedge
+  (\store_{S \cup R},\prog) \redxs (\store,\varnothing) \}
   $$
@@
 
-\subsection{Passive Security}
+\subsection{Honest and Corrupt Views}
 
@@
   $$
-  \houtputs = \{\ \rvl{w} \mid\ \reveal{w}{\be}{\cid} \in \prog
-  \wedge \cid \in H \ \} 
-  \cup \{\ \elab{\mesg{w}}{\cid}\ \mid\  \eassign{\mesg{w}}{\cid}{\be}{\cid'} \in \prog
-  \wedge \cid \in C \wedge \cid' \in H \ \} 
+  \begin{array}{lcl}
+    \houtputs & \defeq
+        & \{\ \rvl{w} \mid\ \reveal{w}{\be}{\cid} \in \prog \wedge \cid \in H \ \}\ \cup \\
+      & & \{\ \elab{\mesg{w}}{\cid}\ \mid\  \eassign{\mesg{w}}{\cid}{\be}{\cid'} \in
+           \prog \wedge \cid \in C \wedge \cid' \in H \ \} 
+  \end{array}
   $$
@@
   $$
-  \cinputs = \{\ \rvl{w} \mid\ \reveal{w}{\be}{\cid} \in \prog
-  \wedge \cid \in C \ \} 
-  \cup \{\ \elab{\mesg{w}}{\cid}\ \mid\  \eassign{\mesg{w}}{\cid}{\be}{\cid'} \in \prog
-  \wedge \cid \in H \wedge \cid' \in C \ \} 
+  \begin{array}{lcl}
+    \cinputs &  \defeq
+        & \{\ \rvl{w} \mid\ \reveal{w}{\be}{\cid} \in \prog \wedge \cid \in C \ \} \ \cup\\
+      & & \{\ \elab{\mesg{w}}{\cid}\ \mid\  \eassign{\mesg{w}}{\cid}{\be}{\cid'} \in
+              \prog \wedge \cid \in H \wedge \cid' \in C \ \}
+  \end{array}
   $$
 \end{definition}
+
+\subsection{Passive Security}
 
@@
 \subsection{Malicious Security}
+
+\adversaryfig
 

### 2024-05-05 revision-1/model.tex
checkpoint

@@
 memories to denote mappings of variables to values (i.e., outcomes),
-so for example given a pmf $\pmf$ we will write $\pmf(\{ \sx{1}{a}
-\mapsto 0, \vx{2}{b} \mapsto 1 \})$ to denote the probability that
-$\sx{1}{a} = 0$ and $\vx{2}{b} = 1$ in the final memory of a
+so for example given a pmf $\pmf$ we will write $\pmf(\{ \elab{\secret{x}}{1}
+\mapsto 0, \elab{\mesg{y}}{2} \mapsto 1 \})$ to denote the (joint) probability that
+$\secret{x}}{1} = 0 \wedge \elab{\mesg{y}}{2} = 1$ in the final memory of a
 program containing these variables.

### 2024-05-05 ppdp24/lang.tex
checkpoint

@@
 \label{section-lang}
-
-\minifedfig
 
@@
 
-\subsection{Syntax} The syntax of $\minifed$, defined in
+\subsection{Syntax}
+
+\minifedfig
+
+The syntax of $\minifed$, defined in
 Figure \ref{fig-minifed}, includes values $v$ and standard
@@
 \end{itemize}
+
 Both messages $\mesg{w}$ and reveals $\rvl{w}$ can be

### 2024-05-05 ppdp24/model.tex
checkpoint

@@
 
-\subsection{Passive Security}
+\subsection{Passive Correctness and Security}
+
+In the passive setting we assume that $H$ and $C$ follow the
+rules of protocols and share messages as expected. A first
+consideration is whether a given protocol is \emph{correct}
+with respect to an ideal functionality. 
+\begin{definition}[Passive Correctness]
+  %Given $\prog$ with output variables $\out{1},\ldots,\out{n}$ and ideal
+  We say that a protocol $\prog$ is \emph{passive correct} for a functionality
+  $\idealf$ iff for all $\store \in \mems(\secrets(\prog))$
+  we have $\progtt(\prog)(\idealf(\store) \mid \store) = 1$.
+  %with $\idealf(\store) = v_1,\ldots,v_n$ we have
+  %$\progtt(\prog)(\out{1} \mapsto v_1,\ldots,\out{n} \mapsto v_n \mid \store) = 1$.
+\end{definition}
 
@@
 protocol execution.
-
-Given inputs $\store$ and ideal
-functionality output $v$, we write $ \prob(\SIM(\store,v) = \store') $
+\begin{definition}
+  Given $\store$, and $v$,we write $ \prob(\SIM(\store,v) = \store') $
   to denote the probability that $\SIM(\store,v)$ returns corrupt views
-$\store'$ as a result. We can then define the probability distribution
-of corrupt views reconstructed by the simulator as follows:
-\begin{definition}
-  Given $\store$, and $v$, we write $\dist(\SIM(\store,v))$ to
+  $\store'$ as a result  we write $\dist(\SIM(\store,v))$ to
   denote the distribution of corrupt views reconstructed by the
@@
 \end{definition}
-Observe that this definition relies on protocol correctness, which
-we can define as follows.
-\begin{definition}[Passive Correctness]
-  Given $\prog$ with output variables $\out{1},\ldots,\out{n}$ and ideal
-  functionality $\idealf$, we say that $\prog$ is \emph{passive correct}
-  for $\idealf$ iff for all $\store \in \mems(\secrets(\prog))$
-  with $\idealf(\store) = v_1,\ldots,v_n$ we have
-  $\progtt(\prog)(\out{1} \mapsto v_1,\ldots,\out{n} \mapsto v_n \mid \store) = 1$.
-\end{definition}
 
 \subsection{Malicious Security}
-
-\adversaryfig
 
@@
 be interpreted to yield a corrupt input. Then we define an evaluation
-relation that incorporates the adversary:
-$$
-\begin{array}{rclr}
-  (\store, \xassign{x}{\be}{\cid};\prog) &\aredx&
-  (\extend{\store}{x}{\lcod{\store,\be}{\cid}}, \prog) & \cid \in H\\
-  (\store, \xassign{x}{\be}{\cid};\prog) &\aredx&
-  (\extend{\store}{x}{\lcod{\arewrite(\store_C,\be)}{\cid}}, \prog) & \cid \in C
-\end{array}
-$$
+relation that incorporates the adversary.
 
-$$
-\begin{array}{rcl@{\qquad}r}
-  (\store,\elab{\assert{\phi(\be)}}{\cid};\prog) &\redx& (\store,\prog) & \text{if\ } \phi(\lcod{\store,\be}{\cid}) \text{\ or\ } \cid \in C\\
-  (\store,\elab{\assert{\phi(\be)}}{\cid};\prog) &\redx& (\store,\bot)  & \text{if\ } \neg\phi(\store,\lcod{\store,\be}{\cid})
-\end{array}
-$$
+\adversaryfig
 

### 2024-05-05 ppdp24/model.tex
checkpoint

@@
 
-In the malicious model we assume that corrupt clients are in the thrall of an
-adversary $\adversary$ who does not necessarily follow the rules of the protocol.
-We model this by positing a $\arewrite$ function which is given a corrupt memory
-$\store_C$ and expression $\be$, and returns a rewritten expression that can
-be interpreted to yield a corrupt input. Then we define an evaluation
-relation that incorporates the adversary.
+In the malicious model we assume that corrupt clients are in the
+thrall of an adversary $\adversary$ who does not necessarily follow
+the rules of the protocol.  We model this by positing a $\arewrite$
+function which is given a corrupt memory $\store_C$ and expression
+$\be$, and returns a rewritten expression that can be interpreted to
+yield a corrupt input. We define the evaluation relation that
+incorporates the adversary in Figure \ref{fig-adversary}.
+
+Another important distinction of the malicious setting is that it
+typically incorporates \emph{abort}. Honest parties implement
+strategies to detect rule-breaking-- aka \emph{cheating}-- by
+using, e.g., message authentication codes with semi-homomorphic
+properties as in BDOZ and SPDZ \cite{XXX}. If cheating is
+detected, the protocol is aborted. To model this, we extend
+$\minifed$ with abort $\bot$ and an \emph{assert} command.
+\begin{definition}
+  We add $\bot$ to our syntax of programs and add the form
+  $\elab{\assert(\phi(\be))}{\cid}$ to the command syntax of $\minifed$,
+  with semantics given in Figure \ref{fig-adversary}. We also extend the
+  domain of memories $\store$ to $\mathbb{F}_p \cup \{ \bot \}$.
+\end{definition}
 

### 2024-05-05 ppdp24/model.tex
checkpoint

@@
 
+\adversaryfig
+
 Another important distinction of the malicious setting is that it
@@
 detected, the protocol is aborted. To model this, we extend
-$\minifed$ with abort $\bot$ and an \emph{assert} command.
+$\minifed$ with an \ttt{assert} command and extend the
+range of memories with $\bot$. Note that the adversary is
+free to ignore their own assertions.
 \begin{definition}
-  We add $\bot$ to our syntax of programs and add the form
-  $\elab{\assert(\phi(\be))}{\cid}$ to the command syntax of $\minifed$,
-  with semantics given in Figure \ref{fig-adversary}. We also extend the
-  domain of memories $\store$ to $\mathbb{F}_p \cup \{ \bot \}$.
+  We add assertions of the form $\elab{\assert{\phi(\be)}}{\cid}$ to the command
+  syntax of $\minifed$, where $\phi$ is a decidable predicate on
+  $\mathbb{F}_p$ and with operational semantics given in Figure
+  \ref{fig-adversary}. We also extend the range of memories $\store$
+  to $\mathbb{F}_p \cup \{ \bot \}$.
 \end{definition}
 
-\adversaryfig
-
-Given this we define the $\adversary$ distribution of $\prog$. 
+It is necessary to add $\bot$ to the range of memories since
+the possibility of abort needs to be reflected in adversrarial
+runs of a protocol. We can define $\aruns(\prog,\adversary)$
+as the ``prefix'' memories that result from possibly-aborting
+protocols, but we also need to ``pad out'' the memories
+of partial runs with $\bot$, as we defin in $\botruns(\prog,\adversary)$,
+to properly reflect the contents of views and outputs even in case of abort. 
 \begin{definition}
@@
   $$
-  \aruns(\prog) \defeq \{ \store \mid \preproc(\store)  \wedge
-  ((\store_{S\cup R},\prog) \aredx^* (\store,\varnothing) \vee (\store_{S \cup R},\prog) \aredx^* (\store,\bot)) \}
+    \aruns(\prog) \defeq 
+    \{ \store \mid \preproc(\store)  \wedge ((\store_{S\cup R},\prog) \aredx^* (\store,\varnothing)\}% \vee (\store_{S \cup R},\prog) \aredx^* (\store,\bot)) \}
   $$
-  and also define:
+  and also define the following which pads out undefined views and outputs with $\bot$:
   $$
-  \botruns(\prog) \defeq \{ \store\{ x_1 \mapsto \bot, \ldots, x_n \mapsto \bot \} \mid
-  \store \in \aruns(\prog) \wedge \{ x_1,\ldots,x_n\} = (V \cup O) - \dom(\store) \}
+  \begin{array}{l}
+    \botruns(\prog) \defeq \\
+    \qquad \{ \store\{ x_1 \mapsto \bot, \ldots, x_n \mapsto \bot \} \mid \\
+    \qquad \phantom{\{} \store \in \aruns(\prog) \wedge \{ x_1,\ldots,x_n\} = (V \cup O) - \dom(\store) \}
+  \end{array}
   $$
@@
 
+Given this preamble, we can define malicious simulation and malicious security
+in a standard manner \cite{evans2018pragmatic}, as follows.
 \begin{definition}[Malicious Simulation]

### 2024-05-05 ppdp24/hyper.tex
checkpoint

@@
 In this Section we formulate probabilistic versions of well-studied
-hyperproperties of confidentiality and integrity, including noninterference,
-gradual release, declassification, and robust declassification.
-We demonstrate a soundness relation between noninterference and
-passive security, and between robust declassification and malicious
-security. We subsequently leverage this relation to enforce
-malicious security using ``traditional'' security type methods
-in Section \ref{section-types}. Previous work has explored
-a similar approach to security type enforcement
-\cite{6266151,almeida2018enforcing} but mainly
-for aspects of passive security.
+hyperproperties of confidentiality and integrity, including
+noninterference, gradual release, declassification, and robust
+declassification.  We follow nomenclature developed in previous work
+on characterizing declassification policies in deterministic settings
+\cite{sabelfeld2009declassification}, but adapt them to our
+probabilistic one.
+
+%We demonstrate a soundness relation between noninterference and
+%passive security, and between robust declassification and malicious
+%security.
+%Some previous work on security type enforcement
+%\cite{6266151,almeida2018enforcing} and has considered similar relationships
+%but mainly for aspects of passive security.
 
@@
   \label{definition-NIMO}
-  We say that a program $\prog$ satisfies \emph{noninterference modulo output}
-  iff for all $H$ and $C$ and 
-  $\store_1,\store_2 \in \mems(S)$ we have:
+  We say that a program $\prog$ with $\iov(\prog) = S \cup V \cup O$
+  satisfies \emph{noninterference modulo output}
+  iff for all $H$ and $C$ and $\store_1 \in \mems(S_C \cup O)$ and $\store_2 \in \mems(\houtputs)$
+  we have:
   $$
-  (\store_1 =_C \store_2 \ \wedge \ 
-  (\condd{\progtt(\prog)}{O}{\store_1} = \condd{\progtt(\prog)}{O}{\store_2}))
-  \implies 
-  (\condd{\progtt(\prog)}{\houtputs}{\store_1} = \condd{\progtt(\prog)}{\houtputs}{\store_2})
+  \condd{\progtt(\prog)}{S_H}{\store_1} = \condd{\progtt(\prog)}{S_H}{\store_1 \cap \store_2}
  $$
-  where $\iov(\prog) = S \cup V \cup O$.
 \end{definition}
+%\begin{definition}[Noninterference modulo output]
+%  \label{definition-NIMO}
+%  We say that a program $\prog$ satisfies \emph{noninterference modulo output}
+%  iff for all $H$ and $C$ and 
+%  $\store_1,\store_2 \in \mems(S)$ we have:
+%  $$
+%  (\store_1 =_C \store_2 \ \wedge \ 
+%  (\condd{\progtt(\prog)}{O}{\store_1} = \condd{\progtt(\prog)}{O}{\store_2}))
+%  \implies 
+%  (\condd{\progtt(\prog)}{\houtputs}{\store_1} = \condd{\progtt(\prog)}{\houtputs}{\store_2})
+%  $$
+%  where $\iov(\prog) = S \cup V \cup O$.
+%\end{definition}
 Intuitively, this conditional noninterference property implies that
-the simulator can just run the protocol in simulation to
+corrupt views give the adversary no better chance of guessing honest
+secrets than the output and corrupt inputs do. So
+the simulator can just arbitrarily pick any honest secrets that
+produce the expected output and run the protocol in simulation to
 reconstruct real world corrupt view distributions. But it requires
@@
 
-\begin{proof}
-  Let $H$ and $C$ be an arbitrary partition of the federation and
-  suppose $\prog$ satisfies noninterference modulo output. Let
-  $\iov(\prog) = (V,S)$ and $\flips(\prog) = F$ with output view $\outv
-  \in V$ and let $\store$ be an arbitrary element of $\mems(S \cup
-  F)$. Then the distribution of adversarial views in the real world
-  is, by definition,
-  $\condd{\progd(\prog)}{V_C}{\store_S}$.
-
-  The simulator is given both $\store_{S_C}$ and
-  $\idealf(\store_S)$.  The simulation $\SIM_C(\store_{S_C},
-  \idealf(\store_S))$ can be defined as follows. First, some $\store'
-  \in \ik(\idealf,\idealf(\store))$ is randomly chosen such that
-  $\store' =_C \store_S$, as is a random tape $\store'' \in
-  \mems(F)$\footnote{The real/ideal model allows consultation of a
-  ``Random Oracle''.}. Then, the run of $(\store' \cup \store'',
-  \prog)$ is evaluated in simulation, yielding $(\store^{\SIM},\varnothing)$.
-  The simulation returns $\store^{\SIM}_{V_C}$ as a result.
-
-  Now, since the random tape is selected in simulation from the same distribution
-  that we assume for the real world, after selection of $\store'$ the
-  probability that any particular $\store^{\SIM}_{V_C}$ is returned is by definition
-  $
-  \progd(\prog)(\store^{\SIM}_{V_C} | \store')
-  $.
-  Furthermore, since we assume that $\prog$ correctly implements $\idealf$, this
-  means
-  $
-  \store_S =_C \store' \wedge
-     \condd{\progd(\prog)}{\{ \outv \}}{\store_S} =
-     \condd{\progd(\prog)}{\{ \outv \}}{\store'}
-  $,
-  so by the definition of noninterference modulo output,
-  any choice of $\store'$ yields the same distribution $\condd{\progd(\prog)}{V_C}{\store_S}$.
-  Therefore by definition
-  $
-   \progd(\SIM_C(\store_{S_C},\idealf(\store))) = \condd{\progd(\prog)}{V_C}{\store_S}
-  $.
-\end{proof}
-
 \subsection{Delimited Release}
@@
 In fact, MPC protocols typically satisfy a \emph{delimited release}
-property\cite{XXX}, where messages exchanged remain probabilistically separable
+property\cite{sabelfeld2009declassification}, where messages exchanged remain probabilistically separable
 from secrets, with only declassification events (reveals and outputs)

### 2024-05-05 ppdp24/hyper.tex
checkpoint

@@
 
-In fact, MPC protocols typically satisfy a \emph{delimited release}
-property\cite{sabelfeld2009declassification}, where messages exchanged remain probabilistically separable
+In practice, MPC protocols typically satisfy a \emph{delimited release}
+property \cite{sabelfeld2009declassification}, where messages exchanged remain probabilistically separable
 from secrets, with only declassification events (reveals and outputs)

### 2024-05-05 ppdp24/hyper.tex
checkpoint

@@
 
-\begin{definition}[Malicious Correctness]
-  A protocol $\prog$ with $\iov(\prog) = S \cup V \cup O$ is \emph{malicious correct} iff
-  $
-  \forall \adversary, \store \in \mems(S_H) \ .\ \support(\progtt(\prog)(O_H|\store)) \supseteq
-    \support(\progtt(\prog,\adversary)(O_H|\store))
-  $.
-\end{definition}
+%\begin{definition}[Malicious Correctness]
+%  A protocol $\prog$ with $\iov(\prog) = S \cup V \cup O$ is \emph{malicious correct} iff
+%  $
+%  \forall \adversary, \store \in \mems(S_H) \ .\ \support(\progtt(\prog)(O_H|\store)) \supseteq
+%    \support(\progtt(\prog,\adversary)(O_H|\store))
+%  $.
+%\end{definition}
 
@@
 
+\begin{lemma}
+  If a protocol $\prog$ with $\iov(\prog) = S \cup V \cup O$ is passive correct for
+  $\idealf$ and
+  has integrity, then for all $H$, $C$, and $\adversary$ and $\store_1 \in \mems(S_H)$,
+  there exists $\store_2 \in \mems(S_C)$ such that:
+  $$
+  \idealf(\store_1 \cap \store_2)_H \cup \{ \bot \} \supseteq
+    \support(\progtt(\prog,\adversary)(O_H|\store_1))
+  $$
+\end{lemma}
+
 \begin{theorem}
-  If a protocol satisfies noninterference modulo output and has
-  integrity, then it is malicious secure.
+  If a protocol satisfies noninterference modulo output
+  and has integrity, then it is malicious secure.
 \end{theorem}
@@
   A protocol satisfies \emph{robust declassification} iff it has integrity and
-  satisfies gradual release. 
+  satisfies delimited release. 
 \end{definition}
@@
   Robust declassification does not imply malicious security, but
-  passive security with robust declassification implies malicious security.
+  security with robust declassification implies malicious security.
 \end{theorem}

### 2024-05-05 ppdp24/hyper.tex
checkpoint

@@
   We say that a protocol $\prog$ with $\iov(\prog) = S \cup V \cup O$ has
-  \emph{integrity} iff for all $H$ and $C$ 
-  it is malicious correct and for all adversaries $\adversary$,
+  \emph{integrity} iff for all $H$, $C$, and $\adversary$,
   if $\store \in \aruns(\prog)$ 

### 2024-05-05 ppdp24/hyper.tex
checkpoint

@@
 
-\subsection{Passive Security and Noninterference}
+\subsection{Conditional Noninterference}
 
@@
 \begin{definition}
-  Given a functionality $\idealf$ and output value $v$, its
-  \emph{kernel}, denoted $\ik(\idealf,v)$ is
+  Given a functionality $\idealf$ and outputs $\store_{\mathit{out}}$, their 
+  \emph{kernel}, denoted $\ik(\idealf,\store_{\mathit{out}})$ is
   $
-  \{ \store\ |\ \idealf(\store) = v \}
+  \{ \store\ |\ \idealf(\store) = \store_{\mathit{out}} \}
   $.
-  We say that $\idealf$ is \emph{pre-imageable} iff $\ik(\idealf, v)$ for all
-  $v$ can be computed tractably.
+  We say that $\idealf$ is \emph{pre-imageable} iff $\ik(\idealf, \store_{\mathit{out}})$ for all
+  $\store_{\mathit{out}}$ can be computed tractably.
 \end{definition}
@@
 
-\subsection{Malicious Security and Robust Declassification}
+\subsection{Integrity and Robust Declassification}
 
@@
 \begin{lemma}
+  \label{lemma-malicious-correct}
   If a protocol $\prog$ with $\iov(\prog) = S \cup V \cup O$ is passive correct for
@@
   where $\iov(\prog) = S \cup V \cup O$, and let $\adversary$ be some
-  adversary.  Given some $\store\in S_H$, by definition of malicious
-  correctness the simulator can extract inputs $\store' \in (S_C)$
-  from $\adversary$ where $\idealf(\store \cup \store')$ is equivalent
-  to the real world outputs. These can then be used in a simulated run
-  of $\prog$ along with $\store \in S_H$ to generate a distribution
-  of corrupt views of $\margd{\progtt(\prog)}{\houtputs}$.
-  By assumption of integrity and Lemma \ref{lemma-integrity},
-  this is the same as the real-world distribution of corrupt views
-  in the presence of $\adversary$. 
+  adversary. Suppose $\store \in \aruns(\prog,\adversary)$.
+  As integrity requires, and as Lemma \ref{lemma-malicious-correct}
+  demonstrates with respect to outputs, the most the adversary can do
+  in the presence of integrity is to elicit the same responses from
+  the honest parties-- via the strategy $\store_{\cinputs}$-- as
+  are elicited from some passive run of the protocol using
+  some $\store' \in \mems(S)$ where $\store'_H = \store_{S_H}$,
+  and perhaps to force an abort after some number of message
+  exchanges.
+
+  Therefore, in simulation, $\adversary$ can provide the simulator
+  with some $\store'_C$ in $\SIM_1$ which its strategy is ``impersonating'',
+  allowing $\idealf(\store')$ to
+  be communicated to $\adversary$ in $\SIM_2$ who can then
+  decide whether or not to abort. In the case of abort, the
+  subset of $\houtputs$ to be defined can be communicated to
+  $\SIM_3$, along with $\idealf(\store')$, via $\Sigma$. 
+  In $\SIM_3$, the simulator can then run $\prog$ in simulation
+  with inputs $\store'_C$ and arbitrary $\store'' \in \mems(S_H)$
+  such that $\store'_C \cap \store'' \in \ik(\idealf,\idealf(\store'))$.
+  The assumption of passive security of $\prog$ implies the result.
 \end{proof}

### 2024-05-05 ppdp24/hyper.tex
checkpoint

@@
 
+\emph{Integrity} is an important hyperproperty in security models that admit
+malicious adversaries. Consistent with formulations in deterministic settings,
+we have already defined protocol confidentiality as the preservation of low equivalence
+(of secrets and views), and now we define protocol integrity as the preservation
+of high equivalence (of secrets and views). Intuitively, this property says
+that any adversarial strategy either ``mimics'' a passive strategy with some
+choice of inputs or causes an abort.
 \begin{definition}[Integrity]
@@
 \end{definition}
-
+A first important observation is that integrity preserves protocol correctness
+for honest outputs, except for the possibility of abort. 
 \begin{lemma}
@@
 \end{lemma}
-
+The following result establishes that integrity implies malicious
+security for protocols that are passive secure (which also subsumes
+correctness). 
 \begin{theorem}
-  If a protocol satisfies noninterference modulo output
-  and has integrity, then it is malicious secure.
+  If a protocol is passive secure and has integrity, then it
+  is malicious secure.
 \end{theorem}
@@
 
+The hyperproperty of robust declassification \cite{930133} similarly
+combines a confidentiality property with integrity to establish that
+malicious actors cannot declassify more information than is intended
+by policy. But in this prior work, this policy is established
+by the declassifications themselves, as in delimited release.
+Thus, we can define a robust declassification property as follows. 
 \begin{definition}[Robust Declassification]
@@
 \end{definition}
-
+However, it is important to note that delimited release, and
+hence robust declassification, are not sufficient to establish
+passive or malicious simulator security, where the declassification
+policy is established by the ideal functionality $\idealf$. 
 \begin{theorem}

### 2024-05-05 ppdp24/bruteforce.tex
checkpoint

\section{Verification of Protocols in $\mathbb{F}_2$}

In $\mathbb{F}_2$ a brute force strategy for computing
$\progtt(\prog)$ for any $\prog$ is to directly compute
$\runs(\prog)$. By querying $\progtt{\prog}$ we can verify any of the
hyperproperties discussed previously, or other properties such as
perfect secrecy \cite{XXX}.  Since $\runs(\prog)$ is exponential in
the size of $\secrets(\prog) \cup \randos(\prog)$, this strategy is
feasible only for smaller protocols. However, we can efficiently
convert protocols in the passive setting (without $\ttt{assert}$) to
stratified Datalog programs, and then extract $\runs(\prog)$ by
calculating Least Herbrand models by parallelization and/or other HPC
acceleration techniques for logic programs \cite{aspis2018linear}.
The rewriting we describe here is to Datalog with negation, with a
negation-as-failure model, though we can also use techniques in
\cite{sakama2017linear} to eliminate negation from resulting programs.

In Figure \ref{fig-solve} we define the algorithm
$\solve{\stores}{\be}{\cid}$ which filters a given $\stores$ to obtain the
subset whose elements satisfy $\be$ (make it true). It's correctness
is characterized as follows.
\begin{lemma}
  \label{lemma-solves}
  %For all $\stores$ and $\be$ with $\vars(\be) \subseteq \dom(\stores)$,
  $(\solve{\stores}{\be}{\cid}) = \{ \store \in \stores \ \mid\ \lcod{\store,\be}{\cid} = 1 \}$.
\end{lemma}

\solvefig

As per standard nomenclature, \emph{atoms} are $\minifed$ variables $x$, \emph{literals}
are atoms or negated atoms, and clause bodies are conjunctions of literals.
A \emph{fact} is a clause with no body. A \emph{Datalog program} is a list of clauses.
$$
\begin{array}{rclr}
  \mathit{body} &::=&  x \mid \neg x \mid x \wedge \mathit{body} \mid \neg x \wedge \mathit{body} \mid \varnothing \\
  \mathit{clause} &::=& x \gets \mathit{body} \mid x \gets
\end{array}
$$

\begin{mathpar}
  \vars\ \secret{w} \cid \defeq \{ \elab{\secret{w}}{\cid} \}

  \vars\ \mesg{w} \cid \defeq \{ \elab{\mesg{w}}{\cid} \}

  \vars\ \rvl{w} \cid \defeq \{ \elab{\rvl{w}}{\cid} \}

  \vars\ \flip{w} \cid \defeq \{ \elab{\flip{w}}{\cid} \}

  \vars\ (\be_1 \exor \be_2)\ \cid \defeq (\vars\ \be_1\ \cid) \cup (\vars\ \be_2\ \cid)
\end{mathpar}

The first step in converting a protocol $\prot$ to a Datalog
program is to apply $\solve$ locally to each command
in $\prog$, obtaining constraints on memories that satisfy each
view in isolation.  
\begin{definition} Define:
$$
{vtt}(\xassign{x}{\be}{\cid}) \defeq (x, (\solve{(\mems(\vars\ \be\ \cid))}{\be}{\cid}))
$$
\end{definition}
Given this definition, the mapping of ${vtt}$ across a program
$\prog$- i.e., $(\mathit{map}\ {vtt}\ \prog)$-  essentially defines the
logic program, modulo some syntactic conversion. We can
accomplish the latter as follows, where $\datalog(\prog)$ defines the
full conversion.
\begin{definition} We define the conversion from memories to
  literals and clause bodies as follows:
\begin{mathpar}
  \logit{x \mapsto 1} = x

  \logit{x \mapsto 0} = \neg x

  \logit{\{ x_1 \mapsto \beta_1, \ldots, x_n \mapsto \beta_n\}} =
  \logit{x_1 \mapsto \beta_1} \wedge \cdots \wedge \logit{x_n \mapsto \beta_n}
\end{mathpar}
Given pairs $(x,\stores)$ in the range of ${vtt}$, we define the conversion
to clauses as follows:
\begin{mathpar}
  \mathit{clauses}(x,\{ \store_1,...,\store_n \}) = x \gets \logit{\store_1} \vee \cdots \vee x \gets \logit{\store_n}
\end{mathpar}
The $\minifed$-to-Datalog conversion is then defined as:
$$
\datalog(\prog) \defeq  \mathit{map}\ \mathit{clauses}\ (\mathit{map}\ {vtt}\ \prog)
$$
\end{definition}

In addition to converting view definitions to logic clauses, we also need to convert
secrets and random tapes. Since we assume given values for these in an arbitrary run of
the program, we can capture these as a ``fact base'' in our program, where
a fact is a clause of the form $x \gets \varnothing$ and means that $x$
is true in any model of the program. 
\begin{definition}
  Given $\store$, let $\{x_1,\ldots,x_n \} =
  \{ x \in \dom(\store) \mid \store(x) = 1 \}$.
  Then define:
  $$
  \mathit{facts}(\store) = x_1 \gets \varnothing, \ldots, x_n \gets \varnothing
  $$
\end{definition}

For any safe $\prog$ with $\iov(\prog) = S \cup V$ and $\flips(\prog) = F$, and
$\store \in \mems(S \cup F)$, it is the case
that $(\mathit{facts}(\store),\datalog(\prog))$ is a \emph{normal}, \emph{stratified}
(in fact, non-recursive) Datalog program, and so has a unique Least Herbrand Model
and is thus amenable to HPC optimization techniques \cite{aspis2018linear}. 
To compute $\runs(\prog)$, and thus $\progd(\prog)$, we compute
the Least Herbrand Model $\store$ of $(\mathit{facts}(\store'),\datalog(\prog))$
for all $\store' \in \mems(S \cup F)$, observing that model computation for
each $\store'$ can be done in parallel. The following establishes
correctness of this approach.
\begin{lemma}
  For all $\prog$ with $\iov(\prog) = S \cup V$ and $\flips(\prog) = F$,
  $\datalog(\prog)$ is a \emph{normal}, \emph{stratified}
  program \cite{aspis2018linear}, and $\store$ is the unique Least Herbrand
  Model of $(\mathit{facts}(\store_{S \cup F}),\datalog(\prog))$
  iff $\store \in \runs(\prog)$.
\end{lemma}

### 2024-05-05 ppdp24/bruteforce.tex
checkpoint

@@
 
-As per standard nomenclature, \emph{atoms} are $\minifed$ variables $x$, \emph{literals}
-are atoms or negated atoms, and clause bodies are conjunctions of literals.
-A \emph{fact} is a clause with no body. A \emph{Datalog program} is a list of clauses.
+We define the syntax of Datalog as follows. As per standard
+nomenclature, \emph{atoms} are $\minifed$ variables $x$,
+\emph{literals} are atoms or negated atoms, and clause bodies are
+conjunctions of literals.  A \emph{fact} is a clause with no body. A
+\emph{Datalog program} is a list of clauses.
 $$
@@
 $$
-
+When translating protocols, we need to extract the variables
+that occur in expressions $\be$ computed by a client $\cid$,
+written $\vars\ \be\ \cid$ where:
 \begin{mathpar}
@@
 \end{mathpar}
-
-The first step in converting a protocol $\prot$ to a Datalog
-program is to apply $\solve$ locally to each command
-in $\prog$, obtaining constraints on memories that satisfy each
-view in isolation.  
+... and so on. Then, to begin converting a protocol $\prot$ to a Datalog
+we define the function ${vtt}$ that applies $\solve$ locally to each command
+in $\prog$, obtaining the subset of memories that result
+in a variable assignment of $1$.  
 \begin{definition} Define:
@@
 \end{definition}
-Given this definition, the mapping of ${vtt}$ across a program
+The mapping of ${vtt}$ across a program
 $\prog$- i.e., $(\mathit{map}\ {vtt}\ \prog)$-  essentially defines the

### 2024-05-05 ppdp24/bruteforce.tex
checkpoint

@@
 perfect secrecy \cite{XXX}.  Since $\runs(\prog)$ is exponential in
-the size of $\secrets(\prog) \cup \randos(\prog)$, this strategy is
+the size of $\secrets(\prog) \cup \flips(\prog)$, this strategy is
 feasible only for smaller protocols. However, we can efficiently
@@
 secrets and random tapes. Since we assume given values for these in an arbitrary run of
-the program, we can capture these as a ``fact base'' in our program, where
-a fact is a clause of the form $x \gets \varnothing$ and means that $x$
-is true in any model of the program. 
+the program, we can capture these as a particular fact base.
 \begin{definition}
@@
   $$
-  \mathit{facts}(\store) = x_1 \gets \varnothing, \ldots, x_n \gets \varnothing
+  \mathit{facts}(\store) = x_1 \gets, \ldots, x_n \gets
   $$
 \end{definition}
-
-For any safe $\prog$ with $\iov(\prog) = S \cup V$ and $\flips(\prog) = F$, and
-$\store \in \mems(S \cup F)$, it is the case
-that $(\mathit{facts}(\store),\datalog(\prog))$ is a \emph{normal}, \emph{stratified}
-(in fact, non-recursive) Datalog program, and so has a unique Least Herbrand Model
-and is thus amenable to HPC optimization techniques \cite{aspis2018linear}. 
-To compute $\runs(\prog)$, and thus $\progd(\prog)$, we compute
-the Least Herbrand Model $\store$ of $(\mathit{facts}(\store'),\datalog(\prog))$
-for all $\store' \in \mems(S \cup F)$, observing that model computation for
-each $\store'$ can be done in parallel. The following establishes
+The following result ties these pieces together and establishes
 correctness of this approach.
 \begin{lemma}
-  For all $\prog$ with $\iov(\prog) = S \cup V$ and $\flips(\prog) = F$,
+  For all $\prog$, 
   $\datalog(\prog)$ is a \emph{normal}, \emph{stratified}
   program \cite{aspis2018linear}, and $\store$ is the unique Least Herbrand
-  Model of $(\mathit{facts}(\store_{S \cup F}),\datalog(\prog))$
+  Model of: $$(\mathit{facts}(\store_{\secrets(\prog) \cup \flips(\prog)}),\datalog(\prog))$$
   iff $\store \in \runs(\prog)$.
 \end{lemma}
+Finally, to compute $\runs(\prog)$, and thus $\progtt(\prog)$, we compute
+the Least Herbrand Model $\store$ of $(\mathit{facts}(\store'),\datalog(\prog))$
+for all $\store' \in \mems(\secrets(\prog) \cup \flips(\prog))$, observing
+that model computation for distinct fact bases can be done in parallel.

### 2024-05-05 ppdp24/bruteforce.tex
checkpoint

@@
 ... and so on. Then, to begin converting a protocol $\prot$ to a Datalog
-we define the function ${vtt}$ that applies $\solve$ locally to each command
+we define the function ${vtt}$ that applies $\solve{}{}{}$ locally to each command
 in $\prog$, obtaining the subset of memories that result

### 2024-05-06 ppdp24/bruteforce.tex
checkpoint

--- 2024-05-05 ppdp24/bruteforce.tex
@@
 $\progtt(\prog)$ for any $\prog$ is to directly compute
-$\runs(\prog)$. By querying $\progtt{\prog}$ we can verify any of the
+$\runs(\prog)$. By querying $\progtt(\prog)$ we can verify any of the
 hyperproperties discussed previously, or other properties such as
@@
 negation-as-failure model, though we can also use techniques in
-\cite{sakama2017linear} to eliminate negation from resulting programs.
+\cite{sakama2017linear} to eliminate negation from resulting programs
 
@@
 $\solve{\stores}{\be}{\cid}$ which filters a given $\stores$ to obtain the
-subset whose elements satisfy $\be$ (make it true). It's correctness
-is characterized as follows.
+subset whose elements satisfy $\be$. In this figure and elsewhere for
+$\mathbb{F}_2$ we use standard names of logical connectives
+for field operations ($\eand$ and $\exor$ for $*$ and $+$ respectively) and
+add $\enot$ and $\eor$ as trivial but convenient extensions. The correctness
+of this operation is characterized as follows.
 \begin{lemma}
@@
 ... and so on. Then, to begin converting a protocol $\prot$ to a Datalog
-we define the function ${vtt}$ that applies $\solve$ locally to each command
+we define the function ${vtt}$ that applies $\solve{\!}{\!}{\!\!}$ locally to each command
 in $\prog$, obtaining the subset of memories that result
@@
 The mapping of ${vtt}$ across a program
-$\prog$- i.e., $(\mathit{map}\ {vtt}\ \prog)$-  essentially defines the
+$\prog$-- i.e., $(\mathit{map}\ {vtt}\ \prog)$--  essentially defines the
 logic program, modulo some syntactic conversion. We can
@@
 \begin{mathpar}
-  \logit{x \mapsto 1} = x
+  \logit{x \mapsto 1} \defeq x
 
-  \logit{x \mapsto 0} = \neg x
+  \logit{x \mapsto 0} \defeq \neg x
 
-  \logit{\{ x_1 \mapsto \beta_1, \ldots, x_n \mapsto \beta_n\}} =
+  \logit{\{ x_1 \mapsto \beta_1, \ldots, x_n \mapsto \beta_n\}} \defeq
   \logit{x_1 \mapsto \beta_1} \wedge \cdots \wedge \logit{x_n \mapsto \beta_n}
@@
 Given pairs $(x,\stores)$ in the range of ${vtt}$, we define the conversion
-to clauses as follows:
-\begin{mathpar}
-  \mathit{clauses}(x,\{ \store_1,...,\store_n \}) = x \gets \logit{\store_1} \vee \cdots \vee x \gets \logit{\store_n}
-\end{mathpar}
+to clauses as  $\mathit{clauses}(x,\{ \store_1,...,\store_n \}) \defeq x \gets \logit{\store_1} \vee \cdots \vee x \gets \logit{\store_n}$.
 The $\minifed$-to-Datalog conversion is then defined as:
@@
   \{ x \in \dom(\store) \mid \store(x) = 1 \}$.
-  Then define:
-  $$
-  \mathit{facts}(\store) = x_1 \gets, \ldots, x_n \gets
-  $$
+  Then define $\mathit{facts}(\store) \defeq x_1 \gets, \ldots, x_n \gets$.
 \end{definition}

### 2024-05-06 ppdp24/metalang.tex
checkpoint

--- 2023-07-12 popl24/metalang.tex
@@
 
+\metaprotfig
+
 \subsection{Semantics}
-
-\begin{fpfig}[t]{$\metaprot$ syntax (T), evaluation contexts (M), and operational semantics (B).}{fig-metaprot}
-    {\small
-    $$
-    \begin{array}{rcl@{\hspace{8mm}}r}
-      \multicolumn{3}{l}{\flab \in \mathrm{Field},\   x \in \mathrm{EVar}, \  f \in \mathrm{FName}}\\[2mm]
-      %x &\in& \mathrm{EVar}\\
-      %f &\in& \mathrm{FName}\\[2mm]
-      e &::=& b \mid \flip{e}{e} \mid \secret{e}{e} \mid \view{e}{e} \mid \oracle{e} \mid \enot\ e \mid e\ \eand\ e \mid e\ \exor\ e \mid & \textit{expressions}\\[0mm]
-      & & \select{e}{e}{e} \mid 
-      \send{\view{e}{e}}{e} \mid \send{\view{e}{e}}{\OT{e}{e}{e}} \mid e;e \mid \\[0mm]
-      & & x \mid \elet{x}{e}{e} \mid f(e,\ldots,e) \mid \{ \flab = e; \ldots; \flab = e \}
-      \mid e.\flab \mid e\concat e \mid (e) \\[2mm]
-      v &::=& w \mid \cid \mid \be \mid \{ \flab = v;\ldots;\flab = v \} 
-      \mid \ttt{()} & \textit{values}\\[2mm]
-           {fn} &::=& f(x,\ldots,x) \{ e \} & \textit{functions}
-    \end{array}
-    $$
-    }
-
-    \rule{130mm}{0.5pt}
-
-    {\small
-    $$
-    \begin{array}{rcl@{\hspace{3mm}}r}
-      E &::=& [\,] \mid \enot\ E \mid E\ \bop\ e \mid v\ \bop\ E \mid  \flip{E}{e} \mid \secret{E}{e} \mid \view{E}{e} \mid \oracle{E} \mid  \\[1mm]
-      & & \flip{\cid}{E} \mid \secret{\cid}{E} \mid \view{\cid}{E} \mid \send{E}{e} \mid \send{\view{\cid}{w}}{E} \mid \OT{E}{e}{e} \\[1mm]
-      & & \mid \OT{v}{E}{e} \mid \OT{v}{v}{E} \mid \select{E}{e}{e} \mid \select{v}{E}{e} \mid \\[1mm]
-      & & \select{v}{v}{E} \mid \elet{x}{E}{e} \mid f(v,\ldots,v,E,e,\ldots,e) \mid \\[1mm]
-      & & \{ \flab = v;\ldots;\flab = v;\flab = E;\flab = e;\ldots;\flab = e \} \mid E.\flab \mid E\concat e \mid v \concat E
-    \end{array}
-    $$
-    \vspace{.5mm}
-
-    \rule{130mm}{0.5pt}
-
-    $$
-    \begin{array}{rcl@{\hspace{10mm}}r}
-      \config{\prog}{\elet{x}{v}{e}} &\redx& \config{\prog}{e[v/x]}\\
-      \config{\prog}{f(v_1,...,v_n)} &\redx&
-      \config{\prog}{e[v_1/x_1,\ldots,v_n/x_n]} & 
-      \codebase(f) = x_1,\ldots,x_n,\ e\\
-      \config{\prog}{\{\ldots; \flab = v; \ldots\}.\flab} &\redx&
-      \config{\prog}{v}\\
-      \config{\prog}{w_1\concat w_2} &\redx& \config{\prog}{w_1w_2}\\
-      \config{\prog}{v;e} &\redx& \config{\prog}{e}\\
-      \config{\prog}{\instr} &\redx& \config{\prog;\instr}{()}\\
-      \config{\prog}{E[e]} &\redx& \config{\prog'}{E[e']} & \text{if}\ \config{\prog}{e} \redx \config{\prog'}{e'} 
-    \end{array}
-    $$
-  }
-\end{fpfig}
 
@@
 and instantiation of compositional program elements.
-
-\subsection{Type Theory and Static Type Safety}
-
-It is desirable to statically enforce safety of both $\metaprot$
-programs and the safety of the $\fedprot$ programs they
-generate. Although safety of the latter could be enforced
-post-generation by a direct analysis, for large programs this can be
-much more expensive and it is also better to not waste time on
-resource-intensive compilation of programs with known errors
-\cite{kreuter2012billion}. Some consequences of safety errors, for example, accidental
-reuse of one-time pads can also undermine security.
-
-The type syntax of $\metaprot$ is defined in Figure
-\ref{fig-metaprot-types}. It includes a weak form of dependency:
-string types $\stringty{e}$ and client types $\cidty{e}$
-are parameterized by expressions $e$ that precisely reflect the
-type of the value. Boolean expression forms have the type
-$\bet{e}$ indexed by expressions $e$ indicating the client
-id type of the expression. The dependency is weak in the
-sense that expressions in types are a strict subset of
-expression forms- for any $\stringty{e}$ the expression $e$
-is either a variable, a string, or a concatenation form,
-and for any $\cidty{e}$ or $\bet{e}$ the expression $e$
-is either a variable or a client id $\cid$. We define
-the equivalence $w_1 \concat w_2 \equiv w_1w_2$ in typings.
-
-Type judgements for expressions are of the form
-$\tjudge{\viewst_1}{\gamma}{e}{\viewst_2}$ where the \emph{view
-effect} $\viewst_1$ denotes the views that have been defined so far,
-and $\viewst_2$ records new views defined as the effect of the
-expression on the residual $\minifed$ program.  Type judgements are
-syntax-directed- selected rules are shown in Figure
-\ref{fig-metaprot-types}
-%, with a more complete set provided in
-%Appendix Figure \ref{fig-metaprot-tjudge}
-and a full implementation is available in our online repository \cite{jpdf-github}. The
-$\TirName{AssignT}$ rule captures the effect of new view
-definitions. The $\TirName{And}$ rule illustrates how program safety
-is enforced, by ensuring that subexpressions of boolean expressions
-have the same owner.
-
-The $\TirName{Appt}$ rule applies to function
-definition and application respectively. Function typings rely on the
-definition of function input type annotations $\tas$ and
-type term substitutions $\sigma$. 
-\begin{definition}
-  A \emph{function input type annotation} $\tas$ is a mapping from
-  function names $f$ to type products $\tau_1 * \cdots * \tau_n$.
-  A \emph{type term substitution} $\sigma$ is a mapping from
-  $\minifed$ variables $x$ to values, where $\sigma(\tau)$ denotes
-  the replacement of occurences of $x$ in $\tau$ with $\sigma(x)$. 
-\end{definition}
-We assume that input type annotations $\tas$ are provided by the
-programmer for all function definitions. This guarantees that
-$\metaprot$ type checking is straightforward and efficient.
-Function types are of the form
-$
-(\tau_1 * \cdots * \tau_n \rightarrow \tau,\viewst)
-$,
-where $\viewst$ denotes the effect of the function on the residual
-program.  The function type can be understood as a dependent $\Pi$
-type, with every term variable bound. When applied, these variables
-are instantiated with a type term substitution $\sigma$. In our
-implementation, we adapt \emph{synthesis} as defined in Dependent ML \cite{10.1145/292540.292560}
-to obtain this $\sigma$- essentially this is a match on the syntactic
-structure of types and expressions.
-
-\begin{fpfig}[t]{Type syntax of $\metaprot$ (T) and selected type judgement rules (B).}{fig-metaprot-types}
-{\small
-  $$
-  \begin{array}{rcl@{\hspace{2mm}}r}
-    \srct &::=& \cidty{e} \mid \stringty{e} \mid \bet{e} \mid  \{ \flab : \srct;\ldots;\flab : \srct \} \mid \tau * \cdots * \tau \rightarrow \tau,\viewst & \gdesc{types}\\ [1mm]
-    %&&  \{ \flab : \srct;\ldots;\flab : \srct \} \mid \tau * \cdots * \tau \rightarrow \tau,\viewst \\[1mm]
-    \viewst  &::=& \view{e}{e};\viewst \mid \varnothing   & \gdesc{view effects}\\[1mm]
-    \Gamma &::=& \Gamma; x : \tau \mid \varnothing & \gdesc{type environments}    
-  \end{array}
-  $$
-}
-
-\rule{130mm}{0.5pt}
-
-{\small
-\begin{mathpar}
-\inferrule[\TirName{SecretT}]
-{\tjudge{\viewst}{\Gamma}{e_1}{\cidty{e_1'}}{\viewst_1}\\
-\tjudge{\viewst_1}{\Gamma}{e_2}{\stringty{e_2'}}{\viewst_2}}
-{\tjudge{\viewst}{\Gamma}{\secret{e_1}{e_2}}{\bet{e_1'}}{\views_2}}
-
-\inferrule[\TirName{AndT}]
-{
-\tjudge{\viewst}{\Gamma}{e_1}{\bet{e}}{\viewst_1}\\
-\tjudge{\viewst_1}{\Gamma}{e_2}{\bet{e}}{\viewst_2}
-}
-{\tjudge{\viewst}{\Gamma}{e_1\ \eand\ e_2}{\bet{e}}{\viewst_2}}
-
-\inferrule[\TirName{AssignT}]
-{
-\tjudge{\viewst}{\Gamma}{e_1}{\cidty{e_1'}}{\viewst_1}\\
-\tjudge{\viewst_1}{\Gamma}{e_2}{\stringty{e_2'}}{\viewst_2}\\
-\tjudge{\viewst_2}{\Gamma}{e_3}{\bet{e_3'}}{\viewst_3}
-}
-{
-\tjudge{\viewst}{\Gamma}{\eassign{\view{e_1}{e_2}}{e_3}}{\unity}{(\viewst_3 ; \view{e_1'}{e_2'} )}
-}
-
-\inferrule[AppT]
-{\Gamma(f) =  
- \tau_1 * \cdots * \tau_n \rightarrow \tau, \viewst_f \\ 
- \tjudge{\viewst}{\Gamma}{e_1}{\sigma(\tau_1)}{\viewst_1}
- \ \cdots\  
- \tjudge{\viewst_{n-1}}{\Gamma}{e_n}{\sigma(\tau_n)}{\viewst_n}}
-{\tjudge{\viewst}{\Gamma}{f(e_1,\ldots,e_n)}{\sigma(\tau)}{(\viewst_n ; \sigma(\viewst_f))}}
-\end{mathpar}
-}
-\end{fpfig}
-
-Top-level type judgements are of the form $\Gamma \vdash \codebase, e
-: \tau, V$, where all the functions in $\codebase$ are well-typed in
-$\Gamma$, the top level view effect $V$ is a set of concrete
-$\fedprot$ views which are constructed by a disjoint union of the views
-in the effect of $e$- the disjointness requirement guarantees that
-views are uniquely defined. Our $\metaprot$ type safety result is
-formulated as follows. In addition to safe execution of the
-metaprogram, it also guarantees safety of the residual $\fedprot$
-program (Definition \ref{definition-safety}).
-\begin{theorem}[$\metaprot$ Type Safety]
-  \label{theorem-metalang-safety}
-  Given $\codebase$, $e$, and $\Gamma$ with $\Gamma \vdash \codebase,e : \unity : V$,
-  then $\config{\varnothing}{e} \redxs \config{\prog}{\ttt{()}}$ where
-  $\prog$ is safe with $\iov(\prog) = S \cup V$ for some $S$.
-\end{theorem}
-
-\subsection{Example: 2-Party GMW}
-\label{section-metalang-gmw}
-
-\input{gmw}
-
-The GMW protocol uses secret sharing to represent data flowing through
-circuits. In the 2-party case, clients 1 and 2 each share their input
-secrets, and use those shares to represent inputs to gates. Outputs
-are also represented as shares. We refer to the pair of shares
-representing any particular value as a \emph{wire value}, and
-we represent them via records of the form
-$
-\ttt{\{c1 = } v_1\ttt{;c2 = } v_2\ttt{\}} 
-$
-where $v_1$ and $v_2$ are client 1 and 2's shares respectively.
-
-For full details of the GMW protocol the reader is referred to
-\cite{evans2018pragmatic}. Our implementation libary is shown in
-Figure \ref{fig-gmw}, with type signatures for the library functions
-also shown. We present the And gate since it is
-an interesting component. The Figure includes the
-following top-level functions:
-\begin{itemize}
-\item \ttt{encodegmw}: This function encodes client 1's and client 2's
-  secret bits called $\ttt{s[1,"s1"]}$ and $\ttt{s[1,"s2"]}$ into two
-  distinct wire values (pairs of shares).
-\item \ttt{andgmw}: This function defines the gate $\ttt{"g"}$, the
-  identifier $\ttt{"g"}$ being used to distinguish randomness used
-  within.  In our version client 1 builds the output table (using
-  \ttt{andtablegmw})and transfers the correct output share to client 2
-  using 1-out-of-4 OT ($\ttt{OT4}$), definable using $\ttt{OT}$, as
-  per standard GMW protocol.
-\item \ttt{decode}: This function decodes and publishes a wire value
-  by $\ttt{xor}$ing the shares. Note that this requires both client 1
-  and 2 to publicize their shares.
-\end{itemize}
-\begin{example}
-  \label{example-gmw-andcircuit}
-The following program uses our GMW library to define
-a circuit with a single And gate and input secrets $\ttt{s1}$ and
-$\ttt{s2}$ from client's 1 and 2 respectively. 
-{\small
-  \begin{verbatimtab}
-  let ss = encodegmw("s1","s2") in v[0,"output"] := decode(andgmw(0,ss.shares1,ss.shares2)) \end{verbatimtab}
-}
-\end{example}
-
-\subsection{Example: YGC}
-\label{section-metalang-ygc}
-
-\begin{fpfig}[t]{Yao's Garbled Circuit component functions.}{fig-ygc-lib}
-{\footnotesize
-\begin{verbatimtab}
-  owl(gid) { { k = flip[2,"gate:" || gid || ".k"]; p = flip[2,"gate:" || gid || ".p"] } }
-
-  encodeygc(sa, sb) {
-    let owl1 = owl(sa) in
-    let owl2 = owl(sb) in
-    v[1,gate: || sa || 1.k] := OT(s[1,sa],owl1.k,(not owl1.k));
-    v[1,gate: || sa || 1.p] := OT(s[1,sa],owl1.p,(not owl1.p));
-    v[1,gate: || sb || 2.k] := select(s[2,sb],owl2.k,(not owl2.k));
-    v[1,gate: || sb || 2.p] := select(s[2,sb],owl2.p,(not owl2.p));
-    { wv1 = { k = v[1,"gate:" || sa || "1.k"]; p = v[1,"gate:" || sa || "1.p"] };
-      wv2 = { k = v[1,"gate:" || sb || "2.k"]; p = v[1,"gate:" || sb || "2.p"] } }
-  }
-
-  andygc(g, ga, gb, wva, wvb) { andgate(g, ga, gb); evalgate(g, wva, wvb) }
-
-  decodeygc(g, wv) { garbledecode(g); evaldecode(wv) }
-\end{verbatimtab}
-}
-\end{fpfig}
-
-Our YGC implementation follows the \emph{point-and-permute} method
-described in \cite{evans2018pragmatic} and elsewhere, to which the
-reader is referred for more in-depth discussion.  In this
-implementation client 2 is the \emph{garbler} and client 1 is the
-\emph{evaluator}. The garbler builds the garbled tables and shares
-them with the evaluator, who then evaluates the gate in an oblivious
-fashion until the final public output is generated through
-decryption.
-
-\emph{Wire labels} are fundamental to YGC, and essentially represent
-gate output values in an encrypted form. In our definition, wire
-labels are represented by records $\ttt{\{ k = }\beta_1\ttt{; p =
-}\beta_2\ttt{ \}}$, where $\ttt{k}$ is the \emph{key bit} and
-$\ttt{p}$ is the \emph{pointer bit}, and $\beta_1$ and $\beta_2$ are
-flips. Flips in each output wire label are owned by the garbler and
-are unique per gate by definition of their identifying string, and the
-representation of $0$ is the negation of $1$. For example, here is the
-representation of 1 and 0 respectively in the output wire label for a
-hypothetical gate 6:
-{\small
-  \begin{mathpar}
-    \ttt{\{ k = flip[2,"gate:6.k"]; p =  flip[2,"gate:6.p"]] \}}
-
-    \ttt{\{ k = not flip[2,"gate:6.k"]; p =  not flip[2,"gate:6.p"]] \}}
-  \end{mathpar}
-}
-In our implementation, gates are wired together using gate
-identifiers, which are strings $w$. Top-level functionality
-shown in Figure \ref{fig-ygc-lib} includes:
-\begin{itemize}
-\item \ttt{owl}: This function returns the output wire label for any
-  gate identified by \ttt{gid}.
-\item \ttt{andgate}: This defines a subprotocol for the garbler
-  to define a garbled gate $\ttt{gid}$ with input wires from gates
-  $\ttt{ga}$ and $\ttt{gb}$. The garbler generates keys and garbles
-  the rows in YGC fashion, them with client $1$ in
-  views in a standard form. For example, the view for
-  a hypothetical gate 6, row 2 garbled truth table is $\ttt{v[1,gate:6tt2]}$.
-  We note that garbled gates of other binary operators can be obtained with
-  replacement of $\ttt{andtable}$ with appropriate garbled table definitions. 
-\item \ttt{evalgate}: This defines a subprotocol for the evaluator to
-  evaluate gate $\ttt{gid}$ given input wire values $\ttt{wva}$ and
-  $\ttt{wvb}$.
-\item \ttt{andygc}: This composes the garbler and evaluator subprotocols
-  for construction and evaluation of an And gate. 
-\item \ttt{garbledecode} and \ttt{evaldecode}: The former function
-  defines the garbler's protocol for encrypting the circuit
-  output from final gate $\ttt{gid}$, and the latter defines
-  the evaluator's output decryption protocol. These are composed with \ttt{decodeygc}.
-\item \ttt{encodegmw}: This defines the initial phase of the protocol,
-  where the evaluator receives the wire value from their own
-  secret $\sx{1}{sa}$ via $\ttt{OT}$, and the garbler communicates
-  the wire value for their own secret $\sx{2}{sb}$ directly.
-\end{itemize}
-We provide more code details in Appendix \ref{section-ygc}, Figures
-\ref{fig-ygc-aux} and \ref{fig-ygc-gates}, and complete code in our
-online repository \cite{jpdf-github}.
-\begin{example}
-  \label{example-ygc-andcircuit}
-The following program uses our YGC library to define
-a circuit with a single and gate and input secrets $\ttt{s1}$ and
-$\ttt{s2}$ from client's 1 and 2 respectively. 
-{\small
-\begin{verbatimtab}
-  let secrets = encode("s1","s2") in
-  v[0,output] := decode(andgg("0", secrets.wv1, secrets.wv2)) \end{verbatimtab}
-}
-%\begin{verbatimtab}
-%  andgate(0,s1,s2);
-%  garbledecode(0);
-%  let secrets = encode(s1,s2) in
-%  v[0,output] := decode(evalgate(0, secrets.wv1, secrets.wv2))
-%\end{verbatimtab}
-\end{example}

### 2024-05-06 ppdp24/metalang.tex
checkpoint

@@
 
-Large practical MPC computations are based on much larger protocols
-than the examples we've considered so far. These larger protocols are
-typically based on compositional units. Examples include GMW circuits
-and Yao's Garbled Circuits (YGC), which are composed of so-called
-garbled gates.  Languages for defined garbled circuits, beginning with
-Fairplay \cite{269581}, treat gates as compositional units that are
-wired together by the programmer to generate a complete circuit. The
-$\fedprot$ language is low-level and does not include abstractions for
-defining composable elements.
+Practical MPC computations protocols are
+typically composed of compositional units. Examples include GMW circuits
+and Yao's Garbled Circuits (YGC), that are composed of so-called
+garbled gates. Languages such as Fairplay \cite{269581} provide gates as
+units of abstraction that are ``wired'' together by the programmer to
+generate a complete circuit.
 
-In this Section, we introduce the $\metaprot$ language which includes
-structured data and function definitions, which are sufficiently
-expressive to define composable protocol elements such as garbled
-gates. The $\metaprot$ language is a \emph{metalanguage}, in the sense
-that it produces $\fedprot$ protocols as a result of computation. That
-is, $\metaprot$ is a high-level language that generates low-level
-protocol code. We endow $\metaprot$ with a type system that enjoys a
-type safety result guaranteeing safety of generated protocols. We
-consider library implementations for GMW and YGC circuits as
-examples. Additional examples are provided in Appendix
-\ref{section-examples-lang}.
+The $\fedprot$ language is low-level and does not include abstractions for
+defining composable elements. So in this Section we introduce the $\metaprot$
+language which that includes structured data and function definitions
+for defining composable protocol elements at a higher level of abstraction. 
+The $\metaprot$ language is a \emph{metalanguage}, where 
+$\fedprot$ protocols are the residuum of computation.
+
+In addition to these declarative benefits of $\metaprot$, component
+definitions support compositional verification of larger programs.
+The techniques in Section \ref{section-bruteforce} are not feasible
+for large boolean circuits, for example, but they are for verification
+of properties of compositional units, as we discuss in Section
+\ref{section-verification}.
 
@@
 
-When we consider the examples of GMW and YGC in detail below, our
+When we consider larger examples, our
 focus will be on developing a codebase that can be used to define
@@
 strings and identifiers can be constructed manually, and expressions
-can occur inside assignments and boolean expression forms, function
+can occur inside assignments and field expression forms, function
 definitions can generalize over $\fedprot$-level patterns to obtain
@@
 where $\prog$ is the $\minifed$ program accumulated during evaluation.
-In this definition we write $e[v/x]$ to denote the substitution of $v$
-for free occurences of $x$ in $e$. The rules are mostly standard,
+In this definition we write $e[\mv/y]$ to denote the substitution of $\mv$
+for free occurences of $y$ in $e$. The rules are mostly standard,
 except when a concrete $\minifed$ assignment is encountered it is added

### 2024-05-06 ppdp24/metalang.tex
checkpoint

@@
 \ref{fig-metaprot}.  It includes a syntax of function
-definitions and records, and values include client ids, identifier
-strings, and boolean expressions. Expression forms allow dynamic
-construction of boolean expression forms and view assignments. When
-$\metaprot$ programs construct an $\fedprot$ assignment, a side effect
-occurs whereby the assignment is added to the end of the $\fedprot$
-program accumulated during evaluation.
-
-Formally, we consider a complete metaprogram to include both a
+definitions and records, and values $\mv$ include client ids $\cid$, identifier
+strings $w$, and expressions $\be$ in field $\mathbb{F}_p$. Expression
+forms allow dynamic construction of field expressions and $\minifed$ commands.
+The construction of a command has the side-effect of adding to the residual
+$\minifed$ protocol. Formally, we consider a complete metaprogram to include both a
 codebase and a ``main'' program that uses the codebase. 
@@
 A \emph{codebase} $\codebase$ is a list of function 
-declarations. We write $ \codebase(f) = x_1,\ldots,x_n,\ e$
+declarations. We write $ \codebase(f) = y_1,\ldots,y_n,\ e$
 iff $f(x_1,\ldots,x_n) \{ e \} \in \codebase$.
@@
 allowing computation within a larger program context, where $E[e]$
-denotes an expression with $e$ in the hole $[]$ of $E$. Evaluation
-contexts include boolean expression forms, allowing generalization
-and instantiation of compositional program elements.
+denotes an expression with $e$ in the hole $[]$ of $E$. The syntax
+of $E$ imposes a left-to-right order of evaluation of subexpressions
+for all forms.

### 2024-05-06 ppdp24/metalang.tex
checkpoint

@@
 definitions and records, and values $\mv$ include client ids $\cid$, identifier
-strings $w$, and expressions $\be$ in field $\mathbb{F}_p$. Expression
-forms allow dynamic construction of field expressions and $\minifed$ commands.
+strings $w$, expressions $\be$ in field $\mathbb{F}_p$, and the unit value $()$.
+Expression forms allow dynamic construction of field expressions and $\minifed$ commands.
 The construction of a command has the side-effect of adding to the residual

### 2024-05-06 ppdp24/bruteforce.tex
checkpoint

@@
-\section{Verification of Protocols in $\mathbb{F}_2$}
+\section{Automated Verification in $\mathbb{F}_2$}
+\label{section-bruteforce}
 
@@
 ... and so on. Then, to begin converting a protocol $\prot$ to a Datalog
-we define the function ${vtt}$ that applies $\solve{\!}{\!}{\!\!}$ locally to each command
+we define the function ${tt}$ that applies $\solve{\!}{\!}{\!\!}$ locally to each command
 in $\prog$, obtaining the subset of memories that result
@@
 $$
-{vtt}(\xassign{x}{\be}{\cid}) \defeq (x, (\solve{(\mems(\vars\ \be\ \cid))}{\be}{\cid}))
+{tt}(\xassign{x}{\be}{\cid}) \defeq (x, (\solve{(\mems(\vars\ \be\ \cid))}{\be}{\cid}))
 $$
 \end{definition}
-The mapping of ${vtt}$ across a program
-$\prog$-- i.e., $(\mathit{map}\ {vtt}\ \prog)$--  essentially defines the
+The mapping of ${tt}$ across a program
+$\prog$-- i.e., $(\mathit{map}\ {tt}\ \prog)$--  essentially defines the
 logic program, modulo some syntactic conversion. We can
@@
 \end{mathpar}
-Given pairs $(x,\stores)$ in the range of ${vtt}$, we define the conversion
+Given pairs $(x,\stores)$ in the range of ${tt}$, we define the conversion
 to clauses as  $\mathit{clauses}(x,\{ \store_1,...,\store_n \}) \defeq x \gets \logit{\store_1} \vee \cdots \vee x \gets \logit{\store_n}$.
@@
 $$
-\datalog(\prog) \defeq  \mathit{map}\ \mathit{clauses}\ (\mathit{map}\ {vtt}\ \prog)
+\datalog(\prog) \defeq  \mathit{map}\ \mathit{clauses}\ (\mathit{map}\ {tt}\ \prog)
 $$
@@
 that model computation for distinct fact bases can be done in parallel. 
+We have used this method to verify the one-time-pad encryption and Shamir
+addition protocols in Section \ref{section-lang-example}, as well as
+larger protocol components as we discuss in Section \ref{section-verification}.

### 2024-05-06 ppdp24/examples.tex
checkpoint

--- 2023-07-12 popl24/metalang.tex
@@
-\section{The $\metaprot$ Metalanguage}
-\label{section-metalang}
-
-Large practical MPC computations are based on much larger protocols
-than the examples we've considered so far. These larger protocols are
-typically based on compositional units. Examples include GMW circuits
-and Yao's Garbled Circuits (YGC), which are composed of so-called
-garbled gates.  Languages for defined garbled circuits, beginning with
-Fairplay \cite{269581}, treat gates as compositional units that are
-wired together by the programmer to generate a complete circuit. The
-$\fedprot$ language is low-level and does not include abstractions for
-defining composable elements.
-
-In this Section, we introduce the $\metaprot$ language which includes
-structured data and function definitions, which are sufficiently
-expressive to define composable protocol elements such as garbled
-gates. The $\metaprot$ language is a \emph{metalanguage}, in the sense
-that it produces $\fedprot$ protocols as a result of computation. That
-is, $\metaprot$ is a high-level language that generates low-level
-protocol code. We endow $\metaprot$ with a type system that enjoys a
-type safety result guaranteeing safety of generated protocols. We
-consider library implementations for GMW and YGC circuits as
-examples. Additional examples are provided in Appendix
-\ref{section-examples-lang}.
-
-\subsection{Syntax}
-
-The syntax of $\metaprot$ is defined in Figure
-\ref{fig-metaprot}.  It includes a syntax of function
-definitions and records, and values include client ids, identifier
-strings, and boolean expressions. Expression forms allow dynamic
-construction of boolean expression forms and view assignments. When
-$\metaprot$ programs construct an $\fedprot$ assignment, a side effect
-occurs whereby the assignment is added to the end of the $\fedprot$
-program accumulated during evaluation.
-
-Formally, we consider a complete metaprogram to include both a
-codebase and a ``main'' program that uses the codebase. 
-\begin{definition}
-A \emph{codebase} $\codebase$ is a list of function 
-declarations. We write $ \codebase(f) = x_1,\ldots,x_n,\ e$
-iff $f(x_1,\ldots,x_n) \{ e \} \in \codebase$.
-A \emph{metaprogram}, aka \emph{metaprotocol}  is a pair of a 
-codebase and expression $\codebase, e$. We may omit
-$\codebase$ if it is clear from context.  
-\end{definition}
-
-When we consider the examples of GMW and YGC in detail below, our
-focus will be on developing a codebase that can be used to define
-arbitrary circuits, i.e., complete and concrete protocols. Since
-strings and identifiers can be constructed manually, and expressions
-can occur inside assignments and boolean expression forms, function
-definitions can generalize over $\fedprot$-level patterns to obtain
-composable program units.
-
-\subsection{Semantics}
-
-\begin{fpfig}[t]{$\metaprot$ syntax (T), evaluation contexts (M), and operational semantics (B).}{fig-metaprot}
-    {\small
-    $$
-    \begin{array}{rcl@{\hspace{8mm}}r}
-      \multicolumn{3}{l}{\flab \in \mathrm{Field},\   x \in \mathrm{EVar}, \  f \in \mathrm{FName}}\\[2mm]
-      %x &\in& \mathrm{EVar}\\
-      %f &\in& \mathrm{FName}\\[2mm]
-      e &::=& b \mid \flip{e}{e} \mid \secret{e}{e} \mid \view{e}{e} \mid \oracle{e} \mid \enot\ e \mid e\ \eand\ e \mid e\ \exor\ e \mid & \textit{expressions}\\[0mm]
-      & & \select{e}{e}{e} \mid 
-      \send{\view{e}{e}}{e} \mid \send{\view{e}{e}}{\OT{e}{e}{e}} \mid e;e \mid \\[0mm]
-      & & x \mid \elet{x}{e}{e} \mid f(e,\ldots,e) \mid \{ \flab = e; \ldots; \flab = e \}
-      \mid e.\flab \mid e\concat e \mid (e) \\[2mm]
-      v &::=& w \mid \cid \mid \be \mid \{ \flab = v;\ldots;\flab = v \} 
-      \mid \ttt{()} & \textit{values}\\[2mm]
-           {fn} &::=& f(x,\ldots,x) \{ e \} & \textit{functions}
-    \end{array}
-    $$
-    }
-
-    \rule{130mm}{0.5pt}
-
-    {\small
-    $$
-    \begin{array}{rcl@{\hspace{3mm}}r}
-      E &::=& [\,] \mid \enot\ E \mid E\ \bop\ e \mid v\ \bop\ E \mid  \flip{E}{e} \mid \secret{E}{e} \mid \view{E}{e} \mid \oracle{E} \mid  \\[1mm]
-      & & \flip{\cid}{E} \mid \secret{\cid}{E} \mid \view{\cid}{E} \mid \send{E}{e} \mid \send{\view{\cid}{w}}{E} \mid \OT{E}{e}{e} \\[1mm]
-      & & \mid \OT{v}{E}{e} \mid \OT{v}{v}{E} \mid \select{E}{e}{e} \mid \select{v}{E}{e} \mid \\[1mm]
-      & & \select{v}{v}{E} \mid \elet{x}{E}{e} \mid f(v,\ldots,v,E,e,\ldots,e) \mid \\[1mm]
-      & & \{ \flab = v;\ldots;\flab = v;\flab = E;\flab = e;\ldots;\flab = e \} \mid E.\flab \mid E\concat e \mid v \concat E
-    \end{array}
-    $$
-    \vspace{.5mm}
-
-    \rule{130mm}{0.5pt}
-
-    $$
-    \begin{array}{rcl@{\hspace{10mm}}r}
-      \config{\prog}{\elet{x}{v}{e}} &\redx& \config{\prog}{e[v/x]}\\
-      \config{\prog}{f(v_1,...,v_n)} &\redx&
-      \config{\prog}{e[v_1/x_1,\ldots,v_n/x_n]} & 
-      \codebase(f) = x_1,\ldots,x_n,\ e\\
-      \config{\prog}{\{\ldots; \flab = v; \ldots\}.\flab} &\redx&
-      \config{\prog}{v}\\
-      \config{\prog}{w_1\concat w_2} &\redx& \config{\prog}{w_1w_2}\\
-      \config{\prog}{v;e} &\redx& \config{\prog}{e}\\
-      \config{\prog}{\instr} &\redx& \config{\prog;\instr}{()}\\
-      \config{\prog}{E[e]} &\redx& \config{\prog'}{E[e']} & \text{if}\ \config{\prog}{e} \redx \config{\prog'}{e'} 
-    \end{array}
-    $$
-  }
-\end{fpfig}
-
-We define a small-step evaluation aka reduction relation $\redx$ in
-Figure \ref{fig-metaprot}.  We write $\redxs$ to denote the
-reflexive, transitive closure of $\redx$. Reduction is defined on
-\emph{configurations} which are pairs of the form $\config{\prog}{e}$,
-where $\prog$ is the $\minifed$ program accumulated during evaluation.
-In this definition we write $e[v/x]$ to denote the substitution of $v$
-for free occurences of $x$ in $e$. The rules are mostly standard,
-except when a concrete $\minifed$ assignment is encountered it is added
-to the end of $\prog$.
-
-The rules rely on a definition of \emph{evaluation contexts} $E$
-allowing computation within a larger program context, where $E[e]$
-denotes an expression with $e$ in the hole $[]$ of $E$. Evaluation
-contexts include boolean expression forms, allowing generalization
-and instantiation of compositional program elements.
-
-\subsection{Type Theory and Static Type Safety}
-
-It is desirable to statically enforce safety of both $\metaprot$
-programs and the safety of the $\fedprot$ programs they
-generate. Although safety of the latter could be enforced
-post-generation by a direct analysis, for large programs this can be
-much more expensive and it is also better to not waste time on
-resource-intensive compilation of programs with known errors
-\cite{kreuter2012billion}. Some consequences of safety errors, for example, accidental
-reuse of one-time pads can also undermine security.
-
-The type syntax of $\metaprot$ is defined in Figure
-\ref{fig-metaprot-types}. It includes a weak form of dependency:
-string types $\stringty{e}$ and client types $\cidty{e}$
-are parameterized by expressions $e$ that precisely reflect the
-type of the value. Boolean expression forms have the type
-$\bet{e}$ indexed by expressions $e$ indicating the client
-id type of the expression. The dependency is weak in the
-sense that expressions in types are a strict subset of
-expression forms- for any $\stringty{e}$ the expression $e$
-is either a variable, a string, or a concatenation form,
-and for any $\cidty{e}$ or $\bet{e}$ the expression $e$
-is either a variable or a client id $\cid$. We define
-the equivalence $w_1 \concat w_2 \equiv w_1w_2$ in typings.
-
-Type judgements for expressions are of the form
-$\tjudge{\viewst_1}{\gamma}{e}{\viewst_2}$ where the \emph{view
-effect} $\viewst_1$ denotes the views that have been defined so far,
-and $\viewst_2$ records new views defined as the effect of the
-expression on the residual $\minifed$ program.  Type judgements are
-syntax-directed- selected rules are shown in Figure
-\ref{fig-metaprot-types}
-%, with a more complete set provided in
-%Appendix Figure \ref{fig-metaprot-tjudge}
-and a full implementation is available in our online repository \cite{jpdf-github}. The
-$\TirName{AssignT}$ rule captures the effect of new view
-definitions. The $\TirName{And}$ rule illustrates how program safety
-is enforced, by ensuring that subexpressions of boolean expressions
-have the same owner.
-
-The $\TirName{Appt}$ rule applies to function
-definition and application respectively. Function typings rely on the
-definition of function input type annotations $\tas$ and
-type term substitutions $\sigma$. 
-\begin{definition}
-  A \emph{function input type annotation} $\tas$ is a mapping from
-  function names $f$ to type products $\tau_1 * \cdots * \tau_n$.
-  A \emph{type term substitution} $\sigma$ is a mapping from
-  $\minifed$ variables $x$ to values, where $\sigma(\tau)$ denotes
-  the replacement of occurences of $x$ in $\tau$ with $\sigma(x)$. 
-\end{definition}
-We assume that input type annotations $\tas$ are provided by the
-programmer for all function definitions. This guarantees that
-$\metaprot$ type checking is straightforward and efficient.
-Function types are of the form
-$
-(\tau_1 * \cdots * \tau_n \rightarrow \tau,\viewst)
-$,
-where $\viewst$ denotes the effect of the function on the residual
-program.  The function type can be understood as a dependent $\Pi$
-type, with every term variable bound. When applied, these variables
-are instantiated with a type term substitution $\sigma$. In our
-implementation, we adapt \emph{synthesis} as defined in Dependent ML \cite{10.1145/292540.292560}
-to obtain this $\sigma$- essentially this is a match on the syntactic
-structure of types and expressions.
-
-\begin{fpfig}[t]{Type syntax of $\metaprot$ (T) and selected type judgement rules (B).}{fig-metaprot-types}
-{\small
-  $$
-  \begin{array}{rcl@{\hspace{2mm}}r}
-    \srct &::=& \cidty{e} \mid \stringty{e} \mid \bet{e} \mid  \{ \flab : \srct;\ldots;\flab : \srct \} \mid \tau * \cdots * \tau \rightarrow \tau,\viewst & \gdesc{types}\\ [1mm]
-    %&&  \{ \flab : \srct;\ldots;\flab : \srct \} \mid \tau * \cdots * \tau \rightarrow \tau,\viewst \\[1mm]
-    \viewst  &::=& \view{e}{e};\viewst \mid \varnothing   & \gdesc{view effects}\\[1mm]
-    \Gamma &::=& \Gamma; x : \tau \mid \varnothing & \gdesc{type environments}    
-  \end{array}
-  $$
-}
-
-\rule{130mm}{0.5pt}
-
-{\small
-\begin{mathpar}
-\inferrule[\TirName{SecretT}]
-{\tjudge{\viewst}{\Gamma}{e_1}{\cidty{e_1'}}{\viewst_1}\\
-\tjudge{\viewst_1}{\Gamma}{e_2}{\stringty{e_2'}}{\viewst_2}}
-{\tjudge{\viewst}{\Gamma}{\secret{e_1}{e_2}}{\bet{e_1'}}{\views_2}}
-
-\inferrule[\TirName{AndT}]
-{
-\tjudge{\viewst}{\Gamma}{e_1}{\bet{e}}{\viewst_1}\\
-\tjudge{\viewst_1}{\Gamma}{e_2}{\bet{e}}{\viewst_2}
-}
-{\tjudge{\viewst}{\Gamma}{e_1\ \eand\ e_2}{\bet{e}}{\viewst_2}}
-
-\inferrule[\TirName{AssignT}]
-{
-\tjudge{\viewst}{\Gamma}{e_1}{\cidty{e_1'}}{\viewst_1}\\
-\tjudge{\viewst_1}{\Gamma}{e_2}{\stringty{e_2'}}{\viewst_2}\\
-\tjudge{\viewst_2}{\Gamma}{e_3}{\bet{e_3'}}{\viewst_3}
-}
-{
-\tjudge{\viewst}{\Gamma}{\eassign{\view{e_1}{e_2}}{e_3}}{\unity}{(\viewst_3 ; \view{e_1'}{e_2'} )}
-}
-
-\inferrule[AppT]
-{\Gamma(f) =  
- \tau_1 * \cdots * \tau_n \rightarrow \tau, \viewst_f \\ 
- \tjudge{\viewst}{\Gamma}{e_1}{\sigma(\tau_1)}{\viewst_1}
- \ \cdots\  
- \tjudge{\viewst_{n-1}}{\Gamma}{e_n}{\sigma(\tau_n)}{\viewst_n}}
-{\tjudge{\viewst}{\Gamma}{f(e_1,\ldots,e_n)}{\sigma(\tau)}{(\viewst_n ; \sigma(\viewst_f))}}
-\end{mathpar}
-}
-\end{fpfig}
-
-Top-level type judgements are of the form $\Gamma \vdash \codebase, e
-: \tau, V$, where all the functions in $\codebase$ are well-typed in
-$\Gamma$, the top level view effect $V$ is a set of concrete
-$\fedprot$ views which are constructed by a disjoint union of the views
-in the effect of $e$- the disjointness requirement guarantees that
-views are uniquely defined. Our $\metaprot$ type safety result is
-formulated as follows. In addition to safe execution of the
-metaprogram, it also guarantees safety of the residual $\fedprot$
-program (Definition \ref{definition-safety}).
-\begin{theorem}[$\metaprot$ Type Safety]
-  \label{theorem-metalang-safety}
-  Given $\codebase$, $e$, and $\Gamma$ with $\Gamma \vdash \codebase,e : \unity : V$,
-  then $\config{\varnothing}{e} \redxs \config{\prog}{\ttt{()}}$ where
-  $\prog$ is safe with $\iov(\prog) = S \cup V$ for some $S$.
-\end{theorem}
+\section{Examples}
 
@@
 \end{example}
-
-\subsection{Example: YGC}
-\label{section-metalang-ygc}
-
-\begin{fpfig}[t]{Yao's Garbled Circuit component functions.}{fig-ygc-lib}
-{\footnotesize
-\begin{verbatimtab}
-  owl(gid) { { k = flip[2,"gate:" || gid || ".k"]; p = flip[2,"gate:" || gid || ".p"] } }
-
-  encodeygc(sa, sb) {
-    let owl1 = owl(sa) in
-    let owl2 = owl(sb) in
-    v[1,gate: || sa || 1.k] := OT(s[1,sa],owl1.k,(not owl1.k));
-    v[1,gate: || sa || 1.p] := OT(s[1,sa],owl1.p,(not owl1.p));
-    v[1,gate: || sb || 2.k] := select(s[2,sb],owl2.k,(not owl2.k));
-    v[1,gate: || sb || 2.p] := select(s[2,sb],owl2.p,(not owl2.p));
-    { wv1 = { k = v[1,"gate:" || sa || "1.k"]; p = v[1,"gate:" || sa || "1.p"] };
-      wv2 = { k = v[1,"gate:" || sb || "2.k"]; p = v[1,"gate:" || sb || "2.p"] } }
-  }
-
-  andygc(g, ga, gb, wva, wvb) { andgate(g, ga, gb); evalgate(g, wva, wvb) }
-
-  decodeygc(g, wv) { garbledecode(g); evaldecode(wv) }
-\end{verbatimtab}
-}
-\end{fpfig}
-
-Our YGC implementation follows the \emph{point-and-permute} method
-described in \cite{evans2018pragmatic} and elsewhere, to which the
-reader is referred for more in-depth discussion.  In this
-implementation client 2 is the \emph{garbler} and client 1 is the
-\emph{evaluator}. The garbler builds the garbled tables and shares
-them with the evaluator, who then evaluates the gate in an oblivious
-fashion until the final public output is generated through
-decryption.
-
-\emph{Wire labels} are fundamental to YGC, and essentially represent
-gate output values in an encrypted form. In our definition, wire
-labels are represented by records $\ttt{\{ k = }\beta_1\ttt{; p =
-}\beta_2\ttt{ \}}$, where $\ttt{k}$ is the \emph{key bit} and
-$\ttt{p}$ is the \emph{pointer bit}, and $\beta_1$ and $\beta_2$ are
-flips. Flips in each output wire label are owned by the garbler and
-are unique per gate by definition of their identifying string, and the
-representation of $0$ is the negation of $1$. For example, here is the
-representation of 1 and 0 respectively in the output wire label for a
-hypothetical gate 6:
-{\small
-  \begin{mathpar}
-    \ttt{\{ k = flip[2,"gate:6.k"]; p =  flip[2,"gate:6.p"]] \}}
-
-    \ttt{\{ k = not flip[2,"gate:6.k"]; p =  not flip[2,"gate:6.p"]] \}}
-  \end{mathpar}
-}
-In our implementation, gates are wired together using gate
-identifiers, which are strings $w$. Top-level functionality
-shown in Figure \ref{fig-ygc-lib} includes:
-\begin{itemize}
-\item \ttt{owl}: This function returns the output wire label for any
-  gate identified by \ttt{gid}.
-\item \ttt{andgate}: This defines a subprotocol for the garbler
-  to define a garbled gate $\ttt{gid}$ with input wires from gates
-  $\ttt{ga}$ and $\ttt{gb}$. The garbler generates keys and garbles
-  the rows in YGC fashion, them with client $1$ in
-  views in a standard form. For example, the view for
-  a hypothetical gate 6, row 2 garbled truth table is $\ttt{v[1,gate:6tt2]}$.
-  We note that garbled gates of other binary operators can be obtained with
-  replacement of $\ttt{andtable}$ with appropriate garbled table definitions. 
-\item \ttt{evalgate}: This defines a subprotocol for the evaluator to
-  evaluate gate $\ttt{gid}$ given input wire values $\ttt{wva}$ and
-  $\ttt{wvb}$.
-\item \ttt{andygc}: This composes the garbler and evaluator subprotocols
-  for construction and evaluation of an And gate. 
-\item \ttt{garbledecode} and \ttt{evaldecode}: The former function
-  defines the garbler's protocol for encrypting the circuit
-  output from final gate $\ttt{gid}$, and the latter defines
-  the evaluator's output decryption protocol. These are composed with \ttt{decodeygc}.
-\item \ttt{encodegmw}: This defines the initial phase of the protocol,
-  where the evaluator receives the wire value from their own
-  secret $\sx{1}{sa}$ via $\ttt{OT}$, and the garbler communicates
-  the wire value for their own secret $\sx{2}{sb}$ directly.
-\end{itemize}
-We provide more code details in Appendix \ref{section-ygc}, Figures
-\ref{fig-ygc-aux} and \ref{fig-ygc-gates}, and complete code in our
-online repository \cite{jpdf-github}.
-\begin{example}
-  \label{example-ygc-andcircuit}
-The following program uses our YGC library to define
-a circuit with a single and gate and input secrets $\ttt{s1}$ and
-$\ttt{s2}$ from client's 1 and 2 respectively. 
-{\small
-\begin{verbatimtab}
-  let secrets = encode("s1","s2") in
-  v[0,output] := decode(andgg("0", secrets.wv1, secrets.wv2)) \end{verbatimtab}
-}
-%\begin{verbatimtab}
-%  andgate(0,s1,s2);
-%  garbledecode(0);
-%  let secrets = encode(s1,s2) in
-%  v[0,output] := decode(evalgate(0, secrets.wv1, secrets.wv2))
-%\end{verbatimtab}
-\end{example}

### 2024-05-06 ppdp24/gmw.tex
checkpoint

--- 2023-11-05 revision-1/gmw.tex
@@
     encodegmw(in1,in2) {
-      v[1, in1 || "out"] := s[1,in1] xor flip[1,in1];
-      v[2, in1 || "out"] := flip[1,in1];
-      v[1, in2 || "out"] := s[2,in2] xor flip[2,in2];
-      v[2, in2 || "out"] := flip[2,in2]
-      { shares1 = { c1 = v[1, in1 || "out"]; c2 = v[2, in1 || "out"] };
-        shares2 = { c1 = v[1, in2 || "out"]; c2 = v[2, in2 || "out"]} } 
+      v[1, in1++"out"] := s[1,in1] xor flip[1,in1];
+      v[2, in1++"out"] := flip[1,in1];
+      v[1, in2++"out"] := s[2,in2] xor flip[2,in2];
+      v[2, in2++"out"] := flip[2,in2]
+      { shares1 = { c1 = v[1, in1++"out"]; c2 = v[2, in1++"out"] };
+        shares2 = { c1 = v[1, in2++"out"]; c2 = v[2, in2++"out"]} } 
     }
@@
     andgmw(g, shares1, shares2) {
-      let r = flip[1,g || ".r"] in
+      let r = flip[1,g++".r"] in
       let table = andtablegmw(shares1.c1, shares2.c1) in
@@
       let r00 =  r xor table.v4 in
-      v[2,g || "out"] := OT4(shares1.c2, shares2.c2, r11, r10, r01, r00);
-      v[1,g || "out"] := r;
-      { c1 = v[1,g || "out"]; c2 = v[2,g || "out"]}
+        v[2,g++"out"] := OT4(shares1.c2, shares2.c2, r11, r10, r01, r00);
+      v[1,g++"out"] := r;
+      { c1 = v[1,g++"out"]; c2 = v[2,g++"out"]}
     }
@@
 }
-
-\rule{130mm}{0.5pt}
-
-{\footnotesize
-  \begin{verbatimtab}
-   encodegmw   : string(gid) * string(gid)
-
-   andtablegmw : { k = bool[i]; p = bool[i] }
-
-   andgmw      : string(gid) *  { c1 = bool[1]; c2 = bool[2] } * { c1 = bool[1]; c2 = bool[2] }
-
-   decodegmw   : { c1 = bool[1]; c2 = bool[2] }  \end{verbatimtab}
-}
 \end{fpfig}

### 2024-05-08 ppdp24/examples.tex
checkpoint

@@
 
-\subsection{Example: 2-Party GMW}
+\subsection{2-Party GMW}
 \label{section-metalang-gmw}
@@
 \end{example}
+
+\subsection{2-Party BDOZ}
+
+\input{bdoz}
+
+\begin{definition}[Cheating Detection]
+  Given $\store \in \aruns(\prog)$ with $\iov(\prog) = S \cup V \cup O$,
+  let $X_H = \{ x \mid x \in (\houtputs \cup O_H) \wedge x \in \dom(\store) \}$,
+  and let $X_C$ be the adversarial inputs to $X_H$.
+  Then \emph{cheating is detected} in $\prog$ iff there exists $\sigma'\in \runs(\prog)$
+  with $\store_{X_C} = \store'_{X_C}$.  
+\end{definition}
+
+\begin{lemma}
+  If cheating is detected in $\prog$, then $\prog$ has integrity.
+\end{lemma}

### 2024-05-09 ppdp24/bruteforce.tex
checkpoint

@@
 
+\begin{comment}
+We immediately note that $\runs(\prog)$ can be obtained by a left-folding
+of $\solvealg$ across $\prog$. 
+\begin{lemma}
+  \label{lemma-cruns}
+  Given $\prog$ where $\iov(\prog) = S \cup V \cup O$ and $\flips(\prog) = R$. Define:
+  \begin{eqnarray*}
+    {tt}\ \ \stores\ (\xassign{x}{\be}{\cid}) &\defeq& \begin{array}{l}
+      \mathrm{let}\ \stores' = \solve{\stores}{\be} \ \mathrm{in}\\
+      \ \ \{\extend{\store}{\view{\cid}{w}}{1} \mid \store \in \stores' \}\ \cup\\
+      \ \ \{\extend{\store}{\view{\cid}{w}}{0} \mid \store \in \stores - \stores' \}\end{array}\\[2mm]
+    \cruns(\prog) &\defeq& \mathit{foldl}\ {tt}\ \mems(S \cup F)\ \prog
+  \end{eqnarray*}
+  Then $\cruns(\prog) = \runs(\prog)$.
+\end{lemma}
+\end{comment}
+
 We define the syntax of Datalog as follows. As per standard
@@
 \begin{array}{rclr}
-  \mathit{body} &::=&  x \mid \neg x \mid x \wedge \mathit{body} \mid \neg x \wedge \mathit{body} \mid \varnothing \\
+  \mathit{body} &::=&  x \mid \neg x \mid x \wedge \mathit{body} \mid \neg x \wedge \mathit{body} \\
   \mathit{clause} &::=& x \gets \mathit{body} \mid x \gets
@@
 that model computation for distinct fact bases can be done in parallel.
-We have used this method to verify the one-time-pad encryption and Shamir
-addition protocols in Section \ref{section-lang-example}, as well as
-larger protocol components as we discuss in Section \ref{section-verification}.
+
+We can then query representations of $\runs(\prog)$ with
+implementations of conditioning and marginalization to automatically
+verify the passive model hyperproperties described in Section
+\ref{section-hyper}-- in particular correctness and noninterference
+modulo output, which imply passive security.  We have used this method
+to verify security in $\mathbb{F}_2$ of Shamir addition as defined in Section
+\ref{section-lang-example}, and single-gate $\eand$ and $\exor$ circuits with GMW
+Beaver triples. We have also
+verified properties compositional properties of gates themselves that establish
+circuit invariants that imply passive security in larger circuits.
+We discuss these examples and proof methods in Section \ref{section-examples}.

### 2024-05-09 ppdp24/examples.tex
checkpoint

--- 2024-05-06 ppdp24/examples.tex
@@
 \section{Examples}
+\label{section-examples}
 
-\subsection{Example: 2-Party GMW}
+\subsection{2-Party GMW}
 \label{section-metalang-gmw}
@@
 \end{example}
+
+\subsection{2-Party BDOZ}
+
+\input{bdoz}
+
+\begin{definition}
+  Given $\prog$ with $\iov(\prog) = S \cup V \cup O$,
+  let $X_H \subseteq \{ x \mid x \in (\houtputs \cup O_H) \wedge x \in \dom(\store) \}$.
+  Then the \emph{adversarial inputs to $X_H$} is the least set
+  $X_C \subseteq \cinputs$ such that $\progtt(\prog) \not\vdash X_C * X_H$.
+\end{definition}
+\begin{definition}[Cheating Detection]
+  \emph{Cheating is detected} in $\prog$ with $\iov(\prog) = S \cup V \cup O$ iff
+  for all  $\store \in \aruns(\prog)$,
+  letting $X_H = \{ x \mid x \in (\houtputs \cup O_H) \wedge x \in \dom(\store) \}$,
+  and letting $X_C$ be the adversarial inputs to $X_H$,
+  there exists $\sigma'\in \runs(\prog)$
+  with $\store_{X_C} = \store'_{X_C}$.  
+\end{definition}
+
+\begin{lemma}
+  If cheating is detected in $\prog$, then $\prog$ has integrity.
+\end{lemma}

### 2024-05-09 ppdp24/gmw.tex
checkpoint

@@
-\begin{fpfig}[t]{2-Party GMW circuit library with And gate (T), and type annotations (B).}{fig-gmw}
+\begin{fpfig}[t]{2-Party GMW circuit library with And gate.}{fig-gmw}
 {\footnotesize
@@
     encodegmw(in1,in2) {
-      v[1, in1++"out"] := s[1,in1] xor flip[1,in1];
-      v[2, in1++"out"] := flip[1,in1];
-      v[1, in2++"out"] := s[2,in2] xor flip[2,in2];
-      v[2, in2++"out"] := flip[2,in2]
+      m[in1++"out"]@1 := (s[in1] xor r[in1])@1;
+      m[in1++"out"]@2 := r[in1]@1;
+      m[in2++"out"]@1 := (s[in2] xor flip[in2])@2;
+      m[in2++"out"]@2 := r[in2]@2
       { shares1 = { c1 = v[1, in1++"out"]; c2 = v[2, in1++"out"] };

### 2024-05-09 ppdp24/metalang.tex
checkpoint

@@
 The $\metaprot$ language is a \emph{metalanguage}, where 
-$\fedprot$ protocols are the residuum of computation.
+$\fedprot$ protocols are the residuum of computation. In addition to these
+declarative benefits of $\metaprot$, component
+definitions support compositional verification of larger programs as we will
+discuss with examples in Section
+\ref{section-examples}.
 
-In addition to these declarative benefits of $\metaprot$, component
-definitions support compositional verification of larger programs.
-The techniques in Section \ref{section-bruteforce} are not feasible
-for large boolean circuits, for example, but they are for verification
-of properties of compositional units, as we discuss in Section
-\ref{section-verification}.
+\metaprotfig
 
@@
 declarations. We write $ \codebase(f) = y_1,\ldots,y_n,\ e$
-iff $f(x_1,\ldots,x_n) \{ e \} \in \codebase$.
+iff $f(y_1,\ldots,y_n) \{ e \} \in \codebase$.
 A \emph{metaprogram}, aka \emph{metaprotocol}  is a pair of a 
@@
 composable program units.
-
-\metaprotfig
-
 \subsection{Semantics}

### 2024-05-09 ppdp24/gmw.tex
checkpoint

@@
   \begin{verbatimtab}
-    encodegmw(in1,in2) {
-      m[in1++"out"]@1 := (s[in1] xor r[in1])@1;
-      m[in1++"out"]@2 := r[in1]@1;
-      m[in2++"out"]@1 := (s[in2] xor flip[in2])@2;
-      m[in2++"out"]@2 := r[in2]@2
-      { shares1 = { c1 = v[1, in1++"out"]; c2 = v[2, in1++"out"] };
-        shares2 = { c1 = v[1, in2++"out"]; c2 = v[2, in2++"out"]} } 
+    encodegmw(in, i1, i2) {
+      m[in]@i2 := (s[in] xor r[in])@i2;
+      m[in]@i1 := r[in1]@2;
+      m[in]
     }
 
-    andtablegmw(b1, b2) {
-      let r11 = (b1 xor true) and (b2 xor true) in
-      let r10 = (b1 xor true) and (b2 xor false) in
-      let r01 = (b1 xor false) and (b2 xor true) in
-      let r00 = (bl xor false) and (b2 xor false) in
+    andtablegmw(b1, b2, r) {
+      let r11 = r xor (b1 xor true) and (b2 xor true) in
+      let r10 = r xor (b1 xor true) and (b2 xor false) in
+      let r01 = r xor (b1 xor false) and (b2 xor true) in
+      let r00 = r xor (bl xor false) and (b2 xor false) in
       { v1 = r11; v2 = r10; v3 = r01; v4 = r00 }
@@
 
-    andgmw(g, shares1, shares2) {
-      let r = flip[1,g++".r"] in
-      let table = andtablegmw(shares1.c1, shares2.c1) in
-      let r11 =  r xor table.v1 in
-      let r10 =  r xor table.v2 in
-      let r01 =  r xor table.v3 in
-      let r00 =  r xor table.v4 in
-        v[2,g++"out"] := OT4(shares1.c2, shares2.c2, r11, r10, r01, r00);
-      v[1,g++"out"] := r;
-      { c1 = v[1,g++"out"]; c2 = v[2,g++"out"]}
+    andgmw(g, v1, v2) {
+      let r = r[g] in
+      let table = andtablegmw(v1,v2,r) in
+      m[g]@2 := OT4(v1,v2,table,2,1);
+      m[g]@1 := r;
+      m[g]
     }
 
-    decodegmw(shares) { v[0,"output"] := shares.c1 xor shares.c2 }   \end{verbatimtab}
+    decodegmw(v) {
+      p[1] := v@1; p[2] := v@2;
+      out@1 := (p[1] + [2])@1;
+      out@2 :=(p[1] + [2])@2
+    }
+  \end{verbatimtab}
 }

### 2024-05-09 ppdp24/examples.tex
checkpoint

@@
 a circuit with a single And gate and input secrets $\ttt{s1}$ and
-$\ttt{s2}$ from client's 1 and 2 respectively. 
-{\small
-  \begin{verbatimtab}
-  let ss = encodegmw("s1","s2") in v[0,"output"] := decode(andgmw(0,ss.shares1,ss.shares2)) \end{verbatimtab}
-}
+$\ttt{s2}$ from client's 1 and 2 respectively:
+$$
+\mathtt{decode(andgmw("gate0",encodegmw("s1"),encodegmw("s2")))}
+$$
 \end{example}

### 2024-05-10 ppdp24/bruteforce.tex
checkpoint

@@
 
-We can then query representations of $\runs(\prog)$ with
+\paragraph{Verifying Security Properties} We can
+query correct representations of $\runs(\prog)$ by using
 implementations of conditioning and marginalization to automatically
@@
 circuit invariants that imply passive security in larger circuits.
-We discuss these examples and proof methods in Section \ref{section-examples}.
+We discuss these examples and proof methods in Section
+\ref{section-examples}.

### 2024-05-10 ppdp24/examples.tex
checkpoint

@@
 $\ttt{s2}$ from client's 1 and 2 respectively:
+\begin{verbatimtab}
+decode(andgmw("gate0",encodegmw("s1"),encodegmw("s2")))
+\end{verbatimtab}
+\end{example}
+
+\begin{lemma}
+  Given:
   $$
-\mathtt{decode(andgmw("gate0",encodegmw("s1"),encodegmw("s2")))}
+  (\varnothing,\andgmw(xy,\mesg{x},\mesg{y}) \redx (\prog,\mesg{xy})
   $$
-\end{example}
+  Then for all
+  $$
+  \store \in \mems(\{ \elab{\secret{y}}{2}, \out{1} \} \cup \iov(\prog))
+  $$
+  such that
+  \begin{mathpar}
+    %\store(\elab{\secret{x}}{1}) = \elab{\mesg{x}}{1} + \elab{\mesg{x}}{2}
+    %
+    \store(\elab{\secret{y}}{2}) = \cod{\store, \elab{\mesg{y}}{1} + \elab{\mesg{y}}{2}}
+
+    \store(\out{1}) = \cod{\store, \elab{\mesg{xy}}{1} + \elab{\mesg{xy}}{2}}
+  \end{mathpar}
+  we have:
+  $$
+  \progtt(\prog)(\elab{\mesg{x}}{1} + \elab{\mesg{x}}{2}|\store_{\{ \secret{y}{2}, \out{1} \}}) =
+  \progtt(\prog)(\elab{\mesg{x}}{1} + \elab{\mesg{x}}{2}|\store)
+  $$
+\end{lemma}
 

### 2024-05-10 ppdp24/gmw.tex
checkpoint

@@
       m[in]@i2 := (s[in] xor r[in])@i2;
-      m[in]@i1 := r[in1]@2;
+      m[in]@i1 := r[in]@2;
       m[in]
@@
       let r00 = r xor (bl xor false) and (b2 xor false) in
-      { v1 = r11; v2 = r10; v3 = r01; v4 = r00 }
+      { row1 = r11; row2 = r10; row3 = r01; row4 = r00 }
     }
@@
       m[g]@2 := OT4(v1,v2,table,2,1);
-      m[g]@1 := r;
+      m[g]@1 := r@1;
       m[g]
@@
     decodegmw(v) {
-      p[1] := v@1; p[2] := v@2;
-      out@1 := (p[1] + [2])@1;
-      out@2 :=(p[1] + [2])@2
+      p["1"] := v@1; p["2"] := v@2;
+      out@1 := (p["1"] + p["2"])@1;
+      out@2 := (p["1"] + p["2"])@2
     }

### 2024-05-10 ppdp24/examples.tex
checkpoint

--- 2024-05-09 ppdp24/examples.tex
@@
 $\ttt{s2}$ from client's 1 and 2 respectively:
+\begin{verbatimtab}
+decode(andgmw("gate0",encodegmw("s1"),encodegmw("s2")))
+\end{verbatimtab}
+\end{example}
+
+\begin{comment}
+\begin{lemma}
+  Given:
   $$
-\mathtt{decode(andgmw("gate0",encodegmw("s1"),encodegmw("s2")))}
+  \begin{array}{c}
+  (\varnothing,\decodegmw(\andgmw(z,\flip{x},\flip{y})) \redxs \\
+  (\prog_1,\decodegmw(\mesg{z})) \redxs (\prog_2,())
+  \end{array}
   $$
-\end{example}
+  Define:
+  \begin{mathpar}
+    X \defeq
+    \{ \elab{\flip{x}}{2}, \elab{\flip{y}}{1}, \elab{\flip{y}}{2}, \rvl{1}, \rvl{2} \}
+
+    O \defeq  \{ \out{1}, \out{2} \}
+  \end{mathpar}
+  Then:
+  $$
+  \sep{\progtt(\prog_1)}{\{ \elab{\mesg{z}}{2} \}}{\{ \elab{\flip{x}}{1}, \elab{\flip{y}}{1}\}}
+  $$
+  and:
+  $$
+  \margd{\progtt(\prog_2)}{(\{ \elab{\flip{x}}{1} \}|\store_{X \cup O})} =
+    \margd{\progtt(\prog_2)}{(\{ \elab{\flip{x}}{1} \}|\store_O)}
+  $$
+\end{lemma}
+
+\begin{lemma}
+  Given:
+  $$
+  \begin{array}{c}
+  (\varnothing,\decodegmw(\andgmw(z,\flip{x},\flip{y})) \redxs \\
+  (\prog_1,\decodegmw(\mesg{z})) \redxs (\prog_2,())
+  \end{array}
+  $$
+  Define:
+  \begin{mathpar}
+    X \defeq
+    \{ \elab{\flip{x}}{2}, \elab{\flip{y}}{1}, \elab{\flip{y}}{2}, \rvl{1}, \rvl{2} \}
+
+    O \defeq  \{ \out{1}, \out{2} \}
+  \end{mathpar}
+  Then:
+  %$$
+  %\sep{\progtt(\prog_1)}{\{ \elab{\mesg{z}}{2} \}}{\{ \elab{\flip{x}}{1}, \elab{\flip{y}}{1}\}}
+  %$$
+  %and:
+  $$
+  \begin{array}{c}
+    \margd{\progtt(\prog_2)}{(\cod{\store(\elab{\flip{x}}{1})\ \exor\ \store(\elab{\flip{x}}{2})}|\store_{X \cup O})} =\\
+    \margd{\progtt(\prog_2)}{(\cod{\store(\elab{\flip{x}}{1})\ \exor\ \store(\elab{\flip{x}}{2})}|\store_O)}
+  \end{array}
+  $$
+  $$
+  \sep{\progtt(\prog_2)}{(\cod{\store(\elab{\flip{x}}{1})\ \exor\ \store(\elab{\flip{x}}{2})}) \cup O}{X}
+  $$
+  $$
+  \sep{\progtt(\prog_2)}{\{ \macgv{x} \} \cup O}{X}
+  $$
+\end{lemma}
+\end{comment}
+
+\begin{lemma}
+  Given:
+  $$
+  \begin{array}{c}
+  (\varnothing,\andgmw(z,\mesg{x},\mesg{y}) \redxs %\\
+  (\prog,\mesg{z})
+  \end{array}
+  $$
+  Define:
+  \begin{mathpar}
+    X \defeq
+    \{ \elab{\mesg{x}}{2}, \macgv{\mesg{y}}, \elab{\mesg{z}}{1}, \elab{\mesg{z}}{2} \}
+  \end{mathpar}
+  Then:
+  $$
+  \sep{\progtt(\prog)}{\{ \macgv{\mesg{x}}, \macgv{\mesg{z}} \}}{X}
+  $$
+\end{lemma}
+
+\begin{lemma}
+  $$ (\varnothing,e) \redxs (\prog,\decodegmw(E[\mesg{x}])) $$
+  $$\iov(\prog) = S \cup M$$
+  $$\sep{\progtt(\prog)}{(S_H \cup \{ \macgv{\mesg{x}} \})}{(M_C \cup \{ \elab{\mesg{x}}{1}, \elab{\mesg{y}}{2} \})}$$
+\end{lemma}
 

### 2024-05-11 ppdp24/examples.tex
checkpoint

--- 2024-05-09 ppdp24/examples.tex
@@
 $\ttt{s2}$ from client's 1 and 2 respectively:
+\begin{verbatimtab}
+decode(andgmw("gate0",encodegmw("s1"),encodegmw("s2")))
+\end{verbatimtab}
+\end{example}
+
+\begin{lemma}
+  If $\condp{\pmf}{X_1}{\detx{X_2}}$ and
+  $\condp{\pmf}{X_2}{\detx{X_3}}$ then $\condp{\pmf}{X_1}{\detx{X_3}}$.
+\end{lemma}
+
+\begin{lemma}
+  Given:
   $$
-\mathtt{decode(andgmw("gate0",encodegmw("s1"),encodegmw("s2")))}
+  \begin{array}{c}
+  (\varnothing,\andgmw(z,\mesg{x},\mesg{y}) \redxs %\\
+  (\prog,\mesg{z})
+  \end{array}
   $$
-\end{example}
+  Then:
+  $$
+  \condsep
+      {\progtt(\prog)}
+      {\{ \macgv{\mesg{x}},\macgv{\mesg{y}} \}}
+      {\{ \macgv{\mesg{z}} \}}
+      {\{ \elab{\mesg{z}}{1},\elab{\mesg{z}}{2} \}}
+  $$
+  $$
+  \conddetx{\progtt(\prog)}{\{ \macgv{\mesg{x}},\macgv{\mesg{y}} \}}{\{ \macgv{\mesg{z}} \}}    
+  $$
+\end{lemma}
+
+\begin{lemma}
+  $$ (\varnothing,e) \redxs (\prog,\decodegmw(E[\mesg{z}])) $$
+  $$\iov(\prog) = S \cup M$$
+  $$\condsep{\progtt(\prog)}{S}{\{\macgv{\mesg{z}}\}}{(M_C \cup \{ \elab{\mesg{z}}{1}, \elab{\mesg{z}}{2} \})}$$
+  $$\conddetx{\progtt(\prog)}{S}{\{\macgv{\mesg{z}}\}}$$
+\end{lemma}
+
+\begin{mathpar}
+  \condsep{\progtt(\prog_1)}{S^1}{\{ \macgv{\mesg{x}} \}}{(M^1_C \cup \{ \elab{\mesg{x}}{1}, \elab{\mesg{x}}{2} \})}
+
+  \condsep{\progtt(\prog_2)}{S^2}{\{ \macgv{\mesg{y}} \}}{(M^2_C \cup \{ \elab{\mesg{y}}{1}, \elab{\mesg{y}}{2} \})}
+
+  \condsep{\progtt(\prog_3)}{S^1 \cup S^2}{\{ \macgv{\mesg{z}} \}}{(M^1_C \cup M^2_C \cup \{ \elab{\mesg{x}}{2}, \elab{\mesg{y}}{2}, \elab{\mesg{z}}{1}, \elab{\mesg{z}}{2} \})}
+\end{mathpar}
 

### 2024-05-12 ppdp24/examples.tex
checkpoint

@@
 
+\begin{definition}
+  For all $\alpha \in \{ \mesg{w}, \flip{w} \}$ define:
+  $$\pmf(\macgv{\alpha} = v) \defeq \sum_{\sigma) \in A} \pmf(\sigma')$$
+  and define:
+  $$\condd{\pmf}{X}{\macgv{\alpha} = v}(\sigma) \defeq  \sum_{\sigma' \in A} \pmf(\sigma \cap \sigma')/\pmf(\sigma')$$
+  where $A = \{ \store \in \mems(\{ \elab{\alpha}{1},\elab{\alpha}{2} \} ) \mid
+      \cod{\store(\elab{\alpha}{1}) + \store(\elab{\alpha}{2})} = v \}$.
+\end{definition}
+
 \begin{lemma}
@@
   $\condp{\pmf}{X_2}{\detx{X_3}}$ then $\condp{\pmf}{X_1}{\detx{X_3}}$.
+\end{lemma}
+
+\begin{lemma}
+  If $\condsep{\pmf}{X_1}{X_2}{X_3}$ and
+  $\condp{\pmf}{X_2}{\detx{X_4}}$ then $\condsep{\pmf}{X_1}{X_4}{X_3}$.
 \end{lemma}
@@
       {\{ \macgv{\mesg{z}} \}}
-      {\{ \elab{\mesg{z}}{1},\elab{\mesg{z}}{2} \}}
+      {\{ \elab{\mesg{x}}{2},\elab{\mesg{y}}{2},\elab{\mesg{z}}{1},\elab{\mesg{z}}{2} \}}
   $$

### 2024-05-12 ppdp24/examples.tex
checkpoint

@@
   and define:
-  $$\condd{\pmf}{X}{\macgv{\alpha} = v}(\sigma) \defeq  \sum_{\sigma' \in A} \pmf(\sigma \cap \sigma')/\pmf(\sigma')$$
+  $$\condd{\pmf}{X}{\macgv{\alpha} = v}(\sigma) \defeq  \sum_{\sigma' \in A} \condd{\pmf}{X}{\sigma'}(\sigma)$$
   where $A = \{ \store \in \mems(\{ \elab{\alpha}{1},\elab{\alpha}{2} \} ) \mid

### 2024-05-12 ppdp24/examples.tex
checkpoint

@@
   For all $\alpha \in \{ \mesg{w}, \flip{w} \}$ define:
-  $$\pmf(\macgv{\alpha} = v) \defeq \sum_{\sigma) \in A} \pmf(\sigma')$$
+  $$\pmf(\macgv{\alpha} = v) \defeq \sum_{\sigma \in A} \pmf(\sigma')$$
   and define:

### 2024-05-12 ppdp24/examples.tex
checkpoint

@@
 
-The GMW protocol uses secret sharing to represent data flowing through
-circuits. In the 2-party case, clients 1 and 2 each share their input
-secrets, and use those shares to represent inputs to gates. Outputs
-are also represented as shares. We refer to the pair of shares
-representing any particular value as a \emph{wire value}, and
-we represent them via records of the form
-$
-\ttt{\{c1 = } v_1\ttt{;c2 = } v_2\ttt{\}} 
-$
-where $v_1$ and $v_2$ are client 1 and 2's shares respectively.
+The GMW protocol is a garbled binary circuit protocol. In this section
+we will assume the 2-party version, though it generalizes to $n$
+parties\cite{XXX}. GMW uses a common technique in MPC, which is to
+represent values $v$ as distributed shares $v_1$ and $v_2$ with
+$v = v_1 \fplus v_2$. This trick maintains secrecy of $v$
+from both parties, and in GMW it is used to maintain the intermediate
+values of internal gate outputs in circuits. In related literature
+the notation $\macgv{x}$ is used to represent the ``true''
+value of $x$ and $[x]$ is often used to represent the share of
+given party.
+
+To capture this convention, which is used in many other protocols, we
+introduce a new naming convention for ``global view'' elements
+$\macgv{\mesg{w}}$, which denote the summed value of
+$\elab{\mesg{w}}{1}$ and $\elab{\mesg{w}}{2}$ in a protocol
+run. This concept integrates program distributions in the
+usual manner, as the probability of the outcome of summation
+of two variables in the distribution.
+\begin{definition}
+  For all $\mesg{w}$ define:
+  $$\pmf(\macgv{\mesg{w}} = v) \defeq \sum_{\sigma \in A} \pmf(\sigma)$$
+  and define:
+  $$\condd{\pmf}{X}{\macgv{\mesg{w}} = v}(\sigma) \defeq  \sum_{\sigma' \in A} \condd{\pmf}{X}{\sigma'}(\sigma)$$
+  where $A = \{ \store \in \mems(\{ \elab{\alpha}{1},\elab{\alpha}{2} \} ) \mid
+      \cod{\store(\elab{\mesg{w}}{1}) + \store(\elab{\mesg{w}}{2})} = v \}$.
+\end{definition}
 
@@
 \cite{evans2018pragmatic}. Our implementation libary is shown in
-Figure \ref{fig-gmw}, with type signatures for the library functions
-also shown. We present the And gate since it is
-an interesting component. The Figure includes the
-following top-level functions:
-\begin{itemize}
-\item \ttt{encodegmw}: This function encodes client 1's and client 2's
-  secret bits called $\ttt{s[1,"s1"]}$ and $\ttt{s[1,"s2"]}$ into two
-  distinct wire values (pairs of shares).
-\item \ttt{andgmw}: This function defines the gate $\ttt{"g"}$, the
-  identifier $\ttt{"g"}$ being used to distinguish randomness used
-  within.  In our version client 1 builds the output table (using
-  \ttt{andtablegmw})and transfers the correct output share to client 2
-  using 1-out-of-4 OT ($\ttt{OT4}$), definable using $\ttt{OT}$, as
-  per standard GMW protocol.
-\item \ttt{decode}: This function decodes and publishes a wire value
-  by $\ttt{xor}$ing the shares. Note that this requires both client 1
-  and 2 to publicize their shares.
-\end{itemize}
-\begin{example}
-  \label{example-gmw-andcircuit}
-The following program uses our GMW library to define
-a circuit with a single And gate and input secrets $\ttt{s1}$ and
+Figure \ref{fig-gmw}, and includes encoding functions, where
+input secrets are split into shares, $\eand$ and $\exor$ gate
+functions, and a decoding function. Note that $\exor$ requires
+no interaction between parties, while $\eand$ necessitates and
+1-of-4 oblivious transfer. The gate computation is
+done entirely in secrect, and the decoding function
+is where the declassification occurs-- both parties reveal
+their shares of the final gate output $\macgv{z}$.
+
+For example, the following program uses our GMW library to define
+a circuit with a single \eand gate and input secrets $\ttt{s1}$ and
 $\ttt{s2}$ from client's 1 and 2 respectively:
 \begin{verbatimtab}
-decode(andgmw("gate0",encodegmw("s1"),encodegmw("s2")))
+         let s1 = encodegmw("s1") in
+         let s2 = encodegmw("s2") in
+         decodegmw(andgmw("z",s1,s2))
 \end{verbatimtab}
-\end{example}
+By convention we will assume that all gates are assigned unique output
+identifiers $\ttt{"z"}$, and that all programs are in the form
+of a sequence of let-bindings followed by a call to $\decodegmw$
+wrapping a circuit.
 
-\begin{definition}
-  For all $\alpha \in \{ \mesg{w}, \flip{w} \}$ define:
-  $$\pmf(\macgv{\alpha} = v) \defeq \sum_{\sigma \in A} \pmf(\sigma')$$
-  and define:
-  $$\condd{\pmf}{X}{\macgv{\alpha} = v}(\sigma) \defeq  \sum_{\sigma' \in A} \condd{\pmf}{X}{\sigma'}(\sigma)$$
-  where $A = \{ \store \in \mems(\{ \elab{\alpha}{1},\elab{\alpha}{2} \} ) \mid
-      \cod{\store(\elab{\alpha}{1}) + \store(\elab{\alpha}{2})} = v \}$.
-\end{definition}
+\paragraph{Oblivious Transfer} \cnote{Here is where we
+  wave hands about OT.}
+
+\subsubsection{Correctness Proof with Verification Tactics}
+
+\begin{lemma}
+  Define:
+  $$
+  \begin{array}{c}
+    \prog_{i} \defeq \\
+    \eassign{\mesg{x}}{1}{\flip{x}}{1}; \eassign{\mesg{x}}{2}{\flip{x}}{2}; \\
+    \eassign{\mesg{y}}{1}{\flip{y}}{1}; \eassign{\mesg{y}}{2}{\flip{y}}{2} 
+  \end{array}
+  $$
+  Given:
+  $$
+  \begin{array}{c}
+  (\prog_i,\andgmw(z,\mesg{x},\mesg{y}) \redxs %\\
+  (\prog,\mesg{z})
+  \end{array}
+  $$
+  Then both of the following conditions hold, where $\iov(\prog) = M$:
+  $$
+  \condsep
+      {\progtt(\prog)}
+      {\{ \macgv{\mesg{x}},\macgv{\mesg{y}} \}}
+      {\{ \macgv{\mesg{z}} \}}
+      {M_{\{2\}} \cup \{ \elab{\mesg{z}}{1},\elab{\mesg{z}}{2} \}}
+  $$
+  $$
+  \conddetx{\progtt(\prog)}{\{ \macgv{\mesg{x}},\macgv{\mesg{y}} \}}{\{ \macgv{\mesg{z}} \}}    
+  $$
+\end{lemma}
 
@@
 \begin{lemma}
-  If $\condsep{\pmf}{X_1}{X_2}{X_3}$ and
-  $\condp{\pmf}{X_2}{\detx{X_4}}$ then $\condsep{\pmf}{X_1}{X_4}{X_3}$.
-\end{lemma}
+  Given:
+  \begin{mathpar}
+    \condsep{\pmf}{X_1}{X_2}{X_3}
 
-\begin{lemma}
-  Given:
-  $$
-  \begin{array}{c}
-  (\varnothing,\andgmw(z,\mesg{x},\mesg{y}) \redxs %\\
-  (\prog,\mesg{z})
-  \end{array}
-  $$
-  Then:
-  $$
-  \condsep
-      {\progtt(\prog)}
-      {\{ \macgv{\mesg{x}},\macgv{\mesg{y}} \}}
-      {\{ \macgv{\mesg{z}} \}}
-      {\{ \elab{\mesg{x}}{2},\elab{\mesg{y}}{2},\elab{\mesg{z}}{1},\elab{\mesg{z}}{2} \}}
-  $$
-  $$
-  \conddetx{\progtt(\prog)}{\{ \macgv{\mesg{x}},\macgv{\mesg{y}} \}}{\{ \macgv{\mesg{z}} \}}    
-  $$
+    \condp{\pmf}{X_1}{\detx{X_2}}
+
+    \condp{\pmf}{X_2}{\detx{X_4}}
+  \end{mathpar}
+    Then $\condsep{\pmf}{X_1}{X_4}{X_3}$.
 \end{lemma}

### 2024-05-12 ppdp24/gmw.tex
checkpoint

@@
-\begin{fpfig}[t]{2-Party GMW circuit library with And gate.}{fig-gmw}
+\begin{fpfig}[t]{2-Party GMW circuit library.}{fig-gmw}
 {\footnotesize
@@
 
-    andgmw(g, v1, v2) {
-      let r = r[g] in
-      let table = andtablegmw(v1,v2,r) in
-      m[g]@2 := OT4(v1,v2,table,2,1);
-      m[g]@1 := r@1;
-      m[g]
+    andgmw(z, x, y) {
+      let r = r[z] in
+      let table = andtablegmw(x,y,r) in
+      m[z]@2 := OT4(x,y,table,2,1);
+      m[z]@1 := r@1;
+      m[z]
     }
 
-    decodegmw(v) {
-      p["1"] := v@1; p["2"] := v@2;
-      out@1 := (p["1"] + p["2"])@1;
-      out@2 := (p["1"] + p["2"])@2
+    xorgmw(z, x, y) { m[z]@1 := (x+y)@1; m[z]@2 := (x+y)@2; m[z] }
+
+    decodegmw(z) {
+      p["1"] := z@1; p["2"] := z@2;
+      out@1 := (p["1"] xor p["2"])@1;
+      out@2 := (p["1"] xor p["2"])@2
     }

### 2024-05-13 ppdp24/examples.tex
checkpoint

@@
 
+\begin{definition}
+  We write $\condsep{\pmf}{X_1}{X_2}{X_3}$ iff for all
+  $\store' \in \mems(X_1)$ and $\store \in \mems(X_2 \cup X_3)$
+  and letting $X = X_1 \cup X_2 \cup X_3$ we have:
+  $$\condd{\pmf}{X}{\store'}(\store) =
+  \condd{\pmf}{X}{\store'}(\store_{X_2}) *
+  \condd{\pmf}{X}{\store'}(\store_{X_3})$$
+  We write $\conddetx{\pmf}{X_1}{X_2}$ iff for all
+  $\store' \in \mems(X_1)$ there exists 
+  $\store \in \mems(X_2)$ such that
+  $\condd{\pmf}{X_1 \cup X_2}{\store'}(\store) = 1$.
+\end{definition}
+
 \begin{lemma}
-  Define:
-  $$
-  \begin{array}{c}
-    \prog_{i} \defeq \\
-    \eassign{\mesg{x}}{1}{\flip{x}}{1}; \eassign{\mesg{x}}{2}{\flip{x}}{2}; \\
-    \eassign{\mesg{y}}{1}{\flip{y}}{1}; \eassign{\mesg{y}}{2}{\flip{y}}{2} 
-  \end{array}
-  $$
+  %Define:
+  %$$
+  %\begin{array}{c}
+  %  \prog_{i} \defeq \\
+  %  \eassign{\mesg{x}}{1}{\flip{x}}{1}; \eassign{\mesg{x}}{2}{\flip{x}}{2}; \\
+  %  \eassign{\mesg{y}}{1}{\flip{y}}{1}; \eassign{\mesg{y}}{2}{\flip{y}}{2} 
+  %\end{array}
+  %$$
   Given:
@@
   \begin{array}{c}
-  (\prog_i,\andgmw(z,\mesg{x},\mesg{y}) \redxs %\\
+  (\varnothing,\andgmw(z,\mesg{x},\mesg{y}) \redxs %\\
   (\prog,\mesg{z})
@@
   $$
-  Then both of the following conditions hold, where $\iov(\prog) = M$:
+  Then both of the following conditions hold for both $\cid \in \{ 1,2 \}$ where $\iov(\prog) = M$:
   $$
@@
       {\{ \macgv{\mesg{z}} \}}
-      {M_{\{2\}} \cup \{ \elab{\mesg{z}}{1},\elab{\mesg{z}}{2} \}}
+      {M_{\{\cid\}} \cup \{ \elab{\mesg{z}}{1},\elab{\mesg{z}}{2} \}}
   $$
@@
   $\condp{\pmf}{X_2}{\detx{X_3}}$ then $\condp{\pmf}{X_1}{\detx{X_3}}$.
+\end{lemma}
+
+\begin{lemma}
+  If $\condp{\pmf}{X_1}{\detx{X_2}}$ and
+  $\condsep{\pmf}{X_2}{X_3}{X_4}$ then $\condsep{\pmf}{X_1}{X_3}{X_4}$.
 \end{lemma}

### 2024-05-13 ppdp24/examples.tex
checkpoint

@@
-\section{Examples}
+\section{Examples and Proof Methods}
 \label{section-examples}
@@
 
-\begin{definition}
+\begin{definition}[Conditional Separation]
   We write $\condsep{\pmf}{X_1}{X_2}{X_3}$ iff for all
@@
   \condd{\pmf}{X}{\store'}(\store_{X_3})$$
+\end{definition}
+
+\begin{definition}[Conditional Determinism]  
   We write $\conddetx{\pmf}{X_1}{X_2}$ iff for all
@@
 \end{definition}
+
+\begin{lemma}
+  \label{lemma-gmwinvariant}
+  $$ (\varnothing,e) \redxs (\prog,\decodegmw(E[\mesg{z}])) $$
+  $$\iov(\prog) = S \cup M$$
+  $$\condsep{\progtt(\prog)}{S}{\{\macgv{\mesg{z}}\}}{(M_C \cup \{ \elab{\mesg{z}}{1}, \elab{\mesg{z}}{2} \})}$$
+  $$\conddetx{\progtt(\prog)}{S}{\{\macgv{\mesg{z}}\}}$$
+\end{lemma}
 
@@
 \begin{lemma}
-  Given:
-  \begin{mathpar}
-    \condsep{\pmf}{X_1}{X_2}{X_3}
-
-    \condp{\pmf}{X_1}{\detx{X_2}}
-
-    \condp{\pmf}{X_2}{\detx{X_4}}
-  \end{mathpar}
-    Then $\condsep{\pmf}{X_1}{X_4}{X_3}$.
+  Given $\condsep{\pmf}{X_1}{X_2}{X_3}$ and $\condp{\pmf}{X_1}{\detx{X_2}}$ and $\condp{\pmf}{X_2}{\detx{X_4}}$,
+%  \begin{mathpar}
+%    \condsep{\pmf}{X_1}{X_2}{X_3}
+%    
+%    \condp{\pmf}{X_1}{\detx{X_2}}
+%
+%    \condp{\pmf}{X_2}{\detx{X_4}}
+%  \end{mathpar}
+    then $\condsep{\pmf}{X_1}{X_4}{X_3}$.
 \end{lemma}
 
-\begin{lemma}
-  $$ (\varnothing,e) \redxs (\prog,\decodegmw(E[\mesg{z}])) $$
-  $$\iov(\prog) = S \cup M$$
-  $$\condsep{\progtt(\prog)}{S}{\{\macgv{\mesg{z}}\}}{(M_C \cup \{ \elab{\mesg{z}}{1}, \elab{\mesg{z}}{2} \})}$$
-  $$\conddetx{\progtt(\prog)}{S}{\{\macgv{\mesg{z}}\}}$$
-\end{lemma}
-
+\begin{proof}[Proof of Lemma \ref{lemma-gmwinvariant}]
 \begin{mathpar}
@@
 \end{mathpar}
+\end{proof}
 

### 2024-05-13 ppdp24/examples.tex
checkpoint

@@
 
-\begin{lemma}
+\begin{lemma}[GMW Invariant]
   \label{lemma-gmwinvariant}
+  Given:
   $$ (\varnothing,e) \redxs (\prog,\decodegmw(E[\mesg{z}])) $$
-  $$\iov(\prog) = S \cup M$$
-  $$\condsep{\progtt(\prog)}{S}{\{\macgv{\mesg{z}}\}}{(M_C \cup \{ \elab{\mesg{z}}{1}, \elab{\mesg{z}}{2} \})}$$
-  $$\conddetx{\progtt(\prog)}{S}{\{\macgv{\mesg{z}}\}}$$
+  Then both of the following conditions hold for all $H$ and $C$ where $\iov(\prog) = S \cup M$:
+  \begin{enumerate}
+    \item $\condsep{\progtt(\prog)}{S}{\{\macgv{\mesg{z}}\}}{(M_C \cup \{ \elab{\mesg{z}}{1}, \elab{\mesg{z}}{2} \})}$
+    \item $\conddetx{\progtt(\prog)}{S}{\{\macgv{\mesg{z}}\}}$
+  \end{enumerate}
 \end{lemma}
 
-\begin{lemma}
+\begin{lemma}[And Gate Tactic]
+  \label{lemma-gmwtactic}
   %Define:
@@
   Then both of the following conditions hold for both $\cid \in \{ 1,2 \}$ where $\iov(\prog) = M$:
-  $$
-  \condsep
+  \begin{enumerate}
+   \item $\condsep
       {\progtt(\prog)}
@@
       {\{ \macgv{\mesg{z}} \}}
-      {M_{\{\cid\}} \cup \{ \elab{\mesg{z}}{1},\elab{\mesg{z}}{2} \}}
-  $$
-  $$
-  \conddetx{\progtt(\prog)}{\{ \macgv{\mesg{x}},\macgv{\mesg{y}} \}}{\{ \macgv{\mesg{z}} \}}    
-  $$
+      {M_{\{\cid\}} \cup \{ \elab{\mesg{z}}{1},\elab{\mesg{z}}{2} \}}$
+    \item
+      $\conddetx{\progtt(\prog)}{\{ \macgv{\mesg{x}},\macgv{\mesg{y}} \}}{\{ \macgv{\mesg{z}} \}}$
+  \end{enumerate}
 \end{lemma}
 
-\begin{lemma}
-  If $\condp{\pmf}{X_1}{\detx{X_2}}$ and
-  $\condp{\pmf}{X_2}{\detx{X_3}}$ then $\condp{\pmf}{X_1}{\detx{X_3}}$.
-\end{lemma}
+\begin{proof}
+Verified automatically using techniques described in Section \ref{section-bruteforce}.  
+\end{proof}
 
 \begin{lemma}
-  If $\condp{\pmf}{X_1}{\detx{X_2}}$ and
-  $\condsep{\pmf}{X_2}{X_3}{X_4}$ then $\condsep{\pmf}{X_1}{X_3}{X_4}$.
-\end{lemma}
-
-\begin{lemma}
-  Given $\condsep{\pmf}{X_1}{X_2}{X_3}$ and $\condp{\pmf}{X_1}{\detx{X_2}}$ and $\condp{\pmf}{X_2}{\detx{X_4}}$,
-%  \begin{mathpar}
-%    \condsep{\pmf}{X_1}{X_2}{X_3}
-%    
-%    \condp{\pmf}{X_1}{\detx{X_2}}
-%
-%    \condp{\pmf}{X_2}{\detx{X_4}}
-%  \end{mathpar}
-    then $\condsep{\pmf}{X_1}{X_4}{X_3}$.
+  \label{lemma-conditioning}
+  Each of the following hold:
+  \begin{enumerate}
+    \item Given $\condp{\pmf}{X_1}{\detx{X_2}}$ and
+      $\condp{\pmf}{X_2}{\detx{X_3}}$, then $\condp{\pmf}{X_1}{\detx{X_3}}$.
+    \item Given $\condp{\pmf}{X_1}{\detx{X_2}}$ and
+      $\condsep{\pmf}{X_2}{X_3}{X_4}$, then $\condsep{\pmf}{X_1}{X_3}{X_4}$.
+    \item Given $\condsep{\pmf}{X_1}{X_2}{X_3}$ and $\condp{\pmf}{X_1}{\detx{X_2}}$
+      and $\condp{\pmf}{X_2}{\detx{X_4}}$, then $\condsep{\pmf}{X_1}{X_4}{X_3}$.
+  \end{enumerate}
 \end{lemma}

### 2024-05-13 ppdp24/examples.tex
checkpoint

@@
 
+Eliminating cheating by adding information theoretic secure MAC codes
+to shares is a fundamental approach realized by protocols such as BDOZ
+\cite{XXX} and SPDZ \cite{XXX}.  These protocols assume a
+pre-processing phase that distributes shares with MAC codes to
+clients.  This integrates well with pre-processed Beaver Triples to
+implement malicious secure, and relatively efficient, multiplication
+\cite{XXX}.  Here we show how to implement malicious secure two party
+multiplication in the BDOZ style via some extensions of $\minifed$.
+
+A field value $v$ is secret shared among 2 clientz in BDOZ as follows.
+Each client $\cid$ gets a pair of the form $(v_\cid,m_\cid)$, where
+$v_\cid$ is a secret share of $v$, i.e., $v = \fcod{v_1 \fplus v_2}$.
+Further, field values $m_\cid$ are MACs of $v_\cid$ that are authenticated
+by the other client's key. In particular, $m_\cid = k + (k_\Delta *
+v_\cid)$ where \emph{keys} $k$ and $k_\Delta$ are known only to $\cid'
+\ne \cid$ and $k_\Delta$. Further, there is a unique \emph{local key}
+$k$ for each shared value, while the \emph{global key} $k_\Delta$ is
+common to all MACs authenticated by $\cid'$. By extending certain
+field operations to these MACed values, a semi-homorphic encryption
+scheme is obtained that is adequate to compute Beaver Triples. For
+example, if local keys $k_1$ and $k_2$ authenticate $\macv_1$ and
+$\macv_2$, then $k_1 \fplus k_2$ authenticates $\macv_1 \macplus
+\macv_2$, where $\macplus$ is addition extended to MACed
+values. Semantics for summation of MACed shares, and multipication of
+shares and plain field values, are given in Figure \ref{fig-bdoz}.
+For more details the reader is referred to \cite{XXX}. We note
+that while we restrict values $v$, $m$, and $k$ to the same
+field in this presentation for simplicity, in general $m$ and
+$k$ can be in extensions of $\mathbb{Z}_p$.
+
+We define variable forms $\macx{\secret{w}}{\cid}$ and
+$\macx{\flip{w}}{\cid}$ to range over local MACed shares of a secret
+and random sample owned by client $\cid$, respectively. We let
+$\mack{x}{}$ denote the local key for authenticating shares of $x$.
+We also assume a global key $\flip{\ttt{delta}}$ on each client.
+Pre-processing is modeled by assuming that shares are provided on
+input random tapes (memories) with the required properties, as defined
+on the bottom of Figure \ref{fig-bdoz}.
+
+Similarly to field values, we can use MACed shares in
+uniform independent distributions as one-time-pads for
+hiding other values, using subtraction over MACed shares.
+So we introduce a function $\macotp$ to do this
+declaratively.
+$$
+\delta(\macotp,v_1,v_2) \defeq \fcod{v_1 \macminus v_2}
+$$
+
+Another key idea in BDOZ is \emph{secure opening}, where distributed
+shares of a value $v$ are communicated and combined to ``open'' the
+same $v$ on each client. This relies on authentication to detect
+cheating in the malicious setting. To implement secure opening we
+define a $\macauth$ function that authenticates shares, and a
+$\macopen$ function that extracts a field value from its shares.
+\begin{mathpar}
+\delta(\macauth, (v, m), k) \defeq
+     (v, m) \text{\ if\ } m = k \fplus (\flip{\ttt{delta}} \ftimes v)
+
+\delta(\macopen, (v_1, m_1), (v_2,m_2)) \defeq
+\fcod{v_1 \fplus v_2}
+\end{mathpar}
+This authentication eliminates cheating and guarantees protocol
+integrity. As we will show in Section \ref{section-types}, this
+can be reflected in a security type for $\macauth$.
+
+Putting these pieces together, we can implement 2-party multiplication with
+Beaver Triples as follows. Recall that Beaver Triples are values
+$(a,b,c)$ such that $c = \fcod{b \ftimes a}$- we implement them here
+as a secret $\secret{c}$ and samples $\flip{a}$ and $\flip{b}$ owned
+by an \emph{oracle} $\Oracle$ that we always assume is honest but is
+otherwise not involved in the protocol. We define pre-processing
+properties of Beaver Triples in Figure \ref{fig-bdoz}. Only one
+Beaver triple can be used per multiplication to ensure secure
+openings, but our scheme can be extended with more Beaver triples
+and reuse of them will be rule out by type linearity as we will
+discuss in Section \ref{section-types}.
+
+First, clients 1 and 2 compute (aka \emph{open}) $\secret{x} \fplus
+\flip{a}$ and $\secret{y} \fplus \flip{b}$ where $\secret{x}$ and
+$\secret{y}$ are 1 and 2's secrets respectively.  Then they both
+perform a secure opening of $\secret{x} \fplus \flip{a}$ and
+$\secret{y} \fplus \flip{b}$, by authenticating each others' shares,
+and record them as $\mesg{d}$ and $\mesg{e}$ respectively. Note
+that the secrets remain hidden since $\flip{a}$ and $\flip{b}$
+are random values.
+
 \begin{definition}

### 2024-05-13 ppdp24/examples.tex
checkpoint

--- 2024-05-13 ppdp24/examples.tex
@@
 
+$$
+\OT{\elab{\be_1}{\cid}}{\be_2}{\be_3}
+$$
+$$
+\begin{array}{l}
+\solve{\stores}{\OT{\elab{\be_1}{\cid_1}}{\be_2}{\be_3}}{\cid_2} = \\
+\qquad ((\solve{\stores}{\be_1}{\cid_1}) \cap 
+(\solve{\stores}{\be_2}{\cid_2})) \cup \\
+\qquad ((\stores - (\solve{\stores}{\be_1}{\cid_1})) \cap
+(\solve{\stores}{\be_3}{\cid_2})
+\end{array}
+$$
+
 \subsubsection{Correctness Proof with Verification Tactics}
 
+As discussed above and in related work \cite{XXX}, probabilistic
+separation conditional on certain variables-- e.g., secret inputs or
+public outputs-- is a key mechanism for reasoning about MPC protocol
+security. Following \cite{XXX}, we define a conditional separation
+relation $\condsep{\pmf}{X_1}{X_2}{X_3}$ to mean that
+under the condition of all value assignments for
+$X_2$ and $X_3$ are separable under $\pmf$ on the condition
+of any value assignment of $X_1$-- i.e., conditionally
+on any $\store \in \mems(X_1)$. 
 \begin{definition}[Conditional Separation]
@@
 \end{definition}
-
+Another key concept needed especially for reasoning about
+circuits is conditional determinism. For example, if
+$\macgv{z}$ is an output of an internal gate, it will
+definitely be computed using random variables, however,
+it \emph{should} be determistic under any set of input
+secrets $S$, since we assume that $\idealf$ is
+deterministic.
 \begin{definition}[Conditional Determinism]  
@@
 
+Given these definitions, we can formulate an invariant
+for circuit computation with respect to internal gates
+as follows. It says that the output of any gate is
+deterministic wrt inputs $S$, and conditionally
+on $S$ the output $\macgv{\mesg{z}}$ remains
+separable from corrupt views and both shares of
+$\macgv{\mesg{z}}$. This last nuance is critical,
+since those shares will in fact be revealed if
+$\macgv{\mesg{z}}$ is decoded as the circuit output. 
 \begin{lemma}[GMW Invariant]
@@
 \end{lemma}
-
+To prove this, we can formulate and automatically prove gate-level
+versions of the invariant. This serves as a proof tactic
+that simplifies the proof of the GMW invariant.
 \begin{lemma}[And Gate Tactic]
@@
 \end{lemma}
-
 \begin{proof}

### 2024-05-13 ppdp24/examples.tex
checkpoint

@@
   \begin{enumerate}
+    \item $\conddetx{\progtt(\prog)}{S}{\{\macgv{\mesg{z}}\}}$
     \item $\condsep{\progtt(\prog)}{S}{\{\macgv{\mesg{z}}\}}{(M_C \cup \{ \elab{\mesg{z}}{1}, \elab{\mesg{z}}{2} \})}$
-    \item $\conddetx{\progtt(\prog)}{S}{\{\macgv{\mesg{z}}\}}$
   \end{enumerate}
@@
 versions of the invariant. This serves as a proof tactic
-that simplifies the proof of the GMW invariant.
+that simplifies the proof of the GMW invariant. It establishes
+that the $\eand$ gate output is deterministic conditional on
+the inputs, and is separable from the output shares and
+either parties' received messages. 
 \begin{lemma}[And Gate Tactic]
@@
   \begin{enumerate}
+  \item
+    $\conddetx{\progtt(\prog)}{\{ \macgv{\mesg{x}},\macgv{\mesg{y}} \}}{\{ \macgv{\mesg{z}} \}}$
    \item $\condsep
@@
       {M_{\{\cid\}} \cup \{ \elab{\mesg{z}}{1},\elab{\mesg{z}}{2} \}}$
-    \item
-      $\conddetx{\progtt(\prog)}{\{ \macgv{\mesg{x}},\macgv{\mesg{y}} \}}{\{ \macgv{\mesg{z}} \}}$
   \end{enumerate}
@@
 
+To properly integrate the local reasoning of Lemma \ref{lemma-gmwtactic} with
+the global reasoning of Lemma \ref{lemma-gmwinvariant}, we can demonstrate
+the following.
 \begin{lemma}
@@
 
+Then we can put the pieces together to prove the invariant, using automated tactics
+for gate-level reasoning.  
 \begin{proof}[Proof of Lemma \ref{lemma-gmwinvariant}]
+  By induction on the length of $(\varnothing,e) \redxs (\prog,\decodegmw(E[\mesg{z}]))$.
+  Encoding establishes the invarian in the basis. The most interesting inductive
+  case is the $\andgmw$ gate. 
+  \paragraph{Case $\andgmw$.} In this case we have:
+  $$
+  \begin{array}{c}
+  (\varnothing,e) \redxs (\prog',\decodegmw(E[\andgmw(z,\mesg{x},\mesg{y})])) \redxs \\
+    (\prog,\decodegmw(E[\mesg{z}]))
+  \end{array}
+  $$
+  The induction hypothesis, together with the assumed uniquenenss of $z$, gives:
   \begin{mathpar}
-  \condsep{\progtt(\prog_1)}{S^1}{\{ \macgv{\mesg{x}} \}}{(M^1_C \cup \{ \elab{\mesg{x}}{1}, \elab{\mesg{x}}{2} \})}
+  \condsep{\progtt(\prog)}{S^1}{\{ \macgv{\mesg{x}} \}}{(M^1_C \cup \{ \elab{\mesg{x}}{1}, \elab{\mesg{x}}{2} \})}
 
-  \condsep{\progtt(\prog_2)}{S^2}{\{ \macgv{\mesg{y}} \}}{(M^2_C \cup \{ \elab{\mesg{y}}{1}, \elab{\mesg{y}}{2} \})}
+  \condsep{\progtt(\prog)}{S^2}{\{ \macgv{\mesg{y}} \}}{(M^2_C \cup \{ \elab{\mesg{y}}{1}, \elab{\mesg{y}}{2} \})}
 
-  \condsep{\progtt(\prog_3)}{S^1 \cup S^2}{\{ \macgv{\mesg{z}} \}}{(M^1_C \cup M^2_C \cup \{ \elab{\mesg{x}}{2}, \elab{\mesg{y}}{2}, \elab{\mesg{z}}{1}, \elab{\mesg{z}}{2} \})}
+  \conddetx{\progtt(\prog)}{S^1}{\{ \macgv{\mesg{x}} \}}
+
+  \conddetx{\progtt(\prog)}{S^1}{\{ \macgv{\mesg{x}} \}}    
   \end{mathpar}
+  This, together with Lemma \ref{lemma-gmwtactic} (1) and Lemma \ref{lemma-conditioning} (1)
+  give:
+  $$
+  \conddetx{\progtt(\prog)}{S^1 \cup S^2}{\{ \macgv{\mesg{z}} \}}
+  $$
+  and Lemma  \ref{lemma-gmwtactic} (2) and Lemma \ref{lemma-conditioning} (2-3) gives:
+  $$
+  \begin{array}{c}
+    \progtt(\prog)|S^1 \cup S^2) \vdash \\
+    {\{ \macgv{\mesg{z}} \}}{(M^1_C \cup M^2_C \cup \{ \elab{\mesg{x}}{2}, \elab{\mesg{y}}{2}, \elab{\mesg{z}}{1}, \elab{\mesg{z}}{2} \})}
+  \end{array}
+  $$
 \end{proof}

### 2024-05-13 ppdp24/examples.tex
checkpoint

--- 2024-05-13 ppdp24/examples.tex
@@
   \begin{enumerate}
+    \item $\conddetx{\progtt(\prog)}{S}{\{\macgv{\mesg{z}}\}}$
     \item $\condsep{\progtt(\prog)}{S}{\{\macgv{\mesg{z}}\}}{(M_C \cup \{ \elab{\mesg{z}}{1}, \elab{\mesg{z}}{2} \})}$
-    \item $\conddetx{\progtt(\prog)}{S}{\{\macgv{\mesg{z}}\}}$
   \end{enumerate}
 \end{lemma}
-
+To prove this, we can formulate and automatically prove gate-level
+versions of the invariant. This serves as a proof tactic
+that simplifies the proof of the GMW invariant. It establishes
+that the $\eand$ gate output is deterministic conditional on
+the inputs, and is separable from the output shares and
+either parties' received messages. 
 \begin{lemma}[And Gate Tactic]
@@
       {M_{\{\cid\}} \cup \{ \elab{\mesg{z}}{1},\elab{\mesg{z}}{2} \}}$
-    \item
-      $\conddetx{\progtt(\prog)}{\{ \macgv{\mesg{x}},\macgv{\mesg{y}} \}}{\{ \macgv{\mesg{z}} \}}$
   \end{enumerate}
 \end{lemma}
-
 \begin{proof}

### 2024-05-13 ppdp24/examples.tex
checkpoint

@@
 
-\paragraph{Oblivious Transfer} \cnote{Here is where we
-  wave hands about OT.}
-
+\paragraph{Oblivious Transfer} A passive secure OT protocol
+based on previous work \cite{XXX} can be defined in $\metaprot$,
+however this protocol assumes some shared randomness. Alternatively,
+a simple passive secure OT can be defined with the addition of
+public key cryptography as a primitive. But given the diversity
+of approaches to OT, we instead assume that OT is abstract with
+respect to its implementation, where calls to OT in $\mathbb{F}_2$
+are of the following form-- given a \emph{choice bit}
+$\be_1$ provided by a receiver $\cid$, the sender
+sends either $\be_2$ or $\be_3$.
 $$
@@
 $$
+Critically, the sender learns nothing about $\be_1$ and the
+receiver learns nothing about the unselected value, so we interpret
+these calls in our implementation in $\mathbb{F}_2$ as follows.
 $$
@@
 
-Eliminating cheating by adding information theoretic secure MAC codes
-to shares is a fundamental approach realized by protocols such as BDOZ
-\cite{XXX} and SPDZ \cite{XXX}.  These protocols assume a
-pre-processing phase that distributes shares with MAC codes to
-clients.  This integrates well with pre-processed Beaver Triples to
-implement malicious secure, and relatively efficient, multiplication
-\cite{XXX}.  Here we show how to implement malicious secure two party
-multiplication in the BDOZ style via some extensions of $\minifed$.
+In a malicious setting, detecting cheating by adding
+information-theoretic secure MAC codes to shares is a fundamental approach
+realized by protocols such as BDOZ \cite{XXX} and SPDZ \cite{XXX}.
+These protocols assume a pre-processing phase that distributes shares
+with MAC codes to clients.  This integrates well with pre-processed
+Beaver Triples to implement malicious secure, and relatively
+efficient, multiplication \cite{XXX}.  Here we show how to implement
+malicious secure two party multiplication in the BDOZ style in
+$\metaprot$.
 

### 2024-05-13 ppdp24/examples.tex
checkpoint

@@
   $$\condd{\pmf}{X}{\macgv{\mesg{w}} = v}(\sigma) \defeq  \sum_{\sigma' \in A} \condd{\pmf}{X}{\sigma'}(\sigma)$$
-  where $A = \{ \store \in \mems(\{ \elab{\alpha}{1},\elab{\alpha}{2} \} ) \mid
-      \cod{\store(\elab{\mesg{w}}{1}) + \store(\elab{\mesg{w}}{2})} = v \}$.
+  where $A$ is:
+  $$\{ \store \in \mems(\{ \elab{\mesg{w}}{1},\elab{\mesg{w}}{2} \} ) \mid
+      \fcod{\store(\elab{\mesg{w}}{1}) + \store(\elab{\mesg{w}}{2})} = v \}$$
 \end{definition}
@@
 
-A field value $v$ is secret shared among 2 clientz in BDOZ as follows.
+A field value $v$ is secret shared among 2 client in BDOZ in
+the same manner as in GMW.
 Each client $\cid$ gets a pair of the form $(v_\cid,m_\cid)$, where

### 2024-05-13 ppdp24/model.tex
checkpoint

@@
   $$
-  \runs(\prog) \defeq \{ \store \mid \preproc(\store) \wedge
+  \begin{array}{c}
+    \runs(\prog) \defeq \\
+    \{ \store \mid \exists \store_1 \in \mems(R) . 
+    \exists \store_2 . \preproc(\store_2) \wedge
     %(\dom(\store) = \iov(\prog) \cup R) \wedge
-  (\store_{S \cup R},\prog) \redxs (\store,\varnothing) \}
+    (\store_1 \cap \store_2,\prog) \redxs (\store,\varnothing) \}
+  \end{array}
   $$
@@
   $$
+  By default, $\preproc(\store) \iff \dom(\store) = S$, i.e.,
+  the initial memory contains all input secrets in a uniform
+  marginal distribution.
+
   In some cases, we will also be concerned with the (joint)
@@
   $$
-    \aruns(\prog) \defeq 
-    \{ \store \mid \preproc(\store)  \wedge ((\store_{S\cup R},\prog) \aredx^* (\store,\varnothing)\}% \vee (\store_{S \cup R},\prog) \aredx^* (\store,\bot)) \}
+  \begin{array}{c}
+    \aruns(\prog) \defeq \\
+    \{ \store \mid \exists \store_1 \in \mems(R) . 
+    \exists \store_2 . \preproc(\store_2) \wedge
+    %(\dom(\store) = \iov(\prog) \cup R) \wedge
+    (\store_1 \cap \store_2,\prog) \aredxs (\store,\varnothing) \}
+  \end{array}
   $$

### 2024-05-14 ppdp24/examples.tex
checkpoint

@@
 
+\begin{definition}[BDOZ preprocessing]
+  Define:
+  \begin{mathpar}
+    \mathit{shares} \defeq
+    \{ \elab{\mesg{w}}{\cid} | \cid \in \{ 1, 2 \} \wedge w \in \{ a,b,c,x,y \}  \}
+
+    \mathit{macs} \defeq  \{ \elab{\mesg{w\ttt{mac}}}{\cid} | \cid \in \{ 1, 2 \} \wedge w \in \{ a,b,c,x,y \}  \}
+
+    \mathit{keys} \defeq  \begin{array}{l}\{ \elab{\mesg{w\ttt{k}}}{\cid} | \cid \in \{ 1, 2 \} \wedge w \in
+    \{ a,b,c,x,y \}  \} \cup \\ \{ \elab{\mesg{\ttt{delta}}}{\cid} | \cid \in \{ 1, 2 \} \} \end{array}
+  \end{mathpar}
+  Then a memory $\store$ satisfies BDOZ preprocessing iff:
+  $$\dom(\store) = \mathit{shares}  \cup \mathit{macs} \cup \mathit{keys}$$
+  and, writing $\store(\macgv{\mesg{w}})$ to denote
+  $\fcod{\store(\elab{\mesg{w}}{1}) + \store(\elab{\mesg{w}}{2})}$,
+  the following conditions hold for all $\cid \in \{1,2\}$ and $w \in \{ a,b,c,x,y\}$:
+\begin{mathpar}
+  \store(\macgv{\mesg{w@\cid}}) = \store(\elab{\secret{w}}{\cid})
+
+  \fcod{\store(\macgv{\mesg{a}}) * \store(\macgv{\mesg{b}})} = \store(\macgv{\mesg{c}})
+
+  \lcod{\store, \mesg{w\ttt{mac}}}{\cid} =
+  \lcod{\store, \mesg{wk} + \mesg{\ttt{delta}} * \mesg{w}}{\cid'}  \qquad \cid \ne \cid'  
+\end{mathpar}
+\end{definition}
+
 \begin{definition}

### 2024-05-14 ppdp24/example-bdoz.tex
checkpoint

\subsection{2-Party BDOZ and Integrity Enforcement}
\label{section-example-bdoz}

\input{bdoz}

In a malicious setting, ``detecting cheating'' by adding
information-theoretic secure MAC codes to shares is a fundamental
approach realized by protocols such as BDOZ \cite{XXX} and SPDZ
\cite{XXX}.  These protocols assume a pre-processing phase that
distributes shares with MAC codes to clients.  This integrates well
with pre-processed Beaver Triples to implement malicious secure, and
relatively efficient, multiplication \cite{XXX}. Recall
that Beaver triples are values $a,b,c$ with $\fcod{a * b} = c$,
unique per multiplication gate, that are secret shared with clients
during pre-processing. Here we consider the 2-party version.

A field value $v$ is secret shared among 2 clients in BDOZ in the same
manner as in GMW.  Each client $\cid$ gets a pair of the form
$(v_\cid,m_\cid)$, where $v_\cid$ is a share of $v$ reconstructed by
addition, i.e., $v = \fcod{v_1 \fplus v_2}$, and $m_\cid$ is a MAC of
$v_\cid$.  More precisely, $m_\cid = k + (k_\Delta * v_\cid)$ where
\emph{keys} $k$ and $k_\Delta$ are known only to $\cid' \ne \cid$ and
$k_\Delta$. The \emph{local key} $k$ is unique per MAC, while the
\emph{global key} $k_\Delta$ is common to all MACs authenticated by
$\cid'$. This is a semi-homomorphic encryption scheme that supports
addition of shares and multiplication of shares by a constant
\cite{XXX}. For more details the reader is referred to \cite{XXX}. We
note that while we restrict values $v$, $m$, and $k$ to the same field
in this presentation for simplicity, in general $m$ and $k$ can be in
extensions of $\mathbb{Z}_p$.

We can capture both the preliminary distribution of Beaver triples and BDOZ shares
as a pre-processing predicate that establishes conditions for initial
memories (see Definition \ref{def-progtt}).  Here we assume two input
secrets $\elab{\secret{x}}{1}$ and $\elab{\secret{y}}{2}$ and a single
Beaver Triple to compute $\elab{\secret{x}}{1} \ftimes
\elab{\secret{y}}{2}$, but we can extend this for additional gates.
As for GMW, we use $\macgv{\mesg{w}}$ to refer to secret-shared values
reconstructed with addition, where by convention shares are message
values $\elab{\mesg{w}}{\cid}$ for all $\cid$.
\begin{definition}[BDOZ preprocessing]
  Define:
  \begin{mathpar}
    \mathit{shares} \defeq
    \{ \elab{\mesg{w}}{\cid}\ |\ \cid \in \{ 1, 2 \} \wedge w \in \{ a,b,c,x,y \}  \}

    \mathit{macs} \defeq  \{ \elab{\mesg{w\ttt{mac}}}{\cid}\ |\ \cid \in \{ 1, 2 \} \wedge w \in \{ a,b,c,x,y \}  \}

    \mathit{keys} \defeq  \begin{array}{l}\{ \elab{\mesg{w\ttt{k}}}{\cid}\ |\ \cid \in \{ 1, 2 \} \wedge w \in
    \{ a,b,c,x,y \}  \} \cup \\ \{ \elab{\mesg{\ttt{delta}}}{\cid}\ |\ \cid \in \{ 1, 2 \} \} \end{array}
  \end{mathpar}
  Then a memory $\store$ satisfies BDOZ preprocessing iff:
  $$\dom(\store) = \{ \elab{\secret{x}}{1}, \elab{\secret{y}}{2} \} \cup \mathit{shares}
  \cup \mathit{macs} \cup \mathit{keys}$$
  and, writing $\store(\macgv{\mesg{w}})$ to denote
  $\fcod{\store(\elab{\mesg{w}}{1}) + \store(\elab{\mesg{w}}{2})}$,
  the following conditions hold:
  \begin{mathpar}
    \store(\macgv{\mesg{x}}) = \store(\elab{\secret{x}}{1})

    \store(\macgv{\mesg{y}}) = \store(\elab{\secret{y}}{2})

    \fcod{\store(\macgv{\mesg{a}}) * \store(\macgv{\mesg{b}})} = \store(\macgv{\mesg{c}})
  \end{mathpar}
  and for all $\cid,\cid' \in \{1,2\}$ with $\cid \ne \cid'$ and $w \in \{ a,b,c,x,y\}$:
  $$\lcod{\store, \mesg{w\ttt{mac}}}{\cid} =
  \lcod{\store, \mesg{wk} + \mesg{\ttt{delta}} * \mesg{w}}{\cid'}$$
\end{definition}

With these conventions, our BDOZ library is defined in Figure \ref{fig-bdoz}.

\input{beaver}

\begin{definition}[BDOZ preprocessing]
  Define:
  \begin{mathpar}
    \mathit{shares} \defeq
    \{ \elab{\mesg{w}}{\cid}\ |\ \cid \in \{ 1, 2 \} \wedge w \in \{ a,b,c,x,y \}  \}

    \mathit{macs} \defeq  \{ \elab{\mesg{w\ttt{mac}}}{\cid}\ |\ \cid \in \{ 1, 2 \} \wedge w \in \{ a,b,c,x,y \}  \}

    \mathit{keys} \defeq  \begin{array}{l}\{ \elab{\mesg{w\ttt{k}}}{\cid}\ |\ \cid \in \{ 1, 2 \} \wedge w \in
    \{ a,b,c,x,y \}  \} \cup \\ \{ \elab{\mesg{\ttt{delta}}}{\cid}\ |\ \cid \in \{ 1, 2 \} \} \end{array}
  \end{mathpar}
  Then a memory $\store$ satisfies BDOZ preprocessing iff:
  $$\dom(\store) = \{ \elab{\secret{x}}{1}, \elab{\secret{y}}{2} \} \cup \mathit{shares}
  \cup \mathit{macs} \cup \mathit{keys}$$
  and, writing $\store(\macgv{\mesg{w}})$ to denote
  $\fcod{\store(\elab{\mesg{w}}{1}) + \store(\elab{\mesg{w}}{2})}$,
  the following conditions hold:
  \begin{mathpar}
    \store(\macgv{\mesg{x}}) = \store(\elab{\secret{x}}{1})

    \store(\macgv{\mesg{y}}) = \store(\elab{\secret{y}}{2})

    \fcod{\store(\macgv{\mesg{a}}) * \store(\macgv{\mesg{b}})} = \store(\macgv{\mesg{c}})
  \end{mathpar}
  and for all $\cid,\cid' \in \{1,2\}$ with $\cid \ne \cid'$ and $w \in \{ a,b,c,x,y\}$:
  $$\lcod{\store, \mesg{w\ttt{mac}}}{\cid} =
  \lcod{\store, \mesg{wk} + \mesg{\ttt{delta}} * \mesg{w}}{\cid'}$$
\end{definition}

\begin{definition}
  Given $\prog$ with $\iov(\prog) = S \cup V \cup O$,
  let $X_H \subseteq \{ x \mid x \in (\houtputs \cup O_H) \wedge x \in \dom(\store) \}$.
  Then the \emph{adversarial inputs to $X_H$} is the least set
  $X_C \subseteq \cinputs$ such that $\progtt(\prog) \not\vdash X_C * X_H$.
\end{definition}
\begin{definition}[Cheating Detection]
  \emph{Cheating is detected} in $\prog$ with $\iov(\prog) = S \cup V \cup O$ iff
  for all  $\store \in \aruns(\prog)$,
  letting $X_H = \{ x \mid x \in (\houtputs \cup O_H) \wedge x \in \dom(\store) \}$,
  and letting $X_C$ be the adversarial inputs to $X_H$,
  there exists $\sigma'\in \runs(\prog)$
  with $\store_{X_C} = \store'_{X_C}$.  
\end{definition}

\begin{lemma}
  If cheating is detected in $\prog$, then $\prog$ has integrity.
\end{lemma}

### 2024-05-14 ppdp24/example-gmw.tex
checkpoint

--- 2024-05-13 ppdp24/examples.tex
@@
-\section{Examples and Proof Methods}
-\label{section-examples}
-
-\subsection{2-Party GMW}
+\section{2-Party GMW and Passive Security Proof Tactics}
 \label{section-metalang-gmw}
+\label{section-example-gmw}
 
@@
 
-The GMW protocol is a garbled binary circuit protocol. In this section
-we will assume the 2-party version, though it generalizes to $n$
+As an extended example of our language and security model, and how the
+automated techniques in Section \ref{section-bruteforce} can serve
+as tactics integrated with PSL/Lilac-style proofs, we consider GMW
+circuits.  The GMW protocol is a garbled binary circuit protocol.  We
+will assume the 2-party version, though it generalizes to $n$
 parties\cite{XXX}. GMW uses a common technique in MPC, which is to
-represent values $v$ as distributed shares $v_1$ and $v_2$ with
-$v = v_1 \fplus v_2$. This trick maintains secrecy of $v$
-from both parties, and in GMW it is used to maintain the intermediate
-values of internal gate outputs in circuits. In related literature
-the notation $\macgv{x}$ is used to represent the ``true''
-value of $x$ and $[x]$ is often used to represent the share of
-given party.
+represent values $v$ as distributed shares $v_1$ and $v_2$ with $v =
+v_1 \fplus v_2$. This trick maintains secrecy of $v$ from both
+parties, and in GMW it is used to maintain the intermediate values of
+internal gate outputs in circuits. In related literature the notation
+$\macgv{x}$ is used to represent the ``true'' value of $x$ and $[x]$
+is often used to represent the share of given party.
 
@@
 
-\subsubsection{Correctness Proof with Verification Tactics}
+\subsection{Correctness Proof with Verification Tactics}
 
@@
 \end{proof}
-
-\subsection{2-Party BDOZ}
-
-\input{bdoz}
-
-In a malicious setting, detecting cheating by adding
-information-theoretic secure MAC codes to shares is a fundamental approach
-realized by protocols such as BDOZ \cite{XXX} and SPDZ \cite{XXX}.
-These protocols assume a pre-processing phase that distributes shares
-with MAC codes to clients.  This integrates well with pre-processed
-Beaver Triples to implement malicious secure, and relatively
-efficient, multiplication \cite{XXX}.  Here we show how to implement
-malicious secure two party multiplication in the BDOZ style in
-$\metaprot$.
-
-A field value $v$ is secret shared among 2 client in BDOZ in
-the same manner as in GMW.
-Each client $\cid$ gets a pair of the form $(v_\cid,m_\cid)$, where
-$v_\cid$ is a secret share of $v$, i.e., $v = \fcod{v_1 \fplus v_2}$.
-Further, field values $m_\cid$ are MACs of $v_\cid$ that are authenticated
-by the other client's key. In particular, $m_\cid = k + (k_\Delta *
-v_\cid)$ where \emph{keys} $k$ and $k_\Delta$ are known only to $\cid'
-\ne \cid$ and $k_\Delta$. Further, there is a unique \emph{local key}
-$k$ for each shared value, while the \emph{global key} $k_\Delta$ is
-common to all MACs authenticated by $\cid'$. By extending certain
-field operations to these MACed values, a semi-homorphic encryption
-scheme is obtained that is adequate to compute Beaver Triples. For
-example, if local keys $k_1$ and $k_2$ authenticate $\macv_1$ and
-$\macv_2$, then $k_1 \fplus k_2$ authenticates $\macv_1 \macplus
-\macv_2$, where $\macplus$ is addition extended to MACed
-values. Semantics for summation of MACed shares, and multipication of
-shares and plain field values, are given in Figure \ref{fig-bdoz}.
-For more details the reader is referred to \cite{XXX}. We note
-that while we restrict values $v$, $m$, and $k$ to the same
-field in this presentation for simplicity, in general $m$ and
-$k$ can be in extensions of $\mathbb{Z}_p$.
-
-We define variable forms $\macx{\secret{w}}{\cid}$ and
-$\macx{\flip{w}}{\cid}$ to range over local MACed shares of a secret
-and random sample owned by client $\cid$, respectively. We let
-$\mack{x}{}$ denote the local key for authenticating shares of $x$.
-We also assume a global key $\flip{\ttt{delta}}$ on each client.
-Pre-processing is modeled by assuming that shares are provided on
-input random tapes (memories) with the required properties, as defined
-on the bottom of Figure \ref{fig-bdoz}.
-
-Similarly to field values, we can use MACed shares in
-uniform independent distributions as one-time-pads for
-hiding other values, using subtraction over MACed shares.
-So we introduce a function $\macotp$ to do this
-declaratively.
-$$
-\delta(\macotp,v_1,v_2) \defeq \fcod{v_1 \macminus v_2}
-$$
-
-Another key idea in BDOZ is \emph{secure opening}, where distributed
-shares of a value $v$ are communicated and combined to ``open'' the
-same $v$ on each client. This relies on authentication to detect
-cheating in the malicious setting. To implement secure opening we
-define a $\macauth$ function that authenticates shares, and a
-$\macopen$ function that extracts a field value from its shares.
-\begin{mathpar}
-\delta(\macauth, (v, m), k) \defeq
-     (v, m) \text{\ if\ } m = k \fplus (\flip{\ttt{delta}} \ftimes v)
-
-\delta(\macopen, (v_1, m_1), (v_2,m_2)) \defeq
-\fcod{v_1 \fplus v_2}
-\end{mathpar}
-This authentication eliminates cheating and guarantees protocol
-integrity. As we will show in Section \ref{section-types}, this
-can be reflected in a security type for $\macauth$.
-
-Putting these pieces together, we can implement 2-party multiplication with
-Beaver Triples as follows. Recall that Beaver Triples are values
-$(a,b,c)$ such that $c = \fcod{b \ftimes a}$- we implement them here
-as a secret $\secret{c}$ and samples $\flip{a}$ and $\flip{b}$ owned
-by an \emph{oracle} $\Oracle$ that we always assume is honest but is
-otherwise not involved in the protocol. We define pre-processing
-properties of Beaver Triples in Figure \ref{fig-bdoz}. Only one
-Beaver triple can be used per multiplication to ensure secure
-openings, but our scheme can be extended with more Beaver triples
-and reuse of them will be rule out by type linearity as we will
-discuss in Section \ref{section-types}.
-
-First, clients 1 and 2 compute (aka \emph{open}) $\secret{x} \fplus
-\flip{a}$ and $\secret{y} \fplus \flip{b}$ where $\secret{x}$ and
-$\secret{y}$ are 1 and 2's secrets respectively.  Then they both
-perform a secure opening of $\secret{x} \fplus \flip{a}$ and
-$\secret{y} \fplus \flip{b}$, by authenticating each others' shares,
-and record them as $\mesg{d}$ and $\mesg{e}$ respectively. Note
-that the secrets remain hidden since $\flip{a}$ and $\flip{b}$
-are random values.
-
-\begin{definition}
-  Given $\prog$ with $\iov(\prog) = S \cup V \cup O$,
-  let $X_H \subseteq \{ x \mid x \in (\houtputs \cup O_H) \wedge x \in \dom(\store) \}$.
-  Then the \emph{adversarial inputs to $X_H$} is the least set
-  $X_C \subseteq \cinputs$ such that $\progtt(\prog) \not\vdash X_C * X_H$.
-\end{definition}
-\begin{definition}[Cheating Detection]
-  \emph{Cheating is detected} in $\prog$ with $\iov(\prog) = S \cup V \cup O$ iff
-  for all  $\store \in \aruns(\prog)$,
-  letting $X_H = \{ x \mid x \in (\houtputs \cup O_H) \wedge x \in \dom(\store) \}$,
-  and letting $X_C$ be the adversarial inputs to $X_H$,
-  there exists $\sigma'\in \runs(\prog)$
-  with $\store_{X_C} = \store'_{X_C}$.  
-\end{definition}
-
-\begin{lemma}
-  If cheating is detected in $\prog$, then $\prog$ has integrity.
-\end{lemma}

### 2024-05-14 ppdp24/example-bdoz.tex
checkpoint

@@
 With these conventions, our BDOZ library is defined in Figure \ref{fig-bdoz}.
+In $\metaprot$ we represent BDOZ share, MAC pairs as records:
+$$
+\ttt{\{ share = }v\ttt{;  mac =}\ m \ttt{\}}
+$$
+and we define $\ttt{macsum}$ for addition of shares,
+$\ttt{maccsum}$ for addition of a share and a constant, and
+$\ttt{macctimes}$ for multiplication of a share and a constant
+in the BDOZ encryption scheme \cite{XXX}. The $\ttt{auth}$
+function implements the MAC check as an $\assert$.
+
+We also define a function $\ttt{secopen}$ to implement ``secure
+opening''.  In this standard subprotocol, the value
+$\macgv{\secret{w_1}} \fplus \macgv{\mesg{w_2}}$ is reconstructed as
+$\mesg{w_3}$, by each client $\cid_2$ communicating
+$\lcod{\mesg{w_1} + \mesg{w_2}}{\cid_2}$ to $\cid_1$.  Assuming that
+$\macgv{\mesg{w_2}}$ is in an independent uniform distribution,
+this perfectly hides $\secret{w_1}$. In a mutiplication gate
+either $a$ or $b$ of a Beaver triple are used in secure opening,
+so, e.g., given $
+(\varnothing,\ttt{secopen}(a, x, d, 2, 1) \redxs (\prog,())$
+we have both:
+\begin{mathpar}
+  \conddetx{\progtt(\prog)}{\{\macgv{\mesg{a}},\elab{\secret{x}}{1} \}}{\{ \elab{\mesg{d}}{2} \}}
+
+  \sep{\progtt(\prog)}{\{ \elab{\mesg{d}}{2} \}}{\{ \elab{\secret{x}}{1}  \}}
+\end{mathpar}
+Furthermore, client 2's authentication of the sum of shares with the
+sum of their keys detects any attempted cheating by $1$.
+Finally, the $\ttt{secreveal}$ function
+is very similar to $\decodegmw$, except with the addition
+of authentication of revealed shares to ensure malicious security.
 
@@
 
-\begin{definition}[BDOZ preprocessing]
-  Define:
-  \begin{mathpar}
-    \mathit{shares} \defeq
-    \{ \elab{\mesg{w}}{\cid}\ |\ \cid \in \{ 1, 2 \} \wedge w \in \{ a,b,c,x,y \}  \}
+The full protocol for malicious secure product of secrets $x$ (that
+is, $\elab{\secret{x}}{1}$) and $y$ (that is, $\elab{\secret{y}}{2}$)
+using Beaver triple $a,b,c$ is defined in Figure \ref{fig-bdoz}. Both
+parties interact in secure opening of $x \fplus a$ and $y + b$,
+followed by local computation to compute their share of the $x * y$
+using only local addition and product as per standard protocol
+\cite{XXX}. The non-interactive reconstruction of the local authentication
+key for both the secure openings and the final reveal is enabled by the
+semi-homorphic properties of the BDOZ scheme.
 
-    \mathit{macs} \defeq  \{ \elab{\mesg{w\ttt{mac}}}{\cid}\ |\ \cid \in \{ 1, 2 \} \wedge w \in \{ a,b,c,x,y \}  \}
-
-    \mathit{keys} \defeq  \begin{array}{l}\{ \elab{\mesg{w\ttt{k}}}{\cid}\ |\ \cid \in \{ 1, 2 \} \wedge w \in
-    \{ a,b,c,x,y \}  \} \cup \\ \{ \elab{\mesg{\ttt{delta}}}{\cid}\ |\ \cid \in \{ 1, 2 \} \} \end{array}
-  \end{mathpar}
-  Then a memory $\store$ satisfies BDOZ preprocessing iff:
-  $$\dom(\store) = \{ \elab{\secret{x}}{1}, \elab{\secret{y}}{2} \} \cup \mathit{shares}
-  \cup \mathit{macs} \cup \mathit{keys}$$
-  and, writing $\store(\macgv{\mesg{w}})$ to denote
-  $\fcod{\store(\elab{\mesg{w}}{1}) + \store(\elab{\mesg{w}}{2})}$,
-  the following conditions hold:
-  \begin{mathpar}
-    \store(\macgv{\mesg{x}}) = \store(\elab{\secret{x}}{1})
-
-    \store(\macgv{\mesg{y}}) = \store(\elab{\secret{y}}{2})
-
-    \fcod{\store(\macgv{\mesg{a}}) * \store(\macgv{\mesg{b}})} = \store(\macgv{\mesg{c}})
-  \end{mathpar}
-  and for all $\cid,\cid' \in \{1,2\}$ with $\cid \ne \cid'$ and $w \in \{ a,b,c,x,y\}$:
-  $$\lcod{\store, \mesg{w\ttt{mac}}}{\cid} =
-  \lcod{\store, \mesg{wk} + \mesg{\ttt{delta}} * \mesg{w}}{\cid'}$$
-\end{definition}
+\subsection{Cheating Detection and Integrity}
 

### 2024-05-14 ppdp24/example-gmw.tex
checkpoint

@@
       and $\condp{\pmf}{X_2}{\detx{X_4}}$, then $\condsep{\pmf}{X_1}{X_4}{X_3}$.
+    %\item Given $\condsep{\pmf}{X_1 \cup X_2}{X_3}{X_4}$ and $\condp{\pmf}{X_1 \cup X_2}{\detx{X_2}}$,
+    %  then $\condsep{\pmf}{X_1}{X_2 \cup X_3}{X_4}$.
   \end{enumerate}
@@
   By induction on the length of $(\varnothing,e) \redxs (\prog,\decodegmw(E[\mesg{z}]))$.
-  Encoding establishes the invarian in the basis. The most interesting inductive
+  Encoding establishes the invariant in the basis. The most interesting inductive
   case is the $\andgmw$ gate. 
@@
 \end{proof}
+
+\begin{theorem}
+  If $e$ is a GMW circuit protocol in $\metaprot$ with $(\varnothing,e) \redxs (\prog,())$
+  then $\prog$ satisfies noninterference modulo output. 
+\end{theorem}
+
+\begin{proof}
+  Given that $e$ is a GMW circuit protocol, then by definition we have:
+  $$
+  (\varnothing,e) \redxs (\prog',\decodegmw(\mesg{z})) \redxs (\prog,())
+  $$
+  where by Lemma \ref{lemma-gmwinvariant} and definition of $\decodegmw$,
+  letting $\iov(\prog) = S \cup M \cup P \cup O$ we
+  have:
+  \begin{mathpar}
+    \conddetx{\progtt(\prog)}{S}{\{\macgv{\mesg{z}}\}}
+
+    \condsep{\progtt(\prog)}{S}{\{\macgv{\mesg{z}}\}}{(M_C \cup \{ \elab{\mesg{z}}{1}, \elab{\mesg{z}}{2} \})}
+  \end{mathpar}
+  so by definition of $\decodegmw$ we have:
+  \begin{mathpar}
+    \conddetx{\progtt(\prog)}{S}{O}
+
+    \condsep{\progtt(\prog)}{S}{O}{(M \cup P)}
+  \end{mathpar}
+\end{proof}

### 2024-05-14 ppdp24/example-gmw.tex
checkpoint

--- 2024-05-14 ppdp24/example-gmw.tex
@@
       and $\condp{\pmf}{X_2}{\detx{X_4}}$, then $\condsep{\pmf}{X_1}{X_4}{X_3}$.
+    \item Given $\condsep{\pmf}{X_1 \cup X_2}{X_3}{X_4}$ and $\condp{\pmf}{X_1 \cup X_2}{\detx{X_2}}$,
+      then $\condsep{\pmf}{X_1}{X_2 \cup X_3}{X_4}$.
   \end{enumerate}
@@
 \end{proof}
+
+\begin{theorem}
+  If $e$ is a GMW circuit protocol in $\metaprot$ with $(\varnothing,e) \redxs (\prog,())$
+  then $\prog$ satisfies noninterference modulo output. 
+\end{theorem}
+
+\begin{proof}
+  Given that $e$ is a GMW circuit protocol, then by definition we have:
+  $$
+  (\varnothing,e) \redxs (\prog',\decodegmw(\mesg{z})) \redxs (\prog,())
+  $$
+  where by Lemma \ref{lemma-gmwinvariant} and definition of $\decodegmw$,
+  for any $H$ and $C$ letting $\iov(\prog) = S \cup M \cup P \cup O$ we
+  have:
+  \begin{mathpar}
+    \conddetx{\progtt(\prog)}{S}{\{\macgv{\mesg{z}}\}}
+
+    \condsep{\progtt(\prog)}{S}{\{\macgv{\mesg{z}}\}}{(M_C \cup \{ \elab{\mesg{z}}{1}, \elab{\mesg{z}}{2} \})}
+  \end{mathpar}
+  so also by definition of $\decodegmw$ we have:
+  \begin{mathpar}
+    \conddetx{\progtt(\prog)}{S_H \cup S_C}{O}
+
+    \condsep{\progtt(\prog)}{S_H \cup S_C}{O}{(M_C \cup P)}
+  \end{mathpar}
+  and this by Lemma \ref{lemma-conditioning} (4) we have:
+  $$
+  \condsep{\progtt(\prog)}{S_C}{(O \cup S_H)}{(M_C \cup P)}
+  $$
+  which implies the result.
+\end{proof}

### 2024-05-14 ppdp24/bruteforce.tex
checkpoint

@@
 hyperproperties discussed previously, or other properties such as
-perfect secrecy \cite{XXX}.  Since $\runs(\prog)$ is exponential in
+perfect secrecy \cite{XXX}.  A basic method to do this is
+to calculate a truth table for the given protocol. Since
+$\runs(\prog)$ is exponential in
 the size of $\secrets(\prog) \cup \flips(\prog)$, this strategy is
-feasible only for smaller protocols. However, we can efficiently
+feasible only for smaller protocols. However, to support some
+scaling we can efficiently
 convert protocols in the passive setting (without $\ttt{assert}$) to
@@
 negation-as-failure model, though we can also use techniques in
-\cite{sakama2017linear} to eliminate negation from resulting programs
+\cite{sakama2017linear} to eliminate negation from resulting programs.
+
+\subsection{Computing a Truth Table}
 
@@
 
-\begin{comment}
 We immediately note that $\runs(\prog)$ can be obtained by a left-folding
@@
       \ \ \{\extend{\store}{\view{\cid}{w}}{1} \mid \store \in \stores' \}\ \cup\\
-      \ \ \{\extend{\store}{\view{\cid}{w}}{0} \mid \store \in \stores - \stores' \}\end{array}\\[2mm]
-    \cruns(\prog) &\defeq& \mathit{foldl}\ {tt}\ \mems(S \cup F)\ \prog
+      \ \ \{\extend{\store}{\view{\cid}{w}}{0} \mid \store \in \stores - \stores' \}\end{array}%\\[2mm]
+    %\cruns(\prog) &\defeq& \mathit{foldl}\ {tt}\ \mems(S \cup R)\ \prog
   \end{eqnarray*}
-  Then $\cruns(\prog) = \runs(\prog)$.
+  Then assuming default preprocessing, $\mathit{foldl}\ {tt}\ \mems(S \cup R)\ \prog = \runs(\prog)$.
 \end{lemma}
-\end{comment}
+However, this method does not take advantage of parallellization,
+in that elements of $\runs(\prog)$ can be calculated independently.
+
+\subsection{Conversion to Stratified Datalog}
 
@@
 $$
-{tt}(\xassign{x}{\be}{\cid}) \defeq (x, (\solve{(\mems(\vars\ \be\ \cid))}{\be}{\cid}))
+\bodies(\xassign{x}{\be}{\cid}) \defeq (x, (\solve{(\mems(\vars\ \be\ \cid))}{\be}{\cid}))
 $$
 \end{definition}
-The mapping of ${tt}$ across a program
-$\prog$-- i.e., $(\mathit{map}\ {tt}\ \prog)$--  essentially defines the
+The mapping of $\bodies$ across a program
+$\prog$-- i.e., $(\mathit{map}\ \bodies\ \prog)$--  essentially defines the
 logic program, modulo some syntactic conversion. We can
@@
 \end{mathpar}
-Given pairs $(x,\stores)$ in the range of ${tt}$, we define the conversion
+Given pairs $(x,\stores)$ in the range of $\bodies$, we define the conversion
 to clauses as  $\mathit{clauses}(x,\{ \store_1,...,\store_n \}) \defeq x \gets \logit{\store_1} \vee \cdots \vee x \gets \logit{\store_n}$.
@@
 $$
-\datalog(\prog) \defeq  \mathit{map}\ \mathit{clauses}\ (\mathit{map}\ {tt}\ \prog)
+\datalog(\prog) \defeq  \mathit{map}\ \mathit{clauses}\ (\mathit{map}\ \bodies\ \prog)
 $$
@@
 We discuss these examples and proof methods in Section
-\ref{section-examples}.
+\ref{section-example-gmw}.

### 2024-05-14 ppdp24/example-bdoz.tex
checkpoint

@@
 
+We can carry out similar proofs of passive security for the protocol in
+Figure \ref{fig-beaver} as for GMW, even using automated tactics for
+the protocol in $\mathbb{F}_2$. But in the case of BDOZ we are also
+concerned with malicious security. To demonstrate this, we can
+demonstrate that the protocol satisfies integrity in the sense of
+Definition \ref{def-integrity}. To do so, we observe that it satisfies
+a stronger property, that we call cheating detection. Intuitively,
+integrity says that the only thing that the adversary can do in the
+malicious model is to elicit the same responses from honest parties
+that an honest run of the protocol would elicit. Cheating detection
+says that the adversary can only execute the protocol honestly, or
+or else gets caught (and abort).
+
+Focusing in, we identify the adversarial inputs as the messages
+sent from the adversary to honest parties, on which honest responses
+to the adversary may depend. We want to say that these are the messages
+that must be legitimate.
 \begin{definition}
@@
 \end{definition}
+Now, we can characterize protocols with cheating detection as those where
+adversarial inputs to honest reponses must themselves be constructed honestly. 
 \begin{definition}[Cheating Detection]
@@
 
+It is straightforward to demonstrate that cheating detection has integrity,
+since only the ``passive'' adversary can elicit a response from honest parties. 
 \begin{lemma}
@@
 \end{lemma}
+
+In the case of BDOZ, cheating detection is accomplished by the information-theoretic
+security of the encryption scheme\cite{XXX}. Furthermore, the symmetry of
+the protocol in Figure \ref{fig-beaver} ensures that both parties will authenticate
+shares, so it is robust to corruption of either party.

### 2024-05-14 ppdp24/model.tex
checkpoint

@@
 \begin{definition}
+  \label{def-progtt}
   \label{def-progd}
@@
   $$
-  Then the \emph{basic distribution of $\prog$}, written $\progtt(\prog)$, is
+  By default, $\preproc(\store) \iff \dom(\store) = S$, i.e.,
+  the initial memory contains all input secrets in a uniform
+  marginal distribution. Then the \emph{basic distribution of $\prog$}, written $\progtt(\prog)$, is
   defined such that for all $\store \in \mems(\iov(\prog) \cup R)$:
@@
   $$
-  By default, $\preproc(\store) \iff \dom(\store) = S$, i.e.,
-  the initial memory contains all input secrets in a uniform
-  marginal distribution.
 
@@
   $$
-  and also define the following which pads out undefined views and outputs with $\bot$:
+  where by default, $\preproc(\store) \iff \dom(\store) = S$, and also define the following
+  which pads out undefined views and outputs with $\bot$:
   $$

### 2024-05-14 ppdp24/example-bdoz.tex
checkpoint

@@
 parties interact in secure opening of $x \fplus a$ and $y + b$,
-followed by local computation to compute their share of the $x * y$
-using only local addition and product as per standard protocol
+followed by the non-interactive computation of shares of $x * y$
+for secure reveals as per standard protocol
 \cite{XXX}. The non-interactive reconstruction of the local authentication

### 2024-05-15 ppdp24/main.tex
checkpoint

@@
 \usepackage{booktabs,array}
+\usepackage{rotating}
 
@@
 
+\acmConference[PPDP]{Principles and Practice of Declarative Programming}{2024}{Milan}
+
 \begin{document}
 
-\title{Language-Based Low-Level MPC Security: $\metaprot$ to $\minifed$}
+\title{Language-Based Security for Low-Level MPC}
 
@@
 \begin{abstract}
-Secure Multi-Party Computation (MPC) protocols support data privacy in
-important modern distributed applications. Security for MPC protocols
-is superficially similar to probabilistic noninterference, but differs
-in a subtle but fundamental way, and approaches for verifying
-noninterference cannot naturally extend to MPC security.  Currently,
-proof methods for MPC protocols are well-studied but manual and thus
-tedious and error-prone, and are also non-standardized and unfamiliar
-to most PL theorists.  Our goal is to leverage connections between the
-security model of MPC and trace-based hyperproperties to obtain
-automated proof methods for MPC protocol development.  We develop a
-language model with a tightly coupled notion of probabilistic program
-distributions, as a foundation for fully and partially automated
-verification of passive MPC security in protocols including
-arbitrarily large YGC and GMW circuits.
+  Secure Multi-Party Computation (MPC) protocols is an important
+  enabling technology for data privacy in modern distributed
+  applications. Currently, proof methods for low-level MPC protocols
+  are primarily manual and thus tedious and error-prone, and are also
+  non-standardized and unfamiliar to most PL theorists. We develop a
+  new staged programming language for defining a variety of low-level
+  protocols with probabilistic features. We also formulate a
+  collection of confidentiality and integrity hyperproperties for our
+  language model that are familiar from information flow, including
+  conditional noninterference, delimited release, and robust
+  declassification. We demonstrate their relation to standard MPC
+  threat models of passive and malicious security, and how they can be
+  leveraged in security verification of protocols. To prove these
+  properties we develop automated tactics in $\mathbb{F}_2$ that
+  can be integrated with separation-logic style reason. 
 \end{abstract}
@@
 %% the work being presented. Separate the keywords with commas.
-\keywords{Secure multiparty computation, security verification, probabilistic programming, static analysis.}
+\keywords{Secure multiparty computation, security verification, probabilistic programming, programming languages, information flow.}
 
 \maketitle
+
+\input{intro}
 
@@
 
+\input{bruteforce}
+
+\input{metalang}
+
+\input{example-gmw}
+
+\input{example-bdoz}
+
 \bibliographystyle{ACM-Reference-Format}

### 2024-05-15 ppdp24/main.tex
checkpoint

@@
   are primarily manual and thus tedious and error-prone, and are also
-  non-standardized and unfamiliar to most PL theorists. We develop a
-  new staged programming language for defining a variety of low-level
-  protocols with probabilistic features. We also formulate a
-  collection of confidentiality and integrity hyperproperties for our
-  language model that are familiar from information flow, including
-  conditional noninterference, delimited release, and robust
-  declassification. We demonstrate their relation to standard MPC
-  threat models of passive and malicious security, and how they can be
-  leveraged in security verification of protocols. To prove these
-  properties we develop automated tactics in $\mathbb{F}_2$ that
-  can be integrated with separation-logic style reason. 
+  non-standardized and unfamiliar to most PL theorists. As a step
+  towards better language support and language-based enforcement, we
+  develop a new staged PL for defining a variety of low-level
+  probabilistic MPC protocols. We also formulate a collection of
+  confidentiality and integrity hyperproperties for our language model
+  that are familiar from information flow, including conditional
+  noninterference, delimited release, and robust declassification. We
+  demonstrate their relation to standard MPC threat models of passive
+  and malicious security, and how they can be leveraged in security
+  verification of protocols. To prove these properties we develop
+  automated tactics in $\mathbb{F}_2$ that can be integrated with
+  separation-logic style reasoning.
 \end{abstract}

### 2024-05-17 ppdp24/model.tex
tiny tweaks and a note

@@
   to denote the probability that $\SIM(\store,v)$ returns corrupt views
-  $\store'$ as a result  we write $\dist(\SIM(\store,v))$ to
+  $\store'$ as a result. We write $\dist(\SIM(\store,v))$ to
   denote the distribution of corrupt views reconstructed by the
@@
   and for all $\store \in \mems(S)$ there exists a
-  simulation $\SIM$ such that:
+  simulation $\SIM$ such that:\todo{is the order of quantifiers right? Doesn't a single simulator have to work for all inputs?}
   $$

### 2024-05-17 ppdp24/intro.tex
checkpoint
tiny tweaks and a note

--- 2023-11-05 revision-1/intro.tex
@@
 blockchains \cite{ishai2009zero, lu2019honeybadgermpc,
-  gao2022symmeproof, tomaz2020preserving}. MPC methods have been
-developed by the cryptography community for many years, while
-receiving the attention of the programming languages community more
-recently. In contrast, information flow security has received
+  gao2022symmeproof, tomaz2020preserving}. The security semantics of
+MPC include both confidentiality and integrity properties incorporated
+into models such as real/ideal (aka simulator) security and universal
+composability (UC), developed primarily by the cryptography community
+\cite{XXX}.  Related proof methods are well-studied \cite{Lindell2017}
+but mostly manual. Somewhat independently, a significant
+body of work in programming languages has focused on definition and
+enforcement of confidentiality and integrity \emph{hyperproperties}
+\cite{10.5555/1891823.1891830} such as noninterference and delimited
+release \cite{XXX}. But following a tradition of connecting
+cryptographic and PL-based security models \cite{XXX}, recent work has
+also recognized connections between MPC security models and
+hyperpoperties of, e.g., noninterference \cite{XXX}, and even
+leveraged these connections to enforce MPC security through mechanisms
+such as security types \cite{XXX}. Major benefits of this connection
+in an MPC setting include better language abstractions for defining
+protocols and for mechanisation and even automation of security proofs.
+
+The goal of this paper is to develop a PL model for defining a variety of
+low-level probabilistic MPC protocols, to formulate a collection
+of confidentiality and integrity hyperproperties for our model
+model with familiar information flow analogs, and to show how these
+properties can be leveraged for improved proof automation.
+
+The distinction between high- and low-level languages for MPC is
+important. High-level languages such as Wysteria \cite{XXX} and
+Viaduct \cite{XXX} are designed to provide effective programming of
+full applications. These language designs incorporate sophisticated
+verified compilation techniques such as orchestration \cite{XXX} to
+guarantee high-level security properties, and they rely on
+\emph{libraries} of low-level MPC protocols, such as binary and
+arithmetic circuits. These low-level protocols encapsulate
+abstractions such as secret sharing and semi-homomorphic encrypytion,
+and rely on probabilistic progamming methods. So, low-level MPC
+programming and protocol verification is a distinct challenge but
+critical to the general challenge of PL design for MPC, and complementary
+to high-level language design.
+
+Analysis of
+
+have been developed
+
+The security
+threat models and engMPC is mainly the
+product of a significant body of work in cryptography,
+
+information flow security has received
 significant attention in PL theory and practice especially since the
@@
 
+\compfig
+
 \subsection{Related Work}

### 2024-05-17 ppdp24/example-gmw.tex

@@
 
-\paragraph{Oblivious Transfer} A passive secure OT protocol
+\paragraph{Oblivious Transfer} A passive secure oblivious transfer (OT) protocol
 based on previous work \cite{XXX} can be defined in $\metaprot$,

### 2024-05-17 ppdp24/intro.tex
checkpoint

@@
 \cite{10.5555/1891823.1891830} such as noninterference and delimited
-release \cite{XXX}. But following a tradition of connecting
+release \cite{XXX}. Following a tradition of connecting
 cryptographic and PL-based security models \cite{XXX}, recent work has
 also recognized connections between MPC security models and
-hyperpoperties of, e.g., noninterference \cite{XXX}, and even
+hyperpoperties of, e.g., noninterference \cite{8429300}, and even
 leveraged these connections to enforce MPC security through mechanisms
@@
 protocols and for mechanisation and even automation of security proofs.
-
 The goal of this paper is to develop a PL model for defining a variety of
@@
 and rely on probabilistic progamming methods. So, low-level MPC
-programming and protocol verification is a distinct challenge but
-critical to the general challenge of PL design for MPC, and complementary
-to high-level language design.
+programming and protocol verification is a distinct challenge and both
+critical to the general challenge of PL design for MPC and
+complementary to high-level language design.
 
-Analysis of
+The connection between information flow hyperproperties and
+MPC security is also complicated especially at a low level.
+MPC protocols involve communication between a group of distributed
+clients called a \emph{federation} that collaboratively compute
+and publish the result of some known \emph{ideal functionality}
+$\idealf$, maintaining confidentiality of inputs to $\idealf$ without
+the use of a trusted third party. However, since the outputs of
+$\idealf$ are public, some information about inputs is inevitably
+leaked. Thus, the ideal
+functionality establishes a declassification policy \cite{XXX}, which
+is more difficult to enforce than pure noninterference.  And
+subtleties of, e.g., semi-homomorphic encryption are central to both
+confidentiality and integrity properties of protocols and similarly
+difficult to track with coarse-grained security types alone.
 
-have been developed
+Nevertheless, as previous authors have observed \cite{XXX}, low-level
+protocol design patterns such as secret sharing and circuit gate
+structure have compositional properties that can be abstracted and
+leveraged in larger proof contexts. We contribute to this line of work
+by developing an automated verification technique for subprotocols and
+show how it can be integrated as a tactic in a larger security proof.
 
-The security
-threat models and engMPC is mainly the
-product of a significant body of work in cryptography,
+\compfig
 
-information flow security has received
-significant attention in PL theory and practice especially since the
-turn of the century \cite{1159651}, including a menagerie of variants,
-enforcement mechanisms, and programming frameworks. Much of this has
-been enabled by the unified metatheory of \emph{hyperproperties}
-\cite{10.5555/1891823.1891830}, including noninterference, that
-establishes a common conceptual framework for reasoning about and
-implementing systems with information flow security.
+\subsection{Related Work}
+\label{section-related-work}
 
-On its surface, security for MPC protocols seems similar to
-probabilistic noninterference. However, MPC security differs from
-noninterference in a subtle but fundamental way, and approaches for
-verifying noninterference cannot naturally extend to MPC security. Our
-goal is to explore connections between the security model of MPC-- the
-\emph{real/ideal} aka \emph{simulator security} model -- and
-trace-based hyperproperties, and to leverage these connections to
-obtain automated enforcement mechanisms for MPC protocol
-development. Currently, proof methods for MPC protocol development are
-well-studied \cite{Lindell2017} with some previous work leveraging
-connections with hyperproperties \cite{8429300}. But these methods
-require fully manual proof construction and are thus tedious and
-error-prone. Therefore our exploration makes both theoretical and
-practical contributions.
+Our main focus is on PL design and automated and semi-automated
+reasoning about security properties of low-level MPC protocols.  We
+consider related systems in several dimensions, including
+whether they are aimed at low-level design with probabilistic features,
+whether they support reasoning about conditional probabilities which
+are central to real/ideal security, whether they consider MPC
+security through the lens of hyperproperties, and whether they consider
+passive and/or malicious security models. We summarize this comparison
+in Figure \ref{fig-comp}, with the caveat that works vary in the degree
+of development in each dimension.
 
-\subsection{The Analysis Challenge of MPC}
-MPC protocols involve communication
-between a group of distributed participants called a \emph{federation}
-that collaboratively compute and publish the result of some known
-\emph{ideal functionality} $\idealf$, while keeping each party's input
-``secret'', without the use of a trusted third party. This last part
-is critical. For example, if we take $\idealf$ to be the majority vote
-function, a protocol for computing $\idealf$ is MPC-secure if, given
-any set of input votes, it correctly computes and publishes the voting
-result but reveals no other information to the public or to other
-participants. However, by publishing the result, some information
-about individual votes may be implicitly declassified.  For example,
-in the case of a majority vote in a federation of size 3, if the
-motion carries and party 1 has voted no, then party 1 knows exactly
-the votes of parties 2 and 3. This cannot be avoided due to the nature
-of $\idealf$.
+As mentioned above, several high-level languages have been developed
+for writing MPC applications, 
+and frequently exploit the connection between hyperproperties and MPC
+security. Previous work on analysis for the SecreC language
+\cite{almeida2018enforcing,10.1145/2637113.2637119} is concerned with
+properties of complex MPC circuits, in particular a user-friendly
+specification and automated enforcement of declassification bounds in
+programs that use MPC in subprograms. The Wys$^\star$ language
+\cite{wysstar}, based on Wysteria \cite{rastogi2014wysteria}, has
+similar goals and includes a trace-based semantics for reasoning about
+the interactions of MPC protocols. Their compiler also guarantees that
+underlying multi-threaded protocols enforce the single-threaded source
+language semantics. These two lines of work were focused on passive
+security. The Viaduct language \cite{XXX} has a well-developed
+information flow type system that automatically enforces both
+confidentiality and integrity through hyperproperties such as robust
+declassification, in addition to rigorous compilation guarantees
+through orchestration \cite{XXX}. However, these high level
+languages lack probabilistic features and other abstractions of
+low-level protocols, the implementation and security of which
+are typically assumed as a selection of library compnonents.
 
-Security in the MPC setting thus means that protocols cannot reveal
-any secret information other than what is implicitly declassified by
-the publicized output of the ideal functionality. The security model
-also assumes that some subset of participants can be corrupted and
-collude adversarially to possibly infer more secret information. The
-accepted method of demonstrating protocol security in this setting is
-to define a \emph{simulation} algorithm that runs in the ``ideal''
-world which, given just the inputs of corrupted parties and the output
-of the ideal functionality, is able to reconstruct information that
-corrupted parties receive in their so-called \emph{views} of the
-protocol running in the real world.  This implies that adversarial
-views provide no information beyond what is provided by the ideal
-output alone. Simulation is defined probabilistically since MPC
-protocols typically rely on cryptographic and probabilistic methods.
+Various related low-level languages with probabilistic features have
+also been developed. The $\lambda_{\mathrm{obliv}}$ language
+\cite{darais2019language} uses a type system is used to automatically
+enforce so-called probabilistic trace obliviousness.  But similar to
+previous work on oblivious data structures \cite{10.1145/3498713},
+obliviousness is related to pure noninterference, not the relaxed form
+related to passive MPC security. The Haskell-based security type system in
+\cite{XXX} enforces a version of conditioned noninterence that is
+sound for passive security in a probabilisitic low-level setting, but
+they do not consider malicious security. And properties of real/ideal
+passive and malicious security for a probabilistic language have been
+formulated in EasyCrypt \cite{8429300}-- though their proof methods,
+while mechanized, are fully manual, and their formulation of malicious
+security is not as clearly related to robust declassification as the
+one we present in Section \ref{section-hyperprop}.
 
-In Section \ref{section-hyperprop-passive} we formalize real/ideal
-security, and we define and discuss simple examples of MPC protocols
-in Section \ref{section-minicat-examples} and more complex ones in
-Section \ref{section-metalang}. But an immediate and main point in
-relation to security hyperproperties is that, due to the potential for
-information release in MPC, simulator security is \emph{not} a strict
-probabilistic noninterference or trace obliviousness property, as we
-show in Section \ref{section-hyperprop-ni}- rather, the public output
-allows and sets an upper bound on declassification.
+Program logics for probabilistic languages and specifically reasoning
+about properties such as joint probabilistic independence is also
+important related work. Probabilistic Separation Logic (PSL)
+\cite{barthe2019probabilistic} develops a logical framework for
+reasoning about probabilistic independence (aka separation) in
+programs, and they consider several (hyper)properties, such as perfect
+secrecy of one-time-pads and indistinguishability in secret sharing,
+that are critical to MPC. However, their methods are manual, and they
+do not capture notions of conditional independence (separation). This
+latter issue has been addressed in Lilac \cite{XXX}, but the
+application of Lilac-style reasoning to MPC protocols has not
+previously been explored, as we do in Section
+\ref{section-example-gmw}.
+
+Our work also shares many ideas with probabilistic programming
+languages designed to perform (exact or approximate) statistical
+inference~\cite{holtzen2020scaling, carpenter2017stan, wood2014new,
+  bingham2019pyro, albarghouthi2017fairsquare, de2007problog,
+  pfeffer2009figaro, saad2021sppl}. Our setting, however, requires
+verifying properties beyond inference, including conditional
+statistical independence. Recent work by Li et al.~\cite{li2023lilac} proposes a
+manual approach for proving such properties, but does not provide
+automation.
 
@@
 gate implementations.
-
-\compfig
-
-\subsection{Related Work}
-\label{section-related-work}
-
-Our main focus is on automated and semi-automated reasoning about
-security properties of specific protocols, both through whole-program
-analysis and compositional properties of program components that
-support security.
-
-Previous work on analysis for the SecreC language
-\cite{almeida2018enforcing,10.1145/2637113.2637119} is concerned with
-properties of complex MPC circuits, in particular a user-friendly
-specification and automated enforcement of declassification bounds in
-programs that use MPC in subprograms. This work is explicitly
-reminiscent of information flow approaches such as delimited
-information release \cite{10.1007/978-3-540-37621-7_9}, downgrading
-policies \cite{li2005downgrading}, and relaxed noninterference
-\cite{10.1145/1040305.1040319}. However, their program logic assumes
-correctness of the underlying MPC protocols.  The Wys$^\star$ language
-\cite{wysstar}, based on Wysteria \cite{rastogi2014wysteria}, has
-similar goals and includes a trace-based semantics for reasoning about
-the interactions of MPC protocols. Their compiler also guarantees that
-underlying multi-threaded protocols enforce the single-threaded source
-language semantics. However their language ultimately compiles to
-protocols assumed to be secure.  All of these systems leverage type
-systems to obtain automated security analyses.
-
-In the realm of protocol verification, previous work has shown how to
-reduce proofs of simulator security to program equivalence assertions,
-which supports computer-aided verification of MPC protocols in
-EasyCrypt \cite{8429300}. This work also leveraged hyperproperty
-characterizations of simulator security, both passive and
-active. However, their EasyCrpyt security proof constructions are
-entirely manual. Probabilistic Separation Logic
-\cite{barthe2019probabilistic} develops a logical framework for
-reasoning about probabilistic independence in programs.  They consider
-several MPC protocols and show how to systematically deduce, e.g.,
-probabilistic independence of input secrets and adversarial views that
-are critical security properties.  However, their methods are manual. 
-Also, their focus on independence does not address the subtleties of
-dependence that arise during MPC protocol execution, and the authors
-do not formulate or prove simulator security properties in
-\cite{barthe2019probabilistic}. The $\lambda_{\mathrm{obliv}}$
-language \cite{darais2019language} is perhaps most closely related to
-our work. Here, a type system is used to automatically enforce
-so-called probabilistic trace obliviousness and they show how to use
-their system to implement several variants of oblivious RAM. However,
-obliviousness is related to noninterference and does not address
-allowable information leakage as in MPC, and their use of type
-linearity disallows the reuse of sampled probabilistic values which is
-common in the MPC idiom, as our examples will illustrate. Similarly,
-previous work on oblivious data structures \cite{10.1145/3498713},
-while intended for sound integration into MPC protocols, focus on
-expression and enforcement of obliviousness and do not account for
-allowable information leakage.
-
-Our work also shares many ideas with probabilistic programming
-languages designed to perform (exact or approximate)
-inference~\cite{holtzen2020scaling, carpenter2017stan, wood2014new,
-  bingham2019pyro, albarghouthi2017fairsquare, de2007problog,
-  pfeffer2009figaro, saad2021sppl}. Our setting, however, requires
-verifying properties beyond inference, including conditional
-independence. Recent work by Li et al.~\cite{li2023lilac} proposes a
-manual approach for proving such properties, but does not provide
-automation.

### 2024-05-17 ppdp24/intro.tex
checkpoint

@@
 secrecy of one-time-pads and indistinguishability in secret sharing,
-that are critical to MPC. However, their methods are manual, and they
-do not capture notions of conditional independence (separation). This
-latter issue has been addressed in Lilac \cite{XXX}, but the
+that are critical to MPC. However, their methods are manual, and
+don't include conditional independence (separation). This
+latter issue has been addressed in Lilac \cite{XXX}. The
 application of Lilac-style reasoning to MPC protocols has not
@@
 
-\paragraph{Language design.} In Section \ref{section-minicat} we
+\paragraph{Language design.} In Section \ref{section-lang} we
 develop a new probabilistic programming language $\minifed$ for
-defining synchronous distributed protocols over the binary field. The
-syntax and semantics provides a succinct account of \emph{views} and
-(synchronous) messaging between protocol \emph{clients}. In Section
-\ref{section-metalang} we define a new metaprogramming language
-$\metaprot$ that dynamically generates $\minifed$ protocols. It
-includes control and data structures and is able to express logical
-protocol components, and enjoys a type safety result that guarantees
-that generated protocols are semantically well-defined.
+defining synchronous distributed protocols over an arbitrary
+arithmetic field. The syntax and semantics provides a succinct account
+of synchronous messaging between protocol \emph{clients}. In Section
+\ref{section-metalang} we define a metalanguage $\metaprot$ that
+dynamically generates $\minifed$ protocols. It is able to express
+important low-level abstractions, as we illustrate via implementations
+of protocols including Shamir addition (Section \ref{section-lang}),
+GMW boolean circuits (Section \ref{section-example-gmw}), and Beaver
+Triple multiplication gates with BDOZ authentication (Section
+\ref{section-example-bdoz}).
 

### 2024-05-17 ppdp24/hyper.tex
checkpoint

@@
 \section{Security Hyperproperties}
+\label{section-hyper}
 
@@
 \begin{definition}[Integrity]
+  \label{def-integrity}
   We say that a protocol $\prog$ with $\iov(\prog) = S \cup V \cup O$ has
@@
 \begin{theorem}
+  \label{theorem-integrity}
   If a protocol is passive secure and has integrity, then it

### 2024-05-17 ppdp24/intro.tex
checkpoint

@@
 protocol design patterns such as secret sharing and circuit gate
-structure have compositional properties that can be abstracted and
-leveraged in larger proof contexts. We contribute to this line of work
-by developing an automated verification technique for subprotocols and
-show how it can be integrated as a tactic in a larger security proof.
+structure have compositional properties that can be independently
+verified and then leveraged in larger proof contexts. We
+contribute to this line of work by developing an automated
+verification technique for subprotocols and show how it can be
+integrated as a tactic in a larger security proof.
 
@@
 
-\paragraph{Hyperproperty formulation.} In Section \ref{section-pmf} we
-formulate \emph{program distributions}, a formalism for expressing
-dependencies between secrets and views, useful for expressing
-characteristics of protocols and precisely quantifying allowable
-information leakage in terms of protocol outputs. On this basis we can
-formulate security hyperproperties, including \emph{noninterference
-modulo output ($\NIMO$)} that implies passive simulator security. But
-program distributions are flexible and can be used to certify other
-important properties of protocol components, such as probabilistic
-independence, correlation, and distributions under conditioning.
+\paragraph{Hyperproperty formulation.} In Section \ref{section-model} we
+develop our formalism for expressing the joint probability mass function of
+program variables, and give standard definitions of passive and
+malicious real/ideal security in our model. In Section
+\ref{section-hyper}, we formulate a variety of familiar information
+flow properties in our probabilistic setting, including condition
+noninterference, delimited release, and robust declassification, and
+consider the relation between these and real/ideal security.  While it
+has been previously shown that probabilistic conditional
+noninterference is sound for passive security, we formulate new
+properties of integrity which, paired with passive security, imply
+malicious security (Theorem \ref{theorem-integrity}). We observe
+in Section \ref{section-example-bdoz} that authentication mechanisms
+such as BDOZ/SPDZ style MACs enforce a strictly weaker property
+of ``cheating detection''.
 
 \paragraph{Fully and partially automated verification.} In Section
-\ref{section-automation} we develop an automated technique, aka
-certification, for automatically verifying security properties such as
-$\NIMO$ and other supporting properties of protocol components.  While
-certification has high complexity, a compositional approach allows
-certification of program components in isolation.  This enables a
-semi-automated method for large circuits where we (1) define
-certification for components, (2) certify components, (3) prove
-$\NIMO$ based on the certifiedness of components. We develop novel
-methods partially based on separation logic
-\cite{barthe2019probabilistic} to enable (3).  In Section
-\ref{section-composition} we illustrate our method to obtain
-semi-automated verification for the GMW and Yao's Garbled
-Circuit protocols, where certification can be re-used to validate new
-gate implementations.
+\ref{section-bruteforce} we develop a method for automatically
+computing the pmf of $\minifed$ protocols in $\mathbb{F}_2$, that can
+be automatically queried to enforce hyperperproperties of
+security. This method is perfectly accurate but has high complexity;
+we show this can be partially mitigated by conversion of protocols in
+$\mathbb{F}_2$ to stratified Datalog which is amenable to HPC
+acceleration. Furthermore, in Section \ref{section-examples-gmw} we
+consider in detail how this automated technique can be used as a local
+automated tactic for proving security in arbitrarily large GMW
+circuits using conditional probabilistic independence as in
+\cite{XXX}.

### 2024-05-17 ppdp24/intro.tex
checkpoint

@@
 composability (UC), developed primarily by the cryptography community
-\cite{XXX}.  Related proof methods are well-studied \cite{Lindell2017}
+\cite{evans2018pragmatic}.  Related proof methods are well-studied \cite{Lindell2017}
 but mostly manual. Somewhat independently, a significant
@@
 \cite{10.5555/1891823.1891830} such as noninterference and delimited
-release \cite{XXX}. Following a tradition of connecting
-cryptographic and PL-based security models \cite{XXX}, recent work has
+release \cite{sabelfeld2009declassification}. Following a tradition of connecting
+cryptographic and PL-based security models \cite{10.1007/3-540-44929-9_1}, recent work has
 also recognized connections between MPC security models and
@@
 leveraged these connections to enforce MPC security through mechanisms
-such as security types \cite{XXX}. Major benefits of this connection
+such as security types \cite{10.1145/3453483.3454074}. Major benefits of this connection
 in an MPC setting include better language abstractions for defining
@@
 The distinction between high- and low-level languages for MPC is
-important. High-level languages such as Wysteria \cite{XXX} and
-Viaduct \cite{XXX} are designed to provide effective programming of
-full applications. These language designs incorporate sophisticated
-verified compilation techniques such as orchestration \cite{XXX} to
-guarantee high-level security properties, and they rely on
-\emph{libraries} of low-level MPC protocols, such as binary and
-arithmetic circuits. These low-level protocols encapsulate
+important. High-level languages such as Wysteria \cite{} and Viaduct
+\cite{10.1145/3453483.3454074} are designed to provide effective
+programming of full applications. These language designs incorporate
+sophisticated verified compilation techniques such as orchestration
+\cite{viaduct-UC} to guarantee high-level security properties, and
+they rely on \emph{libraries} of low-level MPC protocols, such as
+binary and arithmetic circuits. These low-level protocols encapsulate
 abstractions such as secret sharing and semi-homomorphic encrypytion,
@@
 
-The connection between information flow hyperproperties and
-MPC security is also complicated especially at a low level.
-MPC protocols involve communication between a group of distributed
-clients called a \emph{federation} that collaboratively compute
-and publish the result of some known \emph{ideal functionality}
-$\idealf$, maintaining confidentiality of inputs to $\idealf$ without
-the use of a trusted third party. However, since the outputs of
-$\idealf$ are public, some information about inputs is inevitably
-leaked. Thus, the ideal
-functionality establishes a declassification policy \cite{XXX}, which
-is more difficult to enforce than pure noninterference.  And
-subtleties of, e.g., semi-homomorphic encryption are central to both
-confidentiality and integrity properties of protocols and similarly
-difficult to track with coarse-grained security types alone.
+The connection between information flow hyperproperties and MPC
+security is also complicated especially at a low level.  MPC protocols
+involve communication between a group of distributed clients called a
+\emph{federation} that collaboratively compute and publish the result
+of some known \emph{ideal functionality} $\idealf$, maintaining
+confidentiality of inputs to $\idealf$ without the use of a trusted
+third party. However, since the outputs of $\idealf$ are public, some
+information about inputs is inevitably leaked. Thus, the ideal
+functionality establishes a declassification policy
+\cite{sabelfeld2009declassification}, which is more difficult to
+enforce than pure noninterference.  And subtleties of, e.g.,
+semi-homomorphic encryption are central to both confidentiality and
+integrity properties of protocols and similarly difficult to track
+with coarse-grained security types alone.
 
-Nevertheless, as previous authors have observed \cite{XXX}, low-level
-protocol design patterns such as secret sharing and circuit gate
-structure have compositional properties that can be independently
-verified and then leveraged in larger proof contexts. We
-contribute to this line of work by developing an automated
-verification technique for subprotocols and show how it can be
-integrated as a tactic in a larger security proof.
+Nevertheless, as previous authors have observed
+\cite{5a51987acaa84c43bb4bf5bcc7d01683}, low-level protocol design
+patterns such as secret sharing and circuit gate structure have
+compositional properties that can be independently verified and then
+leveraged in larger proof contexts. We contribute to this line of work
+by developing an automated verification technique for subprotocols and
+show how it can be integrated as a tactic in a larger security proof.
 
@@
 As mentioned above, several high-level languages have been developed
-for writing MPC applications, 
-and frequently exploit the connection between hyperproperties and MPC
-security. Previous work on analysis for the SecreC language
+for writing MPC applications, and frequently exploit the connection
+between hyperproperties and MPC security. Previous work on analysis
+for the SecreC language
 \cite{almeida2018enforcing,10.1145/2637113.2637119} is concerned with
@@
 language semantics. These two lines of work were focused on passive
-security. The Viaduct language \cite{XXX} has a well-developed
+security. The Viaduct language
+\cite{10.1145/3453483.3454074} has a well-developed
 information flow type system that automatically enforces both
@@
 declassification, in addition to rigorous compilation guarantees
-through orchestration \cite{XXX}. However, these high level
+through orchestration \cite{viaduct-UC}. However, these high level
 languages lack probabilistic features and other abstractions of
-low-level protocols, the implementation and security of which
-are typically assumed as a selection of library compnonents.
+low-level protocols, the implementation and security of which are
+typically assumed as a selection of library compnonents.
 
@@
 related to passive MPC security. The Haskell-based security type system in
-\cite{XXX} enforces a version of conditioned noninterence that is
+\cite{6266151} enforces a version of conditioned noninterence that is
 sound for passive security in a probabilisitic low-level setting, but
@@
 don't include conditional independence (separation). This
-latter issue has been addressed in Lilac \cite{XXX}. The
+latter issue has been addressed in Lilac \cite{li2023lilac}. The
 application of Lilac-style reasoning to MPC protocols has not
@@
 circuits using conditional probabilistic independence as in
-\cite{XXX}.
+\cite{li2023lilac}.

### 2024-05-17 ppdp24/intro.tex
checkpoint

@@
 while mechanized, are fully manual, and their formulation of malicious
-security is not as clearly related to robust declassification as the
-one we present in Section \ref{section-hyperprop}.
+security is not as clearly related to robust declassification as is the
+one we present in Section \ref{section-hyper}.
 

### 2024-05-17 ppdp24/intro.tex
checkpoint

@@
 $\mathbb{F}_2$ to stratified Datalog which is amenable to HPC
-acceleration. Furthermore, in Section \ref{section-examples-gmw} we
+acceleration. Furthermore, in Section \ref{section-example-gmw} we
 consider in detail how this automated technique can be used as a local

### 2024-05-17 ppdp24/example-gmw.tex
checkpoint

@@
 \begin{theorem}
+  \label{theorem-gmw}
   If $e$ is a GMW circuit protocol in $\metaprot$ with $(\varnothing,e) \redxs (\prog,())$

### 2024-05-17 ppdp24/intro.tex
checkpoint

@@
 such as BDOZ/SPDZ style MACs enforce a strictly weaker property
-of ``cheating detection''.
+of ``cheating detection'' (Lemma \ref{lemma-cheating}).
 
@@
 circuits using conditional probabilistic independence as in
-\cite{li2023lilac}.
+\cite{li2023lilac} (Lemmas \ref{lemma-gmwtactic} and \ref{lemma-gmwinvariant}
+and Theorem \ref{theorem-gmw}).

### 2024-05-17 ppdp24/main.tex
tiny tweaks and a note
checkpoint

@@
 \usepackage{rotating}
+\usepackage{xcolor}
 
@@
 
-\author{Joseph Near}
+\author{Joseph P. Near}
 \affiliation{
@@
 \begin{abstract}
-  Secure Multi-Party Computation (MPC) protocols is an important
+  Secure Multi-Party Computation (MPC) is an important
   enabling technology for data privacy in modern distributed
@@
   automated tactics in $\mathbb{F}_2$ that can be integrated with
-  separation-logic style reasoning.
+  separation logic-style reasoning.
 \end{abstract}

### 2024-05-18 ppdp24/intro.tex
checkpoint

@@
 composability (UC), developed primarily by the cryptography community
-\cite{evans2018pragmatic}.  Related proof methods are well-studied \cite{Lindell2017}
-but mostly manual. Somewhat independently, a significant
-body of work in programming languages has focused on definition and
-enforcement of confidentiality and integrity \emph{hyperproperties}
-\cite{10.5555/1891823.1891830} such as noninterference and delimited
-release \cite{sabelfeld2009declassification}. Following a tradition of connecting
-cryptographic and PL-based security models \cite{10.1007/3-540-44929-9_1}, recent work has
-also recognized connections between MPC security models and
-hyperpoperties of, e.g., noninterference \cite{8429300}, and even
-leveraged these connections to enforce MPC security through mechanisms
-such as security types \cite{10.1145/3453483.3454074}. Major benefits of this connection
-in an MPC setting include better language abstractions for defining
-protocols and for mechanisation and even automation of security proofs.
-The goal of this paper is to develop a PL model for defining a variety of
-low-level probabilistic MPC protocols, to formulate a collection
-of confidentiality and integrity hyperproperties for our model
-model with familiar information flow analogs, and to show how these
-properties can be leveraged for improved proof automation.
+\cite{evans2018pragmatic}.  Related proof methods are well-studied
+\cite{Lindell2017} but mostly manual. Somewhat independently, a
+significant body of work in programming languages has focused on
+definition and enforcement of confidentiality and integrity
+\emph{hyperproperties} \cite{10.5555/1891823.1891830} such as
+noninterference and delimited release
+\cite{sabelfeld2009declassification}. Following a tradition of
+connecting cryptographic and PL-based security models
+\cite{10.1007/3-540-44929-9_1}, recent work has also recognized
+connections between MPC security models and hyperpoperties of, e.g.,
+noninterference \cite{8429300}, and even leveraged these connections
+to enforce MPC security through mechanisms such as security types
+\cite{10.1145/3453483.3454074}. Major benefits of this connection in
+an MPC setting include better language abstractions for defining
+protocols and for mechanisation and even automation of security
+proofs.  The goal of this paper is to develop a PL model for defining
+a variety of low-level probabilistic MPC protocols, to formulate a
+collection of confidentiality and integrity hyperproperties for our
+model model with familiar information flow analogs, and to show how
+these properties can be leveraged for improved proof automation.
 
 The distinction between high- and low-level languages for MPC is
-important. High-level languages such as Wysteria \cite{} and Viaduct
-\cite{10.1145/3453483.3454074} are designed to provide effective
-programming of full applications. These language designs incorporate
-sophisticated verified compilation techniques such as orchestration
+important. High-level languages such as Wysteria
+\cite{rastogi2014wysteria} and Viaduct \cite{10.1145/3453483.3454074}
+are designed to provide effective programming of full
+applications. These language designs incorporate sophisticated
+verified compilation techniques such as orchestration
 \cite{viaduct-UC} to guarantee high-level security properties, and

### 2024-05-18 ppdp24/model.tex
checkpoint

@@
 \label{section-pmf}
+\label{section-model}
 
 MPC protocols are intended to implement some \emph{ideal
-functionality} $\idealf$. Given a protocol $\prog$ that implements
-$\idealf$, with $\iov(\prog) = S \cup V \cup O$, we will assume
-that the domain of $\idealf$ is $\mems(S)$ and its range is
-$\mems(O)$. Recall the assumption of a partitioning of a federation
-into honest and corrupt subsets $H$ and $C$.  \emph{Simulator
-security} in the MPC setting means that, given $\store \in \mems(S)$,
-a secure protocol $\prog$ does not reveal any more information about
-honest secrets $\store_H$ to parties in $C$ beyond what is
-implicitly declassified by $\idealf(\sigma)$. Security comes in
-\emph{passive} and \emph{malicious} flavors, wherein the adversary
-either follows the rules or not, respectively. Characterization of
-both real world protocol execution and simulation is defined
+functionality} $\idealf$ with per-client outputs. In the $\minifed$
+setting, Given a protocol $\prog$ that implements $\idealf$, with
+$\iov(\prog) = S \cup V \cup O$, the domain of $\idealf$ is $\mems(S)$
+and its range is $\mems(O)$.  Real/ideal security in the MPC
+setting means that, given $\store \in \mems(S)$, a secure protocol
+$\prog$ does not reveal any more information about honest secrets
+$\store_H$ to parties in $C$ beyond what is implicitly declassified by
+$\idealf(\sigma)$. Security comes in \emph{passive} and
+\emph{malicious} flavors, wherein the adversary either follows the
+rules or not, respectively. Characterization of both real world
+protocol execution and simulation is defined
 probabilistically. Following previous work
@@
 \mapsto 0, \elab{\mesg{y}}{2} \mapsto 1 \})$ to denote the (joint) probability that
-$\elab{\secret{x}}{1} = 0 \wedge \elab{\mesg{y}}{2} = 1$ in the final memory of a
-program containing these variables.
+$\elab{\secret{x}}{1} = 0 \wedge \elab{\mesg{y}}{2} = 1$.
 \begin{definition}

### 2024-05-18 ppdp24/model.tex
checkpoint

@@
 that we denote $\progtt(\prog)$. Since $\fedcat$ is deterministic the
-results of any run are determined by the input secrets together with
+results of any run are determined by the input values together with
 the random tape. And since we constrain programs to not overwrite
@@
 
-We consider random tapes to be selected from a uniform distribution
-modulo relevant pre-processing, e.g., Beaver triples for efficient
-multiplication, and/or MACed share distributions as in SPDZ \cite{XXX}.
-We will also assume that any possible initial secrets
-are equally likely when formulating the program distribution.
-We define $\runs(\prog)$ as the set of final memories resulting
-from execution of $\prog$ given any initial memory, and
-treat all elements of $\runs(\prog)$ as equally likely.
-This establishes the basic program distribution that can be
-marginalized and conditioned to quantify input/output
-information dependencies. 
+Our semantics require that random tapes contain values for all program
+values $\elab{\flip{w}}{\cid}$ sampled from a uniform distribution
+over $\mathbb{F}_p$. Input memories also contain input secret values
+and possibly other initial view elements as a result of
+pre-processing, e.g., Beaver triples for efficient multiplication,
+and/or MACed share distributions as in SPDZ \cite{evans2018pragmatic,10.1007/978-3-030-68869-1_3}.
+We will also assume that any possible initial secrets are equally
+likely when formulating the program distribution.  We define
+$\runs(\prog)$ as the set of final memories resulting from execution
+of $\prog$ given any initial memory, and treat all elements of
+$\runs(\prog)$ as equally likely.  This establishes the basic program
+distribution that can be marginalized and conditioned to quantify
+input/output information dependencies.
 %In this

### 2024-05-18 ppdp24/model.tex
checkpoint

@@
 pre-processing, e.g., Beaver triples for efficient multiplication,
-and/or MACed share distributions as in SPDZ \cite{evans2018pragmatic,10.1007/978-3-030-68869-1_3}.
-We will also assume that any possible initial secrets are equally
-likely when formulating the program distribution.  We define
-$\runs(\prog)$ as the set of final memories resulting from execution
-of $\prog$ given any initial memory, and treat all elements of
-$\runs(\prog)$ as equally likely.  This establishes the basic program
-distribution that can be marginalized and conditioned to quantify
-input/output information dependencies.
+and/or MACed share distributions as in BDOZ/SPDZ
+\cite{evans2018pragmatic,10.1007/978-3-030-68869-1_3}.  We will also
+assume that any possible initial secrets are equally likely when
+formulating the program distribution.  We define $\runs(\prog)$ as the
+set of final memories resulting from execution of $\prog$ given any
+initial memory, and treat all elements of $\runs(\prog)$ as equally
+likely.  This establishes the basic program distribution that can be
+marginalized and conditioned to quantify input/output information
+dependencies.
 %In this

### 2024-05-18 ppdp24/model.tex
checkpoint

@@
 and/or MACed share distributions as in BDOZ/SPDZ
-\cite{evans2018pragmatic,10.1007/978-3-030-68869-1_3}.  We will also
-assume that any possible initial secrets are equally likely when
-formulating the program distribution.  We define $\runs(\prog)$ as the
-set of final memories resulting from execution of $\prog$ given any
-initial memory, and treat all elements of $\runs(\prog)$ as equally
-likely.  This establishes the basic program distribution that can be
-marginalized and conditioned to quantify input/output information
-dependencies.
+\cite{evans2018pragmatic,10.1007/978-3-030-68869-1_3}. We define
+$\runs(\prog)$ as the set of final memories resulting from execution
+of $\prog$ given any initial memory, and treat all elements of
+$\runs(\prog)$ as equally likely.  This establishes the basic program
+distribution that can be marginalized and conditioned to quantify
+input/output information dependencies.
 %In this
@@
 
-  In some cases, we will also be concerned with the (joint)
-  probabilities of expression interpretation given a preceding program
-  execution, and we write $\progtt(\prog, \be)$ to denote the program
-  distribution $\progtt(\prog;\itv := \be)$ where $\itv$ is a
-  special variable that is never used in programs.
+  %In some cases, we will also be concerned with the (joint)
+  %probabilities of expression interpretation given a preceding program
+  %execution, and we write $\progtt(\prog, \be)$ to denote the program
+  %distribution $\progtt(\prog;\itv := \be)$ where $\itv$ is a
+  %special variable that is never used in programs.
 \end{definition}
@@
   Given a program $\prog$ with $\iov(\prog) = S \cup M \cup P \cup O$,
-  define $\views(\prog) \defeq M \cup P$, and define $\houtputs$ as the messages in
-  $V = \views(\prog)$ sent from honest
-  to corrupt parties, called \emph{corrupt views}:
+  define $\views(\prog) \defeq M \cup P$, and define $\houtputs$ as
+  the messages and reveals in $V \defeq M \cup P$ sent from honest to corrupt
+  parties, called \emph{corrupt views}:
   $$
@@
   $$
-  and similarly define $\cinputs$ as the messages in $V$ sent from corrupt to honest
+  and similarly define $\cinputs$ as the subset of $V$ sent from corrupt to honest
   parties, called \emph{honest views}:
@@
 corrupt inputs and the output of an ideal functionality, and that
-returns a distribution of corrupt views that is indistinguishable
-from the distribution of corrupt views in the real world
-protocol execution.
+returns a reconstruction of corrupt views that is probabilistically
+indistinguishable from the corrupt views in the real world protocol
+execution.
 \begin{definition}
@@
   functionality $\idealf$, with $\views(\prog) = V$.  Then $\prog$
-  is \emph{passive secure in the simulator model} iff for all
+  is \emph{passive secure in the simulator model} iff there exists
+  ad simulation $\SIM$ such that for all
   partitions of the federation into honest and corrupt sets $H$ and $C$
-  and for all $\store \in \mems(S)$ there exists a
-  simulation $\SIM$ such that:\todo{is the order of quantifiers right? Doesn't a single simulator have to work for all inputs?}
+  and for all $\store \in \mems(S)$:
   $$

### 2024-05-18 ppdp24/model.tex
checkpoint

@@
 
-Another important distinction of the malicious setting is that it
-typically incorporates \emph{abort}. Honest parties implement
-strategies to detect rule-breaking-- aka \emph{cheating}-- by
-using, e.g., message authentication codes with semi-homomorphic
-properties as in BDOZ and SPDZ \cite{XXX}. If cheating is
-detected, the protocol is aborted. To model this, we extend
-$\minifed$ with an \ttt{assert} command and extend the
-range of memories with $\bot$. Note that the adversary is
-free to ignore their own assertions.
+A key technical distinction of the malicious setting is that it
+typically incorporates ``abort''. Honest parties implement strategies
+to detect rule-breaking-- aka \emph{cheating}-- by using, e.g.,
+message authentication codes with semi-homomorphic properties as in
+BDOZ/SPDZ \cite{10.1007/978-3-030-68869-1_3}. If cheating is detected,
+the protocol is aborted. To model this, we extend $\minifed$ with an
+\ttt{assert} command and extend the range of memories with
+$\bot$. Note that the adversary is free to ignore their own
+assertions.
 \begin{definition}

### 2024-05-18 ppdp24/hyper.tex
checkpoint

@@
 corrupt views give the adversary no better chance of guessing honest
-secrets than the output and corrupt inputs do. So
-the simulator can just arbitrarily pick any honest secrets that
-produce the expected output and run the protocol in simulation to
-reconstruct real world corrupt view distributions. But it requires
-that the simulator can tractably ``pre-image'' a given output of
-a functionality $\idealf$, to determine the inputs that
-could have produced it. This pre-image is called a
-\emph{kernel} in recent work \cite{XXX}.
+secrets than just the output and corrupt inputs do. So the simulator
+can just arbitrarily pick any honest secrets that could have produced
+the given outputs and run the protocol in simulation to reconstruct
+real world corrupt views. This requires that the simulator can
+tractably ``pre-image'' a given output of a functionality $\idealf$,
+to determine the inputs that could have produced it. This equivalence
+class is called a \emph{kernel} in recent work \cite{10.1145/3571740}.
 \begin{definition}
@@
 
-Previous work has discussed how MPC security is not noninterference,
-but rather how ideal functional sets an upper bound on
-declassification \cite{6266151,almeida2018enforcing}. Nevertheless,
-probabilistic noninterference is preserved by components of
-cryptographic protocols generally, and can be expressed using
-\emph{probabilistic independence} \cite{darais2019language,barthe2019probabilistic}
-We introduce important notation to express independence:
+Probabilistic noninterference is related to perfect secrecy and is
+preserved by components of cryptographic protocols generally. It can
+be expressed using probabilistic independence, aka separation,
+\cite{darais2019language,barthe2019probabilistic}, and we adopt the
+following notation to express independence:
 \begin{definition}
@@
 
-In practice, MPC protocols typically satisfy a \emph{delimited release}
-property \cite{sabelfeld2009declassification}, where messages exchanged remain probabilistically separable
-from secrets, with only declassification events (reveals and outputs)
-releasing information about honest secrets. A key difference is that while
-these declassification events essentially define the policy in delimited
+In practice, MPC protocols typically satisfy a \emph{delimited
+release} property \cite{sabelfeld2009declassification}, where messages
+exchanged remain probabilistically separable from secrets, with only
+declassification events (reveals and outputs) releasing information
+about honest secrets.  A key difference is that while these
+declassification events essentially define the policy in delimited
 release, the ideal functionality sets the release policy for MPC
-passive security, so its necessary to show that declassification events respect
-these bounds.
+passive security, so its necessary to show that declassification
+events respect these bounds.
 \begin{definition}

### 2024-05-18 ppdp24/intro.tex
checkpoint

@@
 connecting cryptographic and PL-based security models
-\cite{10.1007/3-540-44929-9_1}, recent work has also recognized
-connections between MPC security models and hyperpoperties of, e.g.,
-noninterference \cite{8429300}, and even leveraged these connections
-to enforce MPC security through mechanisms such as security types
-\cite{10.1145/3453483.3454074}. Major benefits of this connection in
-an MPC setting include better language abstractions for defining
-protocols and for mechanisation and even automation of security
-proofs.  The goal of this paper is to develop a PL model for defining
-a variety of low-level probabilistic MPC protocols, to formulate a
-collection of confidentiality and integrity hyperproperties for our
-model model with familiar information flow analogs, and to show how
-these properties can be leveraged for improved proof automation.
+\cite{10.1007/3-540-44929-9_1,10.1145/3571740}, recent work has also
+recognized connections between MPC security models and hyperpoperties
+of, e.g., noninterference \cite{8429300}, and even leveraged these
+connections to enforce MPC security through mechanisms such as
+security types \cite{10.1145/3453483.3454074}. Major benefits of this
+connection in an MPC setting include better language abstractions for
+defining protocols and for mechanisation and even automation of
+security proofs.  The goal of this paper is to develop a PL model for
+defining a variety of low-level probabilistic MPC protocols, to
+formulate a collection of confidentiality and integrity
+hyperproperties for our model model with familiar information flow
+analogs, and to show how these properties can be leveraged for
+improved proof automation.
 

### 2024-05-19 ppdp24/intro.tex
tiny tweak in intro

@@
 abstractions such as secret sharing and semi-homomorphic encrypytion,
-and rely on probabilistic progamming methods. So, low-level MPC
-programming and protocol verification is a distinct challenge and both
+and must be verified by hand. So, low-level MPC
+programming and protocol verification remains a distinct challenge and both
 critical to the general challenge of PL design for MPC and

### 2024-05-19 ppdp24/hyper.tex
checkpoint

@@
 
-\subsection{Delimited Release}
+\subsection{Gradual Release}
 
@@
 
-In practice, MPC protocols typically satisfy a \emph{delimited
+In practice, MPC protocols typically satisfy a \emph{gradual
 release} property \cite{sabelfeld2009declassification}, where messages
@@
 about honest secrets.  A key difference is that while these
-declassification events essentially define the policy in delimited
+declassification events essentially define the policy in gradual
 release, the ideal functionality sets the release policy for MPC
@@
   Given $H,C$, a protocol $\prog$ with $\iov(\prog) = S \cup M \cup P \cup O$
-  satisfies \emph{delimited release} iff
+  satisfies \emph{gradual release} iff
   $\sep{\progtt(\prog)}{M_C}{S_H}$.
@@
 by policy. But in this prior work, this policy is established
-by the declassifications themselves, as in delimited release.
+by the declassifications themselves, as in gradual release.
 Thus, we can define a robust declassification property as follows. 
@@
   A protocol satisfies \emph{robust declassification} iff it has integrity and
-  satisfies delimited release. 
+  satisfies gradual release. 
 \end{definition}
-However, it is important to note that delimited release, and
+However, it is important to note that gradual release, and
 hence robust declassification, are not sufficient to establish

### 2024-05-19 ppdp24/intro.tex
checkpoint

--- 2024-05-18 ppdp24/intro.tex
@@
 \emph{hyperproperties} \cite{10.5555/1891823.1891830} such as
-noninterference and delimited release
-\cite{sabelfeld2009declassification}. Following a tradition of
+noninterference and gradual release
+\cite{4223226,sabelfeld2009declassification}. Following a tradition of
 connecting cryptographic and PL-based security models
@@
 flow properties in our probabilistic setting, including condition
-noninterference, delimited release, and robust declassification, and
+noninterference, gradual release, and robust declassification, and
 consider the relation between these and real/ideal security.  While it

### 2024-05-19 ppdp24/main.tex
checkpoint

@@
   that are familiar from information flow, including conditional
-  noninterference, delimited release, and robust declassification. We
+  noninterference, gradual release, and robust declassification. We
   demonstrate their relation to standard MPC threat models of passive
@@
 
+\begin{comment}
 %%
@@
 \keywords{Secure multiparty computation, security verification, probabilistic programming, programming languages, information flow.}
+\end{comment}
 

### 2024-05-20 ppdp24/intro.tex
tweak to intro

@@
 
+Our work provides a complete methodology for end-to-end verification
+of MPC protocols via three components:
+%
+\begin{enumerate}
+\item A \textbf{low-level language} for defining MPC primitives
+  (\S\ref{section-lang}).
+\item A \textbf{fully-automated verification method} for low-level MPC
+  primitives in $\mathbb{F}_2$ (\S\ref{section-bruteforce}).
+\item A \textbf{partially-automated verification method} for
+  high-level MPC protocols, which leverages automated proofs for
+  low-level primitives (\S\ref{section-example-gmw}).
+\end{enumerate}
+%
+As part of this methodology, we also develop hyperproperties that
+encode MPC security that may be of independent interest. The complete
+methodology enables the verification of real-world MPC protocols like
+GMW~\cite{goldreich2019play}.
+
 \paragraph{Language design.} In Section \ref{section-lang} we
@@
 \ref{section-hyper}, we formulate a variety of familiar information
-flow properties in our probabilistic setting, including condition
+flow properties in our probabilistic setting, including conditional
 noninterference, gradual release, and robust declassification, and

### 2024-05-20 ppdp24/example-gmw.tex
checkpoint
tweak to intro

@@
       and $\condp{\pmf}{X_2}{\detx{X_4}}$, then $\condsep{\pmf}{X_1}{X_4}{X_3}$.
-    \item Given $\condsep{\pmf}{X_1 \cup X_2}{X_3}{X_4}$ and $\condp{\pmf}{X_1 \cup X_2}{\detx{X_2}}$,
+    \item Given $\condsep{\pmf}{X_1 \cup X_2}{X_3}{X_4}$ and $\condp{\pmf}{X_1 \cup X_2}{\detx{X_3}}$,
       then $\condsep{\pmf}{X_1}{X_2 \cup X_3}{X_4}$.
@@
 
-  \conddetx{\progtt(\prog)}{S^1}{\{ \macgv{\mesg{x}} \}}    
+  \conddetx{\progtt(\prog)}{S^2}{\{ \macgv{\mesg{y}} \}}    
   \end{mathpar}
@@
     \progtt(\prog)|S^1 \cup S^2) \vdash \\
-    {\{ \macgv{\mesg{z}} \}}{(M^1_C \cup M^2_C \cup \{ \elab{\mesg{x}}{2}, \elab{\mesg{y}}{2}, \elab{\mesg{z}}{1}, \elab{\mesg{z}}{2} \})}
+    {\{ \macgv{\mesg{z}} \}} * {(M^1_C \cup M^2_C \cup \{ \elab{\mesg{x}}{2}, \elab{\mesg{y}}{2}, \elab{\mesg{z}}{1}, \elab{\mesg{z}}{2} \})}
   \end{array}
@@
   \end{mathpar}
-  so also by definition of $\decodegmw$ we have:
+  and by definition of $\decodegmw$ the value $\{\macgv{\mesg{z}}\}$ is output
+  after the parties publicly reveal $\{ \elab{\mesg{z}}{1}, \elab{\mesg{z}}{2} \}$, so we have:
   \begin{mathpar}
@@
   \end{mathpar}
-  and this by Lemma \ref{lemma-conditioning} (4) we have:
+  and thus by Lemma \ref{lemma-conditioning} (4) we have:
   $$

### 2024-05-21 ppdp24/intro.tex

@@
 
-\compfig
-
-\subsection{Related Work}
-\label{section-related-work}
-
-Our main focus is on PL design and automated and semi-automated
-reasoning about security properties of low-level MPC protocols.  We
-consider related systems in several dimensions, including
-whether they are aimed at low-level design with probabilistic features,
-whether they support reasoning about conditional probabilities which
-are central to real/ideal security, whether they consider MPC
-security through the lens of hyperproperties, and whether they consider
-passive and/or malicious security models. We summarize this comparison
-in Figure \ref{fig-comp}, with the caveat that works vary in the degree
-of development in each dimension.
-
-As mentioned above, several high-level languages have been developed
-for writing MPC applications, and frequently exploit the connection
-between hyperproperties and MPC security. Previous work on analysis
-for the SecreC language
-\cite{almeida2018enforcing,10.1145/2637113.2637119} is concerned with
-properties of complex MPC circuits, in particular a user-friendly
-specification and automated enforcement of declassification bounds in
-programs that use MPC in subprograms. The Wys$^\star$ language
-\cite{wysstar}, based on Wysteria \cite{rastogi2014wysteria}, has
-similar goals and includes a trace-based semantics for reasoning about
-the interactions of MPC protocols. Their compiler also guarantees that
-underlying multi-threaded protocols enforce the single-threaded source
-language semantics. These two lines of work were focused on passive
-security. The Viaduct language
-\cite{10.1145/3453483.3454074} has a well-developed
-information flow type system that automatically enforces both
-confidentiality and integrity through hyperproperties such as robust
-declassification, in addition to rigorous compilation guarantees
-through orchestration \cite{viaduct-UC}. However, these high level
-languages lack probabilistic features and other abstractions of
-low-level protocols, the implementation and security of which are
-typically assumed as a selection of library compnonents.
-
-Various related low-level languages with probabilistic features have
-also been developed. The $\lambda_{\mathrm{obliv}}$ language
-\cite{darais2019language} uses a type system is used to automatically
-enforce so-called probabilistic trace obliviousness.  But similar to
-previous work on oblivious data structures \cite{10.1145/3498713},
-obliviousness is related to pure noninterference, not the relaxed form
-related to passive MPC security. The Haskell-based security type system in
-\cite{6266151} enforces a version of conditioned noninterence that is
-sound for passive security in a probabilisitic low-level setting, but
-they do not consider malicious security. And properties of real/ideal
-passive and malicious security for a probabilistic language have been
-formulated in EasyCrypt \cite{8429300}-- though their proof methods,
-while mechanized, are fully manual, and their formulation of malicious
-security is not as clearly related to robust declassification as is the
-one we present in Section \ref{section-hyper}.
-
-Program logics for probabilistic languages and specifically reasoning
-about properties such as joint probabilistic independence is also
-important related work. Probabilistic Separation Logic (PSL)
-\cite{barthe2019probabilistic} develops a logical framework for
-reasoning about probabilistic independence (aka separation) in
-programs, and they consider several (hyper)properties, such as perfect
-secrecy of one-time-pads and indistinguishability in secret sharing,
-that are critical to MPC. However, their methods are manual, and
-don't include conditional independence (separation). This
-latter issue has been addressed in Lilac \cite{li2023lilac}. The
-application of Lilac-style reasoning to MPC protocols has not
-previously been explored, as we do in Section
-\ref{section-example-gmw}.
-
-Our work also shares many ideas with probabilistic programming
-languages designed to perform (exact or approximate) statistical
-inference~\cite{holtzen2020scaling, carpenter2017stan, wood2014new,
-  bingham2019pyro, albarghouthi2017fairsquare, de2007problog,
-  pfeffer2009figaro, saad2021sppl}. Our setting, however, requires
-verifying properties beyond inference, including conditional
-statistical independence. Recent work by Li et al.~\cite{li2023lilac} proposes a
-manual approach for proving such properties, but does not provide
-automation.
-
 \subsection{Overview and Contributions}
@@
 and Theorem \ref{theorem-gmw}).
+
+\compfig
+
+\subsection{Related Work}
+\label{section-related-work}
+
+Our main focus is on PL design and automated and semi-automated
+reasoning about security properties of low-level MPC protocols. Prior
+work has considered \textbf{automated} verification of
+\textbf{high-level} protocols and \textbf{manual} verification of
+\textbf{low-level} protocols---but none offers the combination of
+automation and low-level support we consider.
+%
+We summarize this comparison in Figure
+\ref{fig-comp}, with the caveat that works vary in the degree of
+development in each dimension.
+
+% We consider related systems in several
+% dimensions, including whether they are aimed at low-level design with
+% probabilistic features, whether they support reasoning about
+% conditional probabilities which are central to real/ideal security,
+% whether they consider MPC security through the lens of
+% hyperproperties, and whether they consider passive and/or malicious
+% security models.
+
+As mentioned above, several high-level languages have been developed
+for writing MPC applications, and frequently exploit the connection
+between hyperproperties and MPC security. Previous work on analysis
+for the SecreC language
+\cite{almeida2018enforcing,10.1145/2637113.2637119} is concerned with
+properties of complex MPC circuits, in particular a user-friendly
+specification and automated enforcement of declassification bounds in
+programs that use MPC in subprograms. The Wys$^\star$ language
+\cite{wysstar}, based on Wysteria \cite{rastogi2014wysteria}, has
+similar goals and includes a trace-based semantics for reasoning about
+the interactions of MPC protocols. Their compiler also guarantees that
+underlying multi-threaded protocols enforce the single-threaded source
+language semantics. These two lines of work were focused on passive
+security. The Viaduct language
+\cite{10.1145/3453483.3454074} has a well-developed
+information flow type system that automatically enforces both
+confidentiality and integrity through hyperproperties such as robust
+declassification, in addition to rigorous compilation guarantees
+through orchestration \cite{viaduct-UC}. However, these high level
+languages lack probabilistic features and other abstractions of
+low-level protocols, the implementation and security of which are
+typically assumed as a selection of library compnonents.
+
+Various related low-level languages with probabilistic features have
+also been developed. The $\lambda_{\mathrm{obliv}}$ language
+\cite{darais2019language} uses a type system is used to automatically
+enforce so-called probabilistic trace obliviousness.  But similar to
+previous work on oblivious data structures \cite{10.1145/3498713},
+obliviousness is related to pure noninterference, not the relaxed form
+related to passive MPC security. Similarly, the Haskell-based security type system in
+\cite{6266151} enforces a version of noninterence that is
+sound for passive security, but does not verify the correctness of declassifications and does not consider malicious security. And properties of real/ideal
+passive and malicious security for a probabilistic language have been
+formulated in EasyCrypt \cite{8429300}-- though their proof methods,
+while mechanized, are fully manual, and their formulation of malicious
+security is not as clearly related to robust declassification as is the
+one we present in Section \ref{section-hyper}.
+
+Program logics for probabilistic languages and specifically reasoning
+about properties such as joint probabilistic independence is also
+important related work. Probabilistic Separation Logic (PSL)
+\cite{barthe2019probabilistic} develops a logical framework for
+reasoning about probabilistic independence (aka separation) in
+programs, and they consider several (hyper)properties, such as perfect
+secrecy of one-time-pads and indistinguishability in secret sharing,
+that are critical to MPC. However, their methods are manual, and
+don't include conditional independence (separation). This
+latter issue has been addressed in Lilac \cite{li2023lilac}. The
+application of Lilac-style reasoning to MPC protocols has not
+previously been explored, as we do in Section
+\ref{section-example-gmw}.
+
+Our work also shares many ideas with probabilistic programming
+languages designed to perform (exact or approximate) statistical
+inference~\cite{holtzen2020scaling, carpenter2017stan, wood2014new,
+  bingham2019pyro, albarghouthi2017fairsquare, de2007problog,
+  pfeffer2009figaro, saad2021sppl}. Our setting, however, requires
+verifying properties beyond inference, including conditional
+statistical independence. Recent work by Li et al.~\cite{li2023lilac} proposes a
+manual approach for proving such properties, but does not provide
+automation.

### 2024-05-21 ppdp24/example-gmw.tex
checkpoint

@@
 will assume the 2-party version, though it generalizes to $n$
-parties\cite{XXX}. GMW uses a common technique in MPC, which is to
+parties\cite{goldreich2019play}. GMW uses a common technique in MPC, which is to
 represent values $v$ as distributed shares $v_1$ and $v_2$ with $v =
@@
   $$ (\varnothing,e) \redxs (\prog,\decodegmw(E[\mesg{z}])) $$
-  Then both of the following conditions hold for all $H$ and $C$ where $\iov(\prog) = S \cup M$:
+  Then all of the following conditions hold for all $H$ and $C$ where $\iov(\prog) = S \cup M$:
   \begin{enumerate}
   \item $\conddetx{\progtt(\prog)}{S}{\{\macgv{\mesg{z}}\}}$
-    \item $\condsep{\progtt(\prog)}{S}{\{\macgv{\mesg{z}}\}}{(M_C \cup \{ \elab{\mesg{z}}{1}, \elab{\mesg{z}}{2} \})}$
+  \item $\condunix{\progtt(\prog)}{S}{M_C}$
+  \item $\condsep{\progtt(\prog)}{S}{\{\macgv{\mesg{z}}\}}{\{ \elab{\mesg{z}}{1}, \elab{\mesg{z}}{2} \})}$
   \end{enumerate}
@@
   $$
-  Then both of the following conditions hold for both $\cid \in \{ 1,2 \}$ where $\iov(\prog) = M$:
+  Then all of the following conditions hold for both $\cid \in \{ 1,2 \}$ where $\iov(\prog) = M$:
   \begin{enumerate}
@@
     $\conddetx{\progtt(\prog)}{\{ \macgv{\mesg{x}},\macgv{\mesg{y}} \}}{\{ \macgv{\mesg{z}} \}}$
+  \item $\condunix{\progtt(\prog)}{\{ \macgv{\mesg{x}},\macgv{\mesg{y}} \}}{\{ \elab{\mesg{z}}{\cid} \}}$
   \item $\condsep
@@
     {\{ \macgv{\mesg{z}} \}}
-      {M_{\{\cid\}} \cup \{ \elab{\mesg{z}}{1},\elab{\mesg{z}}{2} \}}$
+    {\{ \elab{\mesg{z}}{1},\elab{\mesg{z}}{2} \}}$
   \end{enumerate}
@@
   \label{lemma-conditioning}
-  Each of the following hold:
+  Give $\pmf$, each of the following hold for $S,V_1,V_2,V_3 \in \dom(\pmf)$:
   \begin{enumerate}
-    \item Given $\condp{\pmf}{X_1}{\detx{X_2}}$ and
-      $\condp{\pmf}{X_2}{\detx{X_3}}$, then $\condp{\pmf}{X_1}{\detx{X_3}}$.
-    \item Given $\condp{\pmf}{X_1}{\detx{X_2}}$ and
-      $\condsep{\pmf}{X_2}{X_3}{X_4}$, then $\condsep{\pmf}{X_1}{X_3}{X_4}$.
-    \item Given $\condsep{\pmf}{X_1}{X_2}{X_3}$ and $\condp{\pmf}{X_1}{\detx{X_2}}$
-      and $\condp{\pmf}{X_2}{\detx{X_4}}$, then $\condsep{\pmf}{X_1}{X_4}{X_3}$.
-    \item Given $\condsep{\pmf}{X_1 \cup X_2}{X_3}{X_4}$ and $\condp{\pmf}{X_1 \cup X_2}{\detx{X_3}}$,
-      then $\condsep{\pmf}{X_1}{X_2 \cup X_3}{X_4}$.
+    \item Given $\condp{\pmf}{S}{\detx{V_1}}$ and
+      $\condp{\pmf}{V_2}{\detx{V_2}}$, then $\condp{\pmf}{S}{\detx{V_2}}$.
+    \item Given $\condp{\pmf}{S}{\detx{V_1}}$ and
+      $\condp{\pmf}{V_1}{\unix{V_2}}$, then $\condp{\pmf}{S}{\unix{V_2}}$.
+    \item Given $\condp{\pmf}{S}{\detx{V_1}}$ and
+      $\condsep{\pmf}{V_1}{V_2}{V_3}$, then $\condsep{\pmf}{S}{V_2}{V_3}$.
+    %\item Given $\condsep{\pmf}{X_1}{X_2}{X_3}$ and $\condp{\pmf}{X_1}{\detx{X_2}}$
+    %  and $\condp{\pmf}{X_2}{\detx{X_4}}$, then $\condsep{\pmf}{X_1}{X_4}{X_3}$.
+    %\item Given $\condsep{\pmf}{X_1 \cup X_2}{X_3}{X_4}$ and $\condp{\pmf}{X_1 \cup X_2}{\detx{X_3}}$,
+    %  then $\condsep{\pmf}{X_1}{X_2 \cup X_3}{X_4}$.
   \end{enumerate}
@@
   $$
-  The induction hypothesis, together with the assumed uniquenenss of $z$, gives:
-  \begin{mathpar}
-  \condsep{\progtt(\prog)}{S^1}{\{ \macgv{\mesg{x}} \}}{(M^1_C \cup \{ \elab{\mesg{x}}{1}, \elab{\mesg{x}}{2} \})}
 
-  \condsep{\progtt(\prog)}{S^2}{\{ \macgv{\mesg{y}} \}}{(M^2_C \cup \{ \elab{\mesg{y}}{1}, \elab{\mesg{y}}{2} \})}
+  \begin{enumerate}[\hspace{5mm}(x.1)]
+  \item $\conddetx{\progtt(\prog)}{S^1}{\{\macgv{\mesg{x}}\}}$
+  \item $\condunix{\progtt(\prog)}{S^1}{M^1_C}$
+  %\item $\condsep{\progtt(\prog)}{S^1}{\{\macgv{\mesg{x}}\}}{\{ \elab{\mesg{x}}{1}, \elab{\mesg{x}}{2} \})}$
+  \end{enumerate}
 
-  \conddetx{\progtt(\prog)}{S^1}{\{ \macgv{\mesg{x}} \}}
+  \begin{enumerate}[\hspace{5mm}(y.1)]
+  \item $\conddetx{\progtt(\prog)}{S^2}{\{\macgv{\mesg{y}}\}}$
+  \item $\condunix{\progtt(\prog)}{S^2}{M^2_C}$
+  %\item $\condsep{\progtt(\prog)}{S^2}{\{\macgv{\mesg{y}}\}}{\{ \elab{\mesg{y}}{1}, \elab{\mesg{y}}{2} \})}$
+  \end{enumerate}
 
-  \conddetx{\progtt(\prog)}{S^2}{\{ \macgv{\mesg{y}} \}}    
-  \end{mathpar}
-  This, together with Lemma \ref{lemma-gmwtactic} (1) and Lemma \ref{lemma-conditioning} (1)
-  give:
-  $$
-  \conddetx{\progtt(\prog)}{S^1 \cup S^2}{\{ \macgv{\mesg{z}} \}}
-  $$
-  and Lemma  \ref{lemma-gmwtactic} (2) and Lemma \ref{lemma-conditioning} (2-3) gives:
-  $$
-  \begin{array}{c}
-    \progtt(\prog)|S^1 \cup S^2) \vdash \\
-    {\{ \macgv{\mesg{z}} \}} * {(M^1_C \cup M^2_C \cup \{ \elab{\mesg{x}}{2}, \elab{\mesg{y}}{2}, \elab{\mesg{z}}{1}, \elab{\mesg{z}}{2} \})}
-  \end{array}
-  $$
+  \begin{enumerate}[\hspace{5mm}(z.1)]
+  \item
+    $\conddetx{\progtt(\prog)}{\{ \macgv{\mesg{x}},\macgv{\mesg{y}} \}}{\{ \macgv{\mesg{z}} \}}$
+  \item $\condunix{\progtt(\prog)}{\{ \macgv{\mesg{x}},\macgv{\mesg{y}} \}}{\{ \elab{\mesg{z}}{\cid} \}}$
+  \item $\condsep
+    {\progtt(\prog)}
+    {\{ \macgv{\mesg{x}},\macgv{\mesg{y}} \}}
+    {\{ \macgv{\mesg{z}} \}}
+    {\{ \elab{\mesg{z}}{1},\elab{\mesg{z}}{2} \}}$
+  \end{enumerate}
+
+  Given $x.1$ and $y.1$ and assumed uniqueness of random variables
+  used in each gate we have $\conddetx{\progtt(\prog)}{S}{\{
+    \macgv{\mesg{x}},\macgv{\mesg{y}}\}}$, so it follows from $z.1$
+  and Lemma \ref{lemma-conditioning} (1) that
+  $\conddetx{\progtt(\prog)}{S}{\{\macgv{\mesg{z}}\}}$, and also by
+  $z.3$ and Lemma \ref{lemma-conditioning} (3):
+  $$\condsep
+  {\progtt(\prog)}
+  {S}
+  {\{ \macgv{\mesg{z}} \}}
+  {\{ \elab{\mesg{z}}{1},\elab{\mesg{z}}{2} \}}$$
 \end{proof}

### 2024-05-21 ppdp24/example-gmw.tex
checkpoint

@@
     %  and $\condp{\pmf}{X_2}{\detx{X_4}}$, then $\condsep{\pmf}{X_1}{X_4}{X_3}$.
-    %\item Given $\condsep{\pmf}{X_1 \cup X_2}{X_3}{X_4}$ and $\condp{\pmf}{X_1 \cup X_2}{\detx{X_3}}$,
+    %\item Given $\condsep{\pmf}{S_1 \cup S_2}{V_3}{X_4}$ and $\condp{\pmf}{X_1 \cup X_2}{\detx{X_3}}$,
     %  then $\condsep{\pmf}{X_1}{X_2 \cup X_3}{X_4}$.
@@
   $$
-
+  Letting $\iov(\prog) = S \cup M$, by definition and the induction hypothesis there
+  exist $S^1,S^2 \subseteq S$ and $M^1,M^2 \subseteq M$
+  such that:
   \begin{enumerate}[\hspace{5mm}(x.1)]
@@
   \item $\condunix{\progtt(\prog)}{S^1}{M^1_C}$
-  %\item $\condsep{\progtt(\prog)}{S^1}{\{\macgv{\mesg{x}}\}}{\{ \elab{\mesg{x}}{1}, \elab{\mesg{x}}{2} \})}$
+  \item $\condsep{\progtt(\prog)}{S^1}{\{\macgv{\mesg{x}}\}}{\{ \elab{\mesg{x}}{1}, \elab{\mesg{x}}{2} \})}$
   \end{enumerate}
-
+  and also:
   \begin{enumerate}[\hspace{5mm}(y.1)]
@@
   \item $\condunix{\progtt(\prog)}{S^2}{M^2_C}$
-  %\item $\condsep{\progtt(\prog)}{S^2}{\{\macgv{\mesg{y}}\}}{\{ \elab{\mesg{y}}{1}, \elab{\mesg{y}}{2} \})}$
+  \item $\condsep{\progtt(\prog)}{S^2}{\{\macgv{\mesg{y}}\}}{\{ \elab{\mesg{y}}{1}, \elab{\mesg{y}}{2} \})}$
   \end{enumerate}
-
+  and by Lemma \ref{lemma-gmwtactic} we have:
   \begin{enumerate}[\hspace{5mm}(z.1)]
@@
   \end{enumerate}
-
   Given $x.1$ and $y.1$ and assumed uniqueness of random variables
@@
   {\{ \elab{\mesg{z}}{1},\elab{\mesg{z}}{2} \}}$$
+  Additionally Lemma \ref{lemma-conditioning} (2) gives $\condunix{\progtt(\prog)}{S}{\{ \elab{\mesg{z}}{\cid} \}}$,
+  and $x.2$ and $y.2$ give $\condunix{\progtt(\prog)}{S}{(M^1 \cup M^2)_C}$. Facts $x.3$ and $y.3$ together with
+  uniqueness of gate identifiers implies:
+  $$
+  \condunix{\progtt(\prog)}{S}{(M^1 \cup M^2)_C \cup \{  \elab{\mesg{z}}{\cid} \}}
+  $$
+  This implies the result.
 \end{proof}
@@
     \conddetx{\progtt(\prog)}{S}{\{\macgv{\mesg{z}}\}}
-
-    \condsep{\progtt(\prog)}{S}{\{\macgv{\mesg{z}}\}}{(M_C \cup \{ \elab{\mesg{z}}{1}, \elab{\mesg{z}}{2} \})}
+    \condunix{\progtt(\prog)}{S}{M_C}
+    \condsep{\progtt(\prog)}{S}{\{\macgv{\mesg{z}}\}}{\{ \elab{\mesg{z}}{1}, \elab{\mesg{z}}{2} \})}
   \end{mathpar}
@@
   \begin{mathpar} 
-    \conddetx{\progtt(\prog)}{S_H \cup S_C}{O}
+    \condunix{\progtt(\prog)}{S}{M_C}
 
-    \condsep{\progtt(\prog)}{S_H \cup S_C}{O}{(M_C \cup P)}
+    \condsep{\progtt(\prog)}{S}{O}{P}
   \end{mathpar}
-  and thus by Lemma \ref{lemma-conditioning} (4) we have:
-  $$
-  \condsep{\progtt(\prog)}{S_C}{(O \cup S_H)}{(M_C \cup P)}
-  $$
   which implies the result.

### 2024-05-21 ppdp24/example-gmw.tex
checkpoint

@@
 security. Following \cite{XXX}, we define a conditional separation
-relation $\condsep{\pmf}{X_1}{X_2}{X_3}$ to mean that
-under the condition of all value assignments for
-$X_2$ and $X_3$ are separable under $\pmf$ on the condition
-of any value assignment of $X_1$-- i.e., conditionally
-on any $\store \in \mems(X_1)$. 
-\begin{definition}[Conditional Separation]
-  We write $\condsep{\pmf}{X_1}{X_2}{X_3}$ iff for all
-  $\store' \in \mems(X_1)$ and $\store \in \mems(X_2 \cup X_3)$
-  and letting $X = X_1 \cup X_2 \cup X_3$ we have:
-  $$\condd{\pmf}{X}{\store'}(\store) =
-  \condd{\pmf}{X}{\store'}(\store_{X_2}) *
-  \condd{\pmf}{X}{\store'}(\store_{X_3})$$
-\end{definition}
-Another key concept needed especially for reasoning about
-circuits is conditional determinism. For example, if
-$\macgv{z}$ is an output of an internal gate, it will
-definitely be computed using random variables, however,
-it \emph{should} be determistic under any set of input
-secrets $S$, since we assume that $\idealf$ is
-deterministic.
-\begin{definition}[Conditional Determinism]  
-  We write $\conddetx{\pmf}{X_1}{X_2}$ iff for all
+relation $\condsep{\pmf}{X_1}{X_2}{X_3}$ to mean that under the
+condition of all value assignments for $X_2$ and $X_3$ are separable
+under $\pmf$ on the condition of any value assignment of $X_1$-- i.e.,
+conditionally on any $\store \in \mems(X_1)$. Another key concept
+needed especially for reasoning about circuits is conditional
+determinism. For example, if $\macgv{z}$ is an output of an internal
+gate, it will definitely be computed using random variables, however,
+it \emph{should} be determistic under any set of input secrets $S$,
+since we assume that $\idealf$ is deterministic. Conditional
+uniformity is also important, since the gradual release property of
+many protocols means that messages appear in a uniform distribution to
+the adversary.
+\begin{definition}[Conditioning Properties] 
+  Give $\pmf$ and $X_1,X_1,X_3 \subseteq \dom{pmf}$, 
+  we write:
+  \begin{itemize}
+  \item $\condsep{\pmf}{X_1}{X_2}{X_3}$ iff for all
+    $\store' \in \mems(X_1)$ and $\store \in \mems(X_2 \cup X_3)$ we have
+    $\pmf(\store|\store') = \pmf(\store_{X_2}|\store') *  \pmf(\store_{X_3}|\store')$.
+  \item $\conddetx{\pmf}{X_1}{X_2}$ iff for all
     $\store' \in \mems(X_1)$ there exists 
-  $\store \in \mems(X_2)$ such that
-  $\condd{\pmf}{X_1 \cup X_2}{\store'}(\store) = 1$.
+    $\store \in \mems(X_2)$ such that $\pmf(\store|\store') = 1$.
+  \item $\condunix{\pmf}{X_1}{X_2}$ iff for all
+    $\store' \in \mems(X_1)$ and
+    $\store \in \mems(X_2)$ we have
+    $\pmf(\store|\store') = 1/p^{|X_2|}$.
+  \end{itemize}
 \end{definition}
@@
   \item $\condunix{\progtt(\prog)}{S^1}{M^1_C}$
-  \item $\condsep{\progtt(\prog)}{S^1}{\{\macgv{\mesg{x}}\}}{\{ \elab{\mesg{x}}{1}, \elab{\mesg{x}}{2} \})}$
+  %\item $\condsep{\progtt(\prog)}{S^1}{\{\macgv{\mesg{x}}\}}{\{ \elab{\mesg{x}}{1}, \elab{\mesg{x}}{2} \})}$
   \end{enumerate}
@@
   \item $\condunix{\progtt(\prog)}{S^2}{M^2_C}$
-  \item $\condsep{\progtt(\prog)}{S^2}{\{\macgv{\mesg{y}}\}}{\{ \elab{\mesg{y}}{1}, \elab{\mesg{y}}{2} \})}$
+  %\item $\condsep{\progtt(\prog)}{S^2}{\{\macgv{\mesg{y}}\}}{\{ \elab{\mesg{y}}{1}, \elab{\mesg{y}}{2} \})}$
   \end{enumerate}
@@
   Additionally Lemma \ref{lemma-conditioning} (2) gives $\condunix{\progtt(\prog)}{S}{\{ \elab{\mesg{z}}{\cid} \}}$,
-  and $x.2$ and $y.2$ give $\condunix{\progtt(\prog)}{S}{(M^1 \cup M^2)_C}$. Facts $x.3$ and $y.3$ together with
+  and $x.2$ and $y.2$ give $\condunix{\progtt(\prog)}{S}{(M^1 \cup M^2)_C}$, $z.3$ together with
   uniqueness of gate identifiers implies:

### 2024-05-22 ppdp24/bruteforce.tex
checkpoint
tiny tweaks and a note

@@
 
-In $\mathbb{F}_2$ a brute force strategy for computing
+In the binary field $\mathbb{F}_2$, a brute force strategy for computing
 $\progtt(\prog)$ for any $\prog$ is to directly compute
@@
 hyperproperties discussed previously, or other properties such as
-perfect secrecy \cite{XXX}.  A basic method to do this is
+perfect secrecy \cite{barthe2019probabilistic}.  A basic method to do this is
 to calculate a truth table for the given protocol. Since
@@
     {tt}\ \ \stores\ (\xassign{x}{\be}{\cid}) &\defeq& \begin{array}{l}
-      \mathrm{let}\ \stores' = \solve{\stores}{\be} \ \mathrm{in}\\
-      \ \ \{\extend{\store}{\view{\cid}{w}}{1} \mid \store \in \stores' \}\ \cup\\
-      \ \ \{\extend{\store}{\view{\cid}{w}}{0} \mid \store \in \stores - \stores' \}\end{array}%\\[2mm]
+      \mathrm{let}\ \stores' = \solve{\stores}{\be}{\cid}\ \mathrm{in}\\
+      \ \ \{\extend{\store}{x}{1} \mid \store \in \stores' \}\ \cup\\
+      \ \ \{\extend{\store}{x}{0} \mid \store \in \stores - \stores' \}\end{array}%\\[2mm]
     %\cruns(\prog) &\defeq& \mathit{foldl}\ {tt}\ \mems(S \cup R)\ \prog

### 2024-05-22 ppdp24/example-bdoz.tex
checkpoint

@@
 information-theoretic secure MAC codes to shares is a fundamental
-approach realized by protocols such as BDOZ \cite{XXX} and SPDZ
-\cite{XXX}.  These protocols assume a pre-processing phase that
-distributes shares with MAC codes to clients.  This integrates well
-with pre-processed Beaver Triples to implement malicious secure, and
-relatively efficient, multiplication \cite{XXX}. Recall
-that Beaver triples are values $a,b,c$ with $\fcod{a * b} = c$,
-unique per multiplication gate, that are secret shared with clients
-during pre-processing. Here we consider the 2-party version.
+approach realized by protocols such as BDOZ and SPDZ
+\cite{SPDZ1,SPDZ2,BDOZ,10.1007/978-3-030-68869-1_3}.  These protocols
+assume a pre-processing phase that distributes shares with MAC codes
+to clients.  This integrates well with pre-processed Beaver Triples to
+implement malicious secure, and relatively efficient, multiplication
+\cite{evans2018pragmatic}. Recall that Beaver triples are values $a,b,c$ with
+$\fcod{a * b} = c$, unique per multiplication gate, that are secret
+shared with clients during pre-processing. Here we consider the
+2-party version.
 
@@
 $\cid'$. This is a semi-homomorphic encryption scheme that supports
-addition of shares and multiplication of shares by a constant
-\cite{XXX}. For more details the reader is referred to \cite{XXX}. We
-note that while we restrict values $v$, $m$, and $k$ to the same field
-in this presentation for simplicity, in general $m$ and $k$ can be in
-extensions of $\mathbb{Z}_p$.
+addition of shares and multiplication of shares by a constant-- for
+more details the reader is referred to Orsini
+\cite{10.1007/978-3-030-68869-1_3}. We note that while we restrict
+values $v$, $m$, and $k$ to the same field in this presentation for
+simplicity, in general $m$ and $k$ can be in extensions of
+$\mathbb{Z}_p$.
 
@@
 $\ttt{macctimes}$ for multiplication of a share and a constant
-in the BDOZ encryption scheme \cite{XXX}. The $\ttt{auth}$
+in the BDOZ encryption scheme \cite{10.1007/978-3-030-68869-1_3}. The $\ttt{auth}$
 function implements the MAC check as an $\assert$.
@@
 parties interact in secure opening of $x \fplus a$ and $y + b$,
-followed by the non-interactive computation of shares of $x * y$
-for secure reveals as per standard protocol
-\cite{XXX}. The non-interactive reconstruction of the local authentication
-key for both the secure openings and the final reveal is enabled by the
-semi-homorphic properties of the BDOZ scheme.
+followed by the non-interactive computation of shares of $x * y$ for
+secure reveals as per standard protocol
+\cite{10.1007/978-3-030-68869-1_3}. The non-interactive reconstruction
+of the local authentication key for both the secure openings and the
+final reveal is enabled by the semi-homorphic properties of the BDOZ
+scheme.
 
@@
 \begin{lemma}
+  \label{lemma-cheating}
   If cheating is detected in $\prog$, then $\prog$ has integrity.
@@
 In the case of BDOZ, cheating detection is accomplished by the information-theoretic
-security of the encryption scheme\cite{XXX}. Furthermore, the symmetry of
+security of the encryption scheme \cite{evans2018pragmatic}. Furthermore, the symmetry of
 the protocol in Figure \ref{fig-beaver} ensures that both parties will authenticate

### 2024-05-22 ppdp24/example-gmw.tex
checkpoint

@@
 \paragraph{Oblivious Transfer} A passive secure oblivious transfer (OT) protocol
-based on previous work \cite{XXX} can be defined in $\metaprot$,
+based on previous work \cite{barthe2019probabilistic} can be defined in $\metaprot$,
 however this protocol assumes some shared randomness. Alternatively,
@@
 
-As discussed above and in related work \cite{XXX}, probabilistic
+As discussed above and in related work \cite{8429300}, probabilistic
 separation conditional on certain variables-- e.g., secret inputs or
 public outputs-- is a key mechanism for reasoning about MPC protocol
-security. Following \cite{XXX}, we define a conditional separation
+security. Following \cite{barthe2019probabilistic}, we define a conditional separation
 relation $\condsep{\pmf}{X_1}{X_2}{X_3}$ to mean that under the

### 2024-05-22 ppdp24/metalang.tex
checkpoint

@@
 
-The $\fedprot$ language is low-level and does not include abstractions for
-defining composable elements. So in this Section we introduce the $\metaprot$
-language which that includes structured data and function definitions
-for defining composable protocol elements at a higher level of abstraction. 
-The $\metaprot$ language is a \emph{metalanguage}, where 
-$\fedprot$ protocols are the residuum of computation. In addition to these
-declarative benefits of $\metaprot$, component
-definitions support compositional verification of larger programs as we will
-discuss with examples in Section
-\ref{section-examples}.
+The $\fedprot$ language is low-level and does not include abstractions
+for defining composable elements. So in this Section we introduce the
+$\metaprot$ language which that includes structured data and function
+definitions for defining composable protocol elements at a higher
+level of abstraction.  The $\metaprot$ language is a
+\emph{metalanguage}, where $\fedprot$ protocols are the residuum of
+computation. In addition to these declarative benefits of $\metaprot$,
+component definitions support compositional verification of larger
+prortocols as we will discuss with examples in Sections
+\ref{section-example-gmw} and \ref{section-example-bdoz}.
 

### 2024-05-22 ppdp24/conclusion.tex
checkpoint

\section{Conclusion and Future Work}

The $\metaprot/\minifed$ language model developed in this paper, along
with the suite of hyperproperties explored, are a foundation
for defining low-level MPC protocols and exploring methods for security
verification.  There are several promising avenues for future
work. Concurrency will be important to capture common MPC idioms such
as commitment and circuit optimizations, and to consider the UC
security model \cite{evans2018pragmatic,viaduct-UC}. To support better
verification, integration of our tactics with a proof assistant is a
compelling direction that resonate with previous work \cite{8429300}.

Better automated verification is a main goal for future work.
Security types have been effectively leveraged for high-level
programs, for both passive and malicious security. This suggests that
they can be similarly applied at low levels in a way that is
complementary to high-level systems. Generalizing our automated
analysis in $\mathbb{F}_2$ to larger arithmetic fields is also a
compelling technical challenge with potential practical benefits,
e.g., application to arithmetic circuits.

### 2024-05-22 ppdp24/intro.tex
checkpoint

@@
 
-Our work provides a complete methodology for end-to-end verification
+In summary, our work provides a complete methodology for end-to-end verification
 of MPC protocols via three components:
@@
 \item A \textbf{low-level language} for defining MPC primitives
-  (\S\ref{section-lang}).
+  (Section \ref{section-lang}) with an associated metalanguage
+  to ease programming (Section \ref{section-metalang}).
 \item A \textbf{fully-automated verification method} for low-level MPC
-  primitives in $\mathbb{F}_2$ (\S\ref{section-bruteforce}).
+  primitives in $\mathbb{F}_2$ (Section \ref{section-bruteforce}).
 \item A \textbf{partially-automated verification method} for
-  high-level MPC protocols, which leverages automated proofs for
-  low-level primitives (\S\ref{section-example-gmw}).
+  MPC protocols, which leverages automated proofs for
+  low-level primitives (Section \ref{section-example-gmw}).
 \end{enumerate}

### 2024-05-22 ppdp24/hyper.tex
checkpoint

@@
 probabilistic noninterence conditioned on output that is sound
-for passive security. It says that if two low-equivalent secret
-inputs generate the same output, then the distributions of corrupt
-views are the same. 
+for passive security. 
 \begin{definition}[Noninterference modulo output]
@@
 %\end{definition}
-Intuitively, this conditional noninterference property implies that
+This conditional noninterference property implies that
 corrupt views give the adversary no better chance of guessing honest

### 2024-05-22 ppdp24/bruteforce.tex
checkpoint

@@
 In Figure \ref{fig-solve} we define the algorithm
-$\solve{\stores}{\be}{\cid}$ which filters a given $\stores$ to obtain the
-subset whose elements satisfy $\be$. In this figure and elsewhere for
-$\mathbb{F}_2$ we use standard names of logical connectives
-for field operations ($\eand$ and $\exor$ for $*$ and $+$ respectively) and
-add $\enot$ and $\eor$ as trivial but convenient extensions. The correctness
-of this operation is characterized as follows.
+$\solve{\stores}{\be}{\cid}$ which filters a given $\stores$ to obtain
+the subset whose elements satisfy $\be$. In this Figure and elsewhere
+we use logical connectives as field operations ($\eand$ and $\exor$
+for $*$ and $+$ respectively) in $\mathbb{F}_2$ and add $\enot$ and
+$\eor$ as trivial but convenient extensions. The correctness of this
+operation is characterized as follows.
 \begin{lemma}
@@
 \end{mathpar}
-... and so on. Then, to begin converting a protocol $\prot$ to a Datalog
-we define the function ${tt}$ that applies $\solve{\!}{\!}{\!\!}$ locally to each command
+... and so on. Then, to convert a protocol $\prot$ to a Datalog
+we first define the function $\bodies$ that applies $\solvealg$ locally to each command
 in $\prog$, obtaining the subset of memories that result

### 2024-05-22 ppdp24/example-gmw.tex
checkpoint

@@
 will assume the 2-party version, though it generalizes to $n$
-parties\cite{goldreich2019play}. GMW uses a common technique in MPC, which is to
+parties \cite{goldreich2019play}. GMW uses a common technique in MPC, which is to
 represent values $v$ as distributed shares $v_1$ and $v_2$ with $v =
@@
 functions, and a decoding function. Note that $\exor$ requires
-no interaction between parties, while $\eand$ necessitates and
+no interaction between parties, while conjunction necessitates
 1-of-4 oblivious transfer. The gate computation is
@@
 public outputs-- is a key mechanism for reasoning about MPC protocol
-security. Following \cite{barthe2019probabilistic}, we define a conditional separation
-relation $\condsep{\pmf}{X_1}{X_2}{X_3}$ to mean that under the
-condition of all value assignments for $X_2$ and $X_3$ are separable
-under $\pmf$ on the condition of any value assignment of $X_1$-- i.e.,
-conditionally on any $\store \in \mems(X_1)$. Another key concept
-needed especially for reasoning about circuits is conditional
-determinism. For example, if $\macgv{z}$ is an output of an internal
-gate, it will definitely be computed using random variables, however,
-it \emph{should} be determistic under any set of input secrets $S$,
-since we assume that $\idealf$ is deterministic. Conditional
-uniformity is also important, since the gradual release property of
-many protocols means that messages appear in a uniform distribution to
-the adversary.
+security. Following \cite{barthe2019probabilistic}, we define a
+conditional separation relation $\condsep{\pmf}{X_1}{X_2}{X_3}$ to
+mean that $X_2$ and $X_3$ are indpendent in $\pmf$ conditionally on
+any assignment of values to $X_1$-- i.e., conditionally on any $\store
+\in \mems(X_1)$. Another key concept needed especially for reasoning
+about circuits is conditional determinism. For example, if $\macgv{z}$
+is an output of an internal gate, it will definitely be computed using
+random variables, however, it \emph{should} be determistic under any
+set of input secrets $S$, since we assume that $\idealf$ is
+deterministic. Conditional uniformity is also important, since the
+gradual release property of many protocols means that messages appear
+in a uniform distribution to the adversary.
 \begin{definition}[Conditioning Properties] 

### 2024-05-22 ppdp24/example-gmw.tex
checkpoint

@@
 
-Given these definitions, we can formulate an invariant
-for circuit computation with respect to internal gates
-as follows. It says that the output of any gate is
-deterministic wrt inputs $S$, and conditionally
-on $S$ the output $\macgv{\mesg{z}}$ remains
+Given these definitions, we can formulate an invariant for circuit
+computation with respect to internal gates as follows. It says that
+the output of any gate is deterministic given inputs $S$, and
+conditionally on $S$ corrupt views are always in a uniform random
+disrtibution (pure noise), while output $\macgv{\mesg{z}}$ remains
 separable from corrupt views and both shares of
-$\macgv{\mesg{z}}$. This last nuance is critical,
-since those shares will in fact be revealed if
-$\macgv{\mesg{z}}$ is decoded as the circuit output. 
+$\macgv{\mesg{z}}$. This last condition (incidentally missing from PSL
+due to the reliance on more recent innovations in conditioning logic
+\cite{li2023lilac}) is critical since those shares will in fact be
+revealed if $\macgv{\mesg{z}}$ is decoded as the circuit output.
 \begin{lemma}[GMW Invariant]
@@
 \end{lemma}
-To prove this, we can formulate and automatically prove gate-level
-versions of the invariant. This serves as a proof tactic
-that simplifies the proof of the GMW invariant. It establishes
-that the $\eand$ gate output is deterministic conditional on
-the inputs, and is separable from the output shares and
-either parties' received messages. 
+To prove this, we can formulate and automatically prove local,
+gate-level versions of the invariant. This serves as a proof tactic
+that simplifies the proof of the GMW invariant. 
 \begin{lemma}[And Gate Tactic]
@@
 the global reasoning of Lemma \ref{lemma-gmwinvariant}, we can demonstrate
-the following.
+the following properties of conditioning-based reasoning. While a frame rule
+was developed in \cite{li2023lilac}, these properties are distinct and
+useful for reasoning about MPC. We omit proofs for brevity. 
 \begin{lemma}
@@
 Then we can put the pieces together to prove the invariant, using automated tactics
-for gate-level reasoning.  
+for gate-level reasoning. We sketch some elements of this proof to focus on
+novel aspects of our technique. 
 \begin{proof}[Proof of Lemma \ref{lemma-gmwinvariant}]
@@
 \end{proof}
-
+The preceding Lemma, together with some additional observations about decoding, establish correctness
+of arbitrary circuits. 
 \begin{theorem}

### 2024-05-22 ppdp24/example-bdoz.tex
checkpoint

@@
 information-theoretic secure MAC codes to shares is a fundamental
-approach realized by protocols such as BDOZ and SPDZ
+technique pioneered in systems such as BDOZ and SPDZ
 \cite{SPDZ1,SPDZ2,BDOZ,10.1007/978-3-030-68869-1_3}.  These protocols
@@
 Furthermore, client 2's authentication of the sum of shares with the
-sum of their keys detects any attempted cheating by $1$.
-Finally, the $\ttt{secreveal}$ function
-is very similar to $\decodegmw$, except with the addition
-of authentication of revealed shares to ensure malicious security.
+sum of their keys detects any attempted cheating by $1$.  Finally, the
+$\ttt{secreveal}$ function is the same as $\decodegmw$ but with the
+addition of authentication of revealed shares to ensure malicious
+security.
 
@@
 
-We can carry out similar proofs of passive security for the protocol in
-Figure \ref{fig-beaver} as for GMW, even using automated tactics for
-the protocol in $\mathbb{F}_2$. But in the case of BDOZ we are also
-concerned with malicious security. To demonstrate this, we can
-demonstrate that the protocol satisfies integrity in the sense of
-Definition \ref{def-integrity}. To do so, we observe that it satisfies
-a stronger property, that we call cheating detection. Intuitively,
-integrity says that the only thing that the adversary can do in the
-malicious model is to elicit the same responses from honest parties
-that an honest run of the protocol would elicit. Cheating detection
-says that the adversary can only execute the protocol honestly, or
-or else gets caught (and abort).
+We can carry out similar proofs of passive security for the protocol
+in Figure \ref{fig-beaver} as for GMW, even using automated tactics
+for the protocol in $\mathbb{F}_2$. But in the case of BDOZ we are
+also concerned with malicious security. To demonstrate this, we can
+show that the protocol satisfies integrity in the sense of Definition
+\ref{def-integrity}. To do so, we observe that it satisfies a stronger
+property, that we call cheating detection. Cheating detection says
+that the adversary can only execute the protocol honestly, or or else
+gets caught (and abort). Note in particular that this is \emph{not}
+a form of endorsement (the dual of declassificatio), which allows
+the adversary some leeway. Rather, it is a confirmation of high integrity.
 
-Focusing in, we identify the adversarial inputs as the messages
-sent from the adversary to honest parties, on which honest responses
+Focusing in, we identify the \emph{adversarial inputs} as the messages
+sent from the adversary to honest parties on which honest responses
 to the adversary may depend. We want to say that these are the messages
@@
 Now, we can characterize protocols with cheating detection as those where
-adversarial inputs to honest reponses must themselves be constructed honestly. 
+the adversary can only behave honestly/passively, or else cause abort.
 \begin{definition}[Cheating Detection]

### 2024-05-22 ppdp24/intro.tex
checkpoint

@@
 obliviousness is related to pure noninterference, not the relaxed form
-related to passive MPC security. Similarly, the Haskell-based security type system in
-\cite{6266151} enforces a version of noninterence that is
-sound for passive security, but does not verify the correctness of declassifications and does not consider malicious security. And properties of real/ideal
-passive and malicious security for a probabilistic language have been
-formulated in EasyCrypt \cite{8429300}-- though their proof methods,
-while mechanized, are fully manual, and their formulation of malicious
-security is not as clearly related to robust declassification as is the
-one we present in Section \ref{section-hyper}.
+related to passive MPC security. The Haskell-based security type
+system in \cite{6266151} enforces a version of noninterence that is
+sound for passive security, but does not verify the correctness of
+declassifications and does not consider malicious security. And
+properties of real/ideal passive and malicious security for a
+probabilistic language have been formulated in EasyCrypt
+\cite{8429300}-- though their proof methods, while mechanized, are
+fully manual, and their formulation of malicious security is not as
+clearly related to robust declassification as is the one we present in
+Section \ref{section-hyper}.
 

### 2024-05-22 ppdp24/example-bdoz.tex
checkpoint

@@
 as a pre-processing predicate that establishes conditions for initial
-memories (see Definition \ref{def-progtt}).  Here we assume two input
+memories (see Definition \ref{def-aprogtt}).  Here we assume two input
 secrets $\elab{\secret{x}}{1}$ and $\elab{\secret{y}}{2}$ and a single

### 2024-05-22 ppdp24/model.tex
checkpoint

@@
 \begin{definition}
-  \label{def-progd}
-  \label{definition-progd}
+  \label{def-aprogd}
+  \label{def-aprogtt}
+  \label{definition-aprogd}
   Given program $\prog$ with $\iov(\prog) = S \cup V \cup O$ and $\flips(\prog) = R$, and

### 2024-05-22 ppdp24/example-gmw.tex
checkpoint

@@
 For full details of the GMW protocol the reader is referred to
-\cite{evans2018pragmatic}. Our implementation libary is shown in
+\cite{evans2018pragmatic}. Our implementation library is shown in
 Figure \ref{fig-gmw}, and includes encoding functions, where
@@
 1-of-4 oblivious transfer. The gate computation is
-done entirely in secrect, and the decoding function
+done entirely in secret, and the decoding function
 is where the declassification occurs-- both parties reveal
@@
 conditional separation relation $\condsep{\pmf}{X_1}{X_2}{X_3}$ to
-mean that $X_2$ and $X_3$ are indpendent in $\pmf$ conditionally on
+mean that $X_2$ and $X_3$ are independent in $\pmf$ conditionally on
 any assignment of values to $X_1$-- i.e., conditionally on any $\store
@@
 is an output of an internal gate, it will definitely be computed using
-random variables, however, it \emph{should} be determistic under any
+random variables, however, it \emph{should} be deterministic under any
 set of input secrets $S$, since we assume that $\idealf$ is
@@
 conditionally on $S$ corrupt views are always in a uniform random
-disrtibution (pure noise), while output $\macgv{\mesg{z}}$ remains
+distribution (pure noise), while output $\macgv{\mesg{z}}$ remains
 separable from corrupt views and both shares of

### 2024-05-22 ppdp24/intro.tex
checkpoint

@@
 connection in an MPC setting include better language abstractions for
-defining protocols and for mechanisation and even automation of
+defining protocols and for mechanization and even automation of
 security proofs.  The goal of this paper is to develop a PL model for
@@
 binary and arithmetic circuits. These low-level protocols encapsulate
-abstractions such as secret sharing and semi-homomorphic encrypytion,
+abstractions such as secret sharing and semi-homomorphic encryption,
 and must be verified by hand. So, low-level MPC
@@
 low-level protocols, the implementation and security of which are
-typically assumed as a selection of library compnonents.
+typically assumed as a selection of library components.
 
@@
 related to passive MPC security. The Haskell-based security type
-system in \cite{6266151} enforces a version of noninterence that is
+system in \cite{6266151} enforces a version of noninterference that is
 sound for passive security, but does not verify the correctness of

### 2024-05-22 ppdp24/intro.tex
spelling

@@
 \cite{10.1007/3-540-44929-9_1,10.1145/3571740}, recent work has also
-recognized connections between MPC security models and hyperpoperties
+recognized connections between MPC security models and hyperproperties
 of, e.g., noninterference \cite{8429300}, and even leveraged these
@@
 computing the pmf of $\minifed$ protocols in $\mathbb{F}_2$, that can
-be automatically queried to enforce hyperperproperties of
+be automatically queried to enforce hyperproperties of
 security. This method is perfectly accurate but has high complexity;

### 2024-06-17 revision-1/model.tex
checkpoint

@@
 \mapsto 0, \elab{\mesg{y}}{2} \mapsto 1 \})$ to denote the (joint) probability that
-$\secret{x}}{1} = 0 \wedge \elab{\mesg{y}}{2} = 1$ in the final memory of a
+$\secret{x}{1} = 0 \wedge \elab{\mesg{y}}{2} = 1$ in the final memory of a
 program containing these variables.

### 2024-06-24 ppdp24/hyper.tex
checkpoint

@@
 with other work \cite{8429300}, we can formulate a version of
-probabilistic noninterence conditioned on output that is sound
+probabilistic noninterference conditioned on output that is sound
 for passive security. 
@@
   Assume given pre-imageable $\idealf$ and a protocol $\prog$ that
-  correctly implements $\idealf$.  If $\prog$ satisfies noninference modulo output
+  correctly implements $\idealf$.  If $\prog$ satisfies noninterference modulo output
   then $\prog$ is passive secure.
@@
   Robust declassification does not imply malicious security, but
-  security with robust declassification implies malicious security.
+  passive security with robust declassification implies malicious security.
 \end{theorem}

### 2024-06-24 ppdp24/model.tex
checkpoint

@@
   $$
-  \progtt(\store) =  1 / |\runs(\prog)| \ \text{if}\ \store \in \runs(\prog), \text{otherwise}\ 0
+  \progtt(\prog)(\store) =  1 / |\runs(\prog)| \ \text{if}\ \store \in \runs(\prog), \text{otherwise}\ 0
   $$
@@
 It is necessary to add $\bot$ to the range of memories since
-the possibility of abort needs to be reflected in adversrarial
+the possibility of abort needs to be reflected in adversarial
 runs of a protocol. We can define $\aruns(\prog,\adversary)$
@@
 protocols, but we also need to ``pad out'' the memories
-of partial runs with $\bot$, as we defin in $\botruns(\prog,\adversary)$,
+of partial runs with $\bot$, as we define in $\botruns(\prog,\adversary)$,
 to properly reflect the contents of views and outputs even in case of abort. 
@@
   $$
-  \progtt(\prog,\adversary) =  1 / |\botruns(\prog)| \ \text{if}\ \store \in \botruns(\prog), \text{otherwise}\ 0
+  \progtt(\prog,\adversary)(\store) =  1 / |\botruns(\prog)| \ \text{if}\ \store \in \botruns(\prog), \text{otherwise}\ 0
   $$

### 2024-06-25 revision-2/example-bdoz.tex
checkpoint

--- 2024-05-22 ppdp24/example-bdoz.tex
@@
 $\macgv{\mesg{w_2}}$ is in an independent uniform distribution,
-this perfectly hides $\secret{w_1}$. In a mutiplication gate
+this perfectly hides $\secret{w_1}$. In a multiplication gate
 either $a$ or $b$ of a Beaver triple are used in secure opening,
@@
 of the local authentication key for both the secure openings and the
-final reveal is enabled by the semi-homorphic properties of the BDOZ
+final reveal is enabled by the semi-homomorphic properties of the BDOZ
 scheme.
@@
 gets caught (and abort). Note in particular that this is \emph{not}
-a form of endorsement (the dual of declassificatio), which allows
+a form of endorsement (the dual of declassification), which allows
 the adversary some leeway. Rather, it is a confirmation of high integrity.

### 2024-06-25 revision-2/examples.tex
checkpoint

--- 2024-05-13 ppdp24/examples.tex
@@
 In a malicious setting, detecting cheating by adding
-information-theoretic secure MAC codes to shares is a fundamental approach
-realized by protocols such as BDOZ \cite{XXX} and SPDZ \cite{XXX}.
-These protocols assume a pre-processing phase that distributes shares
-with MAC codes to clients.  This integrates well with pre-processed
-Beaver Triples to implement malicious secure, and relatively
-efficient, multiplication \cite{XXX}.  Here we show how to implement
-malicious secure two party multiplication in the BDOZ style in
-$\metaprot$.
+information-theoretic secure MAC codes to shares is a fundamental
+approach realized by protocols such as BDOZ \cite{XXX} and SPDZ
+\cite{XXX}.  These protocols assume a pre-processing phase that
+distributes shares with MAC codes to clients.  This integrates well
+with pre-processed Beaver Triples to implement malicious secure, and
+relatively efficient, multiplication \cite{XXX}. Recall
+that Beaver triples are values $a,b,c$ with $\fcod{a * b = c}$,
+unique per multiplication gate, that are secret shared with clients
+during pre-processing.
 
-A field value $v$ is secret shared among 2 client in BDOZ in
-the same manner as in GMW.
-Each client $\cid$ gets a pair of the form $(v_\cid,m_\cid)$, where
-$v_\cid$ is a secret share of $v$, i.e., $v = \fcod{v_1 \fplus v_2}$.
-Further, field values $m_\cid$ are MACs of $v_\cid$ that are authenticated
-by the other client's key. In particular, $m_\cid = k + (k_\Delta *
-v_\cid)$ where \emph{keys} $k$ and $k_\Delta$ are known only to $\cid'
-\ne \cid$ and $k_\Delta$. Further, there is a unique \emph{local key}
-$k$ for each shared value, while the \emph{global key} $k_\Delta$ is
-common to all MACs authenticated by $\cid'$. By extending certain
-field operations to these MACed values, a semi-homorphic encryption
-scheme is obtained that is adequate to compute Beaver Triples. For
-example, if local keys $k_1$ and $k_2$ authenticate $\macv_1$ and
-$\macv_2$, then $k_1 \fplus k_2$ authenticates $\macv_1 \macplus
-\macv_2$, where $\macplus$ is addition extended to MACed
-values. Semantics for summation of MACed shares, and multipication of
-shares and plain field values, are given in Figure \ref{fig-bdoz}.
-For more details the reader is referred to \cite{XXX}. We note
-that while we restrict values $v$, $m$, and $k$ to the same
-field in this presentation for simplicity, in general $m$ and
-$k$ can be in extensions of $\mathbb{Z}_p$.
+A field value $v$ is secret shared among 2 client in BDOZ in the same
+manner as in GMW.  Each client $\cid$ gets a pair of the form
+$(v_\cid,m_\cid)$, where $v_\cid$ is a share of $v$ reconstructed by
+addition, i.e., $v = \fcod{v_1 \fplus v_2}$, and $m_\cid$ is a MAC of
+$v_\cid$.  More precisely, $m_\cid = k + (k_\Delta * v_\cid)$ where
+\emph{keys} $k$ and $k_\Delta$ are known only to $\cid' \ne \cid$ and
+$k_\Delta$. The \emph{local key} $k$ is unique per MAC, while the
+\emph{global key} $k_\Delta$ is common to all MACs authenticated by
+$\cid'$. This is a semi-homomorphic encryption scheme that supports
+addition of shares and multiplication of shares by a constant
+\cite{XXX}. For more details the reader is referred to \cite{XXX}. We
+note that while we restrict values $v$, $m$, and $k$ to the same field
+in this presentation for simplicity, in general $m$ and $k$ can be in
+extensions of $\mathbb{Z}_p$.
 
-We define variable forms $\macx{\secret{w}}{\cid}$ and
-$\macx{\flip{w}}{\cid}$ to range over local MACed shares of a secret
-and random sample owned by client $\cid$, respectively. We let
-$\mack{x}{}$ denote the local key for authenticating shares of $x$.
-We also assume a global key $\flip{\ttt{delta}}$ on each client.
-Pre-processing is modeled by assuming that shares are provided on
-input random tapes (memories) with the required properties, as defined
-on the bottom of Figure \ref{fig-bdoz}.
+We can capture both the distribution of Beaver triples and
+BDOZ shares as a pre-processing predicate that establishes
+conditions for initial memories (see Definition \ref{def-progttb}).
+As for GMW, we use $\macgv{\mesg{w}}$ to refer to secret-shared
+values reconstructed with addition, where by convention
+shares are message values $\elab{\mesg{w}}{\cid}$ for all $\cid$.
+\begin{definition}[BDOZ preprocessing]
+  Define:
+  \begin{mathpar}
+    \mathit{shares} \defeq
+    \{ \elab{\mesg{w}}{\cid}\ |\ \cid \in \{ 1, 2 \} \wedge w \in \{ a,b,c,x,y \}  \}
 
-Similarly to field values, we can use MACed shares in
-uniform independent distributions as one-time-pads for
-hiding other values, using subtraction over MACed shares.
-So we introduce a function $\macotp$ to do this
-declaratively.
-$$
-\delta(\macotp,v_1,v_2) \defeq \fcod{v_1 \macminus v_2}
-$$
+    \mathit{macs} \defeq  \{ \elab{\mesg{w\ttt{mac}}}{\cid}\ |\ \cid \in \{ 1, 2 \} \wedge w \in \{ a,b,c,x,y \}  \}
+
+    \mathit{keys} \defeq  \begin{array}{l}\{ \elab{\mesg{w\ttt{k}}}{\cid}\ |\ \cid \in \{ 1, 2 \} \wedge w \in
+    \{ a,b,c,x,y \}  \} \cup \\ \{ \elab{\mesg{\ttt{delta}}}{\cid}\ |\ \cid \in \{ 1, 2 \} \} \end{array}
+  \end{mathpar}
+  Then a memory $\store$ satisfies BDOZ preprocessing iff:
+  $$\dom(\store) = \{ \elab{\secret{x}}{1}, \elab{\secret{y}}{2} \} \cup \mathit{shares}
+  \cup \mathit{macs} \cup \mathit{keys}$$
+  and, writing $\store(\macgv{\mesg{w}})$ to denote
+  $\fcod{\store(\elab{\mesg{w}}{1}) + \store(\elab{\mesg{w}}{2})}$,
+  the following conditions hold:
+  \begin{mathpar}
+    \store(\macgv{\mesg{x}}) = \store(\elab{\secret{x}}{1})
+
+    \store(\macgv{\mesg{y}}) = \store(\elab{\secret{y}}{2})
+
+    \fcod{\store(\macgv{\mesg{a}}) * \store(\macgv{\mesg{b}})} = \store(\macgv{\mesg{c}})
+  \end{mathpar}
+  and for all $\cid,\cid' \in \{1,2\}$ with $\cid \ne \cid'$ and $w \in \{ a,b,c,x,y\}$:
+  $$\lcod{\store, \mesg{w\ttt{mac}}}{\cid} =
+  \lcod{\store, \mesg{wk} + \mesg{\ttt{delta}} * \mesg{w}}{\cid'}$$
+\end{definition}
+
+\input{beaver}
 
@@
 
+\begin{definition}[BDOZ preprocessing]
+  Define:
+  \begin{mathpar}
+    \mathit{shares} \defeq
+    \{ \elab{\mesg{w}}{\cid}\ |\ \cid \in \{ 1, 2 \} \wedge w \in \{ a,b,c,x,y \}  \}
+
+    \mathit{macs} \defeq  \{ \elab{\mesg{w\ttt{mac}}}{\cid}\ |\ \cid \in \{ 1, 2 \} \wedge w \in \{ a,b,c,x,y \}  \}
+
+    \mathit{keys} \defeq  \begin{array}{l}\{ \elab{\mesg{w\ttt{k}}}{\cid}\ |\ \cid \in \{ 1, 2 \} \wedge w \in
+    \{ a,b,c,x,y \}  \} \cup \\ \{ \elab{\mesg{\ttt{delta}}}{\cid}\ |\ \cid \in \{ 1, 2 \} \} \end{array}
+  \end{mathpar}
+  Then a memory $\store$ satisfies BDOZ preprocessing iff:
+  $$\dom(\store) = \{ \elab{\secret{x}}{1}, \elab{\secret{y}}{2} \} \cup \mathit{shares}
+  \cup \mathit{macs} \cup \mathit{keys}$$
+  and, writing $\store(\macgv{\mesg{w}})$ to denote
+  $\fcod{\store(\elab{\mesg{w}}{1}) + \store(\elab{\mesg{w}}{2})}$,
+  the following conditions hold:
+  \begin{mathpar}
+    \store(\macgv{\mesg{x}}) = \store(\elab{\secret{x}}{1})
+
+    \store(\macgv{\mesg{y}}) = \store(\elab{\secret{y}}{2})
+
+    \fcod{\store(\macgv{\mesg{a}}) * \store(\macgv{\mesg{b}})} = \store(\macgv{\mesg{c}})
+  \end{mathpar}
+  and for all $\cid,\cid' \in \{1,2\}$ with $\cid \ne \cid'$ and $w \in \{ a,b,c,x,y\}$:
+  $$\lcod{\store, \mesg{w\ttt{mac}}}{\cid} =
+  \lcod{\store, \mesg{wk} + \mesg{\ttt{delta}} * \mesg{w}}{\cid'}$$
+\end{definition}
+
 \begin{definition}

### 2024-06-25 revision-2/lang.tex
checkpoint

--- 2024-05-05 ppdp24/lang.tex
@@
 
-The $\minifed$ language provides a simple model of synchronous
+The $\minifed$ language establishes a basic model of synchronous
 protocols between a federation of \emph{clients} exchanging values in
-the binary field. We will identify clients by natural numbers, and
-federations- finite sets of clients- are always given statically.
-As we will see, our threat model assumes a partition of the federation
-into \emph{honest} $H$ and \emph{corrupt} $C$ subsets.
+the binary field. A model of synchronous communication captures a wide
+range of MPC protocols. Concurrency is out of scope in this work but
+an avenue for future work. The lack of sophisticated control
+structures in $\minifed$ is intentional, since minimizing features
+eases analysis and control abstractions such as function definitions
+can be integrated into a metalanguage that generates $\minifed$
+programs (Section \ref{section-metalang}).
 
-We model probabilistic programming via a \emph{random tape}
-semantics. That is, we will assume that programs can make reference to
-values chosen from a uniform random distributions defined in the
-initial program memory.  Programs aka protocols execute
+We identify clients by natural numbers and federations- finite sets of
+clients- are always given statically.  Our threat model assumes a
+partition of the federation into \emph{honest} $H$ and \emph{corrupt}
+$C$ subsets. We model probabilistic programming via a \emph{random
+tape} semantics. That is, we will assume that programs can make
+reference to values chosen from a uniform random distributions defined
+in the initial program memory.  Programs aka protocols execute
 deterministically given the random tape.
@@
 
-The syntax of $\minifed$, defined in
-Figure \ref{fig-minifed}, includes values $v$ and standard
-operations of addition, subtraction, and multiplication in
-a finite field $\mathbb{F}_p$ where $p$ is some prime. 
-Protocols are given input secret values $\secret{w}$
-as well as random samples $\flip{w}$ on the input
-tape, both of which are distinguished by
-strings $w$. Protocols are sequences of assignment
-commands of three different forms:
+The syntax of $\minifed$, defined in Figure \ref{fig-minifed},
+includes values $v$ and standard operations of addition, subtraction,
+and multiplication in a finite field $\mathbb{F}_p$ where $p$ is some
+prime.  Protocols are given input secret values $\secret{w}$ as well
+as random samples $\flip{w}$ on the input tape, implemented using a
+\emph{memory} as described below (Section
+\ref{section-lang-semantics}) where $w$ is a distinguishing 
+identifier string. Protocols are sequences of assignment commands of three
+different forms:
 \begin{itemize}
@@
   is an \emph{output} where expression $\be$ is computed
-  by client $\cid$ and reported as its output.
+  by client $\cid$ and reported as its output. As a
+  sanity condition we disallow commands
+  $\pubout{\cid_1}{\be}{\cid_2}$ where $\cid_1\ne\cid_2$.
 \end{itemize}
-
+For example, in the following protocol, a client 1
+subtracts a random sample $\flip{y}$ from $\mathbb{F}_p$ from their
+secret value $\secret{x}$ and sends the result to client
+2 as a message $\mesg{z}$:
+$$
+\eassign{\mesg{z}}{2}{(\secret{x} - \flip{y})}{1}
+$$
 Both messages $\mesg{w}$ and reveals $\rvl{w}$ can be
-referenced in expressions, once they've been assigned.
+referenced in expressions once they've been defined.
 This distinction between messages and broadcast public
@@
 $\iov(\prog) \cup \flips(\prog)$. For any set of variables $X$ and
-parties $P$, we write $X_P$ to denote the subset of $X$ owned by any
-party in $P$, in particular we write $X_H$ and $X_C$ to denote the
+clients $I$, we write $X_I$ to denote the subset of $X$ owned by any
+client $\cid \in I$, in particular we write $X_H$ and $X_C$ to denote the
 subsets belonging to honest and corrupt parties, respectively.
@@
 \subsection{Semantics}
+\label{section-lang-semantics}
 
@@
 client $\cid$ as $\lcod{\store,\be}{\cid}$. This interpretation is
-defined in Figure \ref{fig-minifed}. It is also parameterized by
-$\delta$ which defines the semantics of library functions $f$.
-
-The small-step reduction relation $\redx$ is then defined in Figure
-\ref{fig-minifed} to evaluate commands. Reduction is a relation on
-\emph{configurations} $(\store, \prog)$ where all three command forms-
-message send, broadcast, and output- are implemented as updates to the
-memory $\store$. We write $\redxs$ to denote the reflexive, transitive
-closure of\ $\redx$.
+defined in Figure \ref{fig-minifed}. The small-step reduction relation
+$\redx$ is then defined in Figure \ref{fig-minifed} to evaluate
+commands. Reduction is a relation on \emph{configurations} $(\store,
+\prog)$ where all three command forms- message send, broadcast, and
+output- are implemented as updates to the memory $\store$. We write
+$\redxs$ to denote the reflexive, transitive closure of\ $\redx$.
 
 \subsection{Example: Passive Secure Addition}
+\label{section-lang-example}
 
@@
 arithmetic fields to implement secret addition. If a field value $v_1$
-is in a uniform random distribution with other variables in a program,
-then either  $v_1 \fplus v_2$ or $v_1 \fminus v_2$  is an encryption of $v_2$
+is uniformly random, then $v_1 \fminus v_2$ is an encryption of $v_2$
 where $v_1$ is an information theoretically secure one-time-pad, which
-is exploited for secret sharing. Of course, $\fplus$ and $\fminus$ are also
-meaningful operations over any two field values regardless of their distributions.
+is exploited for secret sharing, noting that $v_2$ can be
+reconstructed by summing $v_1$ and $v_3 \defeq v_1 \fminus v_2$.
 
-To sum their secret values $\secret{\cid}$, each client $\cid$ in
-the federation $\{ 1, 2, 3 \}$  samples a value $\locflip$
-that can be used as a one-time pad with another
-random sample $\flip{x}$ and $\secret{\cid}$. This yields
-two secret shares communicated as messages to the other clients,
-while each client keeps $\locflip$ as its own share.
+In $\minifed$, to privately sum secret values $\secret{\cid}$, each
+client $\cid$ in the federation $\{ 1, 2, 3 \}$ samples a value
+$\locflip$ that can be used as a one-time pad with another random
+sample $\flip{x}$ and $\secret{\cid}$. This yields two secret shares
+communicated as messages to the other clients, while each client keeps
+$\locflip$ as its own share.
 $$
 \begin{array}{lll}
-  \elab{\mesg{s1}}{2} &:=& \elab{(\locflip \fminus \flip{x} \fminus \secret{1})}{1} \\ 
+  \elab{\mesg{s1}}{2} &:=& \elab{(\secret{1} \fminus \locflip \fminus \flip{x})}{1} \\ 
   \elab{\mesg{s1}}{3} &:=& \elab{\flip{x}}{1} \\ 
-  \elab{\mesg{s2}}{1} &:=& \elab{(\locflip \fminus \flip{x} \fminus \secret{2})}{2} \\ 
+  \elab{\mesg{s2}}{1} &:=& \elab{(\secret{2} \fminus \locflip \fminus \flip{x})}{2} \\ 
   \elab{\mesg{s2}}{3} &:=& \elab{\flip{x}}{2} \\ 
-  \elab{\mesg{s3}}{1} &:=& \elab{(\locflip \fminus \flip{x} \fminus \secret{3})}{3} \\ 
+  \elab{\mesg{s3}}{1} &:=& \elab{(\secret{3} \fminus \locflip \fminus \flip{x})}{3} \\ 
   \elab{\mesg{s3}}{2} &:=& \elab{\flip{x}}{3}
@@
 $$
-Due to field properties of $\fminus$ this scheme guarantees that messages
+This scheme guarantees that messages
 are viewed as random noise by any observer 

### 2024-06-25 revision-2/m.tex
checkpoint

\section{Confidentiality Types for $\minicat$}

\newcommand{\sx}[2]{\elab{\secret{#1}}{#2}}
\newcommand{\mx}[2]{\elab{\mesg{#1}}{#2}} 
%\newcommand{\px}[2]{\elab{\rvl{#1}}{#2}} 
\newcommand{\rx}[2]{\elab{\flip{#1}}{#2}}
\newcommand{\ox}[2]{\elab{\out{#1}}{#2}}
\newcommand{\signals}{\leadsto}

\newcommand{\tj}[5]{#1,#2 \vdash_{#3} #4 : #5}
\newcommand{\cty}[2]{c(#1,#2)}
\newcommand{\setit}[1]{\{ #1 \}}
\newcommand{\ty}{T}

\begin{definition}
  A \emph{ciphertype} is of the form $\cty{x}{\ty}$ where $\ty$ is a \emph{type} which is a set
  of either either variables $x$ or ciphertypes. A \emph{type environment} $\Gamma$ is a finite
  list of type bindings. 
\end{definition}

Intuitively, a type is the set of variables on which a given value may
depend, and a ciphertype $\cty{x}{\ty}$ denotes a value with type
$\ty$ that has been encrypted using $x$ as a one-time-pad. So for
example, the following is the type of $\mx{s1}{2}$ in the example in Section
\ref{section-lang-example}:
$$
\setit{\cty{\elab{\locflip}{1}}{\setit{\cty{\rx{x}{1}}{\setit{\sx{1}{1}}}}}}
$$
and the following is the type of $\mx{s1}{3}$:
$$
\setit{\rx{x}{1}}
$$

Further, the union of the types of multiple values represents their
joint dependencies. We make the conservative approximation that any
dependency on a one-time-pad $x$ ``unlocks'' dependencies of a value
encrypted with $x$. To capture a notion of closure under this
interpretation we define the type rewrite relation $\signals$.
\begin{mathpar}
  (\setit{x,\cty{x}{\ty'}} \cup \ty) \signals (\setit{x} \cup \ty' \cup \ty)

  \inferrule
  {\ty_1 \signals \ty_2 \\ \ty_2 \signals \ty_3}
  {\ty_1 \signals \ty_3}

  \ty \signals \ty
\end{mathpar}

Then, a set of values is considered high security if it has any
dependencies on honest secrets revealed by $\signals$. This allows us
to associate \emph{security labels} $\hilab$ and $\lolab$ with types.
Note that this implicitly assumes that encrypted values, random
samples, and corrupt secrets are all low security.
$$
\seclab(\ty) =
\begin{cases}
  \hilab \text{\ if\ } \ty \signals \{ \sx{w}{\cid} \} \cup \ty' \text{\ and\ } \cid \in H \\
  \lolab \text{\ otherwise}
\end{cases}
$$

Expression type judgements are of the form
$\tj{\Gamma}{R}{\cid}{\be}{\ty}$, where $\Gamma$ is a classical type
environment and $R$ is a linear type environment.  The latter ensures
that a random variable $\elab{\flip{w}}{\cid}$ is only used once for
this purpose, though it may be used more than once as a classical type
that reflects its dependency. To help enforce this in the analysis We
write $R_1;R_2$ to denote the union of $R_1$ and $R_2$ if $R_1$ and
$R_2$ are disjoint, otherwise $R_1;R_2$ is undefined. The type of
programs $\Gamma,R \vdash \prog : \Gamma'$ can be viewed as a Hoare
Type where $\Gamma$ represents dependency preconditions in memory and
$\Gamma'$ are dependencies resulting from execution of $\prog$.

The most interesting rule is $\TirName{Share}$ which generates
a ciphertype, given a random sample used as a one-time-pad that
from which the encrypted value is either added or subtracted.
Additionally, the linear use of the sample is registered in
the judgement. 
\begin{mathpar}
  \inferrule[Value] {}
            {\tj{\Gamma}{\varnothing}{\cid}{v}{\varnothing}}

  \inferrule[Secret]
  {}
  {\tj{\Gamma}{\varnothing}{\cid}{\secret{w}}{\setit{\sx{w}{\cid}}}}

  \inferrule[PubM]
  {}
  {\tj{\Gamma}{\varnothing}{\cid}{\rvl{w}}{\setit{\rvl{w}}}}

  \inferrule[Rando]
  {}
  {\tj{\Gamma}{\varnothing}{\cid}{\flip{w}}{\setit{\rx{w}{\cid}}}}

  \inferrule[Mesg]
  {}
  {\tj{\Gamma}{\varnothing}{\cid}{\mesg{w}}{\Gamma(\mx{w}{\cid})}}

  \inferrule[Share]
  {\tj{\Gamma}{R}{\cid}{\be}{\ty} \\ \oplus \in \{ \fplus,\fminus,\ftimes \}}
  {\tj{\Gamma}{R;\rx{w}{\cid}}{\cid}{\flip{w} \oplus \be}{\setit{\cty{\rx{w}{\cid}}{\ty}}}}

  \inferrule[Binop]
  {\tj{\Gamma}{R_1}{\cid}{\be_1}{\ty_1} \\ \tj{\Gamma}{R_1}{\cid}{\be_1}{\ty_1} \\
              \oplus \in \{ \fplus,\fminus,\ftimes \}}
  {\tj{\Gamma}{R_1;R_2}{\cid}{\be_1 \oplus \be_2}{\ty_1 \cup \ty_2}}

  \inferrule[Command]
  {\tj{\Gamma}{R_1}{\cid}{\be}{\ty} \\ \Gamma; x : \ty,R_2 \vdash \prog : \Gamma'}
  {\Gamma,R_1;R_2 \vdash x := \elab{\be}{\cid};\prog : \Gamma'}

  \inferrule[Term]
  {}
  {\Gamma,\varnothing \vdash \varnothing: \Gamma}          
\end{mathpar}

\begin{theorem}
  Given $\varnothing,R \vdash \prog : \Gamma$ and:
  $$\seclab(\bigcup_{\mx{w}{\cid} \in X} \Gamma(\mx{w}{\cid}) \cup \flips(\prog)_C) = \lolab$$
  for $X \defeq \dom(\Gamma)_C$.
  Then $\prog$ satisfies gradual release.
\end{theorem}

$$
\mx{x}{1} : \setit{\rx{x}{1}}

\mx{x}{2} : \setit{\cty{\rx{x}{1}}{\sx{x}{1}}}
$$

$$
\mx{c}{1} : \setit{\rx{c}{1}}

\mx{c}{2} : \setit{ \cty{\rx{c}{1}}{\setit{\rx{a}{1},\rx{a}{2},\rx{b}{1},\rx{b}{2}}} }
$$

### 2024-06-25 revision-2/metalang.tex
checkpoint

--- 2024-05-22 ppdp24/metalang.tex
@@
 component definitions support compositional verification of larger
-prortocols as we will discuss with examples in Sections
+protocols as we will discuss with examples in Sections
 \ref{section-example-gmw} and \ref{section-example-bdoz}.
@@
 In this definition we write $e[\mv/y]$ to denote the substitution of $\mv$
-for free occurences of $y$ in $e$. The rules are mostly standard,
+for free occurrences of $y$ in $e$. The rules are mostly standard,
 except when a concrete $\minifed$ assignment is encountered it is added

### 2024-06-25 revision-2/minitypes.tex
checkpoint

--- 2024-06-25 revision-2/m.tex
@@
 \newcommand{\ty}{T}
+\newcommand{\eqs}{C}
+\newcommand{\toeq}[1]{\lfloor #1 \rfloor}
+\newcommand{\autheq}[1]{\phi_{\mathrm{auth}}(#1)}
+\newcommand{\upgrade}[1]{\uparrow #1}
 
@@
 
-  \inferrule[PubM]
-  {}
-  {\tj{\Gamma}{\varnothing}{\cid}{\rvl{w}}{\setit{\rvl{w}}}}
-
   \inferrule[Rando]
@@
 
+  \inferrule[PubM]
+  {}
+  {\tj{\Gamma}{\varnothing}{\cid}{\rvl{w}}{\Gamma(\rvl{w})}}
+
   \inferrule[Share]
-  {\tj{\Gamma}{R}{\cid}{\be}{\ty} \\ \oplus \in \{ \fplus,\fminus,\ftimes \}}
-  {\tj{\Gamma}{R;\rx{w}{\cid}}{\cid}{\flip{w} \oplus \be}{\setit{\cty{\rx{w}{\cid}}{\ty}}}}
+  {\tj{\Gamma}{R}{\cid}{\be}{\ty} \\ \oplus \in \{ \fplus,\fminus \}}
+  {\tj{\Gamma}{R;\rx{w}{\cid}}{\cid}{\be \oplus \flip{w}}{\setit{\cty{\rx{w}{\cid}}{\ty}}}}
 
@@
 
-\begin{theorem}
-  Given $\varnothing,R \vdash \prog : \Gamma$ and:
-  $$\seclab(\bigcup_{\mx{w}{\cid} \in X} \Gamma(\mx{w}{\cid}) \cup \flips(\prog)_C) = \lolab$$
-  for $X \defeq \dom(\Gamma)_C$.
+Then a main result is that protocols where the collective messages received by the
+adversary are low security satisfy gradual release.
+\begin{theorem}[Confidentiality Type Safety]
+  Given $\Gamma_1,R \vdash \prog : \Gamma_2$ where $\Gamma_1$ is sound for
+  preprocessing and:
+  $$\seclab(\bigcup_{\mx{w}{\cid} \in X} \Gamma_2(\mx{w}{\cid}) \cup \flips(\prog)_C) = \lolab$$
+  for $X \defeq \dom(\Gamma_2)_C$.
   Then $\prog$ satisfies gradual release.
@@
 
-$$
+This definition accounts for preprocessing of memories. In the default
+case where initial memories just contain secrets and the random tape,
+we can take $\Gamma_1$ to be $\varnothing$ In cases of nontrivial
+pre-processing we need to define $\Gamma_1$ \cnote{and also a formal
+  notion of soundness}. For example, in a 2-client Beaver Triple
+scheme, we imagine that client 1's secret $x$ is shared as
+messages $\mx{x}{1}$ and $\mx{x}{x}$ in the initial memory, obtained
+via the ``0-time-pad'' use of the random value $\rx{x}{1}$.
+So in this case $\Gamma_1$ would contain:
+\begin{mathpar}
 \mx{x}{1} : \setit{\rx{x}{1}}
 
-\mx{x}{2} : \setit{\cty{\rx{x}{1}}{\sx{x}{1}}}
+\mx{x}{2} : \setit{\cty{\rx{x}{1}}{\setit{\sx{x}{1}}}}
+\end{mathpar}
+Similarly, we imagine that clients 1 and 2 possess respective
+shares of random values $a$ and $b$ on their random tape:
 $$
-
+\rx{a}{1}\quad  \rx{a}{2}\quad \rx{b}{1}\quad \rx{b}{2} 
 $$
+as well as shares $\mx{c}{1}$ and $\mx{c}{2}$ of a value
+$c \defeq a * b$, i.e.:
+$$c \defeq
+(\rx{a}{1} \fplus \rx{a}{2}) \ftimes (\rx{b}{1} \fplus \rx{b}{2})$$
+so $\Gamma_1$ contains:
+\begin{mathpar}
   \mx{c}{1} : \setit{\rx{c}{1}}
@@
 \mx{c}{2} : \setit{ \cty{\rx{c}{1}}{\setit{\rx{a}{1},\rx{a}{2},\rx{b}{1},\rx{b}{2}}} }
-$$
+\end{mathpar}
+
+\section{Confidentiality and Integrity Types for $\minifed$}
+
+\begin{mathpar}
+  \inferrule[Value] {}
+            {\tj{\Gamma}{\varnothing}{\cid}{v}{\varnothing}}
+
+  \inferrule[Secret]
+  {}
+  {\tj{\Gamma}{\varnothing}{\cid}{\secret{w}}{\setit{\sx{w}{\cid}}}}
+
+  \inferrule[Rando]
+  {}
+  {\tj{\Gamma}{\varnothing}{\cid}{\flip{w}}{\setit{\rx{w}{\cid}}}}
+
+  \inferrule[Mesg]
+  {}
+  {\tj{\Gamma}{\varnothing}{\cid}{\mesg{w}}{\Gamma(\mx{w}{\cid})}}
+
+  \inferrule[PubM]
+  {}
+  {\tj{\Gamma}{\varnothing}{\cid}{\rvl{w}}{\Gamma(\rvl{w})}}
+
+  \inferrule[Share]
+  {\tj{\Gamma}{R}{\cid}{\be}{\ty} \\ \oplus \in \{ \fplus,\fminus \}}
+  {\tj{\Gamma}{R;\rx{w}{\cid}}{\cid}{\be \oplus \flip{w}}{\setit{\cty{\rx{w}{\cid}}{\ty}}}}
+
+  \inferrule[Binop]
+  {\tj{\Gamma}{R_1}{\cid}{\be_1}{\ty_1} \\ \tj{\Gamma}{R_1}{\cid}{\be_1}{\ty_1} \\
+              \oplus \in \{ \fplus,\fminus,\ftimes \}}
+  {\tj{\Gamma}{R_1;R_2}{\cid}{\be_1 \oplus \be_2}{\ty_1 \cup \ty_2}}
+
+  \inferrule[Command]
+  {\tj{\Gamma}{R_1}{\cid}{\be}{\ty} \\ \Gamma; x : \ty,R_2, \eqs \cup \{ x = \toeq{\elab{\be}{\cid}} \} \vdash \prog : \Gamma'}
+  {\Gamma,R_1;R_2, \eqs \vdash x := \elab{\be}{\cid};\prog : \Gamma'}
+
+  \inferrule[Assert]
+  {\eqs \vdash \toeq{\elab{\be_1}{\cid}} = \toeq{\elab{\be_2}{\cid}} 
+              \\ \Gamma; R, \eqs \vdash \prog : \Gamma'}
+  {\Gamma,R,\eqs \vdash \elab{\assert{\be_1 = \be_2}}{\cid};\prog : \Gamma'}
+
+  \inferrule[Auth]
+  {\eqs \vdash \toeq{\autheq{\mesg{w}}}
+              \\ \Gamma; \elab{\mesg{w}}{\cid} : \upgrade{\Gamma(\elab{\mesg{w}}{\cid})}, R, \eqs \vdash \prog : \Gamma'}
+  {\Gamma,R,\eqs \vdash \elab{\assert{\autheq{\mesg{w}}}}{\cid};\prog : \Gamma'}
+
+  \inferrule[Term]
+  {}
+  {\Gamma,\varnothing \vdash \varnothing: \Gamma,\varnothing}          
+\end{mathpar}

### 2024-06-29 ppdp24/response.txt
checkpoint

Many thanks to the reviewers for their thoughtful comments. To be succinct we restrict our responses to main issues raised.

Reviewer 1 comments:

"The definition of sigma \cap sigma' does not line up with its usage in many places. The intersection is defined so that the domain of the intersection is the intersection of the domains. This is corroborated by the claim that sigma \sub sigma' iff (sigma \cap sigma') = sigma...However, in numerous locations (sigma \cap sigma') seems to be used in place of the union (sigma \cup sigma') which would be correct. Here, one would define (sigma \cup sigma')(x) to be sigma(x) if x \in dom(\sigma) and sigma'(x) if x \in dom(\sigma') with the condition that they agree on their intersection (with the union being undefined otherwise). Note that this satisfies sigma \sub sigma' iff (sigma \cup sigma') = sigma'."

Our definition of sigma \cap sigma' is in fact the one that the reviewer identifies as correct, not the incorrect one the reviewer suggests. In Section 2.2 we say "we write sigma \cap sigma to denote the combination of sigma and sigma' assuming sigma(x) = sigma'(x) for all x in dom(sigma) \cap dom(sigma'), otherwise sigma cap sigma' is undefined." Perhaps the word "combination" here was misleading but we meant "unioning" in the sense the reviewer means. It is also possible that the use of \cap instead of another symbol is confusing, and we can certainly clarify this in a revision and/or use a different symbol. However the reviewer's definition is the one intended in our text.

"Def. 3.3 also has typing problems. In the numerator, P_X specifically has domain mems(X), but its argument (sigma \cap sigma') is in mems(Y) with Y \sub X. I believe the definition would be correct if the union is used instead."

Our previous comment (about sigma \cap sigma') applies to this also.

"The text in Section 3.1 is confusing at best, and I believe actually incorrect (in part due to the intersection/union confusion). First, for a PMF P, P should be viewed as a function mems(X) -> R for some parameter X. This is implicit in Def. 3.1, but making it explicit would help elucidate the issues with definitions 3.2 and 3.3."

Our previous comment applies to this in part, especially wrt correctness. Additionally, in Definition 3.3 we introduce syntactic sugaring P(sigma) to denotes P_X(sigma) where dom(X) is given (as X), which we believe is conventional. However we agree with the reviewer that subsequent definitions would be clarified by being explicit (about X) in Definitions 3.1-3.

"In Def. 3.2, the implied domain of P is mems(Y) for some X \sub Y. The marginal should be a restriction to a smaller/shorter memory. Then sigma' is given as being in mems(X - dom(dom(P))). It's not clear what dom(dom(P)) means. If dom(P) = mems(Y) and dom(mems(Y)) = Y, then it might mean Y. But then X - Y is empty since X \sub Y. Probably the proper definition should be..."

Yes we meant "mems(dom(dom(P)) - X)" (not "mems(X - dom(dom(P)))"), with that edit the definition comports with the one the reviewer suggests. The meaning of dom(dom(P)) is determined by the definition of memories in Section 2.2 and Definition 3.1 (though we can clarify this in the definition).

"Fig. 2, middle: Are the outer brackets on lines 2-4 needed? In any case, their meaning is never defined (though one can infer it)."

The meaning of the outer brackets is defined in Section 2.2 in the paragraph just before Section 2.3.

"One technical error is in Secton 2.3. The subtractions are all backwards."

Yes this was a typo, we did have the order right in Section 2.1.

"Ultimately, these technical issues make it exceedingly difficult to judge the coreectness."

Overall, it seems that a misunderstanding of our definition of sigma \cap \sigma' led to an impression of more general incorrectness. We hope that our above comments have clarified that our definition aligns with the one the reviewer states to be correct, since further we believe that the Definitions in Section 3.1 are more or less standard modulo some specialized notation. We can certainly clarify and improve this notation, e.g., by using a different symbol for \cap and highlighting our Definition.

Reviewer 2 comments:

"- Page 8, Section 6. Why do you call Prelude a "metalanguage"? A metalanguage is a language for defining languages (example: BNF-like formats such as yacc or bison). In fact, Prelude is a language which can be translated into Overture programs, and the "semantics" defined in Figure 5 is the translation from Prelude to Overture. Indeed, Prelude has no real operational semantics acting on a memory; only Overture has it."

Our use of the term "metalanguage" was based on previous "staged" or "metaprogramming" languages such as MetaML, where programs produce programs. The semantics of Prelude (Figure 5) specify that Prelude programs yield Overture programs as side effects of computation, hence our use of the term. We now recognize that the reviewer's definition is popular so we can change "metalanguage" to "staged language" to avoid misleading readers.

### 2024-06-29 ppdp24/response.txt
checkpoint

@@
-Many thanks to the reviewers for their thoughtful comments. To be succinct we restrict our responses to main issues raised.
+Many thanks to the reviewers for their thoughtful comments. We will consider and address all of them in future iterations of this work. To be succinct we restrict our responses to main issues raised.
 
@@
 
-"Ultimately, these technical issues make it exceedingly difficult to judge the coreectness."
+"Ultimately, these technical issues make it exceedingly difficult to judge the coreectness..."
 
@@
 We hope that our above comments have clarified that our definition aligns with the one the reviewer states to be correct, since further we believe that the Definitions in Section 3.1 are more or less standard modulo some specialized notation.
-We can certainly clarify and improve this notation, e.g., by using a different symbol for \cap and highlighting our Definition.
+We can certainly clarify and improve this notation , e.g., by using a different symbol for \cap and highlighting our Definition.
 

### 2024-07-15 ppdp24/example-bdoz.tex
checkpoint

@@
 simplicity, in general $m$ and $k$ can be in extensions of
-$\mathbb{Z}_p$.
+$\mathbb{F}_p$.
 

### 2024-07-15 ppdp24/hyper.tex
checkpoint

@@
   $\idealf$ and
-  has integrity, then for all $H$, $C$, and $\adversary$ and $\store_1 \in \mems(S_H)$,
-  there exists $\store_2 \in \mems(S_C)$ such that:
+  has integrity, then for all $H$, $C$, $\adversary$, $\store_1 \in \mems(S_H)$,
+  $\ox{\cid} \in (O_H)$, and $\mv \in \mathbb{F}_p$, if:
   $$
-  \idealf(\store_1 \cap \store_2)_H \cup \{ \bot \} \supseteq
-    \support(\progtt(\prog,\adversary)(O_H|\store_1))
+  \progtt(\prog,\adversary)(\{ \ox{\cid} \mapsto \mv \} \mid \store_1) > 0
+  $$
+  then exists $\store_2 \in \mems(S_C)$ such that:
+  $$
+  \idealf(\store_1 \uplus \store_2)(\ox{\cid}) = \mv
   $$

### 2024-07-15 ppdp24/lang.tex
checkpoint

@@
 message sends, public broadcast, and client outputs. Memories $\store$ are finite
-(partial) mapping from variables $x$ to values $v \in \mathbb{Z}_p$. The \emph{domain} of a
-memory is written $\dom(\store)$ and is the finite set of variables on
-which the memory is defined. We write $\store\{ x \mapsto v\}$ for
-$x\not\in\dom(\store)$ to denote the memory $\store'$ such that
+(partial) mapping from variables $x$ to values $v \in \mathbb{F}_p$.
+The \emph{domain} of a memory is written $\dom(\store)$ and is
+the finite set of variables on which the memory is defined.
+We write $\store\{ x \mapsto v\}$ for $x\not\in\dom(\store)$
+to denote the memory $\store'$ such that
 $\store'(x) = v$ and otherwise $\store'(y) = \store(y)$ for all $y
@@
 $\dom(\store) \subseteq \dom(\store')$ and $\store(x) =
-\store'(x)$ for all $x \in \dom(\store)$. We write $\store \cap
+\store'(x)$ for all $x \in \dom(\store)$. We write $\store \uplus
 \store'$ to denote the combination of $\store$ and $\store'$
 assuming $\store(x) = \store'(x)$ for all $x \in \dom(\store)
-\cap \dom(\store')$, otherwise $\store \cap \store'$ is undefined.
-We write $\store \subseteq \store'$ iff $\store \cap \store'
+\cap \dom(\store')$, otherwise $\store \uplus \store'$ is undefined.
+\cnote{Check to make sure all $\cap$ and $\cup$s have been revised
+with $\uplus$.}
+We write $\store \subseteq \store'$ iff $\store \uplus \store'
 = \store$.
@@
 the standard interpretation of $\be$ in the arithmetic field
-$\mathbb{Z}_{p}$. With the introduction of variables to expressions,
+$\mathbb{F}_{p}$. With the introduction of variables to expressions,
 we need to interpret variables with respect to a specific memory, and

### 2024-07-15 ppdp24/model.tex
checkpoint

@@
 
-We define joint probability mass functions (pmfs) in the standard
-manner, though following \cite{barthe2019probabilistic} we use
-memories to denote mappings of variables to values (i.e., outcomes),
-so for example given a pmf $\pmf$ we will write $\pmf(\{ \elab{\secret{x}}{1}
-\mapsto 0, \elab{\mesg{y}}{2} \mapsto 1 \})$ to denote the (joint) probability that
-$\elab{\secret{x}}{1} = 0 \wedge \elab{\mesg{y}}{2} = 1$.
+We define discrete joint probability mass functions (pmfs) in a
+standard manner but develop some notations that are useful for our
+presentation. Firstly, whereas distributions typically map
+realizations of random variables to values in $[0..1]$, for
+convenience and following \cite{barthe2019probabilistic} we use
+memories to represent realizations, so for example given a pmf $\pmf$
+over variables $\{ \sx{x}{1}, \mx{y}{2} \}$ we write $\pmf(\{
+\elab{\secret{x}}{1} \mapsto 0, \elab{\mesg{y}}{2} \mapsto 1 \})$ to
+denote the (joint) probability that $\elab{\secret{x}}{1} = 0 \wedge
+\elab{\mesg{y}}{2} = 1$. Recall from Section
+\ref{section-lang-semantics} that $\uplus$ denotes the combination of
+memories, so for example $\{ \elab{\secret{x}}{1} \mapsto 0\} \uplus
+\{\elab{\mesg{y}}{2} \mapsto 1 \} = \{ \elab{\secret{x}}{1} \mapsto 0,
+\elab{\mesg{y}}{2} \mapsto 1 \}$.
 \begin{definition}
   A \emph{probability mass function} $\pmf$ is a function
-  mapping memories in $\mems(X)$ for some $X$ to values in $\mathbb{F}_p$, such that:
+  mapping memories in $\mems(X)$ for given variables $X$ to
+  values in $\mathbb{R}$ such that:
   $$
@@
   $$
+  %The \emph{support} of a distribution is the set of realizations
+  %of random variables with non-zero probability:
+  %$\support(\pmf) \defeq \{ (v_1,\ldots,v_n) \mid
+  %\pmf(\{ x_1 \mapsto v_1, \ldots, x_n \mapsto v_n\}) > 0 \} $
 \end{definition}
@@
 \begin{definition}
-  Given $\pmf$ the \emph{marginal distribution} of variables $X$
-  in $\pmf$, denoted $\margd{\pmf}{X}$, is defined as follows:
+  Given $\pmf$ with $\dom(\pmf) = \mems(Y)$, the \emph{marginal distribution}
+  of variables $X \subseteq Y$ in $\pmf$, denoted $\margd{\pmf}{X}$, is defined as follows:
   $$
   \forall \store \in \mems(X) \quad . \quad \margd{\pmf}{X}(\store) =
-  \sum_{\store' \in \mems(X-\dom(\dom(\pmf)))} \pmf(\store \cap \store')
+  \sum_{\store' \in \mems(Y-X)} \pmf(\store \uplus \store')
   $$
@@
 \begin{definition}
-  Given $\margd{\pmf}{X}$, let $\stores$ be a set of memories with the
-  same domain $Y \subseteq X$. Then the \emph{conditional distribution given
-  $\stores$}  denoted
-  $\condd{\pmf}{X}{\stores}$ is a distribution with domain $X$ where for all
-  $\store \in \mems(X)$:
+  Given $\pmf$ with $\dom(\pmf) = \mems(Z)$, the \emph{conditional distribution}
+  of $X$ given $Y$ where $X \cup Y \subseteq Z$ and $X \cap Y = \varnothing$, denoted $\condd{\pmf}{X}{Y}$,
+  is defined as follows:
   $$
-  \condd{\pmf}{X}{\stores}(\store) =
-  (\sum_{\store' \in \stores} \margd{\pmf}{X}(\store \cap \store')) /
-  (\sum_{\store' \in \stores} \margd{\pmf}{Y}(\store'))
+  \forall \store \in \mems(X \cup Y)\ .\ 
+  \condd{\pmf}{X}{Y}(\store) =
+  \begin{cases}
+    0 \text{\ if\ } \margd{\pmf}{Y}(\store_Y) = 0\\
+    \margd{\pmf}{X \cup Y}(\store) / \margd{\pmf}{Y}(\store_Y)\ \text{\ otherwise}
+  \end{cases}
   $$
-  where $\margd{\pmf}{X}(\store \cap \store')) = 0$ if $\store \cap \store'$ is undefined.
 \end{definition}
-To recover familiar notation we allow the syntactic
-sugarings $\condd{\pmf}{X}{\store}  \defeq \condd{\pmf}{X}{\{ \store\}}$, and
-$\pmf(\store)  \defeq \margd{\pmf}{X}(\store)$ and $\pmf(\store|\stores) \defeq
-\condd{\pmf}{X}{\stores}(\store)$ where $\dom(\store) = X$.
-%\begin{eqnarray*}
-%  \condd{\pmf}{X}{\store}  &\defeq& \condd{\pmf}{X}{\{ \store\}}\\
-%  \pmf(\store)  &\defeq& \margd{\pmf}{X}(\store)  \qquad \dom(\store) = X\\
-%  \pmf(\store|\stores)  &\defeq& \condd{\pmf}{X}{\stores}(\store) \qquad \dom(\store) = X
-%\end{eqnarray*}
-
-We also define the \emph{support} of a distribution in the usual manner-
-it is the set of values a set of variables can take on with non-zero
-probability.
-\begin{definition}[Support]
-  $\support(\pmf) \defeq \{ (v_1,\ldots,v_n) \mid
-  \pmf(x_1 \mapsto v_1, \ldots, x_n \mapsto v_n) > 0 \} $
+We also define some convenient syntactic sugarings. The first will allow us to
+compare marginal distributions under different realization conditions
+(as in, e.g., Definition \ref{definition-NIMO}), the others are standard.
+\begin{definition}
+  Given $\pmf$, for all $\store_2 \in \mems(Y)$ and $\store_1 \in \mems(X)$ define:
+  \begin{enumerate}
+  \item $\condd{\pmf}{X}{\store_2}(\store_1) \defeq \condd{\pmf}{X}{Y}(\store_1 \uplus \store_2)$
+  \item $\pmf(\store_1)  \defeq \margd{\pmf}{X}(\store_1)$ 
+  \item $\pmf(\store_1|\store_2) \defeq \condd{\pmf}{X}{Y}(\store_1 \uplus \store_2)$
+  \end{enumerate}
 \end{definition}

### 2024-07-16 ppdp24/hyper.tex
checkpoint

@@
   $$
-  \condd{\progtt(\prog)}{S_H}{\store_1} = \condd{\progtt(\prog)}{S_H}{\store_1 \cap \store_2}
+  \condd{\progtt(\prog)}{S_H}{\store_1} = \condd{\progtt(\prog)}{S_H}{\store_1 \uplus \store_2}
  $$
@@
   with inputs $\store'_C$ and arbitrary $\store'' \in \mems(S_H)$
-  such that $\store'_C \cap \store'' \in \ik(\idealf,\idealf(\store'))$.
+  such that $\store'_C \uplus \store'' \in \ik(\idealf,\idealf(\store'))$.
   The assumption of passive security of $\prog$ implies the result.

### 2024-07-16 ppdp24/example-gmw.tex
checkpoint

@@
   $$\pmf(\macgv{\mesg{w}} = v) \defeq \sum_{\sigma \in A} \pmf(\sigma)$$
-  and define:
+  and also for all $\store \in \mems(X)$ define:
   $$\condd{\pmf}{X}{\macgv{\mesg{w}} = v}(\sigma) \defeq  \sum_{\sigma' \in A} \condd{\pmf}{X}{\sigma'}(\sigma)$$

### 2024-07-16 ppdp24/example-gmw.tex
checkpoint

--- 2024-05-22 ppdp24/example-gmw.tex
@@
   $$\pmf(\macgv{\mesg{w}} = v) \defeq \sum_{\sigma \in A} \pmf(\sigma)$$
-  and define:
-  $$\condd{\pmf}{X}{\macgv{\mesg{w}} = v}(\sigma) \defeq  \sum_{\sigma' \in A} \condd{\pmf}{X}{\sigma'}(\sigma)$$
+  and also for all $\store' \in \mems(X)$ define:
+  $$\condd{\pmf}{X}{\macgv{\mesg{w}} = v}(\sigma) \defeq  \sum_{\sigma \in A} \condd{\pmf}{X}{\sigma}(\sigma')$$
   where $A$ is:

### 2024-07-16 ppdp24/model.tex
checkpoint

@@
 \begin{definition}
-  Given $\pmf$ with $\dom(\pmf) = \mems(Y)$, the \emph{marginal distribution}
-  of variables $X \subseteq Y$ in $\pmf$, denoted $\margd{\pmf}{X}$, is defined as follows:
+  Given $\pmf$ with $\dom(\pmf) = \mems(X_2)$, the \emph{marginal distribution}
+  of variables $X_1 \subseteq X_2$ in $\pmf$, denoted $\margd{\pmf}{X_1}$, is defined as follows:
   $$
-  \forall \store \in \mems(X) \quad . \quad \margd{\pmf}{X}(\store) =
-  \sum_{\store' \in \mems(Y-X)} \pmf(\store \uplus \store')
+  \forall \store \in \mems(X_1) \quad . \quad \margd{\pmf}{X_1}(\store) =
+  \sum_{\store' \in \mems(X_2-X_1)} \pmf(\store \uplus \store')
   $$
@@
 \begin{definition}
-  Given $\pmf$ with $\dom(\pmf) = \mems(Z)$, the \emph{conditional distribution}
-  of $X$ given $Y$ where $X \cup Y \subseteq Z$ and $X \cap Y = \varnothing$, denoted $\condd{\pmf}{X}{Y}$,
+  Given $\pmf$ with $\dom(\pmf) = \mems(X_3)$, the \emph{conditional distribution}
+  of $X_1$ given $X_2$ where $X_1 \cup X_2 \subseteq X_3$ and $X_1 \cap X_2 = \varnothing$, denoted $\condd{\pmf}{X_1}{X_2}$,
   is defined as follows:
   $$
-  \forall \store \in \mems(X \cup Y)\ .\ 
-  \condd{\pmf}{X}{Y}(\store) =
+  \forall \store \in \mems(X_1 \cup X_2)\ .\ 
+  \condd{\pmf}{X_1}{X_2}(\store) =
   \begin{cases}
-    0 \text{\ if\ } \margd{\pmf}{Y}(\store_Y) = 0\\
-    \margd{\pmf}{X \cup Y}(\store) / \margd{\pmf}{Y}(\store_Y)\ \text{\ otherwise}
+    0 \text{\ if\ } \margd{\pmf}{X_2}(\store_X_2) = 0\\
+    \margd{\pmf}{X_1 \cup X_2}(\store) / \margd{\pmf}{X_2}(\store_X_2)\ \text{\ otherwise}
   \end{cases}
@@
 \begin{definition}
-  Given $\pmf$, for all $\store_2 \in \mems(Y)$ and $\store_1 \in \mems(X)$ define:
+  Given $\pmf$, for all $\store_2 \in \mems(X_2)$ and $\store_1 \in \mems(X_1)$ define:
   \begin{enumerate}
-  \item $\condd{\pmf}{X}{\store_2}(\store_1) \defeq \condd{\pmf}{X}{Y}(\store_1 \uplus \store_2)$
-  \item $\pmf(\store_1)  \defeq \margd{\pmf}{X}(\store_1)$ 
-  \item $\pmf(\store_1|\store_2) \defeq \condd{\pmf}{X}{Y}(\store_1 \uplus \store_2)$
+  \item $\condd{\pmf}{X_1}{\store_2}(\store_1) \defeq \condd{\pmf}{X_1}{X_2}(\store_1 \uplus \store_2)$
+  \item $\pmf(\store_1)  \defeq \margd{\pmf}{X_1}(\store_1)$ 
+  \item $\pmf(\store_1|\store_2) \defeq \condd{\pmf}{X_1}{X_2}(\store_1 \uplus \store_2)$
   \end{enumerate}
@@
     %(\dom(\store) = \iov(\prog) \cup R) \wedge
-    (\store_1 \cap \store_2,\prog) \redxs (\store,\varnothing) \}
+    (\store_1 \uplus \store_2,\prog) \redxs (\store,\varnothing) \}
   \end{array}
@@
     %(\dom(\store) = \iov(\prog) \cup R) \wedge
-    (\store_1 \cap \store_2,\prog) \aredxs (\store,\varnothing) \}
+    (\store_1 \uplus \store_2,\prog) \aredxs (\store,\varnothing) \}
   \end{array}

### 2024-07-16 ppdp24/model.tex
checkpoint

@@
   Given $\pmf$ with $\dom(\pmf) = \mems(X_2)$, the \emph{marginal distribution}
-  of variables $X_1 \subseteq X_2$ in $\pmf$, denoted $\margd{\pmf}{X_1}$, is defined as follows:
+  of variables $X_1 \subseteq X_2$ in $\pmf$ is denoted $\margd{\pmf}{X_1}$ and defined as follows:
   $$
-  \forall \store \in \mems(X_1) \quad . \quad \margd{\pmf}{X_1}(\store) =
+  \forall \store \in \mems(X_1) \quad . \quad \margd{\pmf}{X_1}(\store) \defeq
   \sum_{\store' \in \mems(X_2-X_1)} \pmf(\store \uplus \store')
@@
 \begin{definition}
-  Given $\pmf$ with $\dom(\pmf) = \mems(X_3)$, the \emph{conditional distribution}
-  of $X_1$ given $X_2$ where $X_1 \cup X_2 \subseteq X_3$ and $X_1 \cap X_2 = \varnothing$, denoted $\condd{\pmf}{X_1}{X_2}$,
-  is defined as follows:
+  Given $\pmf$, the \emph{conditional distribution}
+  of $X_1$ given $X_2$ where $X_1 \cup X_2 \subseteq \dom(\pmf)$ and $X_1 \cap X_2 = \varnothing$
+  is denoted $\condd{\pmf}{X_1}{X_2}$ and defined as follows:
   $$
   \forall \store \in \mems(X_1 \cup X_2)\ .\ 
-  \condd{\pmf}{X_1}{X_2}(\store) =
+  \condd{\pmf}{X_1}{X_2}(\store) \defeq
   \begin{cases}
-    0 \text{\ if\ } \margd{\pmf}{X_2}(\store_X_2) = 0\\
-    \margd{\pmf}{X_1 \cup X_2}(\store) / \margd{\pmf}{X_2}(\store_X_2)\ \text{\ otherwise}
+    0 \text{\ if\ } \margd{\pmf}{X_2}(\store_{X_2}) = 0\\
+    \margd{\pmf}{X_1 \cup X_2}(\store) / \margd{\pmf}{X_2}(\store_{X_2})\ \text{\ o.w.}
   \end{cases}
@@
 \begin{definition}
-  Given $\pmf$, for all $\store_2 \in \mems(X_2)$ and $\store_1 \in \mems(X_1)$ define:
+  Given $\pmf$, for all $\store_1 \in \mems(X_1)$ and $\store_2 \in \mems(X_2)$ define:
   \begin{enumerate}

### 2024-07-16 ppdp24/example-bdoz.tex
checkpoint

@@
 A field value $v$ is secret shared among 2 clients in BDOZ in the same
-manner as in GMW.  Each client $\cid$ gets a pair of the form
-$(v_\cid,m_\cid)$, where $v_\cid$ is a share of $v$ reconstructed by
-addition, i.e., $v = \fcod{v_1 \fplus v_2}$, and $m_\cid$ is a MAC of
-$v_\cid$.  More precisely, $m_\cid = k + (k_\Delta * v_\cid)$ where
-\emph{keys} $k$ and $k_\Delta$ are known only to $\cid' \ne \cid$ and
-$k_\Delta$. The \emph{local key} $k$ is unique per MAC, while the
-\emph{global key} $k_\Delta$ is common to all MACs authenticated by
-$\cid'$. This is a semi-homomorphic encryption scheme that supports
-addition of shares and multiplication of shares by a constant-- for
-more details the reader is referred to Orsini
-\cite{10.1007/978-3-030-68869-1_3}. We note that while we restrict
-values $v$, $m$, and $k$ to the same field in this presentation for
-simplicity, in general $m$ and $k$ can be in extensions of
-$\mathbb{F}_p$.
+manner as in GMW, but with the addition of a separate MAC values.  Each
+client $\cid$ gets a pair of values $(v_\cid,m_\cid)$, where $v_\cid$
+is a share of $v$ reconstructed by addition, i.e., $v = \fcod{v_1
+  \fplus v_2}$, and $m_\cid$ is a MAC of $v_\cid$.  More precisely,
+$m_\cid = k + (k_\Delta * v_\cid)$ where \emph{keys} $k$ and
+$k_\Delta$ are known only to $\cid' \ne \cid$ and $k_\Delta$. The
+\emph{local key} $k$ is unique per MAC, while the \emph{global key}
+$k_\Delta$ is common to all MACs authenticated by $\cid'$. This is a
+semi-homomorphic encryption scheme that supports addition of shares
+and multiplication of shares by a constant-- for more details the
+reader is referred to Orsini \cite{10.1007/978-3-030-68869-1_3}. We
+note that while we restrict values $v$, $m$, and $k$ to the same field
+in this presentation for simplicity, in general $m$ and $k$ can be in
+extensions of $\mathbb{F}_p$.
 
-We can capture both the preliminary distribution of Beaver triples and BDOZ shares
-as a pre-processing predicate that establishes conditions for initial
-memories (see Definition \ref{def-aprogtt}).  Here we assume two input
-secrets $\elab{\secret{x}}{1}$ and $\elab{\secret{y}}{2}$ and a single
-Beaver Triple to compute $\elab{\secret{x}}{1} \ftimes
-\elab{\secret{y}}{2}$, but we can extend this for additional gates.
-As for GMW, we use $\macgv{\mesg{w}}$ to refer to secret-shared values
-reconstructed with addition, where by convention shares are message
-values $\elab{\mesg{w}}{\cid}$ for all $\cid$.
+We can capture both the preliminary distribution of Beaver triples and
+BDOZ shares as a pre-processing predicate that establishes conditions
+for initial memories (see Definition \ref{def-aprogtt}).  Here we
+assume two input secrets $\elab{\secret{x}}{1}$ and
+$\elab{\secret{y}}{2}$ and a single Beaver Triple to compute
+$\elab{\secret{x}}{1} \ftimes \elab{\secret{y}}{2}$, but we can extend
+this for additional gates.  As for GMW, we use $\macgv{\mesg{w}}$ to
+refer to secret-shared values reconstructed with addition, but for
+BDOZ by convention each share of $\macgv{\mesg{w}}$ is represented as
+$\elab{\mesg{w\ttt{s}}}{\cid}$, the MAC of which is represented as a
+$\elab{\mesg{w\ttt{m}}}{\cid}$ for all $\cid$, 
+and each client holds a key $\elab{\mesg{w\ttt{k}}}{\cid}$ for
+authentication of the other's share.
 \begin{definition}[BDOZ preprocessing]
@@
     \mathit{shares} \defeq
-    \{ \elab{\mesg{w}}{\cid}\ |\ \cid \in \{ 1, 2 \} \wedge w \in \{ a,b,c,x,y \}  \}
+    \{ \elab{\mesg{w\ttt{s}}}{\cid}\ |\ \cid \in \{ 1, 2 \} \wedge w \in \{ a,b,c,x,y \}  \}
 
-    \mathit{macs} \defeq  \{ \elab{\mesg{w\ttt{mac}}}{\cid}\ |\ \cid \in \{ 1, 2 \} \wedge w \in \{ a,b,c,x,y \}  \}
+    \mathit{macs} \defeq  \{ \elab{\mesg{w\ttt{m}}}{\cid}\ |\ \cid \in \{ 1, 2 \} \wedge w \in \{ a,b,c,x,y \}  \}
 
@@
   and, writing $\store(\macgv{\mesg{w}})$ to denote
-  $\fcod{\store(\elab{\mesg{w}}{1}) + \store(\elab{\mesg{w}}{2})}$,
+  $\fcod{\store(\elab{\mesg{w\ttt{s}}}{1}) + \store(\elab{\mesg{w\ttt{s}}}{2})}$,
   the following conditions hold:
@@
   and for all $\cid,\cid' \in \{1,2\}$ with $\cid \ne \cid'$ and $w \in \{ a,b,c,x,y\}$:
-  $$\lcod{\store, \mesg{w\ttt{mac}}}{\cid} =
-  \lcod{\store, \mesg{wk} + \mesg{\ttt{delta}} * \mesg{w}}{\cid'}$$
+  $$
+  \store(\mx{w\ttt{m}}{\cid}) =
+    \fcod{\store(\mx{w\ttt{k}}{\cid'}) + \store(\mx{\ttt{delta}}{\cid'}) * \store(\mx{w\ttt{s}}{\cid})}
+  $$
+  %$$\lcod{\store, \mesg{w\ttt{m}}}{\cid} =
+  %\fcod{\lcod{\store, \mesg{w\ttt{k}} + \mesg{\ttt{delta}} * \mesg{w\ttt{m}}}{\cid'}$$
 \end{definition}
 
-With these conventions, our BDOZ library is defined in Figure \ref{fig-bdoz}.
-In $\metaprot$ we represent BDOZ share, MAC pairs as records:
-$$
-\ttt{\{ share = }v\ttt{;  mac =}\ m \ttt{\}}
-$$
-and we define $\ttt{macsum}$ for addition of shares,
-$\ttt{maccsum}$ for addition of a share and a constant, and
-$\ttt{macctimes}$ for multiplication of a share and a constant
-in the BDOZ encryption scheme \cite{10.1007/978-3-030-68869-1_3}. The $\ttt{auth}$
-function implements the MAC check as an $\assert$.
+With these conventions, a BDOZ library is defined in Figure
+\ref{fig-bdoz} and a multiplication protocol to compute $\sx{x}{1} *
+\sx{y}{2}$ is defined in Figure \ref{fig-beaver}. The latter is
+malicious secure assuming that client 1 is trusted. Initially, in
+calls to $\ttt{sum\_she}$ each party computes the sum of their shares
+of $\sx{x}{1} + \macgv{\mesg{a}}$ and $\sx{y}{2} + \macgv{\mesg{b}}$,
+and non-interactively computes the associated MACs and keys in the
+SHE authentication scheme. These values are then
+securely opened on client 1, which authenticates client
+2's share via a call to $\ttt{auth}$ that $\ttt{assert}$s
+the BDOZ authentication property. Note that the
+opening of $\sx{y}{2} + \macgv{\mesg{b}}$ reveals no
+information to client 1 about $\sx{y}{2}$ since $\macgv{\mesg{b}}$ is
+in a unform random distribution by assumption. It is
+also malicious secure since cheating is detected
+by the $\ttt{assert}$.
 
-We also define a function $\ttt{secopen}$ to implement ``secure
-opening''.  In this standard subprotocol, the value
-$\macgv{\secret{w_1}} \fplus \macgv{\mesg{w_2}}$ is reconstructed as
-$\mesg{w_3}$, by each client $\cid_2$ communicating
-$\lcod{\mesg{w_1} + \mesg{w_2}}{\cid_2}$ to $\cid_1$.  Assuming that
-$\macgv{\mesg{w_2}}$ is in an independent uniform distribution,
-this perfectly hides $\secret{w_1}$. In a multiplication gate
-either $a$ or $b$ of a Beaver triple are used in secure opening,
-so, e.g., given $
-(\varnothing,\ttt{secopen}(a, x, d, 2, 1) \redxs (\prog,())$
-we have both:
-\begin{mathpar}
-  \conddetx{\progtt(\prog)}{\{\macgv{\mesg{a}},\elab{\secret{x}}{1} \}}{\{ \elab{\mesg{d}}{2} \}}
-
-  \sep{\progtt(\prog)}{\{ \elab{\mesg{d}}{2} \}}{\{ \elab{\secret{x}}{1}  \}}
-\end{mathpar}
-Furthermore, client 2's authentication of the sum of shares with the
-sum of their keys detects any attempted cheating by $1$.  Finally, the
-$\ttt{secreveal}$ function is the same as $\decodegmw$ but with the
-addition of authentication of revealed shares to ensure malicious
-security.
+Following this, each party non-interactively calculates their share of
+$\sx{x}{1} \ftimes \sx{y}{2}$, while client 2 calculates the MAC of
+their share, and reveals both their share and MAC to client 1. These
+are reveals since information about $\sx{y}{2}$ is exposed by
+the multiplication.  Client 1 also calculates the appropriate key and
+authenticates client 2's share following the BDOZ SHE MAC scheme
+\cite{10.1007/978-3-030-68869-1_3} and the standard Beaver protocol
+\cite{evans2018pragmatic}. Hence the protocol is correct and
+malicious secure given trusted client 1. The authentication scheme
+can also be run symmetrically on client 2.
 
 \input{beaver}
-
-The full protocol for malicious secure product of secrets $x$ (that
-is, $\elab{\secret{x}}{1}$) and $y$ (that is, $\elab{\secret{y}}{2}$)
-using Beaver triple $a,b,c$ is defined in Figure \ref{fig-bdoz}. Both
-parties interact in secure opening of $x \fplus a$ and $y + b$,
-followed by the non-interactive computation of shares of $x * y$ for
-secure reveals as per standard protocol
-\cite{10.1007/978-3-030-68869-1_3}. The non-interactive reconstruction
-of the local authentication key for both the secure openings and the
-final reveal is enabled by the semi-homomorphic properties of the BDOZ
-scheme.
 

### 2024-07-17 ppdp24/intro.tex
checkpoint

@@
 formulate a collection of confidentiality and integrity
-hyperproperties for our model model with familiar information flow
+hyperproperties for our model with familiar information flow
 analogs, and to show how these properties can be leveraged for
@@
 also been developed. The $\lambda_{\mathrm{obliv}}$ language
-\cite{darais2019language} uses a type system is used to automatically
+\cite{darais2019language} uses a type system to automatically
 enforce so-called probabilistic trace obliviousness.  But similar to

### 2024-07-17 ppdp24/lang.tex
checkpoint

@@
 provide random tape and secret inputs to protocols, and also record
-message sends, public broadcast, and client outputs. Memories $\store$ are finite
-(partial) mapping from variables $x$ to values $v \in \mathbb{F}_p$.
-The \emph{domain} of a memory is written $\dom(\store)$ and is
-the finite set of variables on which the memory is defined.
-We write $\store\{ x \mapsto v\}$ for $x\not\in\dom(\store)$
-to denote the memory $\store'$ such that
-$\store'(x) = v$ and otherwise $\store'(y) = \store(y)$ for all $y
-\in \dom(\store)$. We write $\store \subseteq \store'$ iff
-$\dom(\store) \subseteq \dom(\store')$ and $\store(x) =
-\store'(x)$ for all $x \in \dom(\store)$. We write $\store \uplus
-\store'$ to denote the combination of $\store$ and $\store'$
-assuming $\store(x) = \store'(x)$ for all $x \in \dom(\store)
-\cap \dom(\store')$, otherwise $\store \uplus \store'$ is undefined.
-\cnote{Check to make sure all $\cap$ and $\cup$s have been revised
-with $\uplus$.}
-We write $\store \subseteq \store'$ iff $\store \uplus \store'
-= \store$.
+message sends, public broadcast, and client outputs. Memories $\store$
+are finite (partial) mappings from variables $x$ to values $v \in
+\mathbb{F}_p$.  The \emph{domain} of a memory is written
+$\dom(\store)$ and is the finite set of variables on which the memory
+is defined.  We write $\store\{ x \mapsto v\}$ for
+$x\not\in\dom(\store)$ to denote the memory $\store'$ such that
+$\store'(x) = v$ and otherwise $\store'(y) = \store(y)$ for all $y \in
+\dom(\store)$. We write $\store \subseteq \store'$ iff $\dom(\store)
+\subseteq \dom(\store')$ and $\store(x) = \store'(x)$ for all $x \in
+\dom(\store)$. We write $\store \uplus \store'$ to denote the
+combination of $\store$ and $\store'$ assuming $\store(x) =
+\store'(x)$ for all $x \in \dom(\store) \cap \dom(\store')$, otherwise
+$\store \uplus \store'$ is undefined.  \cnote{Check to make sure all
+  $\cap$ and $\cup$s have been revised with $\uplus$.}  We write
+$\store \subseteq \store'$ iff $\store \uplus \store' = \store$.
 

### 2024-07-17 ppdp24/model.tex
checkpoint

@@
 setting, Given a protocol $\prog$ that implements $\idealf$, with
-$\iov(\prog) = S \cup V \cup O$, the domain of $\idealf$ is $\mems(S)$
-and its range is $\mems(O)$.  Real/ideal security in the MPC
+$\iov(\prog) = S \cup M \cup P \cup O$, the domain of $\idealf$
+is $\mems(S)$ and its range is $\mems(O)$.  Real/ideal security in the MPC
 setting means that, given $\store \in \mems(S)$, a secure protocol
@@
   is \emph{passive secure in the simulator model} iff there exists
-  ad simulation $\SIM$ such that for all
+  a simulation $\SIM$ such that for all
   partitions of the federation into honest and corrupt sets $H$ and $C$
-  and for all $\store \in \mems(S)$:
+  and for all $\store \in \mems(\secrets(\prog))$:
   $$
@@
 
-In the malicious model we assume that corrupt clients are in the
-thrall of an adversary $\adversary$ who does not necessarily follow
+In the malicious model we assume that corrupt clients are in
+thrall to an adversary $\adversary$ who does not necessarily follow
 the rules of the protocol.  We model this by positing a $\arewrite$
@@
 the possibility of abort needs to be reflected in adversarial
-runs of a protocol. We can define $\aruns(\prog,\adversary)$
+runs of a protocol. We can define $\aruns(\prog)$
 as the ``prefix'' memories that result from possibly-aborting
 protocols, but we also need to ``pad out'' the memories
-of partial runs with $\bot$, as we define in $\botruns(\prog,\adversary)$,
+of partial runs with $\bot$, as we define in $\botruns(\prog)$,
 to properly reflect the contents of views and outputs even in case of abort. 

### 2024-07-17 ppdp24/bruteforce.tex
checkpoint

@@
 \end{lemma}
-However, this method does not take advantage of parallellization,
+However, this method does not take advantage of parallelization,
 in that elements of $\runs(\prog)$ can be calculated independently.
@@
 \end{mathpar}
-... and so on. Then, to convert a protocol $\prot$ to a Datalog
+... and so on. Then, to convert a protocol $\prog$ to a Datalog
 we first define the function $\bodies$ that applies $\solvealg$ locally to each command

### 2024-07-17 ppdp24/example-bdoz.tex
checkpoint

@@
 \begin{definition}[Cheating Detection]
-  \emph{Cheating is detected} in $\prog$ with $\iov(\prog) = S \cup V \cup O$ iff
+  \emph{Cheating is detectable} in $\prog$ with $\iov(\prog) = S \cup V \cup O$ iff
   for all  $\store \in \aruns(\prog)$,
@@
   \label{lemma-cheating}
-  If cheating is detected in $\prog$, then $\prog$ has integrity.
+  If cheating is detectable in $\prog$, then $\prog$ has integrity.
 \end{lemma}

### 2024-07-17 ppdp24/example-gmw.tex
checkpoint

@@
 \begin{verbatimtab}
-         let s1 = encodegmw("s1") in
-         let s2 = encodegmw("s2") in
+         let s1 = encodegmw("s1",1,2) in
+         let s2 = encodegmw("s2",2,1) in
          decodegmw(andgmw("z",s1,s2))
@@
 \begin{definition}[Conditioning Properties] 
-  Give $\pmf$ and $X_1,X_1,X_3 \subseteq \dom{pmf}$, 
+  Given $\pmf$ and $X_1,X_2,X_3 \subseteq \dom(\pmf)$, 
   we write:
@@
   \label{lemma-conditioning}
-  Give $\pmf$, each of the following hold for $S,V_1,V_2,V_3 \in \dom(\pmf)$:
+  Given $\pmf$, each of the following hold for $S,V_1,V_2,V_3 \in \dom(\pmf)$:
   \begin{enumerate}
     \item Given $\condp{\pmf}{S}{\detx{V_1}}$ and
-      $\condp{\pmf}{V_2}{\detx{V_2}}$, then $\condp{\pmf}{S}{\detx{V_2}}$.
+      $\condp{\pmf}{V_1}{\detx{V_2}}$, then $\condp{\pmf}{S}{\detx{V_2}}$.
     \item Given $\condp{\pmf}{S}{\detx{V_1}}$ and

### 2024-07-17 ppdp24/example-gmw.tex
checkpoint

@@
 For example, the following program uses our GMW library to define
-a circuit with a single \eand gate and input secrets $\ttt{s1}$ and
+a circuit with a single \eand\ gate and input secrets $\ttt{s1}$ and
 $\ttt{s2}$ from client's 1 and 2 respectively:

### 2024-07-17 ppdp24/intro.tex
checkpoint

@@
 \item A \textbf{low-level language} for defining MPC primitives
-  (Section \ref{section-lang}) with an associated metalanguage
-  to ease programming (Section \ref{section-metalang}).
+  (Section \ref{section-lang}) with an associated \textbf{metalanguage}
+  to ease programming (Section \ref{section-metalang})\footnote{By metalanguage
+  we mean a multi-stage aka metaprogramming language where code is a value, as
+  in, e.g., MetaML \cite{TAHA2000211}.}. 
 \item A \textbf{fully-automated verification method} for low-level MPC

### 2024-07-17 ppdp24/intro.tex
checkpoint

@@
 \ref{section-bruteforce} we develop a method for automatically
-computing the pmf of $\minifed$ protocols in $\mathbb{F}_2$, that can
-be automatically queried to enforce hyperproperties of
-security. This method is perfectly accurate but has high complexity;
-we show this can be partially mitigated by conversion of protocols in
-$\mathbb{F}_2$ to stratified Datalog which is amenable to HPC
-acceleration. Furthermore, in Section \ref{section-example-gmw} we
-consider in detail how this automated technique can be used as a local
-automated tactic for proving security in arbitrarily large GMW
+computing the probability mass function (pmf) of $\minifed$ protocols
+in $\mathbb{F}_2$, that can be automatically queried to enforce
+hyperproperties of security. This method is perfectly accurate but has
+high complexity; we show this can be partially mitigated by conversion
+of protocols in $\mathbb{F}_2$ to stratified Datalog which is amenable
+to HPC acceleration. Furthermore, in Section \ref{section-example-gmw}
+we consider in detail how this automated technique can be used as a
+local automated tactic for proving security in arbitrarily large GMW
 circuits using conditional probabilistic independence as in
-\cite{li2023lilac} (Lemmas \ref{lemma-gmwtactic} and \ref{lemma-gmwinvariant}
-and Theorem \ref{theorem-gmw}).
+\cite{li2023lilac} (Lemmas \ref{lemma-gmwtactic} and
+\ref{lemma-gmwinvariant} and Theorem \ref{theorem-gmw}).
 

### 2024-07-17 ppdp24/model.tex
checkpoint

@@
 \begin{definition}[Corrupt and Honest Views]
-  Given a program $\prog$ with $\iov(\prog) = S \cup M \cup P \cup O$,
-  define $\views(\prog) \defeq M \cup P$, and define $\houtputs$ as
-  the messages and reveals in $V \defeq M \cup P$ sent from honest to corrupt
+  We let $V$ range over \emph{views} which are sets of messages
+  and reveals. Given a program $\prog$ with $\iov(\prog) = S \cup M \cup P \cup O$,
+  define $\views(\prog) \defeq V \defeq M \cup P$, and define $\houtputs$ as
+  the messages and reveals in $V$ sent from honest to corrupt
   parties, called \emph{corrupt views}:

### 2024-07-17 ppdp24/model.tex
checkpoint

@@
   and reveals. Given a program $\prog$ with $\iov(\prog) = S \cup M \cup P \cup O$,
-  define $\views(\prog) \defeq V \defeq M \cup P$, and define $\houtputs$ as
-  the messages and reveals in $V$ sent from honest to corrupt
+  define $\views(\prog) \defeq M \cup P$, and define $\houtputs$ as
+  the messages and reveals in $V = M \cup P$ sent from honest to corrupt
   parties, called \emph{corrupt views}:

### 2024-07-17 ppdp24/hyper.tex
checkpoint

@@
   Given a functionality $\idealf$ and outputs $\store_{\mathit{out}}$, their 
-  \emph{kernel}, denoted $\ik(\idealf,\store_{\mathit{out}})$ is
+  \emph{kernel}, denoted $\kernel{\idealf}{\store_{\mathit{out}}}$ is
   $
@@
   $.
-  We say that $\idealf$ is \emph{pre-imageable} iff $\ik(\idealf, \store_{\mathit{out}})$ for all
+  We say that $\idealf$ is \emph{pre-imageable} iff $\kernel{\idealf}{\store_{\mathit{out}}}$ for all
   $\store_{\mathit{out}}$ can be computed tractably.
@@
 %  \pmf(\{ x \mapsto 1\}\ |\ \{ y \mapsto 1 \}) = 1$.
-  We write $\sep{\pmf}{X}{Y}$ iff for all
-    $\store \in \mems(X \cup Y)$ we have
-  $\margd{\pmf}{X \cup Y}(\store) =
-  \pmf(\store_X) * \pmf(\store_Y)$
+  We write $\sep{\pmf}{X_1}{X_2}$ iff for all
+    $\store \in \mems(X_1 \cup X_2)$ we have
+  $\margd{\pmf}{X_1 \cup X_2}(\store) =
+  \pmf(\store_{X_1}) * \pmf(\store_{X_2})$
 \end{definition}
@@
   where $\iov(\prog) = S \cup V \cup O$, and let $\adversary$ be some
-  adversary. Suppose $\store \in \aruns(\prog,\adversary)$.
+  adversary. Suppose $\store \in \aruns(\prog)$.
   As integrity requires, and as Lemma \ref{lemma-malicious-correct}
@@
   with inputs $\store'_C$ and arbitrary $\store'' \in \mems(S_H)$
-  such that $\store'_C \uplus \store'' \in \ik(\idealf,\idealf(\store'))$.
+  such that $\store'_C \uplus \store'' \in \kernel{\idealf}{\idealf(\store')}$.
   The assumption of passive security of $\prog$ implies the result.

### 2024-07-17 ppdp24/metalang.tex
checkpoint

@@
 for defining composable elements. So in this Section we introduce the
-$\metaprot$ language which that includes structured data and function
+$\metaprot$ language that includes structured data and function
 definitions for defining composable protocol elements at a higher

### 2024-07-17 ppdp24/model.tex
checkpoint

@@
     simulator some $\store' \in \mems(S_C)$, and the simulator consults an
-    oracle to compute $\idealf(\store \cup \store') \in \mems(O)$.
+    oracle to compute $\idealf(\store \uplus \store') \in \mems(O)$.
   \item In the second phase $\SIM_2$, the simulator is given the corrupt
-    outputs $\idealf(\store \cup \store')_C$, which are again given to
+    outputs $\idealf(\store \uplus \store')_C$, which are again given to
     $\adversary$, who decides either to abort or not. If so, then the
@@
     and arbitrary internal state $\Sigma$.
-    Otherwise the simulator is given $\sigma_{\mathit{out}} \defeq \idealf(\store \cup \store')_H$
+    Otherwise the simulator is given $\sigma_{\mathit{out}} \defeq \idealf(\store \uplus \store')_H$
     and $\Sigma$.
@@
     finally outputs
-    $\store_{\mathit{out}} \cup \store_{\mathit{views}}$ for some
+    $\store_{\mathit{out}} \uplus \store_{\mathit{views}}$ for some
     calculated $\store_{\mathit{views}} \in \mems(\houtputs)$.

### 2024-07-17 ppdp24/bruteforce.tex
checkpoint

@@
 
-In Figure \ref{fig-solve} we define the algorithm
-$\solve{\stores}{\be}{\cid}$ which filters a given $\stores$ to obtain
-the subset whose elements satisfy $\be$. In this Figure and elsewhere
-we use logical connectives as field operations ($\eand$ and $\exor$
-for $*$ and $+$ respectively) in $\mathbb{F}_2$ and add $\enot$ and
-$\eor$ as trivial but convenient extensions. The correctness of this
-operation is characterized as follows.
+Letting $\stores$ denote sets of memories, in Figure \ref{fig-solve}
+we define the algorithm $\solve{\stores}{\be}{\cid}$ which filters a
+given $\stores$ to obtain the subset whose elements satisfy $\be$. In
+this Figure and elsewhere we use logical connectives as field
+operations ($\eand$ and $\exor$ for $*$ and $+$ respectively) in
+$\mathbb{F}_2$ and add $\enot$ and $\eor$ as trivial but convenient
+extensions. The correctness of this operation is characterized as
+follows.
 \begin{lemma}

### 2024-07-18 ppdp24/example-gmw.tex
checkpoint

@@
 \begin{verbatimtab}
-         let s1 = encodegmw("s1",1,2) in
-         let s2 = encodegmw("s2",2,1) in
+         let s1 = encodegmw("s1",2,1) in
+         let s2 = encodegmw("s2",1,2) in
          decodegmw(andgmw("z",s1,s2))
@@
 By convention we will assume that all gates are assigned unique output
-identifiers $\ttt{"z"}$, and that all programs are in the form
+identifiers, and that all programs are in the form
 of a sequence of let-bindings followed by a call to $\decodegmw$

### 2024-07-18 ppdp24/gmw.tex
checkpoint

@@
-\begin{fpfig}[t]{2-Party GMW circuit library.}{fig-gmw}
+\begin{fpfig}[t]{2-party GMW circuit library.}{fig-gmw}
 {\footnotesize
@@
       m[in]@i2 := (s[in] xor r[in])@i2;
-      m[in]@i1 := r[in]@2;
+      m[in]@i1 := r[in]@i2;
       m[in]
@@
 
-    xorgmw(z, x, y) { m[z]@1 := (x+y)@1; m[z]@2 := (x+y)@2; m[z] }
+    xorgmw(z, x, y)
+    { m[z]@1 := (x xor y)@1; m[z]@2 := (x xor y)@2; m[z] }
 

### 2024-07-18 revision-2/eqs.py
checkpoint

@@ -21,14 +21,46 @@ s.add(o == p1 + p2)
 s.check()
 
 # correctness (should be unsat)
-s.add(x + y != o)
-s.check()
-
+s.check(x + y != o)
 
 # 3-party addition in F7
 
 s.resetAssertions()
 
-m1, f11, x = FiniteFieldElems('m1 f11 x', 7)
-m2, f21, y = FiniteFieldElems('m1 f21 y', 7)
-p1, p2, o = FiniteFieldElems('p1 p2 o', 7)
+m21, m31, s1, rl1, r1 = FiniteFieldElems('m21 m31 s1 rl1 r1', 7)
+m12, m32, s2, rl2, r2 = FiniteFieldElems('m12 m32 s2 rl2 r2', 7)
+m13, m23, s3, rl3, r3 = FiniteFieldElems('m13 m23 s3 rl3 r3', 7)
+p1, p2, p3, o = FiniteFieldElems('p1 p2 p3 o', 7)
+
+s.add(m12 == s1 - rl1 - r1)
+s.add(m13 == r1)
+s.add(m21 == s2 - rl2 - r2)
+s.add(m23 == r2)
+s.add(m31 == s3 - rl3 - r3)
+s.add(m32 == r3)
+s.add(p1 == rl1 + m21 + m31)
+s.add(p2 == rl2 + m12 + m32)
+s.add(p3 == rl3 + m13 + m23)
+s.add(o == p1 + p2 + p3)
+
+# sanity check (should be sat)
+s.check()
+
+# correctness (should be unsat)
+s.check(o != s1 + s2 + s3)
+
+# BDOZ/SPDZ SHE scheme 
+
+s.resetAssertions()
+
+mac1,mac2,m1,m2,k1,k2,d,a = FiniteFieldElems('mac1 mac2 m1 m2 k1 k2 d a', 7)
+
+s.add(mac1 == k1 + (d * m1))
+s.add(mac2 == k2 + (d * m2))
+
+# HE for sum of shares,(should be unsat)
+s.check(mac1 + mac2 != (k1 + k2) + (d * (m1 + m2)))
+
+# HE for multiplication of share and public value a (should be unsat)
+s.check(mac1 * a != (k1 * a) + (d * (m1 * a)))
+

### 2024-07-18 revision-2/eqs.py
checkpoint

@@ -21,7 +21,7 @@ s.add(o == p1 + p2)
 s.check()
 
 # correctness (should be unsat)
-s.check(x + y != o)
+s.check(x + y != o)  
 
 # 3-party addition in F7
 

### 2024-07-21 ppdp24/lang.tex
checkpoint

@@
 provide random tape and secret inputs to protocols, and also record
-message sends, public broadcast, and client outputs. Memories $\store$
+message sends, public broadcast, and client outputs.
+\begin{definition}
+  Memories $\store$
 are finite (partial) mappings from variables $x$ to values $v \in
@@
 \subseteq \dom(\store')$ and $\store(x) = \store'(x)$ for all $x \in
-\dom(\store)$. We write $\store \uplus \store'$ to denote the
-combination of $\store$ and $\store'$ assuming $\store(x) =
-\store'(x)$ for all $x \in \dom(\store) \cap \dom(\store')$, otherwise
-$\store \uplus \store'$ is undefined.  \cnote{Check to make sure all
-  $\cap$ and $\cup$s have been revised with $\uplus$.}  We write
-$\store \subseteq \store'$ iff $\store \uplus \store' = \store$.
-
-Given a set of variables $X$, we write $\store_X$ to denote the
-memory $\store$ restricted to the domain $X$, and we define
-$\mems(X)$ as the set of all memories with domain $X$:
+\dom(\store)$. Given any $\store$ and $\store'$ with
+$\store(x) = \store'(x)$ for all $x \in \dom(\store) \cap \dom(\store')$,
+we write $\store \uplus \store'$ to denote the memory
+with domain $X = \dom(\store) \cup \dom(\store')$ such
+that:
+$$
+\forall x \in X .
+(\store \uplus \store')(x) =
+\begin{cases} \store(x) \text{\ if\ } x\in\dom(\store) \\ \store'(x) \text{\ otherwise\ }\end{cases} 
+$$
+%We write $\store \subseteq \store'$ iff $\store \uplus \store' = \store$.
+\end{definition}
+In our subsequent presentation we will often want to consider arbitrary
+memories that range over particular variables and to restrict
+memories to particular subsets of their domain:
+\begin{definition}
+  Given a set of variables $X$ and memory $\store$, we write
+  $\store_X$ to denote the memory with $\dom(\store_X) = X$ and
+  $\store_X(x) = \store(x)$ for all $x \in X$. We define $\mems(X)$ as
+  the set of all memories with domain $X$:
   $$
@@
   $$
-Thus, given a protocol $\prog$, the set of all random tapes for
-$\prog$ is $\mems(\flips(\prog))$.
+\end{definition}
+So for example, given a protocol $\prog$, the set of all random tapes for
+$\prog$ is $\mems(\flips(\prog))$, and the memory $\store_{\secrets(\prog)}$
+is $\store$ restricted to the secrets in $\prog$.
 %We let $\stores$ range

### 2024-07-21 ppdp24/model.tex
checkpoint

@@
     simulator is given $\sigma_{\mathit{out}} \defeq \{ \out{\cid} \mapsto \bot \mid \cid \in H \}$
-    and arbitrary internal state $\Sigma$.
+    and arbitrary internal state $\varsigma$.
     Otherwise the simulator is given $\sigma_{\mathit{out}} \defeq \idealf(\store \uplus \store')_H$
-    and $\Sigma$.
-  \item In the third phase $\SIM_3$, given $\store_{\mathit{out}}$ and $\Sigma$, the simulator
+    and $\varsigma$.
+  \item In the third phase $\SIM_3$, given $\store_{\mathit{out}}$ and $\varsigma$, the simulator
     finally outputs

### 2024-07-21 ppdp24/lang.tex
checkpoint

@@
 \eassign{\mesg{z}}{2}{(\secret{x} - \flip{y})}{1}
-$$
-Both messages $\mesg{w}$ and reveals $\rvl{w}$ can be
-referenced in expressions once they've been defined.
-This distinction between messages and broadcast public
-reveal is consistent with previous formulations, e.g.,
-\cite{6266151}.
-
+$$ Both messages $\mesg{w}$ and reveals $\rvl{w}$ can be referenced in
+expressions once they've been defined.  This distinction between
+messages and broadcast public reveal is consistent with previous
+formulations, e.g., \cite{6266151}. To identify and distinguish
+between collections of variables in protocols we introduce the
+following notation.
+\begin{definition}
 We let $x$ range over \emph{variables} which are identifiers where
@@
 $\rvl{w}$, and $O$ ranges over sets of output variables $\out{\cid}$.
-Given a program $\prog$, we write $\iov(\prog)$ to denote the set $S
-\cup M \cup P \cup O$ of variables in $\prog$ with ownership made
-explicit and $\secrets(\prog)$ to denote $S$, and we write
-$\flips(\prog)$ to denote the set $R$ of random samplings in $\prog$
-with ownership made explicit. We write $\vars(\prog)$ to denote
-$\iov(\prog) \cup \flips(\prog)$. For any set of variables $X$ and
-clients $I$, we write $X_I$ to denote the subset of $X$ owned by any
-client $\cid \in I$, in particular we write $X_H$ and $X_C$ to denote the
-subsets belonging to honest and corrupt parties, respectively.
+Given a program $\prog$, we write $\iov(\prog)$ to denote the
+particular set $S \cup M \cup P \cup O$ of variables in $\prog$ and
+$\secrets(\prog)$ to denote $S$, and we write $\flips(\prog)$ to
+denote the particular set $R$ of random samplings in $\prog$. We write
+$\vars(\prog)$ to denote $\iov(\prog) \cup \flips(\prog)$. For any set
+of variables $X$ and clients $I$, we write $X_I$ to denote the subset
+of $X$ owned by any client $\cid \in I$, in particular we write $X_H$
+and $X_C$ to denote the subsets belonging to honest and corrupt
+parties, respectively.
+\end{definition}
 

### 2024-07-21 ppdp24/example-bdoz.tex
checkpoint

@@
 information to client 1 about $\sx{y}{2}$ since $\macgv{\mesg{b}}$ is
-in a unform random distribution by assumption. It is
-also malicious secure since cheating is detected
+in a uniform random distribution by assumption. It is
+also malicious secure since any cheating is detected
 by the $\ttt{assert}$.

### 2024-07-22 code/SMT/eqs.py
checkpoint

@@ -1,0 +1,66 @@
+from cvc5.pythonic import *
+
+s = Solver()
+
+# 2-party addition in F7
+
+s.resetAssertions()
+
+m1, f11, x = FiniteFieldElems('m1 f11 x', 7)
+m2, f21, y = FiniteFieldElems('m1 f21 y', 7)
+p1, p2, o = FiniteFieldElems('p1 p2 o', 7)
+
+# constraints from the protocol
+s.add(m2 == x - f11)
+s.add(m1 == y - f21)
+s.add(p1 == m1 + f11)
+s.add(p2 == m2 + f21)
+s.add(o == p1 + p2)
+
+# sanity check (should be sat)
+s.check()
+
+# correctness (should be unsat)
+s.check(x + y != o)  
+
+# 3-party addition in F7
+
+s.resetAssertions()
+
+m21, m31, s1, rl1, r1 = FiniteFieldElems('m21 m31 s1 rl1 r1', 7)
+m12, m32, s2, rl2, r2 = FiniteFieldElems('m12 m32 s2 rl2 r2', 7)
+m13, m23, s3, rl3, r3 = FiniteFieldElems('m13 m23 s3 rl3 r3', 7)
+p1, p2, p3, o = FiniteFieldElems('p1 p2 p3 o', 7)
+
+s.add(m12 == s1 - rl1 - r1)
+s.add(m13 == r1)
+s.add(m21 == s2 - rl2 - r2)
+s.add(m23 == r2)
+s.add(m31 == s3 - rl3 - r3)
+s.add(m32 == r3)
+s.add(p1 == rl1 + m21 + m31)
+s.add(p2 == rl2 + m12 + m32)
+s.add(p3 == rl3 + m13 + m23)
+s.add(o == p1 + p2 + p3)
+
+# sanity check (should be sat)
+s.check()
+
+# correctness (should be unsat)
+s.check(o != s1 + s2 + s3)
+
+# BDOZ/SPDZ SHE scheme 
+
+s.resetAssertions()
+
+mac1,mac2,m1,m2,k1,k2,d,a = FiniteFieldElems('mac1 mac2 m1 m2 k1 k2 d a', 7)
+
+s.add(mac1 == k1 + (d * m1))
+s.add(mac2 == k2 + (d * m2))
+
+# HE for sum of shares,(should be unsat)
+s.check(mac1 + mac2 != (k1 + k2) + (d * (m1 + m2)))
+
[…]

### 2024-08-18 csf25/lang.tex
checkpoint

\section{$\minicat$ Syntax and Semantics}

$$
    \begin{array}{rcl@{\hspace{2mm}}r}
      \multicolumn{4}{l}{v \in \mathbb{F}_p,\ w \in \mathrm{String},\ \cid \in \mathrm{Clients} \subset  \mathbb{N} }\\[2mm] %, \bop \in \{ \eand, \eor, \exor \}} \\[2mm]
      \be &::=& \flip{w} \mid \secret{w} \mid \mesg{w} \mid \rvl{w} \mid & \textit{expressions}\\
      & & v \mid \be \fminus \be \mid \be \fplus \be \mid \be \ftimes \be \\[2mm]
      x &::=& \elab{\flip{w}}{\cid} \mid \elab{\secret{w}}{\cid} \mid \elab{\mesg{w}}{\cid} \mid \rvl{w} \mid \out{\cid} & \textit{variables} \\[2mm]
      %& &  \select{\be}{\be}{\be} \mid \ctxt{v}{k} \mid \key{w} \mid \sk{\be}(\be) \mid \pk{\be}{\be}(\be) \mid \pk{\be}{\be} \\[2mm]
      %& &  \select{\fp(\be)}{\be}{\be} \ctxt{v,\be}{k}  \mid \sk{\be}(\be) \mid \pk{\be}{\be}(\be) \mid \pk{\be}{\be} \\[2mm]
      \prog &::=& \eassign{\mesg{w}}{\cid}{\be}{\cid} \mid
      \reveal{w}{e}{\cid} \mid \pubout{\cid}{\be}{\cid} \mid \prog;\prog & \textit{protocols} 
    \end{array}
$$

\bigskip

 $$
  %\begin{array}{c@{\hspace{5mm}}c}
  \begin{array}{rcl}
    \lcod{\store, v}{\cid} &=& v\\
    \lcod{\store, \be_1 \fplus \be_2}{\cid} &=& \fcod{\lcod{\store, \be_1}{\cid} \fplus \lcod{\store, \be_2}{\cid}}\\ 
    \lcod{\store, \be_1 \fminus \be_2}{\cid} &=& \fcod{\lcod{\store, \be_1}{\cid} \fminus \lcod{\store, \be_2}{\cid}}\\ 
    \lcod{\store, \be_1 \ftimes \be_2}{\cid} &=& \fcod{\lcod{\store, \be_1}{\cid} \ftimes \lcod{\store, \be_2}{\cid}}\\
  %\end{array} 
  %\begin{array}{rcl}
    \lcod{\store, \flip{w}}{\cid} &=& \store(\elab{\flip{w}}{\cid})\\
    \lcod{\store, \secret{w}}{\cid} &=& \store(\elab{\secret{w}}{\cid})\\
    \lcod{\store, \mesg{w}}{\cid} &=& \store(\elab{\mesg{w}}{\cid})\\
    \lcod{\store, \rvl{w}}{\cid} &=& \store(\rvl{w})\\
    %\lcod{\store, f\,\be_1\,\cdots\, \be_n}{\cid} &=& \delta(f,\lcod{\store, \be_1}{\cid},\ldots,\lcod{\store,\be_n}{\cid})
  \end{array}
  %\end{array}
  $$

\bigskip

\begin{mathpar}
  (\store, \xassign{x}{\be}{\cid}) \redx \extend{\store}{x}{\lcod{\store,\be}{\cid}}

  \inferrule
      {(\store_1,\prog_1) \redx \store_2 \\ (\store_2,\prog_2) \redx \store_3 }
      {(\store_1,\prog_1;\prog_2) \redx \store_3}
      %(\store, \eassign{\mesg{w}}{\cid_1}{\be}{\cid_2};\prog) \redx (\extend{\store}{\mesg{w}_{\cid_1}}{\lcod{\store,\be}{\cid_2}}, \prog)    
      %(\store, \reveal{w}{\be}{\cid};\prog) \redx (\extend{\store}{\rvl{w}}{\lcod{\store,\be}{\cid}}, \prog)   
      %(\store, \pubout{\cid}{\be}{\cid};\prog) \redx (\extend{\store}{\out{\cid}}{\lcod{\store,\be}{\cid}}, \prog)
\end{mathpar}

\subsection{$\minicat$ Adversarial Semantics}

$$
    \begin{array}{rcl@{\hspace{2mm}}r}
      \prog &::=& \cdots \mid \assert{\be = \be}
    \end{array}
$$

$$
\begin{array}{rclr}
  (\store, \xassign{x}{\be}{\cid}) &\aredx&
  \extend{\store}{x}{\lcod{\store,\be}{\cid}} & \cid \in H\\
  (\store, \xassign{x}{\be}{\cid}) &\aredx&
  \extend{\store}{x}{\lcod{\arewrite(\store_C,\be)}{\cid}} & \cid \in C
\end{array}
$$

$$
\begin{array}{rcl@{\qquad}r}
  (\store,\elab{\assert{\be_1 = \be_2}}{\cid}) &\aredx& \store & \text{if\ }
  \lcod{\store,\be_1}{\cid} = \lcod{\store,\be_2}{\cid}  \text{\ or\ } \cid \in C\\
  (\store,\elab{\assert{\be_1 = \be_2}}{\cid}) &\aredx& \abort & \text{if\ }
  \lcod{\store,\be_1}{\cid} \ne \lcod{\store,\be_2}{\cid}  \text{\ or\ } \cid \in C
\end{array}
$$

\begin{mathpar}
  \inferrule
      {(\store_1,\prog_1) \aredx \store_2 \\ (\store_2,\prog_2) \aredx \store_3 }
      {(\store_1,\prog_1;\prog_2) \aredx \store_3}

  \inferrule
      {(\store_1,\prog_1) \aredx \abort}
      {(\store_1,\prog_1;\prog_2) \aredx \abort}

  \inferrule
      {(\store_1,\prog_1) \aredx \store_2 \\ (\store_2,\prog_2) \aredx \abort }
      {(\store_1,\prog_1;\prog_2) \aredx \abort}
\end{mathpar}

### 2024-08-18 csf25/smt.tex
checkpoint

\section{$\minicat$ Constraint Verification}

\subsection{Constraint Satisfiability Modulo Finite Fields}

$$
\begin{array}{rcl}
  \phi &::=& x \mid \phi \fplus \phi \mid \phi \fminus \phi \mid \phi \ftimes \phi \\
  \eqs &::=& \phi \eop \phi \mid \eqs \wedge \eqs 
\end{array}
$$

We write $\eqs_1 \models \eqs_2$ iff every model of $E_1$ is a model of $E_2$. Note that
this relation is reflexive and transitive.

\begin{mathpar}
  \toeq{x} = x

  \toeq{\elab{\be_1 \fplus \be_2}{\cid}} = \toeq{\elab{\be_2}{\cid}} \fplus \toeq{\elab{\be_1}{\cid}}

  \toeq{\elab{\be_1 \fminus \be_2}{\cid}} = \toeq{\elab{\be_2}{\cid}} \fminus \toeq{\elab{\be_1}{\cid}}

  \toeq{\elab{\be_1 \ftimes \be_2}{\cid}} = \toeq{\elab{\be_2}{\cid}} \ftimes \toeq{\elab{\be_1}{\cid}}
\end{mathpar}

\begin{mathpar}
  \toeq{\OT{\be_1}{\cid_1}{\be_2}{\be_3}{\cid_2}} =
  (\toeq{\elab{\be_1}{\cid_1}} \wedge \toeq{\elab{\be_3}{\cid_2}}) \vee
  (\neg\toeq{\elab{\be_1}{\cid_1}} \wedge \toeq{\elab{\be_2}{\cid_2}}) 
\end{mathpar}

\begin{mathpar}
  \toeq{\xassign{x}{\be}{\cid}} = x \eop \toeq{\elab{\be}{\cid}}

  \toeq{\assert{\be_1 = \be_2}{\cid}} =  \toeq{\elab{\be_1}{\cid}} \eop \toeq{\elab{\be_2}{\cid}}

  \toeq{\prog_1;\prog_2} = \toeq{\prog_1} \wedge \toeq{\prog_2} 
\end{mathpar}

The motivating idea is that we can interpret any protocol $\prog$ as a set
of equality constraints $\toeq{\prog}$ and use an SMT solver to verify
properties relevant to correctness, confidentiality, and integrity.
Further, we can leverage entailment relation is critical for efficiency--
we can use annotations to obtain a weakened precondition for relevant properties.
That is, given $\prog$, program annotations or other cues can be used
to find a minimal $\eqs$ with $\toeq{\prog} \models \eqs$ for verifying
correctness and security.

\subsubsection{Example: Correctness of 3-Party Addition}

$$
\begin{array}{lll}
  \elab{\mesg{s1}}{2} &:=& \elab{(\secret{1} \fminus \locflip \fminus \flip{x})}{1} \\ 
  \elab{\mesg{s1}}{3} &:=& \elab{\flip{x}}{1} \\ 
  \elab{\mesg{s2}}{1} &:=& \elab{(\secret{2} \fminus \locflip \fminus \flip{x})}{2} \\ 
  \elab{\mesg{s2}}{3} &:=& \elab{\flip{x}}{2} \\ 
  \elab{\mesg{s3}}{1} &:=& \elab{(\secret{3} \fminus \locflip \fminus \flip{x})}{3} \\ 
  \elab{\mesg{s3}}{2} &:=& \elab{\flip{x}}{3}\\
  \rvl{1} &:=& \elab{(\locflip \fplus \mesg{s2} \fplus \mesg{s3})}{1} \\ 
  \rvl{2} &:=& \elab{(\mesg{s1} \fplus \locflip \fplus \mesg{s3})}{2} \\
  \rvl{3} &:=& \elab{(\mesg{s1} \fplus \mesg{s2} \fplus \locflip)}{3} \\
  \out{1} &:=& \elab{(\rvl{1} \fplus \rvl{2} + \rvl{3})}{1}\\
  \out{2} &:=& \elab{(\rvl{1} \fplus \rvl{2} + \rvl{3})}{2}\\
  \out{3} &:=& \elab{(\rvl{1} \fplus \rvl{2} + \rvl{3})}{3}
\end{array}
$$
Letting $\prog$ be this protocol, we can verify correctness
as:
$$
\toeq{\prog}\ \models\ \out{3} \eop \sx{1}{1} \fplus \sx{2}{2} \fplus \sx{3}{3}
$$

### 2024-08-18 csf25/main.tex
checkpoint

--- 2024-05-17 ppdp24/main.tex
@@
-\documentclass[sigconf]{acmart}
+\documentclass[conference]{IEEEtran}
+\IEEEoverridecommandlockouts
+% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
 
@@
 \usepackage{amstext}
+\usepackage{amssymb}
 \usepackage{fp-frame}
@@
 \input{macros}
-\input{figures}
-
-\acmConference[PPDP]{Principles and Practice of Declarative Programming}{2024}{Milan}
+%\input{figures}
 
@@
 
-\title{Language-Based Security for Low-Level MPC}
-
-\author{Christian Skalka}
-\affiliation{
-  \institution{University of Vermont}
-  \city{}
-  \country{}
-%  \city{Burlington}
-%  \country{USA}
-}
-\email{ceskalka@uvm.edu}
-
-\author{Joseph P. Near}
-\affiliation{
-  \institution{University of Vermont}
-  \city{}
-  \country{}
-%  \city{Burlington}
-%  \country{USA}
-}
-\email{jnear@uvm.edu}
-
-\begin{abstract}
-  Secure Multi-Party Computation (MPC) is an important
-  enabling technology for data privacy in modern distributed
-  applications. Currently, proof methods for low-level MPC protocols
-  are primarily manual and thus tedious and error-prone, and are also
-  non-standardized and unfamiliar to most PL theorists. As a step
-  towards better language support and language-based enforcement, we
-  develop a new staged PL for defining a variety of low-level
-  probabilistic MPC protocols. We also formulate a collection of
-  confidentiality and integrity hyperproperties for our language model
-  that are familiar from information flow, including conditional
-  noninterference, delimited release, and robust declassification. We
-  demonstrate their relation to standard MPC threat models of passive
-  and malicious security, and how they can be leveraged in security
-  verification of protocols. To prove these properties we develop
-  automated tactics in $\mathbb{F}_2$ that can be integrated with
-  separation logic-style reasoning.
-\end{abstract}
-
-%%
-%% The code below is generated by the tool at http://dl.acm.org/ccs.cfm.
-%% Please copy and paste the code instead of the example below.
-%%
-\begin{CCSXML}
-<ccs2012>
-   <concept>
-       <concept_id>10002978.10002986.10002990</concept_id>
-       <concept_desc>Security and privacy~Logic and verification</concept_desc>
-       <concept_significance>500</concept_significance>
-       </concept>
-   <concept>
-       <concept_id>10003752.10003753.10003757</concept_id>
-       <concept_desc>Theory of computation~Probabilistic computation</concept_desc>
-       <concept_significance>300</concept_significance>
-       </concept>
-   <concept>
-       <concept_id>10003752.10003790.10003806</concept_id>
-       <concept_desc>Theory of computation~Programming logic</concept_desc>
-       <concept_significance>500</concept_significance>
-       </concept>
- </ccs2012>
-\end{CCSXML}
-
-\ccsdesc[500]{Security and privacy~Logic and verification}
-\ccsdesc[500]{Theory of computation~Probabilistic computation}
-\ccsdesc[500]{Theory of computation~Programming logic}
-
-%%
-%% Keywords. The author(s) should pick words that accurately describe
-%% the work being presented. Separate the keywords with commas.
-\keywords{Secure multiparty computation, security verification, probabilistic programming, programming languages, information flow.}
-
-\maketitle
-
-\input{intro}
-
 \input{lang}
 
-\input{model}
+\input{smt}
 
-\input{hyper}
+\input{ctj}
 
-\input{bruteforce}
+\input{itj}
 
@@
 
-\input{example-gmw}
+\input{examples}
 
-\input{example-bdoz}
-
-\bibliographystyle{ACM-Reference-Format}
+\bibliographystyle{IEEEtran}
 \bibliography{logic-bibliography,secure-computation-bibliography}

### 2024-08-18 csf25/lang.tex
checkpoint

@@
-\section{$\minicat$ Syntax and Semantics}
+\section{$\minicat$ Syntax and Operational Semantics}
 
+The $\minifed$ language establishes a basic model of synchronous
+protocols between a federation of \emph{clients} exchanging values in
+the binary field. A model of synchronous communication captures a wide
+range of MPC protocols. Concurrency is out of scope in this work but
+an avenue for future work. The lack of sophisticated control
+structures in $\minifed$ is intentional, since minimizing features
+eases analysis and control abstractions such as function definitions
+can be integrated into a metalanguage that generates $\minifed$
+programs (Section \ref{section-metalang}).
+
+We identify clients by natural numbers and federations- finite sets of
+clients- are always given statically.  Our threat model assumes a
+partition of the federation into \emph{honest} $H$ and \emph{corrupt}
+$C$ subsets. We model probabilistic programming via a \emph{random
+tape} semantics. That is, we will assume that programs can make
+reference to values chosen from a uniform random distributions defined
+in the initial program memory.  Programs aka protocols execute
+deterministically given the random tape.
+
+\subsection{Syntax}
+
+\minicatsyntaxfig
+
+The syntax of $\minifed$, defined in Figure \ref{fig-minicat-syntax},
+includes values $v$ and standard operations of addition, subtraction,
+and multiplication in a finite field $\mathbb{F}_p$ where $p$ is some
+prime.  Protocols are given input secret values $\secret{w}$ as well
+as random samples $\flip{w}$ on the input tape, implemented using a
+\emph{memory} as described below (Section
+\ref{section-lang-semantics}) where $w$ is a distinguishing 
+identifier string. Protocols are sequences of assignment commands of three
+different forms:
+\begin{itemize}
+\item $\eassign{\mesg{w}}{\cid_2}{\be}{\cid_1}$: This
+  is a \emph{message send} where expression $\be$ is computed
+  by client $\cid_1$ and sent to client $\cid_2$ as message
+  $\mesg{w}$.
+\item $\reveal{w}{\be}{\cid}$: This
+  is a \emph{public reveal} where expression $\be$ is computed
+  by client $\cid$ and broadcast to the federation, typically
+  to communicate intermediate results for use in final output
+  computations.
+\item $\pubout{\cid}{\be}{\cid}$: This
+  is an \emph{output} where expression $\be$ is computed
+  by client $\cid$ and reported as its output. As a
+  sanity condition we disallow commands
+  $\pubout{\cid_1}{\be}{\cid_2}$ where $\cid_1\ne\cid_2$.
+\end{itemize}
+For example, in the following protocol, a client 1
+subtracts a random sample $\flip{y}$ from $\mathbb{F}_p$ from their
+secret value $\secret{x}$ and sends the result to client
+2 as a message $\mesg{z}$:
 $$
-    \begin{array}{rcl@{\hspace{2mm}}r}
-      \multicolumn{4}{l}{v \in \mathbb{F}_p,\ w \in \mathrm{String},\ \cid \in \mathrm{Clients} \subset  \mathbb{N} }\\[2mm] %, \bop \in \{ \eand, \eor, \exor \}} \\[2mm]
-      \be &::=& \flip{w} \mid \secret{w} \mid \mesg{w} \mid \rvl{w} \mid & \textit{expressions}\\
-      & & v \mid \be \fminus \be \mid \be \fplus \be \mid \be \ftimes \be \\[2mm]
-      x &::=& \elab{\flip{w}}{\cid} \mid \elab{\secret{w}}{\cid} \mid \elab{\mesg{w}}{\cid} \mid \rvl{w} \mid \out{\cid} & \textit{variables} \\[2mm]
-      %& &  \select{\be}{\be}{\be} \mid \ctxt{v}{k} \mid \key{w} \mid \sk{\be}(\be) \mid \pk{\be}{\be}(\be) \mid \pk{\be}{\be} \\[2mm]
-      %& &  \select{\fp(\be)}{\be}{\be} \ctxt{v,\be}{k}  \mid \sk{\be}(\be) \mid \pk{\be}{\be}(\be) \mid \pk{\be}{\be} \\[2mm]
-      \prog &::=& \eassign{\mesg{w}}{\cid}{\be}{\cid} \mid
-      \reveal{w}{e}{\cid} \mid \pubout{\cid}{\be}{\cid} \mid \prog;\prog & \textit{protocols} 
-    \end{array}
+\eassign{\mesg{z}}{2}{(\secret{x} - \flip{y})}{1}
+$$ Both messages $\mesg{w}$ and reveals $\rvl{w}$ can be referenced in
+expressions once they've been defined.  This distinction between
+messages and broadcast public reveal is consistent with previous
+formulations, e.g., \cite{6266151}. To identify and distinguish
+between collections of variables in protocols we introduce the
+following notation.
+\begin{definition}
+We let $x$ range over \emph{variables} which are identifiers where
+client ownership is specified- e.g.,
+$\elab{\mesg{\mathit{foo}}}{\cid}$ is a message $\mathit{foo}$ that
+was sent to $\cid$. We let $X$ range over sets of variables, and more
+specifically, $S$ ranges over sets of secret variables
+$\elab{\secret{w}}{\cid}$, $R$ ranges over sets of random variables
+$\elab{\flip{w}}{\cid}$, $M$ ranges over sets of message variables
+$\elab{\mesg{w}}{\cid}$, $P$ ranges over sets of public variables
+$\rvl{w}$, and $O$ ranges over sets of output variables $\out{\cid}$.
+Given a program $\prog$, we write $\iov(\prog)$ to denote the
+particular set $S \cup M \cup P \cup O$ of variables in $\prog$ and
+$\secrets(\prog)$ to denote $S$, and we write $\flips(\prog)$ to
+denote the particular set $R$ of random samplings in $\prog$. We write
+$\vars(\prog)$ to denote $\iov(\prog) \cup \flips(\prog)$. For any set
+of variables $X$ and clients $I$, we write $X_I$ to denote the subset
+of $X$ owned by any client $\cid \in I$, in particular we write $X_H$
+and $X_C$ to denote the subsets belonging to honest and corrupt
+parties, respectively.
+\end{definition}
+
+\subsection{Semantics}
+\label{section-lang-semantics}
+
+\minicatredxfig
+
+\emph{Memories} are fundamental to the semantics of $\fedcat$ and
+provide random tape and secret inputs to protocols, and also record
+message sends, public broadcast, and client outputs.
+\begin{definition}
+  Memories $\store$
+are finite (partial) mappings from variables $x$ to values $v \in
+\mathbb{F}_p$.  The \emph{domain} of a memory is written
+$\dom(\store)$ and is the finite set of variables on which the memory
+is defined.  We write $\store\{ x \mapsto v\}$ for
+$x\not\in\dom(\store)$ to denote the memory $\store'$ such that
+$\store'(x) = v$ and otherwise $\store'(y) = \store(y)$ for all $y \in
+\dom(\store)$. We write $\store \subseteq \store'$ iff $\dom(\store)
+\subseteq \dom(\store')$ and $\store(x) = \store'(x)$ for all $x \in
+\dom(\store)$. Given any $\store$ and $\store'$ with
+$\store(x) = \store'(x)$ for all $x \in \dom(\store) \cap \dom(\store')$,
+we write $\store \uplus \store'$ to denote the memory
+with domain $X = \dom(\store) \cup \dom(\store')$ such
+that:
 $$
+\forall x \in X .
+(\store \uplus \store')(x) =
+\begin{cases} \store(x) \text{\ if\ } x\in\dom(\store) \\ \store'(x) \text{\ otherwise\ }\end{cases} 
+$$
+%We write $\store \subseteq \store'$ iff $\store \uplus \store' = \store$.
+\end{definition}
+In our subsequent presentation we will often want to consider arbitrary
+memories that range over particular variables and to restrict
+memories to particular subsets of their domain:
+\begin{definition}
+  Given a set of variables $X$ and memory $\store$, we write
+  $\store_X$ to denote the memory with $\dom(\store_X) = X$ and
+  $\store_X(x) = \store(x)$ for all $x \in X$. We define $\mems(X)$ as
+  the set of all memories with domain $X$:
+  $$
+  \mems(X) \defeq \{ \store \mid \dom(\store) = X \}
+  $$
+\end{definition}
+So for example, given a protocol $\prog$, the set of all random tapes for
+$\prog$ is $\mems(\flips(\prog))$, and the memory $\store_{\secrets(\prog)}$
+is $\store$ restricted to the secrets in $\prog$.
+%We let $\stores$ range
+%over sets of memories with the same domain, and abusing notation
+%we write $\dom(\stores)$ to denote the common domain,
+%and $\stores_X \defeq \{ \store_X | \store \in \stores \}$.
 
-\bigskip
-
- $$
-  %\begin{array}{c@{\hspace{5mm}}c}
-  \begin{array}{rcl}
-    \lcod{\store, v}{\cid} &=& v\\
-    \lcod{\store, \be_1 \fplus \be_2}{\cid} &=& \fcod{\lcod{\store, \be_1}{\cid} \fplus \lcod{\store, \be_2}{\cid}}\\ 
-    \lcod{\store, \be_1 \fminus \be_2}{\cid} &=& \fcod{\lcod{\store, \be_1}{\cid} \fminus \lcod{\store, \be_2}{\cid}}\\ 
-    \lcod{\store, \be_1 \ftimes \be_2}{\cid} &=& \fcod{\lcod{\store, \be_1}{\cid} \ftimes \lcod{\store, \be_2}{\cid}}\\
-  %\end{array} 
-  %\begin{array}{rcl}
-    \lcod{\store, \flip{w}}{\cid} &=& \store(\elab{\flip{w}}{\cid})\\
-    \lcod{\store, \secret{w}}{\cid} &=& \store(\elab{\secret{w}}{\cid})\\
-    \lcod{\store, \mesg{w}}{\cid} &=& \store(\elab{\mesg{w}}{\cid})\\
-    \lcod{\store, \rvl{w}}{\cid} &=& \store(\rvl{w})\\
-    %\lcod{\store, f\,\be_1\,\cdots\, \be_n}{\cid} &=& \delta(f,\lcod{\store, \be_1}{\cid},\ldots,\lcod{\store,\be_n}{\cid})
-  \end{array}
-  %\end{array}
-  $$
-
-\bigskip
-
-\begin{mathpar}
-  (\store, \xassign{x}{\be}{\cid}) \redx \extend{\store}{x}{\lcod{\store,\be}{\cid}}
-
-  \inferrule
-      {(\store_1,\prog_1) \redx \store_2 \\ (\store_2,\prog_2) \redx \store_3 }
-      {(\store_1,\prog_1;\prog_2) \redx \store_3}
-      %(\store, \eassign{\mesg{w}}{\cid_1}{\be}{\cid_2};\prog) \redx (\extend{\store}{\mesg{w}_{\cid_1}}{\lcod{\store,\be}{\cid_2}}, \prog)    
-      %(\store, \reveal{w}{\be}{\cid};\prog) \redx (\extend{\store}{\rvl{w}}{\lcod{\store,\be}{\cid}}, \prog)   
-      %(\store, \pubout{\cid}{\be}{\cid};\prog) \redx (\extend{\store}{\out{\cid}}{\lcod{\store,\be}{\cid}}, \prog)
-\end{mathpar}
+Given a variable-free expression $\be$, we write $\cod{\be}$ to denote
+the standard interpretation of $\be$ in the arithmetic field
+$\mathbb{F}_{p}$. With the introduction of variables to expressions,
+we need to interpret variables with respect to a specific memory, and
+all variables used in an expression must belong to a specified client.
+Thus, we denote interpretation of expressions $\be$ computed on a
+client $\cid$ as $\lcod{\store,\be}{\cid}$. This interpretation is
+defined in Figure \ref{fig-minifed}. The small-step reduction relation
+$\redx$ is then defined in Figure \ref{fig-minifed} to evaluate
+commands. Reduction is a relation on \emph{configurations} $(\store,
+\prog)$ where all three command forms- message send, broadcast, and
+output- are implemented as updates to the memory $\store$. We write
+$\redxs$ to denote the reflexive, transitive closure of\ $\redx$.
 
@@
 
-$$
-\begin{array}{rclr}
-  (\store, \xassign{x}{\be}{\cid}) &\aredx&
-  \extend{\store}{x}{\lcod{\store,\be}{\cid}} & \cid \in H\\
-  (\store, \xassign{x}{\be}{\cid}) &\aredx&
-  \extend{\store}{x}{\lcod{\arewrite(\store_C,\be)}{\cid}} & \cid \in C
-\end{array}
-$$
+\minicataredxfig
 
-$$
-\begin{array}{rcl@{\qquad}r}
-  (\store,\elab{\assert{\be_1 = \be_2}}{\cid}) &\aredx& \store & \text{if\ }
-  \lcod{\store,\be_1}{\cid} = \lcod{\store,\be_2}{\cid}  \text{\ or\ } \cid \in C\\
-  (\store,\elab{\assert{\be_1 = \be_2}}{\cid}) &\aredx& \abort & \text{if\ }
-  \lcod{\store,\be_1}{\cid} \ne \lcod{\store,\be_2}{\cid}  \text{\ or\ } \cid \in C
-\end{array}
-$$
+In the malicious model we assume that corrupt clients are in
+thrall to an adversary $\adversary$ who does not necessarily follow
+the rules of the protocol.  We model this by positing a $\arewrite$
+function which is given a corrupt memory $\store_C$ and expression
+$\be$, and returns a rewritten expression that can be interpreted to
+yield a corrupt input. We define the evaluation relation that
+incorporates the adversary in Figure \ref{fig-adversary}.
 
-\begin{mathpar}
-  \inferrule
-      {(\store_1,\prog_1) \aredx \store_2 \\ (\store_2,\prog_2) \aredx \store_3 }
-      {(\store_1,\prog_1;\prog_2) \aredx \store_3}
-
-  \inferrule
-      {(\store_1,\prog_1) \aredx \abort}
-      {(\store_1,\prog_1;\prog_2) \aredx \abort}
-
-  \inferrule
-      {(\store_1,\prog_1) \aredx \store_2 \\ (\store_2,\prog_2) \aredx \abort }
-      {(\store_1,\prog_1;\prog_2) \aredx \abort}
-\end{mathpar}
+A key technical distinction of the malicious setting is that it
+typically incorporates ``abort''. Honest parties implement strategies
+to detect rule-breaking-- aka \emph{cheating}-- by using, e.g.,
+message authentication codes with semi-homomorphic properties as in
+BDOZ/SPDZ \cite{10.1007/978-3-030-68869-1_3}. If cheating is detected,
+the protocol is aborted. To model this, we extend $\minifed$ with an
+\ttt{assert} command and extend the range of memories with
+$\bot$. Note that the adversary is free to ignore their own
+assertions.
+\begin{definition}
+  We add assertions of the form $\elab{\assert{\phi(\be)}}{\cid}$ to the command
+  syntax of $\minifed$, where $\phi$ is a decidable predicate on
+  $\mathbb{F}_p$ and with operational semantics given in Figure
+  \ref{fig-adversary}. We also extend the range of memories $\store$
+  to $\mathbb{F}_p \cup \{ \bot \}$.
+\end{definition}

### 2024-08-19 csf25/lang.tex
checkpoint

@@
 protocols between a federation of \emph{clients} exchanging values in
-the binary field. A model of synchronous communication captures a wide
-range of MPC protocols. Concurrency is out of scope in this work but
-an avenue for future work. The lack of sophisticated control
-structures in $\minifed$ is intentional, since minimizing features
-eases analysis and control abstractions such as function definitions
-can be integrated into a metalanguage that generates $\minifed$
-programs (Section \ref{section-metalang}).
-
-We identify clients by natural numbers and federations- finite sets of
+the binary field. We identify clients by natural numbers and federations- finite sets of
 clients- are always given statically.  Our threat model assumes a
@@
 \end{itemize}
-For example, in the following protocol, a client 1
-subtracts a random sample $\flip{y}$ from $\mathbb{F}_p$ from their
-secret value $\secret{x}$ and sends the result to client
-2 as a message $\mesg{z}$:
-$$
-\eassign{\mesg{z}}{2}{(\secret{x} - \flip{y})}{1}
-$$ Both messages $\mesg{w}$ and reveals $\rvl{w}$ can be referenced in
-expressions once they've been defined.  This distinction between
+The distinction between
 messages and broadcast public reveal is consistent with previous
@@
 client $\cid$ as $\lcod{\store,\be}{\cid}$. This interpretation is
-defined in Figure \ref{fig-minifed}. The small-step reduction relation
-$\redx$ is then defined in Figure \ref{fig-minifed} to evaluate
-commands. Reduction is a relation on \emph{configurations} $(\store,
-\prog)$ where all three command forms- message send, broadcast, and
-output- are implemented as updates to the memory $\store$. We write
-$\redxs$ to denote the reflexive, transitive closure of\ $\redx$.
+defined in Figure \ref{fig-minicat-redx}, along with the big-step
+reduction relation $\redx$ to evaluate commands. Reduction is a mapping
+fromon \emph{configurations} $(\store,\prog)$ to final stores
+where all three command forms- message send, broadcast, and
+output- are implemented as updates to the memory $\store$.
 

### 2024-08-19 csf25/lang.tex
checkpoint

@@
 
+\cnote{Add comments about OT.}
+
 \subsection{Semantics}
@@
 
-$$
-    \begin{array}{rcl@{\hspace{2mm}}r}
-      \prog &::=& \cdots \mid \assert{\be = \be}
-    \end{array}
-$$
-
 \minicataredxfig
@@
 yield a corrupt input. We define the evaluation relation that
-incorporates the adversary in Figure \ref{fig-adversary}.
+incorporates the adversary in Figure \ref{fig-minicat-aredx}.
 
@@
 the protocol is aborted. To model this, we extend $\minifed$ with an
-\ttt{assert} command and extend the range of memories with
+\ttt{assert} command:
+$$
+    \begin{array}{rcl@{\hspace{2mm}}r}
+      \prog &::=& \cdots \mid \assert{\be = \be}
+    \end{array}
+$$
+and extend the range of memories with
 $\bot$. Note that the adversary is free to ignore their own

### 2024-08-22 code/cvc5mvn/src/main/java/plaid/App.java
Update classes to convert protocols in Overture to CVC5 constraints

@@ -1,70 +1,152 @@
 package plaid;
 
 import java.math.BigInteger;
-import java.util.List;
+import java.util.*;
 
 import org.antlr.v4.runtime.*;
 import org.antlr.v4.runtime.tree.*;
 
 import io.github.cvc5.*;
 import java.math.BigInteger;
-import java.util.ArrayList;
-import java.util.Arrays;
 import java.util.List;
 
-/**
- * Hello world!
- *
- */
 public class App 
 {
+    public static Map<String, Term> createLookup(Iterable<Memory> memories){
+        HashMap<String, Term> results = new HashMap<>();
+
+        // iterate over memories
+        for(Memory memory : memories){
+            results.put(memory.getName(), memory.getTerm());
+        }
+
+        return results;
+    }
+
     public static void main( String[] args ) throws Exception
     {
-        // antlr demo
+        // antlr&cvc5 demo
+        TermManager termManager = new TermManager();
+        Solver solver = new Solver(termManager);
+        solver.resetAssertions();
+        solver.setLogic("ALL");
+        Sort f7 = termManager.mkFiniteFieldSort("7", 10);
+
+
+        // 3-party addition
+        String protocol = """
+                m["s1"]@2 := ((s["1"] + -r["local"]) + -r["x"])@1;
+                m["s1"]@3 := r["x"]@1;
+                m["s2"]@1 := ((s["2"] + -r["local"]) + -r["x"])@2;
+                m["s2"]@3 := r["x"]@2;
+                m["s3"]@1 := ((s["3"] + -r["local"]) + -r["x"])@3;
+                m["s3"]@2 := r["x"]@3;
+                p["1"] := ((r["local"] + m["s2"]) + m["s3"])@1;
+                p["2"] := ((m["s1"] + r["local"]) + m["s3"])@2;
+                p["3"] := ((m["s1"] + m["s2"]) + r["local"])@3;
+                out@1 := ((p["1"] + p["2"]) + p["3"])@1;
+                out@2 := ((p["1"] + p["2"]) + p["3"])@2;
+                out@3 := ((p["1"] + p["2"])+ p["3"])@3
+                
+                """;
+
[…]

### 2024-08-22 code/cvc5mvn/src/main/java/plaid/OvertureConstraintListener.java

@@ -1,0 +1,108 @@
+package plaid;
+
+import io.github.cvc5.*;
+import org.antlr.v4.runtime.RuleContext;
+import java.util.List;
+
+/*
+ * Identifies expressions in Overture and converts them into constraint terms in a cvc5 solver
+ */
+public class OvertureConstraintListener extends OvertureBaseVisitor<Term> {
+    private final TermManager termManager;
+    private final Sort prime;
+    private Iterable<Memory> memories;
+
+    // constructor
+    public OvertureConstraintListener(TermManager termManager, Sort sort, Iterable<Memory> memories){
+        this.termManager = termManager;
+        this.prime = sort;
+        this.memories = memories;
+    }
+
+    @Override
+    protected Term aggregateResult(Term aggregate, Term nextResult) {
+        return aggregate == null ? nextResult : aggregate;
+    }
+    @Override
+    public Term visitValueExpr(OvertureParser.ValueExprContext ctx) {
+        try {
+            return termManager.mkFiniteFieldElem(ctx.VALUE().getText(), prime, 10);
+        } catch (CVC5ApiException e) {
+            throw new RuntimeException(e);
+        }
+    }
+
+    @Override
+    public Term visitMinusExpr(OvertureParser.MinusExprContext ctx) {
+        Term expr = visit(ctx.expression());
+        try {
+            return termManager.mkTerm(Kind.FINITE_FIELD_MULT, expr, termManager.mkFiniteFieldElem("-1", prime, 10));
+        } catch (CVC5ApiException e) {
+            throw new RuntimeException(e);
+        }
+    }
+
+    @Override
+    public Term visitPlusExpr(OvertureParser.PlusExprContext ctx) {
+        Term expr1 = visit(ctx.expression(0));
+        Term expr2 = visit(ctx.expression(1));
+
+        return termManager.mkTerm(Kind.FINITE_FIELD_ADD, expr1, expr2);
+
+    }
+
+    @Override
+    public Term visitTimesExpr(OvertureParser.TimesExprContext ctx) {
+        Term expr1 = visit(ctx.expression(0));
+        Term expr2 = visit(ctx.expression(1));
+
+        return termManager.mkTerm(Kind.FINITE_FIELD_MULT, expr1, expr2);
+    }
+
+    @Override
+    public Term visitParenExpr(OvertureParser.ParenExprContext ctx) {
+        return visit(ctx.expression());
+    }
+
+    @Override
[…]

### 2024-08-22 code/cvc5mvn/src/main/java/plaid/SolverExtension.java

@@ -1,0 +1,39 @@
+package plaid;
+import io.github.cvc5.*;
+import org.antlr.v4.runtime.*;
+import org.antlr.v4.runtime.tree.*;
+import java.util.*;
+
+
+public class SolverExtension {
+    public static Iterable<Memory> addOvertureProtocolConst(Solver solver, Sort prime, String protocol) throws CVC5ApiException{
+        //create a parse tree with protocol
+        ANTLRInputStream input = new ANTLRInputStream(protocol);
+        OvertureLexer lexer = new OvertureLexer(input);
+        CommonTokenStream tokens = new CommonTokenStream(lexer);
+        OvertureParser parser = new OvertureParser(tokens);
+        parser.setBuildParseTree(true);
+        OvertureParser.ProtocolContext tree = parser.protocol();
+
+        //create memories from there
+        TermManager termManager = solver.getTermManager();
+        OvertureConstListener listener = new OvertureConstListener(termManager, prime);
+        new ParseTreeWalker().walk(listener, tree);
+        Iterable<Memory> memoryList = listener.memories();
+
+        System.out.println(memoryList);
+        OvertureConstraintListener overtureConstraintListener = new OvertureConstraintListener(termManager, prime, memoryList);
+
+        // iterate over commands in the parse tree
+        for (OvertureParser.CommandContext commandContext : tree.command()){
+            System.out.println(overtureConstraintListener.visit(commandContext));
+            solver.assertFormula(overtureConstraintListener.visit(commandContext));
+
+        }
+            // for each command, create a constraint (constraint visitor)
+
+        return memoryList;
+
+
+    }
+}

### 2024-08-22 code/cvc5mvn/src/test/java/plaid/SolverExtensionTest.java

@@ -1,0 +1,20 @@
+package plaid;
+
+import org.junit.Test;
+import io.github.cvc5.*;
+
+public class SolverExtensionTest {
+    @Test
+    public void test() throws CVC5ApiException {
+        TermManager termManager = new TermManager();
+        Solver solver = new Solver(termManager);
+
+        Sort f7 = termManager.mkFiniteFieldSort("7", 10);
+        String protocol = "m[\"foo\"]@2 := (s[\"x\"] + -r[\"y\"])@1";
+
+        //SolverExtension.addOvertureProtocolConst(solver, f7, protocol);
+        //System.out.println(solver.checkSat());
+
+        //System.out.println(termManager.mkFiniteFieldElem("7", f7, 10));
+    }
+}

### 2024-08-22 csf25/ctj.tex

\section{Confidentiality Types}

$$
\begin{array}{rcl@{\hspace{2mm}}r}
  t &::=& x \mid \cty{x}{T} \\
  \ty & \in & 2^{t}\\
  \Gamma &::=& \varnothing \mid \Gamma; x : \ty
\end{array} 
$$

\begin{definition}
  $R_1;R_2 = R_1 \cup R_2$ iff $R_1 \cap R_2 = \varnothing$.
\end{definition}

\begin{mathpar}
  \inferrule[DepTy]
  {}
  {\eqj{\varnothing}{\eqs}{\phi}{\vars(\phi)}}

  \inferrule[Encode]
  {\eqs \models \phi \eop \phi' \oplus \rx{w}{\cid} \\
   \oplus \in \{ \fplus,\fminus \}\\
   \eqj{R}{\eqs}{\phi'}{\ty}}
  {\eqj{R;\{ \rx{w}{\cid} \}}{\eqs}{\phi}{\setit{\cty{\rx{w}{\cid}}{\ty}}}}
\end{mathpar}

\begin{mathpar}
  \inferrule[Send]
            {\eqj{R}{\eqs}{\phi}{\ty}}
            {\cpj{R}{\eqs}{x \eop \phi}{(x : \ty)}}

  \inferrule[Seq]
            {\cpj{R_1}{\eqs}{\phi_1}{\Gamma_1}\\
             \cpj{R_2}{\eqs}{\phi_2}{\Gamma_2}}
            {\cpj{R_1;R_2}{\eqs}{\phi_1 \wedge \phi_2}{\Gamma_1;\Gamma_2}}
\end{mathpar}

\begin{definition}
  Given preprocessing predicate $\eqspre$ and protocol $\prog$ we say
  $\cpj{R}{\eqs}{\eqspre \wedge \toeq{\prog}}{\Gamma}$ is \emph{valid} iff it is derivable and
  $\eqspre \wedge \toeq{\prog} \models \eqs$.
\end{definition}

\begin{mathpar}
  \inferrule
      {\cid \in C}
      {\leakj{\Gamma}{C}{\Gamma(\mx{w}{\cid})}}

  \inferrule
      {\leakj{\Gamma}{C}{T_1 \cup T_2}}
      {\leakj{\Gamma}{C}{T_1}}

  \inferrule
      {\leakj{\Gamma}{C}{\setit{ \mx{w}{\cid} }}}
      {\leakj{\Gamma}{C}{\Gamma(\mx{w}{\cid})}}

  \inferrule
      {\leakj{\Gamma}{C}{\setit{ \rx{w}{\cid} }} \\ \leakj{\Gamma}{C}{\setit{ \cty{\rx{w}{\cid}}{\ty} }} }
      {\leakj{\Gamma}{C}{\ty}}
\end{mathpar}

\begin{theorem}
  If $\cpj{R}{\eqs}{\eqspre \wedge \toeq{\prog}}{\Gamma}$ is valid and there exists no $H,C$ 
  and  $\sx{w}{\cid}$ for $\cid \in H$ with $\leakj{\Gamma}{C}{\setit{\sx{w}{\cid}}}$,
  then $\prog$ satisfies gradual release.
\end{theorem}

\subsection{Examples}

\begin{verbatimtab}
m[s1]@2 := (s[1] - r[local] - r[x])@1
m[s1]@3 := r[x]@1

// m[s1]@2 : { c(r[x]@1, { c(r[local]@1, {s[1]@1} ) }
// m[s1]@3 : { r[x]@1 }
\end{verbatimtab}

\begin{verbatimtab}
m[x]@1 := s2(s[x],-r[x],r[x])@2

// m[x]@1 == s[x]@2 + -r[x]@2 
// m[x]@1 : { c(r[x]@2, { s[x]@2 }) }

m[y]@1 := OT(s[y]@1,-r[y],r[y])@2

// m[y]@1 == s[y]@1 + -r[y]@2
// m[y]@1 : { c(r[y]@2, { s[y]@1 }) } 
\end{verbatimtab}

### 2024-08-22 csf25/lang.tex

@@
 $$
-and extend the range of memories with
-$\bot$. Note that the adversary is free to ignore their own
-assertions.
-\begin{definition}
-  We add assertions of the form $\elab{\assert{\phi(\be)}}{\cid}$ to the command
-  syntax of $\minifed$, where $\phi$ is a decidable predicate on
-  $\mathbb{F}_p$ and with operational semantics given in Figure
-  \ref{fig-adversary}. We also extend the range of memories $\store$
-  to $\mathbb{F}_p \cup \{ \bot \}$.
-\end{definition}
+
+The big-step adversarial semantics relation $\aredx$ is then defined
+in Figure \ref{fig-minicat-aredx} as a mapping from configurations to
+$\stores$ or $\bot$.
+
+As we will discuss, this form of asserting equality between expressions
+is sufficient to capture MAC checking and support integrity in protocols.

### 2024-08-22 csf25/smt.tex
checkpoint

@@
 \section{$\minicat$ Constraint Verification}
+
+In previous work we have shown that the semantics of $\minicat$ in
+$\mathbb{F}_2$ can be implemented with Datalog \cite{skalka-near-ppdp24}, which is
+essentially a form of constraint programming. In this paper we extend
+this idea to arbitrary prime fields by using a more general form of
+SMT constraint programming, specifically Satisfiability Modulo Finite
+Fields \cite{SMFF}. As we will show in subsequent sections, this
+interpretation supports correctness guarantees, and also static type
+analyses for enforcing confidentiality and integrity properties.
 
@@
 
-We write $\eqs_1 \models \eqs_2$ iff every model of $E_1$ is a model of $E_2$. Note that
-this relation is reflexive and transitive.
+\begin{mathpar}
+  \store(\phi_1 \fplus \phi_2) = \cod{\store(\phi_1) \fplus \store(\phi_2)}
+
+  \store(\phi_1 \ftimes \phi_2) = \cod{\store(\phi_1) \ftimes \store(\phi_2)}
+
+  \store(\phi_1 \fminus \phi_2) = \cod{\store(\phi_1) \fminus \store(\phi_2)}
+\end{mathpar}
+
+\begin{definition}
+  Given $\eqs$ of the form $\phi_1 \eop \phi_1' \wedge \cdots \wedge \phi_n \eop
+  \phi_n'$ and $\store$, we say that $\store$ is a \emph{model} of
+  $\eqs$, written $\store \models \eqs$, iff $\store(\phi_i) =
+  \store(\phi_i')$ for all $1 \le i \le n$. We write $\eqs_1 \models
+  \eqs_2$ iff  $\store \models E_1$ implies $\store \models
+  E_2$ for all $\store$, and note this relation is reflexive and transitive.
+\end{definition}
+
+\begin{theorem}
+  $\eqs \models \phi \eop \phi'$ iff $\eqs \wedge \phi \not\eop \phi'$ is
+  not satisfiable.
+\end{theorem}
 
@@
 \begin{mathpar}
-  \toeq{\OT{\be_1}{\cid_1}{\be_2}{\be_3}{\cid_2}} =
+  \toeq{\elab{\OT{\be_1}{\cid_1}{\be_2}{\be_3}}{\cid_2}} =
   (\toeq{\elab{\be_1}{\cid_1}} \wedge \toeq{\elab{\be_3}{\cid_2}}) \vee

### 2024-08-24 csf25/lang.tex
checkpoint

@@
 $$
-
+As we will discuss, this form of asserting equality between expressions
+is sufficient to capture MAC checking and support integrity in protocols.
 The big-step adversarial semantics relation $\aredx$ is then defined
@@
 $\stores$ or $\bot$.
-
-As we will discuss, this form of asserting equality between expressions
-is sufficient to capture MAC checking and support integrity in protocols.

### 2024-08-24 csf25/smt.tex
checkpoint

@@
 In previous work we have shown that the semantics of $\minicat$ in
-$\mathbb{F}_2$ can be implemented with Datalog \cite{skalka-near-ppdp24}, which is
-essentially a form of constraint programming. In this paper we extend
-this idea to arbitrary prime fields by using a more general form of
-SMT constraint programming, specifically Satisfiability Modulo Finite
-Fields \cite{SMFF}. As we will show in subsequent sections, this
-interpretation supports correctness guarantees, and also static type
-analyses for enforcing confidentiality and integrity properties.
+$\mathbb{F}_2$ can be implemented with Datalog
+\cite{skalka-near-ppdp24}, which is a form of constraint
+programming. In this paper we extend this idea to arbitrary prime
+fields by using a more general form of SMT constraint programming,
+specifically Satisfiability Modulo Finite Fields \cite{SMFF}. As we
+will show in subsequent sections, this interpretation supports
+correctness guarantees, and also static type analyses for enforcing
+confidentiality and integrity properties.
 
@@
 
+In Figure \ref{fig-eqs} we define a syntax of SMT-style constraints
+$\eqs$, where constraint terms $\phi$ are similar to expressions $\be$
+except the can ``cross party lines''. This turns out to be very useful
+to express correctness properties-- for example, in the GMW protocol
+wire values in circuits are represented by reconstructive shares
+\cite{evans2018pragmatic}.  If by convention shares of values $n$ are
+represented by $\mesg{n}$ on clients, then assuming two clients
+$\setit{1,2}$ the reconstructed value can be expressed as $\mx{n}{1} +
+\mx{n}{2}$.  So, while summing values across clients is disallowed in
+$\minicat$ protocols, we can express properties of shared values
+via constraints.
+
+\begin{fpfig}[t]{Syntax of Constraints and Memory Application Extended to
+  SMT expressions}{fig-eqs}
+\small{
 $$
@@
   \store(\phi_1 \fminus \phi_2) = \cod{\store(\phi_1) \fminus \store(\phi_2)}
-\end{mathpar}
+\end{mathpar}}
+\end{fpfig}
 
+This language of constraints has a direct interpretation in the Finite
+Fields theory of cvc5 \cite{SMFF}. We can leverage this to implement a
+critical \emph{entailment} property, written $\eqs_1 \models \eqs_2$.
+Our entailment relation is based on model satisfiability in a standard
+sense, except we represent models as memories $\store$. To this end
+we extend application of memories to terms $\store(\phi)$ in
+Figure \ref{fig-eqs}, and we formally define entailment as follows.
 \begin{definition}
@@
 
+Given this Definition, the following Theorem is well-known and commonly
+applied in SMT systems to implement our notion of entailment. 
 \begin{theorem}
@@
 
+\subsection{Programs as Constraint Systems}
+
+\begin{fpfig}[t]{Interpretation of $\minicat$ expressions (T) and programs (B) as
+  constraints}{fig-toeq}
+\small{
 \begin{mathpar}
@@
 \begin{mathpar}
-  \toeq{\elab{\OT{\be_1}{\cid_1}{\be_2}{\be_3}}{\cid_2}} =
-  (\toeq{\elab{\be_1}{\cid_1}} \wedge \toeq{\elab{\be_3}{\cid_2}}) \vee
-  (\neg\toeq{\elab{\be_1}{\cid_1}} \wedge \toeq{\elab{\be_2}{\cid_2}}) 
-\end{mathpar}
-
-\begin{mathpar}
   \toeq{\xassign{x}{\be}{\cid}} = x \eop \toeq{\elab{\be}{\cid}}
@@
   \toeq{\prog_1;\prog_2} = \toeq{\prog_1} \wedge \toeq{\prog_2} 
-\end{mathpar}
+\end{mathpar}}
+\end{fpfig}
 
-The motivating idea is that we can interpret any protocol $\prog$ as a set
-of equality constraints $\toeq{\prog}$ and use an SMT solver to verify
-properties relevant to correctness, confidentiality, and integrity.
-Further, we can leverage entailment relation is critical for efficiency--
+A central idea of our approach is that we can interpret any protocol
+$\prog$ as a set of equality constraints (denoted $\toeq{\prog}$) and use an SMT
+solver to verify properties relevant to correctness, confidentiality,
+and integrity. Further, we can leverage entailment relation is critical for efficiency--
 we can use annotations to obtain a weakened precondition for relevant properties.
@@
 
+The mapping $\toeq{\cdot}$ from programs $\prog$ to contraints
+is defined in Figure \ref{fig-toeq}. The interpretion of OT is ommitted
+from this figure which is general-- for $\mathbb{F}_2$ the interpretation
+is:
+\begin{mathpar}
+  \toeq{\elab{\OT{\be_1}{\cid_1}{\be_2}{\be_3}}{\cid_2}} =
+  (\toeq{\elab{\be_1}{\cid_1}} \ftimes \toeq{\elab{\be_3}{\cid_2}}) \fplus
+  (\neg\toeq{\elab{\be_1}{\cid_1}} \ftimes \toeq{\elab{\be_2}{\cid_2}}) 
+\end{mathpar}
+The correctness of this interpretation is characterized as follows.  
+\begin{theorem}
+  $\store$ is a model of $\eqspre \wedge \toeq{\prog}$ iff $\store \in \runs(\prog)$.
+  \cnote{Need to define $\runs$ somewhere.}
+\end{theorem}
+
 \subsubsection{Example: Correctness of 3-Party Addition}
+Consider the following 3-party Shamir Addition protocol.
 
@@
 Letting $\prog$ be this protocol, we can verify correctness
-as:
+as follows-- the output is the sum of the three input secrets:
 $$
-\toeq{\prog}\ \models\ \out{3} \eop \sx{1}{1} \fplus \sx{2}{2} \fplus \sx{3}{3}
+\toeq{\prog}\ \models\ \out{1} \eop \sx{1}{1} \fplus \sx{2}{2} \fplus \sx{3}{3}
 $$

### 2024-08-26 csf25/ctj.tex
checkpoint

@@
 \section{Confidentiality Types}
+
+It is well-known that adding or subtracting a sample from a uniform
+distribution in a finite field yields a value in a uniform
+distribution, meaning that samples can be used as one-time-pads with
+perfect secrecy \cite{barthe2019probabilistic,darais2019language}.
+In our type system for confidentiality, we aim to approximate
+distributions by tracking which program variables an expression
+may depend on. But we also want to capture this fundamental mechanism
+of encryption.
+
+For example, given a message send such as:
+$$
+\xassign{\mx{x}{1}}{\secret{x} \ftimes \flip{x}}{2}
+$$
+we would assign the following type to denote that $\mx{x}{1}$ is dependent on
+both $\sx{x}{2}$ and $\rx{x}{2}$:
+$$
+\mx{x}{1} : \setit{\sx{x}{2}, \rx{x}{2}}
+$$
+But in the case of 2-party reconstructive sharing:
+$$
+\xassign{\mx{x}{1}}{\secret{x} \fminus \flip{x}}{2}
+$$
+we want the type of $\mx{x}{1}$ to be an ``encrypted'' type
+that is by itself independent of $\sx{x}{2}$. However, we also
+want to track the dependence of $\mx{x}{2}$ and $\rx{x}{2}$.
+In short, we assign $\mx{x}{1}$ a ``ciphertext type'':
+$$
+\mx{x}{1} : \cty{\rx{x}{2}}{\sx{x}{2}}
+$$
+indicating that $\mx{x}{1}$ is an encryption of a value
+dependent on $\sx{x}{2}$ with $\rx{x}{2}$ and as long as client 1
+does not subsequently receive messages that are dependent on $\rx{x}{2}$, the
+value $\sx{x}{2}$ remains hidden.
+
+Of course, there are other methods for encrypting values
+in MPC protocols-- but we can observe that many are algebraically
+equivalent to this fundamental schema. For example, in
+Yao's Garbled Circuits (YGC) in $\mathbb{F}_{2}$, the ``garbler'' represents
+encrypted wire values as random samples, and the ``evaluator''
+obtains the wire value for its input via OT. The evaluator shares 
+the initial encrypted wire value of one of its input secrets as
+a selection bit-- assuming that client
+2 is the garbler and client 1 is the evaluator, we can
+write this as:
+$$
+\xassign{\mx{x}{1}}{\mux{\secret{x}}{\neg\flip{x}}{\flip{x}}}{2}
+$$
+where for all $\be_1,\be_2,\be_3$:
+$$
+\mux{\be_1}{\be_2}{\be_3} \defeq (\neg\be_1 \ftimes \be_2) \fplus (\be_1 \ftimes \be_3)
+$$
+and letting this protocol be $\prog$ the following is valid:
+\begin{mathpar}
+%\toeq{\xassign{\mx{x}{1}}{\OT{\secret{x}}{1}{\neg\flip{x}}{\flip{x}}}{2}}
+   \toeq{\prog} \models \mx{x}{1} \eop \sx{x}{2} \fplus \rx{x}{2}
+\end{mathpar}
+This allows us to assign the same ciphertext type to $\mx{x}{1}$ as
+above-- that is, $\cty{\rx{x}{2}}{\sx{x}{2}}$.
 
@@
 \end{verbatimtab}
+
+$$
+\select{x}{y}{z}
+$$

### 2024-08-26 csf25/ctj.tex
checkpoint

@@
 $$
-\mx{x}{1} : \cty{\rx{x}{2}}{\sx{x}{2}}
+\mx{x}{1} : \cty{\rx{x}{2}}{\setit{\sx{x}{2}}}
 $$
-indicating that $\mx{x}{1}$ is an encryption of a value
-dependent on $\sx{x}{2}$ with $\rx{x}{2}$ and as long as client 1
-does not subsequently receive messages that are dependent on $\rx{x}{2}$, the
-value $\sx{x}{2}$ remains hidden.
+indicating syntactically that $\mx{x}{1}$ is an encryption of a
+value dependent on $\sx{x}{2}$ with $\rx{x}{2}$ and as long as client
+1 does not subsequently receive messages that are dependent on
+$\rx{x}{2}$, the value $\sx{x}{2}$ remains hidden.
 
-Of course, there are other methods for encrypting values
-in MPC protocols-- but we can observe that many are algebraically
-equivalent to this fundamental schema. For example, in
-Yao's Garbled Circuits (YGC) in $\mathbb{F}_{2}$, the ``garbler'' represents
-encrypted wire values as random samples, and the ``evaluator''
-obtains the wire value for its input via OT. The evaluator shares 
-the initial encrypted wire value of one of its input secrets as
-a selection bit-- assuming that client
-2 is the garbler and client 1 is the evaluator, we can
-write this as:
+Of course, there are other methods for encrypting values in MPC
+protocols-- but we can observe that many are algebraically equivalent
+to this fundamental schema. For example, in Yao's Garbled Circuits
+(YGC) in $\mathbb{F}_{2}$, the ``garbler'' represents encrypted wire
+values as a random sample (for 1) or its negation (for 0), and the
+``evaluator'' obtains the wire value for its input via OT. The
+evaluator shares the initial encrypted wire value of one of its input
+secrets as a selection bit-- assuming that client 2 is the garbler and
+client 1 is the evaluator, we can write this as:
 $$
@@
 %\toeq{\xassign{\mx{x}{1}}{\OT{\secret{x}}{1}{\neg\flip{x}}{\flip{x}}}{2}}
-   \toeq{\prog} \models \mx{x}{1} \eop \sx{x}{2} \fplus \rx{x}{2}
+   \toeq{\prog} \models \mx{x}{1} \eop \neg\sx{x}{2} \fplus \rx{x}{2}
 \end{mathpar}
 This allows us to assign the same ciphertext type to $\mx{x}{1}$ as
-above-- that is, $\cty{\rx{x}{2}}{\sx{x}{2}}$.
+above-- that is, $\cty{\rx{x}{2}}{\setit{\sx{x}{2}}}$. Leveraging the
+equivalence this way has been previously well-studied
+\cite{barthe2019probabilistic}.  Since this is a one-time-pad scheme,
+it is also important that we ensure that samples are used at most once
+for encryption. As in previous work \cite{darais2019language} we can
+use a form of type linearity to this end.
 
-$$
-\begin{array}{rcl@{\hspace{2mm}}r}
-  t &::=& x \mid \cty{x}{T} \\
-  \ty & \in & 2^{t}\\
-  \Gamma &::=& \varnothing \mid \Gamma; x : \ty
-\end{array} 
-$$
-
+The syntax of types $\ty$ and type environments $\Gamma$ are presented
+in Figure \ref{fig-cpj}\cnote{need to fix set notations}. We also
+define type judgements for expressions and programs by direct
+interpretation as terms $\phi$ and constraints $\eqs$ respectively,
+via the encoding $\toeq{cdot}$. The \TirName{Encode} rule captures
+the use ``fresh'' samples for encryption discussed above. Derivations
+enforce linearity by requiring that variables used for encryption
+are added to judgements, and guaranteed to not be used elsewhere
+via a form of union that ensures disjointness of combined sets. 
 \begin{definition}
@@
 \end{definition}
+The \TirName{DepTy} rule handles all other syntactic forms and just
+records the dependencies on variables in the term.  The \TirName{Send}
+rule records the types of messages, reveals, and outputs in
+enviroments, and the \TirName{Seq} rule combines the types of
+sequenced programs.  Given this we define validity of program type
+judgements as follows. \cnote{Define $\eqspre$.}
+\begin{definition}
+  Given preprocessing predicate $\eqspre$ and protocol $\prog$ we say
+  $\cpj{R}{\eqs}{\eqspre \wedge \toeq{\prog}}{\Gamma}$ is \emph{valid} iff it is derivable and
+  $\eqspre \wedge \toeq{\prog} \models \eqs$.
+\end{definition}
+It is important to note that this presentation of derivations is
+logical, not algorithmic-- in particular it is not syntax-directed
+due to the \TirName{Encode} rule.
 
+\begin{fpfig}[t]{Syntax and Derivation Rules for $\minicat$ Confidentiality Types}{fig-cpj}
+\small{
+$$
+\begin{array}{rcl@{\hspace{3mm}}l}
+  t &::=& x \mid \cty{x}{T} \\
+  \ty & \in & 2^{t} & \gdesc{types}\\
+  \Gamma &::=& \varnothing \mid \Gamma; x : \ty & \gdesc{type environments}
+\end{array} 
+$$
+\medskip
 \begin{mathpar}
@@
 \end{mathpar}
+}
+\end{fpfig}
 
-\begin{definition}
-  Given preprocessing predicate $\eqspre$ and protocol $\prog$ we say
-  $\cpj{R}{\eqs}{\eqspre \wedge \toeq{\prog}}{\Gamma}$ is \emph{valid} iff it is derivable and
-  $\eqspre \wedge \toeq{\prog} \models \eqs$.
-\end{definition}
-
+\begin{fpfig}[t]{Dependencies in Corrupt Views: Derivation Rules}{fig-leakj}
+\small{
 \begin{mathpar}
@@
 \end{mathpar}
+}
+\end{fpfig}
 
@@
 
+\begin{comment}
 \subsection{Examples}
@@
 \end{verbatimtab}
-
-$$
-\select{x}{y}{z}
-$$
+\end{comment}

### 2024-08-30 ppdp24/slides.tex
checkpoint

% Inbuilt themes in beamer
\documentclass{beamer}

% Theme choice:
\usetheme{CambridgeUS}
\usecolortheme{default}

\input{slide-macros}

% Title page details: 
\title{Language-Based Security for Low-Level MPC}

\author{Christian Skalka and Joe Near}

\date{September 10, 2024}

\logo{
\includegraphics[width=3cm]{UVM_Logo_Primary_Horiz_G.png}
}

\beamertemplatenavigationsymbolsempty

\begin{document}

% Title page frame
\begin{frame}
    \titlepage 
\end{frame}

% Remove logo from the next slides
\logo{}

% % Outline frame
% \begin{frame}{Outline}
%     \tableofcontents
% \end{frame}

\section{Ideal Functionality}
\begin{frame}{Ideal Functionality}
  $$
  \idealf(\sx{1}{1},\sx{2}{2},\sx{3}{3}) =
  \sx{1}{1}\ \exor\ \sx{2}{2}\ \exor\ \sx{3}{3}
  $$
  Suppose the output of the protocol is 1 and that party 3 is corrupted, so
  the adversary knows that $\sx{3}{3} = 0$.

  \medskip

  This unavoidably declassifies some information about the secrets of
  parties 1 and 2-- either:
  $$
  \sx{1}{1} = 0 \wedge \sx{2}{2} = 1
  $$
  or:
  $$
  \sx{1}{1} = 1 \wedge \sx{2}{2} = 0
  $$
  with equal probability.
\end{frame}

\section{Conditional Noninterference}
\begin{frame}{Conditional Noninterference}

  \begin{block}{Bounded Declassification}
    The ideal functionality defines a declassification ``limit''. Any protocol $\prog$
    should not reveal any more through broadcasts and DMs.
  \end{block}

  \begin{block}{Adversarial Views}
    In a run of the protocol, in addition to the output and secrets of corrupt
    parties, the adversary has access to all broadcasts and messages received
    by corrupt parties.
  \end{block}

  \begin{alertblock}{Conditional Noninterference}
    The probability of honest secrets conditioned on the protocol output
    should not be changed by conditioning on adversarial views. 
  \end{alertblock}

\end{frame}

\section{Secret Sharing}
\begin{frame}{Secret Sharing}

  \begin{block}{Reconstructive Sharing}
    $$
    \begin{array}{lll}
      \elab{\mesg{s1}}{2} &:=& \elab{(\secret{1} \fminus \locflip \fminus \flip{x})}{1} \\ 
      \elab{\mesg{s1}}{3} &:=& \elab{\flip{x}}{1} \\ 
      \elab{\mesg{s2}}{1} &:=& \elab{(\secret{2} \fminus \locflip \fminus \flip{x})}{2} \\ 
      \elab{\mesg{s2}}{3} &:=& \elab{\flip{x}}{2} \\ 
      \elab{\mesg{s3}}{1} &:=& \elab{(\secret{3} \fminus \locflip \fminus \flip{x})}{3} \\ 
      \elab{\mesg{s3}}{2} &:=& \elab{\flip{x}}{3}
    \end{array}
    $$
  \end{block}

\end{frame}

% Lists frame
\section{Lists in Beamer}
\begin{frame}{Lists in Beamer}

This is an unordered list:
\begin{itemize}
    \item Item 1
    \item Item 2
    \item Item 3
\end{itemize}

and this is an ordered list:
\begin{enumerate}
    \item Item 1
    \item Item 2
    \item Item 3
\end{enumerate}

\end{frame}

% Blocks frame
\section{Blocks in Beamer}
\begin{frame}{Blocks in Beamer}
    \begin{block}{Standard Block}
        This is a standard block.
    \end{block}
    \begin{alertblock}{Alert Message}
        This block presents alert message.
    \end{alertblock}
    \begin{exampleblock}{An example of typesetting tool}
        Example: MS Word, \LaTeX{}
    \end{exampleblock}
\end{frame}

\end{document}

### 2024-08-30 ppdp24/slides.tex
checkpoint

@@
 % \end{frame}
+
+\section{What's an MPC Protocol}
+\begin{frame}
+
+  A \cemph{protocol} $\prog$ implements an \emph{ideal functionality} $\idealf$,
+  defined in terms of operations in a prime field $\mathbb{F}_p$, where
+  the inputs are the secrets of clients.
+
+  \medskip
+
+  Low-level, probabilistic protocols are defined in terms of basic variable types:
+  \begin{itemize}
+  \item $\secret{x}$: secret values in $\mathbb{F}_p$.
+  \item $\mesg{x}$: messages received during the protocol.
+  \item $\reveal{x}$: broadcast messages.
+  \item $\flip{x}$: a uniformly random sample from $\mathbb{F}_p$ (implemented with
+    a random tape semantics).
+  \item $\pubout$: a protocol output
+  \end{itemize}
+  We can localize variables and expressions to \emph{clients}, and any information
+  exchanged between clients must be through messaging.
+  \begin{block}{Example}
+    $$
+    \mx{x}{1} := \elab{(\secret{x} - \flip{x})}{2}
+    $$
+  \end{block}
+\end{frame}
 

### 2024-08-30 ppdp24/slides.tex
checkpoint

--- 2024-08-30 ppdp24/slides.tex
@@
 
-\section{Ideal Functionality}
-\begin{frame}{Ideal Functionality}
+\section{What's an MPC Protocol}
+
+\begin{frame}{The Ideal Functionality}
+  A \emph{protocol} $\prog$ implements an \emph{ideal functionality} $\idealf$,
+  defined in terms of operations in a prime field $\mathbb{F}_p$.
+  \begin{itemize}
+  \item The inputs are \cemph{secret values} of \cemph{clients} in a \emph{federation}
+    $\setit{1,...,j}$.
+  \item The \cemph{output} is made publicly available by $\prog$
+  \end{itemize}
+  The output may leak secret information to an \cemph{adversary} trying to
+  guess secret values.  Consider:
+  $$
+  \idealf(x,y) = x * y
+  $$
+  If the output is odd, the adversary knows both secrets must be.  
+\end{frame}
+
+\begin{frame}{Protocol Implementation}
+
+  Low-level, probabilistic protocols are defined in terms of basic variable types:
+  \begin{itemize}
+  \item $\secret{x}$: secret values in $\mathbb{F}_p$.
+  \item $\mesg{x}$: messages received during the protocol.
+  \item $\flip{x}$: uniformly random samples from $\mathbb{F}_p$ (implemented with
+    a random tape semantics).
+  \end{itemize}
+  We can localize variables and expressions to \emph{clients}, and any information
+  exchanged between clients must be through messaging.
+  \begin{exampleblock}{Toy Example}
+    $$
+    \begin{array}{rcl@{\qquad}l}
+      \mx{x}{1} &:=& \elab{(\secret{x} - \flip{x})}{2}  & \textit{(unicast)}\\
+      \mx{y}{*} &:=& \elab{(\flip{y} + \mesg{x})}{1}  & \textit{(broadcast)}
+    \end{array}
+    $$ 
+  \end{exampleblock}
+\end{frame}
+
+\begin{frame}{(Passive) Threat Model}
+  The adversary wants to guess secrets. In addition to knowing the public
+  output, the adversary may also corrupt some subset of clients $C$:
+  \begin{itemize}
+  \item adversary has access to the secrets of clients in $C$.
+  \item adversary has access to messages received by members of $C$ during execution of $\prog$,
+    aka \cemph{adversarial views}.
+  \end{itemize}
+
+  \begin{alertblock}{Bounded Declassification}
+    The ideal functionality defines a declassification ``limit'', any protocol $\prog$
+    should not reveal any more through adversarial views.
+  \end{alertblock}
+\end{frame}
+
+\begin{frame}{Declassification of the Ideal Functionality}
+
+  \begin{exampleblock}{Example: 3-Party xor}
   $$
@@
   $$
-  Suppose the output of the protocol is 1 and that party 3 is corrupted, so
-  the adversary knows that $\sx{3}{3} = 0$.
-
-  \medskip
-
-  This unavoidably declassifies some information about the secrets of
-  parties 1 and 2-- either:
+  Suppose in a run of $\prog$ the output of $\prog$ is 1 and $C = \{ 3 \}$ and
+  the adversary knows $\sx{3}{3} = 0$, and therefore either:
   $$
@@
   $$
-  with equal probability.
+  with equal probability. \cemph{Adversarial views from the run of $\prog$ should not
+    allow the adversary better guesses.}
+  \end{exampleblock}
+
 \end{frame}
 
-\section{Conditional Noninterference}
-\begin{frame}{Conditional Noninterference}
+\section{Conditional Probabilistic Noninterference}
+\begin{frame}{Conditional Probabilistic Noninterference}
 
-  \begin{block}{Bounded Declassification}
-    The ideal functionality defines a declassification ``limit''. Any protocol $\prog$
-    should not reveal any more through broadcasts and DMs.
-  \end{block}
-
-  \begin{block}{Adversarial Views}
-    In a run of the protocol, in addition to the output and secrets of corrupt
-    parties, the adversary has access to all broadcasts and messages received
-    by corrupt parties.
-  \end{block}
+  \begin{itemize}
+    \item The set of all protocol runs under all possible secrets and random tapes
+      induces a probability distribution of program variables.
+    \item Conditioning on the output and corrupt secrets =  declassification
+      bounds.
+    \item Conditioning on adversarial views tells us whether they leak information.
+  \end{itemize}
 
@@
 
+  \begin{itemize}
+  \item Sound for the standard \emph{passive simulator security} model.
+  \item A \emph{hyperproperty} that accommodates traditional PL-based security
+    mechanisms and theory. 
+  \end{itemize}
+
 \end{frame}
 
-\section{Secret Sharing}
+\section{Shamir Addition}
 \begin{frame}{Secret Sharing}
+
+  To implement 3-party summation (xor in $\mathbb{F}_2)$, we can use \emph{Shamir addition}.
+  Each party first generates \emph{reconstructive shares}.
 
@@
   \end{block}
+  \begin{itemize}
+  \item \emph{algebraically} we have $\elab{\mesg{s1}}{2} + \elab{\mesg{s1}}{3} + \elab{\locflip}{1}$, etc.
+  \item \emph{probabilistically} any 2 out of three shares are in a uniform random distribution.
+  \end{itemize}
 
@@
 
-% Lists frame
-\section{Lists in Beamer}
-\begin{frame}{Lists in Beamer}
+\begin{frame}{Public Reveal}
 
-This is an unordered list:
   \begin{itemize}
-    \item Item 1
-    \item Item 2
-    \item Item 3
+   \item \emph{algebraically} the sum of all shares = $\sx{1}{1} + \sx{2}{2} + \sx{3}{3}$    
   \end{itemize}
 
-and this is an ordered list:
-\begin{enumerate}
-    \item Item 1
-    \item Item 2
-    \item Item 3
-\end{enumerate}
+  \begin{block}{Broadcast Sum-of-Shares}
+    $$
+    \begin{array}{lll}
+      \mx{ss1}{*} &:=& \elab{(\locflip \fplus \mesg{s2} \fplus \mesg{s3})}{1} \\ 
+      \mx{ss2}{*} &:=& \elab{(\mesg{s1} \fplus \locflip \fplus \mesg{s3})}{2} \\
+      \mx{ss3}{*} &:=& \elab{(\mesg{s1} \fplus \mesg{s2} \fplus \locflip)}{3} 
+    \end{array}
+    $$
+  \end{block}
+
+  \begin{itemize}
+  \item Messages $\mx{ss1}$ through $\mx{ss3}$ are \emph{not} probabilistically independent of inputs
+    secrets, but ...
+  \item They \emph{are} independent of secrets conditioned on the output!
+  \end{itemize}
 
@@
 
-% Blocks frame
-\section{Blocks in Beamer}
-\begin{frame}{Blocks in Beamer}
-    \begin{block}{Standard Block}
-        This is a standard block.
+\begin{frame}{Output}
+
+  The output phase does not reveal additional information, but is
+  algebraically correct.
+  \begin{block}{Client Outputs}
+    $$
+    \begin{array}{lll}
+      \out{1} &:=& \elab{(\mesg{ss1} \fplus \mesg{ss2} + \mesg{ss3})}{1}\\
+      \out{2} &:=& \elab{(\mesg{ss1} \fplus \mesg{ss2} + \mesg{ss3})}{2}\\
+      \out{3} &:=& \elab{(\mesg{ss1} \fplus \mesg{ss2} + \mesg{ss3})}{3}
+    \end{array}
+    $$
   \end{block}
-    \begin{alertblock}{Alert Message}
-        This block presents alert message.
-    \end{alertblock}
-    \begin{exampleblock}{An example of typesetting tool}
-        Example: MS Word, \LaTeX{}
-    \end{exampleblock}
+
 \end{frame}

### 2024-08-31 ppdp24/slides.tex
checkpoint

@@
   $$
-  with equal probability. \cemph{Adversarial views from the run of $\prog$ should not
+  with \cemph{equal probability of .5}. \cemph{Adversarial views from the run of $\prog$ should not
     allow the adversary better guesses.}
@@
       bounds.
-    \item Conditioning on adversarial views tells us whether they leak information.
+    \item Conditioning on adversarial views tells us whether they leak information--
+      if it changes the distribution of honest secrets.
   \end{itemize}
@@
 
-\section{Shamir Addition}
-\begin{frame}{Secret Sharing}
+\section{Example: Shamir Addition}
+\begin{frame}{Example: Shamir Addition}
 
-  To implement 3-party summation (xor in $\mathbb{F}_2)$, we can use \emph{Shamir addition}.
+  To implement 3-party summation (xor in $\mathbb{F}_2$), we can use \emph{Shamir addition}.
   Each party first generates \emph{reconstructive shares}.
@@
 
-\begin{frame}{Public Reveal}
+\begin{frame}{Example contd.: Public Broadcast Reveal}
 
@@
 
-\begin{frame}{Output}
+\begin{frame}{Example contd.: Output}
 
@@
 
+  %\begin{exampleblock}{Example Bad Protocol}
+  %  If 1 is honest and accidentally broadcasts its local share:
+  %  $$
+  %  \mx{leak}{*} := \elab{\locflip}{1}
+  %  $$
+  %  conditioning on adversarial views now yields probability 1 for
+  %  either $\sx{1}{1} = 0 \wedge \sx{2}{2} = 1$ or
+  %  $\sx{1}{1} = 1 \wedge \sx{2}{2} = 0$.
+  %\end{exampleblock}
+
 \end{frame}

### 2024-09-02 ppdp24/slides.tex
checkpoint

@@
 \usecolortheme{default}
+
+\usepackage[style=verbose,backend=biber]{biblatex}
+\addbibresource{secure-computation-bibliography.bib}
 
@@
     $\setit{1,...,j}$.
-  \item The \cemph{output} is made publicly available by $\prog$
+  \item The \cemph{output} is made \emph{publicly available} by $\prog$
   \end{itemize}
   The output may leak secret information to an \cemph{adversary} trying to
-  guess secret values.  Consider:
+  guess secret values.
+
+  \begin{exampleblock}{Consider:}
   $$
@@
   $$
-  If the output is odd, the adversary knows both secrets must be.  
+  If the output is odd, the adversary knows both $x$ and $y$ must be, even
+  if $x$ and $y$ remain hidden.
+  \end{exampleblock}
+
 \end{frame}
 
-\begin{frame}{Protocol Implementation}
+\begin{frame}{Protocol Implementation in Overture}
 
-  Low-level, probabilistic protocols are defined in terms of basic variable types:
+  Low-level, probabilistic protocols are defined in terms of basic variable types
+  (identifiers $x$ distinguish them):
   \begin{itemize}
@@
   exchanged between clients must be through messaging.
-  \begin{exampleblock}{Toy Example}
+  \begin{exampleblock}{Toy Example:}
     $$
@@
   The adversary wants to guess secrets. In addition to knowing the public
-  output, the adversary may also corrupt some subset of clients $C$:
+  output, the adversary may also corrupt \emph{any} subset of clients $C$:
   \begin{itemize}
@@
 
+\section{Models of a Protocol}
+
+\begin{frame}{Computing Protocol Distributions}
+
+  The probability distribution of a protocol $\prog$ (as a pmf) can be precisely computed
+  by interpretation as a \cemph{constraint program}.
+  \begin{itemize}
+  \item Interpretation as a Datalog program-- set of Least Herbrand Models.
+    \begin{itemize}
+      \item Random tapes represented as fact base.
+    \end{itemize}
+  \item Interpretation as SMT constraints \footnote{Ozdemir, A., Kremer, G., Tinelli, C., Barrett, C.
+(2023).
+\emph{Satisfiability Modulo Finite Fields}.
+In: Enea, C., Lal, A.
+(eds) Computer Aided Verification.
+CAV 2023.
+}-- set of all models.
+  \end{itemize}
+  Set of models can be queried to enforce hyperproperties-- but poor scaling.
+
+  \begin{alertblock}{Hypothesis: Composable Properties}
+    Compositional properties of protocol components can be independently verified, tractably.
+  \end{alertblock}
+
+\end{frame}
+
 \end{document}

### 2024-09-02 ppdp24/slides.tex
checkpoint

@@
 % \end{frame}
+
+\section{Overview}
+
+\begin{frame}{Background}
+
+  \cemph{Secure Multiparty Computation (MPC)} is a paradigm for
+  \cemph{confidential computing}.
+  \begin{itemize}
+  \item Guarantees data privacy.
+  \item Software-based protocols (vs., e.g., enclaves).
+  \item No trusted 3rd party.
+  \end{itemize}
+  Many existing deployments: voting, auctions, satellite trajectories, COVID exposure notification, ...\footnote{https://mpc.cs.berkeley.edu}
+
+  \medskip
+
+  Various language frameworks, including MPyC, Fairplay, Viaduct.
+\end{frame}
+
+\begin{frame}{Contributions}
+
+  Existing HLLs rely on low-level MPC protocols with correctness
+  assumed.
+  \begin{itemize}
+  \item Low-level protocols possess distinct features, e.g., probabilistic sampling.
+  \item Existing proof techniques manual, error prone, complex security model.
+  \end{itemize}
+  Our contributions:
+  \begin{itemize}
+  \item Language design for low-level protocol definitions.
+  \item (Re)formulation of security mode as \cemph{hyperoperties} (e.g., noninterference).
+  \item Automated proof techniques.
+  \end{itemize}
+
+\end{frame}
 
@@
     \end{itemize}
-  \item Interpretation as SMT constraints \footnote{Ozdemir, A., Kremer, G., Tinelli, C., Barrett, C.
+  \item Interpretation as SMT constraints\footnote{Ozdemir, A., Kremer, G., Tinelli, C., Barrett, C.
 (2023).

### 2024-09-02 ppdp24/slides.tex
checkpoint

@@
   \begin{itemize}
-  \item Interpretation as a Datalog program-- set of Least Herbrand Models.
+  \item $\mathbb{F}_2$ interpretation as a Datalog program-- set of Least Herbrand Models.
     \begin{itemize}
@@
     \end{itemize}
-  \item Interpretation as SMT constraints\footnote{Ozdemir, A., Kremer, G., Tinelli, C., Barrett, C.
+  \item $\mathbb{F}_p$ interpretation as SMT constraints\footnote{Ozdemir, A., Kremer, G., Tinelli, C., Barrett, C.
 (2023).

### 2024-09-02 ppdp24/slides.tex
checkpoint

@@
   \item Language design for low-level protocol definitions.
-  \item (Re)formulation of security mode as \cemph{hyperoperties} (e.g., noninterference).
+  \item (Re)formulation of security mode as \cemph{hyperproperties} (e.g., noninterference).
   \item Automated proof techniques.
@@
 
+\begin{frame}{Circuits and Gates}
+
+  Low-level protocols such as YGC and GMW implement $\idealf$ as a \cemph{circuit}.
+  \begin{itemize}
+  \item Composed by combing-- aka ``wiring''-- \cemph{gates}.
+  \item Each gate standard definition, supports primitive $\fplus$ and $\ftimes$.
+  \item Two input wires, one output wire, values remain encrypted on wire. 
+  \end{itemize}
+
+\end{frame}
+
 \end{document}

### 2024-09-02 ppdp24/slides.tex
checkpoint

@@
 
+\section{Compositional Properties of Circuits}
+
 \begin{frame}{Circuits and Gates}
@@
 
+  \begin{exampleblock}{2-Party GMW}
+    Rep.~invariant: each client in $\setit{1,2}$ holds a share of wire values.
+    \begin{itemize}
+    \item $\ttt{andgmw}(x,y,z)$:
+      \begin{itemize}
+      \item Input shares $\mx{x}{1}$, $\mx{x}{2}$ and $\mx{y}{1}$, $\mx{y}{2}$--
+        wire value for $x$ is $\mx{x}{1} + \mx{x}{2}$, etc.
+      \item Output shares $\mx{z}{1}$, $\mx{z}{2}$ received during gate execution.
+      \end{itemize}
+    \end{itemize}
+
+  \end{exampleblock}
+
+\end{frame}
+
+\begin{frame}
+
+  \begin{exampleblock}{2-Party GMW (Contd.)}
+    \begin{itemize}
+    \item $\ttt{andgmw}(x,y,z)$:
+      \begin{itemize}
+      \item Picking dummy for $x,y,z$, we can automatically generate a
+        pmf under all possible wire inputs.
+      \item Pmf can be queried to verify critical properties-- e.g., the
+        uniform distribution of output shares under any input conditions. 
+      \end{itemize}
+      Once properties of gates are verified, we can show that rep.~invariants
+      are maintained in any circuit.
+    \end{itemize}
+
+  \end{exampleblock}
+
 \end{frame}

### 2024-09-02 ppdp24/slides.tex
checkpoint

@@
       are maintained in any circuit.
+      \begin{itemize}
+      \item Leverage logics for reasoning about probabilistic programming: Probabilistic
+        Separation Logic, Lilac.
+      \item Manual proofs, but automated tactics and applicability to any circuit.
+      \end{itemize}
     \end{itemize}
@@
 
+\section{Conclusion and Future Work}
+
+\logo{
+\includegraphics[width=3cm]{UVM_Logo_Primary_Horiz_G.png}
+}
+
+\begin{frame}{Future Work}
+
+  \begin{itemize}
+  \item Generalize verification to arbitrary prime fields.
+  \item Automated type system to eliminate manual proof elements.
+  \item Support for concurrency.
+  \end{itemize}
+
+\end{frame}
+
 \end{document}

### 2024-09-02 ppdp24/slides.tex
checkpoint

@@
   \begin{itemize}
-  \item Messages $\mx{ss1}$ through $\mx{ss3}$ are \emph{not} probabilistically independent of inputs
-    secrets, but ...
+  \item Messages $\mx{ss1}$ through $\mx{ss3}$ are \emph{not} probabilistically independent of inputs, but ...
   \item They \emph{are} independent of secrets conditioned on the output! 
@@
 
+  \begin{alertblock}
+    Any $\mx{ss1}{*},\mx{ss2}{*},\mx{ss3}{*}$ such that:
+    $$
+    \mx{ss1}{*} + \mx{ss2}{*} + \mx{ss3}{*} = \sx{1}{1} + \sx{2}{2}  + \sx{3}{3}
+    $$
+    are \cemph{equally likely}.
+  \end{alertblock}
   The output phase does not reveal additional information, but is

### 2024-09-02 ppdp24/slides.tex
checkpoint

@@
 
-  \begin{alertblock}
-    Any $\mx{ss1}{*},\mx{ss2}{*},\mx{ss3}{*}$ such that:
+  \begin{itemize}
+    \item Any $\mx{ss1}{*},\mx{ss2}{*},\mx{ss3}{*}$ such that:
     $$
-    \mx{ss1}{*} + \mx{ss2}{*} + \mx{ss3}{*} = \sx{1}{1} + \sx{2}{2}  + \sx{3}{3}
+    \mx{ss1}{*} + \ \mx{ss2}{*} + \ \mx{ss3}{*} = \sx{1}{1} + \sx{2}{2}  + \sx{3}{3}
     $$
     are \cemph{equally likely}.
-  \end{alertblock}
-  The output phase does not reveal additional information, but is
-  algebraically correct.
+  \end{itemize}
+
   \begin{block}{Client Outputs}

### 2024-09-03 ppdp24/slides.tex
checkpoint

@@
   \item (Re)formulation of security mode as \cemph{hyperproperties} (e.g., noninterference).
-  \item Automated proof techniques.
+  \item Semi-automated proof techniques.
+    \begin{itemize}
+    \item Automatically verify small program components, manually combine for whole-program proofs. 
+    \end{itemize}
   \end{itemize}
@@
 
-  \begin{exampleblock}{Example: 3-Party xor}
+  \begin{exampleblock}{Example: 3-Party xor in $\mathbb{F}_2$}
   $$
@@
   \begin{itemize}
-  \item Messages $\mx{ss1}$ through $\mx{ss3}$ are \emph{not} probabilistically independent of inputs, but ...
-  \item They \emph{are} independent of secrets conditioned on the output!
+  \item Messages $\mx{ss1}$ through $\mx{ss3}$ are \cemph{not probabilistically independent of inputs}, but ...
+  \item They \emph{are} \cemph{independent of secrets conditioned on the output}!
   \end{itemize}
@@
 \section{Compositional Properties of Circuits}
+
+\begin{frame}{Prelude and Component Definitions}
+
+  The Prelude language allows modular definition of Overture protocol components.
+  \begin{itemize}
+  \item Function definitions, function calls, structured data.
+  \item A \emph{staged} language that generates a complete Overture protocol
+    as a side effect.
+  \end{itemize}
+  In addition to programmer convenience, Prelude supports verification:
+  \begin{itemize}
+  \item Prelude functions define Overture (sub)protocols that can be automatically verified.
+  \item Large protocols can be verified manually by composing function properties. 
+  \end{itemize}
+
+  \begin{exampleblock}{Example: Summing Shares}
+    $$
+    \ttt{sum}(x,y,z,i) \{ \mx{z}{i} := \elab{\mesg{x} \fplus \mesg{y}}{i} \} 
+    $$
+  \end{exampleblock}
+\end{frame}
 

### 2024-09-03 ppdp24/slides.tex
checkpoint

@@
 
-  \begin{exampleblock}{Example: 3-Party xor in $\mathbb{F}_2$}
+  \begin{exampleblock}{Example: 3-Party $\fplus$ in $\mathbb{F}_2$ (where $\fplus$ is $\exor$ and $\ftimes$ is $\wedge$)}
   $$
   \idealf(\sx{1}{1},\sx{2}{2},\sx{3}{3}) =
-  \sx{1}{1}\ \exor\ \sx{2}{2}\ \exor\ \sx{3}{3}
+  \sx{1}{1}\ \fplus\ \sx{2}{2}\ \fplus\ \sx{3}{3}
   $$
   Suppose in a run of $\prog$ the output of $\prog$ is 1 and $C = \{ 3 \}$ and
-  the adversary knows $\sx{3}{3} = 0$, and therefore either:
+  the adversary knows $\sx{3}{3} = 0$.
+  Therefore the adversary knows that either:
   $$
-  \sx{1}{1} = 0 \wedge \sx{2}{2} = 1
+  \sx{1}{1} = 0 \text{\ and\ } \sx{2}{2} = 1
   $$
@@
   $$
-  \sx{1}{1} = 1 \wedge \sx{2}{2} = 0
+  \sx{1}{1} = 1 \text{\ and\ } \sx{2}{2} = 0
   $$
-  with \cemph{equal probability of .5}. \cemph{Adversarial views from the run of $\prog$ should not
+  with equal probability of .5. \cemph{Adversarial views from the run of $\prog$ should not
     allow the adversary better guesses.}
@@
   \end{itemize}
+
+  \begin{exampleblock}{Example: Local Summation}
+    $$
+    \ttt{sum}(x,y,z,i) \{ \mx{z}{i} := \elab{(\mesg{x} \fplus \mesg{y})}{i} \} 
+    $$
+  \end{exampleblock}
+
   In addition to programmer convenience, Prelude supports verification:
@@
 
-  \begin{exampleblock}{Example: Summing Shares}
-    $$
-    \ttt{sum}(x,y,z,i) \{ \mx{z}{i} := \elab{\mesg{x} \fplus \mesg{y}}{i} \} 
-    $$
-  \end{exampleblock}
 \end{frame}
@@
 
-  \begin{exampleblock}{2-Party GMW}
+  \begin{exampleblock}{Example: 2-Party GMW in $\mathbb{F}_2$}
     Rep.~invariant: each client in $\setit{1,2}$ holds a share of wire values.
@@
       \item Output shares $\mx{z}{1}$, $\mx{z}{2}$ received during gate execution.
+        $$
+        \mx{z}{1} + \mx{z}{2} = (\mx{x}{1} + \mx{x}{2}) * (\mx{y}{1} + \mx{y}{2})
+        $$
       \end{itemize}
     \end{itemize}    
-
   \end{exampleblock}
@@
       \item Picking dummy for $x,y,z$, we can automatically generate a
-        pmf under all possible wire inputs.
+        generic ``and'' gate $\prog_{\wedge}$ in Overture
+      \item Automatically calculate the pmf of $\prog_{\wedge}$.
       \item Pmf can be queried to verify critical properties-- e.g., the

### 2024-09-03 ppdp24/slides.tex
checkpoint

@@
   $$
-  with equal probability of .5. \cemph{Adversarial views from the run of $\prog$ should not
+  with equal joint probability of .5. \cemph{Adversarial views from the run of $\prog$ should not
     allow the adversary better guesses.}
@@
     Rep.~invariant: each client in $\setit{1,2}$ holds a share of wire values.
-    \begin{itemize}
-    \item $\ttt{andgmw}(x,y,z)$:
+
+    \medskip
+
+    $\ttt{andgmw}(x,y,z)$:
+
     \begin{itemize}
@@
     \end{itemize}
-    \end{itemize}    
+
   \end{exampleblock}
@@
   \begin{exampleblock}{2-Party GMW (Contd.)}
+
+    Picking dummy $x,y,z$, we can execute $\ttt{andgmw}(x,y,z)$
+    to generate a generic ``and'' gate $\prog_{\wedge}$ in Overture.
     \begin{itemize}
-    \item $\ttt{andgmw}(x,y,z)$:
-      \begin{itemize}
-      \item Picking dummy for $x,y,z$, we can automatically generate a
-        generic ``and'' gate $\prog_{\wedge}$ in Overture
     \item Automatically calculate the pmf of $\prog_{\wedge}$.
@@
     \end{itemize}
-      Once properties of gates are verified, we can show that rep.~invariants
+    Once properties of gates are verified, we can prove that rep.~invariants
     are maintained in any circuit.
     \begin{itemize}
+    \item Manual proofs, but automated tactics and applicability to any circuit.
     \item Leverage logics for reasoning about probabilistic programming: Probabilistic
       Separation Logic, Lilac.
-      \item Manual proofs, but automated tactics and applicability to any circuit.
-      \end{itemize}
     \end{itemize}

### 2024-09-03 ppdp24/slides.tex
checkpoint

@@
     aka \cemph{adversarial views}.
+  \item adversary cannot break the rules of the protol\footnote{Malicious security also
+  considered separately in our work.}.
   \end{itemize}
@@
 
-  To implement 3-party summation (xor in $\mathbb{F}_2$), we can use \emph{Shamir addition}.
+  To implement 3-party summation in any $\mathbb{F}_p$, we can use \emph{Shamir addition}.
   Each party first generates \emph{reconstructive shares}.
@@
   \begin{itemize}
-  \item \emph{algebraically} we have $\elab{\mesg{s1}}{2} + \elab{\mesg{s1}}{3} + \elab{\locflip}{1}$, etc.
+  \item $\elab{\mesg{s1}}{2}, \elab{\mesg{s1}}{3}, \elab{\locflip}{1}$ are the \emph{shares} of $\sx{1}{1}$,
+    etc.
+  \item \emph{algebraically} $\sx{1}{1} = \elab{\mesg{s1}}{2} + \elab{\mesg{s1}}{3} + \elab{\locflip}{1}$.
   \item \emph{probabilistically} any 2 out of three shares are in a uniform random distribution.
@@
 
-  \begin{exampleblock}{2-Party GMW (Contd.)}
+  \begin{exampleblock}{Semi-Automatically Verifying 2-Party GMW}
 

### 2024-09-08 ppdp24/slides.tex
checkpoint

@@
   \item Language design for low-level protocol definitions.
-  \item (Re)formulation of security mode as \cemph{hyperproperties} (e.g., noninterference).
+  \item (Re)formulation of security model as \cemph{hyperproperties} (e.g., noninterference).
   \item Semi-automated proof techniques.
@@
   \end{itemize}
-  The output may leak secret information to an \cemph{adversary} trying to
+  \cemph{The output may leak secret information to an adversary} trying to
   guess secret values.
@@
     aka \cemph{adversarial views}.
-  \item adversary cannot break the rules of the protol\footnote{Malicious security also
+  \item adversary cannot break the rules of the protocol\footnote{Malicious security also
   considered separately in our work.}.
@@
 
-  \begin{exampleblock}{Example: 3-Party $\fplus$ in $\mathbb{F}_2$ (where $\fplus$ is $\exor$ and $\ftimes$ is $\wedge$)}
+  \begin{exampleblock}{Example: 3-Party $\fplus$} %in $\mathbb{F}_2$ (where $\fplus$ is $\exor$ and $\ftimes$ is $\wedge$)}
   $$
@@
 
-\section{Example: Shamir Addition}
-\begin{frame}{Example: Shamir Addition}
+\section{Example: Additive Sharing}
+\begin{frame}{Example: Additive Secret Shares}
 
-  To implement 3-party summation in any $\mathbb{F}_p$, we can use \emph{Shamir addition}.
+  To implement 3-party summation in any $\mathbb{F}_p$, we can use \emph{additive sharing}.
   Each party first generates \emph{reconstructive shares}.

### 2024-09-08 ppdp24/slides.tex
checkpoint

@@
 
-  In addition to programmer convenience, Prelude supports verification:
-  \begin{itemize}
-  \item Prelude functions define Overture (sub)protocols that can be automatically verified.
-  \item Large protocols can be verified manually by composing function properties. 
-  \end{itemize}
+  In addition to programmer convenience, Prelude supports function-level verification.
+  %Prelude supports verification:
+  %\begin{itemize}
+  %\item Prelude functions define Overture (sub)protocols that can be automatically verified.
+  %\item Large protocols can be verified manually by composing function properties. 
+  %\end{itemize}
 
@@
 
-\begin{frame}{Circuits and Gates}
+\begin{frame}{Circuits and Gate Paradigms}
 
-  Low-level protocols such as YGC and GMW implement $\idealf$ as a \cemph{circuit}.
+  Goldreich-Micali-Wigderson (GMW) protocol implements $\idealf$ as a \cemph{circuit}.
   \begin{itemize}
@@
     \end{itemize}
-    Once properties of gates are verified, we can prove that rep.~invariants
-    are maintained in any circuit.
+    Gate verification serves as a \emph{tactic} for manual circuit-level
+    proofs. 
     \begin{itemize}
-    \item Manual proofs, but automated tactics and applicability to any circuit.
     \item Leverage logics for reasoning about probabilistic programming: Probabilistic

### 2024-09-08 code/cvc5mvn/src/main/java/plaid/App.java
Prelude grammar

@@ -90,7 +90,21 @@ public class App
         System.out.println(r_unsat);
         */
         
-
+        // prelude demo
+        String program = """
+                         decodegmw(z) {
+                            p["1"] := m[z]@1; p["2"] := m[z]@2;
+                            out@1 := (p["1"] + p["2"])@1;
+                            out@2 := (p["1"] + p["2"])@2
+                         }
+                         """;
+        ANTLRInputStream input = new ANTLRInputStream(program);
+        PreludeLexer lexer = new PreludeLexer(input);
+        CommonTokenStream tokens = new CommonTokenStream(lexer);
+        PreludeParser parser = new PreludeParser(tokens);
+        parser.setBuildParseTree(true);
+        PreludeParser.ProgramContext pc = parser.program();
+        System.out.println(pc.toStringTree(parser));
         /*
         //String prog = "p[\"foo\"] * 12 + m[\"bar\"]";
         String prog = "m[\"foo\"]@2 := (s[\"x\"] + -r[\"y\"])@1";

### 2024-09-08 code/cvc5mvn/src/main/java/plaid/SolverExtension.java

@@ -21,12 +21,12 @@ public class SolverExtension {
         new ParseTreeWalker().walk(listener, tree);
         Iterable<Memory> memoryList = listener.memories();
 
-        System.out.println(memoryList);
+        //System.out.println(memoryList);
         OvertureConstraintListener overtureConstraintListener = new OvertureConstraintListener(termManager, prime, memoryList);
 
         // iterate over commands in the parse tree
         for (OvertureParser.CommandContext commandContext : tree.command()){
-            System.out.println(overtureConstraintListener.visit(commandContext));
+           // System.out.println(overtureConstraintListener.visit(commandContext));
             solver.assertFormula(overtureConstraintListener.visit(commandContext));
 
         }

### 2024-09-10 csf25/ctj.tex
checkpoint

@@
 
+\cpjfig
+
 The syntax of types $\ty$ and type environments $\Gamma$ are presented
@@
 
-\begin{fpfig}[t]{Syntax and Derivation Rules for $\minicat$ Confidentiality Types}{fig-cpj}
-\small{
+\subsection{Detecting Leakage Through Dependencies}
+
+\leakjfig
+
+Given a valid typing, we need to conservatively approximate the
+information accessible by the adversary, who will have control over
+all corrupt clients in $C$ given a partitioning $H,C$. In particular,
+if the adversary has access to an encypted value and another value
+that is dependent on its one-time-pad key-- i.e., access to values
+with types of the forms:
 $$
-\begin{array}{rcl@{\hspace{3mm}}l}
-  t &::=& x \mid \cty{x}{T} \\
-  \ty & \in & 2^{t} & \gdesc{types}\\
-  \Gamma &::=& \varnothing \mid \Gamma; x : \ty & \gdesc{type environments}
-\end{array} 
+\setit{\cty{\rx{x}{\cid}}{\ty}} \qquad \setit{\rx{x}{\cid}, \ldots} 
 $$
-\medskip
-\begin{mathpar}
-  \inferrule[DepTy]
-  {}
-  {\eqj{\varnothing}{\eqs}{\phi}{\vars(\phi)}}
-
-  \inferrule[Encode]
-  {\eqs \models \phi \eop \phi' \oplus \rx{w}{\cid} \\
-   \oplus \in \{ \fplus,\fminus \}\\
-   \eqj{R}{\eqs}{\phi'}{\ty}}
-  {\eqj{R;\{ \rx{w}{\cid} \}}{\eqs}{\phi}{\setit{\cty{\rx{w}{\cid}}{\ty}}}}
-\end{mathpar}
-
-\begin{mathpar}
-  \inferrule[Send]
-            {\eqj{R}{\eqs}{\phi}{\ty}}
-            {\cpj{R}{\eqs}{x \eop \phi}{(x : \ty)}}
-
-  \inferrule[Seq]
-            {\cpj{R_1}{\eqs}{\phi_1}{\Gamma_1}\\
-             \cpj{R_2}{\eqs}{\phi_2}{\Gamma_2}}
-            {\cpj{R_1;R_2}{\eqs}{\phi_1 \wedge \phi_2}{\Gamma_1;\Gamma_2}}
-\end{mathpar}
-}
-\end{fpfig}
-
-\begin{fpfig}[t]{Dependencies in Corrupt Views: Derivation Rules}{fig-leakj}
-\small{
-\begin{mathpar}
-  \inferrule
-      {\cid \in C}
-      {\leakj{\Gamma}{C}{\Gamma(\mx{w}{\cid})}}
-
-  \inferrule
-      {\leakj{\Gamma}{C}{T_1 \cup T_2}}
-      {\leakj{\Gamma}{C}{T_1}}
-
-  \inferrule
-      {\leakj{\Gamma}{C}{\setit{ \mx{w}{\cid} }}}
-      {\leakj{\Gamma}{C}{\Gamma(\mx{w}{\cid})}}
-
-  \inferrule
-      {\leakj{\Gamma}{C}{\setit{ \rx{w}{\cid} }} \\ \leakj{\Gamma}{C}{\setit{ \cty{\rx{w}{\cid}}{\ty} }} }
-      {\leakj{\Gamma}{C}{\ty}}
-\end{mathpar}
-}
-\end{fpfig}
+then we can conservatively assume that the adversary can decrypt the
+value of type $\ty$.
 

### 2024-09-10 ppdp24/slides.tex
checkpoint

@@
   A \emph{protocol} $\prog$ implements an \emph{ideal functionality} $\idealf$,
-  defined in terms of operations in a prime field $\mathbb{F}_p$.
+  defined in terms of operations $(+,-,*)$ in a prime field $\mathbb{F}_p$.
   \begin{itemize}
@@
 
-  \begin{exampleblock}{Example: 3-Party $\fplus$} %in $\mathbb{F}_2$ (where $\fplus$ is $\exor$ and $\ftimes$ is $\wedge$)}
+  \begin{exampleblock}{Example: 3-Party $\fplus$ in $\mathbb{F}_2$ (where $\fplus$ is $\exor$ and $\ftimes$ is $\wedge$)}
   $$
@@
       bounds.
-    \item Conditioning on adversarial views tells us whether they leak information--
-      if it changes the distribution of honest secrets.
+    \item Conditioning on adversarial views = distribution of honest secrets.
   \end{itemize}
@@
     should not be changed by conditioning on adversarial views.
-  \end{alertblock}
 
@@
   \item Sound for the standard \emph{passive simulator security} model.
-  \item A \emph{hyperproperty} that accommodates traditional PL-based security
-    mechanisms and theory. 
+  %\item A \emph{hyperproperty} that accommodates traditional PL-based security
+  %  mechanisms and theory. 
   \end{itemize}
+  \end{alertblock}
 

### 2024-09-11 code/cvc5mvn/src/main/java/plaid/App.java
updated Prelude grammar to get a '@i' appended to a function call

@@ -91,52 +91,16 @@ public class App
         */
         
         // prelude demo
-        // String program = """
-        //                 decodegmw(z) {
-        //                    p["1"] := m[z]@1; p["2"] := m[z]@2;
-        //                    out@1 := (p["1"] + p["2"])@1;
-        //                    out@2 := (p["1"] + p["2"])@2
-        //                 }
-        //                 """;
-
         String program =
-//                """
-//                in(x, y) { x ++ y }
-//
-//                bar(x, y) { m[foo(x,y)] }
-//
-//                baz(x, y, i) { bar(x,y) := (s[foo(x,y)] + r[foo(y,x)])@i }
-//
-//                main() { baz("foo","bar",1); out@1 := bar("foo","bar")@1 }
-//                """;
                 """
-                        encodegmw(pin, i1, i2) {
-                        m[pin]@i2 := (s[pin] + r[pin])@i1;
-                        m[pin]@i1 := r[pin]@i1
-                        }
-                        andtablegmw(x, y, z) {
-                        let r11 = r[z] + (m[x] + true) * (m[y] + true) in
-                        let r10 = r[z] + (m[x] + true) * (m[y] + false) in
-                        let r01 = r[z] + (m[x] + false) * (m[y] + true) in
-                        let r00 = r[z] + (m[x] + false) * (m[y] + false) in
-                        { row1 = r11; row2 = r10; row3 = r01; row4 = r00 }
-                        }
-                        
-                        decodegmw(z) {
-                        p["1"] := m[z]@1; p["2"] := m[z]@2;
-                        out@1 := (p["1"] + p["2"])@1;
-                        out@2 := (p["1"] + p["2"])@2
-                        }
-                        main() {
-                        encodegmw("x",2,1);
-                        encodegmw("y",2,1);
-                        encodegmw("z",1,2);
-                        andgmw("g1","x","z");
-                        xorgmw("g2","g1","y");
-                        decodegmw("g2")
-                        }
-                        """;
+                foo(x, y) { x ++ y }
 
+                bar(x, y) { m[foo(x,y)] }
+
+                baz(x, y, i) { bar(x,y)@i := (s[foo(x,y)] + r[foo(y,x)])@i }
+
+                main() { baz("foo","bar",1); out@1 := bar("foo","bar")@1 }
+                """;
 
         ANTLRInputStream input = new ANTLRInputStream(program);
         PreludeLexer lexer = new PreludeLexer(input);

### 2024-09-30 esop25/cpj.tex
checkpoint

--- 2024-09-10 csf25/ctj.tex
@@
 
-\leakjfig
-
-Given a valid typing, we need to conservatively approximate the
-information accessible by the adversary, who will have control over
-all corrupt clients in $C$ given a partitioning $H,C$. In particular,
-if the adversary has access to an encypted value and another value
-that is dependent on its one-time-pad key-- i.e., access to values
-with types of the forms:
+Given a valid typing, we can leverage structural type information to
+conservatively approximate the information accessible by the
+adversary who has control over all corrupt clients in $C$ given
+a partitioning $H,C$. In particular, if the adversary has access to an
+encypted value and another value that is dependent on its one-time-pad
+key-- i.e., access to values with types of the forms:
 $$
@@
 then we can conservatively assume that the adversary can decrypt the
-value of type $\ty$.
+value of type $\ty$ and thus have access to those values.
 
+In Figure \ref{fig-leakj} we define derivation rules for the
+``leakage'' judgement $\leakj{\Gamma}{C}{\ty}$ which intuitively says
+that given a protocol with type environment $\Gamma$ and corrupt
+clients $C$, the adversary may be able to observe dependencies on
+values of type $\ty$ in their adversarial messages.  The real aim of
+this relation is to determine if
+$\leakj{\Gamma}{C}{\setit{\sx{w}{\cid}}}$ with $\cid \in H$-- that is,
+if the adversary may observe dependencies on honest secrets.
+
+Our main type safety result combines typing and leakage judgements,
+and guarantees gradual release of protocols-- that is, that adversarial
+messages reflect no dependencies on honest secrets.
+\cnote{Need to define gradual release... should probably have a section
+  where we summarize hyperproperties.}
 \begin{theorem}
@@
 \end{theorem}
+
+\leakjfig
+
+Of course, it is important to recall from \cite{skalka-near-ppdp24}
+that gradual release does \emph{not} imply passive security, since it
+only concerns adversarial messages, but not public reveals of outputs
+where dependencies on input secrects can occur even in secure
+protocols.
 

### 2024-09-30 esop25/examples.tex
checkpoint

\section{Extended Examples}

\subsection{Confidentiality Examples}

\begin{verbatimtab}
andtableygc(g,x,y)
{
   let table = (~r[g],~r[g],~r[g],r[g])
   in permute4(r[x],r[y],table)
}

m[x]@1 := s2(s[x],r[x],~r[x])@2;
m[x]@1 as s[x]@2 xor r[x]@2;

// m[x]@1 : { c(r[x]@2, { s[x]@2 }) }

m[y]@1 := OT(s[y]@1,r[y],~r[y])@2;
m[y]@1 as s[y]@1 xor r[y]@2;

// m[y]@1 : { c(r[y]@2, { s[y]@1 }) }

m[ag]@1 := OT4(m[x]@1, m[y]@1, andtable(ag,r[x],r[y]))@2;
m[ag]@1 as  ~((r[x]@2 = m[x]@1) and (r[y]@2 = m[y]@1)) xor r[ag]@2;

// m[ag]@1 : { c(r[ag]@2, {r[x]@2, r[y]@2, m[x]@1,  m[y]@1} }

p[o] := OT2(m[ag]@1, perm2(r[ag],(false,true)))@2

// p[o] : { c(r[ag]@2, {r[x]@2, r[y]@2,  m[x]@1,  m[y]@1}), r[ag]@2  }

out@1 := p[o]@1

// out@1 == s[x] and s[y]
\end{verbatimtab}

\begin{verbatimtab}
    encodegmw(in, i1, i2) {
      m[in]@i2 := (s[in] xor r[in])@i1;
      m[in]@i1 := r[in]@i1
    }

    andtablegmw(x, y, z) {
      let r11 = r[z] xor (m[x] xor true) and (m[y] xor true) in
      let r10 = r[z] xor (m[x] xor true) and (m[y] xor false) in
      let r01 = r[z] xor (m[x] xor false) and (m[y] xor true) in
      let r00 = r[z] xor (m[x] xor false) and (m[y] xor false) in
      { row1 = r11; row2 = r10; row3 = r01; row4 = r00 }
    }

    andgmw(z, x, y) {
      let table = andtablegmw(x,y,z) in
      m[z]@2 := OT4(m[x],m[y],table,2,1);
      m[z]@2 as ~((m[x]@1 xor m[x]@2) and (m[y]@1 xor m[y]@2)) xor r[z]@1);
      m[z]@1 := r[z]@1
    }

    // and gate correctness postcondition
    {} andgmw { m[z]@1 xor m[z]@2 == (m[x]@1 xor m[x]@2) and (m[y]@1 xor m[y]@2) }

    // and gate type
    andgmw :
     Pi z,x,y .
     {}
     { { r[z]@1 },
       (m[z]@1 : { r[z]@1 }; m[z]@2 : {c(r[z]@1, { m[x]@1, m[x]@2, m[y]@1, m[y]@2 })} ),
       m[z]@1 xor m[z]@2 == (m[x]@1 xor m[x]@2) and (m[y]@1 xor m[y]@2)}

    xorgmw(z, x, y) {
      m[z]@1 := (m[x] xor m[y])@1; m[z]@2 := (m[x] xor m[y])@2;
    }

    decodegmw(z) {
      p["1"] := m[z]@1; p["2"] := m[z]@2;
      out@1 := (p["1"] xor p["2"])@1;
      out@2 := (p["1"] xor p["2"])@2
    }

    prot() {
      encodegmw("x",2,1);
      encodegmw("y",2,1);
      encodegmw("z",1,2);
      andgmw("g1","x","z");
      xorgmw("g2","g1","y");
      decodegmw("g2")
    }

    {} prot { out@1 == (s["x"]@1 and s["z"]@2) xor s["y"]@1 }
\end{verbatimtab}

\subsection{Integrity Examples}

\begin{verbatimtab}
  secopen(w1,w2,w3,i1,i2) {
      pre(m[w1++"m"]@i2 == m[w1++"k"]@i1 + (m["delta"]@i1 * m[w1++"s"]@i2 /\
          m[w1++"m"]@i2 == m[w1++"k"]@i1 + (m["delta"]@i1 * m[w1++"s"]@i2));
      let locsum =  macsum(macshare(w1), macshare(w2)) in
      m[w3++"s"]@i1 := (locsum.share)@i2;
      m[w3++"m"]@i1 := (locsum.mac)@i2;
      auth(m[w3++"s"],m[w3++"m"],mack(w1) + mack(w2),i1);
      m[w3]@i1 := (m[w3++"s"] + (locsum.share))@i1
  }

  _open(x,i1,i2){
    m[x++"exts"]@i1 := m[x++"s"]@i2;
    m[x++"extm"]@i1 := m[x++"m"]@i2;
    assert(m[x++"extm"] == m[x++"k"] + (m["delta"] * m[x++"exts"]));
    m[x]@i1 := (m[x++"exts"] + m[x++"s"])@i2
  }`

  _sum(z, x, y,i1,i2) {
      pre(m[x++"m"]@i2 == m[x++"k"]@i1 + (m["delta"]@i1 * m[x++"s"]@i2 /\
          m[y++"m"]@i2 == m[y++"k"]@i1 + (m["delta"]@i1 * m[y++"s"]@i2));
      m[z++"s"]@i2 := (m[x++"s"] + m[y++"s"])@i2;
      m[z++"m"]@i2 := (m[x++"m"] + m[y++"m"])@i2;
      m[z++"k"]@i1 := (m[x++"k"] + m[y++"k"])@i1;
      post(m[z++"m"]@i2 == m[z++"k"]@i1 + (m["delta"]@i1 * m[z++"s"]@i2)
  }

  sum(z,x,y) { _sum(z,x,y,1,2);_sum(z,x,y,2,1) }

  open(x) { _open(x,1,2); _open(x,2,1) }

  sum("a","x","d");
  open("d");
  sum("b","y","e");
  open("e");
  let xys =
      macsum(macctimes(macshare("b"), m["d"]),
             macsum(macctimes(macshare("a"), m["e"]),
                    macshare("c")))
  let xyk = mack("b") * m["d"] + mack("a") * m["e"] + mack("c")

  secopen("a","x","d",1,2);
    secopen("a","x","d",2,1);
    secopen("b","y","e",1,2);
    secopen("b","y","e",2,1);
    let xys =
      macsum(macctimes(macshare("b"), m["d"]),
             macsum(macctimes(macshare("a"), m["e"]),
                    macshare("c")))
    in
    let xyk = mack("b") * m["d"] + mack("d") * m["d"] + mack("c")               
    in
    secreveal(xys,xyk,"1",1,2);
    secreveal(maccsum(xys,m["d"] * m["e"]),
              xyk - m["d"] * m["e"],
              "2",2,1);
    out@1 := (p[1] + p[2])@1;
    out@2 := (p[1] + p[2])@2;
\end{verbatimtab}

### 2024-09-30 esop25/ipj.tex
checkpoint

\section{Integrity Types}

\subsection{BDOZ MAC Integrity}

In a malicious setting, detecting cheating by adding
information-theoretic secure MAC codes to shares is a fundamental
technique pioneered in systems such as BDOZ and SPDZ
\cite{SPDZ1,SPDZ2,BDOZ,10.1007/978-3-030-68869-1_3}.  These protocols
assume a pre-processing phase that distributes shares with MAC codes
to clients.  This integrates well with pre-processed Beaver Triples to
implement malicious secure, and relatively efficient, multiplication
\cite{evans2018pragmatic}.

A field value $v$ is secret shared among 2 clients in BDOZ with
accompanying MAC values.  Each client $\cid$ gets a pair of values
$(v_\cid,m_\cid)$, where $v_\cid$ is a share of $v$ reconstructed by
addition, i.e., $v = \fcod{v_1 \fplus v_2}$, and $m_\cid$ is a MAC of
$v_\cid$.  More precisely, $m_\cid = k + (k_\Delta * v_\cid)$ where
\emph{keys} $k$ and $k_\Delta$ are known only to $\cid' \ne \cid$ and
$k_\Delta$. The \emph{local key} $k$ is unique per MAC, while the
\emph{global key} $k_\Delta$ is common to all MACs authenticated by
$\cid'$. This is a semi-homomorphic encryption scheme that supports
addition of shares and multiplication of shares by a constant-- for
more details the reader is referred to Orsini
\cite{10.1007/978-3-030-68869-1_3}. We note that while we restrict
values $v$, $m$, and $k$ to the same field in this presentation for
simplicity, in general $m$ and $k$ can be in extensions of
$\mathbb{F}_p$.

As in \cite{skalka-near-ppdp24}, we use $\macgv{\mesg{w}}$ to refer to
secret-shared values reconstructed with addition, and by
convention each share of $\macgv{\mesg{w}}$ is represented as
$\elab{\mesg{w\ttt{s}}}{\cid}$, the MAC of which is represented as a
$\elab{\mesg{w\ttt{m}}}{\cid}$ for all $\cid$, and each client holds a
key $\elab{\mesg{w\ttt{k}}}{\cid}$ for authentication of the other's
share. Each client also holds a global key $\mesg{\ttt{delta}}$. For
any such share identified by string $w$, the BDOZ MAC scheme is defined 
by the equality predicate $\macbdoz{w}$:
$$
\macbdoz{w} \defeq
\mesg{w\ttt{m}} = \mesg{w\ttt{k}} \fplus \ttt{(}\mesg{\ttt{delta}} \ftimes
\mesg{w\ttt{s}}\ttt{)}
$$

\subsection{Typing and Integrity Labeling}

\ipjfig

The integrity of values during protocol executions depends on the
partitioning $H,C$, since any value received from a member of
$C$ is initially considered to be low integrity. But as for
confidentiality types, we want our basic type analysis to be
generalizable to arbitrary $H,C$\cnote{Add related text to that
  section.}. So integrity type judgements are of the form
$\ipj{\eqs}{\prog}{\Delta}$, where $\eqs$ plays a similar role as in
confidentiality types, as a possibly weakened expression of algebraic
properties of $\prog$, and encironments $\Delta$ record integrity
information. Environments are ordered lists, since order of evaluation
is important for tracking integrity-- validating low
integrity information through MAC checking allows \emph{subsequent}
treatment of it as high integrity, but not \emph{before}.

The typing of protocols depends on a type judgement for expressions
of the form $\itj{\cid}{\be}{\ity{\cid}{V}}$, which records
messages and reveals $V$ used in the construction of $\be$.
In an assignment, this source is recorded in $\Delta$-- so for
example, letting:
$$
\prog\ \defeq\ (\xassign{\mx{x}{1}}{\secret{x} - \flip{x}}{2};\ \rvl{x} := \mx{x}{1})
$$
the following judgement is valid:
$$
\ipj{\toeq{\prog}}{\prog}{(\mx{x}{1} : \ity{2}{\varnothing}; \rvl{x} : \ity{1}{\mx{x}{1}})}
$$
More formally, we defined validity as follows:
\begin{definition}
  Given pre-processing predicate $\eqspre$ and protocol $\prog$, 
  we say $\ipj{\eqs}{\prog}{\Delta}$ is \emph{valid}
  iff it is derivable and $\eqspre \wedge \toeq{\prog} \models E$.
\end{definition}

In the above example, note that environment implicitly records
that $\rvl{x}$ depends on data from client 1, which in turn
depends on data from client 2. The rule for MAC
checking allows an overwrite of the dependency. We explore
the application of this in more detail in Section
\ref{section-examples}.

\cheatjfig

In all cases, we want to determine the integrity level of computed
values based on assumed partitionins $H,C$. In the previous example,
we want to say that $\rvl{x}$ has low integrity if either client 1 or
2 is corrupt.  To this end we define \emph{security labelings}
$\seclev$ which are mappings from variables $x$ to security levels
$\latel$. By default, given a partitioning $H,C$ this labeling maps
messages to the integrity level of the owner.
\begin{definition}  
  Given $H,C$,
  define $\seclev_{H,C}$ such that for all $\mx{w}{\cid}$ we
  have $\seclev_{H,C}(\mx{w}{\cid} ) = \hilab$ if $\cid \in H$ and $\lolab$
  otherwise.
\end{definition}
During execution of a protocol, the integrity of messages and reveals
affected by values sent by other parties, as recorded in $\Delta$
obtained from its typing. We additionally need to determine
whether this impacts integrity, which we accomplish through
the inductively defined rewrite relation $\cheatj{\Delta}{H,C}{\seclev}$,
axiomatized in Figure \ref{fig-cheatj}.
Our main type safety result for integrity is then given as follows.
\begin{theorem}
  Given pre-processing predicate $\eqspre$ and protocol $\prog$ with
  $\views(\prog) = V$, if
  $\ipj{\eqs}{\prog}{\Delta}$ is valid
  and for all $H,C$ with $\cheatj{\Delta}{H,C}{\seclev}$ 
  we have $\seclev(x) = \hilab$ for all $x \in \houtputs$, then cheating
  is detectable in $\prog$.
\end{theorem}

### 2024-09-30 esop25/metalang.tex
checkpoint

\section{Compositional Type Verification in $\metaprot$}

\subsection{Syntax and Semantics}

$$
\begin{array}{rcl}
  %\notg{x} &::=& \elab{\flip{e}}{e} \mid \elab{\secret{e}}{e} \mid \elab{\mesg{e}}{e} \mid \rvl{e} \mid \out{e}\\[2mm]
  \multicolumn{3}{l}{\flab \in \mathrm{Field},\   y \in \mathrm{EVar}, \  f \in \mathrm{FName}}\\[1mm]
  %x &\in& \mathrm{EVar}\\
  %f &\in& \mathrm{FName}\\[2mm]
  e &::=& \mv \mid \flip{e} \mid \secret{e} \mid \mesg{e} \mid \rvl{e} \mid \outkw \mid e \bop e \mid
  \elet{y}{e}{e} \mid \\
  & & f(e,\ldots,e) \mid \{ \flab = e; \ldots; \flab = e \} \mid e.\flab \mid \elab{e}{e} \mid y\\
  %  & \textit{expressions}\\
  \cmd &::=& %\msend{e}{e}{e}{e} \mid \reveal{e}{e}{e} \mid \pubout{e}{e}{e} \mid
  \assign{e}{e} \mid \elab{\assert{e = e}}{e} \mid f(e,\ldots,e) \mid  \cmd;\cmd \\[2mm] %\pre{\eqs} \mid \post{\eqs} \\[1mm]
  \bop &::=& \fplus \mid \fminus \mid \ftimes \mid \concat  \\[1mm]% \textit{operators}\\[2mm]
  \mv &::=& w \mid \cid \mid \be \mid x \mid \{ \flab = \mv;\ldots;\flab = \mv \} 
  \\ % \mid \ttt{()} \\[1mm] %& \textit{values}\\[2mm]
  \mathit{fn} &::=& f(y,\ldots,y) \{ e \} \mid  f(y,\ldots,y) \{ \cmd \} %\\[1mm]%& \textit{functions}
  %\phi &::=& \elab{\flip{e}}{e} \mid \elab{\secret{e}}{e} \mid \elab{\mesg{e}}{e} \mid \rvl{e} \mid \out{e} \mid \phi \fplus \phi \mid \phi \fminus \phi \mid \phi \ftimes \phi \\
  %\eqs &::=& \phi \eop \phi \mid \eqs \wedge \eqs 
\end{array}
$$

\begin{mathpar}
  \inferrule
      {e[\mv/y] \redx \mv'}
      {\elet{y}{\mv}{e} \redx \mv'}

  \inferrule
      {\codebase(f) = y_1,\ldots,y_n,\ e \\ e_1 \redx \mv_1 \cdots e_n \redx \mv_n \\
        e[\mv_1/y_1]\cdots[\mv_n/y_n] \redx \mv}
      {f(e_1,\ldots,e_n) \redx \mv}

  \inferrule
      {e_1 \redx \mv_1 \\ \cdots \\ e_n \redx \mv_n }
      {\{ \flab_1 = e_1; \ldots; \flab_n = e_n \} \redx \{ \flab_1 = \mv_1; \ldots; \flab_n = \mv_n \} }

  \inferrule
      {e \redx \{\ldots; \flab = \mv; \ldots\}}
      {e.\flab \redx \mv}

  \inferrule
      {e_1 \redx w_1 \\ e_2 \redx w_2}
      {e_1 \concat e_2 \redx w_1w_2}

  \inferrule
      {e \redx w}
      {\mesg{e} \redx \mesg{w}}

  \inferrule
      {e_1 \redx \be_1 \\ e_2 \redx \be_2}     
      {e_1 \fplus e_2 \redx \be_1 \fplus \be_2}

  \inferrule
      {e_1 \redx \be \\ e_2 \redx \cid}
      {\elab{e_1}{e_2} \redx \elab{\be}{\cid}}
\end{mathpar}

\begin{mathpar}
  \inferrule
      {e_1 \redx x \\ e_2 \redx \elab{\be}{\cid}}
      {\assign{e_1}{e_2} \redx \xassign{x}{\be}{\cid}}

  \inferrule
      {e_1 \redx \prog_1 \\ e_2 \redx \prog_2}
      {e_1;e_2 \redx \prog_1;\prog_2}

  \inferrule
      {e_1 \redx \be_1 \\ e_2 \redx \be_2 \\ e_3 \redx \cid}
      {\elab{\assert{e_1 = e_2}}{e_3} \redx \elab{\assert{\be_1 = \be_2}}{\cid}}

  \inferrule
      {\codebase(f) = y_1,\ldots,y_n, \instr \\
        e_1 \redx \mv_1 \ \cdots \ e_n \redx \mv_n \\
        \subn = [\mv_1/y_1]\cdots[\mv_n/y_n] \\
        \subn(\instr) \redx \prog
      }
      {f(e_1,\ldots,e_n) \redx \prog}
\end{mathpar}

\subsection{Dependent Hoare Type Theory}

$$
\cmd ::= \cdots \mid \eqcast{\mx{e}{e}}{\notg{\phi}} 
$$

$$
\begin{array}{rcl}
  \notg{\phi} &::=& e \mid \notg{\phi} \fplus \notg{\phi} \mid \notg{\phi} \fminus \notg{\phi} \mid \notg{\phi} \ftimes \notg{\phi} \\
  \notg{\eqs} &::=& \notg{\phi} \eop \notg{\phi} \mid \notg{\eqs} \wedge \notg{\eqs} \\
  \notg{t} &::=& e \mid \cty{e}{\notg{\ty}} \\
  \notg{\ty} & \in & 2^{\notg{t}}\\
  \notg{\Gamma} &::=& \varnothing \mid \notg\Gamma; e : \notg{\ty}\\
  \notg\Delta &::=& \varnothing \mid \notg\Delta; e : \ity{e}{\notg{V}}\\
  \notg{X} &\in& 2^{e}
\end{array}
$$

\begin{mathpar}
  \inferrule
      {\notg{\phi_1} \redx \phi_1 \\ \notg{\phi_2} \redx \phi_2}     
      {\notg{\phi_1} \ftimes \notg{\phi_2} \redx \phi_1 \ftimes \phi_2}

  \inferrule
      {\notg{\phi_1} \redx \phi_1 \\ \notg{\phi_2} \redx \phi_2}
      {\notg{\phi_1} \eop \notg{\phi_2} \redx \phi_1 \eop \phi_2}

  \inferrule
      {\notg{\eqs_1} \redx \eqs_1 \\ \notg{\eqs_2} \redx \eqs_2 }
      {\notg{\eqs_1} \wedge \notg{\eqs_2} \redx \eqs_1 \wedge \eqs_2}
\end{mathpar}

\begin{mathpar}
  \inferrule
      {e \redx x \\ \notg{\ty} \redx \ty}
      {\cty{e}{\notg{\ty}} \redx \cty{x}{\ty}}

  \inferrule
      {\notg{t_1} \redx t_1 \\ \cdots \\ \notg{t_n} \redx t_n}
      {\setit{\notg{t_1},\ldots,\notg{t_n}} \redx \setit{ t_1,\ldots,t_n }}

  \inferrule
      {\notg{\Gamma} \redx \Gamma \\ e \redx x \\ \notg{\ty} \redx \ty }
      {\notg{\Gamma}; e : \notg{\ty} \redx \Gamma; x : \ty }

  \inferrule
      {\notg{\Delta} \redx \Delta \\ e_1 \redx x  \\ e_2 \redx \cid \\ \notg{V} \redx V}
      {\notg{\Delta}; e_1 : \ity{e_2}{\notg{V}} \redx \Delta; x : \ity{\cid}{V} }

  \inferrule
      {\notg{\eqs_1} \redx \eqs_1 \\ \notg{\Gamma} \redx \\ \notg{R} \redx R
        \\ \notg{\Delta} \redx \Delta \\ \notg{\eqs_2} \redx \eqs_2}
      {\hty{\notg{\eqs_1}}{\notg{\Gamma}}{\notg{R}}{\notg{\Delta}}{\notg{\eqs_2}} \redx
        \hty{\eqs_1}{\Gamma}{R}{\Delta}{\eqs_2}}
\end{mathpar}

\begin{mathpar}
  \atj{x}{\varnothing}{\setit{x}}

  \inferrule
  {\atj{\phi}{R}{\ty} \\ \rx{w}{\cid}\not\in R \\ \oplus \in \setit{\fplus,\fminus}}
  {\atj{\phi \oplus \rx{w}{\cid}}{R \cup \setit{\rx{w}{\cid}}}{\setit{\cty{\rx{w}{\cid}}{\ty}}}}

  \inferrule
  {\atj{\phi_1}{R_1}{\ty_1} \\
   \atj{\phi_2}{R_2}{\ty_2} \\ \oplus \in \{ \fplus,\fminus,\ftimes \}}
  {\atj{\phi_1 \oplus \phi_2}{R_1;R_2}{\ty_1 \cup \ty_2}}
\end{mathpar}

\begin{mathpar}
  \inferrule[Mesg]
            {e_1 \redx x \\ e_2 \redx \be \\ e_3 \redx \cid \\ \atj{\toeq{\elab{\be}{\cid}}}{R_2}{\ty} \\
              \itj{\cid}{\be}{V}}
            {\mtj{\xassign{e_1}{e_2}{e_3}}{\eqs}{(x:\ty)}{R_1;R_2}{(x : \ity{\cid}{V})}{\eqs \wedge x \eop \toeq{\elab{\be}{\cid}}}}

  \inferrule[Encode]
            {e_1 \redx w \\ e_2 \redx \cid \\ \notg{\phi} \redx \phi \\
              \eqs \models \toeq{\elab{\be}{\cid}} \eop \phi\\
              \atj{\phi}{R}{\ty}}
            {\mtj{\eqcast{\mx{e_1}{e_2}}{\notg{\phi}}}{\eqs}{(\mx{w}{\cid}:\ty)}{R}{\varnothing}{\eqs}}

  \inferrule[App]
            {\tsig(f) = \dht{y_1,\ldots,y_n}{\notg{\eqs_1}}{\notg{\Gamma}}{\notg{R}}{\notg{\Delta}}{\notg{\eqs_2}} \\
              e_1 \redx \mv_1\ \cdots\ e_n \redx \mv_n \\
              \subn = [\mv_1/y_1]\cdots[\mv_n/y_n] \\
              \subn(\hty{\notg{\eqs_1}}{\notg{\Gamma}}{\notg{R}}{\notg{\Delta}}{\notg{\eqs_2}}) \redx
                    \hty{\eqs_1}{\Gamma}{R}{\Delta}{\eqs_2} \\
              \eqs \models \eqs_1}
            {\mtj{f(e_1,\ldots,e_n)}{\eqs}{\Gamma}{R}{\Delta}{\eqs \wedge \eqs_2}}

  \inferrule[Seq]          
            {\mtj{\prog_1}{\eqs_1}{\Gamma_1}{R_1}{\Delta_1}{\eqs_2} \\
             \mtj{\prog_2}{\eqs_2}{\Gamma_2}{R_2}{\Delta_2}{\eqs_3}}
            {\mtj{\prog_1;\prog_2}{\eqs_1}{\Gamma_1;\Gamma_2}{R_1;R_2}{\Delta_1;\Delta_2}{\eqs_3}}
\end{mathpar}

\begin{mathpar}
  \inferrule[Sig]
            {\codebase(f) = y_1,\ldots,y_n, \instr \\
              \subn = [\mv_1/y_1]\cdots[\mv_n/y_n] \\
              \subn(\hty{\notg{\eqs_1}}{\notg{\Gamma}}{\notg{R}}{\notg{\Delta}}{\notg{\eqs_2}}) \redx
                    \hty{\eqs_1}{\Gamma}{R}{\Delta}{\eqs_2} \\
              \mtj{\subn(\instr)}{\eqs_1}{\Gamma}{R}{\Delta}{\eqs} \\
              \eqs \models \eqs_2}
            {f : \dht{y_1,\ldots,y_n}{\notg{\eqs_1}}{\notg{\Gamma}}{\notg{R}}{\notg{\Delta}}{\notg{\eqs_2}}}
\end{mathpar}

\begin{definition}
  $\tsig$ is \emph{verified} iff $f : \tsig(f)$ is valid for all $f \in \dom(\tsig)$.
\end{definition}

\begin{theorem}
  Given preprocessing predicate $\eqspre$, program $\instr$, and verified $\tsig$, if
  the judgement $\mtj{\instr}{\eqspre}{\Gamma}{R}{\Delta}{\eqs}$ is derivable then
  $\instr \redx \prog$ and:
  \begin{enumerate}
  \item $\cpj{R}{\eqs}{\eqspre \wedge \toeq{\prog}}{\Gamma}$ is valid.
  \item $\ipj{\eqs}{\prog}{\Delta}$ is valid.
  \end{enumerate}
\end{theorem}

### 2024-09-30 code/cvc5mvn/src/test/java/plaid/antlr/FullProgramParseTest.java
assert successful parse

@@ -3,6 +3,8 @@ package plaid.antlr;
 import org.junit.Test;
 import plaid.ast.Program;
 
+import static org.junit.Assert.assertNotNull;
+
 public class FullProgramParseTest {
 
     /**
@@ -66,5 +68,6 @@ public class FullProgramParseTest {
                 """);
 
         System.out.print(program);
+        assertNotNull(program);
     }
 }

### 2024-10-02 code/cvc5mvn/src/main/java/plaid/antlr/Loader.java
javadoc

@@ -2,7 +2,6 @@ package plaid.antlr;
 
 import org.antlr.v4.runtime.ANTLRInputStream;
 import org.antlr.v4.runtime.CommonTokenStream;
-import org.antlr.v4.runtime.tree.ParseTree;
 import org.antlr.v4.runtime.tree.ParseTreeWalker;
 import plaid.PreludeLexer;
 import plaid.PreludeParser;
@@ -13,8 +12,17 @@ import plaid.ast.PreludeCommand;
 import plaid.ast.PreludeExpression;
 import plaid.ast.Program;
 
+/**
+ * Support for converting Prelude source code into an abstract syntax tree.
+ */
 public class Loader {
 
+    /**
+     * Creates a ANTLR4 parser for Prelude source code.
+     *
+     * @param src Source code
+     * @return ANTLR4 parser
+     */
     public static PreludeParser createParser(String src) {
         ANTLRInputStream input = new ANTLRInputStream(src);
         PreludeLexer lexer = new PreludeLexer(input);
@@ -24,28 +32,64 @@ public class Loader {
         return parser;
     }
 
+    /**
+     * Converts an ANTLR4 context into an abstract syntax tree for an expression.
+     *
+     * @param ctx ANTLR4 context
+     * @return Abstract syntax tree for an expression
+     */
     public static PreludeExpression toExpression(ExprContext ctx) {
         return new ExpressionVisitor().visit(ctx);
     }
 
+    /**
+     * Converts Prelude source code into an abstract syntax tree for an expression.
+     *
+     * @param src Source code
+     * @return Abstract syntax tree for an expression
+     */
     public static PreludeExpression toExpression(String src) {
         return toExpression(createParser(src).expr());
     }
 
+    /**
+     * Converts an ANTLR4 context into an abstract syntax tree for a command.
+     *
+     * @param ctx ANTLR4 context
+     * @return Abstract syntax tree for a command
+     */
     public static PreludeCommand toCommand(CommandContext ctx) {
         return new CommandVisitor().visit(ctx);
     }
 
+    /**
+     * Converts Prelude source code into an abstract syntax tree for a command.
+     *
+     * @param src Source code
+     * @return Abstract syntax tree for a command
+     */
     public static PreludeCommand toCommand(String src) {
         return toCommand(createParser(src).command());
     }
 
+    /**
+     * Converts an ANTLR4 context into an abstract syntax tree for a Prelude program.
+     *
+     * @param ctx ANTLR4 context
[…]

### 2024-10-02 esop25/ipj.tex
checkpoint

@@
   Given $H,C$,
-  define $\seclev_{H,C}$ such that for all $\mx{w}{\cid}$ we
-  have $\seclev_{H,C}(\mx{w}{\cid} ) = \hilab$ if $\cid \in H$ and $\lolab$
-  otherwise.
+  define $\seclev_{H,C}$ such that $\seclev_{H,C}(\mx{w}{\cid} ) = \hilab$
+  if $\cid \in H$  and $\lolab$
+  otherwise for all $\mx{w}{\cid}$.
 \end{definition}

### 2024-10-02 esop25/metalang.tex
checkpoint

@@
 \section{Compositional Type Verification in $\metaprot$}
+
+Practical MPC computations protocols are
+typically composed of compositional units. Examples include GMW circuits
+and Yao's Garbled Circuits (YGC), that are composed of so-called
+garbled gates. Languages such as Fairplay \cite{269581} provide gates as
+units of abstraction that are ``wired'' together by the programmer to
+generate a complete circuit.
+
+The $\fedprot$ language is low-level and does not include abstractions
+for defining composable elements. So in this Section we introduce the
+$\metaprot$ language that includes structured data and function
+definitions for defining composable protocol elements at a higher
+level of abstraction.  The $\metaprot$ language is a
+\emph{metalanguage}, where $\fedprot$ protocols are the residuum of
+computation. In addition to these declarative benefits of $\metaprot$,
+component definitions support compositional verification of larger
+protocols as we will discuss with examples in Section
+\ref{section-examples}.
+
+\metaprotsyntaxfig
 
@@
 
-$$
-\begin{array}{rcl}
-  %\notg{x} &::=& \elab{\flip{e}}{e} \mid \elab{\secret{e}}{e} \mid \elab{\mesg{e}}{e} \mid \rvl{e} \mid \out{e}\\[2mm]
-  \multicolumn{3}{l}{\flab \in \mathrm{Field},\   y \in \mathrm{EVar}, \  f \in \mathrm{FName}}\\[1mm]
-  %x &\in& \mathrm{EVar}\\
-  %f &\in& \mathrm{FName}\\[2mm]
-  e &::=& \mv \mid \flip{e} \mid \secret{e} \mid \mesg{e} \mid \rvl{e} \mid \outkw \mid e \bop e \mid
-  \elet{y}{e}{e} \mid \\
-  & & f(e,\ldots,e) \mid \{ \flab = e; \ldots; \flab = e \} \mid e.\flab \mid \elab{e}{e} \mid y\\
-  %  & \textit{expressions}\\
-  \cmd &::=& %\msend{e}{e}{e}{e} \mid \reveal{e}{e}{e} \mid \pubout{e}{e}{e} \mid
-  \assign{e}{e} \mid \elab{\assert{e = e}}{e} \mid f(e,\ldots,e) \mid  \cmd;\cmd \\[2mm] %\pre{\eqs} \mid \post{\eqs} \\[1mm]
-  \bop &::=& \fplus \mid \fminus \mid \ftimes \mid \concat  \\[1mm]% \textit{operators}\\[2mm]
-  \mv &::=& w \mid \cid \mid \be \mid x \mid \{ \flab = \mv;\ldots;\flab = \mv \} 
-  \\ % \mid \ttt{()} \\[1mm] %& \textit{values}\\[2mm]
-  \mathit{fn} &::=& f(y,\ldots,y) \{ e \} \mid  f(y,\ldots,y) \{ \cmd \} %\\[1mm]%& \textit{functions}
-  %\phi &::=& \elab{\flip{e}}{e} \mid \elab{\secret{e}}{e} \mid \elab{\mesg{e}}{e} \mid \rvl{e} \mid \out{e} \mid \phi \fplus \phi \mid \phi \fminus \phi \mid \phi \ftimes \phi \\
-  %\eqs &::=& \phi \eop \phi \mid \eqs \wedge \eqs 
-\end{array}
-$$
+The syntax of $\metaprot$ is defined in Figure
+\ref{fig-metaprot}.  It includes a syntax of function
+definitions and records, and values $\mv$ include client ids $\cid$, identifier
+strings $w$, expressions $\be$ in field $\mathbb{F}_p$, and the unit value $()$.
+Expression forms allow dynamic construction of field expressions and $\minifed$ commands.
+The construction of a command has the side-effect of adding to the residual
+$\minifed$ protocol. Formally, we consider a complete metaprogram to include both a
+codebase and a ``main'' program that uses the codebase. 
+\begin{definition}
+A \emph{codebase} $\codebase$ is a list of function 
+declarations. We write $ \codebase(f) = y_1,\ldots,y_n,\ e$
+iff $f(y_1,\ldots,y_n) \{ e \} \in \codebase$.
+A \emph{metaprogram}, aka \emph{metaprotocol}  is a pair of a 
+codebase and expression $\codebase, e$. We may omit
+$\codebase$ if it is clear from context.  
+\end{definition}
 
-\begin{mathpar}
-  \inferrule
-      {e[\mv/y] \redx \mv'}
-      {\elet{y}{\mv}{e} \redx \mv'}
+We define a small-step evaluation aka reduction relation $\redx$ in
+Figure \ref{fig-metaprot}.  We write $\redxs$ to denote the
+reflexive, transitive closure of $\redx$. Reduction is defined on
+\emph{configurations} which are pairs of the form $\config{\prog}{e}$,
+where $\prog$ is the $\minifed$ program accumulated during evaluation.
+In this definition we write $e[\mv/y]$ to denote the substitution of $\mv$
+for free occurrences of $y$ in $e$. The rules are mostly standard,
+except when a concrete $\minifed$ assignment is encountered it is added
+to the end of $\prog$.
 
-  \inferrule
-      {\codebase(f) = y_1,\ldots,y_n,\ e \\ e_1 \redx \mv_1 \cdots e_n \redx \mv_n \\
-        e[\mv_1/y_1]\cdots[\mv_n/y_n] \redx \mv}
-      {f(e_1,\ldots,e_n) \redx \mv}
+\metaprotexprsemanticsfig
 
-  \inferrule
-      {e_1 \redx \mv_1 \\ \cdots \\ e_n \redx \mv_n }
-      {\{ \flab_1 = e_1; \ldots; \flab_n = e_n \} \redx \{ \flab_1 = \mv_1; \ldots; \flab_n = \mv_n \} }
+The rules rely on a definition of \emph{evaluation contexts} $E$
+allowing computation within a larger program context, where $E[e]$
+denotes an expression with $e$ in the hole $[]$ of $E$. The syntax
+of $E$ imposes a left-to-right order of evaluation of subexpressions
+for all forms.
 
-  \inferrule
-      {e \redx \{\ldots; \flab = \mv; \ldots\}}
-      {e.\flab \redx \mv}
-
-  \inferrule
-      {e_1 \redx w_1 \\ e_2 \redx w_2}
-      {e_1 \concat e_2 \redx w_1w_2}
-
-  \inferrule
-      {e \redx w}
-      {\mesg{e} \redx \mesg{w}}
-
-  \inferrule
-      {e_1 \redx \be_1 \\ e_2 \redx \be_2}     
-      {e_1 \fplus e_2 \redx \be_1 \fplus \be_2}
-
-  \inferrule
-      {e_1 \redx \be \\ e_2 \redx \cid}
-      {\elab{e_1}{e_2} \redx \elab{\be}{\cid}}
-\end{mathpar}
-
-\begin{mathpar}
-  \inferrule
-      {e_1 \redx x \\ e_2 \redx \elab{\be}{\cid}}
-      {\assign{e_1}{e_2} \redx \xassign{x}{\be}{\cid}}
-
-  \inferrule
-      {e_1 \redx \prog_1 \\ e_2 \redx \prog_2}
-      {e_1;e_2 \redx \prog_1;\prog_2}
-
-  \inferrule
-      {e_1 \redx \be_1 \\ e_2 \redx \be_2 \\ e_3 \redx \cid}
-      {\elab{\assert{e_1 = e_2}}{e_3} \redx \elab{\assert{\be_1 = \be_2}}{\cid}}
-
-  \inferrule
-      {\codebase(f) = y_1,\ldots,y_n, \instr \\
-        e_1 \redx \mv_1 \ \cdots \ e_n \redx \mv_n \\
-        \subn = [\mv_1/y_1]\cdots[\mv_n/y_n] \\
-        \subn(\instr) \redx \prog
-      }
-      {f(e_1,\ldots,e_n) \redx \prog}
-\end{mathpar}
+\metaprotinstrsemanticsfig
 
 \subsection{Dependent Hoare Type Theory}
+
+\subsubsection{$\minicat$ expression type algorithm}
+
+\atjfig
+
+\begin{lemma}
+  If $\atj{\toeq{\phi}}{R}{\ty}$ then $\eqj{R}{\eqs}{\phi}{\ty}$ for any $\eqs$.
+\end{lemma}
 
@@
 \begin{mathpar}
-  \atj{x}{\varnothing}{\setit{x}}
-
-  \inferrule
-  {\atj{\phi}{R}{\ty} \\ \rx{w}{\cid}\not\in R \\ \oplus \in \setit{\fplus,\fminus}}
-  {\atj{\phi \oplus \rx{w}{\cid}}{R \cup \setit{\rx{w}{\cid}}}{\setit{\cty{\rx{w}{\cid}}{\ty}}}}
-
-  \inferrule
-  {\atj{\phi_1}{R_1}{\ty_1} \\
-   \atj{\phi_2}{R_2}{\ty_2} \\ \oplus \in \{ \fplus,\fminus,\ftimes \}}
-  {\atj{\phi_1 \oplus \phi_2}{R_1;R_2}{\ty_1 \cup \ty_2}}
-\end{mathpar}
-
-\begin{mathpar}
   \inferrule[Mesg]

### 2024-10-02 esop25/smt.tex
checkpoint

--- 2024-08-24 csf25/smt.tex
@@
 
-\begin{fpfig}[t]{Syntax of Constraints and Memory Application Extended to
+\begin{fpfig}[t]{Syntax of constraints and memory application extended to
   SMT expressions}{fig-eqs}
@@
 
-  \toeq{\assert{\be_1 = \be_2}{\cid}} =  \toeq{\elab{\be_1}{\cid}} \eop \toeq{\elab{\be_2}{\cid}}
+  \toeq{\elab{\assert{\be_1 = \be_2}}{\cid}} =  \toeq{\elab{\be_1}{\cid}} \eop \toeq{\elab{\be_2}{\cid}}
 
@@
 \subsubsection{Example: Correctness of 3-Party Addition}
-Consider the following 3-party Shamir Addition protocol.
+Consider the following 3-party additive secret sharing protocol.
 

### 2024-10-02 esop25/metalang.tex
checkpoint

@@
 The syntax of $\metaprot$ is defined in Figure
-\ref{fig-metaprot}.  It includes a syntax of function
-definitions and records, and values $\mv$ include client ids $\cid$, identifier
-strings $w$, expressions $\be$ in field $\mathbb{F}_p$, and the unit value $()$.
-Expression forms allow dynamic construction of field expressions and $\minifed$ commands.
-The construction of a command has the side-effect of adding to the residual
-$\minifed$ protocol. Formally, we consider a complete metaprogram to include both a
-codebase and a ``main'' program that uses the codebase. 
+\ref{fig-metaprotsyntax}.  It includes a syntax of values $\mv$ that
+include client ids $\cid$, identifier strings $w$, expressions $\be$
+in field $\mathbb{F}_p$, record values, and $\minicat$ variables
+$x$. $\metaprot$ expression forms allow dynamic construction of these
+values. $\metaprot$ \emph{instruction} forms allow dynamic
+construction of $\minicat$ protocols $\prog$, incorporating expression
+evaluation. The syntax also supports definitions of functions that
+compute values $\mv$ and, as a distinct form, functions that compute
+protocols.  Formally, we consider a complete metaprogram to include
+both a codebase and a ``main'' program that uses the codebase.
 \begin{definition}
@@
 declarations. We write $ \codebase(f) = y_1,\ldots,y_n,\ e$
-iff $f(y_1,\ldots,y_n) \{ e \} \in \codebase$.
-A \emph{metaprogram}, aka \emph{metaprotocol}  is a pair of a 
-codebase and expression $\codebase, e$. We may omit
-$\codebase$ if it is clear from context.  
+if $f(y_1,\ldots,y_n) \{ e \} \in \codebase$, and we
+write  $ \codebase(f) = y_1,\ldots,y_n,\ \instr$
+if $f(y_1,\ldots,y_n) \{ \instr \} \in \codebase$.
+%A \emph{metaprogram}, aka \emph{metaprotocol}  is a pair of a 
+%codebase and expression $\codebase, e$. We may omit
+%$\codebase$ if it is clear from context.  
 \end{definition}
 
-We define a small-step evaluation aka reduction relation $\redx$ in
-Figure \ref{fig-metaprot}.  We write $\redxs$ to denote the
-reflexive, transitive closure of $\redx$. Reduction is defined on
-\emph{configurations} which are pairs of the form $\config{\prog}{e}$,
-where $\prog$ is the $\minifed$ program accumulated during evaluation.
-In this definition we write $e[\mv/y]$ to denote the substitution of $\mv$
-for free occurrences of $y$ in $e$. The rules are mostly standard,
-except when a concrete $\minifed$ assignment is encountered it is added
-to the end of $\prog$.
+We define a big-step evaluation relation $\redx$ in Figures
+\ref{fig-metaprotexprsemantics} and \ref{fig-metaprotinstrsemantics}
+for expressions and instructions, respectively.  In this definition we
+write $e[\mv/y]$ and $\instr[\mv/y]$ to denote the substitution of
+$\mv$ for free occurrences of $y$ in $e$ or $\instr$ respectively. The
+rules are mostly standard.
 
 \metaprotexprsemanticsfig
-
-The rules rely on a definition of \emph{evaluation contexts} $E$
-allowing computation within a larger program context, where $E[e]$
-denotes an expression with $e$ in the hole $[]$ of $E$. The syntax
-of $E$ imposes a left-to-right order of evaluation of subexpressions
-for all forms.
 

### 2024-10-03 esop25/cpj.tex
checkpoint

@@
 \section{Confidentiality Types}
+\label{section-cpj}
 
@@
 
-For example, given a message send such as:
+For example, given some string $y$ and a message send such as:
 $$
-\xassign{\mx{x}{1}}{\secret{x} \ftimes \flip{x}}{2}
+\xassign{\mx{w}{1}}{\secret{w} \ftimes \flip{w}}{2}
 $$
-we would assign the following type to denote that $\mx{x}{1}$ is dependent on
-both $\sx{x}{2}$ and $\rx{x}{2}$:
+we would assign the following type to denote that $\mx{w}{1}$ is dependent on
+both $\sx{w}{2}$ and $\rx{w}{2}$:
 $$
-\mx{x}{1} : \setit{\sx{x}{2}, \rx{x}{2}}
+\mx{w}{1} : \setit{\sx{w}{2}, \rx{w}{2}}
 $$
@@
 $$
-\xassign{\mx{x}{1}}{\secret{x} \fminus \flip{x}}{2}
+\xassign{\mx{w}{1}}{\secret{w} \fminus \flip{w}}{2}
 $$
-we want the type of $\mx{x}{1}$ to be an ``encrypted'' type
-that is by itself independent of $\sx{x}{2}$. However, we also
-want to track the dependence of $\mx{x}{2}$ and $\rx{x}{2}$.
-In short, we assign $\mx{x}{1}$ a ``ciphertext type'':
+we want the type of $\mx{w}{1}$ to be an ``encrypted'' type
+that is by itself independent of $\sx{w}{2}$. However, we also
+want to track the dependence of $\mx{w}{2}$ and $\rx{w}{2}$.
+In short, we assign $\mx{w}{1}$ a ``ciphertext type'':
 $$
-\mx{x}{1} : \cty{\rx{x}{2}}{\setit{\sx{x}{2}}}
+\mx{w}{1} : \cty{\rx{w}{2}}{\setit{\sx{w}{2}}}
 $$
-indicating syntactically that $\mx{x}{1}$ is an encryption of a
-value dependent on $\sx{x}{2}$ with $\rx{x}{2}$ and as long as client
+indicating syntactically that $\mx{w}{1}$ is an encryption of a
+value dependent on $\sx{w}{2}$ with $\rx{w}{2}$ and as long as client
 1 does not subsequently receive messages that are dependent on
-$\rx{x}{2}$, the value $\sx{x}{2}$ remains hidden.
+$\rx{w}{2}$, the value $\sx{w}{2}$ remains hidden.
 
@@
 $$
-\xassign{\mx{x}{1}}{\mux{\secret{x}}{\neg\flip{x}}{\flip{x}}}{2}
+\xassign{\mx{w}{1}}{\mux{\secret{w}}{\neg\flip{w}}{\flip{w}}}{2}
 $$
@@
 \begin{mathpar}
-%\toeq{\xassign{\mx{x}{1}}{\OT{\secret{x}}{1}{\neg\flip{x}}{\flip{x}}}{2}}
-   \toeq{\prog} \models \mx{x}{1} \eop \neg\sx{x}{2} \fplus \rx{x}{2}
+%\toeq{\xassign{\mx{w}{1}}{\OT{\secret{w}}{1}{\neg\flip{w}}{\flip{w}}}{2}}
+   \toeq{\prog} \models \mx{w}{1} \eop \neg\sx{w}{2} \fplus \rx{w}{2}
 \end{mathpar}
-This allows us to assign the same ciphertext type to $\mx{x}{1}$ as
-above-- that is, $\cty{\rx{x}{2}}{\setit{\sx{x}{2}}}$. Leveraging the
+This allows us to assign the same ciphertext type to $\mx{w}{1}$ as
+above-- that is, $\cty{\rx{w}{2}}{\setit{\sx{w}{2}}}$. Leveraging the
 equivalence this way has been previously well-studied
@@
 $$
-\setit{\cty{\rx{x}{\cid}}{\ty}} \qquad \setit{\rx{x}{\cid}, \ldots} 
+\setit{\cty{\rx{w}{\cid}}{\ty}} \qquad \setit{\rx{w}{\cid}, \ldots} 
 $$

### 2024-10-03 esop25/ipj.tex
checkpoint

@@
 \section{Integrity Types}
+\label{section-ipj}
 
@@
 In an assignment, this source is recorded in $\Delta$-- so for
-example, letting:
+example, for some string $w$ letting:
 $$
-\prog\ \defeq\ (\xassign{\mx{x}{1}}{\secret{x} - \flip{x}}{2};\ \rvl{x} := \mx{x}{1})
+\prog\ \defeq\ (\xassign{\mx{w}{1}}{\secret{w} - \flip{w}}{2};\ \rvl{w} := \mx{w}{1})
 $$
@@
 $$
-\ipj{\toeq{\prog}}{\prog}{(\mx{x}{1} : \ity{2}{\varnothing}; \rvl{x} : \ity{1}{\mx{x}{1}})}
+\ipj{\toeq{\prog}}{\prog}{(\mx{w}{1} : \ity{2}{\varnothing}; \rvl{w} : \ity{1}{\mx{w}{1}})}
 $$
@@
 In the above example, note that environment implicitly records
-that $\rvl{x}$ depends on data from client 1, which in turn
+that $\rvl{w}$ depends on data from client 1, which in turn
 depends on data from client 2. The rule for MAC
@@
 values based on assumed partitionins $H,C$. In the previous example,
-we want to say that $\rvl{x}$ has low integrity if either client 1 or
+we want to say that $\rvl{w}$ has low integrity if either client 1 or
 2 is corrupt.  To this end we define \emph{security labelings}

### 2024-10-03 esop25/metalang.tex
checkpoint

@@
 $\mv$ for free occurrences of $y$ in $e$ or $\instr$ respectively. The
-rules are mostly standard.
+rules are mostly standard. Note that we do not include an evaluation
+rule for the form $\eqcast{e}{e}{\notg{\phi}}$ which only has
+relevance for the type system and which we assume is erased from
+programs prior to evaluation. We defer discussion of this form,
+as well as the syntactic category $\notg{\phi}$, to the next
+Section.
 
@@
 \subsection{Dependent Hoare Type Theory}
+
+Our first goal with a $\metaprot$ type theory is to define an
+algorithmic system that is sound for both confidentiality and
+integrity typings as defined in Sections \ref{section-cpj} and
+\ref{section-ipj}. Returning to the example in Section
+\ref{section-cpj}, note that the key equivalence of the $\ttt{mux}$
+expression with a one-time-pad encryption we observe there
+is not trivial. To ensure that this sort of equivalence
+can be picked up by the type system, we introduce a
+an annotation form that allows the programmer to provide
+the needed hint.
+
+Consider the following $\ttt{encode}$ function definition that
+generalizes this encoding for any identifier $y$ and sender, receiver
+pair $s,r$:
+$$
+\begin{array}{l}
+\ttt{encode}(y,s,r) \{\\
+\quad \xassign{\mx{y}{r}}{\mux{\secret{y}}{\neg\flip{y}}{\flip{y}}}{s};\\
+\quad \eqcast{\mx{y}{r}}{\neg\sx{y}{s} \fplus \rx{y}{s}}\\
+\}
+\end{array}
+$$
+
+Furthermore, we want to minimize the amount of SMT solving needed for
+type verification.  Our approach is to support pre- and post-condition
+annotations on $\metaprot$ functions that can be independently
+verified, allowing significant reduction of SMT overhead in
+whole-program analysis.
 

### 2024-10-03 esop25/metalang.tex
checkpoint

@@
 
+\metaprotexprsemanticsfig
+
 We define a big-step evaluation relation $\redx$ in Figures
@@
 
-\metaprotexprsemanticsfig
-
 \metaprotinstrsemanticsfig
@@
 
-Our first goal with a $\metaprot$ type theory is to define an
+Our first goal in the $\metaprot$ type theory is to define an
 algorithmic system that is sound for both confidentiality and
@@
 generalizes this encoding for any identifier $y$ and sender, receiver
-pair $s,r$:
+pair $s,r$. The second line in the body is a hint that
+the message $\mx{y}{r}$ can be considered equivalent to
+$\neg\sx{y}{s} \fplus \rx{y}{s}$:
 $$
@@
 $$
+This hint can be validated using SMT, and then the
+syntactic structure of $\neg\sx{y}{s} \fplus \rx{y}{s}$
+allows its immediate interpretation as a one-time-pad encoding.
 
-Furthermore, we want to minimize the amount of SMT solving needed for
-type verification.  Our approach is to support pre- and post-condition
-annotations on $\metaprot$ functions that can be independently
-verified, allowing significant reduction of SMT overhead in
-whole-program analysis.
+Our second goal in the $\metaprot$ type system is to minimize the
+amount of SMT solving needed for type verification.  Returning to the
+$\ttt{encode}$ example, type checking in our systems only requires the
+hint to be verified once, with the guarantee that $\ttt{encode}$ can
+be applied anywhere without needing to re-verify the hint in
+application instances. To verify the hint in $\ttt{encode}$, we can
+just choose arbitrary values $w,\cid_1,\cid_2$ for $y$, $s$, and $r$,
+evaluate $\ttt{encode}(w,\cid_1,\cid_2) \redx \prog$, and then verify:
+$$\toeq{\prog} \models \mx{w}{\cid_2} \eop
+\neg\sx{w}{\cid_1} \fplus \rx{w}{\cid_1}$$ 
+Since $\ttt{encode}$ is closed, validity is guaranteed for
+any instance of $y$, $s$, and $r$.
+
+We generalize these benefits of compositional verification by allowing
+pre- and post-condition annotations on $\metaprot$ functions. For
+example, consider a GMW-style, 2-party ``and-gate'' function
+$\ttt{andgmw}(x,y,z)$. In this protocol, each party $\cid
+\in \{1,2\}$ holds an additive secret share $\mx{x}{\cid}$ of
+values identified by $x$ and $y$, and at the end of the
+protocol each hold a secret share $\mx{z}{\cid}$, where:
+$$
+\mx{z}{1} \fplus \mx{z}{2} \eop (\mx{x}{1} \fplus \mx{x}{2}) \ftimes (\mx{y}{1} \fplus \mx{y}{2})
+$$
+We provide details of $\ttt{andgmw}$ and other GMW protocol
+elements in Section \ref{section-examples}.  Our point now is that,
+similarly to the $\ttt{encode}$ example, we can verify this
+post-condition once as a correctness property for $\ttt{andgmw}$, and
+then integrate instances of it into circuit correctness properties
+with the guarantee that each instance also holds for any
+$\ttt{andgmw}$ gate.  Since the program logic of $\ttt{andgmw}$ is
+non-trivial, and typical circuits can use up to thousands of gates,
+this has significant practical benefits by greatly reducing SMT overhead
+in whole-program analysis.
 
@@
 
+A core element of $\metaprot$ type checking is type checking
+of $\minicat$ expressions. The integrity type system presented in Section
+\ref{section-ipj} is already algorithmic and ready to use. But confidentiality
+typing presented in Section \ref{section-cpj} is not syntax-directed
+due to the \TirName{Encode} rule. But as described above, by introducing
+hint annotations we can ``cast'' any relevant expression form into
+the simplest equivalent syntactic form for one-time-pad encoding.
+Thus, in Figure \ref{fig-atj} we present the algorithm for
+confidentiality type judgements, where we eliminate the need for
+integrated SMT solving by assuming this sort of casting.
+
 \atjfig
 
+The following Lemma establishes correctness of algorithmic confidentiality
+type checking, and makes explicit that SMT checking is eliminated in the
+judgement.
 \begin{lemma}
@@
 \end{lemma}
-
-$$
-\cmd ::= \cdots \mid \eqcast{\mx{e}{e}}{\notg{\phi}}
-$$
 

### 2024-10-03 esop25/metalang.tex
checkpoint

@@
 
-\subsubsection{$\minicat$ expression type algorithm}
+\subsubsection{$\minicat$ expression type algorithm.}
 
@@
 \end{lemma}
+
+\subsubsection{Dependent Hoare types for instructions.}
+
+Hoare-style types for instructions have the following form:
+$$
+\hty{\eqs_1}{\Gamma}{R}{\Delta}{\eqs_2}
+$$
+Here, $\eqs_1$ and $\eqs_2$ are the pre- and post- conditions
+respectively, $\Gamma$ is the confidentiality type environment of the
+protocol resulting from execution of the instruction, which is sound
+wrt confidentiality typing, $R$ are the one-time-pads consumed in the
+confidentialty typing, and $\Delta$ is the sound integrity type
+environment.
+
+To obtain high precision in the type system, and to generalize
+functions typings, we also introduce a form of type dependency,
+specifically dependence on $\metaprot$ expressions. Dependent
+$\Pi$ types have the form:
+$$
+\dht{y_1,\ldots,y_n}{\notg{\eqs_1}}{\notg{\Gamma}}{\notg{R}}{\notg{\Delta}}{\notg{\eqs_2}}
+$$
+where $y_1,\ldots,y_n$ range over values $\mv$ and each of
+$\notg{\eqs_1}$ etc.~may contain expressions with free occurences
+of $y_1,\ldots,y_n$-- the syntax of these forms is in Figure
+\ref{fig-notg}. These $\Pi$ types are assigned to functions
+and instantiated at application points.
+
+For example, here is a valid typing of $\ttt{encode}$.  It says that
+under any precondition, evaluating $\ttt{encode}$ results in a cipher
+type for the encoded message $\mx{x}{r}$, which consumes the
+one-time-pad $\rx{x}{s}$, and the integrity of $\mx{x}{r}$ is
+determined by the integrity of $\rx{x}{s}$ and $\sx{x}{s}$. The
+postcondition expresses the key confidentiality property of
+$\mx{x}{r}$, but also may be practically useful for correctness
+properties since it is a simpler expression form than the $\ttt{mux}$
+form:
+$$
+\ttt{encode} :
+\begin{array}[t]{ll}
+  \Pi x,s,r . & \{ \} \quad \mx{x}{r} : \cty{\rx{x}{s}}{\sx{x}{s}}, \{ \rx{x}{s} \} \\
+  & \phantom{\{ \}} \quad \mx{x}{r} : \ity{s}{\{ \rx{x}{s}, \sx{x}{s} \}} \\
+  & \{ \mx{x}{r} \eop \neg\sx{y}{s} \fplus \rx{y}{s} \}
+\end{array}
+$$
 

### 2024-10-03 esop25/metalang.tex
checkpoint

@@
 rules are mostly standard. Note that we do not include an evaluation
-rule for the form $\eqcast{e}{e}{\notg{\phi}}$ which only has
+rule for the form $\eqcast{\mx{e}{e}}{\notg{\phi}}$ which only has
 relevance for the type system and which we assume is erased from
@@
 
+\notgfig
+
 To obtain high precision in the type system, and to generalize
@@
 $$
-
-$$
-\begin{array}{rcl}
-  \notg{\phi} &::=& e \mid \notg{\phi} \fplus \notg{\phi} \mid \notg{\phi} \fminus \notg{\phi} \mid \notg{\phi} \ftimes \notg{\phi} \\
-  \notg{\eqs} &::=& \notg{\phi} \eop \notg{\phi} \mid \notg{\eqs} \wedge \notg{\eqs} \\
-  \notg{t} &::=& e \mid \cty{e}{\notg{\ty}} \\
-  \notg{\ty} & \in & 2^{\notg{t}}\\
-  \notg{\Gamma} &::=& \varnothing \mid \notg\Gamma; e : \notg{\ty}\\
-  \notg\Delta &::=& \varnothing \mid \notg\Delta; e : \ity{e}{\notg{V}}\\
-  \notg{X} &\in& 2^{e}
-\end{array}
-$$
-
-\begin{mathpar}
-  \inferrule
-      {\notg{\phi_1} \redx \phi_1 \\ \notg{\phi_2} \redx \phi_2}     
-      {\notg{\phi_1} \ftimes \notg{\phi_2} \redx \phi_1 \ftimes \phi_2}
-
-  \inferrule
-      {\notg{\phi_1} \redx \phi_1 \\ \notg{\phi_2} \redx \phi_2}
-      {\notg{\phi_1} \eop \notg{\phi_2} \redx \phi_1 \eop \phi_2}
-
-  \inferrule
-      {\notg{\eqs_1} \redx \eqs_1 \\ \notg{\eqs_2} \redx \eqs_2 }
-      {\notg{\eqs_1} \wedge \notg{\eqs_2} \redx \eqs_1 \wedge \eqs_2}
-\end{mathpar}
-
-\begin{mathpar}
-  \inferrule
-      {e \redx x \\ \notg{\ty} \redx \ty}
-      {\cty{e}{\notg{\ty}} \redx \cty{x}{\ty}}
-
-  \inferrule
-      {\notg{t_1} \redx t_1 \\ \cdots \\ \notg{t_n} \redx t_n}
-      {\setit{\notg{t_1},\ldots,\notg{t_n}} \redx \setit{ t_1,\ldots,t_n }}
-
-  \inferrule
-      {\notg{\Gamma} \redx \Gamma \\ e \redx x \\ \notg{\ty} \redx \ty }
-      {\notg{\Gamma}; e : \notg{\ty} \redx \Gamma; x : \ty }
-
-  \inferrule
-      {\notg{\Delta} \redx \Delta \\ e_1 \redx x  \\ e_2 \redx \cid \\ \notg{V} \redx V}
-      {\notg{\Delta}; e_1 : \ity{e_2}{\notg{V}} \redx \Delta; x : \ity{\cid}{V} }
-
-  \inferrule
-      {\notg{\eqs_1} \redx \eqs_1 \\ \notg{\Gamma} \redx \\ \notg{R} \redx R
-        \\ \notg{\Delta} \redx \Delta \\ \notg{\eqs_2} \redx \eqs_2}
-      {\hty{\notg{\eqs_1}}{\notg{\Gamma}}{\notg{R}}{\notg{\Delta}}{\notg{\eqs_2}} \redx
-        \hty{\eqs_1}{\Gamma}{R}{\Delta}{\eqs_2}}
-\end{mathpar}
 

### 2024-10-03 esop25/metalang.tex
checkpoint

@@
 rules are mostly standard. Note that we do not include an evaluation
-rule for the form $\eqcast{\mx{e}{e}}{\notg{\phi}}$ which only has
-relevance for the type system and which we assume is erased from
-programs prior to evaluation. We defer discussion of this form,
-as well as the syntactic category $\notg{\phi}$, to the next
-Section.
+rule for the form $\eqcast{\mx{e}{e}}{\notg{\phi}}$ which is a type
+annotation that we assume is erased from programs prior to
+evaluation. We defer discussion of this form, as well as the syntactic
+category $\notg{\phi}$, to the next Section.
 
@@
 
-\begin{mathpar}
-  \inferrule[Mesg]
-            {e_1 \redx x \\ e_2 \redx \be \\ e_3 \redx \cid \\ \atj{\toeq{\elab{\be}{\cid}}}{R_2}{\ty} \\
-              \itj{\cid}{\be}{V}}
-            {\mtj{\xassign{e_1}{e_2}{e_3}}{\eqs}{(x:\ty)}{R_1;R_2}{(x : \ity{\cid}{V})}{\eqs \wedge x \eop \toeq{\elab{\be}{\cid}}}}
-
-  \inferrule[Encode]
-            {e_1 \redx w \\ e_2 \redx \cid \\ \notg{\phi} \redx \phi \\
-              \eqs \models \toeq{\elab{\be}{\cid}} \eop \phi\\
-              \atj{\phi}{R}{\ty}}
-            {\mtj{\eqcast{\mx{e_1}{e_2}}{\notg{\phi}}}{\eqs}{(\mx{w}{\cid}:\ty)}{R}{\varnothing}{\eqs}}
-
-  \inferrule[App]
-            {\tsig(f) = \dht{y_1,\ldots,y_n}{\notg{\eqs_1}}{\notg{\Gamma}}{\notg{R}}{\notg{\Delta}}{\notg{\eqs_2}} \\
-              e_1 \redx \mv_1\ \cdots\ e_n \redx \mv_n \\
-              \subn = [\mv_1/y_1]\cdots[\mv_n/y_n] \\
-              \subn(\hty{\notg{\eqs_1}}{\notg{\Gamma}}{\notg{R}}{\notg{\Delta}}{\notg{\eqs_2}}) \redx
-                    \hty{\eqs_1}{\Gamma}{R}{\Delta}{\eqs_2} \\
-              \eqs \models \eqs_1}
-            {\mtj{f(e_1,\ldots,e_n)}{\eqs}{\Gamma}{R}{\Delta}{\eqs \wedge \eqs_2}}
-
-  \inferrule[Seq]          
-            {\mtj{\prog_1}{\eqs_1}{\Gamma_1}{R_1}{\Delta_1}{\eqs_2} \\
-             \mtj{\prog_2}{\eqs_2}{\Gamma_2}{R_2}{\Delta_2}{\eqs_3}}
-            {\mtj{\prog_1;\prog_2}{\eqs_1}{\Gamma_1;\Gamma_2}{R_1;R_2}{\Delta_1;\Delta_2}{\eqs_3}}
-\end{mathpar}
-
-\begin{mathpar}
-  \inferrule[Sig]
-            {\codebase(f) = y_1,\ldots,y_n, \instr \\
-              \subn = [\mv_1/y_1]\cdots[\mv_n/y_n] \\
-              \subn(\hty{\notg{\eqs_1}}{\notg{\Gamma}}{\notg{R}}{\notg{\Delta}}{\notg{\eqs_2}}) \redx
-                    \hty{\eqs_1}{\Gamma}{R}{\Delta}{\eqs_2} \\
-              \mtj{\subn(\instr)}{\eqs_1}{\Gamma}{R}{\Delta}{\eqs} \\
-              \eqs \models \eqs_2}
-            {f : \dht{y_1,\ldots,y_n}{\notg{\eqs_1}}{\notg{\Gamma}}{\notg{R}}{\notg{\Delta}}{\notg{\eqs_2}}}
-\end{mathpar}
+\mtjfig
 

### 2024-10-03 esop25/metalang.tex
checkpoint

@@
 $$
+A key property of this example and our type system generally is that once
+the $\Pi$ type is verified, typing any application of it requires verification
+of the precondition instance, but not the postcondition instance.
+
+\subsection{Algorithmic Type Validity in $\metaprot$}
 
@@
 
+We equate types up to evaluation of subexpressions as defined in
+Figure \ref{fig-notg}. Since expression evaluation is total we can
+just evaluate types of closed instructions, which have closed types,
+to obtain Hoare typings for instructions. In Figure \ref{fig-mtj} we
+define the type derivation rules for $\metaprot$. In the
+\TirName{Mesg} rule we use algorithmic confidentiality and integrity
+type checking in a straightforward manner.  In the \TirName{Encode}
+rule, we verify the hint given in the type annotation, and then use it
+for type checking.  In the \TirName{App} rule we verify preconditions
+of the given function type and instantiate the type and postconditions
+with the given expression arguments.  The \TirName{Sig} rule we
+specify how to verify $\Pi$ types for functions. Note that the values
+$\mv_1,\ldots,\mv_n$ chosen to instantiate the function parameters are
+chosen arbitrarily\cnote{Might need to be a little more careful about
+  this}.
+
+We allow the programmer to specify a type signature $\tsig$ that
+maps functions in $\codebase$ to valid dependent Hoare
+typings. Define:
 \begin{definition}
@@
 \end{definition}
+In practice, the programmer only needs to specify pre-~and post-conditions,
+since the rest of the type is reconstructed by type checking.
 
-\begin{theorem}
+The following Theorem establishes our main result, which
+is the $\metaprot$ type checking is sound with respect to both
+confidentiality and integrity typings. Since these typings
+imply security hyperproperties of confidentiality and integrity,
+$\metaprot$ type checking implicitly enforces them.
+\begin{theorem}[Soundness of $\metaprot$ type checking]
   Given preprocessing predicate $\eqspre$, program $\instr$, and verified $\tsig$, if
-  the judgement $\mtj{\instr}{\eqspre}{\Gamma}{R}{\Delta}{\eqs}$ is derivable then
+  the judgement: $$\mtj{\instr}{\eqspre}{\Gamma}{R}{\Delta}{\eqs}$$ is derivable then
   $\instr \redx \prog$ and:

### 2024-10-03 code/cvc5mvn/src/main/java/plaid/antlr/ExpressionVisitor.java
minus operator unary

@@ -2,6 +2,8 @@ package plaid.antlr;
 
 import plaid.PreludeBaseVisitor;
 import plaid.PreludeParser.FunctionCallExprContext;
+import plaid.PreludeParser.MinusExprContext;
+import plaid.PreludeParser.PlusExprContext;
 import plaid.ast.ConcatExpr;
 import plaid.ast.FieldExpr;
 import plaid.ast.FieldSelectExpr;
@@ -23,21 +25,20 @@ import plaid.ast.TimesExpr;
 import java.util.stream.Collectors;
 
 import static plaid.PreludeParser.AtExprContext;
-import static plaid.PreludeParser.FieldExprContext;
-import static plaid.PreludeParser.LetExprContext;
-import static plaid.PreludeParser.IdentExprContext;
 import static plaid.PreludeParser.ConcatExprContext;
+import static plaid.PreludeParser.FieldExprContext;
 import static plaid.PreludeParser.FieldSelectExprContext;
+import static plaid.PreludeParser.IdentExprContext;
+import static plaid.PreludeParser.LetExprContext;
 import static plaid.PreludeParser.MessageExprContext;
+import static plaid.PreludeParser.NumContext;
 import static plaid.PreludeParser.OutputExprContext;
 import static plaid.PreludeParser.ParenPExprContext;
-import static plaid.PreludeParser.PlusMinusExprContext;
 import static plaid.PreludeParser.PublicExprContext;
 import static plaid.PreludeParser.RandomExprContext;
 import static plaid.PreludeParser.SecretExprContext;
 import static plaid.PreludeParser.StrContext;
 import static plaid.PreludeParser.TimesExprContext;
-import static plaid.PreludeParser.NumContext;
 
 public class ExpressionVisitor extends PreludeBaseVisitor<PreludeExpression> {
 
@@ -124,10 +125,13 @@ public class ExpressionVisitor extends PreludeBaseVisitor<PreludeExpression> {
     }
 
     @Override
-    public PreludeExpression visitPlusMinusExpr(PlusMinusExprContext ctx) {
-        PreludeExpression e0 = visit(ctx.expr(0));
-        PreludeExpression e1 = visit(ctx.expr(1));
-        return ctx.pmop().getText().equals("+") ? new PlusExpr(e0, e1) : new MinusExpr(e0, e1);
+    public PlusExpr visitPlusExpr(PlusExprContext ctx) {
[…]

### 2024-10-03 code/cvc5mvn/src/main/java/plaid/ast/MessageExpr.java

@@ -25,12 +25,12 @@ public class MessageExpr implements MemoryExpr {
         if (this == o) return true;
         if (o == null || getClass() != o.getClass()) return false;
         MessageExpr that = (MessageExpr) o;
-        return Objects.equals(e, that.e);
+        return Objects.equals(e, that.e) && Objects.equals(i, that.i);
     }
 
     @Override
     public int hashCode() {
-        return Objects.hashCode(e);
+        return Objects.hash(e, i);
     }
 
     @Override

### 2024-10-03 code/cvc5mvn/src/main/java/plaid/ast/MinusExpr.java

@@ -5,20 +5,14 @@ import java.util.Objects;
 
 public class MinusExpr implements PreludeExpression {
 
-    private final PreludeExpression e1;
-    private final PreludeExpression e2;
+    private final PreludeExpression e;
 
-    public MinusExpr(PreludeExpression e1, PreludeExpression e2) {
-        this.e1 = e1;
-        this.e2 = e2;
+    public MinusExpr(PreludeExpression e) {
+        this.e = e;
     }
 
-    public PreludeExpression getE1() {
-        return e1;
-    }
-
-    public PreludeExpression getE2() {
-        return e2;
+    public PreludeExpression getE() {
+        return e;
     }
 
     @Override
@@ -26,17 +20,17 @@ public class MinusExpr implements PreludeExpression {
         if (this == o) return true;
         if (o == null || getClass() != o.getClass()) return false;
         MinusExpr minusExpr = (MinusExpr) o;
-        return Objects.equals(e1, minusExpr.e1) && Objects.equals(e2, minusExpr.e2);
+        return Objects.equals(e, minusExpr.e);
     }
 
     @Override
     public int hashCode() {
-        return Objects.hash(e1, e2);
+        return Objects.hashCode(e);
     }
 
     @Override
     public Iterable<Node> children() {
-        return List.of(e1, e2);
+        return List.of(e);
     }
 
 }

### 2024-10-03 code/cvc5mvn/src/test/java/plaid/antlr/ExpressionVisitorTest.java

@@ -9,6 +9,7 @@ import plaid.ast.Identifier;
 import plaid.ast.LetExpr;
 import plaid.ast.MessageExpr;
 import plaid.ast.MinusExpr;
+import plaid.ast.Node;
 import plaid.ast.Num;
 import plaid.ast.OutputExpr;
 import plaid.ast.PlusExpr;
@@ -58,16 +59,12 @@ public class ExpressionVisitorTest {
     }
 
     /**
-     * Subtraction and addition have equal precedence.
+     * Negation happens before party index.
      */
     @Test
-    public void additionAssoc() {
-        assertEquals(
-                new PlusExpr(new MinusExpr(new Num(3), new Num(2)), new Num(1)),
-                ast("3 - 2 + 1"));
-        assertEquals(
-                new MinusExpr(new PlusExpr(new Num(3), new Num(2)), new Num(1)),
-                ast("3 + 2 - 1"));
+    public void minusFirst() {
+        Node expr = ast("-m[z]@1");
+        assertEquals(new MinusExpr(new MessageExpr(new Identifier("z"), new Num(1))), expr);
     }
 
     /**

### 2024-10-03 code/cvc5mvn/src/main/java/plaid/antlr/CommandVisitor.java
let commands

@@ -5,11 +5,13 @@ import plaid.PreludeParser.AssertCommandContext;
 import plaid.PreludeParser.AssignCommandContext;
 import plaid.PreludeParser.CommandListContext;
 import plaid.PreludeParser.FunctionCallCommandContext;
+import plaid.PreludeParser.LetCommandContext;
 import plaid.ast.AssertCommand;
 import plaid.ast.AssignCommand;
 import plaid.ast.CommandList;
 import plaid.ast.FunctionCallCommand;
 import plaid.ast.Identifier;
+import plaid.ast.LetCommand;
 import plaid.ast.PreludeCommand;
 
 public class CommandVisitor extends PreludeBaseVisitor<PreludeCommand> {
@@ -41,6 +43,14 @@ public class CommandVisitor extends PreludeBaseVisitor<PreludeCommand> {
     public CommandList visitCommandList(CommandListContext ctx) {
         // TODO Either unit test or eliminate
         return new CommandList(ctx.command().stream().map(this::visit).toList());
+    }
+
+    @Override
+    public LetCommand visitLetCommand(LetCommandContext ctx) {
+        return new LetCommand(
+                new Identifier(ctx.ident().getText()),
+                Loader.toExpression(ctx.expr()),
+                Loader.toCommand(ctx.command()));
     }
 
 }

### 2024-10-03 code/cvc5mvn/src/main/java/plaid/ast/LetCommand.java

@@ -1,0 +1,16 @@
+package plaid.ast;
+
+import java.util.List;
+
+public record LetCommand(
+        Identifier identifier,
+        PreludeExpression expr,
+        PreludeCommand command)
+    implements PreludeCommand {
+
+    @Override
+    public Iterable<Node> children() {
+        return List.of(identifier, expr, command);
+    }
+
+}

### 2024-10-07 esop25/cpj.tex
checkpoint

@@
 
+\subsubsection{Type Safety}
+
+\cnote{Define (conditional) inputs in hyperproperties section.}
+
+\begin{definition}
+  \label{definition-sound}
+  Given $\prog$ with $\iov(\prog) = S \cup M \cup P \cup O$
+  and $\flips(\prog) = R$, we say that
+  \emph{$\Gamma$ is sound for $\prog$} iff for all $M' \subseteq M$
+      and $X \subseteq S \cup R$, the following
+      conditions hold:
+  \begin{enumerate}[\hspace{5mm}i.]
+    \item  if $X$ are inputs to $M'$ then
+      there exists $\ty$ with $\leakj{\Gamma}{M'}{\ty}$ and $X \subseteq \ty$.
+    \item  for all  $R' \subseteq R$, if
+      $X$ are $R'$ are conditional inputs to $M'$ then $\leakj{\Gamma}{M'}{\ty}$
+      and $\leakclose{\Gamma}{\ty \cup R'}{\ty'}$ and $X \subseteq \ty'$.
+  \end{enumerate}
+\end{definition}
+
+\begin{lemma}
+  \label{lemma-cpjsound}
+  If $\cpj{R}{\eqs}{\eqspre \wedge \toeq{\prog}}{\Gamma}$ is valid then $\Gamma$ is
+  sound for $\prog$.
+\end{lemma}
+\begin{proof}
+  Unrolling Definition \ref{definition-sound}, proceed by induction on $M'$.
+  In case $M' = \varnothing$ the result follows trivially. Otherwise XXX
+\end{proof}
+
 Our main type safety result combines typing and leakage judgements,
@@
 \begin{theorem}
-  If $\cpj{R}{\eqs}{\eqspre \wedge \toeq{\prog}}{\Gamma}$ is valid and there exists no $H,C$ 
-  and  $\sx{w}{\cid}$ for $\cid \in H$ with $\leakj{\Gamma}{C}{\setit{\sx{w}{\cid}}}$,
-  then $\prog$ satisfies gradual release.
+  Given $\prog$ with $\views(\prog) = M \cup P$, if $\cpj{R}{\eqs}{\eqspre \wedge \toeq{\prog}}{\Gamma}$
+  is valid and there exists no $H,C$ such that $\leakj{\Gamma}{M_C}{T}$ with $T$ closed
+  and $S_H \cap T \ne \varnothing$, then $\prog$ satisfies gradual release.
 \end{theorem}
+\begin{proof}
+  Immediate by Lemma \ref{lemma-cpjsound} and definition of gradual release.
+\end{proof}
 

### 2024-10-07 esop25/cpj.tex
checkpoint

@@
   Unrolling Definition \ref{definition-sound}, proceed by induction on $M'$.
-  In case $M' = \varnothing$ the result follows trivially. Otherwise XXX
+  In case $M' = \varnothing$ the result follows trivially. Otherwise
+  $M' = M_0 \cup \{ \mx{w}{\cid} \}$ and wlog $\Gamma = \Gamma_0; \mx{w}{\cid} : \ty_0$
+  and $\prog = \prog_0;\xassign{\mx{w}{\cid}{\be}{\cid_0}$, where
+    $\cpj{R_0}{\eqs}{\eqspre \wedge \toeq{\prog_0}}{\Gamma_0}$ is valid.    
 \end{proof}
@@
   Given $\prog$ with $\views(\prog) = M \cup P$, if $\cpj{R}{\eqs}{\eqspre \wedge \toeq{\prog}}{\Gamma}$
-  is valid and there exists no $H,C$ such that $\leakj{\Gamma}{M_C}{T}$ with $T$ closed
-  and $S_H \cap T \ne \varnothing$, then $\prog$ satisfies gradual release.
+  is valid and for all $H,C$ we have $\leakj{\Gamma}{M_C}{T}$ with $T$ closed
+  and $S_H \cap T = \varnothing$, then $\prog$ satisfies gradual release.
 \end{theorem}
 \begin{proof}
-  Immediate by Lemma \ref{lemma-cpjsound} and definition of gradual release.
+  Immediate by Lemma \ref{lemma-cpjsound}. \qed
 \end{proof}

### 2024-10-07 esop25/cpj.tex
checkpoint

@@
   $M' = M_0 \cup \{ \mx{w}{\cid} \}$ and wlog $\Gamma = \Gamma_0; \mx{w}{\cid} : \ty_0$
-  and $\prog = \prog_0;\xassign{\mx{w}{\cid}{\be}{\cid_0}$, where
+  and $\prog = \prog_0;\xassign{\mx{w}{\cid}}{\be}{\cid_0}$, where
   $\cpj{R_0}{\eqs}{\eqspre \wedge \toeq{\prog_0}}{\Gamma_0}$ is valid.    

### 2024-10-07 esop25/cpj.tex
checkpoint

--- 2024-10-07 esop25/cpj.tex
@@
 
+\begin{lemma}
+  \label{lemma-eqsprogsound}
+  If $\toeq{\prog} \models \eqs$ and $\eqs \models \eqs'$, then
+  for all $\store \in \runs(\prog)$ we have $\store \models \eqs'$.
+\end{lemma}
+
 \begin{definition}
@@
   $M' = M_0 \cup \{ \mx{w}{\cid} \}$ and wlog $\Gamma = \Gamma_0; \mx{w}{\cid} : \ty_0$
-  and $\prog = \prog_0;\xassign{\mx{w}{\cid}{\be}{\cid_0}$, where
-    $\cpj{R_0}{\eqs}{\eqspre \wedge \toeq{\prog_0}}{\Gamma_0}$ is valid.    
+  and $\prog = \prog_0;\xassign{\mx{w}{\cid}}{\be}{\cid_0}$, where $R = R_0;R_1$ and:
+  \begin{mathpar}
+    \cpj{R_0}{\eqs}{\eqspre \wedge \toeq{\prog_0}}{\Gamma_0}
+
+    \eqj{R_1}{\eqs}{\toeq{\elab{\be}{\cid}}}{\ty_0}
+  \end{mathpar}
+  are both valid, and by the induction hypothesis $\Gamma_0$ is sound for $\prog_0$.     
 \end{proof}

### 2024-10-07 esop25/ipj.tex
checkpoint

@@
 
+\subsection{Assigning Integrity Labels}
+
 In all cases, we want to determine the integrity level of computed
@@
 axiomatized in Figure \ref{fig-cheatj}.
+
+\subsubsection{Type Safety}
+
+\begin{definition}
+  Given $H,C$ and $\prog$ with $\views(\prog) = V$ and $\secrets(\prog) = S$,
+  $\adversary$, and $\store \in \aruns(\prog)$, we say $\store$ is
+  \emph{abberant} if $\neg\exists \store' \in \runs(\prog) .
+  \store_{S_H \cup V} = \store'_{S_H \cup V}$, in which case we say
+  any $x \in V$ \emph{has an abberation}.
+\end{definition}
+
+\begin{lemma}
+  If $\prog$ does not have integrity for some $H,C$, then it has
+  abberant $\store \in \aruns(\prog)$ for some $\adversary$. 
+\end{lemma}
+
+\begin{lemma}
+  Given pre-processing predicate $\eqspre$ and protocol $\prog$ with
+  $\views(\prog) = V$, if 
+  $\ipj{\eqs}{\prog}{\Delta}$ is valid
+  and $x\in V$ has an abberation for some $H,C$,
+  then $\cheatj{\Delta}{H,C}{\seclev}$ and $\seclev(x) = \lolab$.
+\end{lemma}
+
 Our main type safety result for integrity is then given as follows.

### 2024-10-07 esop25/ipj.tex
checkpoint

@@
   and for all $H,C$ with $\cheatj{\Delta}{H,C}{\seclev}$ 
-  we have $\seclev(x) = \hilab$ for all $x \in \houtputs$, then cheating
-  is detectable in $\prog$.
+  we have $\seclev(x) = \hilab$ for all $x \in \houtputs$, then $\prog$
+  has integrity.
 \end{theorem}

### 2024-10-07 esop25/cpj.tex
checkpoint

@@
 
-\subsubsection{Type Safety}
+\subsubsection{Type Soundness}
 
@@
   are both valid, and by the induction hypothesis $\Gamma_0$ is sound for $\prog_0$.
+  We can then proceed by a second induction on the derivation of $\ty_0$, where there
+  are two subcases. XXX
 \end{proof}
 
-Our main type safety result combines typing and leakage judgements,
+Our main type soundness result combines typing and leakage judgements,
 and guarantees gradual release of protocols-- that is, that adversarial

### 2024-10-07 esop25/ipj.tex
checkpoint

@@
   \emph{abberant} if $\neg\exists \store' \in \runs(\prog) .
-  \store_{S_H \cup V} = \store'_{S_H \cup V}$, in which case we say
+  \store_{S_H \cup V} = \store'_{S_H \cup V}$, and if
+   $\neg\exists \store' \in \runs(\prog) .
+  \store_{S_H \cup \{ x \}} = \store'_{S_H \cup \{ x \}}$ we say
   any $x \in V$ \emph{has an abberation}.
@@
 
-Our main type safety result for integrity is then given as follows.
+Our main type soundness result for integrity is then given as follows.
 \begin{theorem}
@@
   we have $\seclev(x) = \hilab$ for all $x \in \houtputs$, then $\prog$
-  has integrity.
+  has integrity for all $H,C$.
 \end{theorem}

### 2024-10-07 esop25/ipj.tex
checkpoint

@@
   $\adversary$, and $\store \in \aruns(\prog)$, we say $\store$ is
-  \emph{abberant} if $\neg\exists \store' \in \runs(\prog) .
-  \store_{S_H \cup V} = \store'_{S_H \cup V}$, and if
-   $\neg\exists \store' \in \runs(\prog) .
-  \store_{S_H \cup \{ x \}} = \store'_{S_H \cup \{ x \}}$ we say
-  any $x \in V$ \emph{has an abberation}.
+  \emph{aberrant through $V' \subseteq V$} if $\neg\exists \store' \in \runs(\prog) .
+  \store_{S_H \cup V'} = \store'_{S_H \cup V'}$, and if there exists
+  such $\adversary$ and $\store$ we say any
+  $x \in V'$ is \emph{in an aberration}.
 \end{definition}
@@
   If $\prog$ does not have integrity for some $H,C$, then it has
-  abberant $\store \in \aruns(\prog)$ for some $\adversary$. 
+  aberrant $\store \in \aruns(\prog)$ for some $\adversary$. 
+\end{lemma}
+
+\begin{lemma}
+  Given $\prog$ and $H,C$, if $x$ is in an aberration then one of
+  the two inductively defined conditions holds:
+  \begin{enumerate}[\hspace{5mm}i.]
+  \item There exists $\xassign{x}{\be}{\cid} \in \prog$ with $\cid \in C$.
+  \item There exists $\xassign{x}{\be}{\cid} \in \prog$ with $\cid \in H$ and
+    $x' \in \vars(\be)$ such that $x'$ is in an aberration.
+  \end{enumerate}
 \end{lemma}
@@
   $\ipj{\eqs}{\prog}{\Delta}$ is valid
-  and $x\in V$ has an abberation for some $H,C$,
+  and $x\in V$ has an aberration for some $H,C$,
   then $\cheatj{\Delta}{H,C}{\seclev}$ and $\seclev(x) = \lolab$.

### 2024-10-07 code/cvc5mvn/src/test/java/plaid/antlr/ExpressionVisitorTest.java
underscores in identifiers

@@ -199,4 +199,20 @@ public class ExpressionVisitorTest {
                 new Identifier("b"), new Num(1))), ast("{a=0; b=1}"));
     }
 
+    /**
+     * Identifiers can be mixed case.
+     */
+    @Test
+    public void mixedCaseIdentifiers() {
+        assertEquals(new Identifier("aA"), ast("aA"));
+    }
+
+    /**
+     * Identifiers can start with underscores.
+     */
+    @Test
+    public void underscoreIdentifiers() {
+        assertEquals(new Identifier("_1"), ast("_1"));
+    }
+
 }

### 2024-10-09 esop25/cpj.tex
checkpoint

@@
 want to track the dependence of $\mx{w}{2}$ and $\rx{w}{2}$.
-In short, we assign $\mx{w}{1}$ a ``ciphertext type'':
+In short, we assign $\mx{w}{1}$ a ``ciphertext type'', aka
+a \emph{ciphertype}:
 $$
@@
 \end{mathpar}
-This allows us to assign the same ciphertext type to $\mx{w}{1}$ as
+This allows us to assign the same ciphertype to $\mx{w}{1}$ as
 above-- that is, $\cty{\rx{w}{2}}{\setit{\sx{w}{2}}}$. Leveraging the
@@
 interpretation as terms $\phi$ and constraints $\eqs$ respectively,
-via the encoding $\toeq{cdot}$. The \TirName{Encode} rule captures
+via the encoding $\toeq{\cdot}$. The \TirName{Encode} rule captures
 the use ``fresh'' samples for encryption discussed above. Derivations

### 2024-10-09 esop25/examples.tex
checkpoint

@@
 \section{Extended Examples}
+
+We consider several examples that illustrate the versatility of the $\TirName{Encode}$
+typing rule for confidentiality, the use of postconditions for correctness,
+and the use of both pre-~and postconditions for integrity.
 
@@
 
+The $\ttt{encode}$ function example in Section \ref{section-metalangty}
+illustrates the use of type annotations to track confidentiality. In
+2-party Yao's Garbled Circuits (YGC), a \emph{garbler} prepares a circuit encoding
+that is shared with a \emph{evaluator}, and the $\ttt{encode}$ function
+defines the method the garbler uses to share their input secret encodings
+with the evaluator. The example illustrates how confidentiality is reflected
+in its (cipher)type.
+
+In the 2-party GMW protocol \cite{evans2018pragmatic}, another garbled
+binary circuit protocol, values $v$ are encoded as reconstructive
+shares $v_1$ and $v_2$ held by each party with $v = v_1 \fplus v_2$.
+We will follow the convention that shared values are identified by
+strings $w$ and encoded as shares $\mx{w}{1}$ and $\mx{w}{2}$.  As for
+YGC, ciphertypes reflect the confidentiality of GMW input encodings as
+defined in the $\ttt{encodegmw}$ function in Figure \ref{fig-gmw}. No
+programmer annotation is needed given the syntax of the function
+body. More interestingly, the $\ttt{andgmw}$ gate definition shows how
+a programmer hint can express the relevant confidentiality property of
+the output share $\mx{z}{2}$ on client 2. The non-trivial equivalence
+can be verified by SMT once during verification, and subsequently
+confidentiality is expressed in the dependent type of the function
+(as for $\ttt{encode}$).
+
+Note also that the $\ttt{andgmw}$ function is decorated with a
+postcondition that expresses the correctness property of the
+function. Although not strictly necessary for confidentiality, this
+annotation can at least help eliminate bugs, and also can be used
+compositionally for whole-program correctness properties. As for
+confidentiality hints, this postcondition needs to be verified only
+once at the point of definition and subsequently is guaranteed to hold
+for any application.
+
+\input{gmw}
+
+In the 2-party GMW protocol \cite{evans2018pragmatic}, another garbled
+binary circuit protocol, values $v$ are encoded as reconstructive
+shares $v_1$ and $v_2$ held by each party with $v = v_1 \fplus v_2$.
+We will follow the convention that shared values are identified by
+strings $w$ and encoded as shares $\mx{w}{1}$ and $\mx{w}{2}$.  As for
+YGC, ciphertypes reflect the confidentiality of GMW input encodings as
+defined in the $\ttt{encodegmw}$ function in Figure \ref{fig-gmw}. No
+programmer annotation is needed given the syntax of the function
+body. More interestingly, the $\ttt{andgmw}$ gate definition shows how
+a programmer hint can express the relevant confidentiality property of
+the output share $\mx{z}{2}$ on client 2. The non-trivial equivalence
+can be verified by SMT once during verification, and subsequently
+confidentiality is expressed in the dependent type of the function
+(as for $\ttt{encode}$).
+
+Note also that the $\ttt{andgmw}$ function is decorated with a
+postcondition that expresses the correctness property of the
+function. Although not strictly necessary for confidentiality, this
+annotation can at least help eliminate bugs, and also can be used
+compositionally for whole-program correctness properties. As for
+confidentiality hints, this postcondition needs to be verified only
+once at the point of definition and subsequently is guaranteed to hold
+for any application.
+
+\subsubsection{Integrity Examples}
+
+\input{bdoz}
+
+We can capture both the preliminary distribution of Beaver triples and
+BDOZ shares as a pre-processing predicate that establishes conditions
+for initial memories (see Definition \ref{def-aprogtt}).  Here we
+assume two input secrets $\elab{\secret{x}}{1}$ and
+$\elab{\secret{y}}{2}$ and a single Beaver Triple to compute
+$\elab{\secret{x}}{1} \ftimes \elab{\secret{y}}{2}$, but we can extend
+this for additional gates.  As for GMW, we use
+$\macgv{\mesg{w}}$\cnote{need to update this} to refer to
+secret-shared values reconstructed with addition, but for BDOZ by
+convention each share of $\macgv{\mesg{w}}$ is represented as
+$\elab{\mesg{w\ttt{s}}}{\cid}$, the MAC of which is represented as a
+$\elab{\mesg{w\ttt{m}}}{\cid}$ for all $\cid$, and each client holds a
+key $\elab{\mesg{w\ttt{k}}}{\cid}$ for authentication of the other's
+share.
+$$
+\begin{array}{l}
+\mx{\ttt{"xs"}}{2} \eop \sx{\ttt{"x"}}{1} \fplus \rx{\ttt{"x"}}{1} \wedge 
+\mx{\ttt{"xs"}}{1} \eop \rx{\ttt{"x"}}{1} \wedge \\
+\mx{\ttt{"ys"}}{1} \eop \sx{\ttt{"y"}}{2} \fplus \rx{\ttt{"y"}}{2} \wedge 
+\mx{\ttt{"ys"}}{2} \eop \rx{\ttt{"y"}}{2} 
+\end{array}
+$$
+
+$$
+\begin{array}{l}
+\mx{\ttt{"delta"}}{1} \eop \rx{\ttt{"delta"}}{1} \wedge
+\mx{\ttt{"xk"}}{1} \eop \rx{\ttt{"xk"}}{1} \wedge\\
+\mx{\ttt{"xm"}}{2} \eop \mx{\ttt{"xk"}}{1} \fplus (\mx{\ttt{"delta"}}{1} * \mx{"xs"}{2})
+\end{array}
+$$
+
+$$
+\begin{array}{l}
+\mx{\ttt{"g1as"}}{1} \eop \rx{\ttt{"g1as"}}{1} \wedge 
+\mx{\ttt{"g1bs"}}{1} \eop \rx{\ttt{"g1bs"}}{1} \wedge\\
+\mx{\ttt{"g1as"}}{2} \eop \rx{\ttt{"g1as"}}{2} \wedge 
+\mx{\ttt{"g1bs"}}{2} \eop \rx{\ttt{"g1bs"}}{2} \wedge\\
+\mx{\ttt{"g1cs"}}{1} \fplus \mx{\ttt{"g1cs"}}{2} \eop \\
+\quad (\mx{\ttt{"g1as"}}{1} \fplus \mx{\ttt{"g1bs"}}{2}) \ftimes
+      (\mx{\ttt{"g1bs"}}{1} \fplus \mx{\ttt{"g1bs"}}{2}) 
+\end{array}
+$$
+
+%%%%%%%%%%%%%%%%%%
+%%% ARCHIVAL BELOW
+%%%%%%%%%%%%%%%%%%
+\begin{comment}
 \begin{verbatimtab}
@@
 \end{verbatimtab}
+
+\end{comment}

### 2024-10-09 esop25/ipj.tex
checkpoint

@@
 As in \cite{skalka-near-ppdp24}, we use $\macgv{\mesg{w}}$ to refer to
-secret-shared values reconstructed with addition, and by
+secret-shared values reconstructed with addition\cnote{needed?}, and by
 convention each share of $\macgv{\mesg{w}}$ is represented as

### 2024-10-09 esop25/smt.tex
checkpoint

@@
 except the can ``cross party lines''. This turns out to be very useful
-to express correctness properties-- for example, in the GMW protocol
-wire values in circuits are represented by reconstructive shares
-\cite{evans2018pragmatic}.  If by convention shares of values $n$ are
-represented by $\mesg{n}$ on clients, then assuming two clients
-$\setit{1,2}$ the reconstructed value can be expressed as $\mx{n}{1} +
-\mx{n}{2}$.  So, while summing values across clients is disallowed in
-$\minicat$ protocols, we can express properties of shared values
-via constraints.
+to express correctness properties-- for example, in the
+Goldreich-Micali-Wigderson (GMW) protocol wire values in circuits are
+represented by reconstructive shares \cite{evans2018pragmatic}.  If by
+convention shares of values $n$ are represented by $\mesg{n}$ on
+clients, then assuming two clients $\setit{1,2}$ the reconstructed
+value can be expressed as $\mx{n}{1} + \mx{n}{2}$.  So, while summing
+values across clients is disallowed in $\minicat$ protocols, we can
+express properties of shared values via constraints.
 

### 2024-10-09 esop25/examples.tex
checkpoint

@@
 \section{Extended Examples}
+\label{section-examples}
 
-We consider several examples that illustrate the versatility of the $\TirName{Encode}$
+We consider examples that illustrate the versatility of the $\TirName{Encode}$
 typing rule for confidentiality, the use of postconditions for correctness,
@@
 \mx{\ttt{"xk"}}{1} \eop \rx{\ttt{"xk"}}{1} \wedge\\
-\mx{\ttt{"xm"}}{2} \eop \mx{\ttt{"xk"}}{1} \fplus (\mx{\ttt{"delta"}}{1} * \mx{"xs"}{2})
+\mx{\ttt{"xm"}}{2} \eop \mx{\ttt{"xk"}}{1} \fplus (\mx{\ttt{"delta"}}{1} * \mx{\ttt{"xs"}}{2})
 \end{array}
@@
 \begin{array}{l}
-\mx{\ttt{"g1as"}}{1} \eop \rx{\ttt{"g1as"}}{1} \wedge 
-\mx{\ttt{"g1bs"}}{1} \eop \rx{\ttt{"g1bs"}}{1} \wedge\\
-\mx{\ttt{"g1as"}}{2} \eop \rx{\ttt{"g1as"}}{2} \wedge 
-\mx{\ttt{"g1bs"}}{2} \eop \rx{\ttt{"g1bs"}}{2} \wedge\\
-\mx{\ttt{"g1cs"}}{1} \fplus \mx{\ttt{"g1cs"}}{2} \eop \\
-\quad (\mx{\ttt{"g1as"}}{1} \fplus \mx{\ttt{"g1bs"}}{2}) \ftimes
-      (\mx{\ttt{"g1bs"}}{1} \fplus \mx{\ttt{"g1bs"}}{2}) 
+\mx{\ttt{"g1as"}}{1} \eop \rx{\ttt{"g1as"}}{1}\ \wedge 
+\mx{\ttt{"g1bs"}}{1} \eop \rx{\ttt{"g1bs"}}{1}\ \wedge\\
+\mx{\ttt{"g1as"}}{2} \eop \rx{\ttt{"g1as"}}{2}\ \wedge 
+\mx{\ttt{"g1bs"}}{2} \eop \rx{\ttt{"g1bs"}}{2}\ \wedge\\
+\mx{\ttt{"g1cs"}}{1} \eop \\
+\qquad ((\mx{\ttt{"g1as"}}{1} \fplus \mx{\ttt{"g1bs"}}{2})\ \ftimes\\
+\qquad\phantom{(}(\mx{\ttt{"g1bs"}}{1} \fplus \mx{\ttt{"g1bs"}}{2})) \fminus \rx{\ttt{"g1cs"}}{2}\ \wedge\\
+\mx{\ttt{"g1cs"}}{2} \eop \rx{\ttt{"g1cs"}}{2}
 \end{array}
+$$
+
+$$
+\ttt{mult}(\ttt{"g1"},\ttt{"x"},\ttt{"y"})
 $$

### 2024-10-09 esop25/lang.tex
checkpoint

--- 2024-08-24 csf25/lang.tex
@@
 \end{definition}
-
-\cnote{Add comments about OT.}
+An example of a $\minicat$ protocol, for additive secret
+sharing, is discussed in Section \ref{section-smt-toeq}.
 
@@
 \begin{definition}
-  Memories $\store$
-are finite (partial) mappings from variables $x$ to values $v \in
-\mathbb{F}_p$.  The \emph{domain} of a memory is written
+Memories $\store$ are finite (partial) mappings from variables $x$ to
+values $v \in \mathbb{F}_p$.  The \emph{domain} of a memory is written
 $\dom(\store)$ and is the finite set of variables on which the memory
@@
 \subseteq \dom(\store')$ and $\store(x) = \store'(x)$ for all $x \in
-\dom(\store)$. Given any $\store$ and $\store'$ with
-$\store(x) = \store'(x)$ for all $x \in \dom(\store) \cap \dom(\store')$,
-we write $\store \uplus \store'$ to denote the memory
-with domain $X = \dom(\store) \cup \dom(\store')$ such
-that:
+\dom(\store)$. Given any $\store$ and $\store'$ with $\store(x) =
+\store'(x)$ for all $x \in \dom(\store) \cap \dom(\store')$, we write
+$\store \uplus \store'$ to denote the memory with domain $X =
+\dom(\store) \cup \dom(\store')$ such that:
 $$
@@
 \end{definition}
-So for example, given a protocol $\prog$, the set of all random tapes for
-$\prog$ is $\mems(\flips(\prog))$, and the memory $\store_{\secrets(\prog)}$
+So for example, the set of all random tapes for
+for a protocol $\prog$ is $\mems(\flips(\prog))$, and the memory $\store_{\secrets(\prog)}$
 is $\store$ restricted to the secrets in $\prog$.
@@
 output- are implemented as updates to the memory $\store$.
+
+\paragraph{Oblivious Transfer} A passive secure oblivious transfer (OT) protocol
+based on previous work \cite{barthe2019probabilistic} can be defined
+in $\minicat$, albeit assuming some shared randomness. Alternatively,
+a simple passive secure OT can be defined with the addition of public
+key cryptography as a primitive. But given the diversity of approaches
+to OT, we instead assume that OT is abstract with respect to its
+implementation, where calls to OT in $\mathbb{F}_2$ are of the form
+$\OT{\elab{\be_1}{\cid}}{\be_2}{\be_3}$ -- given a \emph{choice bit}
+$\be_1$ provided by a receiver $\cid$, the sender sends either $\be_2$
+or $\be_3$.  Critically, the sender learns nothing about $\be_1$ and
+the receiver learns nothing about the unselected value. We return to
+the formal interpretation of OT in Section \ref{section-smt}.
 
@@
     \begin{array}{rcl@{\hspace{2mm}}r}
-      \prog &::=& \cdots \mid \assert{\be = \be}
+      \prog &::=& \cdots \mid \elab{\assert{\be = \be}}{\cid}
     \end{array}
@@
 $\stores$ or $\bot$.
+
+\subsubsection{Protocol Runs}
+
+Our semantics require that random tapes contain values for all program
+values $\elab{\flip{w}}{\cid}$ sampled from a uniform distribution
+over $\mathbb{F}_p$. Input memories also contain input secret values
+and possibly other initial view elements as a result of
+pre-processing, e.g., Beaver triples for efficient multiplication,
+and/or MACed share distributions as in BDOZ/SPDZ
+\cite{evans2018pragmatic,10.1007/978-3-030-68869-1_3}. We define
+$\runs(\prog)$ as the set of final memories resulting from execution
+of $\prog$ given any initial memory, possibly augmented with a
+preprocessed memory (a memory that satsfies a predicate
+$\preproc$), and $\aruns(\prog)$ is the analog in the malicious
+setting. This formulation will be fundamental to our consideration
+of probability distributions and hyperproperties of protocols.
+%In this
+%setting, given a program $\prog$ with $\iov(\prog) = S \cup V$ and
+%$\flips(\prog) = F$ we will consider all $\store \in \mems(S \cup V
+%\cup F)$ such that $ \config{\store_{S \cup F}}{\prog} \redxs
+%\config{\store_}{\varnothing} $ to be equally probable, establishing
+%the basic distribution of the program. %From this, we can immediately
+%derive the marginal distribution of $S \cup V$ to reason about
+%dependencies between secrets and views.
+\begin{definition}
+  \label{def-runs}
+  Given $\prog$ with $\secrets(\prog) = S$ and $\flips(\prog) = R$ and
+  predicate $\preproc$ on memories, define:
+  $$
+  \begin{array}{c}
+    \runs(\prog) \defeq 
+    \{ \store \mid \exists \store_1 \in \mems(S \cup R) . 
+    \exists \store_2 . \preproc(\store_2) \wedge
+    %(\dom(\store) = \iov(\prog) \cup R) \wedge
+    (\store_1 \uplus \store_2,\prog) \redxs \store \}
+  \end{array}
+  $$
+  And similarly for any $\adversary$ define:
+  $$
+  \begin{array}{c}
+    \aruns(\prog) \defeq 
+    \{ \store \mid \exists \store_1 \in \mems(S \cup R) . 
+    \exists \store_2 . \preproc(\store_2) \wedge
+    %(\dom(\store) = \iov(\prog) \cup R) \wedge
+    (\store_1 \uplus \store_2,\prog) \aredxs \store \}
+  \end{array}
+  $$
+\end{definition}

### 2024-10-09 esop25/lang.tex
checkpoint

@@
 \end{definition}
-An example of a $\minicat$ protocol, for additive secret
-sharing, is discussed in Section \ref{section-smt-toeq}.
+An example of a $\minicat$ protocol, for additive secret sharing, is
+discussed in Section \ref{section-smt-toeq}.  We disallow overwriting
+of variables, since assignment is intended to represent distinct
+message sends.
 
@@
 of $\prog$ given any initial memory, possibly augmented with a
-preprocessed memory (a memory that satsfies a predicate
-$\preproc$), and $\aruns(\prog)$ is the analog in the malicious
-setting. This formulation will be fundamental to our consideration
-of probability distributions and hyperproperties of protocols.
+preprocessed memory (a memory that satsfies a predicate $\preproc$),
+and $\aruns(\prog)$ is the analog in the malicious setting. Since we
+constrain programs to not overwrite any variable, we are assured that
+final memories contain both a complete record of all initial secrets
+as well as views resulting from communicated information. This
+formulation will be fundamental to our consideration of probability
+distributions and hyperproperties of protocols.
 %In this

### 2024-10-09 esop25/smt.tex
checkpoint

@@
 \section{$\minicat$ Constraint Verification}
+\label{section-smt}
 
@@
   SMT expressions}{fig-eqs}
-\small{
 $$
 \begin{array}{rcl}
-  \phi &::=& x \mid \phi \fplus \phi \mid \phi \fminus \phi \mid \phi \ftimes \phi \\
-  \eqs &::=& \phi \eop \phi \mid \eqs \wedge \eqs 
+  \phi ::= x \mid \phi \fplus \phi \mid \phi \fminus \phi \mid \phi \ftimes \phi  \qquad
+  \eqs ::= \phi \eop \phi \mid \eqs \wedge \eqs 
 \end{array}
@@
   \store(\phi_1 \fminus \phi_2) = \cod{\store(\phi_1) \fminus \store(\phi_2)}
-\end{mathpar}}
+\end{mathpar}
 \end{fpfig}
@@
 \subsection{Programs as Constraint Systems}
+\label{section-smt-toeq}
 
@@
 \end{mathpar}
-
 \begin{mathpar}
@@
 \end{mathpar}
-The correctness of this interpretation is characterized as follows.  
+In general we will assume that any preprocessing predicate is defined
+via a constraint $\eqspre$, i.e., $\preproc(\store) \iff \store
+\models \eqspre$ for all $\store$.  The correctness of the
+SMT interpretation of programs is characterized as follows.
 \begin{theorem}
   $\store$ is a model of $\eqspre \wedge \toeq{\prog}$ iff $\store \in \runs(\prog)$.
-  \cnote{Need to define $\runs$ somewhere.}
 \end{theorem}
@@
 Consider the following 3-party additive secret sharing protocol.
-
+{\footnotesize
 $$
+\begin{array}{c@{\qquad}c}
 \begin{array}{lll}
@@
   \elab{\mesg{s3}}{1} &:=& \elab{(\secret{3} \fminus \locflip \fminus \flip{x})}{3} \\ 
-  \elab{\mesg{s3}}{2} &:=& \elab{\flip{x}}{3}\\
+  \elab{\mesg{s3}}{2} &:=& \elab{\flip{x}}{3}
+\end{array} & 
+\begin{array}{lll}
   \rvl{1} &:=& \elab{(\locflip \fplus \mesg{s2} \fplus \mesg{s3})}{1} \\ 
@@
 \end{array}
-$$
+\end{array}
+$$}
 Letting $\prog$ be this protocol, we can verify correctness

### 2024-10-09 esop25/cpj.tex
checkpoint

@@
 may depend on. But we also want to capture this fundamental mechanism
-of encryption.
+of encryption. As we will show, SMT can be used in this and
+other contexts to support confidentiality properties of protocols.
 
@@
 (YGC) in $\mathbb{F}_{2}$, the ``garbler'' represents encrypted wire
-values as a random sample (for 1) or its negation (for 0), and the
-``evaluator'' obtains the wire value for its input via OT. The
-evaluator shares the initial encrypted wire value of one of its input
-secrets as a selection bit-- assuming that client 2 is the garbler and
-client 1 is the evaluator, we can write this as:
+values as a random sample (for 1) or its negation (for 0), and shares
+its secret input encoding with the ``evaluator'' by using
+it as a selection bit. That is, assuming that client 2 is the garbler and
+client 1 is the evaluator, we can define:
 $$
@@
 The syntax of types $\ty$ and type environments $\Gamma$ are presented
-in Figure \ref{fig-cpj}\cnote{need to fix set notations}. We also
+in Figure \ref{fig-cpj}. We also
 define type judgements for expressions and programs by direct
@@
 are added to judgements, and guaranteed to not be used elsewhere
-via a form of union that ensures disjointness of combined sets. 
+via a form of union that ensures disjointness of combined sets:
 \begin{definition}
-  $R_1;R_2 = R_1 \cup R_2$ iff $R_1 \cap R_2 = \varnothing$.
+  $R_1;R_2 = R_1 \cup R_2$ if $R_1 \cap R_2 = \varnothing$ and is otherwise
+  undefined.
 \end{definition}
@@
 sequenced programs.  Given this we define validity of program type
-judgements as follows. \cnote{Define $\eqspre$.}
+judgements as follows. 
 \begin{definition}

### 2024-10-09 esop25/ipj.tex
checkpoint

--- 2024-10-07 esop25/ipj.tex
@@
 \label{section-ipj}
-
-\subsection{BDOZ MAC Integrity}
 
@@
 The integrity of values during protocol executions depends on the
-partitioning $H,C$, since any value received from a member of
-$C$ is initially considered to be low integrity. But as for
-confidentiality types, we want our basic type analysis to be
-generalizable to arbitrary $H,C$\cnote{Add related text to that
-  section.}. So integrity type judgements are of the form
+partitioning $H,C$, since any value received from a member of $C$ is
+initially considered to be low integrity. But as for confidentiality
+types, we want our basic type analysis to be generalizable to
+arbitrary $H,C$. So integrity type judgements are of the form
 $\ipj{\eqs}{\prog}{\Delta}$, where $\eqs$ plays a similar role as in
@@
 information. Environments are ordered lists, since order of evaluation
-is important for tracking integrity-- validating low
-integrity information through MAC checking allows \emph{subsequent}
-treatment of it as high integrity, but not \emph{before}.
+is important for tracking integrity-- validating low integrity
+information through MAC checking allows \emph{subsequent} treatment of
+it as high integrity, but not \emph{before}. Note that MAC checking is
+\emph{not} a form of endorsement as it is usually defined (as a dual
+of declassification \cite{sabelfeld2009declassification}), since a
+successful check guarantees no deviation from the protocol.
 
@@
 $$
-More formally, we defined validity as follows:
+More formally, we define validity as follows:
 \begin{definition}
@@
 
-In the above example, note that environment implicitly records
+In the above example, note that the environment implicitly records
 that $\rvl{w}$ depends on data from client 1, which in turn
@@
 \end{definition}
-During execution of a protocol, the integrity of messages and reveals
-affected by values sent by other parties, as recorded in $\Delta$
-obtained from its typing. We additionally need to determine
-whether this impacts integrity, which we accomplish through
+But the values of messages and reveals may be
+affected by values sent by other parties, which is
+recorded in the $\Delta$ obtained from its typing. To determine
+whether this impacts integrity for any given $H,C$, we apply
 the inductively defined rewrite relation $\cheatj{\Delta}{H,C}{\seclev}$,
-axiomatized in Figure \ref{fig-cheatj}.
+axiomatized in Figure \ref{fig-cheatj}. The composition
+of typing and this latter rewriting obtains essentially
+a traditional taint analysis, modulo MAC checking.
 

### 2024-10-09 esop25/examples.tex
checkpoint

@@
 \subsection{Confidentiality Examples}
+
+\input{gmw}
 
@@
 In the 2-party GMW protocol \cite{evans2018pragmatic}, another garbled
-binary circuit protocol, values $v$ are encoded as reconstructive
-shares $v_1$ and $v_2$ held by each party with $v = v_1 \fplus v_2$.
-We will follow the convention that shared values are identified by
+binary circuit protocol, values are encrypted in a manner similar
+to BDOZ as described in Section \ref{section-ipj}. In our definition
+of GMW we use the convention that shared values $\macgv{w}$ are identified by
 strings $w$ and encoded as shares $\mx{w}{1}$ and $\mx{w}{2}$.  As for
@@
 
-\input{gmw}
-
-In the 2-party GMW protocol \cite{evans2018pragmatic}, another garbled
-binary circuit protocol, values $v$ are encoded as reconstructive
-shares $v_1$ and $v_2$ held by each party with $v = v_1 \fplus v_2$.
-We will follow the convention that shared values are identified by
-strings $w$ and encoded as shares $\mx{w}{1}$ and $\mx{w}{2}$.  As for
-YGC, ciphertypes reflect the confidentiality of GMW input encodings as
-defined in the $\ttt{encodegmw}$ function in Figure \ref{fig-gmw}. No
-programmer annotation is needed given the syntax of the function
-body. More interestingly, the $\ttt{andgmw}$ gate definition shows how
-a programmer hint can express the relevant confidentiality property of
-the output share $\mx{z}{2}$ on client 2. The non-trivial equivalence
-can be verified by SMT once during verification, and subsequently
-confidentiality is expressed in the dependent type of the function
-(as for $\ttt{encode}$).
-
-Note also that the $\ttt{andgmw}$ function is decorated with a
-postcondition that expresses the correctness property of the
-function. Although not strictly necessary for confidentiality, this
-annotation can at least help eliminate bugs, and also can be used
-compositionally for whole-program correctness properties. As for
-confidentiality hints, this postcondition needs to be verified only
-once at the point of definition and subsequently is guaranteed to hold
-for any application.
-
-\subsubsection{Integrity Examples}
+\subsection{Integrity Examples}
 
@@
 
-We can capture both the preliminary distribution of Beaver triples and
-BDOZ shares as a pre-processing predicate that establishes conditions
-for initial memories (see Definition \ref{def-aprogtt}).  Here we
-assume two input secrets $\elab{\secret{x}}{1}$ and
-$\elab{\secret{y}}{2}$ and a single Beaver Triple to compute
-$\elab{\secret{x}}{1} \ftimes \elab{\secret{y}}{2}$, but we can extend
-this for additional gates.  As for GMW, we use
-$\macgv{\mesg{w}}$\cnote{need to update this} to refer to
-secret-shared values reconstructed with addition, but for BDOZ by
-convention each share of $\macgv{\mesg{w}}$ is represented as
-$\elab{\mesg{w\ttt{s}}}{\cid}$, the MAC of which is represented as a
-$\elab{\mesg{w\ttt{m}}}{\cid}$ for all $\cid$, and each client holds a
-key $\elab{\mesg{w\ttt{k}}}{\cid}$ for authentication of the other's
-share.
-$$
+Returning to the example of malicious-secure 2-party BDOZ arithmetic
+circuits begun in Section \ref{section-ipj}, in Figure \ref{fig-bdozsum} we
+define $\ttt{sum}$ and $\ttt{open}$ functions. The latter implements
+``secure opening''-- each party sends it's local share of a
+global value $\macgv{w}$ along with its MAC to the other party,
+which is authenticated via the $\macbdoz{w}$ check (in \ttt{\_open}),
+and then each party reconstructs $\macgv{w}$.
+
+The $\ttt{sum}$ function implements an addition gate. Note that this
+is done non-interactively-- each party just sums its local shares
+of the two values. The pre- and postcondition annotations
+on $\ttt{sum}$ express the additive homomorphism associated
+with this encryption scheme-- the sum of MACs of the input
+shares is a valid MAC for the output share on each client,
+which can be checked using the sum of keys of the input shares.
+
+In BDOZ a pre-processing phase is assumed where initial input secrets
+are shared along with their associated MACs and keys. This can be
+expressed in $\eqspre$ for input secrets $\sx{\ttt{"x"}}{1}$ and
+$\sx{\ttt{"y"}}{2}$, for example, which would subsume the following
+constraints on shares:
+{\footnotesize$$
 \begin{array}{l}
-\mx{\ttt{"xs"}}{2} \eop \sx{\ttt{"x"}}{1} \fplus \rx{\ttt{"x"}}{1} \wedge 
+\mx{\ttt{"xs"}}{2} \eop \sx{\ttt{"x"}}{1} \fminus \rx{\ttt{"x"}}{1} \wedge 
 \mx{\ttt{"xs"}}{1} \eop \rx{\ttt{"x"}}{1} \wedge \\
-\mx{\ttt{"ys"}}{1} \eop \sx{\ttt{"y"}}{2} \fplus \rx{\ttt{"y"}}{2} \wedge 
+\mx{\ttt{"ys"}}{1} \eop \sx{\ttt{"y"}}{2} \fminus \rx{\ttt{"y"}}{2} \wedge 
 \mx{\ttt{"ys"}}{2} \eop \rx{\ttt{"y"}}{2} 
 \end{array}
-$$
-
-$$
+$$}
+and the following constraint on keys and MACs for authentication
+of $\sx{\ttt{"x"}}{1}$ (and similarly for $\sx{\ttt{"y"}}{2}$):
+{\footnotesize$$
 \begin{array}{l}
@@
 \end{array}
-$$
+$$}
+Given this, a malicious secure opening of $\sx{\ttt{"x"}}{1} +
+\sx{\ttt{"y"}}{2}$ would be obtained as
+$\ttt{sum}(\ttt{"z"},\ttt{"x"},\ttt{"y"}); \ttt{open}(\ttt{"z"})$.
 
-$$
+A common approach to implementing multiplication gates in a BDOZ
+setting is to use \emph{Beaver Triples}. Recall that Beaver triples
+are values $a,b,c$ with $\fcod{a * b} = c$, unique per multiplication
+gate, that are secret shared with clients during pre-processing.
+In our encoding we assume the additional convention that each gate
+output identifier distinguishes the Beaver triple, so for
+example the share of the $a$ value for a gate $\ttt{"g1"}$
+is identified by $\ttt{"g1as"}$, etc., and $\eqspre$ would
+subsume the following constraints for the $a,b,c$ values of
+gate $\ttt{"g1"}$.
+{\footnotesize$$
 \begin{array}{l}
@@
 \end{array}
-$$
+$$}
+The definition of the $\ttt{mult}$ gate is presented in Figure
+\ref{fig-bdozmult}, where we present just client 1's side of the
+protocol $\ttt{\_mult1}$ (client 2's side is nearly symmetric but with
+a small variation-- refer to
+\cite{evans2018pragmatic,10.1007/978-3-030-68869-1_3} for more
+details).  We could call this function on input secrets
+$\sx{\ttt{"x"}}{1}$ and $\sx{\ttt{"y"}}{2}$ in gate $\ttt{"g1"}$ as
+$\ttt{mult}(\ttt{"g1"},\ttt{"x"},\ttt{"y"})$, for example, or embed
+this gate internally in a circuit.
 
-$$
-\ttt{mult}(\ttt{"g1"},\ttt{"x"},\ttt{"y"})
-$$
+As for $\ttt{\_sum}$, the postcondition of the $\ttt{\_mult1}$
+function expresses the relevant semi-homorphic encryption property of
+the resulting share, MAC, and key after gate execution-- specifically,
+it preserves the BDOZ authentication property. The precondition that
+input shares must possess this property is enforced by the two calls
+to $\ttt{sum}$ within the body of $\ttt{mult}$ and need not be made
+explicit. Finally, the postcondition of $\ttt{mult}$ expresses the
+correctness property of the multiplication gate.
 

### 2024-10-10 esop25/intro.tex
checkpoint

--- 2024-07-17 ppdp24/intro.tex
@@
 important modern, distributed applications such as privacy-preserving
-machine learning \cite{li2021privacy, knott2021crypten,
-  koch2020privacy, liu2020privacy} and Zero-Knowledge proofs in
-blockchains \cite{ishai2009zero, lu2019honeybadgermpc,
-  gao2022symmeproof, tomaz2020preserving}. The security semantics of
-MPC include both confidentiality and integrity properties incorporated
-into models such as real/ideal (aka simulator) security and universal
-composability (UC), developed primarily by the cryptography community
-\cite{evans2018pragmatic}.  Related proof methods are well-studied
-\cite{Lindell2017} but mostly manual. Somewhat independently, a
-significant body of work in programming languages has focused on
-definition and enforcement of confidentiality and integrity
-\emph{hyperproperties} \cite{10.5555/1891823.1891830} such as
-noninterference and gradual release
-\cite{4223226,sabelfeld2009declassification}. Following a tradition of
-connecting cryptographic and PL-based security models
+machine learning
+\cite{li2021privacy,knott2021crypten,koch2020privacy,liu2020privacy}
+and Zero-Knowledge proofs in blockchains
+\cite{ishai2009zero,lu2019honeybadgermpc,gao2022symmeproof,tomaz2020preserving}.
+The security semantics of MPC include both confidentiality and
+integrity properties incorporated into models such as real/ideal (aka
+simulator) security and universal composability (UC), developed
+primarily by the cryptography community \cite{evans2018pragmatic}.
+Related proof methods are well-studied \cite{Lindell2017} but mostly
+manual. Somewhat independently, a significant body of work in
+programming languages has focused on definition and enforcement of
+confidentiality and integrity \emph{hyperproperties}
+\cite{10.5555/1891823.1891830} such as noninterference and gradual
+release \cite{4223226,sabelfeld2009declassification}. Following a
+tradition of connecting cryptographic and PL-based security models
 \cite{10.1007/3-540-44929-9_1,10.1145/3571740}, recent work has also
@@
 formulate a collection of confidentiality and integrity
-hyperproperties for our model with familiar information flow
-analogs, and to show how these properties can be leveraged for
-improved proof automation.
+hyperproperties for our model with familiar information flow analogs,
+and to show how these properties can be leveraged for improved proof
+automation.
 
@@
 
+\testcompfig
+
 \paragraph{Hyperproperty formulation.} In Section \ref{section-model} we
@@
 \ref{lemma-gmwinvariant} and Theorem \ref{theorem-gmw}).
-
-\compfig
 
@@
 languages designed to perform (exact or approximate) statistical
-inference~\cite{holtzen2020scaling, carpenter2017stan, wood2014new,
-  bingham2019pyro, albarghouthi2017fairsquare, de2007problog,
-  pfeffer2009figaro, saad2021sppl}. Our setting, however, requires
+inference~\cite{holtzen2020scaling,carpenter2017stan,wood2014new,bingham2019pyro,albarghouthi2017fairsquare,de2007problog,pfeffer2009figaro,saad2021sppl}. Our setting, however, requires
 verifying properties beyond inference, including conditional

### 2024-10-10 esop25/frontmatter.tex
checkpoint

\titlerunning{Security Types for Low-Level MPC}

\title{SMT-Boosted Security Types for Low-Level MPC}

%\author{
%Christian Skalka
%\and
%Joe Near}

%\institute{
%University of Vermont, Burlington VT 05405, USA 
%\email{ceskalka@uvm.edu,jnear@uvm.edu}
%}

\author{Author names withheld for double-blind reviewing.}
\institute{}

\newcommand{\theabstract}{
  \begin{abstract}
  Secure Multi-Party Computation (MPC) is an important
  enabling technology for data privacy in modern distributed
  applications. We devevelop a new type theory to automatically
  enforce correctness, confidentiality, and integrity properties
  of protocols written in the \emph{Prelude/Overture} language
  framework. Judgements in the type theory are predicated on
  SMT verifications in a theory of finite fields. 
\end{abstract} }

### 2024-10-10 esop25/intro.tex
checkpoint

@@
 Related proof methods are well-studied \cite{Lindell2017} but mostly
-manual. Somewhat independently, a significant body of work in
-programming languages has focused on definition and enforcement of
-confidentiality and integrity \emph{hyperproperties}
-\cite{10.5555/1891823.1891830} such as noninterference and gradual
-release \cite{4223226,sabelfeld2009declassification}. Following a
-tradition of connecting cryptographic and PL-based security models
-\cite{10.1007/3-540-44929-9_1,10.1145/3571740}, recent work has also
-recognized connections between MPC security models and hyperproperties
-of, e.g., noninterference \cite{8429300}, and even leveraged these
-connections to enforce MPC security through mechanisms such as
-security types \cite{10.1145/3453483.3454074}. Major benefits of this
-connection in an MPC setting include better language abstractions for
-defining protocols and for mechanization and even automation of
-security proofs.  The goal of this paper is to develop a PL model for
-defining a variety of low-level probabilistic MPC protocols, to
-formulate a collection of confidentiality and integrity
-hyperproperties for our model with familiar information flow analogs,
-and to show how these properties can be leveraged for improved proof
-automation.
+manual. To support automated reasoning, recent work has reformulated
+MPC security semantics as \emph{hyperproperties}
+\cite{8429300,10.1145/3453483.3454074,skalka-near-ppdp24}.
+Separately, recent research in the SMT community has developed
+theories of finite fields \cite{SMFF} with clear relevance to
+verification of cryptographic schemes that rely on field
+arithmetic. The goal of this paper is to combine these two approaches
+in an automated type system for verifying correctness and security
+properties of MPC protocols.
 
@@
 
-The connection between information flow hyperproperties and MPC
-security is also complicated especially at a low level.  MPC protocols
-involve communication between a group of distributed clients called a
-\emph{federation} that collaboratively compute and publish the result
-of some known \emph{ideal functionality} $\idealf$, maintaining
-confidentiality of inputs to $\idealf$ without the use of a trusted
-third party. However, since the outputs of $\idealf$ are public, some
-information about inputs is inevitably leaked. Thus, the ideal
-functionality establishes a declassification policy
-\cite{sabelfeld2009declassification}, which is more difficult to
-enforce than pure noninterference.  And subtleties of, e.g.,
-semi-homomorphic encryption are central to both confidentiality and
-integrity properties of protocols and similarly difficult to track
-with coarse-grained security types alone.
-
-Nevertheless, as previous authors have observed
-\cite{5a51987acaa84c43bb4bf5bcc7d01683}, low-level protocol design
-patterns such as secret sharing and circuit gate structure have
-compositional properties that can be independently verified and then
-leveraged in larger proof contexts. We contribute to this line of work
-by developing an automated verification technique for subprotocols and
-show how it can be integrated as a tactic in a larger security proof.
-
 \subsection{Overview and Contributions}
 
-In summary, our work provides a complete methodology for end-to-end verification
-of MPC protocols via three components:
-%
 \begin{enumerate}
@@
 \end{enumerate}
-%
-As part of this methodology, we also develop hyperproperties that
-encode MPC security that may be of independent interest. The complete
-methodology enables the verification of real-world MPC protocols like
-GMW~\cite{goldreich2019play}.
-
-\paragraph{Language design.} In Section \ref{section-lang} we
-develop a new probabilistic programming language $\minifed$ for
-defining synchronous distributed protocols over an arbitrary
-arithmetic field. The syntax and semantics provides a succinct account
-of synchronous messaging between protocol \emph{clients}. In Section
-\ref{section-metalang} we define a metalanguage $\metaprot$ that
-dynamically generates $\minifed$ protocols. It is able to express
-important low-level abstractions, as we illustrate via implementations
-of protocols including Shamir addition (Section \ref{section-lang}),
-GMW boolean circuits (Section \ref{section-example-gmw}), and Beaver
-Triple multiplication gates with BDOZ authentication (Section
-\ref{section-example-bdoz}).
-
-\testcompfig
-
-\paragraph{Hyperproperty formulation.} In Section \ref{section-model} we
-develop our formalism for expressing the joint probability mass function of
-program variables, and give standard definitions of passive and
-malicious real/ideal security in our model. In Section
-\ref{section-hyper}, we formulate a variety of familiar information
-flow properties in our probabilistic setting, including conditional
-noninterference, gradual release, and robust declassification, and
-consider the relation between these and real/ideal security.  While it
-has been previously shown that probabilistic conditional
-noninterference is sound for passive security, we formulate new
-properties of integrity which, paired with passive security, imply
-malicious security (Theorem \ref{theorem-integrity}). We observe
-in Section \ref{section-example-bdoz} that authentication mechanisms
-such as BDOZ/SPDZ style MACs enforce a strictly weaker property
-of ``cheating detection'' (Lemma \ref{lemma-cheating}).
-
-\paragraph{Fully and partially automated verification.} In Section
-\ref{section-bruteforce} we develop a method for automatically
-computing the probability mass function (pmf) of $\minifed$ protocols
-in $\mathbb{F}_2$, that can be automatically queried to enforce
-hyperproperties of security. This method is perfectly accurate but has
-high complexity; we show this can be partially mitigated by conversion
-of protocols in $\mathbb{F}_2$ to stratified Datalog which is amenable
-to HPC acceleration. Furthermore, in Section \ref{section-example-gmw}
-we consider in detail how this automated technique can be used as a
-local automated tactic for proving security in arbitrarily large GMW
-circuits using conditional probabilistic independence as in
-\cite{li2023lilac} (Lemmas \ref{lemma-gmwtactic} and
-\ref{lemma-gmwinvariant} and Theorem \ref{theorem-gmw}).
 
@@
 
+\cnote{This sections needs updating with, including with
+  \cite{skalka-near-ppdp24,SMFF}}.
 Our main focus is on PL design and automated and semi-automated
@@
 We summarize this comparison in Figure
-\ref{fig-comp}, with the caveat that works vary in the degree of
+\ref{fig-comp-wrap}, with the caveat that works vary in the degree of
 development in each dimension.
@@
 % security models.
+
+\compwrapfig
 

### 2024-10-10 esop25/cpj.tex
checkpoint

@@
   We can then proceed by a second induction on the derivation of $\ty_0$, where there
-  are two subcases. XXX
+  are two subcases. \cnote{complete me.}
 \end{proof}
@@
 \begin{theorem}
+  \label{theorem-cpj}
   Given $\prog$ with $\views(\prog) = M \cup P$, if $\cpj{R}{\eqs}{\eqspre \wedge \toeq{\prog}}{\Gamma}$

### 2024-10-10 esop25/intro.tex
checkpoint

@@
 
-\begin{enumerate}
-\item A \textbf{low-level language} for defining MPC primitives
-  (Section \ref{section-lang}) with an associated \textbf{metalanguage}
-  to ease programming (Section \ref{section-metalang})\footnote{By metalanguage
-  we mean a multi-stage aka metaprogramming language where code is a value, as
-  in, e.g., MetaML \cite{TAHA2000211}.}. 
-\item A \textbf{fully-automated verification method} for low-level MPC
-  primitives in $\mathbb{F}_2$ (Section \ref{section-bruteforce}).
-\item A \textbf{partially-automated verification method} for
-  MPC protocols, which leverages automated proofs for
-  low-level primitives (Section \ref{section-example-gmw}).
+Our work is based on the $\metaprot$/$\minicat$ framework developed in
+previous work \cite{skalka-near-ppdp24}. In Section \ref{section-lang}
+we recall the $\minicat$ language model and key definitions. In
+Section \ref{section-model} we recall formulations of program distributions,
+and key probabilistic hyperproperties of MPC in $\minicat$. On this
+basis we make the following contributions:
+\begin{enumerate}[i.]
+\item In Section \ref{section-smt}, an interpretation of protocols in
+  $\minicat$ as SMT constraints (Theorem \ref{theorem-toeq}) and
+  methodology for verifying correctness properties.
+\item In Section \ref{section-cpj}, a confidentiality type system for
+  $\minicat$ protocols with a soundness property guaranteeing that
+  information is released only through explicit declassifications
+  (Theorem \ref{theorem-cpj}).
+\item In Section \ref{section-ipj}, an integrity type system for
+  $\minicat$ protocols with a soundness property guaranteeing
+  robustness in malicious settings (Theorem \ref{theorem-ipj}).
+\item In Section \ref{section-metalang}, a dependent Hoare type system and
+  algorithm for the $\metaprot$ language that is sound for the
+  confidentiality and integrity type systems (Theorem
+  \ref{theorem-mtj}), with independently verified, compositional $\Pi$
+  types for functions.
 \end{enumerate}
+In all cases, our type systems are boosted by verification mechanisms
+provided by satisfiability modulo finite fields \cite{SMFF}. In
+Section \ref{section-examples} we develop and discuss extended example
+applications of our type analyses to real protocols including
+the Goldreich-Micali-Wigderson (GMW), Bendlin-Damgard-Orland-Zakarias
+(BDOZ), and Yao's Garbled Circuits (YGC) protocols.
 
@@
 
-\compwrapfig
-
 As mentioned above, several high-level languages have been developed
@@
 typically assumed as a selection of library components.
+
+\compwrapfig
 

### 2024-10-10 esop25/ipj.tex
checkpoint

@@
 \begin{theorem}
+  \label{theorem-ipj}
   Given pre-processing predicate $\eqspre$ and protocol $\prog$ with

### 2024-10-10 esop25/metalang.tex
checkpoint

@@
 \section{Compositional Type Verification in $\metaprot$}
+\label{section-metalang}
 
@@
 \subsection{Dependent Hoare Type Theory}
+\label{section-metalangty}
 
@@
 \begin{theorem}[Soundness of $\metaprot$ type checking]
+  \label{theorem-mtj}
   Given preprocessing predicate $\eqspre$, program $\instr$, and verified $\tsig$, if

### 2024-10-10 esop25/model.tex
checkpoint

--- 2024-07-17 ppdp24/model.tex
@@
 \section{Security Model}
-\label{section-pmf}
 \label{section-model}
@@
 setting, Given a protocol $\prog$ that implements $\idealf$, with
-$\iov(\prog) = S \cup M \cup P \cup O$, the domain of $\idealf$
-is $\mems(S)$ and its range is $\mems(O)$.  Real/ideal security in the MPC
-setting means that, given $\store \in \mems(S)$, a secure protocol
-$\prog$ does not reveal any more information about honest secrets
-$\store_H$ to parties in $C$ beyond what is implicitly declassified by
+$\iov(\prog) = S \cup M \cup P \cup O$, the domain of $\idealf$ is
+$\mems(S)$ and its range is $\mems(O)$. Security in the MPC setting
+means that, given $\store \in \mems(S)$, a secure protocol $\prog$
+does not reveal any more information about honest secrets $\store_H$
+to parties in $C$ beyond what is implicitly declassified by
 $\idealf(\sigma)$. Security comes in \emph{passive} and
@@
 rules or not, respectively. Characterization of both real world
-protocol execution and simulation is defined
-probabilistically. Following previous work
-\cite{barthe2019probabilistic} we use probability mass functions to
-express joint dependencies between input and output variables, as a
-metric of information leakage.
+protocol execution and simulation is defined probabilistically.  In
+this work we will focus on the enforcement of hyperproperties of
+passive and malicious security developed in related work
+\cite{skalka-near-ppdp24}, that were shown to be sound for real/ideal
+security properties that are traditionally used for MPC.  As in that
+work we use probability mass functions to express joint dependencies
+between input and output variables, as a metric of information
+leakage.
 
-\subsection{Probability Mass Functions}
+\subsection{Probability Mass Functions (pmfs)}
 
-We define discrete joint probability mass functions (pmfs) in a
-standard manner but develop some notations that are useful for our
-presentation. Firstly, whereas distributions typically map
-realizations of random variables to values in $[0..1]$, for
-convenience and following \cite{barthe2019probabilistic} we use
-memories to represent realizations, so for example given a pmf $\pmf$
-over variables $\{ \sx{x}{1}, \mx{y}{2} \}$ we write $\pmf(\{
+Whereas pmfs typically map \emph{realizations} of variables in a joint
+distribution to values in $[0..1]$, for convenience and following
+\cite{barthe2019probabilistic,skalka-near-ppdp24} we use memories to
+represent realizations, so for example given a pmf $\pmf$ over
+variables $\{ \sx{x}{1}, \mx{y}{2} \}$ we write $\pmf(\{
 \elab{\secret{x}}{1} \mapsto 0, \elab{\mesg{y}}{2} \mapsto 1 \})$ to
@@
 \end{definition}
+As in previous work \cite{darais2019language,barthe2019probabilistic,skalka-near-ppdp24}, 
+probabilistic independence, aka separation, is an important concept and we adopt the
+following standard notation to express separation:
+\begin{definition}
+%  We write $\vc{\pmf}{x}{y}$ iff $\pmf(\{ x \mapsto 0\}\ |\ \{ y \mapsto 0 \}) =
+%  \pmf(\{ x \mapsto 1\}\ |\ \{ y \mapsto 1 \}) = 1$.
+  We write $\sep{\pmf}{X_1}{X_2}$ iff for all
+    $\store \in \mems(X_1 \cup X_2)$ we have
+  $\margd{\pmf}{X_1 \cup X_2}(\store) =
+  \pmf(\store_{X_1}) * \pmf(\store_{X_2})$
+\end{definition}
 
 \subsection{Basic Distribution of a Protocol}
-Now we can define the probability distribution of a program $\prog$,
-that we denote $\progtt(\prog)$. Since $\fedcat$ is deterministic the
-results of any run are determined by the input values together with
-the random tape. And since we constrain programs to not overwrite
-views, we are assured that \emph{final} memories contain both a
-complete record of all initial secrets as well as views resulting from
-communicated information.
 
-Our semantics require that random tapes contain values for all program
-values $\elab{\flip{w}}{\cid}$ sampled from a uniform distribution
-over $\mathbb{F}_p$. Input memories also contain input secret values
-and possibly other initial view elements as a result of
-pre-processing, e.g., Beaver triples for efficient multiplication,
-and/or MACed share distributions as in BDOZ/SPDZ
-\cite{evans2018pragmatic,10.1007/978-3-030-68869-1_3}. We define
-$\runs(\prog)$ as the set of final memories resulting from execution
-of $\prog$ given any initial memory, and treat all elements of
-$\runs(\prog)$ as equally likely.  This establishes the basic program
-distribution that can be marginalized and conditioned to quantify
-input/output information dependencies.
+We treat all elements of $\runs(\prog)$ as equally likely.  This
+establishes the basic program distribution that can be marginalized
+and conditioned to quantify input/output information dependencies.
 %In this
@@
   \label{definition-progd}
-  Given $\prog$ with $\secrets(\prog) = S$ and $\flips(\prog) = R$ and
-  pre-processing predicate $\preproc$ on memories, define:
-  $$
-  \begin{array}{c}
-    \runs(\prog) \defeq \\
-    \{ \store \mid \exists \store_1 \in \mems(R) . 
-    \exists \store_2 . \preproc(\store_2) \wedge
-    %(\dom(\store) = \iov(\prog) \cup R) \wedge
-    (\store_1 \uplus \store_2,\prog) \redxs (\store,\varnothing) \}
-  \end{array}
-  $$
-  By default, $\preproc(\store) \iff \dom(\store) = S$, i.e.,
-  the initial memory contains all input secrets in a uniform
-  marginal distribution. Then the \emph{basic distribution of $\prog$}, written $\progtt(\prog)$, is
-  defined such that for all $\store \in \mems(\iov(\prog) \cup R)$:
+  The \emph{basic distribution of $\prog$}, written $\progtt(\prog)$, is
+  defined such that for all $\store \in \mems(\vars(\prog))$:
   $$
@@
   $$
-
-  %In some cases, we will also be concerned with the (joint)
-  %probabilities of expression interpretation given a preceding program
-  %execution, and we write $\progtt(\prog, \be)$ to denote the program
-  %distribution $\progtt(\prog;\itv := \be)$ where $\itv$ is a
-  %special variable that is never used in programs.
+  For any $\adversary$ the basic \emph{$\adversary$ distribution of $\prog$}, written
+  $\progtt(\prog,\adversary)$, is
+  defined such that for all $\store \in \mems(\iov(\prog) \cup R)$:
+  $$
+  \progtt(\prog,\adversary)(\store) =  1 / |\botruns(\prog)| \ \text{if}\ \store \in \botruns(\prog), \text{otherwise}\ 0
+  $$
+  where $\botruns$ pads out undefined views and outputs with $\bot$:
+  $$
+  \begin{array}{c}
+    \botruns(\prog) \defeq \\
+    \{ \store\{ x_1 \mapsto \bot, \ldots, x_n \mapsto \bot \} \mid 
+    \store \in \aruns(\prog) \wedge \{ x_1,\ldots,x_n\} = (V \cup O) - \dom(\store) \}
+  \end{array}
+  $$
 \end{definition}
@@
 
-\subsection{Passive Correctness and Security}
+\subsection{Hyperproperties of Confidentiality and Integrity}
 
-In the passive setting we assume that $H$ and $C$ follow the
-rules of protocols and share messages as expected. A first
-consideration is whether a given protocol is \emph{correct}
-with respect to an ideal functionality. 
-\begin{definition}[Passive Correctness]
-  %Given $\prog$ with output variables $\out{1},\ldots,\out{n}$ and ideal
-  We say that a protocol $\prog$ is \emph{passive correct} for a functionality
-  $\idealf$ iff for all $\store \in \mems(\secrets(\prog))$
-  we have $\progtt(\prog)(\idealf(\store) \mid \store) = 1$.
-  %with $\idealf(\store) = v_1,\ldots,v_n$ we have
-  %$\progtt(\prog)(\out{1} \mapsto v_1,\ldots,\out{n} \mapsto v_n \mid \store) = 1$.
+Here we restate hyperproperties developed in previous work
+\cite{skalka-near-ppdp24}, to which the reader is referred for a more
+thorough discussion of their relation to real/ideal security.  As
+usual, using hyperproperties for security semantics eases proofs of
+correctness for our type systems.
+
+\subsubsection{Confidentiality}
+
+Since MPC protocols release some information about secrets through
+outputs of $\idealf$, they do not enjoy strict noninterference.
+Public reveals and protocol outputs are fundamentally forms of
+declassification. But the following property of probabilistic
+noninterference \emph{conditioned on output} is sound for passive
+security \cite{skalka-near-ppdp24}.
+\begin{definition}[Noninterference modulo output]
+  \label{definition-NIMO}
+  We say that a program $\prog$ with $\iov(\prog) = S \cup V \cup O$
+  satisfies \emph{noninterference modulo output}
+  iff for all $H$ and $C$ and $\store_1 \in \mems(S_C \cup O)$ and $\store_2 \in \mems(\houtputs)$
+  we have:
+  $$
+  \condd{\progtt(\prog)}{S_H}{\store_1} = \condd{\progtt(\prog)}{S_H}{\store_1 \uplus \store_2}
+ $$
+\end{definition}
+This conditional noninterference property implies that
+corrupt views give the adversary no better chance of guessing honest
+secrets than just the output and corrupt inputs do.
+
+In practice, MPC protocols typically satisfy a \emph{gradual
+release} property \cite{sabelfeld2009declassification}, where messages
+exchanged remain probabilistically separable from secrets, with only
+declassification events (reveals and outputs) releasing information
+about honest secrets.  
+\begin{definition}
+  Given $H,C$, a protocol $\prog$ with $\iov(\prog) = S \cup M \cup P \cup O$
+  satisfies \emph{gradual release} iff
+  $\sep{\progtt(\prog)}{M_C}{S_H}$.
 \end{definition}
 
-In the passive setting the simulator must construct a probabilistic
-algorithm $\SIM$, aka a \emph{simulation}, that is parameterized by
-corrupt inputs and the output of an ideal functionality, and that
-returns a reconstruction of corrupt views that is probabilistically
-indistinguishable from the corrupt views in the real world protocol
-execution.
-\begin{definition}
-  Given $\store$, and $v$,we write $ \prob(\SIM(\store,v) = \store') $
-  to denote the probability that $\SIM(\store,v)$ returns corrupt views
-  $\store'$ as a result. We write $\dist(\SIM(\store,v))$ to
-  denote the distribution of corrupt views reconstructed by the
-  simulation, where for
-  all $\store' \in \mems(V)$:
+\subsubsection{Integrity}
+
+\emph{Integrity} is an important hyperproperty in security models that
+admit malicious adversaries. Consistent with formulations in
+deterministic settings, we define protocol integrity as the
+preservation of high equivalence (of secrets and views). Intuitively,
+this property says that any adversarial strategy either ``mimics'' a
+passive strategy with some choice of inputs or causes an abort.
+\begin{definition}[Integrity]
+  \label{def-integrity}
+  We say that a protocol $\prog$ with $\iov(\prog) = S \cup V \cup O$ has
+  \emph{integrity} iff for all $H$, $C$, and $\adversary$,
+  if $\store \in \aruns(\prog)$ 
+  then there exists $\store' \in \mems(S)$ with $\store_{S_H} = \store'_{S_H} $ and:
     $$
-  \dist(\SIM(\store,v))(\store')\ \defeq\ \prob(\SIM(\store,v) = \store') 
+    \condd{\progtt(\prog,\adversary)}{X}{\store_{S_H \cup \cinputs}} =
+    \condd{\progtt(\prog)}{X}{\store'}
     $$ 
+  where $X \defeq (\houtputs \cup O_H) \cap \dom(\store)$. 
 \end{definition}
-Then we can define passive security in the real/ideal
-model as follows. 
-\begin{definition}[Passive Security]
-  Assume given a program $\prog$ that correctly implements an ideal
-  functionality $\idealf$, with $\views(\prog) = V$.  Then $\prog$
-  is \emph{passive secure in the simulator model} iff there exists
-  a simulation $\SIM$ such that for all
-  partitions of the federation into honest and corrupt sets $H$ and $C$
-  and for all $\store \in \mems(\secrets(\prog))$:
-  $$
-  \dist(\SIM(\store_{C},\idealf(\store))) = \condd{\progtt(\prog)}{\houtputs}{\store}
-  $$
-\end{definition}
-
-\subsection{Malicious Security}
-
-In the malicious model we assume that corrupt clients are in
-thrall to an adversary $\adversary$ who does not necessarily follow
-the rules of the protocol.  We model this by positing a $\arewrite$
-function which is given a corrupt memory $\store_C$ and expression
-$\be$, and returns a rewritten expression that can be interpreted to
-yield a corrupt input. We define the evaluation relation that
-incorporates the adversary in Figure \ref{fig-adversary}.
-
-\adversaryfig
-
-A key technical distinction of the malicious setting is that it
-typically incorporates ``abort''. Honest parties implement strategies
-to detect rule-breaking-- aka \emph{cheating}-- by using, e.g.,
-message authentication codes with semi-homomorphic properties as in
-BDOZ/SPDZ \cite{10.1007/978-3-030-68869-1_3}. If cheating is detected,
-the protocol is aborted. To model this, we extend $\minifed$ with an
-\ttt{assert} command and extend the range of memories with
-$\bot$. Note that the adversary is free to ignore their own
-assertions.
-\begin{definition}
-  We add assertions of the form $\elab{\assert{\phi(\be)}}{\cid}$ to the command
-  syntax of $\minifed$, where $\phi$ is a decidable predicate on
-  $\mathbb{F}_p$ and with operational semantics given in Figure
-  \ref{fig-adversary}. We also extend the range of memories $\store$
-  to $\mathbb{F}_p \cup \{ \bot \}$.
-\end{definition}
-
-It is necessary to add $\bot$ to the range of memories since
-the possibility of abort needs to be reflected in adversarial
-runs of a protocol. We can define $\aruns(\prog)$
-as the ``prefix'' memories that result from possibly-aborting
-protocols, but we also need to ``pad out'' the memories
-of partial runs with $\bot$, as we define in $\botruns(\prog)$,
-to properly reflect the contents of views and outputs even in case of abort. 
-\begin{definition}
-  \label{def-aprogd}
-  \label{def-aprogtt}
-  \label{definition-aprogd}
-  Given program $\prog$ with $\iov(\prog) = S \cup V \cup O$ and $\flips(\prog) = R$, and
-  any assumed pre-processing predicate $\preproc$ on memories, define:
-  $$
-  \begin{array}{c}
-    \aruns(\prog) \defeq \\
-    \{ \store \mid \exists \store_1 \in \mems(R) . 
-    \exists \store_2 . \preproc(\store_2) \wedge
-    %(\dom(\store) = \iov(\prog) \cup R) \wedge
-    (\store_1 \uplus \store_2,\prog) \aredxs (\store,\varnothing) \}
-  \end{array}
-  $$
-  where by default, $\preproc(\store) \iff \dom(\store) = S$, and also define the following
-  which pads out undefined views and outputs with $\bot$:
-  $$
-  \begin{array}{l}
-    \botruns(\prog) \defeq \\
-    \qquad \{ \store\{ x_1 \mapsto \bot, \ldots, x_n \mapsto \bot \} \mid \\
-    \qquad \phantom{\{} \store \in \aruns(\prog) \wedge \{ x_1,\ldots,x_n\} = (V \cup O) - \dom(\store) \}
-  \end{array}
-  $$
-  Then the \emph{$\adversary$ distribution of $\prog$}, written $\progtt(\prog,\adversary)$, is
-  defined such that for all $\store \in \mems(\iov(\prog) \cup R)$:
-  $$
-  \progtt(\prog,\adversary)(\store) =  1 / |\botruns(\prog)| \ \text{if}\ \store \in \botruns(\prog), \text{otherwise}\ 0
-  $$
-\end{definition}
-
-Given this preamble, we can define malicious simulation and malicious security
-in a standard manner \cite{evans2018pragmatic}, as follows.
-\begin{definition}[Malicious Simulation]
-  Given a protocol $\prog$ with $\iov(\prog) = S \cup V \cup O$, honest and corrupt 
-  clients $H$ and $C$, adversary $\adversary$, and honest inputs
-  $\store \in \mems(S_H)$, the \emph{malicious simulation}  $\SIM(\store)$ has three phases:
-  \begin{enumerate}
-  \item In the first phase $\SIM_1$, $\adversary$ gives the
-    simulator some $\store' \in \mems(S_C)$, and the simulator consults an
-    oracle to compute $\idealf(\store \cup \store') \in \mems(O)$.
-  \item In the second phase $\SIM_2$, the simulator is given the corrupt
-    outputs $\idealf(\store \cup \store')_C$, which are again given to
-    $\adversary$, who decides either to abort or not. If so, then the
-    simulator is given $\sigma_{\mathit{out}} \defeq \{ \out{\cid} \mapsto \bot \mid \cid \in H \}$
-    and arbitrary internal state $\Sigma$.
-    Otherwise the simulator is given $\sigma_{\mathit{out}} \defeq \idealf(\store \cup \store')_H$
-    and $\Sigma$.
-  \item In the third phase $\SIM_3$, given $\store_{\mathit{out}}$ and $\Sigma$, the simulator
-    finally outputs
-    $\store_{\mathit{out}} \cup \store_{\mathit{views}}$ for some
-    calculated $\store_{\mathit{views}} \in \mems(\houtputs)$.
-  \end{enumerate}
-\end{definition}
-
-\begin{definition}[Malicious Security]
-  We write $\dist(\SIM(\store))$ to
-  denote the distribution of honest outputs and corrupt views reconstructed by the
-  malicious simulation, where for
-  all $\store'$:
-  $$
-  \dist(\SIM(\store))(\store')\ \defeq\ \prob(\SIM(\store) = \store') 
-  $$
-  Then a protocol $\prog$ with $\iov(\prog) = S \cup V \cup O$ is \emph{malicious
-  secure} iff for all $H$, $C$, $\adversary$, and $\store \in \mems(S_H)$:
-  $$
-  \dist(\SIM(\store)) = \condd{\progtt(\prog,\adversary)}{\houtputs \cup O_H}{\store}
-  $$  
-\end{definition}
+As demonstrated in previous work \cite{skalka-near-ppdp24}, integrity
+plus noninteference modulo output implies malicious security in the
+real/ideal model, and integrity plus gradual release is a
+probabilistic form of robust declassification
+\cite{sabelfeld2009declassification}.

### 2024-10-10 esop25/lang.tex
checkpoint

@@
 \section{$\minicat$ Syntax and Operational Semantics}
+\label{section-lang}
 
@@
 implementation, where calls to OT in $\mathbb{F}_2$ are of the form
-$\OT{\elab{\be_1}{\cid}}{\be_2}{\be_3}$ -- given a \emph{choice bit}
+$\OT{\be_1}{\cid}{\be_2}{\be_3}$ -- given a \emph{choice bit}
 $\be_1$ provided by a receiver $\cid$, the sender sends either $\be_2$

### 2024-10-10 esop25/smt.tex
checkpoint

@@
 
-In previous work we have shown that the semantics of $\minicat$ in
+In previous work it was shown that the semantics of $\minicat$ in
 $\mathbb{F}_2$ can be implemented with Datalog
@@
 \begin{theorem}
+  \label{theorem-toeq}
   $\store$ is a model of $\eqspre \wedge \toeq{\prog}$ iff $\store \in \runs(\prog)$.

### 2024-10-10 esop25/cpj.tex
checkpoint

@@
 
-\cnote{Define (conditional) inputs in hyperproperties section.}
+\begin{definition}
+  Given $\prog$ with $\iov(\prog) = S \cup V \cup O$, we say that
+  $X \subseteq S \cup \flips(\prog)$ \emph{interfere with} 
+  $M \subseteq V$ iff $\progd{\prog} \not\vdash X * M$.
+\end{definition}
 
 \begin{lemma}
-  \label{lemma-eqsprogsound}
-  If $\toeq{\prog} \models \eqs$ and $\eqs \models \eqs'$, then
-  for all $\store \in \runs(\prog)$ we have $\store \models \eqs'$.
+  Given $\prog$ with $\iov(\prog) = S \cup V \cup O$ and $\flips(\prog) = R$,
+  for all $x \in S \cup R$ and $M \subseteq V$ exactly one of the following conditions holds:
+  \begin{enumerate}[\hspace{5mm}i.]
+  \item $\setit{x}$ interferes with $M$.
+  \item There exists $X \subseteq S \cup R$ such that $\setit{x} \cup X$ interferes
+    with $M$ but neither $X$ nor $\setit{x}$ alone do, in which case we say
+    that $\setit{x}$ interferes with $M$ conditionally on $X$.
+  \item Neither condition (i) nor (ii) hold.
+  \end{enumerate}
 \end{lemma}
+
+%\begin{lemma}
+%  \label{lemma-eqsprogsound}
+%  If $\toeq{\prog} \models \eqs$ and $\eqs \models \eqs'$, then
+%  for all $\store \in \runs(\prog)$ we have $\store \models \eqs'$.
+%\end{lemma}
 
@@
   \emph{$\Gamma$ is sound for $\prog$} iff for all $M' \subseteq M$
-      and $X \subseteq S \cup R$, the following
+      and $x \in S \cup R$, the following
       conditions hold:
   \begin{enumerate}[\hspace{5mm}i.]
-    \item  if $X$ are inputs to $M'$ then
-      there exists $\ty$ with $\leakj{\Gamma}{M'}{\ty}$ and $X \subseteq \ty$.
-    \item  for all  $R' \subseteq R$, if
-      $X$ are $R'$ are conditional inputs to $M'$ then $\leakj{\Gamma}{M'}{\ty}$
-      and $\leakclose{\Gamma}{\ty \cup R'}{\ty'}$ and $X \subseteq \ty'$.
+    \item  if $\setit{x}$ interferes with $M'$ then
+      there exists $\ty$ with $\leakj{\Gamma}{M'}{\ty}$ and $x \in \ty$.
+    \item  if there exists  $X \subseteq S \cup R$ such that
+      $\setit{x}$ interferes with $M'$ conditionally on $X$,
+      then $\leakj{\Gamma}{M'}{\ty}$ and
+      $\leakclose{\Gamma}{\ty \cup X}{\ty'}$ and
+      $\setit{x} \in \ty'$.
   \end{enumerate}
@@
   We can then proceed by a second induction on the derivation of $\ty_0$, where there
-  are two subcases. \cnote{complete me.}
+  are two subcases defined by the $\TirName{DepTy}$ and $\TirName{Encode}$ rules.
+  In the former case the result is immediate since all variables occuring in
+  $\toeq{\elab{\be}{\cid}}$ are elements of $\ty_0$ by definition. In the latter case,
+  we have:
+  \begin{mathpar}
+    \eqs \models \toeq{\elab{\be}{\cid}} \eop \phi' \fminus \rx{w}{\cid}
+
+    \ty_0 = \cty{\rx{w}{\cid}}{\ty_1}
+
+    R_1 = R_2;\setit{\rx{w}{\cid}}
+
+    \eqj{R_2}{\eqs}{R_2}{\ty_1}
+  \end{mathpar}
+  These facts imply that $\rx{w}{\cid}$ is used as a one-time-pad to encode
+  $\phi$ and introduces interference conditional on $\rx{w}{\cid}$
+  \cite{barthe2019probabilistic}. Observing that:
+  $$
+  \leakclose{\Gamma}{\setit{\cty{\rx{w}{\cid}}{\ty_1}, \rx{w}{\cid}}}{\ty'}
+  $$
+  The result in this subcase follows by the (second) induction hypothesis. \qed
 \end{proof}
@@
 messages reflect no dependencies on honest secrets.
-\cnote{Need to define gradual release... should probably have a section
-  where we summarize hyperproperties.}
 \begin{theorem}

### 2024-10-10 esop25/cpj.tex
checkpoint

@@
 \begin{lemma}
+  \label{lemma-interference}
   Given $\prog$ with $\iov(\prog) = S \cup V \cup O$ and $\flips(\prog) = R$,
@@
 \begin{proof}
-  Immediate by Lemma \ref{lemma-cpjsound}. \qed
+  Immediate by Lemmas \ref{lemma-cpjsound} and \ref{lemma-interference}. \qed
 \end{proof}

### 2024-10-10 code/cvc5mvn/src/main/java/plaid/ast/AssertCommand.java
Prelude AST with prettyPrint

@@ -38,4 +38,11 @@ public class AssertCommand implements  PreludeCommand{
         return List.of(e1, e2);
     }
 
+    @Override
+    public String toString() {
+        return "AssertCommand{" +
+                "e1=" + e1 +
+                ", e2=" + e2 +
+                '}';
+    }
 }

### 2024-10-10 code/cvc5mvn/src/main/java/plaid/ast/AssignCommand.java

@@ -38,4 +38,11 @@ public class AssignCommand implements PreludeCommand{
         return List.of(e1, e2);
     }
 
+    @Override
+    public String toString() {
+        return "AssignCommand{" +
+                "e1=" + e1 +
+                ", e2=" + e2 +
+                '}';
+    }
 }

### 2024-10-10 code/cvc5mvn/src/main/java/plaid/ast/AtExpr.java

@@ -38,4 +38,11 @@ public class AtExpr implements PreludeExpression{
         return List.of(e1, e2);
     }
 
+    @Override
+    public String toString() {
+        return "AtExpr{" +
+                "e1=" + e1 +
+                ", e2=" + e2 +
+                '}';
+    }
 }

### 2024-10-10 code/cvc5mvn/src/main/java/plaid/ast/CommandFunction.java

@@ -50,4 +50,12 @@ public class CommandFunction implements PreludeFunction{
         return result;
     }
 
+    @Override
+    public String toString() {
+        return "CommandFunction{" +
+                "fname=" + fname +
+                ", y=" + y +
+                ", c=" + c +
+                '}';
+    }
 }

### 2024-10-10 code/cvc5mvn/src/main/java/plaid/ast/CommandList.java

@@ -33,4 +33,10 @@ public class CommandList implements PreludeCommand{
         return new ArrayList<>(commands);
     }
 
+    @Override
+    public String toString() {
+        return "CommandList{" +
+                "commands=" + commands +
+                '}';
+    }
 }

### 2024-10-10 code/cvc5mvn/src/main/java/plaid/ast/ConcatExpr.java

@@ -39,4 +39,11 @@ public class ConcatExpr implements PreludeExpression {
         return List.of(e1, e2);
     }
 
+    @Override
+    public String toString() {
+        return "ConcatExpr{" +
+                "e1=" + e1 +
+                ", e2=" + e2 +
+                '}';
+    }
 }

### 2024-10-10 code/cvc5mvn/src/main/java/plaid/ast/ExprFunction.java

@@ -50,4 +50,12 @@ public class ExprFunction implements PreludeFunction{
         return result;
     }
 
+    @Override
+    public String toString() {
+        return "ExprFunction{" +
+                "fname=" + fname +
+                ", y=" + y +
+                ", e=" + e +
+                '}';
+    }
 }

### 2024-10-10 code/cvc5mvn/src/main/java/plaid/ast/FieldExpr.java

@@ -38,4 +38,11 @@ public class FieldExpr implements PreludeExpression{
         result.addAll(elements.values());
         return result;
     }
+
+    @Override
+    public String toString() {
+        return "FieldExpr{" +
+                "elements=" + elements +
+                '}';
+    }
 }

### 2024-10-10 code/cvc5mvn/src/main/java/plaid/ast/FieldSelectExpr.java

@@ -38,4 +38,11 @@ public class FieldSelectExpr implements PreludeExpression{
         return List.of(e, l);
     }
 
+    @Override
+    public String toString() {
+        return "FieldSelectExpr{" +
+                "e=" + e +
+                ", l=" + l +
+                '}';
+    }
 }

### 2024-10-10 code/cvc5mvn/src/main/java/plaid/ast/FunctionCallCommand.java

@@ -42,4 +42,11 @@ public class FunctionCallCommand implements PreludeCommand{
         return result;
     }
 
+    @Override
+    public String toString() {
+        return "FunctionCallCommand{" +
+                "fname=" + fname +
+                ", functionCalls=" + functionCalls +
+                '}';
+    }
 }

### 2024-10-10 code/cvc5mvn/src/main/java/plaid/ast/FunctionCallExpr.java

@@ -5,39 +5,40 @@ import java.util.Objects;
 
 public class FunctionCallExpr implements PreludeExpression{
     private final Identifier fname;
-    private final List<PreludeExpression> functionCalls;
+    private final List<PreludeExpression> parameters;
 
     public FunctionCallExpr(Identifier fname, List<PreludeExpression> functionCalls){
         this.fname = fname;
-        this.functionCalls = functionCalls;
+        this.parameters = functionCalls;
     }
 
     public Identifier getFname() {
         return fname;
     }
 
-    public List<PreludeExpression> getFunctionCalls() {
-        return functionCalls;
+    public List<PreludeExpression> getParameters() {
+        return parameters;
     }
+
 
     @Override
     public boolean equals(Object o) {
         if (this == o) return true;
         if (o == null || getClass() != o.getClass()) return false;
         FunctionCallExpr that = (FunctionCallExpr) o;
-        return Objects.equals(fname, that.fname) && Objects.equals(functionCalls, that.functionCalls);
+        return Objects.equals(fname, that.fname) && Objects.equals(parameters, that.parameters);
     }
 
     @Override
     public int hashCode() {
-        return Objects.hash(fname, functionCalls);
+        return Objects.hash(fname, parameters);
     }
 
     @Override
     public Iterable<Node> children() {
         List<Node> result = new ArrayList<>();
         result.add(fname);
-        result.addAll(functionCalls);
+        result.addAll(parameters);
         return result;
     }
 

### 2024-10-10 code/cvc5mvn/src/main/java/plaid/ast/Identifier.java

@@ -33,4 +33,10 @@ public class Identifier implements PreludeExpression {
         return List.of();
     }
 
+    @Override
+    public String toString() {
+        return "Identifier{" +
+                "name='" + name + '\'' +
+                '}';
+    }
 }

### 2024-10-10 code/cvc5mvn/src/main/java/plaid/ast/LetExpr.java

@@ -48,4 +48,12 @@ public class LetExpr implements PreludeExpression{
         return List.of(y, e1, e2);
     }
 
+    @Override
+    public String toString() {
+        return "LetExpr{" +
+                "y=" + y +
+                ", e1=" + e1 +
+                ", e2=" + e2 +
+                '}';
+    }
 }

### 2024-10-10 code/cvc5mvn/src/main/java/plaid/ast/MinusExpr.java

@@ -33,4 +33,10 @@ public class MinusExpr implements PreludeExpression {
         return List.of(e);
     }
 
+    @Override
+    public String toString() {
+        return "MinusExpr{" +
+                "e=" + e +
+                '}';
+    }
 }

### 2024-10-10 code/cvc5mvn/src/main/java/plaid/ast/Num.java

@@ -34,4 +34,11 @@ public class Num implements OvertureExpression {
         return List.of();
     }
 
+
+    @Override
+    public String toString() {
+        return "Num{" +
+                "num=" + num +
+                '}';
+    }
 }

### 2024-10-10 code/cvc5mvn/src/main/java/plaid/ast/OvertureValue.java

@@ -1,4 +1,5 @@
 package plaid.ast;
 
+@Deprecated
 public class OvertureValue {
 }

### 2024-10-10 code/cvc5mvn/src/main/java/plaid/ast/PlusExpr.java

@@ -39,4 +39,11 @@ public class PlusExpr implements PreludeExpression {
         return List.of(e1, e2);
     }
 
+    @Override
+    public String toString() {
+        return "PlusExpr{" +
+                "e1=" + e1 +
+                ", e2=" + e2 +
+                '}';
+    }
 }

### 2024-10-10 code/cvc5mvn/src/main/java/plaid/ast/Program.java

@@ -22,6 +22,23 @@ public class Program implements Node {
         return exprFunctions;
     }
 
+    public ExprFunction resolveExprFunction(PreludeExpression functionName){
+        for (ExprFunction exprFunction : exprFunctions) {
+            if(exprFunction.getFname().equals(functionName)){
+                return exprFunction;
+            }
+        }
+        return null;
+    }
+
+    public CommandFunction resolveCommandFunction(PreludeExpression functionName){
+        for (CommandFunction commandFunction : commandFunctions) {
+            if(commandFunction.getFname().equals(functionName)){
+                return commandFunction;
+            }
+        }
+        return null;
+    }
     @Override
     public boolean equals(Object o) {
         if (this == o) return true;
@@ -43,4 +60,11 @@ public class Program implements Node {
         return result;
     }
 
+    @Override
+    public String toString() {
+        return "Program{" +
+                "commandFunctions=" + commandFunctions +
+                ", exprFunctions=" + exprFunctions +
+                '}';
+    }
 }

### 2024-10-10 code/cvc5mvn/src/main/java/plaid/ast/Str.java

@@ -33,4 +33,10 @@ public class Str extends Value {
         return List.of();
     }
 
+    @Override
+    public String toString() {
+        return "Str{" +
+                "str='" + str + '\'' +
+                '}';
+    }
 }

### 2024-10-10 code/cvc5mvn/src/main/java/plaid/ast/TimesExpr.java

@@ -39,4 +39,11 @@ public class TimesExpr implements PreludeExpression{
         return List.of(e1, e2);
     }
 
+    @Override
+    public String toString() {
+        return "TimesExpr{" +
+                "e1=" + e1 +
+                ", e2=" + e2 +
+                '}';
+    }
 }

### 2024-10-10 code/cvc5mvn/src/main/java/plaid/ast/Value.java

@@ -1,5 +1,6 @@
 package plaid.ast;
 
+@Deprecated
 // value alternative to prelude expression
 abstract public class Value implements PreludeExpression {
     // what we want to do here?

### 2024-10-10 code/cvc5mvn/src/test/java/plaid/antlr/FullProgramParseTest.java

@@ -40,7 +40,7 @@ public class FullProgramParseTest {
                     m[n]@i2 := (s[n] + r[n])@i1;
                     m[n]@i1 := r[n]@i1
                 }
-                
+
                 andgmw(z, x, y) {
                    let table = andtablegmw(x,y,z,1) in
                    m[x]@1 := m[x]@2;
@@ -48,17 +48,17 @@ public class FullProgramParseTest {
                    m[z]@2 := mux4(m[x], m[y], table.row1, table.row2, table.row3, table.row4)@1;
                    m[z]@1 := r[z]@1
                 }
-                
+
                 xorgmw(z, x, y) {
                         m[z]@1 := (m[x] + m[y])@1; m[z]@2 := (m[x] + m[y])@2
                 }
-                
+
                 decodegmw(z) {
                         p["1"] := m[z]@1; p["2"] := m[z]@2;
                         out@1 := (p["1"] + p["2"])@1;
                         out@2 := (p["1"] + p["2"])@2
                 }
-                
+
                 main(){
                     encodegmw("x",2,1);
                     encodegmw("y",2,1);
@@ -67,10 +67,53 @@ public class FullProgramParseTest {
                     xorgmw("g2","g1","y");
                     decodegmw("g2")
                 }
-                
+
                 """);
 
         System.out.print(program);
+        assertNotNull(program);
+    }
+
+    @Test
+    public void chrisExample(){
+        Program program = Loader.toProgram("""
+                    cmdfunctions: 
+                    _open(x,i1,i2){
+                      m[x++"exts"]@i1 := m[x++"s"]@i2;
+                      m[x++"extm"]@i1 := m[x++"m"]@i2;
+                      assert(m[x++"extm"] = m[x++"k"] + (m["delta"] * m[x++"exts"]))@i1 ;
+                      m[x]@i1 := (m[x++"exts"] + m[x++"s"])@i2
+                    }
+                
+                    _sum(z, x, y,i1,i2) {
+                        m[z++"s"]@i2 := (m[x++"s"] + m[y++"s"])@i2;
+                        m[z++"m"]@i2 := (m[x++"m"] + m[y++"m"])@i2;
+                        m[z++"k"]@i1 := (m[x++"k"] + m[y++"k"])@i1
+                    }
+                
+                    sum(z,x,y) { _sum(z,x,y,1,2); _sum(z,x,y,2,1) }
+                
[…]

### 2024-10-11 esop25/frontmatter.tex
checkpoint

@@
   \begin{abstract}
-  Secure Multi-Party Computation (MPC) is an important
-  enabling technology for data privacy in modern distributed
-  applications. We devevelop a new type theory to automatically
-  enforce correctness, confidentiality, and integrity properties
-  of protocols written in the \emph{Prelude/Overture} language
-  framework. Judgements in the type theory are predicated on
-  SMT verifications in a theory of finite fields. 
+  Secure Multi-Party Computation (MPC) is an important enabling
+  technology for data privacy in modern distributed applications. We
+  devevelop a new type theory to automatically enforce correctness,
+  confidentiality, and integrity properties of protocols written in
+  the \emph{Prelude/Overture} language framework. Judgements in the
+  type theory are predicated on SMT verifications in a theory of
+  finite fields, which supports precise and efficient analysis. Our
+  approach is compositional, scalable, and generalizes to arbitrary
+  prime fields for data and key sizes.
 \end{abstract} }

### 2024-10-11 esop25/metalang.tex
checkpoint

@@
 \begin{lemma}
+  \label{lemma-atj-sound}
   If $\atj{\toeq{\phi}}{R}{\ty}$ then $\eqj{R}{\eqs}{\phi}{\ty}$ for any $\eqs$.
@@
 $\mv_1,\ldots,\mv_n$ chosen to instantiate the function parameters are
-chosen arbitrarily\cnote{Might need to be a little more careful about
-  this}.
+chosen ``fresh'', meaning that any strings and client identifiers
+occurring as substrings in $\mv_1,\ldots,\mv_n$ are unique and
+not used in any program source code. This ensures generality
+of verification as demonstrated in Lemma \ref{lemma-eqs-notg}.
 
@@
 \end{definition}
-In practice, the programmer only needs to specify pre-~and post-conditions,
-since the rest of the type is reconstructed by type checking.
+In practice, to define $\tsig$ the programmer only needs to specify
+pre-~and postconditions, as we illustrate in Section
+\ref{section-examples}, since the rest of the type is reconstructed by
+type checking.
+
+\subsection{Type Checking Soundness}
+
+\begin{lemma}
+  \label{lemma-fresh}
+  Given $w$ and $\cid$ with $\fresh(w,\cid)$, if $\notg{\eqs}[w/y_1,\cid/y_n]$
+  is not satisfiable, then there exists no $w',\cid'$ such that
+  $\notg{\eqs}[w'/y_1,\cid'/y_n]$ is satisifiable.
+\end{lemma}
+
+\begin{proof}
+  Let $\eqs = \notg{\eqs}[w/y_1,\cid/y_n]$ and $\eqs' =
+  \notg{\eqs}[w'/y_1,\cid'/y_n]$. If $w',\cid'$ are fresh or chosen
+  such that $\eqs'$ is an $\alpha$-renaming of $\eqs$, the result is
+  immediate.  Otherwise, it must be the case that at least two
+  distinct variables $x_1$ and $x_2$ in $\eqs$ are renamed as a single
+  variable $x$ in $\eqs'$, given freshness of $w,\cid$. Let
+  $\mathit{rename}$ be the function that maps variables in $\eqs$ to
+  variables in $\eqs'$ such that $\mathit{rename}(\eqs) = \eqs'$. Let
+  $\store'$ be a model of $\eqs'$, and define $\store$ such that:
+  $$
+  \store(x) \defeq \store'(\mathit{rename}(x)) 
+  $$
+  Then $\store$ must be a model of $\eqs$, which is a contradiction. \qed
+\end{proof}
+
+\begin{lemma}
+  \label{lemma-eqs-notg}
+  For all $\fresh(\mv_1,\ldots,\mv_n)$, $\notg{E_1}$, $\notg{E_2}$, assuming: 
+  \begin{mathpar}
+    \notg{\eqs_1}[\mv_1/y_1]\cdots[\mv_n/y_n]  \redx \eqs_1
+
+    \notg{\eqs_2}[\mv_1/y_1]\cdots[\mv_n/y_n]  \redx \eqs_2
+
+    E_1 \models E_2
+  \end{mathpar}
+  then there exists no $\mv_1',\ldots,\mv_n'$ such that:
+    \begin{mathpar}
+    \notg{\eqs_1}[\mv_1'/y_1]\cdots[\mv_n'/y_n]  \redx \eqs_1
+
+    \notg{\eqs_2}[\mv_1'/y_1]\cdots[\mv_n'/y_n]  \redx \eqs_2
+
+    E_1' \not\models E_2'
+  \end{mathpar}  
+\end{lemma}
+
+\begin{proof}
+  Given the assumptions, suppose that $E_1' \not\models E_2'$
+  is an $\alpha$-renaming of $E_1 \models E_2$. In this
+  case the consequence follows immediately. Otherwise,
+  suppose on the contrary we can witness
+  $E_1' \not\models E_2'$. By definition this means that
+  $E_1' \wedge \neg E_2'$ is satisfiable
+  but $E_1 \wedge \neg E_2$ is not, which is a contradiction
+  by Lemma \ref{lemma-eqs-notg}. \qed 
+\end{proof}
 
@@
 \end{theorem}
+
+\begin{proof}
+  (Sketch) The result follows by induction on derivation of the
+  main type judgement for $\instr$. The \TirName{Mesg} case follows
+  by Lemma \ref{lemma-atj-sound}. The \TirName{Encode} and \TirName{App}
+  cases follow by definition of the \TirName{Sig} rule and Lemma
+  \ref{lemma-eqs-notg}. 
+\end{proof}

### 2024-10-11 esop25/smt.tex
checkpoint

@@
 
-In Figure \ref{fig-eqs} we define a syntax of SMT-style constraints
-$\eqs$, where constraint terms $\phi$ are similar to expressions $\be$
-except the can ``cross party lines''. This turns out to be very useful
+We introduce the following syntax of SMT-style constraints
+for $\minicat$:
+$$
+\begin{array}{rcl}
+  \phi ::= x \mid \phi \fplus \phi \mid \phi \fminus \phi \mid \phi \ftimes \phi  \qquad
+  \eqs ::= \phi \eop \phi \mid \eqs \wedge \eqs \mid \neg\eqs
+\end{array}
+$$
+%
+%In Figure \ref{fig-eqs} we define a syntax of SMT-style constraints
+%$\eqs$, where
+Note that constraint terms $\phi$ are similar to expressions $\be$
+except that $\phi$ can ``cross party lines''. This turns out to be very useful
 to express correctness properties-- for example, in the
@@
 
-\begin{fpfig}[t]{Syntax of constraints and memory application extended to
-  SMT expressions}{fig-eqs}
-$$
-\begin{array}{rcl}
-  \phi ::= x \mid \phi \fplus \phi \mid \phi \fminus \phi \mid \phi \ftimes \phi  \qquad
-  \eqs ::= \phi \eop \phi \mid \eqs \wedge \eqs 
-\end{array}
-$$
+%\begin{fpfig}[t]{Syntax of constraints}{fig-eqs}
+%$$
+%\begin{array}{rcl}
+%  \phi ::= x \mid \phi \fplus \phi \mid \phi \fminus \phi \mid \phi \ftimes \phi  \qquad
+%  \eqs ::= \phi \eop \phi \mid \eqs \wedge \eqs \mid \neg\eqs
+%\end{array}
+%$$
+%
+%\begin{mathpar}
+%  \store(\phi_1 \fplus \phi_2) = \cod{\store(\phi_1) \fplus \store(\phi_2)}
+%  
+%  \store(\phi_1 \ftimes \phi_2) = \cod{\store(\phi_1) \ftimes \store(\phi_2)}
+%  
+%  \store(\phi_1 \fminus \phi_2) = \cod{\store(\phi_1) \fminus \store(\phi_2)}
+%\end{mathpar}
+%\end{fpfig}
 
-\begin{mathpar}
-  \store(\phi_1 \fplus \phi_2) = \cod{\store(\phi_1) \fplus \store(\phi_2)}
-
-  \store(\phi_1 \ftimes \phi_2) = \cod{\store(\phi_1) \ftimes \store(\phi_2)}
-
-  \store(\phi_1 \fminus \phi_2) = \cod{\store(\phi_1) \fminus \store(\phi_2)}
-\end{mathpar}
-\end{fpfig}
-
-This language of constraints has a direct interpretation in the Finite
+This language of constraints has an obvious direct interpretation in the Finite
 Fields theory of cvc5 \cite{SMFF}. We can leverage this to implement a
 critical \emph{entailment} property, written $\eqs_1 \models \eqs_2$.
-Our entailment relation is based on model satisfiability in a standard
-sense, except we represent models as memories $\store$. To this end
-we extend application of memories to terms $\store(\phi)$ in
-Figure \ref{fig-eqs}, and we formally define entailment as follows.
+Our entailment relation is based on satisfiability in a standard
+sense, where we represent models as memories $\store$ (mappings
+from variables to values).
 \begin{definition}
-  Given $\eqs$ of the form $\phi_1 \eop \phi_1' \wedge \cdots \wedge \phi_n \eop
-  \phi_n'$ and $\store$, we say that $\store$ is a \emph{model} of
-  $\eqs$, written $\store \models \eqs$, iff $\store(\phi_i) =
-  \store(\phi_i')$ for all $1 \le i \le n$. We write $\eqs_1 \models
+  We write $\store \models \eqs$ if $\store$ satisfies, aka is a model
+  of, $\eqs$. We write $\eqs_1 \models
   \eqs_2$ iff  $\store \models E_1$ implies $\store \models
@@
 \end{definition}
-
-Given this Definition, the following Theorem is well-known and commonly
-applied in SMT systems to implement our notion of entailment. 
+Given this Definition, the following Theorem is well-known and a fundamental
+technique in SMT to implement our (common) notion of entailment. 
 \begin{theorem}
-  $\eqs \models \phi \eop \phi'$ iff $\eqs \wedge \phi \not\eop \phi'$ is
+  $\eqs \models \eqs'$ iff $\eqs \wedge \neg\eqs'$ is
   not satisfiable.

### 2024-10-11 esop25/metalang.tex
checkpoint

@@
   but $E_1 \wedge \neg E_2$ is not, which is a contradiction
-  by Lemma \ref{lemma-eqs-notg}. \qed 
+  by Lemma \ref{lemma-fresh}. \qed 
 \end{proof}

### 2024-10-11 esop25/cpj.tex
checkpoint

@@
 The syntax of types $\ty$ and type environments $\Gamma$ are presented
-in Figure \ref{fig-cpj}. We also
+in Figure \ref{fig-cpj}\footnote{In general we use $2^{\chi}$ to denote
+the powerset of terms in any grammatical sort $\chi$.}. We also
 define type judgements for expressions and programs by direct
@@
   $X \subseteq S \cup \flips(\prog)$ \emph{interfere with} 
-  $M \subseteq V$ iff $\progd{\prog} \not\vdash X * M$.
+  $M \subseteq V$ iff $\progd(\prog) \not\vdash X * M$.
 \end{definition}

### 2024-10-11 esop25/ipj.tex
checkpoint

@@
 \begin{definition}
-  Given $H,C$ and $\prog$ with $\views(\prog) = V$ and $\secrets(\prog) = S$,
-  $\adversary$, and $\store \in \aruns(\prog)$, we say $\store$ is
-  \emph{aberrant through $V' \subseteq V$} if $\neg\exists \store' \in \runs(\prog) .
-  \store_{S_H \cup V'} = \store'_{S_H \cup V'}$, and if there exists
-  such $\adversary$ and $\store$ we say any
-  $x \in V'$ is \emph{in an aberration}.
+  Given $H,C$ and $\prog$ with $\secrets(\prog) = S$ and $x \in \views(\prog)$,
+  we say $x$ has an \emph{aberration} iff
+  $\exists \adversary, \store \in \botruns(\prog)$ such that $\store(x) \ne \bot$
+  and $\neg\exists \store' \in \runs(\prog) .
+  \store_{S_H \cup \setit{x}} = \store'_{S_H \cup \setit{x}}$.
 \end{definition}
@@
 \begin{lemma}
-  If $\prog$ does not have integrity for some $H,C$, then it has
-  aberrant $\store \in \aruns(\prog)$ for some $\adversary$. 
+  If $\prog$ with $\views(\prog) = V$ does not have integrity for some
+  $H,C$, then there exists $x \in \houtputs$ with an
+  aberration.
 \end{lemma}
@@
 \begin{lemma}
-  Given $\prog$ and $H,C$, if $x$ is in an aberration then one of
+  Given $\prog$ and $H,C$, if $x$ has an aberration then one of
   the two inductively defined conditions holds:
@@
   \item There exists $\xassign{x}{\be}{\cid} \in \prog$ with $\cid \in H$ and
-    $x' \in \vars(\be)$ such that $x'$ is in an aberration.
+    $x' \in \vars(\be)$ such that $x'$ has an aberration.
   \end{enumerate}
@@
   $\ipj{\eqs}{\prog}{\Delta}$ is valid
-  and $x\in V$ has an aberration for some $H,C$,
+  and $x$ has an aberration for some $H,C$,
   then $\cheatj{\Delta}{H,C}{\seclev}$ and $\seclev(x) = \lolab$.

### 2024-10-12 esop25/cpj.tex
checkpoint

@@
       $\leakclose{\Gamma}{\ty \cup X}{\ty'}$ and
-      $\setit{x} \in \ty'$.
+      $x \in \ty'$.
   \end{enumerate}

### 2024-10-12 esop25/ipj.tex
checkpoint

@@
 \begin{definition}
-  Given $H,C$ and $\prog$ with $\secrets(\prog) = S$ and $x \in \views(\prog)$,
-  we say $x$ has an \emph{aberration} iff
-  $\exists \adversary, \store \in \botruns(\prog)$ such that $\store(x) \ne \bot$
-  and $\neg\exists \store' \in \runs(\prog) .
-  \store_{S_H \cup \setit{x}} = \store'_{S_H \cup \setit{x}}$.
+  Given $H,C$ and $\prog$ with $\iov(\prog) = S \cup V \cup O$,
+  we say $x \in V \cup O$ has an \emph{aberration} iff
+  $\exists \adversary, \store \in \botruns(\prog)$ such that
+  $\store(x) \ne \bot$ and $\neg\exists \store' \in \runs(\prog)$
+  with  $\store_{S_H} = \store'_{S_H}$ and
+  $\store(x) = \lcod{\store',\be}{\cid}$ where
+  $\xassign{x}{\be}{\cid} \in \prog$.
 \end{definition}
@@
   \label{theorem-ipj}
-  Given pre-processing predicate $\eqspre$ and protocol $\prog$ with
-  $\views(\prog) = V$, if
+  Given  $\eqspre$ and $\prog$ with
+  $\iov(\prog) = S \cup V \cup O$, if
   $\ipj{\eqs}{\prog}{\Delta}$ is valid
   and for all $H,C$ with $\cheatj{\Delta}{H,C}{\seclev}$ 
-  we have $\seclev(x) = \hilab$ for all $x \in \houtputs$, then $\prog$
+  we have $\seclev(x) = \hilab$ for all $x \in \houtputs \cup O_H$, then $\prog$
   has integrity for all $H,C$.

### 2024-10-13 esop25/cpj.tex
checkpoint

@@
 
-\begin{definition}
-  Given $\prog$ with $\iov(\prog) = S \cup V \cup O$, we say that
-  $X \subseteq S \cup \flips(\prog)$ \emph{interfere with} 
-  $M \subseteq V$ iff $\progd(\prog) \not\vdash X * M$.
-\end{definition}
+%\begin{definition}
+%  Given $\prog$ and $X_1,X_2 \subseteq \vars{\prog}$, we say that
+%  $X_1$ \emph{interferes with} $X_2$
+%  iff $\progd(\prog) \not\vdash X_1 * X_2$.
+%\end{definition}
 
@@
   \begin{enumerate}[\hspace{5mm}i.]
-  \item $\setit{x}$ interferes with $M$.
-  \item There exists $X \subseteq S \cup R$ such that $\setit{x} \cup X$ interferes
-    with $M$ but neither $X$ nor $\setit{x}$ alone do, in which case we say
-    that $\setit{x}$ interferes with $M$ conditionally on $X$.
-  \item Neither condition (i) nor (ii) hold.
+  \item $\notsep{\progd(\prog)}{\setit{x}}{M}$
+  \item There exists $X \subseteq S \cup R$ such that
+    $\notsep{\progd(\prog)}{\setit{x} \cup X}{M}$ while
+    $\sep{\progd(\prog)}{\setit{x}}{M}$ and $\sep{\progd(\prog)}{X}{M}$,
+    in which case we say that $x$ encodes $X$ in $M$
+  \item Neither condition (i) nor (ii) hold, implying $\sep{\progd(\prog)}{\setit{x}}{M}$.
   \end{enumerate}
@@
   \begin{enumerate}[\hspace{5mm}i.]
-    \item  if $\setit{x}$ interferes with $M'$ then
+  \item  if $\notsep{\progd(\prog)}{\setit{x}}{M'}$
     there exists $\ty$ with $\leakj{\Gamma}{M'}{\ty}$ and $x \in \ty$.
     \item  if there exists  $X \subseteq S \cup R$ such that
-      $\setit{x}$ interferes with $M'$ conditionally on $X$,
+      $\setit{x}$ encodes $X$ in $M'$, 
       then $\leakj{\Gamma}{M'}{\ty}$ and
@@
   These facts imply that $\rx{w}{\cid}$ is used as a one-time-pad to encode
-  $\phi$ and introduces interference conditional on $\rx{w}{\cid}$
+  $\vars{\phi}$ in $M'$ 
   \cite{barthe2019probabilistic}. Observing that:

### 2024-10-13 esop25/ipj.tex
checkpoint

@@
 \begin{lemma}
-  If $\prog$ with $\views(\prog) = V$ does not have integrity for some
-  $H,C$, then there exists $x \in \houtputs$ with an
-  aberration.
+  Given protocol $\prog$ with
+  $\views(\prog) = V$, if 
+  $\ipj{\eqs}{\prog}{\Delta}$ is valid
+  for some $\eqs$ and $x \in \cinputs$ has an
+  aberration for some $H,C$, then $\cheatj{\Delta}{H,C}{\seclev}$
+  and $\seclev(x) = \lolab$.
 \end{lemma}
@@
 \begin{lemma}
-  Given $\prog$ and $H,C$, if $x$ has an aberration then one of
-  the two inductively defined conditions holds:
-  \begin{enumerate}[\hspace{5mm}i.]
-  \item There exists $\xassign{x}{\be}{\cid} \in \prog$ with $\cid \in C$.
-  \item There exists $\xassign{x}{\be}{\cid} \in \prog$ with $\cid \in H$ and
-    $x' \in \vars(\be)$ such that $x'$ has an aberration.
-  \end{enumerate}
+  If $\prog$ with $\iov(\prog) = S \cup V \cup O$ does not have
+  integrity for some $H,C$, then there exists $\setit{x} \cup V'
+  \subseteq \cinputs$ where $\setit{x}$ has an aberration and
+  $\notcondsep{\progd(\prog)}{S_H}{\setit{x} \cup V'}{\houtputs \cup O_H}$.
 \end{lemma}
@@
 \begin{lemma}
-  Given pre-processing predicate $\eqspre$ and protocol $\prog$ with
-  $\views(\prog) = V$, if 
-  $\ipj{\eqs}{\prog}{\Delta}$ is valid
-  and $x$ has an aberration for some $H,C$,
-  then $\cheatj{\Delta}{H,C}{\seclev}$ and $\seclev(x) = \lolab$.
+  Given $\prog$ with $\iov(\prog) = S \cup V \cup O$ and
+  $\ipj{\eqs}{\prog}{\Delta}$ valid for some $\eqs$.  For all $x \in
+  \cinputs$ with an aberration, if
+  $\notcondsep{\progd(\prog)}{S_H}{\setit{x} \cup V'}{\houtputs \cup O_H}$
+  then there exists $y \in \houtputs \cup O_H$ such that
+  $\cheatj{\Delta}{H,C}{\seclev}$ and $\seclev(y) = \lolab$.
 \end{lemma}
+
+%\begin{lemma}
+%  Given $\prog$ and $H,C$, if $x$ has an aberration then one of
+%  the two inductively defined conditions holds:
+%  \begin{enumerate}[\hspace{5mm}i.]
+%  \item There exists $\xassign{x}{\be}{\cid} \in \prog$ with $\cid \in C$.
+%  \item There exists $\xassign{x}{\be}{\cid} \in \prog$ with $\cid \in H$ and
+%    $x' \in \vars(\be)$ such that $x'$ has an aberration.
+%  \end{enumerate}
+%\end{lemma}
 

### 2024-10-13 esop25/model.tex
checkpoint

@@
 \end{definition}
-As in previous work \cite{darais2019language,barthe2019probabilistic,skalka-near-ppdp24}, 
+As in previous work \cite{darais2019language,barthe2019probabilistic,skalka-near-ppdp24,li2023lilac}, 
 probabilistic independence, aka separation, is an important concept and we adopt the
-following standard notation to express separation:
+following standard notation to express conditional and unconditional separation:
 \begin{definition}
@@
 %  \pmf(\{ x \mapsto 1\}\ |\ \{ y \mapsto 1 \}) = 1$.
-  We write $\sep{\pmf}{X_1}{X_2}$ iff for all
-    $\store \in \mems(X_1 \cup X_2)$ we have
-  $\margd{\pmf}{X_1 \cup X_2}(\store) =
-  \pmf(\store_{X_1}) * \pmf(\store_{X_2})$
+  We write $\condsep{\pmf}{X_1}{X_2}{X_3}$ iff for all
+    $\store \in \mems(X_1 \cup X_2)$ and $\store' \in \mems(X_3)$ we have
+  $\margd{\pmf}{X_1 \cup X_2}(\store|\store') =
+  \pmf(\store_{X_1}|\store') * \pmf(\store_{X_2}|\store')$. If $X_1$ is empty
+  we write $\sep{\pmf}{X_2}{X_3}$.
 \end{definition}

### 2024-10-13 esop25/cpj.tex
checkpoint

@@
     $\sep{\progd(\prog)}{\setit{x}}{M}$ and $\sep{\progd(\prog)}{X}{M}$,
-    in which case we say that $x$ encodes $X$ in $M$
+    and we say that $x$ encodes $X$ in $M$.
   \item Neither condition (i) nor (ii) hold, implying $\sep{\progd(\prog)}{\setit{x}}{M}$.

### 2024-10-14 esop25/intro.tex
add a bit in introduction

@@
 Our work is based on the $\metaprot$/$\minicat$ framework developed in
-previous work \cite{skalka-near-ppdp24}. In Section \ref{section-lang}
-we recall the $\minicat$ language model and key definitions. In
+previous work \cite{skalka-near-ppdp24}. The prior work developed the 
+language and semantics, plus two mechanisms for verifying security: an
+automated approach for protocols in $\mathbb{F}_2$ (the binary field),
+based on enumerating all possible executions of the protocol, that scales 
+only to very small protocols; and a manual approach based on a program logic.
+%
+In Section \ref{section-lang} we recall the $\minicat$ language model and key definitions. In
 Section \ref{section-model} we recall formulations of program distributions,
-and key probabilistic hyperproperties of MPC in $\minicat$. On this
-basis we make the following contributions:
+and key probabilistic hyperproperties of MPC in $\minicat$.
+
+Together, our contributions \emph{automate} the verification of security
+properties in $\metaprot$/$\minicat$. In this work, we build on~\cite{skalka-near-ppdp24}
+by developing automatic verification approaches that work for arbitrary
+finite fields and scale to large protocols. To accomplish this, we develop new
+type systems for compositional verification, and use SMT to automate the checks
+that were accomplished by enumeration in prior work. Specifically, we make the
+following contributions:
+%
 \begin{enumerate}[i.]

### 2024-10-14 esop25/lang.tex
checkpoint

@@
 \end{definition}
-An example of a $\minicat$ protocol, for additive secret sharing, is
-discussed in Section \ref{section-smt-toeq}.  We disallow overwriting
-of variables, since assignment is intended to represent distinct
-message sends.
+In all cases we disallow overwriting of variables, since assignment is
+intended to represent distinct message sends. An example of a
+$\minicat$ protocol, for additive secret sharing, is discussed in
+Section \ref{section-smt-toeq}.
 
@@
 reduction relation $\redx$ to evaluate commands. Reduction is a mapping
-fromon \emph{configurations} $(\store,\prog)$ to final stores
+from \emph{configurations} $(\store,\prog)$ to final stores
 where all three command forms- message send, broadcast, and
@@
 to OT, we instead assume that OT is abstract with respect to its
-implementation, where calls to OT in $\mathbb{F}_2$ are of the form
-$\OT{\be_1}{\cid}{\be_2}{\be_3}$ -- given a \emph{choice bit}
-$\be_1$ provided by a receiver $\cid$, the sender sends either $\be_2$
-or $\be_3$.  Critically, the sender learns nothing about $\be_1$ and
-the receiver learns nothing about the unselected value. We return to
+implementation, and any use is of the form
+$\mx{w}{\cid_1} := \elab{\OT{\be_1}{\cid_1}{\be_2}{\be_3}}{\cid_2}$ --
+given a \emph{choice bit} $\be_1$ provided by a receiver $\cid_1$, the
+sender $\cid_2$ sends either $\be_2$
+or $\be_3$.  Critically, $\cid_2$ learns nothing about $\be_1$ and
+$\cid_1$ learns nothing about the unselected value. We return to
 the formal interpretation of OT in Section \ref{section-smt}.

### 2024-10-14 esop25/model.tex
checkpoint

@@
 
-Whereas pmfs typically map \emph{realizations} of variables in a joint
-distribution to values in $[0..1]$, for convenience and following
+Pmfs map \emph{realizations} of variables in a joint
+distribution to values in $[0..1]$. For convenience and following
 \cite{barthe2019probabilistic,skalka-near-ppdp24} we use memories to
@@
     $\store \in \mems(X_1 \cup X_2)$ and $\store' \in \mems(X_3)$ we have
-  $\margd{\pmf}{X_1 \cup X_2}(\store|\store') =
+  $\pmf(\store|\store') =
   \pmf(\store_{X_1}|\store') * \pmf(\store_{X_2}|\store')$. If $X_1$ is empty
@@
 from corrupt clients. We call the former \emph{corrupt views}, and
-the latter \emph{honest views}. Generally we let $V$ range over sets
-of views.
+the latter \emph{honest views}. We let $V$ range over sets
+of views, i.e., subsets of $M \cup P$.
 \begin{definition}[Corrupt and Honest Views]
@@
 \cite{skalka-near-ppdp24}, to which the reader is referred for a more
-thorough discussion of their relation to real/ideal security.  As
-usual, using hyperproperties for security semantics eases proofs of
+thorough discussion of their relation to real/ideal security. Using
+hyperproperties for security semantics eases proofs of
 correctness for our type systems.
@@
 \end{definition}
-As demonstrated in previous work \cite{skalka-near-ppdp24}, integrity
-plus noninteference modulo output implies malicious security in the
-real/ideal model, and integrity plus gradual release is a
-probabilistic form of robust declassification
+Integrity plus noninteference modulo output implies malicious security
+in the real/ideal model \cite{skalka-near-ppdp24}, and integrity plus
+gradual release is a probabilistic form of robust declassification
 \cite{sabelfeld2009declassification}.

### 2024-10-14 esop25/intro.tex
checkpoint

@@
 
-Secure Multi-Party Computation (MPC) protocols support data privacy in
-important modern, distributed applications such as privacy-preserving
-machine learning
+Data privacy is a critical concern in distributed applications
+including privacy-preserving machine learning
 \cite{li2021privacy,knott2021crypten,koch2020privacy,liu2020privacy}
-and Zero-Knowledge proofs in blockchains
+blockchains
 \cite{ishai2009zero,lu2019honeybadgermpc,gao2022symmeproof,tomaz2020preserving}.
-The security semantics of MPC include both confidentiality and
-integrity properties incorporated into models such as real/ideal (aka
-simulator) security and universal composability (UC), developed
-primarily by the cryptography community \cite{evans2018pragmatic}.
-Related proof methods are well-studied \cite{Lindell2017} but mostly
-manual. To support automated reasoning, recent work has reformulated
-MPC security semantics as \emph{hyperproperties}
-\cite{8429300,10.1145/3453483.3454074,skalka-near-ppdp24}.
-Separately, recent research in the SMT community has developed
-theories of finite fields \cite{SMFF} with clear relevance to
-verification of cryptographic schemes that rely on field
-arithmetic. The goal of this paper is to combine these two approaches
-in an automated type system for verifying correctness and security
-properties of MPC protocols.
+Secure multiparty computation (MPC) is a key software-based enabling
+technology for these applications. MPC include both confidentiality
+and integrity properties, formulated as real/ideal (aka simulator)
+security and universal composability (UC) \cite{evans2018pragmatic},
+with well-studied manual proof methods \cite{Lindell2017}.  MPC
+security semantics has also been reformulated as
+\emph{hyperproperties}
+\cite{8429300,10.1145/3453483.3454074,skalka-near-ppdp24} to support
+automated language-based proof methods. And recent research in the SMT
+community has developed theories of finite fields \cite{SMFF} with
+clear relevance to verification of cryptographic schemes that rely on
+field arithmetic. The goal of this paper is to combine these two
+approaches in an automated type system for verifying correctness and
+security properties of MPC protocols.
 
-The distinction between high- and low-level languages for MPC is
-important. High-level languages such as Wysteria
-\cite{rastogi2014wysteria} and Viaduct \cite{10.1145/3453483.3454074}
-are designed to provide effective programming of full
-applications. These language designs incorporate sophisticated
-verified compilation techniques such as orchestration
-\cite{viaduct-UC} to guarantee high-level security properties, and
-they rely on \emph{libraries} of low-level MPC protocols, such as
-binary and arithmetic circuits. These low-level protocols encapsulate
-abstractions such as secret sharing and semi-homomorphic encryption,
-and must be verified by hand. So, low-level MPC
-programming and protocol verification remains a distinct challenge and both
-critical to the general challenge of PL design for MPC and
-complementary to high-level language design.
+Our focus is on low-level MPC protocols. Previous high-level
+MPC-enabled languages such as Wysteria \cite{rastogi2014wysteria} and
+Viaduct \cite{10.1145/3453483.3454074} are designed to provide
+effective programming of full applications, and incorporate
+sophisticated compilation techniques such as orchestration
+\cite{viaduct-UC} to guarantee high-level security properties. But
+these approaches rely on \emph{libraries} of low-level MPC protocols,
+such as binary and arithmetic circuits. These low-level protocols are
+probabilistic, encapsulate abstractions such as secret sharing and
+semi-homomorphic encryption, and are complementary to high-level
+language design.
 

### 2024-10-14 esop25/intro.tex
checkpoint

@@
 
-\cnote{This sections needs updating with, including with
-  \cite{skalka-near-ppdp24,SMFF}}.
-Our main focus is on PL design and automated and semi-automated
-reasoning about security properties of low-level MPC protocols. Prior
-work has considered \textbf{automated} verification of
-\textbf{high-level} protocols and \textbf{manual} verification of
-\textbf{low-level} protocols---but none offers the combination of
-automation and low-level support we consider.
-%
-We summarize this comparison in Figure
-\ref{fig-comp-wrap}, with the caveat that works vary in the degree of
-development in each dimension.
+The goal of the $\minicat/\metaprot$ framework is to automate
+correctness and security verification of low-level protocols.  Other
+prior work has considered automated verification of high-level
+protocols and manual verification of low-level protocols.  We
+summarize this comparison in Figure \ref{fig-comp-wrap}.
 
@@
 
-As mentioned above, several high-level languages have been developed
-for writing MPC applications, and frequently exploit the connection
-between hyperproperties and MPC security. Previous work on analysis
-for the SecreC language
-\cite{almeida2018enforcing,10.1145/2637113.2637119} is concerned with
-properties of complex MPC circuits, in particular a user-friendly
-specification and automated enforcement of declassification bounds in
-programs that use MPC in subprograms. The Wys$^\star$ language
-\cite{wysstar}, based on Wysteria \cite{rastogi2014wysteria}, has
-similar goals and includes a trace-based semantics for reasoning about
-the interactions of MPC protocols. Their compiler also guarantees that
-underlying multi-threaded protocols enforce the single-threaded source
-language semantics. These two lines of work were focused on passive
-security. The Viaduct language
-\cite{10.1145/3453483.3454074} has a well-developed
-information flow type system that automatically enforces both
-confidentiality and integrity through hyperproperties such as robust
-declassification, in addition to rigorous compilation guarantees
-through orchestration \cite{viaduct-UC}. However, these high level
-languages lack probabilistic features and other abstractions of
-low-level protocols, the implementation and security of which are
-typically assumed as a selection of library components.
+The most closely related work is \cite{skalka-near-ppdp24}, where
+foundations for $\minicat/\metaprot$ are introduced including a
+verification algorithm. However, this algorithm only works for
+$\mathbb{F}_2$-- e.g., binary circuits-- and does not scale due to
+exponential complexity, necessitating semi-automated proof
+techniqued. In contrast, our type systems scale to arbitrary prime
+fields and whole program analysis for larger circuits (protocols). So
+our work is a significant advancement of $\minicat/\metaprot$.  Out
+work is also closely related to satisfiability modulo finite fields
+\cite{SMFF} but only in the sense that we apply their method as we
+discuss in Section \ref{section-smt} and subsequently.
+
+Other low-level languages with probabilistic features for privacy
+preserving protocols have been proposed. The
+$\lambda_{\mathrm{obliv}}$ language \cite{darais2019language} uses a
+type system to automatically enforce so-called probabilistic trace
+obliviousness.  But similar to previous work on oblivious data
+structures \cite{10.1145/3498713}, obliviousness is related to pure
+noninterference, not the relaxed form related to passive MPC
+security. The Haskell-based security type system in \cite{6266151}
+enforces a version of noninterference that is sound for passive
+security, but does not consider malicious security. And properties of
+real/ideal passive and malicious security for a probabilistic language
+have been (manually) formulated in EasyCrypt \cite{8429300}.
 
@@
 
-Various related low-level languages with probabilistic features have
-also been developed. The $\lambda_{\mathrm{obliv}}$ language
-\cite{darais2019language} uses a type system to automatically
-enforce so-called probabilistic trace obliviousness.  But similar to
-previous work on oblivious data structures \cite{10.1145/3498713},
-obliviousness is related to pure noninterference, not the relaxed form
-related to passive MPC security. The Haskell-based security type
-system in \cite{6266151} enforces a version of noninterference that is
-sound for passive security, but does not verify the correctness of
-declassifications and does not consider malicious security. And
-properties of real/ideal passive and malicious security for a
-probabilistic language have been formulated in EasyCrypt
-\cite{8429300}-- though their proof methods, while mechanized, are
-fully manual, and their formulation of malicious security is not as
-clearly related to robust declassification as is the one we present in
-Section \ref{section-hyper}.
+Several high-level languages have been developed for writing MPC
+applications. Previous work on analysis for the
+SecreC language \cite{almeida2018enforcing,10.1145/2637113.2637119} is
+concerned with properties of complex MPC circuits, in particular a
+user-friendly specification and automated enforcement of
+declassification bounds in programs that use MPC in subprograms. The
+Wys$^\star$ language \cite{wysstar}, based on Wysteria
+\cite{rastogi2014wysteria}, has similar goals and includes a
+trace-based semantics for reasoning about the interactions of MPC
+protocols. Their compiler also guarantees that underlying
+multi-threaded protocols enforce the single-threaded source language
+semantics. These two lines of work were focused on passive
+security. The Viaduct language \cite{10.1145/3453483.3454074} has a
+well-developed information flow type system that automatically
+enforces both confidentiality and integrity through hyperproperties
+such as robust declassification, in addition to rigorous compilation
+guarantees through orchestration \cite{viaduct-UC}. However, these
+high level languages lack probabilistic features and other
+abstractions of low-level protocols, the implementation and security
+of which are typically assumed as a selection of library components.
 
@@
 secrecy of one-time-pads and indistinguishability in secret sharing,
-that are critical to MPC. However, their methods are manual, and
-don't include conditional independence (separation). This
-latter issue has been addressed in Lilac \cite{li2023lilac}. The
-application of Lilac-style reasoning to MPC protocols has not
-previously been explored, as we do in Section
-\ref{section-example-gmw}.
-
-Our work also shares many ideas with probabilistic programming
-languages designed to perform (exact or approximate) statistical
-inference~\cite{holtzen2020scaling,carpenter2017stan,wood2014new,bingham2019pyro,albarghouthi2017fairsquare,de2007problog,pfeffer2009figaro,saad2021sppl}. Our setting, however, requires
-verifying properties beyond inference, including conditional
-statistical independence. Recent work by Li et al.~\cite{li2023lilac} proposes a
-manual approach for proving such properties, but does not provide
-automation.
+that are critical to MPC. Lilac \cite{li2023lilac} extends this line
+of work with formalisms for reasoning about conditional independence,
+which is critical for MPC \cite{skalka-near-ppdp24}.

### 2024-10-14 esop25/intro.tex
checkpoint

@@
 In all cases, our type systems are boosted by verification mechanisms
-provided by satisfiability modulo finite fields \cite{SMFF}. In
+provided by Satisfiability Modulo Finite Fields \cite{SMFF}. In
 Section \ref{section-examples} we develop and discuss extended example
@@
 our work is a significant advancement of $\minicat/\metaprot$.  Out
-work is also closely related to satisfiability modulo finite fields
+work is also closely related to Satisfiability Modulo Finite Fields
 \cite{SMFF} but only in the sense that we apply their method as we

### 2024-10-14 esop25/smt.tex
checkpoint

@@
 fields by using a more general form of SMT constraint programming,
-specifically Satisfiability Modulo Finite Fields \cite{SMFF}. As we
+Satisfiability Modulo Finite Fields \cite{SMFF}. As we
 will show in subsequent sections, this interpretation supports
@@
 Note that constraint terms $\phi$ are similar to expressions $\be$
-except that $\phi$ can ``cross party lines''. This turns out to be very useful
+except that $\phi$ can ``cross party lines''. This is needed
 to express correctness properties-- for example, in the
@@
 \begin{definition}
-  We write $\store \models \eqs$ if $\store$ satisfies, aka is a model
+  We write $\store \models \eqs$ iff $\store$ satisfies, aka is a model
   of, $\eqs$. We write $\eqs_1 \models
@@
 \begin{theorem}
-  $\eqs \models \eqs'$ iff $\eqs \wedge \neg\eqs'$ is
+  $\eqs_1 \models \eqs_2$ iff $\eqs_1 \wedge \neg\eqs_2$ is
   not satisfiable.

### 2024-10-14 esop25/intro.tex
related work table

@@
 Secure multiparty computation (MPC) is a key software-based enabling
-technology for these applications. MPC include both confidentiality
+technology for these applications. MPC protocols provide both confidentiality
 and integrity properties, formulated as real/ideal (aka simulator)

### 2024-10-14 esop25/smt.tex
checkpoint

@@
 
-This language of constraints has an obvious direct interpretation in the Finite
-Fields theory of cvc5 \cite{SMFF}. We can leverage this to implement a
-critical \emph{entailment} property, written $\eqs_1 \models \eqs_2$.
-Our entailment relation is based on satisfiability in a standard
-sense, where we represent models as memories $\store$ (mappings
-from variables to values).
+This language of constraints has an obvious direct interpretation in
+Satisfiability Modulo Finite Fields \cite{SMFF}. We can leverage this
+to implement a critical \emph{entailment} property, written $\eqs_1
+\models \eqs_2$.  Our entailment relation is based on satisfiability
+in a standard sense, where we represent models as memories $\store$
+(mappings from variables to values).
 \begin{definition}

### 2024-10-14 esop25/cpj.tex
checkpoint

@@
 $$
-we want the type of $\mx{w}{1}$ to be an ``encrypted'' type
-that is by itself independent of $\sx{w}{2}$. However, we also
-want to track the dependence of $\mx{w}{2}$ and $\rx{w}{2}$.
-In short, we assign $\mx{w}{1}$ a ``ciphertext type'', aka
-a \emph{ciphertype}:
+we want the type of $\mx{w}{1}$ to be an ``encrypted'' type that is
+by itself independent of $\sx{w}{2}$. However, we also want to track
+the \emph{joint} dependence of $\mx{w}{2}$ on $\rx{w}{2}$ and
+$\sx{w}{2}$, in case information about $\rx{w}{2}$ is subsequently
+leaked.  In short, we assign $\mx{w}{1}$ a ``ciphertext type'', aka a
+\emph{ciphertype}:
 $$
@@
 indicating syntactically that $\mx{w}{1}$ is an encryption of a
-value dependent on $\sx{w}{2}$ with $\rx{w}{2}$ and as long as client
-1 does not subsequently receive messages that are dependent on
-$\rx{w}{2}$, the value $\sx{w}{2}$ remains hidden.
+value dependent on $\sx{w}{2}$ with $\rx{w}{2}$.
 
@@
 $$
-where for all $\be_1,\be_2,\be_3$:
+where for all $\be_1,\be_2,\be_3$ (with $\neg$ denoting negation in $\mathbb{F}_2$):
 $$
@@
 \end{mathpar}
-This allows us to assign the same ciphertype to $\mx{w}{1}$ as
-above-- that is, $\cty{\rx{w}{2}}{\setit{\sx{w}{2}}}$. Leveraging the
-equivalence this way has been previously well-studied
-\cite{barthe2019probabilistic}.  Since this is a one-time-pad scheme,
-it is also important that we ensure that samples are used at most once
-for encryption. As in previous work \cite{darais2019language} we can
-use a form of type linearity to this end.
+This allows us to assign the same ciphertype to $\mx{w}{1}$ as above--
+that is, $\cty{\rx{w}{2}}{\setit{\sx{w}{2}}}$.  Since this is a
+one-time-pad scheme, it is also important that we ensure that samples
+are used at most once for encryption. As in previous work 
+\cite{darais2019language} we use type linearity to enforce this.
 
@@
 The syntax of types $\ty$ and type environments $\Gamma$ are presented
-in Figure \ref{fig-cpj}\footnote{In general we use $2^{\chi}$ to denote
-the powerset of terms in any grammatical sort $\chi$.}. We also
+in Figure \ref{fig-cpj}\footnote{In general we use $2^{\chi}$ to
+denote the powerset of terms in any grammatical sort $\chi$.}. We also
 define type judgements for expressions and programs by direct
 interpretation as terms $\phi$ and constraints $\eqs$ respectively,
-via the encoding $\toeq{\cdot}$. The \TirName{Encode} rule captures
-the use ``fresh'' samples for encryption discussed above. Derivations
-enforce linearity by requiring that variables used for encryption
-are added to judgements, and guaranteed to not be used elsewhere
-via a form of union that ensures disjointness of combined sets:
+via the encoding $\toeq{\cdot}$. The \TirName{Encode} rule enforces
+linearity by requiring that variables used for encryption are added to
+judgements, and guaranteed to not be used elsewhere via a form of
+union that ensures disjointness of combined sets:
 \begin{definition}
@@
 \end{definition}
-The \TirName{DepTy} rule handles all other syntactic forms and just
-records the dependencies on variables in the term.  The \TirName{Send}
-rule records the types of messages, reveals, and outputs in
-enviroments, and the \TirName{Seq} rule combines the types of
-sequenced programs.  Given this we define validity of program type
-judgements as follows. 
+The \TirName{DepTy} rule handles all cases and just records the
+dependencies on variables in the term, a conservative approximation of
+dependencies.  The \TirName{Send} rule records the types of messages,
+reveals, and outputs in enviroments, and the \TirName{Seq} rule
+combines the types of sequenced programs.  Given this we define
+validity of program type judgements as follows.
 \begin{definition}
@@
   These facts imply that $\rx{w}{\cid}$ is used as a one-time-pad to encode
-  $\vars{\phi}$ in $M'$ 
+  $\vars(\phi)$ in $M'$ 
   \cite{barthe2019probabilistic}. Observing that:

### 2024-10-14 esop25/appendix.tex
checkpoint

--- 2024-10-14 esop25/cpj.tex
@@
-\section{Confidentiality Types}
-\label{section-cpj}
+\section*{Appendix}
 
-It is well-known that adding or subtracting a sample from a uniform
-distribution in a finite field yields a value in a uniform
-distribution, meaning that samples can be used as one-time-pads with
-perfect secrecy \cite{barthe2019probabilistic,darais2019language}.
-In our type system for confidentiality, we aim to approximate
-distributions by tracking which program variables an expression
-may depend on. But we also want to capture this fundamental mechanism
-of encryption. As we will show, SMT can be used in this and
-other contexts to support confidentiality properties of protocols.
-
-For example, given some string $y$ and a message send such as:
-$$
-\xassign{\mx{w}{1}}{\secret{w} \ftimes \flip{w}}{2}
-$$
-we would assign the following type to denote that $\mx{w}{1}$ is dependent on
-both $\sx{w}{2}$ and $\rx{w}{2}$:
-$$
-\mx{w}{1} : \setit{\sx{w}{2}, \rx{w}{2}}
-$$
-But in the case of 2-party reconstructive sharing:
-$$
-\xassign{\mx{w}{1}}{\secret{w} \fminus \flip{w}}{2}
-$$
-we want the type of $\mx{w}{1}$ to be an ``encrypted'' type that is
-by itself independent of $\sx{w}{2}$. However, we also want to track
-the \emph{joint} dependence of $\mx{w}{2}$ on $\rx{w}{2}$ and
-$\sx{w}{2}$, in case information about $\rx{w}{2}$ is subsequently
-leaked.  In short, we assign $\mx{w}{1}$ a ``ciphertext type'', aka a
-\emph{ciphertype}:
-$$
-\mx{w}{1} : \cty{\rx{w}{2}}{\setit{\sx{w}{2}}}
-$$
-indicating syntactically that $\mx{w}{1}$ is an encryption of a
-value dependent on $\sx{w}{2}$ with $\rx{w}{2}$.
-
-Of course, there are other methods for encrypting values in MPC
-protocols-- but we can observe that many are algebraically equivalent
-to this fundamental schema. For example, in Yao's Garbled Circuits
-(YGC) in $\mathbb{F}_{2}$, the ``garbler'' represents encrypted wire
-values as a random sample (for 1) or its negation (for 0), and shares
-its secret input encoding with the ``evaluator'' by using
-it as a selection bit. That is, assuming that client 2 is the garbler and
-client 1 is the evaluator, we can define:
-$$
-\xassign{\mx{w}{1}}{\mux{\secret{w}}{\neg\flip{w}}{\flip{w}}}{2}
-$$
-where for all $\be_1,\be_2,\be_3$ (with $\neg$ denoting negation in $\mathbb{F}_2$):
-$$
-\mux{\be_1}{\be_2}{\be_3} \defeq (\neg\be_1 \ftimes \be_2) \fplus (\be_1 \ftimes \be_3)
-$$
-and letting this protocol be $\prog$ the following is valid:
-\begin{mathpar}
-%\toeq{\xassign{\mx{w}{1}}{\OT{\secret{w}}{1}{\neg\flip{w}}{\flip{w}}}{2}}
-   \toeq{\prog} \models \mx{w}{1} \eop \neg\sx{w}{2} \fplus \rx{w}{2}
-\end{mathpar}
-This allows us to assign the same ciphertype to $\mx{w}{1}$ as above--
-that is, $\cty{\rx{w}{2}}{\setit{\sx{w}{2}}}$.  Since this is a
-one-time-pad scheme, it is also important that we ensure that samples
-are used at most once for encryption. As in previous work 
-\cite{darais2019language} we use type linearity to enforce this.
-
-\cpjfig
-
-The syntax of types $\ty$ and type environments $\Gamma$ are presented
-in Figure \ref{fig-cpj}\footnote{In general we use $2^{\chi}$ to
-denote the powerset of terms in any grammatical sort $\chi$.}. We also
-define type judgements for expressions and programs by direct
-interpretation as terms $\phi$ and constraints $\eqs$ respectively,
-via the encoding $\toeq{\cdot}$. The \TirName{Encode} rule enforces
-linearity by requiring that variables used for encryption are added to
-judgements, and guaranteed to not be used elsewhere via a form of
-union that ensures disjointness of combined sets:
-\begin{definition}
-  $R_1;R_2 = R_1 \cup R_2$ if $R_1 \cap R_2 = \varnothing$ and is otherwise
-  undefined.
-\end{definition}
-The \TirName{DepTy} rule handles all cases and just records the
-dependencies on variables in the term, a conservative approximation of
-dependencies.  The \TirName{Send} rule records the types of messages,
-reveals, and outputs in enviroments, and the \TirName{Seq} rule
-combines the types of sequenced programs.  Given this we define
-validity of program type judgements as follows.
-\begin{definition}
-  Given preprocessing predicate $\eqspre$ and protocol $\prog$ we say
-  $\cpj{R}{\eqs}{\eqspre \wedge \toeq{\prog}}{\Gamma}$ is \emph{valid} iff it is derivable and
-  $\eqspre \wedge \toeq{\prog} \models \eqs$.
-\end{definition}
-It is important to note that this presentation of derivations is
-logical, not algorithmic-- in particular it is not syntax-directed
-due to the \TirName{Encode} rule.
-
-\subsection{Detecting Leakage Through Dependencies}
-
-Given a valid typing, we can leverage structural type information to
-conservatively approximate the information accessible by the
-adversary who has control over all corrupt clients in $C$ given
-a partitioning $H,C$. In particular, if the adversary has access to an
-encypted value and another value that is dependent on its one-time-pad
-key-- i.e., access to values with types of the forms:
-$$
-\setit{\cty{\rx{w}{\cid}}{\ty}} \qquad \setit{\rx{w}{\cid}, \ldots} 
-$$
-then we can conservatively assume that the adversary can decrypt the
-value of type $\ty$ and thus have access to those values.
-
-In Figure \ref{fig-leakj} we define derivation rules for the
-``leakage'' judgement $\leakj{\Gamma}{C}{\ty}$ which intuitively says
-that given a protocol with type environment $\Gamma$ and corrupt
-clients $C$, the adversary may be able to observe dependencies on
-values of type $\ty$ in their adversarial messages.  The real aim of
-this relation is to determine if
-$\leakj{\Gamma}{C}{\setit{\sx{w}{\cid}}}$ with $\cid \in H$-- that is,
-if the adversary may observe dependencies on honest secrets.
-
-\subsubsection{Type Soundness}
+\subsubsection{Confidentiality Type Soundness}
 
@@
 
-Our main type soundness result combines typing and leakage judgements,
-and guarantees gradual release of protocols-- that is, that adversarial
-messages reflect no dependencies on honest secrets.
-\begin{theorem}
-  \label{theorem-cpj}
-  Given $\prog$ with $\views(\prog) = M \cup P$, if $\cpj{R}{\eqs}{\eqspre \wedge \toeq{\prog}}{\Gamma}$
-  is valid and for all $H,C$ we have $\leakj{\Gamma}{M_C}{T}$ with $T$ closed
-  and $S_H \cap T = \varnothing$, then $\prog$ satisfies gradual release.
-\end{theorem}
-\begin{proof}
+\begin{proof}[Theorem \ref{theorem-cpj}]
   Immediate by Lemmas \ref{lemma-cpjsound} and \ref{lemma-interference}. \qed
@@
 
-\leakjfig
+\subsubsection{Integrity Type Soundness}
 
-Of course, it is important to recall from \cite{skalka-near-ppdp24}
-that gradual release does \emph{not} imply passive security, since it
-only concerns adversarial messages, but not public reveals of outputs
-where dependencies on input secrects can occur even in secure
-protocols.
+\begin{definition}
+  Given $H,C$ and $\prog$ with $\iov(\prog) = S \cup V \cup O$,
+  we say $x \in V \cup O$ has an \emph{aberration} iff
+  $\exists \adversary, \store \in \botruns(\prog)$ such that
+  $\store(x) \ne \bot$ and $\neg\exists \store' \in \runs(\prog)$
+  with  $\store_{S_H} = \store'_{S_H}$ and
+  $\store(x) = \lcod{\store',\be}{\cid}$ where
+  $\xassign{x}{\be}{\cid} \in \prog$.
+\end{definition}
 
-\begin{comment}
-\subsection{Examples}
+\begin{lemma}
+  Given protocol $\prog$ with
+  $\views(\prog) = V$, if 
+  $\ipj{\eqs}{\prog}{\Delta}$ is valid
+  for some $\eqs$ and $x \in \cinputs$ has an
+  aberration for some $H,C$, then $\cheatj{\Delta}{H,C}{\seclev}$
+  and $\seclev(x) = \lolab$.
+\end{lemma}
 
-\begin{verbatimtab}
-m[s1]@2 := (s[1] - r[local] - r[x])@1
-m[s1]@3 := r[x]@1
+\begin{lemma}
+  If $\prog$ with $\iov(\prog) = S \cup V \cup O$ does not have
+  integrity for some $H,C$, then there exists $\setit{x} \cup V'
+  \subseteq \cinputs$ where $\setit{x}$ has an aberration and
+  $\notcondsep{\progd(\prog)}{S_H}{\setit{x} \cup V'}{\houtputs \cup O_H}$.
+\end{lemma}
 
-// m[s1]@2 : { c(r[x]@1, { c(r[local]@1, {s[1]@1} ) }
-// m[s1]@3 : { r[x]@1 }
-\end{verbatimtab}
-
-\begin{verbatimtab}
-m[x]@1 := s2(s[x],-r[x],r[x])@2
-
-// m[x]@1 == s[x]@2 + -r[x]@2 
-// m[x]@1 : { c(r[x]@2, { s[x]@2 }) }
-
-m[y]@1 := OT(s[y]@1,-r[y],r[y])@2
-
-// m[y]@1 == s[y]@1 + -r[y]@2
-// m[y]@1 : { c(r[y]@2, { s[y]@1 }) } 
-\end{verbatimtab}
-\end{comment}
+\begin{lemma}
+  Given $\prog$ with $\iov(\prog) = S \cup V \cup O$ and
+  $\ipj{\eqs}{\prog}{\Delta}$ valid for some $\eqs$.  For all $x \in
+  \cinputs$ with an aberration, if
+  $\notcondsep{\progd(\prog)}{S_H}{\setit{x} \cup V'}{\houtputs \cup O_H}$
+  then there exists $y \in \houtputs \cup O_H$ such that
+  $\cheatj{\Delta}{H,C}{\seclev}$ and $\seclev(y) = \lolab$.
+\end{lemma}

### 2024-10-14 esop25/cpj.tex
checkpoint

@@
 
-\subsubsection{Type Soundness}
-
-%\begin{definition}
-%  Given $\prog$ and $X_1,X_2 \subseteq \vars{\prog}$, we say that
-%  $X_1$ \emph{interferes with} $X_2$
-%  iff $\progd(\prog) \not\vdash X_1 * X_2$.
-%\end{definition}
-
-\begin{lemma}
-  \label{lemma-interference}
-  Given $\prog$ with $\iov(\prog) = S \cup V \cup O$ and $\flips(\prog) = R$,
-  for all $x \in S \cup R$ and $M \subseteq V$ exactly one of the following conditions holds:
-  \begin{enumerate}[\hspace{5mm}i.]
-  \item $\notsep{\progd(\prog)}{\setit{x}}{M}$
-  \item There exists $X \subseteq S \cup R$ such that
-    $\notsep{\progd(\prog)}{\setit{x} \cup X}{M}$ while
-    $\sep{\progd(\prog)}{\setit{x}}{M}$ and $\sep{\progd(\prog)}{X}{M}$,
-    and we say that $x$ encodes $X$ in $M$.
-  \item Neither condition (i) nor (ii) hold, implying $\sep{\progd(\prog)}{\setit{x}}{M}$.
-  \end{enumerate}
-\end{lemma}
-
-%\begin{lemma}
-%  \label{lemma-eqsprogsound}
-%  If $\toeq{\prog} \models \eqs$ and $\eqs \models \eqs'$, then
-%  for all $\store \in \runs(\prog)$ we have $\store \models \eqs'$.
-%\end{lemma}
-
-\begin{definition}
-  \label{definition-sound}
-  Given $\prog$ with $\iov(\prog) = S \cup M \cup P \cup O$
-  and $\flips(\prog) = R$, we say that
-  \emph{$\Gamma$ is sound for $\prog$} iff for all $M' \subseteq M$
-      and $x \in S \cup R$, the following
-      conditions hold:
-  \begin{enumerate}[\hspace{5mm}i.]
-  \item  if $\notsep{\progd(\prog)}{\setit{x}}{M'}$
-    there exists $\ty$ with $\leakj{\Gamma}{M'}{\ty}$ and $x \in \ty$.
-    \item  if there exists  $X \subseteq S \cup R$ such that
-      $\setit{x}$ encodes $X$ in $M'$, 
-      then $\leakj{\Gamma}{M'}{\ty}$ and
-      $\leakclose{\Gamma}{\ty \cup X}{\ty'}$ and
-      $x \in \ty'$.
-  \end{enumerate}
-\end{definition}
-
-\begin{lemma}
-  \label{lemma-cpjsound}
-  If $\cpj{R}{\eqs}{\eqspre \wedge \toeq{\prog}}{\Gamma}$ is valid then $\Gamma$ is
-  sound for $\prog$.
-\end{lemma}
-\begin{proof}
-  Unrolling Definition \ref{definition-sound}, proceed by induction on $M'$.
-  In case $M' = \varnothing$ the result follows trivially. Otherwise
-  $M' = M_0 \cup \{ \mx{w}{\cid} \}$ and wlog $\Gamma = \Gamma_0; \mx{w}{\cid} : \ty_0$
-  and $\prog = \prog_0;\xassign{\mx{w}{\cid}}{\be}{\cid_0}$, where $R = R_0;R_1$ and:
-  \begin{mathpar}
-    \cpj{R_0}{\eqs}{\eqspre \wedge \toeq{\prog_0}}{\Gamma_0}
-
-    \eqj{R_1}{\eqs}{\toeq{\elab{\be}{\cid}}}{\ty_0}
-  \end{mathpar}
-  are both valid, and by the induction hypothesis $\Gamma_0$ is sound for $\prog_0$.
-  We can then proceed by a second induction on the derivation of $\ty_0$, where there
-  are two subcases defined by the $\TirName{DepTy}$ and $\TirName{Encode}$ rules.
-  In the former case the result is immediate since all variables occuring in
-  $\toeq{\elab{\be}{\cid}}$ are elements of $\ty_0$ by definition. In the latter case,
-  we have:
-  \begin{mathpar}
-    \eqs \models \toeq{\elab{\be}{\cid}} \eop \phi' \fminus \rx{w}{\cid}
-
-    \ty_0 = \cty{\rx{w}{\cid}}{\ty_1}
-
-    R_1 = R_2;\setit{\rx{w}{\cid}}
-
-    \eqj{R_2}{\eqs}{R_2}{\ty_1}
-  \end{mathpar}
-  These facts imply that $\rx{w}{\cid}$ is used as a one-time-pad to encode
-  $\vars(\phi)$ in $M'$ 
-  \cite{barthe2019probabilistic}. Observing that:
-  $$
-  \leakclose{\Gamma}{\setit{\cty{\rx{w}{\cid}}{\ty_1}, \rx{w}{\cid}}}{\ty'}
-  $$
-  The result in this subcase follows by the (second) induction hypothesis. \qed
-\end{proof}
-
 Our main type soundness result combines typing and leakage judgements,
 and guarantees gradual release of protocols-- that is, that adversarial
-messages reflect no dependencies on honest secrets.
+messages reflect no dependencies on honest secrets. Proof details are
+provided in an Appendix.
 \begin{theorem}
@@
 \end{theorem}
-\begin{proof}
-  Immediate by Lemmas \ref{lemma-cpjsound} and \ref{lemma-interference}. \qed
-\end{proof}
 
@@
 
-Of course, it is important to recall from \cite{skalka-near-ppdp24}
-that gradual release does \emph{not} imply passive security, since it
-only concerns adversarial messages, but not public reveals of outputs
-where dependencies on input secrects can occur even in secure
-protocols.
+It is important to recall from \cite{skalka-near-ppdp24} that gradual
+release does \emph{not} imply passive security, since it only concerns
+adversarial messages, but not public reveals of outputs where
+dependencies on input secrects can occur even in secure
+protocols. Nevertheless, gradual release is an important
+confidentiality property for MPC (and other distributed) protocols.
 

### 2024-10-14 esop25/ipj.tex
checkpoint

--- 2024-10-12 esop25/ipj.tex
@@
 
-In a malicious setting, detecting cheating by adding
-information-theoretic secure MAC codes to shares is a fundamental
-technique pioneered in systems such as BDOZ and SPDZ
-\cite{SPDZ1,SPDZ2,BDOZ,10.1007/978-3-030-68869-1_3}.  These protocols
-assume a pre-processing phase that distributes shares with MAC codes
-to clients.  This integrates well with pre-processed Beaver Triples to
-implement malicious secure, and relatively efficient, multiplication
-\cite{evans2018pragmatic}.
+Our integrity type system is essentially a taint analysis, with
+machinery to allow consideration of any $H,C$ partitioning.  We also
+include rules to consider MAC codes to prevent cheating. We consider a
+semi-homomorphic encryption scheme that is used in BDOZ and SPDZ
+\cite{SPDZ1,SPDZ2,BDOZ,10.1007/978-3-030-68869-1_3} and is applicable
+to various species of circuits. However other MAC schemes could be
+considered with a modified derivation rule incorporated with our taint
+analysis framework. BDOZ/SPDZ assumes a pre-processing phase that
+distributes shares with MAC codes to clients.  This integrates well
+with pre-processed Beaver Triples to implement malicious secure, and
+relatively efficient, multiplication \cite{evans2018pragmatic}.
 
-A field value $v$ is secret shared among 2 clients in BDOZ with
-accompanying MAC values.  Each client $\cid$ gets a pair of values
-$(v_\cid,m_\cid)$, where $v_\cid$ is a share of $v$ reconstructed by
-addition, i.e., $v = \fcod{v_1 \fplus v_2}$, and $m_\cid$ is a MAC of
-$v_\cid$.  More precisely, $m_\cid = k + (k_\Delta * v_\cid)$ where
-\emph{keys} $k$ and $k_\Delta$ are known only to $\cid' \ne \cid$ and
-$k_\Delta$. The \emph{local key} $k$ is unique per MAC, while the
-\emph{global key} $k_\Delta$ is common to all MACs authenticated by
-$\cid'$. This is a semi-homomorphic encryption scheme that supports
-addition of shares and multiplication of shares by a constant-- for
-more details the reader is referred to Orsini
-\cite{10.1007/978-3-030-68869-1_3}. We note that while we restrict
-values $v$, $m$, and $k$ to the same field in this presentation for
-simplicity, in general $m$ and $k$ can be in extensions of
-$\mathbb{F}_p$.
+In BDOZ, a field value $v$ is secret shared among 2 clients 
+with accompanying MAC values.  Each client $\cid$ gets a pair of
+values $(v_\cid,m_\cid)$, where $v_\cid$ is a share of $v$
+reconstructed by addition, i.e., $v = \fcod{v_1 \fplus v_2}$, and
+$m_\cid$ is a MAC of $v_\cid$.  More precisely, $m_\cid = k +
+(k_\Delta * v_\cid)$ where \emph{keys} $k$ and $k_\Delta$ are known
+only to $\cid' \ne \cid$ and $k_\Delta$. The \emph{local key} $k$ is
+unique per MAC, while the \emph{global key} $k_\Delta$ is common to
+all MACs authenticated by $\cid'$. This is a semi-homomorphic
+encryption scheme that supports addition of shares and multiplication
+of shares by a constant-- for more details the reader is referred to
+Orsini \cite{10.1007/978-3-030-68869-1_3}. We note that while we
+restrict values $v$, $m$, and $k$ to the same field in this
+presentation for simplicity, in general $m$ and $k$ can be in
+extensions of $\mathbb{F}_p$.
+
+\ipjfig
 
@@
 
-\ipjfig
-
 The integrity of values during protocol executions depends on the
@@
 types, we want our basic type analysis to be generalizable to
-arbitrary $H,C$. So integrity type judgements are of the form
-$\ipj{\eqs}{\prog}{\Delta}$, where $\eqs$ plays a similar role as in
-confidentiality types, as a possibly weakened expression of algebraic
-properties of $\prog$, and encironments $\Delta$ record integrity
-information. Environments are ordered lists, since order of evaluation
-is important for tracking integrity-- validating low integrity
-information through MAC checking allows \emph{subsequent} treatment of
-it as high integrity, but not \emph{before}. Note that MAC checking is
-\emph{not} a form of endorsement as it is usually defined (as a dual
-of declassification \cite{sabelfeld2009declassification}), since a
-successful check guarantees no deviation from the protocol.
+arbitrary $H,C$. So integrity type judgements, defined in Figure
+\ref{fig-ipj}, are of the form $\ipj{\eqs}{\prog}{\Delta}$, where
+$\eqs$ plays a similar role as in confidentiality types, as a possibly
+weakened expression of algebraic properties of $\prog$, and
+environments $\Delta$ record dependency information that can be
+specialized to a given $H,C$. Integrity environments are ordered lists
+since order of evaluation is important for tracking integrity--
+validating low integrity information through MAC checking allows
+subsequent treatment of it as high integrity, but not before. Note
+that MAC checking is \emph{not} a form of endorsement as it is usually
+defined (as a dual of declassification
+\cite{sabelfeld2009declassification}), since a successful check
+guarantees no deviation from the protocol.
 
 The typing of protocols depends on a type judgement for expressions
-of the form $\itj{\cid}{\be}{\ity{\cid}{V}}$, which records
+of the form $\itj{\cid}{\be}{V}$, which records
 messages and reveals $V$ used in the construction of $\be$.
-In an assignment, this source is recorded in $\Delta$-- so for
+This source is recorded in $\Delta$ for message sends-- so for
 example, for some string $w$ letting:
@@
 \end{definition}
-But the values of messages and reveals may be
-affected by values sent by other parties, which is
-recorded in the $\Delta$ obtained from its typing. To determine
-whether this impacts integrity for any given $H,C$, we apply
-the inductively defined rewrite relation $\cheatj{\Delta}{H,C}{\seclev}$,
-axiomatized in Figure \ref{fig-cheatj}. The composition
-of typing and this latter rewriting obtains essentially
-a traditional taint analysis, modulo MAC checking.
-
-\subsubsection{Type Soundness}
-
-\begin{definition}
-  Given $H,C$ and $\prog$ with $\iov(\prog) = S \cup V \cup O$,
-  we say $x \in V \cup O$ has an \emph{aberration} iff
-  $\exists \adversary, \store \in \botruns(\prog)$ such that
-  $\store(x) \ne \bot$ and $\neg\exists \store' \in \runs(\prog)$
-  with  $\store_{S_H} = \store'_{S_H}$ and
-  $\store(x) = \lcod{\store',\be}{\cid}$ where
-  $\xassign{x}{\be}{\cid} \in \prog$.
-\end{definition}
-
-\begin{lemma}
-  If $\prog$ with $\views(\prog) = V$ does not have integrity for some
-  $H,C$, then there exists $x \in \houtputs$ with an
-  aberration.
-\end{lemma}
-
-\begin{lemma}
-  Given $\prog$ and $H,C$, if $x$ has an aberration then one of
-  the two inductively defined conditions holds:
-  \begin{enumerate}[\hspace{5mm}i.]
-  \item There exists $\xassign{x}{\be}{\cid} \in \prog$ with $\cid \in C$.
-  \item There exists $\xassign{x}{\be}{\cid} \in \prog$ with $\cid \in H$ and
-    $x' \in \vars(\be)$ such that $x'$ has an aberration.
-  \end{enumerate}
-\end{lemma}
-
-\begin{lemma}
-  Given pre-processing predicate $\eqspre$ and protocol $\prog$ with
-  $\views(\prog) = V$, if 
-  $\ipj{\eqs}{\prog}{\Delta}$ is valid
-  and $x$ has an aberration for some $H,C$,
-  then $\cheatj{\Delta}{H,C}{\seclev}$ and $\seclev(x) = \lolab$.
-\end{lemma}
-
+But the values of messages and reveals may be affected by values sent
+by other parties, which is recorded in the $\Delta$ obtained from its
+typing. To determine whether this impacts integrity for any given
+$H,C$, we apply the inductively defined rewrite relation
+$\cheatj{\Delta}{H,C}{\seclev}$, axiomatized in Figure
+\ref{fig-cheatj}. The composition of typing and this latter rewriting
+obtains essentially a traditional taint analysis, modulo MAC checking.
 Our main type soundness result for integrity is then given as follows.

### 2024-10-14 esop25/metalang.tex
checkpoint

@@
 
-Practical MPC computations protocols are
-typically composed of compositional units. Examples include GMW circuits
-and Yao's Garbled Circuits (YGC), that are composed of so-called
-garbled gates. Languages such as Fairplay \cite{269581} provide gates as
-units of abstraction that are ``wired'' together by the programmer to
-generate a complete circuit.
-
-The $\fedprot$ language is low-level and does not include abstractions
-for defining composable elements. So in this Section we introduce the
-$\metaprot$ language that includes structured data and function
-definitions for defining composable protocol elements at a higher
-level of abstraction.  The $\metaprot$ language is a
-\emph{metalanguage}, where $\fedprot$ protocols are the residuum of
-computation. In addition to these declarative benefits of $\metaprot$,
-component definitions support compositional verification of larger
-protocols as we will discuss with examples in Section
-\ref{section-examples}.
+The $\metaprot$ language \cite{skalka-near-ppdp24} includes structured
+data and function definitions for defining composable protocol
+elements at a higher level of abstraction.  The $\metaprot$ language
+is a \emph{metalanguage} aka metaprogramming language, where a
+$\fedprot$ protocols is the result of computation. In addition to
+these declarative benefits of $\metaprot$, component definitions
+support compositional verification of larger protocols as we will
+discuss with extended examples in Section \ref{section-examples}.
 
@@
 both a codebase and a ``main'' program that uses the codebase.
+We disallow recursion, mainly to guarantee decidability
+of type dependence (Section \ref{section-metalangty}).
 \begin{definition}
@@
   $$
-  \store(x) \defeq \store'(\mathit{rename}(x)) 
+  \forall x \in \dom(\vars(E)) . \store(x) \defeq \store'(\mathit{rename}(x)) 
   $$

### 2024-10-14 esop25/appendix.tex
checkpoint

@@
 \end{lemma}
+
+\subsection{Type Checking Soundness}
+
+\begin{lemma}
+  \label{lemma-fresh}
+  Given $w$ and $\cid$ with $\fresh(w,\cid)$, if $\notg{\eqs}[w/y_1,\cid/y_n]$
+  is not satisfiable, then there exists no $w',\cid'$ such that
+  $\notg{\eqs}[w'/y_1,\cid'/y_n]$ is satisifiable.
+\end{lemma}
+
+\begin{proof}
+  Let $\eqs = \notg{\eqs}[w/y_1,\cid/y_n]$ and $\eqs' =
+  \notg{\eqs}[w'/y_1,\cid'/y_n]$. If $w',\cid'$ are fresh or chosen
+  such that $\eqs'$ is an $\alpha$-renaming of $\eqs$, the result is
+  immediate.  Otherwise, it must be the case that at least two
+  distinct variables $x_1$ and $x_2$ in $\eqs$ are renamed as a single
+  variable $x$ in $\eqs'$, given freshness of $w,\cid$. Let
+  $\mathit{rename}$ be the function that maps variables in $\eqs$ to
+  variables in $\eqs'$ such that $\mathit{rename}(\eqs) = \eqs'$. Let
+  $\store'$ be a model of $\eqs'$, and define $\store$ such that:
+  $$
+  \forall x \in \dom(\vars(E)) . \store(x) \defeq \store'(\mathit{rename}(x)) 
+  $$
+  Then $\store$ must be a model of $\eqs$, which is a contradiction. \qed
+\end{proof}
+
+\begin{lemma}
+  \label{lemma-eqs-notg}
+  For all $\fresh(\mv_1,\ldots,\mv_n)$, $\notg{E_1}$, $\notg{E_2}$, assuming: 
+  \begin{mathpar}
+    \notg{\eqs_1}[\mv_1/y_1]\cdots[\mv_n/y_n]  \redx \eqs_1
+
+    \notg{\eqs_2}[\mv_1/y_1]\cdots[\mv_n/y_n]  \redx \eqs_2
+
+    E_1 \models E_2
+  \end{mathpar}
+  then there exists no $\mv_1',\ldots,\mv_n'$ such that:
+    \begin{mathpar}
+    \notg{\eqs_1}[\mv_1'/y_1]\cdots[\mv_n'/y_n]  \redx \eqs_1
+
+    \notg{\eqs_2}[\mv_1'/y_1]\cdots[\mv_n'/y_n]  \redx \eqs_2
+
+    E_1' \not\models E_2'
+  \end{mathpar}  
+\end{lemma}
+
+\begin{proof}
+  Given the assumptions, suppose that $E_1' \not\models E_2'$
+  is an $\alpha$-renaming of $E_1 \models E_2$. In this
+  case the consequence follows immediately. Otherwise,
+  suppose on the contrary we can witness
+  $E_1' \not\models E_2'$. By definition this means that
+  $E_1' \wedge \neg E_2'$ is satisfiable
+  but $E_1 \wedge \neg E_2$ is not, which is a contradiction
+  by Lemma \ref{lemma-fresh}. \qed 
+\end{proof}
+
+\begin{proof}[Theorem \ref{theorem-mtj}]
+  (Sketch) The result follows by induction on derivation of the
+  main type judgement for $\instr$. The \TirName{Mesg} case follows
+  by Lemma \ref{lemma-atj-sound}. The \TirName{Encode} and \TirName{App}
+  cases follow by definition of the \TirName{Sig} rule and Lemma
+  \ref{lemma-eqs-notg}. 
+\end{proof}

### 2024-10-14 esop25/metalang.tex
checkpoint

@@
 data and function definitions for defining composable protocol
-elements at a higher level of abstraction.  The $\metaprot$ language
-is a \emph{metalanguage} aka metaprogramming language, where a
-$\fedprot$ protocols is the result of computation. In addition to
-these declarative benefits of $\metaprot$, component definitions
-support compositional verification of larger protocols as we will
-discuss with extended examples in Section \ref{section-examples}.
+elements at a higher level of abstraction than $\minicat$.  The
+$\metaprot$ language is a \emph{metalanguage} aka metaprogramming
+language, where a $\fedprot$ protocols is the result of
+computation. In addition to these declarative benefits of $\metaprot$,
+component definitions support compositional verification of larger
+protocols. Separate verification of well-designed components results
+in confidentiality and integrity properties to be recorded in their
+types, allowing for significant reduction of SMT verification in whole
+program analysis, as we will discuss with extended examples in Section
+\ref{section-examples}.
 
@@
 compute values $\mv$ and, as a distinct form, functions that compute
-protocols.  Formally, we consider a complete metaprogram to include
+protocols $\prog$.  Formally, we consider a complete metaprogram to include
 both a codebase and a ``main'' program that uses the codebase.
@@
 integrity typings as defined in Sections \ref{section-cpj} and
-\ref{section-ipj}. Returning to the example in Section
-\ref{section-cpj}, note that the key equivalence of the $\ttt{mux}$
-expression with a one-time-pad encryption we observe there
-is not trivial. To ensure that this sort of equivalence
-can be picked up by the type system, we introduce a
-an annotation form that allows the programmer to provide
-the needed hint.
+\ref{section-ipj}. Returning to the YGC secret encoding example in
+Section \ref{section-cpj}, note that the key equivalence of the
+$\ttt{mux}$ expression with a one-time-pad encryption we observe there
+is not trivial. To ensure that this sort of equivalence can be picked
+up by the type system, we introduce an $\ttt{as}$ annotation form that
+allows the programmer to provide the needed hint.
 
 Consider the following $\ttt{encode}$ function definition that
-generalizes this encoding for any identifier $y$ and sender, receiver
+generalizes this YGC encoding for any identifier $y$ and sender, receiver
 pair $s,r$. The second line in the body is a hint that
@@
 application instances. To verify the hint in $\ttt{encode}$, we can
-just choose arbitrary values $w,\cid_1,\cid_2$ for $y$, $s$, and $r$,
-evaluate $\ttt{encode}(w,\cid_1,\cid_2) \redx \prog$, and then verify:
+just choose arbitrary ``fresh'' values $w,\cid_1,\cid_2$ for $y$, $s$,
+and $r$, evaluate $\ttt{encode}(w,\cid_1,\cid_2) \redx \prog$, and
+then verify:
 $$\toeq{\prog} \models \mx{w}{\cid_2} \eop
@@
 
+\atjfig
+
 \subsubsection{$\minicat$ expression type algorithm.}
@@
 due to the \TirName{Encode} rule. But as described above, by introducing
-hint annotations we can ``cast'' any relevant expression form into
+hint annotations we can ``coerce'' any relevant expression form into
 the simplest equivalent syntactic form for one-time-pad encoding.
@@
 integrated SMT solving by assuming this sort of casting.
-
-\atjfig
 
@@
 $$
-Here, $\eqs_1$ and $\eqs_2$ are the pre- and post- conditions
+Here, $\eqs_1$ and $\eqs_2$ are the pre- and postconditions
 respectively, $\Gamma$ is the confidentiality type environment of the
@@
 one-time-pad $\rx{x}{s}$, and the integrity of $\mx{x}{r}$ is
-determined by the integrity of $\rx{x}{s}$ and $\sx{x}{s}$. The
+determined by the security level of $s$. The
 postcondition expresses the key confidentiality property of
@@
 \begin{array}[t]{ll}
-  \Pi x,s,r . & \{ \} \quad \mx{x}{r} : \cty{\rx{x}{s}}{\sx{x}{s}}, \{ \rx{x}{s} \} \\
-  & \phantom{\{ \}} \quad \mx{x}{r} : \ity{s}{\{ \rx{x}{s}, \sx{x}{s} \}} \\
+  \Pi x,s,r . & \{ \}\quad \mx{x}{r} : \cty{\rx{x}{s}}{\sx{x}{s}}, \{ \rx{x}{s} \} \ \cdot\\
+  & \phantom{\{ \}} \quad \mx{x}{r} : \ity{s}{\setit{}} \\
   & \{ \mx{x}{r} \eop \neg\sx{y}{s} \fplus \rx{y}{s} \}
@@
 Figure \ref{fig-notg}. Since expression evaluation is total we can
-just evaluate types of closed instructions, which have closed types,
-to obtain Hoare typings for instructions. In Figure \ref{fig-mtj} we
-define the type derivation rules for $\metaprot$. In the
-\TirName{Mesg} rule we use algorithmic confidentiality and integrity
-type checking in a straightforward manner.  In the \TirName{Encode}
-rule, we verify the hint given in the type annotation, and then use it
-for type checking.  In the \TirName{App} rule we verify preconditions
-of the given function type and instantiate the type and postconditions
-with the given expression arguments.  The \TirName{Sig} rule we
-specify how to verify $\Pi$ types for functions. Note that the values
-$\mv_1,\ldots,\mv_n$ chosen to instantiate the function parameters are
-chosen ``fresh'', meaning that any strings and client identifiers
-occurring as substrings in $\mv_1,\ldots,\mv_n$ are unique and
-not used in any program source code. This ensures generality
-of verification as demonstrated in Lemma \ref{lemma-eqs-notg}.
+just evaluate types of closed instructions, which are guaranteed to
+have closed types, to obtain Hoare typings for instructions. In Figure
+\ref{fig-mtj} we define the type derivation rules for $\metaprot$. In
+the \TirName{Mesg} rule we use algorithmic confidentiality and
+integrity type checking in a straightforward manner.  In the
+\TirName{Encode} rule, we verify the hint given in the type
+annotation, and then use it for type checking.  In the \TirName{App}
+rule we verify preconditions of the given function type and
+instantiate the type and postconditions with the given expression
+arguments.  The \TirName{Sig} rule we specify how to verify $\Pi$
+types for functions. Note that the values $\mv_1,\ldots,\mv_n$ chosen
+to instantiate the function parameters are chosen ``fresh'', meaning
+that any strings and client identifiers occurring as substrings in
+$\mv_1,\ldots,\mv_n$ are unique and not used in any program source
+code. This ensures generality of verification as demonstrated in Lemma
+\ref{lemma-eqs-notg} in the Appendix.
 
@@
 
-\subsection{Type Checking Soundness}
-
-\begin{lemma}
-  \label{lemma-fresh}
-  Given $w$ and $\cid$ with $\fresh(w,\cid)$, if $\notg{\eqs}[w/y_1,\cid/y_n]$
-  is not satisfiable, then there exists no $w',\cid'$ such that
-  $\notg{\eqs}[w'/y_1,\cid'/y_n]$ is satisifiable.
-\end{lemma}
-
-\begin{proof}
-  Let $\eqs = \notg{\eqs}[w/y_1,\cid/y_n]$ and $\eqs' =
-  \notg{\eqs}[w'/y_1,\cid'/y_n]$. If $w',\cid'$ are fresh or chosen
-  such that $\eqs'$ is an $\alpha$-renaming of $\eqs$, the result is
-  immediate.  Otherwise, it must be the case that at least two
-  distinct variables $x_1$ and $x_2$ in $\eqs$ are renamed as a single
-  variable $x$ in $\eqs'$, given freshness of $w,\cid$. Let
-  $\mathit{rename}$ be the function that maps variables in $\eqs$ to
-  variables in $\eqs'$ such that $\mathit{rename}(\eqs) = \eqs'$. Let
-  $\store'$ be a model of $\eqs'$, and define $\store$ such that:
-  $$
-  \forall x \in \dom(\vars(E)) . \store(x) \defeq \store'(\mathit{rename}(x)) 
-  $$
-  Then $\store$ must be a model of $\eqs$, which is a contradiction. \qed
-\end{proof}
-
-\begin{lemma}
-  \label{lemma-eqs-notg}
-  For all $\fresh(\mv_1,\ldots,\mv_n)$, $\notg{E_1}$, $\notg{E_2}$, assuming: 
-  \begin{mathpar}
-    \notg{\eqs_1}[\mv_1/y_1]\cdots[\mv_n/y_n]  \redx \eqs_1
-
-    \notg{\eqs_2}[\mv_1/y_1]\cdots[\mv_n/y_n]  \redx \eqs_2
-
-    E_1 \models E_2
-  \end{mathpar}
-  then there exists no $\mv_1',\ldots,\mv_n'$ such that:
-    \begin{mathpar}
-    \notg{\eqs_1}[\mv_1'/y_1]\cdots[\mv_n'/y_n]  \redx \eqs_1
-
-    \notg{\eqs_2}[\mv_1'/y_1]\cdots[\mv_n'/y_n]  \redx \eqs_2
-
-    E_1' \not\models E_2'
-  \end{mathpar}  
-\end{lemma}
-
-\begin{proof}
-  Given the assumptions, suppose that $E_1' \not\models E_2'$
-  is an $\alpha$-renaming of $E_1 \models E_2$. In this
-  case the consequence follows immediately. Otherwise,
-  suppose on the contrary we can witness
-  $E_1' \not\models E_2'$. By definition this means that
-  $E_1' \wedge \neg E_2'$ is satisfiable
-  but $E_1 \wedge \neg E_2$ is not, which is a contradiction
-  by Lemma \ref{lemma-fresh}. \qed 
-\end{proof}
-
-The following Theorem establishes our main result, which
-is the $\metaprot$ type checking is sound with respect to both
-confidentiality and integrity typings. Since these typings
-imply security hyperproperties of confidentiality and integrity,
-$\metaprot$ type checking implicitly enforces them.
+The following Theorem establishes our main result, that the
+$\metaprot$ type checking is sound with respect to both
+confidentiality and integrity typings. Since these typings imply
+security hyperproperties of confidentiality and integrity, $\metaprot$
+type checking implicitly enforces them. Proof details are given in the
+Appendix.
 \begin{theorem}[Soundness of $\metaprot$ type checking]
@@
 \end{theorem}
-
-\begin{proof}
-  (Sketch) The result follows by induction on derivation of the
-  main type judgement for $\instr$. The \TirName{Mesg} case follows
-  by Lemma \ref{lemma-atj-sound}. The \TirName{Encode} and \TirName{App}
-  cases follow by definition of the \TirName{Sig} rule and Lemma
-  \ref{lemma-eqs-notg}. 
-\end{proof}

### 2024-10-14 esop25/examples.tex
checkpoint

@@
 
-We consider examples that illustrate the versatility of the $\TirName{Encode}$
-typing rule for confidentiality, the use of postconditions for correctness,
-and the use of both pre-~and postconditions for integrity.
+In Section \ref{section-smt} we gave the example of 3-party additive
+sharing for any field $\fieldp{p}$. In Sections \ref{section-ipj} and
+\ref{section-metalangty} we discussed the example of YGC input encoding.
+Now we consider examples that illustrate the versatility of the
+$\TirName{Encode}$ typing rule for confidentiality, the use of
+postconditions for correctness, and the use of both pre-~and
+postconditions for integrity.
+
+Using the Satisfiability Modulo Fields theory in cvc5 on an Apple M1
+processor, we have verified correctness of additive sharing protocols
+in fields up to size $\fieldp{2^{31} - 1}$ (an approximation of 32-bit
+integers) which takes ~16ms to compute for 3 parties and not
+significantly more than for, e.g., $\fieldp{2}$. We have also verified
+the semi-homorphic encryption properties leveraged in the BDOZ circuit
+library functions (Figures \ref{fig-bdozsum} and \ref{fig-bdozmult}) in
+fields up to size $\fieldp{2^{31} - 1}$ which takes \cnote{TBD}ms for the
+property used for the postcondition of the $\ttt{mult}$ gate. This is
+categorically better performance than could be achieved using previous
+brute-force methods in $\fieldp{2}$ \cite{skalka-near-ppdp24}.
 
@@
 
-The $\ttt{encode}$ function example in Section \ref{section-metalangty}
-illustrates the use of type annotations to track confidentiality. In
-2-party Yao's Garbled Circuits (YGC), a \emph{garbler} prepares a circuit encoding
-that is shared with a \emph{evaluator}, and the $\ttt{encode}$ function
-defines the method the garbler uses to share their input secret encodings
-with the evaluator. The example illustrates how confidentiality is reflected
-in its (cipher)type.
-
 In the 2-party GMW protocol \cite{evans2018pragmatic}, another garbled
-binary circuit protocol, values are encrypted in a manner similar
-to BDOZ as described in Section \ref{section-ipj}. In our definition
-of GMW we use the convention that shared values $\macgv{w}$ are identified by
-strings $w$ and encoded as shares $\mx{w}{1}$ and $\mx{w}{2}$.  As for
-YGC, ciphertypes reflect the confidentiality of GMW input encodings as
-defined in the $\ttt{encodegmw}$ function in Figure \ref{fig-gmw}. No
-programmer annotation is needed given the syntax of the function
-body. More interestingly, the $\ttt{andgmw}$ gate definition shows how
-a programmer hint can express the relevant confidentiality property of
-the output share $\mx{z}{2}$ on client 2. The non-trivial equivalence
-can be verified by SMT once during verification, and subsequently
-confidentiality is expressed in the dependent type of the function
-(as for $\ttt{encode}$).
+binary circuit protocol, values are encrypted in a manner similar to
+BDOZ as described in Section \ref{section-ipj}. In our definition of
+GMW we use the convention that shared values $\macgv{\mesg{w}}$ are
+identified by strings $w$ and encoded as shares $\mx{w}{1}$ and
+$\mx{w}{2}$.  As for YGC, ciphertypes reflect the confidentiality of
+GMW input encodings as defined in the $\ttt{encodegmw}$ function in
+Figure \ref{fig-gmw}. No programmer annotation is needed given the
+syntax of the function body. More interestingly, the $\ttt{andgmw}$
+gate definition shows how a programmer hint can express the relevant
+confidentiality property of the output share $\mx{z}{2}$ on client
+2. The non-trivial equivalence can be verified by SMT once during
+verification, and subsequently confidentiality is expressed in the
+dependent type of the function (as for $\ttt{encode}$).
 
@@
 ``secure opening''-- each party sends it's local share of a
-global value $\macgv{w}$ along with its MAC to the other party,
+global value $\macgv{\mesg{w}}$ along with its MAC to the other party,
 which is authenticated via the $\macbdoz{w}$ check (in \ttt{\_open}),
-and then each party reconstructs $\macgv{w}$.
+and then each party reconstructs $\macgv{\mesg{w}}$.
 
@@
 expressed in $\eqspre$ for input secrets $\sx{\ttt{"x"}}{1}$ and
-$\sx{\ttt{"y"}}{2}$, for example, which would subsume the following
+$\sx{\ttt{"y"}}{2}$, for example, which subsume the following
 constraints on shares:
@@
 setting is to use \emph{Beaver Triples}. Recall that Beaver triples
-are values $a,b,c$ with $\fcod{a * b} = c$, unique per multiplication
-gate, that are secret shared with clients during pre-processing.
-In our encoding we assume the additional convention that each gate
-output identifier distinguishes the Beaver triple, so for
-example the share of the $a$ value for a gate $\ttt{"g1"}$
-is identified by $\ttt{"g1as"}$, etc., and $\eqspre$ would
-subsume the following constraints for the $a,b,c$ values of
-gate $\ttt{"g1"}$.
-{\footnotesize$$
+are values $a,b,c$ with $a$ and $b$ chosen randomly and $c = a * b$,
+unique per multiplication gate, that are secret shared with clients
+during pre-processing.  In our encoding we assume the additional
+convention that each gate output identifier distinguishes the Beaver
+triple, so for example the share of the $a$ value for a gate
+$\ttt{"g1"}$ is identified by $\ttt{"g1as"}$, etc., and $\eqspre$
+subsumes the following constraints for the $a,b,c$ values of gate
+$\ttt{"g1"}$.  {\footnotesize$$
 \begin{array}{l}

### 2024-10-14 code/SMT/eqs.py
Update eqs.py with timing logic.

@@ -1,4 +1,5 @@
 from cvc5.pythonic import *
+import time
 
 s = Solver()
 
@@ -6,9 +7,9 @@ s = Solver()
 
 s.resetAssertions()
 
-m1, f11, x = FiniteFieldElems('m1 f11 x', 7)
-m2, f21, y = FiniteFieldElems('m2 f21 y', 7)
-p1, p2, o = FiniteFieldElems('p1 p2 o', 7)
+m1, f11, x = FiniteFieldElems('m1 f11 x', 2147483647)
+m2, f21, y = FiniteFieldElems('m2 f21 y', 2147483647)
+p1, p2, o = FiniteFieldElems('p1 p2 o', 2147483647)
 
 # constraints from the protocol
 s.add(m2 == x - f11)
@@ -20,17 +21,21 @@ s.add(o == p1 + p2)
 # sanity check (should be sat)
 s.check()
 
+start_c1 = time.perf_counter()
+
 # correctness (should be unsat)
 s.check(x + y != o)  
+
+stop_c1 = time.perf_counter()
 
 # 3-party addition in F7
 
 s.resetAssertions()
 
-m21, m31, s1, rl1, r1 = FiniteFieldElems('m21 m31 s1 rl1 r1', 7)
-m12, m32, s2, rl2, r2 = FiniteFieldElems('m12 m32 s2 rl2 r2', 7)
-m13, m23, s3, rl3, r3 = FiniteFieldElems('m13 m23 s3 rl3 r3', 7)
-p1, p2, p3, o = FiniteFieldElems('p1 p2 p3 o', 7)
+m21, m31, s1, rl1, r1 = FiniteFieldElems('m21 m31 s1 rl1 r1', 2147483647)
+m12, m32, s2, rl2, r2 = FiniteFieldElems('m12 m32 s2 rl2 r2', 2147483647)
+m13, m23, s3, rl3, r3 = FiniteFieldElems('m13 m23 s3 rl3 r3', 2147483647)
+p1, p2, p3, o = FiniteFieldElems('p1 p2 p3 o', 2147483647)
 
 s.add(m12 == s1 - rl1 - r1)
 s.add(m13 == r1)
@@ -46,8 +51,13 @@ s.add(o == p1 + p2 + p3)
 # sanity check (should be sat)
 s.check()
 
+start_c2 = time.perf_counter()
+
 # correctness (should be unsat)
+
 s.check(o != s1 + s2 + s3)
+
[…]

### 2024-10-14 esop25/frontmatter.tex
add a bit in introduction
checkpoint

@@
   technology for data privacy in modern distributed applications. We
-  devevelop a new type theory to automatically enforce correctness,
+  develop a new type theory to automatically enforce correctness,
   confidentiality, and integrity properties of protocols written in
@@
   finite fields, which supports precise and efficient analysis. Our
-  approach is compositional, scalable, and generalizes to arbitrary
+  approach is automated, compositional, scalable, and generalizes to arbitrary
   prime fields for data and key sizes.

### 2024-10-14 esop25/examples.tex
checkpoint

@@
 in fields up to size $\fieldp{2^{31} - 1}$ (an approximation of 32-bit
-integers) which takes ~16ms to compute for 3 parties and not
+integers) which takes ~7ms to compute for 3 parties and not
 significantly more than for, e.g., $\fieldp{2}$. We have also verified
@@
 library functions (Figures \ref{fig-bdozsum} and \ref{fig-bdozmult}) in
-fields up to size $\fieldp{2^{31} - 1}$ which takes \cnote{TBD}ms for the
+fields up to size $\fieldp{2^{31} - 1}$ which take ~8ms for the
 property used for the postcondition of the $\ttt{mult}$ gate. This is
-categorically better performance than could be achieved using previous
+categorically better performance than can be achieved using previous
 brute-force methods in $\fieldp{2}$ \cite{skalka-near-ppdp24}.
@@
 explicit. Finally, the postcondition of $\ttt{mult}$ expresses the
-correctness property of the multiplication gate.
+correctness property of the multiplication gate. In any case,
+integrity of any circuit constructed from the $\ttt{sum}$ and
+$\ttt{mult}$ library functions will require little MPC.
 

### 2024-10-14 esop25/appendix.tex
checkpoint

@@
 
+The following Lemma enumerates critical properties of dependence of
+views on input secrets and random variables, and follows by standard
+properties of pmfs.
 \begin{lemma}
@@
 
+The following Definition characterizes the crucial representation invariant
+preserved by confidentiality type checking. 
 \begin{definition}
@@
 \end{definition}
-
+Now we can prove that this invariant is preserved by type checking.
 \begin{lemma}
@@
 \end{proof}
-
+On the basis of the above, proving the main result is straightforward.
 \begin{proof}[Theorem \ref{theorem-cpj}]
@@
 
+A key observation is that, to undermine integrity adversarial strategies
+must be different from passive strategies, in particular at least one
+adversarial input must be different than any used in a passive strategy
+(an abberation). 
 \begin{definition}
@@
 \end{definition}
-
+It is straightforward to prove that if an abberation can
+exist, then it will be assigned $\lolab$ integrity in our
+analysis. 
 \begin{lemma}
+  \label{lemma-aberration-low}
   Given protocol $\prog$ with
@@
 \end{lemma}
-
+\begin{proof}
+  Given $H,C$, \emph{every} adversarial input is assigned $\lolab$
+  integrity by definition of the $\TirName{Mesg}$ rule and the
+  \emph{leak} judgement.  The exception to this is messages that pass
+  the $\TirName{MAC}$ check-- success of this check guarantees that
+  the message is \emph{not} an aberration, so it can safely be
+  switched to $\hilab$ integrity. \qed
+\end{proof}
+The following Lemma follows trivially by Definition \ref{def-integrity},
+since without an aberration adversarial inputs are the same as some passive
+strategy.
 \begin{lemma}
+  \label{lemma-undermined-logic}
   If $\prog$ with $\iov(\prog) = S \cup V \cup O$ does not have
@@
 \begin{lemma}
+  \label{lemma-undermined}
   Given $\prog$ with $\iov(\prog) = S \cup V \cup O$ and
@@
 \end{lemma}
+\begin{proof}
+  By Lemma \ref{lemma-aberration-low} and definition of the
+  confidentiality type analysis, since if $\setit{x} \cup V'$
+  interferes with $\houtputs \cup O_H$ then $\setit{x}$
+  must occur in the definition of some $y \in \houtputs \cup O_H$,
+  yielding $\seclev(y) = \lolab$ in the integrity label assignment
+  judgement defined in Figure \ref{fig-cheatj}. \qed
+\end{proof}
+On the basis of the above we can prove our main result. 
+\begin{proof}[Theorem \ref{theorem-ipj}]
+  By Lemmas \ref{lemma-undermined} and \ref{lemma-undermined-logic}. \qed
+\end{proof}
 
@@
 
+Soundness of the $\metaprot$ type analysis follows mostly by a
+straightforward mapping to the confidentiality and integrity type
+systems. The most interesting rules are signature verification
+($\TirName{Sig}$) and function application ($\TirName{App}$). 
+Here the result hinges on the selection of ``fresh'' variables
+that guarantee generalization of constraint entailment.
+First we obtain a key result about instantiation and
+satisfiability of constraints. 
 \begin{lemma}
@@
 \end{proof}
-
+Now we can extend the preceding Lemma to constraint entailment,
+incorporating evaluation of expressions withing constraints. 
 \begin{lemma}
@@
 \end{proof}
-
+Now we can consider the main result.
 \begin{proof}[Theorem \ref{theorem-mtj}]

### 2024-10-14 esop25/appendix.tex
checkpoint

@@
   are two subcases defined by the $\TirName{DepTy}$ and $\TirName{Encode}$ rules.
-  In the former case the result is immediate since all variables occuring in
+  In the former case the result is immediate since all variables occurring in
   $\toeq{\elab{\be}{\cid}}$ are elements of $\ty_0$ by definition. In the latter case,
@@
 adversarial input must be different than any used in a passive strategy
-(an abberation). 
+(an aberration). 
 \begin{definition}
@@
 \end{definition}
-It is straightforward to prove that if an abberation can
+It is straightforward to prove that if an aberration can
 exist, then it will be assigned $\lolab$ integrity in our
@@
   is not satisfiable, then there exists no $w',\cid'$ such that
-  $\notg{\eqs}[w'/y_1,\cid'/y_n]$ is satisifiable.
+  $\notg{\eqs}[w'/y_1,\cid'/y_n]$ is satisfiable.
 \end{lemma}
@@
 Now we can extend the preceding Lemma to constraint entailment,
-incorporating evaluation of expressions withing constraints. 
+incorporating evaluation of expressions within constraints. 
 \begin{lemma}

### 2024-10-14 esop25/cpj.tex
checkpoint

@@
 dependencies.  The \TirName{Send} rule records the types of messages,
-reveals, and outputs in enviroments, and the \TirName{Seq} rule
+reveals, and outputs in environments, and the \TirName{Seq} rule
 combines the types of sequenced programs.  Given this we define
@@
 a partitioning $H,C$. In particular, if the adversary has access to an
-encypted value and another value that is dependent on its one-time-pad
+encrypted value and another value that is dependent on its one-time-pad
 key-- i.e., access to values with types of the forms:
@@
 adversarial messages, but not public reveals of outputs where
-dependencies on input secrects can occur even in secure
+dependencies on input secrets can occur even in secure
 protocols. Nevertheless, gradual release is an important

### 2024-10-14 esop25/examples.tex
checkpoint

@@
 
-Using the Satisfiability Modulo Fields theory in cvc5 on an Apple M1
-processor, we have verified correctness of additive sharing protocols
-in fields up to size $\fieldp{2^{31} - 1}$ (an approximation of 32-bit
-integers) which takes ~7ms to compute for 3 parties and not
-significantly more than for, e.g., $\fieldp{2}$. We have also verified
-the semi-homorphic encryption properties leveraged in the BDOZ circuit
-library functions (Figures \ref{fig-bdozsum} and \ref{fig-bdozmult}) in
-fields up to size $\fieldp{2^{31} - 1}$ which take ~8ms for the
-property used for the postcondition of the $\ttt{mult}$ gate. This is
-categorically better performance than can be achieved using previous
-brute-force methods in $\fieldp{2}$ \cite{skalka-near-ppdp24}.
+To evaluate our system especially regarding performance, using the
+Satisfiability Modulo Fields theory in cvc5 on an Apple M1 processor,
+we have verified correctness of additive sharing protocols in fields
+up to size $\fieldp{2^{31} - 1}$ (an approximation of 32-bit integers)
+which takes $\sim$7ms to compute for 3 parties and not significantly more
+than for, e.g., $\fieldp{2}$. We have also verified the semi-homomorphic
+encryption properties leveraged in the BDOZ circuit library functions
+(Figures \ref{fig-bdozsum} and \ref{fig-bdozmult}) in fields up to
+size $\fieldp{2^{31} - 1}$ which take $\sim$8ms for the property used for
+the postcondition of the $\ttt{mult}$ gate. This is categorically
+better performance than can be achieved using previous brute-force
+methods in $\fieldp{2}$ \cite{skalka-near-ppdp24}.
 
@@
 $$}
+
 The definition of the $\ttt{mult}$ gate is presented in Figure
@@
 As for $\ttt{\_sum}$, the postcondition of the $\ttt{\_mult1}$
-function expresses the relevant semi-homorphic encryption property of
+function expresses the relevant semi-homomorphic encryption property of
 the resulting share, MAC, and key after gate execution-- specifically,
@@
 integrity of any circuit constructed from the $\ttt{sum}$ and
-$\ttt{mult}$ library functions will require little MPC.
+$\ttt{mult}$ library functions will require little SMT overhead
+due to compositional verification.
 

### 2024-10-14 esop25/intro.tex
checkpoint

@@
 exponential complexity, necessitating semi-automated proof
-techniqued. In contrast, our type systems scale to arbitrary prime
+techniques. In contrast, our type systems scale to arbitrary prime
 fields and whole program analysis for larger circuits (protocols). So

### 2024-10-14 esop25/lang.tex
checkpoint

@@
 of $\prog$ given any initial memory, possibly augmented with a
-preprocessed memory (a memory that satsfies a predicate $\preproc$),
+preprocessed memory (a memory that satisfies a predicate $\preproc$),
 and $\aruns(\prog)$ is the analog in the malicious setting. Since we

### 2024-10-14 esop25/metalang.tex
checkpoint

@@
 We generalize these benefits of compositional verification by allowing
-pre- and post-condition annotations on $\metaprot$ functions. For
+pre-~and postcondition annotations on $\metaprot$ functions. For
 example, consider a GMW-style, 2-party ``and-gate'' function
@@
 similarly to the $\ttt{encode}$ example, we can verify this
-post-condition once as a correctness property for $\ttt{andgmw}$, and
+postcondition once as a correctness property for $\ttt{andgmw}$, and
 then integrate instances of it into circuit correctness properties
@@
 wrt confidentiality typing, $R$ are the one-time-pads consumed in the
-confidentialty typing, and $\Delta$ is the sound integrity type
+confidentiality typing, and $\Delta$ is the sound integrity type
 environment.
@@
 where $y_1,\ldots,y_n$ range over values $\mv$ and each of
-$\notg{\eqs_1}$ etc.~may contain expressions with free occurences
+$\notg{\eqs_1}$ etc.~may contain expressions with free occurrences
 of $y_1,\ldots,y_n$-- the syntax of these forms is in Figure

### 2024-10-14 esop25/model.tex
checkpoint

@@
 \end{definition}
-Integrity plus noninteference modulo output implies malicious security
+Integrity plus noninterference modulo output implies malicious security
 in the real/ideal model \cite{skalka-near-ppdp24}, and integrity plus

### 2024-10-14 esop25/smt.tex
checkpoint

@@
 
-  \toeq{\elab{\be_1 \fplus \be_2}{\cid}} = \toeq{\elab{\be_2}{\cid}} \fplus \toeq{\elab{\be_1}{\cid}}
+  \toeq{\elab{\be_1 \fplus \be_2}{\cid}} = (\toeq{\elab{\be_1}{\cid}} \fplus \toeq{\elab{\be_2}{\cid}})
 
-  \toeq{\elab{\be_1 \fminus \be_2}{\cid}} = \toeq{\elab{\be_2}{\cid}} \fminus \toeq{\elab{\be_1}{\cid}}
+  \toeq{\elab{\be_1 \fminus \be_2}{\cid}} = (\toeq{\elab{\be_1}{\cid}} \fminus \toeq{\elab{\be_2}{\cid}})
 
-  \toeq{\elab{\be_1 \ftimes \be_2}{\cid}} = \toeq{\elab{\be_2}{\cid}} \ftimes \toeq{\elab{\be_1}{\cid}}
+  \toeq{\elab{\be_1 \ftimes \be_2}{\cid}} = (\toeq{\elab{\be_1}{\cid}} \ftimes \toeq{\elab{\be_2}{\cid}})
 \end{mathpar}
 \begin{mathpar}
-  \toeq{\xassign{x}{\be}{\cid}} = x \eop \toeq{\elab{\be}{\cid}}
+  \toeq{\xassign{x}{\be}{\cid}} = (x \eop \toeq{\elab{\be}{\cid}})
 
-  \toeq{\elab{\assert{\be_1 = \be_2}}{\cid}} =  \toeq{\elab{\be_1}{\cid}} \eop \toeq{\elab{\be_2}{\cid}}
+  \toeq{\elab{\assert{\be_1 = \be_2}}{\cid}} =  (\toeq{\elab{\be_1}{\cid}} \eop \toeq{\elab{\be_2}{\cid}})
 
-  \toeq{\prog_1;\prog_2} = \toeq{\prog_1} \wedge \toeq{\prog_2} 
+  \toeq{\prog_1;\prog_2} = (\toeq{\prog_1} \wedge \toeq{\prog_2})
 \end{mathpar}}
@@
 
-The mapping $\toeq{\cdot}$ from programs $\prog$ to contraints
-is defined in Figure \ref{fig-toeq}. The interpretion of OT is ommitted
-from this figure which is general-- for $\mathbb{F}_2$ the interpretation
-is:
+The mapping $\toeq{\cdot}$ from programs $\prog$ to constraints is
+defined in Figure \ref{fig-toeq}. The interpretation of OT is omitted
+from this figure which is general. For $\mathbb{F}_2$ (where we allow
+negation $\neg$ of expressions) the interpretation is:
 \begin{mathpar}

### 2025-01-26 esop25/examples.tex
checkpoint

@@
 gate definition shows how a programmer hint can express the relevant
-confidentiality property of the output share $\mx{z}{2}$ on client
-2. The non-trivial equivalence can be verified by SMT once during
-verification, and subsequently confidentiality is expressed in the
-dependent type of the function (as for $\ttt{encode}$).
+confidentiality property of the output share $\mx{z}{2}$ on client 2,
+using the $\ttt{as}$ casting. The non-trivial equivalence can be
+verified by SMT once during verification, and subsequently
+confidentiality is expressed in the dependent type of the function (as
+for $\ttt{encode}$).
 
@@
 which is authenticated via the $\macbdoz{w}$ check (in \ttt{\_open}),
-and then each party reconstructs $\macgv{\mesg{w}}$.
+and then each party reconstructs $\macgv{\mesg{w}}$. The preconditions
+of $\ttt{open}$ specify the authentication requirements.
 
-The $\ttt{sum}$ function implements an addition gate. Note that this
-is done non-interactively-- each party just sums its local shares
+The $\ttt{sum}$ function implements an addition gate. This is
+non-interactive-- each party just sums its local shares
 of the two values. The pre- and postcondition annotations
@@
 $$}
-Given this, a malicious secure opening of $\sx{\ttt{"x"}}{1} +
+Given these global preconditions, a malicious secure opening of $\sx{\ttt{"x"}}{1} +
 \sx{\ttt{"y"}}{2}$ would be obtained as
-$\ttt{sum}(\ttt{"z"},\ttt{"x"},\ttt{"y"}); \ttt{open}(\ttt{"z"})$.
+$\ttt{sum}(\ttt{"z"},\ttt{"x"},\ttt{"y"}); \ttt{open}(\ttt{"z"})$,
+which type checks.
 
@@
 triple, so for example the share of the $a$ value for a gate
-$\ttt{"g1"}$ is identified by $\ttt{"g1as"}$, etc., and $\eqspre$
-subsumes the following constraints for the $a,b,c$ values of gate
-$\ttt{"g1"}$.  {\footnotesize$$
-\begin{array}{l}
-\mx{\ttt{"g1as"}}{1} \eop \rx{\ttt{"g1as"}}{1}\ \wedge 
-\mx{\ttt{"g1bs"}}{1} \eop \rx{\ttt{"g1bs"}}{1}\ \wedge\\
-\mx{\ttt{"g1as"}}{2} \eop \rx{\ttt{"g1as"}}{2}\ \wedge 
-\mx{\ttt{"g1bs"}}{2} \eop \rx{\ttt{"g1bs"}}{2}\ \wedge\\
-\mx{\ttt{"g1cs"}}{1} \eop \\
-\qquad ((\mx{\ttt{"g1as"}}{1} \fplus \mx{\ttt{"g1bs"}}{2})\ \ftimes\\
-\qquad\phantom{(}(\mx{\ttt{"g1bs"}}{1} \fplus \mx{\ttt{"g1bs"}}{2})) \fminus \rx{\ttt{"g1cs"}}{2}\ \wedge\\
-\mx{\ttt{"g1cs"}}{2} \eop \rx{\ttt{"g1cs"}}{2}
-\end{array}
-$$}
+$\ttt{"g1"}$ is identified by $\ttt{"g1as"}$, etc.
+
+%, and $\eqspre$
+%subsumes the following constraints for the $a,b,c$ values of gate
+%$\ttt{"g1"}$.  {\footnotesize$$
+%\begin{array}{l}
+%\mx{\ttt{"g1as"}}{1} \eop \rx{\ttt{"g1as"}}{1}\ \wedge 
+%\mx{\ttt{"g1bs"}}{1} \eop \rx{\ttt{"g1bs"}}{1}\ \wedge\\
+%\mx{\ttt{"g1as"}}{2} \eop \rx{\ttt{"g1as"}}{2}\ \wedge 
+%\mx{\ttt{"g1bs"}}{2} \eop \rx{\ttt{"g1bs"}}{2}\ \wedge\\
+%\mx{\ttt{"g1cs"}}{1} \eop \\
+%\qquad ((\mx{\ttt{"g1as"}}{1} \fplus \mx{\ttt{"g1bs"}}{2})\ \ftimes\\
+%\qquad\phantom{(}(\mx{\ttt{"g1bs"}}{1} \fplus \mx{\ttt{"g1bs"}}{2})) \fminus \rx{\ttt{"g1cs"}}{2}\ \wedge\\
+%\mx{\ttt{"g1cs"}}{2} \eop \rx{\ttt{"g1cs"}}{2}
+%\end{array}
+%$$}
 
 The definition of the $\ttt{mult}$ gate is presented in Figure
-\ref{fig-bdozmult}, where we present just client 1's side of the
-protocol $\ttt{\_mult1}$ (client 2's side is nearly symmetric but with
-a small variation-- refer to
-\cite{evans2018pragmatic,10.1007/978-3-030-68869-1_3} for more
-details).  We could call this function on input secrets
-$\sx{\ttt{"x"}}{1}$ and $\sx{\ttt{"y"}}{2}$ in gate $\ttt{"g1"}$ as
-$\ttt{mult}(\ttt{"g1"},\ttt{"x"},\ttt{"y"})$, for example, or embed
-this gate internally in a circuit.
-
-As for $\ttt{\_sum}$, the postcondition of the $\ttt{\_mult1}$
-function expresses the relevant semi-homomorphic encryption property of
+\ref{fig-bdozmult},
+%where we present just client 1's side of the
+%protocol $\ttt{\_mult1}$ (client 2's side is nearly symmetric but with
+%a small variation-- refer to
+%\cite{evans2018pragmatic,10.1007/978-3-030-68869-1_3} for more
+%details).
+%We could call this function on input secrets
+%$\sx{\ttt{"x"}}{1}$ and $\sx{\ttt{"y"}}{2}$ in gate $\ttt{"g1"}$ as
+%$\ttt{mult}(\ttt{"g1"},\ttt{"x"},\ttt{"y"})$, for example, or embed
+%this gate internally in a circuit.
+%
+As for $\ttt{sum}$, the preconditions of $\ttt{mult}$ express
+the expected MAC properties of input shares, as well as the
+expected Beaver Triple property, and its postcondition
+fexpresses the semi-homomorphic encryption property of
 the resulting share, MAC, and key after gate execution-- specifically,
-it preserves the BDOZ authentication property. The precondition that
-input shares must possess this property is enforced by the two calls
-to $\ttt{sum}$ within the body of $\ttt{mult}$ and need not be made
-explicit. Finally, the postcondition of $\ttt{mult}$ expresses the
+it preserves the BDOZ authentication property. Finally, the
+postcondition of $\ttt{mult}$ expresses the
 correctness property of the multiplication gate. In any case,
@@
 $\ttt{mult}$ library functions will require little SMT overhead
-due to compositional verification.
+due to compositional verification, once their pre- and postconditions
+are verified by type checking.
 

### 2025-01-26 esop25/intro.tex
checkpoint

@@
 \cite{li2021privacy,knott2021crypten,koch2020privacy,liu2020privacy}
-blockchains
+and blockchains
 \cite{ishai2009zero,lu2019honeybadgermpc,gao2022symmeproof,tomaz2020preserving}.
@@
 \cite{8429300,10.1145/3453483.3454074,skalka-near-ppdp24} to support
-automated language-based proof methods. And recent research in the SMT
+automated language-based proof methods. Recent research in the SMT
 community has developed theories of finite fields \cite{SMFF} with
@@
 prior work has considered automated verification of high-level
-protocols and manual verification of low-level protocols.  We
-summarize this comparison in Figure \ref{fig-comp-wrap}.
+protocols and manual verification of low-level protocols.
+We summarize this comparison in Figure \ref{fig-comp-wrap}.
 
@@
 fields and whole program analysis for larger circuits (protocols). So
-our work is a significant advancement of $\minicat/\metaprot$.  Out
+our work is a significant advancement of $\minicat/\metaprot$.  Our
 work is also closely related to Satisfiability Modulo Finite Fields
@@
 that are critical to MPC. Lilac \cite{li2023lilac} extends this line
-of work with formalisms for reasoning about conditional independence,
-which is critical for MPC \cite{skalka-near-ppdp24}.
+of work with formalisms for conditional independence which is 
+also particularly important for MPC \cite{skalka-near-ppdp24}.

### 2025-01-26 esop25/lang.tex
checkpoint

@@
 So for example, the set of all random tapes for
-for a protocol $\prog$ is $\mems(\flips(\prog))$, and the memory $\store_{\secrets(\prog)}$
+a protocol $\prog$ is $\mems(\flips(\prog))$, and the memory $\store_{\secrets(\prog)}$
 is $\store$ restricted to the secrets in $\prog$.
@@
 In the malicious model we assume that corrupt clients are in
-thrall to an adversary $\adversary$ who does not necessarily follow
+thrall to an adversary $\adversary$ that does not necessarily follow
 the rules of the protocol.  We model this by positing a $\arewrite$
@@
 in Figure \ref{fig-minicat-aredx} as a mapping from configurations to
-$\stores$ or $\bot$.
+a store $\store$ or $\bot$.
 

### 2025-01-26 esop25/metalang.tex
checkpoint

@@
 $\metaprot$ language is a \emph{metalanguage} aka metaprogramming
-language, where a $\fedprot$ protocols is the result of
+language, where an $\fedprot$ protocol is the result of
 computation. In addition to these declarative benefits of $\metaprot$,
@@
 values. $\metaprot$ \emph{instruction} forms allow dynamic
-construction of $\minicat$ protocols $\prog$, incorporating expression
+construction of $\minicat$ protocols $\prog$ that incorporate expression
 evaluation. The syntax also supports definitions of functions that
@@
 
+\mtjfnfig
+
 We allow the programmer to specify a type signature $\tsig$ that

### 2025-01-26 esop25/model.tex
checkpoint

@@
   \begin{cases}
-    0 \text{\ if\ } \margd{\pmf}{X_2}(\store_{X_2}) = 0\\
-    \margd{\pmf}{X_1 \cup X_2}(\store) / \margd{\pmf}{X_2}(\store_{X_2})\ \text{\ o.w.}
+    \begin{array}{ll}
+      0 & \text{\ if\ } \margd{\pmf}{X_2}(\store_{X_2}) = 0\\
+      \margd{\pmf}{X_1 \cup X_2}(\store) / \margd{\pmf}{X_2}(\store_{X_2})\ & \text{\ o.w.}
+     \end{array}
   \end{cases}
@@
 
-Here we restate hyperproperties developed in previous work
-\cite{skalka-near-ppdp24}, to which the reader is referred for a more
-thorough discussion of their relation to real/ideal security. Using
-hyperproperties for security semantics eases proofs of
-correctness for our type systems.
+In this section we restate hyperproperties developed in previous work
+\cite{skalka-near-ppdp24}. The reader is referred to that paper for a more
+thorough discussion of their relation to real/ideal security. In this
+paper, we leverage that relation to establish security properties in
+our type system.
 

### 2025-01-26 esop25/smt.tex
checkpoint

@@
   \eqs_2$ iff  $\store \models E_1$ implies $\store \models
-  E_2$ for all $\store$, and note this relation is reflexive and transitive.
+  E_2$ for all $\store$, and note that this relation is a preorder.
 \end{definition}
@@
 solver to verify properties relevant to correctness, confidentiality,
-and integrity. Further, we can leverage entailment relation is critical for efficiency--
+and integrity. Further, we can leverage entailment for tractability--
 we can use annotations to obtain a weakened precondition for relevant properties.

### 2025-01-26 esop25/intro.tex
checkpoint

@@
 field arithmetic. The goal of this paper is to combine these two
-approaches in an automated type system for verifying correctness and
+approaches in a decidable type system for verifying correctness and
 security properties of MPC protocols.
@@
 structures \cite{10.1145/3498713}, obliviousness is related to pure
-noninterference, not the relaxed form related to passive MPC
-security. The Haskell-based security type system in \cite{6266151}
-enforces a version of noninterference that is sound for passive
-security, but does not consider malicious security. And properties of
-real/ideal passive and malicious security for a probabilistic language
-have been (manually) formulated in EasyCrypt \cite{8429300}.
+noninterference, not the conditional form related to passive MPC
+security (Definition \ref{definition-NIMO}). The Haskell-based security
+type system in \cite{6266151} enforces a version of noninterference
+that is sound for passive security, but does not consider malicious
+security. And properties of real/ideal passive and malicious security
+for a probabilistic language have been (manually) formulated in
+EasyCrypt \cite{8429300}.
 

### 2025-01-26 esop25/model.tex
checkpoint

@@
 functionality} $\idealf$ with per-client outputs. In the $\minifed$
-setting, Given a protocol $\prog$ that implements $\idealf$, with
+setting, given a protocol $\prog$ that implements $\idealf$, with
 $\iov(\prog) = S \cup M \cup P \cup O$, the domain of $\idealf$ is

### 2025-01-26 esop25/smt.tex
checkpoint

@@
 \subsubsection{Example: Correctness of 3-Party Addition}
-Consider the following 3-party additive secret sharing protocol.
-{\footnotesize
+Consider the following 3-party additive secret sharing protocol, where
+each party generates two additive secret shares
+\cite{evans2018pragmatic} for their respective secrets. Each party
+publicly reveals the sums all their shares, and the sum of those is
+the protocol output.  {\footnotesize
 $$

### 2025-01-26 esop25/smt.tex
checkpoint

@@
 publicly reveals the sums all their shares, and the sum of those is
-the protocol output.  {\footnotesize
+the protocol output for each party.  {\footnotesize
 $$
@@
 Letting $\prog$ be this protocol, we can verify correctness
-as follows-- the output is the sum of the three input secrets:
+as follows-- that is, the output sum of all shares is the sum of the
+three input secrets:
 $$

### 2025-01-26 esop25/metalang.tex
checkpoint

@@
 
-To obtain high precision in the type system, and to generalize
-functions typings, we also introduce a form of type dependency,
+To obtain a more complete type system, and to generalize
+function typings, we also introduce a form of type dependency,
 specifically dependence on $\metaprot$ expressions. Dependent

### 2025-01-29 esop25/camera-ready/conclusion.tex
checkpoint

\section{Conclusion and Future Work}

In this paper we have defined confidentiality and integrity type
systems for the $\metaprot/\minifed$ language model. These type
systems resemble previous security type systems in probabilistic
settings, but their accuracy is boosted by the integral use of SMT
verification.  Our work significantly improves automation,
scalability, and flexibility of static analysis in the
$\metaprot/\minifed$ as compared to prior work.

We see several promising avenues for future work. Concurrency will be
important to capture common MPC idioms such as commitment and circuit
optimizations and to consider the UC security model
\cite{evans2018pragmatic,viaduct-UC}. We also hope to improve our type
system to enforce the declassification bounds determined by ideal
functionalities, beyond gradual release.

### 2025-01-29 esop25/camera-ready/examples.tex
checkpoint

--- 2025-01-26 esop25/examples.tex
@@
-\section{Extended Examples}
+\section{Extended Example: GMW}
 \label{section-examples}
@@
 methods in $\fieldp{2}$ \cite{skalka-near-ppdp24}.
-
-\subsection{Confidentiality Examples}
 
@@
 for any application.
-
-\subsection{Integrity Examples}
-
-\input{bdoz}
-
-Returning to the example of malicious-secure 2-party BDOZ arithmetic
-circuits begun in Section \ref{section-ipj}, in Figure \ref{fig-bdozsum} we
-define $\ttt{sum}$ and $\ttt{open}$ functions. The latter implements
-``secure opening''-- each party sends it's local share of a
-global value $\macgv{\mesg{w}}$ along with its MAC to the other party,
-which is authenticated via the $\macbdoz{w}$ check (in \ttt{\_open}),
-and then each party reconstructs $\macgv{\mesg{w}}$. The preconditions
-of $\ttt{open}$ specify the authentication requirements.
-
-The $\ttt{sum}$ function implements an addition gate. This is
-non-interactive-- each party just sums its local shares
-of the two values. The pre- and postcondition annotations
-on $\ttt{sum}$ express the additive homomorphism associated
-with this encryption scheme-- the sum of MACs of the input
-shares is a valid MAC for the output share on each client,
-which can be checked using the sum of keys of the input shares.
-
-In BDOZ a pre-processing phase is assumed where initial input secrets
-are shared along with their associated MACs and keys. This can be
-expressed in $\eqspre$ for input secrets $\sx{\ttt{"x"}}{1}$ and
-$\sx{\ttt{"y"}}{2}$, for example, which subsume the following
-constraints on shares:
-{\footnotesize$$
-\begin{array}{l}
-\mx{\ttt{"xs"}}{2} \eop \sx{\ttt{"x"}}{1} \fminus \rx{\ttt{"x"}}{1} \wedge 
-\mx{\ttt{"xs"}}{1} \eop \rx{\ttt{"x"}}{1} \wedge \\
-\mx{\ttt{"ys"}}{1} \eop \sx{\ttt{"y"}}{2} \fminus \rx{\ttt{"y"}}{2} \wedge 
-\mx{\ttt{"ys"}}{2} \eop \rx{\ttt{"y"}}{2} 
-\end{array}
-$$}
-and the following constraint on keys and MACs for authentication
-of $\sx{\ttt{"x"}}{1}$ (and similarly for $\sx{\ttt{"y"}}{2}$):
-{\footnotesize$$
-\begin{array}{l}
-\mx{\ttt{"delta"}}{1} \eop \rx{\ttt{"delta"}}{1} \wedge
-\mx{\ttt{"xk"}}{1} \eop \rx{\ttt{"xk"}}{1} \wedge\\
-\mx{\ttt{"xm"}}{2} \eop \mx{\ttt{"xk"}}{1} \fplus (\mx{\ttt{"delta"}}{1} * \mx{\ttt{"xs"}}{2})
-\end{array}
-$$}
-Given these global preconditions, a malicious secure opening of $\sx{\ttt{"x"}}{1} +
-\sx{\ttt{"y"}}{2}$ would be obtained as
-$\ttt{sum}(\ttt{"z"},\ttt{"x"},\ttt{"y"}); \ttt{open}(\ttt{"z"})$,
-which type checks.
-
-A common approach to implementing multiplication gates in a BDOZ
-setting is to use \emph{Beaver Triples}. Recall that Beaver triples
-are values $a,b,c$ with $a$ and $b$ chosen randomly and $c = a * b$,
-unique per multiplication gate, that are secret shared with clients
-during pre-processing.  In our encoding we assume the additional
-convention that each gate output identifier distinguishes the Beaver
-triple, so for example the share of the $a$ value for a gate
-$\ttt{"g1"}$ is identified by $\ttt{"g1as"}$, etc.
-
-%, and $\eqspre$
-%subsumes the following constraints for the $a,b,c$ values of gate
-%$\ttt{"g1"}$.  {\footnotesize$$
-%\begin{array}{l}
-%\mx{\ttt{"g1as"}}{1} \eop \rx{\ttt{"g1as"}}{1}\ \wedge 
-%\mx{\ttt{"g1bs"}}{1} \eop \rx{\ttt{"g1bs"}}{1}\ \wedge\\
-%\mx{\ttt{"g1as"}}{2} \eop \rx{\ttt{"g1as"}}{2}\ \wedge 
-%\mx{\ttt{"g1bs"}}{2} \eop \rx{\ttt{"g1bs"}}{2}\ \wedge\\
-%\mx{\ttt{"g1cs"}}{1} \eop \\
-%\qquad ((\mx{\ttt{"g1as"}}{1} \fplus \mx{\ttt{"g1bs"}}{2})\ \ftimes\\
-%\qquad\phantom{(}(\mx{\ttt{"g1bs"}}{1} \fplus \mx{\ttt{"g1bs"}}{2})) \fminus \rx{\ttt{"g1cs"}}{2}\ \wedge\\
-%\mx{\ttt{"g1cs"}}{2} \eop \rx{\ttt{"g1cs"}}{2}
-%\end{array}
-%$$}
-
-The definition of the $\ttt{mult}$ gate is presented in Figure
-\ref{fig-bdozmult},
-%where we present just client 1's side of the
-%protocol $\ttt{\_mult1}$ (client 2's side is nearly symmetric but with
-%a small variation-- refer to
-%\cite{evans2018pragmatic,10.1007/978-3-030-68869-1_3} for more
-%details).
-%We could call this function on input secrets
-%$\sx{\ttt{"x"}}{1}$ and $\sx{\ttt{"y"}}{2}$ in gate $\ttt{"g1"}$ as
-%$\ttt{mult}(\ttt{"g1"},\ttt{"x"},\ttt{"y"})$, for example, or embed
-%this gate internally in a circuit.
-%
-As for $\ttt{sum}$, the preconditions of $\ttt{mult}$ express
-the expected MAC properties of input shares, as well as the
-expected Beaver Triple property, and its postcondition
-fexpresses the semi-homomorphic encryption property of
-the resulting share, MAC, and key after gate execution-- specifically,
-it preserves the BDOZ authentication property. Finally, the
-postcondition of $\ttt{mult}$ expresses the
-correctness property of the multiplication gate. In any case,
-integrity of any circuit constructed from the $\ttt{sum}$ and
-$\ttt{mult}$ library functions will require little SMT overhead
-due to compositional verification, once their pre- and postconditions
-are verified by type checking.
-
-%%%%%%%%%%%%%%%%%%
-%%% ARCHIVAL BELOW
-%%%%%%%%%%%%%%%%%%
-\begin{comment}
-\begin{verbatimtab}
-andtableygc(g,x,y)
-{
-   let table = (~r[g],~r[g],~r[g],r[g])
-   in permute4(r[x],r[y],table)
-}
-
-m[x]@1 := s2(s[x],r[x],~r[x])@2;
-m[x]@1 as s[x]@2 xor r[x]@2;
-
-// m[x]@1 : { c(r[x]@2, { s[x]@2 }) }
-
-m[y]@1 := OT(s[y]@1,r[y],~r[y])@2;
-m[y]@1 as s[y]@1 xor r[y]@2;
-
-// m[y]@1 : { c(r[y]@2, { s[y]@1 }) }
-
-m[ag]@1 := OT4(m[x]@1, m[y]@1, andtable(ag,r[x],r[y]))@2;
-m[ag]@1 as  ~((r[x]@2 = m[x]@1) and (r[y]@2 = m[y]@1)) xor r[ag]@2;
-
-// m[ag]@1 : { c(r[ag]@2, {r[x]@2, r[y]@2, m[x]@1,  m[y]@1} }
-
-p[o] := OT2(m[ag]@1, perm2(r[ag],(false,true)))@2
-
-// p[o] : { c(r[ag]@2, {r[x]@2, r[y]@2,  m[x]@1,  m[y]@1}), r[ag]@2  }
-
-out@1 := p[o]@1
-
-// out@1 == s[x] and s[y]
-\end{verbatimtab}
-
-\begin{verbatimtab}
-    encodegmw(in, i1, i2) {
-      m[in]@i2 := (s[in] xor r[in])@i1;
-      m[in]@i1 := r[in]@i1
-    }
-
-    andtablegmw(x, y, z) {
-      let r11 = r[z] xor (m[x] xor true) and (m[y] xor true) in
-      let r10 = r[z] xor (m[x] xor true) and (m[y] xor false) in
-      let r01 = r[z] xor (m[x] xor false) and (m[y] xor true) in
-      let r00 = r[z] xor (m[x] xor false) and (m[y] xor false) in
-      { row1 = r11; row2 = r10; row3 = r01; row4 = r00 }
-    }
-
-    andgmw(z, x, y) {
-      let table = andtablegmw(x,y,z) in
-      m[z]@2 := OT4(m[x],m[y],table,2,1);
-      m[z]@2 as ~((m[x]@1 xor m[x]@2) and (m[y]@1 xor m[y]@2)) xor r[z]@1);
-      m[z]@1 := r[z]@1
-    }
-
-    // and gate correctness postcondition
-    {} andgmw { m[z]@1 xor m[z]@2 == (m[x]@1 xor m[x]@2) and (m[y]@1 xor m[y]@2) }
-
-    // and gate type
-    andgmw :
-     Pi z,x,y .
-     {}
-     { { r[z]@1 },
-       (m[z]@1 : { r[z]@1 }; m[z]@2 : {c(r[z]@1, { m[x]@1, m[x]@2, m[y]@1, m[y]@2 })} ),
-       m[z]@1 xor m[z]@2 == (m[x]@1 xor m[x]@2) and (m[y]@1 xor m[y]@2)}
-
-    xorgmw(z, x, y) {
-      m[z]@1 := (m[x] xor m[y])@1; m[z]@2 := (m[x] xor m[y])@2;
-    }
-
-    decodegmw(z) {
-      p["1"] := m[z]@1; p["2"] := m[z]@2;
-      out@1 := (p["1"] xor p["2"])@1;
-      out@2 := (p["1"] xor p["2"])@2
-    }
-
-    prot() {
-      encodegmw("x",2,1);
-      encodegmw("y",2,1);
-      encodegmw("z",1,2);
-      andgmw("g1","x","z");
-      xorgmw("g2","g1","y");
-      decodegmw("g2")
-    }
-
-    {} prot { out@1 == (s["x"]@1 and s["z"]@2) xor s["y"]@1 }
-\end{verbatimtab}
-
-\subsection{Integrity Examples}
-
-\begin{verbatimtab}
-  secopen(w1,w2,w3,i1,i2) {
-      pre(m[w1++"m"]@i2 == m[w1++"k"]@i1 + (m["delta"]@i1 * m[w1++"s"]@i2 /\
-          m[w1++"m"]@i2 == m[w1++"k"]@i1 + (m["delta"]@i1 * m[w1++"s"]@i2));
-      let locsum =  macsum(macshare(w1), macshare(w2)) in
-      m[w3++"s"]@i1 := (locsum.share)@i2;
-      m[w3++"m"]@i1 := (locsum.mac)@i2;
-      auth(m[w3++"s"],m[w3++"m"],mack(w1) + mack(w2),i1);
-      m[w3]@i1 := (m[w3++"s"] + (locsum.share))@i1
-  }
-
-  _open(x,i1,i2){
-    m[x++"exts"]@i1 := m[x++"s"]@i2;
-    m[x++"extm"]@i1 := m[x++"m"]@i2;
-    assert(m[x++"extm"] == m[x++"k"] + (m["delta"] * m[x++"exts"]));
-    m[x]@i1 := (m[x++"exts"] + m[x++"s"])@i2
-  }`
-
-  _sum(z, x, y,i1,i2) {
-      pre(m[x++"m"]@i2 == m[x++"k"]@i1 + (m["delta"]@i1 * m[x++"s"]@i2 /\
-          m[y++"m"]@i2 == m[y++"k"]@i1 + (m["delta"]@i1 * m[y++"s"]@i2));
-      m[z++"s"]@i2 := (m[x++"s"] + m[y++"s"])@i2;
-      m[z++"m"]@i2 := (m[x++"m"] + m[y++"m"])@i2;
-      m[z++"k"]@i1 := (m[x++"k"] + m[y++"k"])@i1;
-      post(m[z++"m"]@i2 == m[z++"k"]@i1 + (m["delta"]@i1 * m[z++"s"]@i2)
-  }
-
-  sum(z,x,y) { _sum(z,x,y,1,2);_sum(z,x,y,2,1) }
-
-  open(x) { _open(x,1,2); _open(x,2,1) }
-
-  sum("a","x","d");
-  open("d");
-  sum("b","y","e");
-  open("e");
-  let xys =
-      macsum(macctimes(macshare("b"), m["d"]),
-             macsum(macctimes(macshare("a"), m["e"]),
-                    macshare("c")))
-  let xyk = mack("b") * m["d"] + mack("a") * m["e"] + mack("c")
-
-  secopen("a","x","d",1,2);
-    secopen("a","x","d",2,1);
-    secopen("b","y","e",1,2);
-    secopen("b","y","e",2,1);
-    let xys =
-      macsum(macctimes(macshare("b"), m["d"]),
-             macsum(macctimes(macshare("a"), m["e"]),
-                    macshare("c")))
-    in
-    let xyk = mack("b") * m["d"] + mack("d") * m["d"] + mack("c")               
-    in
-    secreveal(xys,xyk,"1",1,2);
-    secreveal(maccsum(xys,m["d"] * m["e"]),
-              xyk - m["d"] * m["e"],
-              "2",2,1);
-    out@1 := (p[1] + p[2])@1;
-    out@2 := (p[1] + p[2])@2;
-\end{verbatimtab}
-
-\end{comment}

### 2025-01-29 esop25/camera-ready/examples.tex
checkpoint

@@
-\section{Extended Example: GMW}
+\section{Examples and Evaluation}
 \label{section-examples}
@@
 
+\paragraph{GMW Input Encoding and And Gate}
 In the 2-party GMW protocol \cite{evans2018pragmatic}, another garbled

### 2025-01-29 esop25/camera-ready/intro.tex
checkpoint

--- 2025-01-26 esop25/intro.tex
@@
 protocols and manual verification of low-level protocols.
-We summarize this comparison in Figure \ref{fig-comp-wrap}.
+%We summarize this comparison in Figure \ref{fig-comp-wrap}.
 
@@
 
-\compwrapfig
+% \compwrapfig
 
@@
 also particularly important for MPC \cite{skalka-near-ppdp24}.
+
+The Cryptol and SAW tools \cite{10.1007/978-3-319-48869-1_5} allow
+programmers to program security protocols that are verified using
+SMT. However they are used for cryptographic protocols more generally,
+and are not designed specifically for MPC as our framework is.

### 2025-01-29 esop25/camera-ready/smt.tex
checkpoint

--- 2025-01-26 esop25/smt.tex
@@
 Consider the following 3-party additive secret sharing protocol, where
-each party generates two additive secret shares
-\cite{evans2018pragmatic} for their respective secrets. Each party
-publicly reveals the sums all their shares, and the sum of those is
-the protocol output for each party.  {\footnotesize
+each party generates three additive, information theoretically secure
+secret shares by subtracting random field values used as
+one-time-pads. Two shares are communicated to the other clients and
+one remains local. This technique is information-theoretically secure
+\cite{evans2018pragmatic} due to the one-time use of random
+values-- that is, $< 3$ shares reveal nothing about the
+secret, but the sum of all three \emph{is} the secret. Each party
+publicly reveals the sums all their shares and then outputs the sum of the reveals.
+{\footnotesize
 $$

### 2025-01-30 esop25/camera-ready/intro.tex
checkpoint

@@
 provided by Satisfiability Modulo Finite Fields \cite{SMFF}. In
-Section \ref{section-examples} we develop and discuss extended example
-applications of our type analyses to real protocols including
-the Goldreich-Micali-Wigderson (GMW), Bendlin-Damgard-Orland-Zakarias
-(BDOZ), and Yao's Garbled Circuits (YGC) protocols.
+Section \ref{section-examples} we develop and discuss evaluation
+and implementation of the the Goldreich-Micali-Wigderson (GMW)
+protocol. In an extended version of this paper we also discuss
+the Bendlin-Damgard-Orland-Zakarias
+(BDOZ) protocol \cite{skalka2025smtboostedsecuritytypeslowlevel}.
 

### 2025-01-30 esop25/camera-ready/ipj.tex
checkpoint

--- 2024-10-14 esop25/ipj.tex
@@
 depends on data from client 2. The rule for MAC
-checking allows an overwrite of the dependency. We explore
-the application of this in more detail in Section
-\ref{section-examples}.
+checking allows an overwrite of the dependency. An implementation
+of a malicious-secure arithmetic circuit that
+applies this rule is provided in extended version of
+this paper \ref{skalka2025smtboostedsecuritytypeslowlevel}.
 
@@
 In all cases, we want to determine the integrity level of computed
-values based on assumed partitionins $H,C$. In the previous example,
+values based on assumed partitionings $H,C$. In the previous example,
 we want to say that $\rvl{w}$ has low integrity if either client 1 or
@@
 Our main type soundness result for integrity is then given as follows.
+Proof details are given in \cite{skalka2025smtboostedsecuritytypeslowlevel}.
 \begin{theorem}

### 2025-01-30 esop25/camera-ready/metalang.tex
checkpoint

--- 2025-01-26 esop25/metalang.tex
@@
 types, allowing for significant reduction of SMT verification in whole
-program analysis, as we will discuss with extended examples in Section
-\ref{section-examples}.
+program analysis, as we illustrate with extended examples in Section
+\ref{section-examples} and in an extended version of this paper
+\cite{skalka2025smtboostedsecuritytypeslowlevel}.
 
@@
 code. This ensures generality of verification as demonstrated in Lemma
-\ref{lemma-eqs-notg} in the Appendix.
+\ref{lemma-eqs-notg} in an extended version of this paper \cite{skalka2025smtboostedsecuritytypeslowlevel}.
 
@@
 pre-~and postconditions, as we illustrate in Section
-\ref{section-examples}, since the rest of the type is reconstructed by
-type checking.
+\ref{section-examples} and \cite{skalka2025smtboostedsecuritytypeslowlevel},
+since the rest of the type is reconstructed by type checking.
 
@@
 security hyperproperties of confidentiality and integrity, $\metaprot$
-type checking implicitly enforces them. Proof details are given in the
-Appendix.
+type checking implicitly enforces them. Proof details are given \cite{skalka2025smtboostedsecuritytypeslowlevel}.
 \begin{theorem}[Soundness of $\metaprot$ type checking]

### 2025-01-30 esop25/camera-ready/examples.tex
checkpoint

@@
 Now we consider examples that illustrate the versatility of the
-$\TirName{Encode}$ typing rule for confidentiality, the use of
-postconditions for correctness, and the use of both pre-~and
-postconditions for integrity.
+$\TirName{Encode}$ typing rule for confidentiality and the use of
+postconditions for correctness. In an extended version of this paper
+we also exemplify the use of both pre-~and
+postconditions for integrity \cite{skalka2025smtboostedsecuritytypeslowlevel}.
 
@@
 encryption properties leveraged in the BDOZ circuit library functions
-(Figures \ref{fig-bdozsum} and \ref{fig-bdozmult}) in fields up to
+defined in \cite{skalka2025smtboostedsecuritytypeslowlevel} in fields up to
 size $\fieldp{2^{31} - 1}$ which take $\sim$8ms for the property used for
@@
 
-\paragraph{GMW Input Encoding and And Gate}
+\paragraph{GMW Input Encoding and And Gate.}
 In the 2-party GMW protocol \cite{evans2018pragmatic}, another garbled
 binary circuit protocol, values are encrypted in a manner similar to
-BDOZ as described in Section \ref{section-ipj}. In our definition of
-GMW we use the convention that shared values $\macgv{\mesg{w}}$ are
-identified by strings $w$ and encoded as shares $\mx{w}{1}$ and
-$\mx{w}{2}$.  As for YGC, ciphertypes reflect the confidentiality of
-GMW input encodings as defined in the $\ttt{encodegmw}$ function in
-Figure \ref{fig-gmw}. No programmer annotation is needed given the
-syntax of the function body. More interestingly, the $\ttt{andgmw}$
-gate definition shows how a programmer hint can express the relevant
-confidentiality property of the output share $\mx{z}{2}$ on client 2,
-using the $\ttt{as}$ casting. The non-trivial equivalence can be
-verified by SMT once during verification, and subsequently
-confidentiality is expressed in the dependent type of the function (as
-for $\ttt{encode}$).
+using secret sharing as described in Sections
+\ref{section-smt-additive-sharing} and \ref{section-ipj}. In our
+definition of GMW we use the convention that shared values
+$\macgv{\mesg{w}}$ are identified by strings $w$ and encoded as shares
+$\mx{w}{1}$ and $\mx{w}{2}$.  Ciphertypes reflect the
+confidentiality of GMW input encodings as defined in the
+$\ttt{encodegmw}$ function in Figure \ref{fig-gmw}. No programmer
+annotation is needed given the syntax of the function body. More
+interestingly, the $\ttt{andgmw}$ gate definition shows how a
+programmer hint can express the relevant confidentiality property of
+the output share $\mx{z}{2}$ on client 2, using the $\ttt{as}$
+casting. The non-trivial equivalence can be verified by SMT once
+during verification, and subsequently confidentiality is expressed in
+the dependent type of the function (as for $\ttt{encode}$).
 

### 2025-01-30 esop25/camera-ready/ipj.tex
checkpoint

@@
 depends on data from client 2. The rule for MAC
-checking allows an overwrite of the dependency. An implementation
-of a malicious-secure arithmetic circuit that
-applies this rule is provided in extended version of
-this paper \ref{skalka2025smtboostedsecuritytypeslowlevel}.
+checking allows an overwrite of the dependency. We implement
+a malicious-secure arithmetic circuit that
+applies this rule in extended version of
+this paper \cite{skalka2025smtboostedsecuritytypeslowlevel}.
 

### 2025-01-30 esop25/camera-ready/metalang.tex
checkpoint

@@
 $\mv_1,\ldots,\mv_n$ are unique and not used in any program source
-code. This ensures generality of verification as demonstrated in Lemma
-\ref{lemma-eqs-notg} in an extended version of this paper \cite{skalka2025smtboostedsecuritytypeslowlevel}.
+code. This ensures generality of verification as demonstrated
+in an extended version of this paper \cite{skalka2025smtboostedsecuritytypeslowlevel}.
 

### 2025-01-30 esop25/camera-ready/smt.tex
checkpoint

@@
 \subsubsection{Example: Correctness of 3-Party Addition}
+\label{section-smt-additive-sharing}
 Consider the following 3-party additive secret sharing protocol, where

### 2025-04-24 code/cvc5mvn/src/main/java/plaid/logic/ConstraintAnalyzer.java
simplify constraint analyzer data structure and control flow

@@ -10,7 +10,7 @@ public class ConstraintAnalyzer {
     private final Program program; // provide a program to resolve function
     private final ConstraintEvaluator evaluator;
     private final GenEntailVerifier verifier;
-    private final List<Map<PreludeExpression, Constraints>> functionConstraints; // data structure to store inferred pre/post conditions by FN rule
+    private final Map<PreludeExpression, Constraints> functionConstraints = new HashMap<>(); // data structure to store inferred pre/post conditions by FN rule
     
     private Map<Identifier, PreludeExpression> binding(List<TypedIdentifier> formal, List<PreludeExpression> actual){
         // for each actual parameter, bind it to corresponding formal parameter 
@@ -25,14 +25,12 @@ public class ConstraintAnalyzer {
         this.program = program;
         this.evaluator = new ConstraintEvaluator(program);
         this.verifier = new GenEntailVerifier(program, order);
-        functionConstraints = new LinkedList<>();
-        functionConstraints.add(new HashMap<>());
     }
 
     /**
      * calculate precondition and postcondition for function, FN rule
      */
-    public Constraints inferPrePostFN(CommandFunction function) throws CVC5ApiException{
+    public Constraints inferPrePostFN(CommandFunction function) {
         // infer the precondition and postcondition for the function using inferPrePostCmd
         Constraints constraints = inferPrePostCmd(function.c());
         System.out.println("ASDF inferred pre for : " + function.fname() + ScalaFunctions.prettyPrint(constraints.getPre()));
@@ -40,7 +38,7 @@ public class ConstraintAnalyzer {
         // if the function does not have annotated pre/post conditions
         if (function.precond() == null && function.postcond() == null ){
             // then store inferred pre/post conditions and return them
-            functionConstraints.add(Map.of(function.fname(), constraints));
+            functionConstraints.put(function.fname(), constraints);
             return constraints;
         }
         else {
@@ -57,7 +55,7 @@ public class ConstraintAnalyzer {
                     : new AndConstraintExpr(constraints.getPre(), function.postcond());
             // take typings
             if (verifier.genEntails(function.typedVariables(), pre, post)){
-                functionConstraints.add(Map.of(function.fname(), new Constraints(function.precond(), function.postcond())));
+                functionConstraints.put(function.fname(), new Constraints(function.precond(), function.postcond()));
                 return new Constraints(function.precond(), function.postcond());
             }
             else{
@@ -71,7 +69,7 @@ public class ConstraintAnalyzer {
     /**
[…]

### 2025-04-24 esop25/slides.tex
checkpoint

--- 2024-09-10 ppdp24/slides.tex
@@
 
-\input{slide-macros}
+\usepackage{mathpartir}
+\usepackage{comment}
+
+% \input{slide-macros}
+\input{macros}
 
 % Title page details: 
-\title{Language-Based Security for Low-Level MPC}
+\title{SMT-Boosted Security Types for Low-Level MPC}
 
@@
 
-\date{September 10, 2024}
+\date{ESOP, May 6, 2025}
 
@@
   \begin{itemize}
-  \item Guarantees data privacy.
-  \item Software-based protocols (vs., e.g., enclaves).
+  \item Supports data privacy.
+  \item Uses software-based protocols (vs., e.g., enclaves).
   \item No trusted 3rd party.
@@
   \begin{itemize}
-  \item Low-level protocols possess distinct features, e.g., probabilistic sampling.
-  \item Existing proof techniques manual, error prone, complex security model.
+  \item Distinct features, e.g., probabilistic sampling.
+  \item Existing proof techniques mostly manual.
   \end{itemize}
@@
   \item Language design for low-level protocol definitions.
-  \item (Re)formulation of security model as \cemph{hyperproperties} (e.g., noninterference).
-  \item Semi-automated proof techniques.
+  \item Automated type analysis for security properties, with correctness based on
+    \cemph{hyperproperties}\footnote{Skalka and Near, PPDP 2024}.
+  \item Our type analysis leverages an \cemph{SMT finite fields theory}\footnote{Ozdemir et. al, CAV 2023} for:
     \begin{itemize}
-    \item Automatically verify small program components, manually combine for whole-program proofs. 
+    \item \cemph{Correctness}
+    \item \cemph{Confidentiality}
+    \item \cemph{Integrity}
     \end{itemize}
@@
 
-\begin{frame}{The Ideal Functionality}
+\begin{frame}{What's an MPC Protocol?}
   A \emph{protocol} $\prog$ implements an \emph{ideal functionality} $\idealf$,
@@
     $\setit{1,...,j}$.
-  \item The \cemph{output} is made \emph{publicly available} by $\prog$
+  \item Clients are distributed and communicate via \cemph{message passing}.
+  \item Any client can be corrupted by the adversary, who can then see their messages.
+  \item The \cemph{output} is made \emph{publicly available} by $\prog$.
   \end{itemize}
   \cemph{The output may leak secret information to an adversary} trying to
-  guess secret values.
+  guess secret values, since it depends on inputs.
 
-  \begin{exampleblock}{Consider:}
+  \medskip
+
+  \cemph{Confidentiality} means that protocol messages don't enable the adversary
+  to make better guesses than the output alone.
+
+\end{frame}
+
+\begin{frame}{Declassification of the Ideal Functionality}
+
+  \begin{exampleblock}{Example: 3-Party $\fplus$ in $\mathbb{F}_2$ (where $\fplus$ is $\exor$ and $\ftimes$ is $\wedge$)}
   $$
-  \idealf(x,y) = x * y
+    \idealf =
+    \mathit{secret}@{\mathit{client_1}} + \mathit{secret}@{\mathit{client_2}} + 
+      \mathit{secret}@{\mathit{client_3}}
   $$
-  If the output is odd, the adversary knows both $x$ and $y$ must be, even
-  if $x$ and $y$ remain hidden.
+  Suppose in a run of $\prog$ and $\mathit{client_3}$ is
+  corrupted, with $\mathit{secret}@{\mathit{client_3}} = 0$. 
+  In case the output is 1, the adversary knows that either:
+  $$
+  \mathit{secret}@{\mathit{client_1}} = 0 \text{\ and\ } \mathit{secret}@{\mathit{client_2}} = 1
+  $$
+  or:
+  $$
+  \mathit{secret}@{\mathit{client_1}} = 1 \text{\ and\ } \mathit{secret}@{\mathit{client_2}} = 0
+  $$
+  with equal joint probability of .5.
+
+  \medskip
+
+  \cemph{Messages sent to $\mathit{client_3}$ during
+    a run of $\prog$ should not
+    allow the adversary better guesses.}
   \end{exampleblock}
@@
   \end{itemize}
-  We can localize variables and expressions to \emph{clients}, and any information
+  We localize computation to \emph{clients}, and any information
   exchanged between clients must be through messaging.
-  \begin{exampleblock}{Toy Example:}
-    $$
-    \begin{array}{rcl@{\qquad}l}
-      \mx{x}{1} &:=& \elab{(\secret{x} - \flip{x})}{2}  & \textit{(unicast)}\\
-      \mx{y}{*} &:=& \elab{(\flip{y} + \mesg{x})}{1}  & \textit{(broadcast)}
-    \end{array}
-    $$ 
-  \end{exampleblock}
-\end{frame}
-
-\begin{frame}{(Passive) Threat Model}
-  The adversary wants to guess secrets. In addition to knowing the public
-  output, the adversary may also corrupt \emph{any} subset of clients $C$:
-  \begin{itemize}
-  \item adversary has access to the secrets of clients in $C$.
-  \item adversary has access to messages received by members of $C$ during execution of $\prog$,
-    aka \cemph{adversarial views}.
-  \item adversary cannot break the rules of the protocol\footnote{Malicious security also
-  considered separately in our work.}.
-  \end{itemize}
-
-  \begin{alertblock}{Bounded Declassification}
-    The ideal functionality defines a declassification ``limit'', any protocol $\prog$
-    should not reveal any more through adversarial views.
-  \end{alertblock}
-\end{frame}
-
-\begin{frame}{Declassification of the Ideal Functionality}
-
-  \begin{exampleblock}{Example: 3-Party $\fplus$ in $\mathbb{F}_2$ (where $\fplus$ is $\exor$ and $\ftimes$ is $\wedge$)}
-  $$
-  \idealf(\sx{1}{1},\sx{2}{2},\sx{3}{3}) =
-  \sx{1}{1}\ \fplus\ \sx{2}{2}\ \fplus\ \sx{3}{3}
-  $$
-  Suppose in a run of $\prog$ the output of $\prog$ is 1 and $C = \{ 3 \}$ and
-  the adversary knows $\sx{3}{3} = 0$.
-  Therefore the adversary knows that either:
-  $$
-  \sx{1}{1} = 0 \text{\ and\ } \sx{2}{2} = 1
-  $$
-  or:
-  $$
-  \sx{1}{1} = 1 \text{\ and\ } \sx{2}{2} = 0
-  $$
-  with equal joint probability of .5. \cemph{Adversarial views from the run of $\prog$ should not
-    allow the adversary better guesses.}
-  \end{exampleblock}
-
-\end{frame}
-
-\section{Conditional Probabilistic Noninterference}
-\begin{frame}{Conditional Probabilistic Noninterference}
-
-  \begin{itemize}
-    \item The set of all protocol runs under all possible secrets and random tapes
-      induces a probability distribution of program variables.
-    \item Conditioning on the output and corrupt secrets =  declassification
-      bounds.
-    \item Conditioning on adversarial views = distribution of honest secrets.
-  \end{itemize}
-
-  \begin{alertblock}{Conditional Noninterference}
-    The probability of honest secrets conditioned on the protocol output
-    should not be changed by conditioning on adversarial views.
-
-  \begin{itemize}
-  \item Sound for the standard \emph{passive simulator security} model.
-  %\item A \emph{hyperproperty} that accommodates traditional PL-based security
-  %  mechanisms and theory. 
-  \end{itemize}
-  \end{alertblock}
-
+  %\begin{exampleblock}{Toy Example:}
+  %  $$
+  %  \begin{array}{rcl@{\qquad}l}
+  %    \mx{x}{1} &:=& \elab{(\secret{x} - \flip{x})}{2}  & \textit{(unicast)}\\
+  %    \mx{y}{*} &:=& \elab{(\flip{y} + \mesg{x})}{1}  & \textit{(broadcast)}
+  %  \end{array}
+  %  $$ 
+  %\end{exampleblock}
 \end{frame}
@@
     etc.
-  \item \emph{algebraically} $\sx{1}{1} = \elab{\mesg{s1}}{2} + \elab{\mesg{s1}}{3} + \elab{\locflip}{1}$.
+  \item \emph{algebraically} $\sx{1}{1} \eop \elab{\mesg{s1}}{2} + \elab{\mesg{s1}}{3} + \elab{\locflip}{1}$.
   \item \emph{probabilistically} any 2 out of three shares are in a uniform random distribution.
@@
 
-\begin{frame}{Example contd.: Public Broadcast Reveal}
+\begin{frame}{Additive Homomorphism of Shares}
 
-  \begin{itemize}
-   \item \emph{algebraically} the sum of all shares = $\sx{1}{1} + \sx{2}{2} + \sx{3}{3}$    
-  \end{itemize}
+  To sum the secrets \emph{in secret}:
 
-  \begin{block}{Broadcast Sum-of-Shares}
+  \begin{block}{Additive Homomorphism: Sum-of-Shares}
     $$
     \begin{array}{lll}
-      \mx{ss1}{*} &:=& \elab{(\locflip \fplus \mesg{s2} \fplus \mesg{s3})}{1} \\ 
-      \mx{ss2}{*} &:=& \elab{(\mesg{s1} \fplus \locflip \fplus \mesg{s3})}{2} \\
-      \mx{ss3}{*} &:=& \elab{(\mesg{s1} \fplus \mesg{s2} \fplus \locflip)}{3} 
+      \mx{ss1}{1} &:=& \elab{(\locflip \fplus \mesg{s2} \fplus \mesg{s3})}{1} \\ 
+      \mx{ss2}{2} &:=& \elab{(\mesg{s1} \fplus \locflip \fplus \mesg{s3})}{2} \\
+      \mx{ss3}{3} &:=& \elab{(\mesg{s1} \fplus \mesg{s2} \fplus \locflip)}{3} 
     \end{array}
@@
 
+  Additive homorphism:
   \begin{itemize}
-  \item Messages $\mx{ss1}$ through $\mx{ss3}$ are \cemph{not probabilistically independent of inputs}, but ...
-  \item They \emph{are} \cemph{independent of secrets conditioned on the output}!
+  \item \emph{Algebraically}, $\mx{ss1}{1} + \mx{ss1}{2} + \mx{ss1}{1} \eop \sx{1}{1} + \sx{2}{2} + \sx{3}{3}$ 
+  \item \emph{Probabilistically}, protocol step is noninteractive and each of $\mx{ss1}{i}$ remain in a uniform distribution.
   \end{itemize}
@@
 
+\begin{comment}
 \begin{frame}{Example contd.: Output}
@@
     $$
-    \mx{ss1}{*} + \ \mx{ss2}{*} + \ \mx{ss3}{*} = \sx{1}{1} + \sx{2}{2}  + \sx{3}{3}
+    \mx{ss1}{*} + \ \mx{ss2}{*} + \ \mx{ss3}{*} \eop \sx{1}{1} + \sx{2}{2}  + \sx{3}{3}
     $$
@@
 \end{frame}
+\end{comment}
 
-\section{Models of a Protocol}
+\section{SMT Boosting: Correctness}
 
-\begin{frame}{Computing Protocol Distributions}
+\begin{frame}{Verifying Correctness}
 
-  The probability distribution of a protocol $\prog$ (as a pmf) can be precisely computed
-  by interpretation as a \cemph{constraint program}.
-  \begin{itemize}
-  \item $\mathbb{F}_2$ interpretation as a Datalog program-- set of Least Herbrand Models.
-    \begin{itemize}
-      \item Random tapes represented as fact base.
-    \end{itemize}
-  \item $\mathbb{F}_p$ interpretation as SMT constraints\footnote{Ozdemir, A., Kremer, G., Tinelli, C., Barrett, C.
-(2023).
-\emph{Satisfiability Modulo Finite Fields}.
-In: Enea, C., Lal, A.
-(eds) Computer Aided Verification.
-CAV 2023.
-}-- set of all models.
-  \end{itemize}
-  Set of models can be queried to enforce hyperproperties-- but poor scaling.
-
-  \begin{alertblock}{Hypothesis: Composable Properties}
-    Compositional properties of protocol components can be independently verified, tractably.
-  \end{alertblock}
+Letting $\prog$ be the above protocol:
+$$
+\toeq{\prog} \wedge \ttt{out} \eop \mx{ss1}{1} + \mx{ss1}{2} + \mx{ss1}{1} \models\ \ttt{out} \eop \sx{1}{1} \fplus \sx{2}{2} \fplus \sx{3}{3}
+$$
 
@@
 
-\section{Compositional Properties of Circuits}
+\section{SMT Boosting: Confidentiality}
 
-\begin{frame}{Prelude and Component Definitions}
+\begin{frame}{Verifying Confidentiality}
 
-  The Prelude language allows modular definition of Overture protocol components.
-  \begin{itemize}
-  \item Function definitions, function calls, structured data.
-  \item A \emph{staged} language that generates a complete Overture protocol
-    as a side effect.
-  \end{itemize}
-
-  \begin{exampleblock}{Example: Local Summation}
 $$    
-    \ttt{sum}(x,y,z,i) \{ \mx{z}{i} := \elab{(\mesg{x} \fplus \mesg{y})}{i} \} 
+  \inferrule[Encode]
+  {\eqs \models \phi \eop \phi' \oplus \rx{w}{\cid} \\
+   \oplus \in \{ \fplus,\fminus \}\\
+   \eqj{R}{\eqs}{\phi'}{\ty}}
+  {\eqj{R;\{ \rx{w}{\cid} \}}{\eqs}{\phi}{\setit{\cty{\rx{w}{\cid}}{\ty}}}}
 $$
-  \end{exampleblock}
-
-  In addition to programmer convenience, Prelude supports function-level verification.
-  %Prelude supports verification:
-  %\begin{itemize}
-  %\item Prelude functions define Overture (sub)protocols that can be automatically verified.
-  %\item Large protocols can be verified manually by composing function properties. 
-  %\end{itemize}
-
-\end{frame}
-
-\begin{frame}{Circuits and Gate Paradigms}
-
-  Goldreich-Micali-Wigderson (GMW) protocol implements $\idealf$ as a \cemph{circuit}.
-  \begin{itemize}
-  \item Composed by combing-- aka ``wiring''-- \cemph{gates}.
-  \item Each gate standard definition, supports primitive $\fplus$ and $\ftimes$.
-  \item Two input wires, one output wire, values remain encrypted on wire. 
-  \end{itemize}
-
-  \begin{exampleblock}{Example: 2-Party GMW in $\mathbb{F}_2$}
-    Rep.~invariant: each client in $\setit{1,2}$ holds a share of wire values.
-
-    \medskip
-
-    $\ttt{andgmw}(x,y,z)$:
-
-    \begin{itemize}
-    \item Input shares $\mx{x}{1}$, $\mx{x}{2}$ and $\mx{y}{1}$, $\mx{y}{2}$--
-      wire value for $x$ is $\mx{x}{1} + \mx{x}{2}$, etc.
-    \item Output shares $\mx{z}{1}$, $\mx{z}{2}$ received during gate execution.
-      $$
-      \mx{z}{1} + \mx{z}{2} = (\mx{x}{1} + \mx{x}{2}) * (\mx{y}{1} + \mx{y}{2})
-      $$
-    \end{itemize}
-
-  \end{exampleblock}
-
-\end{frame}
-
-\begin{frame}
-
-  \begin{exampleblock}{Semi-Automatically Verifying 2-Party GMW}
-
-    Picking dummy $x,y,z$, we can execute $\ttt{andgmw}(x,y,z)$
-    to generate a generic ``and'' gate $\prog_{\wedge}$ in Overture.
-    \begin{itemize}
-    \item Automatically calculate the pmf of $\prog_{\wedge}$.
-    \item Pmf can be queried to verify critical properties-- e.g., the
-      uniform distribution of output shares under any input conditions. 
-    \end{itemize}
-    Gate verification serves as a \emph{tactic} for manual circuit-level
-    proofs. 
-    \begin{itemize}
-    \item Leverage logics for reasoning about probabilistic programming: Probabilistic
-      Separation Logic, Lilac.
-    \end{itemize}
-
-  \end{exampleblock}
 

### 2025-04-24 code/cvc5mvn/src/main/java/plaid/logic/ConstraintAnalyzer.java
cleanup messages

@@ -10,7 +10,7 @@ public class ConstraintAnalyzer {
     private final Program program; // provide a program to resolve function
     private final ConstraintEvaluator evaluator;
     private final GenEntailVerifier verifier;
-    private final Map<PreludeExpression, Constraints> functionConstraints = new HashMap<>(); // data structure to store inferred pre/post conditions by FN rule
+    private final Map<Identifier, Constraints> functionConstraints = new HashMap<>(); // data structure to store inferred pre/post conditions by FN rule
     
     private Map<Identifier, PreludeExpression> binding(List<TypedIdentifier> formal, List<PreludeExpression> actual){
         // for each actual parameter, bind it to corresponding formal parameter 
@@ -33,8 +33,8 @@ public class ConstraintAnalyzer {
     public Constraints inferPrePostFN(CommandFunction function) {
         // infer the precondition and postcondition for the function using inferPrePostCmd
         Constraints constraints = inferPrePostCmd(function.c());
-        System.out.println("ASDF inferred pre for : " + function.fname() + ScalaFunctions.prettyPrint(constraints.getPre()));
-        System.out.println("ASDF inferred post for: " + function.fname().toString() + ScalaFunctions.prettyPrint(constraints.getPost()));
+        System.out.println("Inferred pre for " + function.fname().name() + ": " + ScalaFunctions.prettyPrint(constraints.getPre()));
+        System.out.println("Inferred post for " + function.fname().name() + ": "  + ScalaFunctions.prettyPrint(constraints.getPost()));
         // if the function does not have annotated pre/post conditions
         if (function.precond() == null && function.postcond() == null ){
             // then store inferred pre/post conditions and return them

### 2025-04-24 esop25/slides.tex
checkpoint

@@
 
-  \begin{block}{Reconstructive Sharing}
+  \begin{block}{Reconstructive Sharing: Protocol $\prog_{\mathit{sum3}}$}
     $$
@@
   \end{block}
+  $\elab{\mesg{s1}}{2}, \elab{\mesg{s1}}{3}, \elab{\locflip}{1}$ are the \cemph{secret shares} of $\sx{1}{1}$,
+    etc.
   \begin{itemize}
-  \item $\elab{\mesg{s1}}{2}, \elab{\mesg{s1}}{3}, \elab{\locflip}{1}$ are the \emph{shares} of $\sx{1}{1}$,
-    etc.
   \item \emph{algebraically} $\sx{1}{1} \eop \elab{\mesg{s1}}{2} + \elab{\mesg{s1}}{3} + \elab{\locflip}{1}$.
@@
 
-  To sum the secrets \emph{in secret}:
-
   \begin{block}{Additive Homomorphism: Sum-of-Shares}     
+  To sum the input secrets \emph{in secret} extend $\prog_{\mathit{sum3}}$:
     $$
@@
 
-  Additive homorphism:
+  Additive homorphism-- each $\mx{ssi}{i}$ for $i \in \setit{1,2,3}$ is a share:
   \begin{itemize}
-  \item \emph{Algebraically}, $\mx{ss1}{1} + \mx{ss1}{2} + \mx{ss1}{1} \eop \sx{1}{1} + \sx{2}{2} + \sx{3}{3}$ 
-  \item \emph{Probabilistically}, protocol step is noninteractive and each of $\mx{ss1}{i}$ remain in a uniform distribution.
+  \item \emph{Algebraically}, $\mx{ss1}{1} + \mx{ss1}{2} + \mx{ss1}{3} \eop \sx{1}{1} + \sx{2}{2} + \sx{3}{3}$ 
+  \item \emph{Probabilistically}, protocol step is noninteractive and each $\mx{ssi}{i}$ remains in a uniform distribution.
   \end{itemize}
@@
 
+\section{Using SMT}
+
+\begin{frame}{Using Satisfiability Modulo Finite Fields (SMFF)}
+
+  SMFF supports equational constraint definitions $\eqs$ over a finite field $\mathcal{F}_{p}$
+  with a mature solver, in the style of SMT.
+
+  \medskip
+  Overture protocols $\prog$ have a straightforward interpretation as SMFF constraints $\toeq{\prog}$ , e.g.,
+  the first two lines of $\prog_{\mathit{sum3}}$ translate as:
+  $$
+  %\toeq{
+  %  \begin{array}{c}
+  %  \begin{array}{lll}
+  %    \elab{\mesg{s1}}{2} &:=& \elab{(\secret{1} \fminus \locflip \fminus \flip{x})}{1} \\ 
+  %    \elab{\mesg{s1}}{3} &:=& \elab{\flip{x}}{1} 
+  %  \end{array}\\
+  %  = \\
+    \begin{array}{lll}
+      \elab{\mesg{s1}}{2} &\eop& \elab{\secret{1}}{1} \fminus \elab{\locflip}{1} \fminus \elab{\flip{x}}{1}\quad \wedge \\ 
+      \elab{\mesg{s1}}{3} &\eop& \elab{\flip{x}}{1} 
+    \end{array}
+    %\end{array}
+    %}
+  $$
+
+\end{frame}
+
 \section{SMT Boosting: Correctness}
@@
 $$
-\toeq{\prog} \wedge \ttt{out} \eop \mx{ss1}{1} + \mx{ss1}{2} + \mx{ss1}{1} \models\ \ttt{out} \eop \sx{1}{1} \fplus \sx{2}{2} \fplus \sx{3}{3}
+\begin{array}{c}
+  \toeq{\prog} \wedge \ttt{out} \eop \mx{ss1}{1} + \mx{ss1}{2} + \mx{ss1}{3}\\ \models\\
+  \ttt{out} \eop \sx{1}{1} \fplus \sx{2}{2} \fplus \sx{3}{3}
+\end{array}
 $$

### 2025-04-26 esop25/slides.tex
checkpoint

@@
 \usepackage{comment}
+\usepackage{fancyvrb}
+\usepackage{moreverb}
 
@@
 \input{macros}
+
+\newcommand{\sumthree}{\prog_{\mathit{sum3}}}
 
@@
   \begin{itemize}
-  \item \emph{algebraically} $\sx{1}{1} \eop \elab{\mesg{s1}}{2} + \elab{\mesg{s1}}{3} + \elab{\locflip}{1}$.
+  \item \emph{algebraically} $\elab{\locflip}{1} + \elab{\mesg{s1}}{2} + \elab{\mesg{s1}}{3} \eop \sx{1}{1}$.
   \item \emph{probabilistically} any 2 out of three shares are in a uniform random distribution.
@@
 
-\begin{frame}{Using Satisfiability Modulo Finite Fields (SMFF)}
+\begin{frame}{cvc5 Finite Fields Theory}
 
-  SMFF supports equational constraint definitions $\eqs$ over a finite field $\mathcal{F}_{p}$
-  with a mature solver, in the style of SMT.
+  The cvc5 Finite Fields Theory supports automatate satisfiability checking of equational
+  constraint definitions $\eqs$ over prime fields $\mathcal{F}_{p}$.
 
   \medskip
+
   Overture protocols $\prog$ have a straightforward interpretation as SMFF constraints $\toeq{\prog}$ , e.g.,
@@
 
+ Verifiable \emph{entailment} $\eqs_1 \models \eqs_2$ means that any model of $\eqs_1$ is a model
+ of $\eqs_2$. Models of $\prog$ correspond exactly to valid runs of $\prog$...
+
 \end{frame}
@@
 
-\begin{frame}{Verifying Correctness}
+\begin{frame}{Verifying Protocol Correctness}
 
-Letting $\prog$ be the above protocol:
+  Entailment is a mechanism for verifying protocol properties. If an interpretation
+  of a protocol entails a property, it holds in all protocol runs. 
 $$
 \begin{array}{c}
-  \toeq{\prog} \wedge \ttt{out} \eop \mx{ss1}{1} + \mx{ss1}{2} + \mx{ss1}{3}\\ \models\\
+  \toeq{\sumthree} \wedge \ttt{out} \eop \mx{ss1}{1} + \mx{ss1}{2} + \mx{ss1}{3}\\ \models\\
   \ttt{out} \eop \sx{1}{1} \fplus \sx{2}{2} \fplus \sx{3}{3}
@@
 
+Correctness of subroutines (e.g., addition and multiplication gates in encrypted circuits)
+or the full protocol wrt $\idealf$ can be checked automatically.
 \end{frame}
@@
 
+\end{frame}
+
+\begin{frame}[fragile]{Modular Typing in Prelude}
+  \small{
+  \begin{verbatimtab}
+SHARED(x) { (m[x]@1 + m[x]@2) }
+
+andtablegmw(x, y, z) {
+  let r11 = (r[z] + (m[x] + 1) * (m[y] + 1)) in
+  ... // elided
+  { r1 = r11; r2 = r10; r3 = r01; r4 = r00 }
+}
+
+andgmw(z:string, x:string, y:string) {
+  let tab = andtablegmw(x,y,z) in
+  m[z]@2 := OT4((m[x], m[y])@2, tab.r4, tab.r3, tab.r2, tab.r1)@1;
+  m[z]@1 := r[z]@1
+}
+postcondition: ( m[z]@1 == r[z]@1 AND
+                 m[z]@2 == (SHARED(x) * SHARED(y)) + r[z]@1) ) 
+  \end{verbatimtab}
+  }
 \end{frame}

### 2025-04-26 esop25/slides.tex
checkpoint

@@
   \end{itemize}
+  \emph{Computations} are normal field operations on variables-- addition (+), multiplication (*),
+  and additive inverse (-).
+
+  \medskip
+
   We localize computation to \emph{clients}, and any information
@@
 
+Verifying equational properties 
 $$    

### 2025-04-26 esop25/slides.tex
checkpoint

@@
 
-Verifying equational properties 
+  Verifying equational properties also supports verification of confidentiality.
+  It is well-known that expressions of the form:
+  $$
+  e - \rx{w}{\cid}
+  $$
+  are a means of encrypting $e$ using $\rx{w}{\cid}$ as a one-time-pad in any
+  size field. This can be leveraged in typing. 
   $$    
   \inferrule[Encode]
-  {\eqs \models \phi \eop \phi' \oplus \rx{w}{\cid} \\
-   \oplus \in \{ \fplus,\fminus \}\\
-   \eqj{R}{\eqs}{\phi'}{\ty}}
-  {\eqj{R;\{ \rx{w}{\cid} \}}{\eqs}{\phi}{\setit{\cty{\rx{w}{\cid}}{\ty}}}}
+  {\eqs \models \toeq{e} \eop \toeq{e'} - \rx{w}{\cid}\\
+   \eqj{R}{\eqs}{e'}{\ty}}
+  {\eqj{R;\{ \rx{w}{\cid} \}}{\eqs}{e}{\setit{\cty{\rx{w}{\cid}}{\ty}}}}
   $$
+  \begin{itemize}
+  \item The \emph{ciphertype} $\cty{\rx{w}{\cid}}{\ty}$ indicates the computation
+    results in a value encrypted with $\rx{w}{\cid}$.
+  \item $\ty$ is a conservative approximation of the dependencies in $e'$.
+  \item The \emph{randomness environment} $R;\{ \rx{w}{\cid} \}$ enforces
+    linearity.
+  \end{itemize}
 

### 2025-04-26 esop25/slides.tex
checkpoint

@@
   \end{itemize}
-  Many existing deployments: voting, auctions, satellite trajectories, COVID exposure notification, ...\footnote{https://mpc.cs.berkeley.edu}
+  Many existing deployments: voting, auctions, COVID exposure notification, federated deep learning...\footnote{https://mpc.cs.berkeley.edu}
 

### 2025-04-26 esop25/slides.tex
checkpoint

@@
 
-\begin{frame}[fragile]{Modular Typing in Prelude}
+\section{SMT Boosting: Confidentiality}
+
+\begin{frame}{Verifying Confidentiality}
+
+  Applying this rule to, e.g., $\sumthree$ is straightforward given it's syntactic
+  form, but applies more generally.
+
+  \medskip
+
+  Yao's Garbled Circuits (YGC) in $\mathbb{F}_{2}$ wire input encoding, using
+  $\rx{w}{2}$ as encrypted wire value:
+$$
+\xassign{\mx{w}{1}}{\mux{\secret{w}}{\neg\flip{w}}{\flip{w}}}{2}
+$$
+where for all $\be_1,\be_2,\be_3$ (with $\neg$ denoting negation in $\mathbb{F}_2$):
+$$
+\mux{\be_1}{\be_2}{\be_3} \defeq (\neg\be_1 \ftimes \be_2) \fplus (\be_1 \ftimes \be_3)
+$$
+and letting this protocol be $\prog$ the following is valid:
+\begin{mathpar}
+%\toeq{\xassign{\mx{w}{1}}{\OT{\secret{w}}{1}{\neg\flip{w}}{\flip{w}}}{2}}
+   \toeq{\prog} \models \mx{w}{1} \eop \neg\sx{w}{2} - \rx{w}{2}
+\end{mathpar}
+\end{frame}
+
+\begin{frame}{Compositional Typing in Prelude}
+
+The $\metaprot$ language is a \emph{metalanguage} aka metaprogramming
+language, where an $\fedprot$ protocol is the result of
+computation.
+\begin{itemize}
+\item Supports compositional verification of larger
+  protocols.
+\item Confidentiality and integrity properties to be recorded in pre/post
+  conditions.
+\item Encapsulation and compositional re-use of function property verification.
+\end{itemize}
+
+$$
+\begin{array}{l}
+\ttt{encode}(y,s,r) \{\\
+\quad \xassign{\mx{y}{r}}{\mux{\secret{y}}{\neg\flip{y}}{\flip{y}}}{s};\\
+% \quad \eqcast{\mx{y}{r}}{\neg\sx{y}{s} \fplus \rx{y}{s}}\\
+\}\\
+\ttt{postcondition:} ( \mx{y}{r} \eop \neg\sx{y}{s} \fplus \rx{y}{s} )
+\end{array}
+$$
+
+\end{frame}
+
+\begin{frame}[fragile]{Compositional Typing in Prelude}
   \small{
@@
 
+\begin{frame}[fragile]{Compositional Integrity Typing in Prelude}
+  \small{
+  \begin{verbatimtab}
+MACOK(x, i1, i2)
+  { m[x++"m"]@i2 ==
+    m[x++"k"]@i1 + (m["delta"]@i1 * m[x++"s"]@i2) }
+
+MACed(x) { MACOK(x, 1, 2) AND MACOK(x, 2, 1) }
+
+_sum(z:string, x:string, y:string, i:cid){
+    m[z++"s"]@i := (m[x++"s"] + m[y++"s"])@i;
+    m[z++"m"]@i := (m[x++"m"] + m[y++"m"])@i;
+    m[z++"k"]@i := (m[x++"k"] + m[y++"k"])@i
+}
+
+precondition: ( MACed(x) AND MACed(y) )
+sum(z:string, x:string, y:string){ _sum(z,x,y,1); _sum(z,x,y,2) }
+postcondition: ( MACed(z) )
+  \end{verbatimtab}
+  }
+\end{frame}
+
 \end{document}

### 2025-04-28 esop25/new-cpj.tex
checkpoint

\newcommand{\Shares}{\mathcal{S}}
\newcommand{\sharet}[2]{\mathit{share}(#1,#2)}
\newcommand{\sid}{\texttt{s}}
\newcommand{\sexp}{s}
\newcommand{\keys}{\mathcal{K}}
\newcommand{\rkey}{\kappa}
\renewcommand{\unity}[2]{U(#1,#2)}
\newcommand{\uncorr}[2]{#1\bowtie#2}

\section{New Confidentiality Typing}

\begin{mathpar}
  \inferrule[Mesg]
            {}
            {\Gamma, \varnothing, \eqs \vdash \mx{w}{\cid} : \Gamma(\mx{w}{\cid})}

  \inferrule[Rand]
            {}
            {\Gamma, \varnothing, \eqs \vdash \rx{w}{\cid} : \unity{\rx{w}{\cid}}{\setit{\rx{w}{\cid}}}}

  \inferrule[Ideal]
            {}
            {\Gamma, \varnothing, \eqs \vdash f(\sx{w_1}{\cid_1},\ldots,\sx{w_n}{\cid_n}) :
              \setit{f(\sx{w_1}{\cid_1},\ldots,\sx{w_n}{\cid_n})}}

  \inferrule[Enc]
            {\Gamma, R, \eqs \vdash \phi : \ty}
            {\Gamma, R;\rx{w}{\cid}, \eqs \vdash \phi \fminus \rx{w}{\cid} : \unity{\rx{w}{\cid}}{\ty}}

  \inferrule[HeAdd]
      {\Gamma, R_1, \eqs \vdash \phi_1 : \setit{\unity{\rkey_1}{\ty_1}} \\
        \Gamma, R_2, \eqs \vdash \phi_2 : \setit{\unity{\rkey_2}{\ty_2}} \\
        \uncorr{\rkey_1}{\rkey_2}
      }
      {\Gamma, R_1;R_2, \eqs \vdash \phi_1 \fplus \phi_2 : \setit{\unity{\rkey_1 \oplus \rkey_2}{\ty_1 \cup \ty_2}}}

  \inferrule[Join]
      {\Gamma, R_1, \eqs \vdash \phi_1 : \ty_1 \\ \Gamma, R_2, \eqs \vdash \phi_2 : \ty_2}
      {\Gamma, R_1;R_2, \eqs \vdash \phi_1 \bop \phi_2 : \ty_1 \cup \ty_2}

  \inferrule[Eq]
            {\eqs \models \phi \eop \phi' \\ \Gamma, R, \eqs \vdash \phi' : \ty}
            {\Gamma, R, \eqs \vdash \phi : \ty}          
\end{mathpar}

\begin{mathpar}
  \inferrule[Mesg]
      {\Gamma, R, \eqs \vdash \toeq{\elab{e}{\cid}} : \ty}
      {\Gamma, \eqs \vdash \xassign{x}{e}{\cid} : (\Gamma;x:\ty),R}

  \inferrule[Seq]
      {\Gamma_1,\eqs \vdash \cmd_1 : \Gamma_2,R_1 \\
        \Gamma_2,\eqs \vdash \cmd_2 : \Gamma_3,R_2}
      {\Gamma_1, \eqs \vdash \cmd_1;\cmd_2 : \Gamma_3,R_1;R_2}
\end{mathpar}

\begin{comment}
\section{Deprecated}

\begin{mathpar}
  \inferrule[Mesg]
            {}
            {\Gamma, \eqs \vdash x : \Gamma(x)}

  \inferrule[Key]
            {\sid:[\ldots,x_k,\ldots] \in \keys}
            {\Gamma, \eqs \vdash x_k : \setit{\sharet{\sid}{k}}}

  \inferrule[Secret]
            {}
            {\Gamma, \eqs \vdash \sx{w}{\cid} : \setit{\sx{w}{\cid}}}

  \inferrule[Join]
            {\Gamma, \eqs \vdash \phi_1 : \ty_1 \\ \Gamma, \eqs \vdash \phi_2 : \ty_2}
            {\Gamma, \eqs \vdash \phi_1 \bop \phi_2 : \ty_1 \cup \ty_2}

  \inferrule[Eq]
            {\eqs \models \phi \eop \phi' \\ \Gamma, \eqs \vdash \phi' : \ty}
            {\Gamma, \eqs \vdash \phi : \ty}

  \inferrule[Share]
            {\Gamma, \eqs \vdash x_1 : \sharet{\sid}{1} \\ \cdots \\ \Gamma, \eqs \vdash x_{n-1} : \sharet{\sid}{n-1} }
            {\Gamma, \eqs \vdash \phi - x_1 - \cdots - x_{n-1} : [\phi]_\sid}

  \inferrule[HeAdd]
      {\Gamma, \eqs \vdash \phi_1 : \setit{\sharet{\sexp_1}{k}} \\
       \Gamma, \eqs \vdash \phi_2 : \setit{\sharet{\sexp_2}{k}} }
      {\Gamma, \eqs \vdash \phi_1 \fplus \phi_2 : \setit{\sharet{\sexp_1 \oplus \sexp_2}{k}}}
\end{mathpar}

\begin{mathpar}
  \inferrule[Enc]
      {\Gamma, \eqs \vdash \toeq{\elab{e}{\cid}} : [\phi]_{\sid}}
      {\Gamma,\Shares,\eqs \vdash \xassign{x}{e}{\cid} : (\Gamma;x:\sharet{\sid}{n}),(\Shares;[\phi]_\sid)}

  \inferrule[Mesg]
      {\Gamma, \eqs \vdash \toeq{\elab{e}{\cid}} : \ty}
      {\Gamma,\Shares,\eqs \vdash \xassign{x}{e}{\cid} : (\Gamma;x:\ty),\Shares}

  \inferrule[Seq]
      {\Gamma_1,\Shares_1,\eqs \vdash \prog_1 : \Gamma_2,\Shares_2 \\
        \Gamma_2,\Shares_2,\eqs \vdash \prog_2 : \Gamma_3,\Shares_3}
      {\Gamma_1,\Shares_1,\eqs \vdash \prog_1;\prog_2 : \Gamma_3,\Shares_3}
\end{mathpar}

\begin{mathpar}
  \inferrule
      {}
      {\Shares \vdash T \cup \setit{\sx{w}{\cid}},C \leadsto T, C.\sx{w}{\cid}}

  \inferrule
      {}
      {\Shares \vdash T \cup \setit{\sharet{\sexp}{1},\ldots,\sharet{\sexp}{n}},C \leadsto T, C.\Shares(\sexp)}

  \inferrule
      {\setit{\sharet{\sexp_1 \oplus \sexp_2}{j}, \sharet{\sexp_1}{k}, \sharet{\sexp_2}{k}} \subseteq T}
      {\Shares \vdash T,C \leadsto T \cup \setit{\sharet{\sexp_1 \oplus \sexp_2}{k}},C}

  \inferrule
      {\Shares \vdash T_1,C_1 \leadsto T_2,C_2 \\ \Shares \vdash T_2,C_2 \leadsto T_3,C_3 }
      {\Shares \vdash T_1,C_1 \leadsto T_3,C_3 }
\end{mathpar}
\end{comment}

### 2025-04-28 esop25/slides.tex
checkpoint

@@
 
-  The cvc5 Finite Fields Theory supports automatate satisfiability checking of equational
+  The cvc5 Finite Fields Theory supports automated satisfiability checking of equational
   constraint definitions $\eqs$ over prime fields $\mathcal{F}_{p}$.
@@
   \begin{itemize}
-  \item Generalize verification to arbitrary prime fields.
-  \item Automated type system to eliminate manual proof elements.
+  \item In this paper: typing verifies noninterference until declassification.
+    \begin{itemize}
+    \item Goal: accurately verify declassification bounds.
+    \end{itemize}
+  \item Automated weakest precondition/stongest postcondition inference.
+  \item Scalability evaluation of large circuits.
   \item Support for concurrency.

### 2025-04-28 esop25/slides.tex
checkpoint

@@
  Verifiable \emph{entailment} $\eqs_1 \models \eqs_2$ means that any model of $\eqs_1$ is a model
- of $\eqs_2$. Models of $\prog$ correspond exactly to valid runs of $\prog$...
+ of $\eqs_2$. Models of $\prog$ correspond exactly to valid runs of $\prog$, therefore...
 
@@
 
-  Entailment is a mechanism for verifying protocol properties. If an interpretation
+  ... entailment is a mechanism for verifying protocol properties. If an interpretation
   of a protocol entails a property, it holds in all protocol runs. 
@@
   $$
-  e - \rx{w}{\cid}
+  (e - \rx{w}){\cid}
   $$

### 2025-05-04 esop25/slides.tex
checkpoint

@@
 
-  Existing HLLs rely on low-level MPC protocols with correctness
+  Existing high-level PLs rely on low-level MPC protocols with correctness
   assumed.
@@
   \item Any client can be corrupted by the adversary, who can then see their messages.
-  \item The \cemph{output} is made \emph{publicly available} by $\prog$.
+  \item The \cemph{output} $=\idealf$ is made \emph{publicly available} by $\prog$.
   \end{itemize}
-  \cemph{The output may leak secret information to an adversary} trying to
+  \cemph{$\idealf$ may expose some information to an adversary} trying to
   guess secret values, since it depends on inputs.
@@
 
-  \cemph{Confidentiality} means that protocol messages don't enable the adversary
-  to make better guesses than the output alone.
+  \cemph{Confidentiality} means that the protocol doesn't enable the adversary
+  to make better guesses than $\idealf$.
 
@@
 
-  Low-level, probabilistic protocols are defined in terms of basic variable types
-  (identifiers $x$ distinguish them):
+  In our language Overture, low-level, probabilistic protocols are defined in
+  terms of basic variable types (identifiers $x$ distinguish them):
   \begin{itemize}
@@
   The cvc5 Finite Fields Theory supports automated satisfiability checking of equational
-  constraint definitions $\eqs$ over prime fields $\mathcal{F}_{p}$.
+  constraint definitions $\eqs$ over prime fields $\mathbb{F}_{p}$.
 
@@
 
+\section{SMT Boosting: Integrity}
 \begin{frame}[fragile]{Compositional Integrity Typing in Prelude}

### 2025-05-05 esop25/slides.tex
checkpoint

@@
   \cemph{Secure Multiparty Computation (MPC)} is a paradigm for
-  \cemph{confidential computing}.
+  \cemph{confidential computing} in distributed settings.
   \begin{itemize}
@@
 
-  Existing high-level PLs rely on low-level MPC protocols with correctness
+  Existing high-level PLs rely on low-level MPC protocols with correctness and security
   assumed.
   \begin{itemize}
-  \item Distinct features, e.g., probabilistic sampling.
-  \item Existing proof techniques mostly manual.
+  \item Distinct low-level features, e.g., probabilistic sampling.
+  \item Existing low-level proof techniques mostly manual.
   \end{itemize}
@@
   \item Clients are distributed and communicate via \cemph{message passing}.
-  \item Any client can be corrupted by the adversary, who can then see their messages.
+  \item Any client can be corrupted by the \cemph{adversary}, who can then see their messages.
   \item The \cemph{output} $=\idealf$ is made \emph{publicly available} by $\prog$.
@@
   \cemph{Confidentiality} means that the protocol doesn't enable the adversary
-  to make better guesses than $\idealf$.
+  to make better guesses than $\idealf$.\footnote{Passive adversary assumed for this talk.}
 
@@
 
-\begin{frame}{cvc5 Finite Fields Theory}
+\begin{frame}{SMT Finite Fields Theory}
 
-  The cvc5 Finite Fields Theory supports automated satisfiability checking of equational
+  The SMT Finite Fields Theory supports automated satisfiability checking of equational
   constraint definitions $\eqs$ over prime fields $\mathbb{F}_{p}$.
@@
 
-  Overture protocols $\prog$ have a straightforward interpretation as SMFF constraints $\toeq{\prog}$ , e.g.,
+  Overture protocols $\prog$ have a straightforward interpretation as equality constraints $\toeq{\prog}$ , e.g.,
   the first two lines of $\prog_{\mathit{sum3}}$ translate as:
@@
 
- Verifiable \emph{entailment} $\eqs_1 \models \eqs_2$ means that any model of $\eqs_1$ is a model
- of $\eqs_2$. Models of $\prog$ correspond exactly to valid runs of $\prog$, therefore...
+    Random variables (e.g., $\rx{x}{1}$) unconstrained in the interpretation.
+
+    \medskip
+
+    \cemph{Models of $\toeq{\prog}$ correspond exactly to runs of $\prog$}.
 
@@
 
-  ... entailment is a mechanism for verifying protocol properties. If an interpretation
+    Verifiable \emph{entailment} $\eqs_1 \models \eqs_2$ means that any model of $\eqs_1$ is a model
+    of $\eqs_2$.
+
+    \medskip
+    Entailment is a mechanism for verifying protocol properties. If an interpretation
     of a protocol entails a property, it holds in all protocol runs. 
@@
   $$
-  (e - \rx{w}){\cid}
+  (e - \flip{w}){\cid}
   $$
-  are a means of encrypting $e$ using $\rx{w}{\cid}$ as a one-time-pad in any
+  are a means of encrypting $\elab{e}{\cid}$ using $\rx{w}{\cid}$ as a one-time-pad in any
   size field. This can be leveraged in typing. 

### 2025-05-07 esop25/new-cpj.tex
checkpoint

@@
-\newcommand{\Shares}{\mathcal{S}}
-\newcommand{\sharet}[2]{\mathit{share}(#1,#2)}
-\newcommand{\sid}{\texttt{s}}
-\newcommand{\sexp}{s}
-\newcommand{\keys}{\mathcal{K}}
-\newcommand{\rkey}{\kappa}
-\renewcommand{\unity}[2]{U(#1,#2)}
-\newcommand{\uncorr}[2]{#1\bowtie#2}
+\renewcommand{\unity}[2]{U(#1, #2)}
+\renewcommand{\ty}{T}
+\newcommand{\eset}{D}
+\newcommand{\dty}[2]{\langle #1, #2 \rangle}
+\newcommand{\sharet}[1]{\mathrm{shares}(#1)}
+\newcommand{\cvec}[1]{[#1]}
+\newcommand{\secopent}{\mathrm{secopen}}
+\newcommand{\declasst}[1]{\mathrm{declass}(#1)}
+\renewcommand{\view}{\mathit{view}}
+\newcommand{\revealed}{\mathit{revealed}}
 
-\section{New Confidentiality Typing}
+\section{Declassification Typing}
+
+$$
+\begin{array}{rcl@{\hspace{4mm}}r}
+\eset &\in& 2^{\phi} & \gdesc{Dependency Approximations}\\ 
+\ty &::=& \dty{R}{\eset} & \gdesc{Probabilistic Types} \\
+\Gamma &::=& \varnothing \mid \Gamma; x : \ty & \gdesc{Type Environments}
+\end{array}
+$$
+
+\subsection{Primitive Typing Rules}
 
@@
             {}
-            {\Gamma, \varnothing, \eqs \vdash \mx{w}{\cid} : \Gamma(\mx{w}{\cid})}
+            {\Gamma, \eqs \vdash \mx{w}{\cid} : \Gamma(\mx{w}{\cid})}
 
@@
             {}
-            {\Gamma, \varnothing, \eqs \vdash \rx{w}{\cid} : \unity{\rx{w}{\cid}}{\setit{\rx{w}{\cid}}}}
+            {\Gamma, \eqs \vdash \rx{w}{\cid} : \dty{\setit{\rx{w}{\cid}}}{\varnothing}}
 
@@
             {}
-            {\Gamma, \varnothing, \eqs \vdash f(\sx{w_1}{\cid_1},\ldots,\sx{w_n}{\cid_n}) :
-              \setit{f(\sx{w_1}{\cid_1},\ldots,\sx{w_n}{\cid_n})}}
+            {\Gamma, \eqs \vdash f(\sx{w_1}{\cid_1},\ldots,\sx{w_n}{\cid_n}) :
+              \dty{\varnothing}{\setit{f(\sx{w_1}{\cid_1},\ldots,\sx{w_n}{\cid_n})}}}
 
-  \inferrule[Enc]
-            {\Gamma, R, \eqs \vdash \phi : \ty}
-            {\Gamma, R;\rx{w}{\cid}, \eqs \vdash \phi \fminus \rx{w}{\cid} : \unity{\rx{w}{\cid}}{\ty}}
-
-  \inferrule[HeAdd]
-      {\Gamma, R_1, \eqs \vdash \phi_1 : \setit{\unity{\rkey_1}{\ty_1}} \\
-        \Gamma, R_2, \eqs \vdash \phi_2 : \setit{\unity{\rkey_2}{\ty_2}} \\
-        \uncorr{\rkey_1}{\rkey_2}
-      }
-      {\Gamma, R_1;R_2, \eqs \vdash \phi_1 \fplus \phi_2 : \setit{\unity{\rkey_1 \oplus \rkey_2}{\ty_1 \cup \ty_2}}}
+  \inferrule[Invert]
+            {\Gamma, \eqs \vdash \phi : \ty}
+            {\Gamma, \eqs \vdash -\phi : \ty}
 
   \inferrule[Join]
-      {\Gamma, R_1, \eqs \vdash \phi_1 : \ty_1 \\ \Gamma, R_2, \eqs \vdash \phi_2 : \ty_2}
-      {\Gamma, R_1;R_2, \eqs \vdash \phi_1 \bop \phi_2 : \ty_1 \cup \ty_2}
+      {\Gamma, \eqs \vdash \phi_1 : \dty{R_1}{\eset_1}\\
+       \Gamma, \eqs \vdash \phi_2 : \dty{R_2}{\eset_2}\\
+        R \defeq (R_1 - (R_2 \cup \eset_2) \cup (R_2 - (R_1 \cup \eset_1)) \\ 
+        \eset \defeq (R_1 \cup R_2 \cup \eset_1 \cup \eset_2) - R }
+      {\Gamma, \eqs \vdash \phi_1 \fplus \phi_2 : \dty{R}{\eset}}
+
+  \inferrule[Smear]
+      {\Gamma, \eqs \vdash \phi_1 : \dty{R_1}{\eset_1}\\
+       \Gamma, \eqs \vdash \phi_2 : \dty{R_2}{\eset_2}}
+      {\Gamma, \eqs \vdash \phi_1 \bop \phi_2 : \dty{\varnothing}{R_1 \cup R_2 \cup \eset_1 \cup \eset_2}}
 
   \inferrule[Eq]
-            {\eqs \models \phi \eop \phi' \\ \Gamma, R, \eqs \vdash \phi' : \ty}
-            {\Gamma, R, \eqs \vdash \phi : \ty}          
+            {\eqs \models \phi \eop \phi' \\ \Gamma, \eqs \vdash \phi' : \ty}
+            {\Gamma, \eqs \vdash \phi : \ty}          
 \end{mathpar}
@@
   \inferrule[Mesg]
-      {\Gamma, R, \eqs \vdash \toeq{\elab{e}{\cid}} : \ty}
-      {\Gamma, \eqs \vdash \xassign{x}{e}{\cid} : (\Gamma;x:\ty),R}
+      {\Gamma, \eqs \vdash \toeq{\elab{e}{\cid}} : \ty}
+      {\Gamma, \eqs \vdash \xassign{x}{e}{\cid} : (\Gamma;x:\ty)}
 
   \inferrule[Seq]
-      {\Gamma_1,\eqs \vdash \cmd_1 : \Gamma_2,R_1 \\
-        \Gamma_2,\eqs \vdash \cmd_2 : \Gamma_3,R_2}
-      {\Gamma_1, \eqs \vdash \cmd_1;\cmd_2 : \Gamma_3,R_1;R_2}
+      {\Gamma_1,\eqs \vdash \cmd_1 : \Gamma_2 \\
+        \Gamma_2,\eqs \vdash \cmd_2 : \Gamma_3}
+      {\Gamma_1, \eqs \vdash \cmd_1;\cmd_2 : \Gamma_3}
 \end{mathpar}
 
-\begin{comment}
-\section{Deprecated}
+\begin{definition}[Adversarial Views]
+  Given $\prog, \eqs, \Gamma$ with $\toeq{\prog} \models \eqs$ and $\varnothing, \eqs \vdash \prog : \Gamma$.
+  Then for all $H,C$, define:
+  $$
+  \view(\Gamma,\eqs,C,\prog) \defeq \bigcup_{X \in (\pow(\vars(\prog)_C)) - \varnothing} \ty
+     \qquad \text{where}\ \Gamma,\eqs \vdash \Sigma_{x \in X} x : \ty
+  $$
+\end{definition}
+
+\begin{definition}[Revelation]
+  $$
+  \begin{array}{ll}
+    \revealed(\ty,H) \defeq & \setit{f(\sx{w_1}{\cid_1},\ldots,\sx{w_n}{\cid_n}) \mid  \\
+      & \qquad \dty{\varnothing}{\eset} \in \ty\ \wedge\\
+      & \qquad f(\sx{w_1}{\cid_1},\ldots,\sx{w_n}{\cid_n}) \in \eset\ \wedge \\
+      & \qquad \exists 1 \le i \le n . \cid_i \in H} 
+  \end{array}
+  $$
+\end{definition}
+
+\begin{theorem}[Security Typing]
+  Given $\prog, \eqs, \Gamma$ with $\toeq{\prog} \models \eqs$ and $\varnothing, \eqs \vdash \prog : \Gamma$.
+  The $\prog$ satisfies {NIMO} for ideal functionality $\idealf$ if for all $H,C$, we have:
+  $$
+  \revealed(\view(\Gamma,\eqs,C,\prog),H) = \setit{\idealf}
+  $$
+\end{theorem}
+
+\subsection{Strategies}
 
 \begin{mathpar}
-  \inferrule[Mesg]
-            {}
-            {\Gamma, \eqs \vdash x : \Gamma(x)}
+  \inferrule[Sharing]
+      {
+        \eqs \models \mx{w_1}{\cid_1} \fplus \cdots \fplus \mx{w_n}{\cid_n} \eop \phi \\
+        \\
+        \exists R\ .\ (\forall M \in \pow(\setit{\mx{w_1}{\cid_1},\ldots,\mx{w_n}{\cid_n}})\ .\ 0 < |M| < n \implies \\
+          \qquad \Gamma, \eqs \vdash \Sigma_{x \in M} x : \unity{R'}{\eset} \ \wedge \ R' \subseteq R)
+      }   
+      {\Gamma,R,\eqs \vdash \cvec{\mx{w_1}{\cid_1},\ldots,\mx{w_n}{\cid_n}} : \sharet{\phi}}
 
-  \inferrule[Key]
-            {\sid:[\ldots,x_k,\ldots] \in \keys}
-            {\Gamma, \eqs \vdash x_k : \setit{\sharet{\sid}{k}}}
+  \inferrule[Homomorphic Addition]
+      {
+       \Gamma,R_1,\eqs \vdash \cvec{\mx{w^1_1}{\cid_1},\ldots,\mx{w^1_n}{\cid_n}} : \sharet{\phi_1} \\
+       \Gamma,R_2,\eqs \vdash \cvec{\mx{w^2_1}{\cid_1},\ldots,\mx{w^2_n}{\cid_n}} : \sharet{\phi_2} \\
+       \eqs \models \mx{w_1}{\cid_1} \eop \mx{w^1_1}{\cid_1} \fplus \mx{w^2_1}{\cid_1} 
+       \ \ldots\ 
+       \eqs \models \mx{w_n}{\cid_n} \eop \mx{w^1_n}{\cid_n} \fplus \mx{w^2_n}{\cid_n} 
+      }
+      {
+        \Gamma, R_1;R_2, \eqs \vdash \cvec{\mx{w_1}{\cid_1},\ldots,\mx{w_n}{\cid_n}} : \sharet{\phi_1 \fplus \phi_2}
+      }
 
-  \inferrule[Secret]
-            {}
-            {\Gamma, \eqs \vdash \sx{w}{\cid} : \setit{\sx{w}{\cid}}}
+ \inferrule[Secure Opening]
+      {
+        \Gamma,R_1,\eqs \vdash \cvec{\mx{w_1}{\cid_1},\ldots,\mx{w_n}{\cid_n}} : \sharet{\phi} \\
+        \eqs \models \px{w_1} \eop \mx{w_1}{\cid_1}
+        \ \ldots\
+        \eqs \models \px{w_n} \eop \mx{w_n}{\cid_n}\\
+        \Gamma,\eqs \vdash \phi : \unity{R_2}{\eset}
+      }
+      {
+        \Gamma, R_1;R_2, \eqs \vdash \cvec{\px{w_1},\ldots,\px{w_n}} : \secopent
+      }
 
-  \inferrule[Join]
-            {\Gamma, \eqs \vdash \phi_1 : \ty_1 \\ \Gamma, \eqs \vdash \phi_2 : \ty_2}
-            {\Gamma, \eqs \vdash \phi_1 \bop \phi_2 : \ty_1 \cup \ty_2}
-
-  \inferrule[Eq]
-            {\eqs \models \phi \eop \phi' \\ \Gamma, \eqs \vdash \phi' : \ty}
-            {\Gamma, \eqs \vdash \phi : \ty}
-
-  \inferrule[Share]
-            {\Gamma, \eqs \vdash x_1 : \sharet{\sid}{1} \\ \cdots \\ \Gamma, \eqs \vdash x_{n-1} : \sharet{\sid}{n-1} }
-            {\Gamma, \eqs \vdash \phi - x_1 - \cdots - x_{n-1} : [\phi]_\sid}
-
-  \inferrule[HeAdd]
-      {\Gamma, \eqs \vdash \phi_1 : \setit{\sharet{\sexp_1}{k}} \\
-       \Gamma, \eqs \vdash \phi_2 : \setit{\sharet{\sexp_2}{k}} }
-      {\Gamma, \eqs \vdash \phi_1 \fplus \phi_2 : \setit{\sharet{\sexp_1 \oplus \sexp_2}{k}}}
+ \inferrule[Declassification]
+      {
+        \Gamma,R,\eqs \vdash \cvec{\mx{w_1}{\cid_1},\ldots,\mx{w_n}{\cid_n}} : \sharet{\phi} \\
+        \eqs \models \px{w_1} \eop \mx{w_1}{\cid_1}
+        \ \ldots\
+        \eqs \models \px{w_n} \eop \mx{w_n}{\cid_n}
+      }
+      {
+        \Gamma, R, \eqs \vdash \cvec{\px{w_1},\ldots,\px{w_n}} : \declasst{\phi}
+      }
 \end{mathpar}
-
-\begin{mathpar}
-  \inferrule[Enc]
-      {\Gamma, \eqs \vdash \toeq{\elab{e}{\cid}} : [\phi]_{\sid}}
-      {\Gamma,\Shares,\eqs \vdash \xassign{x}{e}{\cid} : (\Gamma;x:\sharet{\sid}{n}),(\Shares;[\phi]_\sid)}
-
-  \inferrule[Mesg]
-      {\Gamma, \eqs \vdash \toeq{\elab{e}{\cid}} : \ty}
-      {\Gamma,\Shares,\eqs \vdash \xassign{x}{e}{\cid} : (\Gamma;x:\ty),\Shares}
-
-  \inferrule[Seq]
-      {\Gamma_1,\Shares_1,\eqs \vdash \prog_1 : \Gamma_2,\Shares_2 \\
-        \Gamma_2,\Shares_2,\eqs \vdash \prog_2 : \Gamma_3,\Shares_3}
-      {\Gamma_1,\Shares_1,\eqs \vdash \prog_1;\prog_2 : \Gamma_3,\Shares_3}
-\end{mathpar}
-
-\begin{mathpar}
-  \inferrule
-      {}
-      {\Shares \vdash T \cup \setit{\sx{w}{\cid}},C \leadsto T, C.\sx{w}{\cid}}
-
-  \inferrule
-      {}
-      {\Shares \vdash T \cup \setit{\sharet{\sexp}{1},\ldots,\sharet{\sexp}{n}},C \leadsto T, C.\Shares(\sexp)}
-
-  \inferrule
-      {\setit{\sharet{\sexp_1 \oplus \sexp_2}{j}, \sharet{\sexp_1}{k}, \sharet{\sexp_2}{k}} \subseteq T}
-      {\Shares \vdash T,C \leadsto T \cup \setit{\sharet{\sexp_1 \oplus \sexp_2}{k}},C}
-
-  \inferrule
-      {\Shares \vdash T_1,C_1 \leadsto T_2,C_2 \\ \Shares \vdash T_2,C_2 \leadsto T_3,C_3 }
-      {\Shares \vdash T_1,C_1 \leadsto T_3,C_3 }
-\end{mathpar}
-\end{comment}

### 2025-05-09 esop25/new-cpj.tex
checkpoint

@@
 \newcommand{\revealed}{\mathit{revealed}}
+\newcommand{\sty}{\varsigma}
+\newcommand{\noise}{\mathrm{noise}}
+\newcommand{\owners}{\mathrm{owners}}
+\newcommand{\federation}{I_{\mathit{fed}}}
 
@@
 
-\begin{definition}[Adversarial Views]
+\begin{definition}[Views]
   Given $\prog, \eqs, \Gamma$ with $\toeq{\prog} \models \eqs$ and $\varnothing, \eqs \vdash \prog : \Gamma$.
-  Then for all $H,C$, define:
+  Define:
   $$
-  \view(\Gamma,\eqs,C,\prog) \defeq \bigcup_{X \in (\pow(\vars(\prog)_C)) - \varnothing} \ty
-     \qquad \text{where}\ \Gamma,\eqs \vdash \Sigma_{x \in X} x : \ty
+  \view(\Gamma,\eqs,X) \defeq \bigcup_{X' \in (\pow(X) - \varnothing)} \ty
+     \qquad \text{where}\ \Gamma,\eqs \vdash \Sigma_{x \in X'} x : \ty
   $$
@@
   Given $\prog, \eqs, \Gamma$ with $\toeq{\prog} \models \eqs$ and $\varnothing, \eqs \vdash \prog : \Gamma$.
-  The $\prog$ satisfies {NIMO} for ideal functionality $\idealf$ if for all $H,C$, we have:
+  Then $\prog$ satisfies {NIMO} for ideal functionality $\idealf$ if for all $H,C$, we have:
   $$
-  \revealed(\view(\Gamma,\eqs,C,\prog),H) = \setit{\idealf}
+  \revealed(\view(\Gamma,\eqs,\vars(\prog)_C),H) = \setit{\idealf}
   $$
@@
 
-\subsection{Strategies}
+\subsection{Strategy Typings}
+
+$$
+\inferrule
+    {\Gamma, \eqs \vdash \phi : \dty{R}{\eset} \\ R - R_I \neq \varnothing}
+    {\Gamma, \eqs \vdash \phi : \noise(R,I)}
+$$
 
@@
         \\
-        \exists R\ .\ (\forall M \in \pow(\setit{\mx{w_1}{\cid_1},\ldots,\mx{w_n}{\cid_n}})\ .\ 0 < |M| < n \implies \\
-          \qquad \Gamma, \eqs \vdash \Sigma_{x \in M} x : \unity{R'}{\eset} \ \wedge \ R' \subseteq R)
+        \exists R_2\ .\ (\forall M \in \pow(\setit{\mx{w_1}{\cid_1},\ldots,\mx{w_n}{\cid_n}})\ .\ 0 < |M| < n \implies \\
+          \qquad \Gamma, \eqs \vdash \Sigma_{x \in M} x : \noise(R,\owners(M)) \ \wedge \ R \subseteq R_2)
       }   
-      {\Gamma,R,\eqs \vdash \cvec{\mx{w_1}{\cid_1},\ldots,\mx{w_n}{\cid_n}} : \sharet{\phi}}
+      {\Gamma,\eqs \vdash R_1,\Delta \leadsto R_1;R_2, \Delta; \cvec{\mx{w_1}{\cid_1},\ldots,\mx{w_n}{\cid_n}} : \sharet{\phi}}
 
@@
       {
-       \Gamma,R_1,\eqs \vdash \cvec{\mx{w^1_1}{\cid_1},\ldots,\mx{w^1_n}{\cid_n}} : \sharet{\phi_1} \\
-       \Gamma,R_2,\eqs \vdash \cvec{\mx{w^2_1}{\cid_1},\ldots,\mx{w^2_n}{\cid_n}} : \sharet{\phi_2} \\
+       \Delta(\cvec{\mx{w^1_1}{\cid_1},\ldots,\mx{w^1_n}{\cid_n}}) = \sharet{\phi_1} \\
+       \Delta(\cvec{\mx{w^2_1}{\cid_1},\ldots,\mx{w^2_n}{\cid_n}}) = \sharet{\phi_2} \\
        \eqs \models \mx{w_1}{\cid_1} \eop \mx{w^1_1}{\cid_1} \fplus \mx{w^2_1}{\cid_1} 
@@
       {
-        \Gamma, R_1;R_2, \eqs \vdash \cvec{\mx{w_1}{\cid_1},\ldots,\mx{w_n}{\cid_n}} : \sharet{\phi_1 \fplus \phi_2}
+        \Gamma, \eqs \vdash R,\Delta \leadsto R,\Delta;\cvec{\mx{w_1}{\cid_1},\ldots,\mx{w_n}{\cid_n}} : \sharet{\phi_1 \fplus \phi_2}
       }
@@
       {
-        \Gamma,R_1,\eqs \vdash \cvec{\mx{w_1}{\cid_1},\ldots,\mx{w_n}{\cid_n}} : \sharet{\phi} \\
+        \Delta(\cvec{\mx{w_1}{\cid_1},\ldots,\mx{w_n}{\cid_n}}) = \sharet{\phi} \\
         \eqs \models \px{w_1} \eop \mx{w_1}{\cid_1}
@@
         \eqs \models \px{w_n} \eop \mx{w_n}{\cid_n}\\
-        \Gamma,\eqs \vdash \phi : \unity{R_2}{\eset}
+        \Gamma,\eqs \vdash \phi : \noise(R_2, \federation)
       }
       {
-        \Gamma, R_1;R_2, \eqs \vdash \cvec{\px{w_1},\ldots,\px{w_n}} : \secopent
+        \Gamma, \eqs \vdash R_1;\Delta \leadsto R_1;R_2, \Delta; \cvec{\px{w_1},\ldots,\px{w_n}} : \secopent
       }
@@
       {
-        \Gamma,R,\eqs \vdash \cvec{\mx{w_1}{\cid_1},\ldots,\mx{w_n}{\cid_n}} : \sharet{\phi} \\
+        \Delta(\cvec{\mx{w_1}{\cid_1},\ldots,\mx{w_n}{\cid_n}}) =  \sharet{\phi} \\
         \eqs \models \px{w_1} \eop \mx{w_1}{\cid_1}
@@
       {
-        \Gamma, R, \eqs \vdash \cvec{\px{w_1},\ldots,\px{w_n}} : \declasst{\phi}
+        \Gamma, \eqs \vdash R, \Delta \leadsto R, \Delta;\cvec{\px{w_1},\ldots,\px{w_n}} : \declasst{\phi}
       }
 \end{mathpar}
+
+\begin{theorem}
+  Given $\prog, \eqs, \Gamma$ with $\toeq{\prog} \models \eqs$ and $\varnothing, \eqs \vdash \prog : \Gamma$.
+  Suppose also $\Gamma,\eqs \vdash \varnothing,\varnothing \Rightarrow \Delta$, and define:
+  $$\eset = \{ \phi \mid \exists \nu \in \dom(\Delta) . \Delta(\nu) \mapsto \declasst{\phi} \}$$.
+  Then for any $H,C$, with:
+  $$
+  X_C \defeq \{ x \mid x \in \vars(\prog)_C \wedge \neg \exists \nu \in \dom(\Delta) . x \in \nu \}
+  $$
+  we have:
+  $$
+  \eset \cup  \revealed(\view(\Gamma,\eqs,X_C),H) = \revealed(\view(\Gamma,\eqs,\vars(\prog)_C),H)
+  $$
+\end{theorem}

### 2025-05-10 esop25/new-cpj.tex
checkpoint

@@
 \newcommand{\federation}{I_{\mathit{fed}}}
+\newcommand{\tjoin}{\bowtie}
 
@@
 
+The types of expressions $\elab{\be}{\cid}$ reflect their dependencies which
+may include variables from clients beside $\cid$, so typings are in terms
+of interpretations $\toeq{\elab{\be}{\cid}}$. Types are of the form $\dty{R}{\eset}$,
+where $\eset$ is a set of expressions $\phi$, and intuitively the typing
+$\toeq{\elab{\be}{\cid}} : \dty{R}{\eset}$ means that the marginal distribution
+of its variable dependencies is equivalent to the (uniform) marginal distribution of
+variables in $R$. If $R$ is empty this means that the distribution is dependent
+on the marginal distribution of variables in $\eset$. So for example:
+\begin{mathpar}
+  \sx{w}{\cid} : \dty{\varnothing}{\setit{\sx{w}{\cid}}}
+
+  \rx{w}{\cid} : \dty{\setit{\rx{w}{\cid}}}{\varnothing}
+
+  \sx{w}{\cid} - \rx{w}{\cid} : \dty{\setit{\rx{w}{\cid}}}{\setit{\sx{w}{\cid}}}
+\end{mathpar}
+The typing rules are an approximation of dependencies, so wherever dependencies
+may be revealed, the system (conservatively) reflects this. So for example:
+\begin{mathpar}
+  \sx{w}{\cid} \ftimes \rx{w}{\cid} : \dty{\varnothing}{\setit{\rx{w}{\cid}, \sx{w}{\cid}}}
+\end{mathpar}
 \subsection{Primitive Typing Rules}
+
+The primitive typing rules are precise and expressive and reflect no particular
+implementation strategies, they just track dependencies as reflected in types.
+The most important rules are for binary operators, where dependencies merge
+in interesting ways especially with summation. In particular, if two
+randomized values are summed, they remain randomized modulo disjointness
+of randomness. Otherwise, overlapping randomness is merged into the dependencies.
+For example:
+\begin{eqnarray*}
+& (\sx{w}{\cid_1} - \rx{w}{\cid_1}) + (\sx{w}{\cid_2} - \rx{w}{\cid_2}) : \\
+& \dty{\setit{\rx{w}{\cid_1},\rx{w}{\cid_2}}}{\setit{\sx{w}{\cid_2},\sx{w}{\cid_1}}}
+\end{eqnarray*}
+but:
+$$
+(\sx{w}{\cid_1} - \rx{w}{\cid_1}) + \rx{w}{\cid_1} :
+\dty{\varnothing}{\setit{\rx{w}{\cid_1},\sx{w}{\cid_1}}}
+$$
+
+\begin{definition}[Type Join]
+  Given $\dty{R_1}{\eset_1}$ and $\dty{R_2}{\eset_2}$, define
+  $\dty{R_1}{\eset_1} \tjoin \dty{R_2}{\eset_2} \defeq \dty{R}{\eset}$ where
+  $R = (R_1 - (R_2 \cup \eset_2) \cup (R_2 - (R_1 \cup \eset_1))$ 
+  and $\eset = (R_1 \cup R_2 \cup \eset_1 \cup \eset_2) - R $.
+  %\begin{eqnarray*}
+  %  R &\defeq& (R_1 - (R_2 \cup \eset_2) \cup (R_2 - (R_1 \cup \eset_1))\\
+  %  \eset &\defeq& (R_1 \cup R_2 \cup \eset_1 \cup \eset_2) - R 
+  %\end{eqnarray*}
+\end{definition}
 
@@
       {\Gamma, \eqs \vdash \phi_1 : \dty{R_1}{\eset_1}\\
-       \Gamma, \eqs \vdash \phi_2 : \dty{R_2}{\eset_2}\\
-        R \defeq (R_1 - (R_2 \cup \eset_2) \cup (R_2 - (R_1 \cup \eset_1)) \\ 
-        \eset \defeq (R_1 \cup R_2 \cup \eset_1 \cup \eset_2) - R }
-      {\Gamma, \eqs \vdash \phi_1 \fplus \phi_2 : \dty{R}{\eset}}
+       \Gamma, \eqs \vdash \phi_2 : \dty{R_2}{\eset_2}}%\\
+        %R \defeq (R_1 - (R_2 \cup \eset_2) \cup (R_2 - (R_1 \cup \eset_1)) \\ 
+        %\eset \defeq (R_1 \cup R_2 \cup \eset_1 \cup \eset_2) - R }
+      {\Gamma, \eqs \vdash \phi_1 \fplus \phi_2 : \dty{R_1}{\eset_1} \tjoin \dty{R_2}{\eset_2}}
 
@@
   Suppose also $\Gamma,\eqs \vdash \varnothing,\varnothing \Rightarrow \Delta$, and define:
-  $$\eset = \{ \phi \mid \exists \nu \in \dom(\Delta) . \Delta(\nu) \mapsto \declasst{\phi} \}$$.
+  $$\eset \defeq \{ \phi \mid \exists \nu \in \dom(\Delta) . \Delta(\nu) \mapsto \declasst{\phi} \}$$
+  Let $\iov(\prog) = S \cup M \cup P$. 
+  Then for any $H,C$, if for all $x \in (M \cup P)_C$ there exists
+  $\nu \in \dom(\Delta)$ with $x \in \nu$, we have:
+  $$
+  D = \revealed(\view(\Gamma,\eqs,\vars(\prog)_C),H)
+  $$
+\end{theorem}
+
+\begin{lemma}
+  Given $\prog, \eqs, \Gamma$ with $\toeq{\prog} \models \eqs$ and $\varnothing, \eqs \vdash \prog : \Gamma$.
+  Suppose also $\Gamma,\eqs \vdash \varnothing,\varnothing \Rightarrow \Delta$, and define:
+  $$\eset = \{ \phi \mid \exists \nu \in \dom(\Delta) . \Delta(\nu) \mapsto \declasst{\phi} \}$$
   Then for any $H,C$, with:
   $$
-  X_C \defeq \{ x \mid x \in \vars(\prog)_C \wedge \neg \exists \nu \in \dom(\Delta) . x \in \nu \}
+  X \defeq \{ x \mid x \in \vars(\prog)_C \wedge \neg \exists \nu \in \dom(\Delta) . x \in \nu \}
   $$
@@
   $$
-  \eset \cup  \revealed(\view(\Gamma,\eqs,X_C),H) = \revealed(\view(\Gamma,\eqs,\vars(\prog)_C),H)
+  \eset \cup  \revealed(\view(\Gamma,\eqs,X),H) = \revealed(\view(\Gamma,\eqs,\vars(\prog)_C),H)
   $$
-\end{theorem}
+\end{lemma}

### 2025-05-10 esop25/new-cpj.tex
checkpoint

@@
 $$
-
+To join types appropriately, we make the following definition.
 \begin{definition}[Type Join]
@@
 \end{definition}
+Then the rules for typing are as follows. The most interesting are the rules
+for summation that allow joining types and possibly retaining uniformity,
+and the rule for binary operations (including multiplication) that
+revert to the approximation of distributions of dependent expressions.
 

### 2025-05-10 esop25/new-cpj.tex
checkpoint

@@
 
-\section{Declassification Typing}
+\section{$\minifed$ Declassification Typing}
 
@@
   \inferrule[Seq]
-      {\Gamma_1,\eqs \vdash \cmd_1 : \Gamma_2 \\
-        \Gamma_2,\eqs \vdash \cmd_2 : \Gamma_3}
-      {\Gamma_1, \eqs \vdash \cmd_1;\cmd_2 : \Gamma_3}
+      {\Gamma_1,\eqs \vdash \prog_1 : \Gamma_2 \\
+        \Gamma_2,\eqs \vdash \prog_2 : \Gamma_3}
+      {\Gamma_1, \eqs \vdash \prog_1;\prog_2 : \Gamma_3}
 \end{mathpar}
-
+We note that typing tracks uniformity of marginal distributions of independent messages,
+but not their joint distributions. In order to capture this, we define a notion of
+dependency closures in views:  
 \begin{definition}[Views]
@@
 \end{definition}
-
+The goal of the analysis is to show that views contain only noise and the
+result of the ideal functionality. To this end we define a notion of revelation, where
+functions of honest secrets are available in the clear in adversarial
+views:
 \begin{definition}[Revelation]
@@
 \end{definition}
-
+Now we can state our security typing correctness result, which says that the only
+information revealed in adversarial views (besides noise) is the ideal functionality. 
 \begin{theorem}[Security Typing]
@@
 \subsection{Strategy Typings}
+
+The type system presented above is in a logical form and is not syntax-directed.
+Also, the definition of $\view$ involves a powerset construction with exponential
+complexity in the size of the program.
+
+As a step towards an algorithm, we can define specific strategies and analysis
+based on our type system that both reduces overall complexity and leverages SMT
+solving. This will form the foundation of a decidable and scalable type analysis
+for $\metaprot$.  We define \emph{vectors} of messages $\nu$ that participate in strategies,
+e.g., the collective shares of secret values.
+\begin{definition}
+  A \emph{message vector} $\nu$ is either a vector of distinct
+  broadcast messages $\cvec{\px{w_1},\ldots,\px{w_n}}$, or of unicast
+  messages $\cvec{\mx{w_1}{\cid_1},\ldots,\mx{w_n}{\cid_n}}$ where
+  each $\cid_i$ for $i \in \setit{\cid_1,\ldots,\cid_n}$ is distinct.  
+\end{definition}
+
+\emph{Strategy typings} and environments are then defined via the following syntax.
+$$
+\begin{array}{rcl@{\hspace{4mm}}r}
+\sty &::=& \sharet{\phi} \mid \declasst{\phi} \mid \secopent & \gdesc{Strategy Types} \\[1mm]
+\Delta &::=& \varnothing \mid \Delta; \nu : \sty & \gdesc{Strategy Type Environments}
+\end{array}
+$$
 

### 2025-06-05 code/bristol/translate.py
start on bristol circuits

@@ -1,0 +1,50 @@
+# TODO Add parameter for file path
+file_path = 'test.txt'
+
+print(f'test(prefix, [inputs...])')
+with open(file_path, 'r') as file:
+    # Number of gates, number of wires
+    # TODO Maybe validate?
+    header_line_1 = file.readline()
+    tokens = header_line_1.strip().split(' ')
+    num_gates = int(tokens[0])
+    num_wires = int(tokens[1])
+
+    # Number of inputs, number of bits per input
+    header_line_2 = file.readline()
+    tokens = header_line_2.strip().split(' ')
+    num_inputs = int(tokens[0])
+    assert len(tokens) == num_inputs + 1
+    input_lengths = [int(x) for x in tokens[1:]]
+
+    # Number of outputs, number of bits per output
+    header_line_3 = file.readline()
+    tokens = header_line_3.strip().split(' ')
+    num_outputs = int(tokens[0])
+    assert len(tokens) == num_outputs + 1
+    output_lengths = [int(x) for x in tokens[1:]]
+
+    # Should be blank
+    header_line_4 = file.readline()
+    assert header_line_4.strip() == ''
+
+    gate_counter = 0
+    while True:
+        gate_counter += 1
+        line = file.readline()
+        if not line: break
+        line = line.strip().split(' ')
+        gate_type = line[-1]
+        n_in_wires = int(line[0])
+        in_wires = [int(x) for x in line[2:n_in_wires + 2]]
+        n_out_wires = int(line[1])
+        out_wires = [int(x) for x in line[n_in_wires + 2:-1]]
+        if gate_type == 'AND':
+            print(f'andgmw(prefix ++ "_w{out_wires[0]}", prefix ++ "_w{in_wires[0]}", prefix ++ "_w{in_wires[1]}")', end='')
+        # TODO Add support for other gate types
+        else:
+            raise ValueError(f'Unknown gate type: {gate_type}')
+        if gate_counter < num_gates:
+            print(';')
+        else:
+            print('')

### 2025-06-05 esop25/sharety.tex
checkpoint

\section{Share Types}

\newcommand{\sharety}{\mathrm{shares}}
\newcommand{\pubty}{\mathrm{public}}
\newcommand{\genshares}{\mathrm{genshares}}
\newcommand{\recon}{\mathrm{recon}}
\newcommand{\xvec}{\nu}

\begin{eqnarray*}
  \genshares(\phi,y,\cid) &\defeq& \setit{\rx{y}{\cid}}, [\phi - \rx{y}{\cid}, \rx{y}{\cid}]\\
  \recon(\phi_1,\phi_2) &\defeq& \phi_1 \fplus \phi_2
\end{eqnarray*}

\begin{mathpar}
  \inferrule[Share]
      {\eqs \models \genshares(\phi_1,y,\cid) \eop R_2,[x,\phi_2] }
      {\Gamma,R_1,\eqs \vdash [x] : R_1;R_2, \Gamma; [x,\phi_2] : \sharety}

  \inferrule[Open]
      {\Gamma,R,\eqs \vdash [p_1,p_2] : \sharety}
      {\Gamma,R,\eqs \vdash [p_1,p_2] : R, \Gamma; [\recon(p_1,p_2)] : \pubty}

  \inferrule[Seq]
      {\Gamma_1,\eqs,R_1 \vdash \xvec_1 : R_2,\Gamma_2 \\
        \Gamma_2,\eqs,R_2 \vdash \xvec_2 : R_3,\Gamma_3}
      {\Gamma_1, \eqs,R_1 \vdash \xvec_1\xvec_2 : R_3\Gamma_3}
\end{mathpar}

\begin{mathpar}
  \inferrule[Hyp]
      {}
      {\Gamma,\eqs \vdash \xvec : \Gamma(\xvec)}

  \inferrule[HEAdd]
      {\Gamma,\eqs \vdash [\phi^1_1,\phi^1_2] : \sharety\\
        \Gamma,\eqs \vdash [\phi^2_1,\phi^2_2] : \sharety}
      {\Gamma,\eqs \vdash [\phi^1_1\fplus\phi^2_1,\phi^1_2\fplus\phi^2_2] : \sharety}

  \inferrule[HEAddPub]
      {\Gamma,\eqs \vdash [\phi] : \pubty\\
        \Gamma,\eqs \vdash [\phi_1,\phi_2] : \sharety}
      {\Gamma,\eqs \vdash [\phi_1\fplus\phi,\phi_2] : \sharety}

  \inferrule[HEMultPub]
      {\Gamma,\eqs \vdash [\phi] : \pubty\\
        \Gamma,\eqs \vdash [\phi_1,\phi_2] : \sharety}
      {\Gamma,\eqs \vdash [\phi_1\ftimes\phi,\phi_2\ftimes\phi] : \sharety}
\end{mathpar}

### 2025-06-10 esop25/sharety.tex
checkpoint

@@
 \newcommand{\recon}{\mathrm{recon}}
-\newcommand{\xvec}{\nu}
+\newcommand{\xvec}{\upsilon}
+\newcommand{\xdefs}{\mathcal{D}}
+\newcommand{\kwas}{\ \mathrm{as}\ }
+\newcommand{\shtrip}[3]{#1 \Vdash #2 \rhd #3}
+% \newcommand{\shtrip}{\htrip}
+\newcommand{\redxsub}{\varrho}
+
+\subsection{Logical Rules for $\minifed$}
+
+The logical typing rules enforce confidentiality. They impose
+coordination across clients and are defined in terms of vectors of
+expressions $\xvec$. Critically \emph{\bf all rules are defined modulo
+equivalence of vectors given $\eqs$}. We start by defining share
+generation and reconstruction. While we assume additive shares in this
+work, the type system is generally (intentionally) agnostic wrt to
+these abstractions, and so could be swapped out for, e.g., Shamir
+shares.
 
@@
 
+Generally the type system works more like an information flow type system
+than our previous work. The following rules define allowable interactions
+between clients. The idea is that all interactive messages must be categorized
+as either shares or public values in order for the protocol to be well-typed.
 \begin{mathpar}
   \inferrule[Share]
-      {\eqs \models \genshares(\phi_1,y,\cid) \eop R_2,[x,\phi_2] }
-      {\Gamma,R_1,\eqs \vdash [x] : R_1;R_2, \Gamma; [x,\phi_2] : \sharety}
+      {\eqs \models \genshares(\phi_1,y,\cid) \eop R_2,[\mx{z}{\cid},\phi_2] }
+      {\Gamma,R_1,\eqs \vdash [\mx{z}{\cid}] : R_1;R_2, \Gamma; [\mx{z}{\cid},\phi_2] : \sharety}
 
   \inferrule[Open]
-      {\Gamma,R,\eqs \vdash [p_1,p_2] : \sharety}
-      {\Gamma,R,\eqs \vdash [p_1,p_2] : R, \Gamma; [\recon(p_1,p_2)] : \pubty}
+      {\Gamma, \eqs \vdash [\px{y},\px{z}] : \sharety}
+      {\Gamma,R,\eqs \vdash [\px{y},\px{z}] : R, \Gamma; [\recon(\px{y},\px{z})] : \pubty}
 
@@
         \Gamma_2,\eqs,R_2 \vdash \xvec_2 : R_3,\Gamma_3}
-      {\Gamma_1, \eqs,R_1 \vdash \xvec_1\xvec_2 : R_3\Gamma_3}
+      {\Gamma_1, \eqs,R_1 \vdash \xvec_1\xvec_2 : R_3,\Gamma_3}
 \end{mathpar}
 
+In addition to the interactive rules, we have rules for local computation.
+Again, these are all modulo equivalence of vectors under $\eqs$. These rules
+allow standard homorphic encryption operations, and in particular ensure
+that all openings are secure in a NIMO sense. These can be combined with
+the Hoare logic to obtain a precise bound on information release, e.g.,
+that outputs are secure openings of the ideal functionality.
 \begin{mathpar}
@@
       {\Gamma,\eqs \vdash [\phi_1\ftimes\phi,\phi_2\ftimes\phi] : \sharety}
+
+  \inferrule[PubOp]
+      {\Gamma,\eqs \vdash [\phi_1] : \pubty\\
+        \Gamma,\eqs \vdash [\phi_2] : \pubty}
+      {\Gamma,\eqs \vdash [\phi_1\bop\phi_2] : \pubty}
 \end{mathpar}
+
+\subsection{Algorithmic Rules for $\metaprot$}
+
+The $\minifed$ type system is undecidable given validity up to equivalence of expression
+vectors, so as in previous work we want a decidable algorithm for $\metaprot$ that must
+rely on type annotations and explicit type preconditions. For example, a GMW or BGW
+sum gate should assume that the inputs are shares, and assert that the outputs are shares.
+To this end we extend the syntax of commands to include type annotations:
+$$
+\cmd ::= \cdots \mid [e,e] \kwas \genshares(\notg{\phi},e,e) \mid [e,e] : \sharety \mid
+     [e,e] : \pubty \mid [e] : \pubty
+$$
+These annotations need to be verified, and in general may rely on an SMT equivalence.
+Currently the only examples of the latter I've found are for the genshares case, but
+coercion forms for others could easily be added. The relevant equivalence for
+genshares can be inferred as a precondition in the Hoare logic, as follows.
+\begin{mathpar}
+  %\inferrule[Assert]
+  %    {}
+  %    {\htrip{\mx{e^1_1}{e^1_2} \eop \phi - \rx{y}{\cid} \wedge
+  %            \mx{e^2_1}{e^2_2} \eop \rx{y}{\cid}}
+  %           {[\mx{e^1_1}{e^1_2},\mx{e^2_1}{e^2_2}] \kwas \genshares(\phi,y,\cid)}{\eqtrue}}
+  \inferrule[CoerceShares]
+      {}
+      {\htrip{e_1 \eop \notg{\phi} - \rx{e}{e'} \wedge e_2 \eop \rx{e}{e'}}
+             {[e_1,e_2] \kwas \genshares(\notg{\phi},e,e')}{\eqtrue}}
+
+  %\inferrule[CoerceShares]
+  %    {}
+  %    {\htrip{e \eop e_1 \fplus e_2}{[e] \kwas \recon(e_1,e_2)}{\eqtrue}} 
+\end{mathpar}
+The typing rules are then given as follows. First, the type annotations, which
+rely on local computation typing which is purely syntax-directed given an empty
+constraint. $\xdefs$ is a mapping from $x$s to $\phi$s (their ``definitions'')
+which is just the explicit definition of $x$ in the protocol (see rule
+\TirName{Defn}).
+\begin{mathpar}
+  \inferrule[CoerceGenshare]
+      {e_1 \redx \mx{w_1}{\cid_1} \\ e_2 \redx \mx{w_2}{\cid_2} \\ e \redx w \\ e' \redx \cid}
+      {\shtrip{\Gamma, R, \xdefs}
+        {[e_1,e_2] \kwas \genshares(\notg{\phi},e,e')}
+        {\Gamma;[\mx{w_1}{\cid_1},\mx{w_2}{\cid_2}] : \sharety, R;\setit{\rx{w}{\cid}}, \xdefs} }
+
+  \inferrule[LocalShare]
+      {e_1 \redx \mx{w_1}{\cid_1} \\ e_2 \redx \mx{w_2}{\cid_2}  \\\\
+        \Gamma,\eqtrue \vdash [\xdefs(\mx{w_1}{\cid_1} ), \xdefs(\mx{w_2}{\cid_2})] : \sharety}
+      {\shtrip{\Gamma, R, \xdefs}{[e_1,e_2] : \sharety}{\Gamma;[\mx{w_1}{\cid_1},\mx{w_2}{\cid_2}] : \sharety,
+          R, \xdefs}}
+
+  \inferrule[Open]
+      {e_1 \redx \px{w_1} \\ e_2 \redx \px{w_2}  \\
+        \Gamma,\eqtrue \vdash [\xdefs(\px{w_1}), \xdefs(\px{w_2})] : \sharety}
+      {\shtrip{\Gamma, R, \xdefs}{[e_1,e_2] : \pubty}{\Gamma;[\px{w_1} \fplus \px{w_2}] : \pubty,
+          R, \xdefs}}
+
+  \inferrule[Public]
+      {e \redx \mx{w}{\cid} \\ \Gamma,\eqtrue \vdash [\xdefs(\mx{w}{\cid})] : \pubty}
+      {\shtrip{\Gamma, R, \xdefs}{[e] : \pubty}{\Gamma;[\mx{w}{\cid}] : \pubty,
+          R, \xdefs}}
+
+  \inferrule[Defn]
+      {e_1 \redx \mx{w}{\cid_1} \\ e_2 \redx \elab{\be}{\cid_2}}
+      {\shtrip{\Gamma, R, \xdefs}{e_1 := e_2}{\Gamma, R, \xdefs[\mx{w}{\cid_1} \mapsto \toeq{\elab{\be}{\cid_2}}]}}
+\end{mathpar}
+The remaining rules are as follows, where ``evaluating substitution''
+$\redxsub(\notg{R}),\redxsub(\notg{\Gamma})$ is defined to impose evaluation
+\cnote{probably should do this for $\peq$ as well}. I.e., given $\redxsub = [e_1/y_1 \cdots e_n/y_n]$:
+    $$
+    \redxsub(\notg{\Gamma}) = \Gamma \ \text{where}\ \notg{\Gamma}[e_1/y_1 \cdots e_n/y_n] \redx \Gamma
+    $$
+and so on.:
+\begin{mathpar}
+  \inferrule[App]
+      {f : \Pi y_1,\ldots,y_n. \notg{\Gamma_1} \rightarrow (\notg{R'},\notg{\Gamma_2}) \\
+       \redxsub = [e_1/y_1 \cdots e_n/y_n]}
+      {\shtrip{\Gamma;\redxsub(\notg{\Gamma_1}), R, \xdefs}{f(e_1,\ldots,e_n)}{\Gamma;\redxsub(\notg{\Gamma_2}), R;\redxsub(\notg{R'}), \xdefs}}
+
+  \inferrule[Seq]    
+      {\shtrip{\Gamma_1,R_1,\xdefs_1}{\cmd_1}{\Gamma_2,R_2,\xdefs_2} \\
+       \shtrip{\Gamma_2,R_2,\xdefs_2}{\cmd_2}{\Gamma_3,R_3,\xdefs_3}}
+      {\shtrip{\Gamma_1,R_1,\xdefs_1}{\cmd_1;\cmd_2}{\Gamma_3,R_3,\xdefs_3}}
+
+  \inferrule[Let]
+      {\shtrip{\Gamma_1,R_1,\xdefs_1}{\cmd[e/y]}{\Gamma_2,R_2,\xdefs_2}}
+      {\shtrip{\Gamma_1,R_1,\xdefs_1}{\elet{y}{e}{\cmd}}{\Gamma_2,R_2,\xdefs_2}}
+\end{mathpar}
+The \TirName{App} rule relies on function typings, which are a dependent type form
+$\Pi y_1,\ldots,y_n. \notg{\Gamma_1} \rightarrow (\notg{R},\notg{\Gamma_2})$ where
+$y_1,\ldots,y_n$ are the expression parameters, $\notg{\Gamma_1}$ is the
+generalized typing precondition, and $(\notg{R},\notg{\Gamma_2})$ is the generalized
+typing postconditions including any randomness consumed by OTP encryption.
+Function typings are computed as follows (compositionally):
+$$
+\inferrule[Fn]
+    {\codebase(f) = y_1 : \tau_1, \ldots, y_n : \tau_n, \cmd \\ \mv_1,\ldots,\mv_n = \fresh(\tau_1,\ldots,\tau_n) \\
+      \redxsub = [\mv_1/y_1 \cdots \mv_n/y_n] \\
+     \shtrip{\redxsub(\notg{\Gamma_1}), \varnothing, \varnothing}{\cmd}{\redxsub(\notg{\Gamma_2}), \redxsub(\notg{R}),\xdefs}}
+    {f : \Pi y_1,\ldots,y_n. \notg{\Gamma_1} \rightarrow (\notg{R},\notg{\Gamma_2})}
+ $$
+
+Note that in rule $\TirName{Fn}$, the precondition $\notg{\Gamma_1}$
+and the typing annotations that may contribute to
+$(\notg{R},\notg{\Gamma_2})$ need to be specified somewhere, a style
+choice. Currently they are being written as additional clauses to the
+\ttt{precondition} and \ttt{postcondition} of functions, e.g.:
+\begin{verbatimtab}
+precondition: ([m[x++"s"]@1,m[x++"s"]@2] : shares;
+               [m[y++"s"]@1,m[y++"s"]@2] : shares)
+sum(z:string, x:string, y:string){ _sum(z,x,y,1); _sum(z,x,y,2) }
+postcondition: ( [m[z++"s"]@1,m[z++"s"]@2] : shares),
+               ( SHARED(z) == SHARED(x) + SHARED(y) )
+\end{verbatimtab}
+This example illustrates how we can write both confidentiality and correctness properties--
+or not include either as desired. That is, the confidentiality type system will ensure
+gradual release, without any SMT pre or postconditions.

### 2025-06-10 esop25/sharety.tex
checkpoint

@@
 \end{mathpar}
-The remaining rules are as follows, where ``evaluating substitution''
-$\redxsub(\notg{R}),\redxsub(\notg{\Gamma})$ is defined to impose evaluation
-\cnote{probably should do this for $\peq$ as well}. I.e., given $\redxsub = [e_1/y_1 \cdots e_n/y_n]$:
+The remaining rules are as follows, with the rule for application
+being the most interesting as usual. Since we allow dependent
+generalization in function types, as in previous work (and as in the
+Hoare logic) we equate expressions up to the result of their
+evaluation. To make rules a little more succint we define ``evaluating
+substitution'' $\redxsub(\notg{R}),\redxsub(\notg{\Gamma})$ to also
+impose evaluation \cnote{probably should do this for $\peq$ as
+  well}. I.e., given $\redxsub = [e_1/y_1 \cdots e_n/y_n]$:
     $$

### 2025-10-17 code/SMT/eqs.py
checkpoint

@@ -30,7 +30,7 @@ s.check()
 # start_c1 = time.perf_counter()
 
 # correctness (should be unsat)
-s.check(x + y != o)  
+s.check(x + y != o)
 
 # stop_c1 = time.perf_counter()
 

### 2025-10-29 code/cvc5mvn/src/main/scala/plaid/logic/Evaluator.scala
documentation

@@ -4,6 +4,7 @@ import plaid.ast.*
 
 import scala.jdk.CollectionConverters.*
 
+/** Support for reducing an AST using variable substitution, expanding function calls, etc. */
 case class Evaluator(
   program: Program,
   bindings: Map[Identifier, Expr] = Map()) {
@@ -14,6 +15,7 @@ case class Evaluator(
     case (Str(s1), ConcatExpr(Str(s2), e2)) => ConcatExpr(Str(s1 + s2), e2)
     case _ => ConcatExpr(e1, e2)
 
+  /** Reduces an expression as much as possible in the context of this evaluator's Program and bindings. */
   def expression(e: Expr): Expr = e match
     case str: Str => str
     case num: Num => num
@@ -57,6 +59,7 @@ case class Evaluator(
     case _ =>
       throw new IllegalArgumentException("Bad Expression")
 
+  /** Reduces a command as much as possible in the context of this evaluator's Program and bindings. */
   def command(cmd: Cmd): Cmd = cmd match
     case AssignCmd(e1, e2) => AssignCmd(expression(e1), expression(e2))
     case CallCmd(fn, parms) =>
@@ -72,6 +75,7 @@ case class Evaluator(
     case AssertCmd(e1, e2, e3) => AssertCmd(expression(e1), expression(e2), expression(e3))
     case _ => throw new IllegalArgumentException("Bad instruction")
 
+  /** Reduces a constraint as much as possible in the context of this evaluator's Program and bindings. */
   def constraint(c: Constraint): Constraint = c match
     case AndConstraint(e1, e2) => AndConstraint(constraint(e1), constraint(e2))
     case NotConstraint(e) => NotConstraint(constraint(e))

### 2025-10-30 toplas26/lang.tex
checkpoint

--- 2025-01-26 esop25/lang.tex
@@
 \end{definition}
+
+\paragraph{Sanity conditions on protocols} To simplify
+the presentation we will disregard protocols that are semantically
+ill-defined. That is, we generally assume that for any
+configuration $(\store, \prog)$ there exists $\store$
+with $(\store, \prog) \redxs \store$. This rules out
+use of undefined variables and, in the adversarial semantics,
+programs consisting of just a failing $\assert$.

### 2025-10-30 toplas26/smt.tex
checkpoint

--- 2025-01-26 esop25/smt.tex
@@
-\section{$\minicat$ Constraint Verification}
+\section{$\minicat$ SMT-Based Hoare Logic}
 \label{section-smt}
@@
 fields by using a more general form of SMT constraint programming,
-Satisfiability Modulo Finite Fields \cite{SMFF}. As we
-will show in subsequent sections, this interpretation supports
-correctness guarantees, and also static type analyses for enforcing
-confidentiality and integrity properties.
+Satisfiability Modulo Finite Fields \cite{SMFF}. As we will show in
+subsequent sections, this interpretation supports automated
+enforcement of correctness guarantees, and also static type analyses
+for enforcing confidentiality and integrity properties.
+In this section we formulate a Hoare Logic for $\minicat$
+as a foundation for protocol verification.
 
@@
 
-We introduce the following syntax of SMT-style constraints
-for $\minicat$:
+We introduce the following syntax of SMT-style constraints for
+$\minicat$:
 $$
@@
   \phi ::= x \mid \phi \fplus \phi \mid \phi \fminus \phi \mid \phi \ftimes \phi  \qquad
-  \eqs ::= \phi \eop \phi \mid \eqs \wedge \eqs \mid \neg\eqs
+  \eqs ::= \phi \eop \phi \mid \eqs \wedge \eqs \mid \eqs \vee \eqs \mid \neg\eqs \mid \eqs \impl \eqs
 \end{array}
@@
   constraints}{fig-toeq}
-\small{
 \begin{mathpar}
@@
   \toeq{\prog_1;\prog_2} = (\toeq{\prog_1} \wedge \toeq{\prog_2})
-\end{mathpar}}
+\end{mathpar}
 \end{fpfig}
@@
 via a constraint $\eqspre$, i.e., $\preproc(\store) \iff \store
-\models \eqspre$ for all $\store$.  The correctness of the
-SMT interpretation of programs is characterized as follows.
-\begin{theorem}
-  \label{theorem-toeq}
-  $\store$ is a model of $\eqspre \wedge \toeq{\prog}$ iff $\store \in \runs(\prog)$.
-\end{theorem}
+\models \eqspre$ for all $\store$. Intuitively, a protocol satisfies
+a correctness (post)condition $\eqs$ if its SMT interpretation
+entails $\eqs$.
+
+%The correctness of the
+%SMT interpretation of programs is characterized as follows.
+%\begin{theorem}
+%  \label{theorem-toeq}
+%  $\store$ is a model of $\eqspre \wedge \toeq{\prog}$ iff $\store \in \runs(\prog)$.
+%\end{theorem}
 
@@
 publicly reveals the sums all their shares, and the sum of those is
-the protocol output for each party.  {\footnotesize
+the protocol output for each party.  
 $$
@@
 \end{array}
-$$}
+$$
 Letting $\prog$ be this protocol, we can verify correctness
@@
 $$
+
+\noindent We can establish a logical identity between the models of a program and
+its (non-adversarial) runs as follows.   
+\begin{theorem}
+  \label{theorem-toeq}
+  $\sigma \models \eqspre \wedge \toeq{\prog}$ iff $\store \in \runs(\prog)$.
+\end{theorem}
+\noindent This result follows as a consequence of a more general Lemma
+that incorporates a notion of preconditions. 
+\begin{lemma}
+  \label{lemma-toeq}
+  Given $\eqs, \prog, \store$ with $\store \models \eqs$.
+  Then $(\store,\prog) \redxs \store'$ iff $\store' \models
+  \eqs \wedge \toeq{\prog}$.
+\end{lemma}
+\cnote{Prove me?}
+
+\subsection{$\minicat$ Hoare Logic}
+
+When reasoning about MPC protocols it is important to reason about
+postconditions, e.g., to verify correctness conditions. It is also
+important to reason about preconditions of protocols, e.g., to express
+properties of assumed initialization protocols such as Beaver Triples.
+To support this reasoning we introduce a Hoare logic for $\minicat$,
+where $\htrip{\eqs_1}{\prog}{\eqs_2}$ means that $\eqs_1,\eqs_2$ are
+the preconditions and postconditions, respectively, of $\prog$.
+A semantic notion of validity for Hoare triples is defined as
+follows:
+
+\minicathtripfig
+
+\begin{definition}
+$\htrip{\eqs_1}{\prog}{\eqs_2}$ is \emph{valid} iff $\eqs_1 \wedge \toeq{\prog} \models \eqs_2$.
+\end{definition}
+
+The following Lemma establishes that this is a meaningful (and
+standard) definition. It follows by Lemma \ref{lemma-toeq}
+and transitivity of $\models$.
+\begin{lemma}
+  \label{lemma-hoare}
+  Given $\eqs_1, \eqs_2, \prog, \store$ with $\store \models \eqs_1$
+  and $\eqs_1 \wedge \toeq{\prog} \models \eqs_2$. Then
+  $(\store,\prog) \redxs \store'$ implies $\store' \models
+  \eqs_2$.
+\end{lemma}
+
+Now we can introduce a validity-preserving derivation system for the
+$\minicat$ Hoare logic. Since the language is computationally simple
+the Hoare logic can be relatively simple and mostly standard.
+The derivation rules are specified in Figure \ref{fig-minicathtrip}.
+They are mostly familiar except perhaps for \TirName{Assert}, which
+just enforces the expected equality as a precondition, and
+\TirName{Hyp}. It is unknown to the authors and not immediately
+obvious whether \TirName{Hyp} is derivable from other rules, but
+it is validity preserving as we show below. Its main use in this
+paper will be to enable an automated Hoare logic check we develop
+later (Section \ref{section-metahtrip}).
+\begin{lemma}
+  If $\htrip{\eqs_1}{\prog}{\eqs_2}$ is derivable, then it is valid.
+\end{lemma}
+
+\begin{proof}
+  Assume $\htrip{\eqs_1}{\prog}{\eqs_2}$ is derivable without \TirName{Hyp}.
+  Then $\eqs_1 \wedge \toeq{\prog} \models \eqs_2$, since all the other rules
+  are well-known to be validity preserving. 
+  Now since $\eqs_1 \wedge \toeq{\prog} \models \eqs_2$ by assumption, we have
+  $
+  \eqs_1 \wedge (\eqs_2 \impl \eqs) \wedge \toeq{\prog} \models \eqs_2 
+  $,
+  and clearly:
+  $$
+  \eqs_1 \wedge (\eqs_2 \impl \eqs) \wedge \toeq{\prog} \models \eqs_2 \wedge (\eqs_2 \impl \eqs)
+  $$
+  and since $\eqs_2 \wedge (\eqs_2 \impl \eqs)$ iff
+  $\eqs_2 \wedge \eqs$, we have
+  $\eqs_1 \wedge (\eqs_2 \impl \eqs) \wedge \toeq{\prog} \models \eqs_2 \wedge \eqs$.
+\end{proof}

### 2025-10-30 toplas26/metalang.tex
checkpoint

\section{The $\metaprot$ Language and Automated Hoare Logic}
\label{section-metalang}

The $\metaprot$ language \cite{skalka-near-ppdp24} includes structured
data and function definitions for defining composable protocol
elements at a higher level of abstraction than $\minicat$.  The
$\metaprot$ language is a \emph{metalanguage} aka metaprogramming
language, where an $\fedprot$ protocol is the result of
computation. In addition to these declarative benefits of $\metaprot$,
component definitions support compositional verification of larger
protocols. Separate verification of well-designed components results
in confidentiality and integrity properties to be recorded in their
types, allowing for significant reduction of SMT verification in whole
program analysis, as we will discuss with extended examples in Section
\ref{section-examples}.

\metaprotsyntaxfig

\subsection{Syntax and Semantics}

The syntax of $\metaprot$ is defined in Figure
\ref{fig-metaprotsyntax}.  It includes a syntax of values $\mv$ that
include client ids $\cid$, identifier strings $w$, expressions $\be$
in field $\mathbb{F}_p$, record values, and $\minicat$ variables
$x$. $\metaprot$ expression forms allow dynamic construction of these
values. $\metaprot$ \emph{instruction} forms allow dynamic
construction of $\minicat$ protocols $\prog$ that incorporate expression
evaluation. The syntax also supports definitions of functions that
compute values $\mv$ and, as a distinct form, functions that compute
protocols $\prog$.  Formally, we consider a complete metaprogram to include
both a codebase and a ``main'' program that uses the codebase.
We disallow recursion, mainly to guarantee decidability
of type dependence (Section \ref{section-metalangty}).
\begin{definition}
A \emph{codebase} $\codebase$ is a list of function 
declarations. We write $ \codebase(f) = y_1,\ldots,y_n,\ e$
if $f(y_1,\ldots,y_n) \{ e \} \in \codebase$, and we
write  $ \codebase(f) = y_1 : \tau_1, \ldots, y_n : \tau_n, \instr$
if $f(y_1 : \tau_1, \ldots, y_n : \tau_n) \{ \instr \} \in \codebase$.
%A \emph{metaprogram}, aka \emph{metaprotocol}  is a pair of a 
%codebase and expression $\codebase, e$. We may omit
%$\codebase$ if it is clear from context.  
\end{definition}

\metaprotexprsemanticsfig

We define a big-step evaluation relation $\redx$ in Figures
\ref{fig-metaprotexprsemantics} and \ref{fig-metaprotinstrsemantics}
for expressions and instructions, respectively.  In this definition we
write $e[\mv/y]$ and $\instr[\mv/y]$ to denote the substitution of
$\mv$ for free occurrences of $y$ in $e$ or $\instr$ respectively. The
rules are mostly standard. Note that we do not include an evaluation
rule for the form $\eqcast{\mx{e}{e}}{\notg{\phi}}$ which is a type
annotation that we assume is erased from programs prior to
evaluation. We defer discussion of this form, as well as the syntactic
category $\notg{\phi}$, to the next Section.

\metaprotinstrsemanticsfig

\subsection{$\metaprot$ Automated Abstract Hoare Logic}

\begin{verbatimtab}
precondition: ( m[x++"foo"]@1 == 3 )
bar(x:string){ m[x++"bar"]@2 := (m[x++"foo"] + m[x++"baz"])@1 }
postcondition: ( [m[z++"bar"]@2 == m[x++"baz"]@1 + 3 )
\end{verbatimtab}

\peqfig

\begin{lemma}
  \label{lemma-genentails}
  If $\forall y_1:\tau_1,\ldots,y_n:\tau_n . \peq_1 \models \peq_2$ and
  $\peq_1[\mv_1/y_1]\cdots[\mv_n/y_n] \redx \eqs_1$ and
  $\peq_2[\mv_1/y_1]\cdots[\mv_n/y_n] \redx \eqs_2$, then
  $\eqs_1 \models \eqs_2$.  
\end{lemma}

\begin{fpfig}[t]{Algorithmic $\metaprot$ Abstract Hoare Triple Deduction Rules}{fig-metahtrip}
\begin{mathpar}
\inferrule[GenEntails]
          {\mv_1,\ldots,\mv_n = \fresh(\tau_1,\ldots,\tau_n) \\ \peq_1[\mv_1/y_n \cdots \mv_n/y_n] \redx \eqs_1 \\
           \peq_2[\mv_1/y_n \cdots \mv_n/y_n] \redx \eqs_2 \\ \eqs_1 \models \eqs_2}
          {\forall y_1:\tau_1,\ldots,y_n:\tau_n . \peq_1 \models \peq_2 }

  \inferrule[Mesg]
            {}
            {\htrip{\eqtrue}{\xassign{e_1}{e_2}{e_3}}{e_1 \eop \toeq{\elab{e_2}{e_3}}}}

  %\inferrule[Encode]
  %          {\mx{e_1}{e_2} \redx x \\ \notg{\phi} \redx \phi \\
  %            \eqs \models x \eop \phi\\
  %            \atj{\phi}{R}{\ty}}
  %          {\mtj{\eqcast{\mx{e_1}{e_2}}{\notg{\phi}}}{\eqs}{(x : \ty)}{R}{\varnothing}{\eqs}}
  %
  \inferrule[Assert]
            {}
            {\htrip{\toeq{\elab{e_1}{e_3}} \eop \toeq{\elab{e_2}{e_3}}}{\elab{\assert{e_1 = e_2}}{e_3}}{\eqtrue}}

  \inferrule[Seq]          
            {\htrip{\peq_1^1}{\cmd_1}{\peq_2^1} \\ \htrip{\peq_1^2}{\cmd_2}{\peq_2^2}}
            {\htrip{\peq_1^1 \wedge (\peq_2^1 \impl \peq_1^2)}{\cmd_1;\cmd_2}{\peq_2^1 \wedge \peq_2^2}}

  \inferrule[Let]
            {\htrip{\peq_1}{\cmd[e/y]}{\peq_2}}
            {\htrip{\peq_1}{\elet{y}{e}{\cmd}}{\peq_2}}

  \inferrule[App]
            {\htrip{\peq_1}{f(y_1:\tau_1,\ldots,y_n:\tau_n)}{\peq_2}}
            {\htrip{\peq_1[e_1/y_1 \cdots e_n/y_n]}{f(e_1,\ldots,e_n)}{\peq_2[e_1/y_1 \cdots e_n/y_n]}}

  \inferrule[Fn]
            {\codebase(f) = y_1 : \tau_1, \ldots, y_n : \tau_n, \cmd \\ \htrip{\peq_1}{\cmd}{\peq_2}}
            {\htrip{\peq_1}{f(y_1:\tau_1,\ldots,y_n:\tau_n)}{\peq_2}}

  \inferrule[Hardpack]
            {\precond(f) = \peq_1 \\ \postcond(f) = \peq_2 \\
              \htrip{\peq_1'}{f(y_1:\tau_1,\ldots,y_n:\tau_n)}{\peq_2'} \\
              \forall y_1:\tau_1,\ldots,y_n:\tau_n . 
            \peq_1 \models \peq_1'\ \text{and}\ \peq_1 \wedge \peq_2' \models \peq_2}
            {\htrip{\peq_1}{f(y_1:\tau_1,\ldots,y_n:\tau_n)}{\peq_2}}
\end{mathpar}
\end{fpfig}

\begin{lemma}
  \label{lemma-closure}
  If $\htrip{\peq_1}{\prog}{\peq_2}$ and $\instr[\mv_1/y_1]\cdots[\mv_n/y_n] \redx \prog$
  then $\peq_1[\mv_1/y_1]\cdots[\mv_n/y_n] \redx \eqs_1$ and
  $\peq_2[\mv_1/y_1]\cdots[\mv_n/y_n] \redx \eqs_2$ for some $\eqs_1, \eqs_2$. 
\end{lemma}

\begin{lemma}[$\metaprot$ Hoare Triple Correctness]
  \label{lemma-htrip}
  If $\htrip{\peq_1}{\cmd}{\peq_2}$ with $\peq_1 \redx
  \eqs_1$, $\peq_2 \redx \eqs_2$, and $\cmd \redx
  \prog$, then $\htrip{\eqs_1}{\prog}{\eqs_2}$.
\end{lemma}

\begin{proof}
  By structural induction on the derivation of $\cmd \redx \prog$ and
  case analysis on $\cmd$.

  \textit{Case} $\cmd = (\xassign{e_1}{e_2}{e_3})$. In this case by definition
  and inversion of $\redx$ and $\TirName{Assign}$ we have:
  \begin{mathpar}
    e_1 \redx x

    e_2 \redx \be

    e_3 \redx \cid

    \prog = (\xassign{x}{\be}{\cid})

    \peq_1 \redx \true

    \peq_2  \redx x \eop \toeq{\elab{\be}{\cid}}
  \end{mathpar}
  Now by \TirName{Assign} we have
  $\htrip{\toeq{\elab{\be}{\cid}} \eop \toeq{\elab{\be}{\cid}}}{\xassign{x}{\be}{\cid}}{x \eop \toeq{\elab{\be}{\cid}}}$, so
  we have
  $\htrip{\true}{\xassign{x}{\be}{\cid}}{x \eop
    \toeq{\elab{\be}{\cid}}}$ by \TirName{Consequence}.

  \textit{Case} $\cmd = \elab{\assert{e_1 = e_2}}{e_3}$. In this case by definition we have:
  \begin{mathpar}
    e_1 \redx \be_1

    e_2 \redx \be_2

    e_3 \redx \cid

    \prog = \elab{\assert{\be_1 = \be_2}}{\cid}

    \peq_1 \redx \toeq{\elab{\be_1}{\cid}} \eop \toeq{\elab{\be_2}{\cid}}

    \peq_2 \redx \true
  \end{mathpar}
  But we have $\htrip{\toeq{\elab{\be_1}{\cid}} \eop \toeq{\elab{\be_2}{\cid}}}{\elab{\assert{\be_1 = \be_2}}{\cid}}{\toeq{\elab{\be_1}{\cid}} \eop \toeq{\elab{\be_2}{\cid}}}$ by $\TirName{Assert}$, so
  $\htrip{\toeq{\elab{\be_1}{\cid}} \eop \toeq{\elab{\be_2}{\cid}}}{\elab{\assert{\be_1 = \be_2}}{\cid}}{\true}$ by $\TirName{Consequence}$.

  \textit{Case} $\cmd = \cmd_1;\cmd_2$. In this case by assumption and
  definition and inversion of $\redx$ and $\TirName{Seq}$ we have:
  \begin{mathpar}
    \cmd_1 \redx \prog_1

    \cmd_2 \redx \prog_2

    \prog = \prog_1;\prog_2

    \peq_1 \redx \eqs_1^1 \wedge (\eqs_2^1 \impl \eqs_1^2)

    \peq_2 \redx \eqs_2^1 \wedge \eqs_2^2
  \end{mathpar}
  and by inversion of the $\TirName{Seq}$ rule we have:
  \begin{mathpar}
    \htrip{\peq_1^1}{\cmd_1}{\peq_2^1}

    \htrip{\peq_1^2}{\cmd_2}{\peq_2^2}
  \end{mathpar}
  and by the induction hypothesis we have:
  \begin{mathpar}
    \htrip{\eqs_1^1}{\cmd_1}{\eqs_2^1}

    \htrip{\eqs_1^2}{\cmd_2}{\eqs_2^2}
  \end{mathpar}
  But by \TirName{Hyp} we have:
  $\htrip{\eqs_1^1 \wedge (\eqs_2^1 \impl \eqs_1^2)}{\cmd_1}{\eqs_2^1 \wedge \eqs_1^2}$
  and by sanity conditions \cnote{need to clarify these} we have
  $\vars(\eqs_2^1) \cap \avars(\prog_2) = \varnothing$ so by \TirName{Frame} we
  have
  $\htrip{\eqs_2^1 \wedge \eqs_1^2}{\cmd_2}{\eqs_2^1 \wedge \eqs_2^2}$.
  Thus by $\TirName{Seq}$ we have:
  $$\htrip{\eqs_1^1 \wedge (\eqs_2^1 \impl \eqs_1^2)}{\prog_1;\prog_2}{\eqs_2^1 \wedge \eqs_2^2}$$

  \textit{Case} $\cmd = f(e_1,\ldots,e_n)$. In this case by assumption and
  definition and inversion of $\TirName{App}$ and $\redx$ we have:
  \begin{mathpar}
    \codebase(f) = y_1,\ldots,y_n, \instr

    e_1 \redx \mv_1 \cdots e_n \redx \mv_n

    \instr[\mv_1/y_1]\cdots[\mv_n/y_n] \redx \prog

    \peq_1 = \peq^0_1[\mv_1/y_1]\cdots[\mv_n/y_n]

    \peq_2 = \peq^0_2[\mv_1/y_1]\cdots[\mv_n/y_n]

    \htrip{\peq^0_1}{f(y_1:\tau_1,\ldots,y_n:\tau_n)}{\peq^0_2}
  \end{mathpar}
  Note that sanity conditions on pre- and post-condition annotations
  means there exists $\eqs_1$ and $\eqs_2$ where $\peq_1 \redx
  \eqs_1$ and $\peq_2 \redx \eqs_2$. Now, the judgement
  $\htrip{\peq^0_1}{f(y_1:\tau_1,\ldots,y_n:\tau_n)}{\peq^0_2}$
  follows by one or two subcases. Where it follows by \TirName{Fn} the
  result is immediate by definition and the induction
  hypothesis. Otherwise it follows by \TirName{Hardpack} and we
  proceed as follows.

  By inversion of \TirName{Hardpack} we have
  $\htrip{\peq_1'}{f(y_1:\tau_1,\ldots,y_n:\tau_n)}{\peq_2'}$ by
  \TirName{Fn} where $\forall y_1:\tau_1,\ldots,y_n:\tau_n . \peq^0_1
  \models \peq_1'$. Then $\peq_1'[\mv_1/y_1]\cdots[\mv_n/y_n] \redx
  \eqs_1'$ and $\peq_2'[\mv_1/y_1]\cdots[\mv_n/y_n] \redx
  \eqs_2'$ by Lemma \ref{lemma-closure}. 
  By Lemma \ref{lemma-genentails} we have $\eqs_1
  \models \eqs_1'$, so by \TirName{Consequence} we have
  $\htrip{\eqs_1}{\prog}{\eqs_2'}$, and by \TirName{Hyp}
  we have:
  $$\htrip{\eqs_1 \wedge (\eqs_2' \impl \eqs_2)}{\prog}{\eqs_2' \wedge \eqs_2}$$

  Furthermore, also by inversion of \TirName{Hardpack} we have
  $\forall y_1:\tau_1,\ldots,y_n:\tau_n . \peq^0_1 \wedge 
  \peq_2' \models \peq^0_2$. 
  By Lemma \ref{lemma-genentails} we thus have $\eqs_1
  \wedge \eqs_2' \models \eqs_2$, which implies
  $\eqs_1 \models \eqs_2' \impl \eqs_2$.
  Thus by \TirName{Consequence} we have $\htrip{\eqs_1}{\prog}{\eqs_2}$.

  Finally, \textit{case} $\prog = \elet{y}{e}{\cmd}$ follows trivially
  by the induction hypothesis, which covers all cases to prove
  the result.
\end{proof}

\begin{theorem}
  If $\htrip{\eqs_1}{\ttt{main()}}{\eqs_2}$ as a consequence of
  \TirName{HardPack} and $\ttt{main}() \redx \prog$ then
  $\htrip{\eqs_1}{\prog}{\eqs_2}$.
\end{theorem}

### 2025-11-01 toplas26/metalang.tex
checkpoint

@@
 component definitions support compositional verification of larger
-protocols. Separate verification of well-designed components results
-in confidentiality and integrity properties to be recorded in their
-types, allowing for significant reduction of SMT verification in whole
-program analysis, as we will discuss with extended examples in Section
+protocols. This includes an automated verification of algebraic
+program properties expressed as Hoare logic assertions that we
+introduce in this Section. Furthermore, we can compositionally verify
+confidentiality and integrity properties expressed as types, as we
+will show in Section \ref{section-sharety}. We present and discuss
+extended examples of $\metaprot$ programs in Section
 \ref{section-examples}.
@@
 $x$. $\metaprot$ expression forms allow dynamic construction of these
-values. $\metaprot$ \emph{instruction} forms allow dynamic
+values. Note in particular that $\concat$ is string concatentation.
+$\metaprot$ \emph{instructions} allow dynamic
 construction of $\minicat$ protocols $\prog$ that incorporate expression
@@
 compute values $\mv$ and, as a distinct form, functions that compute
-protocols $\prog$.  Formally, we consider a complete metaprogram to include
+protocols $\prog$.  The latter requires type annotations on arguments
+which are restricted to being either strings, client ids, or
+records-- we use this type information to support Hoare logic
+automation as we discuss below in Section \ref{section-metahtrip}.
+
+Formally, we consider a complete metaprogram to include
 both a codebase and a ``main'' program that uses the codebase.
@@
 $\mv$ for free occurrences of $y$ in $e$ or $\instr$ respectively. The
-rules are mostly standard. Note that we do not include an evaluation
-rule for the form $\eqcast{\mx{e}{e}}{\notg{\phi}}$ which is a type
+rules are mostly standard.
+
+Note that we do not include an evaluation
+rule for the form $\eqcast{\mx{e}{e}}{e}$ which is a type
 annotation that we assume is erased from programs prior to
-evaluation. We defer discussion of this form, as well as the syntactic
-category $\notg{\phi}$, to the next Section.
+evaluation. We defer discussion of this form to Section
+\ref{section-sharety}
+
+\paragraph{Sanity conditions on programs.} We assume that
+any $\metaprot$ codebase $\codebase$ includes a function
+$\ttt{main}()$. And as for $\minicat$, we assume that $\metaprot$
+programs are semantically well-formed-- that is, that
+$\ttt{main}() \redx \prog$ and $\prog$ is well-formed.
 
@@
 
-\subsection{$\metaprot$ Automated Abstract Hoare Logic}
+\subsection{Abstract pre and postcondition annotations}
 
+$\metaprot$ instruction functions can optionally be annotated with
+pre and postconditions. To model this in our theory we posit
+functions $\precond$  and $\postcond$ 
+that specify conditions for instruction functions $f \in \dom(\codebase)$.
+Note that $\precond(f)$ or $\precond(f)$ may be undefined if no annotation
+is given for $f$. In actual code, we write pre and postconditions directly before and
+after functions to define them. Here is a simple example function $\ttt{bar}$
+with. Note that the pre and postconditions are abstract with respect to the
+input string $\ttt{x}$, and that string concatenation  
+can be used in pre and postcondition constraints:
 \begin{verbatimtab}
@@
 \end{verbatimtab}
+Since expressions can occur in $\metaprot$ constraints, they are not
+SMT constraints, but abstractions of constraints parameterized by
+the formal parameters of the function they annotate. We denote these
+\emph{abstract constraints} as $\eqs$ with the syntax given
+in Figure \ref{fig-peq}. When the function is applied, the pre and postconditions
+are instantiated as SMT constraints, via evaluating the expressions
+within the constraints. The evaluation semantics of abstract constraints
+are also given in Figure \ref{fig-peq}.
 
 \peqfig
+
+In the case of \ttt{main}, since it has no arguments, the
+pre and postconditions must be concrete SMT expressions.
+For example, given the \ttt{bar} function above, we
+could define a full program as follows, where \ttt{T}
+is the true constraint. 
+\begin{verbatimtab}
+precondition: ( T )
+main(){
+  m["thisfoo"]@1 := 3; m["thisbaz"]@1 := 2; bar("this")
+}
+postcondition: ( [m["thisbar"]@2 == 5 )
+\end{verbatimtab}
+There are no preconditions to this \ttt{main} because they are
+all established by the program, and the postcondition follows
+as a consequence of the application \ttt{bar("this")}.
+
+The use of pre and postconditions in \ttt{main} allows verification of
+whole-program (protocol) correctness. Further, the use of annotations
+on function definitions supports \emph{compositional} verification of
+program components. In general postconditions can serve as
+pre-verified properties of functions that can be reused in, and reduce
+overhead of, whole protocol verification.
+
+\subsection{$\metaprot$ Automated Abstract Hoare Logic}
+\label{section-metahtrip}
+
+In the $\metaprot$ theory, we introduce Hoare logic assertions
+of the form $\htrip{\peq_1}{\cmd}{\peq_2}$. The correctness
+condition of this assertion is compositional, in that instantiated
+abstract Hoare triples evaluate to valid $\minicat$ Hoare triples.
+This is formally elaborated below in Lemma
+\ref{lemma-metahtrip} and Corollaries \ref{cor-metahtrip} and
+\ref{cor-mainhtrip} below.
+
+In Figure \ref{fig-metahtrip} we give the derivation rules for the
+abstract Hoare logic. These are based on $\minicat$ Hoare logic, but
+are algorithmic. A main predicate for these rules is the
+$\TirName{GenEntails}$ rule, which encodes a generalized version of
+entailment. By using fresh variables to instantiate abstract
+constraints, we guarantee entailment for any instantiation, as
+follows.
 
@@
 
-\begin{fpfig}[t]{Algorithmic $\metaprot$ Abstract Hoare Triple Deduction Rules}{fig-metahtrip}
+\metahtripfig
+
+The \TirName{HardPack} rule uses \TirName{GenEntails} to ensure that
+the annotated pre and postconditions are correct for a given function,
+if they are defined. It is in fact not necessary for \emph{any} pre
+and postconditions to be defined, but their intent is to enable
+fine-tuning of postconditions for improving verification efficiency as
+discussed above and later in Section \ref{section-examples}.  While
+not relevant to the theory, an important implementation detail is that
+verification of any given function $f$ only needs to be performed
+once, and can then be memoized for other applications using
+the $\TirName{App}$ rule as a consequence of Lemma \ref{lemma-genentails}
+(and more precisely of Corollary \ref{cor-metahtrip}).
+
+The \TirName{Assign} and \TirName{Assert} rules propagate equations
+induced by assignments and assertions to post and preconditions
+respectively.  Note that without the use of \TirName{HardPack}, the
+entire protocol would accumulate in postconditions of \ttt{main}. The
+\TirName{Seq} rule relies on the correctness of our \TirName{Hyp} rule
+to join pre and postconditions as we discuss in the relevant case of
+Lemma \ref{lemma-metahtrip}, which is the foundational correctness
+result for the analysis.
+
+\paragraph{Pre and post sanity conditions.} We require that
+annotated pre and postconditions on functions will evaluate to SMT
+constraints when instantiated with function arguments, and refer only
+to variables that occur in their bodies. Deduction rules other than
+\TirName{HardPack} enforce this by definition.  For any $f$ with
+$\codebase(f) = y_1 : \tau_1, \ldots, y_n : \tau_n, \cmd$,
+$\precond(f) = \peq_1$, and $\postcond(f) = \peq_2$, we require that
+for all $[\mv_1/y_1]\cdots[\mv_n/y_n]$ there exists $\eqs_1$, $\eqs_2$
+where:
 \begin{mathpar}
-\inferrule[GenEntails]
-          {\mv_1,\ldots,\mv_n = \fresh(\tau_1,\ldots,\tau_n) \\ \peq_1[\mv_1/y_n \cdots \mv_n/y_n] \redx \eqs_1 \\
-           \peq_2[\mv_1/y_n \cdots \mv_n/y_n] \redx \eqs_2 \\ \eqs_1 \models \eqs_2}
-          {\forall y_1:\tau_1,\ldots,y_n:\tau_n . \peq_1 \models \peq_2 }
+  \cmd[\mv_1/y_1]\cdots[\mv_n/y_n] \redx \prog
 
-  \inferrule[Mesg]
-            {}
-            {\htrip{\eqtrue}{\xassign{e_1}{e_2}{e_3}}{e_1 \eop \toeq{\elab{e_2}{e_3}}}}
+  \peq_1[\mv_1/y_1]\cdots[\mv_n/y_n] \redx \eqs_1
 
-  %\inferrule[Encode]
-  %          {\mx{e_1}{e_2} \redx x \\ \notg{\phi} \redx \phi \\
-  %            \eqs \models x \eop \phi\\
-  %            \atj{\phi}{R}{\ty}}
-  %          {\mtj{\eqcast{\mx{e_1}{e_2}}{\notg{\phi}}}{\eqs}{(x : \ty)}{R}{\varnothing}{\eqs}}
-  %
-  \inferrule[Assert]
-            {}
-            {\htrip{\toeq{\elab{e_1}{e_3}} \eop \toeq{\elab{e_2}{e_3}}}{\elab{\assert{e_1 = e_2}}{e_3}}{\eqtrue}}
+  \peq_2[\mv_1/y_1]\cdots[\mv_n/y_n] \redx \eqs_2
+\end{mathpar}
+and $\vars(\eqs_1) \cup \vars(\eqs_2) \subseteq \vars(\prog)$.
 
-  \inferrule[Seq]          
-            {\htrip{\peq_1^1}{\cmd_1}{\peq_2^1} \\ \htrip{\peq_1^2}{\cmd_2}{\peq_2^2}}
-            {\htrip{\peq_1^1 \wedge (\peq_2^1 \impl \peq_1^2)}{\cmd_1;\cmd_2}{\peq_2^1 \wedge \peq_2^2}}
+\subsection{Abstact Hoare Logic Soundnes}
 
-  \inferrule[Let]
-            {\htrip{\peq_1}{\cmd[e/y]}{\peq_2}}
-            {\htrip{\peq_1}{\elet{y}{e}{\cmd}}{\peq_2}}
-
-  \inferrule[App]
-            {\htrip{\peq_1}{f(y_1:\tau_1,\ldots,y_n:\tau_n)}{\peq_2}}
-            {\htrip{\peq_1[e_1/y_1 \cdots e_n/y_n]}{f(e_1,\ldots,e_n)}{\peq_2[e_1/y_1 \cdots e_n/y_n]}}
-
-  \inferrule[Fn]
-            {\codebase(f) = y_1 : \tau_1, \ldots, y_n : \tau_n, \cmd \\ \htrip{\peq_1}{\cmd}{\peq_2}}
-            {\htrip{\peq_1}{f(y_1:\tau_1,\ldots,y_n:\tau_n)}{\peq_2}}
-
-  \inferrule[Hardpack]
-            {\precond(f) = \peq_1 \\ \postcond(f) = \peq_2 \\
-              \htrip{\peq_1'}{f(y_1:\tau_1,\ldots,y_n:\tau_n)}{\peq_2'} \\
-              \forall y_1:\tau_1,\ldots,y_n:\tau_n . 
-            \peq_1 \models \peq_1'\ \text{and}\ \peq_1 \wedge \peq_2' \models \peq_2}
-            {\htrip{\peq_1}{f(y_1:\tau_1,\ldots,y_n:\tau_n)}{\peq_2}}
-\end{mathpar}
-\end{fpfig}
-
+We begin by showing that the analysis preserves the property that in
+any $\htrip{\peq_1}{\prog}{\peq_2}$, fully instantiating $\prog$
+also fully instantiates $\peq_1$ and $\peq_2$. The result
+follows by sanity conditions on pre and postcondition annotations,
+the definition of analysis, and induction on $\cmd$.
 \begin{lemma}
@@
 
+Now we can prove our main result in a general form. The proof
+follows mainly by appeal to the $\minicat$ Hoare logic deduction
+system.
 \begin{lemma}[$\metaprot$ Hoare Triple Correctness]
-  \label{lemma-htrip}
+  \label{lemma-metahtrip}
   If $\htrip{\peq_1}{\cmd}{\peq_2}$ with $\peq_1 \redx
@@
   $\htrip{\eqs_1^1 \wedge (\eqs_2^1 \impl \eqs_1^2)}{\cmd_1}{\eqs_2^1 \wedge \eqs_1^2}$
-  and by sanity conditions \cnote{need to clarify these} we have
+  and by sanity conditions we have
   $\vars(\eqs_2^1) \cap \avars(\prog_2) = \varnothing$ so by \TirName{Frame} we
@@
 
-\begin{theorem}
-  If $\htrip{\eqs_1}{\ttt{main()}}{\eqs_2}$ as a consequence of
-  \TirName{HardPack} and $\ttt{main}() \redx \prog$ then
+As a consequence of the preceding Lemma, we can state important
+Corollaries that characterize, first, the compositionality of function
+constraints, and second, the correctness of whole program
+analysis. These both follow immediately by reasoning in the
+\TirName{App} case of Lemma \ref{lemma-metahtrip}.
+
+\begin{corollary}[Compositionality of Abstract Function Constraints]
+  \label{cor-metahtrip}
+  Given:
+  \begin{mathpar}
+    \htrip{\peq_1}{f(y_1:\tau_1,\ldots,y_n:\tau_n)}{\peq_2}
+
+    \codebase(f) = y_1 : \tau_1, \ldots, y_n : \tau_n, \cmd
+
+    \cmd[\mv_1/y_1]\cdots[\mv_n/y_n] \redx \prog
+
+    \peq_1[\mv_1/y_1]\cdots[\mv_n/y_n] \redx \eqs_1
+
+    \peq_2[\mv_1/y_1]\cdots[\mv_n/y_n] \redx \eqs_2
+  \end{mathpar}
+  Then $\htrip{\eqs_1}{\prog}{\eqs_2}$ is valid.
+\end{corollary}
+
+\begin{corollary}
+  \label{cor-mainhtrip}
+  If $\htrip{\peq_1}{\ttt{main()}}{\peq_2}$ and $\ttt{main}() \redx \prog$ and
+  $\peq_1 \redx \eqs_1$ and $\peq_2 \redx \eqs_2$ then
   $\htrip{\eqs_1}{\prog}{\eqs_2}$.
-\end{theorem}
+\end{corollary}

### 2025-11-05 code/cvc5mvn/src/main/scala/plaid/prelude/App.scala
contracts

@@ -4,7 +4,7 @@ import picocli.CommandLine
 import picocli.CommandLine.{Command, Option, Parameters}
 import plaid.prelude.antlr.Loader
 import plaid.prelude.ast.{Identifier, dependencyOrdered, expandAll, resolve}
-import plaid.prelude.logic.ConstraintAnalyzer
+import plaid.prelude.logic.contracts
 
 import java.io.File
 import java.nio.file.Files
@@ -28,11 +28,8 @@ class App extends Runnable {
     val exprFns = ast.exprFuncs.expandAll()
     val constraintFns = ast.constraintFuncs.expandAll(exprFns)
     val cmdFns = ast.cmdFuncs.expandAll(exprFns, constraintFns)
-
-    val analyzer = new ConstraintAnalyzer(ast, fieldSize)
-    val constraints = analyzer.inferPrePostFN(ast.cmdFuncs.resolve(Identifier("main")))
-    println(s"The precondition for main: ${constraints.precondition.prettyPrint()}")
-    println(s"The postcondition for main: ${constraints.postcondition.prettyPrint()}")
+    val contracts = cmdFns.contracts()
+    // TODO Check and optionally display
 }
 
 object App {

### 2025-11-05 code/cvc5mvn/src/main/scala/plaid/prelude/ast/Constraint.scala

@@ -28,7 +28,7 @@ sealed trait Constraint extends Node {
    * constraint. All expressions in all constraints should be expanded before
    * this is used, but it would probably work anyway.
    */
-  def expand(ctx: List[ConstraintFunc], bindings: Map[Identifier, Expr] = Map()): Constraint = transform {
+  def expand(ctx: List[ConstraintFunc] = List(), bindings: Map[Identifier, Expr] = Map()): Constraint = transform {
     case CallConstraint(id, parms) =>
       val f = ctx.resolve(id)
       val formalParms = f.parms

### 2025-11-05 code/cvc5mvn/src/main/scala/plaid/prelude/ast/Expr.scala

@@ -52,12 +52,12 @@ sealed trait Expr extends Node {
   /**
    * Move party indexes directly around the memories they enclose.
    */
-  def indexParties(partyIndex: Expr): Expr = transform {
-    case AtExpr(e1, e2) => Some(e1.indexParties(e2))
-    case e: RandomExpr => Some(AtExpr(e, partyIndex))
-    case e: SecretExpr => Some(AtExpr(e, partyIndex))
-    case e: MessageExpr => Some(AtExpr(e, partyIndex))
-    case e: OutputExpr => Some(AtExpr(e, partyIndex))
+  def indexParties(partyIndex: Option[Expr] = None): Expr = transform {
+    case AtExpr(e1, e2) => Some(e1.indexParties(Some(e2)))
+    case e: RandomExpr => Some(AtExpr(e, partyIndex.get))
+    case e: SecretExpr => Some(AtExpr(e, partyIndex.get))
+    case e: MessageExpr => Some(AtExpr(e, partyIndex.get))
+    case e: OutputExpr => Some(AtExpr(e, partyIndex.get))
     case other => None
   }
 }

### 2025-11-05 code/cvc5mvn/src/main/scala/plaid/prelude/ast/Func.scala

@@ -49,7 +49,6 @@ extension (trg: List[ConstraintFunc])
 extension (trg: List[CmdFunc])
   /** Expand all the constraint functions in this list. */
   def expandAll(exprCtx: List[ExprFunc], constraintCtx: List[ConstraintFunc]): List[CmdFunc] = trg
-    .dependencyOrdered { _.cmdDependencies() }
     .map { _.expand(exprCtx, constraintCtx) }
 
 case class ExprFunc(id: Identifier, parms: List[Identifier], body: Expr) extends Func {

### 2025-11-05 code/cvc5mvn/src/main/scala/plaid/prelude/logic/ConstraintAnalyzer.scala

@@ -1,148 +1,0 @@
-package plaid.prelude.logic
-
-import plaid.prelude.ast.*
-
-import scala.annotation.tailrec
-import scala.collection.mutable
-
-class ConstraintAnalyzer(val program: Program, order: String) {
-
-  private val verifier = new GenEntailVerifier(program, order)
-  private val functionConstraints = new mutable.HashMap[Identifier, Constraints]()
-
-  private def binding(formal: List[TypedIdentifier], actual: List[Expr]): Map[Identifier, Expr] =
-    formal.map(_.y).zip(actual).toMap
-
-  /** Calculate precondition and postcondition for a function (FN rule) */
-  def inferPrePostFN(function: CmdFunc): Constraints = {
-//    val evaluator = Evaluator(program)
-//    val constraints = inferPrePostCmd(function.c, evaluator)
-//
-//    if function.precond == null && function.postcond == null then
-//      functionConstraints.put(function.fname, constraints)
-//      constraints
-//    else
-//      val e1 = Option(function.precond).getOrElse(TrueConstraint())
-//      val e2 = Option(function.postcond).getOrElse(TrueConstraint())
-//      val e1p = Option(constraints.precondition).getOrElse(TrueConstraint())
-//      val e2p = Option(constraints.postcondition).getOrElse(TrueConstraint())
-//
-//      if verifier.genEntails(function.typedVariables, e1, e1p) &&
-//        verifier.genEntails(function.typedVariables, AndConstraint(e1, e2p), e2) then
-//        functionConstraints.put(function.fname, Constraints(function.precond.get, function.postcond.get))
-//        Constraints(function.precond.get, function.postcond.get)
-//      else
-        throw new RuntimeException(s"The hardpack does not hold for ${prettyPrint(function.id)}")
-  }
-
-  /** Recursively infer pre/postconditions for a command */
-  def inferPrePostCmd(command: Cmd, evaluator: Evaluator): Constraints = command match
-    case assignCmd: AssignCmd =>
-      Constraints(
-        TrueConstraint(),
-        EqualConstraint(
-          evaluator.expression(assignCmd.e1),
-          PartyIndexer.expression(evaluator.expression(assignCmd.e2), null)
-        )
-      )
-
-    case assertCmd: AssertCmd =>
-      Constraints(
-        EqualConstraint(
[…]

### 2025-11-05 code/cvc5mvn/src/main/scala/plaid/prelude/logic/Constraints.scala

@@ -1,5 +1,0 @@
-package plaid.prelude.logic
-
-import plaid.prelude.ast.Constraint
-
-case class Constraints(var precondition: Constraint, var postcondition: Constraint)

### 2025-11-05 code/cvc5mvn/src/main/scala/plaid/prelude/logic/Contract.scala

@@ -1,0 +1,60 @@
+package plaid.prelude.logic
+
+import plaid.prelude.ast.{AndConstraint, AssertCmd, AssignCmd, CallCmd, Cmd, CmdFunc, Constraint, EqualConstraint, Identifier, LetCmd, TrueConstraint, dependencyOrdered, OTExpr, OTFourExpr}
+
+case class Entailment(cmd: Cmd, a: Constraint, b: Constraint)
+
+/**
+ * A context that is built up as commands are processed in succession. It
+ * includes as a single constraint the conjunction of the postconditions of all
+ * the commands that have been processed so far, and a collection of all the
+ * entailments that have to hold based on all the commands that have been
+ * processed so far.
+ */
+case class HoareContext(cons: Constraint = TrueConstraint(), ent: List[Entailment] = List()) {
+  /** Add a new postcondition in the constraint of this Hoare context. */
+  private def include(c: Constraint): HoareContext = copy(cons = AndConstraint(cons, c))
+  /** Add a new entailment to this Hoare context. */
+  private def include(e: Entailment): HoareContext = copy(ent = e :: ent)
+  /** Include a new command in this Hoare context. */
+  def include(cmd: Cmd, contracts: List[Contract]): HoareContext = cmd match
+    case AssertCmd(e1, e2, party) =>
+      val a = e1.indexParties(Some(party))
+      val b = e2.indexParties(Some(party))
+      include(Entailment(cmd, cons, EqualConstraint(a, b)))
+    case AssignCmd(dest, OTExpr(e1, i1, e2, e3)) => ???
+    case AssignCmd(dest, OTFourExpr(s1, s2, i1, e1, e2, e3, e4)) => ???
+    case AssignCmd(e1, e2) => include(EqualConstraint(e1.indexParties(), e2.indexParties()))
+    case CallCmd(id, parms) =>
+      val contract = contracts.lookup(id)
+      val formalParms = contract.f.typedVariables.map(_.y)
+      val actualParms = parms.map(_.indexParties())
+      val bindings = formalParms.zip(actualParms).toMap
+      val pre = contract.pre.expand(bindings = bindings)
+      val post = contract.post.expand(bindings = bindings)
+      include(Entailment(cmd, cons, pre)).include(post)
+    case other => throw Exception(s"Unsupported command $other")
+}
+
+/**
+ * Tracks for a function what its (possibly inferred) preconditions and
[…]

### 2025-11-05 code/cvc5mvn/src/main/scala/plaid/prelude/logic/GenEntailVerifier.scala

@@ -3,7 +3,6 @@ package plaid.prelude.logic
 import io.github.cvc5.{Sort, TermManager}
 import plaid.prelude.ast.*
 import plaid.prelude.cvc.{TermFactory, Verifier}
-import plaid.prelude.eval.ConstraintChecker
 
 import java.util
 import java.util.TreeMap
@@ -51,7 +50,7 @@ class GenEntailVerifier(val program: Program, order: String) {
     val groundE2: Constraint = evaluator.constraint(e2)
 
     // check if ground constraints are fully instantiated
-    if ConstraintChecker.checkConstraint(groundE1) && ConstraintChecker.checkConstraint(groundE2) then
+    if WellFormed.checkConstraint(groundE1) && WellFormed.checkConstraint(groundE2) then
       val termFactory = new TermFactory(termManager, sort)
       val groundE1Term = termFactory.constraintToTerm(groundE1)
       val groundE2Term = termFactory.constraintToTerm(groundE2)

### 2025-11-05 code/cvc5mvn/src/main/scala/plaid/prelude/logic/Unique.scala

@@ -1,4 +1,4 @@
-package plaid.prelude.ext
+package plaid.prelude.logic
 
 import plaid.prelude.ast.{Identifier, Num, Str}
 

### 2025-11-05 code/cvc5mvn/src/main/scala/plaid/prelude/logic/WellFormed.scala

@@ -1,10 +1,12 @@
-package plaid.prelude.eval
+package plaid.prelude.logic
 
 import plaid.prelude.ast.*
 
 // check if a (final/evaluated) constraint is valid
-object ConstraintChecker {
+object WellFormed {
 
+  // TODO Apply constraints
+  
   /**
    * valid constraint node type
    */

### 2025-11-05 code/cvc5mvn/src/test/scala/plaid/prelude/eval/WellFormedTest.scala

@@ -3,14 +3,15 @@ package plaid.prelude.eval
 import org.junit.Test
 import org.junit.Assert.{assertFalse, assertTrue}
 import plaid.prelude.antlr.Loader
+import plaid.prelude.logic.WellFormed
 
-class ConstraintCheckerTest {
+class WellFormedTest {
 
   private def assertPasses(src: String): Unit =
-    assertTrue(ConstraintChecker.checkConstraint(Loader.constraint(src)))
+    assertTrue(WellFormed.checkConstraint(Loader.constraint(src)))
 
   private def assertFails(src: String): Unit =
-    assertFalse(ConstraintChecker.checkConstraint(Loader.constraint(src)))
+    assertFalse(WellFormed.checkConstraint(Loader.constraint(src)))
 
   /** Memories should always come with party indexes */
   @Test

### 2025-11-05 code/cvc5mvn/src/main/scala/plaid/prelude/App.scala
add missing function children, fix descendants bug

@@ -4,7 +4,7 @@ import io.github.cvc5.TermManager
 import picocli.CommandLine
 import picocli.CommandLine.{Command, Option, Parameters}
 import plaid.prelude.antlr.Loader
-import plaid.prelude.ast.{Identifier, dependencyOrdered, expandAll, resolve}
+import plaid.prelude.ast.expandAll
 import plaid.prelude.cvc.TermFactory
 import plaid.prelude.logic.{contracts, verificationFailures}
 
@@ -31,7 +31,8 @@ class App extends Runnable {
     val constraintFns = ast.constraintFuncs.expandAll(exprFns)
     val cmdFns = ast.cmdFuncs.expandAll(exprFns, constraintFns)
     val contracts = cmdFns.contracts()
-    
+
+    contracts.foreach(x => println(x))
     contracts.foreach(x =>
       val termManager = TermManager()
       val termFactory = TermFactory(termManager, fieldSize)

### 2025-11-05 code/cvc5mvn/src/main/scala/plaid/prelude/ast/Func.scala

@@ -33,7 +33,7 @@ extension [F <: Func](trg: List[F])
     if trg.isEmpty then Nil
     else
       val ids = trg.map(_.id).toSet
-      val first = trg.filter(x => f(x).intersect(ids).isEmpty)
+      val first = trg.filter(x => f(x).forall(ids.contains))
       if first.isEmpty then throw Exception("Cyclic function calls detected")
       first ++ trg.filterNot(first.contains).dependencyOrdered(f)
 

### 2025-11-05 code/cvc5mvn/src/main/scala/plaid/prelude/ast/Node.scala

@@ -30,12 +30,15 @@ trait Node {
     case CallCmd(id, parms) => id :: parms
     case LetCmd(y, e, c) => y :: e :: c
     case RecordType(elements) => elements.flatten((x, y) => List(x, y))
+    case ExprFunc(id, parms, body) => id :: body :: parms
+    case ConstraintFunc(id, parms, body) => id :: body :: parms
+    case CmdFunc(id, parms, body, precond, postcond) => id :: body ++ parms ++ precond.toList ++ postcond.toList
     case _ => List()
 
   /**
    * Collect all the descendants of this node.
    */
-  def descendants(): Iterable[Node] = children().flatMap(_.descendants())
+  def descendants(): Iterable[Node] = children() ++ children().flatMap(_.descendants())
 
   /**
    * Provides a human-readable string representation of an AST node and its decedents.

### 2025-11-05 code/cvc5mvn/src/main/scala/plaid/prelude/logic/Verifier.scala

@@ -11,7 +11,7 @@ extension (trg: Contract)
       val b = x.b.expand(bindings = bindings)
 
       if !WellFormed.checkConstraint(a) || !WellFormed.checkConstraint(b) then
-        throw Exception(s"Constraints must be ground")
+        throw Exception(s"Constraints must be ground: $a and $b")
 
       // TODO Reorganize the CVC5 stuff, this is weird
       val verifier = new VerifierTmp(cvc)

### 2025-11-05 code/cvc5mvn/src/test/scala/plaid/prelude/antlr/LoaderTest.scala

@@ -1,0 +1,23 @@
+package plaid.prelude.antlr
+
+import org.junit.Assert.assertEquals
+import org.junit.Test
+
+class LoaderTest {
+
+  /** Program sections can appear in any order. */
+  @Test
+  def sectionOrderAgnostic(): Unit = {
+    val exprSection = "exprfunctions: exprfn() { 12 }"
+    val constraintSection = "constraintfunctions: constraintfn() { T }"
+    val cmdSection = "cmdfunctions: cmdfn() { m[\"x\"]@1 := m[\"x\"]@2 }"
+    val prog1 = Loader.program(exprSection + constraintSection + cmdSection)
+    val prog2 = Loader.program(constraintSection + exprSection + cmdSection)
+    val prog3 = Loader.program(cmdSection + exprSection + constraintSection)
+    val prog4 = Loader.program(cmdSection + constraintSection + exprSection)
+    assertEquals(prog1, prog2)
+    assertEquals(prog2, prog3)
+    assertEquals(prog3, prog4)
+  }
+
+}

### 2025-11-05 code/cvc5mvn/src/test/scala/plaid/prelude/ast/ExprTest.scala

@@ -1,0 +1,25 @@
+package plaid.prelude.ast
+
+import org.junit.Assert.assertEquals
+import org.junit.Test
+import plaid.prelude.antlr.Loader
+
+class ExprTest {
+
+  /** Functions with no calls have no dependencies. */
+  @Test
+  def dependenciesWithoutCalls(): Unit = {
+    val src = "exprfunctions: f(i) { m[\"x\" ++ i] }"
+    val fn = Loader.program(src).exprFuncs.head
+    assertEquals(Set(), fn.exprDependencies())
+  }
+
+  /** Functions consisting of a call has just that dependency. */
+  @Test
+  def dependenciesJustCall(): Unit = {
+    val src = "exprfunctions: f(i) { g(i) }"
+    val fn = Loader.program(src).exprFuncs.head
+    assertEquals(Set(Identifier("g")), fn.exprDependencies())
+  }
+
+}

### 2025-11-05 code/cvc5mvn/src/main/scala/plaid/prelude/App.scala
fix dependency sort order

@@ -4,7 +4,9 @@ import io.github.cvc5.TermManager
 import picocli.CommandLine
 import picocli.CommandLine.{Command, Option, Parameters}
 import plaid.prelude.antlr.Loader
-import plaid.prelude.ast.expandAll
+import plaid.prelude.ast.ListExprFuncExt.expandAll
+import plaid.prelude.ast.ListConstraintFuncExt.expandAll
+import plaid.prelude.ast.ListCmdFuncExt.expandAll
 import plaid.prelude.cvc.TermFactory
 import plaid.prelude.logic.{contracts, verificationFailures}
 

### 2025-11-05 code/cvc5mvn/src/main/scala/plaid/prelude/ast/Func.scala

@@ -29,30 +29,42 @@ extension [F <: Func](trg: List[F])
   /** Look up a function by its identifier. */
   def resolve(id: Identifier): F = trg.find(x => id == x.id).get
   /** Sort in such a way that there are no forward dependencies */
-  def dependencyOrdered(f: F => Set[Identifier]): List[F] =
+  def dependencyOrderedUsing(f: F => Set[Identifier]): List[F] =
     if trg.isEmpty then Nil
     else
       val ids = trg.map(_.id).toSet
-      val first = trg.filter(x => f(x).forall(ids.contains))
+      val first = trg.filter(x => f(x).forall(y => !ids.contains(y)))
       if first.isEmpty then throw Exception("Cyclic function calls detected")
-      first ++ trg.filterNot(first.contains).dependencyOrdered(f)
+      first ++ trg.filterNot(first.contains).dependencyOrderedUsing(f)
 
-extension (trg: List[ExprFunc])
-  /** Expand all the expression functions in this list. */
-  def expandAll(): List[ExprFunc] = trg
-    .dependencyOrdered { _.exprDependencies() }
-    .foldLeft(List()) { (acc, f) => f.expand(acc) :: acc }
+object ListExprFuncExt {
+  extension (trg: List[ExprFunc])
+    /** Sort in such a way that there are no forward dependencies */
+    def dependencyOrdered(): List[ExprFunc] = trg.dependencyOrderedUsing(_.exprDependencies())
+    /** Expand all the expression functions in this list. */
+    def expandAll(): List[ExprFunc] = trg
+      .dependencyOrdered()
+      .foldLeft(List()) { (acc, f) => f.expand(acc) :: acc }
+}
 
-extension (trg: List[ConstraintFunc])
-  /** Expand all the constraint functions in this list. */
-  def expandAll(exprCtx: List[ExprFunc]): List[ConstraintFunc] = trg
-    .dependencyOrdered { _.constraintDependencies() }
-    .foldLeft(List()) { (acc, f) => f.expand(exprCtx, acc) :: acc }
+object ListConstraintFuncExt {
+  extension (trg: List[ConstraintFunc])
+    /** Sort in such a way that there are no forward dependencies */
+    def dependencyOrdered(): List[ConstraintFunc] = trg.dependencyOrderedUsing(_.constraintDependencies())
+    /** Expand all the constraint functions in this list. */
[…]

### 2025-11-05 code/cvc5mvn/src/main/scala/plaid/prelude/logic/Contract.scala

@@ -1,6 +1,7 @@
 package plaid.prelude.logic
 
-import plaid.prelude.ast.{AndConstraint, AssertCmd, AssignCmd, CallCmd, Cmd, CmdFunc, Constraint, EqualConstraint, Identifier, LetCmd, Node, OTExpr, OTFourExpr, TrueConstraint, dependencyOrdered}
+import plaid.prelude.ast.ListCmdFuncExt.dependencyOrdered
+import plaid.prelude.ast.{AndConstraint, AssertCmd, AssignCmd, CallCmd, Cmd, CmdFunc, Constraint, EqualConstraint, Identifier, Node, OTExpr, OTFourExpr, TrueConstraint}
 
 case class Entailment(src: Node, a: Constraint, b: Constraint)
 
@@ -50,7 +51,7 @@ extension (trg: List[Contract])
 extension (trg: List[CmdFunc])
   /** Calculate contracts for all the commands (which must be expanded). */
   def contracts(): List[Contract] = trg
-    .dependencyOrdered { _.cmdDependencies() }
+    .dependencyOrdered()
     .foldLeft(List()) { (acc, x) => x.contract(acc) :: acc}
 
 extension (trg: CmdFunc)

### 2025-11-05 code/cvc5mvn/src/test/scala/plaid/prelude/ast/CmdFuncTest.scala

@@ -1,0 +1,25 @@
+package plaid.prelude.ast
+
+import org.junit.Assert.assertEquals
+import org.junit.Test
+import plaid.prelude.antlr.Loader
+
+class CmdFuncTest {
+
+  /** Functions with no calls have no dependencies. */
+  @Test
+  def dependenciesWithoutCalls(): Unit = {
+    val src = "cmdfunctions: f() { m[\"x\"]@1 := 3 }"
+    val fn = Loader.program(src).cmdFuncs.head
+    assertEquals(Set(), fn.cmdDependencies())
+  }
+
+  /** Functions consisting of a call has just that dependency. */
+  @Test
+  def cmdfunctions(): Unit = {
+    val src = "cmdfunctions: f() { g() }"
+    val fn = Loader.program(src).cmdFuncs.head
+    assertEquals(Set(Identifier("g")), fn.cmdDependencies())
+  }
+
+}

### 2025-11-05 code/cvc5mvn/src/test/scala/plaid/prelude/ast/ConstraintFuncTest.scala

@@ -1,0 +1,25 @@
+package plaid.prelude.ast
+
+import org.junit.Assert.assertEquals
+import org.junit.Test
+import plaid.prelude.antlr.Loader
+
+class ConstraintFuncTest {
+
+  /** Functions with no calls have no dependencies. */
+  @Test
+  def dependenciesWithoutCalls(): Unit = {
+    val src = "constraintfunctions: f(i) { m[\"x\"] == 3 }"
+    val fn = Loader.program(src).constraintFuncs.head
+    assertEquals(Set(), fn.constraintDependencies())
+  }
+
+  /** Functions consisting of a call has just that dependency. */
+  @Test
+  def dependenciesJustCall(): Unit = {
+    val src = "constraintfunctions: f(i) { g(i) }"
+    val fn = Loader.program(src).constraintFuncs.head
+    assertEquals(Set(Identifier("g")), fn.constraintDependencies())
+  }
+
+}

### 2025-11-05 code/cvc5mvn/src/test/scala/plaid/prelude/ast/ExprFuncTest.scala

@@ -1,0 +1,45 @@
+package plaid.prelude.ast
+
+import org.junit.Assert.assertEquals
+import org.junit.Test
+import plaid.prelude.antlr.Loader
+import plaid.prelude.ast.ListExprFuncExt.dependencyOrdered
+
+class ExprFuncTest {
+
+  /** Functions with no calls have no dependencies. */
+  @Test
+  def dependenciesWithoutCalls(): Unit =
+    val src = "exprfunctions: f(i) { m[\"x\" ++ i] }"
+    val fn = Loader.program(src).exprFuncs.head
+    assertEquals(Set(), fn.exprDependencies())
+
+  /** Functions consisting of a call has just that dependency. */
+  @Test
+  def dependenciesJustCall(): Unit =
+    val src = "exprfunctions: f(i) { g(i) }"
+    val fn = Loader.program(src).exprFuncs.head
+    assertEquals(Set(Identifier("g")), fn.exprDependencies())
+
+  /** A chain of functions that call each other have the correct forward dependency order when sorted. */
+  @Test
+  def dependencyOrderChain(): Unit =
+    val src = "exprfunctions: f() { g() } g() { h() } h() { 0 }"
+    val List(f, g, h) = Loader.program(src).exprFuncs
+    val sortedOrder = List(h, g, f)
+    assertEquals(sortedOrder, List(f, g, h).dependencyOrdered())
+    assertEquals(sortedOrder, List(g, f, h).dependencyOrdered())
+    assertEquals(sortedOrder, List(h, f, g).dependencyOrdered())
+    assertEquals(sortedOrder, List(h, g, f).dependencyOrdered())
+
+  /** A function that has multiple dependencies appears after both when sorted. */
+  @Test
+  def dependencyOrderMultiple(): Unit =
+    val src = "exprfunctions: f() { g() ++ h() } g() { 0 } h() { 0 }"
+    val List(f, g, h) = Loader.program(src).exprFuncs
+    assertEquals(f, List(f, g, h).dependencyOrdered().last)
+    assertEquals(f, List(g, f, h).dependencyOrdered().last)
+    assertEquals(f, List(h, f, g).dependencyOrdered().last)
+    assertEquals(f, List(h, g, f).dependencyOrdered().last)
+
+}

### 2025-11-05 code/cvc5mvn/src/test/scala/plaid/prelude/ast/ExprTest.scala

@@ -1,25 +1,0 @@
-package plaid.prelude.ast
-
-import org.junit.Assert.assertEquals
-import org.junit.Test
-import plaid.prelude.antlr.Loader
-
-class ExprTest {
-
-  /** Functions with no calls have no dependencies. */
-  @Test
-  def dependenciesWithoutCalls(): Unit = {
-    val src = "exprfunctions: f(i) { m[\"x\" ++ i] }"
-    val fn = Loader.program(src).exprFuncs.head
-    assertEquals(Set(), fn.exprDependencies())
-  }
-
-  /** Functions consisting of a call has just that dependency. */
-  @Test
-  def dependenciesJustCall(): Unit = {
-    val src = "exprfunctions: f(i) { g(i) }"
-    val fn = Loader.program(src).exprFuncs.head
-    assertEquals(Set(Identifier("g")), fn.exprDependencies())
-  }
-
-}

### 2025-11-05 code/cvc5mvn/src/main/scala/plaid/prelude/ast/Cmd.scala
fix expansion bug

@@ -7,13 +7,13 @@ sealed trait Cmd extends Node {
    * for looking up expression functions, and bindings are for any identifiers
    * that should be replaced by expressions.
    */
-  def expandExpr(ctx: List[ExprFunc], bindings: Map[Identifier, Expr] = Map()): Cmd = this match
+  def expand(ctx: List[ExprFunc], bindings: Map[Identifier, Expr] = Map()): Cmd = this match
     case AssertCmd(e1, e2, e3) => AssertCmd(e1.expand(ctx, bindings), e2.expand(ctx, bindings), e3.expand(ctx, bindings))
     case AssignCmd(e1, e2) => AssignCmd(e1.expand(ctx, bindings), e2.expand(ctx, bindings))
     case CallCmd(id, parms) => CallCmd(id, parms.map(_.expand(ctx, bindings)))
     case LetCmd(y, e, c) =>
       val b = bindings.updated(y, e.expand(ctx, bindings))
-      LetCmd(y, e, c.map(_.expandExpr(ctx, b)))
+      LetCmd(y, e, c.map(_.expand(ctx, b)))
 }
 
 case class AssertCmd(e1: Expr, e2: Expr, e3: Expr) extends Cmd

### 2025-11-05 code/cvc5mvn/src/main/scala/plaid/prelude/ast/Constraint.scala

@@ -13,27 +13,20 @@ sealed trait Constraint extends Node {
     case other => f(other).getOrElse(other))
 
   /**
-   * Expand all the expressions contained inside this constraint. The ctx is
-   * for looking up expression functions, and bindings are for any identifiers
-   * that should be replaced by expressions.
-   */
-  def expandExpr(ctx: List[ExprFunc], bindings: Map[Identifier, Expr] = Map()): Constraint = transform {
-    case CallConstraint(id, parms) => Some(CallConstraint(id, parms.map(_.expand(ctx, bindings))))
-    case EqualConstraint(e1, e2) => Some(EqualConstraint(e1.expand(ctx, bindings), e2.expand(ctx, bindings)))
-    case _ => None
-  }
-
-  /**
    * Recursively inline any constraint function calls contained inside this
    * constraint. All expressions in all constraints should be expanded before
    * this is used, but it would probably work anyway.
    */
-  def expand(ctx: List[ConstraintFunc] = List(), bindings: Map[Identifier, Expr] = Map()): Constraint = transform {
+  def expand(
+      exprCtx: List[ExprFunc] = List(),
+      constraintCtx: List[ConstraintFunc] = List(),
+      bindings: Map[Identifier, Expr] = Map()): Constraint = transform {
+    case EqualConstraint(e1, e2) => Some(EqualConstraint(e1.expand(exprCtx, bindings), e2.expand(exprCtx, bindings)))
     case CallConstraint(id, parms) =>
-      val f = ctx.resolve(id)
+      val f = constraintCtx.resolve(id)
       val formalParms = f.parms
-      val actualParms = parms.map(_.expand(bindings = bindings))
-      Some(f.body.expand(ctx, Map.from(formalParms.zip(actualParms))))
+      val actualParms = parms.map(_.expand(exprCtx, bindings))
+      Some(f.body.expand(exprCtx, constraintCtx, Map.from(formalParms.zip(actualParms))))
     case _ => None
   }
 }

### 2025-11-05 code/cvc5mvn/src/main/scala/plaid/prelude/ast/Func.scala

@@ -75,13 +75,13 @@ case class ExprFunc(id: Identifier, parms: List[Identifier], body: Expr) extends
 case class ConstraintFunc(id: Identifier, parms: List[Identifier], body: Constraint) extends Func {
   /** Expand expression and constraint function calls in this constraint function. */
   def expand(exprCtx: List[ExprFunc], constraintCtx: List[ConstraintFunc]): ConstraintFunc =
-    copy(body = body.expandExpr(exprCtx).expand(constraintCtx))
+    copy(body = body.expand(exprCtx, constraintCtx))
 }
 
 case class CmdFunc(id: Identifier, typedVariables: List[TypedIdentifier], body: List[Cmd], precond: Option[Constraint], postcond: Option[Constraint]) extends Func {
   /** Expand expression and constraint function calls in this constraint function. */
-  def expand(exprCtx: List[ExprFunc], constraintCtx: List[ConstraintFunc]): CmdFunc = copy(
-    body = body.map(_.expandExpr(exprCtx)),
-    precond = precond.map(_.expand(constraintCtx)),
-    postcond = postcond.map(_.expand(constraintCtx)))
+  def expand(exprCtx: List[ExprFunc] = Nil, constraintCtx: List[ConstraintFunc] = Nil): CmdFunc = copy(
+    body = body.map(_.expand(exprCtx)),
+    precond = precond.map(_.expand(exprCtx, constraintCtx)),
+    postcond = postcond.map(_.expand(exprCtx, constraintCtx)))
 }

### 2025-11-05 code/cvc5mvn/src/test/scala/plaid/prelude/ast/CmdFuncTest.scala

@@ -8,18 +8,28 @@ class CmdFuncTest {
 
   /** Functions with no calls have no dependencies. */
   @Test
-  def dependenciesWithoutCalls(): Unit = {
+  def dependenciesWithoutCalls(): Unit =
     val src = "cmdfunctions: f() { m[\"x\"]@1 := 3 }"
     val fn = Loader.program(src).cmdFuncs.head
     assertEquals(Set(), fn.cmdDependencies())
-  }
 
   /** Functions consisting of a call has just that dependency. */
   @Test
-  def cmdfunctions(): Unit = {
+  def dependenciesJustCall(): Unit =
     val src = "cmdfunctions: f() { g() }"
     val fn = Loader.program(src).cmdFuncs.head
     assertEquals(Set(Identifier("g")), fn.cmdDependencies())
-  }
 
+  /** When a function gets expanded, so do its preconditions and postconditions. */
+  @Test
+  def expandPreAndPost(): Unit =
+    val src ="""
+      cmdfunctions:
+      precondition: ("ab" == "a" ++ "b")
+      f() { m["x"]@1 := 3 }
+      postcondition: ("y" ++ "z" == "yz") """
+
+    val fn = Loader.program(src).cmdFuncs.head.expand()
+    assertEquals(Loader.constraint(""" "ab" == "ab" """), fn.precond.get)
+    assertEquals(Loader.constraint(""" "yz" == "yz" """), fn.postcond.get)
 }

### 2025-11-05 code/cvc5mvn/src/test/scala/plaid/prelude/ast/ConstraintTest.scala

@@ -1,0 +1,16 @@
+package plaid.prelude.ast
+
+import org.junit.Assert.assertEquals
+import org.junit.Test
+import plaid.prelude.antlr.Loader
+
+class ConstraintTest {
+
+  /** Expanding simplifies embedded expressions. */
+  @Test
+  def expandPreAndPost(): Unit =
+    val src = """ "ab" == "a" ++ "b" """
+    val constraint = Loader.constraint(src).expand()
+    assertEquals(Loader.constraint(""" "ab" == "ab" """), constraint)
+
+}

### 2025-11-05 code/cvc5mvn/src/test/scala/plaid/prelude/ast/ExprTest.scala

@@ -1,0 +1,16 @@
+package plaid.prelude.ast
+
+import org.junit.Assert.assertEquals
+import org.junit.Test
+import plaid.prelude.antlr.Loader
+
+class ExprTest {
+
+  /** Expanding simplifies the concatenation of two string literals. */
+  @Test
+  def simplifyLiteralConcat(): Unit =
+    val src = """ "a" ++ "b" """
+    val expr = Loader.expression(src).expand()
+    assertEquals(Loader.expression(""" "ab" """), expr)
+
+}

### 2025-11-05 toplas26/sty.tex
checkpoint

\section{Share Types}
\label{section-sty}

It is well-known that adding or subtracting a sample from a uniform
distribution in a finite field yields a value in a uniform
distribution, meaning that samples can be used as one-time-pads with
perfect secrecy \cite{barthe2019probabilistic,darais2019language}.

$$
\begin{array}{rclr}
  \sty &::=& \sharety \mid \pubty \mid \outty & \textit{share types} \\
  \stt &::=& [\phi,\phi] \mid \pubx{\phi} & \textit{share terms} 
\end{array}
$$

\stylocalfig

\begin{mathpar}
  \inferrule[Share]
      {\eqs \models \mx{z}{\cid_2} \eop \phi - \rx{y}{\cid_1} \\ \cid_1 \ne \cid_2}
      {\eqs \vdash \Gamma,R \stredx \Gamma \cup \setit{[\mx{z}{\cid_2}, \rx{y}{\cid_1}] : \sharety}, R \uplus \setit{\rx{y}{\cid_1}}}

  \inferrule[SecOpen]
      {
        \Gamma, \eqs \vdash [\px{x},\px{y}] : \sharety\\
        \eqs \models  \px{x} + \px{y} \eop \phi' - \rx{y}{\oid}
      }
      {\eqs \vdash \Gamma,R \stredx \Gamma \cup \setit{\pubx{\px{x} + \px{y}}  : \pubty}, R \uplus \setit{\rx{y}{\cid_o}}}

  \inferrule[Reveal]
      {
        \Gamma, \eqs \vdash [\px{x},\px{y}] : \sharety
      }
      {\eqs \vdash \Gamma,R \stredx \Gamma \cup \setit{\pubx{\px{x} + \px{y}}  : \outty}, R}

  \inferrule[Output]
      {
        \Gamma, \eqs \vdash \pubx{\out{\cid}} : \outty
      }
      {\eqs \vdash \Gamma,R \stredx \Gamma \uplus \setit{\pubx{\out{\cid}} : \outty}, R}
\end{mathpar}

### 2025-11-05 toplas26/sty.tex
checkpoint

@@
 $$
-\begin{array}{rclr}
+\begin{array}{rcl@{\hspace{10mm}}l}
   \sty &::=& \sharety \mid \pubty \mid \outty & \textit{share types} \\
@@
 
+\subsection{Non-interactive typing rules}
+
 \stylocalfig
 
-\begin{mathpar}
-  \inferrule[Share]
-      {\eqs \models \mx{z}{\cid_2} \eop \phi - \rx{y}{\cid_1} \\ \cid_1 \ne \cid_2}
-      {\eqs \vdash \Gamma,R \stredx \Gamma \cup \setit{[\mx{z}{\cid_2}, \rx{y}{\cid_1}] : \sharety}, R \uplus \setit{\rx{y}{\cid_1}}}
+\subsection{Interactive typing rules}
 
-  \inferrule[SecOpen]
-      {
-        \Gamma, \eqs \vdash [\px{x},\px{y}] : \sharety\\
-        \eqs \models  \px{x} + \px{y} \eop \phi' - \rx{y}{\oid}
-      }
-      {\eqs \vdash \Gamma,R \stredx \Gamma \cup \setit{\pubx{\px{x} + \px{y}}  : \pubty}, R \uplus \setit{\rx{y}{\cid_o}}}
+\styinteractivefig
 
-  \inferrule[Reveal]
-      {
-        \Gamma, \eqs \vdash [\px{x},\px{y}] : \sharety
-      }
-      {\eqs \vdash \Gamma,R \stredx \Gamma \cup \setit{\pubx{\px{x} + \px{y}}  : \outty}, R}
+\subsection{Share Type Soundness}
 
-  \inferrule[Output]
-      {
-        \Gamma, \eqs \vdash \pubx{\out{\cid}} : \outty
-      }
-      {\eqs \vdash \Gamma,R \stredx \Gamma \uplus \setit{\pubx{\out{\cid}} : \outty}, R}
-\end{mathpar}
+\begin{definition}
+  We write $\eqs,\Gamma,R : {OK}$ iff both of the following conditions hold:
+  \begin{enumerate}[\hspace{5mm}i.]
+  \item For all $([\mx{z}{\cid}, \rx{y}{\cid'}] : \sharety) \in \Gamma$,
+    we have $\eqs \models \mx{z}{\cid} \eop \phi - \rx{x}{\cid'}$ for
+    some $\phi$ and unique $\rx{x}{\cid'} \in R$, with $\cid \ne \cid'$.
+  \item For all $([\phi] : \pubty) \in \Gamma$,
+    we have $\eqs \models \phi \eop \phi' - \rx{x}{\oid}$ for
+    some $\phi'$ and unique $\rx{x}{\oid} \in R$.
+  \end{enumerate}
+\end{definition}
+
+\begin{lemma}
+  Given $\eqs,\Gamma,R  : {OK}$.
+  and $\eqs  \vdash \Gamma,R \stredx R',\Gamma'$.
+  Then $\eqs, \Gamma',R' : {OK}$.
+\end{lemma}
+
+\begin{definition} Define $\ov(\Gamma)$ as follows:
+\begin{eqnarray*}
+\ov(\varnothing) &=& \varnothing \\[-1.5mm]
+\ov(\setit{[\mx{z}{\cid}, \rx{y}{\cid'}] : \sharety} \cup \Gamma) &=& \setit{\mx{z}{\cid}} \cup \ov(\Gamma)\\[-1.5mm]
+\ov(\setit{[\phi + \px{x}] : \pubty} \cup \Gamma) &=& \setit{\px{x}} \cup \ov(\Gamma)\\[-1.5mm]
+\ov(\setit{[\phi] : \outty} \cup \Gamma) &=& \setit{\phi} \cup \ov(\Gamma)
+\end{eqnarray*}
+\end{definition}
+
+\begin{lemma}
+  Assume given $\Gamma,R, \eqspre \wedge \toeq{\prog} : {OK}_\cid$  and
+  $\iov(\prog) = S \cup V \cup O$
+  and $\Gamma, \eqspre \wedge \toeq{\prog} \vdash [\phi_1,\phi_2] : \sharety$.
+  Then, letting $X \defeq \ov(\Gamma) \cap \realviews{\setit{\cid_1}}{\setit{\cid_2}}$:
+  $$
+  \condd{\progtt(\prog)}{S_{\setit{\cid_1}}}{S_{\setit{\cid_2}}  \cup X \cup \setit{\phi_1 + \phi_2}}
+  = 
+  \condd{\progtt(\prog)}{S_{\setit{\cid_1}}}{S_{\setit{\cid_2}} \cup X \cup \setit{\phi_1 + \phi_2, \phi_1}}
+  $$
+\end{lemma}
+
+\begin{lemma}
+  Assume given $\eqspre \wedge \toeq{\prog}, \Gamma,R: {OK}_\cid$  and
+  $\iov(\prog) = S \cup V \cup O$
+  and $\Gamma, \eqspre \wedge \toeq{\prog} \vdash [\phi_1,\phi_2] : \sharety$.
+  Then, letting $X \defeq \ov(\Gamma) \cap \realviews{\setit{\cid_1}}{\setit{\cid_2}}$ for
+  $\cid_1 \ne \cid_2$:
+  $$
+  \condd{\progtt(\prog)}{S_{\setit{\cid_1}}}{S_{\setit{\cid_2}}  \cup X \cup \setit{\phi_1 + \phi_2}}
+  = 
+  \condd{\progtt(\prog)}{S_{\setit{\cid_1}}}{S_{\setit{\cid_2}} \cup X \cup \setit{\phi_1 + \phi_2, \phi_1}}
+  $$
+\end{lemma}
+
+\begin{lemma}
+  Assume given $\eqspre \wedge \toeq{\prog}, \Gamma_1,R_1: {OK}_\cid$  and
+  $\iov(\prog) = S \cup V \cup O$
+  and $\eqspre \wedge \toeq{\prog} \vdash \Gamma_1,R_1 \stredx \Gamma_2,R_2$.
+  Letting $X_1 \defeq \ov(\Gamma_1) \cap \realviews{\setit{\cid_1}}{\setit{\cid_2}}$,
+  for $\cid_1 \ne \cid_2$ assume also:
+  $$
+  \condd{\progtt(\prog)}{S_{\setit{\cid_1}}}{S_{\setit{\cid_2}}  \cup X_1 \cup O}
+  = 
+  \condd{\progtt(\prog)}{S_{\setit{\cid_1}}}{S_{\setit{\cid_2}} \cup O}
+  $$
+  Then, letting $X_2 \defeq \ov(\Gamma_2) \cap \realviews{\setit{\cid_1}}{\setit{\cid_2}}$,
+  we have:
+  $$
+  \condd{\progtt(\prog)}{S_{\setit{\cid_1}}}{S_{\setit{\cid_2}}  \cup X_2 \cup O}
+  = 
+  \condd{\progtt(\prog)}{S_{\setit{\cid_1}}}{S_{\setit{\cid_2}} \cup O}
+  $$
+\end{lemma}
+
+\begin{theorem}[Share Type Soundness]
+  Assume given $\prog$ with 
+  $\iov(\prog) = S \cup V \cup O$
+  and $\eqspre \wedge \toeq{\prog} \vdash \varnothing,\varnothing \stredx^*
+  \Gamma,R$ where $\realviews{\setit{\cid_1}}{\setit{\cid_2}} \cup O
+  \subseteq \ov(\Gamma)$ for all $\cid_1$, $\cid_2$ with $\cid_1 \ne \cid_2$.
+  Then $\prog$ satisfies noninterference modulo output.
+\end{theorem}

### 2025-11-07 toplas26/smt.tex
checkpoint

@@
 \subsubsection{Example: Correctness of 3-Party Addition}
+\label{section-smt-example}
 Consider the following 3-party additive secret sharing protocol, where
@@
 \begin{proof}
+  \cnote{de-obfuscate.}
   Assume $\htrip{\eqs_1}{\prog}{\eqs_2}$ is derivable without \TirName{Hyp}.

### 2025-11-07 toplas26/sty.tex
checkpoint

@@
 
-It is well-known that adding or subtracting a sample from a uniform
-distribution in a finite field yields a value in a uniform
-distribution, meaning that samples can be used as one-time-pads with
-perfect secrecy \cite{barthe2019probabilistic,darais2019language}.
+Our approach to static analysis is to enforce the use of sound
+techniques in semi-homorphic encryption in protocols.  As discussed in
+the example in Section \ref{section-smt-example}, protocols integrate
+interactive elements, such as the sharing of one-time-pad (OTP)
+encrypted secrets, and noninteractive elements, such as addition (of
+shares). We approach this by introducing distinct judgement forms for
+interactive and noninteractive protocol elements. Our type system
+as presented will assume 2-client protocols, but this is for
+simplicity and the analysis can be generalized to more parties.
 
+Both interactive and non-interactive judgements leverage the use
+of SMT interpretations as described in Section \ref{section-smt}.
+We introduce a \emph{share type} language $\sty$, classifying
+\emph{share terms}. We represent share terms using standard notation
+\cite{10.1007/978-3-030-68869-1_3}. A share term $[\phi_1,\phi_2]$
+represents a secret-shared value where client 1 has $\phi_1$ and
+client 2 has $\phi_2$, and this shared value can be reconstructed
+as $\phi_1 + \phi_2$.\cnote{Comment on generation and
+  reconstruction in general somewhere.} A share term $\pubx{\phi}$
+denotes a publicly shared value.
 $$
@@
 
-\subsection{Non-interactive typing rules}
+\subsection{Non-Interactive Typing Rules}
 
@@
 
-\subsection{Interactive typing rules}
+\subsection{Interactive Typing Rules}
+
+It is well-known that adding or subtracting a sample from a uniform
+distribution in a finite field yields a value in a uniform
+distribution, meaning that samples can be used as one-time-pads with
+perfect secrecy \cite{barthe2019probabilistic,darais2019language}.
 
 \styinteractivefig
+
+\subsection{Share Typing Example} To illustrate details of the
+type system we return to the example of additive sharing in the context
+of a 2-client protocol. We first focus on initial secret sharing steps--
+let $\prog$ be the following protocol in $\fieldp{2}$ (noting
+that $-$ is equivalent to xor in $\fieldp{2}$):
+$$
+\begin{array}{lll}
+  \elab{\mesg{s1}}{2} &:=& \elab{(\secret{1} \fminus \locflip)}{1}; \\
+  \elab{\mesg{s2}}{1} &:=& \elab{(\secret{2} \fminus \locflip)}{2} 
+\end{array}
+$$
+We can then assert:
+$$
+\toeq{\prog} \vdash \varnothing,\varnothing \stredx
+\setit{[\mx{s2}{1}, \elab{\locflip}{2}] : \sharety,
+  [\mx{s1}{2}, \elab{\locflip}{1}] : \sharety}, \setit{\elab{\locflip}{1},\elab{\locflip}{2}}
+$$
+Suppose the protocol goes on to reveal the negation of the
+sum (xor-ing) of the inputs:
+$$
+\begin{array}{lll}
+  \px{1} &:=& \elab{\mesg{s2} + \locflip + 1}{1}; \\
+  \px{2} &:=& \elab{\mesg{s1} + \locflip}{2}\\
+  \out{1} &:=& \elab{\px{1} + \px{2}}{1}\\
+  \out{2} &:=& \elab{\px{1} + \px{2}}{2}
+\end{array}
+$$
+Letting this extended protocol be $\prog'$, and letting:
+$$
+\begin{array}{rcl}
+\Gamma &\defeq& \setit{[\mx{s2}{1}, \elab{\locflip}{2}] : \sharety,
+  [\mx{s1}{2}, \elab{\locflip}{1}] : \sharety}\\
+R &\defeq&  \setit{\elab{\locflip}{1},\elab{\locflip}{2}}
+\end{array}
+$$
+observe we can then obtain share typings for noninteractive components for
+addition of shares by \TirName{Shared} and \TirName{HeAdd}:
+$$
+\Gamma, \toeq{\prog'} \vdash
+[\mx{s2}{1} + \elab{\locflip}{1}, \mx{s1}{2} + \elab{\locflip}{2}] : \sharety
+$$
+and hence for negation by \TirName{HeAddPub}:
+$$
+\Gamma, \toeq{\prog'} \vdash
+[\mx{s2}{1} + \elab{\locflip}{1} + 1, \mx{s1}{2} + \elab{\locflip}{2}] : \sharety
+$$
+This implies:
+$$
+\toeq{\prog'} \vdash \varnothing,\varnothing \stredx
+\Gamma \cup \setit{\pubx{\px{1} + \px{2}} : \outty}, R 
+$$
+Hence by \TirName{PubEntails} we have:
+$$
+\begin{array}{l}
+\Gamma \cup \setit{\pubx{\px{1} + \px{2}} : \outty}, \toeq{\prog'} \vdash
+\pubx{\out{1}} : \outty \\
+\Gamma \cup \setit{\pubx{\px{1} + \px{2}} : \outty}, \toeq{\prog'} \vdash
+\pubx{\out{2}} : \outty
+\end{array}
+$$
 
@@
 \ov(\setit{[\mx{z}{\cid}, \rx{y}{\cid'}] : \sharety} \cup \Gamma) &=& \setit{\mx{z}{\cid}} \cup \ov(\Gamma)\\[-1.5mm]
-\ov(\setit{[\phi + \px{x}] : \pubty} \cup \Gamma) &=& \setit{\px{x}} \cup \ov(\Gamma)\\[-1.5mm]
-\ov(\setit{[\phi] : \outty} \cup \Gamma) &=& \setit{\phi} \cup \ov(\Gamma)
+\ov(\setit{[\px{x} + \px{y}] : \pubty} \cup \Gamma) &=& \setit{\px{x},\px{y}} \cup \ov(\Gamma)\\[-1.5mm]
+\ov(\setit{[\out{\cid}] : \outty} \cup \Gamma) &=& \setit{\out{\cid}} \cup \ov(\Gamma)
 \end{eqnarray*}

### 2025-11-07 toplas26/sty.tex
checkpoint

@@
 \end{array}
-$$
+$$ We proceed by formally describing the type analysis, then
+discussing an illustrative example, and finally by demonstrating share
+type soundness (Theorem \ref{theorem-sty}).
 
 \subsection{Non-Interactive Typing Rules}
+
+Our non-interactive typing rules are presented in Figure
+\ref{fig-stylocal}. Judgmements are of the form $\Gamma, \eqs \vdash
+\stt : \sty$, where $\Gamma$ is a set of \emph{share type bindings} of
+the form $\stt : \sty$. In the context of typing a protocol $\prog$
+soundness requires that the constraint $\eqs$ must be entailed by
+$\eqspre \wedge \toeq{\prog}$ (see, e.g., Lemma
+\ref{lemma-sty-noninteractive-sound}).
+
+The syntax-directed rules all known semi-homomorphic operations,
+including addition of shares (\TirName{HEAdd}), and both addition and
+multiplication of shares with a public value (\TirName{HEAddPub} and
+\TirName{HEMultPub}, respectively) \cite{10.1007/978-3-030-68869-1_3}.
+These rules also leverage SMT entailment in the \TirName{ShareEntails}
+and \TirName{PubEntails} rules, since these operations should be
+allowable up to algebraic equivalence in the context of a protocol.
 
@@
 \begin{lemma}
+  \label{lemma-sty-noninteractive-sound}
   Assume given $\eqspre \wedge \toeq{\prog}, \Gamma,R: {OK}_\cid$  and
@@
 \begin{theorem}[Share Type Soundness]
+  \label{theorem-sty}
   Assume given $\prog$ with 

### 2025-11-07 toplas26/sty.tex
checkpoint

@@
 perfect secrecy \cite{barthe2019probabilistic,darais2019language}.
+
+Of course, there are other methods for encrypting values in MPC
+protocols-- but we can observe that many are algebraically equivalent
+to this fundamental schema. For example, in Yao's Garbled Circuits
+(YGC) in $\mathbb{F}_{2}$, the ``garbler'' represents encrypted wire
+values as a random sample (for 1) or its negation (for 0), and shares
+its secret input encoding with the ``evaluator'' by using
+it as a selection bit. That is, assuming that client 2 is the garbler and
+client 1 is the evaluator, we can define:
+$$
+\xassign{\mx{w}{1}}{\mux{\secret{w}}{\neg\flip{w}}{\flip{w}}}{2}
+$$
+where for all $\be_1,\be_2,\be_3$ (with $\neg$ denoting negation in $\mathbb{F}_2$):
+$$
+\mux{\be_1}{\be_2}{\be_3} \defeq (\neg\be_1 \ftimes \be_2) \fplus (\be_1 \ftimes \be_3)
+$$
+and letting this protocol be $\prog$ the following is valid:
+\begin{mathpar}
+%\toeq{\xassign{\mx{w}{1}}{\OT{\secret{w}}{1}{\neg\flip{w}}{\flip{w}}}{2}}
+   \toeq{\prog} \models \mx{w}{1} \eop \neg\sx{w}{2} \fplus \rx{w}{2}
+\end{mathpar}
+This allows us to assign the same ciphertype to $\mx{w}{1}$ as above--
+that is, $\cty{\rx{w}{2}}{\setit{\sx{w}{2}}}$.  Since this is a
+one-time-pad scheme, it is also important that we ensure that samples
+are used at most once for encryption. As in previous work 
+\cite{darais2019language} we use type linearity to enforce this.
 

### 2025-11-07 toplas26/sty.tex
checkpoint

@@
 \end{array}
-$$ We proceed by formally describing the type analysis, then
+$$
+We proceed by formally describing the type analysis, then
 discussing an illustrative example, and finally by demonstrating share
-type soundness (Theorem \ref{theorem-sty}).
+type soundness in Theorem \ref{theorem-sty}. This result shows that
+our analysis enforces noninterference modulo output, in contrast to
+previous systems that only enforce a gradual release property
+(\cite{skalka-near-esop25}). By connecting this analysis with
+protocol correctness properties enforced by the Hoare logic
+discussed in Section \ref{section-smt}, the output can also
+be connected to an intended ideal functionality.
 
 \subsection{Non-Interactive Typing Rules}
+
+\stylocalfig
 
@@
 allowable up to algebraic equivalence in the context of a protocol.
-
-\stylocalfig
 

### 2025-11-07 toplas26/sty.tex
checkpoint

--- 2025-11-07 toplas26/sty.tex
@@
 \end{array}
-$$ We proceed by formally describing the type analysis, then
+$$
+Intuitively, the meaning of share types $\sty$ are as follows.
+When we ascribe the type $[\phi_1,\phi_2] : \sharety$, this
+means that $[\phi_1,\phi_2]$ are shares of their
+reconstruction, which we can represent as $\phi_1 + \phi_2$
+as discussed in Section \ref{???}. 
+In particular, each of $\phi_1$ and $\phi_2$ independently
+reveal nothing about $\phi_1 + \phi_2$. Likewise,
+$\pubx{\phi} : \outty$ means that $\phi$ has been
+publicly \emph{revealed} via broadcast shares, and our
+type system requires that $\phi \eop \phi_1 + \phi_2$ with
+$[\phi_1,\phi_2] : \sharety$. This implies that the
+process of revealing $\phi$ reveals nothing besides $\phi$.
+The ascription $\pubx{\phi} : \pubty$ is similar, with the
+additional meaning that $\phi$ is a publicly-known value
+independent of any client inputs. This includes known
+constant values used in protocols, and more interestingly
+\emph{secure openings}. Secure openings are commonly used
+to interactively compute public values that are masked with
+OTPs secret-shared with clients in initialization protocols,
+as for example in Beaver Triple multiplication which
+we discuss in Section \ref{section-examples}.
+
+\stylocalfig
+
+We proceed by formally describing the type analysis, then
 discussing an illustrative example, and finally by demonstrating share
-type soundness (Theorem \ref{theorem-sty}).
+type soundness in Theorem \ref{theorem-sty}. This result shows that
+our analysis enforces noninterference modulo output, in contrast to
+the previous system for $\minicat$ that only enforce a gradual release
+property \cite{skalka-near-esop25}. By connecting this analysis with
+protocol correctness properties enforced by the Hoare logic discussed
+in Section \ref{section-smt}, the output can also be connected to an
+intended ideal functionality.\cnote{Update intro grid accordingly.}
 
@@
 
-The syntax-directed rules all known semi-homomorphic operations,
+The syntax-directed rules are all known semi-homomorphic operations,
 including addition of shares (\TirName{HEAdd}), and both addition and
@@
 
-\stylocalfig
-
 \subsection{Interactive Typing Rules}
 
+As discussed in Section \ref{???}, a \emph{secret sharing scheme}
+includes functions $\genshares$ and $\recon$ which generate
+and reconstruct shares, respectively.
 It is well-known that adding or subtracting a sample from a uniform
 distribution in a finite field yields a value in a uniform
-distribution, meaning that samples can be used as one-time-pads with
+distribution, meaning that samples can be used as OTPs with
 perfect secrecy \cite{barthe2019probabilistic,darais2019language}.
+So we can define:
+\begin{eqnarray*}
+  \genshares(\phi) &\defeq& [\phi - \rx{y}{\cid_2}, \rx{y}{\cid_2}]\\
+  \recon([\phi_1,\phi_2]) &\defeq& \phi_1 \fplus \phi_2
+\end{eqnarray*}
+where we require that $\genshares$ consumes the randomness
+$\setit{\rx{y}{\cid_2}}$ for use as an OTP.
 
 Of course, there are other methods for encrypting values in MPC
-protocols-- but we can observe that many are algebraically equivalent
+protocols that can be shown to be algebraically equivalent
 to this fundamental schema. For example, in Yao's Garbled Circuits
@@
 \end{mathpar}
+For this reason SMT entailment is also an important enabling technology
+for the interactive rules.
+
+\styinteractivefig
+
+The interactive share typing rules are defined in Figure \ref{fig-styinteractive},
+with judgements of the form $\eqs \vdash \Gamma_1,R_1, \stredx \Gamma_2,R_2$.
+Each such judgement establishes a type classification for interactively-defined
+variables-- that is, unicast and broadcast messages exchanged between
+clients.
+
 This allows us to assign the same ciphertype to $\mx{w}{1}$ as above--
@@
 \cite{darais2019language} we use type linearity to enforce this.
-
-\styinteractivefig
 

### 2025-11-07 code/cvc5mvn/src/main/scala/plaid/prelude/App.scala
magic functions

@@ -34,7 +34,7 @@ class App extends Runnable {
     var totalTests = 0
     var totalFailures = 0
 
-    contracts.foreach(x =>
+    contracts.filter(_.f.body.nonEmpty).foreach(x =>
       println(s"\n*** ${x.f.id.name}")
       val termManager = TermManager()
       val termFactory = TermFactory(termManager, fieldSize)
@@ -45,7 +45,7 @@ class App extends Runnable {
         val test = if ent.origin == x.f then "(postcondition)" else ent.origin.prettyPrint()
         val status = if failures.contains(ent) then "FAIL" else "PASS"
         println(s"[$status] $test")))
-    println(s"Ran $totalTests tests\nHad $totalFailures failures")
+    println(s"\nRan $totalTests tests\nHad $totalFailures failures")
 }
 
 object App {

### 2025-11-07 code/cvc5mvn/src/main/scala/plaid/prelude/logic/Contract.scala

@@ -57,5 +57,6 @@ extension (trg: CmdFunc)
     val pre = trg.precond.getOrElse(TrueConstraint()).expand(exprFns, constraintFns)
     val ctx = trg.body.foldLeft(HoareContext(cons = pre)) { (acc, x) => acc.include(x, exprFns, contractCtx) }
     val post = trg.postcond.getOrElse(ctx.cons).expand(exprFns, constraintFns)
-    val overall = Entailment(trg, ctx.cons, post)
-    Contract(trg, pre, post, overall :: ctx.ent)
+    // Empty body for "magic" function that just gets admitted
+    val ent = if trg.body.isEmpty then Nil else Entailment(trg, ctx.cons, post) :: ctx.ent
+    Contract(trg, pre, post, ent)

### 2025-11-07 code/cvc5mvn/src/main/scala/plaid/prelude/logic/WellFormed.scala

@@ -5,8 +5,6 @@ import plaid.prelude.ast.*
 // check if a (final/evaluated) constraint is valid
 object WellFormed {
 
-  // TODO Apply constraints
-  
   /**
    * valid constraint node type
    */

### 2025-11-07 code/cvc5mvn/src/test/scala/plaid/prelude/cvc/CvcUtilsTest.scala

@@ -8,37 +8,32 @@ class CvcUtilsTest {
 
   /** Produces cvc5 constant names for all memory types. */
   @Test
-  def memNames(): Unit = {
+  def memNames(): Unit =
     assertEquals("m_x_4", CvcUtils.getCvcName(MessageExpr(Str("x")), 4))
     assertEquals("o_4", CvcUtils.getCvcName(OutputExpr(), 4))
     assertEquals("p_x", CvcUtils.getCvcName(PublicExpr(Str("x")), 4))
     assertEquals("r_x_4", CvcUtils.getCvcName(RandomExpr(Str("x")), 4))
     assertEquals("s_x_4", CvcUtils.getCvcName(SecretExpr(Str("x")), 4))
-  }
 
   /** Number nodes in the AST evalConstraint to integers. */
   @Test
-  def intEval(): Unit = {
+  def intEval(): Unit =
     val num = Num(3)
     assertEquals(3, CvcUtils.toInt(num))
-  }
 
   /** Nodes in the AST that are not numbers do not evalConstraint to integers. */
   @Test(expected = classOf[IllegalArgumentException])
-  def intEvalUnsupported(): Unit = {
+  def intEvalUnsupported(): Unit =
     CvcUtils.toInt(Str("x"))
-  }
 
   /** String nodes in the AST evalConstraint to strings. */
   @Test
-  def stringEval(): Unit = {
+  def stringEval(): Unit =
     val str = Str("x")
     assertEquals("x", CvcUtils.toString(str))
-  }
 
   /** Nodes in the AST that are not strings do not evalConstraint to strings. */
   @Test(expected = classOf[IllegalArgumentException])
-  def stringEvalUnsupported(): Unit = {
+  def stringEvalUnsupported(): Unit =
     CvcUtils.toString(Num(3))
-  }
 }

### 2025-11-07 toplas26/sty.tex
checkpoint

@@
 $$
-\xassign{\mx{w}{1}}{\mux{\secret{w}}{\neg\flip{w}}{\flip{w}}}{2}
+\xassign{\mx{w}{1}}{\mux{\secret{w}}{\flip{w}}{\neg\flip{w}}}{2}
 $$
@@
 %\toeq{\xassign{\mx{w}{1}}{\OT{\secret{w}}{1}{\neg\flip{w}}{\flip{w}}}{2}}
-   \toeq{\prog} \models \mx{w}{1} \eop \neg\sx{w}{2} \fplus \rx{w}{2}
+   \toeq{\prog} \models \mx{w}{1} \eop \sx{w}{2} \fminus \rx{w}{2}
 \end{mathpar}
@@
 with judgements of the form $\eqs \vdash \Gamma_1,R_1, \stredx \Gamma_2,R_2$.
-Each such judgement establishes a type classification for interactively-defined
+Each such judgement establishes a sound type binding for interactively-defined
 variables-- that is, unicast and broadcast messages exchanged between
-clients.
+clients. In addition to type bindings, judgments track random variables
+used as OTPs, allowing us to use type linearity to enforce one-time
+usage similar to previous related work \cite{darais2019language}.
+To this end, in these rules we define $R_1 \uplus R_2$ as the disjoint
+union of $R_1$ and $R_2$.
 
-This allows us to assign the same ciphertype to $\mx{w}{1}$ as above--
-that is, $\cty{\rx{w}{2}}{\setit{\sx{w}{2}}}$.  Since this is a
-one-time-pad scheme, it is also important that we ensure that samples
-are used at most once for encryption. As in previous work 
-\cite{darais2019language} we use type linearity to enforce this.
+The \TirName{Share} rule generates the binding $[\mx{z}{\cid_2}, \rx{y}{\cid_1}]$
+if this is a secret sharing as defined by $\genshares$. It also adds
+$\rx{y}{\cid_1}$ to the set of variables used as OTPs, ensuring that
+it has not already been used.
 

### 2025-11-07 toplas26/sty.tex
checkpoint

@@
 As discussed in Section \ref{???}, a \emph{secret sharing scheme}
-includes functions $\genshares$ and $\recon$ which generate
+includes functions $(\genshares,\recon)$ which generate
 and reconstruct shares, respectively.
@@
 
-The interactive share typing rules are defined in Figure \ref{fig-styinteractive},
-with judgements of the form $\eqs \vdash \Gamma_1,R_1, \stredx \Gamma_2,R_2$.
-Each such judgement establishes a sound type binding for interactively-defined
+The interactive share typing rules are defined in Figure
+\ref{fig-styinteractive}, with judgements of the form $\eqs \vdash
+\Gamma_1,R_1, \stredx \Gamma_2,R_2$.  Each such judgement establishes
+a sound type binding in $\Gamma_2$ for interactively-defined
 variables-- that is, unicast and broadcast messages exchanged between
-clients. In addition to type bindings, judgments track random variables
-used as OTPs, allowing us to use type linearity to enforce one-time
-usage similar to previous related work \cite{darais2019language}.
-To this end, in these rules we define $R_1 \uplus R_2$ as the disjoint
-union of $R_1$ and $R_2$.
+clients-- assuming the bindings in $\Gamma_1$. In addition to type
+bindings, judgments track random variables used as OTPs, allowing us
+to use type linearity to enforce one-time usage similar to previous
+related work \cite{darais2019language}, where any randomness used in
+the typing is not in $R_1$ and added to $R_2$. To this end, in these
+rules $\uplus$ is disjoint union.
 
-The \TirName{Share} rule generates the binding $[\mx{z}{\cid_2}, \rx{y}{\cid_1}]$
-if this is a secret sharing as defined by $\genshares$. It also adds
-$\rx{y}{\cid_1}$ to the set of variables used as OTPs, ensuring that
-it has not already been used.
+The \TirName{Share} rule generates the binding $[\mx{z}{\cid_2},
+  \rx{y}{\cid_1}] : \sharety$ if this is a secret sharing as defined
+by $\genshares$. It also adds $\rx{y}{\cid_1}$ to the set of variables
+used as OTPs, ensuring that it has not already been used. The
+\TirName{Reveal} rule generates the binding $\pubx{\px{x} + \px{y}} :
+\outty$ with the requirement that $[\px{x},\px{y}]$ are typable as
+shares. The \TirName{SecOpen} rule is similar except it allows the
+assertion $\pubx{\px{x} + \px{y}} : \pubty$ if we can additionally
+show that the shared value is masked by the use of random value
+$\rx{z}{\oid}$ as an OTP. Here we are positing a third party $\oid$
+that secret-shares $\rx{z}{\oid}$ as part of an initialization
+protocol. We illustrate secure openings in practice at more length
+in Section \ref{section-examples}, though we note that appropriate
+properties of initialization protocols can be encoded in $\eqspre$.
 

### 2025-11-07 code/cvc5mvn/src/main/scala/plaid/prelude/ast/Cmd.scala
fix let command bug

@@ -7,13 +7,14 @@ sealed trait Cmd extends Node {
    * for looking up expression functions, and bindings are for any identifiers
    * that should be replaced by expressions.
    */
-  def expand(ctx: List[ExprFunc], bindings: Map[Identifier, Expr] = Map()): Cmd = this match
-    case AssertCmd(e1, e2, e3) => AssertCmd(e1.expand(ctx, bindings), e2.expand(ctx, bindings), e3.expand(ctx, bindings))
-    case AssignCmd(e1, e2) => AssignCmd(e1.expand(ctx, bindings), e2.expand(ctx, bindings))
-    case CallCmd(id, parms) => CallCmd(id, parms.map(_.expand(ctx, bindings)))
+  def expand(ctx: List[ExprFunc], bindings: Map[Identifier, Expr] = Map()): List[Cmd] = this match
+    case AssertCmd(e1, e2, e3) =>
+      List(AssertCmd(e1.expand(ctx, bindings), e2.expand(ctx, bindings), e3.expand(ctx, bindings)))
+    case AssignCmd(e1, e2) => List(AssignCmd(e1.expand(ctx, bindings), e2.expand(ctx, bindings)))
+    case CallCmd(id, parms) => List(CallCmd(id, parms.map(_.expand(ctx, bindings))))
     case LetCmd(y, e, c) =>
       val b = bindings.updated(y, e.expand(ctx, bindings))
-      LetCmd(y, e, c.map(_.expand(ctx, b)))
+      c.flatMap(_.expand(ctx, b))
 }
 
 case class AssertCmd(e1: Expr, e2: Expr, e3: Expr) extends Cmd

### 2025-11-07 code/cvc5mvn/src/main/scala/plaid/prelude/logic/Contract.scala

@@ -18,7 +18,7 @@ case class HoareContext(cons: Constraint = TrueConstraint(), ent: List[Entailmen
   /** Add a new entailment to this Hoare context. */
   private def include(e: Entailment): HoareContext = copy(ent = e :: ent)
   /** Include a new command in this Hoare context. */
-  def include(cmd: Cmd, exprFns: List[ExprFunc], contracts: List[Contract]): HoareContext = cmd.expand(exprFns) match
+  def include(cmd: Cmd, contracts: List[Contract]): HoareContext = cmd match
     case AssertCmd(e1, e2, party) =>
       val a = e1.indexParties(Some(party))
       val b = e2.indexParties(Some(party))
@@ -55,7 +55,9 @@ extension (trg: List[CmdFunc])
 extension (trg: CmdFunc)
   def contract(exprFns: List[ExprFunc], constraintFns: List[ConstraintFunc], contractCtx: List[Contract]): Contract =
     val pre = trg.precond.getOrElse(TrueConstraint()).expand(exprFns, constraintFns)
-    val ctx = trg.body.foldLeft(HoareContext(cons = pre)) { (acc, x) => acc.include(x, exprFns, contractCtx) }
+    val ctx = trg.body
+      .flatMap(_.expand(exprFns))
+      .foldLeft(HoareContext(cons = pre)) { (acc, x) => acc.include(x, contractCtx) }
     val post = trg.postcond.getOrElse(ctx.cons).expand(exprFns, constraintFns)
     // Empty body for "magic" function that just gets admitted
     val ent = if trg.body.isEmpty then Nil else Entailment(trg, ctx.cons, post) :: ctx.ent

### 2025-11-07 code/cvc5mvn/src/test/scala/plaid/prelude/ast/CmdTest.scala

@@ -1,0 +1,35 @@
+package plaid.prelude.ast
+
+import org.junit.Assert.assertEquals
+import org.junit.Test
+import plaid.prelude.antlr.Loader
+
+class CmdTest {
+
+  /** Expressions simplify when an assign command is expanded. */
+  @Test
+  def expandAssignCommand(): Unit =
+    val cmd = Loader.command(""" m["x" ++ "y"]@1 := m["u" ++ "v"]@2 """)
+    val expected = List(Loader.command(""" m["xy"]@1 := m["uv"]@2 """))
+    assertEquals(expected, cmd.expand(Nil))
+
+  /** Expanding let commands substitutes the binding in the let body. */
+  @Test
+  def expandLetCommand(): Unit =
+    val command = Loader.command("""let table = "foo" in m[table]@1 := r["x"]@2""")
+    assertEquals(List(Loader.command("""m["foo"]@1 := r["x"]@2""")), command.expand(Nil))
+
+  /** Expanding covers nested let bindings. */
+  @Test
+  def expandDoubleLetCommand(): Unit =
+    val command = Loader.command("""let table = "foo" in let i = 2 in m[table]@1 := r["x"]@i""")
+    assertEquals(List(Loader.command("""m["foo"]@1 := r["x"]@2""")), command.expand(Nil))
+
+  /** Expressions simplify when an assert command is expanded. */
+  @Test
+  def expandAssertCommand(): Unit =
+    val cmd = Loader.command("""assert(m["a" ++ "b"] = m["c" ++ "d"])@i1""")
+    assertEquals(
+      List(Loader.command("""assert(m["ab"] = m["cd"])@i1""")),
+      cmd.expand(Nil))
+}

### 2025-11-07 code/cvc5mvn/src/test/scala/plaid/prelude/ast/ConstraintTest.scala

@@ -8,22 +8,34 @@ class ConstraintTest {
 
   /** Expanding simplifies embedded expressions. */
   @Test
-  def expandPreAndPost(): Unit =
-    val src = """ "ab" == "a" ++ "b" """
+  def expandEmbeddedExpr(): Unit =
+    val src = """ "a" ++ "b" == "a" ++ "b" """
     val constraint = Loader.constraint(src).expand()
     assertEquals(Loader.constraint(""" "ab" == "ab" """), constraint)
 
   /** Expanding substitutes constraint function calls for the appropriate body with bindings. */
   @Test
   def expandCalls(): Unit =
-    val fn = Loader.program("constraintfunctions: f(x) { x == x }").constraintFuncs.head
+    val fn = Loader.constraintFunc("f(x) { x == x }")
     val constraint = Loader.constraint("NOT f(5)")
-    assertEquals(Loader.constraint(""" NOT 5 == 5 """), constraint.expand(constraintCtx = List(fn)))
+    assertEquals(
+      Loader.constraint("NOT 5 == 5"),
+      constraint.expand(constraintCtx = List(fn)))
 
   /** Constraint function call parameters get expanded when the call gets expanded. */
   @Test
   def callBindingsExpanded(): Unit =
-    val fn = Loader.program("constraintfunctions: f(x) { x == x }").constraintFuncs.head
+    val fn = Loader.constraintFunc("f(x) { x == x }")
     val constraint = Loader.constraint("""NOT f("a" ++ "b")""")
-    assertEquals(Loader.constraint(""" NOT "ab" == "ab" """), constraint.expand(constraintCtx = List(fn)))
+    assertEquals(
+      Loader.constraint(""" NOT "ab" == "ab" """),
+      constraint.expand(constraintCtx = List(fn)))
+
+  /** Constraint function calls to expression functions get expanded. */
+  @Test
+  def constraintsExpandExprFuncs(): Unit =
+    val f = Loader.exprFunc("f(i) {out@i}")
+    val constraint = Loader.constraint("f(1) == 2")
+    val expected = Loader.constraint("out@1 == 2")
+    assertEquals(Loader.constraint("out@1 == 2"), constraint.expand(List(f)))
 }

### 2025-11-07 code/cvc5mvn/src/test/scala/plaid/prelude/ast/ExprTest.scala

@@ -6,6 +6,16 @@ import plaid.prelude.antlr.Loader
 
 class ExprTest {
 
+  /** Identifiers concatenated with literals do not simplify. */
+  @Test
+  def identifierConcat(): Unit =
+    assertEquals(
+      ConcatExpr(Identifier("i"), Str("x")),
+      Loader.expression("i++\"x\""))
+    assertEquals(
+      ConcatExpr(Str("x"), Identifier("i")),
+      Loader.expression("\"x\"++i"))
+
   /** Expanding simplifies the concatenation of two string literals. */
   @Test
   def simplifyLiteralConcat(): Unit =
@@ -27,6 +37,12 @@ class ExprTest {
     val expr = Loader.expression(src).expand()
     assertEquals(Loader.expression(""" i ++ "ab" ++ j """), expr)
 
+  @Test
+  def simplifyGroupedConcat(): Unit =
+    assertEquals(
+      Str("xywz"),
+      Loader.expression("(\"x\"++\"y\")++(\"w\"++\"z\")").expand())
+
   /** Expanding can bind and simplify concatenations at the same time. */
   @Test
   def substituteConcat(): Unit =
@@ -34,4 +50,51 @@ class ExprTest {
     val map = Map(Identifier("i") -> Str("x"), Identifier("j") -> Str("y"))
     val expr = Loader.expression(src).expand(bindings = map)
     assertEquals(Str("xaby"), expr)
+
+  /** If a let binding is unused, the expansion is the let body. */
+  @Test
+  def unusedLetSubstitution(): Unit =
+    assertEquals(Num(3), Loader.expression("let r11 = 3 in r11").expand())
+
+  /** The inner let body in nested let bindings can access both variables. */
+  @Test
+  def doubleLetSubstitution(): Unit =
+    val letExpr = "let r11 = 3 in let r10 = 4 in r11 + r10"
+    assertEquals(
+      PlusExpr(Num(3), Num(4)),
+      Loader.expression(letExpr).expand())
+
+  /** The value of an outer let binding propagates to the body via an inner let binding. */
+  @Test
+  def letPropagation(): Unit =
+    val letExpr = "let y = 3 in let x = y + 1 in x + y"
+    assertEquals(
+      PlusExpr(PlusExpr(Num(3), Num(1)), Num(3)),
+      Loader.expression(letExpr).expand())
+
+  /** An outer let binding variables is shadowed by an inner let binding. */
+  @Test
+  def letShadowing(): Unit =
[…]

### 2025-11-07 code/cvc5mvn/src/test/scala/plaid/prelude/transform/EvaluatorTest.scala

@@ -1,207 +1,0 @@
-package plaid.prelude.transform
-
-import org.junit.Assert.assertEquals
-import org.junit.Test
-import plaid.prelude.antlr.Loader
-import plaid.prelude.ast.*
-
-class EvaluatorTest {
-/*
-  private def evalExpr(src: String, exprFunctions: List[ExprFunc]): Expr = {
-    val ast = Loader.expression(src)
-    val evaluator = Evaluator(Program(List(), exprFunctions, List()))
-    evaluator.expression(ast)
-  }
-
-  private def evalCommand(src: String, commandFunctions: List[CmdFunc]): Cmd = {
-    val ast = Loader.command(src)
-    val evaluator = Evaluator(Program(commandFunctions, List(), List()))
-    evaluator.command(ast)
-  }
-
-  @Test
-  def evalString(): Unit = {
-    assertEquals(Str("foo"), evalExpr("\"foo\"", List()))
-  }
-
-  @Test
-  def literalConcat(): Unit = {
-    assertEquals(Str("xy"), evalExpr("\"x\"++\"y\"", List()))
-  }
-
-  @Test
-  def identifierConcat(): Unit = {
-    assertEquals(
-      ConcatExpr(Identifier("i"), Str("x")),
-      evalExpr("i++\"x\"", List())
-    )
-    assertEquals(
-      ConcatExpr(Str("x"), Identifier("i")),
-      evalExpr("\"x\"++i", List())
-    )
-  }
-
-  @Test
-  def evalGroupedConcat(): Unit = {
-    assertEquals(Str("xywz"), evalExpr("(\"x\"++\"y\")++(\"w\"++\"z\")", List()))
-  }
-
-  @Test
-  def letSubstitution(): Unit = {
-    assertEquals(Num(3), evalExpr("let r11 = 3 in r11", List()))
-  }
-
-  @Test
-  def doubleLetSubstitution(): Unit = {
-    val letExpr = "let r11 = 3 in let r10 = 4 in r11 + r10"
-    assertEquals(PlusExpr(Num(3), Num(4)), evalExpr(letExpr, List()))
-  }
-
-  @Test
-  def evalLetExpr(): Unit = {
-    val letExpr = "let y = 3 in let x = y + 1 in x + y"
-    assertEquals(PlusExpr(PlusExpr(Num(3), Num(1)), Num(3)), evalExpr(letExpr, List()))
-  }
-
-  @Test
-  def evalShadowing(): Unit = {
[…]

### 2025-11-07 toplas26/sty.tex
checkpoint

@@
 
+We write $\stredx^*$ to denote the reflexive, transitive
+closure of $\stredx$. For any protocol $\prog$, the goal of the
+type analysis is to show that all interactive messages can be classified--
+that is, to show that $\eqspre \wedge \toeq{\prog} \vdash
+\varnothing,\varnothing \stredx^* \Gamma,R$ where all
+interactive messages are classified in $\Gamma$.
+
 \subsection{Share Typing Example} To illustrate details of the
@@
   \px{1} &:=& \elab{\mesg{s2} + \locflip + 1}{1}; \\
-  \px{2} &:=& \elab{\mesg{s1} + \locflip}{2}\\
-  \out{1} &:=& \elab{\px{1} + \px{2}}{1}\\
+  \px{2} &:=& \elab{\mesg{s1} + \locflip}{2};\\
+  \out{1} &:=& \elab{\px{1} + \px{2}}{1};\\
   \out{2} &:=& \elab{\px{1} + \px{2}}{2}
@@
 
+Now we can prove our share typing soundness result, showing that share
+typing enforces noninteference modulo output, with our main result
+in Theorem \ref{theorem-sty}. We begin by defining a key well-formedness
+property of typing components that we show is preserved by non-interactive
+typing judgements.
 \begin{definition}
@@
 \end{definition}
-
-\begin{lemma}
-  Assume given $\Gamma,R, \eqspre \wedge \toeq{\prog} : {OK}_\cid$  and
-  $\iov(\prog) = S \cup V \cup O$
-  and $\Gamma, \eqspre \wedge \toeq{\prog} \vdash [\phi_1,\phi_2] : \sharety$.
-  Then, letting $X \defeq \ov(\Gamma) \cap \realviews{\setit{\cid_1}}{\setit{\cid_2}}$:
-  $$
-  \condd{\progtt(\prog)}{S_{\setit{\cid_1}}}{S_{\setit{\cid_2}}  \cup X \cup \setit{\phi_1 + \phi_2}}
-  = 
-  \condd{\progtt(\prog)}{S_{\setit{\cid_1}}}{S_{\setit{\cid_2}} \cup X \cup \setit{\phi_1 + \phi_2, \phi_1}}
-  $$
-\end{lemma}
 

### 2025-11-07 toplas26/sty.tex
checkpoint

@@
 
+Next, we make a simple definition to formalize what we mean by the interactive
+variables in the share type environment $\Gamma$.
 \begin{definition} Define $\ov(\Gamma)$ as follows:
@@
 \end{definition}
-
+Given the above, we can formulate our key Lemma about noninteractive typing
+judgements, which tells us that individual shares in isolation don't reveal anything
+about their reconstruction. This result follows by Theorem \ref{theorem-toeq}
+and known homomorphic properties of secret sharing schemes. 
 \begin{lemma}

### 2025-11-07 toplas26/sty.tex
checkpoint

@@
 
+Based on the preceding Lemma, we can show that noninterence with respect
+to classified interactive variables is maintained by typing judgements. 
 \begin{lemma}
+  \label{lemma-styinteractive}
   Assume given $\eqspre \wedge \toeq{\prog}, \Gamma_1,R_1: {OK}_\cid$  and
@@
 \end{lemma}
-
+Now we can prove our main Theorem, which follows immediately 
+by Lemma \ref{lemma-styinteractive} and induction on $|\Gamma|$
+(i.e., the number of bindings in the environment $\Gamma$ obtained
+by a complete typing). 
 \begin{theorem}[Share Type Soundness]

### 2025-11-08 toplas26/metalang.tex
checkpoint

@@
 \begin{lemma}
+  \label{lemma-fresh}
+  Given $\eqs$ and $w, \cid = \fresh(\mathit{string}, \mathit{cid})$ and
+  some $w',\cid'$ where 
+  $\peq[w/y_1,\cid/y_n] \redx \eqs$ and
+  $\peq[w'/y_1,\cid'/y_n] \redx \eqs'$.
+  Then $\store \models \eqs'$ implies $\store \models \eqs$. 
+\end{lemma}
+
+\begin{proof}
+  If $w',\cid'$ are fresh or chosen the result is
+  trivial.  Otherwise, it must be the case that at least two
+  distinct variables $x_1$ and $x_2$ in $\eqs$ are renamed as a single
+  variable $x$ in $\eqs'$, given freshness of $w,\cid$. Let
+  $\mathit{rename}$ be the function that maps variables in $\eqs$ to
+  variables in $\eqs'$ such that $\mathit{rename}(\eqs) = \eqs'$. Assume
+  $\store' \models \eqs'$, and define $\store$ such that:
+  $$
+  \forall x \in \dom(\vars(E)) . \store(x) \defeq \store'(\mathit{rename}(x)) 
+  $$
+  Then $\store$ must be a model of $\eqs$, which is a contradiction. 
+\end{proof}
+
+\begin{lemma}
   \label{lemma-genentails}
@@
 \end{lemma}
+\begin{proof}
+  By assumption and definition of \TirName{GenEntails} we have
+  $\peq_1[\mv_1'/y_n \cdots \mv_n'/y_n] \redx \eqs_1'$ and
+  $\peq_2[\mv_1'/y_n \cdots \mv_n'/y_n] \redx \eqs_2'$ and $\eqs_1'
+  \models \eqs_2'$ for some $\mv_1',\ldots,\mv_n' =
+  \fresh(\tau_1,\ldots,\tau_n)$.  Now suppose on the contrary that
+  $\eqs_1 \not\models \eqs_2$.  This implies the existence of $\store$
+  such that $\store \models \eqs_1$ and $\store \not\models \eqs_2$.
+  But Lemma \ref{lemma-fresh} implies that $\store \models \eqs_1'$,
+  so by definition we have $\store \models \eqs_2'$ implying $\store
+  \models \eqs_2$ by construction, which is a contradiction.
+\end{proof}
 

### 2025-11-08 toplas26/metaty.tex
checkpoint

\section{$\metaprot$ Type Checking}

\subsection{Share Type Checking with Annotations}

\begin{mathpar}
  \htrip{e_1 \eop e_1 \wedge e_2 \eop e_2'}{[e_1,e_2] \kwas [e_1',e_2']}{\true}

  \htrip{e_1 \eop e_1 \wedge e_2 \eop e_2'}{[e_1,e_2] \kwmask [e_1',e_2']}{\true}

  \htrip{\true}{[e_1,e_2] : \sharety}{\true}

  \htrip{\true}{[e_1,e_2] : \outty}{\true}

  \htrip{\true}{\pubx{e} : \outty}{\true}
\end{mathpar}

\begin{mathpar}
  \inferrule
      {e_1 \redx \mx{w_1}{\cid_1} \\ e_2 \redx \mx{w_2}{\cid_2} \\
        e_1' \redx \phi - \rx{w_3}{\cid_2} \\
        e_2' \redx \rx{w_3}{\cid_2}}
      {[e_1,e_2] \kwas [e_1',e_2'] \redx \pskip}

  \inferrule
      {e_1 \redx \px{x} \\ e_2 \redx \px{y} \\
        e_1' \redx \phi - \rx{w_3}{\oid} \\
        e_2' \redx \rx{w_3}{\oid}}
      {[e_1,e_2] \kwmask [e_1',e_2'] \redx \pskip}

  \inferrule
      {e_1 \redx \mx{w_1}{\cid_1} \\ e_2 \redx \mx{w_2}{\cid_2}}
      {[e_1,e_2] : \sharety \redx \pskip}

  \inferrule
      {e_1 \redx \px{w_1} \\ e_2 \redx \px{w_2}}
      {[e_1,e_2] : \outty \redx \pskip}

  \inferrule
      {e \redx \pubx{\out{\cid}}}
      {\pubx{e}: \outty \redx \pskip}
\end{mathpar}

\begin{mathpar}
  \inferrule[CoerceGenshare]
      {e_1 \redx \mx{w_1}{\cid_1} \\ e_2 \redx \mx{w_2}{\cid_2} \\
        e_1' \redx \phi - \rx{w_3}{\cid_2} \\
        e_2' \redx \rx{w_3}{\cid_2}}
      {\shtrip{\Gamma, R, \xdefs}
        {[e_1,e_2] \kwas [e_1',e_2']}
        {\Gamma \cup \setit{[\mx{w_1}{\cid_1},\mx{w_2}{\cid_2}] : \sharety},
          R \uplus \setit{\rx{w}{\cid}}, \xdefs} }

  \inferrule[CoerceSecureOpen]
      {e_1 \redx \px{x} \\ e_2 \redx \px{y} \\
        e_1' \redx \phi - \rx{w_3}{\oid} \\
        e_2' \redx \rx{w_3}{\oid}}
      {\shtrip{\Gamma, R, \xdefs}
        {[e_1,e_2] \kwmask [e_1',e_2']}
        {\Gamma \cup  \setit{\pubx{\px{x} + \px{y}} : \pubty},
        R \uplus \setit{\rx{w}{\oid}}, \xdefs} }

  \inferrule[LocalShares]
      {e_1 \redx \mx{w_1}{\cid_1} \\ e_2 \redx \mx{w_2}{\cid_2}  \\
        \Gamma \Vdash [\xdefs(\mx{w_1}{\cid_1} ), \xdefs(\mx{w_2}{\cid_2})] : \sharety}
      {\shtrip{\Gamma, R, \xdefs}{[e_1,e_2] : \sharety}
        {\Gamma \cup \setit{[\mx{w_1}{\cid_1},\mx{w_2}{\cid_2}] : \sharety}, R, \xdefs}}

  \inferrule[Reveal]
      {e_1 \redx \px{w_1} \\ e_2 \redx \px{w_2}  \\
        \Gamma \Vdash [\xdefs(\px{w_1}), \xdefs(\px{w_2})] : \sharety}
      {\shtrip{\Gamma, R, \xdefs}{[e_1,e_2] : \outty}
        {\Gamma \cup \setit{\pubx{\px{w_1} \fplus \px{w_2}} : \outty}, R, \xdefs}}

  \inferrule[Output]
      {e \redx \out{\cid} \\
        \Gamma \Vdash \pubx{\xdefs(\out{x})} : \outty}
      {\shtrip{\Gamma, R, \xdefs}{\pubx{\out{\cid}} : \outty}
        {\Gamma \cup \setit{\pubx{\xdefs(\out{\cid})} : \outty}, R, \xdefs}}
\end{mathpar}

\begin{mathpar}
  \inferrule[Defn]
      {e_1 \redx \mx{w}{\cid_1} \\ e_2 \redx \elab{\be}{\cid_2}}
      {\shtrip{\Gamma, R, \xdefs}{e_1 := e_2}{\Gamma, R, \xdefs[\mx{w}{\cid_1} \mapsto \toeq{\elab{\be}{\cid_2}}]}}

  \inferrule[App]
      {f : \Pi y_1,\ldots,y_n. \notg{\Gamma_1} \rightarrow (\notg{\Gamma_2}, \notg{R'}) \\
        \notg{\Gamma_1}[e_1/y_1 \cdots e_n/y_n] \redx \Gamma_1 \\
        \notg{\Gamma_2}[e_1/y_1 \cdots e_n/y_n] \redx \Gamma_2 \\
        \notg{R'}[e_1/y_1 \cdots e_n/y_n] \redx R'        
      }
      {\shtrip{\Gamma \cup \Gamma_1, R, \xdefs}{f(e_1,\ldots,e_n)}{\Gamma \cup \Gamma_1 \cup \Gamma_2, R \uplus R', \xdefs}}

  \inferrule[Seq]    
      {\shtrip{\Gamma_1,R_1,\xdefs_1}{\cmd_1}{\Gamma_2,R_2,\xdefs_2} \\
       \shtrip{\Gamma_2,R_2,\xdefs_2}{\cmd_2}{\Gamma_3,R_3,\xdefs_3}}
      {\shtrip{\Gamma_1,R_1,\xdefs_1}{\cmd_1;\cmd_2}{\Gamma_3,R_3,\xdefs_3}}

  \inferrule[Let]
      {\shtrip{\Gamma_1,R_1,\xdefs_1}{\cmd[e/y]}{\Gamma_2,R_2,\xdefs_2}}
      {\shtrip{\Gamma_1,R_1,\xdefs_1}{\elet{y}{e}{\cmd}}{\Gamma_2,R_2,\xdefs_2}}

  \inferrule[Fn]
            {\codebase(f) = y_1 : \tau_1, \ldots, y_n : \tau_n, \cmd \\
              \mv_1,\ldots,\mv_n = \fresh(\tau_1,\ldots,\tau_n) \\
              \notg{\Gamma_1}[e_1/y_1 \cdots e_n/y_n] \redx \Gamma_1 \\
              \notg{\Gamma_2}[e_1/y_1 \cdots e_n/y_n] \redx \Gamma_2 \\
              \notg{R}[e_1/y_1 \cdots e_n/y_n] \redx R \\  
     \shtrip{\Gamma_1, \varnothing, \varnothing}{\cmd}{\Gamma_2, R,\xdefs}}
    {f : \Pi y_1,\ldots,y_n. \notg{\Gamma_1} \rightarrow (\notg{\Gamma_2}, \notg{R})}
\end{mathpar}

### 2025-11-08 code/cvc5mvn/src/main/scala/plaid/prelude/logic/Unique.scala
test type logic

@@ -1,6 +1,6 @@
 package plaid.prelude.logic
 
-import plaid.prelude.ast.{Identifier, Num, Str}
+import plaid.prelude.ast.{Num, Str}
 
 import java.util.concurrent.atomic.AtomicInteger
 
@@ -8,8 +8,13 @@ import java.util.concurrent.atomic.AtomicInteger
  * Support for generating unique instances of AST objects.
  */
 object Unique {
-  private val partyCounter = new AtomicInteger(0)
-  private val stringCounter = new AtomicInteger(0)
+  private val start = 1
+  private val partyCounter = new AtomicInteger(start)
+  private val stringCounter = new AtomicInteger(start)
+
+  def reset(): Unit =
+    partyCounter.set(start)
+    stringCounter.set(start)
 
   /**
    * Generate a unique party index using a negative number to avoid collisions.

### 2025-11-08 code/cvc5mvn/src/test/scala/plaid/prelude/ast/TypeTest.scala

@@ -1,0 +1,41 @@
+package plaid.prelude.ast
+
+import org.junit.Assert.assertEquals
+import org.junit.Test
+import plaid.prelude.antlr.Loader
+import plaid.prelude.logic.Unique
+
+class TypeTest {
+
+  @Test
+  def freshString(): Unit =
+    Unique.reset()
+    assertEquals(Str("$1"), StringType().freshValue())
+    assertEquals(Str("$2"), StringType().freshValue())
+    assertEquals(Str("$3"), StringType().freshValue())
+
+  @Test
+  def freshParty(): Unit =
+    Unique.reset()
+    assertEquals(Num(-1), PartyIndexType().freshValue())
+    assertEquals(Num(-2), PartyIndexType().freshValue())
+    assertEquals(Num(-3), PartyIndexType().freshValue())
+
+  @Test
+  def freshRecord(): Unit =
+    Unique.reset()
+    val recordType = Loader.typeMarker("{ s : string ; i : cid}")
+    val expected = FieldExpr(Map(
+      Identifier("s") -> Str("$1"),
+      Identifier("i") -> Num(-1)))
+    assertEquals(expected, recordType.freshValue())
+
+  @Test
+  def freshNestedRecord(): Unit =
+    Unique.reset()
+    val recordType = Loader.typeMarker("{t: {s:string; t2:{i:cid}}}")
+    val expected = FieldExpr(Map(Identifier("t") ->
+      FieldExpr(Map(Identifier("s") -> Str("$1"), Identifier("t2") ->
+        FieldExpr(Map(Identifier("i") -> Num(-1)))))))
+    assertEquals(expected, recordType.freshValue())
+}

### 2025-11-08 code/cvc5mvn/src/test/scala/plaid/prelude/logic/ConstraintAnalyzerTest.scala

@@ -219,7 +219,34 @@ class ConstraintAnalyzerTest {
     val expected = Str("abcd")
     assertEquals(expected, actual)
   }
-  
+
+    @Test
+  def genEntailTest(): Unit = {
+    val e1 =
+      """m[z++"s"]@i1 == m[x++"s"]@i1 + m[y++"s"]@i1 AND
+        |m[z++"s"]@i2 == m[x++"s"]@i2 + m[y++"s"]@i2""".stripMargin
+    val e2 =
+      """m[z++"s"]@i1 + m[z++"s"]@i2 ==
+        |m[x++"s"]@i1 + m[y++"s"]@i1 + m[x++"s"]@i2 + m[y++"s"]@i2""".stripMargin
+
+    val program = Program(List(), List(), List())
+    val genEntailVerifier = new Verifier(program, "2")
+
+    val typing = List(
+      TypedIdentifier(Identifier("i2"), PartyIndexType()),
+      TypedIdentifier(Identifier("y"), StringType()),
+      TypedIdentifier(Identifier("x"), StringType()),
+      TypedIdentifier(Identifier("i1"), PartyIndexType()),
+      TypedIdentifier(Identifier("z"), StringType())
+    )
+
+    val result1 = genEntailVerifier.genEntails(typing, Loader.constraint(e1), Loader.constraint(e2))
+    assertTrue(result1)
+
+    val result2 = genEntailVerifier.genEntails(typing, Loader.constraint(e2), Loader.constraint(e1))
+    assertFalse(result2)
+  }
+
  */
 
 }

### 2025-11-08 code/cvc5mvn/src/test/scala/plaid/prelude/logic/verifierTest.scala

@@ -1,99 +1,0 @@
-package plaid.prelude.logic
-
-import io.github.cvc5.CVC5ApiException
-import org.junit.Assert.*
-import org.junit.Test
-import plaid.prelude.antlr.Loader
-import plaid.prelude.ast.*
-
-import java.util
-import java.util.TreeMap
-
-class verifierTest {
-/*
-  @Test
-  @throws[ CVC5ApiException ]
-  def genFreshStringTest(): Unit = {
-    val program = Program(List(), List(), List())
-    val genEntailVerifier = new Verifier(program, "2")
-
-    assertEquals(Str("$1"), genEntailVerifier.genFreshValue(StringType()))
-    assertEquals(Str("$2"), genEntailVerifier.genFreshValue(StringType()))
-    assertEquals(Str("$3"), genEntailVerifier.genFreshValue(StringType()))
-  }
-
-  @Test
-  @throws[ CVC5ApiException ]
-  def genFreshCIDTest(): Unit = {
-    val program = Program(List(), List(), List())
-    val genEntailVerifier = new Verifier(program, "2")
-
-    assertEquals(Num(-1), genEntailVerifier.genFreshValue(PartyIndexType()))
-    assertEquals(Num(-2), genEntailVerifier.genFreshValue(PartyIndexType()))
-    assertEquals(Num(-3), genEntailVerifier.genFreshValue(PartyIndexType()))
-  }
-
-  @Test
-  @throws[ CVC5ApiException ]
-  def genFreshRecordTest(): Unit = {
-    val src = "{ s : string ; i : cid}"
-    val program = Program(List(), List(), List())
-    val genEntailVerifier = new Verifier(program, "2")
-
-    val recordType = Loader.typeMarker(src).asInstanceOf[RecordType]
-    val map = new util.TreeMap[Identifier, Expr]()
-    map.put(Identifier("s"), Str("$2"))
-    map.put(Identifier("i"), Num(-1))
-    val expected = FieldExpr(map)
-
-    assertEquals(expected, genEntailVerifier.genFreshValue(recordType))
-  }
-
-  @Test
-  @throws[ CVC5ApiException ]
-  def nestedRecordType(): Unit = {
-    val src = "{t: {s:string; t2:{i:cid}}}"
-    val program = Program(List(), List(), List())
-    val genEntailVerifier = new Verifier(program, "2")
-
[…]

### 2025-11-08 toplas26/sty.tex
checkpoint

@@
   \begin{enumerate}[\hspace{5mm}i.]
-  \item For all $([\mx{z}{\cid}, \rx{y}{\cid'}] : \sharety) \in \Gamma$,
-    we have $\eqs \models \mx{z}{\cid} \eop \phi - \rx{x}{\cid'}$ for
+  \item For all $([\phi_1,\phi_2] : \sharety) \in \Gamma$,
+    we have $\eqs \models \phi_1 \eop \phi - \rx{x}{\cid'} \wedge
+    \phi_2 \eop \rx{y}{\cid'}$ for
     some $\phi$ and unique $\rx{x}{\cid'} \in R$, with $\cid \ne \cid'$.
-  \item For all $([\phi] : \pubty) \in \Gamma$,
+  \item For all $(\pubx{\phi} : \pubty) \in \Gamma$,
     we have $\eqs \models \phi \eop \phi' - \rx{x}{\oid}$ for
@@
   \label{lemma-sty-noninteractive-sound}
-  Assume given $\eqspre \wedge \toeq{\prog}, \Gamma,R: {OK}_\cid$  and
+  Assume given $\eqspre \wedge \toeq{\prog}, \Gamma,R: {OK}$  and
   $\iov(\prog) = S \cup V \cup O$
@@
   \label{lemma-styinteractive}
-  Assume given $\eqspre \wedge \toeq{\prog}, \Gamma_1,R_1: {OK}_\cid$  and
+  Assume given $\eqspre \wedge \toeq{\prog}, \Gamma_1,R_1: {OK}$  and
   $\iov(\prog) = S \cup V \cup O$

### 2025-11-11 toplas26/metaty.tex
checkpoint

@@
-\section{$\metaprot$ Type Checking}
-
 \subsection{Share Type Checking with Annotations}
 
+Share types rely on SMT entailments via the \TirName{ShareEntails} and \TirName{PubEntails} rules,
+which are not syntax directed. Our goal with the $\metaprot$ language is to automate share type
+checking. To enable this we introduce \emph{share type annotations}
+
+\stafig
+
+\begin{definition}[Abstract Share Type Environments]
+  We write $\notg{\Gamma}$ to denote sets of bindings $\notg{\stt} : \sty$, and
+  $\notg{R}$ to denote sets of expressions (that evaluate to random variables).
+  We define evaluation of these, respectively, as follows:
   \begin{mathpar}
-  \htrip{e_1 \eop e_1 \wedge e_2 \eop e_2'}{[e_1,e_2] \kwas [e_1',e_2']}{\true}
+  \inferrule
+      {\notg{\stt}_1 \redx \stt_1 \cdots \notg{\stt}_n \redx \stt_n}
+      {\setit{\notg{\stt}_1 : \sty_1, \ldots, \notg{\stt}_n : \sty_n} \redx
+        \setit{\stt_1 : \sty_1, \ldots, \stt_n : \sty_n}}
 
-  \htrip{e_1 \eop e_1 \wedge e_2 \eop e_2'}{[e_1,e_2] \kwmask [e_1',e_2']}{\true}
-
-  \htrip{\true}{[e_1,e_2] : \sharety}{\true}
-
-  \htrip{\true}{[e_1,e_2] : \outty}{\true}
-
-  \htrip{\true}{\pubx{e} : \outty}{\true}
+  \inferrule
+      {e_1 \redx \rx{w_1}{\cid_1} \cdots e_1 \redx \rx{w_n}{\cid_n}}
+      {\setit{e_1,\ldots,e_n} \redx
+        \setit{\rx{w_1}{\cid_1}, \ldots, \rx{w_n}{\cid_n}}}
   \end{mathpar}
+\end{definition}
 
 \begin{mathpar}
-  \inferrule
-      {e_1 \redx \mx{w_1}{\cid_1} \\ e_2 \redx \mx{w_2}{\cid_2} \\
-        e_1' \redx \phi - \rx{w_3}{\cid_2} \\
-        e_2' \redx \rx{w_3}{\cid_2}}
-      {[e_1,e_2] \kwas [e_1',e_2'] \redx \pskip}
-
-  \inferrule
-      {e_1 \redx \px{x} \\ e_2 \redx \px{y} \\
-        e_1' \redx \phi - \rx{w_3}{\oid} \\
-        e_2' \redx \rx{w_3}{\oid}}
-      {[e_1,e_2] \kwmask [e_1',e_2'] \redx \pskip}
-
-  \inferrule
-      {e_1 \redx \mx{w_1}{\cid_1} \\ e_2 \redx \mx{w_2}{\cid_2}}
-      {[e_1,e_2] : \sharety \redx \pskip}
-
-  \inferrule
-      {e_1 \redx \px{w_1} \\ e_2 \redx \px{w_2}}
-      {[e_1,e_2] : \outty \redx \pskip}
-
-  \inferrule
-      {e \redx \pubx{\out{\cid}}}
-      {\pubx{e}: \outty \redx \pskip}
-\end{mathpar}
-
-\begin{mathpar}
-  \inferrule[CoerceGenshare]
-      {e_1 \redx \mx{w_1}{\cid_1} \\ e_2 \redx \mx{w_2}{\cid_2} \\
-        e_1' \redx \phi - \rx{w_3}{\cid_2} \\
-        e_2' \redx \rx{w_3}{\cid_2}}
-      {\shtrip{\Gamma, R, \xdefs}
-        {[e_1,e_2] \kwas [e_1',e_2']}
-        {\Gamma \cup \setit{[\mx{w_1}{\cid_1},\mx{w_2}{\cid_2}] : \sharety},
-          R \uplus \setit{\rx{w}{\cid}}, \xdefs} }
-
-  \inferrule[CoerceSecureOpen]
-      {e_1 \redx \px{x} \\ e_2 \redx \px{y} \\
-        e_1' \redx \phi - \rx{w_3}{\oid} \\
-        e_2' \redx \rx{w_3}{\oid}}
-      {\shtrip{\Gamma, R, \xdefs}
-        {[e_1,e_2] \kwmask [e_1',e_2']}
-        {\Gamma \cup  \setit{\pubx{\px{x} + \px{y}} : \pubty},
-        R \uplus \setit{\rx{w}{\oid}}, \xdefs} }
+  \inferrule[Coerce]
+      {e_1 \redx \phi_1 \\ e_2 \redx \phi_2}
+      {\shtrip{\Gamma, R, \xdefs}{e_1 \kwas e_2}{\Gamma, R, \xdefs[\phi_1 \mapsto \phi_2]}}
 
   \inferrule[LocalShares]
-      {e_1 \redx \mx{w_1}{\cid_1} \\ e_2 \redx \mx{w_2}{\cid_2}  \\
-        \Gamma \Vdash [\xdefs(\mx{w_1}{\cid_1} ), \xdefs(\mx{w_2}{\cid_2})] : \sharety}
+      {e_1 \redx x_1 \\ e_2 \redx x_2  \\
+        \Gamma \Vdash [\xdefs(x_1), \xdefs(x_2)] : \sharety}
       {\shtrip{\Gamma, R, \xdefs}{[e_1,e_2] : \sharety}
-        {\Gamma \cup \setit{[\mx{w_1}{\cid_1},\mx{w_2}{\cid_2}] : \sharety}, R, \xdefs}}
+        {\Gamma \cup \setit{[x_1,x_2] : \sharety}, R, \xdefs}}
+
+  \inferrule[GenShares]
+      {e_1 \redx x_1 \\ e_2 \redx x_2  \\ \xdefs(x_1) = \phi - \rx{w}{\cid} \\ \xdefs(x_2) =  \rx{w}{\cid}}
+      {\shtrip{\Gamma, R, \xdefs}{[e_1,e_2] : \sharety}
+        {\Gamma \cup \setit{[x_1,x_2] : \sharety}, R \uplus \setit{\rx{w}{\cid}}, \xdefs}}
+
+  \inferrule[LocalPublic]
+      {e \redx \phi \\ \Gamma \Vdash \pubx{\xdefs(\phi)} : \pubty}
+      {\shtrip{\Gamma, R, \xdefs}{\pubx{e} : \pubty}
+        {\Gamma \cup \setit{\pubx{\phi} : \pubty}, R, \xdefs}}
+
+  \inferrule[MaskedPublic]
+      {e_1 \redx x_1 \\ e_2 \redx x_2  \\
+        \Gamma \Vdash [x_1, x_2] : \sharety \\
+        \xdefs(x_1 + x_2)  = \phi + \rx{w}{\oid}}
+      {\shtrip{\Gamma, R, \xdefs}{\pubx{e_1+e_2} : \pubty}
+        {\Gamma \cup \setit{\pubx{x_1 + x_2} : \pubty}, R \uplus \setit{\rx{w}{\oid}}, \xdefs}}
 
   \inferrule[Reveal]
-      {e_1 \redx \px{w_1} \\ e_2 \redx \px{w_2}  \\
-        \Gamma \Vdash [\xdefs(\px{w_1}), \xdefs(\px{w_2})] : \sharety}
-      {\shtrip{\Gamma, R, \xdefs}{[e_1,e_2] : \outty}
-        {\Gamma \cup \setit{\pubx{\px{w_1} \fplus \px{w_2}} : \outty}, R, \xdefs}}
+      {e_1 \redx x_1 \\ e_2 \redx x_2  \\
+        \Gamma \Vdash [x_1, x_2] : \sharety}
+      {\shtrip{\Gamma, R, \xdefs}{\pubx{e_1 + e_2} : \outty}
+        {\Gamma \cup \setit{\pubx{x_1 \fplus x_2} : \outty}, R, \xdefs}}
 
@@
       {e \redx \out{\cid} \\
-        \Gamma \Vdash \pubx{\xdefs(\out{x})} : \outty}
-      {\shtrip{\Gamma, R, \xdefs}{\pubx{\out{\cid}} : \outty}
-        {\Gamma \cup \setit{\pubx{\xdefs(\out{\cid})} : \outty}, R, \xdefs}}
+       \xdefs(\out{\cid}) = x_1 + x_2 \\
+       \Gamma \Vdash \pubx{x_1 + x_2} : \outty }
+      {\shtrip{\Gamma, R, \xdefs}{\out{\cid} : \outty}
+        {\Gamma \cup \setit{\pubx{\out{\cid}} : \outty}, R, \xdefs}}
 \end{mathpar}
@@
   \inferrule[Defn]
-      {e_1 \redx \mx{w}{\cid_1} \\ e_2 \redx \elab{\be}{\cid_2}}
-      {\shtrip{\Gamma, R, \xdefs}{e_1 := e_2}{\Gamma, R, \xdefs[\mx{w}{\cid_1} \mapsto \toeq{\elab{\be}{\cid_2}}]}}
+      {e_1 \redx x \\ e_2 \redx \elab{\be}{\cid}}
+      {\shtrip{\Gamma, R, \xdefs}{e_1 := e_2}{\Gamma, R, \xdefs[x \mapsto \toeq{\elab{\be}{\cid}}]}}
 
@@
               \mv_1,\ldots,\mv_n = \fresh(\tau_1,\ldots,\tau_n) \\
-              \notg{\Gamma_1}[e_1/y_1 \cdots e_n/y_n] \redx \Gamma_1 \\
-              \notg{\Gamma_2}[e_1/y_1 \cdots e_n/y_n] \redx \Gamma_2 \\
-              \notg{R}[e_1/y_1 \cdots e_n/y_n] \redx R \\  
-     \shtrip{\Gamma_1, \varnothing, \varnothing}{\cmd}{\Gamma_2, R,\xdefs}}
+              \notg{\Gamma_1}[\mv_1/y_1 \cdots \mv_n/y_n] \redx \Gamma_1 \\
+              \notg{\Gamma_2}[\mv_1/y_1 \cdots \mv_n/y_n] \redx \Gamma_2 \\
+              \notg{R}[\mv_1/y_1 \cdots \mv_n/y_n] \redx R \\  
+     \shtrip{\Gamma_1, \varnothing, \varnothing}{\cmd[\mv_1/y_1 \cdots \mv_n/y_n]}{\Gamma_2, R,\xdefs}}
     {f : \Pi y_1,\ldots,y_n. \notg{\Gamma_1} \rightarrow (\notg{\Gamma_2}, \notg{R})}
 \end{mathpar}
+
+\begin{definition} Define $\clean(\Gamma)$ as follows:
+\begin{eqnarray*}
+\clean(\varnothing) &=& \varnothing \\[-1.5mm]
+\clean(\setit{[\mx{z}{\cid}, \mx{y}{\cid'}] : \sharety} \cup \Gamma) &=& \clean(\Gamma)\\[-1.5mm]
+\clean(\setit{\pubx{\mx{z}{\cid}} : \pubty} \cup \Gamma) &=& \clean(\Gamma)\\[-1.5mm]
+\clean(\setit{\stt : \sty} \cup \Gamma) &=& \setit{\stt : \sty} \cup \clean(\Gamma)
+\end{eqnarray*}
+\end{definition}
+
+\begin{lemma}
+  Assume $\cmd \redx \prog$ and $\htrip{\eqs_1 \wedge \toeq{\xdefs_1}}{\prog}{\eqs_2}$ and
+  $\shtrip{\Gamma_1,R_1,\xdefs_1}{\cmd}{\Gamma_2,R_2,\xdefs_2}$ and   
+  $\eqs_1 \wedge \toeq{\xdefs_1},\clean(\Gamma_1),R_1 : {OK}$.
+  Then $\eqs_1 \wedge \toeq{\xdefs_1} \wedge \toeq{\prog} \vdash \clean(\Gamma_1),R_1 \stredx^* \clean(\Gamma_2),R_2$.
+\end{lemma}
+
+\begin{lemma}
+  Assume $\htrip{\eqspre}{\ttt{main()}}{\eqs}$ and $\ttt{main} : \Gamma_1 \rightarrow (\Gamma_2,R)$
+  and $\eqspre,\Gamma_1,R' : {OK}$ with $R' \cap R = \varnothing$. Given $\ttt{main()} \redx \prog$,
+  then $\eqspre \wedge \toeq{\prog} \vdash \varnothing, \varnothing \stredx^* \clean(\Gamma_2),R$.
+\end{lemma}
+
+\begin{theorem}
+  Assume $\htrip{\eqspre}{\ttt{main()}}{\eqs}$ and $\ttt{main} : \Gamma_1 \rightarrow (\Gamma_2,R)$
+  and $\eqspre,\Gamma_1,R' : {OK}$ with $R' \cap R = \varnothing$. Given $\ttt{main()} \redx \prog$,
+  with $\iov(\prog) = S \cup V \cup O$, then $\prog$ satisfies noninterference modulo output if
+  $\realviews{\setit{\cid_1}}{\setit{\cid_2}} \cup O
+  \subseteq \ov(\clean(\Gamma_2))$ for all $\cid_1$, $\cid_2$ with $\cid_1 \ne \cid_2$.
+\end{theorem}

### 2025-11-11 code/cvc5mvn/src/test/scala/plaid/prelude/logic/WellFormedTest.scala
move test to appropriate package

@@ -1,9 +1,8 @@
-package plaid.prelude.eval
+package plaid.prelude.logic
 
-import org.junit.Test
 import org.junit.Assert.{assertFalse, assertTrue}
+import org.junit.Test
 import plaid.prelude.antlr.Loader
-import plaid.prelude.logic.WellFormed
 
 class WellFormedTest {
 
@@ -15,41 +14,35 @@ class WellFormedTest {
 
   /** Memories should always come with party indexes */
   @Test
-  def validMemories(): Unit = {
+  def validMemories(): Unit =
     assertFails("""m["m"]@2 == m["k"]@1 + (m["delta"]@1 * m["s"])@2""")
     assertFails("""out@1 == (m["x"] + m["y"])@1""")
     assertPasses("""m["m"]@2 == m["k"]@1 + (m["delta"]@1 * m["s"]@2)""")
-  }
 
   /** Constraint expression should contain only the types of nodes allowed in the grammar */
   @Test
-  def constraintNodeTypesOnly(): Unit = {
+  def constraintNodeTypesOnly(): Unit =
     assertFails("""{ mesg= m["x"] ; rand = r["x"] }.mesg@1  ==  m["x"]@1""")
     assertFails("""let x = "foo" in m[x]@1 == m[x]@1""")
     assertPasses("""m["x"]@1  ==  m["x"]@1""")
-  }
 
   /** Party indexes should be numbers */
   @Test
-  def partyIndexNumbers(): Unit = {
+  def partyIndexNumbers(): Unit =
     assertFails("""s["w"]@i == p["w"]@1""")
     assertPasses("""s["w"]@1 == p["w"]@1""")
-  }
 
   /** Memory indexes should always be strings */
   @Test
-  def memoryIndexesStrings(): Unit = {
+  def memoryIndexesStrings(): Unit =
     assertFails("""out@1 == m[x ++ "foo"]@1""")
     assertFails("""out@1 == m[x]@1""")
     assertFails("""out@1 == m[m[x ++ "foo"]]@1""")
     assertPasses("""out@1 == m["xfoo"]@1""")
-  }
 
   /** At expressions cannot be nested inside one another */
   @Test
-  def nestedAts(): Unit = {
+  def nestedAts(): Unit =
     assertPasses("""out@1 == (m["x"]@1 + m["y"]@1)""")
     assertFails("""out@1 == (m["x"]@1 + m["y"]@1)@1""")
-  }
-
 }

### 2025-11-11 toplas26/metalang.tex
checkpoint

@@
 pre-verified properties of functions that can be reused in, and reduce
-overhead of, whole protocol verification.
+overhead of, whole protocol verification. A representative example of
+a GMW $\ttt{and}$ gate \cite{evans2018pragmatic} implementation is
+given in Figure \ref{fig-gmwprepost}. Here, the $\ttt{andgmw}$
+function takes as input strings $x$ and $y$ that distinguish values
+secret-shared between clients 1 and 2 as $\mx{x}{1}, \mx{x}{2}$ and
+$\mx{y}{1},\mx{y}{2}$ respectively-- that is, these shares encode
+$\ttt{RECON}(x)$ and $\ttt{RECON}(y)$ The postcondition of the
+protocol defined by the $\ttt{andgmw}$ specifies that it results in a
+value shared between clients 1 and 2 as $\ttt{RECON}(z)$ that is the
+conjunction of $\ttt{RECON}(x)$ and $\ttt{RECON}(y)$. This is
+non-trivial to prove, and expressing this postcondition in a
+compositional manner reduces proof overhead and improves scalability
+to large circuits that may incorporate a multitude of $\ttt{and}$
+gates.
+
+\begin{fpfig}[t]{GMW $\ttt{and}$ gate with postcondition.}{fig-gmwprepost}
+\begin{verbatimtab}
+RECON(x) { (m[x]@1 + m[x]@2) }
+
+andtablegmw(x, y, z) {
+  let r11 = (r[z] + (m[x] + 1) * (m[y] + 1)) in
+  let r10 = (r[z] + (m[x] + 1) * (m[y] + 0)) in
+  let r01 = (r[z] + (m[x] + 0) * (m[y] + 1)) in
+  let r00 = (r[z] + (m[x] + 0) * (m[y] + 0)) in
+  { row1 = r11; row2 = r10; row3 = r01; row4 = r00 }
+}
+
+andgmw(z:string, x:string, y:string) {
+  let table = andtablegmw(x,y,z) in
+  m[z]@2 := OT4((m[x], m[y])@2, table.row4, table.row3, table.row2, table.row1)@1;
+  m[z]@1 := r[z]@1
+}
+postcondition:  ( RECON(z) == RECON(x) * RECON(y))
+\end{verbatimtab}
+\end{fpfig}
 

### 2025-11-11 toplas26/metaty.tex
checkpoint

@@
 
+\begin{fpfig}[t]{GMW $\ttt{and}$ gate with share type annotations.}{fig-gmwsharetypes}
+\begin{verbatimtab}
+andgmw(z:string, x:string, y:string) {
+  let table = andtablegmw(x,y,z) in
+  m[z]@2 := OT4((m[x], m[y])@2, table.row4, table.row3, table.row2, table.row1)@1;
+  m[z]@1 := r[z]@1;
+  m[z]@2 as (RECON(x) * RECON(y)) + r[z]@1;
+  [m[z]@1,m[z]@2] : shares
+}
+postcondition:  ( RECON(z) == RECON(x) * RECON(y))
+\end{verbatimtab}
+\end{fpfig}
+
 \begin{definition} Define $\clean(\Gamma)$ as follows:

### 2025-11-11 code/cvc5mvn/src/test/scala/plaid/prelude/cvc/TermFactoryTest.scala
reinstate most term factory tests

@@ -1,12 +1,18 @@
 package plaid.prelude.cvc
 
-import io.github.cvc5.{CVC5ApiException, Kind, Solver, Term, TermManager}
-import org.junit.Assert.{assertEquals, assertNotEquals, assertNotNull}
+import io.github.cvc5.{Term, TermManager}
+import org.junit.Assert.{assertEquals, assertNotEquals}
 import org.junit.Test
 import plaid.prelude.antlr.Loader
 import plaid.prelude.ast.*
 
 class TermFactoryTest {
+
+  private def mockModel(factory: TermFactory, nameBasedModel: Map[String, Int]): Map[Term, Int] =
+    nameBasedModel.map { case (name, value) =>
+      val term = factory.getMemories.find(_.name == name).get.term
+      term -> value
+    }
 
   /**
    * Two numbers generated by two different term managers are equal. This test
@@ -90,264 +96,121 @@ class TermFactoryTest {
     val factory = new TermFactory(termManager, "7")
     factory.toTerm(expr)
 
-  /*
-
+  /** Numeric expressions convert to finite field values. */
   @Test
-  @throws[Exception]
-  def printModel(): Unit = {
-    val tm = new TermManager()
-    val sort = tm.mkFiniteFieldSort("2", 10)
-    val solver = new Solver(tm)
-    solver.setOption("produce-models", "true")
-    // solver.setOption("finite-model-find", "true")
+  def numericTerms(): Unit =
+    val termManager = new TermManager()
+    val factory = new TermFactory(termManager, "7")
+    val term = factory.toTerm(Num(1))
+    assertEquals("1", term.getFiniteFieldValue)
 
-    solver.setLogic("ALL")
-    solver.resetAssertions()
+  /** Registers memory nodes when they are children of other nodes. */
+  @Test
+  def registerChildren(): Unit =
+    val termManager = new TermManager()
+    val factory = new TermFactory(termManager, "7")
+    val mem = AtExpr(MessageExpr(Str("x")), Num(3))
+    val expr = PlusExpr(Num(6), mem)
+    factory.toTerm(expr)
+    val memories = factory.getMemories
+    assertEquals(1, memories.size)
 
-    val m_x_1 = tm.mkFiniteFieldElem("1", sort, 10)
-    val s_x_2 = tm.mkConst(sort, "s_x_2")
-    val r_x_2 = tm.mkConst(sort, "r_x_2")
[…]

### 2025-11-11 toplas26/metalang.tex
checkpoint

@@
   \label{lemma-fresh}
-  Given $\eqs$ and $w, \cid = \fresh(\mathit{string}, \mathit{cid})$ and
+  Given $\eqs$ and $w, \cid = \fresh(\stringty, \cidty)$ and
   some $w',\cid'$ where 

### 2025-11-11 toplas26/metaty.tex
checkpoint

@@
 checking. To enable this we introduce \emph{share type annotations}
+
+\begin{definition}
+  Define $\Gamma \Vdash \stt : \sty$ iff $\Gamma,\true \Vdash \stt : \sty$ without the
+  use of \TirName{ShareEntails} and \TirName{PubEntails}.
+\end{definition}
+
+\begin{definition}[$\metaprot$ Type Checking Judgements]
+  \emph{Type checking judgements} are of the syntactic form
+  $\shtrip{\Gamma, R, \xdefs}{\cmd}{\Gamma, R, \xdefs}$
+  where $\xdefs : \phi \rightarrow \phi$ are called
+  \emph{definitions}.
+\end{definition}
 
@@
   \inferrule
-      {\notg{\stt}_1 \redx \stt_1 \cdots \notg{\stt}_n \redx \stt_n}
+      {\notg{\stt}_1 \redx \stt_1\ \cdots\ \notg{\stt}_n \redx \stt_n}
       {\setit{\notg{\stt}_1 : \sty_1, \ldots, \notg{\stt}_n : \sty_n} \redx
@@
   \inferrule
-      {e_1 \redx \rx{w_1}{\cid_1} \cdots e_1 \redx \rx{w_n}{\cid_n}}
+      {e_1 \redx \rx{w_1}{\cid_1}\ \cdots\ e_1 \redx \rx{w_n}{\cid_n}}
       {\setit{e_1,\ldots,e_n} \redx
@@
 
-\begin{mathpar}
-  \inferrule[Coerce]
-      {e_1 \redx \phi_1 \\ e_2 \redx \phi_2}
-      {\shtrip{\Gamma, R, \xdefs}{e_1 \kwas e_2}{\Gamma, R, \xdefs[\phi_1 \mapsto \phi_2]}}
+\begin{definition}[$\metaprot$ Function Types]
+  Dependent function types are of the syntactic form
+  $\Pi y_1,\ldots,y_n. \notg{\Gamma} \rightarrow (\notg{\Gamma}, \notg{R})$.
+  where $y_1,\ldots,y_n$ range over $\stringty$ or $\cidty$ types.
+\end{definition}
 
-  \inferrule[LocalShares]
-      {e_1 \redx x_1 \\ e_2 \redx x_2  \\
-        \Gamma \Vdash [\xdefs(x_1), \xdefs(x_2)] : \sharety}
-      {\shtrip{\Gamma, R, \xdefs}{[e_1,e_2] : \sharety}
-        {\Gamma \cup \setit{[x_1,x_2] : \sharety}, R, \xdefs}}
+\shtripstafig
 
-  \inferrule[GenShares]
-      {e_1 \redx x_1 \\ e_2 \redx x_2  \\ \xdefs(x_1) = \phi - \rx{w}{\cid} \\ \xdefs(x_2) =  \rx{w}{\cid}}
-      {\shtrip{\Gamma, R, \xdefs}{[e_1,e_2] : \sharety}
-        {\Gamma \cup \setit{[x_1,x_2] : \sharety}, R \uplus \setit{\rx{w}{\cid}}, \xdefs}}
-
-  \inferrule[LocalPublic]
-      {e \redx \phi \\ \Gamma \Vdash \pubx{\xdefs(\phi)} : \pubty}
-      {\shtrip{\Gamma, R, \xdefs}{\pubx{e} : \pubty}
-        {\Gamma \cup \setit{\pubx{\phi} : \pubty}, R, \xdefs}}
-
-  \inferrule[MaskedPublic]
-      {e_1 \redx x_1 \\ e_2 \redx x_2  \\
-        \Gamma \Vdash [x_1, x_2] : \sharety \\
-        \xdefs(x_1 + x_2)  = \phi + \rx{w}{\oid}}
-      {\shtrip{\Gamma, R, \xdefs}{\pubx{e_1+e_2} : \pubty}
-        {\Gamma \cup \setit{\pubx{x_1 + x_2} : \pubty}, R \uplus \setit{\rx{w}{\oid}}, \xdefs}}
-
-  \inferrule[Reveal]
-      {e_1 \redx x_1 \\ e_2 \redx x_2  \\
-        \Gamma \Vdash [x_1, x_2] : \sharety}
-      {\shtrip{\Gamma, R, \xdefs}{\pubx{e_1 + e_2} : \outty}
-        {\Gamma \cup \setit{\pubx{x_1 \fplus x_2} : \outty}, R, \xdefs}}
-
-  \inferrule[Output]
-      {e \redx \out{\cid} \\
-       \xdefs(\out{\cid}) = x_1 + x_2 \\
-       \Gamma \Vdash \pubx{x_1 + x_2} : \outty }
-      {\shtrip{\Gamma, R, \xdefs}{\out{\cid} : \outty}
-        {\Gamma \cup \setit{\pubx{\out{\cid}} : \outty}, R, \xdefs}}
-\end{mathpar}
-
-\begin{mathpar}
-  \inferrule[Defn]
-      {e_1 \redx x \\ e_2 \redx \elab{\be}{\cid}}
-      {\shtrip{\Gamma, R, \xdefs}{e_1 := e_2}{\Gamma, R, \xdefs[x \mapsto \toeq{\elab{\be}{\cid}}]}}
-
-  \inferrule[App]
-      {f : \Pi y_1,\ldots,y_n. \notg{\Gamma_1} \rightarrow (\notg{\Gamma_2}, \notg{R'}) \\
-        \notg{\Gamma_1}[e_1/y_1 \cdots e_n/y_n] \redx \Gamma_1 \\
-        \notg{\Gamma_2}[e_1/y_1 \cdots e_n/y_n] \redx \Gamma_2 \\
-        \notg{R'}[e_1/y_1 \cdots e_n/y_n] \redx R'        
-      }
-      {\shtrip{\Gamma \cup \Gamma_1, R, \xdefs}{f(e_1,\ldots,e_n)}{\Gamma \cup \Gamma_1 \cup \Gamma_2, R \uplus R', \xdefs}}
-
-  \inferrule[Seq]    
-      {\shtrip{\Gamma_1,R_1,\xdefs_1}{\cmd_1}{\Gamma_2,R_2,\xdefs_2} \\
-       \shtrip{\Gamma_2,R_2,\xdefs_2}{\cmd_2}{\Gamma_3,R_3,\xdefs_3}}
-      {\shtrip{\Gamma_1,R_1,\xdefs_1}{\cmd_1;\cmd_2}{\Gamma_3,R_3,\xdefs_3}}
-
-  \inferrule[Let]
-      {\shtrip{\Gamma_1,R_1,\xdefs_1}{\cmd[e/y]}{\Gamma_2,R_2,\xdefs_2}}
-      {\shtrip{\Gamma_1,R_1,\xdefs_1}{\elet{y}{e}{\cmd}}{\Gamma_2,R_2,\xdefs_2}}
-
-  \inferrule[Fn]
-            {\codebase(f) = y_1 : \tau_1, \ldots, y_n : \tau_n, \cmd \\
-              \mv_1,\ldots,\mv_n = \fresh(\tau_1,\ldots,\tau_n) \\
-              \notg{\Gamma_1}[\mv_1/y_1 \cdots \mv_n/y_n] \redx \Gamma_1 \\
-              \notg{\Gamma_2}[\mv_1/y_1 \cdots \mv_n/y_n] \redx \Gamma_2 \\
-              \notg{R}[\mv_1/y_1 \cdots \mv_n/y_n] \redx R \\  
-     \shtrip{\Gamma_1, \varnothing, \varnothing}{\cmd[\mv_1/y_1 \cdots \mv_n/y_n]}{\Gamma_2, R,\xdefs}}
-    {f : \Pi y_1,\ldots,y_n. \notg{\Gamma_1} \rightarrow (\notg{\Gamma_2}, \notg{R})}
-\end{mathpar}
+\shtripcmdfig
 

### 2025-11-11 toplas26/metaty.tex
checkpoint

@@
 
-Share types rely on SMT entailments via the \TirName{ShareEntails} and \TirName{PubEntails} rules,
-which are not syntax directed. Our goal with the $\metaprot$ language is to automate share type
-checking. To enable this we introduce \emph{share type annotations}
+Share types rely on SMT entailments via the \TirName{ShareEntails} and
+\TirName{PubEntails} rules, which are not syntax directed. Our goal
+with the $\metaprot$ language is to automate share type checking. To
+enable this we introduce \emph{share type annotations}. These
+annotations allow programmer specification of properties that can be
+verified as part of the Hoare triple analysis defined in Section
+\ref{section-metahtrip} and that enable automated checking of types.
 
+A fundamental component of the system is automated share type checking
+of noninteractive computations.  We define an algorithmic, syntax
+directed subset of noninteractive share typing as follows:
 \begin{definition}
@@
 \end{definition}
+
+The syntax of type annotations $\sta$ is defined in Figure
+\ref{fig-sta}, which we extend the language of commands $\cmd$ to
+include. Type annotations are of the form $e_1 \kwas e_2$, which
+defines an SMT equivalence, and $e : \sty$ which are verified using
+the automated $\Vdash$ typing judgement-- the key idea here being that
+explicit equivalences can stand in for implicit \TirName{ShareEntails}
+and \TirName{PubEntails} judgements. In the semantics type annotations
+are interpreted as no-ops, and their integration in Hoare triple
+derivations requires annotated equivalences to be included in
+preconditions.  In an implementation, Hoare triple verification can be
+conducted independently prior to type checking.
+
+\subsubsection{Annotation Type Checking}
 

### 2025-11-11 toplas26/metaty.tex
checkpoint

@@
 
+\stafig
+
 The syntax of type annotations $\sta$ is defined in Figure
@@
 
+Type checking judgements for typing annotations are defined as follows.
+The \emph{definitions} element $\xdefs$ of judgements is an environment
+of SMT equivalences induced either by the program or by annotation. 
 \begin{definition}[$\metaprot$ Type Checking Judgements]
@@
 
-\stafig
+\shtripstafig
+
+All typing judgements are dependent on closed expressions $e$ which we
+equate up to evaluation. Since the semantics of expressions is total,
+we simply evaluate expressions in preconditions of derivation
+rules. The derivation rules for annotations $\sta$ are given in Figure
+\ref{fig-shtripsta}. First, we note that the \TirName{Coerce} rule
+extends definitions to include the equivalence specified by the $e_1
+\kwas e_2$ annotation form-- we assume the validity of the equivalence
+is enforced by Hoare triple verification and so is not checked (again)
+here.  These definitions extend or even overwrite variable definitions
+established by the protocol itself (see the \TirName{Defn} rule in
+Figure \ref{fig-shtripcmd} discussed below).
+
+In the rules for share type annotations of the form $\notg{\stt} :
+\sty$,
+
+\subsubsection{Functions and Application}
 
@@
 \end{definition}
-
-\shtripstafig
 

### 2025-11-11 toplas26/metaty.tex
checkpoint

@@
 
-In the rules for share type annotations of the form $\notg{\stt} :
-\sty$,
+Programmers can also specify bindings for local computations, for
+example if $\mx{x}{1},\mx{x}{2}$ are shares obtained by, e.g., local
+homomorphic addition, the programmer can specify
+$[\mx{x}{1},\mx{x}{2}] : \sharety$. This supports compositionality of
+type checking. Thus, in the rules for share type annotations of the
+form $\notg{\stt} : \sty$, we have different rules for checking local and interactive
+$\sharety$ type ascriptions (\TirName{LocalShares} and \TirName{GenShares},
+respectively), and for checking local and interactive
+$\pubty$ type ascriptions (\TirName{LocalPublic} and \TirName{MaskedPublic},
+respectively). Observe the use of definitions $\xdefs$ to refer to
+SMT equivalences, and the use of $\Vdash$ to verify share typing.
 

### 2025-11-12 toplas26/metaty.tex
checkpoint

@@
 $\pubty$ type ascriptions (\TirName{LocalPublic} and \TirName{MaskedPublic},
-respectively). Observe the use of definitions $\xdefs$ to refer to
-SMT equivalences, and the use of $\Vdash$ to verify share typing.
+respectively). In these rules we use of definitions $\xdefs$ to refer to
+SMT equivalences, and $\Vdash$ to verify share typing.
 
@@
 
+Rules for other commands are defined in Figure \ref{fig-shtripcmd}.
+As anticipated above, the \TirName{Defn} rule automatically adds the
+protocol definition of variables to $\xdefs$ as a default equivalence,
+which may allow automatic share typing without any use of
+\TirName{Coerce}. The most complex rules are for function application
+and definition (\TirName{App} and \TirName{Fn}). These rules allow for
+generalization and instantiation of share type environments and
+thus compositionality in typing. Generalized, aka abstract
+share type environments are dependent on $\stringty$ and
+$\cidty$ types, and we equate them up to evaluation in
+our rules.
 \begin{definition}[Abstract Share Type Environments]
@@
 
+We also introduce dependent $\Pi$ types for functions, that allow
+dependence on $\stringty$ and $\cidty$ types.
 \begin{definition}[$\metaprot$ Function Types]
@@
 \end{definition}
+As defined in \TirName{Fn}, a dependent function type $\Pi
+y_1,\ldots,y_n. \notg{\Gamma_1} \rightarrow (\notg{\Gamma_2},
+\notg{R})$ specifies that, given share typing assumptions in
+$\notg{\Gamma_1}$, application the function results in the bindings in
+$\notg{\Gamma_2}$ by using the random values in $\notg{R}$. It is
+always the case that the type is closed, i.e., only $y_1,\ldots,y_n$
+can occur free in $\notg{\Gamma_1}$, $\notg{\Gamma_2}$ $\notg{R}$.
+Type checking for functions is accomplished by using fresh ``witness''
+values of $\stringty$ and $\cidty$ types to instantiate dependent
+variables, similarly to the \TirName{GenEntails} rule. As defined in
+the \TirName{App} rule, typing ensures that application of the
+function instantiates the dependencies, checks that the instantiation
+of $\notg{\Gamma_1}$ is contained in the share typing environment, and
+adds the instantiation of $\notg{\Gamma_2}$ as an effect, as will as
+the instantiation of $\notg{R}$ with linearity enforced.
 
 \shtripcmdfig
+
+\subsubsection{Example: GMW \ttt{and} Gate}
+
+Returning to the GMW \ttt{and} gate example introduced in Figure \ref{fig-gmwprepost},
+we can extend the definition of \ttt{andgmw} with type annotations to ensure a
+valid share typing, as in Figure \ref{fig-gmwsharetypes}. Here a coercion
+is needed to specify that the gate output are shares obtained using
+$\ttt{r[z]@1}$ as an OTP. The type for $\ttt{andgmw}$ can then be
+obtained as:
+$$
+\ttt{andgmw} : \Pi x,y,z . \varnothing \rightarrow (\setit{[m[z]@1,m[z]@2] : \sharety}, \setit{r[z]@1})  
+$$
+Note that the function does not require any share type preconditions, since
+the (sub)protocol it defines interactively generates new shares using the OTP.
 

### 2025-11-12 toplas26/metaty.tex
checkpoint

@@
 
+When typing functions, a practical programming detail is where the
+share type environment precondition $\notg{\Gamma_1}$ is defined.
+In our implementation, we write these as an additional component
+of \ttt{precondition} annotations, as illustrated in extended
+examples discussed in Section \ref{section-examples}.
+
 \shtripcmdfig
@@
 \end{fpfig}
+
+\subsubsection{Soundness of $\metaprot$ Type Checking}
 

### 2025-11-13 toplas26/metaty.tex
checkpoint

--- 2025-11-12 toplas26/metaty.tex
@@
 adds the instantiation of $\notg{\Gamma_2}$ as an effect, as will as
-the instantiation of $\notg{R}$ with linearity enforced.
+the instantiation of $\notg{R}$ with linearity enforced. A practical
+programming detail is where the share type environment precondition
+$\notg{\Gamma_1}$ is defined.  In our implementation, we write these
+as an additional component of \ttt{precondition} annotations, as
+illustrated in extended examples discussed in Section
+\ref{section-examples}.
 
@@
 Note that the function does not require any share type preconditions, since
-the (sub)protocol it defines interactively generates new shares using the OTP.
+the (sub)protocol it defines interactively generates new shares using an OTP
+($r[z]@1$).
 
@@
 }
-postcondition:  ( RECON(z) == RECON(x) * RECON(y))
+postcondition: (RECON(z) == RECON(x) * RECON(y))
 \end{verbatimtab}
 \end{fpfig}
+
+\subsubsection{Soundness of $\metaprot$ Type Checking}
+
+\begin{lemma}
+  Given $\prog_1,\Gamma_1,R_1 : {OK}$ and $\cmd \redx \prog_2$ and $\htrip{\eqs_1}{\cmd}{\eqs_2}$ with
+  $\eqs_1$ satisfiable and $\toeq{\prog_1} \models \eqs_1 \wedge \toeq{\xdefs_1}$ and
+  $\shtrip{\Gamma_1,R_1,\xdefs_1}{\cmd}{\Gamma_2,R_2,\xdefs_2}$.
+  Then $\prog_1;\prog_2,\Gamma_2,R_2 : {OK}$ and $\toeq{\prog_1;\prog_2} \models \toeq{\xdefs_2}$.
+\end{lemma}
+
+\begin{theorem}
+  Given $\htrip{\eqspre}{\ttt{main()}}{\eqs}$ and $\ttt{main} : \Gamma_1 \rightarrow (\Gamma_2,R)$
+  and $\main() \redx \prog$ and $\progpre,\Gamma_1,R' : {OK}$ with $R' \cap R = \varnothing$
+  and $\toeq{\progpre} \models \eqspre$. Then $(\progpre;\prog),\Gamma_2,R : {OK}$.
+\end{theorem}
+
+% DEPRECATED
+\begin{comment}
+\begin{lemma}
+  Assume $\cmd \redx \prog$ and $\htrip{\eqs_1 \wedge \toeq{\xdefs_1}}{\prog}{\eqs_2}$ and
+  $\shtrip{\Gamma_1,R_1,\xdefs_1}{\cmd}{\Gamma_2,R_2,\xdefs_2}$ and   
+  $\eqs_1 \wedge \toeq{\xdefs_1},\clean(\Gamma_1),R_1 : {OK}$.
+  Then $\eqs_1 \wedge \toeq{\xdefs_1} \wedge \toeq{\prog} \vdash \clean(\Gamma_1),R_1 \stredx^* \clean(\Gamma_2),R_2$.
+\end{lemma}
 
@@
 \end{theorem}
+
+\end{comment}

### 2025-11-13 toplas26/sty.tex
checkpoint

@@
 
-We write $\stredx^*$ to denote the reflexive, transitive
+We write $\stredx^*$ to denote the transitive
 closure of $\stredx$. For any protocol $\prog$, the goal of the
@@
 typing judgements.
+
+\begin{definition}
+  Given $\prog,\Gamma,R$ where $\secrets(\prog) = S$ and:
+  $$
+  \Gamma = \setit{[\phi_1^1,\phi_1^2] : \sharety, \ldots, [\phi_n^1,\phi_n^2] : \sharety}
+  \cup \setit{\pubx{\phi_1} : \pubty,\ldots,\pubx{\phi_j} : \pubty} \cup
+  \setit{\pubx{\phi^o_1} : \outty,\ldots,\pubx{\phi^o_k} : \outty}
+  $$
+  Then we write $\prog,\Gamma,R : {OK}$ iff all of the following conditions
+  hold:
+  \begin{enumerate}[\hspace{5mm}i.]
+  \item For all $a,b \in \setit{1,2}$ we have
+    $
+    \condd{\progtt(\prog)}{S_{\setit{\cid_a}}}{S_{\setit{\cid_b}} \cup \setit{\phi^o_1,\ldots,\phi^o_k}}
+    = 
+    \condd{\progtt(\prog)}{S_{\setit{\cid_a}}}{S_{\setit{\cid_b}} \cup \setit{\phi_1^b,\ldots,\phi_n^b} \cup
+      \setit{\phi_1,\ldots,\phi_j} \setit{\phi^o_1,\ldots,\phi^o_k}}
+    $.
+  \item For all $a,b \in \setit{1,2}$and $0 < i \le n$ we have
+    $\condd{\progtt(\prog)}{S_{\setit{\cid_a}}}{\setit{\phi_i^1 + \phi_i^2}} =
+    \condd{\progtt(\prog)}{S_{\setit{\cid_a}}}{\setit{\phi_i^1 + \phi_i^2, \phi_i^a}}$
+  \item  For all $a,b \in \setit{1,2}$ and $0 < i \le n$, if
+    $\toeq{\prog} \models \phi_i^a \eop \rx{w}{\cid_a} \wedge \phi_i^b \eop \phi - \rx{w}{\cid_a}$ for
+    some $\phi$ and $\rx{w}{\cid_a}$ then $\rx{w}{\cid_a} \in R$.
+  \item For all $0 < i \le j$, if $\toeq{\prog} \models \phi_i \eop \phi + \rx{w}{\cid_{\oid}}$ for
+    some $\phi$  and $\rx{w}{\cid_{\oid}}$ then $\rx{w}{\cid_{\oid}} \in R$.
+  \end{enumerate} 
+\end{definition}
+
+\begin{lemma}
+  \label{lemma-sty-noninteractive-sound}
+  Given $\prog,\Gamma,R$ where $\secrets(\prog) = S$ and:
+  $$
+  \Gamma = \setit{[\phi_1^1,\phi_1^2] : \sharety, \ldots, [\phi_n^1,\phi_n^2] : \sharety}
+  \cup \setit{\pubx{\phi_1} : \pubty,\ldots,\pubx{\phi_j} : \pubty} \cup
+  \setit{\pubx{\phi^o_1} : \outty,\ldots,\pubx{\phi^o_k} : \outty}
+  $$
+  and also $\Gamma, \toeq{\prog} \vdash [\phi^1,\phi^2] : \sharety$
+  Then for all $a,b \in \setit{1,2}$ the following conditions hold:
+    \begin{enumerate}[\hspace{5mm}i.]
+    \item $\condd{\progtt(\prog)}{S_{\setit{\cid_a}}}{S_{\setit{\cid_b}} \cup \setit{\phi_1^b,\ldots,\phi_n^b} \cup
+      \setit{\phi_1,\ldots,\phi_j} \setit{\phi^o_1,\ldots,\phi^o_k}}
+    =
+    \condd{\progtt(\prog)}{S_{\setit{\cid_a}}}{S_{\setit{\cid_b}} \cup \setit{\phi_1^b,\ldots,\phi_n^b,\phi^b} \cup
+       \setit{\phi_1,\ldots,\phi_j} \setit{\phi^o_1,\ldots,\phi^o_k}}$
+    \item $\condd{\progtt(\prog)}{S_{\setit{\cid_a}}}{\setit{\phi_i^1 + \phi_i^2}} =
+      \condd{\progtt(\prog)}{S_{\setit{\cid_a}}}{\setit{\phi_i^1 + \phi_i^2, \phi_i^a}}$
+    \end{enumerate}
+\end{lemma}
+
+\begin{lemma}
+  Given $\prog,\Gamma_1,R_1  : {OK}$
+  and $\toeq{\prog}  \vdash \Gamma_1, R_1 \stredx \Gamma_2, R_2$.
+  Then $\toeq{\prog}, \Gamma_2, R_2 : {OK}$.
+\end{lemma}
+
+% DEPRECATED
+\begin{comment}
 \begin{definition}
@@
 \end{theorem}
+\end{comment}

### 2025-11-13 toplas26/sty.tex
checkpoint

@@
     $.
-  \item For all $a,b \in \setit{1,2}$and $0 < i \le n$ we have
+  \item For all $a,b \in \setit{1,2}$ and $0 < i \le n$ we have
     $\condd{\progtt(\prog)}{S_{\setit{\cid_a}}}{\setit{\phi_i^1 + \phi_i^2}} =
@@
 \end{lemma}
+
+\begin{definition}
+  Given  $\prog,\Gamma,R  : {OK}$ and $\iov(\prog) = S \cup V \cup O$.
+  Then $\Gamma$ is \emph{closed for $\prog$} iff for all $H,C$,
+  for all $x \in \houtputs$ there exists $\phi$ such that either
+  $\Gamma,\toeq{\prog} \vdash [x,\phi] : \sharety$ if $C = \setit{1}$
+  or  $\Gamma,\toeq{\prog} \vdash [\phi,x] : \sharety$ if $C = \setit{2}$,
+  and for all $\out{\cid} \in O$ we have $\Gamma,\toeq{\prog} \vdash \pubx{\out{\cid}} : \outty$.
+\end{definition}
+
+\begin{theorem}
+  If $\toeq{\prog} \vdash \varnothing,\varnothing \stredx \Gamma,R$ and $\Gamma$ is closed
+  for $\prog$, then $\prog$ satisfies noninterference modulo output. 
+\end{theorem}
 

### 2025-11-13 toplas26/examples.tex
checkpoint

--- 2025-01-26 esop25/examples.tex
@@
 
+\begin{verbatimtab}
+  IDEALF() { (s["x"]@1 + s["y"]@2) + (s["i"]@1 * s["j"]@2) }
+
+RECON(x) { (m[x++"s"]@1 + m[x++"s"]@2) }
+
+// These are Beaver triple constraints. Party 3 is the oracle.
+BEAVERTRIP(x) {
+  m[x++"c"]@3 == r[x++"a"]@3 * r[x++"b"]@3 AND
+  m[x++"as"]@1 + m[x++"as"]@2 == r[x++"a"]@3 AND
+  m[x++"bs"]@1 + m[x++"bs"]@2 == r[x++"b"]@3 AND
+  m[x++"cs"]@1 + m[x++"cs"]@2 == m[x++"c"]@3 
+}
+
+sshare(x:string, i1:cid, i2:cid) {
+   m[x++"s"]@i2 := (s[x] + -r[x])@i1;
+   m[x++"s"]@i1 := r[x]@i1
+}
+
+_open(x:string, i1:cid, i2:cid, n:string){
+    p[x++"n"] := m[x]@i2;
+    m[x++"p"]@i1 = p[x++"n"] + m[x]@i1
+}
+
+open(x:string) {
+  _open(x,1,2,"1");
+  _open(x,2,1,"2")
+}
+
+secopen(d:string;x:string,a:string) {
+    sum(d, x, a);				       
+    open(d)
+}
+
+_sum(z:string, x:string, y:string, i:cid) { m[z++"s"]@i := (m[x++"s"] + m[y++"s"])@i }
+
+sum(z:string, x:string, y:string){ _sum(z,x,y,1); _sum(z,x,y,2) }
+postcondition: ( RECON(z) == RECON(x) + RECON(y) )
+
+_mult(z : string, y : string, i : cid) {
+    m[z++"s"]@i := (m[z++"d"] * m[y++"s"] + -m[z++"e"] * m[z++"as"] + m[z++"cs"])@i
+}
+
+precondition: ( BEAVERTRIP(z) )
+mult(z : string, x : string, y : string) {
+    secopen(z++"d", x, z++"a");
+    secopen(z++"e", y, z++"b");
+    _mult(z,y,1); _mult(z,y,2)
+}
+postcondition: ( RECON(z) == RECON(x) * RECON(y) )
+
+precondition: ( BEAVERTRIP("k") ) 
+main() {
+    sshare("x", 1, 2);
+    sshare("y", 2, 1);
+    sshare("i", 1, 2);
+    sshare("j", 2, 1);
+
+    sum("z","x","y");
+    mult("k","i","j");
+    sum("out","z","k");
+    open("out")
+}
+postcondition: ( m["out"]@1 == IDEALF() )
+\end{verbatimtab}
+
 \subsection{Confidentiality Examples}
 
-\input{gmw}
+%\input{gmw}
 

### 2025-11-13 code/cvc5mvn/src/main/scala/plaid/prelude/cvc/TermFactory.scala
mostly cosmetics

@@ -6,45 +6,35 @@ import plaid.prelude.ast.*
 import scala.collection.mutable
 import scala.compiletime.uninitialized
 
-object TermFactory {
-  // TODO Don't we specify the field size?
+class TermFactory(val termManager: TermManager, order: String) {
+  // TODO We manually specify field size, this should be renamed?
   private val DEFAULT_FIELD_SIZE = 10
 
   /** Extract party index for assert commands */
-  private def getPartyIndex(command: Cmd): Integer =
-    command match {
-      case x: AssertCmd => CvcUtils.toInt(x.e3)
-      case _            => null
-    }
-}
+  private def getPartyIndex(command: Cmd): Integer = command match
+    case x: AssertCmd => CvcUtils.toInt(x.e3)
+    case _ => null
 
-class TermFactory(val termManager: TermManager, order: String) {
-
-  val sort: Sort = termManager.mkFiniteFieldSort(order, TermFactory.DEFAULT_FIELD_SIZE)
+  val sort: Sort = termManager.mkFiniteFieldSort(order, DEFAULT_FIELD_SIZE)
   private val memories = new mutable.HashSet[Memory]()
-  private val minusOne: Term = termManager.mkFiniteFieldElem("-1", sort, TermFactory.DEFAULT_FIELD_SIZE)
+  private val minusOne: Term = termManager.mkFiniteFieldElem("-1", sort, DEFAULT_FIELD_SIZE)
   private var partyIndex: Integer = uninitialized
 
-  // initialize solver logic
   private val solver = new Solver(termManager)
-  try solver.setLogic("ALL")
-  catch {
-    case e: CVC5ApiException => throw new RuntimeException(e)
-  }
+  solver.setLogic("ALL")
 
   def getMemories: Set[Memory] = memories.toSet
 
   private def not(term: Term): Term =
-    termManager.mkTerm(Kind.FINITE_FIELD_ADD, term, termManager.mkFiniteFieldElem("1", sort, TermFactory.DEFAULT_FIELD_SIZE))
+    termManager.mkTerm(Kind.FINITE_FIELD_ADD, term, termManager.mkFiniteFieldElem("1", sort, DEFAULT_FIELD_SIZE))
 
-  def toTerm(expr: Expr, idx: Integer): Term = {
+  def toTerm(expr: Expr, idx: Integer): Term =
     partyIndex = idx
[…]

### 2025-11-13 code/cvc5mvn/src/main/scala/plaid/prelude/logic/Verifier.scala

@@ -15,5 +15,5 @@ extension (trg: Contract)
 
       // TODO Reorganize the CVC5 stuff, this is weird
       val verifier = new VerifierTmp(cvc)
-      !verifier.entails(cvc.constraintToTerm(a), cvc.constraintToTerm(b))
+      !verifier.entails(cvc.toTerm(a), cvc.toTerm(b))
     }

### 2025-11-13 code/cvc5mvn/src/test/scala/plaid/prelude/cvc/TermFactoryTest.scala

@@ -146,7 +146,7 @@ class TermFactoryTest {
     val termManager = new TermManager()
     val factory = new TermFactory(termManager, "7")
     val ast = Loader.constraint("m[\"foo\"]@3 == 5")
-    val term = factory.constraintToTerm(ast)
+    val term = factory.toTerm(ast)
     val verifier = new VerifierTmp(factory)
     val actual = verifier.findModelSatisfying(term)
     val expected = Map("m_foo_3" -> 5)
@@ -160,7 +160,7 @@ class TermFactoryTest {
     val ast = AndConstraint(
       EqualConstraint(AtExpr(MessageExpr(Str("x")), Num(3)), AtExpr(RandomExpr(Str("y")), Num(5))),
       EqualConstraint(AtExpr(RandomExpr(Str("y")), Num(5)), Num(1)))
-    val term = factory.constraintToTerm(ast)
+    val term = factory.toTerm(ast)
     val verifier = new VerifierTmp(factory)
     val actual = verifier.findModelSatisfying(term)
     val expected = Map("m_x_3" -> 1, "r_y_5" -> 1)
@@ -172,7 +172,7 @@ class TermFactoryTest {
   def notConstraintTerm(): Unit =
     val termManager = new TermManager()
     val factory = new TermFactory(termManager, "7")
-    val term = factory.constraintToTerm(
+    val term = factory.toTerm(
       NotConstraint(
         EqualConstraint(
           AtExpr(MessageExpr(Str("foo")), Num(3)),
@@ -187,7 +187,7 @@ class TermFactoryTest {
   def trueConstraintTerm(): Unit =
     val termManager = new TermManager()
     val factory = new TermFactory(termManager, "7")
-    val term = factory.constraintToTerm(TrueConstraint())
+    val term = factory.toTerm(TrueConstraint())
     assertEquals(termManager.mkTrue(), term)
 
   /** Creates complex constraint expressions */
@@ -196,7 +196,7 @@ class TermFactoryTest {
     val termManager = new TermManager()
     val factory = new TermFactory(termManager, "7")
     val ast = Loader.constraint("out@1 == out@2 + 1 AND out@2 == out@3 + 2 AND out@3 == 4")
-    val term = factory.constraintToTerm(ast)
+    val term = factory.toTerm(ast)
     val verifier = new VerifierTmp(factory)
     val actual = verifier.findModelSatisfying(term)
     val expected = Map("o_3" -> 4, "o_2" -> 6, "o_1" -> 0)

### 2025-11-13 code/cvc5mvn/src/test/scala/plaid/prelude/cvc/VerifierTmpTest.scala

@@ -15,7 +15,7 @@ class VerifierTmpTest {
     satisfiable(Loader.constraint(src))
 
   private def satisfiable(constraint: Constraint): Boolean =
-    val e = termFactory.constraintToTerm(constraint)
+    val e = termFactory.toTerm(constraint)
     verifier.satisfies(e)
 
   /** Single equality constraints are always satisfiable. */
@@ -36,16 +36,16 @@ class VerifierTmpTest {
   /** Everything entails a universally satisfied constraint. */
   @Test
   def everythingEntails(): Unit =
-    val t = termFactory.constraintToTerm(Loader.constraint("T"))
-    val f = termFactory.constraintToTerm(Loader.constraint("NOT T"))
+    val t = termFactory.toTerm(Loader.constraint("T"))
+    val f = termFactory.toTerm(Loader.constraint("NOT T"))
     assertTrue(verifier.entails(t, t))
     assertTrue(verifier.entails(f, t))
 
   /** Nothing entails an unsatisfiable constraint except another unsatisfiable constraint. */
   @Test
   def nothingEntails(): Unit =
-    val t = termFactory.constraintToTerm(Loader.constraint("T"))
-    val f = termFactory.constraintToTerm(Loader.constraint("NOT T"))
+    val t = termFactory.toTerm(Loader.constraint("T"))
+    val f = termFactory.toTerm(Loader.constraint("NOT T"))
     assertFalse(verifier.entails(t, f))
     assertTrue(verifier.entails(f, f))
 }

### 2025-11-14 toplas26/examples.tex
checkpoint

--- 2025-01-26 esop25/examples.tex
@@
 
+\begin{fpfig}[t]{2-party gate library with Beaver Triple multiplication, circuit definition, and compositional correctness conditions.}{fig-beaver}
+\footnotesize
+\begin{verbatimtab}
+IDEALF() { (s["x"]@1 + s["y"]@2) + (s["i"]@1 * s["j"]@2) }
+
+RECON(x) { (m[x]@1 + m[x]@2) }
+
+BEAVERTRIP(x) {
+  m[x++"c"]@3 == r[x++"a"]@3 * r[x++"b"]@3 AND
+  m[x++"a"]@1 + m[x++"a"]@2 == r[x++"a"]@3 AND
+  m[x++"b"]@1 + m[x++"b"]@2 == r[x++"b"]@3 AND
+  m[x++"c"]@1 + m[x++"c"]@2 == m[x++"c"]@3 
+}
+
+sshare(x:string, i1:cid, i2:cid) { m[x]@i2 := (s[x] + -r[x])@i1; m[x]@i1 := r[x]@i1 }
+
+_open(x:string, i1:cid, i2:cid, n:string){ p[x++"n"] := m[x]@i2; m[x++"p"]@i1 = p[x++"n"] + m[x]@i1 }
+
+open(x:string) { _open(x,1,2,"1"); _open(x,2,1,"2") }
+
+secopen(d:string;x:string,a:string) { sum(d, x, a); open(d) }
+
+_sum(z:string, x:string, y:string, i:cid) { m[z]@i := (m[x] + m[y])@i }
+
+sum(z:string, x:string, y:string){ _sum(z,x,y,1); _sum(z,x,y,2) }
+postcondition: ( RECON(z) == RECON(x) + RECON(y) )
+
+_mult(z : string, y : string, i : cid) {
+    m[z]@i := (m[z++"d"] * m[y] + -m[z++"e"] * m[z++"a"] + m[z++"c"])@i
+}
+
+precondition: ( BEAVERTRIP(z) )
+mult(z : string, x : string, y : string) {
+  secopen(z++"d", x, z++"a"); secopen(z++"e", y, z++"b"); _mult(z,y,1); _mult(z,y,2)
+}
+postcondition: ( RECON(z) == RECON(x) * RECON(y) )
+
+precondition: ( BEAVERTRIP("k") ) 
+main() {
+    sshare("x", 1, 2); sshare("y", 2, 1); sshare("i", 1, 2); sshare("j", 2, 1);
+    sum("z","x","y"); mult("k","i","j"); sum("out","z","k");
+    open("out")
+}
+postcondition: ( m["out"]@1 == IDEALF() )
+\end{verbatimtab}
+\end{fpfig}
+
+\begin{fpfig}[t]{2-party gate library with Beaver Triple multiplication extended with share type annotations.}{fig-beaver-types}
+\footnotesize
+\begin{verbatimtab}
+sshare(x:string, i1:cid, i2:cid) {
+   m[x++"s"]@i1 := (s[x] + -r[x])@i2; m[x++"s"]@i2 := r[x]@i2;
+   [m[x++"s"]@i1,m[x++"s"]@i2] : shares
+}
+
+([m[x]@i1, m[x]@i2] : shares)
+open(x:string) {
+  _open(x,1,2,"1"); _open(x,2,1,"2");
+  [p[x++"1"],p[x++"2"]] : shares;
+  m[d++"p"]@2 as m[d++"p"]@1  
+}
+
+precondition: ([m[x]@i1, m[x]@i2] : shares,
+	       [m[a]@i1, m[a]@i2] : shares)
+secopen(d:string;x:string,a:string) {
+    sum(d, x, a); open(d);
+    m[d++"p"]@1 as RECON(x) + r[x++a]@o;
+    <m[d++"p"]@1> : public;
+    <m[d++"p"]@2> : public
+}
+
+precondition: ([m[x]@1,m[x]@2] : shares;
+               [m[y]@1,m[y]@2] : shares)
+sum(z:string, x:string, y:string) {
+    _sum(z,x,y,1); _sum(z,x,y,2);
+    [m[z]@1,m[z]@2] : shares
+}
+postcondition: ( RECON(z) == RECON(x) + RECON(y) )
+
+precondition: ( [m[x]@1,m[x]@2] : shares;
+                [m[y]@1,m[y]@2] : shares;
+	        [m[x++"c"]@1,m[x++"c"]@2] : shares ),
+	       ( BEAVERTRIP(z) )
+mult(z : string, x : string, y : string) {
+    secopen(z++"d", x, z++"a"); secopen(z++"e", y, z++"b"); _mult(z,y,1); _mult(z,y,2);
+    [m[z]@i1,m[z]@i2] : shares
+}
+postcondition: ( RECON(z) == RECON(x) * RECON(y)) )
+\end{verbatimtab}
+\end{fpfig}
+
+\begin{fpfig}[t]{2-party gate library with Beaver Triple multiplication extended with BDOZ MACs.}{fig-beaver-bdoz}
+\footnotesize
+\begin{verbatimtab}
+MACOK(x, i1, i2) { m[x++"m"]@i2 == m[x++"k"]@i1 + (m["delta"]@i1 * m[x++"s"]@i2) }
+
+MACed(x) { MACOK(x, 1, 2) AND MACOK(x, 2, 1) }
+
+_open(x:string, i1:cid, i2:cid, n:string){
+    p[x++"sn"] := m[x++"s"]@i2;
+    p[x++"mn"] := m[x++"m"]@i2;
+    assert(p[x++"mn"] = m[x++"k"] + (m["delta"] * p[x++"sn"]))@i1;
+    m[x++"p"]@i1 = p[x++"sn"] + m[x]@i1
+}
+
+precondition: ( MACed(x) )
+open(x:string) { _open(x,1,2,"1"); _open(x,2,1,"2") }
+
+_sum(z:string, x:string, y:string, i:cid){
+    m[z++"s"]@i := (m[x++"s"] + m[y++"s"])@i;
+    m[z++"m"]@i := (m[x++"m"] + m[y++"m"])@i;
+    m[z++"k"]@i := (m[x++"k"] + m[y++"k"])@i
+}
+
+precondition: ( MACed(x) AND MACed(y) )
+sum(z:string, x:string, y:string){ _sum(z,x,y,1); _sum(z,x,y,2) }
+postcondition: ( MACed(z) AND RECON(z) == RECON(x) + RECON(y) )
+
+_mult(z : string, y : string, i : cid) {
+   m[z++"s"]@i := (m[z++"d"] * m[y++"s"] + -m[z++"e"] * m[z++"as"] + m[z++"cs"])@i;
+   m[z++"m"]@i := (m[z++"d"] * m[y++"m"] + -m[z++"e"] * m[z++"am"] + m[z++"cm"])@i;
+   m[z++"k"]@i := (m[z++"d"] * m[y++"k"] + -m[z++"e"] * m[z++"ak"] + m[z++"ck"])@i
+}
+
+precondition: ( MACed(x) AND MACed(y) AND
+                MACed(z++"a") AND MACed(z++"b") AND MACed(z++"c") AND BEAVERTRIP(z) )
+mult(z : string, x : string, y : string) {
+    secopen(z++"d", x, z++"a"); secopen(z++"e", y, z++"b"); _mult(z,y,1); _mult(z,y,2);
+}
+postcondition: ( RECON(z) == RECON(x) * RECON(y) )
+
+\end{verbatimtab}
+\end{fpfig}
+
 \subsection{Confidentiality Examples}
 
-\input{gmw}
+%\input{gmw}
 
@@
 
-\input{bdoz}
+% \input{bdoz}
 
@@
 are verified by type checking.
-
-%%%%%%%%%%%%%%%%%%
-%%% ARCHIVAL BELOW
-%%%%%%%%%%%%%%%%%%
-\begin{comment}
-\begin{verbatimtab}
-andtableygc(g,x,y)
-{
-   let table = (~r[g],~r[g],~r[g],r[g])
-   in permute4(r[x],r[y],table)
-}
-
-m[x]@1 := s2(s[x],r[x],~r[x])@2;
-m[x]@1 as s[x]@2 xor r[x]@2;
-
-// m[x]@1 : { c(r[x]@2, { s[x]@2 }) }
-
-m[y]@1 := OT(s[y]@1,r[y],~r[y])@2;
-m[y]@1 as s[y]@1 xor r[y]@2;
-
-// m[y]@1 : { c(r[y]@2, { s[y]@1 }) }
-
-m[ag]@1 := OT4(m[x]@1, m[y]@1, andtable(ag,r[x],r[y]))@2;
-m[ag]@1 as  ~((r[x]@2 = m[x]@1) and (r[y]@2 = m[y]@1)) xor r[ag]@2;
-
-// m[ag]@1 : { c(r[ag]@2, {r[x]@2, r[y]@2, m[x]@1,  m[y]@1} }
-
-p[o] := OT2(m[ag]@1, perm2(r[ag],(false,true)))@2
-
-// p[o] : { c(r[ag]@2, {r[x]@2, r[y]@2,  m[x]@1,  m[y]@1}), r[ag]@2  }
-
-out@1 := p[o]@1
-
-// out@1 == s[x] and s[y]
-\end{verbatimtab}
-
-\begin{verbatimtab}
-    encodegmw(in, i1, i2) {
-      m[in]@i2 := (s[in] xor r[in])@i1;
-      m[in]@i1 := r[in]@i1
-    }
-
-    andtablegmw(x, y, z) {
-      let r11 = r[z] xor (m[x] xor true) and (m[y] xor true) in
-      let r10 = r[z] xor (m[x] xor true) and (m[y] xor false) in
-      let r01 = r[z] xor (m[x] xor false) and (m[y] xor true) in
-      let r00 = r[z] xor (m[x] xor false) and (m[y] xor false) in
-      { row1 = r11; row2 = r10; row3 = r01; row4 = r00 }
-    }
-
-    andgmw(z, x, y) {
-      let table = andtablegmw(x,y,z) in
-      m[z]@2 := OT4(m[x],m[y],table,2,1);
-      m[z]@2 as ~((m[x]@1 xor m[x]@2) and (m[y]@1 xor m[y]@2)) xor r[z]@1);
-      m[z]@1 := r[z]@1
-    }
-
-    // and gate correctness postcondition
-    {} andgmw { m[z]@1 xor m[z]@2 == (m[x]@1 xor m[x]@2) and (m[y]@1 xor m[y]@2) }
-
-    // and gate type
-    andgmw :
-     Pi z,x,y .
-     {}
-     { { r[z]@1 },
-       (m[z]@1 : { r[z]@1 }; m[z]@2 : {c(r[z]@1, { m[x]@1, m[x]@2, m[y]@1, m[y]@2 })} ),
-       m[z]@1 xor m[z]@2 == (m[x]@1 xor m[x]@2) and (m[y]@1 xor m[y]@2)}
-
-    xorgmw(z, x, y) {
-      m[z]@1 := (m[x] xor m[y])@1; m[z]@2 := (m[x] xor m[y])@2;
-    }
-
-    decodegmw(z) {
-      p["1"] := m[z]@1; p["2"] := m[z]@2;
-      out@1 := (p["1"] xor p["2"])@1;
-      out@2 := (p["1"] xor p["2"])@2
-    }
-
-    prot() {
-      encodegmw("x",2,1);
-      encodegmw("y",2,1);
-      encodegmw("z",1,2);
-      andgmw("g1","x","z");
-      xorgmw("g2","g1","y");
-      decodegmw("g2")
-    }
-
-    {} prot { out@1 == (s["x"]@1 and s["z"]@2) xor s["y"]@1 }
-\end{verbatimtab}
-
-\subsection{Integrity Examples}
-
-\begin{verbatimtab}
-  secopen(w1,w2,w3,i1,i2) {
-      pre(m[w1++"m"]@i2 == m[w1++"k"]@i1 + (m["delta"]@i1 * m[w1++"s"]@i2 /\
-          m[w1++"m"]@i2 == m[w1++"k"]@i1 + (m["delta"]@i1 * m[w1++"s"]@i2));
-      let locsum =  macsum(macshare(w1), macshare(w2)) in
-      m[w3++"s"]@i1 := (locsum.share)@i2;
-      m[w3++"m"]@i1 := (locsum.mac)@i2;
-      auth(m[w3++"s"],m[w3++"m"],mack(w1) + mack(w2),i1);
-      m[w3]@i1 := (m[w3++"s"] + (locsum.share))@i1
-  }
-
-  _open(x,i1,i2){
-    m[x++"exts"]@i1 := m[x++"s"]@i2;
-    m[x++"extm"]@i1 := m[x++"m"]@i2;
-    assert(m[x++"extm"] == m[x++"k"] + (m["delta"] * m[x++"exts"]));
-    m[x]@i1 := (m[x++"exts"] + m[x++"s"])@i2
-  }`
-
-  _sum(z, x, y,i1,i2) {
-      pre(m[x++"m"]@i2 == m[x++"k"]@i1 + (m["delta"]@i1 * m[x++"s"]@i2 /\
-          m[y++"m"]@i2 == m[y++"k"]@i1 + (m["delta"]@i1 * m[y++"s"]@i2));
-      m[z++"s"]@i2 := (m[x++"s"] + m[y++"s"])@i2;
-      m[z++"m"]@i2 := (m[x++"m"] + m[y++"m"])@i2;
-      m[z++"k"]@i1 := (m[x++"k"] + m[y++"k"])@i1;
-      post(m[z++"m"]@i2 == m[z++"k"]@i1 + (m["delta"]@i1 * m[z++"s"]@i2)
-  }
-
-  sum(z,x,y) { _sum(z,x,y,1,2);_sum(z,x,y,2,1) }
-
-  open(x) { _open(x,1,2); _open(x,2,1) }
-
-  sum("a","x","d");
-  open("d");
-  sum("b","y","e");
-  open("e");
-  let xys =
-      macsum(macctimes(macshare("b"), m["d"]),
-             macsum(macctimes(macshare("a"), m["e"]),
-                    macshare("c")))
-  let xyk = mack("b") * m["d"] + mack("a") * m["e"] + mack("c")
-
-  secopen("a","x","d",1,2);
-    secopen("a","x","d",2,1);
-    secopen("b","y","e",1,2);
-    secopen("b","y","e",2,1);
-    let xys =
-      macsum(macctimes(macshare("b"), m["d"]),
-             macsum(macctimes(macshare("a"), m["e"]),
-                    macshare("c")))
-    in
-    let xyk = mack("b") * m["d"] + mack("d") * m["d"] + mack("c")               
-    in
-    secreveal(xys,xyk,"1",1,2);
-    secreveal(maccsum(xys,m["d"] * m["e"]),
-              xyk - m["d"] * m["e"],
-              "2",2,1);
-    out@1 := (p[1] + p[2])@1;
-    out@2 := (p[1] + p[2])@2;
-\end{verbatimtab}
-
-\end{comment}

### 2025-11-14 toplas26/frontmatter.tex
checkpoint

--- 2024-10-14 esop25/frontmatter.tex
@@
-\titlerunning{Security Types for Low-Level MPC}
-
 \title{SMT-Boosted Security Types for Low-Level MPC}
 
-%\author{
-%Christian Skalka
-%\and
-%Joe Near}
+\author{Christian Skalka}
+\affiliation{%
+  \institution{University of Vermont}
+  \city{Burlington, Vermont}
+  \country{USA}}
+\email{ceskalka@uvm.edu}
 
-%\institute{
-%University of Vermont, Burlington VT 05405, USA 
-%\email{ceskalka@uvm.edu,jnear@uvm.edu}
-%}
+\author{Joseph P.~Near}
+\affiliation{%
+  \institution{University of Vermont}
+  \city{Burlington, Vermont}
+  \country{USA}}
+\email{jnear@uvm.edu}
 
-\author{Author names withheld for double-blind reviewing.}
-\institute{}
+\author{Steven Baldasty}
+\affiliation{%
+  \institution{University of Vermont}
+  \city{Burlington, Vermont}
+  \country{USA}}
+\email{sbaldast@uvm.edu}
 
-\newcommand{\theabstract}{
+\author{Yeonhee Yeh}
+\affiliation{%
+  \institution{University of Vermont}
+  \city{Burlington, Vermont}
+  \country{USA}}
+\email{yyeh@uvm.edu}
+
 \begin{abstract}
@@
   finite fields, which supports precise and efficient analysis. Our
-  approach is automated, compositional, scalable, and generalizes to arbitrary
-  prime fields for data and key sizes.
-\end{abstract} }
+  approach is automated, compositional, scalable, and generalizes to
+  arbitrary prime fields.
+\end{abstract}

### 2025-11-14 toplas26/examples.tex
checkpoint

@@
 In Section \ref{section-smt} we gave the example of 3-party additive
-sharing for any field $\fieldp{p}$. In Sections \ref{section-ipj} and
-\ref{section-metalangty} we discussed the example of YGC input encoding.
-Now we consider examples that illustrate the versatility of the
-$\TirName{Encode}$ typing rule for confidentiality, the use of
-postconditions for correctness, and the use of both pre-~and
-postconditions for integrity.
+sharing for any field $\fieldp{p}$, in Section \ref{section-sty} we
+discussed the example of YGC input encoding, and in Sections
+\ref{section-metalang} and \ref{section-metaty} we presented and
+discussed the example of gates in the GMW protocol for binary
+circuits. In this section we present an extended example of arithmetic
+circuits with multiplication gates defined using Beaver triples
+\cite{evans2018pragmatic}. We first define a compositional gate
+library in $\metaprot$ with correctness conditions defined using the
+Hoare logic developed in Section \ref{section-metalang}, along with a
+small example circuit.  Then we will extend the library with type
+annotations illustrating the use of share types developed in
+Sections \ref{section-sty} and \ref{section-metaty} to enforce
+confidentiality properties. Last, we will extend the base library
+with additional pre and postcondition annotations to enforce
+integrity properties supported by BDOZ/SPDZ style MACs
+\cite{SPDZ1,SPDZ2,BDOZ,10.1007/978-3-030-68869-1_3}.
 
@@
 Satisfiability Modulo Fields theory in cvc5 on an Apple M1 processor,
-we have verified correctness of additive sharing protocols in fields
-up to size $\fieldp{2^{31} - 1}$ (an approximation of 32-bit integers)
-which takes $\sim$7ms to compute for 3 parties and not significantly more
-than for, e.g., $\fieldp{2}$. We have also verified the semi-homomorphic
-encryption properties leveraged in the BDOZ circuit library functions
-(Figures \ref{fig-bdozsum} and \ref{fig-bdozmult}) in fields up to
-size $\fieldp{2^{31} - 1}$ which take $\sim$8ms for the property used for
-the postcondition of the $\ttt{mult}$ gate. This is categorically
-better performance than can be achieved using previous brute-force
-methods in $\fieldp{2}$ \cite{skalka-near-ppdp24}.
+we have verified correctness of these arithmetic gate libraries and
+circuit in fields up to size $\fieldp{2^{31} - 1}$ (an approximation
+of 32-bit integers) which takes $\sim$7ms to compute for 3 parties and
+not significantly more than for, e.g., $\fieldp{2}$. We have also
+verified the semi-homomorphic encryption properties leveraged in the
+BDOZ circuit library functions in fields up to size $\fieldp{2^{31} -
+  1}$ which take $\sim$8ms for the property used for the postcondition
+of the $\ttt{mult}$ gate. This is categorically better performance
+than can be achieved using previous brute-force methods in
+$\fieldp{2}$ \cite{skalka-near-ppdp24}.
+
+\subsection{Arithmetic Circuits with Beaver Triples}
 
@@
 
-_open(x:string, i1:cid, i2:cid, n:string){ p[x++"n"] := m[x]@i2; m[x++"p"]@i1 = p[x++"n"] + m[x]@i1 }
+_open(x:string, i1:cid, i2:cid, n:string){ p[x++"n"] := m[x]@i2; m[x++"p"]@i1 := p[x++"n"] + m[x]@i1 }
 
@@
     assert(p[x++"mn"] = m[x++"k"] + (m["delta"] * p[x++"sn"]))@i1;
-    m[x++"p"]@i1 = p[x++"sn"] + m[x]@i1
+    m[x++"p"]@i1 := p[x++"sn"] + m[x]@i1
 }
@@
 postcondition: ( RECON(z) == RECON(x) * RECON(y) )
-
 \end{verbatimtab}

### 2025-11-14 toplas26/examples.tex
checkpoint

@@
+\begin{fpfig}[t]{2-party gate library with Beaver Triple multiplication, circuit definition, and compositional correctness conditions.}{fig-beaver}
+\footnotesize
+\begin{verbatimtab}
+IDEALF() { (s["x"]@1 + s["y"]@2) + (s["i"]@1 * s["j"]@2) }
+
+RECON(x) { (m[x]@1 + m[x]@2) }
+
+BEAVERTRIP(x) {
+  m[x++"c"]@3 == r[x++"a"]@3 * r[x++"b"]@3 AND
+  m[x++"a"]@1 + m[x++"a"]@2 == r[x++"a"]@3 AND
+  m[x++"b"]@1 + m[x++"b"]@2 == r[x++"b"]@3 AND
+  m[x++"c"]@1 + m[x++"c"]@2 == m[x++"c"]@3 
+}
+
+genshares(x:string, i1:cid, i2:cid) { m[x]@i2 := (s[x] + -r[x])@i1; m[x]@i1 := r[x]@i1 }
+
+_open(x:string, i1:cid, i2:cid, n:string){ p[x++"n"] := m[x]@i2; m[x++"p"]@i1 := p[x++"n"] + m[x]@i1 }
+
+open(x:string) { _open(x,1,2,"1"); _open(x,2,1,"2") }
+
+secopen(d:string;x:string,a:string) { sum(d, x, a); open(d) }
+
+_sum(z:string, x:string, y:string, i:cid) { m[z]@i := (m[x] + m[y])@i }
+
+sum(z:string, x:string, y:string){ _sum(z,x,y,1); _sum(z,x,y,2) }
+postcondition: ( RECON(z) == RECON(x) + RECON(y) )
+
+_mult(z : string, y : string, i : cid) {
+    m[z]@i := (m[z++"d"] * m[y] + -m[z++"e"] * m[z++"a"] + m[z++"c"])@i
+}
+
+precondition: ( BEAVERTRIP(z) )
+mult(z : string, x : string, y : string) {
+  secopen(z++"d", x, z++"a"); secopen(z++"e", y, z++"b"); _mult(z,y,1); _mult(z,y,2)
+}
+postcondition: ( RECON(z) == RECON(x) * RECON(y) )
+
+precondition: ( BEAVERTRIP("k") ) 
+main() {
+    genshares("x", 1, 2); genshares("y", 2, 1); genshares("i", 1, 2); genshares("j", 2, 1);
+    sum("z","x","y"); mult("k","i","j"); sum("out","z","k");
+    open("out");
+    out@1 = m["out"]@1; out@2 =  m["out"]@2
+
+}
+postcondition: ( out@1 == IDEALF() AND out@2 == IDEALF() )
+\end{verbatimtab}
+\end{fpfig}
+
 \section{Extended Examples}
@@
 \subsection{Arithmetic Circuits with Beaver Triples}
+\label{section-beaver}
 
-\begin{fpfig}[t]{2-party gate library with Beaver Triple multiplication, circuit definition, and compositional correctness conditions.}{fig-beaver}
-\footnotesize
-\begin{verbatimtab}
-IDEALF() { (s["x"]@1 + s["y"]@2) + (s["i"]@1 * s["j"]@2) }
+In Figure \ref{fig-beaver} we define a 2-party arithmetic circuit with
+4 input gates, 3 internal gates, and 1 output gate. The ideal functionality
+is defined as $\ttt{IDEALF}$. As in GMW or BDOZ, circuit wire values
+are represented as additive secret shares between two clients we call
+$\ttt{1}$ and $\ttt{2}$. As in the GMW example discussed in Section
+\ref{section-metaty}, we use string idenfiers $g$ to distinguish
+gates, and gate output wire shares are defined as identifiers
+$\mx{g}{1}$ and $\mx{g}{2}$. The function $\ttt{RECON}$
+represents share reconstruction \emph{in constraints} (not the
+protocols themselves). Input gates are
+implemented using the $\ttt{genshares}$ function as described in
+Section \ref{section-sty}. The $\ttt{open}$ function
+implements a federation reveal of a shared value through
+share broadcast and reconstruction.
 
-RECON(x) { (m[x]@1 + m[x]@2) }
+A \ttt{sum} gate can be defined non-interactively using the
+homomorphic addition. The correctness \ttt{postcondition} is simple
+and perhaps unecessary to accelerate circuit verification, but is
+appealing declaratively.  The multiplication gate \ttt{mult} is more
+complex. We use the common approach of \emph{Beaver Triples}
+\cite{evans2018pragmatic}. Recall that Beaver triples are values
+$a,b,c$ with $a$ and $b$ chosen randomly and $c = a * b$, unique per
+multiplication gate, that are secret-shared with clients during a
+pre-processing phase by an assumed oracular client $\ttt{3}$.
+The uniqueness requirement is necessary because
+$a$ and $b$ are random values that serve as OTP masks for secure
+openings, implemented as \ttt{secopen}. In our encoding we follow the
+convention that each gate output identifier distinguishes the Beaver
+triple, so for example the share of the $a$ value for a gate
+$\ttt{"g"}$ is identified by $\ttt{"ga"}$, etc. The correctness
+property of the multiplication gate, expressed as a
+\ttt{postcondition}, relies on the $\ttt{BEAVERTRIP}$
+precondition. This compositional correctness property can then be used
+in circuit construction, without the need for ``re-proof'' in our
+Hoare logic inference.
 
-BEAVERTRIP(x) {
-  m[x++"c"]@3 == r[x++"a"]@3 * r[x++"b"]@3 AND
-  m[x++"a"]@1 + m[x++"a"]@2 == r[x++"a"]@3 AND
-  m[x++"b"]@1 + m[x++"b"]@2 == r[x++"b"]@3 AND
-  m[x++"c"]@1 + m[x++"c"]@2 == m[x++"c"]@3 
-}
+The circuit itself is constructed in \ttt{main} using these
+gates. This is a small example but arbitrarily large circuits
+can be constructed with this library. The precondition of
+\ttt{main} must require distinct Beaver triples for each
+multiplication gate in the circuit. The postcondition allows
+verification of circuit correctness-- that the protocol
+output is equal to the ideal functionality $\ttt{IDEALF}$.
 
-sshare(x:string, i1:cid, i2:cid) { m[x]@i2 := (s[x] + -r[x])@i1; m[x]@i1 := r[x]@i1 }
-
-_open(x:string, i1:cid, i2:cid, n:string){ p[x++"n"] := m[x]@i2; m[x++"p"]@i1 := p[x++"n"] + m[x]@i1 }
-
-open(x:string) { _open(x,1,2,"1"); _open(x,2,1,"2") }
-
-secopen(d:string;x:string,a:string) { sum(d, x, a); open(d) }
-
-_sum(z:string, x:string, y:string, i:cid) { m[z]@i := (m[x] + m[y])@i }
-
-sum(z:string, x:string, y:string){ _sum(z,x,y,1); _sum(z,x,y,2) }
-postcondition: ( RECON(z) == RECON(x) + RECON(y) )
-
-_mult(z : string, y : string, i : cid) {
-    m[z]@i := (m[z++"d"] * m[y] + -m[z++"e"] * m[z++"a"] + m[z++"c"])@i
-}
-
-precondition: ( BEAVERTRIP(z) )
-mult(z : string, x : string, y : string) {
-  secopen(z++"d", x, z++"a"); secopen(z++"e", y, z++"b"); _mult(z,y,1); _mult(z,y,2)
-}
-postcondition: ( RECON(z) == RECON(x) * RECON(y) )
-
-precondition: ( BEAVERTRIP("k") ) 
-main() {
-    sshare("x", 1, 2); sshare("y", 2, 1); sshare("i", 1, 2); sshare("j", 2, 1);
-    sum("z","x","y"); mult("k","i","j"); sum("out","z","k");
-    open("out")
-}
-postcondition: ( m["out"]@1 == IDEALF() )
-\end{verbatimtab}
-\end{fpfig}
+\subsection{Share Type Annotations.}
 
@@
 \begin{verbatimtab}
-sshare(x:string, i1:cid, i2:cid) {
+genshares(x:string, i1:cid, i2:cid) {
    m[x++"s"]@i1 := (s[x] + -r[x])@i2; m[x++"s"]@i2 := r[x]@i2;
@@
     sum(d, x, a); open(d);
-    m[d++"p"]@1 as RECON(x) + r[x++a]@o;
+    m[d++"p"]@1 as RECON(x) + r[x++a]@3;
     <m[d++"p"]@1> : public;
@@
 \end{fpfig}
+
+In Figure \ref{fig-beaver-types} we extend our previous gate definitions with
+share type annotations to enforce confidentiality. The core idea is that
+wire values must be \ttt{share}s. The \ttt{genshares}
+function includes an annotation specifying that all input gates establish
+the $\ttt{shares}$ invariant for wire values. In the \ttt{sum} function,
+the \ttt{share} type preconditions for input wires are specified. The
+output wire shares can be declared as \ttt{share}s without any coercion
+needed, since given preconditions our typing algorithm ($\Vdash$) is
+able to check the use of homomorphic addition.
+
+We add coercions and typing assertions to the $\ttt{secopen}$ function
+to specify properties of secure openings. After noninteractively
+summing a wire value $\ttt{x}$ and a random Beaver triple value
+$\ttt{x++a}$, the federation can publicly open the value since it is
+masked by the random value. This is expressed using the coercion
+$\ttt{as RECON(x) +  r[x++a]@3}$.
+
+\subsubsection{Malicious-Secure Arithmetic Circuits}
 
@@
 
-\subsection{Confidentiality Examples}
-
-%\input{gmw}
-
-In the 2-party GMW protocol \cite{evans2018pragmatic}, another garbled
-binary circuit protocol, values are encrypted in a manner similar to
-BDOZ as described in Section \ref{section-ipj}. In our definition of
-GMW we use the convention that shared values $\macgv{\mesg{w}}$ are
-identified by strings $w$ and encoded as shares $\mx{w}{1}$ and
-$\mx{w}{2}$.  As for YGC, ciphertypes reflect the confidentiality of
-GMW input encodings as defined in the $\ttt{encodegmw}$ function in
-Figure \ref{fig-gmw}. No programmer annotation is needed given the
-syntax of the function body. More interestingly, the $\ttt{andgmw}$
-gate definition shows how a programmer hint can express the relevant
-confidentiality property of the output share $\mx{z}{2}$ on client 2,
-using the $\ttt{as}$ casting. The non-trivial equivalence can be
-verified by SMT once during verification, and subsequently
-confidentiality is expressed in the dependent type of the function (as
-for $\ttt{encode}$).
-
-Note also that the $\ttt{andgmw}$ function is decorated with a
-postcondition that expresses the correctness property of the
-function. Although not strictly necessary for confidentiality, this
-annotation can at least help eliminate bugs, and also can be used
-compositionally for whole-program correctness properties. As for
-confidentiality hints, this postcondition needs to be verified only
-once at the point of definition and subsequently is guaranteed to hold
-for any application.
-
-\subsection{Integrity Examples}
-
-% \input{bdoz}
-
-Returning to the example of malicious-secure 2-party BDOZ arithmetic
-circuits begun in Section \ref{section-ipj}, in Figure \ref{fig-bdozsum} we
-define $\ttt{sum}$ and $\ttt{open}$ functions. The latter implements
-``secure opening''-- each party sends it's local share of a
-global value $\macgv{\mesg{w}}$ along with its MAC to the other party,
-which is authenticated via the $\macbdoz{w}$ check (in \ttt{\_open}),
-and then each party reconstructs $\macgv{\mesg{w}}$. The preconditions
-of $\ttt{open}$ specify the authentication requirements.
-
-The $\ttt{sum}$ function implements an addition gate. This is
-non-interactive-- each party just sums its local shares
-of the two values. The pre- and postcondition annotations
-on $\ttt{sum}$ express the additive homomorphism associated
-with this encryption scheme-- the sum of MACs of the input
-shares is a valid MAC for the output share on each client,
-which can be checked using the sum of keys of the input shares.
-
 In BDOZ a pre-processing phase is assumed where initial input secrets
@@
 
-A common approach to implementing multiplication gates in a BDOZ
-setting is to use \emph{Beaver Triples}. Recall that Beaver triples
-are values $a,b,c$ with $a$ and $b$ chosen randomly and $c = a * b$,
-unique per multiplication gate, that are secret shared with clients
-during pre-processing.  In our encoding we assume the additional
-convention that each gate output identifier distinguishes the Beaver
-triple, so for example the share of the $a$ value for a gate
-$\ttt{"g1"}$ is identified by $\ttt{"g1as"}$, etc.
-
-%, and $\eqspre$
-%subsumes the following constraints for the $a,b,c$ values of gate
-%$\ttt{"g1"}$.  {\footnotesize$$
-%\begin{array}{l}
-%\mx{\ttt{"g1as"}}{1} \eop \rx{\ttt{"g1as"}}{1}\ \wedge 
-%\mx{\ttt{"g1bs"}}{1} \eop \rx{\ttt{"g1bs"}}{1}\ \wedge\\
-%\mx{\ttt{"g1as"}}{2} \eop \rx{\ttt{"g1as"}}{2}\ \wedge 
-%\mx{\ttt{"g1bs"}}{2} \eop \rx{\ttt{"g1bs"}}{2}\ \wedge\\
-%\mx{\ttt{"g1cs"}}{1} \eop \\
-%\qquad ((\mx{\ttt{"g1as"}}{1} \fplus \mx{\ttt{"g1bs"}}{2})\ \ftimes\\
-%\qquad\phantom{(}(\mx{\ttt{"g1bs"}}{1} \fplus \mx{\ttt{"g1bs"}}{2})) \fminus \rx{\ttt{"g1cs"}}{2}\ \wedge\\
-%\mx{\ttt{"g1cs"}}{2} \eop \rx{\ttt{"g1cs"}}{2}
-%\end{array}
-%$$}
-
 The definition of the $\ttt{mult}$ gate is presented in Figure

### 2025-11-14 toplas26/sty.tex
checkpoint

@@
 as $\phi_1 + \phi_2$.\cnote{Comment on generation and
-  reconstruction in general somewhere.} A share term $\pubx{\phi}$
-denotes a publicly shared value.
+  reconstruction in general somewhere before this section?.} A
+share term $\pubx{\phi}$ denotes a publicly shared value.
 $$
@@
 \begin{theorem}
+  If $\prog,\Gamma,R : {OK}$ and $\Gamma$ is closed
+  for $\prog$, then $\prog$ satisfies noninterference modulo output. 
+\end{theorem}
+
+\begin{theorem}
   If $\toeq{\prog} \vdash \varnothing,\varnothing \stredx \Gamma,R$ and $\Gamma$ is closed

### 2025-11-14 toplas26/examples.tex
checkpoint

@@
-\begin{fpfig}[t]{2-party gate library with Beaver Triple multiplication, circuit definition, and compositional correctness conditions.}{fig-beaver}
-\footnotesize
-\begin{verbatimtab}
-IDEALF() { (s["x"]@1 + s["y"]@2) + (s["i"]@1 * s["j"]@2) }
-
-RECON(x) { (m[x]@1 + m[x]@2) }
-
-BEAVERTRIP(x) {
-  m[x++"c"]@3 == r[x++"a"]@3 * r[x++"b"]@3 AND
-  m[x++"a"]@1 + m[x++"a"]@2 == r[x++"a"]@3 AND
-  m[x++"b"]@1 + m[x++"b"]@2 == r[x++"b"]@3 AND
-  m[x++"c"]@1 + m[x++"c"]@2 == m[x++"c"]@3 
-}
-
-genshares(x:string, i1:cid, i2:cid) { m[x]@i2 := (s[x] + -r[x])@i1; m[x]@i1 := r[x]@i1 }
-
-_open(x:string, i1:cid, i2:cid, n:string){ p[x++"n"] := m[x]@i2; m[x++"p"]@i1 := p[x++"n"] + m[x]@i1 }
-
-open(x:string) { _open(x,1,2,"1"); _open(x,2,1,"2") }
-
-secopen(d:string;x:string,a:string) { sum(d, x, a); open(d) }
-
-_sum(z:string, x:string, y:string, i:cid) { m[z]@i := (m[x] + m[y])@i }
-
-sum(z:string, x:string, y:string){ _sum(z,x,y,1); _sum(z,x,y,2) }
-postcondition: ( RECON(z) == RECON(x) + RECON(y) )
-
-_mult(z : string, y : string, i : cid) {
-    m[z]@i := (m[z++"d"] * m[y] + -m[z++"e"] * m[z++"a"] + m[z++"c"])@i
-}
-
-precondition: ( BEAVERTRIP(z) )
-mult(z : string, x : string, y : string) {
-  secopen(z++"d", x, z++"a"); secopen(z++"e", y, z++"b"); _mult(z,y,1); _mult(z,y,2)
-}
-postcondition: ( RECON(z) == RECON(x) * RECON(y) )
-
-precondition: ( BEAVERTRIP("k") ) 
-main() {
-    genshares("x", 1, 2); genshares("y", 2, 1); genshares("i", 1, 2); genshares("j", 2, 1);
-    sum("z","x","y"); mult("k","i","j"); sum("out","z","k");
-    open("out");
-    out@1 = m["out"]@1; out@2 =  m["out"]@2
-
-}
-postcondition: ( out@1 == IDEALF() AND out@2 == IDEALF() )
-\end{verbatimtab}
-\end{fpfig}
+\input{beaver}
 
@@
 
-\begin{fpfig}[t]{2-party gate library with Beaver Triple multiplication extended with share type annotations.}{fig-beaver-types}
-\footnotesize
-\begin{verbatimtab}
-genshares(x:string, i1:cid, i2:cid) {
-   m[x++"s"]@i1 := (s[x] + -r[x])@i2; m[x++"s"]@i2 := r[x]@i2;
-   [m[x++"s"]@i1,m[x++"s"]@i2] : shares
-}
-
-([m[x]@i1, m[x]@i2] : shares)
-open(x:string) {
-  _open(x,1,2,"1"); _open(x,2,1,"2");
-  [p[x++"1"],p[x++"2"]] : shares;
-  m[d++"p"]@2 as m[d++"p"]@1  
-}
-
-precondition: ([m[x]@i1, m[x]@i2] : shares,
-	       [m[a]@i1, m[a]@i2] : shares)
-secopen(d:string;x:string,a:string) {
-    sum(d, x, a); open(d);
-    m[d++"p"]@1 as RECON(x) + r[x++a]@3;
-    <m[d++"p"]@1> : public;
-    <m[d++"p"]@2> : public
-}
-
-precondition: ([m[x]@1,m[x]@2] : shares;
-               [m[y]@1,m[y]@2] : shares)
-sum(z:string, x:string, y:string) {
-    _sum(z,x,y,1); _sum(z,x,y,2);
-    [m[z]@1,m[z]@2] : shares
-}
-postcondition: ( RECON(z) == RECON(x) + RECON(y) )
-
-precondition: ( [m[x]@1,m[x]@2] : shares;
-                [m[y]@1,m[y]@2] : shares;
-	        [m[x++"c"]@1,m[x++"c"]@2] : shares ),
-	       ( BEAVERTRIP(z) )
-mult(z : string, x : string, y : string) {
-    secopen(z++"d", x, z++"a"); secopen(z++"e", y, z++"b"); _mult(z,y,1); _mult(z,y,2);
-    [m[z]@i1,m[z]@i2] : shares
-}
-postcondition: ( RECON(z) == RECON(x) * RECON(y)) )
-\end{verbatimtab}
-\end{fpfig}
-
 In Figure \ref{fig-beaver-types} we extend our previous gate definitions with
@@
 \subsubsection{Malicious-Secure Arithmetic Circuits}
-
-\begin{fpfig}[t]{2-party gate library with Beaver Triple multiplication extended with BDOZ MACs.}{fig-beaver-bdoz}
-\footnotesize
-\begin{verbatimtab}
-MACOK(x, i1, i2) { m[x++"m"]@i2 == m[x++"k"]@i1 + (m["delta"]@i1 * m[x++"s"]@i2) }
-
-MACed(x) { MACOK(x, 1, 2) AND MACOK(x, 2, 1) }
-
-_open(x:string, i1:cid, i2:cid, n:string){
-    p[x++"sn"] := m[x++"s"]@i2;
-    p[x++"mn"] := m[x++"m"]@i2;
-    assert(p[x++"mn"] = m[x++"k"] + (m["delta"] * p[x++"sn"]))@i1;
-    m[x++"p"]@i1 := p[x++"sn"] + m[x]@i1
-}
-
-precondition: ( MACed(x) )
-open(x:string) { _open(x,1,2,"1"); _open(x,2,1,"2") }
-
-_sum(z:string, x:string, y:string, i:cid){
-    m[z++"s"]@i := (m[x++"s"] + m[y++"s"])@i;
-    m[z++"m"]@i := (m[x++"m"] + m[y++"m"])@i;
-    m[z++"k"]@i := (m[x++"k"] + m[y++"k"])@i
-}
-
-precondition: ( MACed(x) AND MACed(y) )
-sum(z:string, x:string, y:string){ _sum(z,x,y,1); _sum(z,x,y,2) }
-postcondition: ( MACed(z) AND RECON(z) == RECON(x) + RECON(y) )
-
-_mult(z : string, y : string, i : cid) {
-   m[z++"s"]@i := (m[z++"d"] * m[y++"s"] + -m[z++"e"] * m[z++"as"] + m[z++"cs"])@i;
-   m[z++"m"]@i := (m[z++"d"] * m[y++"m"] + -m[z++"e"] * m[z++"am"] + m[z++"cm"])@i;
-   m[z++"k"]@i := (m[z++"d"] * m[y++"k"] + -m[z++"e"] * m[z++"ak"] + m[z++"ck"])@i
-}
-
-precondition: ( MACed(x) AND MACed(y) AND
-                MACed(z++"a") AND MACed(z++"b") AND MACed(z++"c") AND BEAVERTRIP(z) )
-mult(z : string, x : string, y : string) {
-    secopen(z++"d", x, z++"a"); secopen(z++"e", y, z++"b"); _mult(z,y,1); _mult(z,y,2);
-}
-postcondition: ( RECON(z) == RECON(x) * RECON(y) )
-\end{verbatimtab}
-\end{fpfig}
 

### 2025-11-14 toplas26/examples.tex
checkpoint

@@
 masked by the random value. This is expressed using the coercion
-$\ttt{as RECON(x) +  r[x++a]@3}$.
+$\ttt{as RECON(x) +  r[x++a]@3}$. The coercion then allows the
+opened value to be declared $\ttt{public}$.
+
+In the \ttt{mult} function, each client $\ttt{i}$ symmetrically
+computes:
+$$
+\ttt{((m[z++"d"] * m[y]) + (-m[z++"e"] * m[z++"a"]) + m[z++"c"])@i}
+$$
+Since $\ttt{m[z++"d"]@i : public}$ and $\ttt{m[z++"e"]@i : public}$
+are both established by $\ttt{secopen}$, and
+$\ttt{m[y]@i}$, $\ttt{m[z++"a"]@i}$, and $\ttt{m[z++"c"]@i}$ are
+all declared to be shares in preconditions, we have that each
+of $\ttt{m[z++"d"] * m[y]@i}$ and $\ttt{-m[z++"e"] * m[z++"a"]}$
+are shares, so the gate outputs are shares, and this will
+be successfully checked by syntax-directed $\Vdash$ type checking that
+results from the declaration $\ttt{[m[z]@i1,m[z]@i2] : shares}$.
 

### 2025-11-14 toplas26/examples.tex
checkpoint

--- 2025-11-14 toplas26/examples.tex
@@
 masked by the random value. This is expressed using the coercion
-$\ttt{as RECON(x) +  r[x++a]@3}$.
+$\ttt{as RECON(x) +  r[x++a]@3}$. The coercion then allows the
+opened value to be declared $\ttt{public}$.
+
+In the \ttt{mult} function, each client $\ttt{i}$ symmetrically
+computes:
+$$
+\ttt{((m[z++"d"] * m[y]) + (-m[z++"e"] * m[z++"a"]) + m[z++"c"])@i}
+$$
+Since $\ttt{m[z++"d"]@i : public}$ and $\ttt{m[z++"e"]@i : public}$
+are both established by $\ttt{secopen}$, and
+$\ttt{m[y]@i}$, $\ttt{m[z++"a"]@i}$, and $\ttt{m[z++"c"]@i}$ are
+all declared to be shares in preconditions, we have that each
+of the shared computations $\ttt{(m[z++"d"] * m[y])@i}$ and $\ttt{(-m[z++"e"] * m[z++"a"])@i}$
+are shares, so the gate outputs are shares, and this will
+be successfully checked by syntax-directed $\Vdash$ type checking that
+results from the declaration $\ttt{[m[z]@i1,m[z]@i2] : shares}$.
 
@@
 
-In BDOZ a pre-processing phase is assumed where initial input secrets
-are shared along with their associated MACs and keys. This can be
-expressed in $\eqspre$ for input secrets $\sx{\ttt{"x"}}{1}$ and
-$\sx{\ttt{"y"}}{2}$, for example, which subsume the following
-constraints on shares:
-{\footnotesize$$
-\begin{array}{l}
-\mx{\ttt{"xs"}}{2} \eop \sx{\ttt{"x"}}{1} \fminus \rx{\ttt{"x"}}{1} \wedge 
-\mx{\ttt{"xs"}}{1} \eop \rx{\ttt{"x"}}{1} \wedge \\
-\mx{\ttt{"ys"}}{1} \eop \sx{\ttt{"y"}}{2} \fminus \rx{\ttt{"y"}}{2} \wedge 
-\mx{\ttt{"ys"}}{2} \eop \rx{\ttt{"y"}}{2} 
-\end{array}
-$$}
-and the following constraint on keys and MACs for authentication
-of $\sx{\ttt{"x"}}{1}$ (and similarly for $\sx{\ttt{"y"}}{2}$):
-{\footnotesize$$
-\begin{array}{l}
-\mx{\ttt{"delta"}}{1} \eop \rx{\ttt{"delta"}}{1} \wedge
-\mx{\ttt{"xk"}}{1} \eop \rx{\ttt{"xk"}}{1} \wedge\\
-\mx{\ttt{"xm"}}{2} \eop \mx{\ttt{"xk"}}{1} \fplus (\mx{\ttt{"delta"}}{1} * \mx{\ttt{"xs"}}{2})
-\end{array}
-$$}
-Given these global preconditions, a malicious secure opening of $\sx{\ttt{"x"}}{1} +
-\sx{\ttt{"y"}}{2}$ would be obtained as
-$\ttt{sum}(\ttt{"z"},\ttt{"x"},\ttt{"y"}); \ttt{open}(\ttt{"z"})$,
-which type checks.
-
-The definition of the $\ttt{mult}$ gate is presented in Figure
-\ref{fig-bdozmult},
-%where we present just client 1's side of the
-%protocol $\ttt{\_mult1}$ (client 2's side is nearly symmetric but with
-%a small variation-- refer to
-%\cite{evans2018pragmatic,10.1007/978-3-030-68869-1_3} for more
-%details).
-%We could call this function on input secrets
-%$\sx{\ttt{"x"}}{1}$ and $\sx{\ttt{"y"}}{2}$ in gate $\ttt{"g1"}$ as
-%$\ttt{mult}(\ttt{"g1"},\ttt{"x"},\ttt{"y"})$, for example, or embed
-%this gate internally in a circuit.
-%
-As for $\ttt{sum}$, the preconditions of $\ttt{mult}$ express
-the expected MAC properties of input shares, as well as the
-expected Beaver Triple property, and its postcondition
-fexpresses the semi-homomorphic encryption property of
-the resulting share, MAC, and key after gate execution-- specifically,
-it preserves the BDOZ authentication property. Finally, the
-postcondition of $\ttt{mult}$ expresses the
-correctness property of the multiplication gate. In any case,
-integrity of any circuit constructed from the $\ttt{sum}$ and
-$\ttt{mult}$ library functions will require little SMT overhead
-due to compositional verification, once their pre- and postconditions
-are verified by type checking.
+Our Definition \ref{def-integrity} establishes a general semantics of
+integrity, while the $\minicat$ integrity type system presented in
+Section \ref{section-ipj} focuses on the use of BDOZ/SPDZ-style
+MACs. A main concern here is the encorcement of MAC checking which is
+an algebraic property. Here we show how pre and postconditions can be
+used to enforce these properties in our arithmetic circuit library, in
+Figure \ref{fig-beaver-bdoz}. Note that this code is an extension of the
+library presented in Figure \ref{fig-beaver} and does not incorporate
+the type annotations in Figure \ref{fig-beaver-types}, but this is
+just for clarity of the presentation and these features could be combined.

### 2025-11-14 toplas26/examples.tex
checkpoint

@@
 discussed the example of gates in the GMW protocol for binary
-circuits. In this section we present an extended example of arithmetic
-circuits with multiplication gates defined using Beaver triples
+circuits. In this section we present an extended example of 2-party
+arithmetic circuits with multiplication gates defined using Beaver triples
 \cite{evans2018pragmatic}. We first define a compositional gate
@@
 just for clarity of the presentation and these features could be combined.
+
+In this code we use a convention where shared values $\ttt{x}$ are
+represented as a 3-tuple of values, MAC, and keys, where the value and
+MAC of client \ttt{1} are $\ttt{m[x++"s"]@1}$ and $\ttt{m[x++"m"]@1}$
+respectively, and the corresponding key is $\ttt{m[x++"k"]@2}$ is held
+by client \ttt{2}. The correctness condition of this 3-tuple is
+$\ttt{MACOK}(x,2,1)$. The same is true symmetrically, so $\ttt{MACed}$
+establishes both sides of the correctness condition.  In general,
+these are used in gate preconditions to assume that the MAC conditions
+hold for input shares, and output shares are $\ttt{MACed}$. The
+homomorphic properties that preserve $\ttt{MACed}$ness, especially in
+\ttt{mult}, are perhaps even more complex than the algebraic
+correctness of gates, so it is very beneficial to verify them
+compositional especially for large circuits.
+
+Our code is written so that share MACs are checked for when they are
+\ttt{open}ed by both clients, specified by the appropriate
+$\ttt{assert}$ included in the protocol. This ensures malicious
+security of secure openings and outputs. As is standard in BDOZ/SPDZ,
+we assume that input values, MACs, and keys are distributed to clients
+by a pre-processing protocol.

### 2025-11-14 toplas26/metaty.tex
checkpoint

@@
+\section{$\metaprot$ Type Checking}
+
+Now we develop a compositional type system for $\metaprot$ based on 
+$\minicat$ typing and the automated Hoare logic developed in
+Section \ref{section-metalang}. We focus on share types for confidentiality
+which is more complex and general than our integrity type system.
+An algorithmic approach to the latter was developed previously for
+$\metaprot$ \cite{skalka-near-esop25}, and also the core element
+of the system-- BDOZ/SPDZ-style MAC checking-- can be treated as
+algebraic correctness conditions enforceable by Hoare triples
+as we show in Section \ref{section-examples}.
+
 \subsection{Share Type Checking with Annotations}
+\label{section-metaty}
 

### 2025-11-14 toplas26/sty.tex
checkpoint

@@
 \subsection{Share Type Soundness}
+\label{section-sty-sound}
 
@@
 \begin{theorem}
+  \label{theorem-sty}
   If $\prog,\Gamma,R : {OK}$ and $\Gamma$ is closed
@@
 \begin{theorem}
+  \label{theorem-sty}
   If $\toeq{\prog} \vdash \varnothing,\varnothing \stredx \Gamma,R$ and $\Gamma$ is closed

### 2025-11-17 toplas26/model.tex
checkpoint

--- 2025-01-26 esop25/model.tex
@@
   Given $\pmf$, the \emph{conditional distribution}
-  of $X_1$ given $X_2$ where $X_1 \cup X_2 \subseteq \dom(\pmf)$ and $X_1 \cap X_2 = \varnothing$
+  of $X_1$ given $X_2$ where $X_1 \cup X_2 = \dom(\pmf)$ 
   is denoted $\condd{\pmf}{X_1}{X_2}$ and defined as follows:
@@
       0 & \text{\ if\ } \margd{\pmf}{X_2}(\store_{X_2}) = 0\\
-      \margd{\pmf}{X_1 \cup X_2}(\store) / \margd{\pmf}{X_2}(\store_{X_2})\ & \text{\ o.w.}
+      \pmf(\store) / \margd{\pmf}{X_2}(\store_{X_2})\ & \text{\ o.w.}
      \end{array}
@@
 \end{definition}
-We also define some convenient syntactic sugarings. The first will allow us to
-compare marginal distributions under different realization conditions
-(as in, e.g., Definition \ref{definition-NIMO}), the others are standard.
+We also define some convenient syntactic sugaring for ``implicit
+marginalization'' given a realization.
+%The first will allow us to
+%compare marginal distributions under different realization conditions
+%(as in, e.g., Definition \ref{definition-NIMO}), the others are standard.
 \begin{definition}
-  Given $\pmf$, for all $\store_1 \in \mems(X_1)$ and $\store_2 \in \mems(X_2)$ define:
-  \begin{enumerate}
-  \item $\condd{\pmf}{X_1}{\store_2}(\store_1) \defeq \condd{\pmf}{X_1}{X_2}(\store_1 \uplus \store_2)$
-  \item $\pmf(\store_1)  \defeq \margd{\pmf}{X_1}(\store_1)$ 
-  \item $\pmf(\store_1|\store_2) \defeq \condd{\pmf}{X_1}{X_2}(\store_1 \uplus \store_2)$
-  \end{enumerate}
+  Given $\pmf$ and $X \subseteq \dom(\pmf)$, for all $\store \in \mems(X)$
+  define $\pmf(\store)  \defeq \margd{\pmf}{X}(\store)$.
+  %and
+  %$\store_2 \in \mems(X_2)$ define
+  %$\pmf(\store_1)  \defeq \margd{\pmf}{X_1}(\store_1)$ 
+  %\begin{enumerate}
+  %\item $\condd{\pmf}{X_1}{\store_2}(\store_1) \defeq \condd{\pmf}{X_1}{X_2}(\store_1 \uplus \store_2)$
+  %\item $\pmf(\store_1)  \defeq \margd{\pmf}{X_1}(\store_1)$ 
+  %\item $\pmf(\store_1|\store_2) \defeq \condd{\pmf}{X_1}{X_2}(\store_1 \uplus \store_2)$
+  %\end{enumerate}
 \end{definition}
@@
   \begin{array}{c}
-    \botruns(\prog) \defeq \\
+    \botruns(\prog)\ \defeq\ % \\
     \{ \store\{ x_1 \mapsto \bot, \ldots, x_n \mapsto \bot \} \mid 
@@
     \houtputs & \defeq
-        & \{\ \rvl{w} \mid\ \reveal{w}{\be}{\cid} \in \prog \wedge \cid \in H \ \}\ \cup \\
-      & & \{\ \elab{\mesg{w}}{\cid}\ \mid\  \eassign{\mesg{w}}{\cid}{\be}{\cid'} \in
+        & \{\ \rvl{w} \mid\ \reveal{w}{\be}{\cid} \in \prog \wedge \cid \in H \ \}\ \cup\ %\\
+    % & &
+    \{\ \elab{\mesg{w}}{\cid}\ \mid\  \eassign{\mesg{w}}{\cid}{\be}{\cid'} \in
            \prog \wedge \cid \in C \wedge \cid' \in H \ \} 
@@
     \cinputs &  \defeq
-        & \{\ \rvl{w} \mid\ \reveal{w}{\be}{\cid} \in \prog \wedge \cid \in C \ \} \ \cup\\
-      & & \{\ \elab{\mesg{w}}{\cid}\ \mid\  \eassign{\mesg{w}}{\cid}{\be}{\cid'} \in
+        & \{\ \rvl{w} \mid\ \reveal{w}{\be}{\cid} \in \prog \wedge \cid \in C \ \} \ \cup\ %\\
+    %& &
+    \{\ \elab{\mesg{w}}{\cid}\ \mid\  \eassign{\mesg{w}}{\cid}{\be}{\cid'} \in
               \prog \wedge \cid \in H \wedge \cid' \in C \ \}
@@
   satisfies \emph{noninterference modulo output}
-  iff for all $H$ and $C$ and $\store_1 \in \mems(S_C \cup O)$ and $\store_2 \in \mems(\houtputs)$
+  iff for all $H$ and $C$ %and $\store_1 \in \mems(S_C \cup O)$ and $\store_2 \in \mems(\houtputs)$
   we have:
   $$
-  \condd{\progtt(\prog)}{S_H}{\store_1} = \condd{\progtt(\prog)}{S_H}{\store_1 \uplus \store_2}
+  \condd{\progtt(\prog)}{S_H}{S_C \cup O} = \condd{\progtt(\prog)}{S_H}{S_C \cup O \cup \houtputs}
  $$
@@
   \emph{integrity} iff for all $H$, $C$, and $\adversary$,
-  if $\store \in \aruns(\prog)$ 
-  then there exists $\store' \in \mems(S)$ with $\store_{S_H} = \store'_{S_H} $ and:
+  if $\store \in \aruns(\prog)$ we have:
     $$
-    \condd{\progtt(\prog,\adversary)}{X}{\store_{S_H \cup \cinputs}} =
-    \condd{\progtt(\prog)}{X}{\store'}
+    \condd{\progtt(\prog,\adversary)}{X}{S_H \cup \cinputs} =
+    \condd{\progtt(\prog)}{X}{S_H \cup \cinputs}
     $$ 

### 2025-11-17 toplas26/sty.tex
checkpoint

@@
 
+$$
+\mems(\Phi) \defeq \{ \store \mid \dom(\store) = \setit{\ x_{\phi}\ \mid\ \phi \in \Phi\ } \}
+$$
+
+\begin{definition}
+  Given pmf $\pmf$ and $\Phi$ with $\dom(\pmf) = \bigcup_{\phi \in \Phi} \dom(\phi)$,
+  for all $\store \in \mems(\Phi)$ define:
+%$$
+%\pmf(\phi = v) \defeq \sum_{\store \in \mems(X) \ \text{with}\ \store(\phi) = v} \pmf(\store)
+%$$
+  $$
+  \pmf(\store) \defeq \sum_{\store' \in \mems(X)}
+  \begin{cases}
+    \pmf(\store') \ \text{if}\ \forall \phi \in \Phi . \store'(\phi) = \store(x_\phi)  \\
+    0 \ \text{otherwise}
+  \end{cases}
+  $$
+\end{definition}
+
+\begin{definition}
+  Given $\pmf$, the \emph{conditional distribution}
+  of $X$ given $\Phi$ where $\bigcup_{\phi \in \Phi} \dom(\phi) \cup X = \dom(\pmf)$ 
+  is denoted $\condd{\pmf}{X}{\Phi}$ and defined as follows: 
+  $$
+  \forall \store_1 \in \mems(X)\ .\ \forall \store_2 \in \mems(\Phi) \ .\ 
+  \condd{\pmf}{X}{\Phi}(\store_1 \uplus \store_2) \defeq
+  \begin{cases}
+    0 \text{\ if\ } \margd{\pmf}{\Phi}(\store_2) = 0\\
+    \pmf(\store_1 \uplus \store_2) / \pmf(\store_2) \text{\ otherwise}
+  \end{cases}
+  $$
+\end{definition}
+
 Now we can prove our share typing soundness result, showing that share
@@
   \label{lemma-sty-noninteractive-sound}
-  Given $\prog,\Gamma,R$ where $\secrets(\prog) = S$ and:
+  Given $\prog,\Gamma,R : {OK}$ where $\secrets(\prog) = S$ and:
   $$

### 2025-11-17 toplas26/model.tex
checkpoint

@@
   and reveals. Given a program $\prog$ with $\iov(\prog) = S \cup M \cup P \cup O$,
-  define $\views(\prog) \defeq M \cup P$, and define $\houtputs$ as
-  the messages and reveals in $V = M \cup P$ sent from honest to corrupt
-  parties, called \emph{corrupt views}:
-  $$
+  define $\views(\prog) \defeq M \cup P$. Letting $I_1$, $I_2$ be
+  a partitioning of the federation clients, define:
   \begin{array}{lcl}
-    \houtputs & \defeq
-        & \{\ \rvl{w} \mid\ \reveal{w}{\be}{\cid} \in \prog \wedge \cid \in H \ \}\ \cup\ %\\
+    V_{I_1 \rhd I_2} & \defeq
+        & \{\ \rvl{w} \mid\ \reveal{w}{\be}{\cid} \in \prog \wedge \cid \in I_1 \ \}\ \cup\ %\\
     % & &
     \{\ \elab{\mesg{w}}{\cid}\ \mid\  \eassign{\mesg{w}}{\cid}{\be}{\cid'} \in
-           \prog \wedge \cid \in C \wedge \cid' \in H \ \} 
+           \prog \wedge \cid \in I_2 \wedge \cid' \in I_1 \ \} 
   \end{array}
   $$
-  and similarly define $\cinputs$ as the subset of $V$ sent from corrupt to honest
-  parties, called \emph{honest views}:
-  $$
-  \begin{array}{lcl}
-    \cinputs &  \defeq
-        & \{\ \rvl{w} \mid\ \reveal{w}{\be}{\cid} \in \prog \wedge \cid \in C \ \} \ \cup\ %\\
-    %& &
-    \{\ \elab{\mesg{w}}{\cid}\ \mid\  \eassign{\mesg{w}}{\cid}{\be}{\cid'} \in
-              \prog \wedge \cid \in H \wedge \cid' \in C \ \}
-  \end{array}
-  $$
+  Given a partitioning of the federation $H,C$, we call $\houtputs$
+  \emph{corrupt views} and $\hinputs$ \emph{honest views}.
 \end{definition}

### 2025-11-18 toplas26/sty.tex
checkpoint

@@
 \begin{array}{lll}
-  \elab{\mesg{s1}}{2} &:=& \elab{(\secret{1} \fminus \locflip)}{1}; \\
-  \elab{\mesg{s2}}{1} &:=& \elab{(\secret{2} \fminus \locflip)}{2} 
+  \elab{\mesg{s1}}{2} := \elab{(\secret{1} \fminus \locflip)}{1}; \quad
+  \elab{\mesg{s2}}{1} := \elab{(\secret{2} \fminus \locflip)}{2} 
 \end{array}
@@
 \setit{[\mx{s2}{1}, \elab{\locflip}{2}] : \sharety,
-  [\mx{s1}{2}, \elab{\locflip}{1}] : \sharety}, \setit{\elab{\locflip}{1},\elab{\locflip}{2}}
+  [\elab{\locflip}{1}, \mx{s1}{2}] : \sharety}, \setit{\elab{\locflip}{1},\elab{\locflip}{2}}
 $$
@@
 $$
-\begin{array}{lll}
-  \px{1} &:=& \elab{\mesg{s2} + \locflip + 1}{1}; \\
-  \px{2} &:=& \elab{\mesg{s1} + \locflip}{2};\\
-  \out{1} &:=& \elab{\px{1} + \px{2}}{1};\\
-  \out{2} &:=& \elab{\px{1} + \px{2}}{2}
+\begin{array}{l}
+  \px{1} := \elab{\mesg{s2} + \locflip + 1}{1};\quad \px{2} := \elab{\mesg{s1} + \locflip}{2};\\
+  \out{1} := \elab{\px{1} + \px{2}}{1};\quad \out{2} := \elab{\px{1} + \px{2}}{2}
 \end{array}
@@
 \Gamma &\defeq& \setit{[\mx{s2}{1}, \elab{\locflip}{2}] : \sharety,
-  [\mx{s1}{2}, \elab{\locflip}{1}] : \sharety}\\
+  [\elab{\locflip}{1}, \mx{s1}{2}] : \sharety}\\
 R &\defeq&  \setit{\elab{\locflip}{1},\elab{\locflip}{2}}
@@
 
-$$
+Share type soundness (Theorem \ref{theorem-sty}) means that share
+typing enforces noninterference modulo output. To prove this we need
+to assert probabilistic properties in share typings, including for
+functions of protocol variables such as reconstruction of shares. So
+we make standard extensions of notation for probability distributions
+to express probabilities, and conditional probabilities, of functions
+of program variables $\phi$. For this extension we posit that
+memories can incorporate variables $x_\phi$ that assign valuations
+to functions $\phi$.
+\begin{definition}
+  We let $\Phi$ range over sets of SMT formulas $\phi$, and define
+  $
   \mems(\Phi) \defeq \{ \store \mid \dom(\store) = \setit{\ x_{\phi}\ \mid\ \phi \in \Phi\ } \}
-$$
-
-\begin{definition}
+  $.
   Given pmf $\pmf$ and $\Phi$ with $\dom(\pmf) = \bigcup_{\phi \in \Phi} \dom(\phi)$,
@@
   \begin{cases}
-    0 \text{\ if\ } \margd{\pmf}{\Phi}(\store_2) = 0\\
+    0 \text{\ if\ } \pmf(\store_2) = 0\\
     \pmf(\store_1 \uplus \store_2) / \pmf(\store_2) \text{\ otherwise}
@@
 
-Now we can prove our share typing soundness result, showing that share
-typing enforces noninteference modulo output, with our main result
-in Theorem \ref{theorem-sty}. We begin by defining a key well-formedness
-property of typing components that we show is preserved by non-interactive
-typing judgements.
-
+To proceed, we define a well-formedness property of tuples $\prog,\Gamma,R$
+intuitively meaning that all the share types in $\Gamma$ are sound with respect
+to $\prog$, and consumed randomness in $R$ as OTPs for initial sharing.
+Property (i) ensures that views do not reveal information about secrets.
+Property (ii) ensures the key property of sharing-- that individual shares
+don't reveal more about secrets than their reconstruction does. Properties
+(iii) and (iv) ensure that OTPs used for generating shares or for
+secure openings, respectively, are included in $R$.
 \begin{definition}
+  \label{def-OK}
   Given $\prog,\Gamma,R$ where $\secrets(\prog) = S$ and:
@@
 \end{definition}
-
+Next we show that, given well-formed typing environments, share type judgements preserve
+the important properties (i-ii) in Definitions \ref{def-OK}. This result follows from
+known semi-homomorphic properties of non-interactive operations that are syntactically
+verified in typing judgements.
 \begin{lemma}
@@
 \end{lemma}
-
+Next, we show that interactive typing rules preserve well-formedness of type
+environments. This Lemma follows by Lemma \ref{lemma-sty-noninteractive-sound}
+and the security of share generation using OTPs-- that is, shares
+generated by $\genshares$ are in uniform marginal distributions. 
 \begin{lemma}
+  \label{lemma-OK}
   Given $\prog,\Gamma_1,R_1  : {OK}$
@@
 \end{lemma}
-
+In addition to typing judgements, in order to show that a protocol is
+secure, we need to classify all corrupt views as shares, all
+public values as secure openings, and all outputs as reconstructed
+shares. We capture this formally with the following notion of
+type environment closure.
 \begin{definition}
+  \label{def-sty-closed}
   Given  $\prog,\Gamma,R  : {OK}$ and $\iov(\prog) = S \cup V \cup O$.
@@
 \end{definition}
-
+Now, the following is a consequence of Definitions \ref{def-OK} and \ref{def-sty-closed}, and then
+our main Theorem \ref{theorem-sty} follows by Lemma \ref{lemma-OK} and Theorem \ref{theorem-OK}.
 \begin{theorem}
-  \label{theorem-sty}
+  \label{theorem-OK}
   If $\prog,\Gamma,R : {OK}$ and $\Gamma$ is closed
@@
 \end{theorem}
-
 \begin{theorem}

### 2025-11-19 toplas26/metaty.tex
checkpoint

@@
 Share types rely on SMT entailments via the \TirName{ShareEntails} and
-\TirName{PubEntails} rules, which are not syntax directed. Our goal
+\TirName{PubEntails} rules, which are not syntax directed, and likewise
+the interactive typing judgement is nondeterministic. Our goal
 with the $\metaprot$ language is to automate share type checking. To
@@
 
+Demonstrating soundness of $\metaprot$ type checking is similar to $\minicat$--
+we show the ${OK}$ invariant is preserved by the algorithm. By Theorem \ref{theorem-OK}
+soundness follows. Lemma \ref{lemma-metaty} follows by appeal to Lemma \ref{lemma-sty-noninteractive-sound}
+since the algorithmic judgement $\Vdash$ is sound with regard to the noninteractive
+typing rules by definition. Theorem \ref{theorem-metaty} follows as a special
+case of Lemma \ref{lemma-metaty}.
 \begin{lemma}
+  \label{lemma-metaty}
   Given $\prog_1,\Gamma_1,R_1 : {OK}$ and $\cmd \redx \prog_2$ and $\htrip{\eqs_1}{\cmd}{\eqs_2}$ with
@@
 \begin{theorem}
+  \label{theorem-metaty}
   Given $\htrip{\eqspre}{\ttt{main()}}{\eqs}$ and $\ttt{main} : \Gamma_1 \rightarrow (\Gamma_2,R)$
@@
 \end{theorem}
+
+Note that Theorem \ref{theorem-metaty} assumes the existence of
+$\progpre$ with $\progpre,\Gamma_1,R' : {OK}$ and $R' \cap R =
+\varnothing$ and $\toeq{\progpre} \models \eqspre$. The assumption of
+some preprocessing protocol such as Beaver triple or MAC share
+distribitions is typical, and our approach does not require the
+protocol to be defined in detail, just (weakened) postconditions of it
+that are sufficient to verify the protocol being programmed.
+
+Theorem \ref{theorem-metaty} also does not provide a complete recipe
+for obtaining noninterference modulo output-- as suggested by Theorem
+\ref{theorem-sty}, it is also necessary to check closure of
+$\Gamma_2$. There are various ways this can be accomplished for the
+$\metaprot$ analysis-- for example, we evaluate $\ttt{main}$ to obtain
+all program variables and then check closure of $\Gamma_2$ as
+prescribed by Definition \ref{def-sty-closed}. We can also add checks
+to the type system to guarantee that all messages exchanged between
+clients are classified when type checking functions, which is
+a simple addition.
 

### 2025-11-20 code/cvc5mvn/src/main/scala/plaid/prelude/App.scala
bandaid validation issues

@@ -6,7 +6,7 @@ import plaid.prelude.antlr.Loader
 import plaid.prelude.ast.ListConstraintFuncExt.expandAll
 import plaid.prelude.ast.ListExprFuncExt.expandAll
 import plaid.prelude.cvc.TermFactory
-import plaid.prelude.logic.{checkConstraintExpansion, contracts, verificationFailures}
+import plaid.prelude.logic.{contracts, verificationFailures}
 
 import java.io.File
 import java.nio.file.Files
@@ -30,11 +30,6 @@ class App extends Runnable {
     val exprFns = ast.exprFuncs.expandAll()
     val constraintFns = ast.constraintFuncs.expandAll(exprFns)
     val contracts = ast.cmdFuncs.contracts(exprFns, constraintFns)
-  
-    val constraintErrors = contracts.flatMap(_.checkConstraintExpansion())
-    constraintErrors.foreach(x => println(s"ERROR ${x.ctx}: ${x.msg}: ${x.offender}"))
-    if constraintErrors.nonEmpty then
-      throw Exception("Found errors in the constraints, bailing out.")
 
     var totalTests = 0
     var totalFailures = 0

### 2025-11-20 code/cvc5mvn/src/main/scala/plaid/prelude/logic/Verifier.scala

@@ -35,5 +35,13 @@ extension (trg: Contract)
     trg.internals.filter { x =>
       val a = x.a.expand(bindings = bindings)
       val b = x.b.expand(bindings = bindings)
+
+      // TODO Validation shouldn't really go here, it needs to be revisited
+      val ctx = trg.f.id.name
+      val constraintErrors = a.checkProperExpansion(ctx) ++ b.checkProperExpansion(ctx)
+      constraintErrors.foreach(x => println(s"ERROR ${x.ctx}: ${x.msg}: ${x.offender}"))
+      if constraintErrors.nonEmpty then
+        throw Exception("Found errors in the constraints, bailing out.")
+
       !cvc.entails(cvc.toTerm(a), cvc.toTerm(b))
     }

### 2025-11-20 code/cvc5mvn/src/main/scala/plaid/prelude/logic/WellFormed.scala

@@ -61,8 +61,3 @@ extension (trg: Constraint)
     trg.descendants().flatMap(c => checks
       .flatMap { (f, msg) => if !f(c) then List(AstError(ctx, msg, c)) else Nil })
       .toList
-
-extension (trg: Contract)
-  def checkConstraintExpansion(): List[AstError] =
-    val ctx = trg.f.id.name
-    trg.internals.flatMap(x => x.a.checkProperExpansion(ctx) ++ x.b.checkProperExpansion(ctx))

### 2025-11-21 toplas26/examples.tex
checkpoint

@@
-\input{beaver}
-
 \section{Extended Examples}
 \label{section-examples}
+
+\preludetestfig
 
@@
 
-To evaluate our system especially regarding performance, using the
-Satisfiability Modulo Fields theory in cvc5 on an Apple M1 processor,
-we have verified correctness of these arithmetic gate libraries and
-circuit in fields up to size $\fieldp{2^{31} - 1}$ (an approximation
-of 32-bit integers) which takes $\sim$7ms to compute for 3 parties and
-not significantly more than for, e.g., $\fieldp{2}$. We have also
-verified the semi-homomorphic encryption properties leveraged in the
-BDOZ circuit library functions in fields up to size $\fieldp{2^{31} -
-  1}$ which take $\sim$8ms for the property used for the postcondition
-of the $\ttt{mult}$ gate. This is categorically better performance
-than can be achieved using previous brute-force methods in
+\paragraph{Implementation and Evaluation}
+We have implemented the $\metaprot$ language and automated Hoare logic
+using an Antlr parser and Scala backend with cvc5 Java bindings for
+interaction with the Satisfiability Modulo Fields theory.  To evaluate
+performance, we have written and verified gate libraries and small
+circuits using the GMW protocol for binary circuits as excerpted in
+Figure \ref{fig-gmwprepost} in Section \ref{section-metalang}, and for
+passive and malicious secure arithmetic circuits using Beaver Triple
+multiplication as presented in Figures \ref{fig-beaver} and
+\ref{fig-beaver-bdoz} in this Section.  We have considered field sizes
+typically used to encode binary, 16 bit, 32 bit, and 64 bit
+integers. Results are presented in Figure \ref{fig-preludetest}. In
+summary, verification is efficient, largely due to the performance of
+underlying cvc5 verification. It is categorically better performance
+than can be achieved using previous brute-force methods developed in
 $\fieldp{2}$ \cite{skalka-near-ppdp24}.
@@
 \label{section-beaver}
+
+\input{beaver}
 

### 2025-11-21 toplas26/sty.tex
checkpoint

--- 2025-11-17 toplas26/sty.tex
@@
   \item For all $a,b \in \setit{1,2}$ and $0 < i \le n$ we have
-    $\condd{\progtt(\prog)}{S_{\setit{\cid_a}}}{\setit{\phi_i^1 + \phi_i^2}} =
-    \condd{\progtt(\prog)}{S_{\setit{\cid_a}}}{\setit{\phi_i^1 + \phi_i^2, \phi_i^a}}$
+    $\condd{\progtt(\prog)}{S_{\setit{\cid_a}}}{\setit{\recon(\phi_i^1,\phi_i^2)}} =
+    \condd{\progtt(\prog)}{S_{\setit{\cid_a}}}{\setit{\recon(\phi_i^1,\phi_i^2), \phi_i^b}}$
   \item  For all $a,b \in \setit{1,2}$ and $0 < i \le n$, if
-    $\toeq{\prog} \models \phi_i^a \eop \rx{w}{\cid_a} \wedge \phi_i^b \eop \phi - \rx{w}{\cid_a}$ for
-    some $\phi$ and $\rx{w}{\cid_a}$ then $\rx{w}{\cid_a} \in R$.
+    $([x_1,x_2] : \sharety) \in \Gamma$ with $\genshares(\phi,\rx{w}{\cid}) = \phi_1,\phi_2$
+    and $\toeq{\prog} \models x_1 \eop \phi_1 \wedge x_2 \eop \phi_2$ 
+    for some $\phi$ and $\rx{w}{\cid}$ then $\rx{w}{\cid} \in R$.
   \item For all $0 < i \le j$, if $\toeq{\prog} \models \phi_i \eop \phi + \rx{w}{\cid_{\oid}}$ for
@@
 \end{definition}
+With the definition of ${OK}$ in place, we can define what we mean by soundness
+of $\genshares$. In this definition, the first condition establishes that
+initial shares are independent of honest secrets, and the second establishes
+that they reveal no more than their reconstruction. These conditions are
+obviously met by ``vanilla'' additive or Shamir shares which generate
+shares in uniform marginal distributions.
+\begin{definition}
+  \label{def-genshares}
+  We say that $\genshares$ is \emph{sound} iff given any $\prog,\Gamma,R : {OK}$
+  and $\phi$ and $\rx{w}{\cid} \not \in R$ with $\genshares(\phi,\rx{w}{\cid}) =
+  \phi_1,\phi_2$ and $\toeq{\prog} \models x_1 \eop \phi_1 \wedge x_2 \eop \phi_2$
+  with $\secrets(\prog) = S$ and:
+  $$
+  \Gamma = \setit{[\phi_1^1,\phi_1^2] : \sharety, \ldots, [\phi_n^1,\phi_n^2] : \sharety}
+  \cup \setit{\pubx{\phi_1} : \pubty,\ldots,\pubx{\phi_j} : \pubty} \cup
+  \setit{\pubx{\phi^o_1} : \outty,\ldots,\pubx{\phi^o_k} : \outty}
+  $$
+  we have for all $a,b \in \setit{1,2}$:
+  \begin{enumerate}[\hspace{5mm}i.]
+    \item $\condd{\progtt(\prog)}{S_{\setit{\cid_a}}}{S_{\setit{\cid_b}} \cup \setit{\phi_1^b,\ldots,\phi_n^b} \cup
+      \setit{\phi_1,\ldots,\phi_j} \setit{\phi^o_1,\ldots,\phi^o_k}}
+    =
+    \condd{\progtt(\prog)}{S_{\setit{\cid_a}}}{S_{\setit{\cid_b}} \cup \setit{\phi_1^b,\ldots,\phi_n^b,x_b} \cup
+       \setit{\phi_1,\ldots,\phi_j} \setit{\phi^o_1,\ldots,\phi^o_k}}$
+    \item $\condd{\progtt(\prog)}{S_{\setit{\cid_a}}}{\setit{\recon(x_1,x_2)}} =
+      \condd{\progtt(\prog)}{S_{\setit{\cid_a}}}{\setit{\recon(x_1,x_2), x_b}}$
+    \end{enumerate}
+\end{definition}
 
+Next we show that, given well-formed typing environments, share type judgements preserve
+the important properties (i-ii) in Definitions \ref{def-OK}. This result follows from
+known semi-homomorphic properties of non-interactive operations that are syntactically
+verified in typing judgements, and by Definition \ref{def-genshares} and the
+assumption of soundness of $\genshares$.
 \begin{lemma}
@@
        \setit{\phi_1,\ldots,\phi_j} \setit{\phi^o_1,\ldots,\phi^o_k}}$
-    \item $\condd{\progtt(\prog)}{S_{\setit{\cid_a}}}{\setit{\phi_i^1 + \phi_i^2}} =
-      \condd{\progtt(\prog)}{S_{\setit{\cid_a}}}{\setit{\phi_i^1 + \phi_i^2, \phi_i^a}}$
+    \item $\condd{\progtt(\prog)}{S_{\setit{\cid_a}}}{\setit{\recon(\phi_i^1,\phi_i^2)}} =
+      \condd{\progtt(\prog)}{S_{\setit{\cid_a}}}{\setit{\recon(\phi_i^1, \phi_i^2), \phi_i^b}}$
     \end{enumerate}
 \end{lemma}
-
+Next, we show that interactive typing rules preserve well-formedness of type
+environments. This Lemma follows by Lemma \ref{lemma-sty-noninteractive-sound}
+and the security of share generation using OTPs-- that is, shares
+generated by $\genshares$ are in uniform marginal distributions. 
 \begin{lemma}
+  \label{lemma-OK}
   Given $\prog,\Gamma_1,R_1  : {OK}$

### 2025-11-21 toplas26/sty.tex
checkpoint

@@
 
-Our approach to static analysis is to enforce the use of sound
-techniques in semi-homorphic encryption in protocols.  As discussed in
+Our approach to static confidentiality analysis is to enforce the use of sound
+techniques in semi-homomorphic encryption in protocols.  As discussed in
 the example in Section \ref{section-smt-example}, protocols integrate
@@
 encrypted secrets, and noninteractive elements, such as addition (of
-shares). We approach this by introducing distinct judgement forms for
+shares). To comport with this we introduce distinct judgement forms for
 interactive and noninteractive protocol elements. Our type system
 as presented will assume 2-client protocols, but this is for
-simplicity and the analysis can be generalized to more parties.
+clarity of presentation and the analysis can be generalized to more parties.
 
-Both interactive and non-interactive judgements leverage the use
-of SMT interpretations as described in Section \ref{section-smt}.
+Our confidentiality analysis mainly considers protocols in terms of
+their SMT interpretations as described in Section \ref{section-smt}.
 We introduce a \emph{share type} language $\sty$, classifying
@@
 client 2 has $\phi_2$, and this shared value can be reconstructed
-as $\phi_1 + \phi_2$.\cnote{Comment on generation and
-  reconstruction in general somewhere before this section?.} A
-share term $\pubx{\phi}$ denotes a publicly shared value.
+as a public value $\pubx{\phi}$.
 $$
@@
 $$
-Intuitively, the meaning of share types $\sty$ are as follows.
-When we ascribe the type $[\phi_1,\phi_2] : \sharety$, this
-means that $[\phi_1,\phi_2]$ are shares of their
-reconstruction, which we can represent as $\phi_1 + \phi_2$
-as discussed in Section \ref{???}. 
-In particular, each of $\phi_1$ and $\phi_2$ independently
-reveal nothing about $\phi_1 + \phi_2$. Likewise,
-$\pubx{\phi} : \outty$ means that $\phi$ has been
-publicly \emph{revealed} via broadcast shares, and our
-type system requires that $\phi \eop \phi_1 + \phi_2$ with
-$[\phi_1,\phi_2] : \sharety$. This implies that the
-process of revealing $\phi$ reveals nothing besides $\phi$.
-The ascription $\pubx{\phi} : \pubty$ is similar, with the
-additional meaning that $\phi$ is a publicly-known value
-independent of any client inputs. This includes known
-constant values used in protocols, and more interestingly
-\emph{secure openings}. Secure openings are commonly used
-to interactively compute public values that are masked with
-OTPs secret-shared with clients in initialization protocols,
-as for example in Beaver Triple multiplication which
-we discuss in Section \ref{section-examples}.
+\paragraph{Share generation and reconstruction}
+A \emph{secret sharing scheme} includes functions
+$(\genshares,\recon)$ which generate and reconstruct shares,
+respectively. The $\genshares$ function takes some
+input $\phi$ and randomness $\rx{x}{w}$ and produces
+two shares (or as many is needed for the number of clients
+in the federation): 
+$$
+\genshares(\phi,\rx{w}{\cid}) = \phi_1,\phi_2
+$$
+and:
+$$
+\recon(\genshares(\phi,\rx{w}{\cid})) = \phi
+$$
+In order for $\genshares$ to be \emph{sound}, we require
+that both of $\phi_1$ and $\phi_2$ are marginally independent
+of $\phi$, and also of any other values in the protocol.
+This entails the requirement that $\rx{w}{\cid}$ is
+consumed by $\genshares$, e.g., as an OTP.
+We formally characterize soundness of $\genshares$
+in our protocol model in Definition \ref{def-genshares}.
+We emphasize that our type system is agnostic with respect
+to the definition of $\genshares$ and $\recon$ modulo
+soundness-- we use an additive shares encoding in this work,
+but the analysis could apply instead to, e.g., Shamir
+sharing.
+
+More generally, when we ascribe the type $[\phi_1,\phi_2] : \sharety$,
+this means that $[\phi_1,\phi_2]$ are shares of their reconstruction
+$\recon(\phi_1,\phi_2)$, noting that shares can be obtained either by
+$\genshares$ or by noninteractive homomorphic protocols.  Likewise,
+$\pubx{\phi} : \outty$ means that $\phi$ has been publicly
+\emph{revealed} via broadcast shares, and our type system requires
+that $\phi \eop \recon(\phi_1, \phi_2)$ with $[\phi_1,\phi_2] :
+\sharety$. This implies that the process of revealing $\phi$ reveals
+nothing besides $\phi$.  The ascription $\pubx{\phi} : \pubty$ is
+similar, with the additional meaning that $\phi$ is a publicly-known
+value independent of any client inputs. This includes known constant
+values used in protocols, and more interestingly \emph{secure
+openings}. Secure openings are commonly used to interactively compute
+public values that are masked with OTPs secret-shared with clients in
+initialization protocols, as for example in Beaver Triple
+multiplication which we discuss in Section \ref{section-examples}.
 
@@
 
-As discussed in Section \ref{???}, a \emph{secret sharing scheme}
-includes functions $(\genshares,\recon)$ which generate
-and reconstruct shares, respectively.
+The interactive share typing rules are defined in Figure
+\ref{fig-styinteractive}, with judgements of the form $\eqs \vdash
+\Gamma_1,R_1, \stredx \Gamma_2,R_2$.  Each such judgement establishes
+a sound type binding in $\Gamma_2$ for interactively-defined
+variables-- that is, unicast and broadcast messages exchanged between
+clients-- assuming the bindings in $\Gamma_1$. In addition to type
+bindings, judgments track randomness used by $\genshares$, allowing us
+to use type linearity to enforce one-time usage similar to previous
+related work \cite{darais2019language}, where any randomness used in
+the typing is not in $R_1$ and added to $R_2$. To this end, in these
+rules $\uplus$ is disjoint union.
+
+The \TirName{Share} rule generates the binding $[\mx{z}{\cid_2},
+  \rx{y}{\cid_1}] : \sharety$ if this is a secret sharing as defined
+by $\genshares$. It also adds $\rx{y}{\cid_1}$ to the set of variables
+consumed by $\genshares$, ensuring that it has not already been used. The
+\TirName{Reveal} rule generates the binding $\pubx{\recon(\px{x},\px{y})} :
+\outty$ with the requirement that $[\px{x},\px{y}]$ are typable as
+shares. The \TirName{SecOpen} rule is similar except it allows the
+assertion $\pubx{\recon(\px{x},\px{y})} : \pubty$ if we can additionally
+show that the shared value is masked by the use of random value
+$\rx{z}{\oid}$ as an OTP. Here we are positing a third party $\oid$
+that secret-shares $\rx{z}{\oid}$ as part of an initialization
+protocol. We illustrate secure openings in practice at more length
+in Section \ref{section-examples}, though we note that appropriate
+properties of initialization protocols can be encoded in $\eqspre$.
+
+We write $\stredx^*$ to denote the transitive
+closure of $\stredx$. For any protocol $\prog$, the goal of the
+type analysis is to show that all interactive messages can be classified--
+that is, to show that $\eqspre \wedge \toeq{\prog} \vdash
+\varnothing,\varnothing \stredx^* \Gamma,R$ where all
+interactive messages are classified in $\Gamma$.
+
+\styinteractivefig
+
+\subsection{Instantiating $\genshares$}
+
 It is well-known that adding or subtracting a sample from a uniform
@@
 \begin{eqnarray*}
-  \genshares(\phi) &\defeq& [\phi - \rx{y}{\cid_2}, \rx{y}{\cid_2}]\\
-  \recon([\phi_1,\phi_2]) &\defeq& \phi_1 \fplus \phi_2
+  \genshares(\phi,\rx{y}{\cid}) &\defeq& \phi - \rx{y}{\cid}, \rx{y}{\cid}\\
+  \recon(\phi_1,\phi_2) &\defeq& \phi_1 \fplus \phi_2
 \end{eqnarray*}
@@
 
-Of course, there are other methods for encrypting values in MPC
+This approach highlights an imporant interaction between our type
+system and SMT entailment as defined by the $\TirName{ShareEntails}$
+rule. In particular there are other methods for encrypting values in MPC
 protocols that can be shown to be algebraically equivalent
@@
 
-\styinteractivefig
-
-The interactive share typing rules are defined in Figure
-\ref{fig-styinteractive}, with judgements of the form $\eqs \vdash
-\Gamma_1,R_1, \stredx \Gamma_2,R_2$.  Each such judgement establishes
-a sound type binding in $\Gamma_2$ for interactively-defined
-variables-- that is, unicast and broadcast messages exchanged between
-clients-- assuming the bindings in $\Gamma_1$. In addition to type
-bindings, judgments track random variables used as OTPs, allowing us
-to use type linearity to enforce one-time usage similar to previous
-related work \cite{darais2019language}, where any randomness used in
-the typing is not in $R_1$ and added to $R_2$. To this end, in these
-rules $\uplus$ is disjoint union.
-
-The \TirName{Share} rule generates the binding $[\mx{z}{\cid_2},
-  \rx{y}{\cid_1}] : \sharety$ if this is a secret sharing as defined
-by $\genshares$. It also adds $\rx{y}{\cid_1}$ to the set of variables
-used as OTPs, ensuring that it has not already been used. The
-\TirName{Reveal} rule generates the binding $\pubx{\px{x} + \px{y}} :
-\outty$ with the requirement that $[\px{x},\px{y}]$ are typable as
-shares. The \TirName{SecOpen} rule is similar except it allows the
-assertion $\pubx{\px{x} + \px{y}} : \pubty$ if we can additionally
-show that the shared value is masked by the use of random value
-$\rx{z}{\oid}$ as an OTP. Here we are positing a third party $\oid$
-that secret-shares $\rx{z}{\oid}$ as part of an initialization
-protocol. We illustrate secure openings in practice at more length
-in Section \ref{section-examples}, though we note that appropriate
-properties of initialization protocols can be encoded in $\eqspre$.
-
-We write $\stredx^*$ to denote the transitive
-closure of $\stredx$. For any protocol $\prog$, the goal of the
-type analysis is to show that all interactive messages can be classified--
-that is, to show that $\eqspre \wedge \toeq{\prog} \vdash
-\varnothing,\varnothing \stredx^* \Gamma,R$ where all
-interactive messages are classified in $\Gamma$.
-
 \subsection{Share Typing Example} To illustrate details of the
@@
 $$
-We can then assert:
+By $\TirName{GenShares}$ we can then assert:
 $$

### 2025-11-23 toplas26/sty.tex
checkpoint

@@
 
-The \TirName{Share} rule generates the binding $[\mx{z}{\cid_2},
-  \rx{y}{\cid_1}] : \sharety$ if this is a secret sharing as defined
-by $\genshares$. It also adds $\rx{y}{\cid_1}$ to the set of variables
-consumed by $\genshares$, ensuring that it has not already been used. The
+The \TirName{GenShares} rule generates the binding $[x_1,
+  x_2] : \sharety$ if the term is a secret sharing as defined
+by $\genshares$, and adds $\rx{y}{\cid_1}$ to the set of variables
+consumed by $\genshares$, ensuring one-time usage. The
 \TirName{Reveal} rule generates the binding $\pubx{\recon(\px{x},\px{y})} :
@@
 Property (ii) ensures the key property of sharing-- that individual shares
-don't reveal more about secrets than their reconstruction does. Properties
-(iii) and (iv) ensure that OTPs used for generating shares or for
+don't reveal more about secrets than their reconstruction does. Property
+(iii) ensures that public values are independent of outputs or secrets. Properties
+(iv) and (v) ensure that OTPs used for generating shares or for
 secure openings, respectively, are included in $R$.
@@
     \condd{\progtt(\prog)}{S_{\setit{\cid_a}}}{S_{\setit{\cid_b}} \cup \setit{\phi_1^b,\ldots,\phi_n^b} \cup
-      \setit{\phi_1,\ldots,\phi_j} \setit{\phi^o_1,\ldots,\phi^o_k}}
+      %\setit{\phi_1,\ldots,\phi_j}
+      \setit{\phi^o_1,\ldots,\phi^o_k}}
     $.
@@
     \condd{\progtt(\prog)}{S_{\setit{\cid_a}}}{\setit{\recon(\phi_i^1,\phi_i^2), \phi_i^b}}$
+  \item $\sep{\progd(\prog)}{(S \cup O)}{\setit{\phi_1,\ldots,\phi_j}}$
   \item  For all $a,b \in \setit{1,2}$ and $0 < i \le n$, if
@@
     \item $\condd{\progtt(\prog)}{S_{\setit{\cid_a}}}{S_{\setit{\cid_b}} \cup \setit{\phi_1^b,\ldots,\phi_n^b} \cup
-      \setit{\phi_1,\ldots,\phi_j} \setit{\phi^o_1,\ldots,\phi^o_k}}
+      %\setit{\phi_1,\ldots,\phi_j}
+      \setit{\phi^o_1,\ldots,\phi^o_k}}
     =
     \condd{\progtt(\prog)}{S_{\setit{\cid_a}}}{S_{\setit{\cid_b}} \cup \setit{\phi_1^b,\ldots,\phi_n^b,x_b} \cup
-       \setit{\phi_1,\ldots,\phi_j} \setit{\phi^o_1,\ldots,\phi^o_k}}$
+      %\setit{\phi_1,\ldots,\phi_j}
+      \setit{\phi^o_1,\ldots,\phi^o_k}}$
     \item $\condd{\progtt(\prog)}{S_{\setit{\cid_a}}}{\setit{\recon(x_1,x_2)}} =
@@
 known semi-homomorphic properties of non-interactive operations that are syntactically
-verified in typing judgements, and by Definition \ref{def-genshares} and the
-assumption of soundness of $\genshares$.
+verified in typing judgements.
 \begin{lemma}
@@
     \item $\condd{\progtt(\prog)}{S_{\setit{\cid_a}}}{S_{\setit{\cid_b}} \cup \setit{\phi_1^b,\ldots,\phi_n^b} \cup
-      \setit{\phi_1,\ldots,\phi_j} \setit{\phi^o_1,\ldots,\phi^o_k}}
+      %\setit{\phi_1,\ldots,\phi_j}
+      \setit{\phi^o_1,\ldots,\phi^o_k}}
     =
     \condd{\progtt(\prog)}{S_{\setit{\cid_a}}}{S_{\setit{\cid_b}} \cup \setit{\phi_1^b,\ldots,\phi_n^b,\phi^b} \cup
-       \setit{\phi_1,\ldots,\phi_j} \setit{\phi^o_1,\ldots,\phi^o_k}}$
+      % \setit{\phi_1,\ldots,\phi_j}
+      \setit{\phi^o_1,\ldots,\phi^o_k}}$
     \item $\condd{\progtt(\prog)}{S_{\setit{\cid_a}}}{\setit{\recon(\phi_i^1,\phi_i^2)}} =
@@
 \end{lemma}
+\begin{proof}
+  The proof is by induction on the derivation of $\Gamma, \toeq{\prog} \vdash [\phi^1,\phi^2] : \sharety$.
+  In the base case where the judgement follows by the \TirName{Shared} rule, the result
+  follows trivially by Definition \ref{def-OK}. In cases \TirName{HEAdd}, \TirName{HEAddPub}, and
+  \TirName{HEMultPub} the result follows by Definition \ref{def-OK} and known
+  semi-homorphic properties of secret sharing \cite{10.1007/978-3-030-68869-1_3}.
+  Case \TirName{ShareEntails} follows by correctness of the SMT interpretation of
+  protocols as in Theorem \ref{theorem-toeq}.
+\end{proof}
+
 Next, we show that interactive typing rules preserve well-formedness of type
-environments. This Lemma follows by Lemma \ref{lemma-sty-noninteractive-sound}
-and the security of share generation using OTPs-- that is, shares
+environments. This Lemma follows by Lemma \ref{lemma-sty-noninteractive-sound},
+and the security of share generation using OTPs as articulated in
+Definition \ref{def-genshares}-- that is, shares
 generated by $\genshares$ are in uniform marginal distributions. 
@@
 \end{lemma}
+\begin{proof}
+  The proof follows by case analysis on  $\toeq{\prog}  \vdash \Gamma_1, R_1 \stredx \Gamma_2, R_2$.
+  In case \TirName{GenShares} the result follows by Definition \ref{def-genshares},
+  which satisfies relevant conditions (i-ii,iv) of Definition \ref{def-OK}. In case
+  \TirName{SecOpen} the public value must be reconstructed shares so Lemma
+  \ref{lemma-sty-noninteractive-sound} relevant conditions (i-ii), and the
+  the requirement that the public value be masked with an OTP satisfies conditions
+  (iii) and (iv) of Definition \ref{def-OK}. In case \TirName{Reveal} since
+  the output value must be reconstructed shares, Lemma \ref{lemma-sty-noninteractive-sound}
+  satisfies relevant case (ii) of  Definition \ref{def-OK}. 
+\end{proof}
+
 In addition to typing judgements, in order to show that a protocol is
@@
   \label{def-sty-closed}
-  Given  $\prog,\Gamma,R  : {OK}$ and $\iov(\prog) = S \cup V \cup O$.
-  Then $\Gamma$ is \emph{closed for $\prog$} iff for all $H,C$,
+  $\Gamma$ is \emph{closed for $\prog$} iff for all $H,C$,
   for all $x \in \houtputs$ there exists $\phi$ such that either
-  $\Gamma,\toeq{\prog} \vdash [x,\phi] : \sharety$ if $C = \setit{1}$
-  or  $\Gamma,\toeq{\prog} \vdash [\phi,x] : \sharety$ if $C = \setit{2}$,
-  and for all $\out{\cid} \in O$ we have $\Gamma,\toeq{\prog} \vdash \pubx{\out{\cid}} : \outty$.
+  $([x,\phi] : \sharety) \in \Gamma$ if $C = \setit{1}$
+  or  $([\phi,x] : \sharety) \in \Gamma$ if $C = \setit{2}$,
+  and for all $\out{\cid} \in O$ we have $(\pubx{\out{\cid}} : \outty) \in \Gamma$.
 \end{definition}
@@
 \end{theorem}
+\begin{proof}
+  Let $\iov(\prog) = S \cup V \cup O$ and assume some partitioning $H,C$.
+  %For any $\out{\cid} \in O$ with $\Gamma,\toeq{\prog} \vdash \pubx{\out{\cid}} : \outty$
+  %we can assume wlog that $(\out{\cid} : \outty) \in \Gamma$, since we can
+  %add the binding to $\Gamma'$ with  $\prog,\Gamma',R  : {OK}$ by Lemma
+  %\ref{lemma-OK}.
+  By Definitions \ref{def-sty-closed} we know that every $x \in \houtputs$
+  occurs in some $\stt$ with $(\stt, \sharety) \in \Gamma$, and
+  that $(\out{\cid} : \outty) \in \Gamma$ for all $\out{\cid} \in O$.
+  Therefore by Definition \ref{def-OK} (and marginalizing away
+  extraneous $(\stt : \sty) \in \Gamma$) we have:
+  $$
+  \condd{\progd(\prog)}{S_H}{S_C \cup O} = \condd{\progd(\prog)}{S_H}{S_C \cup V_C \cup O}
+  $$
+  which is noninterference modulo output by Definition \ref{definition-NIMO}.
+\end{proof}
+
 \begin{theorem}
@@
 \end{theorem}
+
+\begin{proof}
+  By Lemma \ref{lemma-OK} we have that $\prog,\Gamma,R : {OK}$, so the result follows
+  by Therorem \ref{theorem-OK}.
+\end{proof}
 

### 2025-11-24 toplas26/sty.tex
checkpoint

@@
   \label{def-OK}
-  Given $\prog,\Gamma,R$ where $\secrets(\prog) = S$ and:
-  $$
-  \Gamma = \setit{[\phi_1^1,\phi_1^2] : \sharety, \ldots, [\phi_n^1,\phi_n^2] : \sharety}
-  \cup \setit{\pubx{\phi_1} : \pubty,\ldots,\pubx{\phi_j} : \pubty} \cup
-  \setit{\pubx{\phi^o_1} : \outty,\ldots,\pubx{\phi^o_k} : \outty}
-  $$
+  Given $\prog,\Gamma,R$ where $\iov(\prog) = S \cup M \cup P \cup O$. For all $a,b\in \setit{1,2}$ with $a \ne b$ let:
+  \begin{eqnarray*}
+    %\Phi_a^m & \defeq & \setit{ \phi_a \mid ([\phi_1,\phi_2] : \sharety) \in \Gamma} \\
+    %\Phi_a^p & \defeq & \setit{ \phi_b \mid (\pubx{\recon(\phi_1,\phi_2)} : \sty) \in \Gamma}\\
+    \Phi_a & \defeq & \setit{ \phi_a \mid ([\phi_1,\phi_2] : \sharety) \in \Gamma} \cup
+                        \setit{ \phi_b \mid (\pubx{\recon(\phi_1,\phi_2)} : \sty) \in \Gamma} \\% \Phi_a^m \cup \Phi_a^p \\
+    \Phi^p & \defeq & \setit{ \recon(\phi_1,\phi_2) \mid (\pubx{\recon(\phi_1,\phi_2)} : \pubty) \in \Gamma}\\
+    \Phi^o & \defeq & \setit{ \recon(\phi_1,\phi_2) \mid (\pubx{\recon(\phi_1,\phi_2)} : \outty) \in \Gamma}
+  \end{eqnarray*}
   Then we write $\prog,\Gamma,R : {OK}$ iff all of the following conditions
@@
     $
-    \condd{\progtt(\prog)}{S_{\setit{\cid_a}}}{S_{\setit{\cid_b}} \cup \setit{\phi^o_1,\ldots,\phi^o_k}}
+    \condd{\progtt(\prog)}{S_{\setit{\cid_a}}}{S_{\setit{\cid_b}} \cup \Phi^o}
     = 
-    \condd{\progtt(\prog)}{S_{\setit{\cid_a}}}{S_{\setit{\cid_b}} \cup \setit{\phi_1^b,\ldots,\phi_n^b} \cup
-      %\setit{\phi_1,\ldots,\phi_j}
-      \setit{\phi^o_1,\ldots,\phi^o_k}}
+    \condd{\progtt(\prog)}{S_{\setit{\cid_a}}}{S_{\setit{\cid_b}} \cup  \Phi_a \cup \Phi^o}
     $.
-  \item For all $a,b \in \setit{1,2}$ and $0 < i \le n$ we have
-    $\condd{\progtt(\prog)}{S_{\setit{\cid_a}}}{\setit{\recon(\phi_i^1,\phi_i^2)}} =
-    \condd{\progtt(\prog)}{S_{\setit{\cid_a}}}{\setit{\recon(\phi_i^1,\phi_i^2), \phi_i^b}}$
-  \item $\sep{\progd(\prog)}{(S \cup O)}{\setit{\phi_1,\ldots,\phi_j}}$
-  \item  For all $a,b \in \setit{1,2}$ and $0 < i \le n$, if
-    $([x_1,x_2] : \sharety) \in \Gamma$ with $\genshares(\phi,\rx{w}{\cid}) = \phi_1,\phi_2$
+  \item For all $a,b \in \setit{1,2}$ and $([\phi_1,\phi_2] : \sharety) \in \Gamma$ we have:
+    %$\condsep{\progd(\prog)}{\setit{\recon(\phi_i^1,\phi_i^2)}}{\setit{\phi_i^b}}{\recon(\phi_i^1,\phi_i^2)}$
+    $$\condd{\progtt(\prog)}{S_{\setit{\cid_a}}}{S_{\setit{\cid_b}} \cup  \Phi_a \cup \Phi^o \cup \setit{\recon(\phi_i^1,\phi_i^2)}} =
+    \condd{\progtt(\prog)}{S_{\setit{\cid_a}}}{S_{\setit{\cid_b}}  \cup  \Phi_a \cup \Phi^o \cup \setit{\recon(\phi_i^1,\phi_i^2), \phi_i^b}}$$
+  \item $\sep{\progd(\prog)}{(S \cup O)}{\Phi^p}$
+  \item  For all $a,b \in \setit{1,2}$ and if
+    $([x_1,x_2] : \sharety) \in \Gamma$ with $x_a \in \imesgs{M}{\setit{b}}{\setit{a}}$ then
+    $\genshares(\phi,\rx{w}{\cid}) = \phi_1,\phi_2$
     and $\toeq{\prog} \models x_1 \eop \phi_1 \wedge x_2 \eop \phi_2$ 
-    for some $\phi$ and $\rx{w}{\cid}$ then $\rx{w}{\cid} \in R$.
-  \item For all $0 < i \le j$, if $\toeq{\prog} \models \phi_i \eop \phi + \rx{w}{\cid_{\oid}}$ for
-    some $\phi$  and $\rx{w}{\cid_{\oid}}$ then $\rx{w}{\cid_{\oid}} \in R$.
+    for some $\phi$ and $\rx{w}{\cid} \in R$.
+  \item For all $\phi \in \Phi^p$ if $\toeq{\prog} \models \phi \eop \phi' + \rx{w}{\cid_{\oid}}$ for some $\phi'$
+    and $\rx{w}{\cid_{\oid}}$ then $\rx{w}{\cid_{\oid}} \in R$.
   \end{enumerate} 
 \end{definition}
+%\begin{definition}
+%  \label{def-OK}
+%  Given $\prog,\Gamma,R$ where $\secrets(\prog) = S$ and:
+%  $$
+%  \Gamma = \setit{[\phi_1^1,\phi_1^2] : \sharety, \ldots, [\phi_n^1,\phi_n^2] : \sharety}
+%  \cup \setit{\pubx{\phi_1} : \pubty,\ldots,\pubx{\phi_j} : \pubty} \cup
+%  \setit{\pubx{\phi^o_1} : \outty,\ldots,\pubx{\phi^o_k} : \outty}
+%  $$
+%  Then we write $\prog,\Gamma,R : {OK}$ iff all of the following conditions
+%  hold:
+%  \begin{enumerate}[\hspace{5mm}i.]
+%  \item For all $a,b \in \setit{1,2}$ we have
+%    $
+%    \condd{\progtt(\prog)}{S_{\setit{\cid_a}}}{S_{\setit{\cid_b}} \cup \setit{\phi^o_1,\ldots,\phi^o_k}}
+%    = 
+%    \condd{\progtt(\prog)}{S_{\setit{\cid_a}}}{S_{\setit{\cid_b}} \cup \setit{\phi_1^b,\ldots,\phi_n^b} \cup
+%      %\setit{\phi_1,\ldots,\phi_j}
+%      \setit{\phi^o_1,\ldots,\phi^o_k}}
+%    $.
+%  \item For all $a,b \in \setit{1,2}$ and $0 < i \le n$ we have
+%    %$\condsep{\progd(\prog)}{\setit{\recon(\phi_i^1,\phi_i^2)}}{\setit{\phi_i^b}}{\recon(\phi_i^1,\phi_i^2)}$
+%    $\condd{\progtt(\prog)}{S_{\setit{\cid_a}}}{\setit{\recon(\phi_i^1,\phi_i^2)}} =
+%    \condd{\progtt(\prog)}{S_{\setit{\cid_a}}}{\setit{\recon(\phi_i^1,\phi_i^2), \phi_i^b}}$
+%  \item $\sep{\progd(\prog)}{(S \cup O)}{\setit{\phi_1,\ldots,\phi_j}}$
+%  \item  For all $a,b \in \setit{1,2}$ and $0 < i \le n$, if
+%    $([x_1,x_2] : \sharety) \in \Gamma$ with $\genshares(\phi,\rx{w}{\cid}) = \phi_1,\phi_2$
+%    and $\toeq{\prog} \models x_1 \eop \phi_1 \wedge x_2 \eop \phi_2$ 
+%    for some $\phi$ and $\rx{w}{\cid}$ then $\rx{w}{\cid} \in R$.
+%  \item For all $0 < i \le j$, if $\toeq{\prog} \models \phi_i \eop \phi + \rx{w}{\cid_{\oid}}$ for
+%    some $\phi$  and $\rx{w}{\cid_{\oid}}$ then $\rx{w}{\cid_{\oid}} \in R$.
+%  \end{enumerate} 
+%\end{definition}
 With the definition of ${OK}$ in place, we can define what we mean by soundness
@@
 shares in uniform marginal distributions.
+
 \begin{definition}
@@
 
+%\begin{definition}
+%  \label{def-genshares}
+%  We say that $\genshares$ is \emph{sound} iff given any $\prog,\Gamma,R : {OK}$
+%  and $\phi$ and $\rx{w}{\cid} \not \in R$ with $\genshares(\phi,\rx{w}{\cid}) =
+%  \phi_1,\phi_2$ and $\toeq{\prog} \models x_1 \eop \phi_1 \wedge x_2 \eop \phi_2$
+%  with $\secrets(\prog) = S$ and:
+%  $$
+%  \Gamma = \setit{[\phi_1^1,\phi_1^2] : \sharety, \ldots, [\phi_n^1,\phi_n^2] : \sharety}
+%  \cup \setit{\pubx{\phi_1} : \pubty,\ldots,\pubx{\phi_j} : \pubty} \cup
+%  \setit{\pubx{\phi^o_1} : \outty,\ldots,\pubx{\phi^o_k} : \outty}
+%  $$
+%  we have for all $a,b \in \setit{1,2}$:
+%  \begin{enumerate}[\hspace{5mm}i.]
+%    \item $\condd{\progtt(\prog)}{S_{\setit{\cid_a}}}{S_{\setit{\cid_b}} \cup \setit{\phi_1^b,\ldots,\phi_n^b} \cup
+%      %\setit{\phi_1,\ldots,\phi_j}
+%      \setit{\phi^o_1,\ldots,\phi^o_k}}
+%    =
+%    \condd{\progtt(\prog)}{S_{\setit{\cid_a}}}{S_{\setit{\cid_b}} \cup \setit{\phi_1^b,\ldots,\phi_n^b,x_b} \cup
+%      %\setit{\phi_1,\ldots,\phi_j}
+%      \setit{\phi^o_1,\ldots,\phi^o_k}}$
+%    \item $\condd{\progtt(\prog)}{S_{\setit{\cid_a}}}{\setit{\recon(x_1,x_2)}} =
+%      \condd{\progtt(\prog)}{S_{\setit{\cid_a}}}{\setit{\recon(x_1,x_2), x_b}}$
+%    \end{enumerate}
+%\end{definition}
+
 Next we show that, given well-formed typing environments, share type judgements preserve
@@
   \label{lemma-sty-noninteractive-sound}
-  Given $\prog,\Gamma,R : {OK}$ where $\secrets(\prog) = S$ and:
-  $$
-  \Gamma = \setit{[\phi_1^1,\phi_1^2] : \sharety, \ldots, [\phi_n^1,\phi_n^2] : \sharety}
-  \cup \setit{\pubx{\phi_1} : \pubty,\ldots,\pubx{\phi_j} : \pubty} \cup
-  \setit{\pubx{\phi^o_1} : \outty,\ldots,\pubx{\phi^o_k} : \outty}
-  $$
-  and also $\Gamma, \toeq{\prog} \vdash [\phi^1,\phi^2] : \sharety$
-  Then for all $a,b \in \setit{1,2}$ the following conditions hold:
+  Given $\prog,\Gamma,R : {OK}$ where $\secrets(\prog) = S$. For all $a,b\in \setit{1,2}$ with $a \ne b$ let:
+  \begin{eqnarray*}
+    %\Phi_a^m & \defeq & \setit{ \phi_a \mid ([\phi_1,\phi_2] : \sharety) \in \Gamma} \\
+    %\Phi_a^p & \defeq & \setit{ \phi_b \mid (\pubx{\recon(\phi_1,\phi_2)} : \sty) \in \Gamma}\\
+    \Phi_a & \defeq & \setit{ \phi_a \mid ([\phi_1,\phi_2] : \sharety) \in \Gamma} \cup
+                        \setit{ \phi_b \mid (\pubx{\recon(\phi_1,\phi_2)} : \sty) \in \Gamma} \\% \Phi_a^m \cup \Phi_a^p \\
+    \Phi^o & \defeq & \setit{ \recon(\phi_1,\phi_2) \mid (\pubx{\recon(\phi_1,\phi_2)} : \outty) \in \Gamma}
+  \end{eqnarray*} 
+  and also suppose $\Gamma, \toeq{\prog} \vdash [\phi_1,\phi_2] : \sharety$.
+  Then for all $a,b \in \setit{1,2}$ with $a \ne b$ the following conditions hold:
     \begin{enumerate}[\hspace{5mm}i.]
-    \item $\condd{\progtt(\prog)}{S_{\setit{\cid_a}}}{S_{\setit{\cid_b}} \cup \setit{\phi_1^b,\ldots,\phi_n^b} \cup
-      %\setit{\phi_1,\ldots,\phi_j}
-      \setit{\phi^o_1,\ldots,\phi^o_k}}
+    \item
+    $
+    \condd{\progtt(\prog)}{S_{\setit{\cid_a}}}{S_{\setit{\cid_b}} \cup \Phi^o}
     = 
-    \condd{\progtt(\prog)}{S_{\setit{\cid_a}}}{S_{\setit{\cid_b}} \cup \setit{\phi_1^b,\ldots,\phi_n^b,\phi^b} \cup
-      % \setit{\phi_1,\ldots,\phi_j}
-      \setit{\phi^o_1,\ldots,\phi^o_k}}$
-    \item $\condd{\progtt(\prog)}{S_{\setit{\cid_a}}}{\setit{\recon(\phi_i^1,\phi_i^2)}} =
-      \condd{\progtt(\prog)}{S_{\setit{\cid_a}}}{\setit{\recon(\phi_i^1, \phi_i^2), \phi_i^b}}$
+    \condd{\progtt(\prog)}{S_{\setit{\cid_a}}}{S_{\setit{\cid_b}} \cup  \Phi_a \cup \setit{\phi_a} \cup \Phi^o}
+    $
+  \item 
+    $\condd{\progtt(\prog)}{S_{\setit{\cid_a}}}{S_{\setit{\cid_b}} \cup  \Phi_a \cup \Phi^o \cup \setit{\recon(\phi_1,\phi_2)}} =
+    \condd{\progtt(\prog)}{S_{\setit{\cid_a}}}{S_{\setit{\cid_b}}  \cup  \Phi_a \cup \Phi^o \cup \setit{\recon(\phi_1,\phi_2), \phi_b}}$
     \end{enumerate}

### 2025-11-24 toplas26/sty.tex
checkpoint

@@
   \label{def-genshares}
-  We say that $\genshares$ is \emph{sound} iff given any $\prog,\Gamma,R : {OK}$
-  and $\phi$ and $\rx{w}{\cid} \not \in R$ with $\genshares(\phi,\rx{w}{\cid}) =
-  \phi_1,\phi_2$ and $\toeq{\prog} \models x_1 \eop \phi_1 \wedge x_2 \eop \phi_2$
-  with $\secrets(\prog) = S$ and:
+  We say that $\genshares$ is \emph{sound} iff given any $\prog$ with
+  $\vars(\prog) = S \cup V \cup O \cup R$
+  and $\phi$ and $\rx{w}{\cid} \not \in R$ and $\genshares(\phi,\rx{w}{\cid}) =
+  \phi_1,\phi_2$ we have for all $a \in \setit{1,2}$:
   $$
-  \Gamma = \setit{[\phi_1^1,\phi_1^2] : \sharety, \ldots, [\phi_n^1,\phi_n^2] : \sharety}
-  \cup \setit{\pubx{\phi_1} : \pubty,\ldots,\pubx{\phi_j} : \pubty} \cup
-  \setit{\pubx{\phi^o_1} : \outty,\ldots,\pubx{\phi^o_k} : \outty}
+  \sep{\progd(\prog)}{(S \cup V \cup O \cup \setit{\phi})}{\setit{\phi_a}}
   $$
-  we have for all $a,b \in \setit{1,2}$:
-  \begin{enumerate}[\hspace{5mm}i.]
-    \item $\condd{\progtt(\prog)}{S_{\setit{\cid_a}}}{S_{\setit{\cid_b}} \cup \setit{\phi_1^b,\ldots,\phi_n^b} \cup
-      %\setit{\phi_1,\ldots,\phi_j}
-      \setit{\phi^o_1,\ldots,\phi^o_k}}
-    =
-    \condd{\progtt(\prog)}{S_{\setit{\cid_a}}}{S_{\setit{\cid_b}} \cup \setit{\phi_1^b,\ldots,\phi_n^b,x_b} \cup
-      %\setit{\phi_1,\ldots,\phi_j}
-      \setit{\phi^o_1,\ldots,\phi^o_k}}$
-    \item $\condd{\progtt(\prog)}{S_{\setit{\cid_a}}}{\setit{\recon(x_1,x_2)}} =
-      \condd{\progtt(\prog)}{S_{\setit{\cid_a}}}{\setit{\recon(x_1,x_2), x_b}}$
-    \end{enumerate}
 \end{definition}

### 2025-11-24 toplas26/sty.tex
checkpoint

@@
 
-To proceed, we define a well-formedness property of tuples $\prog,\Gamma,R$
+Now we define what we mean by soundness of $\genshares$. In essence
+we mean perfect secrecy \cite{barthe2019probabilistic} in the context
+of protocols where multiple encodings can occur. These
+conditions are obviously met by, e.g., ``vanilla'' additive or Shamir
+shares which generate shares in uniform marginal distributions.
+\begin{definition}
+  \label{def-genshares}
+  We say that $\genshares$ is \emph{sound} iff given any $\prog$ with
+  $\vars(\prog) = S \cup V \cup O \cup R$
+  and $\phi$ and $\rx{w}{\cid} \not \in R$ and $\genshares(\phi,\rx{w}{\cid}) =
+  \phi_1,\phi_2$ we have for all $a \in \setit{1,2}$:
+  $$
+  \sep{\progd(\prog)}{(S \cup V \cup O \cup \setit{\phi})}{\setit{\phi_a}}
+  $$
+\end{definition}
+
+Next we define a well-formedness property of tuples $\prog,\Gamma,R$
 intuitively meaning that all the share types in $\Gamma$ are sound with respect
 to $\prog$, and consumed randomness in $R$ as OTPs for initial sharing.
-Property (i) ensures that views do not reveal information about secrets.
-Property (ii) ensures the key property of sharing-- that individual shares
+This is a representation invariant that we will show is preserved by
+typing, and under closure conditions (Definition \ref{def-sty-closed})
+guarantees confidentiality (Theorem \ref{theorem-OK})$.
+Property (i) ensures that views do not reveal information about secrets
+consistent with noninterference modulo output.
+Property (ii) ensures a key property of sharing-- that individual shares
 don't reveal more about secrets than their reconstruction does. Property
@@
 %\end{definition}
-With the definition of ${OK}$ in place, we can define what we mean by soundness
-of $\genshares$. In this definition, the first condition establishes that
-initial shares are independent of honest secrets, and the second establishes
-that they reveal no more than their reconstruction. These conditions are
-obviously met by ``vanilla'' additive or Shamir shares which generate
-shares in uniform marginal distributions.
-
-\begin{definition}
-  \label{def-genshares}
-  We say that $\genshares$ is \emph{sound} iff given any $\prog$ with
-  $\vars(\prog) = S \cup V \cup O \cup R$
-  and $\phi$ and $\rx{w}{\cid} \not \in R$ and $\genshares(\phi,\rx{w}{\cid}) =
-  \phi_1,\phi_2$ we have for all $a \in \setit{1,2}$:
-  $$
-  \sep{\progd(\prog)}{(S \cup V \cup O \cup \setit{\phi})}{\setit{\phi_a}}
-  $$
-\end{definition}
 
@@
   which satisfies relevant conditions (i-ii,iv) of Definition \ref{def-OK}. In case
-  \TirName{SecOpen} the public value must be reconstructed shares so Lemma
-  \ref{lemma-sty-noninteractive-sound} relevant conditions (i-ii), and the
-  the requirement that the public value be masked with an OTP satisfies conditions
+  \TirName{SecOpen} the public value must be reconstructed shares, so Lemma
+  \ref{lemma-sty-noninteractive-sound} enforces relevant conditions (i-ii) of Definition
+  \ref{def-OK}, and the requirement that the public value be masked with an OTP satisfies conditions
   (iii) and (iv) of Definition \ref{def-OK}. In case \TirName{Reveal} since

### 2025-11-24 toplas26/sty.tex
checkpoint

@@
 clarity of presentation and the analysis can be generalized to more parties.
+We define the federation as $\setit{1,2}$ (client 1 and client 2).
 
@@
 typing, and under closure conditions (Definition \ref{def-sty-closed})
-guarantees confidentiality (Theorem \ref{theorem-OK})$.
+guarantees confidentiality (Theorem \ref{theorem-OK}).
 Property (i) ensures that views do not reveal information about secrets
@@
 public values as secure openings, and all outputs as reconstructed
-shares. We capture this formally with the following notion of
+shares by adding them to $\Gamma$. We capture this formally with the following notion of
 type environment closure.
@@
 \begin{proof}
-  Let $\iov(\prog) = S \cup V \cup O$ and assume some partitioning $H,C$.
+  Let $\iov(\prog) = S \cup V \cup O$ and assume some partitioning $H,C$
+  with $H = \setit{b}$ and $C = \setit{a}$ for $a,b \in \setit{1,2}$.
   %For any $\out{\cid} \in O$ with $\Gamma,\toeq{\prog} \vdash \pubx{\out{\cid}} : \outty$
@@
   %\ref{lemma-OK}.
-  By Definitions \ref{def-sty-closed} we know that every $x \in \houtputs$
-  occurs in some $\stt$ with $(\stt, \sharety) \in \Gamma$, and
-  that $(\out{\cid} : \outty) \in \Gamma$ for all $\out{\cid} \in O$.
+  Let $\Phi_a$ and $\Phi^o$ be as defined in Definition \ref{def-OK}. 
+  By Definitions \ref{def-sty-closed} we have that $x \in \houtputs$
+  implies $x \in \Phi_a$ and $(\out{\cid} : \outty) \in \Gamma$ for all
+  $\out{\cid} \in O$. 
   Therefore by Definition \ref{def-OK} (and marginalizing away
   extraneous $(\stt : \sty) \in \Gamma$) we have:
+  $$
+  \condd{\progd(\prog)}{S_H}{S_C \cup O} = \condd{\progd(\prog)}{S_H}{S_C \cup V_C \cup \Phi^p \cup O}
+  $$
+  By definition of $\TirName{Output}$ it is also the case that for all
+  $\out{\cid} \in O$ we have $\toeq{\prog} \models o = \phi$ for some
+  $\phi \in \Phi^o$. Therefore we have:
   $$

### 2025-11-24 toplas26/sty.tex
checkpoint

@@
   Let $\Phi_a$ and $\Phi^o$ be as defined in Definition \ref{def-OK}. 
-  By Definitions \ref{def-sty-closed} we have that $x \in \houtputs$
+  By Definition \ref{def-sty-closed} we have that $x \in \houtputs$
   implies $x \in \Phi_a$ and $(\out{\cid} : \outty) \in \Gamma$ for all
@@
   $$
-  \condd{\progd(\prog)}{S_H}{S_C \cup O} = \condd{\progd(\prog)}{S_H}{S_C \cup V_C \cup \Phi^p \cup O}
+  \condd{\progd(\prog)}{S_H}{S_C \cup O} = \condd{\progd(\prog)}{S_H}{S_C \cup \houtputs \cup \Phi^p \cup O}
   $$
@@
   $$
-  \condd{\progd(\prog)}{S_H}{S_C \cup O} = \condd{\progd(\prog)}{S_H}{S_C \cup V_C \cup O}
+  \condd{\progd(\prog)}{S_H}{S_C \cup O} = \condd{\progd(\prog)}{S_H}{S_C \cup \houtputs \cup O}
   $$

### 2025-11-24 toplas26/smt.tex
checkpoint

@@
 \begin{proof}
-  \cnote{de-obfuscate.}
-  Assume $\htrip{\eqs_1}{\prog}{\eqs_2}$ is derivable without \TirName{Hyp}.
-  Then $\eqs_1 \wedge \toeq{\prog} \models \eqs_2$, since all the other rules
-  are well-known to be validity preserving. 
-  Now since $\eqs_1 \wedge \toeq{\prog} \models \eqs_2$ by assumption, we have
-  $
+  By induction on the derivation and cases analysis on the last derivation
+  step. We only need to consider $\TirName{Hyp}$ since other cases are standard.
+  In this case the consequent is of the form
+  $\htrip{\eqs_1 \wedge (\eqs_2 \impl \eqs)}{\prog}{\eqs_2 \wedge \eqs}$
+  where $\htrip{\eqs_1}{\prog}{\eqs_2}$ is derivable by inversion
+  of \TirName{Hyp} and hence valid by the induction hypothesis.
+  Thus $\eqs_1 \wedge \toeq{\prog} \models \eqs_2$,
+  %, and
+  % therefore in particular $\eqs_1 \wedge (\eqs_2 \impl \eqs) \wedge \toeq{\prog}$
+  % iff $\eqs_1 \wedge \eqs \wedge \toeq{\prog}$. Now 
+  and we can strengthen the condition on the rhs of $\models$:
+  $$
   \eqs_1 \wedge (\eqs_2 \impl \eqs) \wedge \toeq{\prog} \models \eqs_2 
-  $,
-  and clearly:
+  $$
+  so by standard reasoning we have:
   $$

### 2025-11-24 toplas26/lang.tex
checkpoint

@@
 protocols between a federation of \emph{clients} exchanging values in
-the binary field. We identify clients by natural numbers and federations- finite sets of
-clients- are always given statically.  Our threat model assumes a
-partition of the federation into \emph{honest} $H$ and \emph{corrupt}
-$C$ subsets. We model probabilistic programming via a \emph{random
-tape} semantics. That is, we will assume that programs can make
-reference to values chosen from a uniform random distributions defined
-in the initial program memory.  Programs aka protocols execute
+a finite field. In order to align with existing SMT theories
+\cite{SMFF} we restrict our consideration to prime fields. We identify
+clients by natural numbers and federations- finite sets of clients-
+are always given statically.  Our threat model assumes a partition of
+the federation into \emph{honest} $H$ and \emph{corrupt} $C$
+subsets. We model probabilistic programming via a random tape
+semantics. That is, we will assume that programs can make reference to
+values chosen from a uniform random distributions defined in the
+initial program memory.  Programs aka protocols execute
 deterministically given the random tape.
@@
 ill-defined. That is, we generally assume that for any
-configuration $(\store, \prog)$ there exists $\store$
+program $\prog$ there exists $\store$
 with $(\store, \prog) \redxs \store$. This rules out

### 2025-11-24 toplas26/intro.tex
attempt at intro

--- 2025-01-26 esop25/intro.tex
@@
 Our work is based on the $\metaprot$/$\minicat$ framework developed in
-previous work \cite{skalka-near-ppdp24}. The prior work developed the 
+previous work \cite{skalka-near-ppdp24, skalka-near-esop25}. The prior work developed the 
 language and semantics, plus two mechanisms for verifying security: an
@@
 
-Together, our contributions \emph{automate} the verification of security
+Together, our contributions \emph{automate} the \emph{compositional} verification of security
 properties in $\metaprot$/$\minicat$. In this work, we build on~\cite{skalka-near-ppdp24}
@@
 finite fields and scale to large protocols. To accomplish this, we develop new
-type systems for compositional verification, and use SMT to automate the checks
+Hoare logics and type systems for compositional verification, and use SMT to automate the checks
 that were accomplished by enumeration in prior work. Specifically, we make the
@@
 %
+% \begin{enumerate}[i.]
+% \item In Section \ref{section-smt}, an interpretation of protocols in
+%   $\minicat$ as SMT constraints (Theorem \ref{theorem-toeq}) and
+%   methodology for verifying correctness properties.
+% \item In Section \ref{section-cpj}, a confidentiality type system for
+%   $\minicat$ protocols with a soundness property guaranteeing that
+%   information is released only through explicit declassifications
+%   (Theorem \ref{theorem-cpj}).
+% \item In Section \ref{section-ipj}, an integrity type system for
+%   $\minicat$ protocols with a soundness property guaranteeing
+%   robustness in malicious settings (Theorem \ref{theorem-ipj}).
+% \item In Section \ref{section-metalang}, a dependent Hoare type system and
+%   algorithm for the $\metaprot$ language that is sound for the
+%   confidentiality and integrity type systems (Theorem
+%   \ref{theorem-mtj}), with independently verified, compositional $\Pi$
+%   types for functions.
+% \end{enumerate}
 \begin{enumerate}[i.]
-\item In Section \ref{section-smt}, an interpretation of protocols in
-  $\minicat$ as SMT constraints (Theorem \ref{theorem-toeq}) and
-  methodology for verifying correctness properties.
-\item In Section \ref{section-cpj}, a confidentiality type system for
-  $\minicat$ protocols with a soundness property guaranteeing that
-  information is released only through explicit declassifications
-  (Theorem \ref{theorem-cpj}).
-\item In Section \ref{section-ipj}, an integrity type system for
-  $\minicat$ protocols with a soundness property guaranteeing
-  robustness in malicious settings (Theorem \ref{theorem-ipj}).
-\item In Section \ref{section-metalang}, a dependent Hoare type system and
-  algorithm for the $\metaprot$ language that is sound for the
-  confidentiality and integrity type systems (Theorem
-  \ref{theorem-mtj}), with independently verified, compositional $\Pi$
-  types for functions.
+\item In Section~\ref{section-smt}, a Hoare logic for Overture protocols, for the compositional verification of both semi-honest and malicious security.
+\item In Section~\ref{section-sty}, a novel share type system, enabling security verification for protocols that use secret sharing as a building block; our approach works for any secret sharing scheme and enforces noninterference modulo output.
+\item In Section~\ref{section-ipj}, an integrity type system for extending semi-honest security to malicious security.
+\item In Section~\ref{section-metalang}, a Hoare logic for Prelude programs with an automated implementation, for the compositional verification of high-level protocol specifications.
+\item In Section~\ref{sec:metapr-type-check}, a compositional type-based verification approach for Prelude programs that combines share types with the automated Hoare logic of Section 7.
 \end{enumerate}
+%
+A prior conference version of this work~\cite{skalka-near-esop25}
+proposed a whole-program analysis for Prelude programs that ensures
+confidentiality, and combined it with an integrity type system to
+prove security of Prelude programs. This extended version adds the
+compositional Hoare logics presented in Sections 4 and 7, adds the
+share type system presented in Section 5, and fully develops the
+compositional type-based verification approach in Section 8. In
+addition, we develop an implementation capable of automating the
+verification of examples given in the prior work, and our extensions
+in this work additionally enable the type-based verification of
+noninterference modulo output for secret sharing protocols that
+leverage Beaver triples (see Section~\ref{section-examples}).
+
 In all cases, our type systems are boosted by verification mechanisms

### 2025-11-24 toplas26/smt.tex
checkpoint

@@
 defined in Figure \ref{fig-toeq}. The interpretation of OT is omitted
-from this figure which is general. For $\mathbb{F}_2$ (where we allow
-negation $\neg$ of expressions) the interpretation is:
+from this figure which is general. For $\mathbb{F}_2$ (using 
+$\neg$ to denote negation) the interpretation is:
 \begin{mathpar}
@@
 a correctness (post)condition $\eqs$ if its SMT interpretation
-entails $\eqs$.
+together with $\eqspre$ entails $\eqs$.
 
@@
 \noindent This result follows as a consequence of a more general Lemma
-that incorporates a notion of preconditions. 
+that incorporates a notion of preconditions:
 \begin{lemma}
@@
 \end{lemma}
-\cnote{Prove me?}
 
@@
 
-When reasoning about MPC protocols it is important to reason about
-postconditions, e.g., to verify correctness conditions. It is also
-important to reason about preconditions of protocols, e.g., to express
-properties of assumed initialization protocols such as Beaver Triples.
-To support this reasoning we introduce a Hoare logic for $\minicat$,
+To reasoning effectively about pre and postconditions we introduce
+a Hoare logic for $\minicat$,
 where $\htrip{\eqs_1}{\prog}{\eqs_2}$ means that $\eqs_1,\eqs_2$ are
@@
 Now we can introduce a validity-preserving derivation system for the
-$\minicat$ Hoare logic. Since the language is computationally simple
-the Hoare logic can be relatively simple and mostly standard.
-The derivation rules are specified in Figure \ref{fig-minicathtrip}.
-They are mostly familiar except perhaps for \TirName{Assert}, which
-just enforces the expected equality as a precondition, and
-\TirName{Hyp}. It is unknown to the authors and not immediately
-obvious whether \TirName{Hyp} is derivable from other rules, but
-it is validity preserving as we show below. Its main use in this
-paper will be to enable an automated Hoare logic check we develop
+$\minicat$ Hoare logic.  The derivation rules are specified in Figure
+\ref{fig-minicathtrip}, where in rule \TirName{Frame} we define
+$\avars(\prog)$ as the variables in $\prog$ that are assigned to.  The
+rules are mostly standard except for \TirName{Assert} which includes
+the expected equality in a precondition, and \TirName{Hyp}.  It is
+possible that \TirName{Hyp} is derivable, but in any case it is
+validity preserving as we show in the following Lemma. Its main use in
+this paper will be to enable an automated Hoare logic check we develop
 later (Section \ref{section-metahtrip}).
@@
   % iff $\eqs_1 \wedge \eqs \wedge \toeq{\prog}$. Now 
-  and we can strengthen the condition on the rhs of $\models$:
-  $$
+  and we can strengthen the condition on the rhs of the entailment as $
   \eqs_1 \wedge (\eqs_2 \impl \eqs) \wedge \toeq{\prog} \models \eqs_2 
-  $$
+  $,
   so by standard reasoning we have:
-  $$
+  $
   \eqs_1 \wedge (\eqs_2 \impl \eqs) \wedge \toeq{\prog} \models \eqs_2 \wedge (\eqs_2 \impl \eqs)
-  $$
+  $,
   and since $\eqs_2 \wedge (\eqs_2 \impl \eqs)$ iff

### 2025-11-24 toplas26/sty.tex
checkpoint

@@
 
-Our approach to static confidentiality analysis is to enforce the use of sound
-techniques in semi-homomorphic encryption in protocols.  As discussed in
-the example in Section \ref{section-smt-example}, protocols integrate
-interactive elements, such as the sharing of one-time-pad (OTP)
-encrypted secrets, and noninteractive elements, such as addition (of
-shares). To comport with this we introduce distinct judgement forms for
-interactive and noninteractive protocol elements. Our type system
-as presented will assume 2-client protocols, but this is for
-clarity of presentation and the analysis can be generalized to more parties.
-We define the federation as $\setit{1,2}$ (client 1 and client 2).
+Our approach to static confidentiality analysis is to enforce the use
+of sound techniques in semi-homomorphic encryption in protocols.  As
+discussed in the example in Section \ref{section-smt-example},
+protocols integrate interactive elements, such as the sharing of
+one-time-pad (OTP) encrypted secrets, and noninteractive elements,
+such as addition (of shares). To comport with this we introduce
+distinct judgement forms for interactive and noninteractive protocol
+elements. Our type system as presented will assume 2-client protocols,
+but this is for clarity of presentation and the analysis can be
+generalized to more parties.  So in this Section we define the
+federation as $\setit{1,2}$ (client 1 and client 2).
 
@@
 A \emph{secret sharing scheme} includes functions
-$(\genshares,\recon)$ which generate and reconstruct shares,
+$(\genshares,\recon)$ that generate and reconstruct shares,
 respectively. The $\genshares$ function takes some
@@
 in Section \ref{section-smt}, the output can also be connected to an
-intended ideal functionality.\cnote{Update intro grid accordingly.}
+intended ideal functionality.
 

### 2025-11-24 toplas26/frontmatter.tex
checkpoint

@@
-\title{SMT-Boosted Security Types for Low-Level MPC}
+\title{SMT-Boosted Security Types and Correctness for Low-Level MPC}
 
@@
 \email{ceskalka@uvm.edu}
-
-\author{Joseph P.~Near}
-\affiliation{%
-  \institution{University of Vermont}
-  \city{Burlington, Vermont}
-  \country{USA}}
-\email{jnear@uvm.edu}
 
@@
 
+\author{Joseph P.~Near}
+\affiliation{%
+  \institution{University of Vermont}
+  \city{Burlington, Vermont}
+  \country{USA}}
+\email{jnear@uvm.edu}
+
 \begin{abstract}
@@
   technology for data privacy in modern distributed applications. We
-  develop a new type theory to automatically enforce correctness,
-  confidentiality, and integrity properties of protocols written in
-  the \emph{Prelude/Overture} language framework. Judgements in the
-  type theory are predicated on SMT verifications in a theory of
-  finite fields, which supports precise and efficient analysis. Our
-  approach is automated, compositional, scalable, and generalizes to
-  arbitrary prime fields.
+  develop a new Hoare logic and type theory to automatically enforce
+  correctness, confidentiality, and integrity properties of protocols
+  written in the \emph{Prelude/Overture} language framework. Both
+  definition and automation of our analyses leverages SMT verification
+  in a theory of finite fields, supporting precision and
+  efficiency. Our approach is compositional, scalable, and generalizes
+  to arbitrary prime fields and secret sharing schemes.
 \end{abstract}

### 2025-11-24 toplas26/intro.tex
checkpoint

@@
 clear relevance to verification of cryptographic schemes that rely on
-field arithmetic. The goal of this paper is to combine these two
-approaches in a decidable type system for verifying correctness and
-security properties of MPC protocols.
+field arithmetic. The goal of this paper is to combine hyperproperty
+formulations of security with SMT verification to enforce correctness
+and security properties of MPC protocols.
 

### 2025-11-24 toplas26/sty.tex
checkpoint

@@
 $$
-In order for $\genshares$ to be \emph{sound}, we require
+In order for $\genshares$ to be \emph{sound} with regard to
+confidentiality, we require
 that both of $\phi_1$ and $\phi_2$ are marginally independent
@@
 soundness requires that the constraint $\eqs$ must be entailed by
-$\eqspre \wedge \toeq{\prog}$ (see, e.g., Lemma
-\ref{lemma-sty-noninteractive-sound}).
-
-The syntax-directed rules are all known semi-homomorphic operations,
+$\toeq{\prog}$. The syntax-directed rules are all known semi-homomorphic operations,
 including addition of shares (\TirName{HEAdd}), and both addition and
@@
 These rules also leverage SMT entailment in the \TirName{ShareEntails}
-and \TirName{PubEntails} rules, since these operations should be
+and \TirName{PubEntails} rules, meaning these operations are
 allowable up to algebraic equivalence in the context of a protocol.
@@
   x_2] : \sharety$ if the term is a secret sharing as defined
-by $\genshares$, and adds $\rx{y}{\cid_1}$ to the set of variables
+by $\genshares$, and adds $\rx{w}{\cid}$ to the set of variables
 consumed by $\genshares$, ensuring one-time usage. The
-\TirName{Reveal} rule generates the binding $\pubx{\recon(\px{x},\px{y})} :
-\outty$ with the requirement that $[\px{x},\px{y}]$ are typable as
+\TirName{Reveal} rule generates the binding $\pubx{\recon(\px{w_1},\px{w_2})} :
+\outty$ with the requirement that $[\px{w_1},\px{w_2}]$ are typable as
 shares. The \TirName{SecOpen} rule is similar except it allows the
-assertion $\pubx{\recon(\px{x},\px{y})} : \pubty$ if we can additionally
+assertion $\pubx{\recon(\px{w_1},\px{w_2})} : \pubty$ if we can additionally
 show that the shared value is masked by the use of random value
-$\rx{z}{\oid}$ as an OTP. Here we are positing a third party $\oid$
-that secret-shares $\rx{z}{\oid}$ as part of an initialization
+$\rx{w}{\oid}$ as an OTP. Here we are positing a third party $\oid$
+that secret-shares $\rx{w}{\oid}$ as part of an initialization
 protocol. We illustrate secure openings in practice at more length
@@
 \begin{eqnarray*}
-  \genshares(\phi,\rx{y}{\cid}) &\defeq& \phi - \rx{y}{\cid}, \rx{y}{\cid}\\
+  \genshares(\phi,\rx{w}{\cid}) &\defeq& \phi - \rx{w}{\cid}, \rx{w}{\cid}\\
   \recon(\phi_1,\phi_2) &\defeq& \phi_1 \fplus \phi_2
 \end{eqnarray*}
-where we require that $\genshares$ consumes the randomness
-$\setit{\rx{y}{\cid_2}}$ for use as an OTP.
+where we require that $\genshares$ consumes the random value
+$\rx{w}{\cid}$ as an OTP.
 
-This approach highlights an imporant interaction between our type
+This approach highlights an important interaction between our type
 system and SMT entailment as defined by the $\TirName{ShareEntails}$
@@
 (YGC) in $\mathbb{F}_{2}$, the ``garbler'' represents encrypted wire
-values as a random sample (for 1) or its negation (for 0), and shares
+values as a random sample (for 0) or its negation (for 1), and shares
 its secret input encoding with the ``evaluator'' by using
@@
 type system we return to the example of additive sharing in the context
-of a 2-client protocol. We first focus on initial secret sharing steps--
+of a 2-client protocol. While the ideal functionality of this simple protocol
+precisely reveals input secrets, it serves as an understandable example
+of the type system mechanics.
+We first focus on initial secret sharing steps--
 let $\prog$ be the following protocol in $\fieldp{2}$ (noting
@@
 typing, and under closure conditions (Definition \ref{def-sty-closed})
-guarantees confidentiality (Theorem \ref{theorem-OK}).
+the invariant guarantees confidentiality (Theorem \ref{theorem-OK}).
 Property (i) ensures that views do not reveal information about secrets

### 2025-11-24 toplas26/intro.tex
checkpoint

--- 2025-11-24 toplas26/intro.tex
@@
 clear relevance to verification of cryptographic schemes that rely on
-field arithmetic. The goal of this paper is to combine these two
-approaches in a decidable type system for verifying correctness and
-security properties of MPC protocols.
+field arithmetic. The goal of this paper is to combine hyperproperty
+formulations of security with SMT verification as a foundation for
+enforcing correctness and security properties of MPC protocols.
 
@@
 automated approach for protocols in $\mathbb{F}_2$ (the binary field),
-based on enumerating all possible executions of the protocol, that scales 
+based on enumerating all possible executions of the protocol that scales 
 only to very small protocols; and a manual approach based on a program logic.

### 2025-11-24 toplas26/intro.tex
checkpoint

@@
 \item In Section~\ref{section-smt}, a Hoare logic for Overture protocols, for the compositional verification of both semi-honest and malicious security.
-\item In Section~\ref{section-sty}, a novel share type system, enabling security verification for protocols that use secret sharing as a building block; our approach works for any secret sharing scheme and enforces noninterference modulo output.
-\item In Section~\ref{section-ipj}, an integrity type system for extending semi-honest security to malicious security.
+\item In Section~\ref{section-sty}, a novel share type system, enabling security verification for protocols that use secret sharing as a building block; our approach works for any secret sharing scheme and enforces passive security (confidentiality).
+\item In Section~\ref{section-ipj}, an integrity type system for extending passive security to malicious security.
 \item In Section~\ref{section-metalang}, a Hoare logic for Prelude programs with an automated implementation, for the compositional verification of high-level protocol specifications.
@@
 compositional Hoare logics presented in Sections 4 and 7, adds the
-share type system presented in Section 5, and fully develops the
-compositional type-based verification approach in Section 8. In
-addition, we develop an implementation capable of automating the
-verification of examples given in the prior work, and our extensions
-in this work additionally enable the type-based verification of
-noninterference modulo output for secret sharing protocols that
-leverage Beaver triples (see Section~\ref{section-examples}).
+share type system presented in Section 5 that enforces a stronger
+security property than the type system in~\cite{skalka-near-esop25},
+and fully develops the compositional type-based verification approach
+in Section 8. In addition, we develop an implementation capable of
+automating the verification of examples given in the prior work, and
+our extensions in this work additionally enable the type-based
+verification of noninterference modulo output for more complex secret
+sharing protocols,e.g., that leverage Beaver triples (see
+Section~\ref{section-examples}).
 
@@
 applications of our type analyses to real protocols including
-the Goldreich-Micali-Wigderson (GMW), Bendlin-Damgard-Orland-Zakarias
-(BDOZ), and Yao's Garbled Circuits (YGC) protocols.
+the Goldreich-Micali-Wigderson (GMW), Ben-Or-Goldwasser-Widgerson (BGW),
+and Bendlin-Damgard-Orland-Zakarias (BDOZ), and report related
+performance metrics.
 
@@
 
-The most closely related work is \cite{skalka-near-ppdp24}, where
-foundations for $\minicat/\metaprot$ are introduced including a
-verification algorithm. However, this algorithm only works for
-$\mathbb{F}_2$-- e.g., binary circuits-- and does not scale due to
-exponential complexity, necessitating semi-automated proof
-techniques. In contrast, our type systems scale to arbitrary prime
-fields and whole program analysis for larger circuits (protocols). So
-our work is a significant advancement of $\minicat/\metaprot$.  Our
-work is also closely related to Satisfiability Modulo Finite Fields
-\cite{SMFF} but only in the sense that we apply their method as we
-discuss in Section \ref{section-smt} and subsequently.
+The current work extends previous work \cite{skalka-near-esop25} as
+discussed above. Key differences as noted in Figure
+\ref{fig-comp-wrap} include a Hoare logic and a more expressive
+confidentiality type systems supporting analysis of more complex
+programs that enforces passive security in the real/ideal model, vs.~a
+weaker gradual release property as in \cite{skalka-near-esop25}.
+Furthermore our Hoare logic analysis has been fully implemented in the
+current work in an Antlr/Scala/cvc5-based framework.
+
+Besides \cite{skalka-near-esop25} the most closely related work is
+\cite{skalka-near-ppdp24}, where foundations for $\minicat/\metaprot$
+are introduced including a verification algorithm. However, this
+algorithm only works for $\mathbb{F}_2$-- e.g., binary circuits-- and
+does not scale due to exponential complexity, necessitating
+semi-automated proof techniques. In contrast, our type systems scale
+to arbitrary prime fields and whole program analysis for larger
+circuits (protocols). So our work is a significant advancement of
+$\minicat/\metaprot$.  Our work is also closely related to
+Satisfiability Modulo Finite Fields \cite{SMFF} in that we use it as
+an enabling technology as discussed in Section \ref{section-smt} and
+subsequently.
 

### 2025-11-24 toplas26/conclusion.tex
checkpoint

\section{Conclusion and Future Work}

In this paper we have developed the $\minicat$ language for defining
low-level MPC protocols, and the $\metaprot$ metaprogramming language
for specifying $\minicat$ protocols compositionally and at a higher
level of abstraction. We have developed a Hoare logic based on
SMT-style verification for $\minicat$, as well as confidentiality and
integrity type systems that enforce passive and malicious security in
the real/ideal model. These establish a foundation for an automated
Hoare logic and type system in $\metaprot$ that is compositional,
scalable, and generalizes to arbitrary prime fields and secret sharing
schemes.

Our automated $\metaprot$ Hoare logic leverages Satisfiability
Modulo Finite Fields for efficiency, scalability, and generalizability
to arbitrary prime fields. We developed examples of real binary and
arithmetic circuit protocols, illustrated the applicability of
our analyses to enforce confidentiality and integrity properties,
and evaluated performance of our implementation on these examples.

### 2025-11-24 toplas26/metalang.tex
checkpoint

@@
 confidentiality and integrity properties expressed as types, as we
-will show in Section \ref{section-sharety}. We present and discuss
-extended examples of $\metaprot$ programs in Section
-\ref{section-examples}.
+will discuss in Sections \ref{section-metaty} and especially
+\ref{section-examples} where we present and discuss
+extended examples of $\metaprot$ programs.
 
@@
 in field $\mathbb{F}_p$, record values, and $\minicat$ variables
-$x$. $\metaprot$ expression forms allow dynamic construction of these
+$x$. $\metaprot$ \emph{expression} forms allow dynamic construction of these
 values. Note in particular that $\concat$ is string concatentation.
@@
 both a codebase and a ``main'' program that uses the codebase.
-We disallow recursion, mainly to guarantee decidability
-of type dependence (Section \ref{section-metalangty}).
+For all function (expression or instruction forms) we disallow
+recursion which is important guarantee decidability
+of type dependence (Section \ref{section-metaty}).
 \begin{definition}
@@
 
-Note that we do not include an evaluation
-rule for the form $\eqcast{\mx{e}{e}}{e}$ which is a type
-annotation that we assume is erased from programs prior to
-evaluation. We defer discussion of this form to Section
-\ref{section-sharety}
+%Note that we do not include an evaluation
+%rule for the form $\eqcast{\mx{e}{e}}{e}$ which is a type
+%annotation that we assume is erased from programs prior to
+%evaluation. We defer discussion of this form to Section
+%\ref{section-sharety}
 
@@
 
-$\metaprot$ instruction functions can optionally be annotated with
-pre and postconditions. To model this in our theory we posit
-functions $\precond$  and $\postcond$ 
-that specify conditions for instruction functions $f \in \dom(\codebase)$.
-Note that $\precond(f)$ or $\precond(f)$ may be undefined if no annotation
-is given for $f$. In actual code, we write pre and postconditions directly before and
-after functions to define them. Here is a simple example function $\ttt{bar}$
-with. Note that the pre and postconditions are abstract with respect to the
-input string $\ttt{x}$, and that string concatenation  
-can be used in pre and postcondition constraints:
+$\metaprot$ instruction functions can optionally be annotated with pre
+and postconditions. To model this in our theory we posit functions
+$\precond$ and $\postcond$ that specify conditions for instruction
+functions $f \in \dom(\codebase)$.  Note that $\precond(f)$ or
+$\precond(f)$ may be undefined if no annotation is given for $f$. In
+actual code, we write pre and postconditions directly before and after
+functions to define them. Here is a simple example function
+$\ttt{bar}$ with pre and postcondition annotations. Note that the
+annotations are abstract with respect to the input string $\ttt{x}$,
+and that string concatenation can be used in pre and postcondition
+constraints:
 \begin{verbatimtab}
@@
 the formal parameters of the function they annotate. We denote these
-\emph{abstract constraints} as $\eqs$ with the syntax given
+\emph{abstract constraints} as $\peq$ with the syntax given
 in Figure \ref{fig-peq}. When the function is applied, the pre and postconditions
@@
   $\mathit{rename}$ be the function that maps variables in $\eqs$ to
-  variables in $\eqs'$ such that $\mathit{rename}(\eqs) = \eqs'$. Assume
-  $\store' \models \eqs'$, and define $\store$ such that:
+  variables in $\eqs'$ such that $\mathit{rename}(\eqs) = \eqs'$. Suppose
+  on the contrary that 
+  $\store' \models \eqs'$ and $\store' \not\models \eqs$, and define $\store$ such that:
   $$
@@
 if they are defined. It is in fact not necessary for \emph{any} pre
-and postconditions to be defined, but their intent is to enable
-fine-tuning of postconditions for improving verification efficiency as
-discussed above and later in Section \ref{section-examples}.  While
-not relevant to the theory, an important implementation detail is that
-verification of any given function $f$ only needs to be performed
-once, and can then be memoized for other applications using
-the $\TirName{App}$ rule as a consequence of Lemma \ref{lemma-genentails}
+and postconditions to be defined, since the analysis will infer valid
+pre and postconditions in their absence by rule \TirName{Fn}, but
+their intent is to enable fine-tuning of pre and postconditions for
+improving verification efficiency as discussed above and later in
+Section \ref{section-examples}.  An important implementation detail is
+that verification of any given function $f$ only needs to be performed
+once, and can then be memoized for other applications using the
+\TirName{App} rule as a consequence of Lemma \ref{lemma-genentails}
 (and more precisely of Corollary \ref{cor-metahtrip}).
@@
   \end{mathpar}
-  and by inversion of the $\TirName{Seq}$ rule we have:
-  \begin{mathpar}
-    \htrip{\peq_1^1}{\cmd_1}{\peq_2^1}
-
-    \htrip{\peq_1^2}{\cmd_2}{\peq_2^2}
-  \end{mathpar}
-  and by the induction hypothesis we have:
-  \begin{mathpar}
-    \htrip{\eqs_1^1}{\cmd_1}{\eqs_2^1}
-
-    \htrip{\eqs_1^2}{\cmd_2}{\eqs_2^2}
-  \end{mathpar}
+  and by inversion of the $\TirName{Seq}$ rule we have $\htrip{\peq_1^1}{\cmd_1}{\peq_2^1}$ and $\htrip{\peq_1^2}{\cmd_2}{\peq_2^2}$,
+  %\begin{mathpar}
+  %  \htrip{\peq_1^1}{\cmd_1}{\peq_2^1}
+  %
+  %  \htrip{\peq_1^2}{\cmd_2}{\peq_2^2}
+  %\end{mathpar}
+  and by the induction hypothesis we have $\htrip{\eqs_1^1}{\cmd_1}{\eqs_2^1}$ and $\htrip{\eqs_1^2}{\cmd_2}{\eqs_2^2}$.
+  %\begin{mathpar}
+  %  \htrip{\eqs_1^1}{\cmd_1}{\eqs_2^1}
+  %
+  %  \htrip{\eqs_1^2}{\cmd_2}{\eqs_2^2}
+  %\end{mathpar}
   But by \TirName{Hyp} we have:
@@
   $\peq_1 \redx \eqs_1$ and $\peq_2 \redx \eqs_2$ then
-  $\htrip{\eqs_1}{\prog}{\eqs_2}$.
+  $\htrip{\eqs_1}{\prog}{\eqs_2}$ is valid.
 \end{corollary}

### 2025-11-24 toplas26/metaty.tex
attempt at intro
checkpoint

@@
 \section{$\metaprot$ Type Checking}
+\label{sec:metapr-type-check}
 
@@
 \begin{definition}
-  Define $\Gamma \Vdash \stt : \sty$ iff $\Gamma,\true \Vdash \stt : \sty$ without the
+  Define $\Gamma \Vdash \stt : \sty$ iff $\Gamma,\true \vdash \stt : \sty$ without the
   use of \TirName{ShareEntails} and \TirName{PubEntails}.
@@
 include. Type annotations are of the form $e_1 \kwas e_2$, which
-defines an SMT equivalence, and $e : \sty$ which are verified using
-the automated $\Vdash$ typing judgement-- the key idea here being that
-explicit equivalences can stand in for implicit \TirName{ShareEntails}
-and \TirName{PubEntails} judgements. In the semantics type annotations
-are interpreted as no-ops, and their integration in Hoare triple
+defines an SMT equivalence-- the key idea here being that explicit
+equivalences can stand in for implicit \TirName{ShareEntails} and
+\TirName{PubEntails} judgements-- and $\notg{\stt} : \sty$ which
+defines a share type assertion verified using the automated $\Vdash$
+typing judgement. In the $\metaprot$ semantics type annotations are
+interpreted as $\pskip$s, while their integration in Hoare triple
 derivations requires annotated equivalences to be included in
@@
 $\pubty$ type ascriptions (\TirName{LocalPublic} and \TirName{MaskedPublic},
-respectively). In these rules we use of definitions $\xdefs$ to refer to
+respectively). In these rules we use $\xdefs$ to recover defined
 SMT equivalences, and $\Vdash$ to verify share typing.
@@
 generalization and instantiation of share type environments and
-thus compositionality in typing. Generalized, aka abstract
+thus compositionality in typing. Generalized aka abstract
 share type environments are dependent on $\stringty$ and
@@
 \end{definition}
-
-We also introduce dependent $\Pi$ types for functions, that allow
+\noindent We also introduce dependent $\Pi$ types for functions, that allow
 dependence on $\stringty$ and $\cidty$ types.

### 2025-11-24 toplas26/ipj.tex
checkpoint

--- 2024-10-14 esop25/ipj.tex
@@
 Our main type soundness result for integrity is then given as follows.
+Proof details are available in the appendix of \cite{skalka-near-esop25}.
 \begin{theorem}
@@
   we have $\seclev(x) = \hilab$ for all $x \in \houtputs \cup O_H$, then $\prog$
-  has integrity for all $H,C$.
+  has integrity for all $H,C$ and $\adversary$.
 \end{theorem}

### 2025-11-24 toplas26/metaty.tex
checkpoint

@@
 \end{definition}
+
 \noindent We also introduce dependent $\Pi$ types for functions, that allow
@@
 \notg{R})$ specifies that, given share typing assumptions in
-$\notg{\Gamma_1}$, application the function results in the bindings in
-$\notg{\Gamma_2}$ by using the random values in $\notg{R}$. It is
-always the case that the type is closed, i.e., only $y_1,\ldots,y_n$
-can occur free in $\notg{\Gamma_1}$, $\notg{\Gamma_2}$ $\notg{R}$.
-Type checking for functions is accomplished by using fresh ``witness''
-values of $\stringty$ and $\cidty$ types to instantiate dependent
-variables, similarly to the \TirName{GenEntails} rule. As defined in
-the \TirName{App} rule, typing ensures that application of the
-function instantiates the dependencies, checks that the instantiation
-of $\notg{\Gamma_1}$ is contained in the share typing environment, and
-adds the instantiation of $\notg{\Gamma_2}$ as an effect, as will as
-the instantiation of $\notg{R}$ with linearity enforced. A practical
-programming detail is where the share type environment precondition
-$\notg{\Gamma_1}$ is defined.  In our implementation, we write these
-as an additional component of \ttt{precondition} annotations, as
-illustrated in extended examples discussed in Section
+$\notg{\Gamma_1}$, application of the function results in the bindings
+in $\notg{\Gamma_2}$ using the random values in $\notg{R}$ for share
+generation or secure openings.  We require that the type is closed,
+i.e., only $y_1,\ldots,y_n$ can occur free in $\notg{\Gamma_1}$,
+$\notg{\Gamma_2}$ $\notg{R}$.  Type checking for functions is
+accomplished by using fresh ``witness'' values of $\stringty$ and
+$\cidty$ types to instantiate dependent variables, similarly to the
+\TirName{GenEntails} rule. As defined in the \TirName{App} rule,
+typing ensures that application of the function instantiates the
+dependencies, checks that the instantiation of $\notg{\Gamma_1}$ is
+contained in the share typing environment, and adds the instantiation
+of $\notg{\Gamma_2}$ as an effect along with instantiation of
+$\notg{R}$ with linearity enforced. A practical programming detail is
+where the share type environment precondition $\notg{\Gamma_1}$ is
+defined by the programmer-- by convention we write these as
+additional components of \ttt{precondition} annotations, as
+illustrated in the extended examples discussed in Section
 \ref{section-examples}.
@@
 valid share typing, as in Figure \ref{fig-gmwsharetypes}. Here a coercion
-is needed to specify that the gate output are shares obtained using
+is needed to specify that the gate output wire values are secret-shared using
 $\ttt{r[z]@1}$ as an OTP. The type for $\ttt{andgmw}$ can then be
@@
   Given $\htrip{\eqspre}{\ttt{main()}}{\eqs}$ and $\ttt{main} : \Gamma_1 \rightarrow (\Gamma_2,R)$
-  and $\main() \redx \prog$ and $\progpre,\Gamma_1,R' : {OK}$ with $R' \cap R = \varnothing$
-  and $\toeq{\progpre} \models \eqspre$. Then $(\progpre;\prog),\Gamma_2,R : {OK}$.
+  and $\main() \redx \prog$ and $\progpre,\Gamma_1,R' : {OK}$ %with $R' \cap R = \varnothing$
+  and $\toeq{\progpre} \models \eqspre$. Then $(\progpre;\prog),\Gamma_2,R \uplus R': {OK}$.
 \end{theorem}
@@
 distribitions is typical, and our approach does not require the
-protocol to be defined in detail, just (weakened) postconditions of it
-that are sufficient to verify the protocol being programmed.
+protocol to be defined by the programmer, but rather just (weakened)
+postconditions of it that are sufficient to verify the protocol being
+programmed.
 
 Theorem \ref{theorem-metaty} also does not provide a complete recipe
-for obtaining noninterference modulo output-- as suggested by Theorem
+for obtaining noninterference modulo output-- as per Theorem
 \ref{theorem-sty}, it is also necessary to check closure of

### 2025-11-24 toplas26/examples.tex
checkpoint

@@
 circuits. In this section we present an extended example of 2-party
-arithmetic circuits with multiplication gates defined using Beaver triples
-\cite{evans2018pragmatic}. We first define a compositional gate
-library in $\metaprot$ with correctness conditions defined using the
-Hoare logic developed in Section \ref{section-metalang}, along with a
-small example circuit.  Then we will extend the library with type
-annotations illustrating the use of share types developed in
-Sections \ref{section-sty} and \ref{section-metaty} to enforce
-confidentiality properties. Last, we will extend the base library
-with additional pre and postcondition annotations to enforce
-integrity properties supported by BDOZ/SPDZ style MACs
-\cite{SPDZ1,SPDZ2,BDOZ,10.1007/978-3-030-68869-1_3}.
+arithmetic circuits with multiplication gates defined using Beaver
+triples \cite{evans2018pragmatic}. We first define a compositional
+gate library in $\metaprot$ with correctness conditions defined using
+the Hoare logic developed in Section \ref{section-metalang}, along
+with a small example circuit (Figure \ref{fig-beaver}).  Then we
+extend the library with type annotations illustrating the use of share
+types developed in Sections \ref{section-sty} and \ref{section-metaty}
+to enforce confidentiality properties (Figure
+\ref{fig-beaver-types}). Lastly, we extend the base library with
+additional pre and postcondition annotations to enforce integrity
+properties supported by BDOZ/SPDZ style MACs
+\cite{SPDZ1,SPDZ2,BDOZ,10.1007/978-3-030-68869-1_3} (Figure
+\ref{fig-beaver-bdoz}).
 
@@
 summary, verification is efficient, largely due to the performance of
-underlying cvc5 verification. It is categorically better performance
+underlying cvc5 verification, with categorically better performance
 than can be achieved using previous brute-force methods developed in
@@
 
-A \ttt{sum} gate can be defined non-interactively using the
+A \ttt{sum} gate can be defined non-interactively using 
 homomorphic addition. The correctness \ttt{postcondition} is simple
@@
 appealing declaratively.  The multiplication gate \ttt{mult} is more
-complex. We use the common approach of \emph{Beaver Triples}
+complex, where we use the common approach of \emph{Beaver Triples}
 \cite{evans2018pragmatic}. Recall that Beaver triples are values
@@
 The circuit itself is constructed in \ttt{main} using these
-gates. This is a small example but arbitrarily large circuits
-can be constructed with this library. The precondition of
-\ttt{main} must require distinct Beaver triples for each
-multiplication gate in the circuit. The postcondition allows
-verification of circuit correctness-- that the protocol
-output is equal to the ideal functionality $\ttt{IDEALF}$.
+gates. This is a small example but arbitrarily large circuits can be
+constructed with this library. The precondition of \ttt{main} requires
+distinct Beaver triples for each multiplication gate in the circuit
+(just one in this example). The postcondition allows verification of
+circuit correctness-- that the protocol output is equal to the ideal
+functionality $\ttt{IDEALF}$.
 

### 2025-11-24 toplas26/examples.tex
checkpoint

@@
 $\ttt{1}$ and $\ttt{2}$. As in the GMW example discussed in Section
-\ref{section-metaty}, we use string idenfiers $g$ to distinguish
+\ref{section-metaty}, we use string identifiers $g$ to distinguish
 gates, and gate output wire shares are defined as identifiers
@@
 homomorphic addition. The correctness \ttt{postcondition} is simple
-and perhaps unecessary to accelerate circuit verification, but is
+and perhaps unnecessary to accelerate circuit verification, but is
 appealing declaratively.  The multiplication gate \ttt{mult} is more
@@
 Section \ref{section-ipj} focuses on the use of BDOZ/SPDZ-style
-MACs. A main concern here is the encorcement of MAC checking which is
+MACs. A main concern here is the enforcement of MAC checking which is
 an algebraic property. Here we show how pre and postconditions can be

### 2025-11-24 toplas26/metaty.tex
checkpoint

@@
 
+\shtripcmdfig
+
 Rules for other commands are defined in Figure \ref{fig-shtripcmd}.
@@
 
-\shtripcmdfig
-
 \subsubsection{Example: GMW \ttt{and} Gate}
@@
 some preprocessing protocol such as Beaver triple or MAC share
-distribitions is typical, and our approach does not require the
+distributions is typical, and our approach does not require the
 protocol to be defined by the programmer, but rather just (weakened)

### 2025-11-24 toplas26/sty.tex
checkpoint

@@
 Our non-interactive typing rules are presented in Figure
-\ref{fig-stylocal}. Judgmements are of the form $\Gamma, \eqs \vdash
+\ref{fig-stylocal}. Judgements are of the form $\Gamma, \eqs \vdash
 \stt : \sty$, where $\Gamma$ is a set of \emph{share type bindings} of
@@
   \TirName{HEMultPub} the result follows by Definition \ref{def-OK} and known
-  semi-homorphic properties of secret sharing \cite{10.1007/978-3-030-68869-1_3}.
+  semi-homomorphic properties of secret sharing \cite{10.1007/978-3-030-68869-1_3}.
   Case \TirName{ShareEntails} follows by correctness of the SMT interpretation of
@@
   By Lemma \ref{lemma-OK} we have that $\prog,\Gamma,R : {OK}$, so the result follows
-  by Therorem \ref{theorem-OK}.
+  by Theorem \ref{theorem-OK}.
 \end{proof}
@@
 
-Based on the preceding Lemma, we can show that noninterence with respect
+Based on the preceding Lemma, we can show that noninteference with respect
 to classified interactive variables is maintained by typing judgements. 

### 2025-11-24 toplas26/examples.tex
checkpoint

@@
 to specify properties of secure openings. After noninteractively
-summing a wire value $\ttt{x}$ and a random Beaver triple value
-$\ttt{x++a}$, the federation can publicly open the value since it is
+summing a wire value $\ttt{RECON(x)}$ and a random Beaver triple value
+$\ttt{r[x++a]@3}$, the federation can publicly open the value since it is
 masked by the random value. This is expressed using the coercion
@@
 $$
-Since $\ttt{m[z++"d"]@i : public}$ and $\ttt{m[z++"e"]@i : public}$
+Since $\ttt{m[z++"d"]@i : public}\ $ and $\ \ttt{m[z++"e"]@i : public}$
 are both established by $\ttt{secopen}$, and

### 2025-11-24 toplas26/examples.tex
checkpoint

--- 2025-11-24 toplas26/examples.tex
@@
 to specify properties of secure openings. After noninteractively
-summing a wire value $\ttt{x}$ and a random Beaver triple value
-$\ttt{x++a}$, the federation can publicly open the value since it is
+summing the secret-shared wire value $\ttt{RECON(x)}$ and Beaver triple value
+$\ttt{r[x++a]@3}$, the federation can publicly open the result since it is
 masked by the random value. This is expressed using the coercion
 $\ttt{as RECON(x) +  r[x++a]@3}$. The coercion then allows the
-opened value to be declared $\ttt{public}$.
+securely opened value to be declared $\ttt{public}$.
 
@@
 $$
-Since $\ttt{m[z++"d"]@i : public}$ and $\ttt{m[z++"e"]@i : public}$
+Since $\ttt{m[z++"d"]@i\,:\,public}\ $ and $\ \ttt{m[z++"e"]@i\,:\,public}$
 are both established by $\ttt{secopen}$, and
@@
 be successfully checked by syntax-directed $\Vdash$ type checking that
-results from the declaration $\ttt{[m[z]@i1,m[z]@i2] : shares}$.
+results from the declaration $\ttt{[m[z]@i1,m[z]@i2]\,:\,shares}$.
 
@@
 MAC of client \ttt{1} are $\ttt{m[x++"s"]@1}$ and $\ttt{m[x++"m"]@1}$
-respectively, and the corresponding key is $\ttt{m[x++"k"]@2}$ is held
-by client \ttt{2}. The correctness condition of this 3-tuple is
-$\ttt{MACOK}(x,2,1)$. The same is true symmetrically, so $\ttt{MACed}$
-establishes both sides of the correctness condition.  In general,
-these are used in gate preconditions to assume that the MAC conditions
-hold for input shares, and output shares are $\ttt{MACed}$. The
+respectively, and the corresponding key $\ttt{m[x++"k"]@2}$ is held by
+client \ttt{2}. The correctness condition of this 3-tuple is
+$\ttt{MACOK}(x,2,1)$. The same is true symmetrically, so the
+$\ttt{MACed}$ predicate establishes both sides of the correctness
+condition.  In general, this predicate is used in gate preconditions
+to assume that the MAC conditions hold for input shares, and in
+postconditions to ensure that output shares are $\ttt{MACed}$. The
 homomorphic properties that preserve $\ttt{MACed}$ness, especially in
 \ttt{mult}, are perhaps even more complex than the algebraic
-correctness of gates, so it is very beneficial to verify them
-compositional especially for large circuits.
+correctness of gates, so it is beneficial to verify them compositional
+especially for large circuits.
 
@@
 we assume that input values, MACs, and keys are distributed to clients
-by a pre-processing protocol.
+by a pre-processing protocol with relevant postconditions
+declared in the preconditions of $\ttt{main}$ (as in Figure
+\ref{fig-beaver} with respect to Beaver triples).

### 2025-11-25 toplas26/sty.tex
checkpoint

@@
   $$
-  \condd{\progd(\prog)}{S_H}{S_C \cup O} = \condd{\progd(\prog)}{S_H}{S_C \cup \houtputs \cup \Phi^p \cup O}
+  \condd{\progd(\prog)}{S_H}{S_C \cup O} = \condd{\progd(\prog)}{S_H}{S_C \cup \houtputs \cup \Phi^o \cup O}
   $$

### 2025-11-25 toplas26/sty.tex
checkpoint

@@
   By definition of $\TirName{Output}$ it is also the case that for all
-  $\out{\cid} \in O$ we have $\toeq{\prog} \models o = \phi$ for some
+  $\out{\cid} \in O$ we have $\toeq{\prog} \models o \eop \phi$ for some
   $\phi \in \Phi^o$. Therefore we have:

### 2025-11-25 toplas26/model.tex
checkpoint

@@
 \end{definition}
-We also define some convenient syntactic sugaring for ``implicit
-marginalization'' given a realization.
+We also define some convenient syntactic sugaring for application of
+pmfs to realizations when their domains don't match.
 %The first will allow us to
@@
 \begin{definition}
-  Given $\pmf$ and $X \subseteq \dom(\pmf)$, for all $\store \in \mems(X)$
-  define $\pmf(\store)  \defeq \margd{\pmf}{X}(\store)$.
+  Given $\pmf$ and $\store$ with $\dom(\pmf) \ne \dom(\store)$,
+  $\pmf(\store) \defeq \margd{\pmf}{\dom(\store) \cap \dom(\pmf)}(\store_{\dom(\store) \cap \dom(\pmf)})$.
+  %\defeq \margd{\pmf}{\dom(\store)}(\store)$
+  %if $\dom(\store) \subset \dom(\pmf)$, and define $\pmf(\store) \defeq \pmf(\store_{\dom(\pmf)})$ if
+  %$\dom(\pmf) \subset \dom(\store)$.
   %and
@@
 \begin{definition}
+  \label{def-condsep}
 %  We write $\vc{\pmf}{x}{y}$ iff $\pmf(\{ x \mapsto 0\}\ |\ \{ y \mapsto 0 \}) =
 %  \pmf(\{ x \mapsto 1\}\ |\ \{ y \mapsto 1 \}) = 1$.
-  We write $\condsep{\pmf}{X_1}{X_2}{X_3}$ iff for all
-    $\store \in \mems(X_1 \cup X_2)$ and $\store' \in \mems(X_3)$ we have
-  $\pmf(\store|\store') =
-  \pmf(\store_{X_1}|\store') * \pmf(\store_{X_2}|\store')$. If $X_1$ is empty
+  We write $\condsep{\pmf}{X_1}{X_2}{X_3}$ iff $\condd{\pmf}{X_2 \cup X_3}{X_1}
+  = \condd{\pmf}{X_2}{X_1} * \condd{\pmf}{X_3}{X_1}$. If $X_1$ is empty
   we write $\sep{\pmf}{X_2}{X_3}$.
@@
   a partitioning of the federation clients, define:
+  $$
   \begin{array}{lcl}
@@
   Given a partitioning of the federation $H,C$, we call $\houtputs$
-  \emph{corrupt views} and $\hinputs$ \emph{honest views}.
+  \emph{corrupt views} and $\cinputs$ \emph{honest views}.
 \end{definition}
@@
   satisfies \emph{noninterference modulo output}
-  iff for all $H$ and $C$ %and $\store_1 \in \mems(S_C \cup O)$ and $\store_2 \in \mems(\houtputs)$
+  iff for all $H$ and $C$ and $\store \in \runs(\prog)$ 
   we have:
   $$
-  \condd{\progtt(\prog)}{S_H}{S_C \cup O} = \condd{\progtt(\prog)}{S_H}{S_C \cup O \cup \houtputs}
+  \condd{\progtt(\prog)}{S_H}{S_C \cup O}(\store) = \condd{\progtt(\prog)}{S_H}{S_C \cup O \cup \houtputs}(\store)
  $$
@@
   \emph{integrity} iff for all $H$, $C$, and $\adversary$,
-  if $\store \in \aruns(\prog)$ we have:
+  if $\store \in \aruns(\prog)$ there exists $\store' \in \runs(\prog)$ with
+  $\store_{S_H \cup X}' = \store_{S_H \cup X}$ and:
     $$
-    \condd{\progtt(\prog,\adversary)}{X}{S_H \cup \cinputs} =
-    \condd{\progtt(\prog)}{X}{S_H \cup \cinputs}
+    \condd{\progtt(\prog,\adversary)}{X}{S_H \cup Y}(\store) =
+    \condd{\progtt(\prog)}{X}{S_H \cup Y}(\store')
     $$ 
-  where $X \defeq (\houtputs \cup O_H) \cap \dom(\store)$. 
+  where $X \defeq (\houtputs \cup O_H) \cap \dom(\store)$ and $Y = \cinputs \cap \dom(\store)$. 
 \end{definition}

### 2025-11-25 toplas26/sty.tex
checkpoint

--- 2025-11-25 toplas26/sty.tex
@@
 conditions are obviously met by, e.g., ``vanilla'' additive or Shamir
-shares which generate shares in uniform marginal distributions.
+shares which generate shares in uniform marginal distributions. Here
+and later we leverage (conditional) probabilistic independence as
+characterized in Definition
+\ref{def-condsep}.
 \begin{definition}
@@
     %\Phi_a^p & \defeq & \setit{ \phi_b \mid (\pubx{\recon(\phi_1,\phi_2)} : \sty) \in \Gamma}\\
-    \Phi_a & \defeq & \setit{ \phi_a \mid ([\phi_1,\phi_2] : \sharety) \in \Gamma} \cup
-                        \setit{ \phi_b \mid (\pubx{\recon(\phi_1,\phi_2)} : \sty) \in \Gamma} \\% \Phi_a^m \cup \Phi_a^p \\
+    \Phi_b & \defeq & \setit{ \phi_b \mid ([\phi_1,\phi_2] : \sharety) \in \Gamma} \cup
+                        \setit{ \phi_a \mid (\pubx{\recon(\phi_1,\phi_2)} : \sty) \in \Gamma} \\% \Phi_a^m \cup \Phi_a^p \\
     \Phi^p & \defeq & \setit{ \recon(\phi_1,\phi_2) \mid (\pubx{\recon(\phi_1,\phi_2)} : \pubty) \in \Gamma}\\
@@
     $
-    \condd{\progtt(\prog)}{S_{\setit{\cid_a}}}{S_{\setit{\cid_b}} \cup \Phi^o}
-    = 
-    \condd{\progtt(\prog)}{S_{\setit{\cid_a}}}{S_{\setit{\cid_b}} \cup  \Phi_a \cup \Phi^o}
+    \condsep{\progtt(\prog)}{(S_{\setit{b}} \cup \Phi^o)}{\Phi_b}{S_{\setit{a}}}
     $.
+    %$
+    %\condd{\progtt(\prog)}{S_{\setit{a}}}{S_{\setit{b}} \cup \Phi^o}
+    %= 
+    %\condd{\progtt(\prog)}{S_{\setit{a}}}{S_{\setit{b}} \cup  \Phi_a \cup \Phi^o}
+    %$.
     \item For all $a,b \in \setit{1,2}$ and $([\phi_1,\phi_2] : \sharety) \in \Gamma$ we have:
-    %$\condsep{\progd(\prog)}{\setit{\recon(\phi_i^1,\phi_i^2)}}{\setit{\phi_i^b}}{\recon(\phi_i^1,\phi_i^2)}$
-    $$\condd{\progtt(\prog)}{S_{\setit{\cid_a}}}{S_{\setit{\cid_b}} \cup  \Phi_a \cup \Phi^o \cup \setit{\recon(\phi_i^1,\phi_i^2)}} =
-    \condd{\progtt(\prog)}{S_{\setit{\cid_a}}}{S_{\setit{\cid_b}}  \cup  \Phi_a \cup \Phi^o \cup \setit{\recon(\phi_i^1,\phi_i^2), \phi_i^b}}$$
+    $$\condsep{\progd(\prog)}{(S_{\setit{b}} \cup  \Phi_b \cup \Phi^o \cup \setit{\recon(\phi_i^1,\phi_i^2)})}{\phi_i^a}{S_{\setit{a}}}$$
+    %$$\condd{\progtt(\prog)}{S_{\setit{a}}}{S_{\setit{b}} \cup  \Phi_a \cup \Phi^o \cup \setit{\recon(\phi_i^1,\phi_i^2)}} =
+    %\condd{\progtt(\prog)}{S_{\setit{a}}}{S_{\setit{b}}  \cup  \Phi_a \cup \Phi^o \cup \setit{\recon(\phi_i^1,\phi_i^2), \phi_i^b}}$$
   \item $\sep{\progd(\prog)}{(S \cup O)}{\Phi^p}$
@@
     %\Phi_a^p & \defeq & \setit{ \phi_b \mid (\pubx{\recon(\phi_1,\phi_2)} : \sty) \in \Gamma}\\
-    \Phi_a & \defeq & \setit{ \phi_a \mid ([\phi_1,\phi_2] : \sharety) \in \Gamma} \cup
-                        \setit{ \phi_b \mid (\pubx{\recon(\phi_1,\phi_2)} : \sty) \in \Gamma} \\% \Phi_a^m \cup \Phi_a^p \\
+    \Phi_b & \defeq & \setit{ \phi_b \mid ([\phi_1,\phi_2] : \sharety) \in \Gamma} \cup
+                        \setit{ \phi_a \mid (\pubx{\recon(\phi_1,\phi_2)} : \sty) \in \Gamma} \\% \Phi_a^m \cup \Phi_a^p \\
     \Phi^o & \defeq & \setit{ \recon(\phi_1,\phi_2) \mid (\pubx{\recon(\phi_1,\phi_2)} : \outty) \in \Gamma}
@@
       $
-    \condd{\progtt(\prog)}{S_{\setit{\cid_a}}}{S_{\setit{\cid_b}} \cup \Phi^o}
-    = 
-    \condd{\progtt(\prog)}{S_{\setit{\cid_a}}}{S_{\setit{\cid_b}} \cup  \Phi_a \cup \setit{\phi_a} \cup \Phi^o}
+      \condsep{\progtt(\prog)}{(S_{\setit{b}} \cup \Phi^o)}{(\Phi_b \cup \setit{\phi_b})}{S_{\setit{a}}}
       $
+    %$
+    %\condd{\progtt(\prog)}{S_{\setit{a}}}{S_{\setit{b}} \cup \Phi^o}
+    %= 
+    %\condd{\progtt(\prog)}{S_{\setit{a}}}{S_{\setit{b}} \cup  \Phi_a \cup \setit{\phi_a} \cup \Phi^o}
+    %$
     \item
-    $\condd{\progtt(\prog)}{S_{\setit{\cid_a}}}{S_{\setit{\cid_b}} \cup  \Phi_a \cup \Phi^o \cup \setit{\recon(\phi_1,\phi_2)}} =
-    \condd{\progtt(\prog)}{S_{\setit{\cid_a}}}{S_{\setit{\cid_b}}  \cup  \Phi_a \cup \Phi^o \cup \setit{\recon(\phi_1,\phi_2), \phi_b}}$
+       $\condsep{\progtt(\prog)}{(S_{\setit{b}} \cup  \Phi_b \cup \Phi^o \cup \setit{\recon(\phi_1,\phi_2)})}{\phi_a}{S_{\setit{a}}}$
+    %$\condd{\progtt(\prog)}{S_{\setit{a}}}{S_{\setit{b}} \cup  \Phi_a \cup \Phi^o \cup \setit{\recon(\phi_1,\phi_2)}} =
+    %\condd{\progtt(\prog)}{S_{\setit{a}}}{S_{\setit{b}}  \cup  \Phi_a \cup \Phi^o \cup \setit{\recon(\phi_1,\phi_2), \phi_b}}$
     \end{enumerate}
@@
   $$
-  \condd{\progd(\prog)}{S_H}{S_C \cup O} = \condd{\progd(\prog)}{S_H}{S_C \cup \houtputs \cup \Phi^o \cup O}
+  \condsep{\progd(\prog)}{(S_C \cup \Phi^o \cup O)}{\houtputs}{S_H}
   $$
   By definition of $\TirName{Output}$ it is also the case that for all
-  $\out{\cid} \in O$ we have $\toeq{\prog} \models o = \phi$ for some
-  $\phi \in \Phi^o$. Therefore we have:
+  $\out{\cid} \in O$ we have $\toeq{\prog} \models o \eop \phi$ for some
+  $\phi \in \Phi^o$. Therefore since we can assume wlog that all revealed
+  values are assigned to outputs, we have
+  $
+  \condsep{\progd(\prog)}{(S_C \cup O)}{\houtputs}{S_H}
+  $,
+  so by definition we have that for all $\store \in \runs(\prog)$: 
   $$
-  \condd{\progd(\prog)}{S_H}{S_C \cup O} = \condd{\progd(\prog)}{S_H}{S_C \cup \houtputs \cup O}
+  \condd{\progd(\prog)}{S_H}{S_C \cup O}(\store) = \condd{\progd(\prog)}{S_H}{S_C \cup \houtputs \cup O}(\store)
   $$

### 2025-11-25 toplas26/examples.tex
checkpoint

@@
 of the shared computations $\ttt{(m[z++"d"] * m[y])@i}$ and $\ttt{(-m[z++"e"] * m[z++"a"])@i}$
-are shares, so the gate outputs are shares, and this will
+are shares, so the gate outputs, which are then sums of shares, are shares, and this will
 be successfully checked by syntax-directed $\Vdash$ type checking that

### 2025-11-25 toplas26/lang.tex
checkpoint

@@
 \item $\eassign{\mesg{w}}{\cid_2}{\be}{\cid_1}$: This
-  is a \emph{message send} where expression $\be$ is computed
+  is a unicast \emph{message send} where expression $\be$ is computed
   by client $\cid_1$ and sent to client $\cid_2$ as message
@@
 use of undefined variables and, in the adversarial semantics,
-programs consisting of just a failing $\assert$.
+programs consisting of just a failing $\ttt{assert}$.

### 2025-11-25 toplas26/model.tex
checkpoint

@@
   A \emph{probability mass function} $\pmf$ is a function
-  mapping memories in $\mems(X)$ for given variables $X$ to
+  mapping memories in $\mems(X)$ (the \emph{domain} of $\pmf$), to
   values in $\mathbb{R}$ such that:
@@
   %\pmf(\{ x_1 \mapsto v_1, \ldots, x_n \mapsto v_n\}) > 0 \} $
+  We write $\rvs(\pmf) = X$ to denote the variables
+  over which $\pmf$ is defined. 
 \end{definition}
@@
 \begin{definition}
-  Given $\pmf$ with $\dom(\pmf) = \mems(X_2)$, the \emph{marginal distribution}
+  Given $\pmf$ with $\rvs(\pmf) = X_2$, the \emph{marginal distribution}
   of variables $X_1 \subseteq X_2$ in $\pmf$ is denoted $\margd{\pmf}{X_1}$ and defined as follows:
@@
   Given $\pmf$, the \emph{conditional distribution}
-  of $X_1$ given $X_2$ where $X_1 \cup X_2 = \dom(\pmf)$ 
+  of $X_1$ given $X_2$ where $X_1 \cup X_2 = \rvs(\pmf)$ 
   is denoted $\condd{\pmf}{X_1}{X_2}$ and defined as follows:
@@
 \begin{definition}
-  Given $\pmf$ and $\store$ with $\dom(\pmf) \ne \dom(\store)$,
-  $\pmf(\store) \defeq \margd{\pmf}{\dom(\store) \cap \dom(\pmf)}(\store_{\dom(\store) \cap \dom(\pmf)})$.
+  Given $\pmf$ and $\store$ with $\rvs(\pmf) \ne \dom(\store)$,
+  $\pmf(\store) \defeq \margd{\pmf}{\dom(\store) \cap \rvs(\pmf)}(\store_{\dom(\store) \cap \rvs(\pmf)})$.
   %\defeq \margd{\pmf}{\dom(\store)}(\store)$
-  %if $\dom(\store) \subset \dom(\pmf)$, and define $\pmf(\store) \defeq \pmf(\store_{\dom(\pmf)})$ if
+  %if $\dom(\store) \subset \rvs(\pmf)$, and define $\pmf(\store) \defeq \pmf(\store_{\rvs(\pmf)})$ if
   %$\dom(\pmf) \subset \dom(\store)$.

### 2025-11-25 toplas26/sty.tex
checkpoint

@@
   Let $\iov(\prog) = S \cup V \cup O$ and assume some partitioning $H,C$
-  with $H = \setit{b}$ and $C = \setit{a}$ for $a,b \in \setit{1,2}$.
+  with $H = \setit{a}$ and $C = \setit{b}$ for $a,b \in \setit{1,2}$.
   %For any $\out{\cid} \in O$ with $\Gamma,\toeq{\prog} \vdash \pubx{\out{\cid}} : \outty$
@@
   %\ref{lemma-OK}.
-  Let $\Phi_a$ and $\Phi^o$ be as defined in Definition \ref{def-OK}. 
+  Let $\Phi_b$ and $\Phi^o$ be as defined in Definition \ref{def-OK}. 
   By Definition \ref{def-sty-closed} we have that $x \in \houtputs$
-  implies $x \in \Phi_a$ and $(\out{\cid} : \outty) \in \Gamma$ for all
+  implies $x \in \Phi_b$ and $(\out{\cid} : \outty) \in \Gamma$ for all
   $\out{\cid} \in O$. 

### 2025-11-25 toplas26/model.tex
checkpoint

@@
 \begin{definition}
-  A \emph{probability mass function} $\pmf$ is a function
-  mapping memories in $\mems(X)$ (the \emph{domain} of $\pmf$), to
-  values in $\mathbb{R}$ such that:
+  A \emph{probability mass function} $\pmf$ over variables $X$
+  (where we define $\rvs(\pmf) \defeq X$) is a function
+  with domain $\mems(X)$ and range $\mathbb{R}$ such that:
   $$
@@
   %\pmf(\{ x_1 \mapsto v_1, \ldots, x_n \mapsto v_n\}) > 0 \} $
-  We write $\rvs(\pmf) = X$ to denote the variables
-  over which $\pmf$ is defined. 
+  %We write $\rvs(\pmf) = X$ to denote the variables
+  %over which $\pmf$ is defined. 
 \end{definition}

### 2025-11-25 toplas26/sty.tex
checkpoint

@@
   $.
-  Given pmf $\pmf$ and $\Phi$ with $\dom(\pmf) = \bigcup_{\phi \in \Phi} \dom(\phi)$,
+  Given pmf $\pmf$ and $\Phi$ with $\rvs(\pmf) = \bigcup_{\phi \in \Phi} \dom(\phi)$,
   for all $\store \in \mems(\Phi)$ define:
@@
   Given $\pmf$, the \emph{conditional distribution}
-  of $X$ given $\Phi$ where $\bigcup_{\phi \in \Phi} \dom(\phi) \cup X = \dom(\pmf)$ 
+  of $X$ given $\Phi$ where $\bigcup_{\phi \in \Phi} \dom(\phi) \cup X = \rvs(\pmf)$ 
   is denoted $\condd{\pmf}{X}{\Phi}$ and defined as follows: 

### 2025-12-01 toplas26/model.tex
checkpoint

@@
   $\progtt(\prog,\adversary)$, is
-  defined such that for all $\store \in \mems(\iov(\prog) \cup R)$:
+  defined such that for all $\store \in \mems(\vars(\prog))$:
   $$
@@
     \{ \store\{ x_1 \mapsto \bot, \ldots, x_n \mapsto \bot \} \mid 
-    \store \in \aruns(\prog) \wedge \{ x_1,\ldots,x_n\} = (V \cup O) - \dom(\store) \}
+    \store \in \aruns(\prog) \wedge \{ x_1,\ldots,x_n\} = \iov(\prog) - \dom(\store) \}
   \end{array}

### 2025-12-07 toplas26/examples.tex
yet another update

@@
-\section{Extended Examples}
+\section{Implementation and Extended Examples}
 \label{section-examples}
@@
 
-\paragraph{Implementation and Evaluation}
+\subsection{Evaluation}
 We have implemented the $\metaprot$ language and automated Hoare logic
 using an Antlr parser and Scala backend with cvc5 Java bindings for
-interaction with the Satisfiability Modulo Fields theory.  To evaluate
-performance, we have written and verified gate libraries and small
-circuits using the GMW protocol for binary circuits as excerpted in
+interaction with proof engines.  To evaluate
+performance, we have written and verified gate libraries
+using the GMW protocol for binary circuits as excerpted in
 Figure \ref{fig-gmwprepost} in Section \ref{section-metalang}, and for
@@
 typically used to encode binary, 16 bit, 32 bit, and 64 bit
-integers. Results are presented in Figure \ref{fig-preludetest}. In
-summary, verification is efficient, largely due to the performance of
-underlying cvc5 verification, with categorically better performance
-than can be achieved using previous brute-force methods developed in
-$\fieldp{2}$ \cite{skalka-near-ppdp24}.
+integers.
+
+In addition to libraries, we have verified 2, 8, and 64-bit vector
+zero-test GMW circuits containing 3, 15, and 127 gates respectively,
+based on ``Bristol Fashion'' circuits \cite{bristol}. By verified, we
+mean automated verification of the zero-test specification expressed
+as a \ttt{postcondition}. We considered both the cvc5 Bit-Vector and
+Satisfiability Modulo Finite Fields (SMFF) theories as proof engines
+for these binary circuits.
+
+Results are presented in Figure \ref{fig-preludetest}. We found that
+verification of library components was efficient even up to large
+field sizes. Verification of the zero-test circuits revealed a clear
+performance advantage for the Bit-Vector theory, which verified
+the 64-bit zero test in GMW in less than 7 seconds. In contrast,
+it took several minutes to verify even the small 2-bit zero-test
+circuit in SMFF, and larger circuits took too long to consider
+for a practical evaluation. In all cases, the SMT-based approach
+supported categorically better performance
+than can be achieved using previous brute-force methods developed for
+$\fieldp{2}$ and binary circuits \cite{skalka-near-ppdp24}.
 

### 2025-12-07 toplas26/intro.tex
yet another update

@@
 and fully develops the compositional type-based verification approach
-in Section 8. In addition, we develop an implementation capable of
+in Section 8. In addition, we develop an algorithm capable of
 automating the verification of examples given in the prior work, and
@@
 
-In all cases, our type systems are boosted by verification mechanisms
-provided by Satisfiability Modulo Finite Fields \cite{SMFF}. In
-Section \ref{section-examples} we develop and discuss extended example
-applications of our type analyses to real protocols including
-the Goldreich-Micali-Wigderson (GMW), Ben-Or-Goldwasser-Widgerson (BGW),
-and Bendlin-Damgard-Orland-Zakarias (BDOZ), and report related
-performance metrics.
+In Section \ref{section-examples} we develop and discuss extended
+example applications of our type analyses to real protocols including
+the Goldreich-Micali-Wigderson (GMW), Ben-Or-Goldwasser-Widgerson
+(BGW), and Bendlin-Damgard-Orland-Zakarias (BDOZ). We also summarize
+our implementation of the $\metaprot$ Hoare logic and evaluate
+verification of both binary and arithmetic gate and circuit
+properties. Our implementation integrates with cvc5 SMT to support
+efficient verification, including Satisfiability Modulo Finite Fields
+(SMFF) \cite{SMFF} for binary and arithmetic circuits, and the
+Bit-Vector theory for binary circuits.
 

### 2025-12-08 toplas26/intro.tex
typo fix

@@
 \ref{fig-comp-wrap} include a Hoare logic and a more expressive
-confidentiality type systems supporting analysis of more complex
+confidentiality type system supporting analysis of more complex
 programs that enforces passive security in the real/ideal model, vs.~a

### 2025-12-11 toplas26/examples.tex
yet another update

@@
 using an Antlr parser and Scala backend with cvc5 Java bindings for
-interaction with proof engines.  To evaluate
-performance, we have written and verified gate libraries
-using the GMW protocol for binary circuits as excerpted in
-Figure \ref{fig-gmwprepost} in Section \ref{section-metalang}, and for
-passive and malicious secure arithmetic circuits using Beaver Triple
-multiplication as presented in Figures \ref{fig-beaver} and
-\ref{fig-beaver-bdoz} in this Section.  We have considered field sizes
-typically used to encode binary, 16 bit, 32 bit, and 64 bit
-integers.
+interaction with proof engines.  To evaluate performance, we have
+written and verified gate libraries using the GMW protocol for binary
+circuits as excerpted in Figure \ref{fig-gmwprepost} in Section
+\ref{section-metalang}, and for passive and malicious secure
+arithmetic circuits using Beaver Triple multiplication as presented in
+Figures \ref{fig-beaver} and \ref{fig-beaver-bdoz} in this Section.
+We have considered field sizes typically used to encode binary, 16
+bit, 32 bit, and 64 bit integers. Results are presented in Figure
+\ref{fig-preludetest}. We found that verification of library
+components was efficient even up to large field sizes. Since typing is
+compositional (see, e.g., Figure \ref{fig-beaver-types}), verification
+of gate properties relevant to \emph{security} will scale to larger
+circuits.
 
-In addition to libraries, we have verified 2, 8, and 64-bit vector
-zero-test GMW circuits containing 3, 15, and 127 gates respectively,
-based on ``Bristol Fashion'' circuits \cite{bristol}. By verified, we
-mean automated verification of the zero-test specification expressed
-as a \ttt{postcondition}. We considered both the cvc5 Bit-Vector and
-Satisfiability Modulo Finite Fields (SMFF) theories as proof engines
-for these binary circuits.
+Circuit \emph{correctness} also benefits from compositional
+verification-- for example, the correctness of a GMW ``and'' gate (as
+in Figure \ref{fig-gmwprepost}) only needs to be proven once, though
+instantiated and integrated into the whole-program analysis needed to
+verify a large circuit specification. We have verified 2, 8, and
+64-bit vector zero-test GMW circuits containing 3, 15, and 127 gates
+respectively, based on ``Bristol Fashion'' circuits \cite{bristol}. By
+verified, we mean automated verification of the zero-test
+specification expressed as a \ttt{postcondition}. We considered both
+the cvc5 Bit-Vector and Satisfiability Modulo Finite Fields (SMFF)
+theories as proof engines for these binary circuits.
 
-Results are presented in Figure \ref{fig-preludetest}. We found that
-verification of library components was efficient even up to large
-field sizes. Verification of the zero-test circuits revealed a clear
-performance advantage for the Bit-Vector theory, which verified
-the 64-bit zero test in GMW in less than 7 seconds. In contrast,
-it took several minutes to verify even the small 2-bit zero-test
-circuit in SMFF, and larger circuits took too long to consider
-for a practical evaluation. In all cases, the SMT-based approach
-supported categorically better performance
-than can be achieved using previous brute-force methods developed for
-$\fieldp{2}$ and binary circuits \cite{skalka-near-ppdp24}.
+Verification of the zero-test circuits was efficient in the Bit-Vector
+theory, which verified the 64-bit zero test in GMW in less than 7
+seconds. However, it took several minutes to verify even the small
+2-bit zero-test circuit in SMFF, and larger circuits took too long to
+consider for a practical evaluation. This could be attributed to the
+relative maturity of the Bit-Vector theory as compared to SMFF.  In
+any case, the SMT-based approach achieves categorically better
+performance than previous brute-force methods developed for binary
+circuits \cite{skalka-near-ppdp24}.
 

### 2025-12-11 toplas26/examples.tex
yet another update

@@
 \label{section-examples}
-
-\preludetestfig
 
@@
 \subsection{Evaluation}
+
 We have implemented the $\metaprot$ language and automated Hoare logic
@@
 circuits.
+
+\preludetestfig
 
